### 2022-10-05 Article.Rmd
Commit 2022-10-05 11:07

---
title:         "Trends of SDR over Thessaloniki"
# author:        "Natsis Athanasios"
# institute:     "AUTH"
# affiliation:   "Laboratory of Atmospheric Physics"
abstract:      "Read "

documentclass:  article
classoption:    a4paper,oneside
fontsize:       10pt
geometry:       "left=0.5in,right=0.5in,top=0.5in,bottom=0.5in"
link-citations: yes
colorlinks:     yes

header-includes:
- \usepackage{caption}
- \usepackage{placeins}
- \captionsetup{font=small}

output:
  bookdown::pdf_document2:
    number_sections:  no
    fig_caption:      no
    keep_tex:         no
    keep_rmd:         yes
    latex_engine:     xelatex
    toc:              yes
    fig_width:        7
    fig_height:       4.5
  html_document:
    toc:        true
    fig_width:  7.5
    fig_height: 5
date: "`r format(Sys.time(), '%F')`"

---

```{r echo=F, include=T}

####_  Document options _####

```
```{r echo=F, include=F}
knitr::opts_chunk$set(comment    = ""      )
knitr::opts_chunk$set(dev        = "pdf"   )
# knitr::opts_chunk$set(dev        = "png"    )
knitr::opts_chunk$set(out.width  = "85%"   )
knitr::opts_chunk$set(fig.align  = "center" )
knitr::opts_chunk$set(cache      =  T       )
# knitr::opts_chunk$set(fig.pos    = '!h'    )

```
```{r include=F, echo=F}
####  Set environment  ####
Sys.setenv(TZ = "UTC")
tic <- Sys.time()
Script.Name <- tryCatch({ funr::sys.script() },
                        error = function(e) { cat(paste("\nUnresolved script name: ", e),"\n\n")
                            return("Climatological_") })
if(!interactive()) {
    pdf(  file = paste0("~/Aerosols/REPORTS/runtime/", basename(sub("\\.R$",".pdf", Script.Name))))
    sink( file = paste0("~/Aerosols/REPORTS/runtime/", basename(sub("\\.R$",".out", Script.Name))), split=TRUE)
    filelock::lock(paste0("~/Aerosols/LOGs/",  basename(sub("\\.R$",".lock", Script.Name))), timeout = 0)
}

par(pch = ".")

## FIXME this is for pdf output
# options(warn=-1) ## hide warnings
# options(warn=2)  ## stop on warnings

```
```{r echo=F, include=T}
####  External code  ####
library(data.table, quietly = T, warn.conflicts = F)
library(pander,     quietly = T, warn.conflicts = F)

## Functions from `https://github.com/thanasisn/IStillBreakStuff/tree/main/FUNCTIONS/R`
source("~/CODE/FUNCTIONS/R/sumNA.R")
source("~/CODE/FUNCTIONS/R/linear_regrassion_capture.R")
source("~/CODE/FUNCTIONS/R/trig_deg.R")

####  Variables  ####
panderOptions('table.alignment.default', 'right')
panderOptions('table.split.table',        120   )

DATA_DIR  <- "/home/athan/DATA/Broad_Band"
CLEARdir  <- "/home/athan/DATA/Broad_Band/CS_id"
tag       <- paste0("Natsis Athanasios LAP AUTH ", strftime(Sys.time(), format = "%b %Y" ))
CS_file   <- "/home/athan/DATA/Common_application/Clear_Sky.Rds"

TEST      <- TRUE

MIN_ELEVA  <- 5  ## use data when sun above that
SZA_BIN    <- 1
MIN_N      <- 4
SEAS_MIN_N <- 3

SUM_THRES  <- 50
SZA_THRES  <- 3
```

## Data selection

```{r echo=F, include=T}

####  Get data from Clear sky id data  ####
input_files <- list.files( path    = CLEARdir,
                           pattern = "Clear_Sky_[0-9]{4}.Rds",
                           full.names = T )

if ( !file.exists(CS_file) | max(file.mtime(input_files)) > file.mtime(CS_file)) {
    cat(paste("Load data from Clear Sky proccess from original\n"))
    DATA <- data.table()
    for (af in input_files) {
        temp <- readRDS(af)
        ## drop some data
        temp$CHP1temp        <- NULL
        temp$CHP1tempSD      <- NULL
        temp$CHP1tempUNC     <- NULL
        temp$CS_ref          <- NULL
        temp$Pressure_Source <- NULL
        temp$chp1TempCF      <- NULL
        temp$wattDIR_tmp_cr  <- NULL
        temp$wattDIR_unc_NT  <- NULL
        temp$wattDIR_unc_WT  <- NULL
        temp$wattHOR_tmp_cr  <- NULL
        temp$wattHOR_unc_NT  <- NULL
        temp$wattHOR_unc_WT  <- NULL

        DATA <- rbind(temp, DATA, fill = TRUE)
        rm(temp)
    }
    ## this is used by old scripts
    myRtools::write_RDS(object = DATA, file = CS_file)
} else {
    cat(paste("Load data from Clear Sky proccess from parsed\n"))
    DATA <- readRDS(CS_file)
}
```

### Filter min elevation
Keep data with Sun elevation above `r MIN_ELEVA`

```{r }
DATA <- DATA[ Elevat >= MIN_ELEVA ]
```

### Bais paper obstacle filter

```{r }
DATA <- DATA[ ! (Azimuth > 35 & Azimuth < 120 & Elevat < 10) ]
```
```{r echo=F, include=T}
```

### Keep only data characterized as 'good' by the Radiation Quality control procedure

```{r echo=F, include=T}
DATA[ QCF_DIR != "good", wattDIR := NA ]
DATA[ QCF_DIR != "good", QCF_DIR := NA ]
DATA[ QCF_GLB != "good", wattGLB := NA ]
DATA[ QCF_GLB != "good", QCF_GLB := NA ]
DATA <- DATA[ ! (is.na(wattDIR) & is.na(wattGLB))  ]

## exclude some data for each instrument
# DATA[ QCF_DIR == "Extremely rare limits min (3)", wattDIR := NA ]
# DATA[ QCF_DIR == "Extremely rare limits min (3)", QCF_DIR := NA ]
#
# dirtb <- table(DATA$QCF_DIR)
# for (aty in names(dirtb)[!names(dirtb) %in% "good"]) {
#     temp <- DATA[ QCF_DIR == aty ]
#     cat(nrow(temp),aty,"\n")
# }
# glbtb <- table(DATA$QCF_GLB)
# for (aty in names(glbtb)[!names(glbtb) %in% "good"]) {
#     temp <- DATA[ QCF_GLB == aty ]
#     cat(nrow(temp),aty,"\n")
# }
```

## Data preparation
### Move measurements to mean earth distance

```{r }
DATA[ , wattDIR_1au := wattDIR * (sun_dist ^ 2) ]
DATA[ , wattGLB_1au := wattGLB * (sun_dist ^ 2) ]
```
```{r echo=F, include=T}
```

### Relative to actual TSI at 1au variable representation

```{r }
DATA[ , DIR_att := wattDIR_1au / tsi_1au_comb ]
DATA[ , GLB_att := wattGLB_1au / tsi_1au_comb ]
```

Using the actual values gives similar trends.

```{r echo=F, include=T}

# #' ### Use original variable representation
# DATA[ , DIR_att := wattDIR_1au ]
# DATA[ , GLB_att := wattGLB_1au ]
# #+ echo=F, include=T

# #' ### Ground effect removal?
# #' Aerosol direct effects on global solar shortwave irradiance at high mountainous station Musala Bulgaria_Nojarov2021.pdf
# DATA$wattGLB_1au <- DATA$wattGLB_1au / cosde(DATA$SZA)
# DATA$wattDIR_1au <- DATA$wattDIR_1au / cosde(DATA$SZA)
# #+ echo=F, include=T
```

### Calculate Bouguer atmospheric transparency
see: Changes in solar radiation and their influence on temperature trend in Estonia 1955 2007_Russak2009.pdf

```{r }
DATA[, DIR_transp := ( wattDIR_1au / tsi_1au_comb ) ^ ( 1 / cosde(SZA) ) ]
```
```{r echo=F, include=T}

## fix noon just in case
DATA[ Azimuth <= 180 , preNoon := TRUE  ]
DATA[ Azimuth >  180 , preNoon := FALSE ]
```

### Split data to Clear Sky and non Clear sky
Method based and adapted to site from: Identification of Periods of Clear Sky
Irradiance in Time Series of GHI Measurements
_Matthew J. Reno and Clifford W. Hansen_.

```{r echo=F, include=T}
DATA_all   <- DATA

wecare     <- grep("CSflag_", names(DATA), value = T)
## use only cm21 flags
wecare     <- grep("_11", wecare, invert = T, value = T)
DATA_Clear <- DATA[ rowSums( DATA[, ..wecare ], na.rm = T ) == 0, ]
```
```{r echo=F, include=T}

## old flags
# DATA_Clear <- DATA_all[ CSflag == 0 ]
rm(DATA)

# #### . . My CSid method   ####
# #'
# #' | CS Flag | Test |
# #' |:-------:|:--------------------------------------------------------------|
# #' |   NA    | Undefined, untested                                           |
# #' |    0    | Passed as clear sky                                           |
# #' |    1    | Mean value of irradiance during the time period (MeanVIP)     |
# #' |    2    | Max Value of Irradiance during the time Period (MaxVIP)       |
# #' |    3    | Variability in irradiance by the length (VIL)                 |
# #' |    4    | Variance of Changes in the Time series (VCT)                  |
# #' |    5    | Variability in the Shape of the irradiance Measurements (VSM) |
# #' |    6    | Low Direct Irradiance limit (LDI)                             |
# #' |    7    | Low Global Irradiance limit (LGI)                             |
# #' |    8    | Too Few CS point for the day (FCS)                            |
# #' |    9    | Too Few data points for the day                               |
# #' |   10    | Missing Data                                                  |
# #' |   11    | Direct irradiance simple threshold                            |

#### 1. Long term anomaly trends ####
```

## 1. Long term anomaly trends

```{r }
##TODO t-test and significance
```

#### Calculate daily means ####

```{r echo=F, include=T}

ALL_daily_mean <- DATA_all[, .(DIR_att       = mean(DIR_att,    na.rm = T),
                               GLB_att       = mean(GLB_att,    na.rm = T),
                               DIR_transp    = mean(DIR_transp, na.rm = T),
                               DIR_att_sd    = sd(  DIR_att,    na.rm = T),
                               GLB_att_sd    = sd(  GLB_att,    na.rm = T),
                               DIR_transp_sd = sd(DIR_transp, na.rm = T),
                               doy           = yday(Date),
                               GLB_att_N     = sum(!is.na(GLB_att)),
                               DIR_att_N     = sum(!is.na(DIR_att))  ),
                           by = .( Date = Day ) ]

CLEAR_daily_mean <- DATA_Clear[, .(DIR_att       = mean(DIR_att,    na.rm = T),
                                   GLB_att       = mean(GLB_att,    na.rm = T),
                                   DIR_transp    = mean(DIR_transp, na.rm = T),
                                   DIR_att_sd    = sd(  DIR_att,    na.rm = T),
                                   GLB_att_sd    = sd(  GLB_att,    na.rm = T),
                                   DIR_transp_sd = sd(DIR_transp, na.rm = T),
                                   doy           = yday(Date),
                                   GLB_att_N     = sum(!is.na(GLB_att)),
                                   DIR_att_N     = sum(!is.na(DIR_att))  ),
                               by = .( Date = Day ) ]
```

#### Margin of error calculation for 0.95 confidence interval ####

```{r echo=T, include=T}
CONF_INTERV <- .95
conf_param  <- 1-(1-CONF_INTERV)/2
suppressWarnings({
ALL_daily_mean[,  DIR_att_EM   := qt(conf_param,df=DIR_att_N -1) * DIR_att_sd    / sqrt(DIR_att_N)]
ALL_daily_mean[,  GLB_att_EM   := qt(conf_param,df=GLB_att_N -1) * GLB_att_sd    / sqrt(GLB_att_N)]
ALL_daily_mean[,  DIR_transp_EM:= qt(conf_param,df=DIR_att_N -1) * DIR_transp_sd / sqrt(DIR_att_N)]
CLEAR_daily_mean[,DIR_att_EM   := qt(conf_param,df=DIR_att_N -1) * DIR_att_sd    / sqrt(DIR_att_N)]
CLEAR_daily_mean[,GLB_att_EM   := qt(conf_param,df=GLB_att_N -1) * GLB_att_sd    / sqrt(GLB_att_N)]
CLEAR_daily_mean[,DIR_transp_EM:= qt(conf_param,df=DIR_att_N -1) * DIR_transp_sd / sqrt(DIR_att_N)]
})
```
```{r echo=F, include=F}
```

#### Exclude means with less than `r SUM_THRES` data points

```{r echo=F, include=T}
ALL_daily_mean[  DIR_att_N <= SUM_THRES, DIR_att    := NA ]
ALL_daily_mean[  GLB_att_N <= SUM_THRES, GLB_att    := NA ]
ALL_daily_mean[  DIR_att_N <= SUM_THRES, DIR_transp := NA ]
CLEAR_daily_mean[DIR_att_N <= SUM_THRES, DIR_att    := NA ]
CLEAR_daily_mean[GLB_att_N <= SUM_THRES, GLB_att    := NA ]
CLEAR_daily_mean[DIR_att_N <= SUM_THRES, DIR_transp := NA ]
```

#### Calculate daily seasonal values ####

```{r echo=F, include=T}

ALL_daily_seas <-
    ALL_daily_mean[, .(DIR_att_seas    = mean(DIR_att,    na.rm = T),
                       GLB_att_seas    = mean(GLB_att,    na.rm = T),
                       DIR_transp_seas = mean(DIR_transp, na.rm = T),
                       DIR_att_sd_seas = sd(  DIR_att,    na.rm = T),
                       GLB_att_sd_seas = sd(  GLB_att,    na.rm = T),
                       GLB_att_N_seas  = sum(!is.na(GLB_att)),
                       DIR_att_N_seas  = sum(!is.na(DIR_att))  ),
                   by = .( doy ) ]

CLEAR_daily_seas <-
    CLEAR_daily_mean[, .(DIR_att_seas    = mean(DIR_att,    na.rm = T),
                         GLB_att_seas    = mean(GLB_att,    na.rm = T),
                         DIR_transp_seas = mean(DIR_transp, na.rm = T),
                         DIR_att_sd_seas = sd(  DIR_att,    na.rm = T),
                         GLB_att_sd_seas = sd(  GLB_att,    na.rm = T),
                         GLB_att_N_seas  = sum(!is.na(GLB_att)),
                         DIR_att_N_seas  = sum(!is.na(DIR_att))  ),
                     by = .( doy ) ]

```
```{r echo=F, include=F}
## ~ Plots longterm  ####
plot( ALL_daily_mean$Date, ALL_daily_mean$DIR_att )
plot( ALL_daily_mean$Date, ALL_daily_mean$GLB_att )
plot( ALL_daily_mean$Date, ALL_daily_mean$DIR_transp )

plot( ALL_daily_mean$doy, ALL_daily_mean$DIR_att )
plot( ALL_daily_mean$doy, ALL_daily_mean$GLB_att )
plot( ALL_daily_mean$doy, ALL_daily_mean$DIR_transp )

plot( ALL_daily_mean$doy, ALL_daily_mean$GLB_att_N)
plot( ALL_daily_mean$doy, ALL_daily_mean$DIR_att_N)

hist(ALL_daily_mean$DIR_att_N)
hist(ALL_daily_mean$GLB_att_N)

plot( CLEAR_daily_mean$Date, CLEAR_daily_mean$DIR_att )
plot( CLEAR_daily_mean$Date, CLEAR_daily_mean$GLB_att )
plot( CLEAR_daily_mean$Date, CLEAR_daily_mean$DIR_transp )

plot( CLEAR_daily_mean$doy,  CLEAR_daily_mean$DIR_att )
plot( CLEAR_daily_mean$doy,  CLEAR_daily_mean$GLB_att )
plot( CLEAR_daily_mean$doy,  CLEAR_daily_mean$DIR_transp )

plot( CLEAR_daily_mean$doy,  CLEAR_daily_mean$GLB_att_N)
plot( CLEAR_daily_mean$doy,  CLEAR_daily_mean$DIR_att_N)

hist( CLEAR_daily_mean$DIR_att_N)
hist( CLEAR_daily_mean$GLB_att_N)

## ~ Plots seasonal ####

plot( ALL_daily_seas$doy, ALL_daily_seas$DIR_att_seas )
plot( ALL_daily_seas$doy, ALL_daily_seas$GLB_att_seas )
plot( ALL_daily_seas$doy, ALL_daily_seas$DIR_transp_seas )
plot( ALL_daily_seas$doy, ALL_daily_seas$GLB_att_N_seas)
plot( ALL_daily_seas$doy, ALL_daily_seas$DIR_att_N_seas)

plot( CLEAR_daily_seas$doy, CLEAR_daily_seas$DIR_att_seas )
plot( CLEAR_daily_seas$doy, CLEAR_daily_seas$GLB_att_seas )
plot( CLEAR_daily_seas$doy, CLEAR_daily_seas$DIR_transp_seas )
plot( CLEAR_daily_seas$doy, CLEAR_daily_seas$GLB_att_N_seas)
plot( CLEAR_daily_seas$doy, CLEAR_daily_seas$DIR_att_N_seas)
```

#### Calculate seasonal anomaly ####

```{r echo=F, include=F}

ALL_daily_seas   <- merge(  ALL_daily_mean, ALL_daily_seas,   by = "doy", all = T)
CLEAR_daily_seas <- merge(CLEAR_daily_mean, CLEAR_daily_seas, by = "doy", all = T)

setorder(ALL_daily_seas,Date)
setorder(CLEAR_daily_seas,Date)

## anomaly
# ALL_daily_seas[   , DIR_att    := DIR_att    - DIR_att_seas    ]
# ALL_daily_seas[   , GLB_att    := GLB_att    - GLB_att_seas    ]
# ALL_daily_seas[   , DIR_transp := DIR_transp - DIR_transp_seas ]
# CLEAR_daily_seas[ , DIR_att    := DIR_att    - DIR_att_seas    ]
# CLEAR_daily_seas[ , GLB_att    := GLB_att    - GLB_att_seas    ]
# CLEAR_daily_seas[ , DIR_transp := DIR_transp - DIR_transp_seas ]

## relative anomaly
```
```{r echo=T, include=T}
ALL_daily_seas[  , DIR_att   := 100*( DIR_att    - DIR_att_seas    ) / DIR_att_seas    ]
ALL_daily_seas[  , GLB_att   := 100*( GLB_att    - GLB_att_seas    ) / GLB_att_seas    ]
ALL_daily_seas[  , DIR_transp:= 100*( DIR_transp - DIR_transp_seas ) / DIR_transp_seas ]
CLEAR_daily_seas[, DIR_att   := 100*( DIR_att    - DIR_att_seas    ) / DIR_att_seas    ]
CLEAR_daily_seas[, GLB_att   := 100*( GLB_att    - GLB_att_seas    ) / GLB_att_seas    ]
CLEAR_daily_seas[, DIR_transp:= 100*( DIR_transp - DIR_transp_seas ) / DIR_transp_seas ]
```
```{r echo=F, include=F}

#### ~ Plot of trends  ####
```

\newpage
## Trends on all sky conditions data

```{r echo=F, include=T}

gather <- data.frame()
timefactor <- 1

plot(ALL_daily_seas$Date, ALL_daily_seas$DIR_att, pch  = ".", xlab = "" )
lm1 <- lm( DIR_att ~ Date , data = ALL_daily_seas)
gather <- rbind(gather,
                data.table(
                    var  = "DIR_att",
                    data = "ALL",
                    linear_regression_capture(lm1)
                )
)
abline(lm1)
fit <- lm1[[1]]
legend('top', lty = 1, bty = "n",
       paste('Y =', signif(fit[1],2),if(fit[2]>0)'+'else'-',signif(abs(fit[2]/timefactor*365),3),'* year'))
title(paste("All day daily Direct"),cex=0.9)

plot(ALL_daily_seas$Date, ALL_daily_seas$GLB_att, pch  = ".", xlab = "" )
lm1 <- lm( GLB_att ~ Date , data = ALL_daily_seas)
gather <- rbind(gather,
                data.table(
                    var  = "GLB_att",
                    data = "ALL",
                    linear_regression_capture(lm1)
                )
)
abline(lm1)
fit <- lm1[[1]]
legend('top', lty = 1, bty = "n",
       paste('Y =', signif(fit[1],2),if(fit[2]>0)'+'else'-',signif(abs(fit[2]/timefactor*365),3),'* year'))
title(paste("All sky daily Global"),cex=0.9)

plot(ALL_daily_seas$Date, ALL_daily_seas$DIR_transp, pch  = ".", xlab = "" )
lm1 <- lm( DIR_transp ~ Date , data = ALL_daily_seas)
gather <- rbind(gather,
                data.table(
                    var  = "DIR_transp",
                    data = "ALL",
                    linear_regression_capture(lm1)
                )
)
abline(lm1)
fit <- lm1[[1]]
legend('top', lty = 1, bty = "n",
       paste('Y =', signif(fit[1],2),if(fit[2]>0)'+'else'-',signif(abs(fit[2]/timefactor*365),3),'* year'))
title(paste("All sky daily Direct transparency"),cex=0.9)
```

\newpage
## Trends on clear sky data

```{r echo=F, include=T}

plot(CLEAR_daily_seas$Date, CLEAR_daily_seas$DIR_att, pch  = ".", xlab = "" )
lm1 <- lm( DIR_att ~ Date , data = CLEAR_daily_seas)
gather <- rbind(gather,
                data.table(
                    var  = "DIR_att",
                    data = "CLEAR",
                    linear_regression_capture(lm1)
                )
)
abline(lm1)
fit <- lm1[[1]]
legend('top', lty = 1, bty = "n",
       paste('Y =', signif(fit[1],2),if(fit[2]>0)'+'else'-',signif(abs(fit[2]/timefactor*365),3),'* year'))
title(paste("Clear Sky daily Direct"),cex=0.9)

plot(CLEAR_daily_seas$Date, CLEAR_daily_seas$GLB_att, pch  = ".", xlab = "" )
lm1 <- lm( GLB_att ~ Date , data = CLEAR_daily_seas)
gather <- rbind(gather,
                data.table(
                    var  = "GBL_att",
                    data = "CLEAR",
                    linear_regression_capture(lm1)
                )
)
abline(lm1)
fit <- lm1[[1]]
legend('top', lty = 1, bty = "n",
       paste('Y =', signif(fit[1],2),if(fit[2]>0)'+'else'-',signif(abs(fit[2]/timefactor*365),3),'* year'))
title(paste("Clear Sky daily Global"),cex=0.9)

plot(CLEAR_daily_seas$Date, CLEAR_daily_seas$DIR_transp, pch  = ".", xlab = "" )
lm1 <- lm( DIR_transp ~ Date , data = CLEAR_daily_mean)
gather <- rbind(gather,
                data.table(
                    var  = "DIR_transp",
                    data = "CLEAR",
                    linear_regression_capture(lm1)
                )
)
abline(lm1)
fit <- lm1[[1]]
legend('top', lty = 1, bty = "n",
       paste('Y =', signif(fit[1],2),if(fit[2]>0)'+'else'-',signif(abs(fit[2]/timefactor*365),3),'* year'))
title(paste("Clear Sky daily Direct transparency"),cex=0.9)

wecare <- grep("intercept", names(gather), value = T, invert = T)
gather <- data.table(gather)
```

## Table of trends

```{r echo=F, include=T}
pprint <- gather[ , ..wecare]

pprint[, slope    := slope/timefactor*365 ]
pprint[, slope.sd := slope.sd/timefactor*365 ]

pander(pprint,
       cap = "Slope is in %/year")
```
```{r echo=F, include=F}

## Test plots of all variables
```
```{r echo=F, include=F}
data_list  <- list(ALL_daily_seas, CLEAR_daily_seas)
data_names <- list("All data points", "All clear sky data points")
by_var     <- "Date"
wecare     <- unique(unlist(lapply(data_list, names)))
wecare     <- grep(by_var, wecare, invert = T, value = T)
for(i in 1:length(data_list)) {
    Dplot <- data_list[[i]]
    cat(paste("\n\n## ", i, "\n\n"))
    for (var in wecare) {
        vect <- Dplot[[var]]
        hist(vect, main = var, breaks = 100)
        plot(Dplot[[by_var]], vect, pch = ".", main = var)
        # plot(Dplot$SZA, vect, pch = ".", main = var)
        # plot(Dplot$SZA, vect, pch = ".", main = var)
        # plot(cosde(Dplot$SZA), vect, pch = ".", main = var)
    }
}
```

**END**

```{r include=T, echo=F}
tac <- Sys.time()
cat(sprintf("%s %s@%s %s %f mins\n\n",Sys.time(),Sys.info()["login"],Sys.info()["nodename"],Script.Name,difftime(tac,tic,units="mins")))
```

### 2022-10-05 Article/Article.Rmd
Commit 2022-10-05 12:07

---
title: Full title of the paper (Capitalized)
author:
  - name: Dominik Leutnant
    affil: 1,2,\ddagger,*
    orcid: 0000-0003-3293-2315
  - name: John Doe
    affil: 2, \dagger, \ddagger
affiliation:
  - num: 1
    address: |
      Muenster University of Applied Sciences - 
      Institute for Infrastructure, Water, Resources, Environment
      Correnstr. 25, 48149 Muenster, Germany
    email: leutnant@fh-muenster.de
  - num: 2
    address: |
      Your department
      Street, City, Country
    email: mail@mail.com
# firstnote to eighthnote
firstnote: |
  Current address: Updated affiliation
secondnote: |
  These authors contributed equally to this work.
correspondence: |
  leutnant@fh-muenster.de; Tel.: +XX-000-00-0000.
journal:      water
type:         article
status:       submit
bibliography: mybibfile.bib
appendix:     appendix.tex
simplesummary: |
  A Simple summary goes here.
abstract: |
  A single paragraph of about 200 words maximum. For research articles, 
  abstracts should give a pertinent overview of the work. We strongly encourage
  authors to use the following style of structured abstracts, but without 
  headings: 1) Background: Place the question addressed in a broad context and
  highlight the purpose of the study; 2) Methods: Describe briefly the main
  methods or treatments applied; 3) Results: Summarize the article's main 
  findings; and 4) Conclusion: Indicate the main conclusions or interpretations. 
  The abstract should be an objective representation of the article, it must not 
  contain results which are not presented and substantiated in the main text and 
  should not exaggerate the main conclusions.
keywords: |
  keyword 1; keyword 2; keyword 3 (list three to ten pertinent keywords specific 
  to the article, yet reasonably common within the subject discipline.).
acknowledgement: |
  All sources of funding of the study should be disclosed. Please clearly 
  indicate grants that you have received in support of your research work. 
  Clearly state if you received funds for covering the costs to publish in open 
  access.
authorcontributions: |
  For research articles with several authors, a short paragraph specifying their 
  individual contributions must be provided. The following statements should be 
  used ``X.X. and Y.Y. conceive and designed the experiments; X.X. performed the 
  experiments; X.X. and Y.Y. analyzed the data; W.W. contributed 
  reagents/materials/analysis tools; Y.Y. wrote the paper.'' Authorship must be
  limited to those who have contributed substantially to the work reported.
conflictsofinterest: |
  Declare conflicts of interest or state 'The authors declare no conflict of 
  interest.' Authors must identify and declare any personal circumstances or
  interest that may be perceived as inappropriately influencing the
  representation or interpretation of reported research results. Any role of the
  funding sponsors in the design of the study; in the collection, analyses or 
  interpretation of data in the writing of the manuscript, or in the decision to 
  publish the results must be declared in this section. If there is no role, 
  please state 'The founding sponsors had no role in the design of the study; 
  in the collection, analyses, or interpretation of data; in the writing of the 
  manuscript, an in the decision to publish the results'.
sampleavailability: |
  Samples of the compounds ...... are available from the authors.
abbreviations:
  - short: MDPI
    long: Multidisciplinary Digital Publishing Institute
  - short: DOAJ
    long: Directory of open access journals
  - short: TLA
    long: Three letter acronym
  - short: LD 
    long: linear dichroism
output: rticles::mdpi_article
---

# Version

This Rmd-skeleton uses the mdpi Latex template published 2019/02. 
However, the official template gets more frequently updated than the 'rticles'
package. Therefore, please make sure prior to paper submission, that you're 
using the most recent .cls, .tex and .bst files 
(available [here](http://www.mdpi.com/authors/latex)).

# Introduction

The introduction should briefly place the study in a broad context and highlight why it is important. It should define the purpose of the work and its significance. The current state of the research field should be reviewed carefully and key publications cited. Please highlight controversial and diverging hypotheses when necessary. Finally, briefly mention the main aim of the work and highlight the principal conclusions. As far as possible, please keep the introduction comprehensible to scientists outside your particular field of research. Citing a journal paper [@bertrand-krajewski_distribution_1998; @leutnant_stormwater_2016]. And now citing a book reference @gujer_systems_2008. Please use the command [@leutnant_stormwater_2016] for the following MDPI journals, which use author-date citation: Administrative Sciences, Arts, Econometrics, Economies, Genealogy, Humanities, IJFS, JRFM, Laws, Religions, Risks, Social Sciences.

# Materials and Methods

Materials and Methods should be described with sufficient details to allow others to replicate and build on published results. Please note that publication of your manuscript implicates that you must make all materials, data, computer code, and protocols associated with the publication available to readers. Please disclose at the submission stage any restrictions on the availability of materials or information. New methods and protocols should be described in detail while well-established methods can be briefly described and appropriately cited.

Research manuscripts reporting large datasets that are deposited in a publicly available database should specify where the data have been deposited and provide the relevant accession numbers. If the accession numbers have not yet been obtained at the time of submission, please state that they will be provided during review. They must be provided prior to publication.

Interventionary studies involving animals or humans, and other studies require ethical approval must list the authority that provided approval and the corresponding ethical approval code.

# Results

This section may be divided by subheadings. It should provide a concise and precise description of the experimental results, their interpretation as well as the experimental conclusions that can be drawn.

## Subsection Heading Here

Subsection text here.

### Subsubsection Heading Here

Bulleted lists look like this:

* First bullet
* Second bullet
* Third bullet

Numbered lists can be added as follows:

1. First item
2. Second item
3. Third item

The text continues here.

All figures and tables should be cited in the main text as Figure 1, Table 1, etc.

\begin{figure}[H]
\centering
\includegraphics[width=3 cm]{logo-mdpi}
\caption{This is a figure, Schemes follow the same formatting. If there are multiple panels, they should be listed as: (\textbf{a}) Description of what is contained in the first panel. (\textbf{b}) Description of what is contained in the second panel. Figures should be placed in the main text near to the first time they are cited. A caption on a single line should be centered.}
\end{figure}

\begin{table}[H]
\caption{This is a table caption. Tables should be placed in the main text near to the first time they are cited.}
\centering
%% \tablesize{} %% You can specify the fontsize here, e.g.  \tablesize{\footnotesize}. If commented out \small will be used.
\begin{tabular}{ccc}
\toprule
\textbf{Title 1}	& \textbf{Title 2}	& \textbf{Title 3}\\
\midrule
entry 1		& data			& data\\
entry 2		& data			& data\\
\bottomrule
\end{tabular}
\end{table}

This is an example of an equation:

\begin{equation}
\mathbb{S}
\end{equation}
<!-- If the documentclass option "submit" is chosen, please insert a blank line before and after any math environment (equation and eqnarray environments). This ensures correct linenumbering. The blank line should be removed when the documentclass option is changed to "accept" because the text following an equation should not be a new paragraph. -->

<!-- Please punctuate equations as regular text. Theorem-type environments (including propositions, lemmas, corollaries etc.) can be formatted as follows: -->
Example of a theorem:
\begin{Theorem}
Example text of a theorem.
\end{Theorem}

The text continues here. Proofs must be formatted as follows:

Example of a proof:
\begin{proof}[Proof of Theorem 1]
Text of the proof. Note that the phrase `of Theorem 1' is optional if it is clear which theorem is being referred to.
\end{proof}
The text continues here.

# Discussion

Authors should discuss the results and how they can be interpreted in perspective of previous studies and of the working hypotheses. The findings and their implications should be discussed in the broadest context possible. Future research directions may also be highlighted.

# Conclusion

This section is not mandatory, but can be added to the manuscript if the discussion is unusually long or complex.

# Patents

This section is not mandatory, but may be added if there are patents resulting from the work reported in this manuscript.

### 2022-10-05 Climatological_DHI_GHI_longterm_trends.R

@@ -1,11 +1,12 @@
 # /* !/usr/bin/env Rscript */
 # /* Copyright (C) 2022 Athanasios Natsis <natsisthanasis@gmail.com> */
 #' ---
-#' title:         ""
-#' author:        "Natsis Athanasios"
-#' institute:     "AUTH"
-#' affiliation:   "Laboratory of Atmospheric Physics"
-#' abstract:      "Read "
+#' title:         "Trends of SDR in Thessaloniki "
+#' author:
+#'   - Natsis Athanasios^[Laboratory of Atmospheric Physics,AUTH, natsisthanasis@gmail.com]
+#'   - Jane Doe^[Institution Two, jane@example.org]
+#' abstract:
+#'   "Study of GHI and DNI radiation for 'clear sky' and all sly conditions."
 #'
 #' documentclass:  article
 #' classoption:    a4paper,oneside
@@ -24,7 +25,6 @@
 #'     number_sections:  no
 #'     fig_caption:      no
 #'     keep_tex:         no
-#'     keep_rmd:         yes
 #'     latex_engine:     xelatex
 #'     toc:              yes
 #'     fig_width:        7
@@ -63,9 +63,9 @@ Script.Name <- tryCatch({ funr::sys.script() },
                         error = function(e) { cat(paste("\nUnresolved script name: ", e),"\n\n")
                             return("Climatological_") })
 if(!interactive()) {
-    pdf(  file = paste0("~/Aerosols/REPORTS/runtime/", basename(sub("\\.R$",".pdf", Script.Name))))
-    sink( file = paste0("~/Aerosols/REPORTS/runtime/", basename(sub("\\.R$",".out", Script.Name))), split=TRUE)
-    filelock::lock(paste0("~/Aerosols/LOGs/",  basename(sub("\\.R$",".lock", Script.Name))), timeout = 0)
+    pdf(  file = paste0("~/MANUSCRIPTS/2022_sdr_trends/runtime/", basename(sub("\\.R$",".pdf", Script.Name))))
+    sink( file = paste0("~/MANUSCRIPTS/2022_sdr_trends/runtime/", basename(sub("\\.R$",".out", Script.Name))), split=TRUE)
+    filelock::lock(paste0("~/MANUSCRIPTS/2022_sdr_trends/runtime//",  basename(sub("\\.R$",".lock", Script.Name))), timeout = 0)
 }
 
 par(pch = ".")

### 2022-10-05 Climatological_DHI_GHI_trends_consistency.R

@@ -1,11 +1,12 @@
 # /* !/usr/bin/env Rscript */
 # /* Copyright (C) 2022 Athanasios Natsis <natsisthanasis@gmail.com> */
 #' ---
-#' title:         ""
-#' author:        "Natsis Athanasios"
-#' institute:     "AUTH"
-#' affiliation:   "Laboratory of Atmospheric Physics"
-#' abstract:      "Read "
+#' title:         "Trends of SDR in Thessaloniki "
+#' author:
+#'   - Natsis Athanasios^[Laboratory of Atmospheric Physics,AUTH, natsisthanasis@gmail.com]
+#'   - Jane Doe^[Institution Two, jane@example.org]
+#' abstract:
+#'   "Study of GHI and DNI radiation for 'clear sky' and all sly conditions."
 #'
 #' documentclass:  article
 #' classoption:    a4paper,oneside

### 2022-10-05 Article.Rmd
Commit 2022-10-05 13:07

@@
 ---
-title:         "Trends of SDR over Thessaloniki"
-# author:        "Natsis Athanasios"
-# institute:     "AUTH"
-# affiliation:   "Laboratory of Atmospheric Physics"
-abstract:      "Read "
+title:         "Trends of SDR in Thessaloniki "
+author:
+  - Natsis Athanasios^[Laboratory of Atmospheric Physics,AUTH, natsisthanasis@gmail.com]
+  - Jane Doe^[Institution Two, jane@example.org]
+abstract:
+  "Study of GHI and DNI radiation for 'clear sky' and all sly conditions."
 
@@
     keep_tex:         no
-    keep_rmd:         yes
     latex_engine:     xelatex
@@
     fig_height:       4.5
+    cache:            yes
   html_document:
@@
 
-```{r echo=F, include=T}
-
-####_  Document options _####
-
-```
 ```{r echo=F, include=F}
@@
 # knitr::opts_chunk$set(dev        = "png"    )
-knitr::opts_chunk$set(out.width  = "85%"   )
+knitr::opts_chunk$set(out.width  = "80%"   )
 knitr::opts_chunk$set(fig.align  = "center" )
-knitr::opts_chunk$set(cache      =  T       )
 # knitr::opts_chunk$set(fig.pos    = '!h'    )
 
-```
-```{r include=F, echo=F}
 ####  Set environment  ####
@@
 tic <- Sys.time()
-Script.Name <- tryCatch({ funr::sys.script() },
-                        error = function(e) { cat(paste("\nUnresolved script name: ", e),"\n\n")
-                            return("Climatological_") })
-if(!interactive()) {
-    pdf(  file = paste0("~/Aerosols/REPORTS/runtime/", basename(sub("\\.R$",".pdf", Script.Name))))
-    sink( file = paste0("~/Aerosols/REPORTS/runtime/", basename(sub("\\.R$",".out", Script.Name))), split=TRUE)
-    filelock::lock(paste0("~/Aerosols/LOGs/",  basename(sub("\\.R$",".lock", Script.Name))), timeout = 0)
-}
 
@@
 
-## FIXME this is for pdf output
-# options(warn=-1) ## hide warnings
-# options(warn=2)  ## stop on warnings
-
-```
-```{r echo=F, include=T}
-####  External code  ####
 library(data.table, quietly = T, warn.conflicts = F)
 library(pander,     quietly = T, warn.conflicts = F)
-
 ## Functions from `https://github.com/thanasisn/IStillBreakStuff/tree/main/FUNCTIONS/R`
@@
 
-TEST      <- TRUE
-
 MIN_ELEVA  <- 5  ## use data when sun above that
-SZA_BIN    <- 1
+SZA_BIN    <- 1  ## bin size of SZA
 MIN_N      <- 4
@@
 
-## Data selection
+## Data description
+
+....... span
+
+...... measuremetns CHP1 CM21
+
+....... screening
+
+....... Quality assurance [cite]
+
+## Data preperation
 
@@
 } else {
-    cat(paste("Load data from Clear Sky proccess from parsed\n"))
+    # cat(paste("Load data from Clear Sky proccess from parsed\n"))
     DATA <- readRDS(CS_file)
 }
+
 ```
 
-### Filter min elevation
-Keep data with Sun elevation above `r MIN_ELEVA`
+Due to the large measurement assertenty near the horizon we exclude all measurements
+with greater than SZA $`r 90-MIN_ELEVA`^\circ$. Moreover, due to nearby hills and buildings
+we remove data with azimuth angle between $35^\circ$ and $120^\circ$ and SZA greater than $`r 90-10`^\circ$.
 
-```{r }
+```{r echo=F}
 DATA <- DATA[ Elevat >= MIN_ELEVA ]
+DATA <- DATA[ ! (Azimuth > 35 & Azimuth < 120 & Elevat < 10) ]
 ```
 
-### Bais paper obstacle filter
-
-```{r }
-DATA <- DATA[ ! (Azimuth > 35 & Azimuth < 120 & Elevat < 10) ]
-```
-```{r echo=F, include=T}
-```
-
-### Keep only data characterized as 'good' by the Radiation Quality control procedure
-
-```{r echo=F, include=T}
-DATA[ QCF_DIR != "good", wattDIR := NA ]
-DATA[ QCF_DIR != "good", QCF_DIR := NA ]
-DATA[ QCF_GLB != "good", wattGLB := NA ]
-DATA[ QCF_GLB != "good", QCF_GLB := NA ]
-DATA <- DATA[ ! (is.na(wattDIR) & is.na(wattGLB))  ]
-
-## exclude some data for each instrument
-# DATA[ QCF_DIR == "Extremely rare limits min (3)", wattDIR := NA ]
-# DATA[ QCF_DIR == "Extremely rare limits min (3)", QCF_DIR := NA ]
-#
-# dirtb <- table(DATA$QCF_DIR)
-# for (aty in names(dirtb)[!names(dirtb) %in% "good"]) {
-#     temp <- DATA[ QCF_DIR == aty ]
-#     cat(nrow(temp),aty,"\n")
-# }
-# glbtb <- table(DATA$QCF_GLB)
-# for (aty in names(glbtb)[!names(glbtb) %in% "good"]) {
-#     temp <- DATA[ QCF_GLB == aty ]
-#     cat(nrow(temp),aty,"\n")
-# }
-```
-
-## Data preparation
-### Move measurements to mean earth distance
-
-```{r }
-DATA[ , wattDIR_1au := wattDIR * (sun_dist ^ 2) ]
-DATA[ , wattGLB_1au := wattGLB * (sun_dist ^ 2) ]
-```
-```{r echo=F, include=T}
-```
-
-### Relative to actual TSI at 1au variable representation
-
-```{r }
-DATA[ , DIR_att := wattDIR_1au / tsi_1au_comb ]
-DATA[ , GLB_att := wattGLB_1au / tsi_1au_comb ]
-```
-
-Using the actual values gives similar trends.
-
-```{r echo=F, include=T}
-
-# #' ### Use original variable representation
-# DATA[ , DIR_att := wattDIR_1au ]
-# DATA[ , GLB_att := wattGLB_1au ]
-# #+ echo=F, include=T
-
-# #' ### Ground effect removal?
-# #' Aerosol direct effects on global solar shortwave irradiance at high mountainous station Musala Bulgaria_Nojarov2021.pdf
-# DATA$wattGLB_1au <- DATA$wattGLB_1au / cosde(DATA$SZA)
-# DATA$wattDIR_1au <- DATA$wattDIR_1au / cosde(DATA$SZA)
-# #+ echo=F, include=T
-```
-
-### Calculate Bouguer atmospheric transparency
-see: Changes in solar radiation and their influence on temperature trend in Estonia 1955 2007_Russak2009.pdf
-
-```{r }
-DATA[, DIR_transp := ( wattDIR_1au / tsi_1au_comb ) ^ ( 1 / cosde(SZA) ) ]
-```
-```{r echo=F, include=T}
-
-## fix noon just in case
-DATA[ Azimuth <= 180 , preNoon := TRUE  ]
-DATA[ Azimuth >  180 , preNoon := FALSE ]
-```
-
-### Split data to Clear Sky and non Clear sky
-Method based and adapted to site from: Identification of Periods of Clear Sky
-Irradiance in Time Series of GHI Measurements
-_Matthew J. Reno and Clifford W. Hansen_.
-
-```{r echo=F, include=T}
-DATA_all   <- DATA
-
-wecare     <- grep("CSflag_", names(DATA), value = T)
-## use only cm21 flags
-wecare     <- grep("_11", wecare, invert = T, value = T)
-DATA_Clear <- DATA[ rowSums( DATA[, ..wecare ], na.rm = T ) == 0, ]
-```
-```{r echo=F, include=T}
-
-## old flags
-# DATA_Clear <- DATA_all[ CSflag == 0 ]
-rm(DATA)
-
-# #### . . My CSid method   ####
-# #'
-# #' | CS Flag | Test |
-# #' |:-------:|:--------------------------------------------------------------|
-# #' |   NA    | Undefined, untested                                           |
-# #' |    0    | Passed as clear sky                                           |
-# #' |    1    | Mean value of irradiance during the time period (MeanVIP)     |
-# #' |    2    | Max Value of Irradiance during the time Period (MaxVIP)       |
-# #' |    3    | Variability in irradiance by the length (VIL)                 |
-# #' |    4    | Variance of Changes in the Time series (VCT)                  |
-# #' |    5    | Variability in the Shape of the irradiance Measurements (VSM) |
-# #' |    6    | Low Direct Irradiance limit (LDI)                             |
-# #' |    7    | Low Global Irradiance limit (LGI)                             |
-# #' |    8    | Too Few CS point for the day (FCS)                            |
-# #' |    9    | Too Few data points for the day                               |
-# #' |   10    | Missing Data                                                  |
-# #' |   11    | Direct irradiance simple threshold                            |
-
-#### 1. Long term anomaly trends ####
-```
-
-## 1. Long term anomaly trends
-
-```{r }
-##TODO t-test and significance
-```
-
-#### Calculate daily means ####
-
-```{r echo=F, include=T}
-
-ALL_daily_mean <- DATA_all[, .(DIR_att       = mean(DIR_att,    na.rm = T),
-                               GLB_att       = mean(GLB_att,    na.rm = T),
-                               DIR_transp    = mean(DIR_transp, na.rm = T),
-                               DIR_att_sd    = sd(  DIR_att,    na.rm = T),
-                               GLB_att_sd    = sd(  GLB_att,    na.rm = T),
-                               DIR_transp_sd = sd(DIR_transp, na.rm = T),
-                               doy           = yday(Date),
-                               GLB_att_N     = sum(!is.na(GLB_att)),
-                               DIR_att_N     = sum(!is.na(DIR_att))  ),
-                           by = .( Date = Day ) ]
-
-CLEAR_daily_mean <- DATA_Clear[, .(DIR_att       = mean(DIR_att,    na.rm = T),
-                                   GLB_att       = mean(GLB_att,    na.rm = T),
-                                   DIR_transp    = mean(DIR_transp, na.rm = T),
-                                   DIR_att_sd    = sd(  DIR_att,    na.rm = T),
-                                   GLB_att_sd    = sd(  GLB_att,    na.rm = T),
-                                   DIR_transp_sd = sd(DIR_transp, na.rm = T),
-                                   doy           = yday(Date),
-                                   GLB_att_N     = sum(!is.na(GLB_att)),
-                                   DIR_att_N     = sum(!is.na(DIR_att))  ),
-                               by = .( Date = Day ) ]
-```
-
-#### Margin of error calculation for 0.95 confidence interval ####
-
-```{r echo=T, include=T}
-CONF_INTERV <- .95
-conf_param  <- 1-(1-CONF_INTERV)/2
-suppressWarnings({
-ALL_daily_mean[,  DIR_att_EM   := qt(conf_param,df=DIR_att_N -1) * DIR_att_sd    / sqrt(DIR_att_N)]
-ALL_daily_mean[,  GLB_att_EM   := qt(conf_param,df=GLB_att_N -1) * GLB_att_sd    / sqrt(GLB_att_N)]
-ALL_daily_mean[,  DIR_transp_EM:= qt(conf_param,df=DIR_att_N -1) * DIR_transp_sd / sqrt(DIR_att_N)]
-CLEAR_daily_mean[,DIR_att_EM   := qt(conf_param,df=DIR_att_N -1) * DIR_att_sd    / sqrt(DIR_att_N)]
-CLEAR_daily_mean[,GLB_att_EM   := qt(conf_param,df=GLB_att_N -1) * GLB_att_sd    / sqrt(GLB_att_N)]
-CLEAR_daily_mean[,DIR_transp_EM:= qt(conf_param,df=DIR_att_N -1) * DIR_transp_sd / sqrt(DIR_att_N)]
-})
-```
-```{r echo=F, include=F}
-```
-
-#### Exclude means with less than `r SUM_THRES` data points
-
-```{r echo=F, include=T}
-ALL_daily_mean[  DIR_att_N <= SUM_THRES, DIR_att    := NA ]
-ALL_daily_mean[  GLB_att_N <= SUM_THRES, GLB_att    := NA ]
-ALL_daily_mean[  DIR_att_N <= SUM_THRES, DIR_transp := NA ]
-CLEAR_daily_mean[DIR_att_N <= SUM_THRES, DIR_att    := NA ]
-CLEAR_daily_mean[GLB_att_N <= SUM_THRES, GLB_att    := NA ]
-CLEAR_daily_mean[DIR_att_N <= SUM_THRES, DIR_transp := NA ]
-```
-
-#### Calculate daily seasonal values ####
-
-```{r echo=F, include=T}
-
-ALL_daily_seas <-
-    ALL_daily_mean[, .(DIR_att_seas    = mean(DIR_att,    na.rm = T),
-                       GLB_att_seas    = mean(GLB_att,    na.rm = T),
-                       DIR_transp_seas = mean(DIR_transp, na.rm = T),
-                       DIR_att_sd_seas = sd(  DIR_att,    na.rm = T),
-                       GLB_att_sd_seas = sd(  GLB_att,    na.rm = T),
-                       GLB_att_N_seas  = sum(!is.na(GLB_att)),
-                       DIR_att_N_seas  = sum(!is.na(DIR_att))  ),
-                   by = .( doy ) ]
-
-CLEAR_daily_seas <-
-    CLEAR_daily_mean[, .(DIR_att_seas    = mean(DIR_att,    na.rm = T),
-                         GLB_att_seas    = mean(GLB_att,    na.rm = T),
-                         DIR_transp_seas = mean(DIR_transp, na.rm = T),
-                         DIR_att_sd_seas = sd(  DIR_att,    na.rm = T),
-                         GLB_att_sd_seas = sd(  GLB_att,    na.rm = T),
-                         GLB_att_N_seas  = sum(!is.na(GLB_att)),
-                         DIR_att_N_seas  = sum(!is.na(DIR_att))  ),
-                     by = .( doy ) ]
-
-```
-```{r echo=F, include=F}
-## ~ Plots longterm  ####
-plot( ALL_daily_mean$Date, ALL_daily_mean$DIR_att )
-plot( ALL_daily_mean$Date, ALL_daily_mean$GLB_att )
-plot( ALL_daily_mean$Date, ALL_daily_mean$DIR_transp )
-
-plot( ALL_daily_mean$doy, ALL_daily_mean$DIR_att )
-plot( ALL_daily_mean$doy, ALL_daily_mean$GLB_att )
-plot( ALL_daily_mean$doy, ALL_daily_mean$DIR_transp )
-
-plot( ALL_daily_mean$doy, ALL_daily_mean$GLB_att_N)
-plot( ALL_daily_mean$doy, ALL_daily_mean$DIR_att_N)
-
-hist(ALL_daily_mean$DIR_att_N)
-hist(ALL_daily_mean$GLB_att_N)
-
-plot( CLEAR_daily_mean$Date, CLEAR_daily_mean$DIR_att )
-plot( CLEAR_daily_mean$Date, CLEAR_daily_mean$GLB_att )
-plot( CLEAR_daily_mean$Date, CLEAR_daily_mean$DIR_transp )
-
-plot( CLEAR_daily_mean$doy,  CLEAR_daily_mean$DIR_att )
-plot( CLEAR_daily_mean$doy,  CLEAR_daily_mean$GLB_att )
-plot( CLEAR_daily_mean$doy,  CLEAR_daily_mean$DIR_transp )
-
-plot( CLEAR_daily_mean$doy,  CLEAR_daily_mean$GLB_att_N)
-plot( CLEAR_daily_mean$doy,  CLEAR_daily_mean$DIR_att_N)
-
-hist( CLEAR_daily_mean$DIR_att_N)
-hist( CLEAR_daily_mean$GLB_att_N)
-
-## ~ Plots seasonal ####
-
-plot( ALL_daily_seas$doy, ALL_daily_seas$DIR_att_seas )
-plot( ALL_daily_seas$doy, ALL_daily_seas$GLB_att_seas )
-plot( ALL_daily_seas$doy, ALL_daily_seas$DIR_transp_seas )
-plot( ALL_daily_seas$doy, ALL_daily_seas$GLB_att_N_seas)
-plot( ALL_daily_seas$doy, ALL_daily_seas$DIR_att_N_seas)
-
-plot( CLEAR_daily_seas$doy, CLEAR_daily_seas$DIR_att_seas )
-plot( CLEAR_daily_seas$doy, CLEAR_daily_seas$GLB_att_seas )
-plot( CLEAR_daily_seas$doy, CLEAR_daily_seas$DIR_transp_seas )
-plot( CLEAR_daily_seas$doy, CLEAR_daily_seas$GLB_att_N_seas)
-plot( CLEAR_daily_seas$doy, CLEAR_daily_seas$DIR_att_N_seas)
-```
-
-#### Calculate seasonal anomaly ####
-
-```{r echo=F, include=F}
-
-ALL_daily_seas   <- merge(  ALL_daily_mean, ALL_daily_seas,   by = "doy", all = T)
-CLEAR_daily_seas <- merge(CLEAR_daily_mean, CLEAR_daily_seas, by = "doy", all = T)
-
-setorder(ALL_daily_seas,Date)
-setorder(CLEAR_daily_seas,Date)
-
-## anomaly
-# ALL_daily_seas[   , DIR_att    := DIR_att    - DIR_att_seas    ]
-# ALL_daily_seas[   , GLB_att    := GLB_att    - GLB_att_seas    ]
-# ALL_daily_seas[   , DIR_transp := DIR_transp - DIR_transp_seas ]
-# CLEAR_daily_seas[ , DIR_att    := DIR_att    - DIR_att_seas    ]
-# CLEAR_daily_seas[ , GLB_att    := GLB_att    - GLB_att_seas    ]
-# CLEAR_daily_seas[ , DIR_transp := DIR_transp - DIR_transp_seas ]
-
-## relative anomaly
-```
-```{r echo=T, include=T}
-ALL_daily_seas[  , DIR_att   := 100*( DIR_att    - DIR_att_seas    ) / DIR_att_seas    ]
-ALL_daily_seas[  , GLB_att   := 100*( GLB_att    - GLB_att_seas    ) / GLB_att_seas    ]
-ALL_daily_seas[  , DIR_transp:= 100*( DIR_transp - DIR_transp_seas ) / DIR_transp_seas ]
-CLEAR_daily_seas[, DIR_att   := 100*( DIR_att    - DIR_att_seas    ) / DIR_att_seas    ]
-CLEAR_daily_seas[, GLB_att   := 100*( GLB_att    - GLB_att_seas    ) / GLB_att_seas    ]
-CLEAR_daily_seas[, DIR_transp:= 100*( DIR_transp - DIR_transp_seas ) / DIR_transp_seas ]
-```
-```{r echo=F, include=F}
-
-#### ~ Plot of trends  ####
-```
-
-\newpage
-## Trends on all sky conditions data
-
-```{r echo=F, include=T}
-
-gather <- data.frame()
-timefactor <- 1
-
-plot(ALL_daily_seas$Date, ALL_daily_seas$DIR_att, pch  = ".", xlab = "" )
-lm1 <- lm( DIR_att ~ Date , data = ALL_daily_seas)
-gather <- rbind(gather,
-                data.table(
-                    var  = "DIR_att",
-                    data = "ALL",
-                    linear_regression_capture(lm1)
-                )
-)
-abline(lm1)
-fit <- lm1[[1]]
-legend('top', lty = 1, bty = "n",
-       paste('Y =', signif(fit[1],2),if(fit[2]>0)'+'else'-',signif(abs(fit[2]/timefactor*365),3),'* year'))
-title(paste("All day daily Direct"),cex=0.9)
-
-plot(ALL_daily_seas$Date, ALL_daily_seas$GLB_att, pch  = ".", xlab = "" )
-lm1 <- lm( GLB_att ~ Date , data = ALL_daily_seas)
-gather <- rbind(gather,
-                data.table(
-                    var  = "GLB_att",
-                    data = "ALL",
-                    linear_regression_capture(lm1)
-                )
-)
-abline(lm1)
-fit <- lm1[[1]]
-legend('top', lty = 1, bty = "n",
-       paste('Y =', signif(fit[1],2),if(fit[2]>0)'+'else'-',signif(abs(fit[2]/timefactor*365),3),'* year'))
-title(paste("All sky daily Global"),cex=0.9)
-
-plot(ALL_daily_seas$Date, ALL_daily_seas$DIR_transp, pch  = ".", xlab = "" )
-lm1 <- lm( DIR_transp ~ Date , data = ALL_daily_seas)
-gather <- rbind(gather,
-                data.table(
-                    var  = "DIR_transp",
-                    data = "ALL",
-                    linear_regression_capture(lm1)
-                )
-)
-abline(lm1)
-fit <- lm1[[1]]
-legend('top', lty = 1, bty = "n",
-       paste('Y =', signif(fit[1],2),if(fit[2]>0)'+'else'-',signif(abs(fit[2]/timefactor*365),3),'* year'))
-title(paste("All sky daily Direct transparency"),cex=0.9)
-```
-
-\newpage
-## Trends on clear sky data
-
-```{r echo=F, include=T}
-
-plot(CLEAR_daily_seas$Date, CLEAR_daily_seas$DIR_att, pch  = ".", xlab = "" )
-lm1 <- lm( DIR_att ~ Date , data = CLEAR_daily_seas)
-gather <- rbind(gather,
-                data.table(
-                    var  = "DIR_att",
-                    data = "CLEAR",
-                    linear_regression_capture(lm1)
-                )
-)
-abline(lm1)
-fit <- lm1[[1]]
-legend('top', lty = 1, bty = "n",
-       paste('Y =', signif(fit[1],2),if(fit[2]>0)'+'else'-',signif(abs(fit[2]/timefactor*365),3),'* year'))
-title(paste("Clear Sky daily Direct"),cex=0.9)
-
-plot(CLEAR_daily_seas$Date, CLEAR_daily_seas$GLB_att, pch  = ".", xlab = "" )
-lm1 <- lm( GLB_att ~ Date , data = CLEAR_daily_seas)
-gather <- rbind(gather,
-                data.table(
-                    var  = "GBL_att",
-                    data = "CLEAR",
-                    linear_regression_capture(lm1)
-                )
-)
-abline(lm1)
-fit <- lm1[[1]]
-legend('top', lty = 1, bty = "n",
-       paste('Y =', signif(fit[1],2),if(fit[2]>0)'+'else'-',signif(abs(fit[2]/timefactor*365),3),'* year'))
-title(paste("Clear Sky daily Global"),cex=0.9)
-
-plot(CLEAR_daily_seas$Date, CLEAR_daily_seas$DIR_transp, pch  = ".", xlab = "" )
-lm1 <- lm( DIR_transp ~ Date , data = CLEAR_daily_mean)
-gather <- rbind(gather,
-                data.table(
-                    var  = "DIR_transp",
-                    data = "CLEAR",
-                    linear_regression_capture(lm1)
-                )
-)
-abline(lm1)
-fit <- lm1[[1]]
-legend('top', lty = 1, bty = "n",
-       paste('Y =', signif(fit[1],2),if(fit[2]>0)'+'else'-',signif(abs(fit[2]/timefactor*365),3),'* year'))
-title(paste("Clear Sky daily Direct transparency"),cex=0.9)
-
-wecare <- grep("intercept", names(gather), value = T, invert = T)
-gather <- data.table(gather)
-```
-
-## Table of trends
-
-```{r echo=F, include=T}
-pprint <- gather[ , ..wecare]
-
-pprint[, slope    := slope/timefactor*365 ]
-pprint[, slope.sd := slope.sd/timefactor*365 ]
-
-pander(pprint,
-       cap = "Slope is in %/year")
-```
-```{r echo=F, include=F}
-
-## Test plots of all variables
-```
-```{r echo=F, include=F}
-data_list  <- list(ALL_daily_seas, CLEAR_daily_seas)
-data_names <- list("All data points", "All clear sky data points")
-by_var     <- "Date"
-wecare     <- unique(unlist(lapply(data_list, names)))
-wecare     <- grep(by_var, wecare, invert = T, value = T)
-for(i in 1:length(data_list)) {
-    Dplot <- data_list[[i]]
-    cat(paste("\n\n## ", i, "\n\n"))
-    for (var in wecare) {
-        vect <- Dplot[[var]]
-        hist(vect, main = var, breaks = 100)
-        plot(Dplot[[by_var]], vect, pch = ".", main = var)
-        # plot(Dplot$SZA, vect, pch = ".", main = var)
-        # plot(Dplot$SZA, vect, pch = ".", main = var)
-        # plot(cosde(Dplot$SZA), vect, pch = ".", main = var)
-    }
-}
-```
+## Data selection
 
@@
 tac <- Sys.time()
-cat(sprintf("%s %s@%s %s %f mins\n\n",Sys.time(),Sys.info()["login"],Sys.info()["nodename"],Script.Name,difftime(tac,tic,units="mins")))
+cat(sprintf("%s %s@%s %f mins\n\n",Sys.time(),Sys.info()["login"],Sys.info()["nodename"],difftime(tac,tic,units="mins")))
 ```

### 2022-10-05 Climatological_DHI_GHI_trends_consistency.R

@@ -163,7 +163,7 @@ DATA <- DATA[ Elevat >= MIN_ELEVA ]
 DATA <- DATA[ ! (Azimuth > 35 & Azimuth < 120 & Elevat < 10) ]
 #+ echo=F, include=T
 
-
+#### ~~ RMD mark ####
 
 #' ### Keep only data characterized as 'good' by the Radiation Quality control procedure
 #+ echo=F, include=T

### 2022-10-05 Article.Rmd
Commit 2022-10-05 14:07

@@
 colorlinks:     yes
+urlcolor:        blue
+
+bibliography:    [references.bib]
+biblio-style:    apalike
 
@@
     number_sections:  no
-    fig_caption:      no
+    fig_caption:      yes
     keep_tex:         no
@@
     fig_height:       4.5
-    cache:            yes
   html_document:
@@
 
-```{r echo=F, include=F}
+```{r echo=F, include=F, cache=T}
 knitr::opts_chunk$set(comment    = ""      )
@@
 
-####  Set environment  ####
-Sys.setenv(TZ = "UTC")
-tic <- Sys.time()
-
-par(pch = ".")
-
-library(data.table, quietly = T, warn.conflicts = F)
-library(pander,     quietly = T, warn.conflicts = F)
-## Functions from `https://github.com/thanasisn/IStillBreakStuff/tree/main/FUNCTIONS/R`
-source("~/CODE/FUNCTIONS/R/sumNA.R")
-source("~/CODE/FUNCTIONS/R/linear_regrassion_capture.R")
-source("~/CODE/FUNCTIONS/R/trig_deg.R")
-
-####  Variables  ####
-panderOptions('table.alignment.default', 'right')
-panderOptions('table.split.table',        120   )
-
-DATA_DIR  <- "/home/athan/DATA/Broad_Band"
-CLEARdir  <- "/home/athan/DATA/Broad_Band/CS_id"
-tag       <- paste0("Natsis Athanasios LAP AUTH ", strftime(Sys.time(), format = "%b %Y" ))
-CS_file   <- "/home/athan/DATA/Common_application/Clear_Sky.Rds"
-
-MIN_ELEVA  <- 5  ## use data when sun above that
-SZA_BIN    <- 1  ## bin size of SZA
-MIN_N      <- 4
-SEAS_MIN_N <- 3
-
-SUM_THRES  <- 50
-SZA_THRES  <- 3
 ```
@@
 
+Data quality for radiation measurements as described by
+CN Long and Y Shi, September 2006, DOE/SC-ARM/TR-074.
+Quality Control Procedures/The QCRad Value Added Product Surface
+Radiation Measurement Quality Control Testing Including Climatology_Long2006.pdf"
+
 ## Data preperation
 
-```{r echo=F, include=T}
+Due to the large measurement assertenty near the horizon we exclude all measurements
+with greater than SZA $85^\circ$. Moreover, due to nearby hills and buildings
+we remove data with azimuth angle between $35^\circ$ and $120^\circ$ and SZA greater than $80^\circ$.
 
-####  Get data from Clear sky id data  ####
-input_files <- list.files( path       = CLEARdir,
-                           pattern    = "Clear_Sky_[0-9]{4}.Rds",
-                           full.names = T )
-
-if ( !file.exists(CS_file) | max(file.mtime(input_files)) > file.mtime(CS_file)) {
-    cat(paste("Load data from Clear Sky proccess from original\n"))
-    DATA <- data.table()
-    for (af in input_files) {
-        temp <- readRDS(af)
-        ## drop some data
-        temp$CHP1temp        <- NULL
-        temp$CHP1tempSD      <- NULL
-        temp$CHP1tempUNC     <- NULL
-        temp$CS_ref          <- NULL
-        temp$Pressure_Source <- NULL
-        temp$chp1TempCF      <- NULL
-        temp$wattDIR_tmp_cr  <- NULL
-        temp$wattDIR_unc_NT  <- NULL
-        temp$wattDIR_unc_WT  <- NULL
-        temp$wattHOR_tmp_cr  <- NULL
-        temp$wattHOR_unc_NT  <- NULL
-        temp$wattHOR_unc_WT  <- NULL
-
-        DATA <- rbind(temp, DATA, fill = TRUE)
-        rm(temp)
-    }
-    ## this is used by old scripts
-    myRtools::write_RDS(object = DATA, file = CS_file)
-} else {
-    # cat(paste("Load data from Clear Sky proccess from parsed\n"))
-    DATA <- readRDS(CS_file)
-}
-
-```
-
-Due to the large measurement assertenty near the horizon we exclude all measurements
-with greater than SZA $`r 90-MIN_ELEVA`^\circ$. Moreover, due to nearby hills and buildings
-we remove data with azimuth angle between $35^\circ$ and $120^\circ$ and SZA greater than $`r 90-10`^\circ$.
-
-```{r echo=F}
-DATA <- DATA[ Elevat >= MIN_ELEVA ]
-DATA <- DATA[ ! (Azimuth > 35 & Azimuth < 120 & Elevat < 10) ]
-```
-
-## Data selection
+## Agregation
 
 **END**
-
-```{r include=T, echo=F}
-tac <- Sys.time()
-cat(sprintf("%s %s@%s %f mins\n\n",Sys.time(),Sys.info()["login"],Sys.info()["nodename"],difftime(tac,tic,units="mins")))
-```

### 2022-10-05 Climatological_DHI_GHI_longterm_trends.R

@@ -438,7 +438,7 @@ CLEAR_daily_seas[, DIR_transp:= 100*( DIR_transp - DIR_transp_seas ) / DIR_trans
 
 #' \newpage
 #' ## Trends on all sky conditions data
-#+ echo=F, include=T
+#+ longtermtrends, echo=F, include=T
 
 gather <- data.frame()
 timefactor <- 1

### 2022-10-05 DHI_GHI_longterm_trends.R
Commit 2022-10-05 15:07

@@ -145,6 +145,11 @@ if ( !file.exists(CS_file) | max(file.mtime(input_files)) > file.mtime(CS_file))
 }
 
 
+#' ### Data range
+#' Time data span `r range(DATA$Date)`
+#'
+
+
 
 
 #' ### Filter min elevation
@@ -438,7 +443,7 @@ CLEAR_daily_seas[, DIR_transp:= 100*( DIR_transp - DIR_transp_seas ) / DIR_trans
 
 #' \newpage
 #' ## Trends on all sky conditions data
-#+ longtermtrends, echo=F, include=T
+#+ longtermtrendsALL, echo=F, include=T
 
 gather <- data.frame()
 timefactor <- 1
@@ -496,7 +501,7 @@ title(paste("All sky daily Direct transparency"),cex=0.9)
 
 #' \newpage
 #' ## Trends on clear sky data
-#+ echo=F, include=T
+#+ longtermtrendsCS, echo=F, include=T
 
 plot(CLEAR_daily_seas$Date, CLEAR_daily_seas$DIR_att, pch  = ".", xlab = "" )
 lm1 <- lm( DIR_att ~ Date , data = CLEAR_daily_seas)
@@ -559,6 +564,7 @@ pprint[, slope.sd := slope.sd/timefactor*365 ]
 
 pander(pprint,
        cap = "Slope is in %/year")
+myRtools::write_dat(pprint, "~/MANUSCRIPTS/2022_sdr_trends/figures/tbl_longterm_trends.dat")
 #+ echo=F, include=F
 
 ## Test plots of all variables

### 2022-10-05 Article.Rmd
Commit 2022-10-05 15:07

@@
 classoption:    a4paper,oneside
-fontsize:       10pt
+fontsize:       12pt
 geometry:       "left=0.5in,right=0.5in,top=0.5in,bottom=0.5in"
@@
 
-....... span
+<!-- time span -->
+Time span for CM21 is 1993-2022 and 2016-2022 for CHP1.
 
-...... measuremetns CHP1 CM21
+...... Measurements CHP1 CM21
 
-....... screening
-
-....... Quality assurance [cite]
-
-Data quality for radiation measurements as described by
-CN Long and Y Shi, September 2006, DOE/SC-ARM/TR-074.
-Quality Control Procedures/The QCRad Value Added Product Surface
-Radiation Measurement Quality Control Testing Including Climatology_Long2006.pdf"
-
-## Data preperation
-
-Due to the large measurement assertenty near the horizon we exclude all measurements
-with greater than SZA $85^\circ$. Moreover, due to nearby hills and buildings
+<!-- screening -->
+All measurements have been manually screened for inconsistencies and obviously 
+problematic data have been excluded.
+<!-- physical quantities -->
+The conversions to irradiance was done using the calibration factor  .......chp1.... cmp21......
+, and adjusting for the dark signal offset daily, using the nights before and after each day.
+<!-- Quality assurance -->
+A radiation data quality assurance procedure was applied adjusted for the site,
+based on methods of Long and Shi [-@long_automated_2008; -@Long2006].
+<!-- Data preparation -->
+Due to the large measurement uncertainty near the horizon we exclude all measurements
+with SZA greater than $85^\circ$. Moreover, due to nearby hills and buildings
 we remove data with azimuth angle between $35^\circ$ and $120^\circ$ and SZA greater than $80^\circ$.
 
-## Agregation
+![this is a caption](images/longtermtrendsALL-2.pdf)
+
+## Aggregation
 

### 2022-10-05 Article.Rmd
Commit 2022-10-05 16:07

@@
 
+<!-- \tiny -->
+```{r echo=F, cap="dfsfdffds fs"}
+pp <- data.table::fread("./figures/tbl_longterm_trends.dat")
+pander::pander(pp)
+```
+<!-- \normalsize -->
+
 ## Aggregation
@@
 **END**
+
+<!-- ```{r trackercomm, echo=F, fig.cap="Διάγραμμα συνδέσεων του tracker, κατάλληλη σειρά σύνδεσης (1-2-3) ή (3-2-1)."} -->
+<!-- knitr::include_graphics("figure/trackercomm.png") -->
+<!-- ``` -->
+<!-- όπως φαίνονται στο Σχήμα \@ref(fig:trackercomm). -->

### 2022-10-05 Article.Rmd
Commit 2022-10-05 17:07

@@
 A radiation data quality assurance procedure was applied adjusted for the site,
-based on methods of Long and Shi [-@long_automated_2008; -@Long2006].
+based on methods of Long and Shi\ [-@long_automated_2008; -@Long2006].
 <!-- Data preparation -->
@@
 pp <- data.table::fread("./figures/tbl_longterm_trends.dat")
-pander::pander(pp)
+pander::pander(pp,
+               caption = "Table of trends and stats")
 ```

### 2022-10-05 Climatological_DHI_GHI_longterm_trends.R

@@ -1,12 +1,11 @@
 # /* !/usr/bin/env Rscript */
 # /* Copyright (C) 2022 Athanasios Natsis <natsisthanasis@gmail.com> */
 #' ---
-#' title:         "Trends of SDR in Thessaloniki "
-#' author:
-#'   - Natsis Athanasios^[Laboratory of Atmospheric Physics,AUTH, natsisthanasis@gmail.com]
-#'   - Jane Doe^[Institution Two, jane@example.org]
-#' abstract:
-#'   "Study of GHI and DNI radiation for 'clear sky' and all sly conditions."
+#' title:         ""
+#' author:        "Natsis Athanasios"
+#' institute:     "AUTH"
+#' affiliation:   "Laboratory of Atmospheric Physics"
+#' abstract:      "Read "
 #'
 #' documentclass:  article
 #' classoption:    a4paper,oneside
@@ -63,9 +62,9 @@ Script.Name <- tryCatch({ funr::sys.script() },
                         error = function(e) { cat(paste("\nUnresolved script name: ", e),"\n\n")
                             return("Climatological_") })
 if(!interactive()) {
-    pdf(  file = paste0("~/MANUSCRIPTS/2022_sdr_trends/runtime/", basename(sub("\\.R$",".pdf", Script.Name))))
-    sink( file = paste0("~/MANUSCRIPTS/2022_sdr_trends/runtime/", basename(sub("\\.R$",".out", Script.Name))), split=TRUE)
-    filelock::lock(paste0("~/MANUSCRIPTS/2022_sdr_trends/runtime//",  basename(sub("\\.R$",".lock", Script.Name))), timeout = 0)
+    pdf(  file = paste0("~/Aerosols/REPORTS/runtime/", basename(sub("\\.R$",".pdf", Script.Name))))
+    sink( file = paste0("~/Aerosols/REPORTS/runtime/", basename(sub("\\.R$",".out", Script.Name))), split=TRUE)
+    filelock::lock(paste0("~/Aerosols/LOGs/",  basename(sub("\\.R$",".lock", Script.Name))), timeout = 0)
 }
 
[…]

### 2022-10-05 Climatological_DHI_GHI.R
Commit 2022-10-05 17:07

@@ -1,0 +1,1021 @@
+# /* !/usr/bin/env Rscript */
+# /* Copyright (C) 2022 Athanasios Natsis <natsisthanasis@gmail.com> */
+#' ---
+#' title:         ""
+#' author:        "Natsis Athanasios"
+#' institute:     "AUTH"
+#' affiliation:   "Laboratory of Atmospheric Physics"
+#' abstract:      "Read "
+#'
+#' documentclass:  article
+#' classoption:    a4paper,oneside
+#' fontsize:       10pt
+#' geometry:       "left=0.5in,right=0.5in,top=0.5in,bottom=0.5in"
+#' link-citations: yes
+#' colorlinks:     yes
+#'
+#' header-includes:
+#' - \usepackage{caption}
+#' - \usepackage{placeins}
+#' - \captionsetup{font=small}
+#'
+#' output:
+#'   bookdown::pdf_document2:
+#'     number_sections:  no
+#'     fig_caption:      no
+#'     keep_tex:         no
+#'     latex_engine:     xelatex
+#'     toc:              yes
+#'     fig_width:        7
+#'     fig_height:       4.5
+#'   html_document:
+#'     toc:        true
+#'     fig_width:  7.5
+#'     fig_height: 5
+#' date: "`r format(Sys.time(), '%F')`"
+#'
+#' ---
+
+#'
+#'
+#+ echo=F, include=T
+
+
+####_  Document options _####
+
+#+ echo=F, include=F
+knitr::opts_chunk$set(comment    = ""      )
+knitr::opts_chunk$set(dev        = "pdf"   )
+# knitr::opts_chunk$set(dev        = "png"    )
+knitr::opts_chunk$set(out.width  = "100%"   )
+knitr::opts_chunk$set(fig.align  = "center" )
+knitr::opts_chunk$set(cache      =  T       )
+# knitr::opts_chunk$set(fig.pos    = '!h'    )
+
+
+## TODO
+## - go to mean distance
+## - do it with clouds cloud trends
+## - do it without clouds aerosols trends
+## - climatologist analysis
+
+
+#+ include=F, echo=F
+####  Set environment  ####
+Sys.setenv(TZ = "UTC")
+tic <- Sys.time()
+Script.Name <- tryCatch({ funr::sys.script() },
+                        error = function(e) { cat(paste("\nUnresolved script name: ", e),"\n\n")
+                            return("Climatological_") })
+if(!interactive()) {
[…]

### 2022-10-05 Climatological_DHI_GHI_trends_consistency.R

@@ -1,0 +1,839 @@
+# /* !/usr/bin/env Rscript */
+# /* Copyright (C) 2022 Athanasios Natsis <natsisthanasis@gmail.com> */
+#' ---
+#' title:         ""
+#' author:        "Natsis Athanasios"
+#' institute:     "AUTH"
+#' affiliation:   "Laboratory of Atmospheric Physics"
+#' abstract:      "Read "
+#'
+#' documentclass:  article
+#' classoption:    a4paper,oneside
+#' fontsize:       10pt
+#' geometry:       "left=0.5in,right=0.5in,top=0.5in,bottom=0.5in"
+#' link-citations: yes
+#' colorlinks:     yes
+#'
+#' header-includes:
+#' - \usepackage{caption}
+#' - \usepackage{placeins}
+#' - \captionsetup{font=small}
+#'
+#' output:
+#'   bookdown::pdf_document2:
+#'     number_sections:  no
+#'     fig_caption:      no
+#'     keep_tex:         no
+#'     latex_engine:     xelatex
+#'     toc:              yes
+#'     fig_width:        7
+#'     fig_height:       4.5
+#'   html_document:
+#'     toc:        true
+#'     fig_width:  7.5
+#'     fig_height: 5
+#' date: "`r format(Sys.time(), '%F')`"
+#'
+#' ---
+
+#'
+#'
+#+ echo=F, include=T
+
+
+####_  Document options _####
+
+#+ echo=F, include=F
+knitr::opts_chunk$set(comment    = ""      )
+knitr::opts_chunk$set(dev        = "pdf"   )
+# knitr::opts_chunk$set(dev        = "png"    )
+knitr::opts_chunk$set(out.width  = "80%"   )
+knitr::opts_chunk$set(fig.align  = "center" )
+knitr::opts_chunk$set(cache      =  T       )
+# knitr::opts_chunk$set(fig.pos    = '!h'    )
+
+
+## TODO
+## - go to mean distance
+## - do it with clouds cloud trends
+## - do it without clouds aerosols trends
+## - climatologist analysis
+
+
+#+ include=F, echo=F
+####  Set environment  ####
+Sys.setenv(TZ = "UTC")
+tic <- Sys.time()
+Script.Name <- tryCatch({ funr::sys.script() },
+                        error = function(e) { cat(paste("\nUnresolved script name: ", e),"\n\n")
+                            return("Climatological_") })
+if(!interactive()) {
[…]

### 2022-10-05 Article.Rmd
Commit 2022-10-05 19:07

@@
 
-<!-- ```{r trackercomm, echo=F, fig.cap="Διάγραμμα συνδέσεων του tracker, κατάλληλη σειρά σύνδεσης (1-2-3) ή (3-2-1)."} -->
-<!-- knitr::include_graphics("figure/trackercomm.png") -->
-<!-- ``` -->
-<!-- όπως φαίνονται στο Σχήμα \@ref(fig:trackercomm). -->
+```{r trackercomm, echo=F, fig.cap="Διάγραμμα συνδέσεων του tracker, κατάλληλη σειρά σύνδεσης (1-2-3) ή (3-2-1)."}
+knitr::include_graphics("images/longtermtrendsALL-1.pdf")
+```
+νονται στο Σχήμα \@ref(fig:trackercomm).

### 2022-10-05 Article.Rmd
Commit 2022-10-05 20:07

@@
 knitr::opts_chunk$set(fig.align  = "center" )
-# knitr::opts_chunk$set(fig.pos    = '!h'    )
+knitr::opts_chunk$set(fig.pos    = '!h'    )
 
@@
 
-![this is a caption](images/longtermtrendsALL-2.pdf)
+```{r trackercomm, echo=F, fig.cap="tracker"}
+knitr::include_graphics("./images/longtermtrendsALL-2.pdf")
+```
+
+This is it \@ref(fig:trackercomm).
+
+\FloatBarrier
 
 <!-- \tiny -->
-```{r echo=F, cap="dfsfdffds fs"}
+```{r echo=F}
 pp <- data.table::fread("./figures/tbl_longterm_trends.dat")
@@
 **END**
-
-```{r trackercomm, echo=F, fig.cap="Διάγραμμα συνδέσεων του tracker, κατάλληλη σειρά σύνδεσης (1-2-3) ή (3-2-1)."}
-knitr::include_graphics("images/longtermtrendsALL-1.pdf")
-```
-νονται στο Σχήμα \@ref(fig:trackercomm).

### 2022-10-06 Article.Rmd
Commit 2022-10-06 08:07

@@
 A radiation data quality assurance procedure was applied adjusted for the site,
-based on methods of Long and Shi\ [-@long_automated_2008; -@Long2006].
+based on methods of Long and Shi\ [-@long_automated_2008; -@Long2006]. Only data characterized with low quality indicators was rejected.
 <!-- Data preparation -->
@@
 with SZA greater than $85^\circ$. Moreover, due to nearby hills and buildings
-we remove data with azimuth angle between $35^\circ$ and $120^\circ$ and SZA greater than $80^\circ$.
+we remove data with azimuth angle between $35^\circ$ and $120^\circ$ and SZA greater than $80^\circ$, as at those cases the Sun can not viewed by the instruments.
+
+## Data process
+We take some steps in order to make the measurements comparable throughout the dataset. 
+<!-- 1au -->
+We adjusted all measurements to the mean distance between Sun and Earth.
+<!-- TSI -->
+And, we make all measurements relative to the ......TSI...... , using an homogenized timeseries of satellite TSI data.
+
+<!-- Clear Sky -->
+We will use two dataset of our measurements, one for all sky conditions (ALL), which is the whole available data, and one filtered for 'clear sky' conditions (CS).
+The latter was created using ..........
+....
 

### 2022-10-06 DHI_GHI_sza_trends.R
Commit 2022-10-06 08:07

@@ -101,6 +101,10 @@ TEST      <- TRUE
 col_DIR_att    <- "#2166ac"
 col_DIR_transp <- "#9970ab"
 col_GLB_att    <- "#1a9850"
+pch_am         <-  1
+pch_pm         <-  4
+pch_ampm       <- 13 ## try 10
+pch_daily      <- 19
 
 MIN_ELEVA  <- 5  ## use data when sun above that
 SZA_BIN    <- 1
@@ -641,8 +645,17 @@ for (DBn in dbs) {
 hist(gather_seas$N[gather_seas$N>50], breaks = 100)
 
 szatrends_seas <- gather_seas
-szatrends_seas$preNoon[ szatrends_seas$preNoon == T ] <- 2
-szatrends_seas$preNoon[ szatrends_seas$preNoon == F ] <- 3
+
+
+## define plot colors
+##TODO
+szatrends_seas$preNoon[ szatrends_seas$preNoon == T, col := 2 ]
+szatrends_seas$preNoon[ szatrends_seas$preNoon == F, col := 3 ]
+szatrends_seas$preNoon[ szatrends_seas$preNoon == T, pch := pch_am ]
+szatrends_seas$preNoon[ szatrends_seas$preNoon == F, pch := pch_pm ]
+
+
+
 szatrends_seas <- data.table(szatrends_seas)
 
 hist(szatrends_seas[DATA==dbs[1],N], breaks = 100)
@@ -655,6 +668,9 @@ hist(szatrends_seas[var==vars[2],N], breaks = 100)
 
 plot(szatrends_seas$SZA,szatrends_seas$N)
 
+
+
+
 #+ echo=F, include=T
 for (ase in season) {
 
@@ -665,7 +681,7 @@ for (ase in season) {
                                        szatrends_seas$var    == avar &
                                        szatrends_seas$Season == ase    , ]
 
-            plot(subdata$SZA, subdata$slope, col = subdata$preNoon, pch = 19)
+            plot(subdata$SZA, subdata$slope, col = subdata$col, pch = subdata$pch)
             abline(h=0)
             title(paste(ase, "Slope",type, avar),cex=0.9)
             legend("top",

### 2022-10-06 Article.Rmd
Commit 2022-10-06 10:17

@@
 <!-- physical quantities -->
-The conversions to irradiance was done using the calibration factor  .......chp1.... cmp21......
-, and adjusting for the dark signal offset daily, using the nights before and after each day.
+The conversions to irradiance was done using the calibration factor
+.......chp1.... cmp21......
+, and adjusting for the dark signal offset daily, using the nights 
+before and after each day.
 <!-- Quality assurance -->
 A radiation data quality assurance procedure was applied adjusted for the site,
-based on methods of Long and Shi\ [-@long_automated_2008; -@Long2006]. Only data characterized with low quality indicators was rejected.
+based on methods of Long and Shi\ [-@long_automated_2008; -@Long2006]. 
+Only data characterized with low quality indicators was rejected.
 <!-- Data preparation -->
-Due to the large measurement uncertainty near the horizon we exclude all measurements
-with SZA greater than $85^\circ$. Moreover, due to nearby hills and buildings
-we remove data with azimuth angle between $35^\circ$ and $120^\circ$ and SZA greater than $80^\circ$, as at those cases the Sun can not viewed by the instruments.
+Due to the large measurement uncertainty near the horizon we exclude
+all measurements with SZA greater than $85^\circ$. Moreover, due to 
+nearby hills and buildings we remove data with azimuth angle between 
+$35^\circ$ and $120^\circ$ and SZA greater than $80^\circ$, as at 
+those cases the Sun can not viewed by the instruments.
 
 ## Data process
-We take some steps in order to make the measurements comparable throughout the dataset. 
+
+We take some steps in order to make the measurements comparable throughout
+the dataset. 
 <!-- 1au -->
@@
 <!-- TSI -->
-And, we make all measurements relative to the ......TSI...... , using an homogenized timeseries of satellite TSI data.
+And, we make all measurements relative to the .........TSI.......... , 
+using an homogenized timeseries of satellite TSI data...........
 
 <!-- Clear Sky -->
-We will use two dataset of our measurements, one for all sky conditions (ALL), which is the whole available data, and one filtered for 'clear sky' conditions (CS).
-The latter was created using ..........
-....
+We will use two dataset of our measurements, one for all sky conditions (ALL),
+which is the whole available data, and one filtered for 'clear sky' 
+conditions (CS). The latter was created, using a method adapted to our 
+site as proposed by @Long2000 and by @Reno2016.
+
+## Aggregation
+
+Both dataset were aggregate in SZA bins by am, pm, am+pm
+
+by daily mean
+
+seasonal daily mean
+
+Deseasonlized
+
+## Longterm trends
 
@@
 
-## Aggregation
+## SZA trends
+
+## Trends consistency
 

### 2022-10-06 DHI_GHI_sza_trends.R

@@ -495,11 +495,9 @@ for (DBn in dbs) {
 hist(gather$N[gather$N>50], breaks = 100)
 
 szatrends <- gather
-szatrends$preNoon[ szatrends$preNoon == T ] <- 2
-szatrends$preNoon[ szatrends$preNoon == F ] <- 3
 
 szatrends <- data.table(szatrends)
-setorder(szatrends_seas,SZA)
+setorder(szatrends,SZA)
 
 szatrends[ var == "DIR_att",    col := col_DIR_att    ]
 szatrends[ var == "GLB_att",    col := col_GLB_att    ]
@@ -520,12 +518,9 @@ hist(szatrends[var==vars[2],N], breaks = 100)
 
 plot(szatrends$SZA,szatrends$N)
 
-## vars to plot
+## stats vars to plot
 wecare <- grep( "^slope|^N",names(szatrends),ignore.case = T, value = T)
 
-wenames <- sub("\\."," ",wecare)
-
-gsub("(\\D)(\\D+)", "\\U\\1\\L\\2", wenames, perl = TRUE)
 
 #+ szatrends, echo=F, include=T
 ## ALL - CS
@@ -536,70 +531,82 @@ for (type in unique(szatrends$DATA)) {
         for (awe in wecare) {
             awename <- gsub("(\\D)(\\D+)", "\\U\\1\\L\\2", sub("\\."," ", awe), perl = TRUE)
 
-            ## selece All/CS and DIR/GLB/trans
+            ## select All/CS and DIR/GLB/trans
             subdata <- szatrends[ szatrends$DATA == type &    ##
                                   szatrends$var  == avar , ]
 
             xlim <- range( subdata$SZA,    na.rm = T )
             ylim <- range( subdata[[awe]], na.rm = T )
 
+            pam  <- subdata[ preNoon == T ]
+            ppm  <- subdata[ preNoon == F ]
+
             plot(1, type="n", xlab="SZA", ylab=awename, xlim=xlim, ylim=ylim )
             abline(h=0)
             title(paste(awename,type, avar),cex=0.9)
 
-            points()
+            # lines(pam$SZA, pam[[awe]], pch =  pch_am, col = pam$col, type = "b")
+            # lines(pam$SZA, ppm[[awe]], pch =  pch_pm, col = pam$col, type = "b")
 
+            lines(pam$SZA, pam[[awe]], pch =  pch_am, col = 2, type = "b")
[…]

### 2022-10-07 _main.Rmd
Commit 2022-10-07 08:17

---
title:         "Trends of SDR in Thessaloniki "
author:
  - Natsis Athanasios^[Laboratory of Atmospheric Physics,AUTH, natsisthanasis@gmail.com]
  - Jane Doe^[Institution Two, jane@example.org]
abstract:
  "Study of GHI and DNI radiation for 'clear sky' and all sly conditions."

documentclass:  article
classoption:    a4paper,oneside
fontsize:       12pt
geometry:       "left=0.5in,right=0.5in,top=0.5in,bottom=0.5in"
link-citations: yes
colorlinks:     yes
urlcolor:        blue

bibliography:    [references.bib]
biblio-style:    apalike

header-includes:
  - \usepackage{caption}
  - \usepackage{placeins}
  - \captionsetup{font=small}

output:
  bookdown::pdf_document2:
    number_sections:  no
    fig_caption:      yes
    keep_tex:         no
    latex_engine:     xelatex
    toc:              yes
    fig_width:        7
    fig_height:       4.5
  html_document:
    toc:        true
    fig_width:  7.5
    fig_height: 5
date: "`r format(Sys.time(), '%F')`"

---

```{r echo=F, include=F, cache=T}
knitr::opts_chunk$set(comment    = ""      )
knitr::opts_chunk$set(dev        = "pdf"   )
# knitr::opts_chunk$set(dev        = "png"    )
knitr::opts_chunk$set(out.width  = "80%"   )
knitr::opts_chunk$set(fig.align  = "center" )
knitr::opts_chunk$set(fig.pos    = '!h'    )

```

## Data description

<!-- time span -->
Time span for CM21 is 1993-2022 and 2016-2022 for CHP1.

...... Measurements CHP1 CM21

<!-- screening -->
All measurements have been manually screened for inconsistencies and obviously 
problematic data have been excluded.
<!-- physical quantities -->
The conversions to irradiance was done using the calibration factor
.......chp1.... cmp21......
, and adjusting for the dark signal offset daily, using the nights 
before and after each day.
<!-- Quality assurance -->
A radiation data quality assurance procedure was applied adjusted for the site,
based on methods of Long and Shi\ [-@long_automated_2008; -@Long2006]. 
Only data characterized with low quality indicators was rejected.
<!-- Data preparation -->
Due to the large measurement uncertainty near the horizon we exclude
all measurements with SZA greater than $85^\circ$. Moreover, due to 
nearby hills and buildings we remove data with azimuth angle between 
$35^\circ$ and $120^\circ$ and SZA greater than $80^\circ$, as at 
those cases the Sun can not viewed by the instruments.

## Data process

We take some steps in order to make the measurements comparable throughout
the dataset. 
<!-- 1au -->
We adjusted all measurements to the mean distance between Sun and Earth.
<!-- TSI -->
And, we make all measurements relative to the .........TSI.......... , 
using an homogenized timeseries of satellite TSI data...........

<!-- Clear Sky -->
We will use two dataset of our measurements, one for all sky conditions (ALL),
which is the whole available data, and one filtered for 'clear sky' 
conditions (CS). The latter was created, using a method adapted to our 
site as proposed by @Long2000 and by @Reno2016.

## Aggregation

Both dataset were aggregate in SZA bins by am, pm, am+pm

by daily mean

seasonal daily mean

Deseasonlized

## Longterm trends

```{r trackercomm, echo=F, fig.cap="tracker"}
knitr::include_graphics("./images/longtermtrendsALL-2.pdf")
```

This is it \@ref(fig:trackercomm).

\FloatBarrier

<!-- \tiny -->
```{r echo=F}
pp <- data.table::fread("./figures/tbl_longterm_trends.dat")
pander::pander(pp,
               caption = "Table of trends and stats")
```
<!-- \normalsize -->

## SZA trends

## Trends consistency

**END**

<!--chapter:end:Article.Rmd-->

```{r }
#!/usr/bin/env Rscript
```

---
title:         "Trends of SDR in Thessaloniki "
author:
  - Natsis Athanasios^[Laboratory of Atmospheric Physics,AUTH, natsisthanasis@gmail.com]
  - Jane Doe^[Institution Two, jane@example.org]
abstract:
  "Study of GHI and DNI radiation for 'clear sky' and all sly conditions."

documentclass:  article
classoption:    a4paper,oneside
fontsize:       10pt
geometry:       "left=0.5in,right=0.5in,top=0.5in,bottom=0.5in"
link-citations: yes
colorlinks:     yes

header-includes:
- \usepackage{caption}
- \usepackage{placeins}
- \captionsetup{font=small}

output:
  bookdown::pdf_document2:
    number_sections:  no
    fig_caption:      no
    keep_tex:         no
    latex_engine:     xelatex
    toc:              yes
    fig_width:        7
    fig_height:       4.5
  html_document:
    toc:        true
    fig_width:  7.5
    fig_height: 5
date: "`r format(Sys.time(), '%F')`"

---

```{r echo=F, include=T}

####_  Document options _####

```
```{r echo=F, include=F}
knitr::opts_chunk$set(comment    = ""      )
knitr::opts_chunk$set(dev        = "pdf"   )
# knitr::opts_chunk$set(dev        = "png"    )
knitr::opts_chunk$set(out.width  = "80%"   )
knitr::opts_chunk$set(fig.align  = "center" )
knitr::opts_chunk$set(cache      =  T       )
# knitr::opts_chunk$set(fig.pos    = '!h'    )

```
```{r include=F, echo=F}
####  Set environment  ####
Sys.setenv(TZ = "UTC")
tic <- Sys.time()
Script.Name <- tryCatch({ funr::sys.script() },
                        error = function(e) { cat(paste("\nUnresolved script name: ", e),"\n\n")
                            return("Climatological_") })
if(!interactive()) {
    pdf(  file = paste0("~/MANUSCRIPTS/2022_sdr_trends/runtime/", basename(sub("\\.R$",".pdf", Script.Name))))
    sink( file = paste0("~/MANUSCRIPTS/2022_sdr_trends/runtime/", basename(sub("\\.R$",".out", Script.Name))), split=TRUE)
    filelock::lock(paste0("~/MANUSCRIPTS/2022_sdr_trends/runtime//",  basename(sub("\\.R$",".lock", Script.Name))), timeout = 0)
}

par(pch = ".")

## FIXME this is for pdf output
# options(warn=-1) ## hide warnings
# options(warn=2)  ## stop on warnings

```
```{r echo=F, include=T}
####  External code  ####
library(data.table, quietly = T, warn.conflicts = F)
library(pander,     quietly = T, warn.conflicts = F)

## Functions from `https://github.com/thanasisn/IStillBreakStuff/tree/main/FUNCTIONS/R`
source("~/CODE/FUNCTIONS/R/sumNA.R")
source("~/CODE/FUNCTIONS/R/linear_regrassion_capture.R")
source("~/CODE/FUNCTIONS/R/trig_deg.R")

options(error = function() {
    if (interactive()) {
        # require("beepr"); beep(10)
        system("mplayer /usr/share/sounds/freedesktop/stereo/dialog-warning.oga", ignore.stdout = T, ignore.stderr = T)
        system("notify-send -u normal -t 30000 'R session' 'An error occured!'")
    }
})

####  Variables  ####
panderOptions('table.alignment.default', 'right')
panderOptions('table.split.table',        120   )

DATA_DIR  <- "/home/athan/DATA/Broad_Band"
CLEARdir  <- "/home/athan/DATA/Broad_Band/CS_id"
tag       <- paste0("Natsis Athanasios LAP AUTH ", strftime(Sys.time(), format = "%b %Y" ))
CS_file   <- "/home/athan/DATA/Common_application/Clear_Sky.Rds"

TEST      <- TRUE

col_DIR_att    <- "#2166ac"
col_DIR_transp <- "#9970ab"
col_GLB_att    <- "#1a9850"
pch_am         <-  1
pch_pm         <-  4
pch_ampm       <- 13 ## try 10
pch_daily      <- 19

MIN_ELEVA  <- 5  ## use data when sun above that
SZA_BIN    <- 1
MIN_N      <- 4
SEAS_MIN_N <- 3

SUM_THRES  <- 50
SZA_THRES  <- 3
```

## Data selection

```{r echo=F, include=T}

####  Get data from Clear sky id data  ####
input_files <- list.files( path    = CLEARdir,
                           pattern = "Clear_Sky_[0-9]{4}.Rds",
                           full.names = T )

if ( !file.exists(CS_file) | max(file.mtime(input_files)) > file.mtime(CS_file)) {
    cat(paste("Load data from Clear Sky proccess from original\n"))
    DATA <- data.table()
    for (af in input_files) {
        temp <- readRDS(af)
        ## drop some data
        temp$CHP1temp        <- NULL
        temp$CHP1tempSD      <- NULL
        temp$CHP1tempUNC     <- NULL
        temp$CS_ref          <- NULL
        temp$Pressure_Source <- NULL
        temp$chp1TempCF      <- NULL
        temp$wattDIR_tmp_cr  <- NULL
        temp$wattDIR_unc_NT  <- NULL
        temp$wattDIR_unc_WT  <- NULL
        temp$wattHOR_tmp_cr  <- NULL
        temp$wattHOR_unc_NT  <- NULL
        temp$wattHOR_unc_WT  <- NULL

        DATA <- rbind(temp, DATA, fill = TRUE)
        rm(temp)
    }
    ## this is used by old scripts
    myRtools::write_RDS(object = DATA, file = CS_file)
} else {
    cat(paste("Load data from Clear Sky proccess from parsed\n"))
    DATA <- readRDS(CS_file)
}
```

### Data range
Time data span `r range(DATA$Date)`

### Filter min elevation
Keep data with Sun elevation above `r MIN_ELEVA`

```{r }
DATA <- DATA[ Elevat >= MIN_ELEVA ]
```

### Bais paper obstacle filter

```{r }
DATA <- DATA[ ! (Azimuth > 35 & Azimuth < 120 & Elevat < 10) ]
```
```{r echo=F, include=T}
```

### Keep only data characterized as 'good' by the Radiation Quality control procedure

```{r echo=F, include=T}
DATA[ QCF_DIR != "good", wattDIR := NA ]
DATA[ QCF_DIR != "good", QCF_DIR := NA ]
DATA[ QCF_GLB != "good", wattGLB := NA ]
DATA[ QCF_GLB != "good", QCF_GLB := NA ]
DATA <- DATA[ ! (is.na(wattDIR) & is.na(wattGLB))  ]

## exclude some data for each instrument
# DATA[ QCF_DIR == "Extremely rare limits min (3)", wattDIR := NA ]
# DATA[ QCF_DIR == "Extremely rare limits min (3)", QCF_DIR := NA ]
#
# dirtb <- table(DATA$QCF_DIR)
# for (aty in names(dirtb)[!names(dirtb) %in% "good"]) {
#     temp <- DATA[ QCF_DIR == aty ]
#     cat(nrow(temp),aty,"\n")
# }
# glbtb <- table(DATA$QCF_GLB)
# for (aty in names(glbtb)[!names(glbtb) %in% "good"]) {
#     temp <- DATA[ QCF_GLB == aty ]
#     cat(nrow(temp),aty,"\n")
# }
```

## Data preparation
### Move measurements to mean earth distance

```{r }
DATA[ , wattDIR_1au := wattDIR * (sun_dist ^ 2) ]
DATA[ , wattGLB_1au := wattGLB * (sun_dist ^ 2) ]
```
```{r echo=F, include=T}
```

### Relative to actual TSI at 1au variable representation

```{r }
DATA[ , DIR_att := wattDIR_1au / tsi_1au_comb ]
DATA[ , GLB_att := wattGLB_1au / tsi_1au_comb ]
```

Using the actual values gives similar trends.

```{r echo=F, include=T}

# #' ### Use original variable representation
# DATA[ , DIR_att := wattDIR_1au ]
# DATA[ , GLB_att := wattGLB_1au ]
# #+ echo=F, include=T

# #' ### Ground effect removal?
# #' Aerosol direct effects on global solar shortwave irradiance at high mountainous station Musala Bulgaria_Nojarov2021.pdf
# DATA$wattGLB_1au <- DATA$wattGLB_1au / cosde(DATA$SZA)
# DATA$wattDIR_1au <- DATA$wattDIR_1au / cosde(DATA$SZA)
# #+ echo=F, include=T
```

### Calculate Bouguer atmospheric transparency
see: Changes in solar radiation and their influence on temperature trend in Estonia 1955 2007_Russak2009.pdf

```{r }
DATA[, DIR_transp := ( wattDIR_1au / tsi_1au_comb ) ^ ( 1 / cosde(SZA) ) ]
```
```{r echo=F, include=T}

## fix noon just in case
DATA[ Azimuth <= 180 , preNoon := TRUE  ]
DATA[ Azimuth >  180 , preNoon := FALSE ]
```

### Split data to Clear Sky and non Clear sky
Method based and adapted to site from: Identification of Periods of Clear Sky
Irradiance in Time Series of GHI Measurements
_Matthew J. Reno and Clifford W. Hansen_.

```{r echo=F, include=T}
DATA_all   <- DATA

wecare     <- grep("CSflag_", names(DATA), value = T)
## use only cm21 flags
wecare     <- grep("_11", wecare, invert = T, value = T)
DATA_Clear <- DATA[ rowSums( DATA[, ..wecare ], na.rm = T ) == 0, ]
```
```{r echo=F, include=T}

## old flags
# DATA_Clear <- DATA_all[ CSflag == 0 ]
rm(DATA)

# #### . . My CSid method   ####
# #'
# #' | CS Flag | Test |
# #' |:-------:|:--------------------------------------------------------------|
# #' |   NA    | Undefined, untested                                           |
# #' |    0    | Passed as clear sky                                           |
# #' |    1    | Mean value of irradiance during the time period (MeanVIP)     |
# #' |    2    | Max Value of Irradiance during the time Period (MaxVIP)       |
# #' |    3    | Variability in irradiance by the length (VIL)                 |
# #' |    4    | Variance of Changes in the Time series (VCT)                  |
# #' |    5    | Variability in the Shape of the irradiance Measurements (VSM) |
# #' |    6    | Low Direct Irradiance limit (LDI)                             |
# #' |    7    | Low Global Irradiance limit (LGI)                             |
# #' |    8    | Too Few CS point for the day (FCS)                            |
# #' |    9    | Too Few data points for the day                               |
# #' |   10    | Missing Data                                                  |
# #' |   11    | Direct irradiance simple threshold                            |

#### 2. Long term by SZA ####
```

## 2. Long term by SZA

```{r }
##TODO t-test and significance
```

#### Calculate daily SZA means ####

```{r echo=F, include=T}

ALL_daily_mean <- DATA_all[, .(DIR_att       = mean(DIR_att,    na.rm = T),
                               GLB_att       = mean(GLB_att,    na.rm = T),
                               DIR_transp    = mean(DIR_transp, na.rm = T),
                               DIR_att_sd    = sd(  DIR_att,    na.rm = T),
                               GLB_att_sd    = sd(  GLB_att,    na.rm = T),
                               DIR_transp_sd = sd(DIR_transp, na.rm = T),
                               doy           = yday(Date),
                               GLB_att_N     = sum(!is.na(GLB_att)),
                               DIR_att_N     = sum(!is.na(DIR_att))  ),
                           by = .(SZA     = (SZA - SZA_BIN / 2 ) %/% SZA_BIN,
                                  Date    = Day,
                                  preNoon = preNoon  ) ]

CLEAR_daily_mean <- DATA_Clear[, .(DIR_att       = mean(DIR_att,    na.rm = T),
                                   GLB_att       = mean(GLB_att,    na.rm = T),
                                   DIR_transp    = mean(DIR_transp, na.rm = T),
                                   DIR_att_sd    = sd(  DIR_att,    na.rm = T),
                                   GLB_att_sd    = sd(  GLB_att,    na.rm = T),
                                   DIR_transp_sd = sd(DIR_transp, na.rm = T),
                                   doy           = yday(Date),
                                   GLB_att_N     = sum(!is.na(GLB_att)),
                                   DIR_att_N     = sum(!is.na(DIR_att))  ),
                               by = .(SZA     = (SZA - SZA_BIN / 2 ) %/% SZA_BIN,
                                      Date    = Day,
                                      preNoon = preNoon ) ]
```

#### Margin of error calculation for 0.95 confidence interval ####

```{r echo=T, include=T}
CONF_INTERV <- .95
conf_param  <- 1-(1-CONF_INTERV)/2
suppressWarnings({
ALL_daily_mean[,   DIR_att_EM   := qt(conf_param,df=DIR_att_N -1) * DIR_att_sd    / sqrt(DIR_att_N)]
ALL_daily_mean[,   GLB_att_EM   := qt(conf_param,df=GLB_att_N -1) * GLB_att_sd    / sqrt(GLB_att_N)]
ALL_daily_mean[,   DIR_transp_EM:= qt(conf_param,df=DIR_att_N -1) * DIR_transp_sd / sqrt(DIR_att_N)]
CLEAR_daily_mean[, DIR_att_EM   := qt(conf_param,df=DIR_att_N -1) * DIR_att_sd    / sqrt(DIR_att_N)]
CLEAR_daily_mean[, GLB_att_EM   := qt(conf_param,df=GLB_att_N -1) * GLB_att_sd    / sqrt(GLB_att_N)]
CLEAR_daily_mean[, DIR_transp_EM:= qt(conf_param,df=DIR_att_N -1) * DIR_transp_sd / sqrt(DIR_att_N)]
})
```
```{r echo=F, include=F}
```

#### Exclude means with less than `r SZA_THRES` data points

```{r echo=F, include=T}
ALL_daily_mean[   DIR_att_N <= SZA_THRES, DIR_att    := NA ]
ALL_daily_mean[   GLB_att_N <= SZA_THRES, GLB_att    := NA ]
ALL_daily_mean[   DIR_att_N <= SZA_THRES, DIR_transp := NA ]
CLEAR_daily_mean[ DIR_att_N <= SZA_THRES, DIR_att    := NA ]
CLEAR_daily_mean[ GLB_att_N <= SZA_THRES, GLB_att    := NA ]
CLEAR_daily_mean[ DIR_att_N <= SZA_THRES, DIR_transp := NA ]
```

#### Calculate daily seasonal values by SZA ####

```{r echo=F, include=T}

ALL_daily_seas <-
    ALL_daily_mean[, .(DIR_att_seas    = mean(DIR_att,    na.rm = T),
                       GLB_att_seas    = mean(GLB_att,    na.rm = T),
                       DIR_transp_seas = mean(DIR_transp, na.rm = T),
                       DIR_att_sd_seas = sd(  DIR_att,    na.rm = T),
                       GLB_att_sd_seas = sd(  GLB_att,    na.rm = T),
                       GLB_att_N_seas  = sum(!is.na(GLB_att)),
                       DIR_att_N_seas  = sum(!is.na(DIR_att))  ),
                   by = .( doy, SZA, preNoon ) ]

CLEAR_daily_seas <-
    CLEAR_daily_mean[, .(DIR_att_seas    = mean(DIR_att,    na.rm = T),
                         GLB_att_seas    = mean(GLB_att,    na.rm = T),
                         DIR_transp_seas = mean(DIR_transp, na.rm = T),
                         DIR_att_sd_seas = sd(  DIR_att,    na.rm = T),
                         GLB_att_sd_seas = sd(  GLB_att,    na.rm = T),
                         GLB_att_N_seas  = sum(!is.na(GLB_att)),
                         DIR_att_N_seas  = sum(!is.na(DIR_att))  ),
                     by = .( doy, SZA, preNoon ) ]

```
```{r echo=F, include=F}
## ~ Plots longterm  ####
plot( ALL_daily_mean$Date, ALL_daily_mean$DIR_att )
plot( ALL_daily_mean$Date, ALL_daily_mean$GLB_att )
plot( ALL_daily_mean$Date, ALL_daily_mean$DIR_transp )

plot( ALL_daily_mean$doy, ALL_daily_mean$DIR_att )
plot( ALL_daily_mean$doy, ALL_daily_mean$GLB_att )
plot( ALL_daily_mean$doy, ALL_daily_mean$DIR_transp )

plot( ALL_daily_mean$doy, ALL_daily_mean$GLB_att_N)
plot( ALL_daily_mean$doy, ALL_daily_mean$DIR_att_N)

hist(ALL_daily_mean$DIR_att_N)
hist(ALL_daily_mean$GLB_att_N)

plot( CLEAR_daily_mean$Date, CLEAR_daily_mean$DIR_att )
plot( CLEAR_daily_mean$Date, CLEAR_daily_mean$GLB_att )
plot( CLEAR_daily_mean$Date, CLEAR_daily_mean$DIR_transp )

plot( CLEAR_daily_mean$doy,  CLEAR_daily_mean$DIR_att )
plot( CLEAR_daily_mean$doy,  CLEAR_daily_mean$GLB_att )
plot( CLEAR_daily_mean$doy,  CLEAR_daily_mean$DIR_transp )

plot( CLEAR_daily_mean$doy,  CLEAR_daily_mean$GLB_att_N)
plot( CLEAR_daily_mean$doy,  CLEAR_daily_mean$DIR_att_N)

hist( CLEAR_daily_mean$DIR_att_N)
hist( CLEAR_daily_mean$GLB_att_N)

## ~ Plots seasonal ####

plot( ALL_daily_seas$doy, ALL_daily_seas$DIR_att_seas )
plot( ALL_daily_seas$doy, ALL_daily_seas$GLB_att_seas )
plot( ALL_daily_seas$doy, ALL_daily_seas$DIR_transp_seas )
plot( ALL_daily_seas$doy, ALL_daily_seas$GLB_att_N_seas)
plot( ALL_daily_seas$doy, ALL_daily_seas$DIR_att_N_seas)

plot( CLEAR_daily_seas$doy, CLEAR_daily_seas$DIR_att_seas )
plot( CLEAR_daily_seas$doy, CLEAR_daily_seas$GLB_att_seas )
plot( CLEAR_daily_seas$doy, CLEAR_daily_seas$DIR_transp_seas )
plot( CLEAR_daily_seas$doy, CLEAR_daily_seas$GLB_att_N_seas)
plot( CLEAR_daily_seas$doy, CLEAR_daily_seas$DIR_att_N_seas)
```

#### Calculate seasonal anomaly ####

```{r echo=F, include=F}

ALL_daily_seas   <- merge(  ALL_daily_mean, ALL_daily_seas,   by = c("doy", "SZA", "preNoon"),  all = T)
CLEAR_daily_seas <- merge(CLEAR_daily_mean, CLEAR_daily_seas, by = c("doy", "SZA", "preNoon"), all = T)

setorder(ALL_daily_seas,Date)
setorder(CLEAR_daily_seas,Date)

## anomaly
# ALL_daily_seas[   , DIR_att    := DIR_att    - DIR_att_seas    ]
# ALL_daily_seas[   , GLB_att    := GLB_att    - GLB_att_seas    ]
# ALL_daily_seas[   , DIR_transp := DIR_transp - DIR_transp_seas ]
# CLEAR_daily_seas[ , DIR_att    := DIR_att    - DIR_att_seas    ]
# CLEAR_daily_seas[ , GLB_att    := GLB_att    - GLB_att_seas    ]
# CLEAR_daily_seas[ , DIR_transp := DIR_transp - DIR_transp_seas ]

## relative anomaly
```
```{r echo=T, include=T}
ALL_daily_seas[  , DIR_att   := 100 * ( DIR_att    - DIR_att_seas    ) / DIR_att_seas    ]
ALL_daily_seas[  , GLB_att   := 100 * ( GLB_att    - GLB_att_seas    ) / GLB_att_seas    ]
ALL_daily_seas[  , DIR_transp:= 100 * ( DIR_transp - DIR_transp_seas ) / DIR_transp_seas ]
CLEAR_daily_seas[, DIR_att   := 100 * ( DIR_att    - DIR_att_seas    ) / DIR_att_seas    ]
CLEAR_daily_seas[, GLB_att   := 100 * ( GLB_att    - GLB_att_seas    ) / GLB_att_seas    ]
CLEAR_daily_seas[, DIR_transp:= 100 * ( DIR_transp - DIR_transp_seas ) / DIR_transp_seas ]
```
```{r echo=F, include=F}

####  Plot of SZA trends for all year ####
```

\newpage
## Plot of SZA trends

```{r echo=F, include=F}
timefactor <- 1
vars <- c("DIR_att", "GLB_att", "DIR_transp")
dbs  <- c("ALL_daily_seas", "CLEAR_daily_seas")

gather <- data.frame()

for (DBn in dbs) {
    DB <- get(DBn)
    for (avar in vars) {
        for (anoon in unique( DB$preNoon)) {
            for (asza in unique( DB$SZA )) {

                dataset <- DB[ SZA == asza & preNoon == anoon ]

                if (sum(!is.na(dataset[[avar]])) <= 1) next()

                lm1 <- lm( dataset[[avar]] ~ dataset$Date )

                gather <- rbind(gather,
                                data.frame(
                                    linear_regression_capture(lm1),
                                    preNoon   = anoon,
                                    SZA       = asza,
                                    DATA      = DBn,
                                    var       = avar,
                                    N         = sum(!is.na(dataset[[avar]]))
                                ))

                # plot(dataset$Date, dataset[[avar]], pch  = "." )
                # abline(lm1)
                # fit <- lm1[[1]]
                # legend('top', lty = 1, bty = "n",
                #        paste('Y =', signif(fit[1],2),if(fit[2]>0)'+'else'-',signif(abs(fit[2]/timefactor*365),3),'* year'))
                # title(paste(DBn, asza, anoon))
            }
        }
    }
}

```
```{r echo=F, include=F}
hist(gather$N[gather$N>50], breaks = 100)

szatrends <- gather

szatrends <- data.table(szatrends)
setorder(szatrends,SZA)

szatrends[ var == "DIR_att",    col := col_DIR_att    ]
szatrends[ var == "GLB_att",    col := col_GLB_att    ]
szatrends[ var == "DIR_transp", col := col_DIR_transp ]
szatrends[ preNoon == T, pch := pch_am ]
szatrends[ preNoon == F, pch := pch_pm ]

hist(szatrends[DATA==dbs[1],N], breaks = 100)
hist(szatrends[DATA==dbs[2],N], breaks = 100)

hist(szatrends[var==vars[1],N], breaks = 100)
hist(szatrends[var==vars[2],N], breaks = 100)

# szatrends <- szatrends[ N > 50]

plot(szatrends$SZA,szatrends$N)

test1 <- szatrends[ DATA == "CLEAR_daily_seas" & var == "DIR_att" ]
test2 <- szatrends[ DATA == "CLEAR_daily_seas" & var == "GLB_att" ]
plot(test1$SZA, test1$N, pch =19)
abline(h=50)
plot(test2$SZA, test2$N, pch =19)
abline(h=300)

# szatrends[ var == "GLB_att"    & N <= 300, slope := NA ]
# szatrends[ var == "DIR_att"    & N <=  50, slope := NA ]
# szatrends[ var == "DIR_transp" & N <=  50, slope := NA ]

## stats vars to plot
wecare <- grep( "^slope|^N",names(szatrends),ignore.case = T, value = T)

```
```{r szatrends, echo=F, include=T}
## ALL - CS
for (type in unique(szatrends$DATA)) {
    ## DIR - GLB - transp
    for (avar in unique(szatrends$var)) {
        ## statistic variable
        for (awe in wecare) {
            awename <- gsub("(\\D)(\\D+)", "\\U\\1\\L\\2", sub("\\."," ", awe), perl = TRUE)

            ## select All/CS and DIR/GLB/trans
            subdata <- szatrends[ szatrends$DATA == type &    ##
                                  szatrends$var  == avar , ]

            xlim <- range( subdata$SZA,    na.rm = T )
            ylim <- range( subdata[[awe]], na.rm = T )

            pam  <- subdata[ preNoon == T ]
            ppm  <- subdata[ preNoon == F ]

            plot(1, type="n", xlab="SZA", ylab=awename, xlim=xlim, ylim=ylim )
            abline(h=0)
            title(paste(awename,type, avar),cex=0.9)

            # lines(pam$SZA, pam[[awe]], pch =  pch_am, col = pam$col, type = "b")
            # lines(pam$SZA, ppm[[awe]], pch =  pch_pm, col = pam$col, type = "b")

            lines(pam$SZA, pam[[awe]], pch =  pch_am, col = 2, type = "b")
            lines(ppm$SZA, ppm[[awe]], pch =  pch_pm, col = 3, type = "b")

            legend("top",
                   legend = c("Morning",       "Evening"),
                   # col    = c(unique(pam$col), unique(ppm$col)),
                   col    = c(2, 3),
                   pch    = c(unique(pam$pch), unique(ppm$pch)), ncol = 2, bty = "n")
        }

#
#         subdata <- szatrends[ szatrends$DATA == type & szatrends$var == avar , ]
#
#         plot(subdata$SZA, subdata$slope, col = subdata$preNoon, pch = 19)
#         abline(h=0)
#         title(paste("Slope",type, avar),cex=0.9)
#         legend("top",
#                legend = c("Morning", "Evening"),
#                col    = c(2 , 3),
#                pch    = 19, ncol = 2, bty = "n")
#
#         plot(subdata$SZA, subdata$slope.sd, col = subdata$preNoon, pch = 19)
#         abline(h=0)
#         title(paste("Slope sd",type, avar),cex=0.9)
#         legend("top",
#                legend = c("Morning", "Evening"),
#                col    = c(2 , 3),
#                pch    = 19, ncol = 2, bty = "n")
#
#         plot(subdata$SZA, subdata$slope.t, col = subdata$preNoon, pch = 19)
#         abline(h=0)
#         title(paste("Slope t",type, avar),cex=0.9)
#         legend("top",
#                legend = c("Morning", "Evening"),
#                col    = c(2 , 3),
#                pch    = 19, ncol = 2, bty = "n")
#
#         plot(subdata$SZA, subdata$slope.p, col = subdata$preNoon, pch = 19)
#         abline(h=0)
#         title(paste("Slope p",type, avar),cex=0.9)
#         legend("top",
#                legend = c("Morning", "Evening"),
#                col    = c(2 , 3),
#                pch    = 19, ncol = 2, bty = "n")
#
#         plot(subdata$SZA, subdata$Rsqrd, col = subdata$preNoon, pch = 19)
#         abline(h=0)
#         title(paste("Slope Rsq",type, avar),cex=0.9)
#         legend("top",
#                legend = c("Morning", "Evening"),
#                col    = c(2 , 3),
#                pch    = 19, ncol = 2, bty = "n")
#
#         plot(subdata$SZA, subdata$N, col = subdata$preNoon, pch = 19)
#         title(paste("N",type, avar),cex=0.9)
#         legend("top",
#                legend = c("Morning", "Evening"),
#                col    = c(2 , 3),
#                pch    = 19, ncol = 2, bty = "n")

    }
}

####  Plot of SZA trends for each season of year ####
```

\newpage
## Plot of SZA trends for each season of year

```{r echo=F, include=F}
timefactor  <- 1  ## to display % per year
vars        <- c("DIR_att", "GLB_att", "DIR_transp")
dbs         <- c("ALL_daily_seas", "CLEAR_daily_seas")
season      <- c("Winter", "Spring", "Summer", "Automn")
gather_seas <- data.frame()

for (DBn in dbs) {
    DB <- get(DBn)

    ## set seasons in each data base
    DB[ month(Date) %in% c(12, 1, 2), Season := "Winter"]
    DB[ month(Date) %in% c( 2, 4, 5), Season := "Spring"]
    DB[ month(Date) %in% c( 6, 7, 8), Season := "Summer"]
    DB[ month(Date) %in% c( 9,10,11), Season := "Automn"]

    for (ase in season) {
        for (avar in vars) {
            for (anoon in unique( DB$preNoon)) {
                for (asza in unique( DB$SZA )) {

                    dataset <- DB[ SZA == asza & preNoon == anoon & season == ase ]

                    if (sum(!is.na(dataset[[avar]])) <= 1) next()

                    lm1 <- lm( dataset[[avar]] ~ dataset$Date )

                    gather_seas <- rbind(gather_seas,
                                    data.frame(
                                        linear_regression_capture(lm1),
                                        preNoon   = anoon,
                                        SZA       = asza,
                                        DATA      = DBn,
                                        var       = avar,
                                        Season    = ase,
                                        N         = sum(!is.na(dataset[[avar]]))
                                    ))

                    # plot(dataset$Date, dataset[[avar]], pch  = "." )
                    # abline(lm1)
                    # fit <- lm1[[1]]
                    # legend('top', lty = 1, bty = "n",
                    #        paste('Y =', signif(fit[1],2),if(fit[2]>0)'+'else'-',signif(abs(fit[2]/timefactor*365),3),'* year'))
                    # title(paste(DBn, asza, anoon))
                }
            }
        }
    }
}

```
```{r echo=F, include=F}
hist(gather_seas$N[gather_seas$N>50], breaks = 100)

szatrends_seas <- data.table(gather_seas)
setorder(szatrends_seas,SZA)

## define plot colors
szatrends_seas[ var == "DIR_att",    col := col_DIR_att    ]
szatrends_seas[ var == "GLB_att",    col := col_GLB_att    ]
szatrends_seas[ var == "DIR_transp", col := col_DIR_transp ]
szatrends_seas[ preNoon == T, pch := pch_am ]
szatrends_seas[ preNoon == F, pch := pch_pm ]

hist(szatrends_seas[DATA==dbs[1],N], breaks = 100)
hist(szatrends_seas[DATA==dbs[2],N], breaks = 100)
hist(szatrends_seas[var==vars[1],N], breaks = 100)
hist(szatrends_seas[var==vars[2],N], breaks = 100)

# szatrends_seas <- szatrends_seas[ N > 50]

plot(szatrends_seas$SZA,szatrends_seas$N)

test <- szatrends_seas[ DATA == "CLEAR_daily_seas" & var == "DIR_att" ]
plot(test$SZA, test$N, pch =19)
abline(h=50/4)

szatrends[ N <= 30, slope := NA]

test1 <- szatrends_seas[ DATA == "CLEAR_daily_seas" & var == "DIR_att" ]
test2 <- szatrends_seas[ DATA == "CLEAR_daily_seas" & var == "GLB_att" ]
plot(test1$SZA, test1$N, pch =19)
abline(h=50/4)
plot(test2$SZA, test2$N, pch =19)
abline(h=300/4)

# szatrends[ var == "GLB_att"    & N <= 300, slope := NA ]
# szatrends[ var == "DIR_att"    & N <=  50, slope := NA ]
# szatrends[ var == "DIR_transp" & N <=  50, slope := NA ]

## stats vars to plot
wecare <- grep( "^slope|^N",names(szatrends_seas),ignore.case = T, value = T)

```
```{r szatrendsseas, echo=F, include=T}
## Winter - Summer ....
for (ase in season) {
    ## ALL - Clear sky
    for (type in unique(szatrends_seas$DATA)) {
        ## DIR - GLB - transp
        for (avar in unique(szatrends_seas$var)) {

            ## statistic variable
            for (awe in wecare) {
                awename <- gsub("(\\D)(\\D+)", "\\U\\1\\L\\2", sub("\\."," ", awe), perl = TRUE)

                ## select All/CS  DIR/GLB/trans winter/summer
                subdata <- szatrends_seas[ DATA   == type &
                                           var    == avar &
                                           Season == ase    , ]

                xlim <- range( subdata$SZA,    na.rm = T )
                ylim <- range( subdata[[awe]], na.rm = T )

                pam  <- subdata[ preNoon == T ]
                ppm  <- subdata[ preNoon == F ]

                plot(1, type="n", xlab="SZA", ylab=awename, xlim=xlim, ylim=ylim )
                abline(h=0)
                title(paste(ase,awename,type, avar),cex=0.9)

                # lines(pam$SZA, pam[[awe]], pch =  pch_am, col = pam$col, type = "b")
                # lines(pam$SZA, ppm[[awe]], pch =  pch_pm, col = pam$col, type = "b")

                lines(pam$SZA, pam[[awe]], pch =  pch_am, col = 2, type = "b")
                lines(ppm$SZA, ppm[[awe]], pch =  pch_pm, col = 3, type = "b")

                legend("top",
                       legend = c("Morning",       "Evening"),
                       # col    = c(unique(pam$col), unique(ppm$col)),
                       col    = c(2, 3),
                       pch    = c(unique(pam$pch), unique(ppm$pch)), ncol = 2, bty = "n")
            }

#
#             subdata <- szatrends_seas[ DATA   == type &
#                                        var    == avar &
#                                        Season == ase    , ]
#
#
#             pam <- subdata[ preNoon == T ]
#             ppm <- subdata[ preNoon == F ]
#
#             plot(subdata$SZA, subdata$slope, col = subdata$col, pch = subdata$pch)
#             abline(h=0)
#             title(paste(ase, "Slope",type, avar),cex=0.9)
#             legend("top",
#                    legend = c("Morning", "Evening"),
#                    col    = c(2 , 3),
#                    pch    = c(pch_am,     pch_pm), ncol = 2, bty = "n")
#
            # plot(subdata$SZA, subdata$slope.sd, col = subdata$preNoon, pch = 19)
            # abline(h=0)
            # title(paste("Slope sd",type, avar),cex=0.9)
            # legend("top",
            #        legend = c("Morning", "Evening"),
            #        col    = c(2 , 3),
            #        pch    = 19, ncol = 2, bty = "n")
            #
            # plot(subdata$SZA, subdata$slope.t, col = subdata$preNoon, pch = 19)
            # abline(h=0)
            # title(paste("Slope t",type, avar),cex=0.9)
            # legend("top",
            #        legend = c("Morning", "Evening"),
            #        col    = c(2 , 3),
            #        pch    = 19, ncol = 2, bty = "n")
            #
            # plot(subdata$SZA, subdata$slope.p, col = subdata$preNoon, pch = 19)
            # abline(h=0)
            # title(paste("Slope p",type, avar),cex=0.9)
            # legend("top",
            #        legend = c("Morning", "Evening"),
            #        col    = c(2 , 3),
            #        pch    = 19, ncol = 2, bty = "n")
            #
            # plot(subdata$SZA, subdata$Rsqrd, col = subdata$preNoon, pch = 19)
            # abline(h=0)
            # title(paste("Slope Rsq",type, avar),cex=0.9)
            # legend("top",
            #        legend = c("Morning", "Evening"),
            #        col    = c(2 , 3),
            #        pch    = 19, ncol = 2, bty = "n")
            #
            # plot(subdata$SZA, subdata$N, col = subdata$preNoon, pch = 19)
            # title(paste("N",type, avar),cex=0.9)
            # legend("top",
            #        legend = c("Morning", "Evening"),
            #        col    = c(2 , 3),
            #        pch    = 19, ncol = 2, bty = "n")

        }
    }

}
```

**END**

```{r include=T, echo=F}
tac <- Sys.time()
cat(sprintf("%s %s@%s %s %f mins\n\n",Sys.time(),Sys.info()["login"],Sys.info()["nodename"],Script.Name,difftime(tac,tic,units="mins")))
```

---
title: DHI_GHI_sza_trends.R
author: athan
date: '2022-10-07'

---

<!--chapter:end:DHI_GHI_sza_trends.spin.Rmd-->

### 2022-10-09 Article.Rmd
Commit 2022-10-09 20:17

@@
     toc:              yes
+    toc_depth:        4
     fig_width:        7
@@
 <!-- physical quantities -->
-The conversions to irradiance was done using the calibration factor
+The conversions to irradiance was done using the available calibrations factors,
+for both instruments. For CHP1........
 .......chp1.... cmp21......
@@
 $35^\circ$ and $120^\circ$ and SZA greater than $80^\circ$, as at 
-those cases the Sun can not viewed by the instruments.
+those cases the Sun was not visible by the instruments.
 
@@
 <!-- Clear Sky -->
-We will use two dataset of our measurements, one for all sky conditions (ALL),
-which is the whole available data, and one filtered for 'clear sky' 
+We using two selection of our measurements dataset, one with all sky conditions (ALL),
+consisting of the whole available data, and one filtered for 'clear sky' 
 conditions (CS). The latter was created, using a method adapted to our 
-site as proposed by @Long2000 and by @Reno2016.
+site, and was proposed by @Long2000 and by @Reno2016.
 
@@
 
-Both dataset were aggregate in SZA bins by am, pm, am+pm
+Both dataset were aggregate in $1^\circ$ SZA bins by am, pm, am+pm
+
+some data was rejected.......
 
@@
 
+daily sum
+
 seasonal daily mean
 
-Deseasonlized
+<!-- Deseasonalization -->
+We remove the natural occurring seasonal variation, by calculating the mean
+values for the corresponding time step, and subtracting them .........
 
-## Longterm trends
+## Long-term trends
 

### 2022-10-09 DHI_GHI_2_sza_trends.R

@@ -27,6 +27,7 @@
 #'     keep_tex:         no
 #'     latex_engine:     xelatex
 #'     toc:              yes
+#'     toc_depth:        4
 #'     fig_width:        7
 #'     fig_height:       4.5
 #'   html_document:
@@ -543,11 +544,15 @@ abline(h=300)
 wecare <- grep( "^slope|^N",names(szatrends),ignore.case = T, value = T)
 
 
-#+ szatrends, echo=F, include=T
+#+ szatrends, echo=F, include=T, results = "asis"
 ## ALL - CS
 for (type in unique(szatrends$DATA)) {
     ## DIR - GLB - transp
     for (avar in unique(szatrends$var)) {
+
+        cat("\n\\newpage\n\n")
+        cat(paste("\n###", type, avar,"\n\n"))
+
         ## statistic variable
         for (awe in wecare) {
             awename <- gsub("(\\D)(\\D+)", "\\U\\1\\L\\2", sub("\\."," ", awe), perl = TRUE)
@@ -577,6 +582,8 @@ for (type in unique(szatrends$DATA)) {
                    # col    = c(unique(pam$col), unique(ppm$col)),
                    col    = c(2, 3),
                    pch    = c(unique(pam$pch), unique(ppm$pch)), ncol = 2, bty = "n")
+
+            cat("\n\n")
         }
 
 #
@@ -743,13 +750,16 @@ abline(h=300/4)
 wecare <- grep( "^slope|^N",names(szatrends_seas),ignore.case = T, value = T)
 
 
-#+ szatrendsseas, echo=F, include=T
+#+ szatrendsseas, echo=F, include=T, results = "asis"
 ## Winter - Summer ....
 for (ase in season) {
     ## ALL - Clear sky
     for (type in unique(szatrends_seas$DATA)) {
         ## DIR - GLB - transp
         for (avar in unique(szatrends_seas$var)) {
+
+            cat("\n\\newpage\n\n")
+            cat(paste("###",ase, type, avar,"\n\n"))
 
             ## statistic variable
             for (awe in wecare) {

### 2022-10-09 Article.Rmd
Commit 2022-10-09 22:17

@@
 
+# Intoduction.
+
+# Data and methodology.
+
+## Preperation.
+
+<!-- measurement -->
+For both CM21 and CHP1 instruments, the mean signal over one minute is recorded every one minute.
+<!-- screening -->
+A manual screening processes, removes inconsistencies and erroneous recordings, that can occur on a long contentious operation, like this one.
+<!-- physical quantities -->
+The conversion to irradiance was done using the available calibrations factors.
+For CHP1, there is only the initial one by the manufacture, and for CM21 there are seven calibration procedure in the span of the period we investigate.
+<!-- Dark signal -->
+We have adjusted the measurements for the dark signal offset, by a daily value, using the nights before and after each day.
+<!-- Quality assurance -->
+A radiation data quality assurance procedure was applied, 
+adjusted for the site, based on methods of Long and Shi\ [-@long_automated_2008; -@Long2006]. 
+Only data characterized as good quality were used in this stud.
+<!-- Data preparation -->
+Due to larger measurement uncertainty near the horizon, we exclude
+all measurements with SZA greater than $85^\circ$. 
+Moreover, due to some obstructions around the site (hills and buildings), we remove data with Azimuth angle between 
+$35^\circ$ and $120^\circ$ and SZA greater than $80^\circ$. At 
+those cases the Sun, were not visible by the instruments.
+
+# Results
+
+# Conclusions
+
+---------------------------
+
 ## Data description
@@
 ...... Measurements CHP1 CM21
-
-<!-- screening -->
-All measurements have been manually screened for inconsistencies and obviously 
-problematic data have been excluded.
-<!-- physical quantities -->
-The conversions to irradiance was done using the available calibrations factors,
-for both instruments. For CHP1........
-.......chp1.... cmp21......
-, and adjusting for the dark signal offset daily, using the nights 
-before and after each day.
-<!-- Quality assurance -->
-A radiation data quality assurance procedure was applied adjusted for the site,
-based on methods of Long and Shi\ [-@long_automated_2008; -@Long2006]. 
-Only data characterized with low quality indicators was rejected.
-<!-- Data preparation -->
-Due to the large measurement uncertainty near the horizon we exclude
-all measurements with SZA greater than $85^\circ$. Moreover, due to 
-nearby hills and buildings we remove data with azimuth angle between 
-$35^\circ$ and $120^\circ$ and SZA greater than $80^\circ$, as at 
-those cases the Sun was not visible by the instruments.
 

### 2022-10-09 Article.Rmd
Commit 2022-10-10 00:17

@@
 
+<!-- time span -->
+Time span for CM21 is 1993-2022 and 2016-2022 for CHP1.
+
+goal
+
+basic consepts
+
 # Data and methodology.
@@
 
+## Data process.
+
+To make the measurements comparable throughout the dataset. 
+<!-- 1au -->
+We adjusted all values to the mean Sun and Earth distance.
+<!-- TSI -->
+And, we made all measurements relative to the TSI at 1au,
+in order to compensate for the the Suns intensity variability,
+using an homogenized timeseries of satellite TSI observations.
+As a result, we processed all radiation data as a fraction of Suns TSI.
+
+<!-- Clear Sky -->
+For this analysis, we need .............
+, we using two selection of our measurements dataset, one with all sky conditions (ALL),
+consisting of the whole available data, and one filtered for 'clear sky' 
+conditions (CS). The latter was created, using a method adapted to our 
+site, and was proposed by @Long2000 and by @Reno2016.
+
 # Results
+
+## Long term trends
+
+## Zenith angle dependance and years seasons.
+
+## Consistency of the trends
 
@@
 ---------------------------
-
-## Data description
-
-<!-- time span -->
-Time span for CM21 is 1993-2022 and 2016-2022 for CHP1.
-
-...... Measurements CHP1 CM21
-
-## Data process
-
-We take some steps in order to make the measurements comparable throughout
-the dataset. 
-<!-- 1au -->
-We adjusted all measurements to the mean distance between Sun and Earth.
-<!-- TSI -->
-And, we make all measurements relative to the .........TSI.......... , 
-using an homogenized timeseries of satellite TSI data...........
-
-<!-- Clear Sky -->
-We using two selection of our measurements dataset, one with all sky conditions (ALL),
-consisting of the whole available data, and one filtered for 'clear sky' 
-conditions (CS). The latter was created, using a method adapted to our 
-site, and was proposed by @Long2000 and by @Reno2016.
 
@@
 
-## SZA trends
-
-## Trends consistency
-
 **END**

### 2022-10-10 DHI_GHI_0_data_input.R
Commit 2022-10-10 18:17

@@ -121,9 +121,21 @@ if ( !file.exists(common_data) | file.mtime(CS_file) > file.mtime(common_data) )
     DATA_all[  , CS_ref_HOR := NULL ]
     DATA_Clear[, CS_ref_HOR := NULL ]
 
+
+    ## 1. long-term
+
+
+
+
+
+
+
+    ## 2.
+
+
+
     ## save work space
     save(list = ls(all = TRUE),file = common_data)
-
 } else {
     cat(paste("\n\nLoad environment and data from: ", common_data,"\n\n"))
     load( file = common_data)

### 2022-10-10 DHI_GHI_1_longterm_trends.R

@@ -121,7 +121,7 @@ options(error = function() {
 #' #### Calculate daily means ####
 #+ echo=F, include=T
 
-ALL_daily_mean <- DATA_all[, .(DIR_att       = mean(DIR_att,    na.rm = T),
+ALL_1_daily_mean <- DATA_all[, .(DIR_att       = mean(DIR_att,    na.rm = T),
                                GLB_att       = mean(GLB_att,    na.rm = T),
                                HOR_att       = mean(HOR_att,    na.rm = T),
                                DIR_transp    = mean(DIR_transp, na.rm = T),
@@ -136,7 +136,7 @@ ALL_daily_mean <- DATA_all[, .(DIR_att       = mean(DIR_att,    na.rm = T),
                            by = .( Date = Day ) ]
 
 
-CLEAR_daily_mean <- DATA_Clear[, .(DIR_att       = mean(DIR_att,    na.rm = T),
+CLEAR_1_daily_mean <- DATA_Clear[, .(DIR_att       = mean(DIR_att,    na.rm = T),
                                    GLB_att       = mean(GLB_att,    na.rm = T),
                                    HOR_att       = mean(HOR_att,    na.rm = T),
                                    DIR_transp    = mean(DIR_transp, na.rm = T),
@@ -153,44 +153,39 @@ CLEAR_daily_mean <- DATA_Clear[, .(DIR_att       = mean(DIR_att,    na.rm = T),
 
 
 #' #### Margin of error calculation for 0.95 confidence interval ####
-#+ echo=T, include=T
+#+ echo=F, include=T
 conf_param  <- 1-(1-Daily_confidence_limit)/2
 suppressWarnings({
-    ALL_daily_mean[,  DIR_att_EM   := qt(conf_param,df=DIR_att_N -1) * DIR_att_sd    / sqrt(DIR_att_N)]
-    ALL_daily_mean[,  HOR_att_EM   := qt(conf_param,df=HOR_att_N -1) * HOR_att_sd    / sqrt(HOR_att_N)]
-    ALL_daily_mean[,  GLB_att_EM   := qt(conf_param,df=GLB_att_N -1) * GLB_att_sd    / sqrt(GLB_att_N)]
[…]

### 2022-10-10 DHI_GHI_0_variables.R
Commit 2022-10-10 18:17

@@ -12,6 +12,10 @@ col_HOR_att              <- "#4244ac"
 col_DIR_transp           <- "#9970ab"
 col_GLB_att              <- "#1a9850"
 
+## https://www.rapidtables.com/calc/time/days-in-year.html
+# Days_of_year             <- 365.25   ## Mean Julian year
+Days_of_year             <- 365.2425 ## Mean Gregorian calendar year
+
 pch_am                   <-  1
 pch_pm                   <-  4
 pch_ampm                 <- 13 ## try 10

### 2022-10-10 DHI_GHI_0_data_input.R
Commit 2022-10-10 19:17

@@ -2,6 +2,8 @@
 
 ## Data input
 
+# file.remove(common_data)
+
 if ( !file.exists(common_data) | file.mtime(CS_file) > file.mtime(common_data) ){
     cat(paste("Create environment and data in: ", common_data))
 
@@ -122,7 +124,90 @@ if ( !file.exists(common_data) | file.mtime(CS_file) > file.mtime(common_data) )
     DATA_Clear[, CS_ref_HOR := NULL ]
 
 
-    ## 1. long-term
+    ####  1. long-term   #######################################################
+
+    ## ~ daily means of all data ####
+    ALL_1_daily_mean <- DATA_all[,.(DIR_att       = mean(DIR_att,    na.rm = T),
+                                    GLB_att       = mean(GLB_att,    na.rm = T),
+                                    HOR_att       = mean(HOR_att,    na.rm = T),
+                                    DIR_transp    = mean(DIR_transp, na.rm = T),
+                                    DIR_att_sd    = sd(  DIR_att,    na.rm = T),
+                                    GLB_att_sd    = sd(  GLB_att,    na.rm = T),
+                                    HOR_att_sd    = sd(  HOR_att,    na.rm = T),
+                                    DIR_transp_sd = sd(  DIR_transp, na.rm = T),
+                                    doy           = yday(Date),
+                                    GLB_att_N     = sum(!is.na(GLB_att)),
+                                    HOR_att_N     = sum(!is.na(HOR_att)),
+                                    DIR_att_N     = sum(!is.na(DIR_att))  ),
+                                 by = .( Date = Day ) ]
+
+    ## ~ daily means of clear sky data ####
+    CLEAR_1_daily_mean <- DATA_Clear[,.(DIR_att       = mean(DIR_att,    na.rm = T),
+                                        GLB_att       = mean(GLB_att,    na.rm = T),
+                                        HOR_att       = mean(HOR_att,    na.rm = T),
+                                        DIR_transp    = mean(DIR_transp, na.rm = T),
+                                        DIR_att_sd    = sd(  DIR_att,    na.rm = T),
[…]

### 2022-10-10 DHI_GHI_0_variables.R

@@ -26,7 +26,7 @@ SZA_BIN                  <-  1
 MIN_N                    <-  4
 SEAS_MIN_N               <-  3
 
-Daily_aggregation_N_lim  <- 50
+Daily_aggregation_N_lim  <- 50         ## was SUM_THRES
 Daily_confidence_limit   <-  0.99
 SZA_THRES                <-  3
 

### 2022-10-10 Article.Rmd
Commit 2022-10-10 18:17
Commit 2022-10-10 19:17

@@
 
+```{r echo=F, include=T}
+## variables we use
+source("~/MANUSCRIPTS/2022_sdr_trends/DHI_GHI_0_variables.R")
+```
+
 # Intoduction.
@@
 <!-- measurement -->
-For both CM21 and CHP1 instruments, the mean signal over one minute is recorded every one minute.
+For both CM21 and CHP1 instruments, the mean signal and the standard deviation over one minute is recorded every one minute.
 <!-- screening -->
@@
 <!-- Quality assurance -->
-A radiation data quality assurance procedure was applied, 
-adjusted for the site, based on methods of Long and Shi\ [-@long_automated_2008; -@Long2006]. 
+A radiation data quality assurance procedure was applied, adjusted for the site, based on methods of Long and Shi\ [-@long_automated_2008; -@Long2006].
 Only data characterized as good quality were used in this stud.
@@
 <!-- TSI -->
-And, we made all measurements relative to the TSI at 1au,
-in order to compensate for the the Suns intensity variability,
-using an homogenized timeseries of satellite TSI observations.
+<!-- And, we made all measurements relative to the TSI at 1au, in order to compensate for the the Suns intensity variability, using an homogenized timeseries of satellite TSI observations.
 As a result, we processed all radiation data as a fraction of Suns TSI.
+-->
 
 <!-- Clear Sky -->
-For this analysis, we need .............
+In this analysis, we need .............
 , we using two selection of our measurements dataset, one with all sky conditions (ALL),

### 2022-10-10 DHI_GHI_2_sza_trends.R

@@ -70,242 +70,54 @@ if(!interactive()) {
 par(pch = ".")
 
 
-## FIXME this is for pdf output
-# options(warn=-1) ## hide warnings
-# options(warn=2)  ## stop on warnings
 
 #+ echo=F, include=T
 ####  External code  ####
 library(data.table, quietly = T, warn.conflicts = F)
 library(pander,     quietly = T, warn.conflicts = F)
 
+panderOptions('table.alignment.default', 'right')
+panderOptions('table.split.table',        120   )
+
 ## Functions from `https://github.com/thanasisn/IStillBreakStuff/tree/main/FUNCTIONS/R`
 source("~/CODE/FUNCTIONS/R/sumNA.R")
 source("~/CODE/FUNCTIONS/R/linear_regrassion_capture.R")
 source("~/CODE/FUNCTIONS/R/trig_deg.R")
+source("~/CODE/FUNCTIONS/R/data.R")
+
+
+## For this project
+source("~/MANUSCRIPTS/2022_sdr_trends/DHI_GHI_0_variables.R")
+source("~/MANUSCRIPTS/2022_sdr_trends/DHI_GHI_0_data_input.R")
 
 
 options(error = function() {
     if (interactive()) {
-        # require("beepr"); beep(10)
         system("mplayer /usr/share/sounds/freedesktop/stereo/dialog-warning.oga", ignore.stdout = T, ignore.stderr = T)
         system("notify-send -u normal -t 30000 'R session' 'An error occured!'")
     }
 })
 
 
-####  Variables  ####
-panderOptions('table.alignment.default', 'right')
-panderOptions('table.split.table',        120   )
-
-DATA_DIR  <- "/home/athan/DATA/Broad_Band"
-CLEARdir  <- "/home/athan/DATA/Broad_Band/CS_id"
-tag       <- paste0("Natsis Athanasios LAP AUTH ", strftime(Sys.time(), format = "%b %Y" ))
-CS_file   <- "/home/athan/DATA/Common_application/Clear_Sky.Rds"
-
-TEST      <- TRUE
-
-col_DIR_att    <- "#2166ac"
-col_DIR_transp <- "#9970ab"
-col_GLB_att    <- "#1a9850"
-pch_am         <-  1
-pch_pm         <-  4
-pch_ampm       <- 13 ## try 10
-pch_daily      <- 19
-
[…]

### 2022-10-13 Article.Rmd
Commit 2022-10-13 13:17

@@
 <!-- 1au -->
-We adjusted all values to the mean Sun and Earth distance.
+We adjusted all 1-minute values to the mean Sun and Earth distance.
 <!-- TSI -->
@@
 -->
+
+<!-- data selection -->
+We select all data with $\text{SZA} < `r 90-MIN_ELEVA`^\circ$, in order to avoid measurements near the horizon. Where the signal is very low and the disturbances more frequent.
+<!-- avoid obstacles -->
+Similar, we exclude values data with Azimuth angle between 
+$35^\circ$ and $120^\circ$ and SZA greater than $80^\circ$. Where, due to some obstructions around the site (hills and buildings) the Sun, were not visible by the instruments. . At 
+those cases.
 

### 2022-10-13 DHI_GHI_1_longterm_trends.R
Commit 2022-10-13 14:17

@@ -260,7 +260,8 @@ for (DBn in dbs) {
             plot(dataset$Date, dataset[[avar]],
                  pch  = ".", col = get(paste0("col_",avar)),
                  xlab = "",
-                 ylab = bquote("Deseas." ~ .(translate(avar)) ~ "[" ~ Watt/m^2 ~ "]" ) )
+                 ylab = bquote("Seasonal Anomaly [%]" ) )
+            # ylab = bquote("Deseas." ~ .(translate(avar)) ~ "[" ~ Watt/m^2 ~ "]" ) )
             abline(lm1)
 
             ## decorations
@@ -292,7 +293,8 @@ for (DBn in dbs) {
         plot(dataset$Date, dataset[[avar]],
              pch  = ".", col = get(paste0("col_",avar)),
              xlab = "",
-             ylab = bquote("Deseas." ~ .(translate(avar)) ~ "[" ~ Watt/m^2 ~ "]" ) )
+             ylab = bquote("Seasonal Anomaly [%]" ) )
+             # ylab = bquote("Deseas." ~ .(translate(avar)) ~ "[" ~ Watt/m^2 ~ "]" ) )
         abline(lm1)
 
         ## decorations
@@ -398,7 +400,8 @@ for (DBn in dbs) {
             plot(dataset$Date, dataset[[avar]],
                  pch  = ".", col = get(paste0("col_",avar)),
                  xlab = "",
-                 ylab = bquote("Deseas." ~ .(translate(avar)) ~ "[" ~ Watt/m^2 ~ "]" ) )
+                 ylab = bquote("Seasonal Anomaly [%]" ) )
+                 # ylab = bquote("Deseas." ~ .(translate(avar)) ~ "[" ~ Watt/m^2 ~ "]" ) )
             abline(lm1)
             ## decorations
             fit <- lm1[[1]]

### 2022-10-13 Article.Rmd
Commit 2022-10-13 14:17

--- 2022-10-10 Article.Rmd
@@
     fig_caption:      yes
-    keep_tex:         no
+    keep_tex:         yes
     latex_engine:     xelatex
@@
 A radiation data quality assurance procedure was applied, adjusted for the site, based on methods of Long and Shi\ [-@long_automated_2008; -@Long2006].
-Only data characterized as good quality were used in this stud.
-<!-- Data preparation -->
+Only data characterized as good quality were used in this study.
+<!-- Data selection -->
 Due to larger measurement uncertainty near the horizon, we exclude
-all measurements with SZA greater than $85^\circ$. 
+all measurements with SZA greater than $`r 90-MIN_ELEVA`^\circ$. 
 Moreover, due to some obstructions around the site (hills and buildings), we remove data with Azimuth angle between 

### 2023-02-13 Article.Rmd
Commit 2023-02-13 20:17

@@
 ---
-title:         "Trends of SDR in Thessaloniki "
+title:         "Trends of downward SDR in Thessaloniki "
 author:
-  - Natsis Athanasios^[Laboratory of Atmospheric Physics,AUTH, natsisthanasis@gmail.com]
-  - Jane Doe^[Institution Two, jane@example.org]
+  - Natsis Athanasios^[Laboratory of Atmospheric Physics, AUTH, natsisthanasis@gmail.com]
+  - Alkiviadis Bais^[Laboratory of Atmospheric Physics, AUTH]
 abstract:
-  "Study of GHI and DNI radiation for 'clear sky' and all sly conditions."
+  "Study of GHI and DNI radiation for 'clear-sky' and all-sky conditions."
 
@@
 
-basic consepts
+basic conceptions
 

### 2023-02-13 DHI_GHI_2_sza_trends.R
Commit 2023-02-13 20:17

@@ -200,9 +200,9 @@ ALL_daily_DEseas   <- merge(  ALL_2_daily_mean, ALL_2_daily_seas,   by = c("doy"
 CLEAR_daily_DEseas <- merge(CLEAR_2_daily_mean, CLEAR_2_daily_seas, by = c("doy", "SZA", "preNoon"), all = T)
 CLOUD_daily_DEseas <- merge(CLOUD_2_daily_mean, CLOUD_2_daily_seas, by = c("doy", "SZA", "preNoon"), all = T)
 
-setorder(ALL_daily_DEseas,Date)
-setorder(CLEAR_daily_DEseas,Date)
-setorder(CLOUD_daily_DEseas,Date)
+setorder(ALL_daily_DEseas,   Date)
+setorder(CLEAR_daily_DEseas, Date)
+setorder(CLOUD_daily_DEseas, Date)
 
 
 ## anomaly
@@ -329,6 +329,7 @@ abline(h=300)
 ## stats vars to plot
 wecare <- grep("^slope|^N", names(szatrends), ignore.case = T, value = T)
 wecare <- grep("^slope\\.t", wecare, ignore.case = T, value = T, invert = T)
+wecare <- grep("slope\\.sd", wecare, ignore.case = T, value = T, invert = T)
 
 
 #+ szatrends, echo=F, include=T, results = "asis"
@@ -420,7 +421,7 @@ for (DBn in dbs) {
 
                     if (sum(!is.na(dataset[[avar]])) <= 1) next()
 
-                    lm1 <- lm( dataset[[avar]] ~ dataset$Date )
+                    lm1 <- lm(dataset[[avar]] ~ dataset$Date)
 
                     gather_seas <- rbind(gather_seas,
                                     data.frame(
@@ -491,8 +492,9 @@ abline(h=300/4)
 
 
 ## stats vars to plot
-wecare <- grep( "^slope|^N",names(szatrends_seas),ignore.case = T, value = T)
-wecare <- grep("^slope\\.t",wecare,ignore.case = T,value = T, invert = T)
+wecare <- grep( "^slope|^N", names(szatrends_seas), ignore.case = T, value = T)
+wecare <- grep("^slope\\.t", wecare, ignore.case = T, value = T, invert = T)
+wecare <- grep("slope\\.sd", wecare, ignore.case = T, value = T, invert = T)
 
 
 #+ szatrendsseas, echo=F, include=T, results = "asis"

### 2023-02-13 DHI_GHI_3_trends_consistency.R

@@ -21,20 +21,20 @@
 #' - \captionsetup{font=small}
 #'
 #' output:
-#'   html_document:
-#'     toc:        true
-#'     keep_md:    yes
-#'     fig_width:  7.5
-#'     fig_height: 5
 #'   bookdown::pdf_document2:
-#'     number_sections:  no
-#'     fig_caption:      no
-#'     keep_tex:         yes
-#'     latex_engine:     xelatex
-#'     toc:              yes
-#'     toc_depth:        4
-#'     fig_width:        7
-#'     fig_height:       4.5
+#'     number_sections: no
+#'     fig_caption:     no
+#'     keep_tex:        no
+#'     latex_engine:    xelatex
+#'     toc:             yes
+#'     toc_depth:       4
+#'     fig_width:       7
+#'     fig_height:      4.5
+#'   html_document:
+#'     toc:             true
+#'     keep_md:         no
+#'     fig_width:       7.5
+#'     fig_height:      5
 #'
 #' date: "`r format(Sys.time(), '%F')`"
 #'
@@ -117,8 +117,9 @@ options(error = function() {
 ## ~ Plots longterm  ####
 
 data_list  <- list(ALL   =   ALL_3_monthly_mean,
-                   CLEAR = CLEAR_3_monthly_mean)
-by_var     <- c("Date","Month","SZA")
+                   CLEAR = CLEAR_3_monthly_mean,
+                   CLOUD = CLOUD_3_monthly_mean)
+by_var     <- c("Date", "Month", "SZA")
 wecare     <- unique(unlist(lapply(data_list, names)))
 wecare     <- grep("HOR|GLB|DIR", wecare, value = T)
 for (i in 1:length(data_list)) {
@@ -153,8 +154,9 @@ for (i in 1:length(data_list)) {
 ## ~ Plots seasonal data ####
 
 data_list  <- list(ALL   =   ALL_3_monthly_seas,
-                   CLEAR = CLEAR_3_monthly_seas)
-by_var     <- c("Month","SZA")
+                   CLEAR = CLEAR_3_monthly_seas,
[…]

### 2023-03-05 Article.Rmd
Commit 2023-03-05 09:17

@@
 author:
-  - Natsis Athanasios^[Laboratory of Atmospheric Physics, AUTH, natsisthanasis@gmail.com]
+  - Natsis Athanasios^[Laboratory of Atmospheric Physics, AUTH, natsisphysicist@gmail.com]
   - Alkiviadis Bais^[Laboratory of Atmospheric Physics, AUTH]
@@
 ```{r trackercomm, echo=F, fig.cap="tracker"}
-knitr::include_graphics("./images/longtermtrendsALL-2.pdf")
+knitr::include_graphics("./images/LongtermTrends-1.pdf")
 ```

### 2023-04-06 DHI_GHI_1_longterm_trends.R
Commit 2023-04-06 14:17

@@ -102,6 +102,10 @@ options(error = function() {
 
 ## __ Flags --------------------------------------------------------------------
 
+DRAFT <- TRUE
+DRAFT <- FALSE
+
+
 ## override plot options
 par(pch = ".")
 
@@ -251,28 +255,29 @@ for (DBn in dbs) {
             ## plot fit line
             abline(lm1, lwd = 2)
 
-            ## Running mean
-            first <- head(which(!is.na(dataset[[avar]])),1)
-            last  <- tail(which(!is.na(dataset[[avar]])),1)
+            if (DRAFT) {
+                ## Running mean
+                first <- head(which(!is.na(dataset[[avar]])),1)
+                last  <- tail(which(!is.na(dataset[[avar]])),1)
 
-            rm <- frollmean(dataset[[avar]][first:last],
-                            round(running_mean_window_days),
-                            na.rm = TRUE,
-                            algo  = "exact",
-                            align = "center")
+                rm <- frollmean(dataset[[avar]][first:last],
+                                round(running_mean_window_days),
+                                na.rm = TRUE,
+                                algo  = "exact",
+                                align = "center")
 
-            # points(dataset$Date, rm, col = "red", cex = 0.5)
-            lines(dataset$Date[first:last], rm, col = "red", lwd = 1.5)
+                # points(dataset$Date, rm, col = "red", cex = 0.5)
+                lines(dataset$Date[first:last], rm, col = "red", lwd = 1.5)
 
 
-            ## LOESS curve
-            vec <- !is.na(dataset[[avar]])
-            FTSE.lo3 <- loess.as(dataset$Date[vec], dataset[[avar]][vec],
-                                 degree = 1,
-                                 criterion = LOESS_CRITERIO, user.span = NULL, plot = F)
-            FTSE.lo.predict3 <- predict(FTSE.lo3, dataset$Date)
-            lines(dataset$Date, FTSE.lo.predict3, col = "cyan", lwd = 2.5)
-
+                ## LOESS curve
+                vec <- !is.na(dataset[[avar]])
+                FTSE.lo3 <- loess.as(dataset$Date[vec], dataset[[avar]][vec],
+                                     degree = 1,
+                                     criterion = LOESS_CRITERIO, user.span = NULL, plot = F)
+                FTSE.lo.predict3 <- predict(FTSE.lo3, dataset$Date)
+                lines(dataset$Date, FTSE.lo.predict3, col = "cyan", lwd = 2.5)
+            }
 
 
[…]

### 2023-04-24 Article.Rmd
Commit 2023-04-06 14:17
Commit 2023-04-24 18:17

@@
 
-# Intoduction.
+# Introduction.
 
 <!-- time span -->
-Time span for CM21 is 1993-2022 and 2016-2022 for CHP1.
+Time span for CM-21 is 1993-2022 and 2016-2022 for CHP-1.
 
@@
 
-## Preperation.
+## Preparation.
+
+We can separate the data preparation in three parts, ..........
 
 <!-- measurement -->
-For both CM21 and CHP1 instruments, the mean signal and the standard deviation over one minute is recorded every one minute.
+For both CM-21 and CHP-1 instruments, the mean signal and the standard deviation over one minute is recorded every one minute.
 <!-- screening -->
@@
 The conversion to irradiance was done using the available calibrations factors.
-For CHP1, there is only the initial one by the manufacture, and for CM21 there are seven calibration procedure in the span of the period we investigate.
+For CHP-1, there is only the initial one by the manufacture, and for CM-21 there are seven calibration procedure in the span of the period we investigate.
 <!-- Dark signal -->
@@
 ```{r trackercomm, echo=F, fig.cap="tracker"}
-knitr::include_graphics("./images/LongtermTrends-1.pdf")
+knitr::include_graphics("./images/LongtermTrends-2.pdf")
 ```

### 2023-04-24 Article.Rmd
Commit 2023-04-24 21:17

@@
 
-We can separate the data preparation in three parts, ..........
+We can separate the data preparation in three parts, a) the acquisition of radiation measurements from the sensor, b) a radiation data quality check, and c) the identification of "clear sky" conditions from the radiometric data.
+
+### Acquisition of radiometric data.
 
 <!-- measurement -->
-For both CM-21 and CHP-1 instruments, the mean signal and the standard deviation over one minute is recorded every one minute.
+We sample the signal of the CM-21 pyranometer with a rate of $1 {Hz}$, and record the average signal and the standard deviation over each minute of the hour.
+The measurements are corrected for the zero offset of the instrument, using the "dark signal" (Sun elevation bellow $10\circ$, for a period of
+
 <!-- screening -->

### 2023-04-24 Article.Rmd
Commit 2023-04-24 22:17

@@
 We sample the signal of the CM-21 pyranometer with a rate of $1 {Hz}$, and record the average signal and the standard deviation over each minute of the hour.
-The measurements are corrected for the zero offset of the instrument, using the "dark signal" (Sun elevation bellow $10\circ$, for a period of
+<!-- dark correction -->
+The measurements are corrected for the zero offset of the instrument, using the "dark signal" (Sun elevation bellow $10^\circ$, for a period of $3 h$).
+<!-- physical quantities -->
+The signal is converted to radiation using a ramped value of the instrument sensitivity, derived from the eight laboratory calibrations of the instrument, during the study period.
+
+### Radiation data quality check
 
 <!-- screening -->
-A manual screening processes, removes inconsistencies and erroneous recordings, that can occur on a long contentious operation, like this one.
-<!-- physical quantities -->
-The conversion to irradiance was done using the available calibrations factors.
-For CHP-1, there is only the initial one by the manufacture, and for CM-21 there are seven calibration procedure in the span of the period we investigate.
-<!-- Dark signal -->
-We have adjusted the measurements for the dark signal offset, by a daily value, using the nights before and after each day.
+A manual screening is performed, to remove inconsistencies and erroneous recordings, that can occur during a long continuous operation.
 <!-- Quality assurance -->
-A radiation data quality assurance procedure was applied, adjusted for the site, based on methods of Long and Shi\ [-@long_automated_2008; -@Long2006].
+The manual screening is aided by a radiation data quality assurance procedure, adjusted for this site, based on methods of Long and Shi\ [-@long_automated_2008; -@Long2006].
+
 Only data characterized as good quality were used in this study.
+.....
+
+### Clear-sky identification
+
+......
+
 <!-- Data selection -->

### 2023-04-25 Article.Rmd
Commit 2023-04-25 10:17

@@
 
-......
-
-<!-- Data selection -->
-Due to larger measurement uncertainty near the horizon, we exclude
-all measurements with SZA greater than $`r 90-MIN_ELEVA`^\circ$. 
-Moreover, due to some obstructions around the site (hills and buildings), we remove data with Azimuth angle between 
-$35^\circ$ and $120^\circ$ and SZA greater than $80^\circ$. At 
-those cases the Sun, were not visible by the instruments.
+In order to estimate the effect of the clouds on the global radiation we create three datasets. 
+The all sky conditions containing all the valid measurements. 
+The clear-sky conditions data and the remainder part the cloudy condition data. 
+To identify the clear-sky conditions we used a method proposed by @Long2000 and by @Reno2016 that was adapted and configured for the site.
 
@@
 
+### Data selection
+
+Due to a significant measurement uncertainty near the horizon, we exclude
+all measurements with SZA greater than $`r 90-MIN_ELEVA`^\circ$. 
+Moreover, due to some obstructions around the site (hills and buildings), we exclude data with Azimuth angle between 
+$35^\circ$ and $120^\circ$ and SZA greater than $80^\circ$.
+At those cases the Sun systematically, is not visible by the instruments.
+
 To make the measurements comparable throughout the dataset. 
 <!-- 1au -->
-We adjusted all 1-minute values to the mean Sun and Earth distance.
+We adjusted all 1-minute radiometric values to the mean Sun - Earth distance.
 <!-- TSI -->
@@
 
-<!-- Clear Sky -->
-In this analysis, we need .............
-, we using two selection of our measurements dataset, one with all sky conditions (ALL),
-consisting of the whole available data, and one filtered for 'clear sky' 
-conditions (CS). The latter was created, using a method adapted to our 
-site, and was proposed by @Long2000 and by @Reno2016.
+### Process and aggregation of radiometric data
 

### 2023-04-25 Article.Rmd
Commit 2023-04-25 11:17

@@
 
+In this study, we present any radiometric data as an percent departure from the corresponding seasonal mean.
+When we use daily aggregated data we compute the mean values for the day and exclude instances with less than `r  Daily_aggregation_N_lim` valid data points.
+Accordingly monthly values are computed from the daily aggregated data, `r Monthly_aggegation_N_lim`
+
 # Results

### 2023-04-25 Article.Rmd
Commit 2023-04-25 12:17
Commit 2023-04-25 13:17

@@
 When we use daily aggregated data we compute the mean values for the day and exclude instances with less than `r  Daily_aggregation_N_lim` valid data points.
-Accordingly monthly values are computed from the daily aggregated data, `r Monthly_aggegation_N_lim`
+Accordingly, monthly values are computed from the daily aggregated data, month with less than `r Monthly_aggegation_N_lim` days are rejected. 
+The daily seasonal values are computed based on the day of year number and the monthly by the calendar month.
+
+<!-- Seasonal -->
+..... Define seasons aggregation if used .......
 

### 2023-04-25 Article.Rmd
Commit 2023-04-25 15:17

@@
 
+....
 Only data characterized as good quality were used in this study.
@@
 
-## Zenith angle dependance and years seasons.
+Using the mean daily Global radiation we produce the trends of each data set, as departure from the seasonal value.
+
+```{r trendALL, echo=F, fig.cap="Daily Global radiation seasonal anomaly and long term trend, for all sky conditions."}
+knitr::include_graphics("./images/LongtermTrends-2.pdf")
+```
+
+```{r trendCLEAR, echo=F, fig.cap="Daily Global radiation seasonal anomaly and long term trend, for clear sky conditions."}
+knitr::include_graphics("./images/LongtermTrends-4.pdf")
+```
+
+```{r trendCLOUD, echo=F, fig.cap="Daily Global radiation seasonal anomaly and long term trend, for cloud sky conditions."}
+knitr::include_graphics("./images/LongtermTrends-6.pdf")
+```
+
+## Zenith angle dependence and years seasons.
 
@@
 Both dataset were aggregate in $1^\circ$ SZA bins by am, pm, am+pm
-
-some data was rejected.......
 
@@
 
-```{r trackercomm, echo=F, fig.cap="tracker"}
-knitr::include_graphics("./images/LongtermTrends-2.pdf")
-```
-
-This is it \@ref(fig:trackercomm).
+This is it \@ref(fig:trendALL).
 
@@
 
-<!-- \tiny -->
+\tiny 
 ```{r echo=F}
@@
 ```
-<!-- \normalsize -->
+\normalsize
+
+Some more text here
 

### 2023-04-27 Article.Rmd
Commit 2023-04-27 20:17

@@
 ---
-title:         "Trends of downward SDR in Thessaloniki "
+title:         "Trends of downward SDR in Thessaloniki"
 author:
@@
 knitr::opts_chunk$set(dev        = "pdf"   )
-# knitr::opts_chunk$set(dev        = "png"    )
-knitr::opts_chunk$set(out.width  = "80%"   )
-knitr::opts_chunk$set(fig.align  = "center" )
+knitr::opts_chunk$set(out.width  = "70%"   )
+knitr::opts_chunk$set(fig.align  = "center")
 knitr::opts_chunk$set(fig.pos    = '!h'    )
@@
 <!-- TSI -->
-<!-- And, we made all measurements relative to the TSI at 1au, in order to compensate for the the Suns intensity variability, using an homogenized timeseries of satellite TSI observations.
+<!-- And, we made all measurements relative to the TSI at 1au, in order to compensate for the Suns intensity variability, using a homogenized timeseries of satellite TSI observations.
 As a result, we processed all radiation data as a fraction of Suns TSI.
@@
 
-In this study, we present any radiometric data as an percent departure from the corresponding seasonal mean.
+In this study, we present any radiometric data as a percent departure from the corresponding seasonal mean.
 When we use daily aggregated data we compute the mean values for the day and exclude instances with less than `r Daily_aggregation_N_lim` valid data points.
@@
 ## Long term trends
+
+\FloatBarrier
 
@@
 
+## Long term trends by season of year
+
+\FloatBarrier
+
+```{r seasonalALL, echo=F, fig.cap="Donditions."}
+knitr::include_graphics("./images/SeasonaltTrends-1.pdf")
+```
+
+```{r seasonalCLEAR, echo=F, fig.cap="Daiditions."}
+knitr::include_graphics("./images/SeasonaltTrends-2.pdf")
+```
+
+```{r seasonalCLOUD, echo=F, fig.cap="Daiditions."}
+knitr::include_graphics("./images/SeasonaltTrends-3.pdf")
+```
+
 ## Zenith angle dependence and years seasons.
 
+\FloatBarrier
+
 ## Consistency of the trends
+
+\FloatBarrier
 
@@
 
-\tiny 
+<!--\tiny -->
 ```{r echo=F}
@@
 ```
-\normalsize
+<!--\normalsize-->
 

### 2023-04-27 Article.Rmd
Commit 2023-04-27 21:17

@@
 
-```{r echo=F, include=F, cache=T}
+```{r echo=F, include=T}
 knitr::opts_chunk$set(comment    = ""      )
@@
 knitr::opts_chunk$set(fig.pos    = '!h'    )
-
+library(pander)
 ```
@@
 
+\footnotesize
+```{r trendtable, echo=F}
+pp <- read.csv("./figures/tbl_longterm_trends.dat",
+               strip.white = TRUE,
+               sep = ";",
+               header = TRUE)
+
+pp$slope.sd           <- NULL
+pp$slope.ConfInt_0.95 <- NULL
+pp$slope.ConfInt_0.99 <- NULL
+pp <- pp[grep("Global" ,pp$var), ]
+
+pander(pp)
+```
+\normalsize
+
 ## Long term trends by season of year

### 2023-04-28 Article.Rmd
Commit 2023-04-28 12:17

@@
 
-```{r trendCLEAR, echo=F, fig.cap="Daily Global radiation seasonal anomaly and long term trend, for clear sky conditions."}
-knitr::include_graphics("./images/LongtermTrends-4.pdf")
-```
+<!-- ```{r trendCLEAR, echo=F, fig.cap="Daily Global radiation seasonal anomaly and long term trend, for clear sky conditions."} -->
+<!-- knitr::include_graphics("./images/LongtermTrends-4.pdf") -->
+<!-- ``` -->
 
@@
                header = TRUE)
-
+pp <- pp[grep("Global" ,pp$var), ]
 pp$slope.sd           <- NULL
@@
 pp$slope.ConfInt_0.99 <- NULL
-pp <- pp[grep("Global" ,pp$var), ]
+pp$var                <- NULL
+rownames(pp) <- NULL
+names(pp) <- c("Trend", "Trend p-value", "Sky condition", "Trend statistical significance")
 
-pander(pp)
+panderOptions("table.alignment.default", "left")
+pander(pp,
+       caption = "Trends of daily means by sky conditions.")
 ```
@@
 ```
+
+\footnotesize
+```{r trendseasontable, echo=F}
+pp <- read.csv("./figures/tbl_longterm_trends_season.dat",
+               strip.white = TRUE,
+               sep = ";",
+               header = TRUE)
+pp <- pp[grep("Global" ,pp$var), ]
+pp$slope.sd           <- NULL
+pp$slope.ConfInt_0.95 <- NULL
+pp$slope.ConfInt_0.99 <- NULL
+pp$var                <- NULL
+rownames(pp) <- NULL
+names(pp) <- c("Trend", "Trend p-value", "Sky condition", "Season", "Trend statistical sign.")
+
+panderOptions("table.alignment.default", "left")
+pander(pp,
+       caption = "Trends of daily means by sky conditions.")
+```
+\normalsize
 

### 2023-05-02 DHI_GHI_0_data_input.R
Commit 2023-05-02 13:17

@@ -20,6 +20,10 @@ D_14_2 <- TRUE
 # D_14   <- TRUE
 # D_13   <- TRUE
 
+TEST <- TRUE
+
+
+
 ## new new implementation with corrected limits
 if (D_14_2) {
     common_data <- common_data_14_2
@@ -58,6 +62,12 @@ if (havetorun) {
                                pattern    = inpatern,
                                full.names = T )
     input_files <- grep("_stats_", input_files, value = TRUE, invert = TRUE)
+
+    if (TEST) {
+        input_files <- sample(input_files, 4)
+    }
+
+
 
     if ( !file.exists(CS_file) | max(file.mtime(input_files)) > file.mtime(CS_file)) {
         cat(paste("Load data from Clear Sky proccess from original\n"))
@@ -213,6 +223,7 @@ if (havetorun) {
     ## old flags usage
     # DATA_Clear <- DATA_all[ CSflag == 0 ]
     rm(DATA)
+    gc()
 
     ## remove unused columns
     rm.cols.DT(DATA_all,   "CSflag_*")
@@ -223,10 +234,10 @@ if (havetorun) {
     DATA_Cloud[, CS_ref_HOR := NULL]
 
 
-    ##.----
+    ## . ----
     #### 1. long-term  ---------------------------------------------------------
 
-    ## _ Daily means of data ---------------------------------------------------
+    ##  Daily means of data ----------------------------------------------------
 
     ALL_1_daily_mean <-
         DATA_all[,.(DIR_att       = mean(DIR_att,     na.rm = T),
@@ -407,15 +418,12 @@ if (havetorun) {
     setorder(CLEAR_1_daily_DESEAS, Date)
     setorder(CLOUD_1_daily_DESEAS, Date)
 
+    ## Using the % departure from seasonal values
 
-    ## _ Daily relative anomaly ------------------------------------------------
-
-    ### Using the % departure from seasonal values
-
-    ALL_1_daily_DESEAS[  , DIR_att_des   := 100*( DIR_att    - DIR_att_seas    ) / DIR_att_seas   ]
-    ALL_1_daily_DESEAS[  , HOR_att_des   := 100*( HOR_att    - HOR_att_seas    ) / HOR_att_seas   ]
-    ALL_1_daily_DESEAS[  , GLB_att_des   := 100*( GLB_att    - GLB_att_seas    ) / GLB_att_seas   ]
[…]

### 2023-05-02 DHI_GHI_1_longterm_trends.R

@@ -354,7 +354,7 @@ myRtools::write_dat(pprint,
 #'
 #' ### Trends for each season of the year
 #'
-#' We calculated monthly means from the daily means
+#' We calculated Seasonla means from the daily means
 #'
 #+ echo=F, include=F
 

### 2023-05-05 Article.Rmd
Commit 2023-04-28 13:17
Commit 2023-05-02 13:17
Commit 2023-05-05 10:17

--- 2023-04-27 Article.Rmd
@@
 
-```{r echo=F, include=F, cache=T}
+```{r echo=F, include=T}
+## document configurations 
 knitr::opts_chunk$set(comment    = ""      )
@@
 knitr::opts_chunk$set(fig.pos    = '!h'    )
-
-```
-
-```{r echo=F, include=T}
+library(pander)
+library(data.table)
+## load data
+load("./data/common_data_14_2.Rda")
 ## variables we use
@@
 <!-- time span -->
-Time span for CM-21 is 1993-2022 and 2016-2022 for CHP-1.
+The global radiation was obtained from the pyranometer CM-21, for the period
+`r min(ALL_1_daily_DESEAS$Date)` -  `r max(ALL_1_daily_DESEAS$Date)`.
+In some algorithms we used direct radiation data for comparisons, obtained by a CM-21 pyrheliometer, with data span
+`r min(ALL_1_daily_DESEAS[!is.na(DIR_att), Date])` - `r max(ALL_1_daily_DESEAS[!is.na(DIR_att), Date])`.
 
-goal
+<!-- goals -->
+Here we investigate the trend on shortwave radiation for the last decades at Thessaloniki, and try to infer a relation to sky conditions (all, clear, cloudy).
 
-basic conceptions
+<!-- basic concepts -->
+The approach of our method to characterize the sky conditions and the definition of each sky condition has some ......μη αντικειμένικότητα
 
@@
 <!-- Seasonal -->
-..... Define seasons aggregation if used .......
+For the seasons of the year trends we grouped the mean daily values by season (December - February: Winter, March - May: Spring, etc.)
 
@@
 
+\FloatBarrier
+
 ## Long term trends
-
-\FloatBarrier
 
@@
 
-## Long term trends by season of year
+\footnotesize
+```{r trendtable, echo=F}
+pp <- read.csv("./figures/tbl_longterm_trends.dat",
+               strip.white = TRUE,
+               sep = ";",
+               header = TRUE)
+pp <- pp[grep("Global" ,pp$var), ]
+pp$slope.sd           <- NULL
+pp$slope.ConfInt_0.95 <- NULL
+pp$slope.ConfInt_0.99 <- NULL
+pp$var                <- NULL
+rownames(pp) <- NULL
+names(pp) <- c("Trend %/year", "Trend p-value", "Sky condition", "Trend statistical significance")
+
+panderOptions("table.alignment.default", "left")
+pander(pp,
+       caption = "Trends of daily means by sky conditions.")
+```
+\normalsize
 
@@
 
-```{r seasonalALL, echo=F, fig.cap="Donditions."}
+## Long term trends by season of year
+
+```{r seasonalALL, echo=F, fig.cap="Trends by season for All-sky condition. Displays monthly means of daily means."}
 knitr::include_graphics("./images/SeasonaltTrends-1.pdf")
@@
 
-```{r seasonalCLEAR, echo=F, fig.cap="Daiditions."}
+```{r seasonalCLEAR, echo=F, fig.cap="Trends by season for Clear-sky condition. Displays monthly means of daily means."}
 knitr::include_graphics("./images/SeasonaltTrends-2.pdf")
@@
 
-```{r seasonalCLOUD, echo=F, fig.cap="Daiditions."}
+```{r seasonalCLOUD, echo=F, fig.cap="Trends by season for Cloudy-sky condition. Displays monthly means of daily means."}
 knitr::include_graphics("./images/SeasonaltTrends-3.pdf")
 ```
+
+\footnotesize
+```{r trendseasontable, echo=F}
+pp <- read.csv("./figures/tbl_longterm_trends_season.dat",
+               strip.white = TRUE,
+               sep = ";",
+               header = TRUE)
+pp <- pp[grep("Global" ,pp$var), ]
+pp$slope.sd           <- NULL
+pp$slope.ConfInt_0.95 <- NULL
+pp$slope.ConfInt_0.99 <- NULL
+pp$var                <- NULL
+rownames(pp) <- NULL
+names(pp) <- c("Trend %/year", "Trend p-value", "Sky condition", "Season", "Trend statistical sign.")
+
+panderOptions("table.alignment.default", "left")
+pander(pp,
+       caption = "Trends of daily means by sky conditions.")
+```
+\normalsize
+
+\FloatBarrier
 
@@
 
-\FloatBarrier
+```{r cm, echo=F, fig.cap="Donditions."}
+knitr::include_graphics("./images/CumulativeDailyCumSum-1.pdf")
+```
 

### 2023-05-05 Article.Rmd
Commit 2023-05-05 15:17

@@
 fontsize:       12pt
-geometry:       "left=0.5in,right=0.5in,top=0.5in,bottom=0.5in"
+geometry:       "left=0.75in,right=0.75in,top=0.75in,bottom=0.75in"
 link-citations: yes

### 2023-05-05 Article.Rmd
Commit 2023-05-05 18:17

--- 2023-05-05 Article.Rmd
@@
 urlcolor:       blue
-
 bibliography:   [references.bib]
 biblio-style:   apalike
+date:           "`r format(Sys.time(), '%F')`"
 
@@
   - \captionsetup{font=small}
+  - \usepackage{setspace}\doublespacing
+  - \usepackage{float}
 
@@
     fig_height:      5
-date: "`r format(Sys.time(), '%F')`"
 

### 2023-05-05 Article.Rmd
Commit 2023-05-05 19:17

@@
     fig_caption:     yes
-    keep_tex:        yes
+    keep_tex:        no
     latex_engine:    xelatex

### 2023-05-05 Article.Rmd
Commit 2023-05-05 21:17

@@
 <!-- basic concepts -->
-The approach of our method to characterize the sky conditions and the definition of each sky condition has some ......μη αντικειμένικότητα
+Our method to characterize the sky conditions and the definition of each sky condition has some subjectivity.
+The algorithm was calibrated with the main focus to identify the presence of clouds on the sky dome, although there are marginal cases that there will be false positives or false negateives.
 

### 2023-05-05 Article.Rmd
Commit 2023-05-05 22:17

@@
 Our method to characterize the sky conditions and the definition of each sky condition has some subjectivity.
-The algorithm was calibrated with the main focus to identify the presence of clouds on the sky dome, although there are marginal cases that there will be false positives or false negateives.
+The algorithm was calibrated with the main focus to identify the presence of clouds on the sky dome, although there are marginal cases that there will be false positives or false negatives identifications.
 
@@
 
-We can separate the data preparation in three parts, a) the acquisition of radiation measurements from the sensor, b) a radiation data quality check, and c) the identification of "clear sky" conditions from the radiometric data.
+We can separate the data preparation in three parts,
+a) the acquisition of radiation measurements from the sensor, 
+b) a radiation data quality check, and 
+c) the identification of "clear sky" conditions from the radiometric data.
 
@@
 <!-- measurement -->
-We sample the signal of the CM-21 pyranometer with a rate of $1 {Hz}$, and record the average signal and the standard deviation over each minute of the hour.
+The signal of the CM-21 pyranometer is sampled with a rate of $1 {Hz}$, and a record the average signal and the standard deviation every one minute is recorded.
 <!-- dark correction -->

### 2023-05-06 Article.Rmd
Commit 2023-05-06 10:17

@@
 
-```{r cm, echo=F, fig.cap="Donditions."}
+```{r cumsumALL, echo=F, fig.cap="Donditions."}
 knitr::include_graphics("./images/CumulativeDailyCumSum-1.pdf")
+```
+
+```{r cumsumCLEAR, echo=F, fig.cap="Donditions."}
+knitr::include_graphics("./images/CumulativeDailyCumSum-5.pdf")
+```
+
+```{r cumsumCLOUD, echo=F, fig.cap="Donditions."}
+knitr::include_graphics("./images/CumulativeDailyCumSum-9.pdf")
 ```

### 2023-05-07 Article.Rmd
Commit 2023-05-06 11:17
Commit 2023-05-07 14:17

@@
 
+......
+
 <!-- time span -->
-The global radiation was obtained from the pyranometer CM-21, for the period
-`r min(ALL_1_daily_DESEAS$Date)` -  `r max(ALL_1_daily_DESEAS$Date)`.
-In some algorithms we used direct radiation data for comparisons, obtained by a CM-21 pyrheliometer, with data span
-`r min(ALL_1_daily_DESEAS[!is.na(DIR_att), Date])` - `r max(ALL_1_daily_DESEAS[!is.na(DIR_att), Date])`.
+The global radiation was obtained from the pyranometer CM-21, for the period from
+`r min(ALL_1_daily_DESEAS$Date)` to `r max(ALL_1_daily_DESEAS$Date)`.
+In some algorithms we auxialiary used direct radiation data for comparisons, obtained by a CM-21 pyrheliometer, with data span from 
+`r min(ALL_1_daily_DESEAS[!is.na(DIR_att), Date])` to `r max(ALL_1_daily_DESEAS[!is.na(DIR_att), Date])`.
+
+......
 
@@
 
+......
+
 <!-- basic concepts -->
@@
 The algorithm was calibrated with the main focus to identify the presence of clouds on the sky dome, although there are marginal cases that there will be false positives or false negatives identifications.
+
+......
 
@@
 
+This is it \@ref(fig:trendALL).
+
 ---------------------------
@@
 
-## Long-term trends
-
-This is it \@ref(fig:trendALL).
-
-\FloatBarrier
-
 <!--\tiny -->

### 2023-05-07 Article.Rmd
Commit 2023-05-07 21:17

@@
 <!-- dark correction -->
-The measurements are corrected for the zero offset of the instrument, using the "dark signal" (Sun elevation bellow $10^\circ$, for a period of $3 h$).
+The measurements are corrected for the zero offset of the instrument, using the "dark signal" (Sun elevation angle bellow $-10^\circ$, for a period of $3 h$).
 <!-- physical quantities -->

### 2023-05-09 Article.Rmd
Commit 2023-05-09 11:17

@@
 Using the mean daily Global radiation we produce the trends of each data set, as departure from the seasonal value.
+Only for All-sky (Fig. \@ref(fig:trendALL)) and Cloudy-sky (Fig. \@ref(fig:trendCLOUD)) conditions the statistical significance of the trends is acceptable (Tab. \@ref(tab:trendtable)).
 
@@
 pander(pp,
-       caption = "Trends of daily means by sky conditions.")
+       caption = "\\label{tab:trendtable}Trends of daily means by sky conditions.")
+# \\label{tab:tab2}
+# knitr::kable(pp, caption = "Trends of daily means by sky conditions.")
 ```
@@
 ## Long term trends by season of year
+
+Similar we have produced the deseaonilised trend for different sky conditions for each season of the year, using the corresponding mean monthly values.
 
@@
 pander(pp,
-       caption = "Trends of daily means by sky conditions.")
+       caption = "\\label{tab:trendseasontable}Trends of daily means by sky conditions for the season of the year.")
 ```

### 2023-05-09 Article.Rmd
Commit 2023-05-09 12:17
Commit 2023-05-10 00:17

@@
 
-panderOptions("table.alignment.default", "left")
+# panderOptions("table.alignment.default", "left")
+panderOptions("table.alignment.default", "center")
 pander(pp,
@@
 
+A method to evaluate changes in the long term trend is to use the cumulative sum of the variable.
+
 ```{r cumsumALL, echo=F, fig.cap="Donditions."}
@@
 
-```{r cumsumCLEAR, echo=F, fig.cap="Donditions."}
+<!-- Not showing this trend -->
+```{r cumsumCLEAR, echo=F, fig.cap="Donditions.", include=FALSE}
 knitr::include_graphics("./images/CumulativeDailyCumSum-5.pdf")

### 2023-05-09 Article.Rmd
Commit 2023-05-10 01:17

@@
 
-```{r trendALL, echo=F, fig.cap="Daily Global radiation seasonal anomaly and long term trend, for all sky conditions."}
+```{r trendALL, echo=F, fig.cap="Anomaly (\\%) of the daily Global radiation relative to climatogical values for 1993 - 2022.
+The straight black line shows the lond term trend for All-sky conditions."}
 knitr::include_graphics("./images/LongtermTrends-2.pdf")
@@
 \normalsize
+
+\FloatBarrier
+
+## Trends by SZA
 

### 2023-05-09 DHI_GHI_2_sza_trends.R

@@ -107,7 +107,7 @@ par(pch = ".")
 
 ## choose to grid some plots
 FIGURESGRID <- TRUE
-# FIGURESGRID <- FALSE
+FIGURESGRID <- FALSE
 
 ## choose loess criterion for span
 LOESS_CRITERIO <-  c("aicc", "gcv")[1]

### 2023-05-10 Article.Rmd
Commit 2023-05-10 08:17

--- 2023-05-09 Article.Rmd
@@
     fig_height:      4.5
-  html_document:
-    toc:             true
-    fig_width:       7.5
-    fig_height:      5
 
@@
 `r min(ALL_1_daily_DESEAS$Date)` to `r max(ALL_1_daily_DESEAS$Date)`.
-In some algorithms we auxialiary used direct radiation data for comparisons, obtained by a CM-21 pyrheliometer, with data span from 
+In some algorithms we auxiliary used direct radiation data for comparisons, obtained by a CM-21 pyrheliometer, with data span from 
 `r min(ALL_1_daily_DESEAS[!is.na(DIR_att), Date])` to `r max(ALL_1_daily_DESEAS[!is.na(DIR_att), Date])`.
@@
 <!-- dark correction -->
-The measurements are corrected for the zero offset of the instrument, using the "dark signal" (Sun elevation angle bellow $-10^\circ$, for a period of $3 h$).
+The measurements were corrected for the zero offset of the instrument, using the "dark signal" (Sun elevation angle bellow $-10^\circ$, for a period of $3 h$).
 <!-- physical quantities -->
@@
 
-```{r trendALL, echo=F, fig.cap="Daily Global radiation seasonal anomaly and long term trend, for all sky conditions."}
+```{r trendALL, echo=F, fig.cap="Anomaly (\\%) of the daily Global radiation relative to climatological values for 1993 - 2022.
+The straight black line shows the long term trend for All-sky conditions."}
 knitr::include_graphics("./images/LongtermTrends-2.pdf")
@@
 
+## Trends by SZA
+
+\FloatBarrier
+
 ## Long term trends by season of year
 
-Similar we have produced the deseaonilised trend for different sky conditions for each season of the year, using the corresponding mean monthly values.
+Similar we have produced the deseasonalized trend for different sky conditions for each season of the year, using the corresponding mean monthly values.
 
@@
 
-<!-- Not showing this trend -->
+<!--
+Not showing this trend
 ```{r cumsumCLEAR, echo=F, fig.cap="Donditions.", include=FALSE}
@@
 ```
+-->
 
@@
 
-Both dataset were aggregate in $1^\circ$ SZA bins by am, pm, am+pm
+Both datasets were aggregate in $1^\circ$ SZA bins by am, pm, am+pm
 

### 2023-05-10 DHI_GHI_0_data_input.R

@@ -242,7 +242,7 @@ if (havetorun) {
     ## . ----
     #### 1. long-term  ---------------------------------------------------------
 
-    ##  Daily means of data ----------------------------------------------------
+    ##  Daily means  -----------------------------------------------------------
 
     ALL_1_daily_mean <-
         DATA_all[,.(DIR_att       = mean(DIR_att,     na.rm = T),
@@ -982,9 +982,9 @@ if (havetorun) {
 
 
     ##.----
-    ####  2. Long term by SZA  #################################################
+    ####  2. Long term by SZA  -------------------------------------------------
 
-    ## _ Calculate daily SZA means ####
+    ## Daily SZA means ---------------------------------------------------------
     ALL_2_daily_mean <-
         DATA_all[, .(DIR_att       = mean(DIR_att,    na.rm = T),
                      HOR_att       = mean(HOR_att,    na.rm = T),
@@ -1094,20 +1094,40 @@ if (havetorun) {
     CLOUD_2_daily_mean[ DIR_att_N <= SZA_aggregation_N_lim, DIR_transp_EM := NA ]
 
 
-    ## _ By season of year -----------------------------------------------------
-    ## add season of year tag
-    DATA_all[  month(Date) %in% c(12, 1, 2), Season := "Winter"]
-    DATA_all[  month(Date) %in% c( 3, 4, 5), Season := "Spring"]
-    DATA_all[  month(Date) %in% c( 6, 7, 8), Season := "Summer"]
-    DATA_all[  month(Date) %in% c( 9,10,11), Season := "Autumn"]
-    DATA_Clear[month(Date) %in% c(12, 1, 2), Season := "Winter"]
-    DATA_Clear[month(Date) %in% c( 3, 4, 5), Season := "Spring"]
-    DATA_Clear[month(Date) %in% c( 6, 7, 8), Season := "Summer"]
-    DATA_Clear[month(Date) %in% c( 9,10,11), Season := "Autumn"]
-    DATA_Cloud[month(Date) %in% c(12, 1, 2), Season := "Winter"]
-    DATA_Cloud[month(Date) %in% c( 3, 4, 5), Season := "Spring"]
[…]

### 2023-05-10 DHI_GHI_1_longterm_trends.R
Commit 2023-05-10 08:17

@@ -47,7 +47,7 @@
 
 #+ echo=F, include=F
 knitr::opts_chunk$set(comment    = ""       )
-knitr::opts_chunk$set(dev        = "pdf"    )
+# knitr::opts_chunk$set(dev        = "pdf"    )
 # knitr::opts_chunk$set(dev        = "png"    )
 knitr::opts_chunk$set(out.width  = "100%"   )
 knitr::opts_chunk$set(fig.align  = "center" )

### 2023-05-10 DHI_GHI_2_sza_trends.R

@@ -47,7 +47,7 @@
 
 #+ echo=F, include=F
 knitr::opts_chunk$set(comment    = ""       )
-knitr::opts_chunk$set(dev        = "pdf"    )
+# knitr::opts_chunk$set(dev        = "pdf"    )
 # knitr::opts_chunk$set(dev        = "png"    )
 knitr::opts_chunk$set(out.width  = "100%"   )
 knitr::opts_chunk$set(fig.align  = "center" )
@@ -101,6 +101,9 @@ options(error = function() {
 })
 
 ## __ Flags --------------------------------------------------------------------
+
+DRAFT <- TRUE
+DRAFT <- FALSE
 
 ## override plot options
 par(pch = ".")
@@ -321,7 +324,7 @@ for (type in unique(szatrends$DATA)) {
 
             abline(h = 0, lty = 3 )
 
-            title(paste(awename, translate(type), translate(avar) ), cex.main = .8)
+            title(paste(awename, translate(type), translate(avar)), cex.main = .8)
 
             lines(pam$SZA, pam[[awe]], pch = pch_am, col = 2, type = "b", lwd = 1, cex = 0.5)
             lines(ppm$SZA, ppm[[awe]], pch = pch_pm, col = 3, type = "b", lwd = 1, cex = 0.5)

### 2023-05-11 Article.Rmd
Commit 2023-05-11 12:17

@@
 pp <- data.table::fread("./figures/tbl_longterm_trends.dat")
+cat('\n\n\\footnotesize\n\n')
 pander::pander(pp,
                caption = "Table of trends and stats")
+cat('\n\n\\normalsize\n\n')
 ```

### 2023-05-15 Article.Rmd
Commit 2023-05-12 21:17
Commit 2023-05-15 10:17

@@
 The manual screening is aided by a radiation data quality assurance procedure, adjusted for this site, based on methods of Long and Shi\ [-@long_automated_2008; -@Long2006].
-
-....
-Only data characterized as good quality were used in this study.
-.....
+Problematic recordings have been excluded from processing, although it is not possible to detect all the bad data, the large number of data and the aggregation scheme we use, can provides us with accurate radiometric measurements.
 
@@
 
+..............
+
 \FloatBarrier
@@
 
-## Zenith angle dependence and years seasons.
+## Zenith angle dependency and seasons of the year.
+
+..............
 
@@
 
-Some more text here
-
 **END**

### 2023-05-15 DHI_GHI_0_variables.R

@@ -34,7 +34,7 @@ translate <- function(x) {
 LAST_DAY                 <- as.Date("2023-04-01")
 FIRST_DAY                <- as.Date("1900-07-01")
 # FIRST_DAY                <- as.Date("2005-01-01") ## data inspected by me
-FIRST_DAY                <- as.Date("2016-04-01") ## start of chp1
+# FIRST_DAY                <- as.Date("2016-04-01") ## start of chp1
 
 #### Paths ####
 CLEARdir                 <- "~/DATA/Broad_Band/CS_id"

### 2023-05-16 Article.Rmd
Commit 2023-05-16 13:17

@@
 ---
-title:         "Trends of downward SDR in Thessaloniki"
+title:         "Trends of SDR in Thessaloniki"
 author:
@@
 
+<!-- SDR or GHI -->
+
 # Introduction.
+
+The shortwave downward solar irradiance (SDR) at the Earth's surface plays s significant role, on the climate.
+Changes of the SDR can be related to major changes on the factors affecting the radiative forcing balance.
+In this study we examine the trends of SDR with the ground-based measurements at Thessaloniki for the period 1993 - 2023.
+
+[@Bais2013]
 

### 2023-05-25 Article.Rmd
Commit 2023-05-16 14:17
Commit 2023-05-25 13:17

@@
 title:         "Trends of SDR in Thessaloniki"
+
+# documentclass:  article
+# classoption:    a4paper,oneside
+# fontsize:       12pt
+# geometry:       "left=0.75in,right=0.75in,top=0.75in,bottom=0.75in"
+# link-citations: yes
+# colorlinks:     yes
+# urlcolor:       blue
+# bibliography:   [references.bib]
+# biblio-style:   apalike
+# date:           "`r format(Sys.time(), '%F')`"
+# 
+# header-includes:
+#   - \usepackage{caption}
+#   - \usepackage{placeins}
+#   - \captionsetup{font=small}
+#   - \usepackage{setspace}\doublespacing
+#   - \usepackage{float}
+
 author:
-  - Natsis Athanasios^[Laboratory of Atmospheric Physics, AUTH, natsisphysicist@gmail.com]
-  - Alkiviadis Bais^[Laboratory of Atmospheric Physics, AUTH]
-abstract:
-  "Study of GHI and DNI radiation for 'clear-sky' and all-sky conditions."
+  - name:        Natsis Athanasios
+    email:       natsisphysicist@gmail.com
+    affiliation: Laboratory of Atmospheric Physics, AUTH
+    correspondingauthor: true
+    footnote: 1
+  - name:        Alkiviadis Bais
+    email:       bob@example.com
+    affiliation: Another University
+address:
+  - code: Some Institute of Technology
+    organization: Big Wig University
+    addressline: 1 main street
+    city: Gotham
+    state: State
+    postcode: 123456
+    country: United States
+  - code: Another University
+    organization: Department
+    addressline: A street 29
+    postcode: 2054 NX
+    city: Manchester,
+    country: The Netherlands
+footnote:
+  - code: 1
+    text: "This is the first author footnote."
+  - code: 2
+    text: "Another author footnote."
+abstract: |
+  Study of GHI and DNI radiation for 'clear-sky' and all-sky conditions.
 
-documentclass:  article
-classoption:    a4paper,oneside
-fontsize:       12pt
-geometry:       "left=0.75in,right=0.75in,top=0.75in,bottom=0.75in"
-link-citations: yes
-colorlinks:     yes
-urlcolor:       blue
-bibliography:   [references.bib]
-biblio-style:   apalike
-date:           "`r format(Sys.time(), '%F')`"
-
-header-includes:
-  - \usepackage{caption}
-  - \usepackage{placeins}
-  - \captionsetup{font=small}
-  - \usepackage{setspace}\doublespacing
-  - \usepackage{float}
+  It consists of two paragraphs.
+keywords: 
+  - keyword1
+  - keyword2
+journal: "An awesome journal"
+date: "`r Sys.Date()`"
+classoption: preprint, 3p, authoryear
+bibliography: references.bib
+linenumbers: false
+numbersections: true
 
 output: 
+  rticles::elsevier_article:
+    keep_tex: true
+    citation_package: natbib
   bookdown::pdf_document2:
@@
     fig_height:      4.5
-
 ---
@@
 knitr::opts_chunk$set(fig.align  = "center")
-knitr::opts_chunk$set(fig.pos    = '!h'    )
+knitr::opts_chunk$set(fig.pos    = '!ht'   )
 library(pander)
@@
 Changes of the SDR can be related to major changes on the factors affecting the radiative forcing balance.
+
 In this study we examine the trends of SDR with the ground-based measurements at Thessaloniki for the period 1993 - 2023.
@@
 
-\FloatBarrier
+<!-- \FloatBarrier -->
 

### 2023-05-25 Article_B138.Rmd
Commit 2023-05-25 18:17

--- 2023-05-25 Article.Rmd
@@
 address:
-  - code: Some Institute of Technology
+  - code:         Aristotle University of Thessaloniki
     organization: Big Wig University
     addressline:  1 main street
-    city: Gotham
+    city:         Thessaloniki
     state: State
     postcode: 123456
-    country: United States
+    country:      Greece
   - code:         Another University

### 2023-05-25 Article.Rmd
Commit 2023-05-25 22:17

@@
     email:       natsisphysicist@gmail.com
-    affiliation: Laboratory of Atmospheric Physics, AUTH
+    affiliation: Laboratory of Atmospheric Physics
     correspondingauthor: true
@@
 address:
-  - code: Some Institute of Technology
-    organization: Big Wig University
+  - code:         Laboratory of Atmospheric Physics
+    organization: Aristotle University of Thessaloniki
     addressline:  1 main street
-    city: Gotham
+    city:         Thessaloniki
     state:        State
     postcode:     123456
-    country: United States
+    country:      Greece
   - code:         Another University
@@
   rticles::elsevier_article:
-    keep_tex: true
+    keep_tex:         false
     citation_package: natbib
@@
 
-<!-- SDR or GHI -->
+<!-- SDR or GHI ? -->
 

### 2023-06-05 Article.Rmd
Commit 2023-06-02 09:17
Commit 2023-06-05 14:17
Commit 2023-06-05 15:17

@@
 Changes of the SDR can be related to major changes on the factors affecting the radiative forcing balance.
+Multiple studies attempt to evaluate the phenomenon and the potential causes [@Ohvril2009; @Schwarz2020; @Wild2009; @Wild2012; @Xia2007; @Zerefos2009].
+Although, there are regional variation in the radiation, with both positive and negative trends, most of the studies agree that a major factor is the interaction of aerosols and clouds.
 
-In this study we examine the trends of SDR with the ground-based measurements at Thessaloniki for the period 1993 - 2023.
+In this study we examine the trends of SDR with ground-based measurements at Thessaloniki for the period 1993 - 2023, as derived from a CM-21 pyranometer.
+We reevaluate and extend the dataset used by @Bais2013, applying a different algorithm for the identification of clear-/cloud-sky instances, and we derive the radiation trends for this period under different sky conditions (all-sky, clear-sky and cloud-sky).
 
-[@Bais2013]
-
-......
-
-<!-- time span -->
-The global radiation was obtained from the pyranometer CM-21, for the period from
-`r min(ALL_1_daily_DESEAS$Date)` to `r max(ALL_1_daily_DESEAS$Date)`.
-In some algorithms we auxiliary used direct radiation data for comparisons, obtained by a CM-21 pyrheliometer, with data span from 
-`r min(ALL_1_daily_DESEAS[!is.na(DIR_att), Date])` to `r max(ALL_1_daily_DESEAS[!is.na(DIR_att), Date])`.
+....describe CS and cloud definition....
 
@@
 # Data and methodology.
+
+<!-- time span -->
+The SDR is equivalent to the whole sky global radiation (GHI), and was obtained with a horizontal leveled, CM-21 pyranometer.
+Data spans the period 
+`r min(ALL_1_daily_DESEAS$Date)` to `r max(ALL_1_daily_DESEAS$Date)`.
+In some algorithms we used as auxiliary data, the direct beam radiation (DNI) for comparisons, obtained by a CHP-1 pyrheliometer, with data availability from 
+`r min(ALL_1_daily_DESEAS[!is.na(DIR_att), Date])` to `r max(ALL_1_daily_DESEAS[!is.na(DIR_att), Date])`.
+
+......
 
@@
 $35^\circ$ and $120^\circ$ and SZA greater than $80^\circ$.
-At those cases the Sun systematically, is not visible by the instruments.
+At those instances, Sun is systematically, not visible by the instruments location.
 

### 2023-06-05 Article.Rmd
Commit 2023-06-05 19:17

@@
     fig_caption:      yes
-    keep_tex:        no
+    keep_tex:         yes
     latex_engine:     xelatex
@@
 
-## Preparation.
+<!-- ## Preparation. -->
 
-We can separate the data preparation in three parts,
-a) the acquisition of radiation measurements from the sensor, 
+There are three distinct step to the creation of this dataset:
+a) the acquisition of radiation measurements from the sensors, 
 b) a radiation data quality check, and 
@@
 
-### Acquisition of radiometric data.
+<!-- ### Acquisition of radiometric data. -->
 
 <!-- measurement -->
-The signal of the CM-21 pyranometer is sampled with a rate of $1 {Hz}$, and a record the average signal and the standard deviation every one minute is recorded.
+For the acquisition of radiometric data,
+the signal of the broadband instruments is sampled with a rate of $1 \text{Hz}$, and the signal mean and standard deviation is recorded for every minute.
 <!-- dark correction -->
-The measurements were corrected for the zero offset of the instrument, using the "dark signal" (Sun elevation angle bellow $-10^\circ$, for a period of $3 h$).
+The measurements are corrected for the zero offset of the instrument signal.
+As reference of the "dark signal" is used a Sun elevation angle bellow $-10^\circ$, for a period of $3 \text{h}$.
 <!-- physical quantities -->
@@
 
-### Radiation data quality check
-
+<!-- ### Radiation data quality check -->
+<!-- The radiation quality is assured with a manual screening, aided by an automatic assessment process. -->
 <!-- screening -->
-A manual screening is performed, to remove inconsistencies and erroneous recordings, that can occur during a long continuous operation.
+A manual screening is performed, to remove inconsistencies and erroneous recordings, that can occur during the long continuous operation of the instrument.
 <!-- Quality assurance -->
 The manual screening is aided by a radiation data quality assurance procedure, adjusted for this site, based on methods of Long and Shi\ [-@long_automated_2008; -@Long2006].
-Problematic recordings have been excluded from processing, although it is not possible to detect all the bad data, the large number of data and the aggregation scheme we use, can provides us with accurate radiometric measurements.
+Problematic recordings have been excluded from further processing. 
+Although it is not possible to detect all the bad data, the large number of data and the aggregation scheme we use, can provide us with accurate radiometric measurements, for the scope of this study.
 
@@
 In order to estimate the effect of the clouds on the global radiation we create three datasets. 
-The all sky conditions containing all the valid measurements. 
-The clear-sky conditions data and the remainder part the cloudy condition data. 
-To identify the clear-sky conditions we used a method proposed by @Long2000 and by @Reno2016 that was adapted and configured for the site.
+The all-sky conditions, containing all the valid measurements. 
+The clear-sky conditions data, and the remainder part the cloudy condition data (cloud-sky). 
+To identify the clear-sky conditions we used the method proposed by @Long2000 and by @Reno2016, that was adapted and configured for the site.
+
+.....
+cs - cloud distinction elaborate
+......
 
@@
 Using the mean daily Global radiation we produce the trends of each data set, as departure from the seasonal value.
-Only for All-sky (Fig. \@ref(fig:trendALL)) and Cloudy-sky (Fig. \@ref(fig:trendCLOUD)) conditions the statistical significance of the trends is acceptable (Tab. \@ref(tab:trendtable)).
+Only for All-sky (Fig. \ref{fig:trendALL}) and Cloudy-sky (Fig. \ref{fig:trendCLOUD}) conditions the statistical significance of the trends is acceptable (Tab. \ref{tab:trendtable}).
 

### 2023-06-06 Article.Rmd
Commit 2023-06-06 12:17

@@
 
-The shortwave downward solar irradiance (SDR) at the Earth's surface plays s significant role, on the climate.
-Changes of the SDR can be related to major changes on the factors affecting the radiative forcing balance.
+The shortwave downward solar irradiance (SDR) at the Earth's surface plays a significant role, on Earths climate.
+Changes of the SDR can be related to major changes on factors affecting the radiative forcing balance and human activities.
 Multiple studies attempt to evaluate the phenomenon and the potential causes [@Ohvril2009; @Schwarz2020; @Wild2009; @Wild2012; @Xia2007; @Zerefos2009].
-Although, there are regional variation in the radiation, with both positive and negative trends, most of the studies agree that a major factor is the interaction of aerosols and clouds.
+Although, there is regional variation of SDR, with both positive and negative trends, most of the studies agree that the major factors are the interaction of aerosols and clouds.
 
-In this study we examine the trends of SDR with ground-based measurements at Thessaloniki for the period 1993 - 2023, as derived from a CM-21 pyranometer.
+In this study, we examine the trends of SDR with ground-based measurements at Thessaloniki for the period 1993 - 2023, as derived from a CM-21 pyranometer.
+<!-- goals -->
 We reevaluate and extend the dataset used by @Bais2013, applying a different algorithm for the identification of clear-/cloud-sky instances, and we derive the radiation trends for this period under different sky conditions (all-sky, clear-sky and cloud-sky).
@@
 ....describe CS and cloud definition....
-
-......
-
-<!-- goals -->
-Here we investigate the trend on shortwave radiation for the last decades at Thessaloniki, and try to infer a relation to sky conditions (all, clear, cloudy).
 
@@
 The SDR is equivalent to the whole sky global radiation (GHI), and was obtained with a horizontal leveled, CM-21 pyranometer.
-Data spans the period 
+SDR data spans the period of 
 `r min(ALL_1_daily_DESEAS$Date)` to `r max(ALL_1_daily_DESEAS$Date)`.

### 2023-06-06 Article.Rmd
Commit 2023-06-06 13:17

@@
 
+... site location and aerosols effect...
+
 ....describe CS and cloud definition....
@@
 
-There are three distinct step to the creation of this dataset:
+There are three distinct steps to the creation of this dataset:
 a) the acquisition of radiation measurements from the sensors, 
 b) a radiation data quality check, and 
-c) the identification of "clear sky" conditions from the radiometric data.
+c) the identification of "clear sky" conditions from the radiometric data
+d) data aggregation and trend analysis.
 
@@
 For the acquisition of radiometric data,
-the signal of the broadband instruments is sampled with a rate of $1 \text{Hz}$, and the signal mean and standard deviation is recorded for every minute.
+the signal of the broadband instruments is sampled with a rate of $1 \text{Hz}$. The mean and standard deviation values are recorded for every minute.
 <!-- dark correction -->
 The measurements are corrected for the zero offset of the instrument signal.
-As reference of the "dark signal" is used a Sun elevation angle bellow $-10^\circ$, for a period of $3 \text{h}$.
+As reference of the "dark signal" is used the Sun elevation angle bellow $-10^\circ$, for a period of $3 \text{h}$.
 <!-- physical quantities -->
-The signal is converted to radiation using a ramped value of the instrument sensitivity, derived from the eight laboratory calibrations of the instrument, during the study period.
+The signal is converted to radiation flux, using a ramped value of the instrument sensitivity, derived from the eight laboratory calibrations of the instrument, during the study period.
 
@@
 <!-- screening -->
-A manual screening is performed, to remove inconsistencies and erroneous recordings, that can occur during the long continuous operation of the instrument.
+A manual screening is performed, to remove inconsistencies and erroneous recordings, that can occur randomly or systematically, during the long continuous operation of the instrument.
 <!-- Quality assurance -->
 The manual screening is aided by a radiation data quality assurance procedure, adjusted for this site, based on methods of Long and Shi\ [-@long_automated_2008; -@Long2006].
-Problematic recordings have been excluded from further processing. 
+Thus, problematic recordings have been excluded from further processing. 
 Although it is not possible to detect all the bad data, the large number of data and the aggregation scheme we use, can provide us with accurate radiometric measurements, for the scope of this study.
 
-### Clear-sky identification
-
-In order to estimate the effect of the clouds on the global radiation we create three datasets. 
+<!-- ### Clear sky id --->
+In order to estimate the effect of the clouds on the SDR we created three datasets, by characterizing each one-minute measurement with a corresponding sky condition. 
 The all-sky conditions, containing all the valid measurements. 
-The clear-sky conditions data, and the remainder part the cloudy condition data (cloud-sky). 
+The clear-sky conditions data, where we have inferred that the sky was almost clear of clouds, and the remainder part the data as cloudy sky conditions data (cloud-sky). 
 To identify the clear-sky conditions we used the method proposed by @Long2000 and by @Reno2016, that was adapted and configured for the site.
@@
 $35^\circ$ and $120^\circ$ and SZA greater than $80^\circ$.
-At those instances, Sun is systematically, not visible by the instruments location.
+At those instances, Sun is systematically, not visible by the instrument's location.
 

### 2023-06-06 Article.Rmd
Commit 2023-06-06 19:17

@@
 
-... site location and aerosols effect...
-
-....describe CS and cloud definition....
-
-......
+... site location description and aerosols effect...
 
@@
 The algorithm was calibrated with the main focus to identify the presence of clouds on the sky dome, although there are marginal cases that there will be false positives or false negatives identifications.
-
-......
 
@@
 
-<!-- ### Clear sky id --->
+<!-- ### Clear sky ID -->
 In order to estimate the effect of the clouds on the SDR we created three datasets, by characterizing each one-minute measurement with a corresponding sky condition. 
@@
 To identify the clear-sky conditions we used the method proposed by @Long2000 and by @Reno2016, that was adapted and configured for the site.
-
-.....
-cs - cloud distinction elaborate
-......
+<!-- CS subjectivity -->
+We have to note, that the definition of what constitutes as clear or cloudy sky, has some subjectivity, due to the used method of characterization.
+As a result, the details of the definition are site specific, it relies on a combination of threshold of comparisons with ideal actinometric models and statistics on different signal behaviours.
 
@@
 
-### Data selection
-
-Due to a significant measurement uncertainty near the horizon, we exclude
+<!-- ### Data selection -->
+Due to a significant measurement uncertainty near the horizon, we have to exclude
 all measurements with SZA greater than $`r 90-MIN_ELEVA`^\circ$. 
-Moreover, due to some obstructions around the site (hills and buildings), we exclude data with Azimuth angle between 
-$35^\circ$ and $120^\circ$ and SZA greater than $80^\circ$.
-At those instances, Sun is systematically, not visible by the instrument's location.
-
+Moreover, due to some obstructions around the site (hills and buildings), we excluded data with Azimuth angle between 
+$35^\circ$ and $120^\circ$ with SZA greater than $80^\circ$.
+At the latter instances, Sun is systematically, not visible by the instrument's location.
+<!-- 1au -->
 To make the measurements comparable throughout the dataset. 
-<!-- 1au -->
 We adjusted all 1-minute radiometric values to the mean Sun - Earth distance.
 <!-- TSI -->
-<!-- And, we made all measurements relative to the TSI at 1au, in order to compensate for the Suns intensity variability, using a homogenized timeseries of satellite TSI observations.
+Subsequently, we made all measurements relative to the TSI at 1au, in order to compensate for the Suns intensity variability, using a homogenized time series of satellite TSI observations.
 As a result, we processed all radiation data as a fraction of Suns TSI.
--->
 
-### Process and aggregation of radiometric data
+..... cite NOAA TSI .....
+
+## Process and aggregation of radiometric data
 

### 2023-06-06 Article.Rmd
Commit 2023-06-06 20:17

@@
     affiliation:  Laboratory of Atmospheric Physics
+    footnote:     1
     correspondingauthor: true
-    footnote:    1
   - name:         Alkiviadis Bais
@@
 <!-- Seasonal -->
-For the seasons of the year trends we grouped the mean daily values by season (December - February: Winter, March - May: Spring, etc.)
+For the seasons of the year trends, we also grouped the mean daily values by season (December - February: Winter, March - May: Spring, etc.)
+
+<!-- Deseasonalization -->
+Finally, for each data set, we remove the natural occurring seasonal variation.
+This is done, by calculating the mean values for the appropriate time step, and subtracting the annual cycle from the actual data.
+Thus, for each data set we have obtained the relative departure of the seasonal mean.
+These values are presented in the following section.
+
+.......................
+
+## Aggregation
+
+Both datasets were aggregate in $1^\circ$ SZA bins by am, pm, am+pm
+
+by daily mean
+
+daily sum
+
+seasonal daily mean
+
+filtering of means
+
+statistical significance
+
+Significance and aggregation scheme and filtering will be noted along with the corresponding results.
 
@@
 
-This is it \@ref(fig:trendALL).
+About:
 
----------------------------
+- long terms trends
+- seasonal trends
+- effect of SZA
+- effect of clouds
 
-## Aggregation
-
-Both datasets were aggregate in $1^\circ$ SZA bins by am, pm, am+pm
-
-by daily mean
-
-daily sum
-
-seasonal daily mean
-
-<!-- Deseasonalization -->
-We remove the natural occurring seasonal variation, by calculating the mean
-values for the corresponding time step, and subtracting them .........
+- Aerosols
 

### 2023-06-07 Article.Rmd
Commit 2023-06-07 15:17

@@
 
-# documentclass:  article
-# classoption:    a4paper,oneside
-# fontsize:       12pt
-# geometry:       "left=0.75in,right=0.75in,top=0.75in,bottom=0.75in"
-# link-citations: yes
-# colorlinks:     yes
-# urlcolor:       blue
-# bibliography:   [references.bib]
-# biblio-style:   apalike
-# date:           "`r format(Sys.time(), '%F')`"
 # 
@@
     affiliation:  Laboratory of Atmospheric Physics
-    footnote:     1
+#    footnote:     1
     correspondingauthor: true
   - name:         Alkiviadis Bais
-    email:        bob@example.com
-    affiliation:  Another University
+    email:        abais@auth.gr 
+    affiliation:  Laboratory of Atmospheric Physics
+    correspondingauthor: false
 address:
@@
     organization: Aristotle University of Thessaloniki
-    addressline:  1 main street
+    addressline:  Campus Box 149
     city:         Thessaloniki
-    state:        State
-    postcode:     123456
+    # state:        State
+    postcode:     54124
     country:      Greece
-  - code:         Another University
-    organization: Department
-    addressline:  A street 29
-    postcode:     2054 NX
-    city:         Manchester,
-    country:      The Netherlands
-footnote:
-  - code:         1
-    text:         "This is the first author footnote."
-  - code:         2
-    text:         "Another author footnote."
+#  - code:         Another University
+#    organization: Department
+#    addressline:  A street 29
+#    postcode:     2054 NX
+#    city:         Manchester,
+#    country:      The Netherlands
+# footnote:
+#   - code:         1
+#     text:         "This is the first author footnote."
+#   - code:         2
+#     text:         "Another author footnote."
 abstract: |
@@
     fig_height:       4.5
+    documentclass:  article
+    classoption:    a4paper,oneside
+    fontsize:       12pt
+# geometry:       "left=0.75in,right=0.75in,top=0.75in,bottom=0.75in"
+# link-citations: yes
+# colorlinks:     yes
+# urlcolor:       blue
+# bibliography:   [references.bib]
+# biblio-style:   apalike
+# date:           "`r format(Sys.time(), '%F')`"
 ---
@@
 
-## Long term trends
+## Long-term trends
 

### 2023-06-07 Article.Rmd
Commit 2023-06-07 16:17

@@
     fontsize:       12pt
-# geometry:       "left=0.75in,right=0.75in,top=0.75in,bottom=0.75in"
-# link-citations: yes
-# colorlinks:     yes
-# urlcolor:       blue
-# bibliography:   [references.bib]
-# biblio-style:   apalike
+    geometry:       "left=0.75in,right=0.75in,top=0.75in,bottom=0.75in"
+    link-citations: yes
+    colorlinks:     yes
+    urlcolor:       blue
+    biblio-style:   apalike
 # date:           "`r format(Sys.time(), '%F')`"

### 2023-06-08 Article.Rmd
Commit 2023-06-08 12:17

@@
 load("./data/common_data_14_2.Rda")
+TSIinfo <- data.table(read.csv("./figures/tbl_tsi_info.dat",
+                    strip.white = TRUE,
+                    sep = ";",
+                    header = TRUE))
 ## variables we use
@@
 ..... cite NOAA TSI .....
+```{r}
+as.Date(TSIinfo[TSI_Source == "NOAA", End])
+```
 

### 2023-06-08 Article.Rmd
Commit 2023-06-08 14:17

@@
 We have to note, that the definition of what constitutes as clear or cloudy sky, has some subjectivity, due to the used method of characterization.
-As a result, the details of the definition are site specific, it relies on a combination of threshold of comparisons with ideal actinometric models and statistics on different signal behaviours.
+As a result, the details of the definition are site specific, it relies on a combination of threshold of comparisons with ideal actinometric models and statistics on different signal behavior.
 
@@
 <!-- ### Data selection -->
-Due to a significant measurement uncertainty near the horizon, we have to exclude
+Due to a significant measurement uncertainty near, the horizon, we have to exclude
 all measurements with SZA greater than $`r 90-MIN_ELEVA`^\circ$. 
@@
 $35^\circ$ and $120^\circ$ with SZA greater than $80^\circ$.
-At the latter instances, Sun is systematically, not visible by the instrument's location.
+On the latter instances, Sun is systematically, not visible by the instrument's location.
 <!-- 1au -->
-To make the measurements comparable throughout the dataset. 
-We adjusted all 1-minute radiometric values to the mean Sun - Earth distance.
+To make the measurements comparable throughout the dataset, we adjusted all 1-minute radiometric values to the mean Sun - Earth distance.
 <!-- TSI -->
-Subsequently, we made all measurements relative to the TSI at 1au, in order to compensate for the Suns intensity variability, using a homogenized time series of satellite TSI observations.
+Subsequently, we made all measurements relative to the TSI at $1 \text{au}$, in order to compensate for the Sun's intensity variability, using a homogenized time series of satellite TSI observations.
+The TSI data is a combination of data from NOAA [@Coddington2005] (`r as.Date(TSIinfo[TSI_Source == "NOAA", Start])` - `r as.Date(TSIinfo[TSI_Source == "NOAA", End])`) and adjusted data from @TSIStsi (`r as.Date(TSIinfo[TSI_Source != "NOAA", Start])` - `r as.Date(TSIinfo[TSI_Source != "NOAA", End])`).
 As a result, we processed all radiation data as a fraction of Suns TSI.
-
-..... cite NOAA TSI .....
-```{r}
-as.Date(TSIinfo[TSI_Source == "NOAA", End])
-```
 
@@
 
-.......................
+........................
 

### 2023-06-08 upload.sh

@@ -1,0 +1,22 @@
+#!/bin/bash
+## created on 2023-02-21
+
+#### enter description here
+
+bwlim=500
+rclone="$HOME/PROGRAMS/rclone"
+config="$HOME/Documents/rclone.conf"
+otheropt=" --checkers=20 --delete-before --stats=300s"
+bwlimit="  --bwlimit=${bwlim}k"
+
+# if [[ "$(hostname)" = "sagan" ]]; then 
+#     echo "Upload all pdfs"
+#     "${rclone}" ${otheropt} ${bwlimit} --config "$config" --include "*.pdf"  sync "$HOME/MANUSCRIPTS/2022_sdr_trends/" "lapauththanasis:/Trends"
+# else
+#     echo "This runs only on sagan"
+# fi
+
+echo "Upload all pdfs"
+"${rclone}" ${otheropt} ${bwlimit} --config "$config" --include "*.pdf" --include "*.odt" copy "$HOME/MANUSCRIPTS/2022_sdr_trends/" "lapauththanasis:/Trends"
+
+exit 0 

### 2023-06-08 Article.Rmd
Commit 2023-06-08 15:17

@@
 keywords: 
-  - keyword1
-  - keyword2
+  - GHI
+  - SDR
+  - Solar Brigthening/Dimming
 journal:        "An awesome journal"
@@
 
-... site location description and aerosols effect...
+...
+
+site location description and aerosols effect
+
+...
 
@@
 
-......
+...
+
+...
 
@@
 
-........................
+........
+
+................
 
@@
 Using the mean daily Global radiation we produce the trends of each data set, as departure from the seasonal value.
-Only for All-sky (Fig. \ref{fig:trendALL}) and Cloudy-sky (Fig. \ref{fig:trendCLOUD}) conditions the statistical significance of the trends is acceptable (Tab. \ref{tab:trendtable}).
+Only for All-sky (Fig. \ref{fig:trendALL}) and Cloudy-sky (Fig. \ref{fig:trendCLOUD}) conditions, the statistical significance of the trends is acceptable (Tab. \ref{tab:trendtable}).
 
@@
 
-..............
+.....
+
+.........
 
@@
 
-```{r cumsumALL, echo=F, fig.cap="Donditions."}
+... different in aggregation ...
+
+```{r cumsumALL, echo=F, fig.cap="Running cumulative sum."}
 knitr::include_graphics("./images/CumulativeDailyCumSum-1.pdf")
@@
 # Conclusions
+
+Our result reaffirm the
 

### 2023-06-08 Article.Rmd
Commit 2023-06-08 17:17

@@
 <!-- TSI -->
-Subsequently, we made all measurements relative to the TSI at $1 \text{au}$, in order to compensate for the Sun's intensity variability, using a homogenized time series of satellite TSI observations.
-The TSI data is a combination of data from NOAA [@Coddington2005] (`r as.Date(TSIinfo[TSI_Source == "NOAA", Start])` - `r as.Date(TSIinfo[TSI_Source == "NOAA", End])`) and adjusted data from @TSIStsi (`r as.Date(TSIinfo[TSI_Source != "NOAA", Start])` - `r as.Date(TSIinfo[TSI_Source != "NOAA", End])`).
-As a result, we processed all radiation data as a fraction of Suns TSI.
+Subsequently, we made all measurements relative to the Total Solar Irradiance (TSI) at $1 \text{au}$, in order to compensate for the Sun's intensity variability, using a homogenized time series of satellite TSI observations.
+The TSI data we use, is a combination of data from NOAA [@Coddington2005] (for `r as.Date(TSIinfo[TSI_Source == "NOAA", Start])` - `r as.Date(TSIinfo[TSI_Source == "NOAA", End])`) and adjusted data from @TSIStsi (for `r as.Date(TSIinfo[TSI_Source != "NOAA", Start])` - `r as.Date(TSIinfo[TSI_Source != "NOAA", End])`).
+As a result, we can analyze all radiation data as a fraction of Sun's TSI.
 

### 2023-06-08 Article.Rmd
Commit 2023-06-08 18:17

@@
 
-```{r trendALL, echo=F, fig.cap="Anomaly (\\%) of the daily Global radiation relative to climatological values for 1993 - 2022.
+```{r trendALL, echo=F, fig.cap="Anomaly (\\%) of the daily Global radiation relative to climatological values for 1993 - 2023.
 The straight black line shows the long term trend for All-sky conditions."}
@@
 
-\footnotesize
+<!-- \footnotesize -->
 ```{r trendtable, echo=F}
@@
 ```
-\normalsize
+<!-- \normalsize -->
 
@@
 
-<!--\tiny -->
-```{r echo=F}
-pp <- data.table::fread("./figures/tbl_longterm_trends.dat")
-cat('\n\n\\footnotesize\n\n')
-pander::pander(pp,
-               caption = "Table of trends and stats")
-cat('\n\n\\normalsize\n\n')
-```
-<!--\normalsize-->
-
+-------------------------------------------------
                    **END**
+-------------------------------------------------

### 2023-06-09 Article.Rmd
Commit 2023-06-09 10:17

@@
 
-## Data process.
+## Data and data selection.
 
@@
 The TSI data we use, is a combination of data from NOAA [@Coddington2005] (for `r as.Date(TSIinfo[TSI_Source == "NOAA", Start])` - `r as.Date(TSIinfo[TSI_Source == "NOAA", End])`) and adjusted data from @TSIStsi (for `r as.Date(TSIinfo[TSI_Source != "NOAA", Start])` - `r as.Date(TSIinfo[TSI_Source != "NOAA", End])`).
-As a result, we can analyze all radiation data as a fraction of Sun's TSI.
+As a result, we can present all radiation data as a fraction of Sun's TSI.
 
-## Process and aggregation of radiometric data
+### Aggregation of radiometric data
 
-In this study, we present any radiometric data as a percent departure from the corresponding seasonal mean.
-When we use daily aggregated data we compute the mean values for the day and exclude instances with less than `r Daily_aggregation_N_lim` valid data points.
-Accordingly, monthly values are computed from the daily aggregated data, month with less than `r Monthly_aggegation_N_lim` days are rejected. 
-The daily seasonal values are computed based on the day of year number and the monthly by the calendar month.
+Before further analysis, we implement an appropriate aggregation scheme on the 1-minute data, to preserve the ....representabiliyt....
+When we use daily mean values we exclude instances with less than `r Daily_aggregation_N_lim` valid data points.
+Accordingly, monthly values are computed from the daily aggregated data, where months with less than `r Monthly_aggegation_N_lim` days are rejected. 
+The daily seasonal values are indexed based on the day of year number, and the monthly by the corresponding calendar month.
 

### 2023-06-09 Article_B140.Rmd
Commit 2023-06-09 13:17

--- 2023-06-09 Article.Rmd
@@
 title:         "Trends of SDR in Thessaloniki"
-
-# 
 # header-includes:
-#   - \usepackage{caption}
-#   - \usepackage{placeins}
-#   - \captionsetup{font=small}
-#   - \usepackage{setspace}\doublespacing
-#   - \usepackage{float}
-
 author:
@@
   Study of GHI and DNI radiation for 'clear-sky' and all-sky conditions.
-
   It consists of two paragraphs.
@@
 numbersections: true
-
 output: 
-  rticles::elsevier_article:
-    keep_tex:         false
-    citation_package: natbib
   bookdown::pdf_document2:
@@
     biblio-style:     apalike
+  rticles::elsevier_article:
+    keep_tex:         false
+    citation_package: natbib
 # date:           "`r format(Sys.time(), '%F')`"
@@
 
-Before further analysis, we implement an appropriate aggregation scheme on the 1-minute data, to preserve the ....representabiliyt....
-When we use daily mean values we exclude instances with less than `r Daily_aggregation_N_lim` valid data points.
-Accordingly, monthly values are computed from the daily aggregated data, where months with less than `r Monthly_aggegation_N_lim` days are rejected. 
+<!-- aggregation -->
+Before further analysis and Deseasonalization, we implement an appropriate aggregation scheme on the 1-minute data. To preserve the representativeness we use the following criteria:
+a) for daily mean values we exclude instances with less than `r Daily_aggregation_N_lim` valid data points,
+b) accordingly, monthly values are computed by daily aggregated data, where months with less than `r Monthly_aggegation_N_lim` days are rejected. 
 The daily seasonal values are indexed based on the day of year number, and the monthly by the corresponding calendar month.
-
 <!-- Seasonal -->
-For the seasons of the year trends, we also grouped the mean daily values by season (December - February: Winter, March - May: Spring, etc.)
-
+For the seasons of the year trends, we grouped the mean daily values by season (December - February: Winter, March - May: Spring, etc.).
 <!-- Deseasonalization -->
@@
 
-## Aggregation
+## SZA Aggregation
 
@@
 Using the mean daily Global radiation we produce the trends of each data set, as departure from the seasonal value.
-Only for All-sky (Fig. \ref{fig:trendALL}) and Cloudy-sky (Fig. \ref{fig:trendCLOUD}) conditions, the statistical significance of the trends is acceptable (Tab. \ref{tab:trendtable}).
+Only for All-sky (Fig.
+\ref{fig:trendALL}) and Cloudy-sky (Fig.
+\ref{fig:trendCLOUD}) conditions, the statistical significance of the trends are acceptable (Tab.
+\ref{tab:trendtable}), to draw some conclusions.
 
-```{r trendALL, echo=F, fig.cap="Anomaly (\\%) of the daily Global radiation relative to climatological values for 1993 - 2023.
-The straight black line shows the long term trend for All-sky conditions."}
+```{r trendALL, echo=F, fig.cap="Anomaly (\\%) of the daily Global radiation relative to climatological values for 1993 - 2023. The black line shows the long term trend for all-sky conditions."}
 knitr::include_graphics("./images/LongtermTrends-2.pdf")
@@
 
-```{r trendCLOUD, echo=F, fig.cap="Daily Global radiation seasonal anomaly and long term trend, for cloud sky conditions."}
+```{r trendCLOUD, echo=F, fig.cap="Anomaly (\\%) of the daily Global radiation relative to climatological values for 1993 - 2023. The black line shows the long term trend for cloud-sky conditions."}
 knitr::include_graphics("./images/LongtermTrends-6.pdf")
@@
 ```{r trendtable, echo=F}
-pp <- read.csv("./figures/tbl_longterm_trends.dat",
+LT <- data.table(read.csv("./figures/tbl_longterm_trends.dat",
                strip.white = TRUE,
                sep = ";",
-               header = TRUE)
-pp <- pp[grep("Global" ,pp$var), ]
-pp$slope.sd           <- NULL
-pp$slope.ConfInt_0.95 <- NULL
-pp$slope.ConfInt_0.99 <- NULL
-pp$var                <- NULL
-rownames(pp) <- NULL
-names(pp) <- c("Trend %/year", "Trend p-value", "Sky condition", "Trend statistical significance")
+               header = TRUE))
+LT <- LT[grep("Global", LT$var), ]
+LT$slope.sd           <- NULL
+LT$slope.ConfInt_0.95 <- NULL
+LT$slope.ConfInt_0.99 <- NULL
+LT$var                <- NULL
+setorder(LT, -slope.stat_sig)
+rownames(LT) <- NULL
+names(LT) <- c("Trend %/year", "Trend p-value", "Sky condition", "Trend statistical significance")
 
-panderOptions("table.alignment.default", "left")
-pander(pp,
+panderOptions("table.alignment.default", "center")
+pander(LT,
        caption = "\\label{tab:trendtable}Trends of daily means by sky conditions.")
 # \\label{tab:tab2}
-# knitr::kable(pp, caption = "Trends of daily means by sky conditions.")
+# knitr::kable(LT, caption = "Trends of daily means by sky conditions.")
+
+## some vars to use
+alltrend   <- signif(LT[`Sky condition` == "All sky cond.",   `Trend %/year`], 2)
+cleartrend <- signif(LT[`Sky condition` == "Clear sky cond.", `Trend %/year`], 2)
+cloudtrend <- signif(LT[`Sky condition` == "Cloudy cond.",    `Trend %/year`], 2)
 ```
@@
 
-\footnotesize
+<!-- \footnotesize -->
 ```{r trendseasontable, echo=F}
@@
 pander(pp,
-       caption = "\\label{tab:trendseasontable}Trends of daily means by sky conditions for the season of the year.")
+       caption = "\\label{tab:trendseasontable}Trends of daily means by sky conditions for the seasons of the year.")
 ```
-\normalsize
+<!-- \normalsize -->
 
@@
 
-Our result reaffirm the
+Our result for all-sky condition ($`r alltrend`\%/year$), reaffirm the previous results of @Bais2013 for the site.
+The increasing of this trend shows that the phenomenon and probably the causes have been amplified.
+
+.....
+Also, the trend of $`r cloudtrend`\%/year$ for cloud-sky condition indicate the major part that clouds play on the
+.....
+
+.......
 

### 2023-06-09 Article.Rmd
Commit 2023-06-09 17:17

@@
 
-site location description and aerosols effect
+Site location description and aerosols effect
 
@@
 
-Before further analysis, we implement an appropriate aggregation scheme on the 1-minute data, to preserve the ....representabiliyt....
-When we use daily mean values we exclude instances with less than `r Daily_aggregation_N_lim` valid data points.
-Accordingly, monthly values are computed from the daily aggregated data, where months with less than `r Monthly_aggegation_N_lim` days are rejected. 
-The daily seasonal values are indexed based on the day of year number, and the monthly by the corresponding calendar month.
-
+<!-- aggregation -->
+Before further analysis and Deseasonalization, we implement an appropriate aggregation scheme on the 1-minute data. To preserve the representativeness we use the following criteria:
+a) for daily mean values we exclude instances with less than `r Daily_aggregation_N_lim` valid data points,
+b) accordingly, monthly values are computed by daily aggregated data, where months with less than `r Monthly_aggegation_N_lim` days are rejected. 
+The daily seasonal values were indexed based on the day of year number, and the monthly by the corresponding calendar month.
 <!-- Seasonal -->
-For the seasons of the year trends, we also grouped the mean daily values by season (December - February: Winter, March - May: Spring, etc.)
-
+For the seasons of the year trends, we grouped the mean daily values by season (December - February: Winter, March - May: Spring, etc.).
 <!-- Deseasonalization -->

### 2023-06-09 Article.Rmd
Commit 2023-06-09 18:17

@@
 <!-- aggregation -->
-Before further analysis and Deseasonalization, we implement an appropriate aggregation scheme on the 1-minute data. To preserve the representativeness we use the following criteria:
+Before further analysis and deseasonalization, we implement an appropriate aggregation scheme on the 1-minute data. To preserve the representativeness we use the following criteria:
 a) for daily mean values we exclude instances with less than `r Daily_aggregation_N_lim` valid data points,
@@
 rownames(LT) <- NULL
-names(LT) <- c("Trend %/year", "Trend p-value", "Sky condition", "Trend statistical significance")
+names(LT) <- c("Trend [%/year]", "Trend p-value", "Sky condition", "Trend statistical signif. [%]")
 
@@
 ## some vars to use
-alltrend   <- signif(LT[`Sky condition` == "All sky cond.",   `Trend %/year`], 2)
-cleartrend <- signif(LT[`Sky condition` == "Clear sky cond.", `Trend %/year`], 2)
-cloudtrend <- signif(LT[`Sky condition` == "Cloudy cond.",    `Trend %/year`], 2)
+alltrend   <- signif(LT[`Sky condition` == "All sky cond.",   `Trend [%/year]`], 2)
+cleartrend <- signif(LT[`Sky condition` == "Clear sky cond.", `Trend [%/year]`], 2)
+cloudtrend <- signif(LT[`Sky condition` == "Cloudy cond.",    `Trend [%/year]`], 2)
 ```
@@
 rownames(pp) <- NULL
-names(pp) <- c("Trend %/year", "Trend p-value", "Sky condition", "Season", "Trend statistical sign.")
+names(pp) <- c("Trend [%/year]", "Trend p-value", "Sky condition", "Season", "Trend statistical signif. [%]")
 
@@
 Our result for all-sky condition ($`r alltrend`\%/year$), reaffirm the previous results of @Bais2013 for the site.
-The increasing of this trend shows that the phenomenon and probably the causes have been amplified.
+The increase of this trend, shows that the phenomenon and probably the causes have been amplified.
 

### 2023-06-11 Article.Rmd
Commit 2023-06-11 20:17

@@
 output: 
+  rticles::elsevier_article:
+    keep_tex:         false
+    citation_package: natbib
   bookdown::pdf_document2:
@@
     biblio-style:     apalike
-  rticles::elsevier_article:
-    keep_tex:         false
-    citation_package: natbib
 # date:           "`r format(Sys.time(), '%F')`"
@@
 
+```{r , echo=F, fig.cap="Trendns."}
+knitr::include_graphics("./images/SzaTrends-5.pdf")
+```
+
+.........
+
+```{r , echo=F, fig.cap="Trendns."}
+knitr::include_graphics("./images/SzaTrends-13.pdf")
+```
+
+.........
+
+```{r , echo=F, fig.cap="Trendns."}
+knitr::include_graphics("./images/SzaTrends-21.pdf")
+```
+
+```{r, fig.cap='Figure 1', fig.subcap=c('(a)', '(b)', '(c)')}
+plot(1:10)
+plot(cars, pch = 19)
+boxplot(Sepal.Width ~ Species, data = iris)
+```
+
 <!-- \FloatBarrier -->

### 2023-06-11 Article.Rmd
Commit 2023-06-11 21:17

--- 2023-06-09 Article.Rmd
@@
 numbersections:       true
+
+header-includes:
+   - \usepackage{subfig}
+
 output: 
+  rticles::elsevier_article:
+    keep_tex:         false
+    citation_package: natbib
   bookdown::pdf_document2:
@@
 
+<!-- ```{r , echo=F, fig.cap="All sky"} -->
+<!-- knitr::include_graphics("./images/SzaTrends-5.pdf") -->
+<!-- ``` -->
+
+<!--
+```{r , echo=F, fig.cap="Clear sky"}
+knitr::include_graphics("./images/SzaTrends-13.pdf")
+```
+-->
+
+<!-- ```{r , echo=F, fig.cap="Cloud sky"}
+knitr::include_graphics("./images/SzaTrends-21.pdf")
+``` -->
+
+See Figures \ref{fig:szatrends}a and \ref{fig:szatrends}b
+
+```{r szatrends, echo = FALSE, fig.cap='two plots', fig.subcap=c('one plot', 'the other one'), out.width='.35\\linewidth', fig.asp=1, fig.ncol = 2}
+knitr::include_graphics(c("./images/SzaTrends-5.pdf", "./images/SzaTrends-21.pdf"))
+```
+
 <!-- \FloatBarrier -->
@@
 
+<!--
 ```{r cumsumALL, echo=F, fig.cap="Running cumulative sum."}
@@
 
+<!--
 ```{r cumsumCLOUD, echo=F, fig.cap="Donditions."}
 knitr::include_graphics("./images/CumulativeDailyCumSum-9.pdf")
+```
+-->
+
+```{r cumsum, echo = FALSE, fig.cap='two plots', fig.subcap=c('one plot', 'the other one'), out.width='.35\\linewidth', fig.asp=1, fig.ncol = 2}
+knitr::include_graphics(c("./images/CumulativeDailyCumSum-1.pdf", "./images/CumulativeDailyCumSum-9.pdf"))
 ```

### 2023-06-12 Article.Rmd
Commit 2023-06-12 10:17
Commit 2023-06-12 12:17

@@
 knitr::opts_chunk$set(fig.align  = "center")
-knitr::opts_chunk$set(fig.pos    = '!ht'   )
+knitr::opts_chunk$set(fig.pos    = 'h!'    )
 library(pander)
@@
 This is done, by calculating the mean values for the appropriate time step, and subtracting the annual cycle from the actual data.
+<!-- SZA Deseasonalization -->
+In addition to the previous aggregation scheme, each dataset was aggregate in $1^\circ$ SZA bins, separate for cases before and after local noon. 
+This gave us an approximation of Sun's SZA contribution to the SDR brightening.
+
 Thus, for each data set we have obtained the relative departure of the seasonal mean.
-These values are presented in the following section.
-
-........
-
-................
-
-## SZA Aggregation
-
-Both datasets were aggregate in $1^\circ$ SZA bins by am, pm, am+pm
-
-by daily mean
-
-daily sum
-
-seasonal daily mean
-
-filtering of means
-
-statistical significance
-
-Significance and aggregation scheme and filtering will be noted along with the corresponding results.
+The statistical significance and aggregation scheme and filtering will be noted along with the corresponding results.
 
@@
 ```{r szatrends, echo = FALSE, fig.cap='two plots', fig.subcap=c('one plot', 'the other one'), out.width='.35\\linewidth', fig.asp=1, fig.ncol = 2}
-knitr::include_graphics(c("./images/SzaTrends-5.pdf", "./images/SzaTrends-21.pdf"))
+knitr::include_graphics(c("./images/SzaTrends-13.pdf", "./images/SzaTrends-21.pdf"))
 ```

### 2023-06-12 Article_B141.Rmd
B141

--- 2023-06-12 Article.Rmd
@@
 
-## Trends by SZA
+## Long term trends by Solar zenith angle.
 
-.....
-
-.........
+Analyzing the long term trends of SDR to bins of SZA, we can see the contribution of the geometry and time in a diurnal level (Figures \ref{fig:szatrends}a and \ref{fig:szatrends}b).
+Although there is a seasonal distribution of SZA than in not presented here.
 
@@
 
-See Figures \ref{fig:szatrends}a and \ref{fig:szatrends}b
-
-```{r szatrends, echo = FALSE, fig.cap='two plots', fig.subcap=c('one plot', 'the other one'), out.width='.35\\linewidth', fig.asp=1, fig.ncol = 2}
+```{r szatrends, echo = FALSE, fig.cap='Distribution of the SDR\'s long term trends by SZA.', fig.subcap=c('Trend distribution for all-sky conditions.', 'Trend distribution for cloud-sky conditions.'), out.width='.34\\linewidth', fig.asp=1, fig.ncol = 2}
 knitr::include_graphics(c("./images/SzaTrends-13.pdf", "./images/SzaTrends-21.pdf"))
@@
 Similar we have produced the deseasonalized trend for different sky conditions for each season of the year, using the corresponding mean monthly values.
+This can give us a better understanding of the annual variability of the trends.
 
@@
 
+......
+
+NOT INCLUDE?
+
 .....
-
-.........
 
@@
 A method to evaluate changes in the long term trend is to use the cumulative sum of the variable.
+
+Figure \ref{fig:cumsum}
 

### 2023-06-13 Article_B141.Rmd
Commit 2023-06-13 11:17

@@
 output:
+
   rticles::elsevier_article:
@@
     biblio-style:     apalike
+  word_document:      
+    toc:              yes  
+  odt_document:       default
+
 # date:           "`r format(Sys.time(), '%F')`"

### 2023-06-13 Article.Rmd
Commit 2023-06-13 18:17

@@
 
-Using the mean daily Global radiation we produce the trends of each data set, as departure from the seasonal value.
+Using the mean daily SDR we produce the trends of each data set, as departure from the seasonal value.
 Only for All-sky (Fig.
@@
 
-```{r trendALL, echo=F, fig.cap="Anomaly (\\%) of the daily Global radiation relative to climatological values for 1993 - 2023. The black line shows the long term trend for all-sky conditions."}
+```{r trendALL, echo=F, fig.cap="Anomaly (\\%) of the daily SDR relative to climatological values for 1993 - 2023. The black line shows the long term trend for all-sky conditions."}
 knitr::include_graphics("./images/LongtermTrends-2.pdf")
@@
 
-<!-- ```{r trendCLEAR, echo=F, fig.cap="Daily Global radiation seasonal anomaly and long term trend, for clear sky conditions."} -->
+<!-- ```{r trendCLEAR, echo=F, fig.cap="Daily SDR seasonal anomaly and long term trend, for clear sky conditions."} -->
 <!-- knitr::include_graphics("./images/LongtermTrends-4.pdf") -->
@@
 
-```{r trendCLOUD, echo=F, fig.cap="Anomaly (\\%) of the daily Global radiation relative to climatological values for 1993 - 2023. The black line shows the long term trend for cloud-sky conditions."}
+```{r trendCLOUD, echo=F, fig.cap="Anomaly (\\%) of the daily SDR relative to climatological values for 1993 - 2023. The black line shows the long term trend for cloud-sky conditions."}
 knitr::include_graphics("./images/LongtermTrends-6.pdf")

### 2023-06-13 old/DHI_GHI_0_data_input.R
Commit 2023-06-13 18:17

@@ -1,1664 +1,0 @@
-
-
-## Data input
-
-## to force a rebuild
-# file.remove(common_data)
-
-require(data.table)
-require(zoo)
-source("~/CODE/FUNCTIONS/R/trig_deg.R")
-source("~/CODE/FUNCTIONS/R/data.R")
-source("~/MANUSCRIPTS/2022_sdr_trends/DHI_GHI_0_variables.R")
-
-####  Run data construction ####################################################
-
-D_14_2 <- FALSE
-D_14   <- FALSE
-D_13   <- FALSE
-
-D_14_2 <- TRUE
-# D_14   <- TRUE
-# D_13   <- TRUE
-
-TEST <- TRUE
-TEST <- FALSE
-
-if (TEST) {
-    warning("Running in TEST mode!!")
-}
-
-
-## new new implementation with corrected limits
-if (D_14_2) {
-    common_data <- common_data_14_2
-    CS_file     <- CS_file_14_2
-    inpatern    <- "Clear_sky_id_Reno-Hansen_apply_v14_2_[0-9]{4}.Rds"
-}
-
-## new implementation with corrected limits
-if (D_14) {
-    common_data <- common_data_14
-    CS_file     <- CS_file_14
-    inpatern    <- "Clear_sky_id_Reno-Hansen_apply_v14_[0-9]{4}.Rds"
-}
-
-## old implementation with corrected limits
-if (D_13) {
-    common_data <- common_data_13
-    CS_file     <- CS_file_13
-    inpatern    <- "Clear_Sky_[0-9]{4}.Rds"
-}
-
-
-
-## check if we need to run data production
-havetorun <- !file.exists(common_data) |
-    file.mtime(CS_file)          > file.mtime(common_data) |
-    file.mtime(variables_fl)     > file.mtime(common_data) |
-    file.mtime(variables_fl)     > file.mtime(common_data) |
-    file.mtime(data_procsess_fl) > file.mtime(common_data)
-
-if (havetorun) {
-    cat(paste("\n !! (Re)Create environment and data input ->", common_data),"\n")
-
-    #### 0. Get data from Clear sky id data  ###################################
-    input_files <- list.files( path       = CLEARdir,
-                               pattern    = inpatern,
-                               full.names = T )
-    input_files <- grep("_stats_", input_files, value = TRUE, invert = TRUE)
-
-    if (TEST) {
[…]

### 2023-06-14 Article.Rmd
Commit 2023-06-14 09:17

@@
 ```{r seasonalALL, echo=F, fig.cap="Trends by season for All-sky condition. Displays monthly means of daily means."}
-knitr::include_graphics("./images/SeasonaltTrends-1.pdf")
+knitr::include_graphics("./images/SeasonalTrends-1.pdf")
 ```
@@
 ```{r seasonalCLEAR, echo=F, fig.cap="Trends by season for Clear-sky condition. Displays monthly means of daily means."}
-knitr::include_graphics("./images/SeasonaltTrends-2.pdf")
+knitr::include_graphics("./images/SeasonalTrends-2.pdf")
 ```
@@
 ```{r seasonalCLOUD, echo=F, fig.cap="Trends by season for Cloudy-sky condition. Displays monthly means of daily means."}
-knitr::include_graphics("./images/SeasonaltTrends-3.pdf")
+knitr::include_graphics("./images/SeasonalTrends-3.pdf")
 ```

### 2023-06-14 DHI_GHI_1_longterm_trends.R
Commit 2023-06-14 09:17

@@ -88,6 +88,7 @@ source("~/CODE/FUNCTIONS/R/data.R")
 ## __ Source initial scripts ---------------------------------------------------
 source("~/MANUSCRIPTS/2022_sdr_trends/DHI_GHI_0_data_input.R")
 source("~/MANUSCRIPTS/2022_sdr_trends/DHI_GHI_0_variables.R")
+source("~/CODE/R_myRtools/myRtools/R/write_.R")
 tic <- Sys.time()
 
 ## notification function
@@ -342,8 +343,9 @@ pander(pprint,
 #' \normalsize
 #+ echo=F, include=T
 
-myRtools::write_dat(pprint,
-                    "~/MANUSCRIPTS/2022_sdr_trends/figures/tbl_longterm_trends.dat")
+write_dat(pprint,
+          "~/MANUSCRIPTS/2022_sdr_trends/figures/tbl_longterm_trends.dat",
+          clean = TRUE)
 
 
 
@@ -662,7 +664,9 @@ pander(pprint,
        cap = "Slope is in %/year")
 #+ echo=F, include=F
 
-myRtools::write_dat(pprint, "~/MANUSCRIPTS/2022_sdr_trends/figures/tbl_longterm_trends_season.dat")
+write_dat(pprint,
+          "~/MANUSCRIPTS/2022_sdr_trends/figures/tbl_longterm_trends_season.dat",
+          clean = TRUE)
 
 
 
@@ -873,7 +877,9 @@ pander(pprint,
 #' \normalsize
 #+ echo=F, include=T
 
-myRtools::write_dat(pprint, "~/MANUSCRIPTS/2022_sdr_trends/figures/tbl_longterm_trends_monthly.dat")
+write_dat(pprint,
+          "~/MANUSCRIPTS/2022_sdr_trends/figures/tbl_longterm_trends_monthly.dat",
+          clean = TRUE)
 
 
 

### 2023-06-14 Article.Rmd
Commit 2023-06-14 10:17

@@
 numbersections:       true
-
 header-includes:
    - \usepackage{subfig}
-
 output: 
-  rticles::elsevier_article:
-    keep_tex:         false
-    citation_package: natbib
   bookdown::pdf_document2:
@@
 <!-- time span -->
-The SDR is equivalent to the whole sky global radiation (GHI), and was obtained with a horizontal leveled, CM-21 pyranometer.
+The SDR is equivalent to the whole sky Global radiation, also refered as Global Horizontal Irradiance (GHI), and was obtained with a horizontal leveled, CM-21 pyranometer.
 SDR data spans the period of 
@@
                header = TRUE))
-LT <- LT[grep("Global", LT$var), ]
+LT <- LT[grep("SDR", LT$var), ]
 LT$slope.sd           <- NULL
@@
                header = TRUE)
-pp <- pp[grep("Global" ,pp$var), ]
+pp <- pp[grep("SDR" ,pp$var), ]
 pp$slope.sd           <- NULL

### 2023-06-14 Article.Rmd
Commit 2023-06-14 11:17

@@
 
+## Clear sky identification
+
+......
+
+expanded discription
+
+....
+
 <!-- ### Clear sky ID -->
@@
 Only for All-sky (Fig.
-\ref{fig:trendALL}) and Cloudy-sky (Fig.
-\ref{fig:trendCLOUD}) conditions, the statistical significance of the trends are acceptable (Tab.
+\ref{fig:trendALL}) <!--and Cloudy-sky (Fig.
+\ref{fig:trendCLOUD})--> conditions, the statistical significance of the trends are acceptable (Tab.
 \ref{tab:trendtable}), to draw some conclusions.
@@
 
-```{r trendCLOUD, echo=F, fig.cap="Anomaly (\\%) of the daily SDR relative to climatological values for 1993 - 2023. The black line shows the long term trend for cloud-sky conditions."}
-knitr::include_graphics("./images/LongtermTrends-6.pdf")
-```
+<!-- ```{r trendCLOUD, echo=F, fig.cap="Anomaly (\\%) of the daily SDR relative to climatological values for 1993 - 2023. The black line shows the long term trend for cloud-sky conditions."} -->
+<!-- knitr::include_graphics("./images/LongtermTrends-6.pdf") -->
+<!-- ``` -->
 
@@
 
-```{r szatrends, echo = FALSE, fig.cap='Distribution of the SDR\'s long term trends by SZA.', fig.subcap=c('Trend distribution for all-sky conditions.', 'Trend distribution for cloud-sky conditions.'), out.width='.34\\linewidth', fig.asp=1, fig.ncol = 2}
-knitr::include_graphics(c("./images/SzaTrends-13.pdf", "./images/SzaTrends-21.pdf"))
+```{r szatrends, echo = FALSE, fig.cap='Distribution of the SDR\'s long term trends by SZA.', fig.subcap=c('Trend distribution for all-sky conditions.', 'Trend distribution for cloud-sky conditions.'), out.width='.30\\linewidth', fig.asp=1, fig.ncol = 2}
+knitr::include_graphics("./images/SzaTrends-13.pdf")
+cat(" space ")     
+knitr::include_graphics("./images/SzaTrends-21.pdf")
+```
+
+```{r szatrends2, results="asis", echo = FALSE, fig.cap='Distribution of the SDR\'s long term trends by SZA.', fig.subcap=c('Trend distribution for all-sky conditions.', 'Trend distribution for cloud-sky conditions.'), out.width='.30\\linewidth', fig.asp=1, fig.ncol = 2}
+knitr::include_graphics("./images/SzaTrends-13.pdf")
+cat("\\ \n")     
+knitr::include_graphics("./images/SzaTrends-21.pdf")
 ```
@@
 <!-- \FloatBarrier -->
-
+<!-- 
 ## Zenith angle dependency and seasons of the year.
-
-......
-
-NOT INCLUDE?
-
-.....
+-->
 

### 2023-06-14 Article.Rmd
Commit 2023-06-14 12:17

@@
 
-```{r szatrends, echo = FALSE, fig.cap='Distribution of the SDR\'s long term trends by SZA.', fig.subcap=c('Trend distribution for all-sky conditions.', 'Trend distribution for cloud-sky conditions.'), out.width='.30\\linewidth', fig.asp=1, fig.ncol = 2}
+```{r szatrends, echo = FALSE, fig.cap='Distribution of the SDR\'s long term trends by SZA.', fig.subcap=c('Trend distribution for all-sky conditions.', 'Trend distribution for cloud-sky conditions.'), out.width='.33\\linewidth', fig.asp=1, fig.ncol = 2}
 knitr::include_graphics("./images/SzaTrends-13.pdf")
-cat(" space ")     
 knitr::include_graphics("./images/SzaTrends-21.pdf")
@@
 knitr::include_graphics("./images/SzaTrends-13.pdf")
-cat("\\ \n")     
+cat("\\ \\ \n")     
 knitr::include_graphics("./images/SzaTrends-21.pdf")

### 2023-06-16 Article.Rmd
Commit 2023-06-16 13:17

@@
 header-includes:
-   - \usepackage{subfig}
+   # - \usepackage{subfig}
+   - \usepackage{subcaption}
 output: 
+  rticles::elsevier_article:
+    keep_tex:         true
+    citation_package: natbib
   bookdown::pdf_document2:
@@
 
-expanded discription
+expanded description
 
@@
 
-```{r szatrends, echo = FALSE, fig.cap='Distribution of the SDR\'s long term trends by SZA.', fig.subcap=c('Trend distribution for all-sky conditions.', 'Trend distribution for cloud-sky conditions.'), out.width='.33\\linewidth', fig.asp=1, fig.ncol = 2}
+```{r szatrends, echo = FALSE, fig.cap='Distribution of the SDR\'s long term trends by SZA.', fig.subcap=c('Trend distribution for all-sky conditions.', 'Trend distribution for cloud-sky conditions.'), out.width='.35\\linewidth', fig.asp=1, fig.ncol = 2}
 knitr::include_graphics("./images/SzaTrends-13.pdf")
-knitr::include_graphics("./images/SzaTrends-21.pdf")
-```
-
-```{r szatrends2, results="asis", echo = FALSE, fig.cap='Distribution of the SDR\'s long term trends by SZA.', fig.subcap=c('Trend distribution for all-sky conditions.', 'Trend distribution for cloud-sky conditions.'), out.width='.30\\linewidth', fig.asp=1, fig.ncol = 2}
-knitr::include_graphics("./images/SzaTrends-13.pdf")
-cat("\\ \\ \n")     
 knitr::include_graphics("./images/SzaTrends-21.pdf")
@@
 
-```{r seasonalALL, echo=F, fig.cap="Trends by season for All-sky condition. Displays monthly means of daily means."}
-knitr::include_graphics("./images/SeasonalTrends-1.pdf")
+```{r seasonalALL, echo=F, fig.cap="Trends by season for all condition. Displaying monthly means of daily means.", out.width="100%"}
+knitr::include_graphics("./images/SeasonalTrendsTogether-1.pdf")
 ```
 
-```{r seasonalCLEAR, echo=F, fig.cap="Trends by season for Clear-sky condition. Displays monthly means of daily means."}
-knitr::include_graphics("./images/SeasonalTrends-2.pdf")
-```
+<!-- ```{r seasonalALL, echo=F, fig.cap="Trends by season for All-sky condition. Displays monthly means of daily means."} -->
+<!-- knitr::include_graphics("./images/SeasonalTrends-1.pdf") -->
+<!-- ``` -->
 
-```{r seasonalCLOUD, echo=F, fig.cap="Trends by season for Cloudy-sky condition. Displays monthly means of daily means."}
-knitr::include_graphics("./images/SeasonalTrends-3.pdf")
-```
+<!-- ```{r seasonalCLEAR, echo=F, fig.cap="Trends by season for Clear-sky condition. Displays monthly means of daily means."} -->
+<!-- knitr::include_graphics("./images/SeasonalTrends-2.pdf") -->
+<!-- ``` -->
+
+<!-- ```{r seasonalCLOUD, echo=F, fig.cap="Trends by season for Cloudy-sky condition. Displays monthly means of daily means."} -->
+<!-- knitr::include_graphics("./images/SeasonalTrends-3.pdf") -->
+<!-- ``` -->
 

### 2023-06-16 DHI_GHI_1_longterm_trends.R

@@ -496,12 +496,11 @@ dbs         <- c(  "ALL_1_D_bySeason_DESEAS",
 
 Seasons     <- c("Winter", "Spring", "Summer", "Autumn")
 
-
-
 ## variable
 for (avar in vars) {
 
-    par(mfrow = c(length(Seasons), length(dbs)))
+    par(mfcol = c(length(Seasons), length(dbs)))
+    cplots <- 0
 
     ## get global ylim
     gylim <- c()
@@ -517,16 +516,17 @@ for (avar in vars) {
         DB <- get(DBn)
 
         for (ase in Seasons) {
+            cplots <- cplots + 1
 
             dataset <- DB[ Season == ase, ]
 
-            if (sum(!is.na(dataset[[avar]])) <= 1) next()
+            # if (sum(!is.na(dataset[[avar]])) <= 1) next()
 
             ## linear model counting years
-            lm2 <- lm( dataset[[avar]] ~ dataset$Year )
+            lm2 <- lm(dataset[[avar]] ~ dataset$Year)
 
             ## plot
-            par("mar" = c(2, 3.6, 2, 0.5))
+            par("mar" = c(2, 3.6, 2, 0))
 
             plot(dataset$Year, dataset[[avar]],
                  ylim     = ylim,
@@ -546,10 +546,12 @@ for (avar in vars) {
             )
             # axis(2, pretty(dataset[[avar]]), las = 2)
             axis(2, pretty(ylim), las = 2)
-            mtext(text = bquote("Anomaly [%]"),
-                  cex  = 0.8,
-                  side = 2,
-                  line = 2.6)
+            if (cplots <= 4) {
+                mtext(text = bquote("Anomaly [%]"),
+                      cex  = 0.8,
+                      side = 2,
+                      line = 2.6)
+            }
             # ylab = bquote("Deseas." ~ .(translate(avar)) ~ "[" ~ Watt/m^2 ~ "]"))
 
             abline(lm2)

### 2023-06-19 Article_B145.Rmd
Commit 2023-06-19 14:17

--- 2023-06-16 Article.Rmd
@@
 
-expanded description
+<!-- CS_id v14.2 inserted here -->
+```{r setup, echo=F, include=T, eval=TRUE}
+## Load Clear Sky id variables
+MS <- data.frame(readRDS("~/CS_id/PARAMS/Clear_sky_id_Reno-Hansen_apply_v14_2.Rds"))
+# export all variables
+for (an in names(MS)) {
+    assign(an, MS[[an]])
+}
+```
+
+<!-- # Detection of clear periods in GHI measurements for SDR trends. -->
+
+As global radiation clear sky reference we are using the Haurwitz’s model, adjusted for our site with a factor of `r signif(alpha,3)` (Eq. \ref{eq:ahau}).
+The selection of a clear sky reference model, was based on SDR observation from the period 2016 – 2021.
+Where, after an iterative optimization of eight simple models (Daneshyar–Paltridge–Proctor, Kasten–Czeplak, Haurwitz, Berger–Duffie, Adnot–Bourges–Campana–Gicquel, Robledo-Soler, Kasten and Ineichen-Perez) with different factors.
+We found, that Haurwitz’s model, adjusted with a factor of `r signif(alpha,3)` has the lower root mean squared error (RMSE). 
+The tried models are described by @Reno2012 and tested by @Reno2016.
+The iterative optimization method, for the selection of the reference is discussed by @Long2000 and @Reno2016.
+
+\begin{equation}
+\text{SDR}_\text{Clear Sky} = `r signif(alpha,3)` \times 1098 \times \cos( \text{SZA} ) \times \exp \left( \frac{ - 0.057}{\cos(\text{SZA})} \right) \label{eq:ahau}
+\end{equation}
+
+The following criteria and thresholds were used to identify clear-sky conditions.
+Each criterion was applied for a running window of $`r nt`$ consecutive measurements/minutes, and the characterization is applied at the center value of the window.
+A data point is consider as under cloud-sky condition if it fails to pass any of the criteria, all the other data points are characterized as clear-sky.
+
+<!-- 1. MeanVIP -->
+### Mean value of irradiance during the time period.
+
+The mean of the measured value $\overline{G}_i$ must be inside an envelope based on the reference model $\text{SDR}_\text{Clear Sky}$ (Eq. \ref{eq:MeanVIP}).
+
+\begin{equation}
+`r 1 - CS_ref_rm_VIP_RelLow` \times \overline{\text{SDR}}_{i\text{Clear Sky}} - `r CS_ref_rm_VIP_LowOff`
+< \overline{G}_i <
+`r 1 + CS_ref_rm_VIP_RelUpp` \times \overline{\text{SDR}}_{i\text{Clear Sky}} + `r CS_ref_rm_VIP_UppOff`
+\label{eq:MeanVIP}
+\end{equation}
+
+<!-- 2. MaxVIP -->
+### Max value of irradiance during the time period.
+
+The running max measured value $M_{Gi} = max[\text{SDR}_{i}]$, is compared to a similar constructed value from the reference $M_{CSi} = max[\text{SDR}_{i\text{Clear Sky}}]$ (Eq. \ref{eq:MaxVIP}).
+
+\begin{equation}
+`r MaxVIP_fct` \times M_{CSi} - `r MaxVIP_off_low`
+< M_{Gi} <
+`r MaxVIP_fct` \times M_{CSi} + `r MaxVIP_off_upp`
+\label{eq:MaxVIP}
+\end{equation}
+
+<!-- 3. VIL -->
+### Variability in irradiance by line length.
+
+The length $L$ (Eq. \ref{eq:VILeq}) of the sequence of line segments connecting the points of the SDR time
+series for the measured values $L$ and similar for the reference $L_{CS}$, must be within the limits of Eq. \ref{eq:VILcr}.
+
+\begin{equation}
+L = \sum_{i=1}^{n-1}\sqrt{\left ( \text{SDR}_{i+1} - \text{SDR}_{i}\right )^2 + \left ( t_{i+1} - t_i \right )^2}
+\label{eq:VILeq}
+\end{equation}
+
+\begin{equation}
+`r MinVIL_fct` \times L_{CSi} - `r offVIL_dwl` < L_i < `r MaxVIL_fct` \times L_{CSi} + `r offVIL_upl`
+\label{eq:VILcr}
+\end{equation}
+
+<!-- 4. VCT -->
+### Variance of Changes in the Time series.
+
+<!-- Standard deviation of rate of change in irradiance. -->
+We calculate the standard deviation $\sigma$ of the slope
+($s$) between sequential points in the time series, normalized by the average SDR during the time interval.
+
+\begin{gather}
+s_i = \frac{\text{SDR}_{i+1} - \text{SDR}_{i}}{t_{i+1} - t_i}, \forall i \in \left \{ 1, 2, \ldots, n-1 \right \} \label{eq:VCT1} \\
+\bar{s} = \frac{1}{n-1} \sum_{i=1}^{n-1} s_i \label{eq:VCT2} \\
+\sigma_i = \frac {\sqrt{\frac{1}{n-1} \sum_{i=1}^{n-1} \left( s_i - \bar{s} \right)^2} } {\bar{G_i}} \label{eq:VCT3}
+\end{gather}
+
+\begin{equation}
+\sigma_i < `r offVCT`
+\label{eq:VCTcr}
+\end{equation}
+
+<!-- 5. VSM -->
+### Variability in the Shape of the irradiance Measurements.
+
+The maximum difference $X$ between the change in measured irradiance and the change in clear sky
+irradiance over each measurement interval.
+
+\begin{gather}
+x_i = \text{SDR}_{i+1} - \text{SDR}_{i} \forall i \in \left \{ 1, 2, \ldots, n-1 \right \} \label{eq:VSM1} \\
+x_{CS,i} = \text{SDR}_{CS,i+1} - \text{SDR}_{CS,i} \forall i \in \left \{ 1, 2, \ldots, n-1 \right \} \label{eq:VSM2} \\
+X_i = \max{\left \{ \left | x_i - x_{CS,i} \right | \right \}} \label{eq:VSM3}
+\end{gather}
+
+\begin{equation}
+X_i < `r offVSM`
+\label{eq:VSMcr}
+\end{equation}
 

### 2023-06-20 Article.Rmd
Commit 2023-06-17 19:17
Commit 2023-06-20 09:17

@@
 knitr::opts_chunk$set(comment    = ""      )
-knitr::opts_chunk$set(dev        = "pdf"   )
+# knitr::opts_chunk$set(dev        = "pdf"   )
 knitr::opts_chunk$set(out.width  = "70%"   )

### 2023-06-20 Article.Rmd
Commit 2023-06-20 10:17

@@
     biblio-style:     apalike
+  bookdown::word_document2: default
 # date:           "`r format(Sys.time(), '%F')`"
@@
 knitr::opts_chunk$set(fig.pos    = 'h!'    )
+options(knitr.graphics.auto_pdf = FALSE)
 library(pander)
@@
 ```{r trendALL, echo=F, fig.cap="Anomaly (\\%) of the daily SDR relative to climatological values for 1993 - 2023. The black line shows the long term trend for all-sky conditions."}
-knitr::include_graphics("./images/LongtermTrends-2.pdf")
+knitr::include_graphics("./images/LongtermTrends-2.png")
 ```

### 2023-06-20 Article.Rmd
Commit 2023-06-20 13:17

@@
 output: 
-  rticles::elsevier_article:
-    keep_tex:         true
-    citation_package: natbib
+  bookdown::word_document2: default
   bookdown::pdf_document2:
@@
     biblio-style:     apalike
-  bookdown::word_document2: default
+  rticles::elsevier_article:
+    keep_tex:         true
+    citation_package: natbib
 # date:           "`r format(Sys.time(), '%F')`"
@@
 <!-- ```{r trendCLEAR, echo=F, fig.cap="Daily SDR seasonal anomaly and long term trend, for clear sky conditions."} -->
-<!-- knitr::include_graphics("./images/LongtermTrends-4.pdf") -->
+<!-- knitr::include_graphics("./images/LongtermTrends-4.png") -->
 <!-- ``` -->
@@
 <!-- ```{r trendCLOUD, echo=F, fig.cap="Anomaly (\\%) of the daily SDR relative to climatological values for 1993 - 2023. The black line shows the long term trend for cloud-sky conditions."} -->
-<!-- knitr::include_graphics("./images/LongtermTrends-6.pdf") -->
+<!-- knitr::include_graphics("./images/LongtermTrends-6.png") -->
 <!-- ``` -->
@@
 <!-- ```{r , echo=F, fig.cap="All sky"} -->
-<!-- knitr::include_graphics("./images/SzaTrends-5.pdf") -->
+<!-- knitr::include_graphics("./images/SzaTrends-5.png") -->
 <!-- ``` -->
@@
 ```{r , echo=F, fig.cap="Clear sky"}
-knitr::include_graphics("./images/SzaTrends-13.pdf")
+knitr::include_graphics("./images/SzaTrends-13.png")
 ```
@@
 <!-- ```{r , echo=F, fig.cap="Cloud sky"}
-knitr::include_graphics("./images/SzaTrends-21.pdf")
+knitr::include_graphics("./images/SzaTrends-21.png")
 ``` -->
@@
 ```{r szatrends, echo = FALSE, fig.cap='Distribution of the SDR\'s long term trends by SZA.', fig.subcap=c('Trend distribution for all-sky conditions.', 'Trend distribution for cloud-sky conditions.'), out.width='.35\\linewidth', fig.asp=1, fig.ncol = 2}
-knitr::include_graphics("./images/SzaTrends-13.pdf")
-knitr::include_graphics("./images/SzaTrends-21.pdf")
+knitr::include_graphics("./images/SzaTrends-13.png")
+knitr::include_graphics("./images/SzaTrends-21.png")
 ```
@@
 ```{r seasonalALL, echo=F, fig.cap="Trends by season for all condition. Displaying monthly means of daily means.", out.width="100%"}
-knitr::include_graphics("./images/SeasonalTrendsTogether-1.pdf")
+knitr::include_graphics("./images/SeasonalTrendsTogether-1.png")
 ```
@@
 <!-- ```{r seasonalALL, echo=F, fig.cap="Trends by season for All-sky condition. Displays monthly means of daily means."} -->
-<!-- knitr::include_graphics("./images/SeasonalTrends-1.pdf") -->
+<!-- knitr::include_graphics("./images/SeasonalTrends-1.png") -->
 <!-- ``` -->
@@
 <!-- ```{r seasonalCLEAR, echo=F, fig.cap="Trends by season for Clear-sky condition. Displays monthly means of daily means."} -->
-<!-- knitr::include_graphics("./images/SeasonalTrends-2.pdf") -->
+<!-- knitr::include_graphics("./images/SeasonalTrends-2.png") -->
 <!-- ``` -->
@@
 <!-- ```{r seasonalCLOUD, echo=F, fig.cap="Trends by season for Cloudy-sky condition. Displays monthly means of daily means."} -->
-<!-- knitr::include_graphics("./images/SeasonalTrends-3.pdf") -->
+<!-- knitr::include_graphics("./images/SeasonalTrends-3.png") -->
 <!-- ``` -->
@@
 ```{r cumsumALL, echo=F, fig.cap="Running cumulative sum."}
-knitr::include_graphics("./images/CumulativeDailyCumSum-1.pdf")
+knitr::include_graphics("./images/CumulativeDailyCumSum-1.png")
 ```
@@
 ```{r cumsumCLEAR, echo=F, fig.cap="Donditions.", include=FALSE}
-knitr::include_graphics("./images/CumulativeDailyCumSum-5.pdf")
+knitr::include_graphics("./images/CumulativeDailyCumSum-5.png")
 ```
@@
 ```{r cumsumCLOUD, echo=F, fig.cap="Donditions."}
-knitr::include_graphics("./images/CumulativeDailyCumSum-9.pdf")
+knitr::include_graphics("./images/CumulativeDailyCumSum-9.png")
 ```
@@
 ```{r cumsum, echo = FALSE, fig.cap='two plots', fig.subcap=c('one plot', 'the other one'), out.width='.35\\linewidth', fig.asp=1, fig.ncol = 2}
-knitr::include_graphics(c("./images/CumulativeDailyCumSum-1.pdf", "./images/CumulativeDailyCumSum-9.pdf"))
+knitr::include_graphics(c("./images/CumulativeDailyCumSum-1.png", "./images/CumulativeDailyCumSum-9.png"))
 ```

### 2023-06-25 DHI_GHI_2_sza_trends.tex
Commit 2023-06-25 21:17

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  10pt,
  a4paper,oneside]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Trends of SDR in Thessaloniki},
  pdfauthor={Natsis Athanasios; Alkiviadis Bais},
  colorlinks=true,
  linkcolor={Maroon},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[left=0.5in,right=0.5in,top=0.5in,bottom=0.5in]{geometry}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{caption}
\usepackage{placeins}
\captionsetup{font=small}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi

\title{Trends of SDR in Thessaloniki}
\author{Natsis Athanasios\footnote{Laboratory of Atmospheric Physics, AUTH, \href{mailto:natsisphysicist@gmail.com}{\nolinkurl{natsisphysicist@gmail.com}}} \and Alkiviadis Bais\footnote{Laboratory of Atmospheric Physics, AUTH}}
\date{2023-06-25}

\begin{document}
\maketitle
\begin{abstract}
Study of GHI and DNI radiation for `clear sky' and all sly conditions.
\end{abstract}

{
\hypersetup{linkcolor=}
\setcounter{tocdepth}{4}
\tableofcontents
}
\begin{verbatim}
fANCOVA 0.6-1 loaded
\end{verbatim}

\begin{verbatim}
Loading required package: zoo
\end{verbatim}

\begin{verbatim}
Attaching package: 'zoo'
\end{verbatim}

\begin{verbatim}
The following objects are masked from 'package:base':

    as.Date, as.Date.numeric
\end{verbatim}

\begin{verbatim}

Load environment and data from:  ~/MANUSCRIPTS/2022_sdr_trends/data/common_data_14_2.Rda 
\end{verbatim}

\hypertarget{long-term-by-sza}{%
\subsection{2. Long term by SZA}\label{long-term-by-sza}}

\hypertarget{data-info}{%
\subsubsection{Data info}\label{data-info}}

Time data span 1993-04-12, 2023-05-31

\hypertarget{process}{%
\subsubsection{Process}\label{process}}

\FloatBarrier

\hypertarget{plot-of-sza-trends}{%
\subsection{Plot of SZA trends}\label{plot-of-sza-trends}}

\newpage

\hypertarget{all-sky-cond.-dir.-beam-irrad.}{%
\paragraph{All sky cond. Dir. Beam Irrad.}\label{all-sky-cond.-dir.-beam-irrad.}}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrends-1} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrends-2} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrends-3} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrends-4} \end{center}

\newpage

\hypertarget{clear-sky-cond.-dir.-beam-irrad.}{%
\paragraph{Clear sky cond. Dir. Beam Irrad.}\label{clear-sky-cond.-dir.-beam-irrad.}}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrends-5} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrends-6} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrends-7} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrends-8} \end{center}

\newpage

\hypertarget{cloudy-cond.-dir.-beam-irrad.}{%
\paragraph{Cloudy cond. Dir. Beam Irrad.}\label{cloudy-cond.-dir.-beam-irrad.}}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrends-9} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrends-10} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrends-11} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrends-12} \end{center}

\newpage

\hypertarget{all-sky-cond.-sdr}{%
\paragraph{All sky cond. SDR}\label{all-sky-cond.-sdr}}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrends-13} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrends-14} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrends-15} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrends-16} \end{center}

\newpage

\hypertarget{clear-sky-cond.-sdr}{%
\paragraph{Clear sky cond. SDR}\label{clear-sky-cond.-sdr}}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrends-17} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrends-18} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrends-19} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrends-20} \end{center}

\newpage

\hypertarget{cloudy-cond.-sdr}{%
\paragraph{Cloudy cond. SDR}\label{cloudy-cond.-sdr}}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrends-21} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrends-22} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrends-23} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrends-24} \end{center}

\hypertarget{plot-of-sza-trends-for-each-season-of-year}{%
\subsection{Plot of SZA trends for each season of year}\label{plot-of-sza-trends-for-each-season-of-year}}

\newpage

\hypertarget{winter-all-sky-cond.-dir.-beam-irrad.}{%
\subsubsection{Winter All sky cond. Dir. Beam Irrad.}\label{winter-all-sky-cond.-dir.-beam-irrad.}}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-1} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-2} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-3} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-4} \end{center}

\newpage

\hypertarget{winter-all-sky-cond.-sdr}{%
\subsubsection{Winter All sky cond. SDR}\label{winter-all-sky-cond.-sdr}}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-5} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-6} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-7} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-8} \end{center}

\newpage

\hypertarget{winter-clear-sky-cond.-dir.-beam-irrad.}{%
\subsubsection{Winter Clear sky cond. Dir. Beam Irrad.}\label{winter-clear-sky-cond.-dir.-beam-irrad.}}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-9} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-10} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-11} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-12} \end{center}

\newpage

\hypertarget{winter-clear-sky-cond.-sdr}{%
\subsubsection{Winter Clear sky cond. SDR}\label{winter-clear-sky-cond.-sdr}}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-13} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-14} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-15} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-16} \end{center}

\newpage

\hypertarget{winter-cloudy-cond.-dir.-beam-irrad.}{%
\subsubsection{Winter Cloudy cond. Dir. Beam Irrad.}\label{winter-cloudy-cond.-dir.-beam-irrad.}}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-17} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-18} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-19} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-20} \end{center}

\newpage

\hypertarget{winter-cloudy-cond.-sdr}{%
\subsubsection{Winter Cloudy cond. SDR}\label{winter-cloudy-cond.-sdr}}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-21} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-22} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-23} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-24} \end{center}

\newpage

\hypertarget{spring-all-sky-cond.-dir.-beam-irrad.}{%
\subsubsection{Spring All sky cond. Dir. Beam Irrad.}\label{spring-all-sky-cond.-dir.-beam-irrad.}}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-25} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-26} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-27} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-28} \end{center}

\newpage

\hypertarget{spring-all-sky-cond.-sdr}{%
\subsubsection{Spring All sky cond. SDR}\label{spring-all-sky-cond.-sdr}}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-29} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-30} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-31} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-32} \end{center}

\newpage

\hypertarget{spring-clear-sky-cond.-dir.-beam-irrad.}{%
\subsubsection{Spring Clear sky cond. Dir. Beam Irrad.}\label{spring-clear-sky-cond.-dir.-beam-irrad.}}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-33} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-34} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-35} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-36} \end{center}

\newpage

\hypertarget{spring-clear-sky-cond.-sdr}{%
\subsubsection{Spring Clear sky cond. SDR}\label{spring-clear-sky-cond.-sdr}}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-37} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-38} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-39} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-40} \end{center}

\newpage

\hypertarget{spring-cloudy-cond.-dir.-beam-irrad.}{%
\subsubsection{Spring Cloudy cond. Dir. Beam Irrad.}\label{spring-cloudy-cond.-dir.-beam-irrad.}}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-41} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-42} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-43} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-44} \end{center}

\newpage

\hypertarget{spring-cloudy-cond.-sdr}{%
\subsubsection{Spring Cloudy cond. SDR}\label{spring-cloudy-cond.-sdr}}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-45} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-46} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-47} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-48} \end{center}

\newpage

\hypertarget{summer-all-sky-cond.-dir.-beam-irrad.}{%
\subsubsection{Summer All sky cond. Dir. Beam Irrad.}\label{summer-all-sky-cond.-dir.-beam-irrad.}}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-49} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-50} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-51} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-52} \end{center}

\newpage

\hypertarget{summer-all-sky-cond.-sdr}{%
\subsubsection{Summer All sky cond. SDR}\label{summer-all-sky-cond.-sdr}}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-53} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-54} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-55} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-56} \end{center}

\newpage

\hypertarget{summer-clear-sky-cond.-dir.-beam-irrad.}{%
\subsubsection{Summer Clear sky cond. Dir. Beam Irrad.}\label{summer-clear-sky-cond.-dir.-beam-irrad.}}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-57} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-58} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-59} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-60} \end{center}

\newpage

\hypertarget{summer-clear-sky-cond.-sdr}{%
\subsubsection{Summer Clear sky cond. SDR}\label{summer-clear-sky-cond.-sdr}}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-61} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-62} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-63} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-64} \end{center}

\newpage

\hypertarget{summer-cloudy-cond.-dir.-beam-irrad.}{%
\subsubsection{Summer Cloudy cond. Dir. Beam Irrad.}\label{summer-cloudy-cond.-dir.-beam-irrad.}}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-65} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-66} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-67} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-68} \end{center}

\newpage

\hypertarget{summer-cloudy-cond.-sdr}{%
\subsubsection{Summer Cloudy cond. SDR}\label{summer-cloudy-cond.-sdr}}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-69} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-70} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-71} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-72} \end{center}

\newpage

\hypertarget{autumn-all-sky-cond.-dir.-beam-irrad.}{%
\subsubsection{Autumn All sky cond. Dir. Beam Irrad.}\label{autumn-all-sky-cond.-dir.-beam-irrad.}}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-73} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-74} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-75} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-76} \end{center}

\newpage

\hypertarget{autumn-all-sky-cond.-sdr}{%
\subsubsection{Autumn All sky cond. SDR}\label{autumn-all-sky-cond.-sdr}}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-77} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-78} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-79} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-80} \end{center}

\newpage

\hypertarget{autumn-clear-sky-cond.-dir.-beam-irrad.}{%
\subsubsection{Autumn Clear sky cond. Dir. Beam Irrad.}\label{autumn-clear-sky-cond.-dir.-beam-irrad.}}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-81} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-82} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-83} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-84} \end{center}

\newpage

\hypertarget{autumn-clear-sky-cond.-sdr}{%
\subsubsection{Autumn Clear sky cond. SDR}\label{autumn-clear-sky-cond.-sdr}}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-85} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-86} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-87} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-88} \end{center}

\newpage

\hypertarget{autumn-cloudy-cond.-dir.-beam-irrad.}{%
\subsubsection{Autumn Cloudy cond. Dir. Beam Irrad.}\label{autumn-cloudy-cond.-dir.-beam-irrad.}}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-89} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-90} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-91} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-92} \end{center}

\newpage

\hypertarget{autumn-cloudy-cond.-sdr}{%
\subsubsection{Autumn Cloudy cond. SDR}\label{autumn-cloudy-cond.-sdr}}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-93} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-94} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-95} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-96} \end{center}

\textbf{END}

\begin{verbatim}
2023-06-25 17:38:40.8 athan@tyler DHI_GHI_1_longterm_trends.R 1.213719 mins
\end{verbatim}

\end{document}

### 2023-06-25 DHI_GHI_3_trends_consistency.tex
Commit 2023-06-25 21:17

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  10pt,
  a4paper,oneside]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Trends of SDR in Thessaloniki},
  pdfauthor={Natsis Athanasios; Alkiviadis Bais},
  colorlinks=true,
  linkcolor={Maroon},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage[left=0.5in,right=0.5in,top=0.5in,bottom=0.5in]{geometry}
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{caption}
\usepackage{placeins}
\captionsetup{font=small}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi

\title{Trends of SDR in Thessaloniki}
\author{Natsis Athanasios\footnote{Laboratory of Atmospheric Physics, AUTH, \href{mailto:natsisphysicist@gmail.com}{\nolinkurl{natsisphysicist@gmail.com}}} \and Alkiviadis Bais\footnote{Laboratory of Atmospheric Physics, AUTH}}
\date{2023-06-25}

\begin{document}
\maketitle
\begin{abstract}
Study of GHI and DNI radiation for `clear sky' and all sly conditions.
\end{abstract}

{
\hypersetup{linkcolor=}
\setcounter{tocdepth}{4}
\tableofcontents
}
\begin{verbatim}
fANCOVA 0.6-1 loaded
\end{verbatim}

\begin{verbatim}
Loading required package: zoo
\end{verbatim}

\begin{verbatim}
Attaching package: 'zoo'
\end{verbatim}

\begin{verbatim}
The following objects are masked from 'package:base':

    as.Date, as.Date.numeric
\end{verbatim}

\begin{verbatim}

Load environment and data from:  ~/MANUSCRIPTS/2022_sdr_trends/data/common_data_14_2.Rda 
\end{verbatim}

\hypertarget{consistency-of-trends}{%
\subsection{3. Consistency of trends}\label{consistency-of-trends}}

\hypertarget{data-info}{%
\subsubsection{Data info}\label{data-info}}

Time data span 1993-04-12, 2023-05-31

Where is a \textbf{running mean the window is 1826.2125 days} or
5 years.

\hypertarget{process}{%
\subsubsection{Process}\label{process}}

\newpage
\FloatBarrier

\hypertarget{whole-day-daily-cumulative-sum}{%
\subsubsection{Whole day daily cumulative sum}\label{whole-day-daily-cumulative-sum}}

All data in this section are referring to daily means.

\newpage

\hypertarget{daily-cum-sums-for-all-sky-cond.}{%
\paragraph{Daily cum sums for All sky cond.}\label{daily-cum-sums-for-all-sky-cond.}}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeDailyCumSum-1} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeDailyCumSum-2} \end{center}

\newpage

\footnotesize

\begin{longtable}[]{@{}
  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1410}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1795}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1795}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1923}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1923}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1154}}@{}}
\toprule
\begin{minipage}[b]{\linewidth}\centering
~
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
PositiveSum
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
NegativeSum
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
PositiveMean
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
NegativeMean
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Sum
\end{minipage} \\
\midrule
\endhead
\textbf{1993} & 2232 & -3205 & 16.06 & -31.43 & -973.5 \\
\textbf{1994} & 4452 & -5575 & 22.83 & -35.29 & -1123 \\
\textbf{1995} & 3407 & -5415 & 21.29 & -30.6 & -2009 \\
\textbf{1996} & 2788 & -7750 & 16.79 & -41.44 & -4962 \\
\textbf{1997} & 4470 & -6190 & 22.81 & -39.17 & -1719 \\
\textbf{1998} & 4367 & -6098 & 22.06 & -40.12 & -1731 \\
\textbf{1999} & 4700 & -5052 & 22.49 & -33.02 & -352.4 \\
\textbf{2000} & 5692 & -3631 & 23.23 & -32.71 & 2061 \\
\textbf{2001} & 4570 & -5386 & 23.08 & -36.89 & -816 \\
\textbf{2002} & 4535 & -5942 & 23.62 & -36.01 & -1407 \\
\textbf{2003} & 4989 & -5994 & 24.46 & -38.67 & -1004 \\
\textbf{2004} & 4499 & -5379 & 21.63 & -36.35 & -880 \\
\textbf{2005} & 4613 & -3980 & 23.66 & -29.26 & 633.2 \\
\textbf{2006} & 4764 & -4385 & 23.58 & -32.97 & 378.7 \\
\textbf{2007} & 4446 & -4126 & 21.9 & -36.19 & 319.5 \\
\textbf{2008} & 5096 & -4847 & 22.06 & -38.16 & 249 \\
\textbf{2009} & 5253 & -5057 & 26.27 & -35.61 & 196.8 \\
\textbf{2010} & 3749 & -4362 & 23.14 & -36.35 & -613.4 \\
\textbf{2011} & 4780 & -4629 & 22.44 & -38.26 & 150.5 \\
\textbf{2012} & 4944 & -4277 & 21.5 & -40.35 & 667.3 \\
\textbf{2013} & 5326 & -4362 & 23.88 & -34.08 & 964.2 \\
\textbf{2014} & 4270 & -5710 & 21.57 & -38.07 & -1440 \\
\textbf{2015} & 5612 & -4511 & 27.11 & -36.09 & 1101 \\
\textbf{2016} & 6263 & -4623 & 27.23 & -35.03 & 1640 \\
\textbf{2017} & 6655 & -4315 & 26.94 & -37.85 & 2340 \\
\textbf{2018} & 5163 & -5085 & 24.59 & -34.83 & 78.02 \\
\textbf{2019} & 5472 & -3791 & 23.39 & -32.97 & 1681 \\
\textbf{2020} & 5989 & -3791 & 25.49 & -33.84 & 2199 \\
\textbf{2021} & 5750 & -4535 & 24.47 & -35.16 & 1215 \\
\textbf{2022} & 6231 & -4079 & 25.33 & -35.47 & 2151 \\
\textbf{2023} & 2958 & -1952 & 32.51 & -34.85 & 1006 \\
\bottomrule
\end{longtable}

\normalsize

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeDailyCumSum-3} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeDailyCumSum-4} \end{center}

\newpage

\hypertarget{daily-cum-sums-for-clear-sky-cond.}{%
\paragraph{Daily cum sums for Clear sky cond.}\label{daily-cum-sums-for-clear-sky-cond.}}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeDailyCumSum-5} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeDailyCumSum-6} \end{center}

\newpage

\footnotesize

\begin{longtable}[]{@{}
  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1410}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1795}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1795}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1923}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1923}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1154}}@{}}
\toprule
\begin{minipage}[b]{\linewidth}\centering
~
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
PositiveSum
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
NegativeSum
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
PositiveMean
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
NegativeMean
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Sum
\end{minipage} \\
\midrule
\endhead
\textbf{1993} & 2814 & -3584 & 21.48 & -33.19 & -770.5 \\
\textbf{1994} & 5696 & -5936 & 28.77 & -38.8 & -240 \\
\textbf{1995} & 4001 & -6110 & 27.98 & -31.66 & -2109 \\
\textbf{1996} & 3804 & -8259 & 25.36 & -40.69 & -4456 \\
\textbf{1997} & 4762 & -6823 & 27.06 & -38.33 & -2060 \\
\textbf{1998} & 4678 & -6597 & 25.85 & -39.04 & -1919 \\
\textbf{1999} & 5594 & -5572 & 28.11 & -34.19 & 21.55 \\
\textbf{2000} & 5954 & -4077 & 26.35 & -31.36 & 1877 \\
\textbf{2001} & 4426 & -5877 & 24.73 & -36.5 & -1451 \\
\textbf{2002} & 5201 & -6534 & 30.06 & -35.9 & -1333 \\
\textbf{2003} & 5346 & -6573 & 27.84 & -39.36 & -1228 \\
\textbf{2004} & 4671 & -5968 & 26.54 & -33.16 & -1297 \\
\textbf{2005} & 5624 & -4485 & 30.9 & -30.31 & 1139 \\
\textbf{2006} & 5677 & -4802 & 30.36 & -32.67 & 875 \\
\textbf{2007} & 5241 & -4427 & 27.3 & -35.42 & 813.7 \\
\textbf{2008} & 5997 & -5223 & 28.29 & -36.02 & 773.4 \\
\textbf{2009} & 5527 & -5536 & 29.87 & -35.26 & -9.141 \\
\textbf{2010} & 3575 & -4751 & 24.66 & -34.68 & -1176 \\
\textbf{2011} & 4797 & -5050 & 24.98 & -35.81 & -253.1 \\
\textbf{2012} & 5341 & -4680 & 25.8 & -36.57 & 660.7 \\
\textbf{2013} & 5225 & -4814 & 25.49 & -32.97 & 410.7 \\
\textbf{2014} & 4806 & -6218 & 25.98 & -38.15 & -1412 \\
\textbf{2015} & 5698 & -4929 & 29.22 & -35.98 & 768.7 \\
\textbf{2016} & 6797 & -5042 & 30.75 & -36.01 & 1755 \\
\textbf{2017} & 6495 & -4752 & 29.12 & -34.44 & 1743 \\
\textbf{2018} & 5927 & -5644 & 31.03 & -34.21 & 283.1 \\
\textbf{2019} & 5870 & -4292 & 28.08 & -30.88 & 1578 \\
\textbf{2020} & 6083 & -4301 & 28.83 & -31.63 & 1781 \\
\textbf{2021} & 6692 & -4976 & 29.88 & -35.54 & 1716 \\
\textbf{2022} & 6803 & -4447 & 28.83 & -35.57 & 2356 \\
\textbf{2023} & 3233 & -2072 & 36.33 & -35.72 & 1161 \\
\bottomrule
\end{longtable}

\normalsize

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeDailyCumSum-7} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeDailyCumSum-8} \end{center}

\newpage

\hypertarget{daily-cum-sums-for-cloudy-cond.}{%
\paragraph{Daily cum sums for Cloudy cond.}\label{daily-cum-sums-for-cloudy-cond.}}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeDailyCumSum-9} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeDailyCumSum-10} \end{center}

\newpage

\footnotesize

\begin{longtable}[]{@{}
  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1410}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1795}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1795}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1923}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1923}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1154}}@{}}
\toprule
\begin{minipage}[b]{\linewidth}\centering
~
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
PositiveSum
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
NegativeSum
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
PositiveMean
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
NegativeMean
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Sum
\end{minipage} \\
\midrule
\endhead
\textbf{1993} & 555.1 & -1190 & 20.56 & -27.05 & -635.2 \\
\textbf{1994} & 661.9 & -1754 & 18.39 & -24.7 & -1092 \\
\textbf{1995} & 640.4 & -1042 & 18.3 & -21.7 & -401.2 \\
\textbf{1996} & 873 & -1718 & 26.46 & -28.17 & -845.2 \\
\textbf{1997} & 1272 & -1106 & 23.56 & -26.98 & 166.2 \\
\textbf{1998} & 1049 & -992.1 & 20.58 & -20.67 & 57.34 \\
\textbf{1999} & 714.5 & -1368 & 17.43 & -20.73 & -653.8 \\
\textbf{2000} & 972.2 & -1160 & 23.15 & -22.74 & -187.7 \\
\textbf{2001} & 1005 & -655.4 & 24.5 & -18.73 & 349.2 \\
\textbf{2002} & 1166 & -1155 & 21.2 & -23.56 & 11.55 \\
\textbf{2003} & 1045 & -1142 & 24.31 & -22.83 & -96.28 \\
\textbf{2004} & 1356 & -646.1 & 27.68 & -19.58 & 710.1 \\
\textbf{2005} & 825.7 & -1101 & 20.14 & -20.78 & -275.8 \\
\textbf{2006} & 941.4 & -1006 & 21.89 & -21.86 & -64.25 \\
\textbf{2007} & 654.3 & -1183 & 21.81 & -25.71 & -528.3 \\
\textbf{2008} & 845 & -1394 & 22.84 & -21.78 & -549 \\
\textbf{2009} & 839.7 & -755.1 & 22.69 & -22.21 & 84.59 \\
\textbf{2010} & 792.5 & -529.7 & 23.31 & -21.19 & 262.8 \\
\textbf{2011} & 1305 & -775.3 & 26.1 & -23.49 & 529.9 \\
\textbf{2012} & 1132 & -1052 & 25.15 & -25.05 & 79.83 \\
\textbf{2013} & 1382 & -776.3 & 27.1 & -18.05 & 605.7 \\
\textbf{2014} & 856.1 & -815.9 & 23.14 & -19.9 & 40.2 \\
\textbf{2015} & 949 & -515.4 & 21.09 & -20.61 & 433.7 \\
\textbf{2016} & 1287 & -1040 & 24.76 & -22.12 & 247.8 \\
\textbf{2017} & 971.9 & -737.2 & 22.6 & -21.06 & 234.7 \\
\textbf{2018} & 1244 & -867.5 & 24.4 & -23.44 & 376.8 \\
\textbf{2019} & 928.5 & -866.3 & 23.81 & -21.13 & 62.18 \\
\textbf{2020} & 1196 & -579.6 & 22.15 & -17.56 & 616.7 \\
\textbf{2021} & 1048 & -844.3 & 16.64 & -16.56 & 203.8 \\
\textbf{2022} & 1112 & -784.9 & 21.8 & -19.14 & 326.9 \\
\textbf{2023} & 131.9 & -203 & 10.99 & -20.3 & -71.12 \\
\bottomrule
\end{longtable}

\normalsize

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeDailyCumSum-11} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeDailyCumSum-12} \end{center}

\newpage
\FloatBarrier

\hypertarget{whole-day-monthly-cumulative-sum}{%
\subsubsection{Whole day monthly cumulative sum}\label{whole-day-monthly-cumulative-sum}}

In this section the monthly values have been calculated from daily mean values.

\newpage

\hypertarget{monthly-cum-sums-for-all-sky-cond.}{%
\paragraph{Monthly cum sums for All sky cond.}\label{monthly-cum-sums-for-all-sky-cond.}}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeMonthlyCumSum-1} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeMonthlyCumSum-2} \end{center}

\newpage

\footnotesize

\begin{longtable}[]{@{}
  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1410}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1795}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1795}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1923}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1923}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1154}}@{}}
\toprule
\begin{minipage}[b]{\linewidth}\centering
~
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
PositiveSum
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
NegativeSum
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
PositiveMean
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
NegativeMean
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Sum
\end{minipage} \\
\midrule
\endhead
\textbf{1993} & 19.08 & -48.52 & 6.36 & -9.704 & -29.44 \\
\textbf{1994} & 26.02 & -74.45 & 5.204 & -10.64 & -48.42 \\
\textbf{1995} & 26.9 & -93 & 5.38 & -15.5 & -66.1 \\
\textbf{1996} & 3.185 & -182.2 & 1.592 & -18.22 & -179 \\
\textbf{1997} & 26.52 & -81.3 & 8.84 & -9.033 & -54.78 \\
\textbf{1998} & 27.93 & -76.09 & 5.586 & -10.87 & -48.16 \\
\textbf{1999} & 7.774 & -24.16 & 1.943 & -3.02 & -16.39 \\
\textbf{2000} & 77.82 & -18.31 & 8.646 & -6.102 & 59.51 \\
\textbf{2001} & 49.51 & -76.54 & 9.901 & -10.93 & -27.03 \\
\textbf{2002} & 38.91 & -84.7 & 12.97 & -9.411 & -45.78 \\
\textbf{2003} & 29.84 & -56.12 & 4.974 & -9.354 & -26.28 \\
\textbf{2004} & 28.81 & -57.29 & 4.802 & -9.548 & -28.48 \\
\textbf{2005} & 33.86 & -14.68 & 4.837 & -3.669 & 19.19 \\
\textbf{2006} & 37.15 & -35.07 & 7.43 & -5.844 & 2.085 \\
\textbf{2007} & 24.02 & -15.11 & 4.805 & -3.022 & 8.915 \\
\textbf{2008} & 50.75 & -31.9 & 7.25 & -6.38 & 18.85 \\
\textbf{2009} & 56.6 & -55.74 & 8.086 & -11.15 & 0.8636 \\
\textbf{2010} & 24.77 & -38.35 & 6.191 & -9.587 & -13.58 \\
\textbf{2011} & 24.73 & -28.67 & 4.122 & -5.733 & -3.935 \\
\textbf{2012} & 45 & -23.46 & 5.624 & -7.819 & 21.54 \\
\textbf{2013} & 64.93 & -36.73 & 8.116 & -9.184 & 28.2 \\
\textbf{2014} & 20.71 & -64.85 & 4.141 & -9.264 & -44.14 \\
\textbf{2015} & 92.01 & -45.21 & 13.14 & -11.3 & 46.8 \\
\textbf{2016} & 73.43 & -18.02 & 8.159 & -6.007 & 55.41 \\
\textbf{2017} & 89.19 & -11.66 & 9.91 & -3.886 & 77.53 \\
\textbf{2018} & 51.06 & -44.16 & 7.294 & -8.832 & 6.901 \\
\textbf{2019} & 61.82 & -17.06 & 8.831 & -3.413 & 44.75 \\
\textbf{2020} & 97.56 & -20.95 & 10.84 & -6.983 & 76.61 \\
\textbf{2021} & 60.92 & -21.37 & 8.703 & -4.274 & 39.55 \\
\textbf{2022} & 94.15 & -18.65 & 11.77 & -4.662 & 75.5 \\
\textbf{2023} & 55.86 & -21.35 & 13.96 & -21.35 & 34.51 \\
\bottomrule
\end{longtable}

\normalsize

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeMonthlyCumSum-3} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeMonthlyCumSum-4} \end{center}

\newpage

\hypertarget{monthly-cum-sums-for-clear-sky-cond.}{%
\paragraph{Monthly cum sums for Clear sky cond.}\label{monthly-cum-sums-for-clear-sky-cond.}}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeMonthlyCumSum-5} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeMonthlyCumSum-6} \end{center}

\newpage

\footnotesize

\begin{longtable}[]{@{}
  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1410}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1795}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1795}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1923}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1923}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1154}}@{}}
\toprule
\begin{minipage}[b]{\linewidth}\centering
~
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
PositiveSum
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
NegativeSum
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
PositiveMean
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
NegativeMean
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Sum
\end{minipage} \\
\midrule
\endhead
\textbf{1993} & 23 & -43.35 & 7.667 & -8.67 & -20.35 \\
\textbf{1994} & 44.44 & -62.1 & 7.406 & -10.35 & -17.67 \\
\textbf{1995} & 32.54 & -101.2 & 5.424 & -16.86 & -68.62 \\
\textbf{1996} & 12.47 & -175.1 & 2.495 & -25.01 & -162.6 \\
\textbf{1997} & 25.04 & -91.64 & 8.348 & -10.18 & -66.6 \\
\textbf{1998} & 25.99 & -80.59 & 5.197 & -11.51 & -54.61 \\
\textbf{1999} & 18.37 & -24 & 4.592 & -3.001 & -5.637 \\
\textbf{2000} & 70.9 & -18.61 & 8.862 & -4.653 & 52.28 \\
\textbf{2001} & 43.97 & -96.18 & 7.328 & -16.03 & -52.21 \\
\textbf{2002} & 39.75 & -83.65 & 9.937 & -10.46 & -43.9 \\
\textbf{2003} & 23.96 & -58.03 & 5.989 & -7.254 & -34.07 \\
\textbf{2004} & 23.1 & -64.67 & 7.7 & -7.186 & -41.57 \\
\textbf{2005} & 72.61 & -8.362 & 8.068 & -2.787 & 64.25 \\
\textbf{2006} & 66.38 & -30.4 & 9.483 & -6.08 & 35.98 \\
\textbf{2007} & 38.48 & -24.91 & 5.497 & -6.227 & 13.57 \\
\textbf{2008} & 69.96 & -32.22 & 8.744 & -8.054 & 37.74 \\
\textbf{2009} & 49.6 & -58.44 & 8.267 & -9.741 & -8.844 \\
\textbf{2010} & 17.47 & -62.67 & 3.495 & -10.45 & -45.2 \\
\textbf{2011} & 21.77 & -41.61 & 5.443 & -5.944 & -19.84 \\
\textbf{2012} & 44.68 & -23.08 & 6.383 & -5.769 & 21.61 \\
\textbf{2013} & 58.06 & -48.54 & 8.295 & -9.708 & 9.523 \\
\textbf{2014} & 17.42 & -60.43 & 3.484 & -8.632 & -43.01 \\
\textbf{2015} & 88.25 & -47.45 & 14.71 & -7.909 & 40.8 \\
\textbf{2016} & 87.36 & -24.92 & 8.736 & -12.46 & 62.44 \\
\textbf{2017} & 80.13 & -22.2 & 11.45 & -4.44 & 57.93 \\
\textbf{2018} & 60.39 & -47.65 & 10.07 & -7.942 & 12.74 \\
\textbf{2019} & 60.79 & -20.92 & 10.13 & -3.487 & 39.87 \\
\textbf{2020} & 94.67 & -32.11 & 15.78 & -5.352 & 62.56 \\
\textbf{2021} & 78.64 & -21.81 & 9.83 & -5.452 & 56.83 \\
\textbf{2022} & 105.8 & -22.44 & 13.22 & -5.609 & 83.35 \\
\textbf{2023} & 59.84 & -21.19 & 14.96 & -21.19 & 38.66 \\
\bottomrule
\end{longtable}

\normalsize

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeMonthlyCumSum-7} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeMonthlyCumSum-8} \end{center}

\newpage

\hypertarget{monthly-cum-sums-for-cloudy-cond.}{%
\paragraph{Monthly cum sums for Cloudy cond.}\label{monthly-cum-sums-for-cloudy-cond.}}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeMonthlyCumSum-9} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeMonthlyCumSum-10} \end{center}

\newpage

\footnotesize

\begin{longtable}[]{@{}
  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1410}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1795}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1795}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1923}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1923}}
  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1154}}@{}}
\toprule
\begin{minipage}[b]{\linewidth}\centering
~
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
PositiveSum
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
NegativeSum
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
PositiveMean
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
NegativeMean
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedleft
Sum
\end{minipage} \\
\midrule
\endhead
\textbf{1996} & 1.041 & -32.43 & 1.041 & -16.21 & -31.39 \\
\textbf{1997} & 30.94 & -13.28 & 15.47 & -6.639 & 17.66 \\
\textbf{1998} & 22.21 & -11.6 & 7.405 & -5.802 & 10.61 \\
\textbf{2000} & 5.152 & -11.35 & 5.152 & -11.35 & -6.197 \\
\textbf{2001} & 17.63 & -11.37 & 17.63 & -11.37 & 6.265 \\
\textbf{2002} & 35.68 & -15.07 & 17.84 & -7.534 & 20.61 \\
\textbf{2003} & 31.3 & -26.37 & 15.65 & -13.19 & 4.926 \\
\textbf{2005} & 0.5237 & -13.4 & 0.5237 & -4.466 & -12.87 \\
\textbf{2006} & 20.63 & -22.47 & 6.878 & -11.24 & -1.839 \\
\textbf{2008} & 1.037 & -22.67 & 0.5185 & -7.558 & -21.64 \\
\textbf{2011} & 9.071 & -7.749 & 4.535 & -7.749 & 1.322 \\
\textbf{2012} & 18.89 & -26.75 & 9.443 & -26.75 & -7.868 \\
\textbf{2013} & 23.84 & -3.169 & 11.92 & -3.169 & 20.67 \\
\textbf{2016} & 15.27 & -12.07 & 7.633 & -4.024 & 3.193 \\
\textbf{2021} & 15.2 & -16.53 & 5.068 & -8.266 & -1.329 \\
\bottomrule
\end{longtable}

\normalsize

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeMonthlyCumSum-11} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeMonthlyCumSum-12} \end{center}

\newpage
\FloatBarrier

\hypertarget{cumulative-sums-by-sza-and-part-of-day}{%
\subsubsection{Cumulative sums by SZA and part of day}\label{cumulative-sums-by-sza-and-part-of-day}}

Use deseasonalized monthly values to calculate cumulative sums
by a SZA bin and a period of day.

\newpage

\FloatBarrier

\hypertarget{monthly-cum-sums-for-all-sky-cond.-by-sza}{%
\paragraph{Monthly cum sums for All sky cond. by SZA}\label{monthly-cum-sums-for-all-sky-cond.-by-sza}}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeMonthlySZACumSum-1} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeMonthlySZACumSum-2} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeMonthlySZACumSum-3} \end{center}

\newpage

\FloatBarrier

\hypertarget{monthly-cum-sums-for-clear-sky-cond.-by-sza}{%
\paragraph{Monthly cum sums for Clear sky cond. by SZA}\label{monthly-cum-sums-for-clear-sky-cond.-by-sza}}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeMonthlySZACumSum-4} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeMonthlySZACumSum-5} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeMonthlySZACumSum-6} \end{center}

\newpage

\FloatBarrier

\hypertarget{monthly-cum-sums-for-cloudy-cond.-by-sza}{%
\paragraph{Monthly cum sums for Cloudy cond. by SZA}\label{monthly-cum-sums-for-cloudy-cond.-by-sza}}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeMonthlySZACumSum-7} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeMonthlySZACumSum-8} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeMonthlySZACumSum-9} \end{center}

\newpage

\FloatBarrier

\hypertarget{monthly-cum-sums-for-all-sky-cond.-by-sza-1}{%
\paragraph{Monthly cum sums for All sky cond. by SZA}\label{monthly-cum-sums-for-all-sky-cond.-by-sza-1}}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeSZASumDir-1} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeSZASumDir-2} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeSZASumDir-3} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeSZASumDir-4} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeSZASumDir-5} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeSZASumDir-6} \end{center}

\newpage

\FloatBarrier

\hypertarget{monthly-cum-sums-for-clear-sky-cond.-by-sza-1}{%
\paragraph{Monthly cum sums for Clear sky cond. by SZA}\label{monthly-cum-sums-for-clear-sky-cond.-by-sza-1}}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeSZASumDir-7} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeSZASumDir-8} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeSZASumDir-9} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeSZASumDir-10} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeSZASumDir-11} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeSZASumDir-12} \end{center}

\newpage

\FloatBarrier

\hypertarget{monthly-cum-sums-for-cloudy-cond.-by-sza-1}{%
\paragraph{Monthly cum sums for Cloudy cond. by SZA}\label{monthly-cum-sums-for-cloudy-cond.-by-sza-1}}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeSZASumDir-13} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeSZASumDir-14} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeSZASumDir-15} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeSZASumDir-16} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeSZASumDir-17} \end{center}

\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeSZASumDir-18} \end{center}

\textbf{END}

\begin{verbatim}
2023-06-25 17:43:04.1 athan@tyler DHI_GHI_1_longterm_trends.R 3.506231 mins
\end{verbatim}

\end{document}

### 2023-06-25 DHI_GHI_2_sza_trends.tex
Commit 2023-06-26 01:17

@@
 \author{Natsis Athanasios\footnote{Laboratory of Atmospheric Physics, AUTH, \href{mailto:natsisphysicist@gmail.com}{\nolinkurl{natsisphysicist@gmail.com}}} \and Alkiviadis Bais\footnote{Laboratory of Atmospheric Physics, AUTH}}
-\date{2023-06-25}
+\date{2023-06-26}
 
@@
 
-Time data span 1993-04-12, 2023-05-31
+Time data span 1993-04-12, 2011-03-31
 
@@
 
-\hypertarget{all-sky-cond.-dir.-beam-irrad.}{%
-\paragraph{All sky cond. Dir. Beam Irrad.}\label{all-sky-cond.-dir.-beam-irrad.}}
+\hypertarget{all-sky-cond.-sdr}{%
+\paragraph{All sky cond. SDR}\label{all-sky-cond.-sdr}}
 
@@
 
-\hypertarget{clear-sky-cond.-dir.-beam-irrad.}{%
-\paragraph{Clear sky cond. Dir. Beam Irrad.}\label{clear-sky-cond.-dir.-beam-irrad.}}
+\hypertarget{clear-sky-cond.-sdr}{%
+\paragraph{Clear sky cond. SDR}\label{clear-sky-cond.-sdr}}
 
@@
 
-\hypertarget{cloudy-cond.-dir.-beam-irrad.}{%
-\paragraph{Cloudy cond. Dir. Beam Irrad.}\label{cloudy-cond.-dir.-beam-irrad.}}
+\hypertarget{cloudy-cond.-sdr}{%
+\paragraph{Cloudy cond. SDR}\label{cloudy-cond.-sdr}}
 
@@
 
-\newpage
-
-\hypertarget{all-sky-cond.-sdr}{%
-\paragraph{All sky cond. SDR}\label{all-sky-cond.-sdr}}
-
-\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrends-13} \end{center}
-
-\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrends-14} \end{center}
-
-\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrends-15} \end{center}
-
-\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrends-16} \end{center}
-
-\newpage
-
-\hypertarget{clear-sky-cond.-sdr}{%
-\paragraph{Clear sky cond. SDR}\label{clear-sky-cond.-sdr}}
-
-\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrends-17} \end{center}
-
-\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrends-18} \end{center}
-
-\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrends-19} \end{center}
-
-\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrends-20} \end{center}
-
-\newpage
-
-\hypertarget{cloudy-cond.-sdr}{%
-\paragraph{Cloudy cond. SDR}\label{cloudy-cond.-sdr}}
-
-\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrends-21} \end{center}
-
-\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrends-22} \end{center}
-
-\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrends-23} \end{center}
-
-\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrends-24} \end{center}
-
 \hypertarget{plot-of-sza-trends-for-each-season-of-year}{%
@@
 
-\hypertarget{winter-all-sky-cond.-dir.-beam-irrad.}{%
-\subsubsection{Winter All sky cond. Dir. Beam Irrad.}\label{winter-all-sky-cond.-dir.-beam-irrad.}}
+\hypertarget{winter-all-sky-cond.-sdr}{%
+\subsubsection{Winter All sky cond. SDR}\label{winter-all-sky-cond.-sdr}}
 
@@
 
-\hypertarget{winter-all-sky-cond.-sdr}{%
-\subsubsection{Winter All sky cond. SDR}\label{winter-all-sky-cond.-sdr}}
+\hypertarget{winter-clear-sky-cond.-sdr}{%
+\subsubsection{Winter Clear sky cond. SDR}\label{winter-clear-sky-cond.-sdr}}
 
@@
 
-\hypertarget{winter-clear-sky-cond.-dir.-beam-irrad.}{%
-\subsubsection{Winter Clear sky cond. Dir. Beam Irrad.}\label{winter-clear-sky-cond.-dir.-beam-irrad.}}
+\hypertarget{winter-cloudy-cond.-sdr}{%
+\subsubsection{Winter Cloudy cond. SDR}\label{winter-cloudy-cond.-sdr}}
 
@@
 
-\hypertarget{winter-clear-sky-cond.-sdr}{%
-\subsubsection{Winter Clear sky cond. SDR}\label{winter-clear-sky-cond.-sdr}}
+\hypertarget{spring-all-sky-cond.-sdr}{%
+\subsubsection{Spring All sky cond. SDR}\label{spring-all-sky-cond.-sdr}}
 
@@
 
-\hypertarget{winter-cloudy-cond.-dir.-beam-irrad.}{%
-\subsubsection{Winter Cloudy cond. Dir. Beam Irrad.}\label{winter-cloudy-cond.-dir.-beam-irrad.}}
+\hypertarget{spring-clear-sky-cond.-sdr}{%
+\subsubsection{Spring Clear sky cond. SDR}\label{spring-clear-sky-cond.-sdr}}
 
@@
 
-\hypertarget{winter-cloudy-cond.-sdr}{%
-\subsubsection{Winter Cloudy cond. SDR}\label{winter-cloudy-cond.-sdr}}
+\hypertarget{spring-cloudy-cond.-sdr}{%
+\subsubsection{Spring Cloudy cond. SDR}\label{spring-cloudy-cond.-sdr}}
 
@@
 
-\hypertarget{spring-all-sky-cond.-dir.-beam-irrad.}{%
-\subsubsection{Spring All sky cond. Dir. Beam Irrad.}\label{spring-all-sky-cond.-dir.-beam-irrad.}}
+\hypertarget{summer-all-sky-cond.-sdr}{%
+\subsubsection{Summer All sky cond. SDR}\label{summer-all-sky-cond.-sdr}}
 
@@
 
-\hypertarget{spring-all-sky-cond.-sdr}{%
-\subsubsection{Spring All sky cond. SDR}\label{spring-all-sky-cond.-sdr}}
+\hypertarget{summer-clear-sky-cond.-sdr}{%
+\subsubsection{Summer Clear sky cond. SDR}\label{summer-clear-sky-cond.-sdr}}
 
@@
 
-\hypertarget{spring-clear-sky-cond.-dir.-beam-irrad.}{%
-\subsubsection{Spring Clear sky cond. Dir. Beam Irrad.}\label{spring-clear-sky-cond.-dir.-beam-irrad.}}
+\hypertarget{summer-cloudy-cond.-sdr}{%
+\subsubsection{Summer Cloudy cond. SDR}\label{summer-cloudy-cond.-sdr}}
 
@@
 
-\hypertarget{spring-clear-sky-cond.-sdr}{%
-\subsubsection{Spring Clear sky cond. SDR}\label{spring-clear-sky-cond.-sdr}}
+\hypertarget{autumn-all-sky-cond.-sdr}{%
+\subsubsection{Autumn All sky cond. SDR}\label{autumn-all-sky-cond.-sdr}}
 
@@
 
-\hypertarget{spring-cloudy-cond.-dir.-beam-irrad.}{%
-\subsubsection{Spring Cloudy cond. Dir. Beam Irrad.}\label{spring-cloudy-cond.-dir.-beam-irrad.}}
+\hypertarget{autumn-clear-sky-cond.-sdr}{%
+\subsubsection{Autumn Clear sky cond. SDR}\label{autumn-clear-sky-cond.-sdr}}
 
@@
 
-\hypertarget{spring-cloudy-cond.-sdr}{%
-\subsubsection{Spring Cloudy cond. SDR}\label{spring-cloudy-cond.-sdr}}
+\hypertarget{autumn-cloudy-cond.-sdr}{%
+\subsubsection{Autumn Cloudy cond. SDR}\label{autumn-cloudy-cond.-sdr}}
 
@@
 
-\newpage
-
-\hypertarget{summer-all-sky-cond.-dir.-beam-irrad.}{%
-\subsubsection{Summer All sky cond. Dir. Beam Irrad.}\label{summer-all-sky-cond.-dir.-beam-irrad.}}
-
-\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-49} \end{center}
-
-\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-50} \end{center}
-
-\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-51} \end{center}
-
-\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-52} \end{center}
-
-\newpage
-
-\hypertarget{summer-all-sky-cond.-sdr}{%
-\subsubsection{Summer All sky cond. SDR}\label{summer-all-sky-cond.-sdr}}
-
-\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-53} \end{center}
-
-\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-54} \end{center}
-
-\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-55} \end{center}
-
-\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-56} \end{center}
-
-\newpage
-
-\hypertarget{summer-clear-sky-cond.-dir.-beam-irrad.}{%
-\subsubsection{Summer Clear sky cond. Dir. Beam Irrad.}\label{summer-clear-sky-cond.-dir.-beam-irrad.}}
-
-\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-57} \end{center}
-
-\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-58} \end{center}
-
-\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-59} \end{center}
-
-\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-60} \end{center}
-
-\newpage
-
-\hypertarget{summer-clear-sky-cond.-sdr}{%
-\subsubsection{Summer Clear sky cond. SDR}\label{summer-clear-sky-cond.-sdr}}
-
-\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-61} \end{center}
-
-\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-62} \end{center}
-
-\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-63} \end{center}
-
-\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-64} \end{center}
-
-\newpage
-
-\hypertarget{summer-cloudy-cond.-dir.-beam-irrad.}{%
-\subsubsection{Summer Cloudy cond. Dir. Beam Irrad.}\label{summer-cloudy-cond.-dir.-beam-irrad.}}
-
-\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-65} \end{center}
-
-\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-66} \end{center}
-
-\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-67} \end{center}
-
-\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-68} \end{center}
-
-\newpage
-
-\hypertarget{summer-cloudy-cond.-sdr}{%
-\subsubsection{Summer Cloudy cond. SDR}\label{summer-cloudy-cond.-sdr}}
-
-\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-69} \end{center}
-
-\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-70} \end{center}
-
-\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-71} \end{center}
-
-\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-72} \end{center}
-
-\newpage
-
-\hypertarget{autumn-all-sky-cond.-dir.-beam-irrad.}{%
-\subsubsection{Autumn All sky cond. Dir. Beam Irrad.}\label{autumn-all-sky-cond.-dir.-beam-irrad.}}
-
-\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-73} \end{center}
-
-\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-74} \end{center}
-
-\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-75} \end{center}
-
-\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-76} \end{center}
-
-\newpage
-
-\hypertarget{autumn-all-sky-cond.-sdr}{%
-\subsubsection{Autumn All sky cond. SDR}\label{autumn-all-sky-cond.-sdr}}
-
-\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-77} \end{center}
-
-\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-78} \end{center}
-
-\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-79} \end{center}
-
-\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-80} \end{center}
-
-\newpage
-
-\hypertarget{autumn-clear-sky-cond.-dir.-beam-irrad.}{%
-\subsubsection{Autumn Clear sky cond. Dir. Beam Irrad.}\label{autumn-clear-sky-cond.-dir.-beam-irrad.}}
-
-\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-81} \end{center}
-
-\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-82} \end{center}
-
-\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-83} \end{center}
-
-\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-84} \end{center}
-
-\newpage
-
-\hypertarget{autumn-clear-sky-cond.-sdr}{%
-\subsubsection{Autumn Clear sky cond. SDR}\label{autumn-clear-sky-cond.-sdr}}
-
-\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-85} \end{center}
-
-\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-86} \end{center}
-
-\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-87} \end{center}
-
-\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-88} \end{center}
-
-\newpage
-
-\hypertarget{autumn-cloudy-cond.-dir.-beam-irrad.}{%
-\subsubsection{Autumn Cloudy cond. Dir. Beam Irrad.}\label{autumn-cloudy-cond.-dir.-beam-irrad.}}
-
-\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-89} \end{center}
-
-\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-90} \end{center}
-
-\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-91} \end{center}
-
-\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-92} \end{center}
-
-\newpage
-
-\hypertarget{autumn-cloudy-cond.-sdr}{%
-\subsubsection{Autumn Cloudy cond. SDR}\label{autumn-cloudy-cond.-sdr}}
-
-\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-93} \end{center}
-
-\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-94} \end{center}
-
-\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-95} \end{center}
-
-\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_2_sza_trends_files/figure-latex/SzaTrendsSeas-96} \end{center}
-
 \textbf{END}
@@
 \begin{verbatim}
-2023-06-25 17:38:40.8 athan@tyler DHI_GHI_1_longterm_trends.R 1.213719 mins
+2023-06-25 21:52:15.8 athan@sagan DHI_GHI_1_longterm_trends.R 0.348333 mins
 \end{verbatim}

### 2023-06-25 DHI_GHI_3_trends_consistency.tex
Commit 2023-06-26 01:17

@@
 \endhead
-\textbf{1993} & 2232 & -3205 & 16.06 & -31.43 & -973.5 \\
-\textbf{1994} & 4452 & -5575 & 22.83 & -35.29 & -1123 \\
-\textbf{1995} & 3407 & -5415 & 21.29 & -30.6 & -2009 \\
-\textbf{1996} & 2788 & -7750 & 16.79 & -41.44 & -4962 \\
-\textbf{1997} & 4470 & -6190 & 22.81 & -39.17 & -1719 \\
-\textbf{1998} & 4367 & -6098 & 22.06 & -40.12 & -1731 \\
-\textbf{1999} & 4700 & -5052 & 22.49 & -33.02 & -352.4 \\
-\textbf{2000} & 5692 & -3631 & 23.23 & -32.71 & 2061 \\
-\textbf{2001} & 4570 & -5386 & 23.08 & -36.89 & -816 \\
-\textbf{2002} & 4535 & -5942 & 23.62 & -36.01 & -1407 \\
-\textbf{2003} & 4989 & -5994 & 24.46 & -38.67 & -1004 \\
-\textbf{2004} & 4499 & -5379 & 21.63 & -36.35 & -880 \\
-\textbf{2005} & 4613 & -3980 & 23.66 & -29.26 & 633.2 \\
-\textbf{2006} & 4764 & -4385 & 23.58 & -32.97 & 378.7 \\
-\textbf{2007} & 4446 & -4126 & 21.9 & -36.19 & 319.5 \\
-\textbf{2008} & 5096 & -4847 & 22.06 & -38.16 & 249 \\
-\textbf{2009} & 5253 & -5057 & 26.27 & -35.61 & 196.8 \\
-\textbf{2010} & 3749 & -4362 & 23.14 & -36.35 & -613.4 \\
-\textbf{2011} & 4780 & -4629 & 22.44 & -38.26 & 150.5 \\
-\textbf{2012} & 4944 & -4277 & 21.5 & -40.35 & 667.3 \\
-\textbf{2013} & 5326 & -4362 & 23.88 & -34.08 & 964.2 \\
-\textbf{2014} & 4270 & -5710 & 21.57 & -38.07 & -1440 \\
-\textbf{2015} & 5612 & -4511 & 27.11 & -36.09 & 1101 \\
-\textbf{2016} & 6263 & -4623 & 27.23 & -35.03 & 1640 \\
-\textbf{2017} & 6655 & -4315 & 26.94 & -37.85 & 2340 \\
-\textbf{2018} & 5163 & -5085 & 24.59 & -34.83 & 78.02 \\
-\textbf{2019} & 5472 & -3791 & 23.39 & -32.97 & 1681 \\
-\textbf{2020} & 5989 & -3791 & 25.49 & -33.84 & 2199 \\
-\textbf{2021} & 5750 & -4535 & 24.47 & -35.16 & 1215 \\
-\textbf{2022} & 6231 & -4079 & 25.33 & -35.47 & 2151 \\
-\textbf{2023} & 2958 & -1952 & 32.51 & -34.85 & 1006 \\
+\textbf{1993} & 2610 & -3026 & 18 & -31.52 & -415.8 \\
+\textbf{1994} & 5058 & -5349 & 24.79 & -35.9 & -291.5 \\
+\textbf{1995} & 3670 & -5098 & 22.38 & -29.47 & -1429 \\
+\textbf{1996} & 3269 & -7514 & 19.23 & -41.06 & -4245 \\
+\textbf{1997} & 5002 & -5946 & 24.88 & -38.87 & -944.7 \\
+\textbf{1998} & 4977 & -5880 & 23.59 & -42.3 & -902.3 \\
+\textbf{1999} & 5124 & -4687 & 23.94 & -31.67 & 437.1 \\
+\textbf{2000} & 6492 & -3419 & 25.97 & -32.25 & 3073 \\
+\textbf{2001} & 4913 & -5082 & 23.97 & -36.56 & -168.3 \\
+\textbf{2002} & 5152 & -5694 & 26.56 & -34.93 & -542.2 \\
+\textbf{2003} & 5420 & -5678 & 25.93 & -37.85 & -257.8 \\
+\textbf{2004} & 5095 & -5125 & 23.81 & -36.09 & -29.19 \\
+\textbf{2005} & 5160 & -3690 & 24.93 & -29.76 & 1469 \\
+\textbf{2006} & 5554 & -4210 & 26.2 & -34.22 & 1344 \\
+\textbf{2007} & 4938 & -3920 & 23.74 & -35.96 & 1019 \\
+\textbf{2008} & 5814 & -4603 & 24.23 & -39.01 & 1212 \\
+\textbf{2009} & 5724 & -4746 & 28.06 & -34.39 & 977.8 \\
+\textbf{2010} & 4347 & -4191 & 26.84 & -34.92 & 156.5 \\
+\textbf{2011} & 1707 & -2171 & 40.65 & -48.24 & -463.5 \\
 \bottomrule
@@
 \endhead
-\textbf{1993} & 2814 & -3584 & 21.48 & -33.19 & -770.5 \\
-\textbf{1994} & 5696 & -5936 & 28.77 & -38.8 & -240 \\
-\textbf{1995} & 4001 & -6110 & 27.98 & -31.66 & -2109 \\
-\textbf{1996} & 3804 & -8259 & 25.36 & -40.69 & -4456 \\
-\textbf{1997} & 4762 & -6823 & 27.06 & -38.33 & -2060 \\
-\textbf{1998} & 4678 & -6597 & 25.85 & -39.04 & -1919 \\
-\textbf{1999} & 5594 & -5572 & 28.11 & -34.19 & 21.55 \\
-\textbf{2000} & 5954 & -4077 & 26.35 & -31.36 & 1877 \\
-\textbf{2001} & 4426 & -5877 & 24.73 & -36.5 & -1451 \\
-\textbf{2002} & 5201 & -6534 & 30.06 & -35.9 & -1333 \\
-\textbf{2003} & 5346 & -6573 & 27.84 & -39.36 & -1228 \\
-\textbf{2004} & 4671 & -5968 & 26.54 & -33.16 & -1297 \\
-\textbf{2005} & 5624 & -4485 & 30.9 & -30.31 & 1139 \\
-\textbf{2006} & 5677 & -4802 & 30.36 & -32.67 & 875 \\
-\textbf{2007} & 5241 & -4427 & 27.3 & -35.42 & 813.7 \\
-\textbf{2008} & 5997 & -5223 & 28.29 & -36.02 & 773.4 \\
-\textbf{2009} & 5527 & -5536 & 29.87 & -35.26 & -9.141 \\
-\textbf{2010} & 3575 & -4751 & 24.66 & -34.68 & -1176 \\
-\textbf{2011} & 4797 & -5050 & 24.98 & -35.81 & -253.1 \\
-\textbf{2012} & 5341 & -4680 & 25.8 & -36.57 & 660.7 \\
-\textbf{2013} & 5225 & -4814 & 25.49 & -32.97 & 410.7 \\
-\textbf{2014} & 4806 & -6218 & 25.98 & -38.15 & -1412 \\
-\textbf{2015} & 5698 & -4929 & 29.22 & -35.98 & 768.7 \\
-\textbf{2016} & 6797 & -5042 & 30.75 & -36.01 & 1755 \\
-\textbf{2017} & 6495 & -4752 & 29.12 & -34.44 & 1743 \\
-\textbf{2018} & 5927 & -5644 & 31.03 & -34.21 & 283.1 \\
-\textbf{2019} & 5870 & -4292 & 28.08 & -30.88 & 1578 \\
-\textbf{2020} & 6083 & -4301 & 28.83 & -31.63 & 1781 \\
-\textbf{2021} & 6692 & -4976 & 29.88 & -35.54 & 1716 \\
-\textbf{2022} & 6803 & -4447 & 28.83 & -35.57 & 2356 \\
-\textbf{2023} & 3233 & -2072 & 36.33 & -35.72 & 1161 \\
+\textbf{1993} & 3054 & -3389 & 22.13 & -33.56 & -334.9 \\
+\textbf{1994} & 6247 & -5727 & 31.24 & -37.93 & 520.4 \\
+\textbf{1995} & 4210 & -5814 & 28.45 & -30.93 & -1604 \\
+\textbf{1996} & 4206 & -8033 & 26.62 & -41.19 & -3827 \\
+\textbf{1997} & 5268 & -6573 & 29.43 & -37.56 & -1304 \\
+\textbf{1998} & 5243 & -6363 & 28.34 & -38.56 & -1119 \\
+\textbf{1999} & 5920 & -5190 & 28.88 & -33.05 & 730.8 \\
+\textbf{2000} & 6699 & -3870 & 29.64 & -29.77 & 2829 \\
+\textbf{2001} & 4720 & -5592 & 26.08 & -35.17 & -872.5 \\
+\textbf{2002} & 5795 & -6316 & 32.56 & -35.68 & -520.3 \\
+\textbf{2003} & 5721 & -6286 & 30.11 & -37.19 & -565.2 \\
+\textbf{2004} & 5192 & -5678 & 28.84 & -32.26 & -485.9 \\
+\textbf{2005} & 6115 & -4224 & 32.7 & -29.54 & 1891 \\
+\textbf{2006} & 6389 & -4608 & 32.43 & -33.63 & 1781 \\
+\textbf{2007} & 5666 & -4235 & 29.21 & -34.43 & 1432 \\
+\textbf{2008} & 6599 & -4942 & 29.59 & -36.88 & 1656 \\
+\textbf{2009} & 6003 & -5238 & 32.28 & -33.58 & 765.2 \\
+\textbf{2010} & 4117 & -4603 & 28.01 & -34.1 & -486.4 \\
+\textbf{2011} & 1740 & -2226 & 42.43 & -49.46 & -486.1 \\
 \bottomrule
@@
 \endhead
-\textbf{1993} & 555.1 & -1190 & 20.56 & -27.05 & -635.2 \\
-\textbf{1994} & 661.9 & -1754 & 18.39 & -24.7 & -1092 \\
-\textbf{1995} & 640.4 & -1042 & 18.3 & -21.7 & -401.2 \\
-\textbf{1996} & 873 & -1718 & 26.46 & -28.17 & -845.2 \\
-\textbf{1997} & 1272 & -1106 & 23.56 & -26.98 & 166.2 \\
-\textbf{1998} & 1049 & -992.1 & 20.58 & -20.67 & 57.34 \\
-\textbf{1999} & 714.5 & -1368 & 17.43 & -20.73 & -653.8 \\
-\textbf{2000} & 972.2 & -1160 & 23.15 & -22.74 & -187.7 \\
-\textbf{2001} & 1005 & -655.4 & 24.5 & -18.73 & 349.2 \\
-\textbf{2002} & 1166 & -1155 & 21.2 & -23.56 & 11.55 \\
-\textbf{2003} & 1045 & -1142 & 24.31 & -22.83 & -96.28 \\
-\textbf{2004} & 1356 & -646.1 & 27.68 & -19.58 & 710.1 \\
-\textbf{2005} & 825.7 & -1101 & 20.14 & -20.78 & -275.8 \\
-\textbf{2006} & 941.4 & -1006 & 21.89 & -21.86 & -64.25 \\
-\textbf{2007} & 654.3 & -1183 & 21.81 & -25.71 & -528.3 \\
-\textbf{2008} & 845 & -1394 & 22.84 & -21.78 & -549 \\
-\textbf{2009} & 839.7 & -755.1 & 22.69 & -22.21 & 84.59 \\
-\textbf{2010} & 792.5 & -529.7 & 23.31 & -21.19 & 262.8 \\
-\textbf{2011} & 1305 & -775.3 & 26.1 & -23.49 & 529.9 \\
-\textbf{2012} & 1132 & -1052 & 25.15 & -25.05 & 79.83 \\
-\textbf{2013} & 1382 & -776.3 & 27.1 & -18.05 & 605.7 \\
-\textbf{2014} & 856.1 & -815.9 & 23.14 & -19.9 & 40.2 \\
-\textbf{2015} & 949 & -515.4 & 21.09 & -20.61 & 433.7 \\
-\textbf{2016} & 1287 & -1040 & 24.76 & -22.12 & 247.8 \\
-\textbf{2017} & 971.9 & -737.2 & 22.6 & -21.06 & 234.7 \\
-\textbf{2018} & 1244 & -867.5 & 24.4 & -23.44 & 376.8 \\
-\textbf{2019} & 928.5 & -866.3 & 23.81 & -21.13 & 62.18 \\
-\textbf{2020} & 1196 & -579.6 & 22.15 & -17.56 & 616.7 \\
-\textbf{2021} & 1048 & -844.3 & 16.64 & -16.56 & 203.8 \\
-\textbf{2022} & 1112 & -784.9 & 21.8 & -19.14 & 326.9 \\
-\textbf{2023} & 131.9 & -203 & 10.99 & -20.3 & -71.12 \\
+\textbf{1993} & 556.6 & -1021 & 21.41 & -23.76 & -464.8 \\
+\textbf{1994} & 616.9 & -1544 & 17.14 & -22.7 & -926.6 \\
+\textbf{1995} & 698 & -899.1 & 17.9 & -21.41 & -201.1 \\
+\textbf{1996} & 957.9 & -1481 & 30.9 & -24.28 & -523.2 \\
+\textbf{1997} & 1305 & -996.4 & 25.59 & -24.3 & 308.4 \\
+\textbf{1998} & 1227 & -761.8 & 24.06 & -18.14 & 465 \\
+\textbf{1999} & 754.4 & -1116 & 16.4 & -18.6 & -361.7 \\
+\textbf{2000} & 1015 & -993.2 & 22.55 & -21.59 & 21.44 \\
+\textbf{2001} & 1047 & -576 & 26.17 & -16.46 & 470.8 \\
+\textbf{2002} & 1282 & -1019 & 23.73 & -21.67 & 263 \\
+\textbf{2003} & 1123 & -1041 & 26.13 & -21.69 & 82.5 \\
+\textbf{2004} & 1395 & -562.8 & 27.9 & -19.41 & 832.3 \\
+\textbf{2005} & 904.3 & -1018 & 18.84 & -22.13 & -113.7 \\
+\textbf{2006} & 1044 & -929.3 & 21.75 & -23.23 & 114.7 \\
+\textbf{2007} & 701.4 & -1082 & 25.05 & -23.03 & -380.8 \\
+\textbf{2008} & 916.9 & -1225 & 22.36 & -21.13 & -308.6 \\
+\textbf{2009} & 894.8 & -575.7 & 22.37 & -19.85 & 319.1 \\
+\textbf{2010} & 790.5 & -469.3 & 23.25 & -19.55 & 321.2 \\
+\textbf{2011} & 96.34 & -14.4 & 16.06 & -3.6 & 81.94 \\
 \bottomrule
@@
 \begin{longtable}[]{@{}
-  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1410}}
-  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1795}}
-  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1795}}
-  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1923}}
-  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1923}}
-  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1154}}@{}}
+  >{\centering\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1392}}
+  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1772}}
+  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1772}}
+  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1899}}
+  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1899}}
+  >{\raggedleft\arraybackslash}p{(\columnwidth - 10\tabcolsep) * \real{0.1266}}@{}}
 \toprule
@@
 \endhead
-\textbf{1993} & 19.08 & -48.52 & 6.36 & -9.704 & -29.44 \\
-\textbf{1994} & 26.02 & -74.45 & 5.204 & -10.64 & -48.42 \\
-\textbf{1995} & 26.9 & -93 & 5.38 & -15.5 & -66.1 \\
-\textbf{1996} & 3.185 & -182.2 & 1.592 & -18.22 & -179 \\
-\textbf{1997} & 26.52 & -81.3 & 8.84 & -9.033 & -54.78 \\
-\textbf{1998} & 27.93 & -76.09 & 5.586 & -10.87 & -48.16 \\
-\textbf{1999} & 7.774 & -24.16 & 1.943 & -3.02 & -16.39 \\
-\textbf{2000} & 77.82 & -18.31 & 8.646 & -6.102 & 59.51 \\
-\textbf{2001} & 49.51 & -76.54 & 9.901 & -10.93 & -27.03 \\
-\textbf{2002} & 38.91 & -84.7 & 12.97 & -9.411 & -45.78 \\
-\textbf{2003} & 29.84 & -56.12 & 4.974 & -9.354 & -26.28 \\
-\textbf{2004} & 28.81 & -57.29 & 4.802 & -9.548 & -28.48 \\
-\textbf{2005} & 33.86 & -14.68 & 4.837 & -3.669 & 19.19 \\
-\textbf{2006} & 37.15 & -35.07 & 7.43 & -5.844 & 2.085 \\
-\textbf{2007} & 24.02 & -15.11 & 4.805 & -3.022 & 8.915 \\
-\textbf{2008} & 50.75 & -31.9 & 7.25 & -6.38 & 18.85 \\
-\textbf{2009} & 56.6 & -55.74 & 8.086 & -11.15 & 0.8636 \\
-\textbf{2010} & 24.77 & -38.35 & 6.191 & -9.587 & -13.58 \\
-\textbf{2011} & 24.73 & -28.67 & 4.122 & -5.733 & -3.935 \\
-\textbf{2012} & 45 & -23.46 & 5.624 & -7.819 & 21.54 \\
-\textbf{2013} & 64.93 & -36.73 & 8.116 & -9.184 & 28.2 \\
-\textbf{2014} & 20.71 & -64.85 & 4.141 & -9.264 & -44.14 \\
-\textbf{2015} & 92.01 & -45.21 & 13.14 & -11.3 & 46.8 \\
-\textbf{2016} & 73.43 & -18.02 & 8.159 & -6.007 & 55.41 \\
-\textbf{2017} & 89.19 & -11.66 & 9.91 & -3.886 & 77.53 \\
-\textbf{2018} & 51.06 & -44.16 & 7.294 & -8.832 & 6.901 \\
-\textbf{2019} & 61.82 & -17.06 & 8.831 & -3.413 & 44.75 \\
-\textbf{2020} & 97.56 & -20.95 & 10.84 & -6.983 & 76.61 \\
-\textbf{2021} & 60.92 & -21.37 & 8.703 & -4.274 & 39.55 \\
-\textbf{2022} & 94.15 & -18.65 & 11.77 & -4.662 & 75.5 \\
-\textbf{2023} & 55.86 & -21.35 & 13.96 & -21.35 & 34.51 \\
+\textbf{1993} & 32.03 & -43.15 & 8.008 & -10.79 & -11.12 \\
+\textbf{1994} & 38.66 & -59.06 & 6.443 & -9.843 & -20.4 \\
+\textbf{1995} & 33.47 & -77.73 & 6.694 & -12.96 & -44.26 \\
+\textbf{1996} & 6.869 & -161.9 & 2.29 & -17.99 & -155.1 \\
+\textbf{1997} & 33.28 & -61.31 & 8.321 & -7.664 & -28.02 \\
+\textbf{1998} & 37.8 & -58.98 & 4.726 & -14.74 & -21.17 \\
+\textbf{1999} & 27.94 & -15.43 & 3.992 & -3.086 & 12.51 \\
+\textbf{2000} & 101.9 & -10.79 & 9.266 & -10.79 & 91.13 \\
+\textbf{2001} & 60.36 & -61.26 & 10.06 & -10.21 & -0.9042 \\
+\textbf{2002} & 51.82 & -69.79 & 12.95 & -8.724 & -17.98 \\
+\textbf{2003} & 35.53 & -34.82 & 7.106 & -4.974 & 0.7085 \\
+\textbf{2004} & 43.12 & -43.88 & 5.39 & -10.97 & -0.7636 \\
+\textbf{2005} & 55.88 & -8.827 & 7.983 & -2.207 & 47.06 \\
+\textbf{2006} & 50.15 & -25.46 & 7.165 & -6.365 & 24.69 \\
+\textbf{2007} & 39.78 & -8.71 & 6.63 & -2.177 & 31.07 \\
+\textbf{2008} & 73.43 & -25.38 & 9.179 & -6.345 & 48.05 \\
+\textbf{2009} & 66.67 & -38.18 & 9.524 & -7.635 & 28.49 \\
+\textbf{2010} & 32.42 & -28.63 & 6.484 & -9.544 & 3.789 \\
+\textbf{2011} & 5.498 & -19.1 & 5.498 & -9.55 & -13.6 \\
 \bottomrule
@@
 \endhead
-\textbf{1993} & 23 & -43.35 & 7.667 & -8.67 & -20.35 \\
-\textbf{1994} & 44.44 & -62.1 & 7.406 & -10.35 & -17.67 \\
-\textbf{1995} & 32.54 & -101.2 & 5.424 & -16.86 & -68.62 \\
-\textbf{1996} & 12.47 & -175.1 & 2.495 & -25.01 & -162.6 \\
-\textbf{1997} & 25.04 & -91.64 & 8.348 & -10.18 & -66.6 \\
-\textbf{1998} & 25.99 & -80.59 & 5.197 & -11.51 & -54.61 \\
-\textbf{1999} & 18.37 & -24 & 4.592 & -3.001 & -5.637 \\
-\textbf{2000} & 70.9 & -18.61 & 8.862 & -4.653 & 52.28 \\
-\textbf{2001} & 43.97 & -96.18 & 7.328 & -16.03 & -52.21 \\
-\textbf{2002} & 39.75 & -83.65 & 9.937 & -10.46 & -43.9 \\
-\textbf{2003} & 23.96 & -58.03 & 5.989 & -7.254 & -34.07 \\
-\textbf{2004} & 23.1 & -64.67 & 7.7 & -7.186 & -41.57 \\
-\textbf{2005} & 72.61 & -8.362 & 8.068 & -2.787 & 64.25 \\
-\textbf{2006} & 66.38 & -30.4 & 9.483 & -6.08 & 35.98 \\
-\textbf{2007} & 38.48 & -24.91 & 5.497 & -6.227 & 13.57 \\
-\textbf{2008} & 69.96 & -32.22 & 8.744 & -8.054 & 37.74 \\
-\textbf{2009} & 49.6 & -58.44 & 8.267 & -9.741 & -8.844 \\
-\textbf{2010} & 17.47 & -62.67 & 3.495 & -10.45 & -45.2 \\
-\textbf{2011} & 21.77 & -41.61 & 5.443 & -5.944 & -19.84 \\
-\textbf{2012} & 44.68 & -23.08 & 6.383 & -5.769 & 21.61 \\
-\textbf{2013} & 58.06 & -48.54 & 8.295 & -9.708 & 9.523 \\
-\textbf{2014} & 17.42 & -60.43 & 3.484 & -8.632 & -43.01 \\
-\textbf{2015} & 88.25 & -47.45 & 14.71 & -7.909 & 40.8 \\
-\textbf{2016} & 87.36 & -24.92 & 8.736 & -12.46 & 62.44 \\
-\textbf{2017} & 80.13 & -22.2 & 11.45 & -4.44 & 57.93 \\
-\textbf{2018} & 60.39 & -47.65 & 10.07 & -7.942 & 12.74 \\
-\textbf{2019} & 60.79 & -20.92 & 10.13 & -3.487 & 39.87 \\
-\textbf{2020} & 94.67 & -32.11 & 15.78 & -5.352 & 62.56 \\
-\textbf{2021} & 78.64 & -21.81 & 9.83 & -5.452 & 56.83 \\
-\textbf{2022} & 105.8 & -22.44 & 13.22 & -5.609 & 83.35 \\
-\textbf{2023} & 59.84 & -21.19 & 14.96 & -21.19 & 38.66 \\
+\textbf{1993} & 33.26 & -39 & 11.09 & -7.8 & -5.739 \\
+\textbf{1994} & 54.63 & -46.3 & 9.105 & -7.716 & 8.331 \\
+\textbf{1995} & 42.76 & -87.74 & 7.127 & -14.62 & -44.98 \\
+\textbf{1996} & 17.19 & -158.6 & 3.437 & -22.66 & -141.5 \\
+\textbf{1997} & 32.75 & -74.28 & 8.188 & -9.285 & -41.52 \\
+\textbf{1998} & 32.38 & -62.16 & 4.625 & -12.43 & -29.79 \\
+\textbf{1999} & 35.78 & -14.45 & 4.473 & -3.612 & 21.34 \\
+\textbf{2000} & 92.01 & -10.09 & 8.365 & -10.09 & 81.92 \\
+\textbf{2001} & 53.16 & -82.67 & 8.86 & -13.78 & -29.51 \\
+\textbf{2002} & 54.62 & -72.14 & 13.65 & -9.017 & -17.52 \\
+\textbf{2003} & 28.23 & -37.42 & 7.057 & -4.678 & -9.193 \\
+\textbf{2004} & 33.27 & -48.89 & 5.545 & -8.149 & -15.62 \\
+\textbf{2005} & 100 & -6.177 & 11.12 & -2.059 & 93.86 \\
+\textbf{2006} & 87.41 & -22.43 & 12.49 & -4.486 & 64.98 \\
+\textbf{2007} & 52.56 & -18.87 & 6.57 & -6.289 & 33.7 \\
+\textbf{2008} & 91.7 & -26.03 & 10.19 & -8.678 & 65.67 \\
+\textbf{2009} & 57.69 & -40.85 & 11.54 & -5.836 & 16.84 \\
+\textbf{2010} & 25.07 & -52.13 & 5.014 & -8.688 & -27.05 \\
+\textbf{2011} & 6.812 & -21.58 & 6.812 & -10.79 & -14.77 \\
 \bottomrule
@@
 \endhead
-\textbf{1996} & 1.041 & -32.43 & 1.041 & -16.21 & -31.39 \\
-\textbf{1997} & 30.94 & -13.28 & 15.47 & -6.639 & 17.66 \\
-\textbf{1998} & 22.21 & -11.6 & 7.405 & -5.802 & 10.61 \\
-\textbf{2000} & 5.152 & -11.35 & 5.152 & -11.35 & -6.197 \\
-\textbf{2001} & 17.63 & -11.37 & 17.63 & -11.37 & 6.265 \\
-\textbf{2002} & 35.68 & -15.07 & 17.84 & -7.534 & 20.61 \\
-\textbf{2003} & 31.3 & -26.37 & 15.65 & -13.19 & 4.926 \\
-\textbf{2005} & 0.5237 & -13.4 & 0.5237 & -4.466 & -12.87 \\
-\textbf{2006} & 20.63 & -22.47 & 6.878 & -11.24 & -1.839 \\
-\textbf{2008} & 1.037 & -22.67 & 0.5185 & -7.558 & -21.64 \\
-\textbf{2011} & 9.071 & -7.749 & 4.535 & -7.749 & 1.322 \\
-\textbf{2012} & 18.89 & -26.75 & 9.443 & -26.75 & -7.868 \\
-\textbf{2013} & 23.84 & -3.169 & 11.92 & -3.169 & 20.67 \\
-\textbf{2016} & 15.27 & -12.07 & 7.633 & -4.024 & 3.193 \\
-\textbf{2021} & 15.2 & -16.53 & 5.068 & -8.266 & -1.329 \\
+\textbf{1996} & 5.182 & -29.29 & 5.182 & -14.64 & -24.11 \\
+\textbf{1997} & 37.41 & -8.749 & 18.7 & -4.375 & 28.66 \\
+\textbf{1998} & 25.8 & -6.183 & 8.599 & -3.092 & 19.61 \\
+\textbf{2000} & 9.461 & -9.806 & 9.461 & -9.806 & -0.3458 \\
+\textbf{2001} & 19.95 & -9.823 & 19.95 & -9.823 & 10.13 \\
+\textbf{2002} & 42.68 & -10.76 & 21.34 & -5.378 & 31.93 \\
+\textbf{2003} & 35.13 & -21.3 & 17.57 & -10.65 & 13.83 \\
+\textbf{2005} & 7.114 & -6.776 & 3.557 & -3.388 & 0.3376 \\
+\textbf{2006} & 27.81 & -19.11 & 9.269 & -9.554 & 8.698 \\
+\textbf{2008} & 5.68 & -15.25 & 2.84 & -5.085 & -9.574 \\
 \bottomrule
@@
 
-\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeSZASumDir-1} \end{center}
-
-\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeSZASumDir-2} \end{center}
-
-\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeSZASumDir-3} \end{center}
-
-\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeSZASumDir-4} \end{center}
-
-\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeSZASumDir-5} \end{center}
-
-\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeSZASumDir-6} \end{center}
-
 \newpage
@@
 \paragraph{Monthly cum sums for Clear sky cond. by SZA}\label{monthly-cum-sums-for-clear-sky-cond.-by-sza-1}}
-
-\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeSZASumDir-7} \end{center}
-
-\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeSZASumDir-8} \end{center}
-
-\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeSZASumDir-9} \end{center}
-
-\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeSZASumDir-10} \end{center}
-
-\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeSZASumDir-11} \end{center}
-
-\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeSZASumDir-12} \end{center}
 
@@
 
-\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeSZASumDir-13} \end{center}
-
-\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeSZASumDir-14} \end{center}
-
-\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeSZASumDir-15} \end{center}
-
-\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeSZASumDir-16} \end{center}
-
-\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeSZASumDir-17} \end{center}
-
-\begin{center}\includegraphics[width=1\linewidth]{DHI_GHI_3_trends_consistency_files/figure-latex/CumulativeSZASumDir-18} \end{center}
-
 \textbf{END}
@@
 \begin{verbatim}
-2023-06-25 17:43:04.1 athan@tyler DHI_GHI_1_longterm_trends.R 3.506231 mins
+2023-06-25 21:53:08.2 athan@sagan DHI_GHI_1_longterm_trends.R 0.759222 mins
 \end{verbatim}

### 2023-06-25 DHI_GHI_1_longterm_trends.R

@@ -24,7 +24,7 @@
 #'   bookdown::pdf_document2:
 #'     number_sections: no
 #'     fig_caption:     no
-#'     keep_tex:        yes
+#'     keep_tex:        no
 #'     latex_engine:    xelatex
 #'     toc:             yes
 #'     toc_depth:       4

### 2023-06-26 Article.tex
Commit 2023-06-26 07:17

\documentclass[preprint, 3p,
authoryear]{elsarticle} %review=doublespace preprint=single 5p=2 column
%%% Begin My package additions %%%%%%%%%%%%%%%%%%%

\usepackage[hyphens]{url}

  \journal{An awesome journal} % Sets Journal name

\usepackage{graphicx}
%%%%%%%%%%%%%%%% end my additions to header

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
% TODO: Currently lineno needs to be loaded after amsmath because of conflict
% https://github.com/latex-lineno/lineno/issues/5
\usepackage{lineno} % add
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\usepackage[]{natbib}
\bibliographystyle{plainnat}

\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={Trends of SDR in Thessaloniki},
            colorlinks=false,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}

\setcounter{secnumdepth}{5}
% Pandoc toggle for numbering sections (defaults to be off)

% tightlist command for lists without linebreak
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

% From pandoc table feature
\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}

\usepackage{subcaption}

\begin{document}

\begin{frontmatter}

  \title{Trends of SDR in Thessaloniki}
    \author[Laboratory of Atmospheric Physics]{Natsis Athanasios%
  \corref{cor1}%
  }
   \ead{natsisphysicist@gmail.com} 
    \author[Laboratory of Atmospheric Physics]{Alkiviadis Bais%
  %
  }
   \ead{abais@auth.gr} 
      \affiliation[Laboratory of Atmospheric Physics]{
    organization={Aristotle University of
Thessaloniki},addressline={Campus Box
149},city={Thessaloniki},postcode={54124},country={Greece},}
    \cortext[cor1]{Corresponding author}

  \begin{abstract}
  Study of GHI and DNI radiation for `clear-sky' and all-sky conditions.
  It consists of two paragraphs.
  \end{abstract}
    \begin{keyword}
    GHI \sep SDR \sep 
    Solar Brigthening/Dimming
  \end{keyword}

 \end{frontmatter}

\hypertarget{introduction.}{%
\section{Introduction.}\label{introduction.}}

The shortwave downward solar irradiance (SDR) at the Earth's surface
plays a significant role, on Earths climate. Changes of the SDR can be
related to major changes on factors affecting the radiative forcing
balance and human activities. Multiple studies attempt to evaluate the
phenomenon and the potential causes
\citep{Ohvril2009, Schwarz2020, Wild2009, Wild2012, Xia2007, Zerefos2009}.
Although, there is regional variation of SDR, with both positive and
negative trends, most of the studies agree that the major factors are
the interaction of aerosols and clouds.

In this study, we examine the trends of SDR with ground-based
measurements at Thessaloniki for the period 1993 - 2023, as derived from
a CM-21 pyranometer. We reevaluate and extend the dataset used by
\citet{Bais2013}, applying a different algorithm for the identification
of clear-/cloud-sky instances, and we derive the radiation trends for
this period under different sky conditions (all-sky, clear-sky and
cloud-sky).

\ldots{}

Site location description and aerosols effect

\ldots{}

Our method to characterize the sky conditions and the definition of each
sky condition has some subjectivity. The algorithm was calibrated with
the main focus to identify the presence of clouds on the sky dome,
although there are marginal cases that there will be false positives or
false negatives identifications.

\hypertarget{data-and-methodology.}{%
\section{Data and methodology.}\label{data-and-methodology.}}

The SDR is equivalent to the whole sky Global radiation, also refered as
Global Horizontal Irradiance (GHI), and was obtained with a horizontal
leveled, CM-21 pyranometer. SDR data spans the period of 1993-04-12 to
2011-03-31. In some algorithms we used as auxiliary data, the direct
beam radiation (DNI) for comparisons, obtained by a CHP-1 pyrheliometer,
with data availability from Inf to -Inf.

\ldots{}

\ldots{}

There are three distinct steps to the creation of this dataset: a) the
acquisition of radiation measurements from the sensors, b) a radiation
data quality check, and c) the identification of ``clear sky''
conditions from the radiometric data d) data aggregation and trend
analysis.

For the acquisition of radiometric data, the signal of the broadband
instruments is sampled with a rate of \(1 \text{Hz}\). The mean and
standard deviation values are recorded for every minute. The
measurements are corrected for the zero offset of the instrument signal.
As reference of the ``dark signal'' is used the Sun elevation angle
bellow \(-10^\circ\), for a period of \(3 \text{h}\). The signal is
converted to radiation flux, using a ramped value of the instrument
sensitivity, derived from the eight laboratory calibrations of the
instrument, during the study period.

A manual screening is performed, to remove inconsistencies and erroneous
recordings, that can occur randomly or systematically, during the long
continuous operation of the instrument. The manual screening is aided by
a radiation data quality assurance procedure, adjusted for this site,
based on methods of Long and
Shi~\citetext{\citeyear{long_automated_2008}; \citeyear{Long2006}}.
Thus, problematic recordings have been excluded from further processing.
Although it is not possible to detect all the bad data, the large number
of data and the aggregation scheme we use, can provide us with accurate
radiometric measurements, for the scope of this study.

\hypertarget{clear-sky-identification}{%
\subsection{Clear sky identification}\label{clear-sky-identification}}

\ldots\ldots{}

As global radiation clear sky reference we are using the Haurwitz's
model, adjusted for our site with a factor of 0.965 (Eq. \ref{eq:ahau}).
The selection of a clear sky reference model, was based on SDR
observation from the period 2016 -- 2021. Where, after an iterative
optimization of eight simple models (Daneshyar--Paltridge--Proctor,
Kasten--Czeplak, Haurwitz, Berger--Duffie,
Adnot--Bourges--Campana--Gicquel, Robledo-Soler, Kasten and
Ineichen-Perez) with different factors. We found, that Haurwitz's model,
adjusted with a factor of 0.965 has the lower root mean squared error
(RMSE). The tried models are described by \citet{Reno2012} and tested by
\citet{Reno2016}. The iterative optimization method, for the selection
of the reference is discussed by \citet{Long2000} and \citet{Reno2016}.

\begin{equation}
\text{SDR}_\text{Clear Sky} = 0.965 \times 1098 \times \cos( \text{SZA} ) \times \exp \left( \frac{ - 0.057}{\cos(\text{SZA})} \right) \label{eq:ahau}
\end{equation}

The following criteria and thresholds were used to identify clear-sky
conditions. Each criterion was applied for a running window of \(11\)
consecutive measurements/minutes, and the characterization is applied at
the center value of the window. A data point is consider as under
cloud-sky condition if it fails to pass any of the criteria, all the
other data points are characterized as clear-sky.

\hypertarget{mean-value-of-irradiance-during-the-time-period.}{%
\subsubsection{Mean value of irradiance during the time
period.}\label{mean-value-of-irradiance-during-the-time-period.}}

The mean of the measured value \(\overline{G}_i\) must be inside an
envelope based on the reference model \(\text{SDR}_\text{Clear Sky}\)
(Eq. \ref{eq:MeanVIP}).

\begin{equation}
0.91 \times \overline{\text{SDR}}_{i\text{Clear Sky}} - 20
< \overline{G}_i <
1.095 \times \overline{\text{SDR}}_{i\text{Clear Sky}} + 30
\label{eq:MeanVIP}
\end{equation}

\hypertarget{max-value-of-irradiance-during-the-time-period.}{%
\subsubsection{Max value of irradiance during the time
period.}\label{max-value-of-irradiance-during-the-time-period.}}

The running max measured value \(M_{Gi} = max[\text{SDR}_{i}]\), is
compared to a similar constructed value from the reference
\(M_{CSi} = max[\text{SDR}_{i\text{Clear Sky}}]\) (Eq. \ref{eq:MaxVIP}).

\begin{equation}
1 \times M_{CSi} - 75
< M_{Gi} <
1 \times M_{CSi} + 75
\label{eq:MaxVIP}
\end{equation}

\hypertarget{variability-in-irradiance-by-line-length.}{%
\subsubsection{Variability in irradiance by line
length.}\label{variability-in-irradiance-by-line-length.}}

The length \(L\) (Eq. \ref{eq:VILeq}) of the sequence of line segments
connecting the points of the SDR time series for the measured values
\(L\) and similar for the reference \(L_{CS}\), must be within the
limits of Eq. \ref{eq:VILcr}.

\begin{equation}
L = \sum_{i=1}^{n-1}\sqrt{\left ( \text{SDR}_{i+1} - \text{SDR}_{i}\right )^2 + \left ( t_{i+1} - t_i \right )^2}
\label{eq:VILeq}
\end{equation}

\begin{equation}
1 \times L_{CSi} - 5 < L_i < 1.3 \times L_{CSi} + 13
\label{eq:VILcr}
\end{equation}

\hypertarget{variance-of-changes-in-the-time-series.}{%
\subsubsection{Variance of Changes in the Time
series.}\label{variance-of-changes-in-the-time-series.}}

We calculate the standard deviation \(\sigma\) of the slope (\(s\))
between sequential points in the time series, normalized by the average
SDR during the time interval.

\begin{gather}
s_i = \frac{\text{SDR}_{i+1} - \text{SDR}_{i}}{t_{i+1} - t_i}, \forall i \in \left \{ 1, 2, \ldots, n-1 \right \} \label{eq:VCT1} \\
\bar{s} = \frac{1}{n-1} \sum_{i=1}^{n-1} s_i \label{eq:VCT2} \\
\sigma_i = \frac {\sqrt{\frac{1}{n-1} \sum_{i=1}^{n-1} \left( s_i - \bar{s} \right)^2} } {\bar{G_i}} \label{eq:VCT3}
\end{gather}

\begin{equation}
\sigma_i < \ensuremath{1.1\times 10^{-4}}
\label{eq:VCTcr}
\end{equation}

\hypertarget{variability-in-the-shape-of-the-irradiance-measurements.}{%
\subsubsection{Variability in the Shape of the irradiance
Measurements.}\label{variability-in-the-shape-of-the-irradiance-measurements.}}

The maximum difference \(X\) between the change in measured irradiance
and the change in clear sky irradiance over each measurement interval.

\begin{gather}
x_i = \text{SDR}_{i+1} - \text{SDR}_{i} \forall i \in \left \{ 1, 2, \ldots, n-1 \right \} \label{eq:VSM1} \\
x_{CS,i} = \text{SDR}_{CS,i+1} - \text{SDR}_{CS,i} \forall i \in \left \{ 1, 2, \ldots, n-1 \right \} \label{eq:VSM2} \\
X_i = \max{\left \{ \left | x_i - x_{CS,i} \right | \right \}} \label{eq:VSM3}
\end{gather}

\begin{equation}
X_i < 7.5
\label{eq:VSMcr}
\end{equation}

\ldots.

In order to estimate the effect of the clouds on the SDR we created
three datasets, by characterizing each one-minute measurement with a
corresponding sky condition. The all-sky conditions, containing all the
valid measurements. The clear-sky conditions data, where we have
inferred that the sky was almost clear of clouds, and the remainder part
the data as cloudy sky conditions data (cloud-sky). To identify the
clear-sky conditions we used the method proposed by \citet{Long2000} and
by \citet{Reno2016}, that was adapted and configured for the site. We
have to note, that the definition of what constitutes as clear or cloudy
sky, has some subjectivity, due to the used method of characterization.
As a result, the details of the definition are site specific, it relies
on a combination of threshold of comparisons with ideal actinometric
models and statistics on different signal behavior.

\hypertarget{data-and-data-selection.}{%
\subsection{Data and data selection.}\label{data-and-data-selection.}}

Due to a significant measurement uncertainty near, the horizon, we have
to exclude all measurements with SZA greater than \(85^\circ\).
Moreover, due to some obstructions around the site (hills and
buildings), we excluded data with Azimuth angle between \(35^\circ\) and
\(120^\circ\) with SZA greater than \(80^\circ\). On the latter
instances, Sun is systematically, not visible by the instrument's
location. To make the measurements comparable throughout the dataset, we
adjusted all 1-minute radiometric values to the mean Sun - Earth
distance. Subsequently, we made all measurements relative to the Total
Solar Irradiance (TSI) at \(1 \text{au}\), in order to compensate for
the Sun's intensity variability, using a homogenized time series of
satellite TSI observations. The TSI data we use, is a combination of
data from NOAA \citep{Coddington2005} (for 1993-04-12 - 2011-03-31) and
adjusted data from \citet{TSIStsi} (for - ). As a result, we can present
all radiation data as a fraction of Sun's TSI.

\hypertarget{aggregation-of-radiometric-data}{%
\subsubsection{Aggregation of radiometric
data}\label{aggregation-of-radiometric-data}}

Before further analysis and deseasonalization, we implement an
appropriate aggregation scheme on the 1-minute data. To preserve the
representativeness we use the following criteria: a) for daily mean
values we exclude instances with less than 180 valid data points, b)
accordingly, monthly values are computed by daily aggregated data, where
months with less than 20 days are rejected. The daily seasonal values
were indexed based on the day of year number, and the monthly by the
corresponding calendar month. For the seasons of the year trends, we
grouped the mean daily values by season (December - February: Winter,
March - May: Spring, etc.). Finally, for each data set, we remove the
natural occurring seasonal variation. This is done, by calculating the
mean values for the appropriate time step, and subtracting the annual
cycle from the actual data. In addition to the previous aggregation
scheme, each dataset was aggregate in \(1^\circ\) SZA bins, separate for
cases before and after local noon. This gave us an approximation of
Sun's SZA contribution to the SDR brightening.

Thus, for each data set we have obtained the relative departure of the
seasonal mean. The statistical significance and aggregation scheme and
filtering will be noted along with the corresponding results.

\hypertarget{results}{%
\section{Results}\label{results}}

\hypertarget{long-term-trends}{%
\subsection{Long-term trends}\label{long-term-trends}}

Using the mean daily SDR we produce the trends of each data set, as
departure from the seasonal value. Only for All-sky (Fig.
\ref{fig:trendALL}) conditions, the statistical significance of the
trends are acceptable (Tab. \ref{tab:trendtable}), to draw some
conclusions.

\begin{figure}[h!]

{\centering \includegraphics[width=0.7\linewidth]{./images/LongtermTrends-2}

}

\caption{Anomaly (\%) of the daily SDR relative to climatological values for 1993 - 2023. The black line shows the long term trend for all-sky conditions.}\label{fig:trendALL}
\end{figure}

\begin{longtable}[]{@{}
  >{\centering\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2048}}
  >{\centering\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1928}}
  >{\centering\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2169}}
  >{\centering\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.3855}}@{}}
\caption{\label{tab:trendtable}Trends of daily means by sky conditions.
}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\centering
Trend {[}\%/year{]}
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Trend p-value
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Sky condition
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Trend statistical signif. {[}\%{]}
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\centering
Trend {[}\%/year{]}
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Trend p-value
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Sky condition
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Trend statistical signif. {[}\%{]}
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
0.4221 & 5.871e-06 & All sky cond. & 100 \\
0.3871 & 0.0001123 & Clear sky cond. & 99.99 \\
0.415 & 0.002237 & Cloudy cond. & 99.78 \\
\end{longtable}

\hypertarget{long-term-trends-by-solar-zenith-angle.}{%
\subsection{Long term trends by Solar zenith
angle.}\label{long-term-trends-by-solar-zenith-angle.}}

Analyzing the long term trends of SDR to bins of SZA, we can see the
contribution of the geometry and time in a diurnal level (Figures
\ref{fig:szatrends}a and \ref{fig:szatrends}b). Although there is a
seasonal distribution of SZA than in not presented here.

\begin{figure}[h!]

{\centering \subfloat[Trend distribution for all-sky conditions.\label{fig:szatrends-1}]{\includegraphics[width=.35\linewidth]{./images/SzaTrends-13} }\subfloat[Trend distribution for cloud-sky conditions.\label{fig:szatrends-2}]{\includegraphics[width=.35\linewidth]{./images/SzaTrends-21} }

}

\caption{Distribution of the SDR's long term trends by SZA.}\label{fig:szatrends}
\end{figure}

\hypertarget{long-term-trends-by-season-of-year}{%
\subsection{Long term trends by season of
year}\label{long-term-trends-by-season-of-year}}

Similar we have produced the deseasonalized trend for different sky
conditions for each season of the year, using the corresponding mean
monthly values. This can give us a better understanding of the annual
variability of the trends.

\begin{figure}[h!]

{\centering \includegraphics[width=1\linewidth]{./images/SeasonalTrendsTogether-1}

}

\caption{Trends by season for all condition. Displaying monthly means of daily means.}\label{fig:seasonalALL}
\end{figure}

\begin{longtable}[]{@{}
  >{\centering\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2361}}
  >{\centering\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2222}}
  >{\centering\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2500}}
  >{\centering\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1250}}@{}}
\caption{\label{tab:trendseasontable}Trends of daily means by sky
conditions for the seasons of the year. (continued
below)}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\centering
Trend {[}\%/year{]}
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Trend p-value
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Sky condition
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Season
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\centering
Trend {[}\%/year{]}
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Trend p-value
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Sky condition
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Season
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
0.8014 & 0.1606 & All sky cond. & Winter \\
-0.1999 & 0.5397 & All sky cond. & Spring \\
0.2777 & 0.09127 & All sky cond. & Summer \\
0.3768 & 0.2006 & All sky cond. & Autumn \\
0.9574 & 0.1081 & Clear sky cond. & Winter \\
-0.4227 & 0.2207 & Clear sky cond. & Spring \\
0.0931 & 0.6638 & Clear sky cond. & Summer \\
0.3233 & 0.3255 & Clear sky cond. & Autumn \\
-0.4354 & 0.4924 & Cloudy cond. & Winter \\
0.8353 & 0.04263 & Cloudy cond. & Spring \\
0.7051 & 0.05457 & Cloudy cond. & Summer \\
0.3708 & 0.2357 & Cloudy cond. & Autumn \\
\end{longtable}

\begin{longtable}[]{@{}
  >{\centering\arraybackslash}p{(\columnwidth - 0\tabcolsep) * \real{0.4444}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\centering
Trend statistical signif. {[}\%{]}
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
83.94 \\
46.03 \\
90.87 \\
79.94 \\
89.19 \\
77.93 \\
33.62 \\
67.45 \\
50.76 \\
95.74 \\
94.54 \\
76.43 \\
\end{longtable}

\hypertarget{consistency-of-the-trends}{%
\subsection{Consistency of the trends}\label{consistency-of-the-trends}}

A method to evaluate changes in the long term trend is to use the
cumulative sum of the variable.

Figure \ref{fig:cumsum}

\ldots{} different in aggregation \ldots{}

\begin{figure}[h!]

{\centering \subfloat[one plot\label{fig:cumsum-1}]{\includegraphics[width=.35\linewidth]{./images/CumulativeDailyCumSum-1} }\subfloat[the other one\label{fig:cumsum-2}]{\includegraphics[width=.35\linewidth]{./images/CumulativeDailyCumSum-9} }

}

\caption{two plots}\label{fig:cumsum}
\end{figure}

\hypertarget{conclusions}{%
\section{Conclusions}\label{conclusions}}

Our result for all-sky condition (\(0.42\%/year\)), reaffirm the
previous results of \citet{Bais2013} for the site. The increase of this
trend, shows that the phenomenon and probably the causes have been
amplified.

\ldots.. Also, the trend of \(0.41\%/year\) for cloud-sky condition
indicate the major part that clouds play on the \ldots..

\ldots\ldots.

About:

\begin{itemize}
\item
  long terms trends
\item
  seasonal trends
\item
  effect of SZA
\item
  effect of clouds
\item
  Aerosols
\end{itemize}

\begin{longtable}[]{@{}c@{}}
\toprule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\textbf{END} \\
\end{longtable}

\bibliography{references.bib}

\end{document}

### 2023-06-26 Article_A.tex
Commit 2023-06-26 07:17

--- 2023-06-26 Article.tex
@@
-\documentclass[preprint, 3p,
-authoryear]{elsarticle} %review=doublespace preprint=single 5p=2 column
-%%% Begin My package additions %%%%%%%%%%%%%%%%%%%
-
-\usepackage[hyphens]{url}
-
-  \journal{An awesome journal} % Sets Journal name
-
-\usepackage{graphicx}
-%%%%%%%%%%%%%%%% end my additions to header
-
+% Options for packages loaded elsewhere
+\PassOptionsToPackage{unicode}{hyperref}
+\PassOptionsToPackage{hyphens}{url}
+%
+\documentclass[
+  preprint, 3p, authoryear]{article}
+\usepackage{amsmath,amssymb}
+\usepackage{lmodern}
+\usepackage{iftex}
+\ifPDFTeX
   \usepackage[T1]{fontenc}
-\usepackage{lmodern}
-\usepackage{amssymb,amsmath}
-% TODO: Currently lineno needs to be loaded after amsmath because of conflict
-% https://github.com/latex-lineno/lineno/issues/5
-\usepackage{lineno} % add
-\usepackage{ifxetex,ifluatex}
-\usepackage{fixltx2e} % provides \textsubscript
-% use upquote if available, for straight quotes in verbatim environments
+  \usepackage[utf8]{inputenc}
+  \usepackage{textcomp} % provide euro and other symbols
+\else % if luatex or xetex
+  \usepackage{unicode-math}
+  \defaultfontfeatures{Scale=MatchLowercase}
+  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
+\fi
+% Use upquote if available, for straight quotes in verbatim environments
 \IfFileExists{upquote.sty}{\usepackage{upquote}}{}
-\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
-  \usepackage[utf8]{inputenc}
-\else % if luatex or xelatex
-  \usepackage{fontspec}
-  \ifxetex
-    \usepackage{xltxtra,xunicode}
-  \fi
-  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
-  \newcommand{\euro}{€}
-\fi
-% use microtype if available
-\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
-\usepackage[]{natbib}
-\bibliographystyle{plainnat}
-
-\ifxetex
-  \usepackage[setpagesize=false, % page size defined by xetex
-              unicode=false, % unicode breaks when used with xetex
-              xetex]{hyperref}
-\else
-  \usepackage[unicode=true]{hyperref}
-\fi
-\hypersetup{breaklinks=true,
-            bookmarks=true,
-            pdfauthor={},
+\IfFileExists{microtype.sty}{% use microtype if available
+  \usepackage[]{microtype}
+  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
+}{}
+\makeatletter
+\@ifundefined{KOMAClassName}{% if non-KOMA class
+  \IfFileExists{parskip.sty}{%
+    \usepackage{parskip}
+  }{% else
+    \setlength{\parindent}{0pt}
+    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
+}{% if KOMA class
+  \KOMAoptions{parskip=half}}
+\makeatother
+\usepackage{xcolor}
+\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
+\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
+\hypersetup{
   pdftitle={Trends of SDR in Thessaloniki},
-            colorlinks=false,
-            urlcolor=blue,
-            linkcolor=magenta,
-            pdfborder={0 0 0}}
-
-\setcounter{secnumdepth}{5}
-% Pandoc toggle for numbering sections (defaults to be off)
-
-% tightlist command for lists without linebreak
-\providecommand{\tightlist}{%
-  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
-
-% From pandoc table feature
+  pdfkeywords={GHI, SDR, Solar Brigthening/Dimming},
+  hidelinks,
+  pdfcreator={LaTeX via pandoc}}
+\urlstyle{same} % disable monospaced font for URLs
+\usepackage[margin=1in]{geometry}
 \usepackage{longtable,booktabs,array}
@@
 \makesavenoteenv{longtable}
+\usepackage{graphicx}
+\makeatletter
+\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
+\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
+\makeatother
+% Scale images if necessary, so that they will not overflow the page
+% margins by default, and it is still possible to overwrite the defaults
+% using explicit options in \includegraphics[width, height, ...]{}
+\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
+% Set default figure placement to htbp
+\makeatletter
+\def\fps@figure{htbp}
+\makeatother
+\setlength{\emergencystretch}{3em} % prevent overfull lines
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+\setcounter{secnumdepth}{5}
+\newlength{\cslhangindent}
+\setlength{\cslhangindent}{1.5em}
+\newlength{\csllabelwidth}
+\setlength{\csllabelwidth}{3em}
+\newlength{\cslentryspacingunit} % times entry-spacing
+\setlength{\cslentryspacingunit}{\parskip}
+\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
+ {% don't indent paragraphs
+  \setlength{\parindent}{0pt}
+  % turn on hanging indent if param 1 is 1
+  \ifodd #1
+  \let\oldpar\par
+  \def\par{\hangindent=\cslhangindent\oldpar}
+  \fi
+  % set entry spacing
+  \setlength{\parskip}{#2\cslentryspacingunit}
+ }%
+ {}
+\usepackage{calc}
+\newcommand{\CSLBlock}[1]{#1\hfill\break}
+\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
+\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
+\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
+\usepackage{subcaption}
+\ifLuaTeX
+  \usepackage{selnolig}  % disable illegal ligatures
+\fi
 
-\usepackage{subcaption}
+\title{Trends of SDR in Thessaloniki}
+\author{true \and true}
+\date{2023-06-26}
 
 \begin{document}
-
-\begin{frontmatter}
-
-  \title{Trends of SDR in Thessaloniki}
-    \author[Laboratory of Atmospheric Physics]{Natsis Athanasios%
-  \corref{cor1}%
-  }
-   \ead{natsisphysicist@gmail.com} 
-    \author[Laboratory of Atmospheric Physics]{Alkiviadis Bais%
-  %
-  }
-   \ead{abais@auth.gr} 
-      \affiliation[Laboratory of Atmospheric Physics]{
-    organization={Aristotle University of
-Thessaloniki},addressline={Campus Box
-149},city={Thessaloniki},postcode={54124},country={Greece},}
-    \cortext[cor1]{Corresponding author}
-
+\maketitle
 \begin{abstract}
@@
 \end{abstract}
-    \begin{keyword}
-    GHI \sep SDR \sep 
-    Solar Brigthening/Dimming
-  \end{keyword}
 
- \end{frontmatter}
-
+{
+\setcounter{tocdepth}{4}
+\tableofcontents
+}
 \hypertarget{introduction.}{%
@@
 
-The shortwave downward solar irradiance (SDR) at the Earth's surface
-plays a significant role, on Earths climate. Changes of the SDR can be
-related to major changes on factors affecting the radiative forcing
-balance and human activities. Multiple studies attempt to evaluate the
-phenomenon and the potential causes
-\citep{Ohvril2009, Schwarz2020, Wild2009, Wild2012, Xia2007, Zerefos2009}.
-Although, there is regional variation of SDR, with both positive and
-negative trends, most of the studies agree that the major factors are
-the interaction of aerosols and clouds.
+The shortwave downward solar irradiance (SDR) at the Earth's surface plays a significant role, on Earths climate.
+Changes of the SDR can be related to major changes on factors affecting the radiative forcing balance and human activities.
+Multiple studies attempt to evaluate the phenomenon and the potential causes (Ohvril et al. 2009; Schwarz et al. 2020; Wild 2009, 2012; Xia et al. 2007; Zerefos et al. 2009).
+Although, there is regional variation of SDR, with both positive and negative trends, most of the studies agree that the major factors are the interaction of aerosols and clouds.
 
-In this study, we examine the trends of SDR with ground-based
-measurements at Thessaloniki for the period 1993 - 2023, as derived from
-a CM-21 pyranometer. We reevaluate and extend the dataset used by
-\citet{Bais2013}, applying a different algorithm for the identification
-of clear-/cloud-sky instances, and we derive the radiation trends for
-this period under different sky conditions (all-sky, clear-sky and
-cloud-sky).
+In this study, we examine the trends of SDR with ground-based measurements at Thessaloniki for the period 1993 - 2023, as derived from a CM-21 pyranometer.
+We reevaluate and extend the dataset used by Bais et al.
+(2013), applying a different algorithm for the identification of clear-/cloud-sky instances, and we derive the radiation trends for this period under different sky conditions (all-sky, clear-sky and cloud-sky).
 
@@
 
-Our method to characterize the sky conditions and the definition of each
-sky condition has some subjectivity. The algorithm was calibrated with
-the main focus to identify the presence of clouds on the sky dome,
-although there are marginal cases that there will be false positives or
-false negatives identifications.
+Our method to characterize the sky conditions and the definition of each sky condition has some subjectivity.
+The algorithm was calibrated with the main focus to identify the presence of clouds on the sky dome, although there are marginal cases that there will be false positives or false negatives identifications.
 
@@
 
-The SDR is equivalent to the whole sky Global radiation, also refered as
-Global Horizontal Irradiance (GHI), and was obtained with a horizontal
-leveled, CM-21 pyranometer. SDR data spans the period of 1993-04-12 to
-2011-03-31. In some algorithms we used as auxiliary data, the direct
-beam radiation (DNI) for comparisons, obtained by a CHP-1 pyrheliometer,
-with data availability from Inf to -Inf.
+The SDR is equivalent to the whole sky Global radiation, also refered as Global Horizontal Irradiance (GHI), and was obtained with a horizontal leveled, CM-21 pyranometer.
+SDR data spans the period of
+1993-04-12 to 2011-03-31.
+In some algorithms we used as auxiliary data, the direct beam radiation (DNI) for comparisons, obtained by a CHP-1 pyrheliometer, with data availability from
+Inf to -Inf.
 
@@
 
-There are three distinct steps to the creation of this dataset: a) the
-acquisition of radiation measurements from the sensors, b) a radiation
-data quality check, and c) the identification of ``clear sky''
-conditions from the radiometric data d) data aggregation and trend
-analysis.
+There are three distinct steps to the creation of this dataset:
+a) the acquisition of radiation measurements from the sensors,
+b) a radiation data quality check, and
+c) the identification of ``clear sky'' conditions from the radiometric data
+d) data aggregation and trend analysis.
 
-For the acquisition of radiometric data, the signal of the broadband
-instruments is sampled with a rate of \(1 \text{Hz}\). The mean and
-standard deviation values are recorded for every minute. The
-measurements are corrected for the zero offset of the instrument signal.
-As reference of the ``dark signal'' is used the Sun elevation angle
-bellow \(-10^\circ\), for a period of \(3 \text{h}\). The signal is
-converted to radiation flux, using a ramped value of the instrument
-sensitivity, derived from the eight laboratory calibrations of the
-instrument, during the study period.
+For the acquisition of radiometric data,
+the signal of the broadband instruments is sampled with a rate of \(1 \text{Hz}\). The mean and standard deviation values are recorded for every minute.
+The measurements are corrected for the zero offset of the instrument signal.
+As reference of the ``dark signal'' is used the Sun elevation angle bellow \(-10^\circ\), for a period of \(3 \text{h}\).
+The signal is converted to radiation flux, using a ramped value of the instrument sensitivity, derived from the eight laboratory calibrations of the instrument, during the study period.
 
-A manual screening is performed, to remove inconsistencies and erroneous
-recordings, that can occur randomly or systematically, during the long
-continuous operation of the instrument. The manual screening is aided by
-a radiation data quality assurance procedure, adjusted for this site,
-based on methods of Long and
-Shi~\citetext{\citeyear{long_automated_2008}; \citeyear{Long2006}}.
+A manual screening is performed, to remove inconsistencies and erroneous recordings, that can occur randomly or systematically, during the long continuous operation of the instrument.
+The manual screening is aided by a radiation data quality assurance procedure, adjusted for this site, based on methods of Long and Shi~(2008, 2006).
 Thus, problematic recordings have been excluded from further processing.
-Although it is not possible to detect all the bad data, the large number
-of data and the aggregation scheme we use, can provide us with accurate
-radiometric measurements, for the scope of this study.
+Although it is not possible to detect all the bad data, the large number of data and the aggregation scheme we use, can provide us with accurate radiometric measurements, for the scope of this study.
 
@@
 
-As global radiation clear sky reference we are using the Haurwitz's
-model, adjusted for our site with a factor of 0.965 (Eq. \ref{eq:ahau}).
-The selection of a clear sky reference model, was based on SDR
-observation from the period 2016 -- 2021. Where, after an iterative
-optimization of eight simple models (Daneshyar--Paltridge--Proctor,
-Kasten--Czeplak, Haurwitz, Berger--Duffie,
-Adnot--Bourges--Campana--Gicquel, Robledo-Soler, Kasten and
-Ineichen-Perez) with different factors. We found, that Haurwitz's model,
-adjusted with a factor of 0.965 has the lower root mean squared error
-(RMSE). The tried models are described by \citet{Reno2012} and tested by
-\citet{Reno2016}. The iterative optimization method, for the selection
-of the reference is discussed by \citet{Long2000} and \citet{Reno2016}.
+As global radiation clear sky reference we are using the Haurwitz's model, adjusted for our site with a factor of 0.965 (Eq. \ref{eq:ahau}).
+The selection of a clear sky reference model, was based on SDR observation from the period 2016 -- 2021.
+Where, after an iterative optimization of eight simple models (Daneshyar--Paltridge--Proctor, Kasten--Czeplak, Haurwitz, Berger--Duffie, Adnot--Bourges--Campana--Gicquel, Robledo-Soler, Kasten and Ineichen-Perez) with different factors.
+We found, that Haurwitz's model, adjusted with a factor of 0.965 has the lower root mean squared error (RMSE).
+The tried models are described by Reno, Hansen, and Stein (2012) and tested by Reno and Hansen (2016).
+The iterative optimization method, for the selection of the reference is discussed by Charles N. Long and Ackerman (2000) and Reno and Hansen (2016).
 
@@
 
-The following criteria and thresholds were used to identify clear-sky
-conditions. Each criterion was applied for a running window of \(11\)
-consecutive measurements/minutes, and the characterization is applied at
-the center value of the window. A data point is consider as under
-cloud-sky condition if it fails to pass any of the criteria, all the
-other data points are characterized as clear-sky.
+The following criteria and thresholds were used to identify clear-sky conditions.
+Each criterion was applied for a running window of \(11\) consecutive measurements/minutes, and the characterization is applied at the center value of the window.
+A data point is consider as under cloud-sky condition if it fails to pass any of the criteria, all the other data points are characterized as clear-sky.
 
 \hypertarget{mean-value-of-irradiance-during-the-time-period.}{%
-\subsubsection{Mean value of irradiance during the time
-period.}\label{mean-value-of-irradiance-during-the-time-period.}}
+\subsubsection{Mean value of irradiance during the time period.}\label{mean-value-of-irradiance-during-the-time-period.}}
 
-The mean of the measured value \(\overline{G}_i\) must be inside an
-envelope based on the reference model \(\text{SDR}_\text{Clear Sky}\)
-(Eq. \ref{eq:MeanVIP}).
+The mean of the measured value \(\overline{G}_i\) must be inside an envelope based on the reference model \(\text{SDR}_\text{Clear Sky}\) (Eq. \ref{eq:MeanVIP}).
 
@@
 \hypertarget{max-value-of-irradiance-during-the-time-period.}{%
-\subsubsection{Max value of irradiance during the time
-period.}\label{max-value-of-irradiance-during-the-time-period.}}
+\subsubsection{Max value of irradiance during the time period.}\label{max-value-of-irradiance-during-the-time-period.}}
 
-The running max measured value \(M_{Gi} = max[\text{SDR}_{i}]\), is
-compared to a similar constructed value from the reference
-\(M_{CSi} = max[\text{SDR}_{i\text{Clear Sky}}]\) (Eq. \ref{eq:MaxVIP}).
+The running max measured value \(M_{Gi} = max[\text{SDR}_{i}]\), is compared to a similar constructed value from the reference \(M_{CSi} = max[\text{SDR}_{i\text{Clear Sky}}]\) (Eq. \ref{eq:MaxVIP}).
 
@@
 \hypertarget{variability-in-irradiance-by-line-length.}{%
-\subsubsection{Variability in irradiance by line
-length.}\label{variability-in-irradiance-by-line-length.}}
+\subsubsection{Variability in irradiance by line length.}\label{variability-in-irradiance-by-line-length.}}
 
-The length \(L\) (Eq. \ref{eq:VILeq}) of the sequence of line segments
-connecting the points of the SDR time series for the measured values
-\(L\) and similar for the reference \(L_{CS}\), must be within the
-limits of Eq. \ref{eq:VILcr}.
+The length \(L\) (Eq. \ref{eq:VILeq}) of the sequence of line segments connecting the points of the SDR time
+series for the measured values \(L\) and similar for the reference \(L_{CS}\), must be within the limits of Eq. \ref{eq:VILcr}.
 
@@
 \hypertarget{variance-of-changes-in-the-time-series.}{%
-\subsubsection{Variance of Changes in the Time
-series.}\label{variance-of-changes-in-the-time-series.}}
+\subsubsection{Variance of Changes in the Time series.}\label{variance-of-changes-in-the-time-series.}}
 
-We calculate the standard deviation \(\sigma\) of the slope (\(s\))
-between sequential points in the time series, normalized by the average
-SDR during the time interval.
+We calculate the standard deviation \(\sigma\) of the slope
+(\(s\)) between sequential points in the time series, normalized by the average SDR during the time interval.
 
@@
 \hypertarget{variability-in-the-shape-of-the-irradiance-measurements.}{%
-\subsubsection{Variability in the Shape of the irradiance
-Measurements.}\label{variability-in-the-shape-of-the-irradiance-measurements.}}
+\subsubsection{Variability in the Shape of the irradiance Measurements.}\label{variability-in-the-shape-of-the-irradiance-measurements.}}
 
-The maximum difference \(X\) between the change in measured irradiance
-and the change in clear sky irradiance over each measurement interval.
+The maximum difference \(X\) between the change in measured irradiance and the change in clear sky
+irradiance over each measurement interval.
 
@@
 
-In order to estimate the effect of the clouds on the SDR we created
-three datasets, by characterizing each one-minute measurement with a
-corresponding sky condition. The all-sky conditions, containing all the
-valid measurements. The clear-sky conditions data, where we have
-inferred that the sky was almost clear of clouds, and the remainder part
-the data as cloudy sky conditions data (cloud-sky). To identify the
-clear-sky conditions we used the method proposed by \citet{Long2000} and
-by \citet{Reno2016}, that was adapted and configured for the site. We
-have to note, that the definition of what constitutes as clear or cloudy
-sky, has some subjectivity, due to the used method of characterization.
-As a result, the details of the definition are site specific, it relies
-on a combination of threshold of comparisons with ideal actinometric
-models and statistics on different signal behavior.
+In order to estimate the effect of the clouds on the SDR we created three datasets, by characterizing each one-minute measurement with a corresponding sky condition.
+The all-sky conditions, containing all the valid measurements.
+The clear-sky conditions data, where we have inferred that the sky was almost clear of clouds, and the remainder part the data as cloudy sky conditions data (cloud-sky).
+To identify the clear-sky conditions we used the method proposed by Charles N. Long and Ackerman (2000) and by Reno and Hansen (2016), that was adapted and configured for the site.
+We have to note, that the definition of what constitutes as clear or cloudy sky, has some subjectivity, due to the used method of characterization.
+As a result, the details of the definition are site specific, it relies on a combination of threshold of comparisons with ideal actinometric models and statistics on different signal behavior.
 
@@
 
-Due to a significant measurement uncertainty near, the horizon, we have
-to exclude all measurements with SZA greater than \(85^\circ\).
-Moreover, due to some obstructions around the site (hills and
-buildings), we excluded data with Azimuth angle between \(35^\circ\) and
-\(120^\circ\) with SZA greater than \(80^\circ\). On the latter
-instances, Sun is systematically, not visible by the instrument's
-location. To make the measurements comparable throughout the dataset, we
-adjusted all 1-minute radiometric values to the mean Sun - Earth
-distance. Subsequently, we made all measurements relative to the Total
-Solar Irradiance (TSI) at \(1 \text{au}\), in order to compensate for
-the Sun's intensity variability, using a homogenized time series of
-satellite TSI observations. The TSI data we use, is a combination of
-data from NOAA \citep{Coddington2005} (for 1993-04-12 - 2011-03-31) and
-adjusted data from \citet{TSIStsi} (for - ). As a result, we can present
-all radiation data as a fraction of Sun's TSI.
+Due to a significant measurement uncertainty near, the horizon, we have to exclude
+all measurements with SZA greater than \(85^\circ\).
+Moreover, due to some obstructions around the site (hills and buildings), we excluded data with Azimuth angle between
+\(35^\circ\) and \(120^\circ\) with SZA greater than \(80^\circ\).
+On the latter instances, Sun is systematically, not visible by the instrument's location.
+To make the measurements comparable throughout the dataset, we adjusted all 1-minute radiometric values to the mean Sun - Earth distance.
+Subsequently, we made all measurements relative to the Total Solar Irradiance (TSI) at \(1 \text{au}\), in order to compensate for the Sun's intensity variability, using a homogenized time series of satellite TSI observations.
+The TSI data we use, is a combination of data from NOAA (Coddington et al. 2005) (for 1993-04-12 - 2011-03-31) and adjusted data from LASP (n.d.) (for - ).
+As a result, we can present all radiation data as a fraction of Sun's TSI.
 
 \hypertarget{aggregation-of-radiometric-data}{%
-\subsubsection{Aggregation of radiometric
-data}\label{aggregation-of-radiometric-data}}
+\subsubsection{Aggregation of radiometric data}\label{aggregation-of-radiometric-data}}
 
-Before further analysis and deseasonalization, we implement an
-appropriate aggregation scheme on the 1-minute data. To preserve the
-representativeness we use the following criteria: a) for daily mean
-values we exclude instances with less than 180 valid data points, b)
-accordingly, monthly values are computed by daily aggregated data, where
-months with less than 20 days are rejected. The daily seasonal values
-were indexed based on the day of year number, and the monthly by the
-corresponding calendar month. For the seasons of the year trends, we
-grouped the mean daily values by season (December - February: Winter,
-March - May: Spring, etc.). Finally, for each data set, we remove the
-natural occurring seasonal variation. This is done, by calculating the
-mean values for the appropriate time step, and subtracting the annual
-cycle from the actual data. In addition to the previous aggregation
-scheme, each dataset was aggregate in \(1^\circ\) SZA bins, separate for
-cases before and after local noon. This gave us an approximation of
-Sun's SZA contribution to the SDR brightening.
+Before further analysis and deseasonalization, we implement an appropriate aggregation scheme on the 1-minute data. To preserve the representativeness we use the following criteria:
+a) for daily mean values we exclude instances with less than 180 valid data points,
+b) accordingly, monthly values are computed by daily aggregated data, where months with less than 20 days are rejected.
+The daily seasonal values were indexed based on the day of year number, and the monthly by the corresponding calendar month.
+For the seasons of the year trends, we grouped the mean daily values by season (December - February: Winter, March - May: Spring, etc.).
+Finally, for each data set, we remove the natural occurring seasonal variation.
+This is done, by calculating the mean values for the appropriate time step, and subtracting the annual cycle from the actual data.
+In addition to the previous aggregation scheme, each dataset was aggregate in \(1^\circ\) SZA bins, separate for cases before and after local noon.
+This gave us an approximation of Sun's SZA contribution to the SDR brightening.
 
-Thus, for each data set we have obtained the relative departure of the
-seasonal mean. The statistical significance and aggregation scheme and
-filtering will be noted along with the corresponding results.
+Thus, for each data set we have obtained the relative departure of the seasonal mean.
+The statistical significance and aggregation scheme and filtering will be noted along with the corresponding results.
 
@@
 
-Using the mean daily SDR we produce the trends of each data set, as
-departure from the seasonal value. Only for All-sky (Fig.
-\ref{fig:trendALL}) conditions, the statistical significance of the
-trends are acceptable (Tab. \ref{tab:trendtable}), to draw some
-conclusions.
+Using the mean daily SDR we produce the trends of each data set, as departure from the seasonal value.
+Only for All-sky (Fig. \ref{fig:trendALL}) conditions, the statistical significance of the trends are acceptable (Tab. \ref{tab:trendtable}), to draw some conclusions.
 
@@
 }\tabularnewline
-\toprule\noalign{}
+\toprule
 \begin{minipage}[b]{\linewidth}\centering
@@
 \end{minipage} \\
-\midrule\noalign{}
+\midrule
 \endfirsthead
-\toprule\noalign{}
+\toprule
 \begin{minipage}[b]{\linewidth}\centering
@@
 \end{minipage} \\
-\midrule\noalign{}
+\midrule
 \endhead
-\bottomrule\noalign{}
-\endlastfoot
 0.4221 & 5.871e-06 & All sky cond. & 100 \\
@@
 0.415 & 0.002237 & Cloudy cond. & 99.78 \\
+\bottomrule
 \end{longtable}
@@
 \hypertarget{long-term-trends-by-solar-zenith-angle.}{%
-\subsection{Long term trends by Solar zenith
-angle.}\label{long-term-trends-by-solar-zenith-angle.}}
+\subsection{Long term trends by Solar zenith angle.}\label{long-term-trends-by-solar-zenith-angle.}}
 
-Analyzing the long term trends of SDR to bins of SZA, we can see the
-contribution of the geometry and time in a diurnal level (Figures
-\ref{fig:szatrends}a and \ref{fig:szatrends}b). Although there is a
-seasonal distribution of SZA than in not presented here.
+Analyzing the long term trends of SDR to bins of SZA, we can see the contribution of the geometry and time in a diurnal level (Figures \ref{fig:szatrends}a and \ref{fig:szatrends}b).
+Although there is a seasonal distribution of SZA than in not presented here.
 
@@
 \hypertarget{long-term-trends-by-season-of-year}{%
-\subsection{Long term trends by season of
-year}\label{long-term-trends-by-season-of-year}}
+\subsection{Long term trends by season of year}\label{long-term-trends-by-season-of-year}}
 
-Similar we have produced the deseasonalized trend for different sky
-conditions for each season of the year, using the corresponding mean
-monthly values. This can give us a better understanding of the annual
-variability of the trends.
+Similar we have produced the deseasonalized trend for different sky conditions for each season of the year, using the corresponding mean monthly values.
+This can give us a better understanding of the annual variability of the trends.
 
@@
   >{\centering\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1250}}@{}}
-\caption{\label{tab:trendseasontable}Trends of daily means by sky
-conditions for the seasons of the year. (continued
-below)}\tabularnewline
-\toprule\noalign{}
+\caption{\label{tab:trendseasontable}Trends of daily means by sky conditions for the seasons of the year. (continued below)}\tabularnewline
+\toprule
 \begin{minipage}[b]{\linewidth}\centering
@@
 \end{minipage} \\
-\midrule\noalign{}
+\midrule
 \endhead
-\bottomrule\noalign{}
-\endlastfoot
 0.8014 & 0.1606 & All sky cond. & Winter \\
@@
 0.3708 & 0.2357 & Cloudy cond. & Autumn \\
+\bottomrule
 \end{longtable}
@@
   >{\centering\arraybackslash}p{(\columnwidth - 0\tabcolsep) * \real{0.4444}}@{}}
-\toprule\noalign{}
+\toprule
 \begin{minipage}[b]{\linewidth}\centering
@@
 \end{minipage} \\
-\midrule\noalign{}
+\midrule
 \endhead
-\bottomrule\noalign{}
-\endlastfoot
 83.94 \\
@@
 76.43 \\
+\bottomrule
 \end{longtable}
@@
 
-A method to evaluate changes in the long term trend is to use the
-cumulative sum of the variable.
+A method to evaluate changes in the long term trend is to use the cumulative sum of the variable.
 
@@
 
-Our result for all-sky condition (\(0.42\%/year\)), reaffirm the
-previous results of \citet{Bais2013} for the site. The increase of this
-trend, shows that the phenomenon and probably the causes have been
-amplified.
+Our result for all-sky condition (\(0.42\%/year\)), reaffirm the previous results of Bais et al. (2013) for the site.
+The increase of this trend, shows that the phenomenon and probably the causes have been amplified.
 
-\ldots.. Also, the trend of \(0.41\%/year\) for cloud-sky condition
-indicate the major part that clouds play on the \ldots..
+\ldots..
+Also, the trend of \(0.41\%/year\) for cloud-sky condition indicate the major part that clouds play on the
+\ldots..
 
@@
 \begin{longtable}[]{@{}c@{}}
-\toprule\noalign{}
+\toprule
 \endhead
-\bottomrule\noalign{}
-\endlastfoot
 \textbf{END} \\
+\bottomrule
 \end{longtable}
 
-\bibliography{references.bib}
+\hypertarget{refs}{}
+\begin{CSLReferences}{1}{0}
+\leavevmode\vadjust pre{\hypertarget{ref-Bais2013}{}}%
+Bais, A.
+F., T.
+Drosoglou, C.
+Meleti, K.
+Tourpali, and N.
+Kouremeti.
+2013.
+{``Changes in Surface Shortwave Solar Irradiance from 1993 to 2011 at Thessaloniki (Greece).''} \emph{International Journal of Climatology} 33 (13): 2871--76.
+\url{https://doi.org/f5dzz5}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-Coddington2005}{}}%
+Coddington, Odele, Judith L.
+Lean, Doug Lindholm, Peter Pilewskie amd Martin Snow, and NOAA CDR Program.
+2005.
+{``{NOAA} Climate Data Record ({CDR}) of Total Solar Irradiance ({TSI}), {NRLTSI} Version 2.
+{D}aily.''} 2005.
+\url{https://doi.org/10.7289/V55B00C1}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-TSIStsi}{}}%
+LASP. n.d. {``{TSIS-1} Total Solar Irradiance - Six Hour Average.''} Accessed June 1, 2023. \url{https://lasp.colorado.edu/tsis/data/}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-Long2006}{}}%
+Long, C.
+N., and Y.
+Shi.
+2006.
+{``The QCRad Value Added Product: Surface Radiation Measurement Quality Control Testing, Including Climatology Configurable Limits.''} DOE/SC-ARM/TR-074.
+Office of Science, Office of Biological; Environmental Research, U.S.
+Department of Energy.
+
+\leavevmode\vadjust pre{\hypertarget{ref-long_automated_2008}{}}%
+---------. 2008. {``An {Automated} {Quality} {Assessment} and {Control} {Algorithm} for {Surface} {Radiation} {Measurements}.''} \emph{The Open Atmospheric Science Journal}, 23--37.
+
+\leavevmode\vadjust pre{\hypertarget{ref-Long2000}{}}%
+Long, Charles N., and Thomas P.
+Ackerman.
+2000.
+{``{Identification of clear skies from broadband pyranometer measurements and calculation of downwelling shortwave cloud effects}.''} \emph{Journal of Geophysical Research} 105 (D12): 15609--26.
+\url{https://doi.org/10.1029/2000JD900077}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-Ohvril2009}{}}%
+Ohvril, Hanno, Hilda Teral, Lennart Neiman, Martin Kannel, Marika Uustare, Mati Tee, Viivi Russak, et al.
+2009.
+{``Global Dimming and Brightening Versus Atmospheric Column Transparency, Europe, 1906--2007.''} \emph{Journal of Geophysical Research} 114 (May).
+\url{https://doi.org/10.1029/2008JD010644}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-Reno2016}{}}%
+Reno, Matthew J., and Clifford W.
+Hansen.
+2016.
+{``{Identification of periods of clear sky irradiance in time series of GHI measurements}.''} \emph{Renewable Energy} 90: 520--31.
+\url{https://doi.org/10.1016/j.renene.2015.12.031}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-Reno2012}{}}%
+Reno, Matthew J., Clifford W. Hansen, and Joshua S. Stein. 2012. {``Global Horizontal Irradiance Clear Sky Models: Implementation and Analysis.''} SAND2012-2389, 1039404. \url{https://doi.org/gq5npv}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-Schwarz2020}{}}%
+Schwarz, M., D.
+Folini, S.
+Yang, R.
+P.
+Allan, and M.
+Wild.
+2020.
+{``Changes in Atmospheric Shortwave Absorption as Important Driver of Dimming and Brightening.''} \emph{Nature Geoscience} 13 (2): 110--15.
+\url{https://doi.org/10.1038/s41561-019-0528-y}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-Wild2009}{}}%
+Wild, Martin. 2009. {``Global Dimming and Brightening: A Review.''} \emph{Journal of Geophysical Research Atmospheres} 114 (12): 1--31. \url{https://doi.org/bcq}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-Wild2012}{}}%
+---------. 2012. {``Enlightening Global Dimming and Brightening.''} \emph{Bulletin of the American Meteorological Society} 93 (1): 27--37. \url{https://doi.org/b3w2h4}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-Xia2007}{}}%
+Xia, Xiangao, Hongbin Chen, Zhanqing Li, Pucai Wang, and Jiankai Wang.
+2007.
+{``Significant Reduction of Surface Solar Irradiance Induced by Aerosols in a Suburban Region in Northeastern China.''} \emph{Journal of Geophysical Research Atmospheres} 112 (22): 1--9.
+\url{https://doi.org/cdtntw}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-Zerefos2009}{}}%
+Zerefos, C.
+S., K.
+Eleftheratos, C.
+Meleti, S.
+Kazadzis, A.
+Romanou, C.
+Ichoku, G.
+Tselioudis, and A.
+Bais.
+2009.
+{``Solar Dimming and Brightening over Thessaloniki, Greece, and Beijing, China.''} \emph{Tellus B: Chemical and Physical Meteorology} 61 (4): 657.
+\url{https://doi.org/10.1111/j.1600-0889.2009.00425.x}.
+
+\end{CSLReferences}
 

### 2023-06-26 DHI_GHI_0_variables.R

@@ -31,8 +31,8 @@ translate <- function(x) {
 
 #### Data range ####
 ## will not include the last/first day
-LAST_DAY                 <- as.Date("2023-06-01")
-# LAST_DAY                 <- as.Date("2011-04-01") ## similar to Bais analysis
+# LAST_DAY                 <- as.Date("2023-06-01")
+LAST_DAY                 <- as.Date("2011-04-01") ## similar to Bais analysis
 FIRST_DAY                <- as.Date("1900-07-01")
 # FIRST_DAY                <- as.Date("2005-01-01") ## data inspected by me
 # FIRST_DAY                <- as.Date("2016-04-01") ## start of chp1

### 2023-06-26 Article.Rmd
Commit 2023-06-26 12:17

@@
 output: 
-  bookdown::word_document2: default
+  rticles::elsevier_article:
+    keep_tex:         true
+    citation_package: natbib
   bookdown::pdf_document2:
@@
     biblio-style:     apalike
-  rticles::elsevier_article:
-    keep_tex:         true
-    citation_package: natbib
+  bookdown::word_document2: default
 # date:           "`r format(Sys.time(), '%F')`"
@@
 ```{r trendALL, echo=F, fig.cap="Anomaly (\\%) of the daily SDR relative to climatological values for 1993 - 2023. The black line shows the long term trend for all-sky conditions."}
-knitr::include_graphics("./images/LongtermTrends-2.png")
+knitr::include_graphics("./images/LongtermTrends-1.png")
 ```

### 2023-06-26 DHI_GHI_2_sza_trends.R

@@ -24,7 +24,7 @@
 #'   bookdown::pdf_document2:
 #'     number_sections: no
 #'     fig_caption:     no
-#'     keep_tex:        no
+#'     keep_tex:        yes
 #'     latex_engine:    xelatex
 #'     toc:             yes
 #'     toc_depth:       4
@@ -176,9 +176,6 @@ ccex_sbs <- 1.3
 # }
 # #+ echo=F, include=F
 # rm(data_list)
-
-
-
 
 
 

### 2023-06-26 DHI_GHI_1_longterm_trends.R
Commit 2023-06-26 12:17

@@ -24,7 +24,7 @@
 #'   bookdown::pdf_document2:
 #'     number_sections: no
 #'     fig_caption:     no
-#'     keep_tex:        no
+#'     keep_tex:        yes
 #'     latex_engine:    xelatex
 #'     toc:             yes
 #'     toc_depth:       4

### 2023-06-26 Article.Rmd
Commit 2023-06-26 13:17

--- 2023-06-20 Article.Rmd
@@
    - \usepackage{subcaption}
+
 output: 
-  bookdown::word_document2: default
+  rticles::elsevier_article:
+    keep_tex:         no
+    citation_package: natbib
   bookdown::pdf_document2:
@@
     fig_caption:      yes
-    keep_tex:         yes
+    keep_tex:         no
     latex_engine:     xelatex
@@
 <!-- SDR or GHI ? -->
+
+Cloud “shrinking” and “optical thinning” in the “dimming” period and a subsequent recovery in the “brightening” period over China
+
+aerosols not only strengthen but also weaken the
+growth of clouds with respect to their coverage and
+optical thickness, depending on the levels of pollution
+and the associated amounts of aerosols as postulated
+in a conceptual framework (Wild 2009a, 2012, Yang
+et al 2012).
 

### 2023-06-26 Article.Rmd
Commit 2023-06-26 18:17

@@
 classoption:     preprint, 3p, authoryear
-bibliography:   references.bib
+# bibliography:    references.bib
+bibliography:    '/home/athan/LIBRARY/A_Atmosphere/A_Atmosphere.bib'
 linenumbers:     false

### 2023-06-26 Article.tex
Commit 2023-06-26 18:17

@@
 
+Cloud ``shrinking'' and ``optical thinning'' in the ``dimming'' period
+and a subsequent recovery in the ``brightening'' period over China
+
+aerosols not only strengthen but also weaken the growth of clouds with
+respect to their coverage and optical thickness, depending on the levels
+of pollution and the associated amounts of aerosols as postulated in a
+conceptual framework (Wild 2009a, 2012, Yang et al 2012).
+
 \hypertarget{introduction.}{%
@@
 leveled, CM-21 pyranometer. SDR data spans the period of 1993-04-12 to
-2011-03-31. In some algorithms we used as auxiliary data, the direct
+2023-05-31. In some algorithms we used as auxiliary data, the direct
 beam radiation (DNI) for comparisons, obtained by a CHP-1 pyrheliometer,
-with data availability from Inf to -Inf.
+with data availability from 2016-04-01 to 2023-05-31.
 
@@
 satellite TSI observations. The TSI data we use, is a combination of
-data from NOAA \citep{Coddington2005} (for 1993-04-12 - 2011-03-31) and
-adjusted data from \citet{TSIStsi} (for - ). As a result, we can present
-all radiation data as a fraction of Sun's TSI.
+data from NOAA \citep{Coddington2005} (for 1993-04-12 - 2023-03-31) and
+adjusted data from \citet{TSIStsi} (for 2023-03-31 - 2023-05-31). As a
+result, we can present all radiation data as a fraction of Sun's TSI.
 
@@
 \endlastfoot
-0.4221 & 5.871e-06 & All sky cond. & 100 \\
-0.3871 & 0.0001123 & Clear sky cond. & 99.99 \\
-0.415 & 0.002237 & Cloudy cond. & 99.78 \\
+0.381 & 3.479e-21 & All sky cond. & 100 \\
+0.3605 & 1.068e-16 & Clear sky cond. & 100 \\
+0.3472 & 8.968e-09 & Cloudy cond. & 100 \\
 \end{longtable}
@@
 \endlastfoot
-0.8014 & 0.1606 & All sky cond. & Winter \\
--0.1999 & 0.5397 & All sky cond. & Spring \\
-0.2777 & 0.09127 & All sky cond. & Summer \\
-0.3768 & 0.2006 & All sky cond. & Autumn \\
-0.9574 & 0.1081 & Clear sky cond. & Winter \\
--0.4227 & 0.2207 & Clear sky cond. & Spring \\
-0.0931 & 0.6638 & Clear sky cond. & Summer \\
-0.3233 & 0.3255 & Clear sky cond. & Autumn \\
--0.4354 & 0.4924 & Cloudy cond. & Winter \\
-0.8353 & 0.04263 & Cloudy cond. & Spring \\
-0.7051 & 0.05457 & Cloudy cond. & Summer \\
-0.3708 & 0.2357 & Cloudy cond. & Autumn \\
+0.7215 & 0.001209 & All sky cond. & Winter \\
+0.1679 & 0.1083 & All sky cond. & Spring \\
+0.128 & 0.08705 & All sky cond. & Summer \\
+0.4239 & 0.006659 & All sky cond. & Autumn \\
+0.8715 & 0.000269 & Clear sky cond. & Winter \\
+0.09719 & 0.3923 & Clear sky cond. & Spring \\
+-0.02369 & 0.7972 & Clear sky cond. & Summer \\
+0.3583 & 0.02996 & Clear sky cond. & Autumn \\
+-0.02581 & 0.9335 & Cloudy cond. & Winter \\
+0.3557 & 0.06914 & Cloudy cond. & Spring \\
+0.442 & 0.004988 & Cloudy cond. & Summer \\
+0.4776 & 0.01955 & Cloudy cond. & Autumn \\
 \end{longtable}
@@
 \endlastfoot
-83.94 \\
-46.03 \\
-90.87 \\
-79.94 \\
-89.19 \\
-77.93 \\
-33.62 \\
-67.45 \\
-50.76 \\
-95.74 \\
-94.54 \\
-76.43 \\
+99.88 \\
+89.17 \\
+91.29 \\
+99.33 \\
+99.97 \\
+60.77 \\
+20.28 \\
+97 \\
+6.646 \\
+93.09 \\
+99.5 \\
+98.04 \\
 \end{longtable}
@@
 
-Our result for all-sky condition (\(0.42\%/year\)), reaffirm the
+Our result for all-sky condition (\(0.38\%/year\)), reaffirm the
 previous results of \citet{Bais2013} for the site. The increase of this
@@
 
-\ldots.. Also, the trend of \(0.41\%/year\) for cloud-sky condition
+\ldots.. Also, the trend of \(0.35\%/year\) for cloud-sky condition
 indicate the major part that clouds play on the \ldots..
@@
 
-\bibliography{references.bib}
+\bibliography{/home/athan/LIBRARY/A\_Atmosphere/A\_Atmosphere.bib}
 

### 2023-06-26 Article.Rmd
Commit 2023-06-26 19:17

--- 2023-06-26 Article.Rmd
@@
 classoption:     preprint, 3p, authoryear
-bibliography:   references.bib
+bibliography:    references2.bib
 linenumbers:     false
@@
 source("~/MANUSCRIPTS/2022_sdr_trends/DHI_GHI_0_variables.R")
+```
+
+```{r}
+library(stringr)
+
+lines   <- paste(readLines("./Article.Rmd"), collapse = "")
+entries <- unique(str_match_all(lines, "@([a-zA-Z0-9]+)[,\\. \\?\\!\\]\\;]")[[1]][, 2])
+entries <- grep("gmail|auth", entries, value = TRUE, invert = TRUE)
+
+bib <- paste(readLines("./references2.bib"), collapse = "\n")
+bib <- unlist(strsplit(bib, "\n@"))
+
+output <- sapply(entries, grep, bib, value = T)
+output <- paste("@", output, sep = "")
+
+writeLines(unlist(output), "./references3.bib")
+
 ```
@@
 <!-- Quality assurance -->
-The manual screening is aided by a radiation data quality assurance procedure, adjusted for this site, based on methods of Long and Shi\ [-@long_automated_2008; -@Long2006].
+The manual screening is aided by a radiation data quality assurance procedure, adjusted for this site, based on methods of Long and Shi\ [-@Long2008; -@Long2006].
 Thus, problematic recordings have been excluded from further processing. 
@@
 Subsequently, we made all measurements relative to the Total Solar Irradiance (TSI) at $1 \text{au}$, in order to compensate for the Sun's intensity variability, using a homogenized time series of satellite TSI observations.
-The TSI data we use, is a combination of data from NOAA [@Coddington2005] (for `r as.Date(TSIinfo[TSI_Source == "NOAA", Start])` - `r as.Date(TSIinfo[TSI_Source == "NOAA", End])`) and adjusted data from @TSIStsi (for `r as.Date(TSIinfo[TSI_Source != "NOAA", Start])` - `r as.Date(TSIinfo[TSI_Source != "NOAA", End])`).
+The TSI data we use, is a combination of data from NOAA [@Coddington2005] (for `r as.Date(TSIinfo[TSI_Source == "NOAA", Start])` - `r as.Date(TSIinfo[TSI_Source == "NOAA", End])`) and adjusted data from @LASP2023 (for `r as.Date(TSIinfo[TSI_Source != "NOAA", Start])` - `r as.Date(TSIinfo[TSI_Source != "NOAA", End])`).
 As a result, we can present all radiation data as a fraction of Sun's TSI.

### 2023-06-26 Article.Rmd
Commit 2023-06-26 20:17

@@
 source("~/MANUSCRIPTS/2022_sdr_trends/DHI_GHI_0_variables.R")
-```
-
-```{r}
-library(stringr)
-
-lines   <- paste(readLines("./Article.Rmd"), collapse = "")
-entries <- unique(str_match_all(lines, "@([a-zA-Z0-9]+)[,\\. \\?\\!\\]\\;]")[[1]][, 2])
-entries <- grep("gmail|auth", entries, value = TRUE, invert = TRUE)
-
-bib <- paste(readLines("./references2.bib"), collapse = "\n")
-bib <- unlist(strsplit(bib, "\n@"))
-
-output <- sapply(entries, grep, bib, value = T)
-output <- paste("@", output, sep = "")
-
-writeLines(unlist(output), "./references3.bib")
-
 ```

### 2023-06-26 gather_citations.R

@@ -1,0 +1,35 @@
+
+
+
+library(stringr)
+
+sourcefile   <- "./Article.Rmd"
+mainreposit  <- "./references2.bib"
+localreposit <- "./references3.bib"
+
+## gather citations
+lines   <- paste(readLines(sourcefile), collapse = "")
+entries <- unique(str_match_all(lines, "@([a-zA-Z0-9]+)[,\\. \\?\\!\\]\\;]")[[1]][, 2])
+entries <- grep("gmail|auth", entries, value = TRUE, invert = TRUE)
+
+bib <- readLines(mainreposit)
+
+grep("Long2008", bib, value = T)
+bib <- paste(bib, collapse = "\n")
+bib <- unlist(strsplit(bib, "\n@"))
+
+output <- sapply(entries, grep, bib, value = T)
+grep("Long2008", output, value = T)
+output <- paste("@", as.character(output), sep = "")
+
+output <- gsub("^@c\\(\"", "@", output)
+output <- gsub("\"\n\\)$",  "", output)
+# output <- gsub("\\\\n",  "\n", output)
+
+
+gsub("^@c\\(\"", "@", grep("Long2008", output, value = T))
+
+gsub("\\\\n\"\n\\)$",  "", grep("Long2008", output, value = T))
+
+
+writeLines(unlist(output), localreposit)

### 2023-06-27 gather_citations.R
Commit 2023-06-27 11:17

@@ -1,55 +1,0 @@
-
-
-
-library(stringr)
-library(bib2df)
-library(RefManageR)
-
-setwd("~/MANUSCRIPTS/2022_sdr_trends/")
-sourcefile   <- "./Article.Rmd"
-mainreposit  <- "./references2.bib"
-localreposit <- "./references3.bib"
-folderlink   <- "./References"
-
-
-## gather citations from file
-lines   <- paste(readLines(sourcefile), collapse = "")
-entries <- unique(str_match_all(lines, "@([a-zA-Z0-9]+)[,\\. \\?\\!\\]\\;]")[[1]][, 2])
-entries <- grep("gmail|auth", entries, value = TRUE, invert = TRUE)
-
-
-## get refs
-bib <- bib2df(mainreposit)
-
-## export sub file
-output <- bib[bib$BIBTEXKEY %in% entries, ]
-filesp <- output$FILE
-output$ABSTRACT         <- NA
-output$OWNER            <- NA
-output$GROUPS           <- NA
-output$MODIFICATIONDATE <- NA
-output$TIMESTAMP        <- NA
-output$FILE             <- NA
-df2bib(output, localreposit)
-
-
-
-# link files
-filesp <- output$FILE
-names(output)
-basepath <- dirname(Sys.readlink(mainreposit))
-
-for (af in strsplit(filesp, ":")) {
-    cat(length(af),"\n")
-    if (length(af) >= 2) {
-        sourcefl <- paste0(basepath,"/", af[2])
-        if (file.exists(sourcefl)) {
-            dir.create(folderlink)
-            file.symlink(sourcefl, paste0(folderlink, "/", basename(af[2])))
-        }
-    }
-}
-
-
-tt <- ReadBib(mainreposit, check = FALSE)
-ss <- ReadBib(localreposit, check = FALSE)

### 2023-06-27 Article.Rmd
Commit 2023-06-27 11:17

@@
 classoption:     preprint, 3p, authoryear
-bibliography:    references2.bib
+bibliography:    Mainreferences.bib
 linenumbers:     false
@@
 
-```{r echo=F, include=T}
+```{r echo=F, include=T, warning=FALSE}
 ## document configurations 
@@
 source("~/MANUSCRIPTS/2022_sdr_trends/DHI_GHI_0_variables.R")
+
+## create local citations file
+source("~/CODE/FUNCTIONS/R/gather_citations.R")
+
+suppressMessages(
+    gather_citations(
+        sourcefile   = "./Article.Rmd",
+        mainreposit  = "./Mainreferences.bib",
+        localreposit = "./references.bib",
+        folderlink   = "./References_links",
+        workingdir   = "~/MANUSCRIPTS/2022_sdr_trends/"
+    )
+)
 ```
@@
 # Introduction.
+
+Describe the variability of the phenomenn
+
+the inpacts of the radiation
+
+the as far studies/coauses
 
@@
 <!-- Quality assurance -->
-The manual screening is aided by a radiation data quality assurance procedure, adjusted for this site, based on methods of Long and Shi\ [-@Long2008; -@Long2006].
+The manual screening is aided by a radiation data quality assurance procedure, adjusted for this site, based on methods of Long and Shi\ [@Long2008a ; @Long2006].
 Thus, problematic recordings have been excluded from further processing. 

### 2023-06-28 Article.Rmd
Commit 2023-06-27 13:17
Commit 2023-06-28 11:17

@@
 The shortwave downward solar irradiance (SDR) at the Earth's surface plays a significant role, on Earths climate.
-Changes of the SDR can be related to major changes on factors affecting the radiative forcing balance and human activities.
+Changes of the SDR can be related to major changes on the Earths energy budget, the climate change mechanisms and the water and carbon cycles [@Wild2009].
+Studies of SDR variability, have identify some distinct SDR trends on different regions of the world and for different periods, using the term 'brightening' for positive trends, and 'dimming' for negative trends.
+
 Multiple studies attempt to evaluate the phenomenon and the potential causes [@Ohvril2009; @Schwarz2020; @Wild2009; @Wild2012; @Xia2007; @Zerefos2009].
@@
 
+similar results with Evidence for Clear‐Sky Dimming and Brightening in Central Europe_Wild2021.pdf
+inclue in discusion
+
 -------------------------------------------------

### 2023-06-28 Article.Rmd
Commit 2023-06-28 12:17

@@
 Changes of the SDR can be related to major changes on the Earths energy budget, the climate change mechanisms and the water and carbon cycles [@Wild2009].
-Studies of SDR variability, have identify some distinct SDR trends on different regions of the world and for different periods, using the term 'brightening' for positive trends, and 'dimming' for negative trends.
+Studies of SDR variability, have identify some distinct SDR trends on different regions of the world and for different time periods, using the term 'brightening' for positive trends, and 'dimming' for negative trends.
+Multiple studies attempt to evaluate the phenomenon and the potential causes [@Ohvril2009; @Schwarz2020; @Wild2009; @Wild2012; @Xia2007; @Zerefos2009; and references therein].
+There is a general consensus that the major factors of the attenuation of SDR is the interaction with the atmospheric aerosols and the clouds.
 
-Multiple studies attempt to evaluate the phenomenon and the potential causes [@Ohvril2009; @Schwarz2020; @Wild2009; @Wild2012; @Xia2007; @Zerefos2009].
-Although, there is regional variation of SDR, with both positive and negative trends, most of the studies agree that the major factors are the interaction of aerosols and clouds.
+- observed long term changes
+  europe / asia africa regions
+
+Although, there is regional variation of SDR, with both positive and negative trends, most of the studies agree that .
 

### 2023-06-28 Article.Rmd
Commit 2023-06-28 19:17

@@
 Multiple studies attempt to evaluate the phenomenon and the potential causes [@Ohvril2009; @Schwarz2020; @Wild2009; @Wild2012; @Xia2007; @Zerefos2009; and references therein].
-There is a general consensus that the major factors of the attenuation of SDR is the interaction with the atmospheric aerosols and the clouds.
+
+@Yang2021 reports a general dimming before 2000 and a brightening afterwards on  multiple station in China.
 
 - observed long term changes
-  europe / asia africa regions
+-  europe / asia africa regions
+- split in trends
 
-Although, there is regional variation of SDR, with both positive and negative trends, most of the studies agree that .
+There is a general consensus, among researchers, that the major factors of the attenuation of SDR is the interaction with the atmospheric aerosols and the clouds.
 

### 2023-06-28 Article.Rmd
Commit 2023-06-28 20:17

--- 2023-06-28 Article.Rmd
@@
     latex_engine:     xelatex
-    toc:              yes
+    toc:              no
     toc_depth:        4
@@
 
-Describe the variability of the phenomenn
+The shortwave downward solar irradiance (SDR) at the Earth's surface plays a significant role, on Earths climate.
+Changes of the SDR can be related to major changes on the Earth's energy budget, the mechanisms of climate change, and water and carbon cycle [@Wild2009].
+Studies of SDR variability, have identified some distinct SDR trends on different regions of the world and for different time periods, using the term 'brightening' for positive trends, and 'dimming' for negative trends.
+There are many cases on the long term records, showing a change of SDR trends direction, occurring roughly at the last decades of the 20th century.
+<!-- Trends -->
+@Yang2021 reports a general dimming before 2000 and a brightening afterwards on multiple station in China. 
+@Wild2021 shows a similar pattern, with the breaking point around 1980 from a 71 years record of a Central Europe station and for Brazil by @Yamasoe2021.
+<!-- @Yamasoe2021 similar dimming and brightening around 1980 in Brazil -->
+@Yuan2021 with AI aided spatial analysis on continent level, containing numerous observational station, reach similar conclusions for the above regions and the global trend.
 
-the inpacts of the radiation
+Although, there is a consensus, among researchers, that the major factors of the attenuation of SDR is the interaction with the atmospheric aerosols and the clouds.
 
-the as far studies/coauses
+Due to the variability of those factors and the interaction between them,
 
-The shortwave downward solar irradiance (SDR) at the Earth's surface plays a significant role, on Earths climate.
-Changes of the SDR can be related to major changes on the Earths energy budget, the climate change mechanisms and the water and carbon cycles [@Wild2009].
-Studies of SDR variability, have identify some distinct SDR trends on different regions of the world and for different time periods, using the term 'brightening' for positive trends, and 'dimming' for negative trends.
+...... cloud formation occurrences and constitution nuclei. cite.....
+
+it is difficult to estimate the SDR trends on each site.
+
 Multiple studies attempt to evaluate the phenomenon and the potential causes [@Ohvril2009; @Schwarz2020; @Wild2009; @Wild2012; @Xia2007; @Zerefos2009; and references therein].
-There is a general consensus that the major factors of the attenuation of SDR is the interaction with the atmospheric aerosols and the clouds.
-
-- observed long term changes
-  europe / asia africa regions
-
-Although, there is regional variation of SDR, with both positive and negative trends, most of the studies agree that .
+<!-- causes say less we don't do that -->
 

### 2023-06-29 Article.Rmd
Commit 2023-06-29 10:17

@@
 <!-- @Yamasoe2021 similar dimming and brightening around 1980 in Brazil -->
-@Yuan2021 with AI aided spatial analysis on continent level, containing numerous observational station, reach similar conclusions for the above regions and the global trend.
+Another approach, using AI aided spatial analysis on continent level, of numerous observational stations data, reach similar conclusions for the above regions and the global trend [@Yuan2021].
 
-Although, there is a consensus, among researchers, that the major factors of the attenuation of SDR is the interaction with the atmospheric aerosols and the clouds.
+There is a consensus, among researchers, that the major factors of the attenuation of SDR is the interaction with the atmospheric aerosols and the clouds.
 
@@
 <!-- causes say less we don't do that -->
+
+There is a also a need to evalutate the trends on different location
 

### 2023-06-29 Article.tex
Commit 2023-06-29 12:17

@@
 plays a significant role, on Earths climate. Changes of the SDR can be
-related to major changes on factors affecting the radiative forcing
-balance and human activities. Multiple studies attempt to evaluate the
-phenomenon and the potential causes
-\citep{Ohvril2009, Schwarz2020, Wild2009, Wild2012, Xia2007, Zerefos2009}.
-Although, there is regional variation of SDR, with both positive and
-negative trends, most of the studies agree that the major factors are
-the interaction of aerosols and clouds.
+related to major changes on the Earth's energy budget, the mechanisms of
+climate change, and water and carbon cycle \citep{Wild2009}. Studies of
+SDR variability, have identified some distinct SDR trends on different
+regions of the world and for different time periods, using the term
+`brightening' for positive trends, and `dimming' for negative trends.
+There are many cases on the long term records, showing a change of SDR
+trends direction, occurring roughly at the last decades of the 20th
+century. \citet{Yang2021} reports a general dimming before 2000 and a
+brightening afterwards on multiple station in China. \citet{Wild2021}
+shows a similar pattern, with the breaking point around 1980 from a 71
+years record of a Central Europe station and for Brazil by
+\citet{Yamasoe2021}. Another approach, using AI aided spatial analysis
+on continent level, of numerous observational stations data, reach
+similar conclusions for the above regions and the global trend
+\citep{Yuan2021}.
+
+There is a consensus, among researchers, that the major factors of the
+attenuation of SDR is the interaction with the atmospheric aerosols and
+the clouds.
+
+The interaction between Aerosols and clouds, among other factors, has
+been analysed with modeling \citep{Li2016, Samset2018}, showing the
+existence of feedback mechanisms between the two.
+
+Similar links have been showed with observational data
+{[}\citet{Ohvril2009}; \citet{Schwarz2020}; {]}
+
+Due to the variability of those factors and the interaction between
+them,
+
+\ldots\ldots{} cloud formation occurrences and constitution nuclei.
+cite\ldots..
+
+it is difficult to estimate the SDR trends on each site.
+
+Multiple studies attempt to evaluate the phenomenon and the potential
+causes {[}; \citet{Wild2012}; \citet{Xia2007}; \citet{Zerefos2009}; and
+references therein{]}.
+
+Investigate multiple interaction \citet{Li2016} \citet{Samset2018} data
+analysis and modeling
+
+There is a also a need to evaluate the trends on different location
 
@@
 a radiation data quality assurance procedure, adjusted for this site,
-based on methods of Long and
-Shi~\citetext{\citeyear{long_automated_2008}; \citeyear{Long2006}}.
-Thus, problematic recordings have been excluded from further processing.
-Although it is not possible to detect all the bad data, the large number
-of data and the aggregation scheme we use, can provide us with accurate
-radiometric measurements, for the scope of this study.
+based on methods of Long and Shi~\citetext{\citealp[
+]{Long2008a}; \citealp{Long2006}}. Thus, problematic recordings have
+been excluded from further processing. Although it is not possible to
+detect all the bad data, the large number of data and the aggregation
+scheme we use, can provide us with accurate radiometric measurements,
+for the scope of this study.
 
@@
 data from NOAA \citep{Coddington2005} (for 1993-04-12 - 2023-03-31) and
-adjusted data from \citet{TSIStsi} (for 2023-03-31 - 2023-05-31). As a
+adjusted data from \citet{LASP2023} (for 2023-03-31 - 2023-05-31). As a
 result, we can present all radiation data as a fraction of Sun's TSI.
@@
 
+similar results with Evidence for Clear‐Sky Dimming and Brightening in
+Central Europe\_Wild2021.pdf inclue in discusion
+
 \begin{longtable}[]{@{}c@{}}
@@
 
-\bibliography{/home/athan/LIBRARY/A\_Atmosphere/A\_Atmosphere.bib}
+\bibliography{Mainreferences.bib}
 

### 2023-06-29 Article_A.tex
Commit 2023-06-29 12:17

@@
 \author{true \and true}
-\date{2023-06-26}
+\date{2023-06-29}
 
@@
 
-{
-\setcounter{tocdepth}{4}
-\tableofcontents
-}
+Cloud ``shrinking'' and ``optical thinning'' in the ``dimming'' period and a subsequent recovery in the ``brightening'' period over China
+
+aerosols not only strengthen but also weaken the
+growth of clouds with respect to their coverage and
+optical thickness, depending on the levels of pollution
+and the associated amounts of aerosols as postulated
+in a conceptual framework (Wild 2009a, 2012, Yang
+et al 2012).
+
 \hypertarget{introduction.}{%
@@
 The shortwave downward solar irradiance (SDR) at the Earth's surface plays a significant role, on Earths climate.
-Changes of the SDR can be related to major changes on factors affecting the radiative forcing balance and human activities.
-Multiple studies attempt to evaluate the phenomenon and the potential causes (Ohvril et al. 2009; Schwarz et al. 2020; Wild 2009, 2012; Xia et al. 2007; Zerefos et al. 2009).
-Although, there is regional variation of SDR, with both positive and negative trends, most of the studies agree that the major factors are the interaction of aerosols and clouds.
+Changes of the SDR can be related to major changes on the Earth's energy budget, the mechanisms of climate change, and water and carbon cycle (Wild 2009).
+Studies of SDR variability, have identified some distinct SDR trends on different regions of the world and for different time periods, using the term `brightening' for positive trends, and `dimming' for negative trends.
+There are many cases on the long term records, showing a change of SDR trends direction, occurring roughly at the last decades of the 20th century.
+Yang et al. (2021) reports a general dimming before 2000 and a brightening afterwards on multiple station in China.
+Wild et al. (2021) shows a similar pattern, with the breaking point around 1980 from a 71 years record of a Central Europe station and for Brazil by Yamasoe et al. (2021).
+Another approach, using AI aided spatial analysis on continent level, of numerous observational stations data, reach similar conclusions for the above regions and the global trend (Yuan, Leirvik, and Wild 2021).
+
+There is a consensus, among researchers, that the major factors of the attenuation of SDR is the interaction with the atmospheric aerosols and the clouds.
+
+The interaction between Aerosols and clouds, among other factors, has been analysed with modeling (Li et al. 2016; Samset et al. 2018), showing the existence of feedback mechanisms between the two.
+
+Similar links have been showed with observational data {[}Ohvril et al. (2009); Schwarz et al. (2020); {]}
+
+Due to the variability of those factors and the interaction between them,
+
+\ldots\ldots{} cloud formation occurrences and constitution nuclei. cite\ldots..
+
+it is difficult to estimate the SDR trends on each site.
+
+Multiple studies attempt to evaluate the phenomenon and the potential causes {[}; Wild (2012); Xia et al. (2007); Zerefos et al. (2009); and references therein{]}.
+
+Investigate multiple interaction Li et al. (2016) Samset et al. (2018) data analysis and modeling
+
+There is a also a need to evaluate the trends on different location
 
@@
 SDR data spans the period of
-1993-04-12 to 2011-03-31.
+1993-04-12 to 2023-05-31.
 In some algorithms we used as auxiliary data, the direct beam radiation (DNI) for comparisons, obtained by a CHP-1 pyrheliometer, with data availability from
-Inf to -Inf.
+2016-04-01 to 2023-05-31.
 
@@
 A manual screening is performed, to remove inconsistencies and erroneous recordings, that can occur randomly or systematically, during the long continuous operation of the instrument.
-The manual screening is aided by a radiation data quality assurance procedure, adjusted for this site, based on methods of Long and Shi~(2008, 2006).
+The manual screening is aided by a radiation data quality assurance procedure, adjusted for this site, based on methods of Long and Shi~(C. Long and Shi 2008 ; C. Long and Shi 2006).
 Thus, problematic recordings have been excluded from further processing.
@@
 The tried models are described by Reno, Hansen, and Stein (2012) and tested by Reno and Hansen (2016).
-The iterative optimization method, for the selection of the reference is discussed by Charles N. Long and Ackerman (2000) and Reno and Hansen (2016).
+The iterative optimization method, for the selection of the reference is discussed by C. N. Long and Ackerman (2000) and Reno and Hansen (2016).
 
@@
 The clear-sky conditions data, where we have inferred that the sky was almost clear of clouds, and the remainder part the data as cloudy sky conditions data (cloud-sky).
-To identify the clear-sky conditions we used the method proposed by Charles N. Long and Ackerman (2000) and by Reno and Hansen (2016), that was adapted and configured for the site.
+To identify the clear-sky conditions we used the method proposed by C. N. Long and Ackerman (2000) and by Reno and Hansen (2016), that was adapted and configured for the site.
 We have to note, that the definition of what constitutes as clear or cloudy sky, has some subjectivity, due to the used method of characterization.
@@
 Subsequently, we made all measurements relative to the Total Solar Irradiance (TSI) at \(1 \text{au}\), in order to compensate for the Sun's intensity variability, using a homogenized time series of satellite TSI observations.
-The TSI data we use, is a combination of data from NOAA (Coddington et al. 2005) (for 1993-04-12 - 2011-03-31) and adjusted data from LASP (n.d.) (for - ).
+The TSI data we use, is a combination of data from NOAA (Coddington et al. 2005) (for 1993-04-12 - 2023-03-31) and adjusted data from LASP (2023) (for 2023-03-31 - 2023-05-31).
 As a result, we can present all radiation data as a fraction of Sun's TSI.
@@
 \endhead
-0.4221 & 5.871e-06 & All sky cond. & 100 \\
-0.3871 & 0.0001123 & Clear sky cond. & 99.99 \\
-0.415 & 0.002237 & Cloudy cond. & 99.78 \\
+0.381 & 3.479e-21 & All sky cond. & 100 \\
+0.3605 & 1.068e-16 & Clear sky cond. & 100 \\
+0.3472 & 8.968e-09 & Cloudy cond. & 100 \\
 \bottomrule
@@
 \endhead
-0.8014 & 0.1606 & All sky cond. & Winter \\
--0.1999 & 0.5397 & All sky cond. & Spring \\
-0.2777 & 0.09127 & All sky cond. & Summer \\
-0.3768 & 0.2006 & All sky cond. & Autumn \\
-0.9574 & 0.1081 & Clear sky cond. & Winter \\
--0.4227 & 0.2207 & Clear sky cond. & Spring \\
-0.0931 & 0.6638 & Clear sky cond. & Summer \\
-0.3233 & 0.3255 & Clear sky cond. & Autumn \\
--0.4354 & 0.4924 & Cloudy cond. & Winter \\
-0.8353 & 0.04263 & Cloudy cond. & Spring \\
-0.7051 & 0.05457 & Cloudy cond. & Summer \\
-0.3708 & 0.2357 & Cloudy cond. & Autumn \\
+0.7215 & 0.001209 & All sky cond. & Winter \\
+0.1679 & 0.1083 & All sky cond. & Spring \\
+0.128 & 0.08705 & All sky cond. & Summer \\
+0.4239 & 0.006659 & All sky cond. & Autumn \\
+0.8715 & 0.000269 & Clear sky cond. & Winter \\
+0.09719 & 0.3923 & Clear sky cond. & Spring \\
+-0.02369 & 0.7972 & Clear sky cond. & Summer \\
+0.3583 & 0.02996 & Clear sky cond. & Autumn \\
+-0.02581 & 0.9335 & Cloudy cond. & Winter \\
+0.3557 & 0.06914 & Cloudy cond. & Spring \\
+0.442 & 0.004988 & Cloudy cond. & Summer \\
+0.4776 & 0.01955 & Cloudy cond. & Autumn \\
 \bottomrule
@@
 \endhead
-83.94 \\
-46.03 \\
-90.87 \\
-79.94 \\
-89.19 \\
-77.93 \\
-33.62 \\
-67.45 \\
-50.76 \\
-95.74 \\
-94.54 \\
-76.43 \\
+99.88 \\
+89.17 \\
+91.29 \\
+99.33 \\
+99.97 \\
+60.77 \\
+20.28 \\
+97 \\
+6.646 \\
+93.09 \\
+99.5 \\
+98.04 \\
 \bottomrule
@@
 
-Our result for all-sky condition (\(0.42\%/year\)), reaffirm the previous results of Bais et al. (2013) for the site.
+Our result for all-sky condition (\(0.38\%/year\)), reaffirm the previous results of Bais et al. (2013) for the site.
 The increase of this trend, shows that the phenomenon and probably the causes have been amplified.
@@
 \ldots..
-Also, the trend of \(0.41\%/year\) for cloud-sky condition indicate the major part that clouds play on the
+Also, the trend of \(0.35\%/year\) for cloud-sky condition indicate the major part that clouds play on the
 \ldots..
@@
 \end{itemize}
+
+similar results with Evidence for Clear‐Sky Dimming and Brightening in Central Europe\_Wild2021.pdf
+inclue in discusion
 
@@
 Coddington, Odele, Judith L.
-Lean, Doug Lindholm, Peter Pilewskie amd Martin Snow, and NOAA CDR Program.
+Lean, Doug Lindholm, Peter Pilewskie, Martin Snow, and NOAA CDR Program.
 2005.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-TSIStsi}{}}%
-LASP. n.d. {``{TSIS-1} Total Solar Irradiance - Six Hour Average.''} Accessed June 1, 2023. \url{https://lasp.colorado.edu/tsis/data/}.
+\leavevmode\vadjust pre{\hypertarget{ref-LASP2023}{}}%
+LASP. 2023. {``TSIS-1 Total Solar Irradiance - Six Hour Average.''} Laboratory for Atmospheric; Space Physics, University of Colorado, Boulder. 2023. \url{https://lasp.colorado.edu/tsis/data/}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-Li2016}{}}%
+Li, Zhanqing, W.
+K.‐M.
+Lau, V.
+Ramanathan, G.
+Wu, Y.
+Ding, M.
+G.
+Manoj, J.
+Liu, et al.
+2016.
+{``Aerosol and Monsoon Climate Interactions over Asia.''} \emph{Reviews of Geophysics} 54 (4): 866--929.
+\url{https://doi.org/10.1002/2015RG000500}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-Long2000}{}}%
+Long, Charles N., and Thomas P.
+Ackerman.
+2000.
+{``Identification of Clear Skies from Broadband Pyranometer Measurements and Calculation of Downwelling Shortwave Cloud Effects.''} \emph{Journal of Geophysical Research: Atmospheres} 105 (D12, D12): 15609--26.
+\url{https://doi.org/10.1029/2000jd900077}.
 
 \leavevmode\vadjust pre{\hypertarget{ref-Long2006}{}}%
-Long, C.
-N., and Y.
+Long, C., and Y.
 Shi.
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-long_automated_2008}{}}%
----------. 2008. {``An {Automated} {Quality} {Assessment} and {Control} {Algorithm} for {Surface} {Radiation} {Measurements}.''} \emph{The Open Atmospheric Science Journal}, 23--37.
-
-\leavevmode\vadjust pre{\hypertarget{ref-Long2000}{}}%
-Long, Charles N., and Thomas P.
-Ackerman.
-2000.
-{``{Identification of clear skies from broadband pyranometer measurements and calculation of downwelling shortwave cloud effects}.''} \emph{Journal of Geophysical Research} 105 (D12): 15609--26.
-\url{https://doi.org/10.1029/2000JD900077}.
+\leavevmode\vadjust pre{\hypertarget{ref-Long2008a}{}}%
+---------. 2008. {``An Automated Quality Assessment and Control Algorithm for Surface Radiation Measurements.''} \emph{The Open Atmospheric Science Journal}, 23--37.
 
@@
 2016.
-{``{Identification of periods of clear sky irradiance in time series of GHI measurements}.''} \emph{Renewable Energy} 90: 520--31.
-\url{https://doi.org/10.1016/j.renene.2015.12.031}.
+{``Identification of Periods of Clear Sky Irradiance in Time Series of GHI Measurements.''} \emph{Renewable Energy} 90: 520--31.
+\url{https://doi.org/gq3sbg}.
 
@@
 Reno, Matthew J., Clifford W. Hansen, and Joshua S. Stein. 2012. {``Global Horizontal Irradiance Clear Sky Models: Implementation and Analysis.''} SAND2012-2389, 1039404. \url{https://doi.org/gq5npv}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-Samset2018}{}}%
+Samset, B.
+H., M.
+Sand, C.
+J.
+Smith, S.
+E.
+Bauer, P.
+M.
+Forster, J.
+S.
+Fuglestvedt, S.
+Osprey, and C.‐F.
+Schleussner.
+2018.
+{``Climate Impacts from a Removal of Anthropogenic Aerosol Emissions.''} \emph{Geophysical Research Letters} 45 (2): 1020--29.
+\url{https://doi.org/10.1002/2017GL076079}.
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-Wild2021}{}}%
+Wild, Martin, Stephan Wacker, Su Yang, and Arturo Sanchez‐Lorenzo.
+2021.
+{``Evidence for Clear‐sky Dimming and Brightening in Central Europe.''} \emph{Geophysical Research Letters} 48 (6).
+\url{https://doi.org/10.1029/2020GL092216}.
+
 \leavevmode\vadjust pre{\hypertarget{ref-Xia2007}{}}%
@@
 \url{https://doi.org/cdtntw}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-Yamasoe2021}{}}%
+Yamasoe, Marcia Akemi, Nilton Manuel Évora Rosário, Samantha Novaes Santos Martins Almeida, and Martin Wild.
+2021.
+{``Fifty-Six Years of Surface Solar Radiation and Sunshine Duration over São Paulo, Brazil: 1961--2016.''} \emph{Atmospheric Chemistry and Physics} 21 (9): 6593--603.
+\url{https://doi.org/10.5194/acp-21-6593-2021}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-Yang2021}{}}%
+Yang, Su, Zijiang Zhou, Yu Yu, and Martin Wild.
+2021.
+{``Cloud {`Shrinking'} and {`Optical Thinning'} in the {`Dimming'} Period and a Subsequent Recovery in the {`Brightening'} Period over China.''} \emph{Environmental Research Letters}, January.
+\url{https://doi.org/10.1088/1748-9326/abdf89}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-Yuan2021}{}}%
+Yuan, Menghan, Thomas Leirvik, and Martin Wild.
+2021.
+{``Global Trends in Downward Surface Solar Radiation from Spatial Interpolated Ground Observations During 1961-2019.''} \emph{Journal of Climate}, September, 1--56.
+\url{https://doi.org/10.1175/JCLI-D-21-0165.1}.
 

### 2023-06-29 Article.Rmd
Commit 2023-06-29 11:17
Commit 2023-06-29 12:17

@@
 
+The interaction between Aerosols and clouds, among other factors, has been analysed with modeling [@Li2016; @Samset2018], showing the existence of feedback mechanisms between the two.
+
+Similar links have been showed with observational data [@Ohvril2009; @Schwarz2020; ]
+
 Due to the variability of those factors and the interaction between them,
@@
 
-Multiple studies attempt to evaluate the phenomenon and the potential causes [@Ohvril2009; @Schwarz2020; @Wild2009; @Wild2012; @Xia2007; @Zerefos2009; and references therein].
+Multiple studies attempt to evaluate the phenomenon and the potential causes [;  @Wild2012; @Xia2007; @Zerefos2009; and references therein].
 <!-- causes say less we don't do that -->
 
-There is a also a need to evalutate the trends on different location
+Investigate multiple interaction @Li2016 @Samset2018 data analysis and modeling
+
+There is a also a need to evaluate the trends on different location
 

### 2023-06-29 Article.Rmd
Commit 2023-06-29 13:17

--- 2023-06-29 Article.Rmd
@@
 <!-- Trends -->
-@Yang2021 reports a general dimming before 2000 and a brightening afterwards on multiple station in China. 
-@Wild2021 shows a similar pattern, with the breaking point around 1980 from a 71 years record of a Central Europe station and for Brazil by @Yamasoe2021.
+A general dimming before 2000 and a brightening afterwards, was found on multiple station in China [@Yang2021]. 
+And a similar pattern was identified, with the breaking point around 1980, from a 71 years record of a Central Europe station [@Wild2021] and for Brazil [@Yamasoe2021].
 <!-- @Yamasoe2021 similar dimming and brightening around 1980 in Brazil -->
@@
 There is a consensus, among researchers, that the major factors of the attenuation of SDR is the interaction with the atmospheric aerosols and the clouds. 
+Those interactions, among other factors, have been analysed with modeling [@Li2016; @Samset2018], showing the existence of feedback mechanisms between the two.
+And similar finds, have been showed in observational data [@Schwarz2020; @Ohvril2009; @Zerefos2009; @Xia2007 and references therein].
+
+<!-- Sun variability -->
+
+@Wild2012 : ~
 
@@
 
-Multiple studies attempt to evaluate the phenomenon and the potential causes [@Ohvril2009; @Schwarz2020; @Wild2009; @Wild2012; @Xia2007; @Zerefos2009; and references therein].
-<!-- causes say less we don't do that -->
-
-There is a also a need to evalutate the trends on different location
+There is a also a need to evaluate the trends on different location
 

### 2023-06-30 DHI_GHI_0_variables.R
Commit 2023-06-30 13:17

@@ -31,11 +31,13 @@ translate <- function(x) {
 
 #### Data range ####
 ## will not include the last/first day
-LAST_DAY                 <- as.Date("2023-06-01")
+# LAST_DAY                 <- as.Date("2023-06-01") ## Cleaned data
 # LAST_DAY                 <- as.Date("2011-04-01") ## similar to Bais analysis
-FIRST_DAY                <- as.Date("1900-07-01")
+LAST_DAY                 <- as.POSIXct("2023-03-31 11:59:30") ## set the date of last NOAA TSI data
+FIRST_DAY                <- as.POSIXct("1900-07-01") ## include all available
 # FIRST_DAY                <- as.Date("2005-01-01") ## data inspected by me
 # FIRST_DAY                <- as.Date("2016-04-01") ## start of chp1
+
 
 #### Paths ####
 CLEARdir                 <- "~/DATA/Broad_Band/CS_id"

### 2023-06-30 Article.Rmd
Commit 2023-06-29 20:17
Commit 2023-06-29 21:17
Commit 2023-06-30 17:17

@@
 
-<!-- Sun variability -->
+<!-- Sun variability: should cite that there is no significant trend found Wild2021 -->
 
@@
 Subsequently, we made all measurements relative to the Total Solar Irradiance (TSI) at $1 \text{au}$, in order to compensate for the Sun's intensity variability, using a homogenized time series of satellite TSI observations.
-The TSI data we use, is a combination of data from NOAA [@Coddington2005] (for `r as.Date(TSIinfo[TSI_Source == "NOAA", Start])` - `r as.Date(TSIinfo[TSI_Source == "NOAA", End])`) and adjusted data from @LASP2023 (for `r as.Date(TSIinfo[TSI_Source != "NOAA", Start])` - `r as.Date(TSIinfo[TSI_Source != "NOAA", End])`).
+The TSI data we use, comes from NOAA [@Coddington2005].
+<!-- The TSI data we use, is a combination of data from NOAA [@Coddington2005] (for `r as.Date(TSIinfo[TSI_Source == "NOAA", Start])` - `r as.Date(TSIinfo[TSI_Source == "NOAA", End])`) and adjusted data from @LASP2023 (for `r as.Date(TSIinfo[TSI_Source != "NOAA", Start])` - `r as.Date(TSIinfo[TSI_Source != "NOAA", End])`).-->
 As a result, we can present all radiation data as a fraction of Sun's TSI.

### 2023-06-30 Article_A.tex
Commit 2023-06-30 17:17

@@
 \author{true \and true}
-\date{2023-06-29}
+\date{2023-06-30}
 
@@
 There are many cases on the long term records, showing a change of SDR trends direction, occurring roughly at the last decades of the 20th century.
-Yang et al. (2021) reports a general dimming before 2000 and a brightening afterwards on multiple station in China.
-Wild et al. (2021) shows a similar pattern, with the breaking point around 1980 from a 71 years record of a Central Europe station and for Brazil by Yamasoe et al. (2021).
+A general dimming before 2000 and a brightening afterwards, was found on multiple station in China (Yang et al. 2021).
+And a similar pattern was identified, with the breaking point around 1980, from a 71 years record of a Central Europe station (Wild et al. 2021) and for Brazil (Yamasoe et al. 2021).
 Another approach, using AI aided spatial analysis on continent level, of numerous observational stations data, reach similar conclusions for the above regions and the global trend (Yuan, Leirvik, and Wild 2021).
@@
 There is a consensus, among researchers, that the major factors of the attenuation of SDR is the interaction with the atmospheric aerosols and the clouds.
+Those interactions, among other factors, have been analysed with modeling (Li et al. 2016; Samset et al. 2018), showing the existence of feedback mechanisms between the two.
+And similar finds, have been showed in observational data (Schwarz et al. 2020; Ohvril et al. 2009; Zerefos et al. 2009; Xia et al. 2007 and references therein).
 
-The interaction between Aerosols and clouds, among other factors, has been analysed with modeling (Li et al. 2016; Samset et al. 2018), showing the existence of feedback mechanisms between the two.
-
-Similar links have been showed with observational data {[}Ohvril et al. (2009); Schwarz et al. (2020); {]}
+Wild (2012) : \textasciitilde{}
 
@@
 it is difficult to estimate the SDR trends on each site.
-
-Multiple studies attempt to evaluate the phenomenon and the potential causes {[}; Wild (2012); Xia et al. (2007); Zerefos et al. (2009); and references therein{]}.
-
-Investigate multiple interaction Li et al. (2016) Samset et al. (2018) data analysis and modeling
 
@@
 SDR data spans the period of
-1993-04-12 to 2023-05-31.
+1993-04-12 to 2023-03-31.
 In some algorithms we used as auxiliary data, the direct beam radiation (DNI) for comparisons, obtained by a CHP-1 pyrheliometer, with data availability from
-2016-04-01 to 2023-05-31.
+2016-04-01 to 2023-03-31.
 
@@
 Subsequently, we made all measurements relative to the Total Solar Irradiance (TSI) at \(1 \text{au}\), in order to compensate for the Sun's intensity variability, using a homogenized time series of satellite TSI observations.
-The TSI data we use, is a combination of data from NOAA (Coddington et al. 2005) (for 1993-04-12 - 2023-03-31) and adjusted data from LASP (2023) (for 2023-03-31 - 2023-05-31).
+The TSI data we use, comes from NOAA (Coddington et al. 2005).
 As a result, we can present all radiation data as a fraction of Sun's TSI.
@@
 \endhead
-0.381 & 3.479e-21 & All sky cond. & 100 \\
-0.3605 & 1.068e-16 & Clear sky cond. & 100 \\
-0.3472 & 8.968e-09 & Cloudy cond. & 100 \\
+0.3996 & 8.665e-23 & All sky cond. & 100 \\
+0.3773 & 7.075e-18 & Clear sky cond. & 100 \\
+0.3594 & 3.318e-09 & Cloudy cond. & 100 \\
 \bottomrule
@@
 0.7215 & 0.001209 & All sky cond. & Winter \\
-0.1679 & 0.1083 & All sky cond. & Spring \\
+0.1537 & 0.1498 & All sky cond. & Spring \\
 0.128 & 0.08705 & All sky cond. & Summer \\
@@
 0.8715 & 0.000269 & Clear sky cond. & Winter \\
-0.09719 & 0.3923 & Clear sky cond. & Spring \\
+0.05881 & 0.6247 & Clear sky cond. & Spring \\
 -0.02369 & 0.7972 & Clear sky cond. & Summer \\
@@
 -0.02581 & 0.9335 & Cloudy cond. & Winter \\
-0.3557 & 0.06914 & Cloudy cond. & Spring \\
+0.4333 & 0.02937 & Cloudy cond. & Spring \\
 0.442 & 0.004988 & Cloudy cond. & Summer \\
@@
 99.88 \\
-89.17 \\
+85.02 \\
 91.29 \\
@@
 99.97 \\
-60.77 \\
+37.53 \\
 20.28 \\
@@
 6.646 \\
-93.09 \\
+97.06 \\
 99.5 \\
@@
 
-Our result for all-sky condition (\(0.38\%/year\)), reaffirm the previous results of Bais et al. (2013) for the site.
+Our result for all-sky condition (\(0.4\%/year\)), reaffirm the previous results of Bais et al. (2013) for the site.
 The increase of this trend, shows that the phenomenon and probably the causes have been amplified.
@@
 \ldots..
-Also, the trend of \(0.35\%/year\) for cloud-sky condition indicate the major part that clouds play on the
+Also, the trend of \(0.36\%/year\) for cloud-sky condition indicate the major part that clouds play on the
 \ldots..
@@
 \url{https://doi.org/10.7289/V55B00C1}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-LASP2023}{}}%
-LASP. 2023. {``TSIS-1 Total Solar Irradiance - Six Hour Average.''} Laboratory for Atmospheric; Space Physics, University of Colorado, Boulder. 2023. \url{https://lasp.colorado.edu/tsis/data/}.
 

### 2023-06-30 DHI_GHI_0_data_input.R
Commit 2023-06-30 18:17

@@ -256,16 +256,16 @@ if (havetorun) {
     ##  Daily means  -----------------------------------------------------------
 
     ALL_1_daily_mean <-
-        DATA_all[,.(DIR_att       = mean(DIR_att,     na.rm = T),
-                    GLB_att       = mean(GLB_att,     na.rm = T),
-                    HOR_att       = mean(HOR_att,     na.rm = T),
-                    DIR_transp    = mean(DIR_transp,  na.rm = T),
-                    tsi1au_att    = mean(tsi_1au_comb,na.rm = T),
-                    DIR_att_sd    = sd(  DIR_att,     na.rm = T),
-                    GLB_att_sd    = sd(  GLB_att,     na.rm = T),
-                    HOR_att_sd    = sd(  HOR_att,     na.rm = T),
-                    DIR_transp_sd = sd(  DIR_transp,  na.rm = T),
-                    tsi1au_att_sd = sd(  tsi_1au_comb,na.rm = T),
+        DATA_all[,.(DIR_att       = mean(DIR_att,      na.rm = T),
+                    GLB_att       = mean(GLB_att,      na.rm = T),
+                    HOR_att       = mean(HOR_att,      na.rm = T),
+                    DIR_transp    = mean(DIR_transp,   na.rm = T),
+                    tsi1au_att    = mean(tsi_1au_comb, na.rm = T),
+                    DIR_att_sd    = sd(  DIR_att,      na.rm = T),
+                    GLB_att_sd    = sd(  GLB_att,      na.rm = T),
+                    HOR_att_sd    = sd(  HOR_att,      na.rm = T),
+                    DIR_transp_sd = sd(  DIR_transp,   na.rm = T),
+                    tsi1au_att_sd = sd(  tsi_1au_comb, na.rm = T),
                     doy           = yday(Date),
                     GLB_att_N     = sum(!is.na(GLB_att)),
                     HOR_att_N     = sum(!is.na(HOR_att)),
@@ -273,14 +273,16 @@ if (havetorun) {
                  by = .( Date = Day ) ]
 
     CLEAR_1_daily_mean <-
-        DATA_Clear[,.(DIR_att       = mean(DIR_att,   na.rm = T),
-                      GLB_att       = mean(GLB_att,   na.rm = T),
-                      HOR_att       = mean(HOR_att,   na.rm = T),
[…]

### 2023-06-30 DHI_GHI_1_longterm_trends.R
Commit 2023-06-30 19:17

@@ -206,7 +206,7 @@ rm(data_list)
 #'
 #+ LongtermTrends, echo=F, include=T, results="asis"
 # vars <- c("HOR_att","DIR_transp", "DIR_att", "GLB_att", "tsi1au_att")
-vars <- c("DIR_att_des", "GLB_att_des")
+vars <- c("DIR_att_des", "GLB_att_des", "tsi1au_att")
 # dbs  <- c("ALL_1_daily_DEseas")
 
 dbs         <- c(  "ALL_1_daily_DESEAS",

### 2023-06-30 Article.Rmd
Commit 2023-06-30 18:17
Commit 2023-06-30 20:17

@@
 
+```{r TSIslope, echo=F}
+TSI <- data.table(read.csv("./figures/tbl_longterm_trends.dat",
+               strip.white = TRUE,
+               sep = ";",
+               header = TRUE))
+TSI <- TSI[grep("TSI", TSI$var), ]
+TSI$slope.sd           <- NULL
+TSI$slope.ConfInt_0.95 <- NULL
+TSI$slope.ConfInt_0.99 <- NULL
+TSI$var                <- NULL
+setorder(TSI, -slope.stat_sig)
+rownames(TSI) <- NULL
+names(TSI) <- c("Trend [%/year]", "Trend p-value", "Sky condition", "Trend statistical signif. [%]")
+TSIall <- TSI["Sky condition" == ""]
+
+```
+
 <!-- # Detection of clear periods in GHI measurements for SDR trends. -->
@@
 Subsequently, we made all measurements relative to the Total Solar Irradiance (TSI) at $1 \text{au}$, in order to compensate for the Sun's intensity variability, using a homogenized time series of satellite TSI observations.
-The TSI data we use, comes from NOAA [@Coddington2005].
+The TSI data we use, comes from NOAA [@Coddington2005], it overlaps with our data, and is interpolated to match with the time step of the data.
 <!-- The TSI data we use, is a combination of data from NOAA [@Coddington2005] (for `r as.Date(TSIinfo[TSI_Source == "NOAA", Start])` - `r as.Date(TSIinfo[TSI_Source == "NOAA", End])`) and adjusted data from @LASP2023 (for `r as.Date(TSIinfo[TSI_Source != "NOAA", Start])` - `r as.Date(TSIinfo[TSI_Source != "NOAA", End])`).-->

### 2023-06-30 Article.Rmd
Commit 2023-06-30 21:17

--- 2023-06-30 Article.Rmd
@@
 Those interactions, among other factors, have been analysed with modeling [@Li2016; @Samset2018], showing the existence of feedback mechanisms between the two.
-And similar finds, have been showed in observational data [@Schwarz2020; @Ohvril2009; @Zerefos2009; @Xia2007 and references therein].
+Similar finds, have been showed in observational data [@Schwarz2020; @Ohvril2009; @Zerefos2009; @Xia2007 and references therein].
+<!-- Sun variability: should cite that there is no significant trend found Wild2021 -->
 
-<!-- Sun variability: should cite that there is no significant trend found Wild2021 -->
+Due to the variability of the phenomenon, and its contributing factors, there is a need to further investigate SDR tendencies, in different sites, to estimate its magnitude, and its relation to the local conditions.
+In this study, we examine the trends of SDR, with ground-based measurements at Thessaloniki for the period 1993 - 2023, as derived from a CM-21 pyranometer.
+<!-- goals -->
+We reevaluated and extended the dataset used by @Bais2013, applying a different algorithm for the identification of clear-/cloud-sky instances [@Reno2016; @Reno2012a], and we derive the radiation trends for the period `r year(min(ALL_1_daily_DESEAS$Date))` to `r year(max(ALL_1_daily_DESEAS$Date))`, under different sky conditions (all-sky, clear-sky and cloud-sky).
 
@@
 
-Due to the variability of those factors and the interaction between them,
-
 ...... cloud formation occurrences and constitution nuclei. cite.....
-
-it is difficult to estimate the SDR trends on each site.
-
-There is a also a need to evaluate the trends on different location
-
-In this study, we examine the trends of SDR with ground-based measurements at Thessaloniki for the period 1993 - 2023, as derived from a CM-21 pyranometer.
-<!-- goals -->
-We reevaluate and extend the dataset used by @Bais2013, applying a different algorithm for the identification of clear-/cloud-sky instances, and we derive the radiation trends for this period under different sky conditions (all-sky, clear-sky and cloud-sky).
 
@@
 ...
-
-<!-- basic concepts -->
-Our method to characterize the sky conditions and the definition of each sky condition has some subjectivity.
-The algorithm was calibrated with the main focus to identify the presence of clouds on the sky dome, although there are marginal cases that there will be false positives or false negatives identifications.
 
@@
 
-<!-- CS_id v14.2 inserted here -->
+<!-- basic concepts -->
+Our method to characterize the sky conditions and the definition of each sky condition has some subjectivity.
+The algorithm was calibrated with the main focus to identify the presence of clouds on the sky dome, although there are marginal cases that there will be false positives or false negatives identifications.
+
 ```{r setup, echo=F, include=T, eval=TRUE}
-## Load Clear Sky id variables
+## Load Clear Sky id v14.2 variables
 MS <- data.frame(readRDS("~/CS_id/PARAMS/Clear_sky_id_Reno-Hansen_apply_v14_2.Rds"))
-# export all variables
+# export all Clear Sky variables
 for (an in names(MS)) {
@@
 }
+## get TSI slope for reference
+TSI <- data.table(read.csv("./figures/tbl_longterm_trends.dat",
+               strip.white = TRUE,
+               sep = ";",
+               header = TRUE))
+TSI <- TSI[grep("TSI", TSI$var), ]
+TSI$slope.sd           <- NULL
+TSI$slope.ConfInt_0.95 <- NULL
+TSI$slope.ConfInt_0.99 <- NULL
+TSI$var                <- NULL
+setorder(TSI, -slope.stat_sig)
+rownames(TSI) <- NULL
+names(TSI) <- c("Trend [%/year]", "Trend p-value", "Sky condition", "Trend statistical signif. [%]")
+TSIskopeAll <- TSI[`Sky condition` == "All sky cond."]
 ```

### 2023-07-02 Article.Rmd
Commit 2023-07-02 11:17

@@
 
-<!-- SDR or GHI ? -->
-
-Cloud “shrinking” and “optical thinning” in the “dimming” period and a subsequent recovery in the “brightening” period over China
-
-aerosols not only strengthen but also weaken the
-growth of clouds with respect to their coverage and
-optical thickness, depending on the levels of pollution
-and the associated amounts of aerosols as postulated
-in a conceptual framework (Wild 2009a, 2012, Yang
-et al 2012).
+<!-- use SDR for GHI -->
 
@@
 Due to the variability of the phenomenon, and its contributing factors, there is a need to further investigate SDR tendencies, in different sites, to estimate its magnitude, and its relation to the local conditions.
-In this study, we examine the trends of SDR, with ground-based measurements at Thessaloniki for the period 1993 - 2023, as derived from a CM-21 pyranometer.
+In this study, we examine the trends of SDR, with ground-based measurements at Thessaloniki, Greece for the period `r year(min(ALL_1_daily_DESEAS$Date))` to `r year(max(ALL_1_daily_DESEAS$Date))`, as derived from a CM-21 pyranometer.
 <!-- goals -->
 We reevaluated and extended the dataset used by @Bais2013, applying a different algorithm for the identification of clear-/cloud-sky instances [@Reno2016; @Reno2012a], and we derive the radiation trends for the period `r year(min(ALL_1_daily_DESEAS$Date))` to `r year(max(ALL_1_daily_DESEAS$Date))`, under different sky conditions (all-sky, clear-sky and cloud-sky).
-
-@Wild2012 : ~
-
-...... cloud formation occurrences and constitution nuclei. cite.....
 
@@
 ...
+
+<!--
+Aerosols interactions
+
+#Yang2021  Cloud “shrinking” and “optical thinning” in the “dimming” period and a subsequent recovery in the “brightening” period over China
+
+Aerosols not only strengthen but also weaken the
+growth of clouds with respect to their coverage and
+optical thickness, depending on the levels of pollution
+and the associated amounts of aerosols as postulated
+in a conceptual framework (Wild 2009a, 2012, Yang
+et al 2012).
+
+...... cloud formation occurrences and constitution nuclei. cite.....
+
+#Wild2012 : simplify interaction nice graph
+
+-->
 

### 2023-07-02 Article.Rmd
Commit 2023-07-02 12:17

@@
 
-The shortwave downward solar irradiance (SDR) at the Earth's surface plays a significant role, on Earths climate.
-Changes of the SDR can be related to major changes on the Earth's energy budget, the mechanisms of climate change, and water and carbon cycle [@Wild2009].
-Studies of SDR variability, have identified some distinct SDR trends on different regions of the world and for different time periods, using the term 'brightening' for positive trends, and 'dimming' for negative trends.
-There are many cases on the long term records, showing a change of SDR trends direction, occurring roughly at the last decades of the 20th century.
+The shortwave downward solar irradiance (SDR) at Earth's surface play a significant role, on it's climate.
+Changes of the SDR can be related to changes on Earth's energy budget, the mechanisms of climate change, and water and carbon cycle [@Wild2009].
+Can also affect, solar and agricultural production, and living organisms.
+Studies of SDR variability, have identified some distinct SDR trends on different regions of the world and for different time periods.
+The term 'brightening' is generally used for periods of positive trends of SDR, and 'dimming' for negative trends.
+There are many cases on the long term records of irradiance, showing a systematic change of SDR's trend slope, occurring roughly at the last decades of the 20th century.
 <!-- Trends -->
-A general dimming before 2000 and a brightening afterwards, was found on multiple station in China [@Yang2021]. 
-And a similar pattern was identified, with the breaking point around 1980, from a 71 years record of a Central Europe station [@Wild2021] and for Brazil [@Yamasoe2021].
+On multiple station in China, a dimming period was reported until about 2000, followed by a brightening period [@Yang2021]. 
+A similar pattern was identified, with the breaking point around 1980, from stations in Central Europe [@Wild2021] and Brazil [@Yamasoe2021].
 <!-- @Yamasoe2021 similar dimming and brightening around 1980 in Brazil -->

### 2023-07-02 Article.Rmd
Commit 2023-07-02 13:17

@@
 Studies of SDR variability, have identified some distinct SDR trends on different regions of the world and for different time periods.
-The term 'brightening' is generally used for periods of positive trends of SDR, and 'dimming' for negative trends.
+The term 'brightening' is generally used to describe periods of positive SDR trend, and 'dimming' for negative trend.
 There are many cases on the long term records of irradiance, showing a systematic change of SDR's trend slope, occurring roughly at the last decades of the 20th century.
-<!-- Trends -->
+<!-- Trends china -->
 On multiple station in China, a dimming period was reported until about 2000, followed by a brightening period [@Yang2021]. 
-A similar pattern was identified, with the breaking point around 1980, from stations in Central Europe [@Wild2021] and Brazil [@Yamasoe2021].
-<!-- @Yamasoe2021 similar dimming and brightening around 1980 in Brazil -->
-Another approach, using AI aided spatial analysis on continent level, of numerous observational stations data, reach similar conclusions for the above regions and the global trend [@Yuan2021].
+<!-- Trends Europe Brazil -->
+A similar pattern was identified, with the breaking point around 1980, for stations in Central Europe [@Wild2021] and Brazil [@Yamasoe2021].
+<!-- Trends Global -->
+Also, on global scale, an AI aided continental level spatial analysis, with data from multiple station, reach similar conclusions for the above regions and for the global trend [@Yuan2021].
 
-There is a consensus, among researchers, that the major factors of the attenuation of SDR is the interaction with the atmospheric aerosols and the clouds. 
-Those interactions, among other factors, have been analysed with modeling [@Li2016; @Samset2018], showing the existence of feedback mechanisms between the two.
+There is a consensus, among researchers, that major factors of SDR attenuation is the interaction with atmospheric aerosols and clouds. 
+Those interactions, among other factors, have been analysed with models [@Li2016; @Samset2018], showing the existence of feedback mechanisms between the two.
 Similar finds, have been showed in observational data [@Schwarz2020; @Ohvril2009; @Zerefos2009; @Xia2007 and references therein].
@@
 
-Due to the variability of the phenomenon, and its contributing factors, there is a need to further investigate SDR tendencies, in different sites, to estimate its magnitude, and its relation to the local conditions.
+Due to the variability of the phenomenon, and its contributing factors, there is a need to further investigate and monitor SDR, in different sites, to estimate its magnitude, and its relation to the local conditions.
 In this study, we examine the trends of SDR, with ground-based measurements at Thessaloniki, Greece for the period `r year(min(ALL_1_daily_DESEAS$Date))` to `r year(max(ALL_1_daily_DESEAS$Date))`, as derived from a CM-21 pyranometer.
@@
 We reevaluated and extended the dataset used by @Bais2013, applying a different algorithm for the identification of clear-/cloud-sky instances [@Reno2016; @Reno2012a], and we derive the radiation trends for the period `r year(min(ALL_1_daily_DESEAS$Date))` to `r year(max(ALL_1_daily_DESEAS$Date))`, under different sky conditions (all-sky, clear-sky and cloud-sky).
-
-...
-
-Site location description and aerosols effect
-
-...
 
@@
 # Data and methodology.
+
+Location
+
+The SDR data used in this study cover the period from
+January 1993 to March 2011, and were measured with
+a Kipp & Zonen CM-21 pyranometer operating contin-
+uously at the Laboratory of Atmospheric Physics of the
+Aristotle University of Thessaloniki ($40^\deg$38"N, $22^\circ$57"E,
+∼80 m above sea level). The monitoring site is located
+near the city centre. The horizon is clear of obstruc-
+◦ ◦
+tions except for the azimuth angles between 35 and 120where obstacles of up to 10 elevation block the direct
+irradiance in the morning for SZA larger than 80 .
+These data points were excluded from the analysis, as
+
+Instruments and callibrations
+
+relation to aerosols
+
+Northern Greece has been selected because
+it is well known to be located at the crossroads of aerosols and
+various pollutants pathways originating in central and Eastern
+Europe (Lelieveld et al., 2002; Zerefos et al., 2002). Thessa-
+loniki, the major city in northern Greece, lies along the exit of air
+pollutants from Europe, in the path of the prevailing long-range
+transport (LRT) trajectories, as we move from central and east-
+ern Europe to the Mediterranean (Zerefos et. al., 2002). On the
 

### 2023-07-02 Article.tex
Commit 2023-07-02 13:17

--- 2023-06-26 Article.tex
@@
 
-Cloud ``shrinking'' and ``optical thinning'' in the ``dimming'' period
-and a subsequent recovery in the ``brightening'' period over China
-
-aerosols not only strengthen but also weaken the growth of clouds with
-respect to their coverage and optical thickness, depending on the levels
-of pollution and the associated amounts of aerosols as postulated in a
-conceptual framework (Wild 2009a, 2012, Yang et al 2012).
-
 \hypertarget{introduction.}{%
@@
 
-The shortwave downward solar irradiance (SDR) at the Earth's surface
-plays a significant role, on Earths climate. Changes of the SDR can be
-related to major changes on factors affecting the radiative forcing
-balance and human activities. Multiple studies attempt to evaluate the
-phenomenon and the potential causes
-\citep{Ohvril2009, Schwarz2020, Wild2009, Wild2012, Xia2007, Zerefos2009}.
-Although, there is regional variation of SDR, with both positive and
-negative trends, most of the studies agree that the major factors are
-the interaction of aerosols and clouds.
+The shortwave downward solar irradiance (SDR) at Earth's surface play a
+significant role, on it's climate. Changes of the SDR can be related to
+changes on Earth's energy budget, the mechanisms of climate change, and
+water and carbon cycle \citep{Wild2009}. Can also affect, solar and
+agricultural production, and living organisms. Studies of SDR
+variability, have identified some distinct SDR trends on different
+regions of the world and for different time periods. The term
+`brightening' is generally used to describe periods of positive SDR
+trend, and `dimming' for negative trend. There are many cases on the
+long term records of irradiance, showing a systematic change of SDR's
+trend slope, occurring roughly at the last decades of the 20th century.
+On multiple station in China, a dimming period was reported until about
+2000, followed by a brightening period \citep{Yang2021}. A similar
+pattern was identified, with the breaking point around 1980, for
+stations in Central Europe \citep{Wild2021} and Brazil
+\citep{Yamasoe2021}. Also, on global scale, an AI aided continental
+level spatial analysis, with data from multiple station, reach similar
+conclusions for the above regions and for the global trend
+\citep{Yuan2021}.
 
-In this study, we examine the trends of SDR with ground-based
-measurements at Thessaloniki for the period 1993 - 2023, as derived from
-a CM-21 pyranometer. We reevaluate and extend the dataset used by
-\citet{Bais2013}, applying a different algorithm for the identification
-of clear-/cloud-sky instances, and we derive the radiation trends for
-this period under different sky conditions (all-sky, clear-sky and
-cloud-sky).
+There is a consensus, among researchers, that major factors of SDR
+attenuation is the interaction with atmospheric aerosols and clouds.
+Those interactions, among other factors, have been analysed with models
+\citep{Li2016, Samset2018}, showing the existence of feedback mechanisms
+between the two. Similar finds, have been showed in observational data
+\citep[ and references
+therein]{Schwarz2020, Ohvril2009, Zerefos2009, Xia2007}.
 
-\ldots{}
-
-Site location description and aerosols effect
-
-\ldots{}
-
-Our method to characterize the sky conditions and the definition of each
-sky condition has some subjectivity. The algorithm was calibrated with
-the main focus to identify the presence of clouds on the sky dome,
-although there are marginal cases that there will be false positives or
-false negatives identifications.
+Due to the variability of the phenomenon, and its contributing factors,
+there is a need to further investigate and monitor SDR, in different
+sites, to estimate its magnitude, and its relation to the local
+conditions. In this study, we examine the trends of SDR, with
+ground-based measurements at Thessaloniki, Greece for the period 1993 to
+2023, as derived from a CM-21 pyranometer. We reevaluated and extended
+the dataset used by \citet{Bais2013}, applying a different algorithm for
+the identification of clear-/cloud-sky instances
+\citep{Reno2016, Reno2012a}, and we derive the radiation trends for the
+period 1993 to 2023, under different sky conditions (all-sky, clear-sky
+and cloud-sky).
 
@@
 
+Location
+
+The SDR data used in this study cover the period from January 1993 to
+March 2011, and were measured with a Kipp \& Zonen CM-21 pyranometer
+operating contin- uously at the Laboratory of Atmospheric Physics of the
+Aristotle University of Thessaloniki (\$40\^{}\circ\$38''N,
+\$22\^{}\circ\$57''E, ∼80 m above sea level). The monitoring site is
+located near the city centre. The horizon is clear of obstruc- ◦ ◦ tions
+except for the azimuth angles between 35 and 120where obstacles of up to
+∼10◦ elevation block the direct irradiance in the morning for SZA larger
+than ∼80◦ . These data points were excluded from the analysis, as
+
+Instruments and callibrations
+
+relation to aerosols
+
+Beijing (40 N, 116 E), which are representative of regions with high
+aerosol load. Northern Greece has been selected because it is well known
+to be located at the crossroads of aerosols and various pollutants
+pathways originating in central and Eastern Europe (Lelieveld et al.,
+2002; Zerefos et al., 2002). Thessa- loniki, the major city in northern
+Greece, lies along the exit of air pollutants from Europe, in the path
+of the prevailing long-range transport (LRT) trajectories, as we move
+from central and east- ern Europe to the Mediterranean (Zerefos et. al.,
+2002). On the
+
 The SDR is equivalent to the whole sky Global radiation, also refered as
@@
 leveled, CM-21 pyranometer. SDR data spans the period of 1993-04-12 to
-2023-05-31. In some algorithms we used as auxiliary data, the direct
+2023-03-31. In some algorithms we used as auxiliary data, the direct
 beam radiation (DNI) for comparisons, obtained by a CHP-1 pyrheliometer,
-with data availability from 2016-04-01 to 2023-05-31.
+with data availability from 2016-04-01 to 2023-03-31.
 
@@
 \ldots\ldots{}
+
+Our method to characterize the sky conditions and the definition of each
+sky condition has some subjectivity. The algorithm was calibrated with
+the main focus to identify the presence of clouds on the sky dome,
+although there are marginal cases that there will be false positives or
+false negatives identifications.
 
@@
 the Sun's intensity variability, using a homogenized time series of
-satellite TSI observations. The TSI data we use, is a combination of
-data from NOAA \citep{Coddington2005} (for 1993-04-12 - 2023-03-31) and
-adjusted data from \citet{TSIStsi} (for 2023-03-31 - 2023-05-31). As a
-result, we can present all radiation data as a fraction of Sun's TSI.
+satellite TSI observations. The TSI data we use, comes from NOAA
+\citep{Coddington2005}, it overlaps with our data, and is interpolated
+to match with the time step of the data. As a result, we can present all
+radiation data as a fraction of Sun's TSI.
 
@@
 
-\caption{Anomaly (\%) of the daily SDR relative to climatological values for 1993 - 2023. The black line shows the long term trend for all-sky conditions.}\label{fig:trendALL}
+\caption{Anomaly (\%) of the daily SDR relative to climatological values for 1993 - 2023. The black line shows the long term trend for all-sky conditions.}(\#fig:trendALL)
 \end{figure}
@@
 \endlastfoot
-0.381 & 3.479e-21 & All sky cond. & 100 \\
-0.3605 & 1.068e-16 & Clear sky cond. & 100 \\
-0.3472 & 8.968e-09 & Cloudy cond. & 100 \\
+0.3996 & 8.665e-23 & All sky cond. & 100 \\
+0.3773 & 7.075e-18 & Clear sky cond. & 100 \\
+0.3594 & 3.318e-09 & Cloudy cond. & 100 \\
 \end{longtable}
@@
 
-{\centering \subfloat[Trend distribution for all-sky conditions.\label{fig:szatrends-1}]{\includegraphics[width=.35\linewidth]{./images/SzaTrends-13} }\subfloat[Trend distribution for cloud-sky conditions.\label{fig:szatrends-2}]{\includegraphics[width=.35\linewidth]{./images/SzaTrends-21} }
+{\centering \subfloat[Trend distribution for all-sky conditions.(\#fig:szatrends-1)]{\includegraphics[width=.35\linewidth]{./images/SzaTrends-13} }\subfloat[Trend distribution for cloud-sky conditions.(\#fig:szatrends-2)]{\includegraphics[width=.35\linewidth]{./images/SzaTrends-21} }
 
@@
 
-\caption{Distribution of the SDR's long term trends by SZA.}\label{fig:szatrends}
+\caption{Distribution of the SDR's long term trends by SZA.}(\#fig:szatrends)
 \end{figure}
@@
 
-\caption{Trends by season for all condition. Displaying monthly means of daily means.}\label{fig:seasonalALL}
+\caption{Trends by season for all condition. Displaying monthly means of daily means.}(\#fig:seasonalALL)
 \end{figure}
@@
 
-{\centering \subfloat[one plot\label{fig:cumsum-1}]{\includegraphics[width=.35\linewidth]{./images/CumulativeDailyCumSum-1} }\subfloat[the other one\label{fig:cumsum-2}]{\includegraphics[width=.35\linewidth]{./images/CumulativeDailyCumSum-9} }
+{\centering \subfloat[one plot(\#fig:cumsum-1)]{\includegraphics[width=.35\linewidth]{./images/CumulativeDailyCumSum-1} }\subfloat[the other one(\#fig:cumsum-2)]{\includegraphics[width=.35\linewidth]{./images/CumulativeDailyCumSum-9} }
 
@@
 
-\caption{two plots}\label{fig:cumsum}
+\caption{two plots}(\#fig:cumsum)
 \end{figure}
@@
 
-Our result for all-sky condition (\(0.38\%/year\)), reaffirm the
-previous results of \citet{Bais2013} for the site. The increase of this
-trend, shows that the phenomenon and probably the causes have been
-amplified.
+Our result for all-sky condition (\(0.4\%/year\)), reaffirm the previous
+results of \citet{Bais2013} for the site. The increase of this trend,
+shows that the phenomenon and probably the causes have been amplified.
 
-\ldots.. Also, the trend of \(0.35\%/year\) for cloud-sky condition
+\ldots.. Also, the trend of \(0.36\%/year\) for cloud-sky condition
 indicate the major part that clouds play on the \ldots..

### 2023-07-02 Article.Rmd
Commit 2023-07-02 14:17

@@
 
+<!-- time span -->
+The SDR is equivalent to the whole sky Global radiation (or Global Horizontal Irradiance, GHI), and was obtained with a horizontal leveled, CM-21 pyranometer, from \mbox{`r min(ALL_1_daily_DESEAS$Date)`} to \mbox{`r max(ALL_1_daily_DESEAS$Date)`}.
+Also, we have measurements of direct beam radiation (DNI) obtained by a collocated CHP-1 pyrheliometer, from \mbox{`r min(ALL_1_daily_DESEAS[!is.na(DIR_att), Date])`} to \mbox{`r max(ALL_1_daily_DESEAS[!is.na(DIR_att), Date])`}
+
+In some algorithms, we used as auxiliary data, the direct beam radiation (DNI) for comparisons, obtained by a CHP-1 pyrheliometer, with data availability from  
+`r min(ALL_1_daily_DESEAS[!is.na(DIR_att), Date])` to `r max(ALL_1_daily_DESEAS[!is.na(DIR_att), Date])`.
+
 Location
 
-The SDR data used in this study cover the period from
-January 1993 to March 2011, and were measured with
-a Kipp & Zonen CM-21 pyranometer operating contin-
-uously at the Laboratory of Atmospheric Physics of the
-Aristotle University of Thessaloniki ($40^\deg$38"N, $22^\circ$57"E,
-∼80 m above sea level). The monitoring site is located
+The SDR data used in this study cover the period from January 1993 to March 2011, and were measured with a Kipp & Zonen CM-21 pyranometer operating continuously at the Laboratory of Atmospheric Physics of the Aristotle University of Thessaloniki ( $40^\circ38'$N, $22^\circ57'$E, 80 m above sea level).
+The monitoring site is located
 near the city centre. The horizon is clear of obstruc-
@@
 ern Europe to the Mediterranean (Zerefos et. al., 2002). On the
-
-<!-- time span -->
-The SDR is equivalent to the whole sky Global radiation, also refered as Global Horizontal Irradiance (GHI), and was obtained with a horizontal leveled, CM-21 pyranometer.
-SDR data spans the period of 
-`r min(ALL_1_daily_DESEAS$Date)` to `r max(ALL_1_daily_DESEAS$Date)`.
-In some algorithms we used as auxiliary data, the direct beam radiation (DNI) for comparisons, obtained by a CHP-1 pyrheliometer, with data availability from 
-`r min(ALL_1_daily_DESEAS[!is.na(DIR_att), Date])` to `r max(ALL_1_daily_DESEAS[!is.na(DIR_att), Date])`.
 
@@
 Using the mean daily SDR we produce the trends of each data set, as departure from the seasonal value.
-Only for All-sky (Fig.
-\ref{fig:trendALL}) <!--and Cloudy-sky (Fig.
-\ref{fig:trendCLOUD})--> conditions, the statistical significance of the trends are acceptable (Tab.
-\ref{tab:trendtable}), to draw some conclusions.
+Only for All-sky (Fig.~\ref{fig:trendALL}) <!--and Cloudy-sky (Fig.
+\ref{fig:trendCLOUD})--> conditions, the statistical significance of the trends are acceptable (Tab.~\ref{tab:trendtable}), to draw some conclusions.
 

### 2023-07-02 Article.Rmd
Commit 2023-07-02 15:17

@@
 
-# Data and methodology.
+# Observational data and methodology.
 
 <!-- time span -->
-The SDR is equivalent to the whole sky Global radiation (or Global Horizontal Irradiance, GHI), and was obtained with a horizontal leveled, CM-21 pyranometer, from \mbox{`r min(ALL_1_daily_DESEAS$Date)`} to \mbox{`r max(ALL_1_daily_DESEAS$Date)`}.
-Also, we have measurements of direct beam radiation (DNI) obtained by a collocated CHP-1 pyrheliometer, from \mbox{`r min(ALL_1_daily_DESEAS[!is.na(DIR_att), Date])`} to \mbox{`r max(ALL_1_daily_DESEAS[!is.na(DIR_att), Date])`}
+The SDR is equivalent to the whole sky Global radiation (or Global Horizontal Irradiance, GHI), and was obtained with a horizontal leveled, CM-21 pyranometer, from 
+<!-- \mbox{ -->
+`r min(ALL_1_daily_DESEAS$Date)`
+<!-- } -->
+to
+<!-- \mbox{ -->
+`r max(ALL_1_daily_DESEAS$Date)`.
+<!-- }. -->
+Also, we have measurements of direct beam radiation (DNI) obtained by a collocated CHP-1 pyrheliometer, from 
+<!-- \mbox{ -->
+`r min(ALL_1_daily_DESEAS[!is.na(DIR_att), Date])`
+<!-- } -->
+to
+<!-- \mbox{ -->
+`r max(ALL_1_daily_DESEAS[!is.na(DIR_att), Date])`.
+<!-- }. -->
+Although, we performed similar analysis to DNI as for SDR, due to the sorter duration of the data, the results are not presented here, as they lack the appropriate statistical significance.
 
-In some algorithms, we used as auxiliary data, the direct beam radiation (DNI) for comparisons, obtained by a CHP-1 pyrheliometer, with data availability from  
-`r min(ALL_1_daily_DESEAS[!is.na(DIR_att), Date])` to `r max(ALL_1_daily_DESEAS[!is.na(DIR_att), Date])`.
+Moreover, the DNI was used as auxiliary data, with the clear sky identification algorithm (CSid), to aid with the selection of the appropriate thresholds. 
+We have to note that, despite the capability of the CSid to use the DNI as an characterization parameter,
+
+.... wi dindnt use it  .....
+
+but was not used as indicator of clear sky instances
 

### 2023-07-02 Article.Rmd
Commit 2023-07-02 20:17

@@
 <!-- }. -->
-Although, we performed similar analysis to DNI as for SDR, due to the sorter duration of the data, the results are not presented here, as they lack the appropriate statistical significance.
+Although, we performed a similar analysis to the DNI as for the SDR, the results are not presented here, as they lack the appropriate statistical significance, due to the sorter duration of the data.
+Moreover, the DNI was used as auxiliary data, with the clear sky identification algorithm (CSid), to aid with the selection of the appropriate thresholds. 
+We have to note also, that despite the capability of the CSid to use the DNI as an characterization parameter, we haven't use it, to avoid any selection bias due to unequal availability of measurements.
 
-Moreover, the DNI was used as auxiliary data, with the clear sky identification algorithm (CSid), to aid with the selection of the appropriate thresholds. 
-We have to note that, despite the capability of the CSid to use the DNI as an characterization parameter,
+<!-- Location -->
 
-.... wi dindnt use it  .....
-
-but was not used as indicator of clear sky instances
-
-Location
-
-The SDR data used in this study cover the period from January 1993 to March 2011, and were measured with a Kipp & Zonen CM-21 pyranometer operating continuously at the Laboratory of Atmospheric Physics of the Aristotle University of Thessaloniki ( $40^\circ38'$N, $22^\circ57'$E, 80 m above sea level).
-The monitoring site is located
-near the city centre. The horizon is clear of obstruc-
-◦ ◦
-tions except for the azimuth angles between 35 and 120where obstacles of up to 10 elevation block the direct
-irradiance in the morning for SZA larger than 80 .
-These data points were excluded from the analysis, as
-
-Instruments and callibrations
-
-relation to aerosols
-
-Northern Greece has been selected because
-it is well known to be located at the crossroads of aerosols and
-various pollutants pathways originating in central and Eastern
-Europe (Lelieveld et al., 2002; Zerefos et al., 2002). Thessa-
-loniki, the major city in northern Greece, lies along the exit of air
-pollutants from Europe, in the path of the prevailing long-range
-transport (LRT) trajectories, as we move from central and east-
-ern Europe to the Mediterranean (Zerefos et. al., 2002). On the
-
-...
-
-...
-
-<!-- ## Preparation. -->
-
+The SDR data were measured with a Kipp & Zonen CM-21 pyranometer operating continuously at the Laboratory of Atmospheric Physics of the Aristotle University of Thessaloniki ($40^\circ38'$N, $22^\circ57'$E, 80m a.s.l.).
+The monitoring site is located near the city centre.
+During the study period the CM-21 pyranometer has been independently calibrated four times.
+<!-- Preparation. -->
 There are three distinct steps to the creation of this dataset:
@@
 
-<!-- ### Acquisition of radiometric data. -->
+<!--
+The horizon is clear of obstructions except for the azimuth angles between 35 and 120 where obstacles of up to 10 elevation block the direct irradiance in the morning for SZA larger than 80.
+These data points were excluded from the analysis, as
+-->
 
-<!-- measurement -->
+<!-- Zerefos
+
+Northern Greece has been selected because it is well known to be located at the crossroads of aerosols and various pollutants pathways originating in central and Eastern
+Europe (Lelieveld et al., 2002; Zerefos et al., 2002). Thessa-
+loniki, the major city in northern Greece, lies along the exit of air
+pollutants from Europe, in the path of the prevailing long-range
+transport (LRT) trajectories, as we move from central and east-
+ern Europe to the Mediterranean (Zerefos et. al., 2002). On the
+-->
+
+<!-- a) Acquisition -->
 For the acquisition of radiometric data,
-the signal of the broadband instruments is sampled with a rate of $1 \text{Hz}$. The mean and standard deviation values are recorded for every minute.
+the signal of the broadband instruments is sampled with a rate of $1 \text{Hz}$. The mean and its standard deviation are recorded for every minute.
 <!-- dark correction -->
@@
 
-<!-- ### Radiation data quality check -->
+<!-- b) Radiation data quality check -->
 <!-- The radiation quality is assured with a manual screening, aided by an automatic assessment process. -->
 <!-- screening -->
-A manual screening is performed, to remove inconsistencies and erroneous recordings, that can occur randomly or systematically, during the long continuous operation of the instrument.
+A manual screening was performed, to remove inconsistencies and erroneous recordings, that can occur randomly or systematically, during the long continuous operation of the instrument.
 <!-- Quality assurance -->
-The manual screening is aided by a radiation data quality assurance procedure, adjusted for this site, based on methods of Long and Shi\ [@Long2008a ; @Long2006].
+The manual screening is aided by a radiation data quality assurance procedure, adjusted for this site, based on methods of Long and Shi\ [-@Long2008a; -@Long2006].
 Thus, problematic recordings have been excluded from further processing. 
@@
 
-## Clear sky identification
+<!-- c) Clear sky identification -->
+In order to estimate the effect of the clouds on the SDR, we created three datasets, by characterizing each one-minute measurement with a corresponding sky condition. 
+The all-sky conditions, containing all the valid measurements. 
+The clear-sky conditions data, correspond to data where we have inferred that the sky was almost clear of clouds, and the remainder part the data was characterized with cloud sky conditions (cloud-sky).
+To identify the clear-sky conditions we used the method proposed by @Long2000 and by @Reno2016, that was adapted and configured for the site.
+
+....
+
+cite iterative process
 
@@
 
+<!-- CS subjectivity -->
+We have to note, that the definition of what constitutes as clear or cloudy sky, has some subjectivity, due to the used method of characterization.
+As a result, the details of the definition are site specific, it relies on a combination of threshold of comparisons with ideal actinometric models and statistics on different signal behavior.
+
 <!-- basic concepts -->
-Our method to characterize the sky conditions and the definition of each sky condition has some subjectivity.
 The algorithm was calibrated with the main focus to identify the presence of clouds on the sky dome, although there are marginal cases that there will be false positives or false negatives identifications.
@@
 <!-- 1. MeanVIP -->
-### Mean value of irradiance during the time period.
+<!-- ### Mean value of irradiance during the time period. -->
 
@@
 <!-- 2. MaxVIP -->
-### Max value of irradiance during the time period.
+<!-- ### Max value of irradiance during the time period. -->
 
@@
 <!-- 3. VIL -->
-### Variability in irradiance by line length.
+<!-- ### Variability in irradiance by line length. -->
 
@@
 <!-- 4. VCT -->
-### Variance of Changes in the Time series.
+<!-- ### Variance of Changes in the Time series. -->
 
@@
 \begin{equation}
+s_i = \frac{\text{SDR}_{i+1} - \text{SDR}_{i}}{t_{i+1} - t_i}, \forall i \in \left \{ 1, 2, \ldots, n-1 \right \} \label{eq:VCT1} \\
+\bar{s} = \frac{1}{n-1} \sum_{i=1}^{n-1} s_i \label{eq:VCT2} \\
+\sigma_i = \frac {\sqrt{\frac{1}{n-1} \sum_{i=1}^{n-1} \left( s_i - \bar{s} \right)^2} } {\bar{G_i}} \label{eq:VCT3}
+\end{equation}
+
+\begin{equation}
 \sigma_i < `r offVCT`
@@
 <!-- 5. VSM -->
-### Variability in the Shape of the irradiance Measurements.
+<!-- ### Variability in the Shape of the irradiance Measurements. -->
 
@@
 \end{equation}
-
-....
-
-<!-- ### Clear sky ID -->
-In order to estimate the effect of the clouds on the SDR we created three datasets, by characterizing each one-minute measurement with a corresponding sky condition. 
-The all-sky conditions, containing all the valid measurements. 
-The clear-sky conditions data, where we have inferred that the sky was almost clear of clouds, and the remainder part the data as cloudy sky conditions data (cloud-sky). 
-To identify the clear-sky conditions we used the method proposed by @Long2000 and by @Reno2016, that was adapted and configured for the site.
-<!-- CS subjectivity -->
-We have to note, that the definition of what constitutes as clear or cloudy sky, has some subjectivity, due to the used method of characterization.
-As a result, the details of the definition are site specific, it relies on a combination of threshold of comparisons with ideal actinometric models and statistics on different signal behavior.
-
-## Data and data selection.
 

### 2023-07-02 Article_B147.Rmd
Commit 2023-07-02 21:17

--- 2023-07-02 Article.Rmd
@@
 author:
-  - name:         Natsis Athanasios
+  - name:         Athanasios Natsis 
     email:        natsisphysicist@gmail.com
@@
 
+<!--
 \begin{gather}
@@
 \end{gather}
+-->
 
 \begin{equation}
-s_i = \frac{\text{SDR}_{i+1} - \text{SDR}_{i}}{t_{i+1} - t_i}, \forall i \in \left \{ 1, 2, \ldots, n-1 \right \} \label{eq:VCT1} \\
-\bar{s} = \frac{1}{n-1} \sum_{i=1}^{n-1} s_i \label{eq:VCT2} \\
-\sigma_i = \frac {\sqrt{\frac{1}{n-1} \sum_{i=1}^{n-1} \left( s_i - \bar{s} \right)^2} } {\bar{G_i}} \label{eq:VCT3}
+s_i = \frac{\text{SDR}_{i+1} - \text{SDR}_{i}}{t_{i+1} - t_i}, \forall i \in \left \{ 1, 2, \ldots, n-1 \right \},\;\;
+\bar{s} = \frac{1}{n-1} \sum_{i=1}^{n-1} s_i,\;\;
+\sigma_i = \frac {\sqrt{\frac{1}{n-1} \sum_{i=1}^{n-1} \left( s_i - \bar{s} \right)^2} } {\bar{G_i}} \label{eq:VCT3b}
 \end{equation}
@@
 Using the mean daily SDR we produce the trends of each data set, as departure from the seasonal value.
-Only for All-sky (Fig.~\ref{fig:trendALL}) <!--and Cloudy-sky (Fig.
-\ref{fig:trendCLOUD})--> conditions, the statistical significance of the trends are acceptable (Tab.~\ref{tab:trendtable}), to draw some conclusions.
+Only for All-sky (Fig.
+\ref{fig:trendALL}) <!--and Cloudy-sky (Fig.
+\ref{fig:trendCLOUD})--> conditions, the statistical significance of the trends are acceptable (Tab.
+\ref{tab:trendtable}), to draw some conclusions.
 

### 2023-07-03 Article.Rmd
Commit 2023-07-03 10:17

@@
 # Introduction.
+
+<!-- something about aerosols: bais mail -->
 

### 2023-07-03 Article.Rmd
Commit 2023-07-03 11:17

@@
 <!-- something about aerosols: bais mail -->
+<!-- something about clouds: bais mail -->
+<!--
+Η εισαγωγή είναι εντάξει. Ίσως θα μπορούσες να βάλεις στη 2η
+παράγραφο ακόμη 2-3 σειρές για τις τάσεις του αεροζόλ και των νεφών
+από τις εργασίες και να καταλήξεις με κάποια στοιχεία για το αεροζόλ
+στη Θεσσαλονίκη από εργασίες δικές μας. Για τα νέφη δεν έχω υπόψη μου
+καμία εργασία αλλά ίσως θα μπορούσε να ειπωθεί κάτι χρησιμοποιώντας
+παγκόσμιες εκτιμήσεις για την περιοχή της Μεσογείου.
+π.χ.https://link.springer.com/article/10.1007/s00382-022-06438-3
+-->
 

### 2023-07-03 Article.Rmd
Commit 2023-07-03 12:17

@@
 
-<!-- \FloatBarrier -->
+## Long-term SDR trends
 
-## Long-term trends
+We calculated the SDR trends, from the mean daily values of the departure from the seasonal values, for the whole time series for the three sky conditions (Table \ref{tab:trendtable}).
 
-Using the mean daily SDR we produce the trends of each data set, as departure from the seasonal value.
-Only for All-sky (Fig.
-\ref{fig:trendALL}) <!--and Cloudy-sky (Fig.
-\ref{fig:trendCLOUD})--> conditions, the statistical significance of the trends are acceptable (Tab.
-\ref{tab:trendtable}), to draw some conclusions.
-
-```{r trendALL, echo=F, fig.cap="Anomaly (\\%) of the daily SDR relative to climatological values for 1993 - 2023. The black line shows the long term trend for all-sky conditions."}
+```{r trendALL, echo=F, out.width='.70\\linewidth', fig.cap="Anomaly (\\%) of the daily SDR relative to climatological values for 1993 - 2023  `r year(min(ALL_1_daily_DESEAS$Date))` to `r year(max(ALL_1_daily_DESEAS$Date))`.
+The black line shows the long term trend for all-sky conditions."}
 knitr::include_graphics("./images/LongtermTrends-2.png")
@@
 
-<!-- \footnotesize -->
+\footnotesize
 ```{r trendtable, echo=F}
@@
 names(LT)    <- c("Trend [%/year]", "Trend p-value", "Sky condition", "Trend statistical signif. [%]")
+## no need all are 100%
+LT$`Trend statistical signif. [%]` <- NULL
 
@@
 pander(LT,
-       caption = "\\label{tab:trendtable}Trends of daily means by sky conditions.")
-# \\label{tab:tab2}
+       caption = "\\label{tab:trendtable}Trends of daily SDR means, for different sky conditions. All cases have a statistical significance above $99.9\\%$")
 # knitr::kable(LT, caption = "Trends of daily means by sky conditions.")
 
-## some vars to use
+## some vars to use later
 alltrend   <- signif(LT[`Sky condition` == "All sky cond.",   `Trend [%/year]`], 2)
@@
 ```
-<!-- \normalsize -->
-
-<!-- \FloatBarrier -->
+\normalsize
 

### 2023-07-03 Article.tex
Commit 2023-07-03 12:17

@@
   \title{Trends of SDR in Thessaloniki}
-    \author[Laboratory of Atmospheric Physics]{Natsis Athanasios%
+    \author[Laboratory of Atmospheric Physics]{Athanasios Natsis%
   \corref{cor1}%
@@
 
-\hypertarget{data-and-methodology.}{%
-\section{Data and methodology.}\label{data-and-methodology.}}
+\hypertarget{observational-data-and-methodology.}{%
+\section{Observational data and
+methodology.}\label{observational-data-and-methodology.}}
 
-Location
+The SDR is equivalent to the whole sky Global radiation (or Global
+Horizontal Irradiance, GHI), and was obtained with a horizontal leveled,
+CM-21 pyranometer, from 1993-04-12 to 2023-03-31. Also, we have
+measurements of direct beam radiation (DNI) obtained by a collocated
+CHP-1 pyrheliometer, from 2016-04-01 to 2023-03-31. Although, we
+performed a similar analysis to the DNI as for the SDR, the results are
+not presented here, as they lack the appropriate statistical
+significance, due to the sorter duration of the data. Moreover, the DNI
+was used as auxiliary data, with the clear sky identification algorithm
+(CSid), to aid with the selection of the appropriate thresholds. We have
+to note also, that despite the capability of the CSid to use the DNI as
+an characterization parameter, we haven't use it, to avoid any selection
+bias due to unequal availability of measurements.
 
-The SDR data used in this study cover the period from January 1993 to
-March 2011, and were measured with a Kipp \& Zonen CM-21 pyranometer
-operating contin- uously at the Laboratory of Atmospheric Physics of the
-Aristotle University of Thessaloniki (\$40\^{}\circ\$38''N,
-\$22\^{}\circ\$57''E, ∼80 m above sea level). The monitoring site is
-located near the city centre. The horizon is clear of obstruc- ◦ ◦ tions
-except for the azimuth angles between 35 and 120where obstacles of up to
-∼10◦ elevation block the direct irradiance in the morning for SZA larger
-than ∼80◦ . These data points were excluded from the analysis, as
-
-Instruments and callibrations
-
-relation to aerosols
-
-Beijing (40 N, 116 E), which are representative of regions with high
-aerosol load. Northern Greece has been selected because it is well known
-to be located at the crossroads of aerosols and various pollutants
-pathways originating in central and Eastern Europe (Lelieveld et al.,
-2002; Zerefos et al., 2002). Thessa- loniki, the major city in northern
-Greece, lies along the exit of air pollutants from Europe, in the path
-of the prevailing long-range transport (LRT) trajectories, as we move
-from central and east- ern Europe to the Mediterranean (Zerefos et. al.,
-2002). On the
-
-The SDR is equivalent to the whole sky Global radiation, also refered as
-Global Horizontal Irradiance (GHI), and was obtained with a horizontal
-leveled, CM-21 pyranometer. SDR data spans the period of 1993-04-12 to
-2023-03-31. In some algorithms we used as auxiliary data, the direct
-beam radiation (DNI) for comparisons, obtained by a CHP-1 pyrheliometer,
-with data availability from 2016-04-01 to 2023-03-31.
-
-\ldots{}
-
-\ldots{}
-
-There are three distinct steps to the creation of this dataset: a) the
-acquisition of radiation measurements from the sensors, b) a radiation
-data quality check, and c) the identification of ``clear sky''
-conditions from the radiometric data d) data aggregation and trend
-analysis.
+The SDR data were measured with a Kipp \& Zonen CM-21 pyranometer
+operating continuously at the Laboratory of Atmospheric Physics of the
+Aristotle University of Thessaloniki (\(40^\circ38'\)N,
+\(22^\circ57'\)E, 80m a.s.l.). The monitoring site is located near the
+city centre. During the study period the CM-21 pyranometer has been
+independently calibrated four times. There are three distinct steps to
+the creation of this dataset: a) the acquisition of radiation
+measurements from the sensors, b) a radiation data quality check, and c)
+the identification of ``clear sky'' conditions from the radiometric data
+d) data aggregation and trend analysis.
 
 For the acquisition of radiometric data, the signal of the broadband
-instruments is sampled with a rate of \(1 \text{Hz}\). The mean and
-standard deviation values are recorded for every minute. The
-measurements are corrected for the zero offset of the instrument signal.
-As reference of the ``dark signal'' is used the Sun elevation angle
-bellow \(-10^\circ\), for a period of \(3 \text{h}\). The signal is
-converted to radiation flux, using a ramped value of the instrument
-sensitivity, derived from the eight laboratory calibrations of the
-instrument, during the study period.
+instruments is sampled with a rate of \(1 \text{Hz}\). The mean and its
+standard deviation are recorded for every minute. The measurements are
+corrected for the zero offset of the instrument signal. As reference of
+the ``dark signal'' is used the Sun elevation angle bellow
+\(-10^\circ\), for a period of \(3 \text{h}\). The signal is converted
+to radiation flux, using a ramped value of the instrument sensitivity,
+derived from the eight laboratory calibrations of the instrument, during
+the study period.
 
-A manual screening is performed, to remove inconsistencies and erroneous
-recordings, that can occur randomly or systematically, during the long
-continuous operation of the instrument. The manual screening is aided by
-a radiation data quality assurance procedure, adjusted for this site,
-based on methods of Long and Shi~\citetext{\citealp[
-]{Long2008a}; \citealp{Long2006}}. Thus, problematic recordings have
-been excluded from further processing. Although it is not possible to
-detect all the bad data, the large number of data and the aggregation
-scheme we use, can provide us with accurate radiometric measurements,
-for the scope of this study.
+A manual screening was performed, to remove inconsistencies and
+erroneous recordings, that can occur randomly or systematically, during
+the long continuous operation of the instrument. The manual screening is
+aided by a radiation data quality assurance procedure, adjusted for this
+site, based on methods of Long and
+Shi~\citetext{\citeyear{Long2008a}; \citeyear{Long2006}}. Thus,
+problematic recordings have been excluded from further processing.
+Although it is not possible to detect all the bad data, the large number
+of data and the aggregation scheme we use, can provide us with accurate
+radiometric measurements, for the scope of this study.
 
-\hypertarget{clear-sky-identification}{%
-\subsection{Clear sky identification}\label{clear-sky-identification}}
+In order to estimate the effect of the clouds on the SDR, we created
+three datasets, by characterizing each one-minute measurement with a
+corresponding sky condition. The all-sky conditions, containing all the
+valid measurements. The clear-sky conditions data, correspond to data
+where we have inferred that the sky was almost clear of clouds, and the
+remainder part the data was characterized with cloud sky conditions
+(cloud-sky). To identify the clear-sky conditions we used the method
+proposed by \citet{Long2000} and by \citet{Reno2016}, that was adapted
+and configured for the site.
+
+\ldots.
+
+cite iterative process
 
@@
 
-Our method to characterize the sky conditions and the definition of each
-sky condition has some subjectivity. The algorithm was calibrated with
-the main focus to identify the presence of clouds on the sky dome,
-although there are marginal cases that there will be false positives or
-false negatives identifications.
+We have to note, that the definition of what constitutes as clear or
+cloudy sky, has some subjectivity, due to the used method of
+characterization. As a result, the details of the definition are site
+specific, it relies on a combination of threshold of comparisons with
+ideal actinometric models and statistics on different signal behavior.
+
+The algorithm was calibrated with the main focus to identify the
+presence of clouds on the sky dome, although there are marginal cases
+that there will be false positives or false negatives identifications.
 
@@
 
-\hypertarget{mean-value-of-irradiance-during-the-time-period.}{%
-\subsubsection{Mean value of irradiance during the time
-period.}\label{mean-value-of-irradiance-during-the-time-period.}}
-
 The mean of the measured value \(\overline{G}_i\) must be inside an
@@
 
-\hypertarget{max-value-of-irradiance-during-the-time-period.}{%
-\subsubsection{Max value of irradiance during the time
-period.}\label{max-value-of-irradiance-during-the-time-period.}}
-
 The running max measured value \(M_{Gi} = max[\text{SDR}_{i}]\), is
@@
 \end{equation}
-
-\hypertarget{variability-in-irradiance-by-line-length.}{%
-\subsubsection{Variability in irradiance by line
-length.}\label{variability-in-irradiance-by-line-length.}}
 
@@
 
-\hypertarget{variance-of-changes-in-the-time-series.}{%
-\subsubsection{Variance of Changes in the Time
-series.}\label{variance-of-changes-in-the-time-series.}}
-
 We calculate the standard deviation \(\sigma\) of the slope (\(s\))
@@
 
-\begin{gather}
-s_i = \frac{\text{SDR}_{i+1} - \text{SDR}_{i}}{t_{i+1} - t_i}, \forall i \in \left \{ 1, 2, \ldots, n-1 \right \} \label{eq:VCT1} \\
-\bar{s} = \frac{1}{n-1} \sum_{i=1}^{n-1} s_i \label{eq:VCT2} \\
-\sigma_i = \frac {\sqrt{\frac{1}{n-1} \sum_{i=1}^{n-1} \left( s_i - \bar{s} \right)^2} } {\bar{G_i}} \label{eq:VCT3}
-\end{gather}
+\begin{equation}
+s_i = \frac{\text{SDR}_{i+1} - \text{SDR}_{i}}{t_{i+1} - t_i}, \forall i \in \left \{ 1, 2, \ldots, n-1 \right \},\;\;
+\bar{s} = \frac{1}{n-1} \sum_{i=1}^{n-1} s_i,\;\;
+\sigma_i = \frac {\sqrt{\frac{1}{n-1} \sum_{i=1}^{n-1} \left( s_i - \bar{s} \right)^2} } {\bar{G_i}} \label{eq:VCT3b}
+\end{equation}
 
@@
 \end{equation}
-
-\hypertarget{variability-in-the-shape-of-the-irradiance-measurements.}{%
-\subsubsection{Variability in the Shape of the irradiance
-Measurements.}\label{variability-in-the-shape-of-the-irradiance-measurements.}}
 
@@
 \end{equation}
-
-\ldots.
-
-In order to estimate the effect of the clouds on the SDR we created
-three datasets, by characterizing each one-minute measurement with a
-corresponding sky condition. The all-sky conditions, containing all the
-valid measurements. The clear-sky conditions data, where we have
-inferred that the sky was almost clear of clouds, and the remainder part
-the data as cloudy sky conditions data (cloud-sky). To identify the
-clear-sky conditions we used the method proposed by \citet{Long2000} and
-by \citet{Reno2016}, that was adapted and configured for the site. We
-have to note, that the definition of what constitutes as clear or cloudy
-sky, has some subjectivity, due to the used method of characterization.
-As a result, the details of the definition are site specific, it relies
-on a combination of threshold of comparisons with ideal actinometric
-models and statistics on different signal behavior.
-
-\hypertarget{data-and-data-selection.}{%
-\subsection{Data and data selection.}\label{data-and-data-selection.}}
 
@@
 
-\hypertarget{long-term-trends}{%
-\subsection{Long-term trends}\label{long-term-trends}}
+\hypertarget{long-term-sdr-trends}{%
+\subsection{Long-term SDR trends}\label{long-term-sdr-trends}}
 
-Using the mean daily SDR we produce the trends of each data set, as
-departure from the seasonal value. Only for All-sky (Fig.
-\ref{fig:trendALL}) conditions, the statistical significance of the
-trends are acceptable (Tab. \ref{tab:trendtable}), to draw some
-conclusions.
+We calculated the SDR trends, from the mean daily values of the
+departure from the seasonal values, for the whole time series for the
+three sky conditions (Table \ref{tab:trendtable}).
 
@@
 
-{\centering \includegraphics[width=0.7\linewidth]{./images/LongtermTrends-2}
+{\centering \includegraphics[width=.70\linewidth]{./images/LongtermTrends-2}
 
@@
 
-\caption{Anomaly (\%) of the daily SDR relative to climatological values for 1993 - 2023. The black line shows the long term trend for all-sky conditions.}(\#fig:trendALL)
+\caption{Anomaly (\%) of the daily SDR relative to climatological values for 1993 - 2023  `r year(min(ALL_1_daily_DESEAS$Date))` to `r year(max(ALL_1_daily_DESEAS$Date))`.
+The black line shows the long term trend for all-sky conditions.}\label{fig:trendALL}
 \end{figure}
 
+\footnotesize
+
 \begin{longtable}[]{@{}
-  >{\centering\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2048}}
-  >{\centering\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1928}}
-  >{\centering\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2169}}
-  >{\centering\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.3855}}@{}}
-\caption{\label{tab:trendtable}Trends of daily means by sky conditions.
-}\tabularnewline
+  >{\centering\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.2361}}
+  >{\centering\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.2222}}
+  >{\centering\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.2500}}@{}}
+\caption{\label{tab:trendtable}Trends of daily SDR means, for different
+sky conditions. All cases have a statistical significance above
+\(99.9\%\) \normalsize}\tabularnewline
 \toprule\noalign{}
@@
 Sky condition
-\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
-Trend statistical signif. {[}\%{]}
 \end{minipage} \\
@@
 \endlastfoot
-0.3996 & 8.665e-23 & All sky cond. & 100 \\
-0.3773 & 7.075e-18 & Clear sky cond. & 100 \\
-0.3594 & 3.318e-09 & Cloudy cond. & 100 \\
+0.3996 & 8.665e-23 & All sky cond. \\
+0.3773 & 7.075e-18 & Clear sky cond. \\
+0.3594 & 3.318e-09 & Cloudy cond. \\
 \end{longtable}
@@
 
-{\centering \subfloat[Trend distribution for all-sky conditions.(\#fig:szatrends-1)]{\includegraphics[width=.35\linewidth]{./images/SzaTrends-13} }\subfloat[Trend distribution for cloud-sky conditions.(\#fig:szatrends-2)]{\includegraphics[width=.35\linewidth]{./images/SzaTrends-21} }
+{\centering \subfloat[Trend distribution for all-sky conditions.\label{fig:szatrends-1}]{\includegraphics[width=.35\linewidth]{./images/SzaTrends-13} }\subfloat[Trend distribution for cloud-sky conditions.\label{fig:szatrends-2}]{\includegraphics[width=.35\linewidth]{./images/SzaTrends-21} }
 
@@
 
-\caption{Distribution of the SDR's long term trends by SZA.}(\#fig:szatrends)
+\caption{Distribution of the SDR's long term trends by SZA.}\label{fig:szatrends}
 \end{figure}
@@
 
-\caption{Trends by season for all condition. Displaying monthly means of daily means.}(\#fig:seasonalALL)
+\caption{Trends by season for all condition. Displaying monthly means of daily means.}\label{fig:seasonalALL}
 \end{figure}
@@
 
-{\centering \subfloat[one plot(\#fig:cumsum-1)]{\includegraphics[width=.35\linewidth]{./images/CumulativeDailyCumSum-1} }\subfloat[the other one(\#fig:cumsum-2)]{\includegraphics[width=.35\linewidth]{./images/CumulativeDailyCumSum-9} }
+{\centering \subfloat[one plot\label{fig:cumsum-1}]{\includegraphics[width=.35\linewidth]{./images/CumulativeDailyCumSum-1} }\subfloat[the other one\label{fig:cumsum-2}]{\includegraphics[width=.35\linewidth]{./images/CumulativeDailyCumSum-9} }
 
@@
 
-\caption{two plots}(\#fig:cumsum)
+\caption{two plots}\label{fig:cumsum}
 \end{figure}

### 2023-07-03 Article_A.tex
Commit 2023-07-03 12:17

@@
 \author{true \and true}
-\date{2023-06-30}
+\date{2023-07-03}
 
@@
 
-Cloud ``shrinking'' and ``optical thinning'' in the ``dimming'' period and a subsequent recovery in the ``brightening'' period over China
-
-aerosols not only strengthen but also weaken the
-growth of clouds with respect to their coverage and
-optical thickness, depending on the levels of pollution
-and the associated amounts of aerosols as postulated
-in a conceptual framework (Wild 2009a, 2012, Yang
-et al 2012).
-
 \hypertarget{introduction.}{%
@@
 
-The shortwave downward solar irradiance (SDR) at the Earth's surface plays a significant role, on Earths climate.
-Changes of the SDR can be related to major changes on the Earth's energy budget, the mechanisms of climate change, and water and carbon cycle (Wild 2009).
-Studies of SDR variability, have identified some distinct SDR trends on different regions of the world and for different time periods, using the term `brightening' for positive trends, and `dimming' for negative trends.
-There are many cases on the long term records, showing a change of SDR trends direction, occurring roughly at the last decades of the 20th century.
-A general dimming before 2000 and a brightening afterwards, was found on multiple station in China (Yang et al. 2021).
-And a similar pattern was identified, with the breaking point around 1980, from a 71 years record of a Central Europe station (Wild et al. 2021) and for Brazil (Yamasoe et al. 2021).
-Another approach, using AI aided spatial analysis on continent level, of numerous observational stations data, reach similar conclusions for the above regions and the global trend (Yuan, Leirvik, and Wild 2021).
+The shortwave downward solar irradiance (SDR) at Earth's surface play a significant role, on it's climate.
+Changes of the SDR can be related to changes on Earth's energy budget, the mechanisms of climate change, and water and carbon cycle (Wild 2009).
+Can also affect, solar and agricultural production, and living organisms.
+Studies of SDR variability, have identified some distinct SDR trends on different regions of the world and for different time periods.
+The term `brightening' is generally used to describe periods of positive SDR trend, and `dimming' for negative trend.
+There are many cases on the long term records of irradiance, showing a systematic change of SDR's trend slope, occurring roughly at the last decades of the 20th century.
+On multiple station in China, a dimming period was reported until about 2000, followed by a brightening period (Yang et al. 2021).
+A similar pattern was identified, with the breaking point around 1980, for stations in Central Europe (Wild et al. 2021) and Brazil (Yamasoe et al. 2021).
+Also, on global scale, an AI aided continental level spatial analysis, with data from multiple station, reach similar conclusions for the above regions and for the global trend (Yuan, Leirvik, and Wild 2021).
 
-There is a consensus, among researchers, that the major factors of the attenuation of SDR is the interaction with the atmospheric aerosols and the clouds.
-Those interactions, among other factors, have been analysed with modeling (Li et al. 2016; Samset et al. 2018), showing the existence of feedback mechanisms between the two.
-And similar finds, have been showed in observational data (Schwarz et al. 2020; Ohvril et al. 2009; Zerefos et al. 2009; Xia et al. 2007 and references therein).
+There is a consensus, among researchers, that major factors of SDR attenuation is the interaction with atmospheric aerosols and clouds.
+Those interactions, among other factors, have been analysed with models (Li et al. 2016; Samset et al. 2018), showing the existence of feedback mechanisms between the two.
+Similar finds, have been showed in observational data (Schwarz et al. 2020; Ohvril et al. 2009; Zerefos et al. 2009; Xia et al. 2007 and references therein).
 
-Wild (2012) : \textasciitilde{}
+Due to the variability of the phenomenon, and its contributing factors, there is a need to further investigate and monitor SDR, in different sites, to estimate its magnitude, and its relation to the local conditions.
+In this study, we examine the trends of SDR, with ground-based measurements at Thessaloniki, Greece for the period 1993 to 2023, as derived from a CM-21 pyranometer.
+We reevaluated and extended the dataset used by Bais et al.
+(2013), applying a different algorithm for the identification of clear-/cloud-sky instances (Reno and Hansen 2016; Reno, Hansen, and Stein 2012a), and we derive the radiation trends for the period 1993 to 2023, under different sky conditions (all-sky, clear-sky and cloud-sky).
 
-Due to the variability of those factors and the interaction between them,
+\hypertarget{observational-data-and-methodology.}{%
+\section{Observational data and methodology.}\label{observational-data-and-methodology.}}
 
-\ldots\ldots{} cloud formation occurrences and constitution nuclei. cite\ldots..
+The SDR is equivalent to the whole sky Global radiation (or Global Horizontal Irradiance, GHI), and was obtained with a horizontal leveled, CM-21 pyranometer, from
+1993-04-12
+to
+2023-03-31.
+Also, we have measurements of direct beam radiation (DNI) obtained by a collocated CHP-1 pyrheliometer, from
+2016-04-01
+to
+2023-03-31.
+Although, we performed a similar analysis to the DNI as for the SDR, the results are not presented here, as they lack the appropriate statistical significance, due to the sorter duration of the data.
+Moreover, the DNI was used as auxiliary data, with the clear sky identification algorithm (CSid), to aid with the selection of the appropriate thresholds.
+We have to note also, that despite the capability of the CSid to use the DNI as an characterization parameter, we haven't use it, to avoid any selection bias due to unequal availability of measurements.
 
-it is difficult to estimate the SDR trends on each site.
-
-There is a also a need to evaluate the trends on different location
-
-In this study, we examine the trends of SDR with ground-based measurements at Thessaloniki for the period 1993 - 2023, as derived from a CM-21 pyranometer.
-We reevaluate and extend the dataset used by Bais et al.
-(2013), applying a different algorithm for the identification of clear-/cloud-sky instances, and we derive the radiation trends for this period under different sky conditions (all-sky, clear-sky and cloud-sky).
-
-\ldots{}
-
-Site location description and aerosols effect
-
-\ldots{}
-
-Our method to characterize the sky conditions and the definition of each sky condition has some subjectivity.
-The algorithm was calibrated with the main focus to identify the presence of clouds on the sky dome, although there are marginal cases that there will be false positives or false negatives identifications.
-
-\hypertarget{data-and-methodology.}{%
-\section{Data and methodology.}\label{data-and-methodology.}}
-
-The SDR is equivalent to the whole sky Global radiation, also refered as Global Horizontal Irradiance (GHI), and was obtained with a horizontal leveled, CM-21 pyranometer.
-SDR data spans the period of
-1993-04-12 to 2023-03-31.
-In some algorithms we used as auxiliary data, the direct beam radiation (DNI) for comparisons, obtained by a CHP-1 pyrheliometer, with data availability from
-2016-04-01 to 2023-03-31.
-
-\ldots{}
-
-\ldots{}
-
+The SDR data were measured with a Kipp \& Zonen CM-21 pyranometer operating continuously at the Laboratory of Atmospheric Physics of the Aristotle University of Thessaloniki (\(40^\circ38'\)N, \(22^\circ57'\)E, 80m a.s.l.).
+The monitoring site is located near the city centre.
+During the study period the CM-21 pyranometer has been independently calibrated four times.
 There are three distinct steps to the creation of this dataset:
@@
 For the acquisition of radiometric data,
-the signal of the broadband instruments is sampled with a rate of \(1 \text{Hz}\). The mean and standard deviation values are recorded for every minute.
+the signal of the broadband instruments is sampled with a rate of \(1 \text{Hz}\). The mean and its standard deviation are recorded for every minute.
 The measurements are corrected for the zero offset of the instrument signal.
@@
 
-A manual screening is performed, to remove inconsistencies and erroneous recordings, that can occur randomly or systematically, during the long continuous operation of the instrument.
-The manual screening is aided by a radiation data quality assurance procedure, adjusted for this site, based on methods of Long and Shi~(C. Long and Shi 2008 ; C. Long and Shi 2006).
+A manual screening was performed, to remove inconsistencies and erroneous recordings, that can occur randomly or systematically, during the long continuous operation of the instrument.
+The manual screening is aided by a radiation data quality assurance procedure, adjusted for this site, based on methods of Long and Shi~(2008, 2006).
 Thus, problematic recordings have been excluded from further processing.
@@
 
-\hypertarget{clear-sky-identification}{%
-\subsection{Clear sky identification}\label{clear-sky-identification}}
+In order to estimate the effect of the clouds on the SDR, we created three datasets, by characterizing each one-minute measurement with a corresponding sky condition.
+The all-sky conditions, containing all the valid measurements.
+The clear-sky conditions data, correspond to data where we have inferred that the sky was almost clear of clouds, and the remainder part the data was characterized with cloud sky conditions (cloud-sky).
+To identify the clear-sky conditions we used the method proposed by C. N. Long and Ackerman (2000) and by Reno and Hansen (2016), that was adapted and configured for the site.
+
+\ldots.
+
+cite iterative process
 
 \ldots\ldots{}
+
+We have to note, that the definition of what constitutes as clear or cloudy sky, has some subjectivity, due to the used method of characterization.
+As a result, the details of the definition are site specific, it relies on a combination of threshold of comparisons with ideal actinometric models and statistics on different signal behavior.
+
+The algorithm was calibrated with the main focus to identify the presence of clouds on the sky dome, although there are marginal cases that there will be false positives or false negatives identifications.
 
@@
 We found, that Haurwitz's model, adjusted with a factor of 0.965 has the lower root mean squared error (RMSE).
-The tried models are described by Reno, Hansen, and Stein (2012) and tested by Reno and Hansen (2016).
+The tried models are described by Reno, Hansen, and Stein (2012b) and tested by Reno and Hansen (2016).
 The iterative optimization method, for the selection of the reference is discussed by C. N. Long and Ackerman (2000) and Reno and Hansen (2016).
@@
 
-\hypertarget{mean-value-of-irradiance-during-the-time-period.}{%
-\subsubsection{Mean value of irradiance during the time period.}\label{mean-value-of-irradiance-during-the-time-period.}}
-
 The mean of the measured value \(\overline{G}_i\) must be inside an envelope based on the reference model \(\text{SDR}_\text{Clear Sky}\) (Eq. \ref{eq:MeanVIP}).
@@
 
-\hypertarget{max-value-of-irradiance-during-the-time-period.}{%
-\subsubsection{Max value of irradiance during the time period.}\label{max-value-of-irradiance-during-the-time-period.}}
-
 The running max measured value \(M_{Gi} = max[\text{SDR}_{i}]\), is compared to a similar constructed value from the reference \(M_{CSi} = max[\text{SDR}_{i\text{Clear Sky}}]\) (Eq. \ref{eq:MaxVIP}).
@@
 \end{equation}
-
-\hypertarget{variability-in-irradiance-by-line-length.}{%
-\subsubsection{Variability in irradiance by line length.}\label{variability-in-irradiance-by-line-length.}}
 
@@
 
-\hypertarget{variance-of-changes-in-the-time-series.}{%
-\subsubsection{Variance of Changes in the Time series.}\label{variance-of-changes-in-the-time-series.}}
-
 We calculate the standard deviation \(\sigma\) of the slope
@@
 \end{equation}
-
-\hypertarget{variability-in-the-shape-of-the-irradiance-measurements.}{%
-\subsubsection{Variability in the Shape of the irradiance Measurements.}\label{variability-in-the-shape-of-the-irradiance-measurements.}}
 
@@
 
-\ldots.
-
-In order to estimate the effect of the clouds on the SDR we created three datasets, by characterizing each one-minute measurement with a corresponding sky condition.
-The all-sky conditions, containing all the valid measurements.
-The clear-sky conditions data, where we have inferred that the sky was almost clear of clouds, and the remainder part the data as cloudy sky conditions data (cloud-sky).
-To identify the clear-sky conditions we used the method proposed by C. N. Long and Ackerman (2000) and by Reno and Hansen (2016), that was adapted and configured for the site.
-We have to note, that the definition of what constitutes as clear or cloudy sky, has some subjectivity, due to the used method of characterization.
-As a result, the details of the definition are site specific, it relies on a combination of threshold of comparisons with ideal actinometric models and statistics on different signal behavior.
-
-\hypertarget{data-and-data-selection.}{%
-\subsection{Data and data selection.}\label{data-and-data-selection.}}
-
 Due to a significant measurement uncertainty near, the horizon, we have to exclude
@@
 Subsequently, we made all measurements relative to the Total Solar Irradiance (TSI) at \(1 \text{au}\), in order to compensate for the Sun's intensity variability, using a homogenized time series of satellite TSI observations.
-The TSI data we use, comes from NOAA (Coddington et al. 2005).
+The TSI data we use, comes from NOAA (Coddington et al. 2005), it overlaps with our data, and is interpolated to match with the time step of the data.
 As a result, we can present all radiation data as a fraction of Sun's TSI.
@@
 
-\hypertarget{long-term-trends}{%
-\subsection{Long-term trends}\label{long-term-trends}}
+\hypertarget{long-term-sdr-trends}{%
+\subsection{Long-term SDR trends}\label{long-term-sdr-trends}}
 
-Using the mean daily SDR we produce the trends of each data set, as departure from the seasonal value.
-Only for All-sky (Fig. \ref{fig:trendALL}) conditions, the statistical significance of the trends are acceptable (Tab. \ref{tab:trendtable}), to draw some conclusions.
+We calculated the SDR trends, from the mean daily values of the departure from the seasonal values, for the whole time series for the three sky conditions (Table \ref{tab:trendtable}).
 
@@
 
-\caption{Anomaly (\%) of the daily SDR relative to climatological values for 1993 - 2023. The black line shows the long term trend for all-sky conditions.}\label{fig:trendALL}
+\caption{Anomaly (\%) of the daily SDR relative to climatological values for 1993 - 2023  `r year(min(ALL_1_daily_DESEAS$Date))` to `r year(max(ALL_1_daily_DESEAS$Date))`.
+The black line shows the long term trend for all-sky conditions.}(\#fig:trendALL)
 \end{figure}
 
+\footnotesize
+
 \begin{longtable}[]{@{}
-  >{\centering\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2048}}
-  >{\centering\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1928}}
-  >{\centering\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2169}}
-  >{\centering\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.3855}}@{}}
-\caption{\label{tab:trendtable}Trends of daily means by sky conditions.
-}\tabularnewline
+  >{\centering\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.2361}}
+  >{\centering\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.2222}}
+  >{\centering\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.2500}}@{}}
+\caption{\label{tab:trendtable}Trends of daily SDR means, for different sky conditions. All cases have a statistical significance above \(99.9\%\)
+\normalsize}\tabularnewline
 \toprule
@@
 \endhead
-0.3996 & 8.665e-23 & All sky cond. & 100 \\
-0.3773 & 7.075e-18 & Clear sky cond. & 100 \\
-0.3594 & 3.318e-09 & Cloudy cond. & 100 \\
+0.3996 & 8.665e-23 & All sky cond. \\
+0.3773 & 7.075e-18 & Clear sky cond. \\
+0.3594 & 3.318e-09 & Cloudy cond. \\
 \bottomrule
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-Reno2012a}{}}%
+Reno, Matthew J., Clifford W.
+Hansen, and Joshua S.
+Stein.
+2012a.
+{``{Global Horizontal Irradiance Clear Sky Models: Implementation and Analysis}.''} \emph{SANDIA REPORT Sand2012-2389 Unlimited Release Printed March 2012}, March, 1--66.
+
 \leavevmode\vadjust pre{\hypertarget{ref-Reno2012}{}}%
-Reno, Matthew J., Clifford W. Hansen, and Joshua S. Stein. 2012. {``Global Horizontal Irradiance Clear Sky Models: Implementation and Analysis.''} SAND2012-2389, 1039404. \url{https://doi.org/gq5npv}.
+---------. 2012b. {``Global Horizontal Irradiance Clear Sky Models: Implementation and Analysis.''} SAND2012-2389, 1039404. \url{https://doi.org/gq5npv}.
 
@@
 Wild, Martin. 2009. {``Global Dimming and Brightening: A Review.''} \emph{Journal of Geophysical Research Atmospheres} 114 (12): 1--31. \url{https://doi.org/bcq}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-Wild2012}{}}%
----------. 2012. {``Enlightening Global Dimming and Brightening.''} \emph{Bulletin of the American Meteorological Society} 93 (1): 27--37. \url{https://doi.org/b3w2h4}.
 

### 2023-07-03 Article.Rmd
Commit 2023-07-03 13:17

@@
 
-```{r trendALL, echo=F, out.width='.70\\linewidth', fig.cap="Anomaly (\\%) of the daily SDR relative to climatological values for 1993 - 2023  `r year(min(ALL_1_daily_DESEAS$Date))` to `r year(max(ALL_1_daily_DESEAS$Date))`.
-The black line shows the long term trend for all-sky conditions."}
+....
+
+more about trend in relation to other trends
+
+......
+
+```{r trendALL, echo=F, out.width='.70\\linewidth', fig.cap=paste0("Anomaly (\\%) of the daily SDR relative to climatological values for ", year(min(ALL_1_daily_DESEAS$Date)), " - ", year(max(ALL_1_daily_DESEAS$Date)), ".
+The black line shows the long term trend for all-sky conditions.")}
 knitr::include_graphics("./images/LongtermTrends-2.png")
@@
 
-<!-- \FloatBarrier -->
-
 ## Long term trends by season of year
@@
 
-<!-- \FloatBarrier -->
 <!-- 
@@
 -->
-
-<!-- \FloatBarrier -->
 

### 2023-07-04 Article.Rmd
Commit 2023-07-03 19:24
Commit 2023-07-04 10:30

@@
 load("./data/common_data_14_2.Rda")
-TSIinfo <- data.table(read.csv("./figures/tbl_tsi_info.dat",
+TSIinfo <- data.table(read.csv("~/MANUSCRIPTS/2022_sdr_trends/figures/tbl_tsi_info.dat",
                     strip.white = TRUE,
@@
 παγκόσμιες εκτιμήσεις για την περιοχή της Μεσογείου.
-π.χ.https://link.springer.com/article/10.1007/s00382-022-06438-3
+π.χ. https://link.springer.com/article/10.1007/s00382-022-06438-3
 -->

### 2023-07-04 Article.Rmd
Commit 2023-07-04 11:12

@@
     gather_citations(
-        sourcefile   = "./Article.Rmd",
-        mainreposit  = "./Mainreferences.bib",
-        localreposit = "./references.bib",
-        folderlink   = "./References_links",
+        sourcefile   = "~/MANUSCRIPTS/2022_sdr_trends/Article.Rmd",
+        mainreposit  = "~/MANUSCRIPTS/2022_sdr_trends/Mainreferences.bib",
+        localreposit = "~/MANUSCRIPTS/2022_sdr_trends/references.bib",
+        folderlink   = "~/MANUSCRIPTS/2022_sdr_trends/References_links",
         workingdir   = "~/MANUSCRIPTS/2022_sdr_trends/"

### 2023-07-04 Article.Rmd
Commit 2023-07-04 12:11
Commit 2023-07-04 13:11

@@
 
-```{r echo=F, include=T, warning=FALSE}
+```{r echo=F, include=T, warning=FALSE, message=FALSE}
+res <- rmarkdown::find_pandoc(dir = '/usr/lib/rstudio/resources/app/bin/quarto/bin/tools')
 ## document configurations 
@@
 ## load data
-load("./data/common_data_14_2.Rda")
+load("~/MANUSCRIPTS/2022_sdr_trends/data/common_data_14_2.Rda")
 TSIinfo <- data.table(read.csv("~/MANUSCRIPTS/2022_sdr_trends/figures/tbl_tsi_info.dat",
@@
 
-...... cloud formation occurrences and constitution nuclei. cite.....
+...... cloud formation occurrences and constitution nuclei. cite .....
 
@@
 
-The SDR data were measured with a Kipp & Zonen CM-21 pyranometer operating continuously at the Laboratory of Atmospheric Physics of the Aristotle University of Thessaloniki ($40^\circ38'$N, $22^\circ57'$E, 80m a.s.l.).
+The SDR data were measured with a Kipp & Zonen CM-21 pyranometer operating continuously at the Laboratory of Atmospheric Physics of the Aristotle University of Thessaloniki ($40^\circ38'$N, $22^\circ57'$E, 80m\ a.s.l.).
 The monitoring site is located near the city centre.
@@
 There are three distinct steps to the creation of this dataset:
-a) the acquisition of radiation measurements from the sensors, 
-b) a radiation data quality check, and 
-c) the identification of "clear sky" conditions from the radiometric data
-d) data aggregation and trend analysis.
+a)\ the acquisition of radiation measurements from the sensors, 
+b)\ a radiation data quality check,
+c)\ the identification of "clear sky" conditions from the radiometric data, and
+d)\ data aggregation and trend analysis.
 
@@
 
-As global radiation clear sky reference we are using the Haurwitz’s model, adjusted for our site with a factor of `r signif(alpha,3)` (Eq. \ref{eq:ahau}).
+As global radiation clear sky reference we are using the Haurwitz’s model, adjusted for our site with a factor of `r signif(alpha,3)` (Eq.\ \ref{eq:ahau}).
 The selection of a clear sky reference model, was based on SDR observation from the period 2016 – 2021.
@@
 
-The running max measured value $M_{Gi} = max[\text{SDR}_{i}]$, is compared to a similar constructed value from the reference $M_{CSi} = max[\text{SDR}_{i\text{Clear Sky}}]$ (Eq. \ref{eq:MaxVIP}).
+The running max measured value $M_{Gi} = max[\text{SDR}_{i}]$, is compared to a similar constructed value from the reference $M_{CSi} = max[\text{SDR}_{i\text{Clear Sky}}]$ (Eq.\ \ref{eq:MaxVIP}).
 
@@
 
-We calculated the SDR trends, from the mean daily values of the departure from the seasonal values, for the whole time series for the three sky conditions (Table \ref{tab:trendtable}).
+We calculated the SDR trends, from the mean daily values of the departure from the seasonal values, for the whole time series for the three sky conditions (Table\ \ref{tab:trendtable}).
 
@@
 
-Analyzing the long term trends of SDR to bins of SZA, we can see the contribution of the geometry and time in a diurnal level (Figures \ref{fig:szatrends}a and \ref{fig:szatrends}b).
+Analyzing the long term trends of SDR to bins of SZA, we can see the contribution of the geometry and time in a diurnal level (Figures\ \ref{fig:szatrends}a and \ref{fig:szatrends}b).
 Although there is a seasonal distribution of SZA than in not presented here.
@@
 
-Figure \ref{fig:cumsum}
+Figure\ \ref{fig:cumsum}
 
@@
 similar results with Evidence for Clear‐Sky Dimming and Brightening in Central Europe_Wild2021.pdf
-inclue in discusion
+include in discussion
 

### 2023-07-05 Article.Rmd
Commit 2023-07-05 19:09
Commit 2023-07-05 20:14

@@
 We calculated the SDR trends, from the mean daily values of the departure from the seasonal values, for the whole time series for the three sky conditions (Table\ \ref{tab:trendtable}).
+In this time period, we could not find any significant break or change in the variability of the trends.
+Other studies for the European region are reporting a change of the SDR slope, around the '80, few years before the start of our records [@Wild2021; @Yuan2021; @Ohmura2009].
 
@@
 pander(LT,
-       caption = "\\label{tab:trendtable}Trends of daily SDR means, for different sky conditions. All cases have a statistical significance above $99.9\\%$")
+       caption = "\\label{tab:trendtable}Trends of daily SDR means, for different sky conditions. All cases have a statistical significance above $99.9\\%$.")
 # knitr::kable(LT, caption = "Trends of daily means by sky conditions.")
@@
 
-```{r cumsum, echo = FALSE, fig.cap='two plots', fig.subcap=c('one plot', 'the other one'), out.width='.35\\linewidth', fig.asp=1, fig.ncol = 2}
-knitr::include_graphics(c("./images/CumulativeDailyCumSum-1.png", "./images/CumulativeDailyCumSum-9.png"))
+```{r cumsum, echo = FALSE, fig.cap='two plots', fig.subcap=c('one plot', 'the other one'), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3}
+knitr::include_graphics(c(
+    "./images/CumulativeDailyCumSum-1.png",    # ALL sky
+    "./images/CumulativeDailyCumSum-5.png",    # CLEAR sky
+    "./images/CumulativeDailyCumSum-9.png"))   # Cloud sky
 ```
@@
 -------------------------------------------------
+\newpage

### 2023-07-06 Article.Rmd
Commit 2023-07-06 22:14

@@
 ```{r echo=F, include=T, warning=FALSE, message=FALSE}
+suppressMessages(
     res <- rmarkdown::find_pandoc(dir = '/usr/lib/rstudio/resources/app/bin/quarto/bin/tools')
+)
 ## document configurations 
@@
 ## load data
-load("~/MANUSCRIPTS/2022_sdr_trends/data/common_data_14_2.Rda")
-TSIinfo <- data.table(read.csv("~/MANUSCRIPTS/2022_sdr_trends/figures/tbl_tsi_info.dat",
+load("./data/common_data_14_2.Rda")
+TSIinfo <- data.table(read.csv("./figures/tbl_tsi_info.dat",
                     strip.white = TRUE,
@@
 ## variables we use
-source("~/MANUSCRIPTS/2022_sdr_trends/DHI_GHI_0_variables.R")
+source("./DHI_GHI_0_variables.R")
 
@@
     gather_citations(
-        sourcefile   = "~/MANUSCRIPTS/2022_sdr_trends/Article.Rmd",
-        mainreposit  = "~/MANUSCRIPTS/2022_sdr_trends/Mainreferences.bib",
-        localreposit = "~/MANUSCRIPTS/2022_sdr_trends/references.bib",
-        folderlink   = "~/MANUSCRIPTS/2022_sdr_trends/References_links",
-        workingdir   = "~/MANUSCRIPTS/2022_sdr_trends/"
+        sourcefile   = "./Article.Rmd",
+        mainreposit  = "./Mainreferences.bib",
+        localreposit = "./references.bib",
+        folderlink   = "./References_links",
+        workingdir   = "./"
     )

### 2023-07-06 DHI_GHI_0_data_input.R

@@ -9,7 +9,7 @@ require(data.table)
 require(zoo)
 source("~/CODE/FUNCTIONS/R/trig_deg.R")
 source("~/CODE/FUNCTIONS/R/data.R")
-source("~/MANUSCRIPTS/2022_sdr_trends/DHI_GHI_0_variables.R")
+source("./DHI_GHI_0_variables.R")
 
 ####  Run data construction ####################################################
 
@@ -50,6 +50,11 @@ if (D_13) {
     inpatern    <- "Clear_Sky_[0-9]{4}.Rds"
 }
 
+## create local folders
+dir.create(dirname(common_data), showWarnings = FALSE)
+dir.create("./figures",          showWarnings = FALSE)
+dir.create("./images",           showWarnings = FALSE)
+dir.create("./runtime",          showWarnings = FALSE)
 
 
 ## check if we need to run data production
@@ -69,10 +74,9 @@ if (havetorun) {
     input_files <- grep("_stats_", input_files, value = TRUE, invert = TRUE)
 
     if (TEST) {
-        input_files <- sample(input_files, 4)
+        warning("\nTEST MODE IS ACTIVE!!\n\n")
+        input_files <- sample(input_files, 3)
     }
-
-
 
     if (TEST | !file.exists(CS_file) | max(file.mtime(input_files)) > file.mtime(CS_file)) {
         cat(paste("Load data from Clear Sky proccess from original\n"))
@@ -122,7 +126,7 @@ if (havetorun) {
 
         ## this is used by old scripts
         setorder(DATA,Date)
-        myRtools::write_RDS(object = DATA, file = CS_file)
+        myRtools::write_RDS(object = DATA, file = CS_file, clean = TRUE)
     } else {
         DATA <- readRDS(CS_file)
     }
@@ -181,8 +185,9 @@ if (havetorun) {
     #'
     TSI_info <- DATA[, .(Start = min(Date),
                          End   = max(Date)), by = TSI_Source]
-    myRtools::write_dat(TSI_info,
-                        "~/MANUSCRIPTS/2022_sdr_trends/figures/tbl_tsi_info.dat")
+    myRtools::write_dat(object = TSI_info,
+                        file   = "./figures/tbl_tsi_info.dat",
+                        clean  = TRUE)
     rm(TSI_info)
 
     #' ## Data preparation
@@ -638,10 +643,6 @@ if (havetorun) {
[…]

### 2023-07-06 DHI_GHI_1_longterm_trends.R

@@ -58,11 +58,12 @@ warning("Don't use cache it breaks computations")
 #+ include=F, echo=F
 ## __ Set environment ----------------------------------------------------------
 Sys.setenv(TZ = "UTC")
-Script.Name <- "~/MANUSCRIPTS/2022_sdr_trends/DHI_GHI_1_longterm_trends.R"
+Script.Name <- "./DHI_GHI_1_longterm_trends.R"
+
 if (!interactive()) {
-    pdf( file = paste0("~/MANUSCRIPTS/2022_sdr_trends/runtime/",  basename(sub("\\.R$",".pdf", Script.Name))))
-    sink(file = paste0("~/MANUSCRIPTS/2022_sdr_trends/runtime/",  basename(sub("\\.R$",".out", Script.Name))), split = TRUE)
-    filelock::lock(paste0("~/MANUSCRIPTS/2022_sdr_trends/runtime/", basename(sub("\\.R$",".lock", Script.Name))), timeout = 0)
+    pdf( file = paste0("./runtime/",  basename(sub("\\.R$",".pdf", Script.Name))))
+    sink(file = paste0("./runtime/",  basename(sub("\\.R$",".out", Script.Name))), split = TRUE)
+    filelock::lock(paste0("./runtime/", basename(sub("\\.R$",".lock", Script.Name))), timeout = 0)
 }
 
 
@@ -86,8 +87,8 @@ source("~/CODE/FUNCTIONS/R/data.R")
 
 
 ## __ Source initial scripts ---------------------------------------------------
-source("~/MANUSCRIPTS/2022_sdr_trends/DHI_GHI_0_data_input.R")
-source("~/MANUSCRIPTS/2022_sdr_trends/DHI_GHI_0_variables.R")
+source("./DHI_GHI_0_data_input.R")
+source("./DHI_GHI_0_variables.R")
 source("~/CODE/R_myRtools/myRtools/R/write_.R")
 tic <- Sys.time()
 
@@ -445,7 +446,7 @@ pander(pprint,
 #+ echo=F, include=T
 
 write_dat(pprint,
-          "~/MANUSCRIPTS/2022_sdr_trends/figures/tbl_longterm_trends.dat",
+          "./figures/tbl_longterm_trends.dat",
           clean = TRUE)
 
 
@@ -768,7 +769,7 @@ pander(pprint,
 #+ echo=F, include=F
 
 write_dat(pprint,
[…]

### 2023-07-06 DHI_GHI_2_sza_trends.R

@@ -58,13 +58,12 @@ warning("Don't use cache it breaks computations")
 #+ include=F, echo=F
 ## __ Set environment ----------------------------------------------------------
 Sys.setenv(TZ = "UTC")
-Script.Name <- tryCatch({ funr::sys.script() },
-                        error = function(e) { cat(paste("\nUnresolved script name: ", e),"\n\n")
-                            return("Climatological_") })
+Script.Name <- "./DHI_GHI_2_sza_trends.R"
+
 if (!interactive()) {
-    pdf( file = paste0("~/MANUSCRIPTS/2022_sdr_trends/runtime/",  basename(sub("\\.R$",".pdf", Script.Name))))
-    sink(file = paste0("~/MANUSCRIPTS/2022_sdr_trends/runtime/",  basename(sub("\\.R$",".out", Script.Name))), split = TRUE)
-    filelock::lock(paste0("~/MANUSCRIPTS/2022_sdr_trends/runtime/", basename(sub("\\.R$",".lock", Script.Name))), timeout = 0)
+    pdf( file = paste0("./runtime/",  basename(sub("\\.R$",".pdf", Script.Name))))
+    sink(file = paste0("./runtime/",  basename(sub("\\.R$",".out", Script.Name))), split = TRUE)
+    filelock::lock(paste0("./runtime/", basename(sub("\\.R$",".lock", Script.Name))), timeout = 0)
 }
 
 
@@ -88,8 +87,8 @@ source("~/CODE/FUNCTIONS/R/data.R")
 
 
 ## __ Source initial scripts ---------------------------------------------------
-source("~/MANUSCRIPTS/2022_sdr_trends/DHI_GHI_0_data_input.R")
-source("~/MANUSCRIPTS/2022_sdr_trends/DHI_GHI_0_variables.R")
+source("./DHI_GHI_0_data_input.R")
+source("./DHI_GHI_0_variables.R")
 tic <- Sys.time()
 
 ## notification function

### 2023-07-08 Article.Rmd
Commit 2023-07-08 20:14

@@
 
+......
+
+Cumulative sum [@Regier2019]
+
+```{r cumsum, echo = FALSE, fig.cap='two plots', fig.subcap=c('one plot', 'the other one'), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3}
+knitr::include_graphics(c(
+    "./images/CumulativeDailyCumSum-1.png",    # ALL sky
+    "./images/CumulativeDailyCumSum-5.png",    # CLEAR sky
+    "./images/CumulativeDailyCumSum-9.png"))   # Cloud sky
+```
+
 ....
@@
 
-```{r cumsum, echo = FALSE, fig.cap='two plots', fig.subcap=c('one plot', 'the other one'), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3}
-knitr::include_graphics(c(
-    "./images/CumulativeDailyCumSum-1.png",    # ALL sky
-    "./images/CumulativeDailyCumSum-5.png",    # CLEAR sky
-    "./images/CumulativeDailyCumSum-9.png"))   # Cloud sky
-```
+<!-- ```{r cumsumSTD, echo = FALSE, fig.cap='two plots', fig.subcap=c('one plot', 'the other one'), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3} -->
+<!-- knitr::include_graphics(c( -->
+<!--     "./images/CumulativeDailyCumSumSTD-1.png",    # ALL sky -->
+<!--     "./images/CumulativeDailyCumSumSTD-2.png",    # CLEAR sky -->
+<!--     "./images/CumulativeDailyCumSumSTD-3.png"))   # Cloud sky -->
+<!-- ``` -->
 

### 2023-07-09 Article.Rmd
Commit 2023-07-08 21:10
Commit 2023-07-09 20:14

@@
 
-## create local citations file
+## create local citations file and links here
 source("~/CODE/FUNCTIONS/R/gather_citations.R")
@@
 We calculated the SDR trends, from the mean daily values of the departure from the seasonal values, for the whole time series for the three sky conditions (Table\ \ref{tab:trendtable}).
-In this time period, we could not find any significant break or change in the variability of the trends.
-Other studies for the European region are reporting a change of the SDR slope, around the '80, few years before the start of our records [@Wild2021; @Yuan2021; @Ohmura2009].
+We present only the plot for the trend under all-sky conditions (Figure\ \ref{fig:trendALL}), as the plots for clear-sky condition and cloud-sky condition, are very similar.
+In the studied time period, there is no significant break or change in the variability of the overall trends.
+Other studies for the European region are reporting a change of the SDR slope, around the 1980, few years before the start of our records [@Wild2021; @Yuan2021; @Ohmura2009].
 
-......
-
-Cumulative sum [@Regier2019]
-
-```{r cumsum, echo = FALSE, fig.cap='two plots', fig.subcap=c('one plot', 'the other one'), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3}
-knitr::include_graphics(c(
-    "./images/CumulativeDailyCumSum-1.png",    # ALL sky
-    "./images/CumulativeDailyCumSum-5.png",    # CLEAR sky
-    "./images/CumulativeDailyCumSum-9.png"))   # Cloud sky
-```
-
-....
-
-more about trend in relation to other trends
-
-......
-
-```{r trendALL, echo=F, out.width='.70\\linewidth', fig.cap=paste0("Anomaly (\\%) of the daily SDR relative to climatological values for ", year(min(ALL_1_daily_DESEAS$Date)), " - ", year(max(ALL_1_daily_DESEAS$Date)), ".
-The black line shows the long term trend for all-sky conditions.")}
-knitr::include_graphics("./images/LongtermTrends-2.png")
-```
-
-<!-- ```{r trendCLEAR, echo=F, fig.cap="Daily SDR seasonal anomaly and long term trend, for clear sky conditions."} -->
-<!-- knitr::include_graphics("./images/LongtermTrends-4.png") -->
-<!-- ``` -->
-
-<!-- ```{r trendCLOUD, echo=F, fig.cap="Anomaly (\\%) of the daily SDR relative to climatological values for 1993 - 2023. The black line shows the long term trend for cloud-sky conditions."} -->
-<!-- knitr::include_graphics("./images/LongtermTrends-6.png") -->
-<!-- ``` -->
-
-\footnotesize
+<!-- \footnotesize -->
 ```{r trendtable, echo=F}
@@
 pander(LT,
-       caption = "\\label{tab:trendtable}Trends of daily SDR means, for different sky conditions. All cases have a statistical significance above $99.9\\%$.")
+       caption = paste0("\\label{tab:trendtable}Trends of daily SDR means, for different sky conditions for ", year(min(ALL_1_daily_DESEAS$Date)), " - ", year(max(ALL_1_daily_DESEAS$Date)), ".
+For all cases the statistical significance is above $99.9\\%$."))
 # knitr::kable(LT, caption = "Trends of daily means by sky conditions.")
-
 ## some vars to use later
@@
 ```
-\normalsize
+<!-- \normalsize -->
+
+```{r trendALL, echo=F, out.width='.70\\linewidth', fig.cap=paste0("Anomaly (\\%) of the daily SDR relative to climatological values for ", year(min(ALL_1_daily_DESEAS$Date)), " - ", year(max(ALL_1_daily_DESEAS$Date)), ".
+The black line shows the long term trend for all-sky conditions.")}
+knitr::include_graphics("./images/LongtermTrends-2.png")
+```
+
+Although, using the cumulative sums plots we can reveal some structure within the trends and some difference between the sky conditions [@Regier2019].
+
+Comparining the conditions
+
+Variability issues
+
+```{r cumsum, echo = FALSE, fig.cap='Running plots of the cumulative sum of the daily SDR for three different sky conditions', fig.subcap=c('All-sky conditions', 'Clear-sky conditions', 'Cloud-sky conditions'), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3}
+knitr::include_graphics(c(
+    "./images/CumulativeDailyCumSum-1.png",    # ALL sky
+    "./images/CumulativeDailyCumSum-5.png",    # CLEAR sky
+    "./images/CumulativeDailyCumSum-9.png"))   # Cloud sky
+```
+
+<!-- ```{r cumsumSTD, echo = FALSE, fig.cap='two plots', fig.subcap=c('one plot', 'the other one'), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3} -->
+<!-- knitr::include_graphics(c( -->
+<!--     "./images/CumulativeDailyCumSumSTD-1.png",    # ALL sky -->
+<!--     "./images/CumulativeDailyCumSumSTD-2.png",    # CLEAR sky -->
+<!--     "./images/CumulativeDailyCumSumSTD-3.png"))   # Cloud sky -->
+<!-- ``` -->
+
+....
+
+more about trend in relation to other trends
+Aerosols and clouds
+
+......
+
+<!-- ```{r trendCLEAR, echo=F, fig.cap="Daily SDR seasonal anomaly and long term trend, for clear sky conditions."} -->
+<!-- knitr::include_graphics("./images/LongtermTrends-4.png") -->
+<!-- ``` -->
+
+<!-- ```{r trendCLOUD, echo=F, fig.cap="Anomaly (\\%) of the daily SDR relative to climatological values for 1993 - 2023. The black line shows the long term trend for cloud-sky conditions."} -->
+<!-- knitr::include_graphics("./images/LongtermTrends-6.png") -->
+<!-- ``` -->
 
@@
 
-<!-- ```{r cumsumSTD, echo = FALSE, fig.cap='two plots', fig.subcap=c('one plot', 'the other one'), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3} -->
-<!-- knitr::include_graphics(c( -->
-<!--     "./images/CumulativeDailyCumSumSTD-1.png",    # ALL sky -->
-<!--     "./images/CumulativeDailyCumSumSTD-2.png",    # CLEAR sky -->
-<!--     "./images/CumulativeDailyCumSumSTD-3.png"))   # Cloud sky -->
-<!-- ``` -->
-
 # Conclusions

### 2023-07-09 Article.Rmd
Commit 2023-07-09 21:09

@@
 Although, using the cumulative sums plots we can reveal some structure within the trends and some difference between the sky conditions [@Regier2019]. 
+On all three plots on Figure\ \ref{fig:cumsum}, we observe a step downward part in the begin of the period until about 2000, an almost steady part until about 2016 for All-sky and Clear-sky conditions (Figures\ \ref{fig:cumsum-1}, \ref{fig:cumsum-2}), and until 2009 for Cloud-sky conditions (Figure\ \ref{fig:cumsum-3}), and finally, a steep upward part until the present.
+For a more uniform trend we expect the cumulative sum plot of the anomaly, to have a symmetric 'V' shape.
+In this case......
+
+suspected stagnend period
 

### 2023-07-10 Article.Rmd
Commit 2023-07-10 12:11

@@
 On all three plots on Figure\ \ref{fig:cumsum}, we observe a step downward part in the begin of the period until about 2000, an almost steady part until about 2016 for All-sky and Clear-sky conditions (Figures\ \ref{fig:cumsum-1}, \ref{fig:cumsum-2}), and until 2009 for Cloud-sky conditions (Figure\ \ref{fig:cumsum-3}), and finally, a steep upward part until the present.
-For a more uniform trend we expect the cumulative sum plot of the anomaly, to have a symmetric 'V' shape.
-In this case......
-
-suspected stagnend period
-
-Comparining the conditions
-
-Variability issues
+For a more uniform trend, we expect the cumulative sum plot of the anomaly, to have a symmetric 'V' shape.
+In our case, for all sky conditions we observe a period where the cumulative remain relative steady.
+This indicates, that the anomaly was evenly distributed around the climatological mean.
+Another distinct feature of the cumulative plots, is the difference, between the sky conditions, of the start of the sums sudden increase, for the latter part of the period.
+This can make us to suspected that clear-Sky conditions SDR and cloud-sky SDR is affected by different mechanisms.
 
@@
 
-## Long term trends by Solar zenith angle.
+## Effects of the solar zenith angle on SDR.
+
+.... cite paper low elevation SDR .....
+
+The relative position of the Sun in the sky, seem to be a major drive of the attenuation of SDR, as it is affect
+
+... Aerosols and clouds ...
 

### 2023-07-10 Article.Rmd
Commit 2023-07-10 17:14

@@
 
-```{r cumsum, echo = FALSE, fig.cap='Running plots of the cumulative sum of the daily SDR for three different sky conditions', fig.subcap=c('All-sky conditions', 'Clear-sky conditions', 'Cloud-sky conditions'), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3}
+```{r cumsum, echo = FALSE, fig.cap='Running cumulative sum plots of the daily SDR for three different sky conditions', fig.subcap=c('All-sky conditions', 'Clear-sky conditions', 'Cloud-sky conditions'), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3}
 knitr::include_graphics(c(
@@
 
+We group the anomaly data in $1^\circ$ SZA bins, and calculate the overall trend for each bin
+
 Analyzing the long term trends of SDR to bins of SZA, we can see the contribution of the geometry and time in a diurnal level (Figures\ \ref{fig:szatrends}a and \ref{fig:szatrends}b).
@@
 
-<!-- ```{r , echo=F, fig.cap="All sky"} -->
-<!-- knitr::include_graphics("./images/SzaTrends-5.png") -->
-<!-- ``` -->
-
-<!--
-```{r , echo=F, fig.cap="Clear sky"}
+```{r szatrends, echo = FALSE, fig.cap='Distribution of the SDR\'s long term trends by SZA.', fig.subcap=c('Trend distribution for all-sky conditions.', 'Trend distribution for cloud-sky conditions.'), out.width='.35\\linewidth', out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3}
 knitr::include_graphics("./images/SzaTrends-13.png")
-```
--->
-
-<!-- ```{r , echo=F, fig.cap="Cloud sky"}
-knitr::include_graphics("./images/SzaTrends-21.png")
-``` -->
-
-```{r szatrends, echo = FALSE, fig.cap='Distribution of the SDR\'s long term trends by SZA.', fig.subcap=c('Trend distribution for all-sky conditions.', 'Trend distribution for cloud-sky conditions.'), out.width='.35\\linewidth', fig.asp=1, fig.ncol = 2}
-knitr::include_graphics("./images/SzaTrends-13.png")
+knitr::include_graphics("./images/SzaTrends-17.png" )
 knitr::include_graphics("./images/SzaTrends-21.png")

### 2023-07-10 Article.Rmd
Commit 2023-07-10 18:14

@@
 load("./data/common_data_14_2.Rda")
+## number of measurements
+SDR_N_All   <- sum(  ALL_1_daily_DESEAS[,GLB_att_N])
+SDR_N_Clear <- sum(CLEAR_1_daily_DESEAS[,GLB_att_N])
+SDR_N_Cloud <- sum(CLOUD_1_daily_DESEAS[,GLB_att_N])
+
 TSIinfo <- data.table(read.csv("./figures/tbl_tsi_info.dat",
@@
 
+Of the total 1-minute measurements
+$`r signif(100 * SDR_N_Clear / SDR_N_All, 3)`\%$ 
+is characterized with clear-sky conditon, and the rest 
+$`r signif(100 * SDR_N_Cloud / SDR_N_All, 3)`\%$
+with cloud-sky conditions.
+
 <!-- ### Data selection -->
@@
 
-We group the anomaly data in $1^\circ$ SZA bins, and calculate the overall trend for each bin
+We group the anomaly data in $1^\circ$ SZA bins, and calculated the overall trend for each bin
 
@@
 
-```{r szatrends, echo = FALSE, fig.cap='Distribution of the SDR\'s long term trends by SZA.', fig.subcap=c('Trend distribution for all-sky conditions.', 'Trend distribution for cloud-sky conditions.'), out.width='.35\\linewidth', out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3}
+```{r szatrends, echo = FALSE, fig.cap='Distribution of the SDR\'s long term trends by SZA.', fig.subcap=c('All-sky conditions.', 'Clear-sky conditions.', 'Cloud-sky conditions.'), out.width='.35\\linewidth', out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3}
 knitr::include_graphics("./images/SzaTrends-13.png")

### 2023-07-10 DHI_GHI_1_longterm_trends.R

@@ -563,7 +563,7 @@ for (DBn in dbs) {
                    paste("Trend: ",
                          if (fit[2] > 0) '+' else '-',
                          signif(abs(fit[2]), 3),
-                         "% per year")
+                         "% / year")
             )
 
         }
@@ -684,11 +684,11 @@ for (avar in vars) {
             ## decorations
             fit <- lm2[[1]]
 
-            legend("bottom", lty = 1, bty = "n", lwd = 2, cex = 1,
+            legend("bottom", lty = 1, bty = "n", lwd = 2, cex = 0.8,
                    paste("Trend: ",
                          if (fit[2] > 0) '+' else '-',
                          signif(abs(fit[2]), 3),
-                         "% per year")
+                         "% / year")
             )
         }
     }

### 2023-07-10 Article.Rmd
Commit 2023-07-10 20:13

@@
 
-As global radiation clear sky reference we are using the Haurwitz’s model, adjusted for our site with a factor of `r signif(alpha,3)` (Eq.\ \ref{eq:ahau}).
+As SDR clear sky reference we are using the radiation model by Haurwitz’s, adjusted for our site with the factor  $a = `r signif(alpha,3)`$ (Eq.\ \ref{eq:ahau}), resulted by an iterative optimization process.
 The selection of a clear sky reference model, was based on SDR observation from the period 2016 – 2021.
-Where, after an iterative optimization of eight simple models (Daneshyar–Paltridge–Proctor, Kasten–Czeplak, Haurwitz, Berger–Duffie, Adnot–Bourges–Campana–Gicquel, Robledo-Soler, Kasten and Ineichen-Perez) with different factors.
-We found, that Haurwitz’s model, adjusted with a factor of `r signif(alpha,3)` has the lower root mean squared error (RMSE). 
-The tried models are described by @Reno2012 and tested by @Reno2016.
-The iterative optimization method, for the selection of the reference is discussed by @Long2000 and @Reno2016.
+Where, after an iterative optimization of eight simple models (Daneshyar–Paltridge–Proctor, Kasten–Czeplak, Haurwitz, Berger–Duffie, Adnot–Bourges–Campana–Gicquel, Robledo-Soler, Kasten and Ineichen-Perez) with a wide range factors.
+We found, that Haurwitz’s model, adjusted with a factor of `r signif(alpha,3)` has one of the lowest root mean squared error (RMSE), while retains most of our data. 
+The tried models are described in more details by @Reno2012 and tested by @Reno2016.
+The iterative optimization method, for the selection of the reference is described by @Long2000 and @Reno2016.
+The target of the optimization was the minimizing of the function .......
+
+\begin{equation}
+f(a)=\frac{1}{n}\sum_{i=1}^{n}( \text{SDR}_{\text{meas},CS,i} - a \times \text{SDR}_{\text{ref},CS,i})^2
+\end{equation}
 
@@
 
+${MSE}={Cost}=\frac{1}{n}\sum_{i=1}^{n}(A_i - B_i)^2$
+
 -------------------------------------------------

### 2023-07-10 Build_150/Article_B150.Rmd
Commit 2023-07-10 21:13

--- 2023-07-10 Article.Rmd
@@
 -->
+....
 
-The shortwave downward solar irradiance (SDR) at Earth's surface play a significant role, on it's climate.
+more about trend in relation to other trends
+Aerosols and clouds
+
+......
+
+The shortwave downward solar irradiance (SDR) at Earth's surface play a significant role, on its climate.
 Changes of the SDR can be related to changes on Earth's energy budget, the mechanisms of climate change, and water and carbon cycle [@Wild2009].
@@
 Moreover, the DNI was used as auxiliary data, with the clear sky identification algorithm (CSid), to aid with the selection of the appropriate thresholds. 
-We have to note also, that despite the capability of the CSid to use the DNI as an characterization parameter, we haven't use it, to avoid any selection bias due to unequal availability of measurements.
+We have to note also, that despite the capability of the CSid to use the DNI as a characterization parameter, we haven't used it, to avoid any selection bias due to unequal availability of measurements.
 
@@
 During the study period the CM-21 pyranometer has been independently calibrated four times.
+........ to verify number ........
 <!-- Preparation. -->
@@
 <!-- dark correction -->
-The measurements are corrected for the zero offset of the instrument signal.
+The measurements were corrected for the zero offset of the instrument signal.
 As reference of the "dark signal" is used the Sun elevation angle bellow $-10^\circ$, for a period of $3 \text{h}$.
@@
 
-....
-
-cite iterative process
-
-......
-
 <!-- CS subjectivity -->
@@
 As a result, the details of the definition are site specific, it relies on a combination of threshold of comparisons with ideal actinometric models and statistics on different signal behavior.
-
 <!-- basic concepts -->
@@
 The iterative optimization method, for the selection of the reference is described by @Long2000 and @Reno2016.
-The target of the optimization was the minimizing of the function .......
+The target of the optimization was the minimization of the function $f(a)$ (Eq.\ \ref{eq:minf}) and was accomplished by the computing function `stats::optimise` [@RCT2023], which is based on work by @Brent1973.
 
 \begin{equation}
-f(a)=\frac{1}{n}\sum_{i=1}^{n}( \text{SDR}_{\text{meas},CS,i} - a \times \text{SDR}_{\text{ref},CS,i})^2
+f(a)=\frac{1}{n}\sum_{i=1}^{n}( \text{SDR}_{\text{meas},CS,i} - a \times \text{SDR}_{\text{ref},CS,i})^2 \label{eq:minf}
 \end{equation}
@@
 
-Of the total 1-minute measurements
-$`r signif(100 * SDR_N_Clear / SDR_N_All, 3)`\%$ 
-is characterized with clear-sky conditon, and the rest 
-$`r signif(100 * SDR_N_Cloud / SDR_N_All, 3)`\%$
-with cloud-sky conditions.
-
 <!-- ### Data selection -->
@@
 
+After the preparation of the data, we have $`r SDR_N_All`$  1-minute measurements, of which,
+$`r signif(100 * SDR_N_Clear / SDR_N_All, 3)`\%$ 
+are characterized with clear-sky condition, and the rest 
+$`r signif(100 * SDR_N_Cloud / SDR_N_All, 3)`\%$
+with cloud-sky conditions.
+
 ### Aggregation of radiometric data
@@
 <!-- Seasonal -->
-For the seasons of the year trends, we grouped the mean daily values by season (December - February: Winter, March - May: Spring, etc.).
+For the seasons of the year trends, we grouped the mean daily values by season (Winter: December - February, Spring: March - May, etc.).
 <!-- Deseasonalization -->
@@
 Although, using the cumulative sums plots we can reveal some structure within the trends and some difference between the sky conditions [@Regier2019]. 
-On all three plots on Figure\ \ref{fig:cumsum}, we observe a step downward part in the begin of the period until about 2000, an almost steady part until about 2016 for All-sky and Clear-sky conditions (Figures\ \ref{fig:cumsum-1}, \ref{fig:cumsum-2}), and until 2009 for Cloud-sky conditions (Figure\ \ref{fig:cumsum-3}), and finally, a steep upward part until the present.
+On all three plots on Figure\ \ref{fig:cumsum}, we observe a step downward part at the start of the period until about 2000, an almost steady part until about 2016 for All-sky and Clear-sky conditions (Figures\ \ref{fig:cumsum-1}, \ref{fig:cumsum-2}), and until 2009 for Cloud-sky conditions (Figure\ \ref{fig:cumsum-3}), and finally, a steep upward part until the present.
 For a more uniform trend, we expect the cumulative sum plot of the anomaly, to have a symmetric 'V' shape.
@@
 
-.... cite paper low elevation SDR .....
+.... cite paper about low elevation SDR .....
 
@@
 <!--
-## Zenith angle dependency and seasons of the year.
--->
-
 ## Consistency of the trends
@@
 ... different in aggregation ...
+-->
 
@@
 
-${MSE}={Cost}=\frac{1}{n}\sum_{i=1}^{n}(A_i - B_i)^2$
-
 -------------------------------------------------

### 2023-07-12 Article.Rmd
Commit 2023-07-12 19:11

@@
 ...... cloud formation occurrences and constitution nuclei. cite .....
+p>        "  Too hard to vim
 
@@
 <!-- time span -->
-The SDR is equivalent to the whole sky Global radiation (or Global Horizontal Irradiance, GHI), and was obtained with a horizontal leveled, CM-21 pyranometer, from 
+The SDR is equivalent to the whole sky Global radiation (or Global Horizontal Irradiance, GHI), and was obtained with a horizontal leveled, CM-21 pyranometer, in the period from
 <!-- \mbox{ -->
@@
 <!-- }. -->
-Also, we have measurements of direct beam radiation (DNI) obtained by a collocated CHP-1 pyrheliometer, from 
+Along with SDR, we obtained the direct beam radiation (DNI) by a collocated CHP-1 pyrheliometer, for the period
 <!-- \mbox{ -->
@@
 <!-- }. -->
-Although, we performed a similar analysis to the DNI as for the SDR, the results are not presented here, as they lack the appropriate statistical significance, due to the sorter duration of the data.
-Moreover, the DNI was used as auxiliary data, with the clear sky identification algorithm (CSid), to aid with the selection of the appropriate thresholds. 
-We have to note also, that despite the capability of the CSid to use the DNI as an characterization parameter, we haven't use it, to avoid any selection bias due to unequal availability of measurements.
+Although, we have performed a similar analysis to the DNI, as for the SDR, the results are not presented here, as they lack the appropriate statistical significance, due to the sorter duration of the data.
+Moreover, the DNI data were used as auxiliary data, with the clear sky identification algorithm (CSid), to aid with the selection of the appropriate thresholds. 
+We have to note, that despite the capability of the CSid algorithm, to use the DNI as a characterization parameter, we haven't used it here, to avoid any selection bias due to unequal availability of the two datasets.
 
@@
 
-The SDR data were measured with a Kipp & Zonen CM-21 pyranometer operating continuously at the Laboratory of Atmospheric Physics of the Aristotle University of Thessaloniki ($40^\circ38'$N, $22^\circ57'$E, 80m\ a.s.l.).
-The monitoring site is located near the city centre.
-During the study period the CM-21 pyranometer has been independently calibrated four times.
+The SDR data were measured with a Kipp & Zonen CM-21 pyranometer operating continuously at the Laboratory of Atmospheric Physics of the Aristotle University of Thessaloniki ($40^\circ\,38'\,$N, $22^\circ\,57'\,$E, $80\,$m\ a.s.l.).
+The monitoring site is located near the city centre, and we expect to be affected by the urban environment.
+During the study period, the CM-21 pyranometer has been independently calibrated ..... **four** .... times.
+........ verify number ........
 <!-- Preparation. -->
-There are three distinct steps to the creation of this dataset:
+There are three distinct steps in the creation of the dataset analysed here:
 a)\ the acquisition of radiation measurements from the sensors, 
@@
 For the acquisition of radiometric data,
-the signal of the broadband instruments is sampled with a rate of $1 \text{Hz}$. The mean and its standard deviation are recorded for every minute.
+the signal of the broadband instruments is sampled with a rate of $1\,\text{Hz}$. The mean and the standard deviation are recorded for every minute.
 <!-- dark correction -->
-The measurements are corrected for the zero offset of the instrument signal.
-As reference of the "dark signal" is used the Sun elevation angle bellow $-10^\circ$, for a period of $3 \text{h}$.
+The measurements are corrected for the zero offset of the instrument dark voltage signal.
+As reference of the "dark signal" is used the Sun elevation angle bellow $-10^\circ$, for a period of $3\,\text{h}$.
 <!-- physical quantities -->
-The signal is converted to radiation flux, using a ramped value of the instrument sensitivity, derived from the eight laboratory calibrations of the instrument, during the study period.
+The signal is converted to radiation flux, using a ramped value of the instrument sensitivity, derived from the ......
+**eight** ......
+laboratory calibrations of the instrument, during the study period.
 
@@
 <!-- screening -->
-A manual screening was performed, to remove inconsistencies and erroneous recordings, that can occur randomly or systematically, during the long continuous operation of the instrument.
+A manual screening was performed, to remove inconsistencies and erroneous recordings, that can occur stochastically or systematically, during the continuous operation of the instrument.
 <!-- Quality assurance -->
-The manual screening is aided by a radiation data quality assurance procedure, adjusted for this site, based on methods of Long and Shi\ [-@Long2008a; -@Long2006].
+The manual screening is aided by a radiation data quality assurance procedure, adjusted for the site, based on methods of Long and Shi\ [-@Long2008a; -@Long2006].
 Thus, problematic recordings have been excluded from further processing. 
-Although it is not possible to detect all the bad data, the large number of data and the aggregation scheme we use, can provide us with accurate radiometric measurements, for the scope of this study.
+Although it is not possible to detect all the bad data, the large number of available data, and the aggregation scheme we use, can provide us with accurate radiometric measurements, for the scope of this study.
 
 <!-- c) Clear sky identification -->
-In order to estimate the effect of the clouds on the SDR, we created three datasets, by characterizing each one-minute measurement with a corresponding sky condition. 
+In order to be able to estimate the effect of the clouds on the SDR, we created three datasets, by characterizing each one-minute measurement with a corresponding sky condition. 
 The all-sky conditions, containing all the valid measurements. 
-The clear-sky conditions data, correspond to data where we have inferred that the sky was almost clear of clouds, and the remainder part the data was characterized with cloud sky conditions (cloud-sky).
-To identify the clear-sky conditions we used the method proposed by @Long2000 and by @Reno2016, that was adapted and configured for the site.
-
-....
-
-cite iterative process
-
-......
+The clear-sky conditions data, corresponding to data, where we have inferred that the sky was almost clear of clouds, and the remainder part of the data, characterized as cloud-sky conditions. 
+To identify the clear-sky conditions we used the method proposed by @Long2000 and by @Reno2016, that was adapted and configured for the site, as the authors suggest.
 
 <!-- CS subjectivity -->
-We have to note, that the definition of what constitutes as clear or cloudy sky, has some subjectivity, due to the used method of characterization.
-As a result, the details of the definition are site specific, it relies on a combination of threshold of comparisons with ideal actinometric models and statistics on different signal behavior.
-
+We have to note, that the definition of what constitutes as clear or cloudy sky, has some subjectivity, in this method of characterization.
+As a result, the details of the definition is site specific, it relies on a combination of thresholds and comparisons with ideal actinometric models and statistical analysis on different signal metrics.
 <!-- basic concepts -->
-The algorithm was calibrated with the main focus to identify the presence of clouds on the sky dome, although there are marginal cases that there will be false positives or false negatives identifications.
+The algorithm was calibrated with the main focus to identify the presence of clouds on the sky dome, although few marginal cases exist that will be false positive or false negative identified, but can not affect the general outcome.
 
@@
 
+For completeness, we will provide a brief overview of the CSid process along with the site specific thresholds.
 As SDR clear sky reference we are using the radiation model by Haurwitz’s, adjusted for our site with the factor  $a = `r signif(alpha,3)`$ (Eq.\ \ref{eq:ahau}), resulted by an iterative optimization process.
-The selection of a clear sky reference model, was based on SDR observation from the period 2016 – 2021.
-Where, after an iterative optimization of eight simple models (Daneshyar–Paltridge–Proctor, Kasten–Czeplak, Haurwitz, Berger–Duffie, Adnot–Bourges–Campana–Gicquel, Robledo-Soler, Kasten and Ineichen-Perez) with a wide range factors.
-We found, that Haurwitz’s model, adjusted with a factor of `r signif(alpha,3)` has one of the lowest root mean squared error (RMSE), while retains most of our data. 
+The selection of this clear sky reference model, was based on SDR observations from the period 2016 – 2021.
+Where, after an iterative optimization between eight simple models (Daneshyar–Paltridge–Proctor, Kasten–Czeplak, Haurwitz, Berger–Duffie, Adnot–Bourges–Campana–Gicquel, Robledo-Soler, Kasten and Ineichen-Perez) with a wide range factors, we found, that Haurwitz’s model, adjusted with a factor of `r signif(alpha,3)` has one of the lowest root mean squared error (RMSE), while retains most of our data (Eq.\ \ref{eq:ahau}).
 The tried models are described in more details by @Reno2012 and tested by @Reno2016.
 The iterative optimization method, for the selection of the reference is described by @Long2000 and @Reno2016.
-The target of the optimization was the minimizing of the function .......
+The target of the optimization was the minimization of the function $f(a)$ (Eq.\ \ref{eq:minf}) and was accomplished by the computing function `stats::optimise` [@RCT2023], which is based on work by @Brent1973.
 
 \begin{equation}
-f(a)=\frac{1}{n}\sum_{i=1}^{n}( \text{SDR}_{\text{meas},CS,i} - a \times \text{SDR}_{\text{ref},CS,i})^2
+f(a)=\frac{1}{n}\sum_{i=1}^{n}( \text{SDR}_{\text{meas},CS,i} - a \times \text{SDR}_{\text{ref},CS,i})^2 \label{eq:minf}
 \end{equation}

### 2023-07-12 Article.Rmd
Commit 2023-07-12 22:10

@@
 
-For completeness, we will provide a brief overview of the CSid process along with the site specific thresholds.
-As SDR clear sky reference we are using the radiation model by Haurwitz’s, adjusted for our site with the factor  $a = `r signif(alpha,3)`$ (Eq.\ \ref{eq:ahau}), resulted by an iterative optimization process.
-The selection of this clear sky reference model, was based on SDR observations from the period 2016 – 2021.
-Where, after an iterative optimization between eight simple models (Daneshyar–Paltridge–Proctor, Kasten–Czeplak, Haurwitz, Berger–Duffie, Adnot–Bourges–Campana–Gicquel, Robledo-Soler, Kasten and Ineichen-Perez) with a wide range factors, we found, that Haurwitz’s model, adjusted with a factor of `r signif(alpha,3)` has one of the lowest root mean squared error (RMSE), while retains most of our data (Eq.\ \ref{eq:ahau}).
-The tried models are described in more details by @Reno2012 and tested by @Reno2016.
-The iterative optimization method, for the selection of the reference is described by @Long2000 and @Reno2016.
-The target of the optimization was the minimization of the function $f(a)$ (Eq.\ \ref{eq:minf}) and was accomplished by the computing function `stats::optimise` [@RCT2023], which is based on work by @Brent1973.
+For completeness, we will provide a brief overview of the CSid process, along with the site specific thresholds.
+As a reference clear sky SDR, we are using the radiation model by Haurwitz’s, adjusted for our site with the factor $a = `r signif(alpha,3)`$ (Eq.\ \ref{eq:ahau}), resulted by an iterative optimization process, as described by @Long2000 and @Reno2016.
+The target of the optimization was the minimization of the function $f(a)$ (Eq.\ \ref{eq:minf}) and was accomplished with the computing function `stats::optimise` [@RCT2023], which is based on work by @Brent1973.
+The optimization and the selection of the clear sky reference model, was performed on SDR observations for the period 2016 – 2021.
+During the optimization, eight simple clear sky radiation models were tested (Daneshyar–Paltridge–Proctor, Kasten–Czeplak, Haurwitz, Berger–Duffie, Adnot–Bourges–Campana–Gicquel, Robledo-Soler, Kasten and Ineichen-Perez), with a wide range of factors.
+We found, that Haurwitz’s model, adjusted with the factor $a = `r signif(alpha,3)`$ has one of the lowest root mean squared error (RMSE), while correlates with large amount of our data.
+Thus, our clear sky reference is the Eq.\ \ref{eq:ahau}.
+The tried models are described in more details by @Reno2012 and evaluated by @Reno2016.
 

### 2023-07-12 Article.Rmd
Commit 2023-07-12 23:09

@@
 For completeness, we will provide a brief overview of the CSid process, along with the site specific thresholds.
-As a reference clear sky SDR, we are using the radiation model by Haurwitz’s, adjusted for our site with the factor $a = `r signif(alpha,3)`$ (Eq.\ \ref{eq:ahau}), resulted by an iterative optimization process, as described by @Long2000 and @Reno2016.
-The target of the optimization was the minimization of the function $f(a)$ (Eq.\ \ref{eq:minf}) and was accomplished with the computing function `stats::optimise` [@RCT2023], which is based on work by @Brent1973.
+As a reference clear sky SDR, we are using the radiation model by Haurwitz’s, adjusted for our site with a factor $a$, resulting by an iterative optimization process, as described by @Long2000 and @Reno2016.
+The target of the optimization was the minimization of the function $f(a)$ (Eq.\ \ref{eq:minf}) and was accomplished with the computing function `stats::optimise` [@RCT2023], based on work by @Brent1973.
 The optimization and the selection of the clear sky reference model, was performed on SDR observations for the period 2016 – 2021.
@@
 
+<!-- cost fuction -->
 \begin{equation}
-f(a)=\frac{1}{n}\sum_{i=1}^{n}( \text{SDR}_{\text{meas},CS,i} - a \times \text{SDR}_{\text{ref},CS,i})^2 \label{eq:minf}
+f(a)=\frac{1}{n}\sum_{i=1}^{n}( \text{SDR}_{\text{CSid},i} - a \times \text{SDR}_{\text{testCSref},i})^2 \label{eq:minf}
 \end{equation}
 
+<!-- adjusted Haurwitz -->
 \begin{equation}
-\text{SDR}_\text{Clear Sky} = `r signif(alpha,3)` \times 1098 \times \cos( \text{SZA} ) \times \exp \left( \frac{ - 0.057}{\cos(\text{SZA})} \right) \label{eq:ahau}
+\text{SDR}_\text{CSref} = a \times \text{SDR}_\text{Haurwitz}  = `r signif(alpha,3)` \times 1098 \times \cos( \text{SZA} ) \times \exp \left( \frac{ - 0.057}{\cos(\text{SZA})} \right) \label{eq:ahau}
 \end{equation}
@@
 Each criterion was applied for a running window of $`r nt`$ consecutive measurements/minutes, and the characterization is applied at the center value of the window.
-A data point is consider as under cloud-sky condition if it fails to pass any of the criteria, all the other data points are characterized as clear-sky.
+A data point is consider as under cloud-sky condition if it fails to pass any of the criteria, the rest data points are characterized as clear-sky.
 
@@
 
-The mean of the measured value $\overline{G}_i$ must be inside an envelope based on the reference model $\text{SDR}_\text{Clear Sky}$ (Eq. \ref{eq:MeanVIP}).
+The mean of the measured value $\overline{\text{SDR}}_i = \text{mean}_{n=i-5}^{n=i+5}[\text{SDR}_{n}]$ must be inside an envelope based on the equivalent mean of the reference model $\overline{\text{SDR}}_\text{CSref} = \text{mean}_{n=i-5}^{n=i+5}[\text{SDR}_{\text{CSref},n}]$ (Eq.
+\ref{eq:MeanVIP}).
 
 \begin{equation}
-`r 1 - CS_ref_rm_VIP_RelLow` \times \overline{\text{SDR}}_{i\text{Clear Sky}} - `r CS_ref_rm_VIP_LowOff`
-< \overline{G}_i <
-`r 1 + CS_ref_rm_VIP_RelUpp` \times \overline{\text{SDR}}_{i\text{Clear Sky}} + `r CS_ref_rm_VIP_UppOff`
+`r 1 - CS_ref_rm_VIP_RelLow` \times \overline{\text{SDR}}_{\text{CSref},i} - `r CS_ref_rm_VIP_LowOff`
+< \overline{SDR}_i <
+`r 1 + CS_ref_rm_VIP_RelUpp` \times \overline{\text{SDR}}_{\text{CSref},i} + `r CS_ref_rm_VIP_UppOff`
 \label{eq:MeanVIP}
@@
 
-The running max measured value $M_{Gi} = max[\text{SDR}_{i}]$, is compared to a similar constructed value from the reference $M_{CSi} = max[\text{SDR}_{i\text{Clear Sky}}]$ (Eq.\ \ref{eq:MaxVIP}).
+The running max measured value $M_{\text{SDR},i} = \max_{n=i-5}^{n=i+5}[\text{SDR}_{n}]$, is compared to a similar constructed value from the reference $M_{CSi} = \max_{n=i-5}^{n=i+5}[\text{SDR}_{\text{CSref},n}]$ (Eq.\ \ref{eq:MaxVIP}).
 
@@
 `r MaxVIP_fct` \times M_{CSi} - `r MaxVIP_off_low`
-< M_{Gi} <
+< M_{\text{SDR},i} <
 `r MaxVIP_fct` \times M_{CSi} + `r MaxVIP_off_upp`
@@
 
-The length $L$ (Eq. \ref{eq:VILeq}) of the sequence of line segments connecting the points of the SDR time
-series for the measured values $L$ and similar for the reference $L_{CS}$, must be within the limits of Eq. \ref{eq:VILcr}.
+The length $L$ of the sequence of line segments, connecting the points of the SDR time series for the measured values (Eq.\ \ref{eq:VILeq}), and similar for the reference $L_{CS}$, must be within the limits of Eq.
+\ref{eq:VILcr}.
 
 \begin{equation}
-L = \sum_{i=1}^{n-1}\sqrt{\left ( \text{SDR}_{i+1} - \text{SDR}_{i}\right )^2 + \left ( t_{i+1} - t_i \right )^2}
+L_i = \sum_{i=1}^{n-1}\sqrt{\left ( \text{SDR}_{i+1} - \text{SDR}_{i}\right )^2 + \left ( t_{i+1} - t_i \right )^2}
 \label{eq:VILeq}
@@
 <!-- Standard deviation of rate of change in irradiance. -->
-We calculate the standard deviation $\sigma$ of the slope
-($s$) between sequential points in the time series, normalized by the average SDR during the time interval.
+The standard deviation $\sigma$ of the slope ($s$) between sequential points in the time series, normalized by the average SDR during the time interval, must be bellow a threshold (Eq.\ \ref{eq:VCTcr}).
 

### 2023-07-12 Article.Rmd
Commit 2023-07-13 00:10

@@
 
-The maximum difference $X$ between the change in measured irradiance and the change in clear sky
-irradiance over each measurement interval.
+The maximum difference $X$ between the change in measured irradiance and the change in clear sky irradiance over each measurement interval, must be lower than a threshold (Eq.\ \ref{eq:VSMcr}).
 
@@
 x_i = \text{SDR}_{i+1} - \text{SDR}_{i} \forall i \in \left \{ 1, 2, \ldots, n-1 \right \} \label{eq:VSM1} \\
-x_{CS,i} = \text{SDR}_{CS,i+1} - \text{SDR}_{CS,i} \forall i \in \left \{ 1, 2, \ldots, n-1 \right \} \label{eq:VSM2} \\
-X_i = \max{\left \{ \left | x_i - x_{CS,i} \right | \right \}} \label{eq:VSM3}
+x_{\text{CSref},i} = \text{SDR}_{\text{CSref},i+1} - \text{SDR}_{\text{CSref},i} \forall i \in \left \{ 1, 2, \ldots, n-1 \right \} \label{eq:VSM2} \\
+X_i = \max{\left \{ \left | x_i - x_{\text{CSref},i} \right | \right \}} \label{eq:VSM3}
 \end{gather}
@@
 <!-- ### Data selection -->
-Due to a significant measurement uncertainty near, the horizon, we have to exclude
-all measurements with SZA greater than $`r 90-MIN_ELEVA`^\circ$. 
+Due to a significant measurement uncertainty near the horizon, we have to exclude all measurements with SZA greater than $`r 90-MIN_ELEVA`^\circ$. 
 Moreover, due to some obstructions around the site (hills and buildings), we excluded data with Azimuth angle between 
@@
 <!-- TSI -->
-Subsequently, we made all measurements relative to the Total Solar Irradiance (TSI) at $1 \text{au}$, in order to compensate for the Sun's intensity variability, using a homogenized time series of satellite TSI observations.
+Subsequently, we made all measurements relative to the Total Solar Irradiance (TSI) at $1\,\text{au}$, in order to compensate for the Sun's intensity variability, using a homogenized time series of satellite TSI observations.
 The TSI data we use, comes from NOAA [@Coddington2005], it overlaps with our data, and is interpolated to match with the time step of the data.
@@
 As a result, we can present all radiation data as a fraction of Sun's TSI. 
-
 After the preparation of the data, we have $`r SDR_N_All`$  1-minute measurements, of which,
@@
 with cloud-sky conditions.
+
+.......!!!.......
 

### 2023-07-13 Build_151/Article_B151.Rmd
Commit 2023-07-13 08:09

--- 2023-07-12 Article.Rmd
@@
 
-.......!!!.......
-
-### Aggregation of radiometric data
-
 <!-- aggregation -->
-Before further analysis and deseasonalization, we implement an appropriate aggregation scheme on the 1-minute data. To preserve the representativeness we use the following criteria:
-a) for daily mean values we exclude instances with less than `r Daily_aggregation_N_lim` valid data points,
-b) accordingly, monthly values are computed by daily aggregated data, where months with less than `r Monthly_aggegation_N_lim` days are rejected. 
-The daily seasonal values were indexed based on the day of year number, and the monthly by the corresponding calendar month.
+In order to investigate the SDR trend, we implement an appropriate aggregation scheme of the 1-minute data to a coarser time scale.
+To preserve the representativeness we are using the following criteria:
+a) for daily mean values we exclude days with less than `r Daily_aggregation_N_lim` valid data points,
+b) accordingly, monthly values are computed by daily means, and months with less than `r Monthly_aggegation_N_lim` days are omitted. 
+To create the daily and monthly seasonal values, we group the data based on the day of year number, and the corresponding calendar month respectively.
 <!-- Seasonal -->
-For the seasons of the year trends, we grouped the mean daily values by season (Winter: December - February, Spring: March - May, etc.).
+For the seasons of the year trends, we grouped the mean daily values by the corresponding season (Winter: December - February, Spring: March - May, etc.).
 <!-- Deseasonalization -->
-Finally, for each data set, we remove the natural occurring seasonal variation.
-This is done, by calculating the mean values for the appropriate time step, and subtracting the annual cycle from the actual data.
+Finally, for each data set, we remove the natural occurring seasonal variation, by calculating the mean values for the appropriate time step, and subtracting the annual cycle from the actual data.
 <!-- SZA Deseasonalization -->
-In addition to the previous aggregation scheme, each dataset was aggregate in $1^\circ$ SZA bins, separate for cases before and after local noon. 
-This gave us an approximation of Sun's SZA contribution to the SDR brightening.
-
-Thus, for each data set we have obtained the relative departure of the seasonal mean.
-The statistical significance and aggregation scheme and filtering will be noted along with the corresponding results.
+To estimate Sun's SZA contribution to the SDR brightening, the 1-minute data were aggregate in $1^\circ$ SZA bins, separated for instances before and after the local noon, and then was deseasonalized as above.
+Thus, for each dataset we have obtained the relative departure of the seasonal mean.
 

### 2023-07-14 Article.Rmd
Commit 2023-07-13 07:10
Commit 2023-07-14 17:11

@@
 
-<!-- time span -->
-The SDR is equivalent to the whole sky Global radiation (or Global Horizontal Irradiance, GHI), and was obtained with a horizontal leveled, CM-21 pyranometer, in the period from
+The SDR data were measured with a Kipp & Zonen CM-21 pyranometer operating continuously at the Laboratory of Atmospheric Physics of the Aristotle University of Thessaloniki 
+($40^\circ\,38'\,$N, $22^\circ\,57'\,$E, $80\,$m\ a.s.l.)
+in the period from 
 <!-- \mbox{ -->
@@
 <!-- }. -->
-Along with SDR, we obtained the direct beam radiation (DNI) by a collocated CHP-1 pyrheliometer, for the period
+The monitoring site is located near the city centre, and we expect to be affected by the urban environment. 
+During the study period, the pyranometer has been independently calibrated three times at the Meteorologisches Observatorium Lindenberg, DWD, when it was verified the stability of the instrument to within better than $0.7\%$ relative to the initial calibration by the manufacturer.
+Along with SDR, the direct beam radiation (DNI) was also measured by a collocated Kipp & Zonen CHP-1 pyrheliometer, for the period
 <!-- \mbox{ -->
@@
 <!-- }. -->
-Although, we have performed a similar analysis to the DNI, as for the SDR, the results are not presented here, as they lack the appropriate statistical significance, due to the sorter duration of the data.
-Moreover, the DNI data were used as auxiliary data, with the clear sky identification algorithm (CSid), to aid with the selection of the appropriate thresholds. 
-We have to note, that despite the capability of the CSid algorithm, to use the DNI as a characterization parameter, we haven't used it here, to avoid any selection bias due to unequal availability of the two datasets.
-
-<!-- Location -->
-
-The SDR data were measured with a Kipp & Zonen CM-21 pyranometer operating continuously at the Laboratory of Atmospheric Physics of the Aristotle University of Thessaloniki ($40^\circ\,38'\,$N, $22^\circ\,57'\,$E, $80\,$m\ a.s.l.).
-The monitoring site is located near the city centre, and we expect to be affected by the urban environment.
-During the study period, the CM-21 pyranometer has been independently calibrated ..... **four** .... times.
-........ verify number ........
-<!-- Preparation. -->
+Although, we have performed a similar analysis to the DNI data the results are not presented here, as they lack the appropriate statistical significance, due to the sorter duration of the data. 
+However, the DNI data were used as auxiliary data, in the clear sky identification algorithm (CSid), which is discussed later, for the selection of the appropriate thresholds.
+It is noted that despite the capability of the CSid algorithm to use the DNI as a characterization parameter, we haven't used it here, to avoid any selection bias due to unequal length of the two datasets.
 There are three distinct steps in the creation of the dataset analysed here:
 a)\ the acquisition of radiation measurements from the sensors,
-b)\ a radiation data quality check,
-c)\ the identification of "clear sky" conditions from the radiometric data, and
-d)\ data aggregation and trend analysis.
+b)\ the data quality check,
+c)\ the identification of "clear sky" conditions from the radiometric data, and d)\ the aggregation of data and trend analysis.
 
@@
 <!-- Zerefos 
-
 Northern Greece has been selected because it is well known to be located at the crossroads of aerosols and various pollutants pathways originating in central and Eastern
@@
 
-<!-- a) Acquisition -->
-For the acquisition of radiometric data,
-the signal of the broadband instruments is sampled with a rate of $1\,\text{Hz}$. The mean and the standard deviation are recorded for every minute.
-<!-- dark correction -->
-The measurements are corrected for the zero offset of the instrument dark voltage signal.
-As reference of the "dark signal" is used the Sun elevation angle bellow $-10^\circ$, for a period of $3\,\text{h}$.
-<!-- physical quantities -->
-The signal is converted to radiation flux, using a ramped value of the instrument sensitivity, derived from the ......
-**eight** ......
-laboratory calibrations of the instrument, during the study period.
+<!-- a) acquisition -->
+For the acquisition of radiometric data, the signal of the pyranometer is sampled with a rate of $1\,\text{Hz}$.
+The mean and the standard deviation of these samples are recorded every minute.
+The measurements are corrected for the zero offset (dark signal in volts).
+The "dark signal" is calculated by averaging all measurements recorded for a period of 
+$3\,\text{h}$ 
+before (morning) or after (evening), when the Sun reaches an elevation angle of
+$-10^\circ$.
+The signal is converted to irradiance using a ramped value of the instrument’s sensitivity between the calibrations.
 
 <!-- b) Radiation data quality check -->
-<!-- The radiation quality is assured with a manual screening, aided by an automatic assessment process. -->
-<!-- screening -->
-A manual screening was performed, to remove inconsistencies and erroneous recordings, that can occur stochastically or systematically, during the continuous operation of the instrument.
-<!-- Quality assurance -->
-The manual screening is aided by a radiation data quality assurance procedure, adjusted for the site, based on methods of Long and Shi\ [-@Long2008a; -@Long2006].
+A manual screening was performed, to remove inconsistent and erroneous recordings that can occur stochastically or systematically, during the continuous operation of the instruments.
+The manual screening is aided by a radiation data quality assurance procedure, adjusted for the site, which is based on the methods of 
+Long and Shi\ [-@Long2008a; -@Long2006].
 Thus, problematic recordings have been excluded from further processing. 
-Although it is not possible to detect all the bad data, the large number of available data, and the aggregation scheme we use, can provide us with accurate radiometric measurements, for the scope of this study.
+Although it is impossible to detect all false data, the large number of available data, and the aggregation scheme we used, ensures the good quality of the radiometric measurements used in this study.
 
@@
 <!-- Standard deviation of rate of change in irradiance. -->
-The standard deviation $\sigma$ of the slope ($s$) between sequential points in the time series, normalized by the average SDR during the time interval, must be bellow a threshold (Eq.\ \ref{eq:VCTcr}).
+The standard deviation $\sigma$ of the slope ($s$) between sequential points in the time series, normalized by the average SDR during the time interval, must be below a threshold (Eq.\ \ref{eq:VCTcr}).
 

### 2023-07-14 Article.Rmd
Commit 2023-07-14 18:14

@@
 library(data.table)
-## load data
+## load data from trends analysis
 load("./data/common_data_14_2.Rda")
-## number of measurements
+## number of measurements for each dataset
 SDR_N_All   <- sum(  ALL_1_daily_DESEAS[,GLB_att_N])
@@
 SDR_N_Cloud <- sum(CLOUD_1_daily_DESEAS[,GLB_att_N])
-
+## TSI data span info
 TSIinfo <- data.table(read.csv("./figures/tbl_tsi_info.dat",
@@
                     header = TRUE))
-## variables we use
+## variables we use in the analysis
 source("./DHI_GHI_0_variables.R")
@@
 source("~/CODE/FUNCTIONS/R/gather_citations.R")
-
 suppressMessages(
@@
 <!-- c) Clear sky identification -->
-In order to be able to estimate the effect of the clouds on the SDR, we created three datasets, by characterizing each one-minute measurement with a corresponding sky condition. 
-The all-sky conditions, containing all the valid measurements. 
-The clear-sky conditions data, corresponding to data, where we have inferred that the sky was almost clear of clouds, and the remainder part of the data, characterized as cloud-sky conditions. 
-To identify the clear-sky conditions we used the method proposed by @Long2000 and by @Reno2016, that was adapted and configured for the site, as the authors suggest.
+In order to be able to estimate the effect of the clouds on the long term variability of SDR, we created three datasets, by characterizing each one-minute measurement with a corresponding sky condition (i.e., all-sky, clear-sky and cloudy-sky).
+To identify the clear-sky conditions we used the method proposed by
+@Long2000
+and by
+@Reno2016,
+which was adapted and configured for the site, as the authors suggest.
 
@@
 As a result, the details of the definition is site specific, it relies on a combination of thresholds and comparisons with ideal actinometric models and statistical analysis on different signal metrics.
-<!-- basic concepts -->
-The algorithm was calibrated with the main focus to identify the presence of clouds on the sky dome, although few marginal cases exist that will be false positive or false negative identified, but can not affect the general outcome.
+The CSid algorithm was calibrated with the main focus to identify the presence of clouds on the sky dome, although few marginal cases exist that have been identified as false positive or false negative, but cannot affect the results of the study.
 

### 2023-07-14 Article.Rmd
Commit 2023-07-14 20:14
Commit 2023-07-14 21:09

@@
    - \usepackage{subcaption}
-
 output: 
-  rticles::elsevier_article:
-    keep_tex:         no
-    citation_package: natbib
+  bookdown::word_document2: default
   bookdown::pdf_document2:
@@
     biblio-style:     apalike
-  bookdown::word_document2: default
+  rticles::elsevier_article:
+    keep_tex:         no
+    citation_package: natbib
 # date:           "`r format(Sys.time(), '%F')`"
@@
 <!-- 1au -->
-To make the measurements comparable throughout the dataset, we adjusted all 1-minute radiometric values to the mean Sun - Earth distance.
+To make the measurements comparable throughout the dataset, we adjusted all one-minute radiometric values to the mean Sun - Earth distance.
 <!-- TSI -->
 Subsequently, we made all measurements relative to the Total Solar Irradiance (TSI) at $1\,\text{au}$, in order to compensate for the Sun's intensity variability, using a homogenized time series of satellite TSI observations.
-The TSI data we use, comes from NOAA [@Coddington2005], it overlaps with our data, and is interpolated to match with the time step of the data.
+The TSI data we use, comes from NOAA [@Coddington2005], and were interpolated to match with the time step of our measurements. 
+The final dataset contains
+$`r SDR_N_All`$
+one-minute measurements, of which,  $`r signif(100 * SDR_N_Clear / SDR_N_All, 3)`\%$
+correspond to clear-sky conditions and 
+$`r signif(100 * SDR_N_Cloud / SDR_N_All, 3)`\%$
+to cloud-sky conditions.
 <!-- The TSI data we use, is a combination of data from NOAA [@Coddington2005] (for `r as.Date(TSIinfo[TSI_Source == "NOAA", Start])` - `r as.Date(TSIinfo[TSI_Source == "NOAA", End])`) and adjusted data from @LASP2023 (for `r as.Date(TSIinfo[TSI_Source != "NOAA", Start])` - `r as.Date(TSIinfo[TSI_Source != "NOAA", End])`).-->
-As a result, we can present all radiation data as a fraction of Sun's TSI. 
-After the preparation of the data, we have $`r SDR_N_All`$  1-minute measurements, of which,
-$`r signif(100 * SDR_N_Clear / SDR_N_All, 3)`\%$ 
-are characterized with clear-sky condition, and the rest 
-$`r signif(100 * SDR_N_Cloud / SDR_N_All, 3)`\%$
-with cloud-sky conditions.
 
 <!-- aggregation -->
-In order to investigate the SDR trend, we implement an appropriate aggregation scheme of the 1-minute data to a coarser time scale.
-To preserve the representativeness we are using the following criteria:
+In order to investigate the SDR trends, we implemented an appropriate aggregation scheme to the 1-minute data to derive series in coarser time scales.
+To preserve the representativeness of the data we used the following criteria: 
 a) for daily mean values we exclude days with less than `r Daily_aggregation_N_lim` valid data points,
-b) accordingly, monthly values are computed by daily means, and months with less than `r Monthly_aggegation_N_lim` days are omitted. 
-To create the daily and monthly seasonal values, we group the data based on the day of year number, and the corresponding calendar month respectively.
+b) accordingly, monthly values were computed from daily means
+only when at least `r Monthly_aggegation_N_lim` days were available.
+To create the daily and monthly climatological means, we averaged the data based on the day of year and calendar month, respectively.
 <!-- Seasonal -->
-For the seasons of the year trends, we grouped the mean daily values by the corresponding season (Winter: December - February, Spring: March - May, etc.).
+For the seasonal means we averaged the mean daily values in each season (Winter: December - February, Spring: March - May, etc.).
 <!-- Deseasonalization -->
-Finally, for each data set, we remove the natural occurring seasonal variation, by calculating the mean values for the appropriate time step, and subtracting the annual cycle from the actual data.
-<!-- SZA Deseasonalization -->
-To estimate Sun's SZA contribution to the SDR brightening, the 1-minute data were aggregate in $1^\circ$ SZA bins, separated for instances before and after the local noon, and then was deseasonalized as above.
-Thus, for each dataset we have obtained the relative departure of the seasonal mean.
+Finally, each data set was deseasonalized by subtracting the corresponding climatological annual cycle (daily or monthly) from the actual data.  
+To estimate SZA contribution to the SDR trends, the one-minute data were aggregated the data in $1^\circ$ SZA bins, separately for the morning and afternoon hours, which were then deseasonalized as mentioned above.
+
+a) for daily mean values we excluded days with less
+than 
+[This is not a good criterion! It must be seasonal and 2 hours of
+data are not enough for the calculation of daily means.\
+How about: "excluded days with more than 180 missing i-minute values
+during daylight hours".]{.comment-start id="6" author="Alkis Bais"
+date="2023-07-14T07:52:00Z"}180 valid data points[]{.comment-end
+id="6"},
 

### 2023-07-15 Article.Rmd
Commit 2023-07-15 20:14

@@
 
-For completeness, we will provide a brief overview of the CSid process, along with the site specific thresholds.
-As a reference clear sky SDR, we are using the radiation model by Haurwitz’s, adjusted for our site with a factor $a$, resulting by an iterative optimization process, as described by @Long2000 and @Reno2016.
+For completeness, we will provide below a brief overview of the
+[??]{.comment-start id="0" author="Alkis Bais"
+date="2023-07-14T07:52:00Z"}CSid algorithm[]{.comment-end id="0"}, along
+with the site specific thresholds.
+To calculate the reference clear sky
+SDR~CSref~ we used the [Not visible equation]{.comment-start id="2"
+author="A N" date="2023-07-15T19:23:34Z"}${}_{}$[[]{.comment-end
+id="2"}]{.insertion author="A N" date="2023-07-15T19:23:34Z"} derive by
+the radiation model of @Haurwitz1945, adjusted for our site with a
+factor $a$, resulted by an iterative optimization process, as described
+by @Long2000 and @Reno2016.
+
 The target of the optimization was the minimization of the function $f(a)$ (Eq.\ \ref{eq:minf}) and was accomplished with the computing function `stats::optimise` [@RCT2023], based on work by @Brent1973.

### 2023-07-15 Article.Rmd
Commit 2023-07-15 21:09

--- 2023-07-14 Article.Rmd
@@
 
-For completeness, we will provide a brief overview of the CSid process, along with the site specific thresholds.
-As a reference clear sky SDR, we are using the radiation model by Haurwitz’s, adjusted for our site with a factor $a$, resulting by an iterative optimization process, as described by @Long2000 and @Reno2016.
-The target of the optimization was the minimization of the function $f(a)$ (Eq.\ \ref{eq:minf}) and was accomplished with the computing function `stats::optimise` [@RCT2023], based on work by @Brent1973.
-The optimization and the selection of the clear sky reference model, was performed on SDR observations for the period 2016 – 2021.
-During the optimization, eight simple clear sky radiation models were tested (Daneshyar–Paltridge–Proctor, Kasten–Czeplak, Haurwitz, Berger–Duffie, Adnot–Bourges–Campana–Gicquel, Robledo-Soler, Kasten and Ineichen-Perez), with a wide range of factors.
-We found, that Haurwitz’s model, adjusted with the factor $a = `r signif(alpha,3)`$ has one of the lowest root mean squared error (RMSE), while correlates with large amount of our data.
-Thus, our clear sky reference is the Eq.\ \ref{eq:ahau}.
-The tried models are described in more details by @Reno2012 and evaluated by @Reno2016.
+For completeness, we will provide below a brief overview of the [....Csid
+term is referenced above....]{.comment-start id="0" author="A N"
+date="2023-07-15T20:00:52Z"}CSid algorithm[]{.comment-end id="0"}, along
+with the site specific thresholds.
+To calculate the reference clear sky
+$\text{SDR}_\text{CSref}$ we used the [....Not visible equation....]{.comment-start id="2"
+author="A N" date="2023-07-15T19:23:34Z"}${}_{}$[[]{.comment-end
+id="2"}]{.insertion author="A N" date="2023-07-15T19:23:34Z"} derive by
+the radiation model of @Haurwitz1945, adjusted for our site with a
+factor $a$, resulted by an iterative optimization process, as described
+by @Long2000 and @Reno2016.
+The target of
+the optimization was the minimization of the function $f(a)$ (Eq.\ \ref{eq:minf}) and
+was accomplished with the computing function [\...notation
+"library::function"....]{.comment-start id="2" author="A N"
+date="2023-07-15T20:08:20Z"}"stats::optimise"[]{.comment-end id="2"}
+of the R language [@RCT2023], with an implementation based on the work of @Brent1973. 
+The optimization and the selection of the clear sky reference model, was performed on SDR observations for the period 2016 - 2021.
+During the optimization, eight simple clear sky radiation models were tested (Daneshyar-Paltridge-Proctor, Kasten-Czeplak, Haurwitz, Berger-Duffie, Adnot-Bourges-Campana-Gicquel, Robledo-Soler, Kasten and Ineichen-Perez), with a wide range of factors.
+These models are described in more details by @Reno2012 and evaluated by @Reno2016.
+We found, that Haurwitz's model, adjusted with the factor $a = `r signif(alpha,3)`$ yields one of the lowest root mean squared errors (RMSE), [....rephrased....]{.comment-start id="3" author="A N" date="2023-07-15T20:32:50Z"}while the whole procedure, manages to characterize large amount of our data[]{.comment-end id="3"}.
+Thus, our clear sky reference is derived by the Eq.\ \ref{eq:ahau}.
 

### 2023-07-15 Article.Rmd
Commit 2023-07-15 22:14

@@
 \begin{equation}
-f(a)=\frac{1}{n}\sum_{i=1}^{n}( \text{SDR}_{\text{CSid},i} - a \times \text{SDR}_{\text{testCSref},i})^2 \label{eq:minf}
+f(a) = \frac{1}{n}\sum_{i=1}^{n} ( \text{SDR}_{\text{CSid},i} - a \times \text{SDR}_{\text{testCSref},i} )^2 \label{eq:minf}
 \end{equation}
+
+where: $n$ is the total number of daylight records, $\text{SDR}_{\text{CSid},i}$ are the records identified as clear sky by CSid, $a$ is a hypothetical adjustment factor, and $\text{SDR}_{\text{testCSref},i}$ is the tested model clear sky radiation.
 
@@
 \end{equation}
+
+where: $\text{SDR}_\text{CSref}$ is the reference clear sky SDR, in $w\;m^{-2}$.
 

### 2023-07-15 Article.Rmd
Commit 2023-07-15 23:09

@@
 
-where: $\text{SDR}_\text{CSref}$ is the reference clear sky SDR, in $w\;m^{-2}$.
+where: $\text{SDR}_\text{CSref}$ is the reference clear sky SDR, in $\text{w}\;\text{m}^{-2}$.
 
-The following criteria and thresholds were used to identify clear-sky conditions.
-Each criterion was applied for a running window of $`r nt`$ consecutive measurements/minutes, and the characterization is applied at the center value of the window.
-A data point is consider as under cloud-sky condition if it fails to pass any of the criteria, the rest data points are characterized as clear-sky.
+The criteria that were used to identify whether a measurement was taken
+under clear-sky conditions are presented below.
+A data point is flagged
+as "clear-sky" if all criteria are satisfied.
+Each criterion was applied
+for a running window of $`r nt`$ consecutive one-minute measurements, and
+the characterization is assigned to the central value of the window.
+Each parameter was calculated both from the measurements and the
+clear-sky model.
+The allowable range of variation is defined by the
+model-derived value of the parameter multiplied by a factor plus an
+offset. [.... .expanded the description ....]{.comment-start id="4"
+author="A N" date="2023-07-15T22:48:13Z"}The factors and offsets were
+determined empirically, by manual inspecting each filters performance on
+selected days, and adjusting them accordingly, during the iterative
+process.[[]{.comment-end id="4"}]{.insertion author="A N"
+date="2023-07-15T22:48:13Z"}
 

### 2023-07-20 Article.Rmd
Commit 2023-07-20 20:14

@@
 These models are described in more details by @Reno2012 and evaluated by @Reno2016.
-We found, that Haurwitz's model, adjusted with the factor $a = `r signif(alpha,3)`$ yields one of the lowest root mean squared errors (RMSE), [....rephrased....]{.comment-start id="3" author="A N" date="2023-07-15T20:32:50Z"}while the whole procedure, manages to characterize large amount of our data[]{.comment-end id="3"}.
+We found, that Haurwitz's model, adjusted with the factor $a = `r signif(alpha,3)`$ yields one of the lowest root mean squared errors (RMSE), [....rephrased....]{.comment-start id="3" author="A N" date="2023-07-15T20:32:50Z"}while the procedure, manages to characterize the majority of our data[]{.comment-end id="3"}.
 Thus, our clear sky reference is derived by the Eq.\ \ref{eq:ahau}.
@@
 
-where: $n$ is the total number of daylight records, $\text{SDR}_{\text{CSid},i}$ are the records identified as clear sky by CSid, $a$ is a hypothetical adjustment factor, and $\text{SDR}_{\text{testCSref},i}$ is the tested model clear sky radiation.
+where: $n$ is the total number of daylight records, $\text{SDR}_{\text{CSid},i}$ are the records identified as clear sky by CSid, $a$ is a hypothetical adjustment factor, and $\text{SDR}_{\text{testCSref},i}$ is the tested model of clear sky radiation.
 
@@
 
-where: $\text{SDR}_\text{CSref}$ is the reference clear sky SDR, in $\text{w}\;\text{m}^{-2}$.
+where: $\text{SDR}_\text{CSref}$ is the reference clear sky SDR, in $\text{w}\,\text{m}^{-2}$.
 
@@
 
-The mean of the measured value $\overline{\text{SDR}}_i = \text{mean}_{n=i-5}^{n=i+5}[\text{SDR}_{n}]$ must be inside an envelope based on the equivalent mean of the reference model $\overline{\text{SDR}}_\text{CSref} = \text{mean}_{n=i-5}^{n=i+5}[\text{SDR}_{\text{CSref},n}]$ (Eq.
+a) Mean of the measured $\overline{\text{SDR}}_i$ <!--= \text{mean}_{n=i-5}^{n=i+5}[\text{SDR}_{n}]$--> and running mean of the reference $\overline{\text{SDR}}_{\text{CSref},i}$ <!-- = \text{mean}_{n=i-5}^{n=i+5}[\text{SDR}_{\text{CSref},n}]$--> (Eq.
 \ref{eq:MeanVIP}).
@@
 
-The running max measured value $M_{\text{SDR},i} = \max_{n=i-5}^{n=i+5}[\text{SDR}_{n}]$, is compared to a similar constructed value from the reference $M_{CSi} = \max_{n=i-5}^{n=i+5}[\text{SDR}_{\text{CSref},n}]$ (Eq.\ \ref{eq:MaxVIP}).
+<!-- The running max measured value $M_{\text{SDR},i} = \max_{n=i-5}^{n=i+5}[\text{SDR}_{n}]$, is compared to a similar constructed value from the reference $M_{CSrefi} = \max_{n=i-5}^{n=i+5}[\text{SDR}_{\text{CSref},n}]$ (Eq.\ \ref{eq:MaxVIP}).
+-->
+
+b) Maximum measured value $M_{\text{}}$, and maximum reference $M_{\text{CSref}}$ (Eq.\ \ref{eq:MaxVIP}).
 
 \begin{equation}
-`r MaxVIP_fct` \times M_{CSi} - `r MaxVIP_off_low`
-< M_{\text{SDR},i} <
-`r MaxVIP_fct` \times M_{CSi} + `r MaxVIP_off_upp`
+`r MaxVIP_fct` \times M_{\text{CSref},i} - `r MaxVIP_off_low`
+< M_{\text{}i} <
+`r MaxVIP_fct` \times {\text{CSref},i} + `r MaxVIP_off_upp`
 \label{eq:MaxVIP}

### 2023-07-20 Article.Rmd
Commit 2023-07-20 21:09

@@
 
-where: $n$ is the total number of daylight records, $\text{SDR}_{\text{CSid},i}$ are the records identified as clear sky by CSid, $a$ is a hypothetical adjustment factor, and $\text{SDR}_{\text{testCSref},i}$ is the tested model of clear sky radiation.
+where: $n$ is the total number of daylight records, $\text{SDR}_{\text{CSid},i}$ are the records identified as clear sky by CSid, $a$ is a hypothetical adjustment factor, and $\text{SDR}_{\text{testCSref},i}$ is any of the tested clear sky radiation models.
 
@@
 
-a) Mean of the measured $\overline{\text{SDR}}_i$ <!--= \text{mean}_{n=i-5}^{n=i+5}[\text{SDR}_{n}]$--> and running mean of the reference $\overline{\text{SDR}}_{\text{CSref},i}$ <!-- = \text{mean}_{n=i-5}^{n=i+5}[\text{SDR}_{\text{CSref},n}]$--> (Eq.
-\ref{eq:MeanVIP}).
+<!--a) Mean of the measured $\overline{\text{SDR}}_i$ = \text{mean}_{n=i-5}^{n=i+5}[\text{SDR}_{n}]$ and running mean of the reference $\overline{\text{SDR}}_{\text{CSref},i}$ = \text{mean}_{n=i-5}^{n=i+5}[\text{SDR}_{\text{CSref},n}]$ (Eq.
+\ref{eq:MeanVIP}).-->
 
+a) Mean of the measured $\overline{\text{SDR}}_i$ (Eq. \ref{eq:MeanVIP}).
 \begin{equation}
@@
 
-b) Maximum measured value $M_{\text{}}$, and maximum reference $M_{\text{CSref}}$ (Eq.\ \ref{eq:MaxVIP}).
-
+b) Maximum measured value $M_{\text{}}$ (Eq.\ \ref{eq:MaxVIP}).
 \begin{equation}
@@
 <!-- ### Variability in irradiance by line length. -->
-
-The length $L$ of the sequence of line segments, connecting the points of the SDR time series for the measured values (Eq.\ \ref{eq:VILeq}), and similar for the reference $L_{CS}$, must be within the limits of Eq.
-\ref{eq:VILcr}.
-
+c) Length $L_i$ of the sequential line segments, connecting the points of the $`r nt`$ SDR values (Eq. \ref{eq:VILcr}).
 \begin{equation}
@@
 \end{equation}
-
 \begin{equation}
-`r MinVIL_fct` \times L_{CSi} - `r offVIL_dwl` < L_i < `r MaxVIL_fct` \times L_{CSi} + `r offVIL_upl`
+`r MinVIL_fct` \times L_{\text{CSref},i} - `r offVIL_dwl` < L_i < `r MaxVIL_fct` L_{\text{CSref},i} + `r offVIL_upl`
 \label{eq:VILcr}
 \end{equation}
+where: $t_i$ is the time each SDR measurement has been measured
 
@@
 <!-- Standard deviation of rate of change in irradiance. -->
-The standard deviation $\sigma$ of the slope ($s$) between sequential points in the time series, normalized by the average SDR during the time interval, must be below a threshold (Eq.\ \ref{eq:VCTcr}).
-
-<!--
-\begin{gather}
-s_i = \frac{\text{SDR}_{i+1} - \text{SDR}_{i}}{t_{i+1} - t_i}, \forall i \in \left \{ 1, 2, \ldots, n-1 \right \} \label{eq:VCT1} \\
+d) Standard deviation $\sigma_i$ of the slope ($s_i$) between the $`r nt`$ sequential points, normalized by the mean $\overline{\text{SDR}}_i$ (Eq.\ \ref{eq:VCT1}).
+<!-- \begin{gather}
 \bar{s} = \frac{1}{n-1} \sum_{i=1}^{n-1} s_i \label{eq:VCT2} \\
 \sigma_i = \frac {\sqrt{\frac{1}{n-1} \sum_{i=1}^{n-1} \left( s_i - \bar{s} \right)^2} } {\bar{G_i}} \label{eq:VCT3}
+\end{gather} -->
+\begin{gather}
+  \sigma_i = \frac {\sqrt{\frac{1}{n-1} \sum_{i=1}^{n-1} \left( s_i - \bar{s} \right)^2}} {\overline{\text{SDR}}_i} \label{eq:VCT1} \\
+  s_i = \frac{\text{SDR}_{i+1} - \text{SDR}_{i}}{t_{i+1} - t_i},\;\;   \bar{s} = \frac{1}{n-1} \sum_{i=1}^{n-1} s_i,\;\;\forall i \in \left \{ 1, 2, \ldots, n-1 \right \}\;\;
 \end{gather}
--->
-
+For this criterion, $\sigma_i$ should be below a certain threshold (Eq.\ \ref{eq:VCTcr}):
 \begin{equation}
-s_i = \frac{\text{SDR}_{i+1} - \text{SDR}_{i}}{t_{i+1} - t_i}, \forall i \in \left \{ 1, 2, \ldots, n-1 \right \},\;\;
-\bar{s} = \frac{1}{n-1} \sum_{i=1}^{n-1} s_i,\;\;
-\sigma_i = \frac {\sqrt{\frac{1}{n-1} \sum_{i=1}^{n-1} \left( s_i - \bar{s} \right)^2} } {\bar{G_i}} \label{eq:VCT3b}
-\end{equation}
-
-\begin{equation}
-\sigma_i < `r offVCT`
-\label{eq:VCTcr}
+  \sigma_i < `r offVCT` \label{eq:VCTcr}
 \end{equation}
@@
 
-The maximum difference $X$ between the change in measured irradiance and the change in clear sky irradiance over each measurement interval, must be lower than a threshold (Eq.\ \ref{eq:VSMcr}).
-
+e) Maximum difference $X_i$ between the change in measured irradiance and the change in clear sky irradiance over each measurement interval.
 \begin{gather}
+  X_i = \max{\left \{ \left | x_i - x_{\text{CSref},i} \right | \right \}} \label{eq:VSM3} \\
   x_i = \text{SDR}_{i+1} - \text{SDR}_{i} \forall i \in \left \{ 1, 2, \ldots, n-1 \right \} \label{eq:VSM1} \\
-x_{\text{CSref},i} = \text{SDR}_{\text{CSref},i+1} - \text{SDR}_{\text{CSref},i} \forall i \in \left \{ 1, 2, \ldots, n-1 \right \} \label{eq:VSM2} \\
-X_i = \max{\left \{ \left | x_i - x_{\text{CSref},i} \right | \right \}} \label{eq:VSM3}
+  x_{\text{CSref},i} = \text{SDR}_{\text{CSref},i+1} - \text{SDR}_{\text{CSref},i} \forall i \in \left \{ 1, 2, \ldots, n-1 \right \} \label{eq:VSM2}
 \end{gather}
-
+For this criterion, $X_i$ should be below a certain threshold (Eq.\ \ref{eq:VSMcr}):
 \begin{equation}
-X_i < `r offVSM`
-\label{eq:VSMcr}
+  X_i < `r offVSM` \label{eq:VSMcr}
 \end{equation}

### 2023-07-29 Article.Rmd
Commit 2023-07-21 15:10
Commit 2023-07-29 08:14

@@
 output: 
-  bookdown::word_document2: default
   bookdown::pdf_document2:
@@
     biblio-style:     apalike
+  bookdown::word_document2: default
   rticles::elsevier_article:
@@
 <!-- ### Variability in irradiance by line length. -->
-c) Length $L_i$ of the sequential line segments, connecting the points of the $`r nt`$ SDR values (Eq. \ref{eq:VILcr}).
+c) Length $L_i$ of the sequential line segments, connecting the points of the $`r nt`$ SDR values (Eq. \ref{eq:VILeq}).
 \begin{equation}
@@
 Subsequently, we made all measurements relative to the Total Solar Irradiance (TSI) at $1\,\text{au}$, in order to compensate for the Sun's intensity variability, using a homogenized time series of satellite TSI observations.
-The TSI data we use, comes from NOAA [@Coddington2005], and were interpolated to match with the time step of our measurements. 
+The TSI data we use are part of the "NOAA Climate Data Record of Total Solar Irradiance" dataset [@Coddington2005].Where the initial daily values, were interpolated to match with the time step of our measurements.
 The final dataset contains
 $`r SDR_N_All`$
-one-minute measurements, of which,  $`r signif(100 * SDR_N_Clear / SDR_N_All, 3)`\%$
-correspond to clear-sky conditions and 
+one-minute measurements, of which, 
+$`r signif(100 * SDR_N_Clear / SDR_N_All, 3)`\%$
+were identified as under clear-sky conditions and subsequently
 $`r signif(100 * SDR_N_Cloud / SDR_N_All, 3)`\%$
-to cloud-sky conditions.
+as under cloud-sky conditions.
 <!-- The TSI data we use, is a combination of data from NOAA [@Coddington2005] (for `r as.Date(TSIinfo[TSI_Source == "NOAA", Start])` - `r as.Date(TSIinfo[TSI_Source == "NOAA", End])`) and adjusted data from @LASP2023 (for `r as.Date(TSIinfo[TSI_Source != "NOAA", Start])` - `r as.Date(TSIinfo[TSI_Source != "NOAA", End])`).-->
@@
 To preserve the representativeness of the data we used the following criteria: 
-a) for daily mean values we exclude days with less than `r Daily_aggregation_N_lim` valid data points,
-b) accordingly, monthly values were computed from daily means
+<!-- a) for daily mean values we exclude days with less than `r Daily_aggregation_N_lim` valid data points, -->
+a) for the daily mean values we accept days with more than `r 100 * All_daily_ratio_lim`% of the daytime measurements, present and valid,
+b) monthly values were computed from daily means
 only when at least `r Monthly_aggegation_N_lim` days were available.
@@
 Finally, each data set was deseasonalized by subtracting the corresponding climatological annual cycle (daily or monthly) from the actual data.
-To estimate SZA contribution to the SDR trends, the one-minute data were aggregated the data in $1^\circ$ SZA bins, separately for the morning and afternoon hours, which were then deseasonalized as mentioned above.
-
-a) for daily mean values we excluded days with less
-than 
-[This is not a good criterion! It must be seasonal and 2 hours of
-data are not enough for the calculation of daily means.\
-How about: "excluded days with more than 180 missing i-minute values
-during daylight hours".]{.comment-start id="6" author="Alkis Bais"
-date="2023-07-14T07:52:00Z"}180 valid data points[]{.comment-end
-id="6"},
+To estimate SZA contribution to the SDR trends, the one-minute data were aggregated in $1^\circ$ SZA bins, separately for the morning and afternoon hours, and then were deseasonalized as mentioned above.
 

### 2023-08-29 Article.Rmd
Commit 2023-08-29 17:09
Commit 2023-08-29 18:14

@@
 names(TSI) <- c("Trend [%/year]", "Trend p-value", "Sky condition", "Trend statistical signif. [%]")
-TSIskopeAll <- TSI[`Sky condition` == "All sky cond."]
+TSIslopeAll <- TSI[`Sky condition` == "All sky cond."]
 ```
@@
 Other studies for the European region are reporting a change of the SDR slope, around the 1980, few years before the start of our records [@Wild2021; @Yuan2021; @Ohmura2009].
+It is interesting to note, that for the observation period, the trend of the TSI is
+$`r format(TSIslopeAll[1,1], scientific = T, digits = 2)`\%/year$.
 
@@
 On all three plots on Figure\ \ref{fig:cumsum}, we observe a step downward part at the start of the period until about 2000, an almost steady part until about 2016 for All-sky and Clear-sky conditions (Figures\ \ref{fig:cumsum-1}, \ref{fig:cumsum-2}), and until 2009 for Cloud-sky conditions (Figure\ \ref{fig:cumsum-3}), and finally, a steep upward part until the present.
-For a more uniform trend, we expect the cumulative sum plot of the anomaly, to have a symmetric 'V' shape.
-In our case, for all sky conditions we observe a period where the cumulative remain relative steady.
-This indicates, that the anomaly was evenly distributed around the climatological mean.
-Another distinct feature of the cumulative plots, is the difference, between the sky conditions, of the start of the sums sudden increase, for the latter part of the period.
+For a uniform trend, we expected the cumulative sum plot of the anomaly to have a symmetric 'V' shape.
+This would indicate, that the anomaly is evenly distributed around the climatological mean.
+In our case, for all sky conditions we observe periods where the cumulative sum remains relative steady, roughly in the middle of the study dataset.
+Another distinct feature of the cumulative plots, is the difference, between the Cloud-sky and the rest of the sky conditions, where they display an supplementary behavior.
+
+<!--
 This can make us to suspected that clear-Sky conditions SDR and cloud-sky SDR is affected by different mechanisms.
+The similarity of the All-sky and Clear-sky conditions, can make us to suspected that on both sky conditions the attenuation of SDR has .
+-->
 

### 2023-08-30 Article.Rmd
Commit 2023-08-30 18:14

@@
 
-.... cite paper about low elevation SDR .....
-
-The relative position of the Sun in the sky, seem to be a major drive of the attenuation of SDR, as it is affect
-
-... Aerosols and clouds ...
-
-We group the anomaly data in $1^\circ$ SZA bins, and calculated the overall trend for each bin
-
-Analyzing the long term trends of SDR to bins of SZA, we can see the contribution of the geometry and time in a diurnal level (Figures\ \ref{fig:szatrends}a and \ref{fig:szatrends}b).
-Although there is a seasonal distribution of SZA than in not presented here.
+The relative position of the Sun in the sky, is a major drive of the attenuation of SDR reaching the ground, as it is affected by the radiation path in the atmosphere and possible human activities in urban environment emitting aerosols [@Wang2021].
+In order to estimate the contribution of the SZA, we group the anomaly data in $1^\circ$ SZA bins, and calculated the overall trend for each bin (Figures\ \ref{fig:szatrends}a and \ref{fig:szatrends}b).
+Although there is a seasonal change of SZA limits than is not represented here, the brightening effect of SDR is stronger for All-sky and Clear-sky condition for low Sun elevation angles.
 

### 2023-08-30 Article.Rmd
Commit 2023-08-30 19:10

@@
 The relative position of the Sun in the sky, is a major drive of the attenuation of SDR reaching the ground, as it is affected by the radiation path in the atmosphere and possible human activities in urban environment emitting aerosols [@Wang2021].
-In order to estimate the contribution of the SZA, we group the anomaly data in $1^\circ$ SZA bins, and calculated the overall trend for each bin (Figures\ \ref{fig:szatrends}a and \ref{fig:szatrends}b).
-Although there is a seasonal change of SZA limits than is not represented here, the brightening effect of SDR is stronger for All-sky and Clear-sky condition for low Sun elevation angles.
+In order to estimate the contribution of the SZA, we group the anomaly data in $1^\circ$ SZA bins, and calculated the overall trend for each bin before noon and after noon (Figure\ \ref{fig:szatrends}).
+Although there is a seasonal change of SZA limits than is not represented here, the brightening effect of SDR is stronger for All-sky and Clear-sky condition for low Sun elevation angles (Figures\ \ref{fig:szatrends}a and \ref{fig:szatrends}b).
+In the case of Cloud-sky condition (Figure\ \ref{fig:szatrends}c), there is no strong correlation between SDR brightening and SZA, except for some low Sun angles, that occur only for a small part of the year, and thus we can not draw more conclusions from this analysis for the over trend.
 
-```{r szatrends, echo = FALSE, fig.cap='Distribution of the SDR\'s long term trends by SZA.', fig.subcap=c('All-sky conditions.', 'Clear-sky conditions.', 'Cloud-sky conditions.'), out.width='.35\\linewidth', out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3}
+```{r szatrends, echo = FALSE, fig.cap='Distribution of the SDR\'s long term trends by SZA before and after noon.
+The solid data points represent values with an acceptable statistical significance ($p<0.005$).', fig.subcap=c('All-sky conditions.', 'Clear-sky conditions.', 'Cloud-sky conditions.'), out.width='.35\\linewidth', out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3}
 knitr::include_graphics("./images/SzaTrends-13.png")

### 2023-08-31 Article.Rmd
Commit 2023-08-31 19:10

@@
 LT$`Trend statistical signif. [%]` <- NULL
+## numeric presision
+LT$`Trend [%/year]` <- signif(LT$`Trend [%/year]`, 3)
+LT$`Trend p-value`  <- signif(LT$`Trend p-value`,  3)
 
@@
 
-Similar we have produced the deseasonalized trend for different sky conditions for each season of the year, using the corresponding mean monthly values.
-This can give us a better understanding of the annual variability of the trends.
+Similar to the long term trends discussed above, we have produced the deseasonalized trend for the different sky conditions, for each season of the year, using the corresponding mean monthly values.
+
+This can give us a better understanding of the annual cycle of the trends.
 
@@
 names(pp) <- c("Trend [%/year]", "Trend p-value", "Sky condition", "Season", "Trend statistical signif. [%]")
+## order by season
+target    <- c("Winter", "Spring", "Summer", "Autumn")
+new_order <- sapply(target, function(x,pp){which(pp$Season == x)}, pp=pp)
+pp        <- pp[new_order,]
+## numeric precision
+pp$`Trend [%/year]` <- signif(pp$`Trend [%/year]`, 3)
+pp$`Trend p-value`  <- signif(pp$`Trend p-value`,  3)
+
+rownames(pp) <- NULL
 
@@
 pander(pp,
-       caption = "\\label{tab:trendseasontable}Trends of daily means by sky conditions for the seasons of the year.")
+       caption = "\\label{tab:trendseasontable}SDR trends of monthly anomaly for each season of the year. The vertical scale is common for all subplots, for easier comparison.")
 ```

### 2023-08-31 Article.Rmd
Commit 2023-08-31 20:14

@@
 rownames(pp)          <- NULL
-names(pp) <- c("Trend [%/year]", "Trend p-value", "Sky condition", "Season", "Trend statistical signif. [%]")
+names(pp) <- c("Trend [%/year]", "Trend p-value", "Sky condition", "Season", "Statistical signif. [%]")
 ## order by season
@@
 pp$`Trend p-value`           <- signif(pp$`Trend p-value`,  3)
+pp$`Statistical signif. [%]` <- signif(pp$`Statistical signif. [%]`,  3)
 

### 2023-09-02 Article.Rmd
Commit 2023-09-02 17:10
Commit 2023-09-02 18:14

@@
 
+While in Spring and Summer the brightening of the SDR seem to be very low, although the statistical significance is now adequate for all cases, there are significant changes in Winter and Autumn
+(Figure\ \ref{fig:seasonalALL} and Table\ \ref{tab:trendseasontable}).
+Especially for winter and
+
 This can give us a better understanding of the annual cycle of the trends.
 
-```{r seasonalALL, echo=F, fig.cap="Trends by season for all condition. Displaying monthly means of daily means.", out.width="100%"}
+```{r seasonalALL, echo=F, fig.cap="SDR trends by season of the year for the three sky conditions.
+Each point is a monthly mean calculated from daily means.
+The vertical scale is common for all subplots, for easier comparison.", out.width="100%"}
 knitr::include_graphics("./images/SeasonalTrendsTogether-1.png")
@@
 new_order <- sapply(target, function(x,pp){which(pp$Season == x)}, pp=pp)
+which(pp)
 pp        <- pp[new_order,]
@@
 pander(pp,
-       caption = "\\label{tab:trendseasontable}SDR trends of monthly anomaly for each season of the year. The vertical scale is common for all subplots, for easier comparison.")
+       justify = c('center', 'center', 'right', 'right', 'center'),
+       split.table = Inf,
+       caption = "\\label{tab:trendseasontable}SDR trends of monthly anomaly for each season of the year.")
 ```

### 2023-09-02 Article.Rmd
Commit 2023-09-02 19:09

@@
 
-Similar to the long term trends discussed above, we have produced the deseasonalized trend for the different sky conditions, for each season of the year, using the corresponding mean monthly values.
-
-While in Spring and Summer the brightening of the SDR seem to be very low, although the statistical significance is now adequate for all cases, there are significant changes in Winter and Autumn
+Similar to the long term trends discussed above, we have produced the deseasonalized trend for the different sky conditions, for each season of the year, using the corresponding mean monthly values
 (Figure\ \ref{fig:seasonalALL} and Table\ \ref{tab:trendseasontable}).
-Especially for winter and
-
-This can give us a better understanding of the annual cycle of the trends.
+For All-sky condition the brightening of SDR is very significant for the Winter, close to the long term trend for the Autumn, and almost four times lower than the long term trend for Sprint and Summer, indicating a relation to the cloud pattern of each season.
+Moreover when examining the Clear- and Cloud-sky conditions the trends between these condition are are almost complementary to each other for Winter and Autumn respectively.
+On the latter cases during Spring and Summer the statistical significance along with the actual trend are very low for a meaningful comparison.
 
@@
 target    <- c("Winter", "Spring", "Summer", "Autumn")
-new_order <- sapply(target, function(x,pp){which(pp$Season == x)}, pp=pp)
-which(pp)
+new_order <- sapply(target, function(x,pp){which(pp$Season == x)}, pp = pp)
 pp        <- pp[new_order,]
+## put all sky first
+A  <- which(pp$`Sky condition` == "All sky cond.")
+B  <- which(pp$`Sky condition` != "All sky cond.")
+pp <- pp[c(A,B),]
 ## numeric precision
@@
 
+## arrange columns
+pp <- pp[,c(
+      "Sky condition",
+      "Season",
+      "Trend [%/year]", 
+      "Trend p-value", 
+      "Statistical signif. [%]"
+      )]
+
 rownames(pp) <- NULL
-
 # panderOptions("table.alignment.default", "left")
@@
 pander(pp,
-       justify = c('center', 'center', 'right', 'right', 'center'),
+       justify = c('right', 'right', 'right', 'right', 'center'),
        split.table = Inf,
@@
 ## Consistency of the trends
-
-A method to evaluate changes in the long term trend is to use the cumulative sum of the variable.
-
-Figure\ \ref{fig:cumsum}
-
-... different in aggregation ...
 -->
@@
 
+Brightening is driven during Winter and Summer
+
 .....

### 2023-09-04 Article.Rmd
Commit 2023-09-04 11:13
Commit 2023-09-04 12:12

@@
 
-Our result for all-sky condition ($`r alltrend`\%/year$), reaffirm the previous results of @Bais2013 for the site.
-The increase of this trend, shows that the phenomenon and probably the causes have been amplified.
+We demonstrate that in the period `r paste(year(min(ALL_1_daily_DESEAS$Date)), " - ", year(max(ALL_1_daily_DESEAS$Date)))`, there is brightening trend of SDR for All-sky conditions ($`r alltrend`\%/year$).
+On a previous study [@Bais2013] for the period 1993 - 2011 the trend was found to be $0.33\%/year$.
+The increase of this trend, indicates that the phenomenon and probably the causes, still persists.
+Accordingly, we found a similar trend for Clear-sky conditions ($`r cleartrend`\%/year$) that further supports that the brightening can be independent from the cloud occurrences.
+We have calculate the effect of TSI variability to be negligible, but unfortunately there is no available data for this period for the Aerosols, in order to estimate their effect on SDR, which is the most probable main cause.
+
+The dimming effect of SDR under Cloud-sky condition ($`r cloudtrend`\%/year$), shows that clouds can also attenuate the amount of SDR reaching the ground, and changes on their occurrences can effect the net energy balance.
+Although, we have no adequate data to investigate long term changes of cloud formation in the region, in order to infer if the SDR trend we observe under Cloud-sky is biased by those changes.
 
 Brightening is driven during Winter and Summer
-
-.....
-Also, the trend of $`r cloudtrend`\%/year$ for cloud-sky condition indicate the major part that clouds play on the
-.....
-
-.......
 

### 2023-09-04 DHI_GHI_1_longterm_trends.R

@@ -575,7 +575,7 @@ for (DBn in dbs) {
         cat("\n \n \\footnotesize \n ")
         cat(
             pander(
-                DB[ order(abs(DB[[avar]]), decreasing = T )[1:5] , ..wca ],
+                DB[order(abs(DB[[avar]]), decreasing = T )[1:5] , ..wca ],
                 cap = "Extreme anomaly values"
             )
         )
@@ -592,9 +592,9 @@ for (DBn in dbs) {
 vars        <- c("GLB_att_des")
 
 ## Monthly aggregation
-dbs         <- c(  "ALL_1_D_bySeason_DESEAS",
-                   "CLEAR_1_D_bySeason_DESEAS",
-                   "CLOUD_1_D_bySeason_DESEAS")
+dbs         <- c("ALL_1_D_bySeason_DESEAS",
+                 "CLEAR_1_D_bySeason_DESEAS",
+                 "CLOUD_1_D_bySeason_DESEAS")
 
 Seasons     <- c("Winter", "Spring", "Summer", "Autumn")
 
@@ -687,7 +687,7 @@ for (avar in vars) {
             legend("bottom", lty = 1, bty = "n", lwd = 2, cex = 0.8,
                    paste("Trend: ",
                          if (fit[2] > 0) '+' else '-',
-                         signif(abs(fit[2]), 3),
+                         signif(abs(fit[2]), 2),
                          "% / year")
             )
         }

### 2023-09-05 Article.Rmd
Commit 2023-09-05 12:14

@@
 -->
-....
 
+<!--
 more about trend in relation to other trends
 Aerosols and clouds
-
-......
+-->
 
@@
 Changes of the SDR can be related to changes on Earth's energy budget, the mechanisms of climate change, and water and carbon cycle [@Wild2009].
-Can also affect, solar and agricultural production, and living organisms.
-Studies of SDR variability, have identified some distinct SDR trends on different regions of the world and for different time periods.
+Can also affect, solar and agricultural production, and all living organisms.
+Studies of SDR variability, have identified some distinct SDR trends on different regions of the world on different time periods.
 The term 'brightening' is generally used to describe periods of positive SDR trend, and 'dimming' for negative trend.
@@
 
-....
-
+<!--
 more about trend in relation to other trends
 Aerosols and clouds
-
-......
+-->
 
@@
 On a previous study [@Bais2013] for the period 1993 - 2011 the trend was found to be $0.33\%/year$.
-The increase of this trend, indicates that the phenomenon and probably the causes, still persists.
+The increase of this trend, indicates that the phenomenon and probably the causes, are still occurring.
 Accordingly, we found a similar trend for Clear-sky conditions ($`r cleartrend`\%/year$) that further supports that the brightening can be independent from the cloud occurrences. 
 We have calculate the effect of TSI variability to be negligible, but unfortunately there is no available data for this period for the Aerosols, in order to estimate their effect on SDR, which is the most probable main cause.
+The attenuation of SDR by Aerosols over Europe have been proposed as major factor by @Wild2021.
+The dimming effect of SDR under Cloud-sky condition ($`r cloudtrend`\%/year$), shows that clouds can also attenuate the amount of SDR reaching the ground, and changes on their occurrences can effect the net energy balance.
+Because we have no adequate data to investigate long term changes of cloud formation in the region, we can not infer if the SDR trend we observe under Cloud-sky is biased by changes in clouds and local climate.
+The observed brightening of SDR over Thessaloniki, is correlated with the SZA (higher SZA, have stronger brightening trends).
+And also, with the seasons of the year, where Winter and Autumn show a significant strong trend, in contrast to Spring and Summer.
 
-The dimming effect of SDR under Cloud-sky condition ($`r cloudtrend`\%/year$), shows that clouds can also attenuate the amount of SDR reaching the ground, and changes on their occurrences can effect the net energy balance.
-Although, we have no adequate data to investigate long term changes of cloud formation in the region, in order to infer if the SDR trend we observe under Cloud-sky is biased by those changes.
-
-Brightening is driven during Winter and Summer
-
-About:
-
-- long terms trends
-- seasonal trends
-- effect of SZA
-- effect of clouds
-
-- Aerosols
-
+<!--
 similar results with Evidence for Clear‐Sky Dimming and Brightening in Central Europe_Wild2021.pdf
 include in discussion
+-->
 
--------------------------------------------------
-                   **END**
--------------------------------------------------
-\newpage
+# References

### 2023-09-05 Article.Rmd
Commit 2023-09-05 13:09

@@
 
-There is a consensus, among researchers, that major factors of SDR attenuation is the interaction with atmospheric aerosols and clouds. 
+There is a consensus, among researchers, that the major factors of SDR attenuation is the interaction of Sun radiation with atmospheric aerosols and clouds. 
 Those interactions, among other factors, have been analysed with models [@Li2016; @Samset2018], showing the existence of feedback mechanisms between the two.
@@
 
-Due to the variability of the phenomenon, and its contributing factors, there is a need to further investigate and monitor SDR, in different sites, to estimate its magnitude, and its relation to the local conditions.
+Due to the variability of the phenomenon, and its contributing factors, there is a constant need to investigate and monitor SDR, in different sites, to estimate its magnitude, and its relation to the local conditions.
 In this study, we examine the trends of SDR, with ground-based measurements at Thessaloniki, Greece for the period `r year(min(ALL_1_daily_DESEAS$Date))` to `r year(max(ALL_1_daily_DESEAS$Date))`, as derived from a CM-21 pyranometer.
@@
 However, the DNI data were used as auxiliary data, in the clear sky identification algorithm (CSid), which is discussed later, for the selection of the appropriate thresholds.
-It is noted that despite the capability of the CSid algorithm to use the DNI as a characterization parameter, we haven't used it here, to avoid any selection bias due to unequal length of the two datasets.
-There are three distinct steps in the creation of the dataset analysed here:
+It is noted that despite the capability of the CSid algorithm to use the DNI as a characterization parameter, we haven't utilized it here, to avoid any selection bias, due to unequal length of the two datasets.
+There are four distinct steps in the creation of the dataset analysed here:
 a)\ the acquisition of radiation measurements from the sensors,
 b)\ the data quality check,
-c)\ the identification of "clear sky" conditions from the radiometric data, and d)\ the aggregation of data and trend analysis.
+c)\ the identification of "clear sky" conditions from the radiometric data, and 
+d)\ the aggregation of data and trend analysis.
 
@@
 The mean and the standard deviation of these samples are recorded every minute.
-The measurements are corrected for the zero offset (dark signal in volts).
+The measurements are corrected for the zero offset ("dark signal" in volts).
 The "dark signal" is calculated by averaging all measurements recorded for a period of 
-$3\,\text{h}$ 
-before (morning) or after (evening), when the Sun reaches an elevation angle of
+$3\,\text{h}$, 
+before (morning) or after (evening) the Sun reaches an elevation angle of
 $-10^\circ$.
-The signal is converted to irradiance using a ramped value of the instrument’s sensitivity between the calibrations.
+The signal is converted to irradiance using a ramped value of the instrument’s sensitivity between calibrations.
 
@@
 <!-- c) Clear sky identification -->
-In order to be able to estimate the effect of the clouds on the long term variability of SDR, we created three datasets, by characterizing each one-minute measurement with a corresponding sky condition (i.e., all-sky, clear-sky and cloudy-sky).
-To identify the clear-sky conditions we used the method proposed by
+In order to be able to estimate the effect of the sky condition on the long term variability of SDR, we created three datasets, by characterizing each one-minute measurement with a corresponding sky condition (i.e., all-sky, clear-sky and cloudy-sky).
+To identify the clear-sky conditions we used a method proposed by
 @Long2000
@@
 <!-- CS subjectivity -->
-We have to note, that the definition of what constitutes as clear or cloudy sky, has some subjectivity, in this method of characterization.
-As a result, the details of the definition is site specific, it relies on a combination of thresholds and comparisons with ideal actinometric models and statistical analysis on different signal metrics.
-The CSid algorithm was calibrated with the main focus to identify the presence of clouds on the sky dome, although few marginal cases exist that have been identified as false positive or false negative, but cannot affect the results of the study.
+We have to note, that the definition of what constitutes as clear or cloudy sky, has some subjectivity, in any method of characterization.
+As a result, the details of the definition are site specific, it rely on a combination of thresholds and comparisons with ideal actinometric models and statistical analysis on different signal metrics.
+The CSid algorithm was calibrated with the main focus, to identify the presence of clouds on the sky dome.
+Although the fine tuning of the procedure, few marginal cases exist, that have been identified manually as false positive or false negative but cannot affect the final results of the study.
 
@@
 
-For completeness, we will provide below a brief overview of the [....Csid
-term is referenced above....]{.comment-start id="0" author="A N"
-date="2023-07-15T20:00:52Z"}CSid algorithm[]{.comment-end id="0"}, along
-with the site specific thresholds.
+For completeness, we will provide below a brief overview of the 
+<!--
+[....Csid term is referenced above....]{.comment-start id="0" author="A N"
+date="2023-07-15T20:00:52Z"} 
+-->
+clear sky identification algorithm (CSid),
+<!--
+[]{.comment-end id="0"} 
+-->
+along with the site specific thresholds.
 To calculate the reference clear sky
-$\text{SDR}_\text{CSref}$ we used the [....Not visible equation....]{.comment-start id="2"
-author="A N" date="2023-07-15T19:23:34Z"}${}_{}$[[]{.comment-end
-id="2"}]{.insertion author="A N" date="2023-07-15T19:23:34Z"} derive by
+$\text{SDR}_\text{CSref}$ we used the $\text{SDR}_\text{Haurwitz}$ derived by
 the radiation model of @Haurwitz1945, adjusted for our site with a
-factor $a$, resulted by an iterative optimization process, as described
+factor $a$ (Eq.\ \ref{eq:ahau}), resulted by an iterative optimization process, as described
 by @Long2000 and @Reno2016.
 The target of
-the optimization was the minimization of the function $f(a)$ (Eq.\ \ref{eq:minf}) and
-was accomplished with the computing function [\...notation
-"library::function"....]{.comment-start id="2" author="A N"
-date="2023-07-15T20:08:20Z"}"stats::optimise"[]{.comment-end id="2"}
-of the R language [@RCT2023], with an implementation based on the work of @Brent1973. 
+the optimization was the minimization of a function $f(a)$ (Eq.\ \ref{eq:minf}) and
+was accomplished with the algorithmic function "optimise", which is an implementation based on the work of @Brent1973, from the library "stats" of the R programming language [@RCT2023]. 
 The optimization and the selection of the clear sky reference model, was performed on SDR observations for the period 2016 - 2021.
@@
 These models are described in more details by @Reno2012 and evaluated by @Reno2016.
-We found, that Haurwitz's model, adjusted with the factor $a = `r signif(alpha,3)`$ yields one of the lowest root mean squared errors (RMSE), [....rephrased....]{.comment-start id="3" author="A N" date="2023-07-15T20:32:50Z"}while the procedure, manages to characterize the majority of our data[]{.comment-end id="3"}.
+We found, that Haurwitz's model, adjusted with the factor $a = `r signif(alpha,3)`$ yields one of the lowest root mean squared errors (RMSE), 
+<!--[....rephrased....]{.comment-start id="3" author="A N" date="2023-07-15T20:32:50Z"}-->
+while the procedure, manages to characterize the majority of the data.
+<!--[]{.comment-end id="3"}--> 
 Thus, our clear sky reference is derived by the Eq.\ \ref{eq:ahau}.

### 2023-09-05 Article.Rmd
Commit 2023-09-05 14:14

@@
 We have to note, that the definition of what constitutes as clear or cloudy sky, has some subjectivity, in any method of characterization.
-As a result, the details of the definition are site specific, it rely on a combination of thresholds and comparisons with ideal actinometric models and statistical analysis on different signal metrics.
+As a result, the details of the definition are site specific, it relies on a combination of thresholds and comparisons with ideal actinometric models and statistical analysis on different signal metrics.
 The CSid algorithm was calibrated with the main focus, to identify the presence of clouds on the sky dome.
-Although the fine tuning of the procedure, few marginal cases exist, that have been identified manually as false positive or false negative but cannot affect the final results of the study.
+Although the fine-tuning of the procedure, few marginal cases exist, that have been identified manually as false positive or false negative but cannot affect the final results of the study.
 
@@
 A data point is flagged
-as "clear-sky" if all criteria are satisfied.
+as "clear-sky" if all criteria are satisfied, otherwise it is considered to be "cloud-sky".
 Each criterion was applied
@@
 the characterization is assigned to the central value of the window.
-Each parameter was calculated both from the measurements and the
-clear-sky model.
+Each parameter, was calculated both from the observations and the
+reference clear sky model, for each comparison.
 The allowable range of variation is defined by the
 model-derived value of the parameter multiplied by a factor plus an
-offset. [.... .expanded the description ....]{.comment-start id="4"
-author="A N" date="2023-07-15T22:48:13Z"}The factors and offsets were
+offset. 
+<!--[.... .expanded the description ....]{.comment-start id="4"
+author="A N" date="2023-07-15T22:48:13Z"}
+-->
+The factors and the offsets were
 determined empirically, by manual inspecting each filters performance on
-selected days, and adjusting them accordingly, during the iterative
-process.[[]{.comment-end id="4"}]{.insertion author="A N"
+selected days, and adjusting them accordingly, during an iterative
+process.
+<!--[[]{.comment-end id="4"}]{.insertion author="A N"
 date="2023-07-15T22:48:13Z"}
+-->
 
@@
 `r 1 - CS_ref_rm_VIP_RelLow` \times \overline{\text{SDR}}_{\text{CSref},i} - `r CS_ref_rm_VIP_LowOff`
-< \overline{SDR}_i <
+< \overline{\text{SDR}}_i <
 `r 1 + CS_ref_rm_VIP_RelUpp` \times \overline{\text{SDR}}_{\text{CSref},i} + `r CS_ref_rm_VIP_UppOff`
@@
 < M_{\text{}i} <
-`r MaxVIP_fct` \times {\text{CSref},i} + `r MaxVIP_off_upp`
+`r MaxVIP_fct` \times M_{\text{CSref},i} + `r MaxVIP_off_upp`
 \label{eq:MaxVIP}
@@
 \begin{equation}
-`r MinVIL_fct` \times L_{\text{CSref},i} - `r offVIL_dwl` < L_i < `r MaxVIL_fct` L_{\text{CSref},i} + `r offVIL_upl`
+`r MinVIL_fct` \times L_{\text{CSref},i} - `r offVIL_dwl` < L_i < `r MaxVIL_fct` \times L_{\text{CSref},i} + `r offVIL_upl`
 \label{eq:VILcr}
@@
 <!-- TSI -->
-Subsequently, we made all measurements relative to the Total Solar Irradiance (TSI) at $1\,\text{au}$, in order to compensate for the Sun's intensity variability, using a homogenized time series of satellite TSI observations.
-The TSI data we use are part of the "NOAA Climate Data Record of Total Solar Irradiance" dataset [@Coddington2005].Where the initial daily values, were interpolated to match with the time step of our measurements.
+Subsequently, we made all measurements relative to the Total Solar Irradiance (TSI) at $1\,\text{au}$, in order to compensate for the Sun's intensity variability, using a <!--homogenized--> time series of satellite TSI observations.
+The TSI data we use are part of the "NOAA Climate Data Record of Total Solar Irradiance" dataset [@Coddington2005]. 
+Where the initial daily values, were interpolated to match with the time step of our measurements. 
 The final dataset contains
@@
 <!-- aggregation -->
-In order to investigate the SDR trends, we implemented an appropriate aggregation scheme to the 1-minute data to derive series in coarser time scales.
+In order to investigate the SDR trends, we implemented an appropriate aggregation scheme to the 1-minute data to derive a series in coarser time scale.
 To preserve the representativeness of the data we used the following criteria: 
@@
 
-We calculated the SDR trends, from the mean daily values of the departure from the seasonal values, for the whole time series for the three sky conditions (Table\ \ref{tab:trendtable}).
+We calculated the SDR trends, from the mean daily values of the departure from the seasonal values, for the three sky conditions (Table\ \ref{tab:trendtable}).
 We present only the plot for the trend under all-sky conditions (Figure\ \ref{fig:trendALL}), as the plots for clear-sky condition and cloud-sky condition, are very similar.
@@
 It is interesting to note, that for the observation period, the trend of the TSI is
-$`r format(TSIslopeAll[1,1], scientific = T, digits = 2)`\%/year$.
+$`r format(TSIslopeAll[1,1], scientific = T, digits = 2)`\%/year$, and thus we can not attribute any major effect on SDR trend to Sun variability.
 
@@
 
-Although, using the cumulative sums plots we can reveal some structure within the trends and some difference between the sky conditions [@Regier2019]. 
-On all three plots on Figure\ \ref{fig:cumsum}, we observe a step downward part at the start of the period until about 2000, an almost steady part until about 2016 for All-sky and Clear-sky conditions (Figures\ \ref{fig:cumsum-1}, \ref{fig:cumsum-2}), and until 2009 for Cloud-sky conditions (Figure\ \ref{fig:cumsum-3}), and finally, a steep upward part until the present.
+Although the trends are quite constant, using plots of the cumulative sums we can reveal some structure within the trends and some differences between the sky conditions [@Regier2019]. 
+On all three plots on Figure\ \ref{fig:cumsum}, we observe a downward part at the start of the period until about 2000, an almost steady part until about 2016 for All-sky and Clear-sky conditions (Figures\ \ref{fig:cumsum-1}, \ref{fig:cumsum-2}), and until 2009 for Cloud-sky conditions (Figure\ \ref{fig:cumsum-3}), and finally, a steep upward part until the present.
 For a uniform trend, we expected the cumulative sum plot of the anomaly to have a symmetric 'V' shape.

### 2023-09-05 Article.Rmd
Commit 2023-09-05 16:09

@@
 Although the trends are quite constant, using plots of the cumulative sums we can reveal some structure within the trends and some differences between the sky conditions [@Regier2019]. 
-On all three plots on Figure\ \ref{fig:cumsum}, we observe a downward part at the start of the period until about 2000, an almost steady part until about 2016 for All-sky and Clear-sky conditions (Figures\ \ref{fig:cumsum-1}, \ref{fig:cumsum-2}), and until 2009 for Cloud-sky conditions (Figure\ \ref{fig:cumsum-3}), and finally, a steep upward part until the present.
+On all three plots on Figure\ \ref{fig:cumsum}, we observe a downward part at the start of the period until about 2000, an almost steady part until about 2016 for All-sky and Clear-sky conditions (Figures\ \ref{fig:cumsum-1}, \ref{fig:cumsum-2}) and until 2009 for Cloud-sky conditions (Figure\ \ref{fig:cumsum-3}), and finally, a steep upward part until the present.
 For a uniform trend, we expected the cumulative sum plot of the anomaly to have a symmetric 'V' shape.
 This would indicate, that the anomaly is evenly distributed around the climatological mean.
-In our case, for all sky conditions we observe periods where the cumulative sum remains relative steady, roughly in the middle of the study dataset.
+In our case, for all sky conditions, we observe periods where the cumulative sum remains relative steady, roughly in the middle of the study dataset.
 Another distinct feature of the cumulative plots, is the difference, between the Cloud-sky and the rest of the sky conditions, where they display an supplementary behavior. 
+The causes for these discrepancies haven't been investigated yet.
 
@@
 
-```{r cumsum, echo = FALSE, fig.cap='Running cumulative sum plots of the daily SDR for three different sky conditions', fig.subcap=c('All-sky conditions', 'Clear-sky conditions', 'Cloud-sky conditions'), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3}
+```{r cumsum, echo = FALSE, fig.cap='Running cumulative sum plots of the daily SDR for three different sky conditions.', fig.subcap=c('All-sky conditions.', 'Clear-sky conditions.', 'Cloud-sky conditions.'), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3}
 knitr::include_graphics(c(
@@
 In order to estimate the contribution of the SZA, we group the anomaly data in $1^\circ$ SZA bins, and calculated the overall trend for each bin before noon and after noon (Figure\ \ref{fig:szatrends}).
-Although there is a seasonal change of SZA limits than is not represented here, the brightening effect of SDR is stronger for All-sky and Clear-sky condition for low Sun elevation angles (Figures\ \ref{fig:szatrends}a and \ref{fig:szatrends}b).
-In the case of Cloud-sky condition (Figure\ \ref{fig:szatrends}c), there is no strong correlation between SDR brightening and SZA, except for some low Sun angles, that occur only for a small part of the year, and thus we can not draw more conclusions from this analysis for the over trend.
+Although there is a seasonal variation of the minimum SZA, that is not examined here.
+The brightening effect of SDR is stronger for All-sky and Clear-sky condition for low Sun elevation angles (Figures\ \ref{fig:szatrends}a and \ref{fig:szatrends}b).
+In the case of Cloud-sky condition (Figure\ \ref{fig:szatrends}c), there is no strong correlation between SDR brightening and SZA, except for some low Sun angles, that occur only for a small part of the year, and thus we can not draw more conclusions from this analysis for the overall trends.
 
@@
 
-Similar to the long term trends discussed above, we have produced the deseasonalized trend for the different sky conditions, for each season of the year, using the corresponding mean monthly values
+Similar to the long term trends discussed above, we have produced the deseasonalized trend for the different sky conditions and for each season of the year, using the corresponding mean monthly values
 (Figure\ \ref{fig:seasonalALL} and Table\ \ref{tab:trendseasontable}).
 For All-sky condition the brightening of SDR is very significant for the Winter, close to the long term trend for the Autumn, and almost four times lower than the long term trend for Sprint and Summer, indicating a relation to the cloud pattern of each season.
-Moreover when examining the Clear- and Cloud-sky conditions the trends between these condition are are almost complementary to each other for Winter and Autumn respectively.
-On the latter cases during Spring and Summer the statistical significance along with the actual trend are very low for a meaningful comparison.
+Moreover when examining the Clear- and Cloud-sky conditions the trends between these condition are almost complementary to each other for Winter and Autumn respectively.
+On the latter cases, during Spring and Summer the statistical significance along with the actual trend are very low for a meaningful comparison.
 
@@
 
-We demonstrate that in the period `r paste(year(min(ALL_1_daily_DESEAS$Date)), " - ", year(max(ALL_1_daily_DESEAS$Date)))`, there is brightening trend of SDR for All-sky conditions ($`r alltrend`\%/year$).
-On a previous study [@Bais2013] for the period 1993 - 2011 the trend was found to be $0.33\%/year$.
+We demonstrate that in the period 
+`r paste(year(min(ALL_1_daily_DESEAS$Date)), " - ", year(max(ALL_1_daily_DESEAS$Date)))`, there is a brightening trend of SDR for All-sky conditions ($`r alltrend`\%/year$).
+On a previous study [@Bais2013], for the period 1993 - 2011 the trend was found to be $0.33\%/year$.
 The increase of this trend, indicates that the phenomenon and probably the causes, are still occurring.
 Accordingly, we found a similar trend for Clear-sky conditions ($`r cleartrend`\%/year$) that further supports that the brightening can be independent from the cloud occurrences. 
-We have calculate the effect of TSI variability to be negligible, but unfortunately there is no available data for this period for the Aerosols, in order to estimate their effect on SDR, which is the most probable main cause.
+We have calculated the effect of TSI variability to be a negligible factor in SDR variability.
+Unfortunately, there is no available data for this period for the Aerosols, in order to estimate their effect on SDR, which is the most probable main cause.
 The attenuation of SDR by Aerosols over Europe have been proposed as major factor by @Wild2021.
@@
 And also, with the seasons of the year, where Winter and Autumn show a significant strong trend, in contrast to Spring and Summer.
+
+Future observation from the colocated pyrheliometer, may provide us with the means to further investigate the variability of Sun radiation on ground level in Thessaloniki.
 

### 2023-09-05 Build_152/Article_B152.Rmd
Commit 2023-09-05 17:10

--- 2023-09-05 Article.Rmd
@@
 <!-- aggregation -->
-In order to investigate the SDR trends, we implemented an appropriate aggregation scheme to the 1-minute data to derive a series in coarser time scale.
+In order to investigate the SDR trends, we implemented an appropriate aggregation scheme to the 1-minute data to derive a series in coarser timescale.
 To preserve the representativeness of the data we used the following criteria: 
@@
 In our case, for all sky conditions, we observe periods where the cumulative sum remains relative steady, roughly in the middle of the study dataset.
-Another distinct feature of the cumulative plots, is the difference, between the Cloud-sky and the rest of the sky conditions, where they display an supplementary behavior. 
+Another distinct feature of the cumulative plots, is the difference, between the Cloud-sky and the rest of the sky conditions, where they display a supplementary behavior. 
 The causes for these discrepancies haven't been investigated yet.
@@
 For All-sky condition the brightening of SDR is very significant for the Winter, close to the long term trend for the Autumn, and almost four times lower than the long term trend for Sprint and Summer, indicating a relation to the cloud pattern of each season.
-Moreover when examining the Clear- and Cloud-sky conditions the trends between these condition are almost complementary to each other for Winter and Autumn respectively.
+Moreover, when examining the Clear- and Cloud-sky conditions the trends between these conditions are almost complementary to each other for Winter and Autumn respectively.
 On the latter cases, during Spring and Summer the statistical significance along with the actual trend are very low for a meaningful comparison.
@@
 The increase of this trend, indicates that the phenomenon and probably the causes, are still occurring.
-Accordingly, we found a similar trend for Clear-sky conditions ($`r cleartrend`\%/year$) that further supports that the brightening can be independent from the cloud occurrences. 
+Accordingly, we found a similar trend for Clear-sky conditions ($`r cleartrend`\%/year$) that further supports that the brightening can be independent of the cloud occurrences. 
 We have calculated the effect of TSI variability to be a negligible factor in SDR variability.
@@
 
+Our finds are in agreement with other researchers for the region.
 Future observation from the colocated pyrheliometer, may provide us with the means to further investigate the variability of Sun radiation on ground level in Thessaloniki. 
+Also, more data (cloud occurrence, aerosols, atmospheric water, etc.), will help us to determine the extent of SDR attenuation by each factor.
 

### 2023-09-12 Article.Rmd
Commit 2023-09-12 12:14

@@
 abstract: |
-  Study of GHI and DNI radiation for 'clear-sky' and all-sky conditions.
-  It consists of two paragraphs.
+  The shortwave downward solar irradiance (SDR).........
+  Human and climate.......
+  Brigthening dimming...............
+  We study 29 years of CHP1 data.........
+  Clear cloud id ................
+  Found trends ......
 keywords: 

### 2023-09-12 Article.Rmd
Commit 2023-09-12 20:14

@@
 \begin{equation}
-\text{SDR}_\text{CSref} = a \times \text{SDR}_\text{Haurwitz} = `r signif(alpha,3)` \times 1098 \times \cos( \text{SZA} ) \times \exp \left( \frac{ - 0.057}{\cos(\text{SZA})} \right) \label{eq:ahau}
+\text{SDR}_\text{CSref} = a \times \text{SDR}_\text{Haurwitz} = `r signif(alpha,3)` \times 1098 \times \cos(\theta) \times \exp \left( \frac{ - 0.057}{\cos(\theta)} \right) \label{eq:ahau}
 \end{equation}
 
-where: $\text{SDR}_\text{CSref}$ is the reference clear sky SDR, in $\text{w}\,\text{m}^{-2}$.
+where: $\text{SDR}_\text{CSref}$ is the reference clear sky SDR, in $\text{w}\,\text{m}^{-2}$ and $\theta$ is the solar zenith angle (SZA).
 
@@
 
-We calculated the SDR trends, from the mean daily values of the departure from the seasonal values, for the three sky conditions (Table\ \ref{tab:trendtable}).
-We present only the plot for the trend under all-sky conditions (Figure\ \ref{fig:trendALL}), as the plots for clear-sky condition and cloud-sky condition, are very similar.
+We calculated the linear SDR trends, from the departures of the mean daily values from the seasonal values, and for the three sky conditions (Table\ \ref{tab:trendtable}).
+In Figure\ \ref{fig:trendALL} we present only the time series under all-sky conditions; the plots for clear-sky and cloud-sky conditions, are very similar.
 In the studied time period, there is no significant break or change in the variability of the overall trends.
@@
 
-```{r trendALL, echo=F, out.width='.70\\linewidth', fig.cap=paste0("Anomaly (\\%) of the daily SDR relative to climatological values for ", year(min(ALL_1_daily_DESEAS$Date)), " - ", year(max(ALL_1_daily_DESEAS$Date)), ".
+```{r trendALL, echo=F, out.width='.70\\linewidth', fig.cap=paste0("Anomaly (\\%) of the daily All-sky SDR, relative to climatological values for ", year(min(ALL_1_daily_DESEAS$Date)), " - ", year(max(ALL_1_daily_DESEAS$Date)), ".
 The black line shows the long term trend for all-sky conditions.")}
@@
 Although the trends are quite constant, using plots of the cumulative sums we can reveal some structure within the trends and some differences between the sky conditions [@Regier2019]. 
-On all three plots on Figure\ \ref{fig:cumsum}, we observe a downward part at the start of the period until about 2000, an almost steady part until about 2016 for All-sky and Clear-sky conditions (Figures\ \ref{fig:cumsum-1}, \ref{fig:cumsum-2}) and until 2009 for Cloud-sky conditions (Figure\ \ref{fig:cumsum-3}), and finally, a steep upward part until the present.
+On all three plots on Figure\ \ref{fig:cumsum}, we observe a downward part at the start of the period until about 2000, an almost steady part until about 2016 for All-sky and clear-sky conditions (Figures\ \ref{fig:cumsum-1}, \ref{fig:cumsum-2}) and until 2009 for cloud-sky conditions (Figure\ \ref{fig:cumsum-3}), and finally, a steep upward part until the present.
 For a uniform trend, we expected the cumulative sum plot of the anomaly to have a symmetric 'V' shape.
@@
 In our case, for all sky conditions, we observe periods where the cumulative sum remains relative steady, roughly in the middle of the study dataset.
-Another distinct feature of the cumulative plots, is the difference, between the Cloud-sky and the rest of the sky conditions, where they display a supplementary behavior. 
+Another distinct feature of the cumulative plots, is the difference, between the cloud-sky and the rest of the sky conditions, where they display a supplementary behavior. 
 The causes for these discrepancies haven't been investigated yet.
@@
 This can make us to suspected that clear-Sky conditions SDR and cloud-sky SDR is affected by different mechanisms.
-The similarity of the All-sky and Clear-sky conditions, can make us to suspected that on both sky conditions the attenuation of SDR has .
+The similarity of the All-sky and clear-sky conditions, can make us to suspected that on both sky conditions the attenuation of SDR has .
 -->
 
-```{r cumsum, echo = FALSE, fig.cap='Running cumulative sum plots of the daily SDR for three different sky conditions.', fig.subcap=c('All-sky conditions.', 'Clear-sky conditions.', 'Cloud-sky conditions.'), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3}
+```{r cumsum, echo = FALSE, fig.cap='.......
+Running cumulative ......
+sum plots of the daily SDR for three different sky conditions.', fig.subcap=c('All-sky conditions.', 'clear-sky conditions.', 'Cloud-sky conditions.'), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3}
 knitr::include_graphics(c(
@@
 
-<!-- ```{r trendCLEAR, echo=F, fig.cap="Daily SDR seasonal anomaly and long term trend, for clear sky conditions."} -->
-<!-- knitr::include_graphics("./images/LongtermTrends-4.png") -->
-<!-- ``` -->
-
-<!-- ```{r trendCLOUD, echo=F, fig.cap="Anomaly (\\%) of the daily SDR relative to climatological values for 1993 - 2023. The black line shows the long term trend for cloud-sky conditions."} -->
-<!-- knitr::include_graphics("./images/LongtermTrends-6.png") -->
-<!-- ``` -->
-
 ## Effects of the solar zenith angle on SDR.
@@
 Although there is a seasonal variation of the minimum SZA, that is not examined here.
-The brightening effect of SDR is stronger for All-sky and Clear-sky condition for low Sun elevation angles (Figures\ \ref{fig:szatrends}a and \ref{fig:szatrends}b).
-In the case of Cloud-sky condition (Figure\ \ref{fig:szatrends}c), there is no strong correlation between SDR brightening and SZA, except for some low Sun angles, that occur only for a small part of the year, and thus we can not draw more conclusions from this analysis for the overall trends.
+The brightening effect of SDR is stronger for All-sky and clear-sky condition for low Sun elevation angles (Figures\ \ref{fig:szatrends}a and \ref{fig:szatrends}b).
+In the case of cloud-sky condition (Figure\ \ref{fig:szatrends}c), there is no strong correlation between SDR brightening and SZA, except for some low Sun angles, that occur only for a small part of the year, and thus we can not draw more conclusions from this analysis for the overall trends.
 
@@
 For All-sky condition the brightening of SDR is very significant for the Winter, close to the long term trend for the Autumn, and almost four times lower than the long term trend for Sprint and Summer, indicating a relation to the cloud pattern of each season.
-Moreover, when examining the Clear- and Cloud-sky conditions the trends between these conditions are almost complementary to each other for Winter and Autumn respectively.
+Moreover, when examining the clear- and cloud-sky conditions the trends between these conditions are almost complementary to each other for Winter and Autumn respectively.
 On the latter cases, during Spring and Summer the statistical significance along with the actual trend are very low for a meaningful comparison.
@@
 
-<!-- ```{r seasonalCLEAR, echo=F, fig.cap="Trends by season for Clear-sky condition. Displays monthly means of daily means."} -->
+<!-- ```{r seasonalCLEAR, echo=F, fig.cap="Trends by season for clear-sky condition. Displays monthly means of daily means."} -->
 <!-- knitr::include_graphics("./images/SeasonalTrends-2.png") -->
@@
 pp        <- pp[new_order,]
-## put all sky first
-A  <- which(pp$`Sky condition` == "All sky cond.")
-B  <- which(pp$`Sky condition` != "All sky cond.")
-pp <- pp[c(A,B),]
+# ## put all sky first
+# A  <- which(pp$`Sky condition` == "All sky cond.")
+# B  <- which(pp$`Sky condition` != "All sky cond.")
+# pp <- pp[c(A,B),]
 ## numeric precision
@@
 The increase of this trend, indicates that the phenomenon and probably the causes, are still occurring.
-Accordingly, we found a similar trend for Clear-sky conditions ($`r cleartrend`\%/year$) that further supports that the brightening can be independent of the cloud occurrences. 
+Accordingly, we found a similar trend for clear-sky conditions ($`r cleartrend`\%/year$) that further supports that the brightening can be independent of the cloud occurrences. 
 We have calculated the effect of TSI variability to be a negligible factor in SDR variability.
@@
 The attenuation of SDR by Aerosols over Europe have been proposed as major factor by @Wild2021.
-The dimming effect of SDR under Cloud-sky condition ($`r cloudtrend`\%/year$), shows that clouds can also attenuate the amount of SDR reaching the ground, and changes on their occurrences can effect the net energy balance.
-Because we have no adequate data to investigate long term changes of cloud formation in the region, we can not infer if the SDR trend we observe under Cloud-sky is biased by changes in clouds and local climate.
+The dimming effect of SDR under cloud-sky condition ($`r cloudtrend`\%/year$), shows that clouds can also attenuate the amount of SDR reaching the ground, and changes on their occurrences can effect the net energy balance.
+Because we have no adequate data to investigate long term changes of cloud formation in the region, we can not infer if the SDR trend we observe under cloud-sky is biased by changes in clouds and local climate.
 The observed brightening of SDR over Thessaloniki, is correlated with the SZA (higher SZA, have stronger brightening trends).
@@
 # References
+
+<div id="refs"></div>
+
+\newpage
+# Appendix
+
+```{r trendCLEAR, echo=F, fig.cap=paste0("Anomaly (\\%) of the daily clear-sky SDR, relative to climatological values for", year(min(ALL_1_daily_DESEAS$Date)), " - ", year(max(ALL_1_daily_DESEAS$Date)), ".
+The black line shows the long term trend for clear-sky conditions.")}
+knitr::include_graphics("./images/LongtermTrends-5.png")
+```
+
+```{r trendCLOUD, echo=F, fig.cap=paste0("Anomaly (\\%) of the daily cloud-sky SDR, relative to climatological values for ", year(min(ALL_1_daily_DESEAS$Date)), " - ", year(max(ALL_1_daily_DESEAS$Date)), ".
+The black line shows the long term trend for cloud-sky conditions.")}
+knitr::include_graphics("./images/LongtermTrends-8.png")
+```

### 2023-09-12 DHI_GHI_3_trends_consistency.R

@@ -47,7 +47,7 @@
 
 #+ echo=F, include=F
 knitr::opts_chunk$set(comment    = ""       )
-knitr::opts_chunk$set(dev        = c("pdf", "png") )
+knitr::opts_chunk$set(dev        = c("pdf", "png"))
 # knitr::opts_chunk$set(dev        = "png"    )
 knitr::opts_chunk$set(out.width  = "100%"   )
 knitr::opts_chunk$set(fig.align  = "center" )

### 2023-09-12 Article.Rmd
Commit 2023-09-12 21:10

@@
 We calculated the linear SDR trends, from the departures of the mean daily values from the seasonal values, and for the three sky conditions (Table\ \ref{tab:trendtable}).
-In Figure\ \ref{fig:trendALL} we present only the time series under all-sky conditions; the plots for clear-sky and cloud-sky conditions, are very similar.
+In Figure\ \ref{fig:trendALL} we present only the time series under all-sky conditions; the plots for clear-sky and cloud-sky conditions, are very similar and shown in the Appendix (Figures\ \ref{fig:trendCLEAR} and \ref{fig:trendCLOUD}).
 In the studied time period, there is no significant break or change in the variability of the overall trends.
-Other studies for the European region are reporting a change of the SDR slope, around the 1980, few years before the start of our records [@Wild2021; @Yuan2021; @Ohmura2009].
+Other studies for the European region report a change of the SDR slope, around 1980, a few years before the start of our records [@Wild2021; @Yuan2021; @Ohmura2009].
 It is interesting to note, that for the observation period, the trend of the TSI is
-$`r format(TSIslopeAll[1,1], scientific = T, digits = 2)`\%/year$, and thus we can not attribute any major effect on SDR trend to Sun variability.
+$`r format(TSIslopeAll[1,1], scientific = T, digits = 2)`\%/year$, and thus we can not attribute any major effect on SDR trend to Solar variability.
 
@@
 ```{r trendALL, echo=F, out.width='.70\\linewidth', fig.cap=paste0("Anomaly (\\%) of the daily All-sky SDR, relative to climatological values for ", year(min(ALL_1_daily_DESEAS$Date)), " - ", year(max(ALL_1_daily_DESEAS$Date)), ".
-The black line shows the long term trend for all-sky conditions.")}
+The black line shows the long term linear trend for all-sky conditions.")}
 knitr::include_graphics("./images/LongtermTrends-2.png")
@@
        split.table = Inf,
-       caption = "\\label{tab:trendseasontable}SDR trends of monthly anomaly for each season of the year.")
+       caption = "\\label{tab:trendseasontable}SDR linear trends of monthly anomalies for each season of the year.")
 ```
@@
 Because we have no adequate data to investigate long term changes of cloud formation in the region, we can not infer if the SDR trend we observe under cloud-sky is biased by changes in clouds and local climate.
-The observed brightening of SDR over Thessaloniki, is correlated with the SZA (higher SZA, have stronger brightening trends).
+The observed brightening of SDR over Thessaloniki, is dependent on SZA (higher SZA, have stronger brightening trends).
 And also, with the seasons of the year, where Winter and Autumn show a significant strong trend, in contrast to Spring and Summer.
 
-Our finds are in agreement with other researchers for the region.
-Future observation from the colocated pyrheliometer, may provide us with the means to further investigate the variability of Sun radiation on ground level in Thessaloniki. 
-Also, more data (cloud occurrence, aerosols, atmospheric water, etc.), will help us to determine the extent of SDR attenuation by each factor.
+Our findings are in agreement with other researchers for the region.
+Future observations from the collocated pyrheliometer, may provide us with the means to further investigate the variability of Solar radiation at ground level in Thessaloniki. 
+Also, more data (cloud occurrence, aerosols, atmospheric water vapour, etc.), will help us to determine the extent of SDR attenuation by each factor.
 
@@
 ```{r trendCLEAR, echo=F, fig.cap=paste0("Anomaly (\\%) of the daily clear-sky SDR, relative to climatological values for", year(min(ALL_1_daily_DESEAS$Date)), " - ", year(max(ALL_1_daily_DESEAS$Date)), ".
-The black line shows the long term trend for clear-sky conditions.")}
+The black line shows the long term linear trend for clear-sky conditions.")}
 knitr::include_graphics("./images/LongtermTrends-5.png")
@@
 ```{r trendCLOUD, echo=F, fig.cap=paste0("Anomaly (\\%) of the daily cloud-sky SDR, relative to climatological values for ", year(min(ALL_1_daily_DESEAS$Date)), " - ", year(max(ALL_1_daily_DESEAS$Date)), ".
-The black line shows the long term trend for cloud-sky conditions.")}
+The black line shows the long term linear trend for cloud-sky conditions.")}
 knitr::include_graphics("./images/LongtermTrends-8.png")

### 2023-09-12 DHI_GHI_0_variables.R

@@ -10,7 +10,7 @@ dict <- list(DIR_att    = 'Dir. Beam Irrad.',
              tsi1au_att = "TSI at 1au",
              ALL        = "All sky cond.",
              CLEAR      = "Clear sky cond.",
-             CLOUD      = "Cloudy cond.")
+             CLOUD      = "Cloudy sky cond.")
 ## function to translate objects names
 # translate <- function(...) as.vector(unlist(dict[c(...) == names(dict)]))
 

### 2023-09-13 Article.Rmd
Commit 2023-09-13 10:10

@@
 LT$`Trend statistical signif. [%]` <- NULL
-## numeric presision
+## numeric precision
 LT$`Trend [%/year]` <- signif(LT$`Trend [%/year]`, 3)
 LT$`Trend p-value`  <- signif(LT$`Trend p-value`,  3)
+
+## some vars to use later
+alltrend   <- signif(LT[`Sky condition` == "All sky cond.",   `Trend [%/year]`], 2)
+cleartrend <- signif(LT[`Sky condition` == "Clear sky cond.", `Trend [%/year]`], 2)
+cloudtrend <- signif(LT[`Sky condition` == "Cloudy cond.",    `Trend [%/year]`], 2)
 
@@
 For all cases the statistical significance is above $99.9\\%$."))
-# knitr::kable(LT, caption = "Trends of daily means by sky conditions.")
-## some vars to use later
-alltrend   <- signif(LT[`Sky condition` == "All sky cond.",   `Trend [%/year]`], 2)
-cleartrend <- signif(LT[`Sky condition` == "Clear sky cond.", `Trend [%/year]`], 2)
-cloudtrend <- signif(LT[`Sky condition` == "Cloudy cond.",    `Trend [%/year]`], 2)
 ```

### 2023-09-13 Article.Rmd
Commit 2023-09-13 11:23

@@
 
+## rename sky conditions
+LT$`Sky condition` <- sub(" cond.", "", LT$`Sky condition`)
+
 panderOptions("table.alignment.default", "center")
@@
 
-Although the trends are quite constant, using plots of the cumulative sums we can reveal some structure within the trends and some differences between the sky conditions [@Regier2019]. 
+Although the trends are quite constant, using plots of the cumulative sum (CUSUM) we can reveal some structure within the trends and some differences between the sky conditions [@Regier2019]. 
 On all three plots on Figure\ \ref{fig:cumsum}, we observe a downward part at the start of the period until about 2000, an almost steady part until about 2016 for All-sky and clear-sky conditions (Figures\ \ref{fig:cumsum-1}, \ref{fig:cumsum-2}) and until 2009 for cloud-sky conditions (Figure\ \ref{fig:cumsum-3}), and finally, a steep upward part until the present.
-For a uniform trend, we expected the cumulative sum plot of the anomaly to have a symmetric 'V' shape.
-This would indicate, that the anomaly is evenly distributed around the climatological mean.
-In our case, for all sky conditions, we observe periods where the cumulative sum remains relative steady, roughly in the middle of the study dataset.
-Another distinct feature of the cumulative plots, is the difference, between the cloud-sky and the rest of the sky conditions, where they display a supplementary behavior. 
+For a uniform trend, we would expect the cumulative sums of the anomalies to have a symmetric 'V' shape.
+This would indicate that the anomalies are evenly distributed around the climatological mean.
+In our case, for all sky conditions, we observe periods where the CUSUMs remain relative steady, roughly around the middle of the study dataset.
+Another distinct feature of CUSUMs, is the difference, between the cloud-sky and the rest of the sky conditions, where they display a supplementary behavior. 
 The causes for these discrepancies haven't been investigated yet.
@@
 
-The relative position of the Sun in the sky, is a major drive of the attenuation of SDR reaching the ground, as it is affected by the radiation path in the atmosphere and possible human activities in urban environment emitting aerosols [@Wang2021].
+The solar zenith angle  is a major driver of the attenuation of SDR reaching the ground, as SDR is affected by the enhancement of the radiation path in the atmosphere, especially in urban environment with human activities emitting aerosols [@Wang2021].
 In order to estimate the contribution of the SZA, we group the anomaly data in $1^\circ$ SZA bins, and calculated the overall trend for each bin before noon and after noon (Figure\ \ref{fig:szatrends}).

### 2023-09-13 DHI_GHI_3_trends_consistency.R

@@ -210,7 +210,7 @@ setorder(CLOUD_1_daily_DESEAS, Date)
 trends <- readRDS("./figures/tbl_longterm_trends.Rds")
 
 
-## !! Remove trends before cumulative sum --------------------------------------
+### !! Remove trends before cumulative sum --------------------------------------
 dsets <- c("ALL_1_daily_DESEAS",
            "CLEAR_1_daily_DESEAS",
            "CLOUD_1_daily_DESEAS")
@@ -229,13 +229,9 @@ for (ad in dsets) {
         ## remove trend from data
         DBS[[avaa]] <- DBS[[avaa]] - SDRtredn
     }
-    ## save new data to data table
-    assign(ad, DBS)
+    ## save new data data table without the trend
+    assign(paste0(ad, "_notrend"), DBS)
 }
-
-
-
-
 
 
 
@@ -261,6 +257,50 @@ CLEAR_1_daily_DESEAS[, DIR_transp_cusumSTD := cumsum(tidyr::replace_na((DIR_tran
 CLOUD_1_daily_DESEAS[, GLB_att_cusumSTD    := cumsum(tidyr::replace_na((GLB_att_des    - mean(GLB_att_des   , na.rm = T)) / sd(GLB_att_des   , na.rm = T), 0))]
 CLOUD_1_daily_DESEAS[, DIR_att_cusumSTD    := cumsum(tidyr::replace_na((DIR_att_des    - mean(DIR_att_des   , na.rm = T)) / sd(DIR_att_des   , na.rm = T), 0))]
 CLOUD_1_daily_DESEAS[, DIR_transp_cusumSTD := cumsum(tidyr::replace_na((DIR_transp_des - mean(DIR_transp_des, na.rm = T)) / sd(DIR_transp_des, na.rm = T), 0))]
+
+
+
+
+## Calculate cumsum with zeroing NA
+  ALL_1_daily_DESEAS_notrend[, GLB_att_cusum    := cumsum(tidyr::replace_na(GLB_att_des,    0))]
+  ALL_1_daily_DESEAS_notrend[, DIR_att_cusum    := cumsum(tidyr::replace_na(DIR_att_des,    0))]
+  ALL_1_daily_DESEAS_notrend[, DIR_transp_cusum := cumsum(tidyr::replace_na(DIR_transp_des, 0))]
[…]

### 2023-09-13 Article.Rmd
Commit 2023-09-13 12:14

@@
 
-```{r trendALL, echo=F, out.width='.70\\linewidth', fig.cap=paste0("Anomaly (\\%) of the daily All-sky SDR, relative to climatological values for ", year(min(ALL_1_daily_DESEAS$Date)), " - ", year(max(ALL_1_daily_DESEAS$Date)), ".
+```{r trendALL, echo=F, out.width='.70\\linewidth', fig.cap=paste0("Anomaly (\\%) of the daily all-sky SDR, relative to climatological values for ", year(min(ALL_1_daily_DESEAS$Date)), " - ", year(max(ALL_1_daily_DESEAS$Date)), ".
 The black line shows the long term linear trend for all-sky conditions.")}
@@
 Although the trends are quite constant, using plots of the cumulative sum (CUSUM) we can reveal some structure within the trends and some differences between the sky conditions [@Regier2019]. 
-On all three plots on Figure\ \ref{fig:cumsum}, we observe a downward part at the start of the period until about 2000, an almost steady part until about 2016 for All-sky and clear-sky conditions (Figures\ \ref{fig:cumsum-1}, \ref{fig:cumsum-2}) and until 2009 for cloud-sky conditions (Figure\ \ref{fig:cumsum-3}), and finally, a steep upward part until the present.
+On all three plots on Figure\ \ref{fig:cumsum}, we observe a downward part at the start of the period until about 2000, an almost steady part until about 2016 for all-sky and clear-sky conditions (Figures\ \ref{fig:cumsum-1}, \ref{fig:cumsum-2}) and until 2009 for cloud-sky conditions (Figure\ \ref{fig:cumsum-3}), and finally, a steep upward part until the present.
 For a uniform trend, we would expect the cumulative sums of the anomalies to have a symmetric 'V' shape.
@@
 This can make us to suspected that clear-Sky conditions SDR and cloud-sky SDR is affected by different mechanisms.
-The similarity of the All-sky and clear-sky conditions, can make us to suspected that on both sky conditions the attenuation of SDR has .
+The similarity of the all-sky and clear-sky conditions, can make us to suspected that on both sky conditions the attenuation of SDR has .
 -->
@@
 Although there is a seasonal variation of the minimum SZA, that is not examined here.
-The brightening effect of SDR is stronger for All-sky and clear-sky condition for low Sun elevation angles (Figures\ \ref{fig:szatrends}a and \ref{fig:szatrends}b).
+The brightening effect of SDR is stronger for all-sky and clear-sky condition for low Sun elevation angles (Figures\ \ref{fig:szatrends}a and \ref{fig:szatrends}b).
 In the case of cloud-sky condition (Figure\ \ref{fig:szatrends}c), there is no strong correlation between SDR brightening and SZA, except for some low Sun angles, that occur only for a small part of the year, and thus we can not draw more conclusions from this analysis for the overall trends.
@@
 
-Similar to the long term trends discussed above, we have produced the deseasonalized trend for the different sky conditions and for each season of the year, using the corresponding mean monthly values
+Similar to the long term trends discussed above, we have calculated the trend of the anomalies for the three different sky conditions and for each season of the year, using the corresponding mean monthly values
 (Figure\ \ref{fig:seasonalALL} and Table\ \ref{tab:trendseasontable}).
-For All-sky condition the brightening of SDR is very significant for the Winter, close to the long term trend for the Autumn, and almost four times lower than the long term trend for Sprint and Summer, indicating a relation to the cloud pattern of each season.
+For all-sky condition the trend in SDR is very significant for the Winter, close to the long term trend for the Autumn, and almost four times lower than the long term trend for Sprint and Summer, indicating a relation to the cloud pattern of each season.
 Moreover, when examining the clear- and cloud-sky conditions the trends between these conditions are almost complementary to each other for Winter and Autumn respectively.
@@
 
-<!-- ```{r seasonalALL, echo=F, fig.cap="Trends by season for All-sky condition. Displays monthly means of daily means."} -->
+<!-- ```{r seasonalALL, echo=F, fig.cap="Trends by season for all-sky condition. Displays monthly means of daily means."} -->
 <!-- knitr::include_graphics("./images/SeasonalTrends-1.png") -->
@@
 pp        <- pp[new_order,]
+## order by sky condition
+pp <- pp[order(pp$`Sky condition`),]
+
 # ## put all sky first
@@
 pp$`Statistical signif. [%]` <- signif(pp$`Statistical signif. [%]`,  3)
+
+## rename sky conditions
+pp$`Sky condition` <- sub(" cond.", "", pp$`Sky condition`)
 
@@
 We demonstrate that in the period 
-`r paste(year(min(ALL_1_daily_DESEAS$Date)), " - ", year(max(ALL_1_daily_DESEAS$Date)))`, there is a brightening trend of SDR for All-sky conditions ($`r alltrend`\%/year$).
+`r paste(year(min(ALL_1_daily_DESEAS$Date)), " - ", year(max(ALL_1_daily_DESEAS$Date)))`, there is a brightening trend of SDR for all-sky conditions ($`r alltrend`\%/year$).
 On a previous study [@Bais2013], for the period 1993 - 2011 the trend was found to be $0.33\%/year$.

### 2023-09-13 DHI_GHI_3_trends_consistency.R

@@ -289,20 +289,6 @@ CLOUD_1_daily_DESEAS_notrend[, DIR_transp_cusumSTD := cumsum(tidyr::replace_na((
 
 
 
-
-
-
-
-
-
-
-
-
-
-
-
-
-
 ## ~ Calculate monthly cum sum -------------------------------------------------
 
 setorder(  ALL_1_D_monthly_DESEAS, Date)

### 2023-09-14 Article.Rmd
Commit 2023-09-14 11:13

@@
 In Figure\ \ref{fig:trendALL} we present only the time series under all-sky conditions; the plots for clear-sky and cloud-sky conditions, are very similar and shown in the Appendix (Figures\ \ref{fig:trendCLEAR} and \ref{fig:trendCLOUD}).
-In the studied time period, there is no significant break or change in the variability of the overall trends.
+In the studied period, there is no significant break or change in the variability of the time series.
 Other studies for the European region report a change of the SDR slope, around 1980, a few years before the start of our records [@Wild2021; @Yuan2021; @Ohmura2009].
@@
 
-```{r cumsum, echo = FALSE, fig.cap='.......
-Running cumulative ......
-sum plots of the daily SDR for three different sky conditions.', fig.subcap=c('All-sky conditions.', 'clear-sky conditions.', 'Cloud-sky conditions.'), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3}
+```{r cumsum, echo = FALSE, fig.cap='Running cumulative sum plots of the daily SDR for the three different sky conditions.', fig.subcap=c('All-sky conditions.', 'Clear-sky conditions.', 'Cloud-sky conditions.'), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3}
 knitr::include_graphics(c(
@@
     "./images/CumulativeDailyCumSum-9.png"))   # Cloud sky
+```
+
+```{r cumsumnotrend, echo = FALSE, fig.cap='Running cumulative sum plots of the daily SDR after removing the long term trend from the daily values, for the three different sky conditions.', fig.subcap=c('All-sky conditions.', 'Clear-sky conditions.', 'Cloud-sky conditions.'), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3}
+knitr::include_graphics(c(
+    "./images/CumulativeDailyCumSumNOtrend-1.png",    # ALL sky
+    "./images/CumulativeDailyCumSumNOtrend-5.png",    # CLEAR sky
+    "./images/CumulativeDailyCumSumNOtrend-9.png"))   # Cloud sky
 ```

### 2023-09-14 DHI_GHI_3_trends_consistency.R

@@ -528,15 +528,17 @@ for (adb in database) {
              xlab = "",
              xlim = xlim, ylim = ylim,
              xaxt = "n",
-             ylab = bquote("Cumulative Sum of Anomaly [%]"))
+             ylab = bquote("Cumulative Sum of Anomaly with long trend removed [%]"))
         axis.Date(1, pdb$Date)
-        # abline(h = 0, lty = 2, lwd = 0.8)
+        abline(h = 0, lty = 2, lwd = 0.8)
 
         ## daily from other DT
         lines(pdb$Date, pdb[[paste0(avar,"_cusum")]], col = col, lwd = 2)
 
         title(paste(sub("_.*","",adb), "mean daily cumulative sum ",
                     translate(avar) ), cex.main = 1)
+
+
 
 
         ## test plot for reference
@@ -552,13 +554,10 @@ for (adb in database) {
                     translate(avar) ), cex.main = 1)
         abline(lm1, lwd = 2)
 
-
         ## running mean
         rm <- frollmean(pdb[[paste0(avar,"_des")]], round(running_mean_window_days),
                         na.rm = TRUE, algo = "exact", align = "center")
         points(pdb$Date, rm, col = "red", cex = 0.4)
-
-
 
         ## LOESS curve
         vec <- !is.na(pdb[[paste0(avar,"_des")]])
@@ -568,7 +567,6 @@ for (adb in database) {
         FTSE.lo.predict3 <- predict(FTSE.lo3, pdb$Date)
         lines(pdb$Date, FTSE.lo.predict3, col = "cyan", lwd = 2.5)
 
-
         ## decorations
         fit <- lm1[[1]]
         legend("top", lty = 1, bty = "n", lwd = 2, cex = 1,
@@ -577,6 +575,7 @@ for (adb in database) {
                      signif(abs(fit[2]) * Days_of_year, 3),
                      "% per year")
         )
+
 
 
         # hist(pdb$GLB_att_cusum)

### 2023-09-15 Article.Rmd
Commit 2023-09-15 13:11

@@
 
+```{r echo=F, eval=TRUE}
+## Prepare table and data
+spp <- read.csv("./figures/tbl_longterm_trends_season.dat",
+               strip.white = TRUE,
+               sep         = ";",
+               header      = TRUE)
+spp <- spp[grep("SDR" ,spp$var), ]
+spp$slope.sd           <- NULL
+spp$slope.ConfInt_0.95 <- NULL
+spp$slope.ConfInt_0.99 <- NULL
+spp$var                <- NULL
+rownames(spp)          <- NULL
+names(spp) <- c("Trend [%/year]", "Trend p-value", "Sky condition", "Season", "Statistical signif. [%]")
+## order by season
+target    <- c("Winter", "Spring", "Summer", "Autumn")
+new_order <- sapply(target, function(x,spp){which(spp$Season == x)}, spp = spp)
+spp        <- spp[new_order,]
+## order by sky condition
+spp <- spp[order(spp$`Sky condition`),]
+
+# ## put all sky first
+# A  <- which(spp$`Sky condition` == "All sky cond.")
+# B  <- which(spp$`Sky condition` != "All sky cond.")
+# spp <- spp[c(A,B),]
+## numeric precision
+spp$`Trend [%/year]`          <- signif(spp$`Trend [%/year]`, 3)
+spp$`Trend p-value`           <- signif(spp$`Trend p-value`,  3)
+spp$`Statistical signif. [%]` <- signif(spp$`Statistical signif. [%]`,  3)
+
+## rename sky conditions
+spp$`Sky condition` <- sub(" cond.", "", spp$`Sky condition`)
+```
+
 Similar to the long term trends discussed above, we have calculated the trend of the anomalies for the three different sky conditions and for each season of the year, using the corresponding mean monthly values
 (Figure\ \ref{fig:seasonalALL} and Table\ \ref{tab:trendseasontable}).
-For all-sky condition the trend in SDR is very significant for the Winter, close to the long term trend for the Autumn, and almost four times lower than the long term trend for Sprint and Summer, indicating a relation to the cloud pattern of each season.
+For all-sky condition the trend in SDR in winter is the largest
+($`r signif(spp$"Trend [%/year]"[spp$Season == "Winter" & spp$"Sky condition" == "All sky"], 2)`\,\%/y$),
+followed by the trend in autumn
+($`r signif(spp$"Trend [%/year]"[spp$Season == "Autumn" & spp$"Sky condition" == "All sky"], 2)`\,\%/y$,
+close to the long term trend) both statistically significant above the $99\,\%$ confidence level. 
+In sprint and summer   , indicating a relation to the cloud pattern of each season.
 Moreover, when examining the clear- and cloud-sky conditions the trends between these conditions are almost complementary to each other for Winter and Autumn respectively.
@@
 The vertical scale is common for all subplots, for easier comparison.", out.width="100%"}
-knitr::include_graphics("./images/SeasonalTrendsTogether-1.png")
+knitr::include_graphics("./images/SeasonalTrendsTogether2-2.png")
 ```
@@
 ```{r trendseasontable, echo=F}
-pp <- read.csv("./figures/tbl_longterm_trends_season.dat",
-               strip.white = TRUE,
-               sep         = ";",
-               header      = TRUE)
-pp <- pp[grep("SDR" ,pp$var), ]
-pp$slope.sd           <- NULL
-pp$slope.ConfInt_0.95 <- NULL
-pp$slope.ConfInt_0.99 <- NULL
-pp$var                <- NULL
-rownames(pp)          <- NULL
-names(pp) <- c("Trend [%/year]", "Trend p-value", "Sky condition", "Season", "Statistical signif. [%]")
-## order by season
-target    <- c("Winter", "Spring", "Summer", "Autumn")
-new_order <- sapply(target, function(x,pp){which(pp$Season == x)}, pp = pp)
-pp        <- pp[new_order,]
-## order by sky condition
-pp <- pp[order(pp$`Sky condition`),]
-
-# ## put all sky first
-# A  <- which(pp$`Sky condition` == "All sky cond.")
-# B  <- which(pp$`Sky condition` != "All sky cond.")
-# pp <- pp[c(A,B),]
-## numeric precision
-pp$`Trend [%/year]`          <- signif(pp$`Trend [%/year]`, 3)
-pp$`Trend p-value`           <- signif(pp$`Trend p-value`,  3)
-pp$`Statistical signif. [%]` <- signif(pp$`Statistical signif. [%]`,  3)
-
-## rename sky conditions
-pp$`Sky condition` <- sub(" cond.", "", pp$`Sky condition`)
 
 ## arrange columns
-pp <- pp[,c(
+spp <- spp[,c(
       "Sky condition",
@@
 
-rownames(pp) <- NULL
+rownames(spp) <- NULL
 # panderOptions("table.alignment.default", "left")
 panderOptions("table.alignment.default", "center")
-pander(pp,
+pander(spp,
        justify = c('right', 'right', 'right', 'right', 'center'),
@@
 We demonstrate that in the period 
-`r paste(year(min(ALL_1_daily_DESEAS$Date)), " - ", year(max(ALL_1_daily_DESEAS$Date)))`, there is a brightening trend of SDR for all-sky conditions ($`r alltrend`\%/year$).
-On a previous study [@Bais2013], for the period 1993 - 2011 the trend was found to be $0.33\%/year$.
+`r paste(year(min(ALL_1_daily_DESEAS$Date)), " - ", year(max(ALL_1_daily_DESEAS$Date)))`, there is a brightening trend of SDR for all-sky conditions ($`r alltrend`\%/y$).
+On a previous study [@Bais2013], for the period 1993 - 2011 the trend was found to be $0.33\%/y$.
 The increase of this trend, indicates that the phenomenon and probably the causes, are still occurring.
-Accordingly, we found a similar trend for clear-sky conditions ($`r cleartrend`\%/year$) that further supports that the brightening can be independent of the cloud occurrences. 
+Accordingly, we found a similar trend for clear-sky conditions ($`r cleartrend`\%/y$) that further supports that the brightening can be independent of the cloud occurrences. 
 We have calculated the effect of TSI variability to be a negligible factor in SDR variability.
@@
 The attenuation of SDR by Aerosols over Europe have been proposed as major factor by @Wild2021.
-The dimming effect of SDR under cloud-sky condition ($`r cloudtrend`\%/year$), shows that clouds can also attenuate the amount of SDR reaching the ground, and changes on their occurrences can effect the net energy balance.
+The dimming effect of SDR under cloud-sky condition ($`r cloudtrend`\%/y$), shows that clouds can also attenuate the amount of SDR reaching the ground, and changes on their occurrences can effect the net energy balance.
 Because we have no adequate data to investigate long term changes of cloud formation in the region, we can not infer if the SDR trend we observe under cloud-sky is biased by changes in clouds and local climate.

### 2023-09-15 Article.Rmd
Commit 2023-09-15 14:14
Commit 2023-09-15 16:13

@@
 The solar zenith angle  is a major driver of the attenuation of SDR reaching the ground, as SDR is affected by the enhancement of the radiation path in the atmosphere, especially in urban environment with human activities emitting aerosols [@Wang2021].
-In order to estimate the contribution of the SZA, we group the anomaly data in $1^\circ$ SZA bins, and calculated the overall trend for each bin before noon and after noon (Figure\ \ref{fig:szatrends}).
+In order to estimate the effect of the SZA on the SDR trends, we grouped the anomaly data in bin of $1^\circ$ SZA, and calculated the overall trend for each bin before noon and after noon (Figure\ \ref{fig:szatrends}).
 Although there is a seasonal variation of the minimum SZA, that is not examined here.
@@
 close to the long term trend) both statistically significant above the $99\,\%$ confidence level. 
-In sprint and summer   , indicating a relation to the cloud pattern of each season.
+In sprint and summer, the trends are much smaller and less significant.
+These seasonal differences indicate a possible relation of the trends in clouds during winter and autumn.
+For clear-skies, the trend in winter is even larger than for all-skies ($`r signif(spp$"Trend [%/year]"[spp$Season == "Winter" & spp$"Sky condition" == "Clear sky"], 2)`\,\%/y$) which is another indication of the effects
 Moreover, when examining the clear- and cloud-sky conditions the trends between these conditions are almost complementary to each other for Winter and Autumn respectively.

### 2023-09-15 Article.Rmd
Commit 2023-09-15 17:10

@@
 cleartrend <- signif(LT[`Sky condition` == "Clear sky cond.",  `Trend [%/year]`], 2)
-cloudtrend <- signif(LT[`Sky condition` == "Cloudy cond.",    `Trend [%/year]`], 2)
+cloudtrend <- signif(LT[`Sky condition` == "Cloudy sky cond.", `Trend [%/year]`], 2)
 
@@
 In order to estimate the effect of the SZA on the SDR trends, we grouped the anomaly data in bin of $1^\circ$ SZA, and calculated the overall trend for each bin before noon and after noon (Figure\ \ref{fig:szatrends}).
-Although there is a seasonal variation of the minimum SZA, that is not examined here.
+Although there is a seasonal variation of the minimum SZA, its possible effect has not been examined here.
 The brightening effect of SDR is stronger for all-sky and clear-sky condition for low Sun elevation angles (Figures\ \ref{fig:szatrends}a and \ref{fig:szatrends}b).
@@
 
-\newpage
 # Appendix

### 2023-09-15 DHI_GHI_1_longterm_trends.R

@@ -250,7 +250,6 @@ for (DBn in dbs) {
                                      unlist(strsplit(avar, split = "_"))[1:2]),
                                    collapse = "_")),
                  cex      = 2,
-                 main     = paste(translate(DBn), translate(avar)),
                  cex.main = 0.8,
                  yaxt     = "n",
                  xlab     = "",
@@ -258,12 +257,18 @@ for (DBn in dbs) {
             )
             axis(2, pretty(dataset[[avar]]), las = 2 )
 
+            if (DRAFT == TRUE) {
+                title(main = paste(translate(DBn), translate(avar)),
+                      cex.main = 0.8 )
+            }
+
+
             # ylab = bquote("Deseas." ~ .(translate(avar)) ~ "[" ~ Watt/m^2 ~ "]" ) )
 
             ## plot fit line
             abline(lm1, lwd = 2)
 
-            if (DRAFT) {
+            if (DRAFT == TRUE) {
                 ## Running mean
                 first <- head(which(!is.na(dataset[[avar]])),1)
                 last  <- tail(which(!is.na(dataset[[avar]])),1)
@@ -287,7 +292,7 @@ for (DBn in dbs) {
             }
 
 
-            ## decorations
+            ## display trend on graph
             fit <- lm1[[1]]
 
             legend("top", lty = 1, bty = "n", lwd = 2, cex = 1,

### 2023-09-15 Article.Rmd
Commit 2023-09-15 21:14

@@
 Although there is a seasonal variation of the minimum SZA, its possible effect has not been examined here.
-The brightening effect of SDR is stronger for all-sky and clear-sky condition for low Sun elevation angles (Figures\ \ref{fig:szatrends}a and \ref{fig:szatrends}b).
+For all-sky and clear-sky conditions the brightening effect of SDR (positive trend) for large SZAs (Figures\ \ref{fig:szatrends}a and \ref{fig:szatrends}b).
 In the case of cloud-sky condition (Figure\ \ref{fig:szatrends}c), there is no strong correlation between SDR brightening and SZA, except for some low Sun angles, that occur only for a small part of the year, and thus we can not draw more conclusions from this analysis for the overall trends.
@@
 Because we have no adequate data to investigate long term changes of cloud formation in the region, we can not infer if the SDR trend we observe under cloud-sky is biased by changes in clouds and local climate.
-The observed brightening of SDR over Thessaloniki, is dependent on SZA (higher SZA, have stronger brightening trends).
+The observed brightening of SDR over Thessaloniki; is dependent on SZA (higher SZA, have stronger brightening trends).
 And also, with the seasons of the year, where Winter and Autumn show a significant strong trend, in contrast to Spring and Summer.
@@
 
-# References
+# References {-}
 
@@
 
-# Appendix
+# Appendix {-}
 

### 2023-09-16 Article.Rmd
Commit 2023-09-16 11:12

@@
 It is interesting to note, that for the observation period, the trend of the TSI is
-$`r format(TSIslopeAll[1,1], scientific = T, digits = 2)`\%/year$, and thus we can not attribute any major effect on SDR trend to Solar variability.
+$`r format(TSIslopeAll[1,1], scientific = T, digits = 2)`\,\%/y$, and thus we can not attribute any major effect on SDR trend to Solar variability.
 
@@
 Each point is a monthly mean calculated from daily means.
-The vertical scale is common for all subplots, for easier comparison.", out.width="100%"}
-knitr::include_graphics("./images/SeasonalTrendsTogether2-2.png")
+The vertical scale is common for all graphs, each column of graphs corresponds to a sky condition and each row to a season of the year.", out.width="100%"}
+knitr::include_graphics("./images/SeasonalTrendsTogether3-2.png")
 ```
@@
 We demonstrate that in the period 
-`r paste(year(min(ALL_1_daily_DESEAS$Date)), " - ", year(max(ALL_1_daily_DESEAS$Date)))`, there is a brightening trend of SDR for all-sky conditions ($`r alltrend`\%/y$).
-On a previous study [@Bais2013], for the period 1993 - 2011 the trend was found to be $0.33\%/y$.
+`r paste(year(min(ALL_1_daily_DESEAS$Date)), " - ", year(max(ALL_1_daily_DESEAS$Date)))`, there is a brightening trend of SDR for all-sky conditions ($`r alltrend`\,\%/y$).
+On a previous study [@Bais2013], for the period 1993 - 2011 the trend was found to be $0.33\,\%/y$.
 The increase of this trend, indicates that the phenomenon and probably the causes, are still occurring.
-Accordingly, we found a similar trend for clear-sky conditions ($`r cleartrend`\%/y$) that further supports that the brightening can be independent of the cloud occurrences. 
+Accordingly, we found a similar trend for clear-sky conditions ($`r cleartrend`\,\%/y$) that further supports that the brightening can be independent of the cloud occurrences. 
 We have calculated the effect of TSI variability to be a negligible factor in SDR variability.
@@
 The attenuation of SDR by Aerosols over Europe have been proposed as major factor by @Wild2021.
-The dimming effect of SDR under cloud-sky condition ($`r cloudtrend`\%/y$), shows that clouds can also attenuate the amount of SDR reaching the ground, and changes on their occurrences can effect the net energy balance.
+The dimming effect of SDR under cloud-sky condition ($`r cloudtrend`\,\%/y$), shows that clouds can also attenuate the amount of SDR reaching the ground, and changes on their occurrences can effect the net energy balance.
 Because we have no adequate data to investigate long term changes of cloud formation in the region, we can not infer if the SDR trend we observe under cloud-sky is biased by changes in clouds and local climate.

### 2023-09-16 Article.Rmd
Commit 2023-09-16 16:10

@@
 
-Although the trends are quite constant, using plots of the cumulative sum (CUSUM) we can reveal some structure within the trends and some differences between the sky conditions [@Regier2019]. 
+Although the trends suggest that raw data have an homogeneous behavior, using plots of the cumulative sum (CUSUM) of the computed anomalies, we can reveal some structure within the trends and some differences between the sky conditions [@Regier2019].
 On all three plots on Figure\ \ref{fig:cumsum}, we observe a downward part at the start of the period until about 2000, an almost steady part until about 2016 for all-sky and clear-sky conditions (Figures\ \ref{fig:cumsum-1}, \ref{fig:cumsum-2}) and until 2009 for cloud-sky conditions (Figure\ \ref{fig:cumsum-3}), and finally, a steep upward part until the present.
@@
 In our case, for all sky conditions, we observe periods where the CUSUMs remain relative steady, roughly around the middle of the study dataset.
-Another distinct feature of CUSUMs, is the difference, between the cloud-sky and the rest of the sky conditions, where they display a supplementary behavior. 
+Another distinct feature of the CUSUMs, is the different pattern of the cloudy-sky dataset which peaks around the middle of the period.
+The CUSUMs of the clear and the all-sky conditions exhibit similar patterns.
 The causes for these discrepancies haven't been investigated yet.

### 2023-09-16 Article.Rmd
Commit 2023-09-16 17:10

@@
 abstract: |
-  The shortwave downward solar irradiance (SDR).........
-  Human and climate.......
-  Brigthening dimming...............
-  We study 29 years of CHP1 data.........
-  Clear cloud id ................
-  Found trends ......
+  The shortwave downward solar irradiance (SDR) is an important factor that drives climate processes, production and can affect all living organisms.
+  While monitoring the long term variability, there are observations of upward and downward SDR trends on different locations around the world for different time periods.
+  Periods of positive tredns are refered as brightening periods and with negative tredns as dimming periods.
+  We studied 29 years of CHP1 data from Thessaloniki, Greece, under three sky conditions (clear sky, cloudy sky and all sky conditions), applying a cloud sky identification algorithm.
+  We found a positive trend for all-sky and clear-sky conditions, and also, investigated the consistency of those trend, the effect of the solar zenith angle, and the variation of the trends for the seasons of the year.
 keywords: 
@@
 
-We calculated the linear SDR trends, from the departures of the mean daily values from the seasonal values, and for the three sky conditions (Table\ \ref{tab:trendtable}).
-In Figure\ \ref{fig:trendALL} we present only the time series under all-sky conditions; the plots for clear-sky and cloud-sky conditions, are very similar and shown in the Appendix (Figures\ \ref{fig:trendCLEAR} and \ref{fig:trendCLOUD}).
-In the studied period, there is no significant break or change in the variability of the time series.
-Other studies for the European region report a change of the SDR slope, around 1980, a few years before the start of our records [@Wild2021; @Yuan2021; @Ohmura2009].
-It is interesting to note, that for the observation period, the trend of the TSI is
-$`r format(TSIslopeAll[1,1], scientific = T, digits = 2)`\,\%/y$, and thus we can not attribute any major effect on SDR trend to Solar variability.
-
-<!-- \footnotesize -->
-```{r trendtable, echo=F}
+```{r echo=F, eval=TRUE}
 LT <- data.table(read.csv("./figures/tbl_longterm_trends.dat",
@@
 ## no need all are 100%
+LTminSSig <- signif(min(LT$`Trend statistical signif. [%]`),4)
 LT$`Trend statistical signif. [%]` <- NULL
@@
 LT$`Sky condition` <- sub(" cond.", "", LT$`Sky condition`)
+```
 
+We calculated the linear SDR trends, from the departures of the mean daily values from the seasonal values, and for the three sky conditions (Table\ \ref{tab:trendtable}).
+In Figure\ \ref{fig:trendALL} we present only the time series under all-sky conditions; the plots for clear-sky and cloud-sky conditions, are very similar and shown in the Appendix (Figures\ \ref{fig:trendCLEAR} and\  \ref{fig:trendCLOUD}).
+We observe a positive trend for all-sky conditions 
+($`r alltrend`\,\%/y$), a very close but lower trend for clear-sky ($`r cleartrend`\,\%/y$) and a negative weaker trend for cloud-sky ($`r cloudtrend`\,\%/y$).
+In the studied period, there is no significant break or change in the variability of the time series.
+Other studies for the European region report a change of the SDR slope, around 1980, a few years before the start of our records [@Wild2021; @Yuan2021; @Ohmura2009].
+It is interesting to note, that for the observation period, the trend of the TSI is
+$`r format(TSIslopeAll[1,1], scientific = T, digits = 2)`\,\%/y$, and thus we can not attribute any major effect on SDR trend to Solar variability.
+
+<!-- \footnotesize -->
+```{r trendtable, echo=F}
 panderOptions("table.alignment.default", "center")
@@
        caption = paste0("\\label{tab:trendtable}Trends of daily SDR means, for different sky conditions for ", year(min(ALL_1_daily_DESEAS$Date)), " - ", year(max(ALL_1_daily_DESEAS$Date)), ".
-For all cases the statistical significance is above $99.9\\%$."))
+For all cases the statistical significance is above $", LTminSSig, "\\,\\%$."))
 ```
@@
 
-Although the trends suggest that raw data have an homogeneous behavior, using plots of the cumulative sum (CUSUM) of the computed anomalies, we can reveal some structure within the trends and some differences between the sky conditions [@Regier2019].
+Although the trends suggest that raw data have a homogeneous behavior, using plots of the cumulative sum (CUSUM) of the computed anomalies, we can reveal some structure within the trends and some differences between the sky conditions [@Regier2019].
 On all three plots on Figure\ \ref{fig:cumsum}, we observe a downward part at the start of the period until about 2000, an almost steady part until about 2016 for all-sky and clear-sky conditions (Figures\ \ref{fig:cumsum-1}, \ref{fig:cumsum-2}) and until 2009 for cloud-sky conditions (Figure\ \ref{fig:cumsum-3}), and finally, a steep upward part until the present.
@@
 The causes for these discrepancies haven't been investigated yet.
+
+.......
+cusums without trends?
+........
 

### 2023-09-17 Build_154/Article_B154.Rmd
Commit 2023-09-17 12:16

--- 2023-09-16 Article.Rmd
@@
 For all-sky and clear-sky conditions the brightening effect of SDR (positive trend) for large SZAs (Figures\ \ref{fig:szatrends}a and \ref{fig:szatrends}b).
-In the case of cloud-sky condition (Figure\ \ref{fig:szatrends}c), there is no strong correlation between SDR brightening and SZA, except for some low Sun angles, that occur only for a small part of the year, and thus we can not draw more conclusions from this analysis for the overall trends.
+The trends in the morning and afternoon hours are more or less consistent with small differences which can be attributed to systematic diurnal variations of aerosols, particularly during the warm period of the year [@Wang2021].
+
+For cloud-sky conditions (Figure\ \ref{fig:szatrends}c),there is no dependence of the trends with SZA.
+
+The trends are variable ranging within about $\pm 0.2\,\%/y$.
+
+Between $55^\circ$ and $75^\circ$ SZA the trends are mostly negative for cloud sky conditions and are possibly associated with stronger attenuation by clouds.
 
@@
 ```
+
+```{r  echo=F, fig.cap="Seasonal SZA"}
+knitr::include_graphics("./images/SzaTrendsSeas-1.png")
+```
+
+```{r  echo=F, fig.cap="Seasonal SZA"}
+knitr::include_graphics("./images/SzaTrendsSeas-2.png")
+```
+
+```{r  echo=F, fig.cap="Seasonal SZA"}
+knitr::include_graphics("./images/SzaTrendsSeas-3.png")
+```
+
+```{r  echo=F, fig.cap="Seasonal SZA"}
+knitr::include_graphics("./images/SzaTrendsSeas-4.png")
+```
+
+```{r  echo=F, fig.cap="Seasonal SZA"}
+knitr::include_graphics("./images/SzaTrendsSeas-5.png")
+```
+
+```{r  echo=F, fig.cap="Seasonal SZA"}
+knitr::include_graphics("./images/SzaTrendsSeas-6.png")
+```
+
+```{r  echo=F, fig.cap="Seasonal SZA"}
+knitr::include_graphics("./images/SzaTrendsSeas-7.png")
+```
+
+```{r  echo=F, fig.cap="Seasonal SZA"}
+knitr::include_graphics("./images/SzaTrendsSeas-8.png")
+```
+
+```{r  echo=F, fig.cap="Seasonal SZA"}
+knitr::include_graphics("./images/SzaTrendsSeas-9.png")
+```
+
+```{r  echo=F, fig.cap="Seasonal SZA"}
+knitr::include_graphics("./images/SzaTrendsSeas-10.png")
+```
+
+```{r  echo=F, fig.cap="Seasonal SZA"}
+knitr::include_graphics("./images/SzaTrendsSeas-11.png")
+```
+
+```{r  echo=F, fig.cap="Seasonal SZA"}
+knitr::include_graphics("./images/SzaTrendsSeas-12.png")
+```

### 2023-09-18 Article.Rmd
Commit 2023-09-18 13:13

@@
   We studied 29 years of CHP1 data from Thessaloniki, Greece, under three sky conditions (clear sky, cloudy sky and all sky conditions), applying a cloud sky identification algorithm.
-  We found a positive trend for all-sky and clear-sky conditions, and also, investigated the consistency of those trend, the effect of the solar zenith angle, and the variation of the trends for the seasons of the year.
+  We found a positive trend for all-sky and clear-sky conditions, and also, investigated the consistency of those trends, the effect of the solar zenith angle, and the variation of the trends for the seasons of the year.
+  We indentified that there are some anomalies in the long term SDR trends, for all sky conditions.
 keywords: 
@@
 
+## order columns
+LT <- LT[,c(
+      "Sky condition",
+      "Season",
+      "Trend [%/year]"
+      )]
+
 ## rename sky conditions
@@
 The causes for these discrepancies haven't been investigated yet.
+
+In order to compare any features within the trends, we created another set of CUSUMs plots, by subtracting the corresponding long term trend from the data, prior the CUSUM calculation (Figure\ \ref{fig:cumsumnotrend}).
+With this approach periods where the CUSUM diverge from zero can be interpreted as systematic variation of SDR, further from the climatological mean.
 
@@
 pander(spp,
-       justify = c('right', 'right', 'right', 'right', 'center'),
+       justify = c('center', 'right', 'right', 'right', 'center'),
        split.table = Inf,

### 2023-09-18 Article.Rmd
Commit 2023-09-18 14:14

@@
       "Sky condition",
-      "Season",
-      "Trend [%/year]"
+      "Trend [%/year]",
+      "Trend p-value"
       )]
@@
       "Trend [%/year]", 
-      "Trend p-value", 
+      # "Trend p-value", 
       "Statistical signif. [%]"

### 2023-09-18 Article.Rmd
Commit 2023-09-18 18:14

@@
 ```{r szatrends, echo = FALSE, fig.cap='Distribution of the SDR\'s long term trends by SZA before and after noon.
-The solid data points represent values with an acceptable statistical significance ($p<0.005$).', fig.subcap=c('All-sky conditions.', 'Clear-sky conditions.', 'Cloud-sky conditions.'), out.width='.35\\linewidth', out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3}
+The solid data points represent values with an acceptable statistical significance ($p<0.005$).', fig.subcap=c('All-sky conditions.', 'Clear-sky conditions.', 'Cloud-sky conditions.'), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3}
 knitr::include_graphics("./images/SzaTrends-13.png")
@@
 ```{r trendseasontable, echo=F}
-
 ## arrange columns
@@
 pander(spp,
-       justify = c('center', 'right', 'right', 'right', 'center'),
+       justify = c('center', 'right', 'right', 'center'),
        split.table = Inf,
@@
 
-```{r  echo=F, fig.cap="Seasonal SZA"}
+```{r  echo=F, fig.cap="Seasonal SZA", out.width='.32\\linewidth'}
 knitr::include_graphics("./images/SzaTrendsSeas-1.png")
@@
 
-```{r  echo=F, fig.cap="Seasonal SZA"}
+```{r  echo=F, fig.cap="Seasonal SZA", out.width='.32\\linewidth'}
 knitr::include_graphics("./images/SzaTrendsSeas-2.png")
@@
 
-```{r  echo=F, fig.cap="Seasonal SZA"}
+```{r  echo=F, fig.cap="Seasonal SZA", out.width='.32\\linewidth'}
 knitr::include_graphics("./images/SzaTrendsSeas-3.png")
@@
 
-```{r  echo=F, fig.cap="Seasonal SZA"}
+```{r  echo=F, fig.cap="Seasonal SZA", out.width='.32\\linewidth'}
 knitr::include_graphics("./images/SzaTrendsSeas-4.png")
@@
 
-```{r  echo=F, fig.cap="Seasonal SZA"}
+```{r  echo=F, fig.cap="Seasonal SZA", out.width='.32\\linewidth'}
 knitr::include_graphics("./images/SzaTrendsSeas-5.png")
@@
 
-```{r  echo=F, fig.cap="Seasonal SZA"}
+```{r  echo=F, fig.cap="Seasonal SZA", out.width='.32\\linewidth'}
 knitr::include_graphics("./images/SzaTrendsSeas-6.png")
@@
 
-```{r  echo=F, fig.cap="Seasonal SZA"}
+```{r  echo=F, fig.cap="Seasonal SZA", out.width='.32\\linewidth'}
 knitr::include_graphics("./images/SzaTrendsSeas-7.png")
@@
 
-```{r  echo=F, fig.cap="Seasonal SZA"}
+```{r  echo=F, fig.cap="Seasonal SZA", out.width='.32\\linewidth'}
 knitr::include_graphics("./images/SzaTrendsSeas-8.png")
@@
 
-```{r  echo=F, fig.cap="Seasonal SZA"}
+```{r  echo=F, fig.cap="Seasonal SZA", out.width='.32\\linewidth'}
 knitr::include_graphics("./images/SzaTrendsSeas-9.png")
@@
 
-```{r  echo=F, fig.cap="Seasonal SZA"}
+```{r  echo=F, fig.cap="Seasonal SZA", out.width='.32\\linewidth'}
 knitr::include_graphics("./images/SzaTrendsSeas-10.png")
@@
 
-```{r  echo=F, fig.cap="Seasonal SZA"}
+```{r  echo=F, fig.cap="Seasonal SZA", out.width='.32\\linewidth'}
 knitr::include_graphics("./images/SzaTrendsSeas-11.png")
@@
 
-```{r  echo=F, fig.cap="Seasonal SZA"}
+```{r  echo=F, fig.cap="Seasonal SZA", out.width='.32\\linewidth'}
 knitr::include_graphics("./images/SzaTrendsSeas-12.png")

### 2023-09-18 Article.Rmd
Commit 2023-09-18 19:11

@@
 library(data.table)
+library(kableExtra)
 ## load data from trends analysis
@@
 For all cases the statistical significance is above $", LTminSSig, "\\,\\%$."))
+
+x_latex <- knitr::kable(LT,
+                        format   = "latex",
+                        booktabs = T,
+                        escape   = T,
+                        # format.args = list(digits = 3),
+                        linesep  = "",
+                        # align    = "ccrr",
+                        caption  = "sssssss")
+x_latex %>%
+    collapse_rows(columns = 1, valign = "middle", latex_hline = "custom") %>%
+    kable_styling(latex_options = c("striped",
+                                    "hold_position")
+                  )
+
 ```
@@
 <!-- \footnotesize -->
+
 ```{r trendseasontable, echo=F}
@@
 # panderOptions("table.alignment.default", "left")
-panderOptions("table.alignment.default", "center")
-pander(spp,
-       justify = c('center', 'right', 'right', 'center'),
-       split.table = Inf,
+# panderOptions("table.alignment.default", "center")
+# pander(spp,
+#        justify = c('center', 'right', 'right', 'center'),
+#        split.table = Inf,
+#        caption = "\\label{tab:trendseasontable}SDR linear trends of monthly anomalies for each season of the year.")
+
+## use kable
+
+x_latex <- knitr::kable(spp,
+                        format   = "latex",
+                        booktabs = T,
+                        escape   = T,
+                        # format.args = list(digits = 3),
+                        linesep  = "",
+                        align    = "ccrr",
                         caption  = "\\label{tab:trendseasontable}SDR linear trends of monthly anomalies for each season of the year.")
+x_latex %>%
+    collapse_rows(columns = 1, valign = "middle", latex_hline = "custom") %>%
+    kable_styling(latex_options = c("striped",
+                                    "hold_position")
+                  ) 
+# position = "float_left"
+
 ```
+
 <!-- \normalsize -->

### 2023-09-18 Article.Rmd
Commit 2023-09-18 20:14

@@
 
-x_latex <- knitr::kable(LT,
-                        format   = "latex",
-                        booktabs = T,
-                        escape   = T,
-                        # format.args = list(digits = 3),
-                        linesep  = "",
-                        # align    = "ccrr",
-                        caption  = "sssssss")
-x_latex %>%
-    collapse_rows(columns = 1, valign = "middle", latex_hline = "custom") %>%
-    kable_styling(latex_options = c("striped",
-                                    "hold_position")
-                  )
+# x_latex <- knitr::kable(LT,
+#                         format   = "latex",
+#                         booktabs = T,
+#                         escape   = T,
+#                         # format.args = list(digits = 3),
+#                         linesep  = "",
+#                         align    = "ccc",
+#                         caption  = "test")
+# x_latex %>%
+#     kable_styling(latex_options = c("hold_position"))
 

### 2023-09-21 Article.Rmd
Commit 2023-09-21 11:10

@@
 library(kableExtra)
+library(dplyr)
 ## load data from trends analysis
@@
 rownames(LT) <- NULL
-names(LT)    <- c("Trend [%/year]", "Trend p-value", "Sky condition", "Trend statistical signif. [%]")
+names(LT)    <- c("Trend [%/year]", "Trend p-value", "Sky condition", "Statistical signif. [%]")
 ## no need all are 100%
-LTminSSig <- signif(min(LT$`Trend statistical signif. [%]`),4)
-LT$`Trend statistical signif. [%]` <- NULL
+LTminSSig <- signif(min(LT$`Statistical signif. [%]`), 4)
+# LT$`Trend statistical signif. [%]` <- NULL
 ## numeric precision
@@
 cloudtrend <- signif(LT[`Sky condition` == "Cloudy sky cond.", `Trend [%/year]`], 2)
+## rename sky conditions
+LT$`Sky condition` <- sub(" cond.", "", LT$`Sky condition`)
 
 ## order columns
-LT <- LT[,c(
+LT1 <- LT[,c(
       "Sky condition",
@@
 
-## rename sky conditions
-LT$`Sky condition` <- sub(" cond.", "", LT$`Sky condition`)
+## order columns
+LT2 <- LT[,c(
+      "Sky condition",
+      "Trend [%/year]",
+      "Statistical signif. [%]"
+      )]
+
 ```
@@
 panderOptions("table.alignment.default", "center")
-pander(LT,
+pander(LT1,
        caption = paste0("\\label{tab:trendtable}Trends of daily SDR means, for different sky conditions for ", year(min(ALL_1_daily_DESEAS$Date)), " - ", year(max(ALL_1_daily_DESEAS$Date)), ".
@@
 
-# x_latex <- knitr::kable(LT,
-#                         format   = "latex",
-#                         booktabs = T,
-#                         escape   = T,
-#                         # format.args = list(digits = 3),
-#                         linesep  = "",
-#                         align    = "ccc",
-#                         caption  = "test")
-# x_latex %>%
-#     kable_styling(latex_options = c("hold_position"))
+# LT2 %>% rename_all( ~ str_c("{ ",.," }"))
+x_latex <- knitr::kable(LT2,
+                        format   = "latex",
+                        booktabs = T,
+                        escape   = T,
+                        format.args = list(digits = 4),
+                        linesep  = "",
+                        align    = "crr",
+                        caption  = "test")
+x_latex %>%
+    kable_styling(
+        latex_options = c("hold_position")
+    )
+        # position      =   "float_left"
 
@@
 Although the trends suggest that raw data have a homogeneous behavior, using plots of the cumulative sum (CUSUM) of the computed anomalies, we can reveal some structure within the trends and some differences between the sky conditions [@Regier2019].
-On all three plots on Figure\ \ref{fig:cumsum}, we observe a downward part at the start of the period until about 2000, an almost steady part until about 2016 for all-sky and clear-sky conditions (Figures\ \ref{fig:cumsum-1}, \ref{fig:cumsum-2}) and until 2009 for cloud-sky conditions (Figure\ \ref{fig:cumsum-3}), and finally, a steep upward part until the present.
+On all three plots on Figure\ \ref{fig:cusum}, we observe a downward part at the start of the period until about 2000, an almost steady part until about 2016 for all-sky and clear-sky conditions (Figures\ \ref{fig:cusum-1}, \ref{fig:cusum-2}) and until 2009 for cloud-sky conditions (Figure\ \ref{fig:cusum-3}), and finally, a steep upward part until the present.
 For a uniform trend, we would expect the cumulative sums of the anomalies to have a symmetric 'V' shape.
@@
 
-In order to compare any features within the trends, we created another set of CUSUMs plots, by subtracting the corresponding long term trend from the data, prior the CUSUM calculation (Figure\ \ref{fig:cumsumnotrend}).
+In order to compare any features within the trends, we created another set of CUSUMs plots, by subtracting the corresponding long term trend from the data, prior the CUSUM calculation (Figure\ \ref{fig:cusumnotrend}).
 With this approach periods where the CUSUM diverge from zero can be interpreted as systematic variation of SDR, further from the climatological mean.
@@
 
-```{r cumsum, echo = FALSE, fig.cap='Running cumulative sum plots of the daily SDR for the three different sky conditions.', fig.subcap=c('All-sky conditions.', 'Clear-sky conditions.', 'Cloud-sky conditions.'), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3}
+```{r cusum, echo = FALSE, fig.cap='Running cumulative sum plots of the daily SDR for the three different sky conditions.', fig.subcap=c('All-sky conditions.', 'Clear-sky conditions.', 'Cloud-sky conditions.'), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3}
 knitr::include_graphics(c(
-    "./images/CumulativeDailyCumSum-1.png",    # ALL sky
-    "./images/CumulativeDailyCumSum-5.png",    # CLEAR sky
-    "./images/CumulativeDailyCumSum-9.png"))   # Cloud sky
+    "./images/CumulativeDailyCuSum-1.png",    # ALL sky
+    "./images/CumulativeDailyCuSum-5.png",    # CLEAR sky
+    "./images/CumulativeDailyCuSum-9.png"))   # Cloud sky
 ```
 
-```{r cumsumnotrend, echo = FALSE, fig.cap='Running cumulative sum plots of the daily SDR after removing the long term trend from the daily values, for the three different sky conditions.', fig.subcap=c('All-sky conditions.', 'Clear-sky conditions.', 'Cloud-sky conditions.'), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3}
+```{r cusumnotrend, echo = FALSE, fig.cap='Running cumulative sum plots of the daily SDR after removing the long term trend from the daily values, for the three different sky conditions.', fig.subcap=c('All-sky conditions.', 'Clear-sky conditions.', 'Cloud-sky conditions.'), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3}
 knitr::include_graphics(c(
-    "./images/CumulativeDailyCumSumNOtrend-1.png",    # ALL sky
-    "./images/CumulativeDailyCumSumNOtrend-5.png",    # CLEAR sky
-    "./images/CumulativeDailyCumSumNOtrend-9.png"))   # Cloud sky
+    "./images/CumulativeDailyCuSumNOtrend-1.png",    # ALL sky
+    "./images/CumulativeDailyCuSumNOtrend-5.png",    # CLEAR sky
+    "./images/CumulativeDailyCuSumNOtrend-9.png"))   # Cloud sky
 ```
 
-<!-- ```{r cumsumSTD, echo = FALSE, fig.cap='two plots', fig.subcap=c('one plot', 'the other one'), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3} -->
+<!-- ```{r cusumSTD, echo = FALSE, fig.cap='two plots', fig.subcap=c('one plot', 'the other one'), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3} -->
 <!-- knitr::include_graphics(c( -->
-<!--     "./images/CumulativeDailyCumSumSTD-1.png",    # ALL sky -->
-<!--     "./images/CumulativeDailyCumSumSTD-2.png",    # CLEAR sky -->
-<!--     "./images/CumulativeDailyCumSumSTD-3.png"))   # Cloud sky -->
+<!--     "./images/CumulativeDailyCuSumSTD-1.png",    # ALL sky -->
+<!--     "./images/CumulativeDailyCuSumSTD-2.png",    # CLEAR sky -->
+<!--     "./images/CumulativeDailyCuSumSTD-3.png"))   # Cloud sky -->
 <!-- ``` -->
@@
 <!--
-```{r cumsumALL, echo=F, fig.cap="Running cumulative sum."}
-knitr::include_graphics("./images/CumulativeDailyCumSum-1.png")
+```{r cusumALL, echo=F, fig.cap="Running cumulative sum."}
+knitr::include_graphics("./images/CumulativeDailyCuSum-1.png")
 ```
@@
 Not showing this trend
-```{r cumsumCLEAR, echo=F, fig.cap="Donditions.", include=FALSE}
-knitr::include_graphics("./images/CumulativeDailyCumSum-5.png")
+```{r cusumCLEAR, echo=F, fig.cap="Donditions.", include=FALSE}
+knitr::include_graphics("./images/CumulativeDailyCuSum-5.png")
 ```
@@
 <!--
-```{r cumsumCLOUD, echo=F, fig.cap="Donditions."}
-knitr::include_graphics("./images/CumulativeDailyCumSum-9.png")
+```{r cusumCLOUD, echo=F, fig.cap="Donditions."}
+knitr::include_graphics("./images/CumulativeDailyCuSum-9.png")
 ```

### 2023-09-21 DHI_GHI_3_trends_consistency.R

@@ -137,11 +137,11 @@ ccex_sbs <- 1.3
 
 ## ____ Scatter plots with SZA all data -----------------------------------------
 data_list <- c(  "ALL_1_D_monthly_DESEAS",
-               "CLEAR_1_D_monthly_DESEAS",
-               "CLOUD_1_D_monthly_DESEAS",
-                     "ALL_1_daily_DESEAS",
-                   "CLEAR_1_daily_DESEAS",
-                   "CLEAR_1_daily_DESEAS")
+                 "CLEAR_1_D_monthly_DESEAS",
+                 "CLOUD_1_D_monthly_DESEAS",
+                 "ALL_1_daily_DESEAS",
+                 "CLEAR_1_daily_DESEAS",
+                 "CLEAR_1_daily_DESEAS")
 
 by_var     <- c("Date", "SZA")
 for (i in data_list) {
@@ -215,8 +215,8 @@ trends <- readRDS("./figures/tbl_longterm_trends.Rds")
 dsets <- c("ALL_1_daily_DESEAS",
            "CLEAR_1_daily_DESEAS",
            "CLOUD_1_daily_DESEAS")
-warning("Removing trends from data before cumsum!!!")
-cat("Removing trends from data before cumsum!!!")
+warning("Removing trends from data before cusum!!!")
+cat("Removing trends from data before cusum!!!")
 for (ad in dsets) {
     DBS  <- get(ad)
     type <- sub("_.*", "", ad)
@@ -237,10 +237,10 @@ for (ad in dsets) {
 
 
 
-## Calculate cumsum with zeroing NA
-  ALL_1_daily_DESEAS[, GLB_att_cusum    := cumsum(tidyr::replace_na(GLB_att_des,    0))]
-  ALL_1_daily_DESEAS[, DIR_att_cusum    := cumsum(tidyr::replace_na(DIR_att_des,    0))]
-  ALL_1_daily_DESEAS[, DIR_transp_cusum := cumsum(tidyr::replace_na(DIR_transp_des, 0))]
+## Calculate cusum with zeroing NA
+ALL_1_daily_DESEAS[, GLB_att_cusum    := cumsum(tidyr::replace_na(GLB_att_des,    0))]
+ALL_1_daily_DESEAS[, DIR_att_cusum    := cumsum(tidyr::replace_na(DIR_att_des,    0))]
[…]

### 2023-09-21 Article.Rmd
Commit 2023-09-21 12:14

@@
 
-In order to compare any features within the trends, we created another set of CUSUMs plots, by subtracting the corresponding long term trend from the data, prior the CUSUM calculation (Figure\ \ref{fig:cusumnotrend}).
+In order to compare any features within the trends, we created another set of CUSUMs plots, by subtracting the corresponding long term trend from the SDR anomaly data, prior to the CUSUM calculation (Figure\ \ref{fig:cusumnotrend}).
 With this approach periods where the CUSUM diverge from zero can be interpreted as systematic variation of SDR, further from the climatological mean.
+It is interesting that for the period 1993...... diverge around zero 
+and after ~ for all-sky and clear-sky
+
+For cloud-sky .....
 
@@
 
-```{r  echo=F, fig.cap="Seasonal SZA", out.width='.32\\linewidth'}
+```{r echo = FALSE, fig.cap='Distribu.', fig.subcap=c('', '', ''), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3}
 knitr::include_graphics("./images/SzaTrendsSeas-1.png")
-```
-
-```{r  echo=F, fig.cap="Seasonal SZA", out.width='.32\\linewidth'}
 knitr::include_graphics("./images/SzaTrendsSeas-2.png")
-```
-
-```{r  echo=F, fig.cap="Seasonal SZA", out.width='.32\\linewidth'}
 knitr::include_graphics("./images/SzaTrendsSeas-3.png")
-```
-
-```{r  echo=F, fig.cap="Seasonal SZA", out.width='.32\\linewidth'}
 knitr::include_graphics("./images/SzaTrendsSeas-4.png")
-```
-
-```{r  echo=F, fig.cap="Seasonal SZA", out.width='.32\\linewidth'}
 knitr::include_graphics("./images/SzaTrendsSeas-5.png")
-```
-
-```{r  echo=F, fig.cap="Seasonal SZA", out.width='.32\\linewidth'}
 knitr::include_graphics("./images/SzaTrendsSeas-6.png")
-```
-
-```{r  echo=F, fig.cap="Seasonal SZA", out.width='.32\\linewidth'}
 knitr::include_graphics("./images/SzaTrendsSeas-7.png")
-```
-
-```{r  echo=F, fig.cap="Seasonal SZA", out.width='.32\\linewidth'}
 knitr::include_graphics("./images/SzaTrendsSeas-8.png")
-```
-
-```{r  echo=F, fig.cap="Seasonal SZA", out.width='.32\\linewidth'}
 knitr::include_graphics("./images/SzaTrendsSeas-9.png")
-```
-
-```{r  echo=F, fig.cap="Seasonal SZA", out.width='.32\\linewidth'}
 knitr::include_graphics("./images/SzaTrendsSeas-10.png")
-```
-
-```{r  echo=F, fig.cap="Seasonal SZA", out.width='.32\\linewidth'}
 knitr::include_graphics("./images/SzaTrendsSeas-11.png")
-```
-
-```{r  echo=F, fig.cap="Seasonal SZA", out.width='.32\\linewidth'}
 knitr::include_graphics("./images/SzaTrendsSeas-12.png")

### 2023-09-21 Article.Rmd
Commit 2023-09-21 17:10

@@
 With this approach periods where the CUSUM diverge from zero can be interpreted as systematic variation of SDR, further from the climatological mean.
-It is interesting that for the period 1993...... diverge around zero 
-and after ~ for all-sky and clear-sky
+When the Anomaly CUSUM is increasing, the added values are above the expected climatological values of SDR trend, respectively when it decreases the SDR trend is lower than the climatological trend.
+
+Overall, for all- and clear-sky conditions we observe that there are periods where the anomalies diverge from the expected climatological value, that each lasts for few only years, and then diverge to the opposite site, creating a variation around zero.
+It is interesting that in the period 1993 - 2016 the anomalies of all-sky and clear sky, have a high variability around zero and after 2016, the range of the variability is about one third of the prior period.
+Even more strong fluctuations exist for cloud-sky conditions, where there are long period with relative low variation, and few periods with very strong divergences.
+We can identify two very positive periods (1997 - 2005 and 2020 - 2023) and one very negative (2005 - 2010).
 

### 2023-09-21 Article.Rmd
Commit 2023-09-21 18:14

--- 2023-09-21 Article.Rmd
@@
 
-# LT2 %>% rename_all( ~ str_c("{ ",.," }"))
+```
+
+```{r trendtable2, echo=F}
 x_latex <- knitr::kable(LT2,
@@
 With this approach periods where the CUSUM diverge from zero can be interpreted as systematic variation of SDR, further from the climatological mean.
-It is interesting that for the period 1993...... diverge around zero 
-and after ~ for all-sky and clear-sky
-
-For cloud-sky .....
-
-.......
-cusums without trends?
-........
+When the Anomaly CUSUM is increasing, the added values are above the expected climatological values of SDR trend, respectively when it decreases the SDR trend is lower than the climatological trend.
+Overall, for all- and clear-sky conditions (Figure\ \ref{fig:cusumnotrend-1} and\ \ref{fig:cusumnotrend-2}) we observe that there are periods where the anomalies diverge from the expected climatological value, that each lasts for few only years, and then diverge to the opposite site, creating a variation around zero.
+It is interesting that in the period 1993 - 2016 the anomalies of all-sky and clear- sky, have a high variability around zero.
+After 2016, the range of the variability is decreased at about one third of the prior period.
+Even more strong fluctuations exist for cloud-sky conditions (Figure\ \ref{fig:cusumnotrend-3}), where there are long periods with relative low variation, and few periods with very strong divergences.
+We can identify two very positive periods (1997 - 2005 and 2020 - 2023) and one very negative (2005 - 2010).
 
@@
 
-<!--
-more about trend in relation to other trends
-Aerosols and clouds
--->
-
 ## Effects of the solar zenith angle on SDR.
@@
 Although there is a seasonal variation of the minimum SZA, its possible effect has not been examined here.
-For all-sky and clear-sky conditions the brightening effect of SDR (positive trend) for large SZAs (Figures\ \ref{fig:szatrends}a and \ref{fig:szatrends}b).
+For all-sky and clear-sky conditions the brightening effect of SDR (positive trend) for large SZAs (Figures\ \ref{fig:szatrends-1} and \ref{fig:szatrends-2}).
 The trends in the morning and afternoon hours are more or less consistent with small differences which can be attributed to systematic diurnal variations of aerosols, particularly during the warm period of the year [@Wang2021].
-
-For cloud-sky conditions (Figure\ \ref{fig:szatrends}c),there is no dependence of the trends with SZA.
-
-The trends are variable ranging within about $\pm 0.2\,\%/y$.
-
-Between $55^\circ$ and $75^\circ$ SZA the trends are mostly negative for cloud sky conditions and are possibly associated with stronger attenuation by clouds.
+For cloud-sky conditions (Figure\ \ref{fig:szatrends-3}), we can not discern any significant dependence of the SDR trend with SZA.
+For SZAs $16^\circ$ - $50^\circ$, the trends are ranging within about $\pm 0.2\,\%/y$, with a weak statistical significance.
+Between $55^\circ$ and $75^\circ$ SZA the trends are mostly negative and are possibly associated with stronger attenuation by clouds.
 
@@
 
-<!--
-## Consistency of the trends
--->
-
-<!--
-```{r cusumALL, echo=F, fig.cap="Running cumulative sum."}
-knitr::include_graphics("./images/CumulativeDailyCuSum-1.png")
-```
--->
-
-<!--
-Not showing this trend
-```{r cusumCLEAR, echo=F, fig.cap="Donditions.", include=FALSE}
-knitr::include_graphics("./images/CumulativeDailyCuSum-5.png")
-```
--->
-
-<!--
-```{r cusumCLOUD, echo=F, fig.cap="Donditions."}
-knitr::include_graphics("./images/CumulativeDailyCuSum-9.png")
-```
--->
-
 # Conclusions
@@
 
-```{r echo = FALSE, fig.cap='Distribu.', fig.subcap=c('', '', ''), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3}
+```{r echo = FALSE, fig.cap='SDR trend ~ SZA', fig.subcap=c('', '', ''), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3}
 knitr::include_graphics("./images/SzaTrendsSeas-1.png")

### 2023-09-21 DHI_GHI_2_sza_trends.R

@@ -338,11 +338,12 @@ for (avar in unique(szatrends$var)) {
             if (DRAFT == TRUE) {
                 par("mar" = c(4,   5,   2,   2))
             } else {
-                par("mar" = c(4.5, 4.5, 0.5, 0.5))
+                par("mar" = c(4, 4.5, 0.5, 0.5))
             }
 
+            ## empty plot
             plot(1, type = "n",
-                 xlab = "Solar zenith angle",
+                 xlab = "",
                  ylab = awename,
                  xlim = xlim,
                  ylim = ylim,
@@ -354,9 +355,10 @@ for (avar in unique(szatrends$var)) {
             ## x axis
             axis(1, at = seq(xlim[1], xlim[2]), labels = NA,
                       tcl = -0.25)
+            title(xlab = bquote("Solar zenith angle (SZA)"),
+                  line = 2.5)
 
-
-
+            ## zero line
             abline(h = 0, lty = 3)
 
             if (DRAFT == TRUE) {
@@ -599,7 +601,7 @@ for (ase in seasons) {
 
 
                 plot(1, type = "n",
-                     xlab = "Solar zenith angle",
+                     xlab = "",
                      ylab = awename,
                      xlim = xlim,
                      ylim = ylim,
@@ -611,7 +613,10 @@ for (ase in seasons) {
                 ## x axis
                 axis(1, at = seq(xlim[1], xlim[2]), labels = NA,
                      tcl = -0.25)
+                title(xlab = bquote("Solar zenith angle (SZA)"),
+                      line = 2.5)
 
+                ## zero line
                 abline(h = 0, lty = 3)
 
                 ## test for some plots
@@ -620,11 +625,6 @@ for (ase in seasons) {
 
                 title(paste(ase, awename, typeP, translate(avar)), cex.main = 0.8,)
 
-                # lines(pam$SZA, pam[[awe]], pch =  pch_am, col = pam$col, type = "b")
-                # lines(pam$SZA, ppm[[awe]], pch =  pch_pm, col = pam$col, type = "b")
-
-                # lines(pam$SZA, pam[[awe]], pch = pch_am, col = 2, type = "b", lwd = 1, cex = .5)
-                # lines(ppm$SZA, ppm[[awe]], pch = pch_pm, col = 3, type = "b", lwd = 1, cex = .5)
 
 
                 ## morning lines

### 2023-10-06 Article.Rmd
Commit 2023-10-06 11:10

@@
 rownames(LT) <- NULL
-names(LT)    <- c("Trend [%/year]", "Trend p-value", "Sky condition", "Statistical signif. [%]")
+names(LT)    <- c("Trend [%/year]", "Trend p-value", "Sky conditions", "Statistical signif. [%]")
 ## no need all are 100%
@@
 ## some vars to use later
-alltrend   <- signif(LT[`Sky condition` == "All sky cond.",    `Trend [%/year]`], 2)
-cleartrend <- signif(LT[`Sky condition` == "Clear sky cond.",  `Trend [%/year]`], 2)
-cloudtrend <- signif(LT[`Sky condition` == "Cloudy sky cond.", `Trend [%/year]`], 2)
+alltrend   <- signif(LT[`Sky conditions` == "All sky cond.",    `Trend [%/year]`], 2)
+cleartrend <- signif(LT[`Sky conditions` == "Clear sky cond.",  `Trend [%/year]`], 2)
+cloudtrend <- signif(LT[`Sky conditions` == "Cloudy sky cond.", `Trend [%/year]`], 2)
 ## rename sky conditions
-LT$`Sky condition` <- sub(" cond.", "", LT$`Sky condition`)
+LT$`Sky conditions` <- sub(" cond.", "", LT$`Sky conditions`)
 
@@
 LT1 <- LT[,c(
-      "Sky condition",
+      "Sky conditions",
       "Trend [%/year]",
@@
 LT2 <- LT[,c(
-      "Sky condition",
+      "Sky conditions",
       "Trend [%/year]",
@@
 We calculated the linear SDR trends, from the departures of the mean daily values from the seasonal values, and for the three sky conditions (Table\ \ref{tab:trendtable}).
-In Figure\ \ref{fig:trendALL} we present only the time series under all-sky conditions; the plots for clear-sky and cloud-sky conditions, are very similar and shown in the Appendix (Figures\ \ref{fig:trendCLEAR} and\  \ref{fig:trendCLOUD}).
+In Figure\ \ref{fig:trendALL} we present only the time series under all-sky conditions; the plots for clear-sky and cloud-sky conditions, are very similar and are shown in the Appendix (Figures\ \ref{fig:trendCLEAR} and\  \ref{fig:trendCLOUD}).
 We observe a positive trend for all-sky conditions 
-($`r alltrend`\,\%/y$), a very close but lower trend for clear-sky ($`r cleartrend`\,\%/y$) and a negative weaker trend for cloud-sky ($`r cloudtrend`\,\%/y$).
+($`r alltrend`\,\%/y$), a very close but lower trend for clear-skies ($`r cleartrend`\,\%/y$) and a negative weaker trend for cloud-skies ($`r cloudtrend`\,\%/y$).
 In the studied period, there is no significant break or change in the variability of the time series.
@@
 <!-- \footnotesize -->
+<!-- ```{r trendtable2, echo=F} -->
+<!-- panderOptions("table.alignment.default", "center") -->
+<!-- pander(LT1, -->
+<!--        caption = paste0("\\label{tab:trendtable2}Trends in SDR daily mean for different sky conditions for the period", year(min(ALL_1_daily_DESEAS$Date)), " - ", year(max(ALL_1_daily_DESEAS$Date)), ".
+For all cases the statistical significance is above $", LTminSSig, "\\,\\%$.")) -->
+<!-- ``` -->
+<!-- \normalsize -->
+
 ```{r trendtable, echo=F}
-panderOptions("table.alignment.default", "center")
-pander(LT1,
-       caption = paste0("\\label{tab:trendtable}Trends of daily SDR means, for different sky conditions for ", year(min(ALL_1_daily_DESEAS$Date)), " - ", year(max(ALL_1_daily_DESEAS$Date)), ".
-For all cases the statistical significance is above $", LTminSSig, "\\,\\%$."))
-
-```
-
-```{r trendtable2, echo=F}
 x_latex <- knitr::kable(LT2,
@@
                         align    = "crr",
-                        caption  = "test")
+                        caption  = paste0("\\label{tab:trendtable}Trends in SDR daily mean for different sky conditions for the period ", year(min(ALL_1_daily_DESEAS$Date)), " - ", year(max(ALL_1_daily_DESEAS$Date)), "."))
 x_latex %>%
@@
         # position      =   "float_left"
-
 ```
-<!-- \normalsize -->
 
 ```{r trendALL, echo=F, out.width='.70\\linewidth', fig.cap=paste0("Anomaly (\\%) of the daily all-sky SDR, relative to climatological values for ", year(min(ALL_1_daily_DESEAS$Date)), " - ", year(max(ALL_1_daily_DESEAS$Date)), ".
-The black line shows the long term linear trend for all-sky conditions.")}
+The black line shows the long term linear trend.")}
 knitr::include_graphics("./images/LongtermTrends-2.png")

### 2023-10-08 Article.Rmd
Commit 2023-10-08 20:11

@@
 
-```{r cusum, echo = FALSE, fig.cap='Running cumulative sum plots of the daily SDR for the three different sky conditions.', fig.subcap=c('All-sky conditions.', 'Clear-sky conditions.', 'Cloud-sky conditions.'), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3}
+```{r cusum, echo = FALSE, fig.cap='Running cumulative sum plots of the daily SDR.', fig.subcap=c('All-sky conditions.', 'Clear-sky conditions.', 'Cloud-sky conditions.'), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3}
 knitr::include_graphics(c(
@@
     "./images/CumulativeDailyCuSum-9.png"))   # Cloud sky
+```
+
+```{r cusummonth, echo = FALSE, fig.cap='Running cumulative sum plots of the monthly SDR.', fig.subcap=c('All-sky conditions.', 'Clear-sky conditions.', 'Cloud-sky conditions.'), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3}
+knitr::include_graphics(c(
+    "./images/CumulativeMonthlyCuSum-1.png",    # ALL sky
+    "./images/CumulativeMonthlyCuSum-5.png",    # CLEAR sky
+    "./images/CumulativeMonthlyCuSum-9.png"))   # Cloud sky
 ```
@@
 ```
+
+<!-- ```{r cusumnotrendmonthly, echo = FALSE, fig.cap='Running cumulative sum plots of the monthly SDR after removing the long term trend from the monthly values, for the three different sky conditions.', fig.subcap=c('All-sky conditions.', 'Clear-sky conditions.', 'Cloud-sky conditions.'), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3} -->
+<!-- knitr::include_graphics(c( -->
+<!--     "./images/CumulativeMonthlyCuSum",    # ALL sky -->
+<!--     "./images/CumulativeMonthlyCuSumNOtrend-5.png",    # CLEAR sky -->
+<!--     "./images/CumulativeMonthlyCuSumNOtrend-9.png"))   # Cloud sky -->
+<!-- ``` -->
 

### 2023-10-08 Article.Rmd
Commit 2023-10-08 21:09

@@
 
-```{r cusumnotrend, echo = FALSE, fig.cap='Running cumulative sum plots of the daily SDR after removing the long term trend from the daily values, for the three different sky conditions.', fig.subcap=c('All-sky conditions.', 'Clear-sky conditions.', 'Cloud-sky conditions.'), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3}
+```{r cusumnotrend, echo = FALSE, fig.cap='Running cumulative sum plots of the daily SDR after removing the long term trend from the daily values.', fig.subcap=c('All-skies.', 'Clear-skies.', 'Cloud-skies.'), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3}
 knitr::include_graphics(c(
@@
 
-<!-- ```{r cusumnotrendmonthly, echo = FALSE, fig.cap='Running cumulative sum plots of the monthly SDR after removing the long term trend from the monthly values, for the three different sky conditions.', fig.subcap=c('All-sky conditions.', 'Clear-sky conditions.', 'Cloud-sky conditions.'), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3} -->
-<!-- knitr::include_graphics(c( -->
-<!--     "./images/CumulativeMonthlyCuSum",    # ALL sky -->
-<!--     "./images/CumulativeMonthlyCuSumNOtrend-5.png",    # CLEAR sky -->
-<!--     "./images/CumulativeMonthlyCuSumNOtrend-9.png"))   # Cloud sky -->
-<!-- ``` -->
+```{r cusumnotrendmonthly, echo = FALSE, fig.cap='Running cumulative sum plots of the monthly SDR after removing the long term trend from the monthly values.', fig.subcap=c('All-skies.', 'Clear-skies.', 'Cloud-skies.'), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3}
+knitr::include_graphics(c(
+    "./images/CumulativeMonthlyCuSumNOtrend-1.png",    # ALL sky
+    "./images/CumulativeMonthlyCuSumNOtrend-5.png",    # CLEAR sky
+    "./images/CumulativeMonthlyCuSumNOtrend-9.png"))   # Cloud sky
+```
 

### 2023-10-08 DHI_GHI_2_sza_trends.R

@@ -723,6 +723,233 @@ for (ase in seasons) {
 }
 
 
+stop()
+
+
+## __ by season in a tight grid ----------------------------------------------
+#+ SzaTrendsSeasTogether, echo=F, include=T
+{
+    vars        <- c("GLB_att_des")
+    avar        <- vars[1]
+    dbs         <- c(  "ALL_1_D_bySeason_DESEAS",
+                       "CLEAR_1_D_bySeason_DESEAS",
+                       "CLOUD_1_D_bySeason_DESEAS")
+    Seasons     <- c("Winter", "Spring", "Summer", "Autumn")
+
+    ## the order must be predefined to match
+    expanded <- expand.grid(Dataset = dbs, Seasons = Seasons, stringsAsFactors = FALSE)
+
+    nf <- layout(
+        matrix(1:30, ncol = 5, byrow = TRUE),
+        widths  = c(0.3,   1,1,1, 0.1),
+        heights = c(0.2, 1,1,1,1, 0.4)
+    )
+    layout.show(nf)
+
+    # 1
+    par("mar"=c(0,0,0,0))
+    plot.new()
+    # 2
+    plot.new()
+    text(x = 0.5, y = 0.5,
+         adj  = c(0.6,0.5),
+         "All skies",    cex = 0.9, font = 2)
+
+    # 3
+    plot.new()
+    text(x = 0.5, y = 0.5,
+         adj  = c(0.5,0.5),
+         "Clear skies",  cex = 0.9, font = 2)
+
+    # 4
+    plot.new()
+    text(x = 0.5, y = 0.5,
+         adj  = c(0.5,0.5),
+         "Cloudy skies", cex = 0.9, font = 2)
+
+    # 5
+    plot.new()
+
+
+
+    for (i  in 6:25) {
+
+        if (i == 6) {
+            plot.new()
+            text(x = 0.1, y = 0.5,
+                 adj  = c(0.5, 0.5),
+                 srt  = 90, "Winter", cex = 0.9, font = 2)
+        }
+        if (i == 10) {
+            plot.new()
+            text(x = 0.5, y = 0.5,
+                 adj  = c(0.5, 0.5),
+                 srt  = 90, "Winter", cex = 0.9, font = 2)
+        }
+
+        if (i == 11) {
+            plot.new()
+            text(x = 0.1, y = 0.5,
+                 adj  = c(0.5,0.5),
+                 srt  = 90, "Summer", cex = 0.9, font = 2)
[…]

### 2023-10-09 Article.Rmd
Commit 2023-10-09 12:14
Commit 2023-10-09 13:11
Commit 2023-10-09 17:11

@@
 
-```{r echo = FALSE, fig.cap='SDR trend ~ SZA', fig.subcap=c('', '', ''), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3}
-knitr::include_graphics("./images/SzaTrendsSeas-1.png")
-knitr::include_graphics("./images/SzaTrendsSeas-2.png")
-knitr::include_graphics("./images/SzaTrendsSeas-3.png")
-knitr::include_graphics("./images/SzaTrendsSeas-4.png")
-knitr::include_graphics("./images/SzaTrendsSeas-5.png")
-knitr::include_graphics("./images/SzaTrendsSeas-6.png")
-knitr::include_graphics("./images/SzaTrendsSeas-7.png")
-knitr::include_graphics("./images/SzaTrendsSeas-8.png")
-knitr::include_graphics("./images/SzaTrendsSeas-9.png")
-knitr::include_graphics("./images/SzaTrendsSeas-10.png")
-knitr::include_graphics("./images/SzaTrendsSeas-11.png")
-knitr::include_graphics("./images/SzaTrendsSeas-12.png")
+<!-- ```{r echo = FALSE, fig.cap='SDR trend ~ SZA', fig.subcap=c('', '', ''), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3} -->
+<!-- knitr::include_graphics("./images/SzaTrendsSeas-1.png") -->
+<!-- knitr::include_graphics("./images/SzaTrendsSeas-2.png") -->
+<!-- knitr::include_graphics("./images/SzaTrendsSeas-3.png") -->
+<!-- knitr::include_graphics("./images/SzaTrendsSeas-4.png") -->
+<!-- knitr::include_graphics("./images/SzaTrendsSeas-5.png") -->
+<!-- knitr::include_graphics("./images/SzaTrendsSeas-6.png") -->
+<!-- knitr::include_graphics("./images/SzaTrendsSeas-7.png") -->
+<!-- knitr::include_graphics("./images/SzaTrendsSeas-8.png") -->
+<!-- knitr::include_graphics("./images/SzaTrendsSeas-9.png") -->
+<!-- knitr::include_graphics("./images/SzaTrendsSeas-10.png") -->
+<!-- knitr::include_graphics("./images/SzaTrendsSeas-11.png") -->
+<!-- knitr::include_graphics("./images/SzaTrendsSeas-12.png") -->
+<!-- ``` -->
+
+```{r echo = FALSE, fig.cap='SDR trend ~ SZA', out.width='100%'}
+knitr::include_graphics("./images/SzaTrendsSeasTogether-2.png")
 ```

### 2023-10-10 Article.Rmd
Commit 2023-10-10 09:14

@@
 
-Although the trends suggest that raw data have a homogeneous behavior, using plots of the cumulative sum (CUSUM) of the computed anomalies, we can reveal some structure within the trends and some differences between the sky conditions [@Regier2019].
+Although the year-to-year variability of the anomalies in Figure \ref{fig:trendALL} shows a rather homogeneous behavior, plots of the cumulative sums (CUSUM) of the anomalies can reveal structures which are different for the three sky conditions [@Regier2019].
 On all three plots on Figure\ \ref{fig:cusum}, we observe a downward part at the start of the period until about 2000, an almost steady part until about 2016 for all-sky and clear-sky conditions (Figures\ \ref{fig:cusum-1}, \ref{fig:cusum-2}) and until 2009 for cloud-sky conditions (Figure\ \ref{fig:cusum-3}), and finally, a steep upward part until the present.
@@
 
-<!-- ```{r echo = FALSE, fig.cap='SDR trend ~ SZA', fig.subcap=c('', '', ''), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3} -->
-<!-- knitr::include_graphics("./images/SzaTrendsSeas-1.png") -->
-<!-- knitr::include_graphics("./images/SzaTrendsSeas-2.png") -->
-<!-- knitr::include_graphics("./images/SzaTrendsSeas-3.png") -->
-<!-- knitr::include_graphics("./images/SzaTrendsSeas-4.png") -->
-<!-- knitr::include_graphics("./images/SzaTrendsSeas-5.png") -->
-<!-- knitr::include_graphics("./images/SzaTrendsSeas-6.png") -->
-<!-- knitr::include_graphics("./images/SzaTrendsSeas-7.png") -->
-<!-- knitr::include_graphics("./images/SzaTrendsSeas-8.png") -->
-<!-- knitr::include_graphics("./images/SzaTrendsSeas-9.png") -->
-<!-- knitr::include_graphics("./images/SzaTrendsSeas-10.png") -->
-<!-- knitr::include_graphics("./images/SzaTrendsSeas-11.png") -->
-<!-- knitr::include_graphics("./images/SzaTrendsSeas-12.png") -->
-<!-- ``` -->
-
 ```{r echo = FALSE, fig.cap='SDR trend ~ SZA', out.width='100%'}

### 2023-10-10 Article.Rmd
Commit 2023-10-10 10:14

@@
 π.χ. https://link.springer.com/article/10.1007/s00382-022-06438-3
--->
-
-<!--
-more about trend in relation to other trends
-Aerosols and clouds
 -->
@@
 In order to compare any features within the trends, we created another set of CUSUMs plots, by subtracting the corresponding long term trend from the SDR anomaly data, prior to the CUSUM calculation (Figure\ \ref{fig:cusumnotrend}).
-With this approach periods where the CUSUM diverge from zero can be interpreted as systematic variation of SDR, further from the climatological mean.
-When the Anomaly CUSUM is increasing, the added values are above the expected climatological values of SDR trend, respectively when it decreases the SDR trend is lower than the climatological trend.
-Overall, for all- and clear-sky conditions (Figure\ \ref{fig:cusumnotrend-1} and\ \ref{fig:cusumnotrend-2}) we observe that there are periods where the anomalies diverge from the expected climatological value, that each lasts for few only years, and then diverge to the opposite site, creating a variation around zero.
+With this approach periods when the CUSUMs diverge from zero can be interpreted as systematic variation of SDR from the climatological mean.
+When the CUSUM is increasing, the added values are above the climatological values of SDR trend and vice versa.
+Overall, for all- and clear-sky conditions (Figure\ \ref{fig:cusumnotrend-1} and\ \ref{fig:cusumnotrend-2}) we observe periods when the anomalies diverge from the climatological value, each lasting for several years.
+The pattern for the all sky and clear sky conditions is very similar, suggesting prevalence in clear skies over Thessaloniki.
 It is interesting that in the period 1993 - 2016 the anomalies of all-sky and clear- sky, have a high variability around zero.
 After 2016, the range of the variability is decreased at about one third of the prior period.
-Even more strong fluctuations exist for cloud-sky conditions (Figure\ \ref{fig:cusumnotrend-3}), where there are long periods with relative low variation, and few periods with very strong divergences.
-We can identify two very positive periods (1997 - 2005 and 2020 - 2023) and one very negative (2005 - 2010).
-
-<!--
-This can make us to suspected that clear-Sky conditions SDR and cloud-sky SDR is affected by different mechanisms.
-The similarity of the all-sky and clear-sky conditions, can make us to suspected that on both sky conditions the attenuation of SDR has .
--->
+For cloud-sky conditions (Figure\ \ref{fig:cusumnotrend-3}) the period 1997-2008 is dominated by positive CUSUMs suggesting a reduced effect of clouds on SDR.
+From 1997 to mid-2000s CUSUMs are increasing, likely due to a continuous decrease in the thickness of clouds, followed by a period of rapid increase (within 3 years) in cloud optical thickness lasting up to 2008.
+The following stable period spans for about 15 years up to 2021 when CUSUMs start increasing again.
 
@@
 The solar zenith angle is a major driver of the attenuation of SDR reaching the ground, as SDR is affected by the enhancement of the radiation path in the atmosphere, especially in urban environment with human activities emitting aerosols [@Wang2021].
-In order to estimate the effect of the SZA on the SDR trends, we grouped the anomaly data in bin of $1^\circ$ SZA, and calculated the overall trend for each bin before noon and after noon (Figure\ \ref{fig:szatrends}).
+In order to estimate the effect of the SZA on the SDR trends, we grouped the anomaly data in bins of $1^\circ$ SZA, and calculated the overall trend for each bin before noon and after noon (Figure\ \ref{fig:szatrends}).
 Although there is a seasonal variation of the minimum SZA, its possible effect has not been examined here.
-For all-sky and clear-sky conditions the brightening effect of SDR (positive trend) for large SZAs (Figures\ \ref{fig:szatrends-1} and \ref{fig:szatrends-2}).
+For all-sky and clear-sky conditions the brightening effect of SDR (positive trend) is stronger for large SZAs (Figures\ \ref{fig:szatrends-1} and \ref{fig:szatrends-2}).
 The trends in the morning and afternoon hours are more or less consistent with small differences which can be attributed to systematic diurnal variations of aerosols, particularly during the warm period of the year [@Wang2021].
 For cloud-sky conditions (Figure\ \ref{fig:szatrends-3}), we can not discern any significant dependence of the SDR trend with SZA.
-For SZAs $16^\circ$ - $50^\circ$, the trends are ranging within about $\pm 0.2\,\%/y$, with a weak statistical significance.
-Between $55^\circ$ and $75^\circ$ SZA the trends are mostly negative and are possibly associated with stronger attenuation by clouds.
+For SZAs $16^\circ$ - $50^\circ$, the trends range within about $\pm 0.2\,\%/y$, with a weak statistical significance.
+Between $55^\circ$ and $75^\circ$ SZA the trends for the period before noon are larger and negative, possibly associated with stronger attenuation by clouds under oblique incidence angles.
 
@@
 (Figure\ \ref{fig:seasonalALL} and Table\ \ref{tab:trendseasontable}).
-For all-sky condition the trend in SDR in winter is the largest
+For all-sky conditions the trend in SDR in winter is the largest
 ($`r signif(spp$"Trend [%/year]"[spp$Season == "Winter" & spp$"Sky condition" == "All sky"], 2)`\,\%/y$),
@@
 close to the long term trend) both statistically significant above the $99\,\%$ confidence level. 
-In sprint and summer, the trends are much smaller and less significant.
-These seasonal differences indicate a possible relation of the trends in clouds during winter and autumn.
-For clear-skies, the trend in winter is even larger than for all-skies ($`r signif(spp$"Trend [%/year]"[spp$Season == "Winter" & spp$"Sky condition" == "Clear sky"], 2)`\,\%/y$) which is another indication of the effects
-Moreover, when examining the clear- and cloud-sky conditions the trends between these conditions are almost complementary to each other for Winter and Autumn respectively.
-On the latter cases, during Spring and Summer the statistical significance along with the actual trend are very low for a meaningful comparison.
+In spring and summer, the trends are much smaller and less statistical significant.
+These seasonal differences indicate a possible relation of the trends to clouds during winter and autumn.
+For clear-skies, the trend in winter is $`r signif(spp$"Trend [%/year]"[spp$Season == "Winter" & spp$"Sky condition" == "Clear sky"], 2)`\,\%/y$, larger than for all-skies ($`r signif(spp$"Trend [%/year]"[spp$Season == "Winter" & spp$"Sky condition" == "All sky"], 2)`\,\%/y$), which is another indication of increasing cloud thickness.
+Moreover, the trends under clear- and cloud-sky conditions are almost complementary to each other, particularly for winter and autumn.
+During spring and summer the statistical significance is very low and the actual trend too small for a meaningful comparison.
 
@@
 We have calculated the effect of TSI variability to be a negligible factor in SDR variability.
-Unfortunately, there is no available data for this period for the Aerosols, in order to estimate their effect on SDR, which is the most probable main cause.
-The attenuation of SDR by Aerosols over Europe have been proposed as major factor by @Wild2021.
+Unfortunately, there is no available data for this period for the aerosols, in order to estimate their effect on SDR, which is the most probable main cause.
+The attenuation of SDR by aerosols over Europe have been proposed as major factor by @Wild2021.
 The dimming effect of SDR under cloud-sky condition ($`r cloudtrend`\,\%/y$), shows that clouds can also attenuate the amount of SDR reaching the ground, and changes on their occurrences can effect the net energy balance.

### 2023-10-13 Article.Rmd
Commit 2023-10-13 13:28

@@
 output: 
+  rticles::elsevier_article:
+    keep_tex:         no
+    citation_package: natbib
+  bookdown::word_document2: default
   bookdown::pdf_document2:
@@
     biblio-style:     apalike
-  bookdown::word_document2: default
-  rticles::elsevier_article:
-    keep_tex:         no
-    citation_package: natbib
 # date:           "`r format(Sys.time(), '%F')`"
@@
 
-```{r cusum, echo = FALSE, fig.cap='Running cumulative sum plots of the daily SDR.', fig.subcap=c('All-sky conditions.', 'Clear-sky conditions.', 'Cloud-sky conditions.'), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3}
-knitr::include_graphics(c(
-    "./images/CumulativeDailyCuSum-1.png",    # ALL sky
-    "./images/CumulativeDailyCuSum-5.png",    # CLEAR sky
-    "./images/CumulativeDailyCuSum-9.png"))   # Cloud sky
-```
+<!-- ```{r cusum, echo = FALSE, fig.cap='Running cumulative sum plots of the daily SDR.', fig.subcap=c('All-sky conditions.', 'Clear-sky conditions.', 'Cloud-sky conditions.'), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3} -->
+<!-- knitr::include_graphics(c( -->
+<!--     "./images/CumulativeDailyCuSum-1.png",    # ALL sky -->
+<!--     "./images/CumulativeDailyCuSum-5.png",    # CLEAR sky -->
+<!--     "./images/CumulativeDailyCuSum-9.png"))   # Cloud sky -->
+<!-- ``` -->
 
@@
 
-```{r cusumnotrend, echo = FALSE, fig.cap='Running cumulative sum plots of the daily SDR after removing the long term trend from the daily values.', fig.subcap=c('All-skies.', 'Clear-skies.', 'Cloud-skies.'), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3}
-knitr::include_graphics(c(
-    "./images/CumulativeDailyCuSumNOtrend-1.png",    # ALL sky
-    "./images/CumulativeDailyCuSumNOtrend-5.png",    # CLEAR sky
-    "./images/CumulativeDailyCuSumNOtrend-9.png"))   # Cloud sky
-```
+<!-- ```{r cusumnotrend, echo = FALSE, fig.cap='Running cumulative sum plots of the daily SDR after removing the long term trend from the daily values.', fig.subcap=c('All-skies.', 'Clear-skies.', 'Cloud-skies.'), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3} -->
+<!-- knitr::include_graphics(c( -->
+<!--     "./images/CumulativeDailyCuSumNOtrend-1.png",    # ALL sky -->
+<!--     "./images/CumulativeDailyCuSumNOtrend-5.png",    # CLEAR sky -->
+<!--     "./images/CumulativeDailyCuSumNOtrend-9.png"))   # Cloud sky -->
+<!-- ``` -->
 

### 2023-10-19 Article.Rmd
Commit 2023-10-19 11:15

@@
 The solid data points represent values with an acceptable statistical significance ($p<0.005$).', fig.subcap=c('All-sky conditions.', 'Clear-sky conditions.', 'Cloud-sky conditions.'), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3}
-knitr::include_graphics("./images/SzaTrends-13.png")
-knitr::include_graphics("./images/SzaTrends-17.png" )
-knitr::include_graphics("./images/SzaTrends-21.png")
+knitr::include_graphics("./images/SzaTrends-1.png")
+knitr::include_graphics("./images/SzaTrends-4.png" )
+knitr::include_graphics("./images/SzaTrends-7.png")
 ```

### 2023-10-19 Article.Rmd
Commit 2023-10-19 18:14

@@
 
-```{r szatrends, echo = FALSE, fig.cap='Distribution of the SDR\'s long term trends by SZA before and after noon.
+```{r szatrends, echo = FALSE, fig.cap='*** Daily *** Distribution of the SDR\'s long term trends by SZA before and after noon.
 The solid data points represent values with an acceptable statistical significance ($p<0.005$).', fig.subcap=c('All-sky conditions.', 'Clear-sky conditions.', 'Cloud-sky conditions.'), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3}
@@
 knitr::include_graphics("./images/SzaTrends-7.png")
+```
+```{r szatrendsmonthly, echo = FALSE, fig.cap='*** Montlhly *** Distribution of the SDR\'s long term trends by SZA before and after noon.
+The solid data points represent values with an acceptable statistical significance ($p<0.005$).', fig.subcap=c('All-sky conditions.', 'Clear-sky conditions.', 'Cloud-sky conditions.'), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3}
+knitr::include_graphics("./images/SzaTrendsMonthly-1.png")
+knitr::include_graphics("./images/SzaTrendsMonthly-4.png" )
+knitr::include_graphics("./images/SzaTrendsMonthly-7.png")
 ```
@@
 
-```{r echo = FALSE, fig.cap='SDR trend ~ SZA', out.width='100%'}
+```{r echo = FALSE, fig.cap='*** Daily *** SDR trend ~ SZA', out.width='100%'}
 knitr::include_graphics("./images/SzaTrendsSeasTogether-2.png")
 ```
+
+```{r echo = FALSE, fig.cap='*** Montly *** SDR trend ~ SZA', out.width='100%'}
+knitr::include_graphics("./images/SzaTrendsSeasTogetherMonthly-2.png")
+```

### 2023-10-19 DHI_GHI_2_sza_trends.R

@@ -191,7 +191,7 @@ ccex_sbs <- 1.3
 #+ echo=F, include=F
 
 
-## __ Calculate SZA ~ Date -----------------------------------------------------
+## __ Calculate trend SZA ~ Day ------------------------------------------------
 
 vars <- c("GLB_att_des")
 
@@ -222,9 +222,6 @@ for (DBn in dbs) {
                                     var     = avar,
                                     N       = sum(!is.na(dataset[[avar]]))
                                 ))
-                ## Climatology plot
-                # plot(dataset$doy, dataset[[sub("_des", "_seas", avar)]], cex = 1, pch = 1 )
-                # title(main = paste("Seasonal values", DBn, asza, sub("_des", "_seas", avar)))
             }
         }
     }
@@ -340,7 +337,7 @@ setorder(szatrends, SZA)
 
 
 
-## covert to trend per year
+##__ Covert to trend per year --------------------------------------------------
 szatrends[, slope    := slope    * Days_of_year ]
 szatrends[, slope.sd := slope.sd * Days_of_year ]
 
@@ -382,10 +379,9 @@ szatrends[preNoon == F, pch := pch_pm ]
 wecare <- grep("^slope|^N",  names(szatrends), ignore.case = T, value = T)
 wecare <- grep("^slope\\.t", wecare, ignore.case = T, value = T, invert = T)
 wecare <- grep("slope\\.sd", wecare, ignore.case = T, value = T, invert = T)
+wecare <- grep("slope.Conf", wecare, ignore.case = T, value = T, invert = T)
 
 vars <- c("GLB_att_des")
-
-## TODO separate plots by direct global
 
 #+ SzaTrends, echo=F, include=T, results = "asis"
 for (avar in vars) {
@@ -521,7 +517,7 @@ for (avar in vars) {
 
 
 # __ Plot daily climatology --------------------------------------------------
-#+ SzaClimaMonthly, echo=F, include=T, results = "asis"
+#+ SzaClimaDaily, echo=F, include=T, results = "asis", out.heigth="30%"
 vars <- c("GLB_att_seas")
 
 dbs  <- c(  "ALL_2_daily_DESEAS",
@@ -562,12 +558,12 @@ for (DBn in dbs) {
             points(pam$doy, pam[[avar]],
                    col  = 2,
                    pch  = pam$pch,
-                   cex  = 3)
+                   cex  = 1)
 
[…]

### 2023-10-20 Article.Rmd
Commit 2023-10-20 11:11

@@
 load("./data/common_data_14_2.Rda")
+load("./data/Input_1_longterm_trends.Rda")
 ## number of measurements for each dataset

### 2023-10-20 DHI_GHI_01_Input_longterm.R

@@ -20,6 +20,7 @@ if (
     file.mtime(raw_input_data) < file.mtime("./DHI_GHI_00_raw_data.R")
 ) {
     source("./DHI_GHI_00_raw_data.R")
+    gc()
 }
 
 ## check current steps
@@ -803,5 +804,5 @@ gc()
 
 # ......................................................................... ----
 ##  Save data ------------------------------------------------------------------
-save(file = I1_longterm, list = ls(pattern = "^ALL_|^CLEAR_|^CLOUD_"))
+save(file = I1_longterm, list = ls(pattern = "^ALL_1_|^CLEAR_1_|^CLOUD_1_"))
 cat(paste("\n Long term proccessed data saved", I1_longterm, "\n\n"))

### 2023-10-20 DHI_GHI_1_longterm_trends_b.R

@@ -98,6 +98,7 @@ if (! file.exists(I1_longterm) |
     file.mtime(I1_longterm) < file.mtime("./DHI_GHI_01_Input_longterm.R") )
 {
     source("./DHI_GHI_01_Input_longterm.R")
+    gc()
 }
 
 ## check current steps

### 2023-10-21 Article.Rmd
Commit 2023-10-21 11:12

--- 2023-10-19 Article.Rmd
@@
 
+<!-- ```{r szatrendsmonthly, echo = FALSE, fig.cap='*** Montlhly *** Distribution of the SDR\'s long term trends by SZA before and after noon.
+The solid data points represent values with an acceptable statistical significance ($p<0.005$).', fig.subcap=c('All-sky conditions.', 'Clear-sky conditions.', 'Cloud-sky conditions.'), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3} -->
+<!-- knitr::include_graphics("./images/SzaTrendsMonthly-1.png") -->
+<!-- knitr::include_graphics("./images/SzaTrendsMonthly-4.png" ) -->
+<!-- knitr::include_graphics("./images/SzaTrendsMonthly-7.png") -->
+<!-- ``` -->
+
 ## Long term trends by season of year
@@
 ```
+
+<!-- ```{r echo = FALSE, fig.cap='*** Montly *** SDR trend ~ SZA', out.width='100%'} -->
+<!-- knitr::include_graphics("./images/SzaTrendsSeasTogetherMonthly-2.png") -->
+<!-- ``` -->

### 2023-10-21 DHI_GHI_2_sza_trends.R

@@ -87,7 +87,7 @@ source("~/CODE/FUNCTIONS/R/data.R")
 
 
 ## __ Source initial scripts ---------------------------------------------------
-source("./DHI_GHI_0_data_input.R")
+# source("./DHI_GHI_0_data_input.R")
 source("./DHI_GHI_0_variables.R")
 
 ## check previous steps
@@ -97,7 +97,7 @@ if (! file.exists(I2_szatrend) |
     file.mtime(I2_szatrend) < file.mtime("./DHI_GHI_02_Input_szatrends.R") )
 {
     # source("./DHI_GHI_02_Input_szatrends.R")
-    tryCatch(source("./DHI_GHI_02_Input_szatrends.R"), exit=function(cond) {
+    tryCatch(source("./DHI_GHI_02_Input_szatrends.R"), exit = function(cond) {
         message( conditionMessage(cond) )
     })
     dummy <- gc()
@@ -931,6 +931,10 @@ setorder(szatrends_seas, SZA)
 hist(gather_seas$N[gather_seas$N > 50], breaks = 100)
 
 
+
+
+
+
 ##__ Covert to trend per year --------------------------------------------------
 szatrends_seas[, slope    := slope    * Days_of_year]
 szatrends_seas[, slope.sd := slope.sd * Days_of_year]
@@ -1115,7 +1119,14 @@ for (ase in seasons) {
 
 
 
-## __ by season in a tight grid ----------------------------------------------
+## __ by season daily in tight grid --------------------------------------------
+
+
+test <- szatrends_seas[DATA == "ALL_2_bySeason_daily_DESEAS" & Season == "Winter"]
+
+setorder(test, N)
+
+
 
 #'
 #' ### Grid of SZA trends for each season of year from daily
@@ -1231,7 +1242,9 @@ for (ase in seasons) {
             expanded <- expanded[-1, ]
 
             ## limit plot p-values
-            p_lim     <- 0.05
+            p_lim    <- 0.05
+            N_lim    <- 85
+
 
             ## select All/CS  DIR/GLB/trans winter/summer
             subdata <- szatrends_seas[ DATA   == kk$Dataset &
@@ -1241,8 +1254,11 @@ for (ase in seasons) {
             ## set symbols for plotting
             subdata[ slope.p  < p_lim & preNoon == TRUE,  pch := 16 ]
[…]

### 2023-10-24 Article.Rmd
Commit 2023-10-24 10:14

@@
 For all-sky and clear-sky conditions the brightening effect of SDR (positive trend) is stronger for large SZAs (Figures\ \ref{fig:szatrends-1} and \ref{fig:szatrends-2}).
-The trends in the morning and afternoon hours are more or less consistent with small differences which can be attributed to systematic diurnal variations of aerosols, particularly during the warm period of the year [@Wang2021].
+The trends in the morning and afternoon hours are more or less consistent with small differences, which can be attributed to systematic diurnal variations of aerosols, particularly during the warm period of the year [@Wang2021].
 For cloud-sky conditions (Figure\ \ref{fig:szatrends-3}), we can not discern any significant dependence of the SDR trend with SZA.
 For SZAs $16^\circ$ - $50^\circ$, the trends range within about $\pm 0.2\,\%/y$, with a weak statistical significance.
-Between $55^\circ$ and $75^\circ$ SZA the trends for the period before noon are larger and negative, possibly associated with stronger attenuation by clouds under oblique incidence angles.
+Between $50^\circ$ and $75^\circ$ SZA the trends for the period before noon are stronger and negative, possibly associated with stronger attenuation by clouds under oblique incidence angles.
 
 ```{r szatrends, echo = FALSE, fig.cap='Distribution of the SDR\'s long term trends by SZA before and after noon.
-The solid data points represent values with an acceptable statistical significance ($p<0.005$).', fig.subcap=c('All-sky conditions.', 'Clear-sky conditions.', 'Cloud-sky conditions.'), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3}
+The solid data points represent values with an acceptable statistical significance ($p<0.005$), the trends were computed with SZA bins of $1^\\circ$ for each day, before and after local noon.', fig.subcap=c('All-sky conditions.', 'Clear-sky conditions.', 'Cloud-sky conditions.'), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3}
 knitr::include_graphics("./images/SzaTrends-1.png")
@@
 
-Similar to the long term trends discussed above, we have calculated the trend of the anomalies for the three different sky conditions and for each season of the year, using the corresponding mean monthly values
+Similar to the long term trends discussed above, we have calculated the trend of the anomalies for the three different sky conditions, and for each season of the year, using the corresponding mean monthly values
 (Figure\ \ref{fig:seasonalALL} and Table\ \ref{tab:trendseasontable}).
@@
 ($`r signif(spp$"Trend [%/year]"[spp$Season == "Autumn" & spp$"Sky condition" == "All sky"], 2)`\,\%/y$,
-close to the long term trend) both statistically significant above the $99\,\%$ confidence level. 
+value similar to the long term trend) both statistically significant above the $99\,\%$ confidence level. 
 In spring and summer, the trends are much smaller and less statistical significant.

### 2023-10-24 Article.Rmd
Commit 2023-10-24 11:10

@@
 
-Although the year-to-year variability of the anomalies in Figure \ref{fig:trendALL} shows a rather homogeneous behavior, plots of the cumulative sums (CUSUM) of the anomalies can reveal structures which are different for the three sky conditions [@Regier2019].
-On all three plots on Figure\ \ref{fig:cusum}, we observe a downward part at the start of the period until about 2000, an almost steady part until about 2016 for all-sky and clear-sky conditions (Figures\ \ref{fig:cusum-1}, \ref{fig:cusum-2}) and until 2009 for cloud-sky conditions (Figure\ \ref{fig:cusum-3}), and finally, a steep upward part until the present.
-For a uniform trend, we would expect the cumulative sums of the anomalies to have a symmetric 'V' shape.
-This would indicate that the anomalies are evenly distributed around the climatological mean.
+Although the year-to-year variability of the anomalies (Figure \ref{fig:trendALL} and Figures \ref{fig:trendCLEAR}, \ref{fig:trendCLOUD} on Appendix), shows a rather homogeneous behavior, plots of the cumulative sums (CUSUM) of the anomalies can reveal different structures within all the three sky conditions [@Regier2019].
+In the cases of all-sky and clear-sky (Figures \ref{fig:cusummonth-1} and \ref{fig:cusummonth-2}), we observe three macroscopic periods.
+A downward part from the start until about 2005, an relative steady part until about 2016 and finally, a steep upward part until the present.
+For cloud-sky (Figure\ \ref{fig:cusummonth-3}), we have a different sequence, it begins with a relative steady part until 1997, followed by a upward part until 2005, a long decline until 2020, with a small positive slope until the present.
+For a uniform trend, we would expect the CUSUMs of the anomalies to have a symmetric 'V' shape.
+This would indicate that the anomalies are evenly distributed around the climatological mean, and for a positive uniform trend, the first half to be below and the other half above the climatological mean.
+
 In our case, for all sky conditions, we observe periods where the CUSUMs remain relative steady, roughly around the middle of the study dataset.
@@
 
-```{r cusummonth, echo = FALSE, fig.cap='Running cumulative sum plots of the monthly SDR.', fig.subcap=c('All-sky conditions.', 'Clear-sky conditions.', 'Cloud-sky conditions.'), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3}
+```{r cusummonth, echo = FALSE, fig.cap='Running cumulative sum plots of the monthly SDR.', fig.subcap=c('All-skies.', 'Clear-skies.', 'Cloud-skies.'), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3}
 knitr::include_graphics(c(
@@
 In order to estimate the effect of the SZA on the SDR trends, we grouped the anomaly data in bins of $1^\circ$ SZA, and calculated the overall trend for each bin before noon and after noon (Figure\ \ref{fig:szatrends}).
-Although there is a seasonal variation of the minimum SZA, its possible effect has not been examined here.
+Although there are seasonal variation of the minimum SZA (see Appendix, Figure\ \ref{fig:SZAtrendSeason}), these dependencies would not be further examined here.
 For all-sky and clear-sky conditions the brightening effect of SDR (positive trend) is stronger for large SZAs (Figures\ \ref{fig:szatrends-1} and \ref{fig:szatrends-2}).
@@
 For clear-skies, the trend in winter is $`r signif(spp$"Trend [%/year]"[spp$Season == "Winter" & spp$"Sky condition" == "Clear sky"], 2)`\,\%/y$, larger than for all-skies ($`r signif(spp$"Trend [%/year]"[spp$Season == "Winter" & spp$"Sky condition" == "All sky"], 2)`\,\%/y$), which is another indication of increasing cloud thickness.
-Moreover, the trends under clear- and cloud-sky conditions are almost complementary to each other, particularly for winter and autumn.
+Moreover, the trends under clear- and cloud-sky conditions are almost complementary to each other, particularly for winter and autumn, where the signal is stronger.
 During spring and summer the statistical significance is very low and the actual trend too small for a meaningful comparison.
@@
 
-```{r echo = FALSE, fig.cap='SDR trend ~ SZA', out.width='100%'}
+```{r SZAtrendSeason, echo = FALSE, fig.cap='SDR trend ~ SZA', out.width='100%'}
 knitr::include_graphics("./images/SzaTrendsSeasTogether-2.png")

### 2023-10-24 Article.Rmd
Commit 2023-10-24 12:14

@@
 This would indicate that the anomalies are evenly distributed around the climatological mean, and for a positive uniform trend, the first half to be below and the other half above the climatological mean.
+In our case, for all-skies and clear-skies, we observe periods where the CUSUMs remain relative steady, roughly around the middle of the study dataset.
+This is an indication that the whole brightening phenomenon has not a smooth development over time.
 
-In our case, for all sky conditions, we observe periods where the CUSUMs remain relative steady, roughly around the middle of the study dataset.
 Another distinct feature of the CUSUMs, is the different pattern of the cloudy-sky dataset which peaks around the middle of the period.
@@
 The increase of this trend, indicates that the phenomenon and probably the causes, are still occurring.
-Accordingly, we found a similar trend for clear-sky conditions ($`r cleartrend`\,\%/y$) that further supports that the brightening can be independent of the cloud occurrences. 
+Accordingly, we found a similar trend for clear-sky conditions ($`r cleartrend`\,\%/y$) that further supports that the brightening can be independent of the cloud occurrences, although cloud formation can also attenuate SDR.
 We have calculated the effect of TSI variability to be a negligible factor in SDR variability.

### 2023-10-24 DHI_GHI_01_Input_longterm.R

@@ -124,6 +124,9 @@ suppressWarnings({
 ALL_1_daily_mean[   DIR_att_N <= Daily_aggregation_N_lim, DIR_att       := NA]
 ALL_1_daily_mean[   GLB_att_N <= Daily_aggregation_N_lim, GLB_att       := NA]
 ALL_1_daily_mean[   HOR_att_N <= Daily_aggregation_N_lim, HOR_att       := NA]
+ALL_1_daily_mean[   DIR_att_N <= Daily_aggregation_N_lim, DIR_att_N     := NA]
+ALL_1_daily_mean[   GLB_att_N <= Daily_aggregation_N_lim, GLB_att_N     := NA]
+ALL_1_daily_mean[   HOR_att_N <= Daily_aggregation_N_lim, HOR_att_N     := NA]
 ALL_1_daily_mean[   DIR_att_N <= Daily_aggregation_N_lim, DIR_transp    := NA]
 ALL_1_daily_mean[   DIR_att_N <= Daily_aggregation_N_lim, DIR_att_sd    := NA]
 ALL_1_daily_mean[   GLB_att_N <= Daily_aggregation_N_lim, GLB_att_sd    := NA]
@@ -137,6 +140,9 @@ ALL_1_daily_mean[   DIR_att_N <= Daily_aggregation_N_lim, DIR_transp_EM := NA]
 CLEAR_1_daily_mean[ DIR_att_N <= Daily_aggregation_N_lim, DIR_att       := NA]
 CLEAR_1_daily_mean[ GLB_att_N <= Daily_aggregation_N_lim, GLB_att       := NA]
 CLEAR_1_daily_mean[ HOR_att_N <= Daily_aggregation_N_lim, HOR_att       := NA]
+CLEAR_1_daily_mean[ DIR_att_N <= Daily_aggregation_N_lim, DIR_att_N     := NA]
+CLEAR_1_daily_mean[ GLB_att_N <= Daily_aggregation_N_lim, GLB_att_N     := NA]
+CLEAR_1_daily_mean[ HOR_att_N <= Daily_aggregation_N_lim, HOR_att_N     := NA]
 CLEAR_1_daily_mean[ DIR_att_N <= Daily_aggregation_N_lim, DIR_transp    := NA]
 CLEAR_1_daily_mean[ DIR_att_N <= Daily_aggregation_N_lim, DIR_att_sd    := NA]
[…]

### 2023-10-24 Article.Rmd
Commit 2023-10-24 13:09

@@
 TSI$var                <- NULL
+TSI$N                  <- NULL
+TSI$Obs                <- NULL
 setorder(TSI, -slope.stat_sig)
@@
 LT$var                <- NULL
+LT$N                  <- NULL
 setorder(LT, -slope.stat_sig)
 rownames(LT) <- NULL
-names(LT)    <- c("Trend [%/year]", "Trend p-value", "Sky conditions", "Statistical signif. [%]")
+names(LT)    <- c("Trend [%/year]", 
+                  "Trend p-value", 
+                  "Sky conditions",
+                  "Observations",
+                  "Statistical signif. [%]")
 ## no need all are 100%
@@
       "Trend [%/year]",
-      "Trend p-value"
+      "Trend p-value",
+      "Observations"
       )]
@@
       "Trend [%/year]",
-      "Statistical signif. [%]"
+      "Statistical signif. [%]",
+      "Observations"
       )]
@@
 This would indicate that the anomalies are evenly distributed around the climatological mean, and for a positive uniform trend, the first half to be below and the other half above the climatological mean.
-In our case, for all-skies and clear-skies, we observe periods where the CUSUMs remain relative steady, roughly around the middle of the study dataset.
-This is an indication that the whole brightening phenomenon has not a smooth development over time.
+In our case, there is more complex evolution of the phenomenon.
+Another distinct feature of the CUSUMs, is the different pattern of the cloudy-sky dataset which peaks around the middle of the period.
+Although, it seems to exist a complementary relation to the CUSUMs of the clear- and all-sky cases, we can not assert that clouds are the main drive for this relation, due to the great difference in the number of observational data between the datasets (Table\ \ref{tab:trendtable}).
 
-Another distinct feature of the CUSUMs, is the different pattern of the cloudy-sky dataset which peaks around the middle of the period.
-The CUSUMs of the clear and the all-sky conditions exhibit similar patterns.
-The causes for these discrepancies haven't been investigated yet.
-
-In order to compare any features within the trends, we created another set of CUSUMs plots, by subtracting the corresponding long term trend from the SDR anomaly data, prior to the CUSUM calculation (Figure\ \ref{fig:cusumnotrend}).
+In order to investigate more any features within the trends, we created another set of CUSUMs plots, by subtracting the corresponding long term trend from the SDR anomaly data, prior to the CUSUM calculation (Figure\ \ref{fig:cusumnotrendmonthly}).
 With this approach periods when the CUSUMs diverge from zero can be interpreted as systematic variation of SDR from the climatological mean.
 When the CUSUM is increasing, the added values are above the climatological values of SDR trend and vice versa.
-Overall, for all- and clear-sky conditions (Figure\ \ref{fig:cusumnotrend-1} and\ \ref{fig:cusumnotrend-2}) we observe periods when the anomalies diverge from the climatological value, each lasting for several years.
+Overall, for all- and clear-sky conditions (Figure\ \ref{fig:cusumnotrendmonthly-1} and\ \ref{fig:cusumnotrendmonthly-2}) we observe periods when the anomalies diverge from the climatological value, each lasting for several years.
 The pattern for the all sky and clear sky conditions is very similar, suggesting prevalence in clear skies over Thessaloniki.
@@
 The dimming effect of SDR under cloud-sky condition ($`r cloudtrend`\,\%/y$), shows that clouds can also attenuate the amount of SDR reaching the ground, and changes on their occurrences can effect the net energy balance.
-Because we have no adequate data to investigate long term changes of cloud formation in the region, we can not infer if the SDR trend we observe under cloud-sky is biased by changes in clouds and local climate.
+Because we have no adequate data to investigate the long term changes of cloud formation in the region, we can not infer if the SDR trend we observe under cloud-sky is biased by changes in clouds and local climate.
 The observed brightening of SDR over Thessaloniki; is dependent on SZA (higher SZA, have stronger brightening trends).
 And also, with the seasons of the year, where Winter and Autumn show a significant strong trend, in contrast to Spring and Summer.
+
+<!-- to conclutions? -->
+In the CUSUMs for all-skies and clear-skies, we observed periods where the CUSUMs remain relative steady, 
+<!-- roughly around the middle of the study dataset,--> with a steep decline previous to that, and a steep incline after.
+This is an indication that the whole brightening phenomenon does not have a smooth development over time.
 

### 2023-10-24 Article.Rmd
Commit 2023-10-24 14:10

@@
 
-In order to investigate more any features within the trends, we created another set of CUSUMs plots, by subtracting the corresponding long term trend from the SDR anomaly data, prior to the CUSUM calculation (Figure\ \ref{fig:cusumnotrendmonthly}).
+```{r cusummonth, echo = FALSE, fig.cap='Running cumulative sum plots of the monthly SDR.', fig.subcap=c('All-skies.', 'Clear-skies.', 'Cloud-skies.'), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3}
+knitr::include_graphics(c(
+    "./images/CumulativeMonthlyCuSum-1.png",    # ALL sky
+    "./images/CumulativeMonthlyCuSum-5.png",    # CLEAR sky
+    "./images/CumulativeMonthlyCuSum-9.png"))   # Cloud sky
+```
+
+In order to investigate more the features within the trends, we created another set of CUSUMs plots, by subtracting the corresponding long term trend from the SDR anomaly data, prior to the CUSUM calculation (Figure\ \ref{fig:cusumnotrendmonthly}).
 With this approach periods when the CUSUMs diverge from zero can be interpreted as systematic variation of SDR from the climatological mean.
@@
 
+```{r cusumnotrendmonthly, echo = FALSE, fig.cap='Running cumulative sum plots of the monthly SDR after removing the long term trend from the monthly values.', fig.subcap=c('All-skies.', 'Clear-skies.', 'Cloud-skies.'), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3}
+knitr::include_graphics(c(
+    "./images/CumulativeMonthlyCuSumNOtrend-1.png",    # ALL sky
+    "./images/CumulativeMonthlyCuSumNOtrend-5.png",    # CLEAR sky
+    "./images/CumulativeMonthlyCuSumNOtrend-9.png"))   # Cloud sky
+```
+
 <!-- ```{r cusum, echo = FALSE, fig.cap='Running cumulative sum plots of the daily SDR.', fig.subcap=c('All-sky conditions.', 'Clear-sky conditions.', 'Cloud-sky conditions.'), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3} -->
@@
 
-```{r cusummonth, echo = FALSE, fig.cap='Running cumulative sum plots of the monthly SDR.', fig.subcap=c('All-skies.', 'Clear-skies.', 'Cloud-skies.'), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3}
-knitr::include_graphics(c(
-    "./images/CumulativeMonthlyCuSum-1.png",    # ALL sky
-    "./images/CumulativeMonthlyCuSum-5.png",    # CLEAR sky
-    "./images/CumulativeMonthlyCuSum-9.png"))   # Cloud sky
-```
-
 <!-- ```{r cusumnotrend, echo = FALSE, fig.cap='Running cumulative sum plots of the daily SDR after removing the long term trend from the daily values.', fig.subcap=c('All-skies.', 'Clear-skies.', 'Cloud-skies.'), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3} -->
@@
 <!-- ``` -->
-
-```{r cusumnotrendmonthly, echo = FALSE, fig.cap='Running cumulative sum plots of the monthly SDR after removing the long term trend from the monthly values.', fig.subcap=c('All-skies.', 'Clear-skies.', 'Cloud-skies.'), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3}
-knitr::include_graphics(c(
-    "./images/CumulativeMonthlyCuSumNOtrend-1.png",    # ALL sky
-    "./images/CumulativeMonthlyCuSumNOtrend-5.png",    # CLEAR sky
-    "./images/CumulativeMonthlyCuSumNOtrend-9.png"))   # Cloud sky
-```
 

### 2023-10-24 DHI_GHI_0_data_input.R

@@ -332,1216 +332,6 @@ if (havetorun) {
 
 
 
-##     # ..................................................................... ----
-##     ####  2. Long term by SZA  -------------------------------------------------
-## 
-## 
-##     ## SZA test ----------------------------------------------------------------
-##     # test_all <- DATA_all[, .(GLB_att       = mean(GLB_att,    na.rm = T),
-##     #                          GLB_att_sd    = sd(  GLB_att,    na.rm = T),
-##     #                          GLB_att_N     = sum(!is.na(GLB_att))),
-##     #                      by = .(SZA     = (SZA - SZA_BIN / 2 ) %/% SZA_BIN,
-##     #                             Date    = as.Date(Date),
-##     #                             preNoon = preNoon ) ]
-##     #
-##     # test_all_year <- DATA_all[, .(GLB_att       = mean(GLB_att,    na.rm = T),
-##     #                               GLB_att_sd    = sd(  GLB_att,    na.rm = T),
-##     #                               GLB_att_N     = sum(!is.na(GLB_att))),
-##     #                           by = .(SZA     = (SZA - SZA_BIN / 2 ) %/% SZA_BIN,
-##     #                                  Year    = year(Date),
-##     #                                  preNoon = preNoon ) ]
-##     #
-##     # test_all_monthly <- DATA_all[, .(GLB_att       = mean(GLB_att,    na.rm = T),
-##     #                                  GLB_att_sd    = sd(  GLB_att,    na.rm = T),
-##     #                                  GLB_att_N     = sum(!is.na(GLB_att))),
-##     #                              by = .(SZA     = (SZA - SZA_BIN / 2 ) %/% SZA_BIN,
-##     #                                     Year    = year(Date),
-##     #                                     Month   = month(Date),
-##     #                                     preNoon = preNoon ) ]
-##     # test_all_monthly[, Date := as.Date(paste(Year, Month, 1), "%Y %m %d") ]
-##     #
-##     # gather1 <- data.table()
-##     # for (asza in unique(test_all_year$SZA)) {
-##     #     for (pday in unique(test_all_year$preNoon)) {
-##     #         subdata <- test_all_year[SZA == asza & preNoon == pday]
[…]

### 2023-10-24 DHI_GHI_2_sza_trends_delete.R

@@ -1,1932 +1,0 @@
-# /* #!/usr/bin/env Rscript */
-# /* Copyright (C) 2022 Athanasios Natsis <natsisphysicist@gmail.com> */
-#' ---
-#' title:         "Trends of SDR in Thessaloniki "
-#' author:
-#'   - Natsis Athanasios^[Laboratory of Atmospheric Physics, AUTH, natsisphysicist@gmail.com]
-#'   - Alkiviadis Bais^[Laboratory of Atmospheric Physics, AUTH]
-#' abstract:
-#'   "Study of GHI and DNI radiation for 'clear sky' and all sly conditions."
-#'
-#' documentclass:  article
-#' classoption:    a4paper,oneside
-#' fontsize:       10pt
-#' geometry:       "left=0.5in,right=0.5in,top=0.5in,bottom=0.5in"
-#' link-citations: yes
-#' colorlinks:     yes
-#'
-#' header-includes:
-#' - \usepackage{caption}
-#' - \usepackage{placeins}
-#' - \captionsetup{font=small}
-#'
-#' output:
-#'   bookdown::pdf_document2:
-#'     number_sections: no
-#'     fig_caption:     no
-#'     keep_tex:        yes
-#'     latex_engine:    xelatex
-#'     toc:             yes
-#'     toc_depth:       4
-#'     fig_width:       7
-#'     fig_height:      4.5
-#'   html_document:
-#'     toc:             true
-#'     keep_md:         yes
-#'     fig_width:       7
-#'     fig_height:      4.5
-#'
-#' date: "`r format(Sys.time(), '%F')`"
-#'
-#' ---
-
-#+ echo=F, include=T
-
-
-## __ Document options ---------------------------------------------------------
-
-#+ echo=F, include=F
-knitr::opts_chunk$set(comment    = ""       )
-knitr::opts_chunk$set(dev        = c("pdf", "png")    )
-# knitr::opts_chunk$set(dev        = "png"    )
-knitr::opts_chunk$set(out.width  = "100%"   )
-knitr::opts_chunk$set(fig.align  = "center" )
-knitr::opts_chunk$set(cache      =  FALSE   )  ## !! breaks calculations
-# knitr::opts_chunk$set(fig.pos    = '!h'    )
-warning("Don't use cache it breaks computations")
-
-#+ include=F, echo=F
-## __ Set environment ----------------------------------------------------------
-Sys.setenv(TZ = "UTC")
[…]

### 2023-10-24 Build_156/Article_B156.Rmd
Commit 2023-10-24 15:09

--- 2023-10-24 Article.Rmd
@@
 
-In order to investigate more the features within the trends, we created another set of CUSUMs plots, by subtracting the corresponding long term trend from the SDR anomaly data, prior to the CUSUM calculation (Figure\ \ref{fig:cusumnotrendmonthly}).
+In order to investigate more, the features within the trends, we created another set of CUSUMs plots by subtracting the corresponding long term trend from the SDR anomaly data, prior to the CUSUM calculation (Figure\ \ref{fig:cusumnotrendmonthly}).
 With this approach periods when the CUSUMs diverge from zero can be interpreted as systematic variation of SDR from the climatological mean.
@@
 Overall, for all- and clear-sky conditions (Figure\ \ref{fig:cusumnotrendmonthly-1} and\ \ref{fig:cusumnotrendmonthly-2}) we observe periods when the anomalies diverge from the climatological value, each lasting for several years.
-The pattern for the all sky and clear sky conditions is very similar, suggesting prevalence in clear skies over Thessaloniki.
-It is interesting that in the period 1993 - 2016 the anomalies of all-sky and clear- sky, have a high variability around zero.
+The pattern for the all-sky and clear-sky conditions is very similar, suggesting prevalence in clear skies over Thessaloniki.
+It is interesting that in the period 1993 - 2016 the anomalies of all-sky and clear-sky, have a high variability around zero.
 After 2016, the range of the variability is decreased at about one third of the prior period.
-For cloud-sky conditions (Figure\ \ref{fig:cusumnotrend-3}) the period 1997-2008 is dominated by positive CUSUMs suggesting a reduced effect of clouds on SDR.
+For cloud-sky conditions (Figure\ \ref{fig:cusumnotrendmonthly-3}) the period 1997 - 2008 is dominated by positive CUSUMs suggesting a reduced effect of clouds on SDR.
 From 1997 to mid-2000s CUSUMs are increasing, likely due to a continuous decrease in the thickness of clouds, followed by a period of rapid increase (within 3 years) in cloud optical thickness lasting up to 2008.
@@
 In order to estimate the effect of the SZA on the SDR trends, we grouped the anomaly data in bins of $1^\circ$ SZA, and calculated the overall trend for each bin before noon and after noon (Figure\ \ref{fig:szatrends}).
-Although there are seasonal variation of the minimum SZA (see Appendix, Figure\ \ref{fig:SZAtrendSeason}), these dependencies would not be further examined here.
+Although there are seasonal dependencies of the minimum SZA (see Appendix, Figure\ \ref{fig:SZAtrendSeason}), these dependencies would not be further examined here.
 For all-sky and clear-sky conditions the brightening effect of SDR (positive trend) is stronger for large SZAs (Figures\ \ref{fig:szatrends-1} and \ref{fig:szatrends-2}).
@@
 ## use kable
-
 x_latex <- knitr::kable(spp,
@@
                         caption  = "\\label{tab:trendseasontable}SDR linear trends of monthly anomalies for each season of the year.")
+
 x_latex %>%
@@
     )
+
 # position = "float_left"
@@
 ```
-
 <!-- \normalsize -->
@@
 The increase of this trend, indicates that the phenomenon and probably the causes, are still occurring.
-Accordingly, we found a similar trend for clear-sky conditions ($`r cleartrend`\,\%/y$) that further supports that the brightening can be independent of the cloud occurrences, although cloud formation can also attenuate SDR.
+Accordingly, we found a similar trend for clear-sky conditions ($`r cleartrend`\,\%/y$) that further supports that the brightening can be independent of the cloud occurrences, although, cloud formation can also attenuate SDR.
 We have calculated the effect of TSI variability to be a negligible factor in SDR variability.
-Unfortunately, there is no available data for this period for the aerosols, in order to estimate their effect on SDR, which is the most probable main cause.
+Unfortunately, there is no available data for this period for the aerosols, in order to directly estimate their effect on SDR, which is the most probable main cause.
 The attenuation of SDR by aerosols over Europe have been proposed as major factor by @Wild2021.
@@
 Because we have no adequate data to investigate the long term changes of cloud formation in the region, we can not infer if the SDR trend we observe under cloud-sky is biased by changes in clouds and local climate.
-The observed brightening of SDR over Thessaloniki; is dependent on SZA (higher SZA, have stronger brightening trends).
-And also, with the seasons of the year, where Winter and Autumn show a significant strong trend, in contrast to Spring and Summer.
-
+The observed brightening of SDR over Thessaloniki; is dependent on SZA (higher SZA, have stronger brightening trends), and also, with the seasons of the year, where Winter and Autumn show a significant strong trend, in contrast to Spring and Summer.
 <!-- to conclutions? -->
-In the CUSUMs for all-skies and clear-skies, we observed periods where the CUSUMs remain relative steady, 
+Using CUSUMs of the trends for all- and clear-skies, we observed periods where the CUSUMs remain relative steady, 
 <!-- roughly around the middle of the study dataset,--> with a steep decline previous to that, and a steep incline after.
-This is an indication that the whole brightening phenomenon does not have a smooth development over time.
+This is an indication that the whole brightening phenomenon does not follow a smooth development over time.
 
@@
 
-```{r SZAtrendSeason, echo = FALSE, fig.cap='SDR trend ~ SZA', out.width='100%'}
+```{r SZAtrendSeason, echo = FALSE, fig.cap='Seasonla distribution of the SDR\'s long term trends by SZA before and after noon.
+The solid data points represent values with an acceptable statistical significance ($p<0.005$), the trends were computed with SZA bins of $1^\\circ$ for each day, before and after local noon, all graphs have the same scale.', out.width='100%'}
 knitr::include_graphics("./images/SzaTrendsSeasTogether-2.png")

### 2023-10-24 Article.Rmd
Commit 2023-10-24 16:14

@@
 
-```{r SZAtrendSeason, echo = FALSE, fig.cap='SDR trend ~ SZA', out.width='100%'}
+```{r SZAtrendSeason, echo = FALSE, fig.cap='Seasonla distribution of the SDR\'s long term trends by SZA before and after noon.
+The solid data points represent values with an acceptable statistical significance ($p<0.005$), the trends were computed with SZA bins of $1^\\circ$ for each day, before and after local noon, all graphs have the same scale.
+Bins with $p<0.005$ or with less than 85 data points may be missing from the view.', out.width='100%'}
 knitr::include_graphics("./images/SzaTrendsSeasTogether-2.png")

### 2023-11-01 Article.Rmd
Commit 2023-11-01 12:14

@@
 
+stop("Observations to days of data!!")
+
 ```
@@
                         align    = "crr",
-                        caption  = paste0("\\label{tab:trendtable}Trends in SDR daily mean for different sky conditions for the period ", year(min(ALL_1_daily_DESEAS$Date)), " - ", year(max(ALL_1_daily_DESEAS$Date)), "."))
+                        caption  = paste0("\\label{tab:trendtable}Trends in SDR daily means for different sky conditions for the period ", year(min(ALL_1_daily_DESEAS$Date)), " - ", year(max(ALL_1_daily_DESEAS$Date)), "."))
 x_latex %>%

### 2023-11-01 Article.Rmd
Commit 2023-11-01 16:14

@@
 output: 
-  rticles::elsevier_article:
-    keep_tex:         no
-    citation_package: natbib
+  rticles::mdpi_article: default
   bookdown::word_document2: default
@@
 LT$var                <- NULL
-LT$N                  <- NULL
+# LT$N                  <- NULL
+LT$Obs                <- NULL
 setorder(LT, -slope.stat_sig)
@@
                   "Sky conditions",
-                  "Observations",
+                  "Days",
                   "Statistical signif. [%]")
@@
       "Trend p-value",
-      "Observations"
+      "Days"
       )]
@@
       "Statistical signif. [%]",
-      "Observations"
+      "Days"
       )]
-
-stop("Observations to days of data!!")
-
 ```

### 2023-11-01 Article.tex
Commit 2023-11-01 16:14

@@
-\documentclass[preprint, 3p,
-authoryear]{elsarticle} %review=doublespace preprint=single 5p=2 column
-%%% Begin My package additions %%%%%%%%%%%%%%%%%%%
+%  LaTeX support: latex@mdpi.com
+%  For support, please attach all files needed for compiling as well as the log file, and specify your operating system, LaTeX version, and LaTeX editor.
 
-\usepackage[hyphens]{url}
+%=================================================================
+% pandoc conditionals added to preserve backwards compatibility with previous versions of rticles
 
-  \journal{An awesome journal} % Sets Journal name
+\documentclass[An awesome
+journal,,,moreauthors,pdftex]{Definitions/mdpi}
 
-\usepackage{graphicx}
-%%%%%%%%%%%%%%%% end my additions to header
+%% Some pieces required from the pandoc template
+\setlist[itemize]{leftmargin=*,labelsep=5.8mm}
+\setlist[enumerate]{leftmargin=*,labelsep=4.9mm}
 
-\usepackage[T1]{fontenc}
-\usepackage{lmodern}
-\usepackage{amssymb,amsmath}
-% TODO: Currently lineno needs to be loaded after amsmath because of conflict
-% https://github.com/latex-lineno/lineno/issues/5
-\usepackage{lineno} % add
-\usepackage{ifxetex,ifluatex}
-\usepackage{fixltx2e} % provides \textsubscript
-% use upquote if available, for straight quotes in verbatim environments
-\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
-\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
-  \usepackage[utf8]{inputenc}
-\else % if luatex or xelatex
-  \usepackage{fontspec}
-  \ifxetex
-    \usepackage{xltxtra,xunicode}
-  \fi
-  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
-  \newcommand{\euro}{€}
-\fi
-% use microtype if available
-\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
-\usepackage[]{natbib}
-\bibliographystyle{plainnat}
+%--------------------
+% Class Options:
+%--------------------
 
-\ifxetex
-  \usepackage[setpagesize=false, % page size defined by xetex
-              unicode=false, % unicode breaks when used with xetex
-              xetex]{hyperref}
-\else
-  \usepackage[unicode=true]{hyperref}
-\fi
-\hypersetup{breaklinks=true,
-            bookmarks=true,
-            pdfauthor={},
-            pdftitle={Trends of SDR in Thessaloniki},
-            colorlinks=false,
-            urlcolor=blue,
-            linkcolor=magenta,
-            pdfborder={0 0 0}}
+%---------
+% article
+%---------
+% The default type of manuscript is "article", but can be replaced by:
+% abstract, addendum, article, book, bookreview, briefreport, casereport, comment, commentary, communication, conferenceproceedings, correction, conferencereport, entry, expressionofconcern, extendedabstract, datadescriptor, editorial, essay, erratum, hypothesis, interestingimage, obituary, opinion, projectreport, reply, retraction, review, perspective, protocol, shortnote, studyprotocol, systematicreview, supfile, technicalnote, viewpoint, guidelines, registeredreport, tutorial
+% supfile = supplementary materials
 
-\setcounter{secnumdepth}{5}
-% Pandoc toggle for numbering sections (defaults to be off)
+%----------
+% submit
+%----------
+% The class option "submit" will be changed to "accept" by the Editorial Office when the paper is accepted.
+This will only make changes to the frontpage (e.g., the logo of the journal will get visible), the headings, and the copyright information.
+Also, line numbering will be removed.
+Journal info and pagination for accepted papers will also be assigned by the Editorial Office.
+
+%------------------
+% moreauthors
+%------------------
+% If there is only one author the class option oneauthor should be used. Otherwise use the class option moreauthors.
+
+%---------
+% pdftex
+%---------
+% The option pdftex is for use with pdfLaTeX. Remove "pdftex" for (1) compiling with LaTeX & dvi2pdf (if eps figures are used) or for (2) compiling with XeLaTeX.
+
+%=================================================================
+% MDPI internal commands - do not modify
+\firstpage{1}
+\makeatletter
+\setcounter{page}{\@firstpage}
+\makeatother
+\pubvolume{1}
+\issuenum{1}
+\articlenumber{0}
+\pubyear{2023}
+\copyrightyear{2023}
+%\externaleditor{Academic Editor: Firstname Lastname}
+\datereceived{ }
+\daterevised{ } % Comment out if no revised date
+\dateaccepted{ }
+\datepublished{ }
+%\datecorrected{} % For corrected papers: "Corrected: XXX" date in the original paper.
+%\dateretracted{} % For corrected papers: "Retracted: XXX" date in the original paper.
+\hreflink{https://doi.org/} % If needed use \linebreak
+%\doinum{}
+%\pdfoutput=1 % Uncommented for upload to arXiv.org
+
+%=================================================================
+% Add packages and commands here.
+The following packages are loaded in our class file: fontenc, inputenc, calc, indentfirst, fancyhdr, graphicx, epstopdf, lastpage, ifthen, float, amsmath, amssymb, lineno, setspace, enumitem, mathpazo, booktabs, titlesec, etoolbox, tabto, xcolor, colortbl, soul, multirow, microtype, tikz, totcount, changepage, attrib, upgreek, array, tabularx, pbox, ragged2e, tocloft, marginnote, marginfix, enotez, amsthm, natbib, hyperref, cleveref, scrextend, url, geometry, newfloat, caption, draftwatermark, seqsplit
+% cleveref: load \crefname definitions after \begin{document}
+
+%=================================================================
+% Please use the following mathematics environments: Theorem, Lemma, Corollary, Proposition, Characterization, Property, Problem, Example, ExamplesandDefinitions, Hypothesis, Remark, Definition, Notation, Assumption
+%% For proofs, please use the proof environment (the amsthm package is loaded by the MDPI class).
+
+%=================================================================
+% Full title of the paper (Capitalized)
+\Title{Trends of SDR in Thessaloniki}
+
+% MDPI internal command: Title for citation in the left column
+\TitleCitation{Trends of SDR in Thessaloniki}
+
+% Author Orchid ID: enter ID or remove command
+%\newcommand{\orcidauthorA}{0000-0000-0000-000X} % Add \orcidA{} behind the author's name
+%\newcommand{\orcidauthorB}{0000-0000-0000-000X} % Add \orcidB{} behind the author's name
+
+% Authors, for the paper (add full first names)
+\Author{Athanasios Natsis$^{}$, Alkiviadis Bais$^{}$}
+
+%\longauthorlist{yes}
+
+% MDPI internal command: Authors, for metadata in PDF
+\AuthorNames{Athanasios Natsis, Alkiviadis Bais}
+
+% MDPI internal command: Authors, for citation in the left column
+
+% Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
+\address{%
+}
+
+% Contact information of the corresponding author
+\corres{Correspondence: }
+
+% Current address and/or shared authorship
+
+% The commands \thirdnote{} till \eighthnote{} are available for further notes
+
+% Simple summary
+
+%\conference{} % An extended version of a conference paper
+
+% Abstract (Do not insert blank lines, i.e. \\)
+\abstract{The shortwave downward solar irradiance (SDR) is an important
+factor that drives climate processes, production and can affect all
+living organisms. While monitoring the long term variability, there are
+observations of upward and downward SDR trends on different locations
+around the world for different time periods. Periods of positive tredns
+are refered as brightening periods and with negative tredns as dimming
+periods. We studied 29 years of CHP1 data from Thessaloniki, Greece,
+under three sky conditions (clear sky, cloudy sky and all sky
+conditions), applying a cloud sky identification algorithm. We found a
+positive trend for all-sky and clear-sky conditions, and also,
+investigated the consistency of those trends, the effect of the solar
+zenith angle, and the variation of the trends for the seasons of the
+year. We indentified that there are some anomalies in the long term SDR
+trends, for all sky conditions.}
+
+% Keywords
+\keyword{GHISDRSolar Brigthening/Dimming}
+
+% The fields PACS, MSC, and JEL may be left empty or commented out if not applicable
+%\PACS{J0101}
+%\MSC{}
+%\JEL{}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Only for the journal Diversity
+%\LSID{\url{http://}}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Only for the journal Applied Sciences
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Only for the journal Data
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Only for the journal Toxins
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Only for the journal Encyclopedia
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Only for the journal Advances in Respiratory Medicine
+%\addhighlights{yes}
+%\renewcommand{\addhighlights}{%
+
+%\noindent This is an obligatory section in “Advances in Respiratory Medicine”, whose goal is to increase the discoverability and readability of the article via search engines and other scholars.
+Highlights should not be a copy of the abstract, but a simple text allowing the reader to quickly and simplified find out what the article is about and what can be cited from it.
+Each of these parts should be devoted up to 2~bullet points.\vspace{3pt}\\
+%\textbf{What are the main findings?}
+% \begin{itemize}[labelsep=2.5mm,topsep=-3pt]
+% \item First bullet.
+% \item Second bullet.
+% \end{itemize}\vspace{3pt}
+%\textbf{What is the implication of the main finding?}
+% \begin{itemize}[labelsep=2.5mm,topsep=-3pt]
+% \item First bullet.
+% \item Second bullet.
+% \end{itemize}
+%}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-% From pandoc table feature
-\usepackage{longtable,booktabs,array}
-\usepackage{calc} % for calculating minipage widths
-% Correct order of tables after \paragraph or \subparagraph
-\usepackage{etoolbox}
-\makeatletter
-\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
-\makeatother
-% Allow footnotes in longtable head/foot
-\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
-\makesavenoteenv{longtable}
-
 \usepackage{subcaption}
+\usepackage{booktabs}
+\usepackage{longtable}
+\usepackage{array}
+\usepackage{multirow}
+\usepackage{wrapfig}
+\usepackage{float}
+\usepackage{colortbl}
+\usepackage{pdflscape}
+\usepackage{tabu}
+\usepackage{threeparttable}
+\usepackage{threeparttablex}
+\usepackage[normalem]{ulem}
+\usepackage{makecell}
+\usepackage{xcolor}
 
@@
 
-\begin{frontmatter}
-
-  \title{Trends of SDR in Thessaloniki}
-    \author[Laboratory of Atmospheric Physics]{Athanasios Natsis%
-  \corref{cor1}%
-  }
-   \ead{natsisphysicist@gmail.com} 
-    \author[Laboratory of Atmospheric Physics]{Alkiviadis Bais%
-  %
-  }
-   \ead{abais@auth.gr} 
-      \affiliation[Laboratory of Atmospheric Physics]{
-    organization={Aristotle University of
-Thessaloniki},addressline={Campus Box
-149},city={Thessaloniki},postcode={54124},country={Greece},}
-    \cortext[cor1]{Corresponding author}
-
-  \begin{abstract}
-  Study of GHI and DNI radiation for `clear-sky' and all-sky conditions.
-  It consists of two paragraphs.
-  \end{abstract}
-    \begin{keyword}
-    GHI \sep SDR \sep 
-    Solar Brigthening/Dimming
-  \end{keyword}
-
- \end{frontmatter}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 The shortwave downward solar irradiance (SDR) at Earth's surface play a
-significant role, on it's climate. Changes of the SDR can be related to
+significant role, on its climate. Changes of the SDR can be related to
 changes on Earth's energy budget, the mechanisms of climate change, and
 water and carbon cycle \citep{Wild2009}. Can also affect, solar and
-agricultural production, and living organisms. Studies of SDR
+agricultural production, and all living organisms. Studies of SDR
 variability, have identified some distinct SDR trends on different
-regions of the world and for different time periods. The term
-`brightening' is generally used to describe periods of positive SDR
-trend, and `dimming' for negative trend. There are many cases on the
-long term records of irradiance, showing a systematic change of SDR's
-trend slope, occurring roughly at the last decades of the 20th century.
-On multiple station in China, a dimming period was reported until about
-2000, followed by a brightening period \citep{Yang2021}. A similar
-pattern was identified, with the breaking point around 1980, for
-stations in Central Europe \citep{Wild2021} and Brazil
-\citep{Yamasoe2021}. Also, on global scale, an AI aided continental
-level spatial analysis, with data from multiple station, reach similar
-conclusions for the above regions and for the global trend
-\citep{Yuan2021}.
+regions of the world on different time periods. The term `brightening'
+is generally used to describe periods of positive SDR trend, and
+`dimming' for negative trend. There are many cases on the long term
+records of irradiance, showing a systematic change of SDR's trend slope,
+occurring roughly at the last decades of the 20th century. On multiple
+station in China, a dimming period was reported until about 2000,
+followed by a brightening period \citep{Yang2021}. A similar pattern was
+identified, with the breaking point around 1980, for stations in Central
+Europe \citep{Wild2021} and Brazil \citep{Yamasoe2021}. Also, on global
+scale, an AI aided continental level spatial analysis, with data from
+multiple station, reach similar conclusions for the above regions and
+for the global trend \citep{Yuan2021}.
 
-There is a consensus, among researchers, that major factors of SDR
-attenuation is the interaction with atmospheric aerosols and clouds.
-Those interactions, among other factors, have been analysed with models
-\citep{Li2016, Samset2018}, showing the existence of feedback mechanisms
-between the two. Similar finds, have been showed in observational data
-\citep[ and references
+There is a consensus, among researchers, that the major factors of SDR
+attenuation is the interaction of Sun radiation with atmospheric
+aerosols and clouds. Those interactions, among other factors, have been
+analysed with models \citep{Li2016, Samset2018}, showing the existence
+of feedback mechanisms between the two. Similar finds, have been showed
+in observational data \citep[ and references
 therein]{Schwarz2020, Ohvril2009, Zerefos2009, Xia2007}.
@@
 Due to the variability of the phenomenon, and its contributing factors,
-there is a need to further investigate and monitor SDR, in different
+there is a constant need to investigate and monitor SDR, in different
 sites, to estimate its magnitude, and its relation to the local
@@
 
-The SDR is equivalent to the whole sky Global radiation (or Global
-Horizontal Irradiance, GHI), and was obtained with a horizontal leveled,
-CM-21 pyranometer, from 1993-04-12 to 2023-03-31. Also, we have
-measurements of direct beam radiation (DNI) obtained by a collocated
-CHP-1 pyrheliometer, from 2016-04-01 to 2023-03-31. Although, we
-performed a similar analysis to the DNI as for the SDR, the results are
-not presented here, as they lack the appropriate statistical
-significance, due to the sorter duration of the data. Moreover, the DNI
-was used as auxiliary data, with the clear sky identification algorithm
-(CSid), to aid with the selection of the appropriate thresholds. We have
-to note also, that despite the capability of the CSid to use the DNI as
-an characterization parameter, we haven't use it, to avoid any selection
-bias due to unequal availability of measurements.
-
 The SDR data were measured with a Kipp \& Zonen CM-21 pyranometer
 operating continuously at the Laboratory of Atmospheric Physics of the
-Aristotle University of Thessaloniki (\(40^\circ38'\)N,
-\(22^\circ57'\)E, 80m a.s.l.). The monitoring site is located near the
-city centre. During the study period the CM-21 pyranometer has been
-independently calibrated four times. There are three distinct steps to
-the creation of this dataset: a) the acquisition of radiation
-measurements from the sensors, b) a radiation data quality check, and c)
-the identification of ``clear sky'' conditions from the radiometric data
-d) data aggregation and trend analysis.
+Aristotle University of Thessaloniki (\(40^\circ\,38'\,\)N,
+\(22^\circ\,57'\,\)E, \(80\,\)m~a.s.l.) in the period from 1993-04-13 to
+2023-04-13. The monitoring site is located near the city centre, and we
+expect to be affected by the urban environment. During the study period,
+the pyranometer has been independently calibrated three times at the
+Meteorologisches Observatorium Lindenberg, DWD, when it was verified the
+stability of the instrument to within better than \(0.7\%\) relative to
+the initial calibration by the manufacturer. Along with SDR, the direct
+beam radiation (DNI) was also measured by a collocated Kipp \& Zonen
+CHP-1 pyrheliometer, for the period 2016-04-01 to 2023-04-13. Although,
+we have performed a similar analysis to the DNI data the results are not
+presented here, as they lack the appropriate statistical significance,
+due to the sorter duration of the data. However, the DNI data were used
+as auxiliary data, in the clear sky identification algorithm (CSid),
+which is discussed later, for the selection of the appropriate
+thresholds. It is noted that despite the capability of the CSid
+algorithm to use the DNI as a characterization parameter, we haven't
+utilized it here, to avoid any selection bias, due to unequal length of
+the two datasets. There are four distinct steps in the creation of the
+dataset analysed here: a)~the acquisition of radiation measurements from
+the sensors, b)~the data quality check, c)~the identification of ``clear
+sky'' conditions from the radiometric data, and d)~the aggregation of
+data and trend analysis.
 
-For the acquisition of radiometric data, the signal of the broadband
-instruments is sampled with a rate of \(1 \text{Hz}\). The mean and its
-standard deviation are recorded for every minute. The measurements are
-corrected for the zero offset of the instrument signal. As reference of
-the ``dark signal'' is used the Sun elevation angle bellow
-\(-10^\circ\), for a period of \(3 \text{h}\). The signal is converted
-to radiation flux, using a ramped value of the instrument sensitivity,
-derived from the eight laboratory calibrations of the instrument, during
-the study period.
+For the acquisition of radiometric data, the signal of the pyranometer
+is sampled with a rate of \(1\,\text{Hz}\). The mean and the standard
+deviation of these samples are recorded every minute. The measurements
+are corrected for the zero offset (``dark signal'' in volts). The ``dark
+signal'' is calculated by averaging all measurements recorded for a
+period of \(3\,\text{h}\), before (morning) or after (evening) the Sun
+reaches an elevation angle of \(-10^\circ\). The signal is converted to
+irradiance using a ramped value of the instrument's sensitivity between
+calibrations.
 
-A manual screening was performed, to remove inconsistencies and
-erroneous recordings, that can occur randomly or systematically, during
-the long continuous operation of the instrument. The manual screening is
-aided by a radiation data quality assurance procedure, adjusted for this
-site, based on methods of Long and
+A manual screening was performed, to remove inconsistent and erroneous
+recordings that can occur stochastically or systematically, during the
+continuous operation of the instruments. The manual screening is aided
+by a radiation data quality assurance procedure, adjusted for the site,
+which is based on the methods of Long and
 Shi~\citetext{\citeyear{Long2008a}; \citeyear{Long2006}}. Thus,
 problematic recordings have been excluded from further processing.
-Although it is not possible to detect all the bad data, the large number
-of data and the aggregation scheme we use, can provide us with accurate
-radiometric measurements, for the scope of this study.
+Although it is impossible to detect all false data, the large number of
+available data, and the aggregation scheme we used, ensures the good
+quality of the radiometric measurements used in this study.
 
-In order to estimate the effect of the clouds on the SDR, we created
-three datasets, by characterizing each one-minute measurement with a
-corresponding sky condition. The all-sky conditions, containing all the
-valid measurements. The clear-sky conditions data, correspond to data
-where we have inferred that the sky was almost clear of clouds, and the
-remainder part the data was characterized with cloud sky conditions
-(cloud-sky). To identify the clear-sky conditions we used the method
-proposed by \citet{Long2000} and by \citet{Reno2016}, that was adapted
-and configured for the site.
-
-\ldots.
-
-cite iterative process
-
-\ldots\ldots{}
+In order to be able to estimate the effect of the sky condition on the
+long term variability of SDR, we created three datasets, by
+characterizing each one-minute measurement with a corresponding sky
+condition (i.e., all-sky, clear-sky and cloudy-sky). To identify the
+clear-sky conditions we used a method proposed by \citet{Long2000} and
+by \citet{Reno2016}, which was adapted and configured for the site, as
+the authors suggest.
 
 We have to note, that the definition of what constitutes as clear or
-cloudy sky, has some subjectivity, due to the used method of
-characterization. As a result, the details of the definition are site
-specific, it relies on a combination of threshold of comparisons with
-ideal actinometric models and statistics on different signal behavior.
+cloudy sky, has some subjectivity, in any method of characterization. As
+a result, the details of the definition are site specific, it relies on
+a combination of thresholds and comparisons with ideal actinometric
+models and statistical analysis on different signal metrics. The CSid
+algorithm was calibrated with the main focus, to identify the presence
+of clouds on the sky dome. Although the fine-tuning of the procedure,
+few marginal cases exist, that have been identified manually as false
+positive or false negative but cannot affect the final results of the
+study.
 
-The algorithm was calibrated with the main focus to identify the
-presence of clouds on the sky dome, although there are marginal cases
-that there will be false positives or false negatives identifications.
-
-As global radiation clear sky reference we are using the Haurwitz's
-model, adjusted for our site with a factor of 0.965 (Eq. \ref{eq:ahau}).
-The selection of a clear sky reference model, was based on SDR
-observation from the period 2016 -- 2021. Where, after an iterative
-optimization of eight simple models (Daneshyar--Paltridge--Proctor,
-Kasten--Czeplak, Haurwitz, Berger--Duffie,
-Adnot--Bourges--Campana--Gicquel, Robledo-Soler, Kasten and
-Ineichen-Perez) with different factors. We found, that Haurwitz's model,
-adjusted with a factor of 0.965 has the lower root mean squared error
-(RMSE). The tried models are described by \citet{Reno2012} and tested by
-\citet{Reno2016}. The iterative optimization method, for the selection
-of the reference is discussed by \citet{Long2000} and \citet{Reno2016}.
+For completeness, we will provide below a brief overview of the clear
+sky identification algorithm (CSid), along with the site specific
+thresholds. To calculate the reference clear sky
+\(\text{SDR}_\text{CSref}\) we used the \(\text{SDR}_\text{Haurwitz}\)
+derived by the radiation model of \citet{Haurwitz1945}, adjusted for our
+site with a factor \(a\) (Eq.~\ref{eq:ahau}), resulted by an iterative
+optimization process, as described by \citet{Long2000} and
+\citet{Reno2016}. The target of the optimization was the minimization of
+a function \(f(a)\) (Eq.~\ref{eq:minf}) and was accomplished with the
+algorithmic function ``optimise'', which is an implementation based on
+the work of \citet{Brent1973}, from the library ``stats'' of the R
+programming language \citep{RCT2023}. The optimization and the selection
+of the clear sky reference model, was performed on SDR observations for
+the period 2016 - 2021. During the optimization, eight simple clear sky
+radiation models were tested (Daneshyar-Paltridge-Proctor,
+Kasten-Czeplak, Haurwitz, Berger-Duffie, Adnot-Bourges-Campana-Gicquel,
+Robledo-Soler, Kasten and Ineichen-Perez), with a wide range of factors.
+These models are described in more details by \citet{Reno2012} and
+evaluated by \citet{Reno2016}. We found, that Haurwitz's model, adjusted
+with the factor \(a = 0.965\) yields one of the lowest root mean squared
+errors (RMSE), while the procedure, manages to characterize the majority
+of the data. Thus, our clear sky reference is derived by the
+Eq.~\ref{eq:ahau}.
 
 \begin{equation}
-\text{SDR}_\text{Clear Sky} = 0.965 \times 1098 \times \cos( \text{SZA} ) \times \exp \left( \frac{ - 0.057}{\cos(\text{SZA})} \right) \label{eq:ahau}
+f(a) = \frac{1}{n}\sum_{i=1}^{n} ( \text{SDR}_{\text{CSid},i} - a \times \text{SDR}_{\text{testCSref},i} )^2 \label{eq:minf}
 \end{equation}
 
-The following criteria and thresholds were used to identify clear-sky
-conditions. Each criterion was applied for a running window of \(11\)
-consecutive measurements/minutes, and the characterization is applied at
-the center value of the window. A data point is consider as under
-cloud-sky condition if it fails to pass any of the criteria, all the
-other data points are characterized as clear-sky.
-
-The mean of the measured value \(\overline{G}_i\) must be inside an
-envelope based on the reference model \(\text{SDR}_\text{Clear Sky}\)
-(Eq. \ref{eq:MeanVIP}).
+where: \(n\) is the total number of daylight records,
+\(\text{SDR}_{\text{CSid},i}\) are the records identified as clear sky
+by CSid, \(a\) is a hypothetical adjustment factor, and
+\(\text{SDR}_{\text{testCSref},i}\) is any of the tested clear sky
+radiation models.
 
 \begin{equation}
-0.91 \times \overline{\text{SDR}}_{i\text{Clear Sky}} - 20
-< \overline{G}_i <
-1.095 \times \overline{\text{SDR}}_{i\text{Clear Sky}} + 30
+\text{SDR}_\text{CSref} = a \times \text{SDR}_\text{Haurwitz} = 0.965 \times 1098 \times \cos(\theta) \times \exp \left( \frac{ - 0.057}{\cos(\theta)} \right) \label{eq:ahau}
+\end{equation}
+
+where: \(\text{SDR}_\text{CSref}\) is the reference clear sky SDR, in
+\(\text{w}\,\text{m}^{-2}\) and \(\theta\) is the solar zenith angle
+(SZA).
+
+The criteria that were used to identify whether a measurement was taken
+under clear-sky conditions are presented below. A data point is flagged
+as ``clear-sky'' if all criteria are satisfied, otherwise it is
+considered to be ``cloud-sky''. Each criterion was applied for a running
+window of \(11\) consecutive one-minute measurements, and the
+characterization is assigned to the central value of the window. Each
+parameter, was calculated both from the observations and the reference
+clear sky model, for each comparison. The allowable range of variation
+is defined by the model-derived value of the parameter multiplied by a
+factor plus an offset. The factors and the offsets were determined
+empirically, by manual inspecting each filters performance on selected
+days, and adjusting them accordingly, during an iterative process.
+
+\begin{enumerate}
+\def\labelenumi{\alph{enumi})}
+\tightlist
+\item
+  Mean of the measured \(\overline{\text{SDR}}_i\) (Eq.
+  \ref{eq:MeanVIP}). \begin{equation}
+  0.91 \times \overline{\text{SDR}}_{\text{CSref},i} - 20
+  < \overline{\text{SDR}}_i <
+  1.095 \times \overline{\text{SDR}}_{\text{CSref},i} + 30
   \label{eq:MeanVIP}
   \end{equation}
+\end{enumerate}
 
-The running max measured value \(M_{Gi} = max[\text{SDR}_{i}]\), is
-compared to a similar constructed value from the reference
-\(M_{CSi} = max[\text{SDR}_{i\text{Clear Sky}}]\) (Eq. \ref{eq:MaxVIP}).
-
+\begin{enumerate}
+\def\labelenumi{\alph{enumi})}
+\setcounter{enumi}{1}
+\tightlist
+\item
+  Maximum measured value \(M_{\text{}}\) (Eq.~\ref{eq:MaxVIP}).
   \begin{equation}
-1 \times M_{CSi} - 75
-< M_{Gi} <
-1 \times M_{CSi} + 75
+  1 \times M_{\text{CSref},i} - 75
+  < M_{\text{}i} <
+  1 \times M_{\text{CSref},i} + 75
   \label{eq:MaxVIP}
   \end{equation}
+\end{enumerate}
 
-The length \(L\) (Eq. \ref{eq:VILeq}) of the sequence of line segments
-connecting the points of the SDR time series for the measured values
-\(L\) and similar for the reference \(L_{CS}\), must be within the
-limits of Eq. \ref{eq:VILcr}.
+\begin{enumerate}
+\def\labelenumi{\alph{enumi})}
+\setcounter{enumi}{2}
+\tightlist
+\item
+  Length \(L_i\) of the sequential line segments, connecting the points
+  of the \(11\) SDR values (Eq. \ref{eq:VILeq}). \begin{equation}
+  L_i = \sum_{i=1}^{n-1}\sqrt{\left ( \text{SDR}_{i+1} - \text{SDR}_{i}\right )^2 + \left ( t_{i+1} - t_i \right )^2}
+  \label{eq:VILeq}
+  \end{equation} \begin{equation}
+  1 \times L_{\text{CSref},i} - 5 < L_i < 1.3 \times L_{\text{CSref},i} + 13
+  \label{eq:VILcr}
+  \end{equation} where: \(t_i\) is the time each SDR measurement has
+  been measured
+\end{enumerate}
 
-\begin{equation}
-L = \sum_{i=1}^{n-1}\sqrt{\left ( \text{SDR}_{i+1} - \text{SDR}_{i}\right )^2 + \left ( t_{i+1} - t_i \right )^2}
-\label{eq:VILeq}
+\begin{enumerate}
+\def\labelenumi{\alph{enumi})}
+\setcounter{enumi}{3}
+\tightlist
+\item
+  Standard deviation \(\sigma_i\) of the slope (\(s_i\)) between the
+  \(11\) sequential points, normalized by the mean
+  \(\overline{\text{SDR}}_i\) (Eq.~\ref{eq:VCT1}). \begin{gather}
+    \sigma_i = \frac {\sqrt{\frac{1}{n-1} \sum_{i=1}^{n-1} \left( s_i - \bar{s} \right)^2}} {\overline{\text{SDR}}_i} \label{eq:VCT1} \\
+    s_i = \frac{\text{SDR}_{i+1} - \text{SDR}_{i}}{t_{i+1} - t_i},\;\;   \bar{s} = \frac{1}{n-1} \sum_{i=1}^{n-1} s_i,\;\;\forall i \in \left \{ 1, 2, \ldots, n-1 \right \}\;\;
+  \end{gather} For this criterion, \(\sigma_i\) should be below a
+  certain threshold (Eq.~\ref{eq:VCTcr}): \begin{equation}
+    \sigma_i < \ensuremath{1.1\times 10^{-4}} \label{eq:VCTcr}
   \end{equation}
+\end{enumerate}
 
-\begin{equation}
-1 \times L_{CSi} - 5 < L_i < 1.3 \times L_{CSi} + 13
-\label{eq:VILcr}
+\begin{enumerate}
+\def\labelenumi{\alph{enumi})}
+\setcounter{enumi}{4}
+\tightlist
+\item
+  Maximum difference \(X_i\) between the change in measured irradiance
+  and the change in clear sky irradiance over each measurement interval.
+  \begin{gather}
+    X_i = \max{\left \{ \left | x_i - x_{\text{CSref},i} \right | \right \}} \label{eq:VSM3} \\
+    x_i = \text{SDR}_{i+1} - \text{SDR}_{i} \forall i \in \left \{ 1, 2, \ldots, n-1 \right \} \label{eq:VSM1} \\
+    x_{\text{CSref},i} = \text{SDR}_{\text{CSref},i+1} - \text{SDR}_{\text{CSref},i} \forall i \in \left \{ 1, 2, \ldots, n-1 \right \} \label{eq:VSM2}
+  \end{gather} For this criterion, \(X_i\) should be below a certain
+  threshold (Eq.~\ref{eq:VSMcr}): \begin{equation}
+    X_i < 7.5 \label{eq:VSMcr}
   \end{equation}
+\end{enumerate}
 
-We calculate the standard deviation \(\sigma\) of the slope (\(s\))
-between sequential points in the time series, normalized by the average
-SDR during the time interval.
-
-\begin{equation}
-s_i = \frac{\text{SDR}_{i+1} - \text{SDR}_{i}}{t_{i+1} - t_i}, \forall i \in \left \{ 1, 2, \ldots, n-1 \right \},\;\;
-\bar{s} = \frac{1}{n-1} \sum_{i=1}^{n-1} s_i,\;\;
-\sigma_i = \frac {\sqrt{\frac{1}{n-1} \sum_{i=1}^{n-1} \left( s_i - \bar{s} \right)^2} } {\bar{G_i}} \label{eq:VCT3b}
-\end{equation}
-
-\begin{equation}
-\sigma_i < \ensuremath{1.1\times 10^{-4}}
-\label{eq:VCTcr}
-\end{equation}
-
-The maximum difference \(X\) between the change in measured irradiance
-and the change in clear sky irradiance over each measurement interval.
-
-\begin{gather}
-x_i = \text{SDR}_{i+1} - \text{SDR}_{i} \forall i \in \left \{ 1, 2, \ldots, n-1 \right \} \label{eq:VSM1} \\
-x_{CS,i} = \text{SDR}_{CS,i+1} - \text{SDR}_{CS,i} \forall i \in \left \{ 1, 2, \ldots, n-1 \right \} \label{eq:VSM2} \\
-X_i = \max{\left \{ \left | x_i - x_{CS,i} \right | \right \}} \label{eq:VSM3}
-\end{gather}
-
-\begin{equation}
-X_i < 7.5
-\label{eq:VSMcr}
-\end{equation}
-
-Due to a significant measurement uncertainty near, the horizon, we have
+Due to a significant measurement uncertainty near the horizon, we have
 to exclude all measurements with SZA greater than \(85^\circ\).
@@
 location. To make the measurements comparable throughout the dataset, we
-adjusted all 1-minute radiometric values to the mean Sun - Earth
+adjusted all one-minute radiometric values to the mean Sun - Earth
 distance. Subsequently, we made all measurements relative to the Total
-Solar Irradiance (TSI) at \(1 \text{au}\), in order to compensate for
-the Sun's intensity variability, using a homogenized time series of
-satellite TSI observations. The TSI data we use, comes from NOAA
-\citep{Coddington2005}, it overlaps with our data, and is interpolated
-to match with the time step of the data. As a result, we can present all
-radiation data as a fraction of Sun's TSI.
+Solar Irradiance (TSI) at \(1\,\text{au}\), in order to compensate for
+the Sun's intensity variability, using a time series of satellite TSI
+observations. The TSI data we use are part of the ``NOAA Climate Data
+Record of Total Solar Irradiance'' dataset \citep{Coddington2005}. Where
+the initial daily values, were interpolated to match with the time step
+of our measurements. The final dataset contains \(6589967\) one-minute
+measurements, of which, \(84.2\%\) were identified as under clear-sky
+conditions and subsequently \(15.8\%\) as under cloud-sky conditions.
 
-\hypertarget{aggregation-of-radiometric-data}{%
-\subsubsection{Aggregation of radiometric
-data}\label{aggregation-of-radiometric-data}}
-
-Before further analysis and deseasonalization, we implement an
-appropriate aggregation scheme on the 1-minute data. To preserve the
-representativeness we use the following criteria: a) for daily mean
-values we exclude instances with less than 180 valid data points, b)
-accordingly, monthly values are computed by daily aggregated data, where
-months with less than 20 days are rejected. The daily seasonal values
-were indexed based on the day of year number, and the monthly by the
-corresponding calendar month. For the seasons of the year trends, we
-grouped the mean daily values by season (December - February: Winter,
-March - May: Spring, etc.). Finally, for each data set, we remove the
-natural occurring seasonal variation. This is done, by calculating the
-mean values for the appropriate time step, and subtracting the annual
-cycle from the actual data. In addition to the previous aggregation
-scheme, each dataset was aggregate in \(1^\circ\) SZA bins, separate for
-cases before and after local noon. This gave us an approximation of
-Sun's SZA contribution to the SDR brightening.
-
-Thus, for each data set we have obtained the relative departure of the
-seasonal mean. The statistical significance and aggregation scheme and
-filtering will be noted along with the corresponding results.
+In order to investigate the SDR trends, we implemented an appropriate
+aggregation scheme to the 1-minute data to derive a series in coarser
+timescale. To preserve the representativeness of the data we used the
+following criteria: a) for the daily mean values we accept days with
+more than 50\% of the daytime measurements, present and valid, b)
+monthly values were computed from daily means only when at least 20 days
+were available. To create the daily and monthly climatological means, we
+averaged the data based on the day of year and calendar month,
+respectively. For the seasonal means we averaged the mean daily values
+in each season (Winter: December - February, Spring: March - May, etc.).
+Finally, each data set was deseasonalized by subtracting the
+corresponding climatological annual cycle (daily or monthly) from the
+actual data. To estimate SZA contribution to the SDR trends, the
+one-minute data were aggregated in \(1^\circ\) SZA bins, separately for
+the morning and afternoon hours, and then were deseasonalized as
+mentioned above.
 
@@
 
-We calculated the SDR trends, from the mean daily values of the
-departure from the seasonal values, for the whole time series for the
-three sky conditions (Table \ref{tab:trendtable}).
+We calculated the linear SDR trends, from the departures of the mean
+daily values from the seasonal values, and for the three sky conditions
+(Table~\ref{tab:trendtable}). In Figure~\ref{fig:trendALL} we present
+only the time series under all-sky conditions; the plots for clear-sky
+and cloud-sky conditions, are very similar and are shown in the Appendix
+(Figures~\ref{fig:trendCLEAR} and~ \ref{fig:trendCLOUD}). We observe a
+positive trend for all-sky conditions (\(0.38\,\%/y\)), a very close but
+lower trend for clear-skies (\(0.35\,\%/y\)) and a negative weaker trend
+for cloud-skies (\(-0.28\,\%/y\)). In the studied period, there is no
+significant break or change in the variability of the time series. Other
+studies for the European region report a change of the SDR slope, around
+1980, a few years before the start of our records
+\citep{Wild2021, Yuan2021, Ohmura2009}. It is interesting to note, that
+for the observation period, the trend of the TSI is \(-2e-04\,\%/y\),
+and thus we can not attribute any major effect on SDR trend to Solar
+variability.
+
+\begin{table}[!h]
+
+\caption{\label{tab:trendtable}\label{tab:trendtable}Trends in SDR daily means for different sky conditions for the period 1993 - 2023.}
+\centering
+\begin{tabular}[t]{crrc}
+\toprule
+Sky conditions & Trend [\%/year] & Statistical signif. [\%] & Days\\
+\midrule
+All sky & 0.376 & 100.00 & 10256\\
+Clear sky & 0.349 & 100.00 & 10256\\
+Cloudy sky & -0.276 & 99.97 & 5067\\
+\bottomrule
+\end{tabular}
+\end{table}
 
@@
 
-\caption{Anomaly (\%) of the daily SDR relative to climatological values for 1993 - 2023  `r year(min(ALL_1_daily_DESEAS$Date))` to `r year(max(ALL_1_daily_DESEAS$Date))`.
-The black line shows the long term trend for all-sky conditions.}\label{fig:trendALL}
+\caption{Anomaly (\%) of the daily all-sky SDR, relative to climatological values for 1993 - 2023. The black line shows the long term linear trend.}\label{fig:trendALL}
 \end{figure}
 
-\footnotesize
-
-\begin{longtable}[]{@{}
-  >{\centering\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.2361}}
-  >{\centering\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.2222}}
-  >{\centering\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.2500}}@{}}
-\caption{\label{tab:trendtable}Trends of daily SDR means, for different
-sky conditions. All cases have a statistical significance above
-\(99.9\%\) \normalsize}\tabularnewline
-\toprule\noalign{}
-\begin{minipage}[b]{\linewidth}\centering
-Trend {[}\%/year{]}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
-Trend p-value
-\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
-Sky condition
-\end{minipage} \\
-\midrule\noalign{}
-\endfirsthead
-\toprule\noalign{}
-\begin{minipage}[b]{\linewidth}\centering
-Trend {[}\%/year{]}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
-Trend p-value
-\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
-Sky condition
-\end{minipage} \\
-\midrule\noalign{}
-\endhead
-\bottomrule\noalign{}
-\endlastfoot
-0.3996 & 8.665e-23 & All sky cond. \\
-0.3773 & 7.075e-18 & Clear sky cond. \\
-0.3594 & 3.318e-09 & Cloudy cond. \\
-\end{longtable}
-
-\hypertarget{long-term-trends-by-solar-zenith-angle.}{%
-\subsection{Long term trends by Solar zenith
-angle.}\label{long-term-trends-by-solar-zenith-angle.}}
-
-Analyzing the long term trends of SDR to bins of SZA, we can see the
-contribution of the geometry and time in a diurnal level (Figures
-\ref{fig:szatrends}a and \ref{fig:szatrends}b). Although there is a
-seasonal distribution of SZA than in not presented here.
+Although the year-to-year variability of the anomalies (Figure
+\ref{fig:trendALL} and Figures \ref{fig:trendCLEAR},
+\ref{fig:trendCLOUD} on Appendix), shows a rather homogeneous behavior,
+plots of the cumulative sums (CUSUM) of the anomalies can reveal
+different structures within all the three sky conditions
+\citep{Regier2019}. In the cases of all-sky and clear-sky (Figures
+\ref{fig:cusummonth-1} and \ref{fig:cusummonth-2}), we observe three
+macroscopic periods. A downward part from the start until about 2005, an
+relative steady part until about 2016 and finally, a steep upward part
+until the present. For cloud-sky (Figure~\ref{fig:cusummonth-3}), we
+have a different sequence, it begins with a relative steady part until
+1997, followed by a upward part until 2005, a long decline until 2020,
+with a small positive slope until the present. For a uniform trend, we
+would expect the CUSUMs of the anomalies to have a symmetric `V' shape.
+This would indicate that the anomalies are evenly distributed around the
+climatological mean, and for a positive uniform trend, the first half to
+be below and the other half above the climatological mean. In our case,
+there is more complex evolution of the phenomenon. Another distinct
+feature of the CUSUMs, is the different pattern of the cloudy-sky
+dataset which peaks around the middle of the period. Although, it seems
+to exist a complementary relation to the CUSUMs of the clear- and
+all-sky cases, we can not assert that clouds are the main drive for this
+relation, due to the great difference in the number of observational
+data between the datasets (Table~\ref{tab:trendtable}).
 
@@
 
-{\centering \subfloat[Trend distribution for all-sky conditions.\label{fig:szatrends-1}]{\includegraphics[width=.35\linewidth]{./images/SzaTrends-13} }\subfloat[Trend distribution for cloud-sky conditions.\label{fig:szatrends-2}]{\includegraphics[width=.35\linewidth]{./images/SzaTrends-21} }
+{\centering \subfloat[All-skies.\label{fig:cusummonth-1}]{\includegraphics[width=.32\linewidth]{./images/CumulativeMonthlyCuSum-1} }\subfloat[Clear-skies.\label{fig:cusummonth-2}]{\includegraphics[width=.32\linewidth]{./images/CumulativeMonthlyCuSum-5} }\subfloat[Cloud-skies.\label{fig:cusummonth-3}]{\includegraphics[width=.32\linewidth]{./images/CumulativeMonthlyCuSum-9} }
 
@@
 
-\caption{Distribution of the SDR's long term trends by SZA.}\label{fig:szatrends}
+\caption{Running cumulative sum plots of the monthly SDR.}\label{fig:cusummonth}
+\end{figure}
+
+In order to investigate more, the features within the trends, we created
+another set of CUSUMs plots by subtracting the corresponding long term
+trend from the SDR anomaly data, prior to the CUSUM calculation
+(Figure~\ref{fig:cusumnotrendmonthly}). With this approach periods when
+the CUSUMs diverge from zero can be interpreted as systematic variation
+of SDR from the climatological mean. When the CUSUM is increasing, the
+added values are above the climatological values of SDR trend and vice
+versa. Overall, for all- and clear-sky conditions
+(Figure~\ref{fig:cusumnotrendmonthly-1}
+and~\ref{fig:cusumnotrendmonthly-2}) we observe periods when the
+anomalies diverge from the climatological value, each lasting for
+several years. The pattern for the all-sky and clear-sky conditions is
+very similar, suggesting prevalence in clear skies over Thessaloniki. It
+is interesting that in the period 1993 - 2016 the anomalies of all-sky
+and clear-sky, have a high variability around zero. After 2016, the
+range of the variability is decreased at about one third of the prior
+period. For cloud-sky conditions
+(Figure~\ref{fig:cusumnotrendmonthly-3}) the period 1997 - 2008 is
+dominated by positive CUSUMs suggesting a reduced effect of clouds on
+SDR. From 1997 to mid-2000s CUSUMs are increasing, likely due to a
+continuous decrease in the thickness of clouds, followed by a period of
+rapid increase (within 3 years) in cloud optical thickness lasting up to
+2008. The following stable period spans for about 15 years up to 2021
+when CUSUMs start increasing again.
+
+\begin{figure}[h!]
+
+{\centering \subfloat[All-skies.\label{fig:cusumnotrendmonthly-1}]{\includegraphics[width=.32\linewidth]{./images/CumulativeMonthlyCuSumNOtrend-1} }\subfloat[Clear-skies.\label{fig:cusumnotrendmonthly-2}]{\includegraphics[width=.32\linewidth]{./images/CumulativeMonthlyCuSumNOtrend-5} }\subfloat[Cloud-skies.\label{fig:cusumnotrendmonthly-3}]{\includegraphics[width=.32\linewidth]{./images/CumulativeMonthlyCuSumNOtrend-9} }
+
+}
+
+\caption{Running cumulative sum plots of the monthly SDR after removing the long term trend from the monthly values.}\label{fig:cusumnotrendmonthly}
+\end{figure}
+
+\hypertarget{effects-of-the-solar-zenith-angle-on-sdr.}{%
+\subsection{Effects of the solar zenith angle on
+SDR.}\label{effects-of-the-solar-zenith-angle-on-sdr.}}
+
+The solar zenith angle is a major driver of the attenuation of SDR
+reaching the ground, as SDR is affected by the enhancement of the
+radiation path in the atmosphere, especially in urban environment with
+human activities emitting aerosols \citep{Wang2021}. In order to
+estimate the effect of the SZA on the SDR trends, we grouped the anomaly
+data in bins of \(1^\circ\) SZA, and calculated the overall trend for
+each bin before noon and after noon (Figure~\ref{fig:szatrends}).
+Although there are seasonal dependencies of the minimum SZA (see
+Appendix, Figure~\ref{fig:SZAtrendSeason}), these dependencies would not
+be further examined here. For all-sky and clear-sky conditions the
+brightening effect of SDR (positive trend) is stronger for large SZAs
+(Figures~\ref{fig:szatrends-1} and \ref{fig:szatrends-2}). The trends in
+the morning and afternoon hours are more or less consistent with small
+differences, which can be attributed to systematic diurnal variations of
+aerosols, particularly during the warm period of the year
+\citep{Wang2021}. For cloud-sky conditions
+(Figure~\ref{fig:szatrends-3}), we can not discern any significant
+dependence of the SDR trend with SZA. For SZAs \(16^\circ\) -
+\(50^\circ\), the trends range within about \(\pm 0.2\,\%/y\), with a
+weak statistical significance. Between \(50^\circ\) and \(75^\circ\) SZA
+the trends for the period before noon are stronger and negative,
+possibly associated with stronger attenuation by clouds under oblique
+incidence angles.
+
+\begin{figure}[h!]
+
+{\centering \subfloat[All-sky conditions.\label{fig:szatrends-1}]{\includegraphics[width=.32\linewidth]{./images/SzaTrends-1} }\subfloat[Clear-sky conditions.\label{fig:szatrends-2}]{\includegraphics[width=.32\linewidth]{./images/SzaTrends-4} }\subfloat[Cloud-sky conditions.\label{fig:szatrends-3}]{\includegraphics[width=.32\linewidth]{./images/SzaTrends-7} }
+
+}
+
+\caption{Distribution of the SDR's long term trends by SZA before and after noon.
+The solid data points represent values with an acceptable statistical significance ($p<0.005$), the trends were computed with SZA bins of $1^\circ$ for each day, before and after local noon.}\label{fig:szatrends}
 \end{figure}
@@
 
-Similar we have produced the deseasonalized trend for different sky
-conditions for each season of the year, using the corresponding mean
-monthly values. This can give us a better understanding of the annual
-variability of the trends.
+Similar to the long term trends discussed above, we have calculated the
+trend of the anomalies for the three different sky conditions, and for
+each season of the year, using the corresponding mean monthly values
+(Figure~\ref{fig:seasonalALL} and Table~\ref{tab:trendseasontable}). For
+all-sky conditions the trend in SDR in winter is the largest
+(\(0.69\,\%/y\)), followed by the trend in autumn (\(0.43\,\%/y\), value
+similar to the long term trend) both statistically significant above the
+\(99\,\%\) confidence level. In spring and summer, the trends are much
+smaller and less statistical significant. These seasonal differences
+indicate a possible relation of the trends to clouds during winter and
+autumn. For clear-skies, the trend in winter is \(0.83\,\%/y\), larger
+than for all-skies (\(0.69\,\%/y\)), which is another indication of
+increasing cloud thickness. Moreover, the trends under clear- and
+cloud-sky conditions are almost complementary to each other,
+particularly for winter and autumn, where the signal is stronger. During
+spring and summer the statistical significance is very low and the
+actual trend too small for a meaningful comparison.
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{./images/SeasonalTrendsTogether-1}
+{\centering \includegraphics[width=1\linewidth]{./images/SeasonalTrendsTogether3-2}
 
@@
 
-\caption{Trends by season for all condition. Displaying monthly means of daily means.}\label{fig:seasonalALL}
+\caption{SDR trends by season of the year for the three sky conditions.
+Each point is a monthly mean calculated from daily means.
+The vertical scale is common for all graphs, each column of graphs corresponds to a sky condition and each row to a season of the year.}\label{fig:seasonalALL}
 \end{figure}
 
-\begin{longtable}[]{@{}
-  >{\centering\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2361}}
-  >{\centering\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2222}}
-  >{\centering\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2500}}
-  >{\centering\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1250}}@{}}
-\caption{\label{tab:trendseasontable}Trends of daily means by sky
-conditions for the seasons of the year. (continued
-below)}\tabularnewline
-\toprule\noalign{}
-\begin{minipage}[b]{\linewidth}\centering
-Trend {[}\%/year{]}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
-Trend p-value
-\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
-Sky condition
-\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
-Season
-\end{minipage} \\
-\midrule\noalign{}
-\endfirsthead
-\toprule\noalign{}
-\begin{minipage}[b]{\linewidth}\centering
-Trend {[}\%/year{]}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
-Trend p-value
-\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
-Sky condition
-\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
-Season
-\end{minipage} \\
-\midrule\noalign{}
-\endhead
-\bottomrule\noalign{}
-\endlastfoot
-0.7215 & 0.001209 & All sky cond. & Winter \\
-0.1537 & 0.1498 & All sky cond. & Spring \\
-0.128 & 0.08705 & All sky cond. & Summer \\
-0.4239 & 0.006659 & All sky cond. & Autumn \\
-0.8715 & 0.000269 & Clear sky cond. & Winter \\
-0.05881 & 0.6247 & Clear sky cond. & Spring \\
--0.02369 & 0.7972 & Clear sky cond. & Summer \\
-0.3583 & 0.02996 & Clear sky cond. & Autumn \\
--0.02581 & 0.9335 & Cloudy cond. & Winter \\
-0.4333 & 0.02937 & Cloudy cond. & Spring \\
-0.442 & 0.004988 & Cloudy cond. & Summer \\
-0.4776 & 0.01955 & Cloudy cond. & Autumn \\
-\end{longtable}
+\begin{table}[!h]
 
-\begin{longtable}[]{@{}
-  >{\centering\arraybackslash}p{(\columnwidth - 0\tabcolsep) * \real{0.4444}}@{}}
-\toprule\noalign{}
-\begin{minipage}[b]{\linewidth}\centering
-Trend statistical signif. {[}\%{]}
-\end{minipage} \\
-\midrule\noalign{}
-\endhead
-\bottomrule\noalign{}
-\endlastfoot
-99.88 \\
-85.02 \\
-91.29 \\
-99.33 \\
-99.97 \\
-37.53 \\
-20.28 \\
-97 \\
-6.646 \\
-97.06 \\
-99.5 \\
-98.04 \\
-\end{longtable}
+\caption{\label{tab:trendseasontable}\label{tab:trendseasontable}SDR linear trends of monthly anomalies for each season of the year.}
+\centering
+\begin{tabular}[t]{ccrr}
+\toprule
+Sky condition & Season & Trend [\%/year] & Statistical signif. [\%]\\
+\midrule
+\cellcolor{gray!6}{} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{0.6860} & \cellcolor{gray!6}{99.9}\\
 
-\hypertarget{consistency-of-the-trends}{%
-\subsection{Consistency of the trends}\label{consistency-of-the-trends}}
+ & Spring & 0.1450 & 81.8\\
 
-A method to evaluate changes in the long term trend is to use the
-cumulative sum of the variable.
+\cellcolor{gray!6}{} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{0.1200} & \cellcolor{gray!6}{89.4}\\
 
-Figure \ref{fig:cumsum}
+\multirow{-4}{*}{\centering\arraybackslash All sky} & Autumn & 0.4310 & 99.3\\
+\cmidrule{1-4}
+\cellcolor{gray!6}{} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{0.8260} & \cellcolor{gray!6}{100.0}\\
 
-\ldots{} different in aggregation \ldots{}
+ & Spring & 0.0613 & 38.6\\
 
-\begin{figure}[h!]
+\cellcolor{gray!6}{} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{-0.0307} & \cellcolor{gray!6}{25.9}\\
 
-{\centering \subfloat[one plot\label{fig:cumsum-1}]{\includegraphics[width=.35\linewidth]{./images/CumulativeDailyCumSum-1} }\subfloat[the other one\label{fig:cumsum-2}]{\includegraphics[width=.35\linewidth]{./images/CumulativeDailyCumSum-9} }
+\multirow{-4}{*}{\centering\arraybackslash Clear sky} & Autumn & 0.3670 & 97.2\\
+\cmidrule{1-4}
+\cellcolor{gray!6}{} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{-0.8820} & \cellcolor{gray!6}{98.9}\\
 
-}
+ & Spring & -0.0991 & 37.2\\
 
-\caption{two plots}\label{fig:cumsum}
-\end{figure}
+\cellcolor{gray!6}{} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{0.0444} & \cellcolor{gray!6}{21.5}\\
+
+\multirow{-4}{*}{\centering\arraybackslash Cloudy sky} & Autumn & -0.4000 & 89.5\\
+\bottomrule
+\end{tabular}
+\end{table}
 
@@
 
-Our result for all-sky condition (\(0.4\%/year\)), reaffirm the previous
-results of \citet{Bais2013} for the site. The increase of this trend,
-shows that the phenomenon and probably the causes have been amplified.
+We demonstrate that in the period 1993 - 2023, there is a brightening
+trend of SDR for all-sky conditions (\(0.38\,\%/y\)). On a previous
+study \citep{Bais2013}, for the period 1993 - 2011 the trend was found
+to be \(0.33\,\%/y\). The increase of this trend, indicates that the
+phenomenon and probably the causes, are still occurring. Accordingly, we
+found a similar trend for clear-sky conditions (\(0.35\,\%/y\)) that
+further supports that the brightening can be independent of the cloud
+occurrences, although, cloud formation can also attenuate SDR. We have
+calculated the effect of TSI variability to be a negligible factor in
+SDR variability. Unfortunately, there is no available data for this
+period for the aerosols, in order to directly estimate their effect on
+SDR, which is the most probable main cause. The attenuation of SDR by
+aerosols over Europe have been proposed as major factor by
+\citet{Wild2021}. The dimming effect of SDR under cloud-sky condition
+(\(-0.28\,\%/y\)), shows that clouds can also attenuate the amount of
+SDR reaching the ground, and changes on their occurrences can effect the
+net energy balance. Because we have no adequate data to investigate the
+long term changes of cloud formation in the region, we can not infer if
+the SDR trend we observe under cloud-sky is biased by changes in clouds
+and local climate. The observed brightening of SDR over Thessaloniki; is
+dependent on SZA (higher SZA, have stronger brightening trends), and
+also, with the seasons of the year, where Winter and Autumn show a
+significant strong trend, in contrast to Spring and Summer. Using CUSUMs
+of the trends for all- and clear-skies, we observed periods where the
+CUSUMs remain relative steady, with a steep decline previous to that,
+and a steep incline after. This is an indication that the whole
+brightening phenomenon does not follow a smooth development over time.
 
-\ldots.. Also, the trend of \(0.36\%/year\) for cloud-sky condition
-indicate the major part that clouds play on the \ldots..
+Our findings are in agreement with other researchers for the region.
+Future observations from the collocated pyrheliometer, may provide us
+with the means to further investigate the variability of Solar radiation
+at ground level in Thessaloniki. Also, more data (cloud occurrence,
+aerosols, atmospheric water vapour, etc.), will help us to determine the
+extent of SDR attenuation by each factor.
 
-\ldots\ldots.
+\hypertarget{references}{%
+\section*{References}\label{references}}
+\addcontentsline{toc}{section}{References}
 
-About:
+\hypertarget{appendix}{%
+\section*{Appendix}\label{appendix}}
+\addcontentsline{toc}{section}{Appendix}
 
-\begin{itemize}
-\item
-  long terms trends
-\item
-  seasonal trends
-\item
-  effect of SZA
-\item
-  effect of clouds
-\item
-  Aerosols
-\end{itemize}
+\begin{figure}[h!]
 
-similar results with Evidence for Clear‐Sky Dimming and Brightening in
-Central Europe\_Wild2021.pdf inclue in discusion
+{\centering \includegraphics[width=0.7\linewidth]{./images/LongtermTrends-5}
 
-\begin{longtable}[]{@{}c@{}}
-\toprule\noalign{}
-\endhead
-\bottomrule\noalign{}
-\endlastfoot
-\textbf{END} \\
-\end{longtable}
+}
 
+\caption{Anomaly (\%) of the daily clear-sky SDR, relative to climatological values for1993 - 2023. The black line shows the long term linear trend for clear-sky conditions.}\label{fig:trendCLEAR}
+\end{figure}
+
+\begin{figure}[h!]
+
+{\centering \includegraphics[width=0.7\linewidth]{./images/LongtermTrends-8}
+
+}
+
+\caption{Anomaly (\%) of the daily cloud-sky SDR, relative to climatological values for 1993 - 2023. The black line shows the long term linear trend for cloud-sky conditions.}\label{fig:trendCLOUD}
+\end{figure}
+
+\begin{figure}[h!]
+
+{\centering \includegraphics[width=1\linewidth]{./images/SzaTrendsSeasTogether-2}
+
+}
+
+\caption{Seasonla distribution of the SDR's long term trends by SZA before and after noon.
+The solid data points represent values with an acceptable statistical significance ($p<0.005$), the trends were computed with SZA bins of $1^\circ$ for each day, before and after local noon, all graphs have the same scale.
+Bins with $p<0.005$ or with less than 85 data points may be missing from the view.}\label{fig:SZAtrendSeason}
+\end{figure}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\vspace{6pt}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% optional
+
+% Only for the journal Methods and Protocols:
+% If you wish to submit a video article, please do so with any other supplementary material.
+% \supplementary{The following supporting information can be downloaded at: \linksupplementary{s1}, Figure S1: title; Table S1: title; Video S1: title.
+A supporting video article is available at doi: link.}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Optional
+
+%% Only for journal Encyclopedia
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Optional
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{adjustwidth}{-\extralength}{0cm}
+
+%\printendnotes[custom] % Un-comment to print a list of endnotes
+
+\reftitle{References}
 \bibliography{Mainreferences.bib}
 
+% If authors have biography, please use the format below
+%\section*{Short Biography of Authors}
+%\bio
+%{\raisebox{-0.35cm}{\includegraphics[width=3.5cm,height=5.3cm,clip,keepaspectratio]{Definitions/author1.pdf}}}
+%{\textbf{Firstname Lastname} Biography of first author}
+%
+%\bio
+%{\raisebox{-0.35cm}{\includegraphics[width=3.5cm,height=5.3cm,clip,keepaspectratio]{Definitions/author2.jpg}}}
+%{\textbf{Firstname Lastname} Biography of second author}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% for journal Sci
+%\reviewreports{\\
+%Reviewer 1 comments and authors’ response\\
+%Reviewer 2 comments and authors’ response\\
+%Reviewer 3 comments and authors’ response
+%}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\PublishersNote{}
+\end{adjustwidth}
+
 \end{document}

### 2023-11-01 MDPI_submition/MDPI_submition.Rmd
Commit 2023-11-01 16:14

---
title: Full title of the paper (Capitalized)
author:
  - name:  Athanasios Natsis
    affil: 1,2,* # \ddagger
    orcid: 0000-0002-5199-4119
  - name: Alkiviadis Bais
    affil: 2, \dagger, # \ddagger
affiliation:
  - num: 1
    address: |
      Aristotle University of Thessaloniki - 
      Laboratory of Atmospheric Physics
      Campus Box 149, 54124 Thessaloniki, Greece
    email: natsisphysicist@gmail.com
  - num: 2
    address: |
      Your department
      Street, City, Country
    email: abais@auth.gr
# author citation list in chicago format
authorcitation: |
  Natsis, A.; Bais, A.
# firstnote to eighthnote
firstnote: |
  Current address: Updated affiliation
# secondnote: |
#   These authors contributed equally to this work.
correspondence: |
  leutnant@fh-muenster.de; Tel.: +XX-000-00-0000.
# document options
journal: notspecified
type:    article
status:  submit
# front matter
simplesummary: |
  A Simple summary goes here.
abstract: |
  A single paragraph of about 200 words maximum. For research articles, 
  abstracts should give a pertinent overview of the work. We strongly encourage
  authors to use the following style of structured abstracts, but without 
  headings: 1) Background: Place the question addressed in a broad context and
  highlight the purpose of the study; 2) Methods: Describe briefly the main
  methods or treatments applied; 3) Results: Summarize the article's main 
  findings; and 4) Conclusion: Indicate the main conclusions or interpretations. 
  The abstract should be an objective representation of the article, it must not 
  contain results which are not presented and substantiated in the main text and 
  should not exaggerate the main conclusions.
# back matter
keywords: |
  keyword 1; keyword 2; keyword 3 (list three to ten pertinent keywords specific 
  to the article, yet reasonably common within the subject discipline.).
acknowledgement: |
  All sources of funding of the study should be disclosed. Please clearly 
  indicate grants that you have received in support of your research work. 
  Clearly state if you received funds for covering the costs to publish in open 
  access.
authorcontributions: |
  For research articles with several authors, a short paragraph specifying their 
  individual contributions must be provided. The following statements should be 
  used ``X.X. and Y.Y. conceive and designed the experiments; X.X. performed the 
  experiments; X.X. and Y.Y. analyzed the data; W.W. contributed 
  reagents/materials/analysis tools; Y.Y. wrote the paper.'' Authorship must be
  limited to those who have contributed substantially to the work reported.
funding: |
  Please add: ``This research received no external funding'' or ``This research 
  was funded by NAME OF FUNDER grant number XXX.'' and  and ``The APC was funded 
  by XXX''. Check carefully that the details given are accurate and use the 
  standard spelling of funding agency names at 
  \url{https://search.crossref.org/funding}, any errors may affect your future 
  funding.
institutionalreview: |
  In this section, you should add the Institutional Review Board Statement and 
  approval number, if relevant to your study. You might choose to exclude 
  this statement if the study did not require ethical approval. Please note 
  that the Editorial Office might ask you for further information. Please add 
  “The study was conducted in accordance with the Declaration of Helsinki, 
  and approved by the Institutional Review Board (or Ethics Committee) of 
  NAME OF INSTITUTE (protocol code XXX and date of approval).” for studies 
  involving humans. OR “The animal study protocol was approved by the 
  Institutional Review Board (or Ethics Committee) of NAME OF INSTITUTE 
  (protocol code XXX and date of approval).” for studies involving animals. 
  OR “Ethical review and approval were waived for this study due to REASON 
  (please provide a detailed justification).” OR “Not applicable” for
   studies not involving humans or animals.
informedconsent: |
  Any research article describing a study involving humans should contain this 
  statement. Please add ``Informed consent was obtained from all subjects 
  involved in the study.'' OR ``Patient consent was waived due to REASON 
  (please provide a detailed justification).'' OR ``Not applicable'' for 
  studies not involving humans. You might also choose to exclude this statement 
  if the study did not involve humans.

  Written informed consent for publication must be obtained from participating 
  patients who can be identified (including by the patients themselves). Please 
  state ``Written informed consent has been obtained from the patient(s) to 
  publish this paper'' if applicable.
dataavailability: |
  We encourage all authors of articles published in MDPI journals to share 
  their research data. In this section, please provide details regarding where 
  data supporting reported results can be found, including links to publicly 
  archived datasets analyzed or generated during the study. Where no new data 
  were created, or where data is unavailable due to privacy or ethical 
  re-strictions, a statement is still required. Suggested Data Availability 
  Statements are available in section “MDPI Research Data Policies” at 
  \url{https://www.mdpi.com/ethics}.
conflictsofinterest: |
  Declare conflicts of interest or state 'The authors declare no conflict of 
  interest.' Authors must identify and declare any personal circumstances or
  interest that may be perceived as inappropriately influencing the
  representation or interpretation of reported research results. Any role of the
  funding sponsors in the design of the study; in the collection, analyses or 
  interpretation of data in the writing of the manuscript, or in the decision to 
  publish the results must be declared in this section. If there is no role, 
  please state 'The founding sponsors had no role in the design of the study; 
  in the collection, analyses, or interpretation of data; in the writing of the 
  manuscript, an in the decision to publish the results'.
sampleavailability: |
  Samples of the compounds ...... are available from the authors.
supplementary: |
 The following supporting information can be downloaded at:  
 \linksupplementary{s1}, Figure S1: title; Table S1: title; Video S1: title.
abbreviations:
  - short: MDPI
    long: Multidisciplinary Digital Publishing Institute
  - short: DOAJ
    long: Directory of open access journals
  - short: TLA
    long: Three letter acronym
  - short: LD 
    long: linear dichroism
bibliography: mybibfile.bib
appendix: appendix.tex
endnotes: false
output: 
  rticles::mdpi_article:
    extra_dependencies: longtable
---

# Version

This Rmd-skeleton uses the mdpi Latex template published 2023-03-25. 
However, the official template gets more frequently updated than the **rticles**
package. Therefore, please make sure prior to paper submission, that you're 
using the most recent .cls, .tex and .bst files 
(available [here](http://www.mdpi.com/authors/latex)).

# Article Header Information

The YAML header includes information needed mainly for formatting the front and 
back matter of the article. Required elements include:

```yaml
title: Title of the paper
author:
  - name: first and last name
    affil: |
      One or more comma seperated numbers corresponding to affilitation
      and one or more  comma seperated symbols corresponding 
      optional notes.
    orcid: optional orcid number
affiliation:  
  - num: 1,..., n for each affiliation
    address: required
    email: required
authorcitation: |
  Lastname, F.
correspondence: |
  email@email.com; Tel.: +XX-000-00-0000.
journal: notspecified
type: article
status: submit
```

Journal options are in Table \ref{tab:mdpinames}. The `status` variable should 
generally not be changed by authors. The `type` variable describes 
the type of of submission and defaults to `article` but can be replaced with any of the ones in Table \ref{tab:mdpitype}

```{r mdpitype, echo = FALSE}
type <- c("abstract, addendum, article, book, bookreview, briefreport, 
casereport, comment, commentary, communication, conferenceproceedings, 
correction, conferencereport, entry, expressionofconcern, extendedabstract, 
datadescriptor, editorial, essay, erratum, hypothesis, interestingimage, 
obituary, opinion, projectreport, reply, retraction, review, perspective, 
protocol, shortnote, studyprotocol, systematicreview, supfile, 
technicalnote, viewpoint, guidelines, registeredreport, tutorial")
knitr::kable(rticles::string_to_table(type, 3), 
             col.names = NULL, format = "latex", booktab = TRUE, 
             caption = "MDPI article types.")
```

## Journal Specific YAML variables

```yaml
# for journal Diversity,
# add the Life Science Identifier using:
lsid: http://zoobank.org/urn:lsid:zoobank.org:act:nnnn

# for journal Applied Sciences
# add featured application
featuredapplication: |
  Authors are encouraged to provide a concise 
  description of the specific application or 
  a potential application of the work. This 
  section is not mandatory.

# for the journal Data
# add dataset doi and license
dataset: https://doi.org/10.1000/182
datasetlicense: CC-BY-4.0

# for the journal Toxins
# add key contributions
keycontributions: |
  The breakthroughs or highlights of the manuscript. 
  Authors can write one or two sentences to describe 
  the most important part of the paper.

# for the journal Encyclopedia
encyclopediadef: |
  For entry manuscripts only: please provide a brief overview
  of the entry title instead of an abstract.
entrylink: The Link to this entry published on the encyclopedia platform.

# for the journal Advances in Respiratory Medicine
# add highlights
addhighlights: |
  This is an obligatory section in “Advances in Respiratory Medicine”, 
  whose goal is to increase the discoverability and readability of the
  article via search engines and other scholars. Highlights should not 
  be a copy of the abstract, but a simple text allowing the reader to 
  quickly and simplified find out what the article is about and what can 
  be cited from it. Each of these parts should be devoted up to 2~bullet 
  points.

```

\startlandscape
```{r, mdpinames, echo=FALSE, results='asis'}
mdpi_journals <- "acoustics, actuators, addictions, admsci, adolescents, aerobiology, aerospace, agriculture, agriengineering, agrochemicals, agronomy, ai, air, algorithms, allergies, alloys, analytica, analytics, anatomia, animals, antibiotics, antibodies, antioxidants, applbiosci, appliedchem, appliedmath, applmech, applmicrobiol, applnano, applsci, aquacj, architecture, arm, arthropoda, arts, asc, asi, astronomy, atmosphere, atoms, audiolres, automation, axioms, bacteria, batteries, bdcc, behavsci, beverages, biochem, bioengineering, biologics, biology, biomass, biomechanics, biomed, biomedicines, biomedinformatics, biomimetics, biomolecules, biophysica, biosensors, biotech, birds, bloods, blsf, brainsci, breath, buildings, businesses, cancers, carbon, cardiogenetics, catalysts, cells, ceramics, challenges, chemengineering, chemistry, chemosensors, chemproc, children, chips, cimb, civileng, cleantechnol, climate, clinpract, clockssleep, cmd, coasts, coatings, colloids, colorants, commodities, compounds, computation, computers, condensedmatter, conservation, constrmater, cosmetics, covid, crops, cryptography, crystals, csmf, ctn, curroncol, cyber, dairy, data, ddc, dentistry, dermato, dermatopathology, designs, devices, diabetology, diagnostics, dietetics, digital, disabilities, diseases, diversity, dna, drones, dynamics, earth, ebj, ecologies, econometrics, economies, education, ejihpe, electricity, electrochem, electronicmat, electronics, encyclopedia, endocrines, energies, eng, engproc, entomology, entropy, environments, environsciproc, epidemiologia, epigenomes, est, fermentation, fibers, fintech, fire, fishes, fluids, foods, forecasting, forensicsci, forests, foundations, fractalfract, fuels, future, futureinternet, futurepharmacol, futurephys, futuretransp, galaxies, games, gases, gastroent, gastrointestdisord, gels, genealogy, genes, geographies, geohazards, geomatics, geosciences, geotechnics, geriatrics, grasses, gucdd, hazardousmatters, healthcare, hearts, hemato, hematolrep, heritage, higheredu, highthroughput, histories, horticulturae, hospitals, humanities, humans, hydrobiology, hydrogen, hydrology, hygiene, idr, ijerph, ijfs, ijgi, ijms, ijns, ijpb, ijtm, ijtpp, ime, immuno, informatics, information, infrastructures, inorganics, insects, instruments, inventions, iot, j, jal, jcdd, jcm, jcp, jcs, jcto, jdb, jeta, jfb, jfmk, jimaging, jintelligence, jlpea, jmmp, jmp, jmse, jne, jnt, jof, joitmc, jor, journalmedia, jox, jpm, jrfm, jsan, jtaer, jvd, jzbg, kidneydial, kinasesphosphatases, knowledge, land, languages, laws, life, liquids, literature, livers, logics, logistics, lubricants, lymphatics, machines, macromol, magnetism, magnetochemistry, make, marinedrugs, materials, materproc, mathematics, mca, measurements, medicina, medicines, medsci, membranes, merits, metabolites, metals, meteorology, methane, metrology, micro, microarrays, microbiolres, micromachines, microorganisms, microplastics, minerals, mining, modelling, molbank, molecules, mps, msf, mti, muscles, nanoenergyadv, nanomanufacturing, nanomaterials, ncrna, ndt, network, neuroglia, neurolint, neurosci, nitrogen, notspecified, nri, nursrep, nutraceuticals, nutrients, obesities, oceans, ohbm, onco, oncopathology, optics, oral, organics, organoids, osteology, oxygen, parasites, parasitologia, particles, pathogens, pathophysiology, pediatrrep, pharmaceuticals, pharmaceutics, pharmacoepidemiology, pharmacy, philosophies, photochem, photonics, phycology, physchem, physics, physiologia, plants, plasma, platforms, pollutants, polymers, polysaccharides, poultry, powders, preprints, proceedings, processes, prosthesis, proteomes, psf, psych, psychiatryint, psychoactives, publications, quantumrep, quaternary, qubs, radiation, reactions, receptors, recycling, regeneration, religions, remotesensing, reports, reprodmed, resources, rheumato, risks, robotics, ruminants, safety, sci, scipharm, sclerosis, seeds, sensors, separations, sexes, signals, sinusitis, skins, smartcities, sna, societies, socsci, software, soilsystems, solar, solids, spectroscj, sports, standards, stats, std, stresses, surfaces, surgeries, suschem, sustainability, symmetry, synbio, systems, targets, taxonomy, technologies, telecom, test, textiles, thalassrep, thermo, tomography, tourismhosp, toxics, toxins, transplantology, transportation, traumacare, traumas, tropicalmed, universe, urbansci, uro, vaccines, vehicles, venereology, vetsci, vibration, virtualworlds, viruses, vision, waste, water, wem, wevj, wind, women, world, youth, zoonoticdis"

knitr::kable(rticles::string_to_table(mdpi_journals, 8),
             col.names = NULL,
             format = "latex",
             booktabs = TRUE, 
             caption = "MDPI journal names.",
             longtable = TRUE)

```
\finishlandscape

# Introduction

The introduction should briefly place the study in a broad context and highlight
why it is important. It should define the purpose of the work and its
significance. The current state of the research field should be reviewed
carefully and key publications cited. Please highlight controversial and
diverging hypotheses when necessary. Finally, briefly mention the main aim of
the work and highlight the principal conclusions. As far as possible, please
keep the introduction comprehensible to scientists outside your particular
field of research. Citing a journal paper [@bertrand-krajewski_distribution_1998;
@leutnant_stormwater_2016]. And now citing a book reference @gujer_systems_2008.
Some MDPI journals use Chicago and others use APA, this template should choose
the correct citation format for you once you specify the journal in the YAML
header.

To use endnotes, change `endnotes: true` in the YAML header, then use
`\endnote{This is an endnote.}`.

# Materials and Methods

Materials and Methods should be described with sufficient details to allow 
others to replicate and build on published results. Please note that publication 
of your manuscript implicates that you must make all materials, data, computer 
code, and protocols associated with the publication available to readers. Please
disclose at the submission stage any restrictions on the availability of 
materials or information. New methods and protocols should be described in 
detail while well-established methods can be briefly described and appropriately
cited.

Research manuscripts reporting large datasets that are deposited in a publicly 
available database should specify where the data have been deposited and provide
the relevant accession numbers. If the accession numbers have not yet been
obtained at the time of submission, please state that they will be provided
during review. They must be provided prior to publication.

Interventionary studies involving animals or humans, and other studies require
ethical approval must list the authority that provided approval and the
corresponding ethical approval code.

# Results

This section may be divided by subheadings. It should provide a concise and
precise description of the experimental results, their interpretation as well
as the experimental conclusions that can be drawn.

## Subsection Heading Here

Subsection text here.

### Subsubsection Heading Here

Bulleted lists look like this:

* First bullet
* Second bullet
* Third bullet

Numbered lists can be added as follows:

1. First item
2. Second item
3. Third item

The text continues here.

## Figures, Tables and Schemes

All figures and tables should be cited in the main text as Figure \ref{fig:fig1}, 
\ref{tab:tab1}, etc. To get cross-reference to figure generated by R chunks 
include the `\\label{}` tag in the `fig.cap` attribute of the R chunk: 
`fig.cap = "Fancy Caption\\label{fig:plot}"`.

```{r fig1, echo=FALSE, fig.width=3, fig.cap="A figure added with a code chunk.\\label{fig:fig1}"}
x = rnorm(10)
y = rnorm(10)
plot(x, y)
```

When making tables using `kable`, it is suggested to use
the `format="latex"` and `tabl.envir="table"` arguments
to ensure table numbering and compatibility with the mdpi
document class.

```{r tab1, echo=FALSE}
knitr::kable(mtcars[1:5, 1:3], format = "latex", 
             booktabs = TRUE, 
             caption = "This is a table caption. Tables should be placed in the 
             main text near to the first time they are~cited.", 
             align = 'ccc', centering = FALSE,
             table.envir = "table", position = "H")
```

For a very wide table, landscape layouts are allowed.

\startlandscape

```{r tab2, echo=FALSE}
df <- data.frame(`Title 1` = c("Entry 1", "Entry 2"),
                 `Title 2` = c("Data", "Data"),
                 `Title 3` = c("Data", "Data"),
                 `Title 4` = c("This cell has some longer content that runs over
                               two lines",
                               "Data"))
knitr::kable(df, format = "latex", 
             booktabs = TRUE, 
             caption = "This is a very wide table", 
             align = 'ccc', centering = FALSE,
             table.envir = "table", position = "H")
```

\finishlandscape

## Formatting of Mathematical Components

This is an example of an equation:

$$
a = 1.
$$

If you want numbered equations use Latex and wrap in the equation environment:

\begin{equation}
a = 1,
\end{equation}

the text following an equation need not be a new paragraph. Please punctuate 
equations as regular text.

```{comment}
If the documentclass option "submit" is chosen, please insert a blank line before and after any math environment (equation and eqnarray environments). This ensures correct linenumbering. The blank line should be removed when the documentclass option is changed to "accept" because the text following an equation should not be a new paragraph.
```

This is the example 2 of equation:

\begin{adjustwidth}{-\extralength}{0cm}
\begin{equation}
a = b + c + d + e + f + g + h + i + j + k + l + m + n + o + p + q + r + s + t + 
u + v + w + x + y + z
\end{equation}
\end{adjustwidth}

Theorem-type environments (including propositions, lemmas, corollaries etc.) 
can be formatted as follows:

Example of a theorem:

::: {.Theorem latex=true}
Example text of a theorem
:::

The text continues here. Proofs must be formatted as follows:

Example of a proof:

::: {.proof latex="[Proof of Theorem1]"}
Text of the proof. Note that the phrase ``of Theorem 1'' is optional if it is 
clear which theorem is being referred to.
:::

The text continues here.

# Discussion

Authors should discuss the results and how they can be interpreted in 
perspective of previous studies and of the working hypotheses. The findings and 
their implications should be discussed in the broadest context possible. Future 
research directions may also be highlighted.

# Conclusion

This section is not mandatory, but can be added to the manuscript if the
discussion is unusually long or complex.

# Patents

This section is not mandatory, but may be added if there are patents resulting
from the work reported in this manuscript.

### 2023-11-01 MDPI_submition/MDPI_submition.tex
Commit 2023-11-01 16:14

%  LaTeX support: latex@mdpi.com
%  For support, please attach all files needed for compiling as well as the log file, and specify your operating system, LaTeX version, and LaTeX editor.

%=================================================================
% pandoc conditionals added to preserve backwards compatibility with previous versions of rticles

\documentclass[notspecified,article,submit,moreauthors,pdftex]{Definitions/mdpi}

%% Some pieces required from the pandoc template
\setlist[itemize]{leftmargin=*,labelsep=5.8mm}
\setlist[enumerate]{leftmargin=*,labelsep=4.9mm}

%--------------------
% Class Options:
%--------------------

%---------
% article
%---------
% The default type of manuscript is "article", but can be replaced by:
% abstract, addendum, article, book, bookreview, briefreport, casereport, comment, commentary, communication, conferenceproceedings, correction, conferencereport, entry, expressionofconcern, extendedabstract, datadescriptor, editorial, essay, erratum, hypothesis, interestingimage, obituary, opinion, projectreport, reply, retraction, review, perspective, protocol, shortnote, studyprotocol, systematicreview, supfile, technicalnote, viewpoint, guidelines, registeredreport, tutorial
% supfile = supplementary materials

%----------
% submit
%----------
% The class option "submit" will be changed to "accept" by the Editorial Office when the paper is accepted. This will only make changes to the frontpage (e.g., the logo of the journal will get visible), the headings, and the copyright information. Also, line numbering will be removed. Journal info and pagination for accepted papers will also be assigned by the Editorial Office.

%------------------
% moreauthors
%------------------
% If there is only one author the class option oneauthor should be used. Otherwise use the class option moreauthors.

%---------
% pdftex
%---------
% The option pdftex is for use with pdfLaTeX. Remove "pdftex" for (1) compiling with LaTeX & dvi2pdf (if eps figures are used) or for (2) compiling with XeLaTeX.

%=================================================================
% MDPI internal commands - do not modify
\firstpage{1}
\makeatletter
\setcounter{page}{\@firstpage}
\makeatother
\pubvolume{1}
\issuenum{1}
\articlenumber{0}
\pubyear{2023}
\copyrightyear{2023}
%\externaleditor{Academic Editor: Firstname Lastname}
\datereceived{ }
\daterevised{ } % Comment out if no revised date
\dateaccepted{ }
\datepublished{ }
%\datecorrected{} % For corrected papers: "Corrected: XXX" date in the original paper.
%\dateretracted{} % For corrected papers: "Retracted: XXX" date in the original paper.
\hreflink{https://doi.org/} % If needed use \linebreak
%\doinum{}
%\pdfoutput=1 % Uncommented for upload to arXiv.org

%=================================================================
% Add packages and commands here. The following packages are loaded in our class file: fontenc, inputenc, calc, indentfirst, fancyhdr, graphicx, epstopdf, lastpage, ifthen, float, amsmath, amssymb, lineno, setspace, enumitem, mathpazo, booktabs, titlesec, etoolbox, tabto, xcolor, colortbl, soul, multirow, microtype, tikz, totcount, changepage, attrib, upgreek, array, tabularx, pbox, ragged2e, tocloft, marginnote, marginfix, enotez, amsthm, natbib, hyperref, cleveref, scrextend, url, geometry, newfloat, caption, draftwatermark, seqsplit
% cleveref: load \crefname definitions after \begin{document}

%=================================================================
% Please use the following mathematics environments: Theorem, Lemma, Corollary, Proposition, Characterization, Property, Problem, Example, ExamplesandDefinitions, Hypothesis, Remark, Definition, Notation, Assumption
%% For proofs, please use the proof environment (the amsthm package is loaded by the MDPI class).

%=================================================================
% Full title of the paper (Capitalized)
\Title{Full title of the paper (Capitalized)}

% MDPI internal command: Title for citation in the left column
\TitleCitation{Full title of the paper (Capitalized)}

% Author Orchid ID: enter ID or remove command
%\newcommand{\orcidauthorA}{0000-0000-0000-000X} % Add \orcidA{} behind the author's name
%\newcommand{\orcidauthorB}{0000-0000-0000-000X} % Add \orcidB{} behind the author's name

% Authors, for the paper (add full first names)
\Author{Athanasios
Natsis$^{1,2,\ddagger,*}$\href{https://orcid.org/0000-0002-5199-4119}
{\orcidicon}, Alkiviadis Bais$^{2, \dagger,}$}

%\longauthorlist{yes}

% MDPI internal command: Authors, for metadata in PDF
\AuthorNames{Athanasios Natsis, Alkiviadis Bais}

% MDPI internal command: Authors, for citation in the left column
%\AuthorCitation{Lastname, F.; Lastname, F.; Lastname, F.}
% If this is a Chicago style journal: Lastname, Firstname, Firstname Lastname, and Firstname Lastname.
\AuthorCitation{Natsis, A.; Bais, A.}

% Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
\address{%
$^{1}$ \quad Aristotle University of Thessaloniki - Laboratory of
Atmospheric Physics Campus Box 149, 54124 Thessaloniki,
Greece; \href{mailto:natsisphysicist@gmail.com}{\nolinkurl{natsisphysicist@gmail.com}}\\
$^{2}$ \quad Your department Street, City,
Country; \href{mailto:abais@auth.gr}{\nolinkurl{abais@auth.gr}}\\
}

% Contact information of the corresponding author
\corres{Correspondence: \href{mailto:leutnant@fh-muenster.de}{\nolinkurl{leutnant@fh-muenster.de}};
Tel.: +XX-000-00-0000.}

% Current address and/or shared authorship
\firstnote{Current address: Updated affiliation}

% The commands \thirdnote{} till \eighthnote{} are available for further notes

% Simple summary
\simplesumm{A Simple summary goes here.}

%\conference{} % An extended version of a conference paper

% Abstract (Do not insert blank lines, i.e. \\)
\abstract{A single paragraph of about 200 words maximum. For research
articles, abstracts should give a pertinent overview of the work. We
strongly encourage authors to use the following style of structured
abstracts, but without headings: 1) Background: Place the question
addressed in a broad context and highlight the purpose of the study; 2)
Methods: Describe briefly the main methods or treatments applied; 3)
Results: Summarize the article's main findings; and 4) Conclusion:
Indicate the main conclusions or interpretations. The abstract should be
an objective representation of the article, it must not contain results
which are not presented and substantiated in the main text and should
not exaggerate the main conclusions.}

% Keywords
\keyword{keyword 1; keyword 2; keyword 3 (list three to ten pertinent
keywords specific to the article, yet reasonably common within the
subject discipline.).}

% The fields PACS, MSC, and JEL may be left empty or commented out if not applicable
%\PACS{J0101}
%\MSC{}
%\JEL{}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Diversity
%\LSID{\url{http://}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Applied Sciences

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Data

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Toxins

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Encyclopedia

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Advances in Respiratory Medicine
%\addhighlights{yes}
%\renewcommand{\addhighlights}{%

%\noindent This is an obligatory section in “Advances in Respiratory Medicine”, whose goal is to increase the discoverability and readability of the article via search engines and other scholars. Highlights should not be a copy of the abstract, but a simple text allowing the reader to quickly and simplified find out what the article is about and what can be cited from it. Each of these parts should be devoted up to 2~bullet points.\vspace{3pt}\\
%\textbf{What are the main findings?}
% \begin{itemize}[labelsep=2.5mm,topsep=-3pt]
% \item First bullet.
% \item Second bullet.
% \end{itemize}\vspace{3pt}
%\textbf{What is the implication of the main finding?}
% \begin{itemize}[labelsep=2.5mm,topsep=-3pt]
% \item First bullet.
% \item Second bullet.
% \end{itemize}
%}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Pandoc syntax highlighting
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\BuiltInTok}[1]{#1}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
\newcommand{\ExtensionTok}[1]{#1}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\ImportTok}[1]{#1}
\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
\newcommand{\NormalTok}[1]{#1}
\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
\newcommand{\RegionMarkerTok}[1]{#1}
\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}

% tightlist command for lists without linebreak
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\usepackage{longtable}

\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\hypertarget{version}{%
\section{Version}\label{version}}

This Rmd-skeleton uses the mdpi Latex template published 2023-03-25.
However, the official template gets more frequently updated than the
\textbf{rticles} package. Therefore, please make sure prior to paper
submission, that you're using the most recent .cls, .tex and .bst files
(available \href{http://www.mdpi.com/authors/latex}{here}).

\hypertarget{article-header-information}{%
\section{Article Header Information}\label{article-header-information}}

The YAML header includes information needed mainly for formatting the
front and back matter of the article. Required elements include:

\begin{Shaded}
\begin{Highlighting}[]
\FunctionTok{title}\KeywordTok{:}\AttributeTok{ Title of the paper}
\FunctionTok{author}\KeywordTok{:}
\AttributeTok{  }\KeywordTok{{-}}\AttributeTok{ }\FunctionTok{name}\KeywordTok{:}\AttributeTok{ first and last name}
\FunctionTok{    affil}\KeywordTok{: }\CharTok{|}
\NormalTok{      One or more comma seperated numbers corresponding to affilitation}
\NormalTok{      and one or more  comma seperated symbols corresponding }
\NormalTok{      optional notes.}
\AttributeTok{    }\FunctionTok{orcid}\KeywordTok{:}\AttributeTok{ optional orcid number}
\FunctionTok{affiliation}\KeywordTok{:}\AttributeTok{  }
\AttributeTok{  }\KeywordTok{{-}}\AttributeTok{ }\FunctionTok{num}\KeywordTok{:}\AttributeTok{ 1,..., n for each affiliation}
\AttributeTok{    }\FunctionTok{address}\KeywordTok{:}\AttributeTok{ required}
\AttributeTok{    }\FunctionTok{email}\KeywordTok{:}\AttributeTok{ required}
\FunctionTok{authorcitation}\KeywordTok{: }\CharTok{|}
\NormalTok{  Lastname, F.}
\FunctionTok{correspondence}\KeywordTok{: }\CharTok{|}
\NormalTok{  email@email.com; Tel.: +XX{-}000{-}00{-}0000.}
\FunctionTok{journal}\KeywordTok{:}\AttributeTok{ notspecified}
\FunctionTok{type}\KeywordTok{:}\AttributeTok{ article}
\FunctionTok{status}\KeywordTok{:}\AttributeTok{ submit}
\end{Highlighting}
\end{Shaded}

Journal options are in Table \ref{tab:mdpinames}. The \texttt{status}
variable should generally not be changed by authors. The \texttt{type}
variable describes the type of of submission and defaults to
\texttt{article} but can be replaced with any of the ones in Table
\ref{tab:mdpitype}

\begin{table}

\caption{\label{tab:mdpitype}MDPI article types.}
\centering
\begin{tabular}[t]{lll}
\toprule
abstract & entry & retraction\\
addendum & expressionofconcern & review\\
article & extendedabstract & perspective\\
book & datadescriptor & protocol\\
bookreview & editorial & shortnote\\
\addlinespace
briefreport & essay & studyprotocol\\
casereport & erratum & systematicreview\\
comment & hypothesis & supfile\\
commentary & interestingimage & technicalnote\\
communication & obituary & viewpoint\\
\addlinespace
conferenceproceedings & opinion & guidelines\\
correction & projectreport & registeredreport\\
conferencereport & reply & tutorial\\
\bottomrule
\end{tabular}
\end{table}

\hypertarget{journal-specific-yaml-variables}{%
\subsection{Journal Specific YAML
variables}\label{journal-specific-yaml-variables}}

\begin{Shaded}
\begin{Highlighting}[]
\CommentTok{\# for journal Diversity,}
\CommentTok{\# add the Life Science Identifier using:}
\FunctionTok{lsid}\KeywordTok{:}\AttributeTok{ http://zoobank.org/urn:lsid:zoobank.org:act:nnnn}

\CommentTok{\# for journal Applied Sciences}
\CommentTok{\# add featured application}
\FunctionTok{featuredapplication}\KeywordTok{: }\CharTok{|}
\NormalTok{  Authors are encouraged to provide a concise }
\NormalTok{  description of the specific application or }
\NormalTok{  a potential application of the work. This }
\NormalTok{  section is not mandatory.}

\CommentTok{\# for the journal Data}
\CommentTok{\# add dataset doi and license}
\FunctionTok{dataset}\KeywordTok{:}\AttributeTok{ https://doi.org/10.1000/182}
\FunctionTok{datasetlicense}\KeywordTok{:}\AttributeTok{ CC{-}BY{-}4.0}

\CommentTok{\# for the journal Toxins}
\CommentTok{\# add key contributions}
\FunctionTok{keycontributions}\KeywordTok{: }\CharTok{|}
\NormalTok{  The breakthroughs or highlights of the manuscript. }
\NormalTok{  Authors can write one or two sentences to describe }
\NormalTok{  the most important part of the paper.}

\CommentTok{\# for the journal Encyclopedia}
\FunctionTok{encyclopediadef}\KeywordTok{: }\CharTok{|}
\NormalTok{  For entry manuscripts only: please provide a brief overview}
\NormalTok{  of the entry title instead of an abstract.}
\FunctionTok{entrylink}\KeywordTok{:}\AttributeTok{ The Link to this entry published on the encyclopedia platform.}

\CommentTok{\# for the journal Advances in Respiratory Medicine}
\CommentTok{\# add highlights}
\FunctionTok{addhighlights}\KeywordTok{: }\CharTok{|}
\NormalTok{  This is an obligatory section in “Advances in Respiratory Medicine”, }
\NormalTok{  whose goal is to increase the discoverability and readability of the}
\NormalTok{  article via search engines and other scholars. Highlights should not }
\NormalTok{  be a copy of the abstract, but a simple text allowing the reader to }
\NormalTok{  quickly and simplified find out what the article is about and what can }
\NormalTok{  be cited from it. Each of these parts should be devoted up to 2\textasciitilde{}bullet }
\NormalTok{  points.}
\end{Highlighting}
\end{Shaded}

\startlandscape

\begin{longtable}[t]{llllllll}
\caption{\label{tab:mdpinames}MDPI journal names.}\\
\toprule
acoustics & biomedinformatics & dentistry & galaxies & jcp & metabolites & philosophies & socsci\\
actuators & biomimetics & dermato & games & jcs & metals & photochem & software\\
addictions & biomolecules & dermatopathology & gases & jcto & meteorology & photonics & soilsystems\\
admsci & biophysica & designs & gastroent & jdb & methane & phycology & solar\\
adolescents & biosensors & devices & gastrointestdisord & jeta & metrology & physchem & solids\\
\addlinespace
aerobiology & biotech & diabetology & gels & jfb & micro & physics & spectroscj\\
aerospace & birds & diagnostics & genealogy & jfmk & microarrays & physiologia & sports\\
agriculture & bloods & dietetics & genes & jimaging & microbiolres & plants & standards\\
agriengineering & blsf & digital & geographies & jintelligence & micromachines & plasma & stats\\
agrochemicals & brainsci & disabilities & geohazards & jlpea & microorganisms & platforms & std\\
\addlinespace
agronomy & breath & diseases & geomatics & jmmp & microplastics & pollutants & stresses\\
ai & buildings & diversity & geosciences & jmp & minerals & polymers & surfaces\\
air & businesses & dna & geotechnics & jmse & mining & polysaccharides & surgeries\\
algorithms & cancers & drones & geriatrics & jne & modelling & poultry & suschem\\
allergies & carbon & dynamics & grasses & jnt & molbank & powders & sustainability\\
\addlinespace
alloys & cardiogenetics & earth & gucdd & jof & molecules & preprints & symmetry\\
analytica & catalysts & ebj & hazardousmatters & joitmc & mps & proceedings & synbio\\
analytics & cells & ecologies & healthcare & jor & msf & processes & systems\\
anatomia & ceramics & econometrics & hearts & journalmedia & mti & prosthesis & targets\\
animals & challenges & economies & hemato & jox & muscles & proteomes & taxonomy\\
\addlinespace
antibiotics & chemengineering & education & hematolrep & jpm & nanoenergyadv & psf & technologies\\
antibodies & chemistry & ejihpe & heritage & jrfm & nanomanufacturing & psych & telecom\\
antioxidants & chemosensors & electricity & higheredu & jsan & nanomaterials & psychiatryint & test\\
applbiosci & chemproc & electrochem & highthroughput & jtaer & ncrna & psychoactives & textiles\\
appliedchem & children & electronicmat & histories & jvd & ndt & publications & thalassrep\\
\addlinespace
appliedmath & chips & electronics & horticulturae & jzbg & network & quantumrep & thermo\\
applmech & cimb & encyclopedia & hospitals & kidneydial & neuroglia & quaternary & tomography\\
applmicrobiol & civileng & endocrines & humanities & kinasesphosphatases & neurolint & qubs & tourismhosp\\
applnano & cleantechnol & energies & humans & knowledge & neurosci & radiation & toxics\\
applsci & climate & eng & hydrobiology & land & nitrogen & reactions & toxins\\
\addlinespace
aquacj & clinpract & engproc & hydrogen & languages & notspecified & receptors & transplantology\\
architecture & clockssleep & entomology & hydrology & laws & nri & recycling & transportation\\
arm & cmd & entropy & hygiene & life & nursrep & regeneration & traumacare\\
arthropoda & coasts & environments & idr & liquids & nutraceuticals & religions & traumas\\
arts & coatings & environsciproc & ijerph & literature & nutrients & remotesensing & tropicalmed\\
\addlinespace
asc & colloids & epidemiologia & ijfs & livers & obesities & reports & universe\\
asi & colorants & epigenomes & ijgi & logics & oceans & reprodmed & urbansci\\
astronomy & commodities & est & ijms & logistics & ohbm & resources & uro\\
atmosphere & compounds & fermentation & ijns & lubricants & onco & rheumato & vaccines\\
atoms & computation & fibers & ijpb & lymphatics & oncopathology & risks & vehicles\\
\addlinespace
audiolres & computers & fintech & ijtm & machines & optics & robotics & venereology\\
automation & condensedmatter & fire & ijtpp & macromol & oral & ruminants & vetsci\\
axioms & conservation & fishes & ime & magnetism & organics & safety & vibration\\
bacteria & constrmater & fluids & immuno & magnetochemistry & organoids & sci & virtualworlds\\
batteries & cosmetics & foods & informatics & make & osteology & scipharm & viruses\\
\addlinespace
bdcc & covid & forecasting & information & marinedrugs & oxygen & sclerosis & vision\\
behavsci & crops & forensicsci & infrastructures & materials & parasites & seeds & waste\\
beverages & cryptography & forests & inorganics & materproc & parasitologia & sensors & water\\
biochem & crystals & foundations & insects & mathematics & particles & separations & wem\\
bioengineering & csmf & fractalfract & instruments & mca & pathogens & sexes & wevj\\
\addlinespace
biologics & ctn & fuels & inventions & measurements & pathophysiology & signals & wind\\
biology & curroncol & future & iot & medicina & pediatrrep & sinusitis & women\\
biomass & cyber & futureinternet & j & medicines & pharmaceuticals & skins & world\\
biomechanics & dairy & futurepharmacol & jal & medsci & pharmaceutics & smartcities & youth\\
biomed & data & futurephys & jcdd & membranes & pharmacoepidemiology & sna & zoonoticdis\\
\addlinespace
biomedicines & ddc & futuretransp & jcm & merits & pharmacy & societies & \\
\bottomrule
\end{longtable}
\finishlandscape

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

The introduction should briefly place the study in a broad context and
highlight why it is important. It should define the purpose of the work
and its significance. The current state of the research field should be
reviewed carefully and key publications cited. Please highlight
controversial and diverging hypotheses when necessary. Finally, briefly
mention the main aim of the work and highlight the principal
conclusions. As far as possible, please keep the introduction
comprehensible to scientists outside your particular field of research.
Citing a journal paper
\citep{bertrand-krajewski_distribution_1998, leutnant_stormwater_2016}.
And now citing a book reference \citet{gujer_systems_2008}. Some MDPI
journals use Chicago and others use APA, this template should choose the
correct citation format for you once you specify the journal in the YAML
header.

To use endnotes, change \texttt{endnotes:\ true} in the YAML header,
then use \texttt{\textbackslash{}endnote\{This\ is\ an\ endnote.\}}.

\hypertarget{materials-and-methods}{%
\section{Materials and Methods}\label{materials-and-methods}}

Materials and Methods should be described with sufficient details to
allow others to replicate and build on published results. Please note
that publication of your manuscript implicates that you must make all
materials, data, computer code, and protocols associated with the
publication available to readers. Please disclose at the submission
stage any restrictions on the availability of materials or information.
New methods and protocols should be described in detail while
well-established methods can be briefly described and appropriately
cited.

Research manuscripts reporting large datasets that are deposited in a
publicly available database should specify where the data have been
deposited and provide the relevant accession numbers. If the accession
numbers have not yet been obtained at the time of submission, please
state that they will be provided during review. They must be provided
prior to publication.

Interventionary studies involving animals or humans, and other studies
require ethical approval must list the authority that provided approval
and the corresponding ethical approval code.

\hypertarget{results}{%
\section{Results}\label{results}}

This section may be divided by subheadings. It should provide a concise
and precise description of the experimental results, their
interpretation as well as the experimental conclusions that can be
drawn.

\hypertarget{subsection-heading-here}{%
\subsection{Subsection Heading Here}\label{subsection-heading-here}}

Subsection text here.

\hypertarget{subsubsection-heading-here}{%
\subsubsection{Subsubsection Heading
Here}\label{subsubsection-heading-here}}

Bulleted lists look like this:

\begin{itemize}
\tightlist
\item
  First bullet
\item
  Second bullet
\item
  Third bullet
\end{itemize}

Numbered lists can be added as follows:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\tightlist
\item
  First item
\item
  Second item
\item
  Third item
\end{enumerate}

The text continues here.

\hypertarget{figures-tables-and-schemes}{%
\subsection{Figures, Tables and
Schemes}\label{figures-tables-and-schemes}}

All figures and tables should be cited in the main text as Figure
\ref{fig:fig1}, \ref{tab:tab1}, etc. To get cross-reference to figure
generated by R chunks include the
\texttt{\textbackslash{}\textbackslash{}label\{\}} tag in the
\texttt{fig.cap} attribute of the R chunk:
\texttt{fig.cap\ =\ "Fancy\ Caption\textbackslash{}\textbackslash{}label\{fig:plot\}"}.

\begin{figure}
\centering
\includegraphics{MDPI_submition_files/figure-latex/fig1-1.pdf}
\caption{A figure added with a code chunk.\label{fig:fig1}}
\end{figure}

When making tables using \texttt{kable}, it is suggested to use the
\texttt{format="latex"} and \texttt{tabl.envir="table"} arguments to
ensure table numbering and compatibility with the mdpi document class.

\begin{table}[H]

\caption{\label{tab:tab1}This is a table caption. Tables should be placed in the 
             main text near to the first time they are~cited.}
\begin{tabular}[t]{lccc}
\toprule
  & mpg & cyl & disp\\
\midrule
Mazda RX4 & 21.0 & 6 & 160\\
Mazda RX4 Wag & 21.0 & 6 & 160\\
Datsun 710 & 22.8 & 4 & 108\\
Hornet 4 Drive & 21.4 & 6 & 258\\
Hornet Sportabout & 18.7 & 8 & 360\\
\bottomrule
\end{tabular}
\end{table}

For a very wide table, landscape layouts are allowed.

\startlandscape

\begin{table}[H]

\caption{\label{tab:tab2}This is a very wide table}
\begin{tabular}[t]{cccc}
\toprule
Title.1 & Title.2 & Title.3 & Title.4\\
\midrule
Entry 1 & Data & Data & This cell has some longer content that runs over
                               two lines\\
Entry 2 & Data & Data & Data\\
\bottomrule
\end{tabular}
\end{table}

\finishlandscape

\hypertarget{formatting-of-mathematical-components}{%
\subsection{Formatting of Mathematical
Components}\label{formatting-of-mathematical-components}}

This is an example of an equation:

\[
a = 1.
\]

If you want numbered equations use Latex and wrap in the equation
environment:

\begin{equation}
a = 1,
\end{equation}

the text following an equation need not be a new paragraph. Please
punctuate equations as regular text.

This is the example 2 of equation:

\begin{adjustwidth}{-\extralength}{0cm}
\begin{equation}
a = b + c + d + e + f + g + h + i + j + k + l + m + n + o + p + q + r + s + t + 
u + v + w + x + y + z
\end{equation}
\end{adjustwidth}

Theorem-type environments (including propositions, lemmas, corollaries
etc.) can be formatted as follows:

Example of a theorem:

\begin{Theorem}
Example text of a theorem

\end{Theorem}

The text continues here. Proofs must be formatted as follows:

Example of a proof:

\begin{proof}[Proof of Theorem1]
Text of the proof. Note that the phrase ``of Theorem 1'\,' is optional
if it is clear which theorem is being referred to.

\end{proof}

The text continues here.

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

Authors should discuss the results and how they can be interpreted in
perspective of previous studies and of the working hypotheses. The
findings and their implications should be discussed in the broadest
context possible. Future research directions may also be highlighted.

\hypertarget{conclusion}{%
\section{Conclusion}\label{conclusion}}

This section is not mandatory, but can be added to the manuscript if the
discussion is unusually long or complex.

\hypertarget{patents}{%
\section{Patents}\label{patents}}

This section is not mandatory, but may be added if there are patents
resulting from the work reported in this manuscript.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\vspace{6pt}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% optional
\supplementary{The following supporting information can be downloaded
at:\\
\linksupplementary{s1}, Figure S1: title; Table S1: title; Video S1:
title.}

% Only for the journal Methods and Protocols:
% If you wish to submit a video article, please do so with any other supplementary material.
% \supplementary{The following supporting information can be downloaded at: \linksupplementary{s1}, Figure S1: title; Table S1: title; Video S1: title. A supporting video article is available at doi: link.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\authorcontributions{For research articles with several authors, a short
paragraph specifying their individual contributions must be provided.
The following statements should be used ``X.X. and Y.Y. conceive and
designed the experiments; X.X. performed the experiments; X.X. and Y.Y.
analyzed the data; W.W. contributed reagents/materials/analysis tools;
Y.Y. wrote the paper.'\,' Authorship must be limited to those who have
contributed substantially to the work reported.}

\funding{Please add:
\texttt{This\ research\ received\ no\ external\ funding\textquotesingle{}\textquotesingle{}\ or}This
research was funded by NAME OF FUNDER grant number XXX.'\,' and and
``The APC was funded by XXX'\,'. Check carefully that the details given
are accurate and use the standard spelling of funding agency names at
\url{https://search.crossref.org/funding}, any errors may affect your
future funding.}

\institutionalreview{In this section, you should add the Institutional
Review Board Statement and approval number, if relevant to your study.
You might choose to exclude this statement if the study did not require
ethical approval. Please note that the Editorial Office might ask you
for further information. Please add ``The study was conducted in
accordance with the Declaration of Helsinki, and approved by the
Institutional Review Board (or Ethics Committee) of NAME OF INSTITUTE
(protocol code XXX and date of approval).'' for studies involving
humans. OR ``The animal study protocol was approved by the Institutional
Review Board (or Ethics Committee) of NAME OF INSTITUTE (protocol code
XXX and date of approval).'' for studies involving animals. OR ``Ethical
review and approval were waived for this study due to REASON (please
provide a detailed justification).'' OR ``Not applicable'' for studies
not involving humans or animals.}

\informedconsent{Any research article describing a study involving
humans should contain this statement. Please add
\texttt{Informed\ consent\ was\ obtained\ from\ all\ subjects\ \ involved\ in\ the\ study.\textquotesingle{}\textquotesingle{}\ OR}Patient
consent was waived due to REASON (please provide a detailed
justification).'\,' OR ``Not applicable'\,' for studies not involving
humans. You might also choose to exclude this statement if the study did
not involve humans.

Written informed consent for publication must be obtained from
participating patients who can be identified (including by the patients
themselves). Please state ``Written informed consent has been obtained
from the patient(s) to publish this paper'\,' if applicable.}

\dataavailability{We encourage all authors of articles published in MDPI
journals to share their research data. In this section, please provide
details regarding where data supporting reported results can be found,
including links to publicly archived datasets analyzed or generated
during the study. Where no new data were created, or where data is
unavailable due to privacy or ethical re-strictions, a statement is
still required. Suggested Data Availability Statements are available in
section ``MDPI Research Data Policies'' at
\url{https://www.mdpi.com/ethics}.}

\acknowledgments{All sources of funding of the study should be
disclosed. Please clearly indicate grants that you have received in
support of your research work. Clearly state if you received funds for
covering the costs to publish in open access.}

\conflictsofinterest{Declare conflicts of interest or state `The authors
declare no conflict of interest.' Authors must identify and declare any
personal circumstances or interest that may be perceived as
inappropriately influencing the representation or interpretation of
reported research results. Any role of the funding sponsors in the
design of the study; in the collection, analyses or interpretation of
data in the writing of the manuscript, or in the decision to publish the
results must be declared in this section. If there is no role, please
state `The founding sponsors had no role in the design of the study; in
the collection, analyses, or interpretation of data; in the writing of
the manuscript, an in the decision to publish the results'.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Optional
\sampleavailability{Samples of the compounds \ldots\ldots{} are
available from the authors.}

%% Only for journal Encyclopedia

\abbreviations{Abbreviations}{
The following abbreviations are used in this manuscript:\\

\noindent
\begin{tabular}{@{}ll}
MDPI & Multidisciplinary Digital Publishing Institute \\
DOAJ & Directory of open access journals \\
TLA & Three letter acronym \\
LD & linear dichroism \\
\end{tabular}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Optional
\input{"appendix.tex"}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{adjustwidth}{-\extralength}{0cm}

%\printendnotes[custom] % Un-comment to print a list of endnotes

\reftitle{References}
\bibliography{mybibfile.bib}

% If authors have biography, please use the format below
%\section*{Short Biography of Authors}
%\bio
%{\raisebox{-0.35cm}{\includegraphics[width=3.5cm,height=5.3cm,clip,keepaspectratio]{Definitions/author1.pdf}}}
%{\textbf{Firstname Lastname} Biography of first author}
%
%\bio
%{\raisebox{-0.35cm}{\includegraphics[width=3.5cm,height=5.3cm,clip,keepaspectratio]{Definitions/author2.jpg}}}
%{\textbf{Firstname Lastname} Biography of second author}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% for journal Sci
%\reviewreports{\\
%Reviewer 1 comments and authors’ response\\
%Reviewer 2 comments and authors’ response\\
%Reviewer 3 comments and authors’ response
%}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\PublishersNote{}
\end{adjustwidth}

\end{document}

### 2023-11-01 template.tex
Commit 2023-11-01 16:14

%  LaTeX support: latex@mdpi.com
%  For support, please attach all files needed for compiling as well as the log file, and specify your operating system, LaTeX version, and LaTeX editor.

%=================================================================
\documentclass[journal,article,submit,pdftex,moreauthors]{Definitions/mdpi}

%--------------------
% Class Options:
%--------------------
%----------
% journal
%----------
% Choose between the following MDPI journals:
% acoustics, actuators, addictions, admsci, adolescents, aerobiology, aerospace, agriculture, agriengineering, agrochemicals, agronomy, ai, air, algorithms, allergies, alloys, analytica, analytics, anatomia, animals, antibiotics, antibodies, antioxidants, applbiosci, appliedchem, appliedmath, applmech, applmicrobiol, applnano, applsci, aquacj, architecture, arm, arthropoda, arts, asc, asi, astronomy, atmosphere, atoms, audiolres, automation, axioms, bacteria, batteries, bdcc, behavsci, beverages, biochem, bioengineering, biologics, biology, biomass, biomechanics, biomed, biomedicines, biomedinformatics, biomimetics, biomolecules, biophysica, biosensors, biotech, birds, bloods, blsf, brainsci, breath, buildings, businesses, cancers, carbon, cardiogenetics, catalysts, cells, ceramics, challenges, chemengineering, chemistry, chemosensors, chemproc, children, chips, cimb, civileng, cleantechnol, climate, clinpract, clockssleep, cmd, coasts, coatings, colloids, colorants, commodities, compounds, computation, computers, condensedmatter, conservation, constrmater, cosmetics, covid, crops, cryptography, crystals, csmf, ctn, curroncol, cyber, dairy, data, ddc, dentistry, dermato, dermatopathology, designs, devices, diabetology, diagnostics, dietetics, digital, disabilities, diseases, diversity, dna, drones, dynamics, earth, ebj, ecologies, econometrics, economies, education, ejihpe, electricity, electrochem, electronicmat, electronics, encyclopedia, endocrines, energies, eng, engproc, entomology, entropy, environments, environsciproc, epidemiologia, epigenomes, est, fermentation, fibers, fintech, fire, fishes, fluids, foods, forecasting, forensicsci, forests, foundations, fractalfract, fuels, future, futureinternet, futurepharmacol, futurephys, futuretransp, galaxies, games, gases, gastroent, gastrointestdisord, gels, genealogy, genes, geographies, geohazards, geomatics, geosciences, geotechnics, geriatrics, grasses, gucdd, hazardousmatters, healthcare, hearts, hemato, hematolrep, heritage, higheredu, highthroughput, histories, horticulturae, hospitals, humanities, humans, hydrobiology, hydrogen, hydrology, hygiene, idr, ijerph, ijfs, ijgi, ijms, ijns, ijpb, ijtm, ijtpp, ime, immuno, informatics, information, infrastructures, inorganics, insects, instruments, inventions, iot, j, jal, jcdd, jcm, jcp, jcs, jcto, jdb, jeta, jfb, jfmk, jimaging, jintelligence, jlpea, jmmp, jmp, jmse, jne, jnt, jof, joitmc, jor, journalmedia, jox, jpm, jrfm, jsan, jtaer, jvd, jzbg, kidneydial, kinasesphosphatases, knowledge, land, languages, laws, life, liquids, literature, livers, logics, logistics, lubricants, lymphatics, machines, macromol, magnetism, magnetochemistry, make, marinedrugs, materials, materproc, mathematics, mca, measurements, medicina, medicines, medsci, membranes, merits, metabolites, metals, meteorology, methane, metrology, micro, microarrays, microbiolres, micromachines, microorganisms, microplastics, minerals, mining, modelling, molbank, molecules, mps, msf, mti, muscles, nanoenergyadv, nanomanufacturing,\gdef\@continuouspages{yes}} nanomaterials, ncrna, ndt, network, neuroglia, neurolint, neurosci, nitrogen, notspecified, %%nri, nursrep, nutraceuticals, nutrients, obesities, oceans, ohbm, onco, %oncopathology, optics, oral, organics, organoids, osteology, oxygen, parasites, parasitologia, particles, pathogens, pathophysiology, pediatrrep, pharmaceuticals, pharmaceutics, pharmacoepidemiology,\gdef\@ISSN{2813-0618}\gdef\@continuous pharmacy, philosophies, photochem, photonics, phycology, physchem, physics, physiologia, plants, plasma, platforms, pollutants, polymers, polysaccharides, poultry, powders, preprints, proceedings, processes, prosthesis, proteomes, psf, psych, psychiatryint, psychoactives, publications, quantumrep, quaternary, qubs, radiation, reactions, receptors, recycling, regeneration, religions, remotesensing, reports, reprodmed, resources, rheumato, risks, robotics, ruminants, safety, sci, scipharm, sclerosis, seeds, sensors, separations, sexes, signals, sinusitis, skins, smartcities, sna, societies, socsci, software, soilsystems, solar, solids, spectroscj, sports, standards, stats, std, stresses, surfaces, surgeries, suschem, sustainability, symmetry, synbio, systems, targets, taxonomy, technologies, telecom, test, textiles, thalassrep, thermo, tomography, tourismhosp, toxics, toxins, transplantology, transportation, traumacare, traumas, tropicalmed, universe, urbansci, uro, vaccines, vehicles, venereology, vetsci, vibration, virtualworlds, viruses, vision, waste, water, wem, wevj, wind, women, world, youth, zoonoticdis
% For posting an early version of this manuscript as a preprint, you may use "preprints" as the journal. Changing "submit" to "accept" before posting will remove line numbers.

%---------
% article
%---------
% The default type of manuscript is "article", but can be replaced by:
% abstract, addendum, article, book, bookreview, briefreport, casereport, comment, commentary, communication, conferenceproceedings, correction, conferencereport, entry, expressionofconcern, extendedabstract, datadescriptor, editorial, essay, erratum, hypothesis, interestingimage, obituary, opinion, projectreport, reply, retraction, review, perspective, protocol, shortnote, studyprotocol, systematicreview, supfile, technicalnote, viewpoint, guidelines, registeredreport, tutorial
% supfile = supplementary materials

%----------
% submit
%----------
% The class option "submit" will be changed to "accept" by the Editorial Office when the paper is accepted. This will only make changes to the frontpage (e.g., the logo of the journal will get visible), the headings, and the copyright information. Also, line numbering will be removed. Journal info and pagination for accepted papers will also be assigned by the Editorial Office.

%------------------
% moreauthors
%------------------
% If there is only one author the class option oneauthor should be used. Otherwise use the class option moreauthors.

%---------
% pdftex
%---------
% The option pdftex is for use with pdfLaTeX. Remove "pdftex" for (1) compiling with LaTeX & dvi2pdf (if eps figures are used) or for (2) compiling with XeLaTeX.

%=================================================================
% MDPI internal commands - do not modify
\firstpage{1}
\makeatletter
\setcounter{page}{\@firstpage}
\makeatother
\pubvolume{1}
\issuenum{1}
\articlenumber{0}
\pubyear{2023}
\copyrightyear{2023}
%\externaleditor{Academic Editor: Firstname Lastname}
\datereceived{ }
\daterevised{ } % Comment out if no revised date
\dateaccepted{ }
\datepublished{ }
%\datecorrected{} % For corrected papers: "Corrected: XXX" date in the original paper.
%\dateretracted{} % For corrected papers: "Retracted: XXX" date in the original paper.
\hreflink{https://doi.org/} % If needed use \linebreak
%\doinum{}
%\pdfoutput=1 % Uncommented for upload to arXiv.org

%=================================================================
% Add packages and commands here. The following packages are loaded in our class file: fontenc, inputenc, calc, indentfirst, fancyhdr, graphicx, epstopdf, lastpage, ifthen, float, amsmath, amssymb, lineno, setspace, enumitem, mathpazo, booktabs, titlesec, etoolbox, tabto, xcolor, colortbl, soul, multirow, microtype, tikz, totcount, changepage, attrib, upgreek, array, tabularx, pbox, ragged2e, tocloft, marginnote, marginfix, enotez, amsthm, natbib, hyperref, cleveref, scrextend, url, geometry, newfloat, caption, draftwatermark, seqsplit
% cleveref: load \crefname definitions after \begin{document}

%=================================================================
% Please use the following mathematics environments: Theorem, Lemma, Corollary, Proposition, Characterization, Property, Problem, Example, ExamplesandDefinitions, Hypothesis, Remark, Definition, Notation, Assumption
%% For proofs, please use the proof environment (the amsthm package is loaded by the MDPI class).

%=================================================================
% Full title of the paper (Capitalized)
\Title{Title}

% MDPI internal command: Title for citation in the left column
\TitleCitation{Title}

% Author Orchid ID: enter ID or remove command
\newcommand{\orcidauthorA}{0000-0000-0000-000X} % Add \orcidA{} behind the author's name
%\newcommand{\orcidauthorB}{0000-0000-0000-000X} % Add \orcidB{} behind the author's name

% Authors, for the paper (add full first names)
\Author{Firstname Lastname $^{1,\dagger,\ddagger}$\orcidA{}, Firstname Lastname $^{2,\ddagger}$ and Firstname Lastname $^{2,}$*}

%\longauthorlist{yes}

% MDPI internal command: Authors, for metadata in PDF
\AuthorNames{Firstname Lastname, Firstname Lastname and Firstname Lastname}

% MDPI internal command: Authors, for citation in the left column
\AuthorCitation{Lastname, F.; Lastname, F.; Lastname, F.}
% If this is a Chicago style journal: Lastname, Firstname, Firstname Lastname, and Firstname Lastname.

% Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
\address{%
$^{1}$ \quad Affiliation 1; e-mail@e-mail.com\\
$^{2}$ \quad Affiliation 2; e-mail@e-mail.com}

% Contact information of the corresponding author
\corres{Correspondence: e-mail@e-mail.com; Tel.: (optional; include country code; if there are multiple corresponding authors, add author initials) +xx-xxxx-xxx-xxxx (F.L.)}

% Current address and/or shared authorship
\firstnote{Current address: Affiliation 3.}
\secondnote{These authors contributed equally to this work.}
% The commands \thirdnote{} till \eighthnote{} are available for further notes

%\simplesumm{} % Simple summary

%\conference{} % An extended version of a conference paper

% Abstract (Do not insert blank lines, i.e. \\)
\abstract{A single paragraph of about 200 words maximum. For research articles, abstracts should give a pertinent overview of the work. We strongly encourage authors to use the following style of structured abstracts, but without headings: (1) Background: place the question addressed in a broad context and highlight the purpose of the study; (2) Methods: describe briefly the main methods or treatments applied; (3) Results: summarize the article's main findings; (4) Conclusions: indicate the main conclusions or interpretations. The abstract should be an objective representation of the article, it must not contain results which are not presented and substantiated in the main text and should not exaggerate the main conclusions.}

% Keywords
\keyword{keyword 1; keyword 2; keyword 3 (List three to ten pertinent keywords specific to the article; yet reasonably common within the subject discipline.)}

% The fields PACS, MSC, and JEL may be left empty or commented out if not applicable
%\PACS{J0101}
%\MSC{}
%\JEL{}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Diversity
%\LSID{\url{http://}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Applied Sciences
%\featuredapplication{Authors are encouraged to provide a concise description of the specific application or a potential application of the work. This section is not mandatory.}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Data
%\dataset{DOI number or link to the deposited data set if the data set is published separately. If the data set shall be published as a supplement to this paper, this field will be filled by the journal editors. In this case, please submit the data set as a supplement.}
%\datasetlicense{License under which the data set is made available (CC0, CC-BY, CC-BY-SA, CC-BY-NC, etc.)}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Toxins
%\keycontribution{The breakthroughs or highlights of the manuscript. Authors can write one or two sentences to describe the most important part of the paper.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Encyclopedia
%\encyclopediadef{For entry manuscripts only: please provide a brief overview of the entry title instead of an abstract.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Advances in Respiratory Medicine
%\addhighlights{yes}
%\renewcommand{\addhighlights}{%

%\noindent This is an obligatory section in “Advances in Respiratory Medicine”, whose goal is to increase the discoverability and readability of the article via search engines and other scholars. Highlights should not be a copy of the abstract, but a simple text allowing the reader to quickly and simplified find out what the article is about and what can be cited from it. Each of these parts should be devoted up to 2~bullet points.\vspace{3pt}\\
%\textbf{What are the main findings?}
% \begin{itemize}[labelsep=2.5mm,topsep=-3pt]
% \item First bullet.
% \item Second bullet.
% \end{itemize}\vspace{3pt}
%\textbf{What is the implication of the main finding?}
% \begin{itemize}[labelsep=2.5mm,topsep=-3pt]
% \item First bullet.
% \item Second bullet.
% \end{itemize}
%}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\setcounter{section}{-1} %% Remove this when starting to work on the template.
\section{How to Use this Template}

The template details the sections that can be used in a manuscript. Note that the order and names of article sections may differ from the requirements of the journal (e.g., the positioning of the Materials and Methods section). Please check the instructions on the authors' page of the journal to verify the correct order and names. For any questions, please contact the editorial office of the journal or support@mdpi.com. For LaTeX-related questions please contact latex@mdpi.com.%\endnote{This is an endnote.} % To use endnotes, please un-comment \printendnotes below (before References). Only journal Laws uses \footnote.

% The order of the section titles is different for some journals. Please refer to the "Instructions for Authors” on the journal homepage.

\section{Introduction}

The introduction should briefly place the study in a broad context and highlight why it is important. It should define the purpose of the work and its significance. The current state of the research field should be reviewed carefully and key publications cited. Please highlight controversial and diverging hypotheses when necessary. Finally, briefly mention the main aim of the work and highlight the principal conclusions. As far as possible, please keep the introduction comprehensible to scientists outside your particular field of research. Citing a journal paper \cite{ref-journal}. Now citing a book reference \cite{ref-book1,ref-book2} or other reference types \cite{ref-unpublish,ref-communication,ref-proceeding}. Please use the command \citep{ref-thesis,ref-url} for the following MDPI journals, which use author--date citation: Administrative Sciences, Arts, Econometrics, Economies, Genealogy, Humanities, IJFS, Journal of Intelligence, Journalism and Media, JRFM, Languages, Laws, Religions, Risks, Social Sciences, Literature.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Materials and Methods}

Materials and Methods should be described with sufficient details to allow others to replicate and build on published results. Please note that publication of your manuscript implicates that you must make all materials, data, computer code, and protocols associated with the publication available to readers. Please disclose at the submission stage any restrictions on the availability of materials or information. New methods and protocols should be described in detail while well-established methods can be briefly described and appropriately cited.

Research manuscripts reporting large datasets that are deposited in a publicly avail-able database should specify where the data have been deposited and provide the relevant accession numbers. If the accession numbers have not yet been obtained at the time of submission, please state that they will be provided during review. They must be provided prior to publication.

Interventionary studies involving animals or humans, and other studies require ethical approval must list the authority that provided approval and the corresponding ethical approval code.
\begin{quote}
This is an example of a quote.
\end{quote}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results}

This section may be divided by subheadings. It should provide a concise and precise description of the experimental results, their interpretation as well as the experimental conclusions that can be drawn.
\subsection{Subsection}
\subsubsection{Subsubsection}

Bulleted lists look like this:
\begin{itemize}
\item	First bullet;
\item	Second bullet;
\item	Third bullet.
\end{itemize}

Numbered lists can be added as follows:
\begin{enumerate}
\item	First item;
\item	Second item;
\item	Third item.
\end{enumerate}

The text continues here.

\subsection{Figures, Tables and Schemes}

All figures and tables should be cited in the main text as Figure~\ref{fig1}, Table~\ref{tab1}, etc.

\begin{figure}[H]
\includegraphics[width=10.5 cm]{Definitions/logo-mdpi}
\caption{This is a figure. Schemes follow the same formatting. If there are multiple panels, they should be listed as: (\textbf{a}) Description of what is contained in the first panel. (\textbf{b}) Description of what is contained in the second panel. Figures should be placed in the main text near to the first time they are cited. A caption on a single line should be centered.\label{fig1}}
\end{figure}
\unskip

\begin{table}[H]
\caption{This is a table caption. Tables should be placed in the main text near to the first time they are~cited.\label{tab1}}
\newcolumntype{C}{>{\centering\arraybackslash}X}
\begin{tabularx}{\textwidth}{CCC}
\toprule
\textbf{Title 1}	& \textbf{Title 2}	& \textbf{Title 3}\\
\midrule
Entry 1		& Data			& Data\\
Entry 2		& Data			& Data \textsuperscript{1}\\
\bottomrule
\end{tabularx}
\noindent{\footnotesize{\textsuperscript{1} Tables may have a footer.}}
\end{table}

The text continues here (Figure~\ref{fig2} and Table~\ref{tab2}).

% Example of a figure that spans the whole page width. The same concept works for tables, too.
\begin{figure}[H]
\begin{adjustwidth}{-\extralength}{0cm}
\centering
\includegraphics[width=15.5cm]{Definitions/logo-mdpi}
\end{adjustwidth}
\caption{This is a wide figure.\label{fig2}}
\end{figure}

\begin{table}[H]
\caption{This is a wide table.\label{tab2}}
	\begin{adjustwidth}{-\extralength}{0cm}
		\newcolumntype{C}{>{\centering\arraybackslash}X}
		\begin{tabularx}{\fulllength}{CCCC}
			\toprule
			\textbf{Title 1}	& \textbf{Title 2}	& \textbf{Title 3}     & \textbf{Title 4}\\
			\midrule
\multirow[m]{3}{*}{Entry 1 *}	& Data			& Data			& Data\\
			  	                   & Data			& Data			& Data\\
			             	      & Data			& Data			& Data\\
                   \midrule
\multirow[m]{3}{*}{Entry 2}    & Data			& Data			& Data\\
			  	                  & Data			& Data			& Data\\
			             	     & Data			& Data			& Data\\
                   \midrule
\multirow[m]{3}{*}{Entry 3}    & Data			& Data			& Data\\
			  	                 & Data			& Data			& Data\\
			             	    & Data			& Data			& Data\\
                  \midrule
\multirow[m]{3}{*}{Entry 4}   & Data			& Data			& Data\\
			  	                 & Data			& Data			& Data\\
			             	    & Data			& Data			& Data\\
			\bottomrule
		\end{tabularx}
	\end{adjustwidth}
	\noindent{\footnotesize{* Tables may have a footer.}}
\end{table}

%\begin{listing}[H]
%\caption{Title of the listing}
%\rule{\columnwidth}{1pt}
%\raggedright Text of the listing. In font size footnotesize, small, or normalsize. Preferred format: left aligned and single spaced. Preferred border format: top border line and bottom border line.
%\rule{\columnwidth}{1pt}
%\end{listing}

Text.

Text.

\subsection{Formatting of Mathematical Components}

This is the example 1 of equation:
\begin{linenomath}
\begin{equation}
a = 1,
\end{equation}
\end{linenomath}
the text following an equation need not be a new paragraph. Please punctuate equations as regular text.
%% If the documentclass option "submit" is chosen, please insert a blank line before and after any math environment (equation and eqnarray environments). This ensures correct linenumbering. The blank line should be removed when the documentclass option is changed to "accept" because the text following an equation should not be a new paragraph.

This is the example 2 of equation:
\begin{adjustwidth}{-\extralength}{0cm}
\begin{equation}
a = b + c + d + e + f + g + h + i + j + k + l + m + n + o + p + q + r + s + t + u + v + w + x + y + z
\end{equation}
\end{adjustwidth}

% Example of a page in landscape format (with table and table footnote).
%\startlandscape
%\begin{table}[H] %% Table in wide page
%\caption{This is a very wide table.\label{tab3}}
%	\begin{tabularx}{\textwidth}{CCCC}
%		\toprule
%		\textbf{Title 1}	& \textbf{Title 2}	& \textbf{Title 3}	& \textbf{Title 4}\\
%		\midrule
%		Entry 1		& Data			& Data			& This cell has some longer content that runs over two lines.\\
%		Entry 2		& Data			& Data			& Data\textsuperscript{1}\\
%		\bottomrule
%	\end{tabularx}
%	\begin{adjustwidth}{+\extralength}{0cm}
%		\noindent\footnotesize{\textsuperscript{1} This is a table footnote.}
%	\end{adjustwidth}
%\end{table}
%\finishlandscape

Please punctuate equations as regular text. Theorem-type environments (including propositions, lemmas, corollaries etc.) can be formatted as follows:
%% Example of a theorem:
\begin{Theorem}
Example text of a theorem.
\end{Theorem}

The text continues here. Proofs must be formatted as follows:

%% Example of a proof:
\begin{proof}[Proof of Theorem 1]
Text of the proof. Note that the phrase ``of Theorem 1'' is optional if it is clear which theorem is being referred to.
\end{proof}
The text continues here.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion}

Authors should discuss the results and how they can be interpreted from the perspective of previous studies and of the working hypotheses. The findings and their implications should be discussed in the broadest context possible. Future research directions may also be highlighted.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions}

This section is not mandatory, but can be added to the manuscript if the discussion is unusually long or complex.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Patents}

This section is not mandatory, but may be added if there are patents resulting from the work reported in this manuscript.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\vspace{6pt}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% optional
%\supplementary{The following supporting information can be downloaded at:  \linksupplementary{s1}, Figure S1: title; Table S1: title; Video S1: title.}

% Only for journal Methods and Protocols:
% If you wish to submit a video article, please do so with any other supplementary material.
% \supplementary{The following supporting information can be downloaded at: \linksupplementary{s1}, Figure S1: title; Table S1: title; Video S1: title. A supporting video article is available at doi: link.}

% Only for journal Hardware:
% If you wish to submit a video article, please do so with any other supplementary material.
% \supplementary{The following supporting information can be downloaded at: \linksupplementary{s1}, Figure S1: title; Table S1: title; Video S1: title.\vspace{6pt}\\
%\begin{tabularx}{\textwidth}{lll}
%\toprule
%\textbf{Name} & \textbf{Type} & \textbf{Description} \\
%\midrule
%S1 & Python script (.py) & Script of python source code used in XX \\
%S2 & Text (.txt) & Script of modelling code used to make Figure X \\
%S3 & Text (.txt) & Raw data from experiment X \\
%S4 & Video (.mp4) & Video demonstrating the hardware in use \\
%... & ... & ... \\
%\bottomrule
%\end{tabularx}
%}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\authorcontributions{For research articles with several authors, a short paragraph specifying their individual contributions must be provided. The following statements should be used ``Conceptualization, X.X. and Y.Y.; methodology, X.X.; software, X.X.; validation, X.X., Y.Y. and Z.Z.; formal analysis, X.X.; investigation, X.X.; resources, X.X.; data curation, X.X.; writing---original draft preparation, X.X.; writing---review and editing, X.X.; visualization, X.X.; supervision, X.X.; project administration, X.X.; funding acquisition, Y.Y. All authors have read and agreed to the published version of the manuscript.'', please turn to the  \href{http://img.mdpi.org/data/contributor-role-instruction.pdf}{CRediT taxonomy} for the term explanation. Authorship must be limited to those who have contributed substantially to the work~reported.}

\funding{Please add: ``This research received no external funding'' or ``This research was funded by NAME OF FUNDER grant number XXX.'' and  and ``The APC was funded by XXX''. Check carefully that the details given are accurate and use the standard spelling of funding agency names at \url{https://search.crossref.org/funding}, any errors may affect your future funding.}

\institutionalreview{In this section, you should add the Institutional Review Board Statement and approval number, if relevant to your study. You might choose to exclude this statement if the study did not require ethical approval. Please note that the Editorial Office might ask you for further information. Please add “The study was conducted in accordance with the Declaration of Helsinki, and approved by the Institutional Review Board (or Ethics Committee) of NAME OF INSTITUTE (protocol code XXX and date of approval).” for studies involving humans. OR “The animal study protocol was approved by the Institutional Review Board (or Ethics Committee) of NAME OF INSTITUTE (protocol code XXX and date of approval).” for studies involving animals. OR “Ethical review and approval were waived for this study due to REASON (please provide a detailed justification).” OR “Not applicable” for studies not involving humans or animals.}

\informedconsent{Any research article describing a study involving humans should contain this statement. Please add ``Informed consent was obtained from all subjects involved in the study.'' OR ``Patient consent was waived due to REASON (please provide a detailed justification).'' OR ``Not applicable'' for studies not involving humans. You might also choose to exclude this statement if the study did not involve humans.

Written informed consent for publication must be obtained from participating patients who can be identified (including by the patients themselves). Please state ``Written informed consent has been obtained from the patient(s) to publish this paper'' if applicable.}

\dataavailability{We encourage all authors of articles published in MDPI journals to share their research data. In this section, please provide details regarding where data supporting reported results can be found, including links to publicly archived datasets analyzed or generated during the study. Where no new data were created, or where data is unavailable due to privacy or ethical restrictions, a statement is still required. Suggested Data Availability Statements are available in section ``MDPI Research Data Policies'' at \url{https://www.mdpi.com/ethics}.}

% Only for journal Nursing Reports
%\publicinvolvement{Please describe how the public (patients, consumers, carers) were involved in the research. Consider reporting against the GRIPP2 (Guidance for Reporting Involvement of Patients and the Public) checklist. If the public were not involved in any aspect of the research add: ``No public involvement in any aspect of this research''.}

% Only for journal Nursing Reports
%\guidelinesstandards{Please add a statement indicating which reporting guideline was used when drafting the report. For example, ``This manuscript was drafted against the XXX (the full name of reporting guidelines and citation) for XXX (type of research) research''. A complete list of reporting guidelines can be accessed via the equator network: \url{https://www.equator-network.org/}.}

\acknowledgments{In this section you can acknowledge any support given which is not covered by the author contribution or funding sections. This may include administrative and technical support, or donations in kind (e.g., materials used for experiments).}

\conflictsofinterest{Declare conflicts of interest or state ``The authors declare no conflict of interest.'' Authors must identify and declare any personal circumstances or interest that may be perceived as inappropriately influencing the representation or interpretation of reported research results. Any role of the funders in the design of the study; in the collection, analyses or interpretation of data; in the writing of the manuscript; or in the decision to publish the results must be declared in this section. If there is no role, please state ``The funders had no role in the design of the study; in the collection, analyses, or interpretation of data; in the writing of the manuscript; or in the decision to publish the results''.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Optional

%% Only for journal Encyclopedia
%\entrylink{The Link to this entry published on the encyclopedia platform.}

\abbreviations{Abbreviations}{
The following abbreviations are used in this manuscript:\\

\noindent
\begin{tabular}{@{}ll}
MDPI & Multidisciplinary Digital Publishing Institute\\
DOAJ & Directory of open access journals\\
TLA & Three letter acronym\\
LD & Linear dichroism
\end{tabular}
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Optional
\appendixtitles{no} % Leave argument "no" if all appendix headings stay EMPTY (then no dot is printed after "Appendix A"). If the appendix sections contain a heading then change the argument to "yes".
\appendixstart
\appendix
\section[\appendixname~\thesection]{}
\subsection[\appendixname~\thesubsection]{}
The appendix is an optional section that can contain details and data supplemental to the main text---for example, explanations of experimental details that would disrupt the flow of the main text but nonetheless remain crucial to understanding and reproducing the research shown; figures of replicates for experiments of which representative data are shown in the main text can be added here if brief, or as Supplementary Data. Mathematical proofs of results not central to the paper can be added as an appendix.

\begin{table}[H]
\caption{This is a table caption.\label{tab5}}
\newcolumntype{C}{>{\centering\arraybackslash}X}
\begin{tabularx}{\textwidth}{CCC}
\toprule
\textbf{Title 1}	& \textbf{Title 2}	& \textbf{Title 3}\\
\midrule
Entry 1		& Data			& Data\\
Entry 2		& Data			& Data\\
\bottomrule
\end{tabularx}
\end{table}

\section[\appendixname~\thesection]{}
All appendix sections must be cited in the main text. In the appendices, Figures, Tables, etc. should be labeled, starting with ``A''---e.g., Figure A1, Figure A2, etc.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{adjustwidth}{-\extralength}{0cm}
%\printendnotes[custom] % Un-comment to print a list of endnotes

\reftitle{References}

% Please provide either the correct journal abbreviation (e.g. according to the “List of Title Word Abbreviations” http://www.issn.org/services/online-services/access-to-the-ltwa/) or the full name of the journal.
% Citations and References in Supplementary files are permitted provided that they also appear in the reference list here.

%=====================================
% References, variant A: external bibliography
%=====================================
%\bibliography{your_external_BibTeX_file}

%=====================================
% References, variant B: internal bibliography
%=====================================
\begin{thebibliography}{999}
% Reference 1
\bibitem[Author1(year)]{ref-journal}
Author~1, T. The title of the cited article. {\em Journal Abbreviation} {\bf 2008}, {\em 10}, 142--149.
% Reference 2
\bibitem[Author2(year)]{ref-book1}
Author~2, L. The title of the cited contribution. In {\em The Book Title}; Editor 1, F., Editor 2, A., Eds.; Publishing House: City, Country, 2007; pp. 32--58.
% Reference 3
\bibitem[Author3(year)]{ref-book2}
Author 1, A.; Author 2, B. \textit{Book Title}, 3rd ed.; Publisher: Publisher Location, Country, 2008; pp. 154--196.
% Reference 4
\bibitem[Author4(year)]{ref-unpublish}
Author 1, A.B.; Author 2, C. Title of Unpublished Work. \textit{Abbreviated Journal Name} year, \textit{phrase indicating stage of publication (submitted; accepted; in press)}.
% Reference 5
\bibitem[Author5(year)]{ref-communication}
Author 1, A.B. (University, City, State, Country); Author 2, C. (Institute, City, State, Country). Personal communication, 2012.
% Reference 6
\bibitem[Author6(year)]{ref-proceeding}
Author 1, A.B.; Author 2, C.D.; Author 3, E.F. Title of presentation. In Proceedings of the Name of the Conference, Location of Conference, Country, Date of Conference (Day Month Year); Abstract Number (optional), Pagination (optional).
% Reference 7
\bibitem[Author7(year)]{ref-thesis}
Author 1, A.B. Title of Thesis. Level of Thesis, Degree-Granting University, Location of University, Date of Completion.
% Reference 8
\bibitem[Author8(year)]{ref-url}
Title of Site. Available online: URL (accessed on Day Month Year).
\end{thebibliography}

% If authors have biography, please use the format below
%\section*{Short Biography of Authors}
%\bio
%{\raisebox{-0.35cm}{\includegraphics[width=3.5cm,height=5.3cm,clip,keepaspectratio]{Definitions/author1.pdf}}}
%{\textbf{Firstname Lastname} Biography of first author}
%
%\bio
%{\raisebox{-0.35cm}{\includegraphics[width=3.5cm,height=5.3cm,clip,keepaspectratio]{Definitions/author2.jpg}}}
%{\textbf{Firstname Lastname} Biography of second author}

% For the MDPI journals use author-date citation, please follow the formatting guidelines on http://www.mdpi.com/authors/references
% To cite two works by the same author: \citeauthor{ref-journal-1a} (\citeyear{ref-journal-1a}, \citeyear{ref-journal-1b}). This produces: Whittaker (1967, 1975)
% To cite two works by the same author with specific pages: \citeauthor{ref-journal-3a} (\citeyear{ref-journal-3a}, p. 328; \citeyear{ref-journal-3b}, p.475). This produces: Wong (1999, p. 328; 2000, p. 475)

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% for journal Sci
%\reviewreports{\\
%Reviewer 1 comments and authors’ response\\
%Reviewer 2 comments and authors’ response\\
%Reviewer 3 comments and authors’ response
%}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\PublishersNote{}
\end{adjustwidth}
\end{document}

### 2023-11-01 Article.Rmd
Commit 2023-11-01 17:09

@@
     biblio-style:     apalike
-# date:           "`r format(Sys.time(), '%F')`"
 ---
@@
 
-<!-- use SDR for GHI -->
-
 # Introduction.
 
-<!-- something about aerosols: bais mail -->
-<!-- something about clouds: bais mail -->
 <!--
@@
 ...... cloud formation occurrences and constitution nuclei. cite .....
-p>        "  Too hard to vim
 

### 2023-11-01 MDPI_submition.Rmd
Commit 2023-11-01 17:09

--- 2023-11-01 MDPI_submition/MDPI_submition.Rmd
@@
   - name:  Athanasios Natsis
-    affil: 1,2,* # \ddagger
+    affil: 1,* # \ddagger
     orcid: 0000-0002-5199-4119
   - name: Alkiviadis Bais
-    affil: 2, \dagger, # \ddagger
+    affil: 2, #\dagger, # \ddagger
 affiliation:
@@
       Aristotle University of Thessaloniki - 
-      Laboratory of Atmospheric Physics
+      Laboratory of Atmospheric Physics,
       Campus Box 149, 54124 Thessaloniki, Greece
@@
       Street, City, Country
-    email: abais@auth.gr
+#     email: abais@auth.gr
 # author citation list in chicago format
@@
 # firstnote to eighthnote
-firstnote: |
-  Current address: Updated affiliation
+# firstnote: |
+#   Current address: Updated affiliation
 # secondnote: |
@@
 correspondence: |
-  leutnant@fh-muenster.de; Tel.: +XX-000-00-0000.
+  natsisphysicist@gmail.com;
+#   Tel.: +XX-000-00-0000.
 # document options
-journal: notspecified
+journal: applsci
 type:    article
@@
 abstract: |
-  A single paragraph of about 200 words maximum. For research articles, 
-  abstracts should give a pertinent overview of the work. We strongly encourage
-  authors to use the following style of structured abstracts, but without 
-  headings: 1) Background: Place the question addressed in a broad context and
-  highlight the purpose of the study; 2) Methods: Describe briefly the main
-  methods or treatments applied; 3) Results: Summarize the article's main 
-  findings; and 4) Conclusion: Indicate the main conclusions or interpretations. 
-  The abstract should be an objective representation of the article, it must not 
-  contain results which are not presented and substantiated in the main text and 
-  should not exaggerate the main conclusions.
+  The shortwave downward solar irradiance (SDR) is an important factor that drives climate processes, production and can affect all living organisms.
+  While monitoring the long term variability, there are observations of upward and downward SDR trends on different locations around the world for different time periods.
+  Periods of positive tredns are refered as brightening periods and with negative tredns as dimming periods.
+  We studied 29 years of CHP1 data from Thessaloniki, Greece, under three sky conditions (clear sky, cloudy sky and all sky conditions), applying a cloud sky identification algorithm.
+  We found a positive trend for all-sky and clear-sky conditions, and also, investigated the consistency of those trends, the effect of the solar zenith angle, and the variation of the trends for the seasons of the year.
+  We indentified that there are some anomalies in the long term SDR trends, for all sky conditions.
 # back matter
 keywords: |
-  keyword 1; keyword 2; keyword 3 (list three to ten pertinent keywords specific 
+  GHI; SDR; Solar Brigthening/Dimming (list three to ten pertinent keywords specific 
   to the article, yet reasonably common within the subject discipline.).
@@
    studies not involving humans or animals.
-informedconsent: |
-  Any research article describing a study involving humans should contain this 
-  statement. Please add ``Informed consent was obtained from all subjects 
-  involved in the study.'' OR ``Patient consent was waived due to REASON 
-  (please provide a detailed justification).'' OR ``Not applicable'' for 
-  studies not involving humans. You might also choose to exclude this statement 
-  if the study did not involve humans.
-
-  Written informed consent for publication must be obtained from participating 
-  patients who can be identified (including by the patients themselves). Please 
-  state ``Written informed consent has been obtained from the patient(s) to 
-  publish this paper'' if applicable.
+# informedconsent: |
+#   Any research article describing a study involving humans should contain this 
+#   statement. Please add ``Informed consent was obtained from all subjects 
+#   involved in the study.'' OR ``Patient consent was waived due to REASON 
+#   (please provide a detailed justification).'' OR ``Not applicable'' for 
+#   studies not involving humans. You might also choose to exclude this statement 
+#   if the study did not involve humans.
+#   
+#   Written informed consent for publication must be obtained from participating 
+#   patients who can be identified (including by the patients themselves). Please 
+#   state ``Written informed consent has been obtained from the patient(s) to 
+#   publish this paper'' if applicable.
 dataavailability: |
@@
   manuscript, an in the decision to publish the results'.
-sampleavailability: |
-  Samples of the compounds ...... are available from the authors.
+# sampleavailability: |
+#   Samples of the compounds ...... are available from the authors.
 supplementary: |
@@
     long: Three letter acronym
-  - short: LD 
-    long: linear dichroism
+  - short: SDR 
+    long: Solar downward radiation
 bibliography: mybibfile.bib
@@
     extra_dependencies:  longtable
+    keep_tex:            yes
+
 ---
@@
 
-The YAML header includes information needed mainly for formatting the front and 
-back matter of the article. Required elements include:
-
-```yaml
-title: Title of the paper
-author:
-  - name: first and last name
-    affil: |
-      One or more comma seperated numbers corresponding to affilitation
-      and one or more  comma seperated symbols corresponding 
-      optional notes.
-    orcid: optional orcid number
-affiliation:  
-  - num: 1,..., n for each affiliation
-    address: required
-    email: required
-authorcitation: |
-  Lastname, F.
-correspondence: |
-  email@email.com; Tel.: +XX-000-00-0000.
-journal: notspecified
-type: article
-status: submit
-```
-
-Journal options are in Table \ref{tab:mdpinames}. The `status` variable should 
-generally not be changed by authors. The `type` variable describes 
-the type of of submission and defaults to `article` but can be replaced with any of the ones in Table \ref{tab:mdpitype}
-
-```{r mdpitype, echo = FALSE}
-type <- c("abstract, addendum, article, book, bookreview, briefreport, 
-casereport, comment, commentary, communication, conferenceproceedings, 
-correction, conferencereport, entry, expressionofconcern, extendedabstract, 
-datadescriptor, editorial, essay, erratum, hypothesis, interestingimage, 
-obituary, opinion, projectreport, reply, retraction, review, perspective, 
-protocol, shortnote, studyprotocol, systematicreview, supfile, 
-technicalnote, viewpoint, guidelines, registeredreport, tutorial")
-knitr::kable(rticles::string_to_table(type, 3), 
-             col.names = NULL, format = "latex", booktab = TRUE, 
-             caption = "MDPI article types.")
-```
-
 ## Journal Specific YAML variables
-
-```yaml
-# for journal Diversity,
-# add the Life Science Identifier using:
-lsid: http://zoobank.org/urn:lsid:zoobank.org:act:nnnn
-
-# for journal Applied Sciences
-# add featured application
-featuredapplication: |
-  Authors are encouraged to provide a concise 
-  description of the specific application or 
-  a potential application of the work. This 
-  section is not mandatory.
-
-# for the journal Data
-# add dataset doi and license
-dataset: https://doi.org/10.1000/182
-datasetlicense: CC-BY-4.0
-
-# for the journal Toxins
-# add key contributions
-keycontributions: |
-  The breakthroughs or highlights of the manuscript. 
-  Authors can write one or two sentences to describe 
-  the most important part of the paper.
-
-# for the journal Encyclopedia
-encyclopediadef: |
-  For entry manuscripts only: please provide a brief overview
-  of the entry title instead of an abstract.
-entrylink: The Link to this entry published on the encyclopedia platform.
-
-# for the journal Advances in Respiratory Medicine
-# add highlights
-addhighlights: |
-  This is an obligatory section in “Advances in Respiratory Medicine”, 
-  whose goal is to increase the discoverability and readability of the
-  article via search engines and other scholars. Highlights should not 
-  be a copy of the abstract, but a simple text allowing the reader to 
-  quickly and simplified find out what the article is about and what can 
-  be cited from it. Each of these parts should be devoted up to 2~bullet 
-  points.
-
-```
 

### 2023-11-01 MDPI_submition.tex
Commit 2023-11-01 17:09

--- 2023-11-01 MDPI_submition/MDPI_submition.tex
@@
 
-\documentclass[notspecified,article,submit,moreauthors,pdftex]{Definitions/mdpi}
+\documentclass[applsci,article,submit,moreauthors,pdftex]{Definitions/mdpi}
 
@@
 \Author{Athanasios
-Natsis$^{1,2,\ddagger,*}$\href{https://orcid.org/0000-0002-5199-4119}
-{\orcidicon}, Alkiviadis Bais$^{2, \dagger,}$}
+Natsis$^{1,*}$\href{https://orcid.org/0000-0002-5199-4119}
+{\orcidicon}, Alkiviadis Bais$^{2,}$}
 
@@
 $^{1}$ \quad Aristotle University of Thessaloniki - Laboratory of
-Atmospheric Physics Campus Box 149, 54124 Thessaloniki,
+Atmospheric Physics, Campus Box 149, 54124 Thessaloniki,
 Greece; \href{mailto:natsisphysicist@gmail.com}{\nolinkurl{natsisphysicist@gmail.com}}\\
-$^{2}$ \quad Your department Street, City,
-Country; \href{mailto:abais@auth.gr}{\nolinkurl{abais@auth.gr}}\\
+$^{2}$ \quad Your department Street, City, Country; \\
 }
@@
 % Contact information of the corresponding author
-\corres{Correspondence: \href{mailto:leutnant@fh-muenster.de}{\nolinkurl{leutnant@fh-muenster.de}};
-Tel.: +XX-000-00-0000.}
+\corres{Correspondence: \href{mailto:natsisphysicist@gmail.com}{\nolinkurl{natsisphysicist@gmail.com}};}
 
 % Current address and/or shared authorship
-\firstnote{Current address: Updated affiliation}
 
@@
 % Abstract (Do not insert blank lines, i.e. \\)
-\abstract{A single paragraph of about 200 words maximum. For research
-articles, abstracts should give a pertinent overview of the work. We
-strongly encourage authors to use the following style of structured
-abstracts, but without headings: 1) Background: Place the question
-addressed in a broad context and highlight the purpose of the study; 2)
-Methods: Describe briefly the main methods or treatments applied; 3)
-Results: Summarize the article's main findings; and 4) Conclusion:
-Indicate the main conclusions or interpretations. The abstract should be
-an objective representation of the article, it must not contain results
-which are not presented and substantiated in the main text and should
-not exaggerate the main conclusions.}
+\abstract{The shortwave downward solar irradiance (SDR) is an important
+factor that drives climate processes, production and can affect all
+living organisms. While monitoring the long term variability, there are
+observations of upward and downward SDR trends on different locations
+around the world for different time periods. Periods of positive tredns
+are refered as brightening periods and with negative tredns as dimming
+periods. We studied 29 years of CHP1 data from Thessaloniki, Greece,
+under three sky conditions (clear sky, cloudy sky and all sky
+conditions), applying a cloud sky identification algorithm. We found a
+positive trend for all-sky and clear-sky conditions, and also,
+investigated the consistency of those trends, the effect of the solar
+zenith angle, and the variation of the trends for the seasons of the
+year. We indentified that there are some anomalies in the long term SDR
+trends, for all sky conditions.}
 
 % Keywords
-\keyword{keyword 1; keyword 2; keyword 3 (list three to ten pertinent
-keywords specific to the article, yet reasonably common within the
-subject discipline.).}
+\keyword{GHI; SDR; Solar Brigthening/Dimming (list three to ten
+pertinent keywords specific to the article, yet reasonably common within
+the subject discipline.).}
 
@@
 
-% Pandoc syntax highlighting
-\usepackage{color}
-\usepackage{fancyvrb}
-\newcommand{\VerbBar}{|}
-\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
-\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
-% Add ',fontsize=\small' for more characters per line
-\usepackage{framed}
-\definecolor{shadecolor}{RGB}{248,248,248}
-\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
-\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{#1}}
-\newcommand{\AnnotationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\AttributeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
-\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
-\newcommand{\BuiltInTok}[1]{#1}
-\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
-\newcommand{\CommentVarTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\ConstantTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
-\newcommand{\ControlFlowTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
-\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{#1}}
-\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
-\newcommand{\DocumentationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\ErrorTok}[1]{\textcolor[rgb]{0.64,0.00,0.00}{\textbf{#1}}}
-\newcommand{\ExtensionTok}[1]{#1}
-\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{#1}}
-\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
-\newcommand{\ImportTok}[1]{#1}
-\newcommand{\InformationTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{#1}}}
-\newcommand{\NormalTok}[1]{#1}
-\newcommand{\OperatorTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
-\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{#1}}
-\newcommand{\PreprocessorTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{#1}}}
-\newcommand{\RegionMarkerTok}[1]{#1}
-\newcommand{\SpecialCharTok}[1]{\textcolor[rgb]{0.81,0.36,0.00}{\textbf{#1}}}
-\newcommand{\SpecialStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\VariableTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{#1}}
-\newcommand{\VerbatimStringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{#1}}
-\newcommand{\WarningTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textbf{\textit{#1}}}}
-
 % tightlist command for lists without linebreak
@@
 
-The YAML header includes information needed mainly for formatting the
-front and back matter of the article. Required elements include:
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\FunctionTok{title}\KeywordTok{:}\AttributeTok{ Title of the paper}
-\FunctionTok{author}\KeywordTok{:}
-\AttributeTok{  }\KeywordTok{{-}}\AttributeTok{ }\FunctionTok{name}\KeywordTok{:}\AttributeTok{ first and last name}
-\FunctionTok{    affil}\KeywordTok{: }\CharTok{|}
-\NormalTok{      One or more comma seperated numbers corresponding to affilitation}
-\NormalTok{      and one or more  comma seperated symbols corresponding }
-\NormalTok{      optional notes.}
-\AttributeTok{    }\FunctionTok{orcid}\KeywordTok{:}\AttributeTok{ optional orcid number}
-\FunctionTok{affiliation}\KeywordTok{:}\AttributeTok{  }
-\AttributeTok{  }\KeywordTok{{-}}\AttributeTok{ }\FunctionTok{num}\KeywordTok{:}\AttributeTok{ 1,..., n for each affiliation}
-\AttributeTok{    }\FunctionTok{address}\KeywordTok{:}\AttributeTok{ required}
-\AttributeTok{    }\FunctionTok{email}\KeywordTok{:}\AttributeTok{ required}
-\FunctionTok{authorcitation}\KeywordTok{: }\CharTok{|}
-\NormalTok{  Lastname, F.}
-\FunctionTok{correspondence}\KeywordTok{: }\CharTok{|}
-\NormalTok{  email@email.com; Tel.: +XX{-}000{-}00{-}0000.}
-\FunctionTok{journal}\KeywordTok{:}\AttributeTok{ notspecified}
-\FunctionTok{type}\KeywordTok{:}\AttributeTok{ article}
-\FunctionTok{status}\KeywordTok{:}\AttributeTok{ submit}
-\end{Highlighting}
-\end{Shaded}
-
-Journal options are in Table \ref{tab:mdpinames}. The \texttt{status}
-variable should generally not be changed by authors. The \texttt{type}
-variable describes the type of of submission and defaults to
-\texttt{article} but can be replaced with any of the ones in Table
-\ref{tab:mdpitype}
-
-\begin{table}
-
-\caption{\label{tab:mdpitype}MDPI article types.}
-\centering
-\begin{tabular}[t]{lll}
-\toprule
-abstract & entry & retraction\\
-addendum & expressionofconcern & review\\
-article & extendedabstract & perspective\\
-book & datadescriptor & protocol\\
-bookreview & editorial & shortnote\\
-\addlinespace
-briefreport & essay & studyprotocol\\
-casereport & erratum & systematicreview\\
-comment & hypothesis & supfile\\
-commentary & interestingimage & technicalnote\\
-communication & obituary & viewpoint\\
-\addlinespace
-conferenceproceedings & opinion & guidelines\\
-correction & projectreport & registeredreport\\
-conferencereport & reply & tutorial\\
-\bottomrule
-\end{tabular}
-\end{table}
-
 \hypertarget{journal-specific-yaml-variables}{%
@@
 variables}\label{journal-specific-yaml-variables}}
-
-\begin{Shaded}
-\begin{Highlighting}[]
-\CommentTok{\# for journal Diversity,}
-\CommentTok{\# add the Life Science Identifier using:}
-\FunctionTok{lsid}\KeywordTok{:}\AttributeTok{ http://zoobank.org/urn:lsid:zoobank.org:act:nnnn}
-
-\CommentTok{\# for journal Applied Sciences}
-\CommentTok{\# add featured application}
-\FunctionTok{featuredapplication}\KeywordTok{: }\CharTok{|}
-\NormalTok{  Authors are encouraged to provide a concise }
-\NormalTok{  description of the specific application or }
-\NormalTok{  a potential application of the work. This }
-\NormalTok{  section is not mandatory.}
-
-\CommentTok{\# for the journal Data}
-\CommentTok{\# add dataset doi and license}
-\FunctionTok{dataset}\KeywordTok{:}\AttributeTok{ https://doi.org/10.1000/182}
-\FunctionTok{datasetlicense}\KeywordTok{:}\AttributeTok{ CC{-}BY{-}4.0}
-
-\CommentTok{\# for the journal Toxins}
-\CommentTok{\# add key contributions}
-\FunctionTok{keycontributions}\KeywordTok{: }\CharTok{|}
-\NormalTok{  The breakthroughs or highlights of the manuscript. }
-\NormalTok{  Authors can write one or two sentences to describe }
-\NormalTok{  the most important part of the paper.}
-
-\CommentTok{\# for the journal Encyclopedia}
-\FunctionTok{encyclopediadef}\KeywordTok{: }\CharTok{|}
-\NormalTok{  For entry manuscripts only: please provide a brief overview}
-\NormalTok{  of the entry title instead of an abstract.}
-\FunctionTok{entrylink}\KeywordTok{:}\AttributeTok{ The Link to this entry published on the encyclopedia platform.}
-
-\CommentTok{\# for the journal Advances in Respiratory Medicine}
-\CommentTok{\# add highlights}
-\FunctionTok{addhighlights}\KeywordTok{: }\CharTok{|}
-\NormalTok{  This is an obligatory section in “Advances in Respiratory Medicine”, }
-\NormalTok{  whose goal is to increase the discoverability and readability of the}
-\NormalTok{  article via search engines and other scholars. Highlights should not }
-\NormalTok{  be a copy of the abstract, but a simple text allowing the reader to }
-\NormalTok{  quickly and simplified find out what the article is about and what can }
-\NormalTok{  be cited from it. Each of these parts should be devoted up to 2\textasciitilde{}bullet }
-\NormalTok{  points.}
-\end{Highlighting}
-\end{Shaded}
 
@@
 
-\informedconsent{Any research article describing a study involving
-humans should contain this statement. Please add
-\texttt{Informed\ consent\ was\ obtained\ from\ all\ subjects\ \ involved\ in\ the\ study.\textquotesingle{}\textquotesingle{}\ OR}Patient
-consent was waived due to REASON (please provide a detailed
-justification).'\,' OR ``Not applicable'\,' for studies not involving
-humans. You might also choose to exclude this statement if the study did
-not involve humans.
-
-Written informed consent for publication must be obtained from
-participating patients who can be identified (including by the patients
-themselves). Please state ``Written informed consent has been obtained
-from the patient(s) to publish this paper'\,' if applicable.}
-
 \dataavailability{We encourage all authors of articles published in MDPI
@@
 %% Optional
-\sampleavailability{Samples of the compounds \ldots\ldots{} are
-available from the authors.}
 
@@
 TLA & Three letter acronym \\
-LD & linear dichroism \\
+SDR & Solar downward radiation \\
 \end{tabular}}

### 2023-11-01 MDPI_article.tex
Commit 2023-11-01 18:14

--- 2023-11-01 MDPI_submition.tex
@@
-%  LaTeX support: latex@mdpi.com
-%  For support, please attach all files needed for compiling as well as the log file, and specify your operating system, LaTeX version, and LaTeX editor.
-
-%=================================================================
-% pandoc conditionals added to preserve backwards compatibility with previous versions of rticles
-
-\documentclass[applsci,article,submit,moreauthors,pdftex]{Definitions/mdpi}
-
-%% Some pieces required from the pandoc template
-\setlist[itemize]{leftmargin=*,labelsep=5.8mm}
-\setlist[enumerate]{leftmargin=*,labelsep=4.9mm}
-
-%--------------------
-% Class Options:
-%--------------------
-
-%---------
-% article
-%---------
-% The default type of manuscript is "article", but can be replaced by:
-% abstract, addendum, article, book, bookreview, briefreport, casereport, comment, commentary, communication, conferenceproceedings, correction, conferencereport, entry, expressionofconcern, extendedabstract, datadescriptor, editorial, essay, erratum, hypothesis, interestingimage, obituary, opinion, projectreport, reply, retraction, review, perspective, protocol, shortnote, studyprotocol, systematicreview, supfile, technicalnote, viewpoint, guidelines, registeredreport, tutorial
-% supfile = supplementary materials
-
-%----------
-% submit
-%----------
-% The class option "submit" will be changed to "accept" by the Editorial Office when the paper is accepted.
-This will only make changes to the frontpage (e.g., the logo of the journal will get visible), the headings, and the copyright information.
-Also, line numbering will be removed.
-Journal info and pagination for accepted papers will also be assigned by the Editorial Office.
-
-%------------------
-% moreauthors
-%------------------
-% If there is only one author the class option oneauthor should be used. Otherwise use the class option moreauthors.
-
-%---------
-% pdftex
-%---------
-% The option pdftex is for use with pdfLaTeX. Remove "pdftex" for (1) compiling with LaTeX & dvi2pdf (if eps figures are used) or for (2) compiling with XeLaTeX.
-
-%=================================================================
-% MDPI internal commands - do not modify
-\firstpage{1}
+% Options for packages loaded elsewhere
+\PassOptionsToPackage{unicode}{hyperref}
+\PassOptionsToPackage{hyphens}{url}
+%
+\documentclass[
+]{article}
+\usepackage{amsmath,amssymb}
+\usepackage{iftex}
+\ifPDFTeX
+  \usepackage[T1]{fontenc}
+  \usepackage[utf8]{inputenc}
+  \usepackage{textcomp} % provide euro and other symbols
+\else % if luatex or xetex
+  \usepackage{unicode-math} % this also loads fontspec
+  \defaultfontfeatures{Scale=MatchLowercase}
+  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
+\fi
+\usepackage{lmodern}
+\ifPDFTeX\else
+  % xetex/luatex font selection
+\fi
+% Use upquote if available, for straight quotes in verbatim environments
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
+\IfFileExists{microtype.sty}{% use microtype if available
+  \usepackage[]{microtype}
+  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
+}{}
 \makeatletter
-\setcounter{page}{\@firstpage}
+\@ifundefined{KOMAClassName}{% if non-KOMA class
+  \IfFileExists{parskip.sty}{%
+    \usepackage{parskip}
+  }{% else
+    \setlength{\parindent}{0pt}
+    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
+}{% if KOMA class
+  \KOMAoptions{parskip=half}}
 \makeatother
-\pubvolume{1}
-\issuenum{1}
-\articlenumber{0}
-\pubyear{2023}
-\copyrightyear{2023}
-%\externaleditor{Academic Editor: Firstname Lastname}
-\datereceived{ }
-\daterevised{ } % Comment out if no revised date
-\dateaccepted{ }
-\datepublished{ }
-%\datecorrected{} % For corrected papers: "Corrected: XXX" date in the original paper.
-%\dateretracted{} % For corrected papers: "Retracted: XXX" date in the original paper.
-\hreflink{https://doi.org/} % If needed use \linebreak
-%\doinum{}
-%\pdfoutput=1 % Uncommented for upload to arXiv.org
-
-%=================================================================
-% Add packages and commands here.
-The following packages are loaded in our class file: fontenc, inputenc, calc, indentfirst, fancyhdr, graphicx, epstopdf, lastpage, ifthen, float, amsmath, amssymb, lineno, setspace, enumitem, mathpazo, booktabs, titlesec, etoolbox, tabto, xcolor, colortbl, soul, multirow, microtype, tikz, totcount, changepage, attrib, upgreek, array, tabularx, pbox, ragged2e, tocloft, marginnote, marginfix, enotez, amsthm, natbib, hyperref, cleveref, scrextend, url, geometry, newfloat, caption, draftwatermark, seqsplit
-% cleveref: load \crefname definitions after \begin{document}
-
-%=================================================================
-% Please use the following mathematics environments: Theorem, Lemma, Corollary, Proposition, Characterization, Property, Problem, Example, ExamplesandDefinitions, Hypothesis, Remark, Definition, Notation, Assumption
-%% For proofs, please use the proof environment (the amsthm package is loaded by the MDPI class).
-
-%=================================================================
-% Full title of the paper (Capitalized)
-\Title{Full title of the paper (Capitalized)}
-
-% MDPI internal command: Title for citation in the left column
-\TitleCitation{Full title of the paper (Capitalized)}
-
-% Author Orchid ID: enter ID or remove command
-%\newcommand{\orcidauthorA}{0000-0000-0000-000X} % Add \orcidA{} behind the author's name
-%\newcommand{\orcidauthorB}{0000-0000-0000-000X} % Add \orcidB{} behind the author's name
-
-% Authors, for the paper (add full first names)
-\Author{Athanasios
-Natsis$^{1,*}$\href{https://orcid.org/0000-0002-5199-4119}
-{\orcidicon}, Alkiviadis Bais$^{2,}$}
-
-%\longauthorlist{yes}
-
-% MDPI internal command: Authors, for metadata in PDF
-\AuthorNames{Athanasios Natsis, Alkiviadis Bais}
-
-% MDPI internal command: Authors, for citation in the left column
-%\AuthorCitation{Lastname, F.; Lastname, F.; Lastname, F.}
-% If this is a Chicago style journal: Lastname, Firstname, Firstname Lastname, and Firstname Lastname.
-\AuthorCitation{Natsis, A.; Bais, A.}
-
-% Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
-\address{%
-$^{1}$ \quad Aristotle University of Thessaloniki - Laboratory of
-Atmospheric Physics, Campus Box 149, 54124 Thessaloniki,
-Greece; \href{mailto:natsisphysicist@gmail.com}{\nolinkurl{natsisphysicist@gmail.com}}\\
-$^{2}$ \quad Your department Street, City, Country; \\
-}
-
-% Contact information of the corresponding author
-\corres{Correspondence: \href{mailto:natsisphysicist@gmail.com}{\nolinkurl{natsisphysicist@gmail.com}};}
-
-% Current address and/or shared authorship
-
-% The commands \thirdnote{} till \eighthnote{} are available for further notes
-
-% Simple summary
-\simplesumm{A Simple summary goes here.}
-
-%\conference{} % An extended version of a conference paper
-
-% Abstract (Do not insert blank lines, i.e. \\)
-\abstract{The shortwave downward solar irradiance (SDR) is an important
-factor that drives climate processes, production and can affect all
-living organisms. While monitoring the long term variability, there are
-observations of upward and downward SDR trends on different locations
-around the world for different time periods. Periods of positive tredns
-are refered as brightening periods and with negative tredns as dimming
-periods. We studied 29 years of CHP1 data from Thessaloniki, Greece,
-under three sky conditions (clear sky, cloudy sky and all sky
-conditions), applying a cloud sky identification algorithm. We found a
-positive trend for all-sky and clear-sky conditions, and also,
-investigated the consistency of those trends, the effect of the solar
-zenith angle, and the variation of the trends for the seasons of the
-year. We indentified that there are some anomalies in the long term SDR
-trends, for all sky conditions.}
-
-% Keywords
-\keyword{GHI; SDR; Solar Brigthening/Dimming (list three to ten
-pertinent keywords specific to the article, yet reasonably common within
-the subject discipline.).}
-
-% The fields PACS, MSC, and JEL may be left empty or commented out if not applicable
-%\PACS{J0101}
-%\MSC{}
-%\JEL{}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Only for the journal Diversity
-%\LSID{\url{http://}}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Only for the journal Applied Sciences
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Only for the journal Data
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Only for the journal Toxins
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Only for the journal Encyclopedia
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Only for the journal Advances in Respiratory Medicine
-%\addhighlights{yes}
-%\renewcommand{\addhighlights}{%
-
-%\noindent This is an obligatory section in “Advances in Respiratory Medicine”, whose goal is to increase the discoverability and readability of the article via search engines and other scholars.
-Highlights should not be a copy of the abstract, but a simple text allowing the reader to quickly and simplified find out what the article is about and what can be cited from it.
-Each of these parts should be devoted up to 2~bullet points.\vspace{3pt}\\
-%\textbf{What are the main findings?}
-% \begin{itemize}[labelsep=2.5mm,topsep=-3pt]
-% \item First bullet.
-% \item Second bullet.
-% \end{itemize}\vspace{3pt}
-%\textbf{What is the implication of the main finding?}
-% \begin{itemize}[labelsep=2.5mm,topsep=-3pt]
-% \item First bullet.
-% \item Second bullet.
-% \end{itemize}
-%}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% tightlist command for lists without linebreak
+\usepackage{xcolor}
+\usepackage[margin=1in]{geometry}
+\usepackage{longtable,booktabs,array}
+\usepackage{calc} % for calculating minipage widths
+% Correct order of tables after \paragraph or \subparagraph
+\usepackage{etoolbox}
+\makeatletter
+\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
+\makeatother
+% Allow footnotes in longtable head/foot
+\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
+\makesavenoteenv{longtable}
+\usepackage{graphicx}
+\makeatletter
+\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
+\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
+\makeatother
+% Scale images if necessary, so that they will not overflow the page
+% margins by default, and it is still possible to overwrite the defaults
+% using explicit options in \includegraphics[width, height, ...]{}
+\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
+% Set default figure placement to htbp
+\makeatletter
+\def\fps@figure{htbp}
+\makeatother
+\setlength{\emergencystretch}{3em} % prevent overfull lines
 \providecommand{\tightlist}{%
   \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+\setcounter{secnumdepth}{5}
+\newlength{\cslhangindent}
+\setlength{\cslhangindent}{1.5em}
+\newlength{\csllabelwidth}
+\setlength{\csllabelwidth}{3em}
+\newlength{\cslentryspacingunit} % times entry-spacing
+\setlength{\cslentryspacingunit}{\parskip}
+\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
+ {% don't indent paragraphs
+  \setlength{\parindent}{0pt}
+  % turn on hanging indent if param 1 is 1
+  \ifodd #1
+  \let\oldpar\par
+  \def\par{\hangindent=\cslhangindent\oldpar}
+  \fi
+  % set entry spacing
+  \setlength{\parskip}{#2\cslentryspacingunit}
+ }%
+ {}
+\usepackage{calc}
+\newcommand{\CSLBlock}[1]{#1\hfill\break}
+\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
+\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
+\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
+\usepackage{booktabs}
+\usepackage{longtable}
+\usepackage{array}
+\usepackage{multirow}
+\usepackage{wrapfig}
+\usepackage{float}
+\usepackage{colortbl}
+\usepackage{pdflscape}
+\usepackage{tabu}
+\usepackage{threeparttable}
+\usepackage{threeparttablex}
+\usepackage[normalem]{ulem}
+\usepackage{makecell}
+\usepackage{xcolor}
+\ifLuaTeX
+  \usepackage{selnolig}  % disable illegal ligatures
+\fi
+\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
+\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
+\urlstyle{same}
+\hypersetup{
+  pdftitle={Full title of the paper (Capitalized)},
+  pdfkeywords={GHI; SDR; Solar Brigthening/Dimming (list three to ten pertinent keywords specific
+to the article, yet reasonably common within the subject discipline.).},
+  hidelinks,
+  pdfcreator={LaTeX via pandoc}}
 
-\usepackage{longtable}
+\title{Full title of the paper (Capitalized)}
+\author{true \and true}
+\date{}
 
 \begin{document}
+\maketitle
+\begin{abstract}
+The shortwave downward solar irradiance (SDR) is an important factor that drives climate processes, production and can affect all living organisms.
+While monitoring the long term variability, there are observations of upward and downward SDR trends on different locations around the world for different time periods.
+Periods of positive tredns are refered as brightening periods and with negative tredns as dimming periods.
+We studied 29 years of CHP1 data from Thessaloniki, Greece, under three sky conditions (clear sky, cloudy sky and all sky conditions), applying a cloud sky identification algorithm.
+We found a positive trend for all-sky and clear-sky conditions, and also, investigated the consistency of those trends, the effect of the solar zenith angle, and the variation of the trends for the seasons of the year.
+We indentified that there are some anomalies in the long term SDR trends, for all sky conditions.
+\end{abstract}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+{
+\setcounter{tocdepth}{2}
+\tableofcontents
+}
+\hypertarget{introduction.}{%
+\section{Introduction.}\label{introduction.}}
+
+The shortwave downward solar irradiance (SDR) at Earth's surface play a significant role, on its climate.
+Changes of the SDR can be related to changes on Earth's energy budget, the mechanisms of climate change, and water and carbon cycle (\textbf{Wild2009?}).
+Can also affect, solar and agricultural production, and all living organisms.
+Studies of SDR variability, have identified some distinct SDR trends on different regions of the world on different time periods.
+The term `brightening' is generally used to describe periods of positive SDR trend, and `dimming' for negative trend.
+There are many cases on the long term records of irradiance, showing a systematic change of SDR's trend slope, occurring roughly at the last decades of the 20th century.
+On multiple station in China, a dimming period was reported until about 2000, followed by a brightening period (\textbf{Yang2021?}).
+A similar pattern was identified, with the breaking point around 1980, for stations in Central Europe (\textbf{Wild2021?}) and Brazil (\textbf{Yamasoe2021?}).
+Also, on global scale, an AI aided continental level spatial analysis, with data from multiple station, reach similar conclusions for the above regions and for the global trend (\textbf{Yuan2021?}).
+
+There is a consensus, among researchers, that the major factors of SDR attenuation is the interaction of Sun radiation with atmospheric aerosols and clouds.
+Those interactions, among other factors, have been analysed with models (\textbf{Li2016?}; \textbf{Samset2018?}), showing the existence of feedback mechanisms between the two.
+Similar finds, have been showed in observational data (\textbf{Schwarz2020?}; \textbf{Ohvril2009?}; \textbf{Zerefos2009?}; \textbf{Xia2007?} and references therein).
+
+Due to the variability of the phenomenon, and its contributing factors, there is a constant need to investigate and monitor SDR, in different sites, to estimate its magnitude, and its relation to the local conditions.
+In this study, we examine the trends of SDR, with ground-based measurements at Thessaloniki, Greece for the period 1993 to 2023, as derived from a CM-21 pyranometer.
+We reevaluated and extended the dataset used by (\textbf{Bais2013?}), applying a different algorithm for the identification of clear-/cloud-sky instances (\textbf{Reno2016?}; \textbf{Reno2012a?}), and we derive the radiation trends for the period 1993 to 2023, under different sky conditions (all-sky, clear-sky and cloud-sky).
 
@@
 This Rmd-skeleton uses the mdpi Latex template published 2023-03-25.
-However, the official template gets more frequently updated than the
-\textbf{rticles} package. Therefore, please make sure prior to paper
-submission, that you're using the most recent .cls, .tex and .bst files
+However, the official template gets more frequently updated than the \textbf{rticles}
+package. Therefore, please make sure prior to paper submission, that you're
+using the most recent .cls, .tex and .bst files
 (available \href{http://www.mdpi.com/authors/latex}{here}).
@@
 \hypertarget{journal-specific-yaml-variables}{%
-\subsection{Journal Specific YAML
-variables}\label{journal-specific-yaml-variables}}
+\subsection{Journal Specific YAML variables}\label{journal-specific-yaml-variables}}
 
@@
 \begin{longtable}[t]{llllllll}
-\caption{\label{tab:mdpinames}MDPI journal names.}\\
+\caption{(\#tab:mdpinames)MDPI journal names.}\\
 \toprule
@@
 
-The introduction should briefly place the study in a broad context and
-highlight why it is important. It should define the purpose of the work
-and its significance. The current state of the research field should be
-reviewed carefully and key publications cited. Please highlight
-controversial and diverging hypotheses when necessary. Finally, briefly
-mention the main aim of the work and highlight the principal
-conclusions. As far as possible, please keep the introduction
-comprehensible to scientists outside your particular field of research.
-Citing a journal paper
-\citep{bertrand-krajewski_distribution_1998, leutnant_stormwater_2016}.
-And now citing a book reference \citet{gujer_systems_2008}. Some MDPI
-journals use Chicago and others use APA, this template should choose the
-correct citation format for you once you specify the journal in the YAML
+The introduction should briefly place the study in a broad context and highlight
+why it is important. It should define the purpose of the work and its
+significance. The current state of the research field should be reviewed
+carefully and key publications cited. Please highlight controversial and
+diverging hypotheses when necessary. Finally, briefly mention the main aim of
+the work and highlight the principal conclusions. As far as possible, please
+keep the introduction comprehensible to scientists outside your particular
+field of research. Citing a journal paper (Bertrand-Krajewski, Chebbo, and Saget 1998; Leutnant, Muschalla, and Uhl 2016). And now citing a book reference Gujer (2008).
+Some MDPI journals use Chicago and others use APA, this template should choose
+the correct citation format for you once you specify the journal in the YAML
 header.
 
-To use endnotes, change \texttt{endnotes:\ true} in the YAML header,
-then use \texttt{\textbackslash{}endnote\{This\ is\ an\ endnote.\}}.
+To use endnotes, change \texttt{endnotes:\ true} in the YAML header, then use
+\texttt{\textbackslash{}endnote\{This\ is\ an\ endnote.\}}.
 
@@
 
-Materials and Methods should be described with sufficient details to
-allow others to replicate and build on published results. Please note
-that publication of your manuscript implicates that you must make all
-materials, data, computer code, and protocols associated with the
-publication available to readers. Please disclose at the submission
-stage any restrictions on the availability of materials or information.
-New methods and protocols should be described in detail while
-well-established methods can be briefly described and appropriately
+Materials and Methods should be described with sufficient details to allow
+others to replicate and build on published results. Please note that publication
+of your manuscript implicates that you must make all materials, data, computer
+code, and protocols associated with the publication available to readers. Please
+disclose at the submission stage any restrictions on the availability of
+materials or information. New methods and protocols should be described in
+detail while well-established methods can be briefly described and appropriately
 cited.
 
-Research manuscripts reporting large datasets that are deposited in a
-publicly available database should specify where the data have been
-deposited and provide the relevant accession numbers. If the accession
-numbers have not yet been obtained at the time of submission, please
-state that they will be provided during review. They must be provided
-prior to publication.
+Research manuscripts reporting large datasets that are deposited in a publicly
+available database should specify where the data have been deposited and provide
+the relevant accession numbers. If the accession numbers have not yet been
+obtained at the time of submission, please state that they will be provided
+during review. They must be provided prior to publication.
 
-Interventionary studies involving animals or humans, and other studies
-require ethical approval must list the authority that provided approval
-and the corresponding ethical approval code.
+Interventionary studies involving animals or humans, and other studies require
+ethical approval must list the authority that provided approval and the
+corresponding ethical approval code.
 
@@
 
-This section may be divided by subheadings. It should provide a concise
-and precise description of the experimental results, their
-interpretation as well as the experimental conclusions that can be
-drawn.
+This section may be divided by subheadings. It should provide a concise and
+precise description of the experimental results, their interpretation as well
+as the experimental conclusions that can be drawn.
 
@@
 \hypertarget{subsubsection-heading-here}{%
-\subsubsection{Subsubsection Heading
-Here}\label{subsubsection-heading-here}}
+\subsubsection{Subsubsection Heading Here}\label{subsubsection-heading-here}}
 
@@
 \hypertarget{figures-tables-and-schemes}{%
-\subsection{Figures, Tables and
-Schemes}\label{figures-tables-and-schemes}}
+\subsection{Figures, Tables and Schemes}\label{figures-tables-and-schemes}}
 
-All figures and tables should be cited in the main text as Figure
-\ref{fig:fig1}, \ref{tab:tab1}, etc. To get cross-reference to figure
-generated by R chunks include the
-\texttt{\textbackslash{}\textbackslash{}label\{\}} tag in the
-\texttt{fig.cap} attribute of the R chunk:
+All figures and tables should be cited in the main text as Figure \ref{fig:fig1},
+\ref{tab:tab1}, etc. To get cross-reference to figure generated by R chunks
+include the \texttt{\textbackslash{}\textbackslash{}label\{\}} tag in the \texttt{fig.cap} attribute of the R chunk:
 \texttt{fig.cap\ =\ "Fancy\ Caption\textbackslash{}\textbackslash{}label\{fig:plot\}"}.
 
-\begin{figure}
-\centering
-\includegraphics{MDPI_submition_files/figure-latex/fig1-1.pdf}
-\caption{A figure added with a code chunk.\label{fig:fig1}}
+\begin{figure}[h!]
+
+{\centering \includegraphics[width=0.7\linewidth]{MDPI_article_files/figure-latex/fig1-1}
+
+}
+
+\caption{A figure added with a code chunk.\label{fig:fig1}}(\#fig:fig1)
 \end{figure}
 
-When making tables using \texttt{kable}, it is suggested to use the
-\texttt{format="latex"} and \texttt{tabl.envir="table"} arguments to
-ensure table numbering and compatibility with the mdpi document class.
+When making tables using \texttt{kable}, it is suggested to use
+the \texttt{format="latex"} and \texttt{tabl.envir="table"} arguments
+to ensure table numbering and compatibility with the mdpi
+document class.
 
@@
 
-\caption{\label{tab:tab1}This is a table caption. Tables should be placed in the 
+\caption{(\#tab:tab1)This is a table caption. Tables should be placed in the 
              main text near to the first time they are~cited.}
@@
 
-\caption{\label{tab:tab2}This is a very wide table}
+\caption{(\#tab:tab2)This is a very wide table}
 \begin{tabular}[t]{cccc}
@@
 \hypertarget{formatting-of-mathematical-components}{%
-\subsection{Formatting of Mathematical
-Components}\label{formatting-of-mathematical-components}}
+\subsection{Formatting of Mathematical Components}\label{formatting-of-mathematical-components}}
 
@@
 
-If you want numbered equations use Latex and wrap in the equation
-environment:
+If you want numbered equations use Latex and wrap in the equation environment:
 
@@
 
-the text following an equation need not be a new paragraph. Please
-punctuate equations as regular text.
+the text following an equation need not be a new paragraph. Please punctuate
+equations as regular text.
 
@@
 
-Theorem-type environments (including propositions, lemmas, corollaries
-etc.) can be formatted as follows:
+Theorem-type environments (including propositions, lemmas, corollaries etc.)
+can be formatted as follows:
 
@@
 
-\begin{proof}[Proof of Theorem1]
-Text of the proof. Note that the phrase ``of Theorem 1'\,' is optional
-if it is clear which theorem is being referred to.
-
+\begin{proof}
+Text of the proof. Note that the phrase ``of Theorem 1'\,' is optional if it is
+clear which theorem is being referred to.
 \end{proof}
@@
 Authors should discuss the results and how they can be interpreted in
-perspective of previous studies and of the working hypotheses. The
-findings and their implications should be discussed in the broadest
-context possible. Future research directions may also be highlighted.
+perspective of previous studies and of the working hypotheses. The findings and
+their implications should be discussed in the broadest context possible. Future
+research directions may also be highlighted.
 
@@
 
-This section is not mandatory, but may be added if there are patents
-resulting from the work reported in this manuscript.
+This section is not mandatory, but may be added if there are patents resulting
+from the work reported in this manuscript.
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\hypertarget{refs}{}
+\begin{CSLReferences}{1}{0}
+\leavevmode\vadjust pre{\hypertarget{ref-bertrand-krajewski_distribution_1998}{}}%
+Bertrand-Krajewski, J.
+L., G.
+Chebbo, and A.
+Saget.
+1998.
+{``Distribution of Pollutant Mass Vs Volume in Stormwater Discharges and the First Flush Phenomenon.''} \emph{Water Research} 32 (8): 2341--56.
+\url{http://www.scopus.com/inward/record.url?eid=2-s2.0-0032146023\&partnerID=40\&md5=d6517b6efa014bc12f3ae70abe71977d}.
 
-\vspace{6pt}
+\leavevmode\vadjust pre{\hypertarget{ref-gujer_systems_2008}{}}%
+Gujer, W. 2008. \emph{Systems {Analysis} for {Water} {Technology}}. Berlin, Heidelberg, Germany: Springer-Verlag.
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% optional
-\supplementary{The following supporting information can be downloaded
-at:\\
-\linksupplementary{s1}, Figure S1: title; Table S1: title; Video S1:
-title.}
+\leavevmode\vadjust pre{\hypertarget{ref-leutnant_stormwater_2016}{}}%
+Leutnant, Dominik, Dirk Muschalla, and Mathias Uhl.
+2016.
+{``Stormwater {Pollutant} {Process} {Analysis} with {Long}-{Term} {Online} {Monitoring} {Data} at {Micro}-{Scale} {Sites}.''} \emph{Water} 8 (7): 299.
+\url{https://doi.org/10.3390/w8070299}.
 
-% Only for the journal Methods and Protocols:
-% If you wish to submit a video article, please do so with any other supplementary material.
-% \supplementary{The following supporting information can be downloaded at: \linksupplementary{s1}, Figure S1: title; Table S1: title; Video S1: title.
-A supporting video article is available at doi: link.}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\authorcontributions{For research articles with several authors, a short
-paragraph specifying their individual contributions must be provided.
-The following statements should be used ``X.X. and Y.Y. conceive and
-designed the experiments; X.X. performed the experiments; X.X. and Y.Y.
-analyzed the data; W.W. contributed reagents/materials/analysis tools;
-Y.Y. wrote the paper.'\,' Authorship must be limited to those who have
-contributed substantially to the work reported.}
-
-\funding{Please add:
-\texttt{This\ research\ received\ no\ external\ funding\textquotesingle{}\textquotesingle{}\ or}This
-research was funded by NAME OF FUNDER grant number XXX.'\,' and and
-``The APC was funded by XXX'\,'. Check carefully that the details given
-are accurate and use the standard spelling of funding agency names at
-\url{https://search.crossref.org/funding}, any errors may affect your
-future funding.}
-
-\institutionalreview{In this section, you should add the Institutional
-Review Board Statement and approval number, if relevant to your study.
-You might choose to exclude this statement if the study did not require
-ethical approval. Please note that the Editorial Office might ask you
-for further information. Please add ``The study was conducted in
-accordance with the Declaration of Helsinki, and approved by the
-Institutional Review Board (or Ethics Committee) of NAME OF INSTITUTE
-(protocol code XXX and date of approval).'' for studies involving
-humans. OR ``The animal study protocol was approved by the Institutional
-Review Board (or Ethics Committee) of NAME OF INSTITUTE (protocol code
-XXX and date of approval).'' for studies involving animals. OR ``Ethical
-review and approval were waived for this study due to REASON (please
-provide a detailed justification).'' OR ``Not applicable'' for studies
-not involving humans or animals.}
-
-\dataavailability{We encourage all authors of articles published in MDPI
-journals to share their research data. In this section, please provide
-details regarding where data supporting reported results can be found,
-including links to publicly archived datasets analyzed or generated
-during the study. Where no new data were created, or where data is
-unavailable due to privacy or ethical re-strictions, a statement is
-still required. Suggested Data Availability Statements are available in
-section ``MDPI Research Data Policies'' at
-\url{https://www.mdpi.com/ethics}.}
-
-\acknowledgments{All sources of funding of the study should be
-disclosed. Please clearly indicate grants that you have received in
-support of your research work. Clearly state if you received funds for
-covering the costs to publish in open access.}
-
-\conflictsofinterest{Declare conflicts of interest or state `The authors
-declare no conflict of interest.' Authors must identify and declare any
-personal circumstances or interest that may be perceived as
-inappropriately influencing the representation or interpretation of
-reported research results. Any role of the funding sponsors in the
-design of the study; in the collection, analyses or interpretation of
-data in the writing of the manuscript, or in the decision to publish the
-results must be declared in this section. If there is no role, please
-state `The founding sponsors had no role in the design of the study; in
-the collection, analyses, or interpretation of data; in the writing of
-the manuscript, an in the decision to publish the results'.}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Optional
-
-%% Only for journal Encyclopedia
-
-\abbreviations{Abbreviations}{
-The following abbreviations are used in this manuscript:\\
-
-\noindent
-\begin{tabular}{@{}ll}
-MDPI & Multidisciplinary Digital Publishing Institute \\
-DOAJ & Directory of open access journals \\
-TLA & Three letter acronym \\
-SDR & Solar downward radiation \\
-\end{tabular}}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Optional
-\input{"appendix.tex"}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{adjustwidth}{-\extralength}{0cm}
-
-%\printendnotes[custom] % Un-comment to print a list of endnotes
-
-\reftitle{References}
-\bibliography{mybibfile.bib}
-
-% If authors have biography, please use the format below
-%\section*{Short Biography of Authors}
-%\bio
-%{\raisebox{-0.35cm}{\includegraphics[width=3.5cm,height=5.3cm,clip,keepaspectratio]{Definitions/author1.pdf}}}
-%{\textbf{Firstname Lastname} Biography of first author}
-%
-%\bio
-%{\raisebox{-0.35cm}{\includegraphics[width=3.5cm,height=5.3cm,clip,keepaspectratio]{Definitions/author2.jpg}}}
-%{\textbf{Firstname Lastname} Biography of second author}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% for journal Sci
-%\reviewreports{\\
-%Reviewer 1 comments and authors’ response\\
-%Reviewer 2 comments and authors’ response\\
-%Reviewer 3 comments and authors’ response
-%}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\PublishersNote{}
-\end{adjustwidth}
+\end{CSLReferences}
 

### 2023-11-01 MDPI_submition.Rmd
Commit 2023-11-01 18:14

@@
     long: Solar downward radiation
-bibliography: mybibfile.bib
+bibliography: references.bib
 appendix:     appendix.tex
@@
 ---
+
+```{r echo=F, include=T, warning=FALSE, message=FALSE}
+suppressMessages(
+    res <- rmarkdown::find_pandoc(dir = '/usr/lib/rstudio/resources/app/bin/quarto/bin/tools')
+)
+## document configurations 
+knitr::opts_chunk$set(comment    = ""      )
+# knitr::opts_chunk$set(dev        = "pdf"   )
+knitr::opts_chunk$set(out.width  = "70%"   )
+knitr::opts_chunk$set(fig.align  = "center")
+knitr::opts_chunk$set(fig.pos    = 'h!'    )
+options(knitr.graphics.auto_pdf = FALSE)
+library(pander)
+library(data.table)
+library(kableExtra)
+library(dplyr)
+## load data from trends analysis
+load("./data/common_data_14_2.Rda")
+load("./data/Input_1_longterm_trends.Rda")
+## number of measurements for each dataset
+SDR_N_All   <- sum(  ALL_1_daily_DESEAS[,GLB_att_N])
+SDR_N_Clear <- sum(CLEAR_1_daily_DESEAS[,GLB_att_N])
+SDR_N_Cloud <- sum(CLOUD_1_daily_DESEAS[,GLB_att_N])
+## TSI data span info
+TSIinfo <- data.table(read.csv("./figures/tbl_tsi_info.dat",
+                    strip.white = TRUE,
+                    sep = ";",
+                    header = TRUE))
+## variables we use in the analysis
+source("./DHI_GHI_0_variables.R")
+
+## create local citations file and links here
+source("~/CODE/FUNCTIONS/R/gather_citations.R")
+suppressMessages(
+    gather_citations(
+        sourcefile   = "./Article.Rmd",
+        mainreposit  = "./Mainreferences.bib",
+        localreposit = "./references.bib",
+        folderlink   = "./References_links",
+        workingdir   = "./"
+    )
+)
+```
+
+# Introduction.
+
+<!--
+Η εισαγωγή είναι εντάξει. Ίσως θα μπορούσες να βάλεις στη 2η
+παράγραφο ακόμη 2-3 σειρές για τις τάσεις του αεροζόλ και των νεφών
+από τις εργασίες και να καταλήξεις με κάποια στοιχεία για το αεροζόλ
+στη Θεσσαλονίκη από εργασίες δικές μας. Για τα νέφη δεν έχω υπόψη μου
+καμία εργασία αλλά ίσως θα μπορούσε να ειπωθεί κάτι χρησιμοποιώντας
+παγκόσμιες εκτιμήσεις για την περιοχή της Μεσογείου.
+π.χ. https://link.springer.com/article/10.1007/s00382-022-06438-3
+-->
+
+The shortwave downward solar irradiance (SDR) at Earth's surface play a significant role, on its climate.
+Changes of the SDR can be related to changes on Earth's energy budget, the mechanisms of climate change, and water and carbon cycle [@Wild2009].
+Can also affect, solar and agricultural production, and all living organisms.
+Studies of SDR variability, have identified some distinct SDR trends on different regions of the world on different time periods.
+The term 'brightening' is generally used to describe periods of positive SDR trend, and 'dimming' for negative trend.
+There are many cases on the long term records of irradiance, showing a systematic change of SDR's trend slope, occurring roughly at the last decades of the 20th century.
+<!-- Trends china -->
+On multiple station in China, a dimming period was reported until about 2000, followed by a brightening period [@Yang2021]. 
+<!-- Trends Europe Brazil -->
+A similar pattern was identified, with the breaking point around 1980, for stations in Central Europe [@Wild2021] and Brazil [@Yamasoe2021].
+<!-- Trends Global -->
+Also, on global scale, an AI aided continental level spatial analysis, with data from multiple station, reach similar conclusions for the above regions and for the global trend [@Yuan2021].
+
+There is a consensus, among researchers, that the major factors of SDR attenuation is the interaction of Sun radiation with atmospheric aerosols and clouds. 
+Those interactions, among other factors, have been analysed with models [@Li2016; @Samset2018], showing the existence of feedback mechanisms between the two.
+Similar finds, have been showed in observational data [@Schwarz2020; @Ohvril2009; @Zerefos2009; @Xia2007 and references therein].
+<!-- Sun variability: should cite that there is no significant trend found Wild2021 -->
+
+Due to the variability of the phenomenon, and its contributing factors, there is a constant need to investigate and monitor SDR, in different sites, to estimate its magnitude, and its relation to the local conditions.
+In this study, we examine the trends of SDR, with ground-based measurements at Thessaloniki, Greece for the period `r year(min(ALL_1_daily_DESEAS$Date))` to `r year(max(ALL_1_daily_DESEAS$Date))`, as derived from a CM-21 pyranometer.
+<!-- goals -->
+We reevaluated and extended the dataset used by @Bais2013, applying a different algorithm for the identification of clear-/cloud-sky instances [@Reno2016; @Reno2012a], and we derive the radiation trends for the period `r year(min(ALL_1_daily_DESEAS$Date))` to `r year(max(ALL_1_daily_DESEAS$Date))`, under different sky conditions (all-sky, clear-sky and cloud-sky).
 

### 2023-11-01 MDPI_submition.tex
Commit 2023-11-01 18:14

@@
 \usepackage{longtable}
+\usepackage{booktabs}
+\usepackage{array}
+\usepackage{multirow}
+\usepackage{wrapfig}
+\usepackage{float}
+\usepackage{colortbl}
+\usepackage{pdflscape}
+\usepackage{tabu}
+\usepackage{threeparttable}
+\usepackage{threeparttablex}
+\usepackage[normalem]{ulem}
+\usepackage{makecell}
+\usepackage{xcolor}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\hypertarget{introduction.}{%
+\section{Introduction.}\label{introduction.}}
+
+The shortwave downward solar irradiance (SDR) at Earth's surface play a
+significant role, on its climate. Changes of the SDR can be related to
+changes on Earth's energy budget, the mechanisms of climate change, and
+water and carbon cycle \citep{Wild2009}. Can also affect, solar and
+agricultural production, and all living organisms. Studies of SDR
+variability, have identified some distinct SDR trends on different
+regions of the world on different time periods. The term `brightening'
+is generally used to describe periods of positive SDR trend, and
+`dimming' for negative trend. There are many cases on the long term
+records of irradiance, showing a systematic change of SDR's trend slope,
+occurring roughly at the last decades of the 20th century. On multiple
+station in China, a dimming period was reported until about 2000,
+followed by a brightening period \citep{Yang2021}. A similar pattern was
+identified, with the breaking point around 1980, for stations in Central
+Europe \citep{Wild2021} and Brazil \citep{Yamasoe2021}. Also, on global
+scale, an AI aided continental level spatial analysis, with data from
+multiple station, reach similar conclusions for the above regions and
+for the global trend \citep{Yuan2021}.
+
+There is a consensus, among researchers, that the major factors of SDR
+attenuation is the interaction of Sun radiation with atmospheric
+aerosols and clouds. Those interactions, among other factors, have been
+analysed with models \citep{Li2016, Samset2018}, showing the existence
+of feedback mechanisms between the two. Similar finds, have been showed
+in observational data \citep[ and references
+therein]{Schwarz2020, Ohvril2009, Zerefos2009, Xia2007}.
+
+Due to the variability of the phenomenon, and its contributing factors,
+there is a constant need to investigate and monitor SDR, in different
+sites, to estimate its magnitude, and its relation to the local
+conditions. In this study, we examine the trends of SDR, with
+ground-based measurements at Thessaloniki, Greece for the period 1993 to
+2023, as derived from a CM-21 pyranometer. We reevaluated and extended
+the dataset used by \citet{Bais2013}, applying a different algorithm for
+the identification of clear-/cloud-sky instances
+\citep{Reno2016, Reno2012a}, and we derive the radiation trends for the
+period 1993 to 2023, under different sky conditions (all-sky, clear-sky
+and cloud-sky).
 
@@
 
-\begin{figure}
-\centering
-\includegraphics{MDPI_submition_files/figure-latex/fig1-1.pdf}
-\caption{A figure added with a code chunk.\label{fig:fig1}}
+\begin{figure}[h!]
+
+{\centering \includegraphics[width=0.7\linewidth]{MDPI_submition_files/figure-latex/fig1-1}
+
+}
+
+\caption{A figure added with a code chunk.\label{fig:fig1}}\label{fig:fig1}
 \end{figure}
@@
 \reftitle{References}
-\bibliography{mybibfile.bib}
+\bibliography{references.bib}
 

### 2023-11-01 Article.Rmd
Commit 2023-11-01 22:14

@@
 \end{equation}
-
 where: $n$ is the total number of daylight records, $\text{SDR}_{\text{CSid},i}$ are the records identified as clear sky by CSid, $a$ is a hypothetical adjustment factor, and $\text{SDR}_{\text{testCSref},i}$ is any of the tested clear sky radiation models.
@@
 \end{equation}
-
 where: $\text{SDR}_\text{CSref}$ is the reference clear sky SDR, in $\text{w}\,\text{m}^{-2}$ and $\theta$ is the solar zenith angle (SZA).

### 2023-11-01 MDPI_article.tex
Commit 2023-11-01 22:14

@@
 The shortwave downward solar irradiance (SDR) at Earth's surface play a significant role, on its climate.
-Changes of the SDR can be related to changes on Earth's energy budget, the mechanisms of climate change, and water and carbon cycle (\textbf{Wild2009?}).
+Changes of the SDR can be related to changes on Earth's energy budget, the mechanisms of climate change, and water and carbon cycle (Wild 2009).
 Can also affect, solar and agricultural production, and all living organisms.
@@
 There are many cases on the long term records of irradiance, showing a systematic change of SDR's trend slope, occurring roughly at the last decades of the 20th century.
-On multiple station in China, a dimming period was reported until about 2000, followed by a brightening period (\textbf{Yang2021?}).
-A similar pattern was identified, with the breaking point around 1980, for stations in Central Europe (\textbf{Wild2021?}) and Brazil (\textbf{Yamasoe2021?}).
-Also, on global scale, an AI aided continental level spatial analysis, with data from multiple station, reach similar conclusions for the above regions and for the global trend (\textbf{Yuan2021?}).
+On multiple station in China, a dimming period was reported until about 2000, followed by a brightening period (Yang et al. 2021).
+A similar pattern was identified, with the breaking point around 1980, for stations in Central Europe (Wild et al. 2021) and Brazil (Yamasoe et al. 2021).
+Also, on global scale, an AI aided continental level spatial analysis, with data from multiple station, reach similar conclusions for the above regions and for the global trend (Yuan, Leirvik, and Wild 2021).
 
 There is a consensus, among researchers, that the major factors of SDR attenuation is the interaction of Sun radiation with atmospheric aerosols and clouds.
-Those interactions, among other factors, have been analysed with models (\textbf{Li2016?}; \textbf{Samset2018?}), showing the existence of feedback mechanisms between the two.
-Similar finds, have been showed in observational data (\textbf{Schwarz2020?}; \textbf{Ohvril2009?}; \textbf{Zerefos2009?}; \textbf{Xia2007?} and references therein).
+Those interactions, among other factors, have been analysed with models (Li et al. 2016; Samset et al. 2018), showing the existence of feedback mechanisms between the two.
+Similar finds, have been showed in observational data (Schwarz et al. 2020; Ohvril et al. 2009; Zerefos et al. 2009; Xia et al. 2007 and references therein).
 
@@
 In this study, we examine the trends of SDR, with ground-based measurements at Thessaloniki, Greece for the period 1993 to 2023, as derived from a CM-21 pyranometer.
-We reevaluated and extended the dataset used by (\textbf{Bais2013?}), applying a different algorithm for the identification of clear-/cloud-sky instances (\textbf{Reno2016?}; \textbf{Reno2012a?}), and we derive the radiation trends for the period 1993 to 2023, under different sky conditions (all-sky, clear-sky and cloud-sky).
+We reevaluated and extended the dataset used by Bais et al.
+(2013), applying a different algorithm for the identification of clear-/cloud-sky instances (Reno and Hansen 2016; Reno, Hansen, and Stein 2012a), and we derive the radiation trends for the period 1993 to 2023, under different sky conditions (all-sky, clear-sky and cloud-sky).
+
+\hypertarget{observational-data-and-methodology.}{%
+\section{Observational data and methodology.}\label{observational-data-and-methodology.}}
+
+The SDR data were measured with a Kipp \& Zonen CM-21 pyranometer operating continuously at the Laboratory of Atmospheric Physics of the Aristotle University of Thessaloniki
+(\(40^\circ\,38'\,\)N, \(22^\circ\,57'\,\)E, \(80\,\)m~a.s.l.)
+in the period from
+1993-04-13
+to
+2023-04-13.
+The monitoring site is located near the city centre, and we expect to be affected by the urban environment.
+During the study period, the pyranometer has been independently calibrated three times at the Meteorologisches Observatorium Lindenberg, DWD, when it was verified the stability of the instrument to within better than \(0.7\%\) relative to the initial calibration by the manufacturer.
+Along with SDR, the direct beam radiation (DNI) was also measured by a collocated Kipp \& Zonen CHP-1 pyrheliometer, for the period
+2016-04-01
+to
+2023-04-13.
+Although, we have performed a similar analysis to the DNI data the results are not presented here, as they lack the appropriate statistical significance, due to the sorter duration of the data.
+However, the DNI data were used as auxiliary data, in the clear sky identification algorithm (CSid), which is discussed later, for the selection of the appropriate thresholds.
+It is noted that despite the capability of the CSid algorithm to use the DNI as a characterization parameter, we haven't utilized it here, to avoid any selection bias, due to unequal length of the two datasets.
+There are four distinct steps in the creation of the dataset analysed here:
+a)~the acquisition of radiation measurements from the sensors,
+b)~the data quality check,
+c)~the identification of ``clear sky'' conditions from the radiometric data, and
+d)~the aggregation of data and trend analysis.
+
+For the acquisition of radiometric data, the signal of the pyranometer is sampled with a rate of \(1\,\text{Hz}\).
+The mean and the standard deviation of these samples are recorded every minute.
+The measurements are corrected for the zero offset (``dark signal'' in volts).
+The ``dark signal'' is calculated by averaging all measurements recorded for a period of
+\(3\,\text{h}\),
+before (morning) or after (evening) the Sun reaches an elevation angle of
+\(-10^\circ\).
+The signal is converted to irradiance using a ramped value of the instrument's sensitivity between calibrations.
+
+A manual screening was performed, to remove inconsistent and erroneous recordings that can occur stochastically or systematically, during the continuous operation of the instruments.
+The manual screening is aided by a radiation data quality assurance procedure, adjusted for the site, which is based on the methods of
+Long and Shi~(2008, 2006).
+Thus, problematic recordings have been excluded from further processing.
+Although it is impossible to detect all false data, the large number of available data, and the aggregation scheme we used, ensures the good quality of the radiometric measurements used in this study.
+
+In order to be able to estimate the effect of the sky condition on the long term variability of SDR, we created three datasets, by characterizing each one-minute measurement with a corresponding sky condition (i.e., all-sky, clear-sky and cloudy-sky).
+To identify the clear-sky conditions we used a method proposed by
+Long and Ackerman (2000)
+and by
+Reno and Hansen (2016),
+which was adapted and configured for the site, as the authors suggest.
+
+We have to note, that the definition of what constitutes as clear or cloudy sky, has some subjectivity, in any method of characterization.
+As a result, the details of the definition are site specific, it relies on a combination of thresholds and comparisons with ideal actinometric models and statistical analysis on different signal metrics.
+The CSid algorithm was calibrated with the main focus, to identify the presence of clouds on the sky dome.
+Although the fine-tuning of the procedure, few marginal cases exist, that have been identified manually as false positive or false negative but cannot affect the final results of the study.
+
+For completeness, we will provide below a brief overview of the
+clear sky identification algorithm (CSid),
+along with the site specific thresholds.
+To calculate the reference clear sky
+\(\text{SDR}_\text{CSref}\) we used the \(\text{SDR}_\text{Haurwitz}\) derived by
+the radiation model of Haurwitz (1945), adjusted for our site with a
+factor \(a\) (Eq.~\ref{eq:ahau}), resulted by an iterative optimization process, as described
+by Long and Ackerman (2000) and Reno and Hansen (2016).
+The target of
+the optimization was the minimization of a function \(f(a)\) (Eq.~\ref{eq:minf}) and
+was accomplished with the algorithmic function ``optimise'', which is an implementation based on the work of Brent (1973), from the library ``stats'' of the R programming language (R Core Team 2023).
+The optimization and the selection of the clear sky reference model, was performed on SDR observations for the period 2016 - 2021.
+During the optimization, eight simple clear sky radiation models were tested (Daneshyar-Paltridge-Proctor, Kasten-Czeplak, Haurwitz, Berger-Duffie, Adnot-Bourges-Campana-Gicquel, Robledo-Soler, Kasten and Ineichen-Perez), with a wide range of factors.
+These models are described in more details by Reno, Hansen, and Stein (2012b) and evaluated by Reno and Hansen (2016).
+We found, that Haurwitz's model, adjusted with the factor \(a = 0.965\) yields one of the lowest root mean squared errors (RMSE),
+while the procedure, manages to characterize the majority of the data.
+Thus, our clear sky reference is derived by the Eq.~\ref{eq:ahau}.
+
+\begin{equation}
+f(a) = \frac{1}{n}\sum_{i=1}^{n} ( \text{SDR}_{\text{CSid},i} - a \times \text{SDR}_{\text{testCSref},i} )^2 \label{eq:minf}
+\end{equation}
+where: \(n\) is the total number of daylight records, \(\text{SDR}_{\text{CSid},i}\) are the records identified as clear sky by CSid, \(a\) is a hypothetical adjustment factor, and \(\text{SDR}_{\text{testCSref},i}\) is any of the tested clear sky radiation models.
+
+\begin{equation}
+\text{SDR}_\text{CSref} = a \times \text{SDR}_\text{Haurwitz} = 0.965 \times 1098 \times \cos(\theta) \times \exp \left( \frac{ - 0.057}{\cos(\theta)} \right) \label{eq:ahau}
+\end{equation}
+where: \(\text{SDR}_\text{CSref}\) is the reference clear sky SDR, in \(\text{w}\,\text{m}^{-2}\) and \(\theta\) is the solar zenith angle (SZA).
+
+The criteria that were used to identify whether a measurement was taken
+under clear-sky conditions are presented below.
+A data point is flagged
+as ``clear-sky'' if all criteria are satisfied, otherwise it is considered to be ``cloud-sky''.
+Each criterion was applied
+for a running window of \(11\) consecutive one-minute measurements, and
+the characterization is assigned to the central value of the window.
+Each parameter, was calculated both from the observations and the
+reference clear sky model, for each comparison.
+The allowable range of variation is defined by the
+model-derived value of the parameter multiplied by a factor plus an
+offset.
+The factors and the offsets were
+determined empirically, by manual inspecting each filters performance on
+selected days, and adjusting them accordingly, during an iterative
+process.
+
+\begin{enumerate}
+\def\labelenumi{\alph{enumi})}
+\tightlist
+\item
+  Mean of the measured \(\overline{\text{SDR}}_i\) (Eq. \ref{eq:MeanVIP}).
+  \begin{equation}
+  0.91 \times \overline{\text{SDR}}_{\text{CSref},i} - 20
+  < \overline{\text{SDR}}_i <
+  1.095 \times \overline{\text{SDR}}_{\text{CSref},i} + 30
+  \label{eq:MeanVIP}
+  \end{equation}
+\end{enumerate}
+
+\begin{enumerate}
+\def\labelenumi{\alph{enumi})}
+\setcounter{enumi}{1}
+\tightlist
+\item
+  Maximum measured value \(M_{\text{}}\) (Eq.~\ref{eq:MaxVIP}).
+  \begin{equation}
+  1 \times M_{\text{CSref},i} - 75
+  < M_{\text{}i} <
+  1 \times M_{\text{CSref},i} + 75
+  \label{eq:MaxVIP}
+  \end{equation}
+\end{enumerate}
+
+\begin{enumerate}
+\def\labelenumi{\alph{enumi})}
+\setcounter{enumi}{2}
+\tightlist
+\item
+  Length \(L_i\) of the sequential line segments, connecting the points of the \(11\) SDR values (Eq. \ref{eq:VILeq}).
+  \begin{equation}
+  L_i = \sum_{i=1}^{n-1}\sqrt{\left ( \text{SDR}_{i+1} - \text{SDR}_{i}\right )^2 + \left ( t_{i+1} - t_i \right )^2}
+  \label{eq:VILeq}
+  \end{equation}
+  \begin{equation}
+  1 \times L_{\text{CSref},i} - 5 < L_i < 1.3 \times L_{\text{CSref},i} + 13
+  \label{eq:VILcr}
+  \end{equation}
+  where: \(t_i\) is the time each SDR measurement has been measured
+\end{enumerate}
+
+\begin{enumerate}
+\def\labelenumi{\alph{enumi})}
+\setcounter{enumi}{3}
+\tightlist
+\item
+  Standard deviation \(\sigma_i\) of the slope (\(s_i\)) between the \(11\) sequential points, normalized by the mean \(\overline{\text{SDR}}_i\) (Eq.~\ref{eq:VCT1}).
+  \begin{gather}
+    \sigma_i = \frac {\sqrt{\frac{1}{n-1} \sum_{i=1}^{n-1} \left( s_i - \bar{s} \right)^2}} {\overline{\text{SDR}}_i} \label{eq:VCT1} \\
+    s_i = \frac{\text{SDR}_{i+1} - \text{SDR}_{i}}{t_{i+1} - t_i},\;\;   \bar{s} = \frac{1}{n-1} \sum_{i=1}^{n-1} s_i,\;\;\forall i \in \left \{ 1, 2, \ldots, n-1 \right \}\;\;
+  \end{gather}
+  For this criterion, \(\sigma_i\) should be below a certain threshold (Eq.~\ref{eq:VCTcr}):
+  \begin{equation}
+    \sigma_i < \ensuremath{1.1\times 10^{-4}} \label{eq:VCTcr}
+  \end{equation}
+\end{enumerate}
+
+\begin{enumerate}
+\def\labelenumi{\alph{enumi})}
+\setcounter{enumi}{4}
+\tightlist
+\item
+  Maximum difference \(X_i\) between the change in measured irradiance and the change in clear sky irradiance over each measurement interval.
+  \begin{gather}
+    X_i = \max{\left \{ \left | x_i - x_{\text{CSref},i} \right | \right \}} \label{eq:VSM3} \\
+    x_i = \text{SDR}_{i+1} - \text{SDR}_{i} \forall i \in \left \{ 1, 2, \ldots, n-1 \right \} \label{eq:VSM1} \\
+    x_{\text{CSref},i} = \text{SDR}_{\text{CSref},i+1} - \text{SDR}_{\text{CSref},i} \forall i \in \left \{ 1, 2, \ldots, n-1 \right \} \label{eq:VSM2}
+  \end{gather}
+  For this criterion, \(X_i\) should be below a certain threshold (Eq.~\ref{eq:VSMcr}):
+  \begin{equation}
+    X_i < 7.5 \label{eq:VSMcr}
+  \end{equation}
+\end{enumerate}
+
+Due to a significant measurement uncertainty near the horizon, we have to exclude all measurements with SZA greater than \(85^\circ\).
+Moreover, due to some obstructions around the site (hills and buildings), we excluded data with Azimuth angle between
+\(35^\circ\) and \(120^\circ\) with SZA greater than \(80^\circ\).
+On the latter instances, Sun is systematically, not visible by the instrument's location.
+To make the measurements comparable throughout the dataset, we adjusted all one-minute radiometric values to the mean Sun - Earth distance.
+Subsequently, we made all measurements relative to the Total Solar Irradiance (TSI) at \(1\,\text{au}\), in order to compensate for the Sun's intensity variability, using a time series of satellite TSI observations.
+The TSI data we use are part of the ``NOAA Climate Data Record of Total Solar Irradiance'' dataset (Coddington et al. 2005).
+Where the initial daily values, were interpolated to match with the time step of our measurements.
+The final dataset contains
+\(6589967\)
+one-minute measurements, of which,
+\(84.2\%\)
+were identified as under clear-sky conditions and subsequently
+\(15.8\%\)
+as under cloud-sky conditions.
+
+In order to investigate the SDR trends, we implemented an appropriate aggregation scheme to the 1-minute data to derive a series in coarser timescale.
+To preserve the representativeness of the data we used the following criteria:
+a) for the daily mean values we accept days with more than 50\% of the daytime measurements, present and valid,
+b) monthly values were computed from daily means
+only when at least 20 days were available.
+To create the daily and monthly climatological means, we averaged the data based on the day of year and calendar month, respectively.
+For the seasonal means we averaged the mean daily values in each season (Winter: December - February, Spring: March - May, etc.).
+Finally, each data set was deseasonalized by subtracting the corresponding climatological annual cycle (daily or monthly) from the actual data.
+To estimate SZA contribution to the SDR trends, the one-minute data were aggregated in \(1^\circ\) SZA bins, separately for the morning and afternoon hours, and then were deseasonalized as mentioned above.
+
+\hypertarget{results}{%
+\section{Results}\label{results}}
+
+\hypertarget{long-term-sdr-trends}{%
+\subsection{Long-term SDR trends}\label{long-term-sdr-trends}}
 
@@
 
-\startlandscape
-
-\begin{longtable}[t]{llllllll}
-\caption{(\#tab:mdpinames)MDPI journal names.}\\
-\toprule
-acoustics & biomedinformatics & dentistry & galaxies & jcp & metabolites & philosophies & socsci\\
-actuators & biomimetics & dermato & games & jcs & metals & photochem & software\\
-addictions & biomolecules & dermatopathology & gases & jcto & meteorology & photonics & soilsystems\\
-admsci & biophysica & designs & gastroent & jdb & methane & phycology & solar\\
-adolescents & biosensors & devices & gastrointestdisord & jeta & metrology & physchem & solids\\
-\addlinespace
-aerobiology & biotech & diabetology & gels & jfb & micro & physics & spectroscj\\
-aerospace & birds & diagnostics & genealogy & jfmk & microarrays & physiologia & sports\\
-agriculture & bloods & dietetics & genes & jimaging & microbiolres & plants & standards\\
-agriengineering & blsf & digital & geographies & jintelligence & micromachines & plasma & stats\\
-agrochemicals & brainsci & disabilities & geohazards & jlpea & microorganisms & platforms & std\\
-\addlinespace
-agronomy & breath & diseases & geomatics & jmmp & microplastics & pollutants & stresses\\
-ai & buildings & diversity & geosciences & jmp & minerals & polymers & surfaces\\
-air & businesses & dna & geotechnics & jmse & mining & polysaccharides & surgeries\\
-algorithms & cancers & drones & geriatrics & jne & modelling & poultry & suschem\\
-allergies & carbon & dynamics & grasses & jnt & molbank & powders & sustainability\\
-\addlinespace
-alloys & cardiogenetics & earth & gucdd & jof & molecules & preprints & symmetry\\
-analytica & catalysts & ebj & hazardousmatters & joitmc & mps & proceedings & synbio\\
-analytics & cells & ecologies & healthcare & jor & msf & processes & systems\\
-anatomia & ceramics & econometrics & hearts & journalmedia & mti & prosthesis & targets\\
-animals & challenges & economies & hemato & jox & muscles & proteomes & taxonomy\\
-\addlinespace
-antibiotics & chemengineering & education & hematolrep & jpm & nanoenergyadv & psf & technologies\\
-antibodies & chemistry & ejihpe & heritage & jrfm & nanomanufacturing & psych & telecom\\
-antioxidants & chemosensors & electricity & higheredu & jsan & nanomaterials & psychiatryint & test\\
-applbiosci & chemproc & electrochem & highthroughput & jtaer & ncrna & psychoactives & textiles\\
-appliedchem & children & electronicmat & histories & jvd & ndt & publications & thalassrep\\
-\addlinespace
-appliedmath & chips & electronics & horticulturae & jzbg & network & quantumrep & thermo\\
-applmech & cimb & encyclopedia & hospitals & kidneydial & neuroglia & quaternary & tomography\\
-applmicrobiol & civileng & endocrines & humanities & kinasesphosphatases & neurolint & qubs & tourismhosp\\
-applnano & cleantechnol & energies & humans & knowledge & neurosci & radiation & toxics\\
-applsci & climate & eng & hydrobiology & land & nitrogen & reactions & toxins\\
-\addlinespace
-aquacj & clinpract & engproc & hydrogen & languages & notspecified & receptors & transplantology\\
-architecture & clockssleep & entomology & hydrology & laws & nri & recycling & transportation\\
-arm & cmd & entropy & hygiene & life & nursrep & regeneration & traumacare\\
-arthropoda & coasts & environments & idr & liquids & nutraceuticals & religions & traumas\\
-arts & coatings & environsciproc & ijerph & literature & nutrients & remotesensing & tropicalmed\\
-\addlinespace
-asc & colloids & epidemiologia & ijfs & livers & obesities & reports & universe\\
-asi & colorants & epigenomes & ijgi & logics & oceans & reprodmed & urbansci\\
-astronomy & commodities & est & ijms & logistics & ohbm & resources & uro\\
-atmosphere & compounds & fermentation & ijns & lubricants & onco & rheumato & vaccines\\
-atoms & computation & fibers & ijpb & lymphatics & oncopathology & risks & vehicles\\
-\addlinespace
-audiolres & computers & fintech & ijtm & machines & optics & robotics & venereology\\
-automation & condensedmatter & fire & ijtpp & macromol & oral & ruminants & vetsci\\
-axioms & conservation & fishes & ime & magnetism & organics & safety & vibration\\
-bacteria & constrmater & fluids & immuno & magnetochemistry & organoids & sci & virtualworlds\\
-batteries & cosmetics & foods & informatics & make & osteology & scipharm & viruses\\
-\addlinespace
-bdcc & covid & forecasting & information & marinedrugs & oxygen & sclerosis & vision\\
-behavsci & crops & forensicsci & infrastructures & materials & parasites & seeds & waste\\
-beverages & cryptography & forests & inorganics & materproc & parasitologia & sensors & water\\
-biochem & crystals & foundations & insects & mathematics & particles & separations & wem\\
-bioengineering & csmf & fractalfract & instruments & mca & pathogens & sexes & wevj\\
-\addlinespace
-biologics & ctn & fuels & inventions & measurements & pathophysiology & signals & wind\\
-biology & curroncol & future & iot & medicina & pediatrrep & sinusitis & women\\
-biomass & cyber & futureinternet & j & medicines & pharmaceuticals & skins & world\\
-biomechanics & dairy & futurepharmacol & jal & medsci & pharmaceutics & smartcities & youth\\
-biomed & data & futurephys & jcdd & membranes & pharmacoepidemiology & sna & zoonoticdis\\
-\addlinespace
-biomedicines & ddc & futuretransp & jcm & merits & pharmacy & societies & \\
-\bottomrule
-\end{longtable}
-\finishlandscape
-
 \hypertarget{introduction}{%
@@
 keep the introduction comprehensible to scientists outside your particular
-field of research. Citing a journal paper (Bertrand-Krajewski, Chebbo, and Saget 1998; Leutnant, Muschalla, and Uhl 2016). And now citing a book reference Gujer (2008).
+field of research. Citing a journal paper (\textbf{bertrand-krajewski\_distribution\_1998?}; \textbf{leutnant\_stormwater\_2016?}). And now citing a book reference (\textbf{gujer\_systems\_2008?}).
 Some MDPI journals use Chicago and others use APA, this template should choose
@@
 
-\hypertarget{results}{%
-\section{Results}\label{results}}
+\hypertarget{results-1}{%
+\section{Results}\label{results-1}}
 
@@
 
-The text continues here. Proofs must be formatted as follows:
-
-Example of a proof:
-
-\begin{proof}
-Text of the proof. Note that the phrase ``of Theorem 1'\,' is optional if it is
-clear which theorem is being referred to.
-\end{proof}
-
 The text continues here.
@@
 
-\hypertarget{patents}{%
-\section{Patents}\label{patents}}
-
-This section is not mandatory, but may be added if there are patents resulting
-from the work reported in this manuscript.
-
 \hypertarget{refs}{}
 \begin{CSLReferences}{1}{0}
-\leavevmode\vadjust pre{\hypertarget{ref-bertrand-krajewski_distribution_1998}{}}%
-Bertrand-Krajewski, J.
-L., G.
-Chebbo, and A.
-Saget.
-1998.
-{``Distribution of Pollutant Mass Vs Volume in Stormwater Discharges and the First Flush Phenomenon.''} \emph{Water Research} 32 (8): 2341--56.
-\url{http://www.scopus.com/inward/record.url?eid=2-s2.0-0032146023\&partnerID=40\&md5=d6517b6efa014bc12f3ae70abe71977d}.
+\leavevmode\vadjust pre{\hypertarget{ref-Bais2013}{}}%
+Bais, A.
+F., T.
+Drosoglou, C.
+Meleti, K.
+Tourpali, and N.
+Kouremeti.
+2013.
+{``Changes in Surface Shortwave Solar Irradiance from 1993 to 2011 at Thessaloniki (Greece).''} \emph{International Journal of Climatology} 33 (13): 2871--76.
+\url{https://doi.org/f5dzz5}.
 
-\leavevmode\vadjust pre{\hypertarget{ref-gujer_systems_2008}{}}%
-Gujer, W. 2008. \emph{Systems {Analysis} for {Water} {Technology}}. Berlin, Heidelberg, Germany: Springer-Verlag.
+\leavevmode\vadjust pre{\hypertarget{ref-Brent1973}{}}%
+Brent, Richard P. 1973. {``Algorithms for Minimization Without Derivatives.''} \emph{PrenticeHall, Englewood Cliffs, NJ}.
 
-\leavevmode\vadjust pre{\hypertarget{ref-leutnant_stormwater_2016}{}}%
-Leutnant, Dominik, Dirk Muschalla, and Mathias Uhl.
+\leavevmode\vadjust pre{\hypertarget{ref-Coddington2005}{}}%
+Coddington, Odele, Judith L.
+Lean, Doug Lindholm, Peter Pilewskie, Martin Snow, and NOAA CDR Program.
+2005.
+{``{NOAA} Climate Data Record ({CDR}) of Total Solar Irradiance ({TSI}), {NRLTSI} Version 2.
+{D}aily.''} 2005.
+\url{https://doi.org/10.7289/V55B00C1}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-Haurwitz1945}{}}%
+Haurwitz, Bernhard. 1945. {``Insolation in {Relation} to {Cloudiness} and {Cloud} {Density}.''} \emph{Journal of Meteorology} 2 (September): 154--66.
+
+\leavevmode\vadjust pre{\hypertarget{ref-Li2016}{}}%
+Li, Zhanqing, W.
+K.‐M.
+Lau, V.
+Ramanathan, G.
+Wu, Y.
+Ding, M.
+G.
+Manoj, J.
+Liu, et al.
 2016.
-{``Stormwater {Pollutant} {Process} {Analysis} with {Long}-{Term} {Online} {Monitoring} {Data} at {Micro}-{Scale} {Sites}.''} \emph{Water} 8 (7): 299.
-\url{https://doi.org/10.3390/w8070299}.
+{``Aerosol and Monsoon Climate Interactions over Asia.''} \emph{Reviews of Geophysics} 54 (4): 866--929.
+\url{https://doi.org/10.1002/2015RG000500}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-Long2000}{}}%
+Long, Charles N., and Thomas P.
+Ackerman.
+2000.
+{``Identification of Clear Skies from Broadband Pyranometer Measurements and Calculation of Downwelling Shortwave Cloud Effects.''} \emph{Journal of Geophysical Research: Atmospheres} 105 (D12, D12): 15609--26.
+\url{https://doi.org/10.1029/2000jd900077}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-Long2006}{}}%
+Long, Charles N., and Y.
+Shi.
+2006.
+{``The QCRad Value Added Product: Surface Radiation Measurement Quality Control Testing, Including Climatology Configurable Limits.''} DOE/SC-ARM/TR-074.
+Office of Science, Office of Biological; Environmental Research, U.S.
+Department of Energy.
+
+\leavevmode\vadjust pre{\hypertarget{ref-Long2008a}{}}%
+---------. 2008. {``An Automated Quality Assessment and Control Algorithm for Surface Radiation Measurements.''} \emph{The Open Atmospheric Science Journal}, 23--37.
+
+\leavevmode\vadjust pre{\hypertarget{ref-Ohvril2009}{}}%
+Ohvril, Hanno, Hilda Teral, Lennart Neiman, Martin Kannel, Marika Uustare, Mati Tee, Viivi Russak, et al.
+2009.
+{``Global Dimming and Brightening Versus Atmospheric Column Transparency, Europe, 1906--2007.''} \emph{Journal of Geophysical Research} 114 (May).
+\url{https://doi.org/10.1029/2008JD010644}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-RCT2023}{}}%
+R Core Team. 2023. \emph{R: A Language and Environment for Statistical Computing}. Vienna, Austria: R Foundation for Statistical Computing. \url{https://www.R-project.org/}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-Reno2016}{}}%
+Reno, Matthew J., and Clifford W.
+Hansen.
+2016.
+{``Identification of Periods of Clear Sky Irradiance in Time Series of GHI Measurements.''} \emph{Renewable Energy} 90: 520--31.
+\url{https://doi.org/gq3sbg}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-Reno2012}{}}%
+Reno, Matthew J., Clifford W.
+Hansen, and Joshua S.
+Stein.
+2012b.
+{``Global Horizontal Irradiance Clear Sky Models: Implementation and Analysis.''} SAND2012-2389, 1039404.
+\url{https://doi.org/gq5npv}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-Reno2012a}{}}%
+---------. 2012a. {``{Global Horizontal Irradiance Clear Sky Models: Implementation and Analysis}.''} \emph{SANDIA REPORT SAND2012-2389 Unlimited Release Printed March 2012}, March, 1--66.
+
+\leavevmode\vadjust pre{\hypertarget{ref-Samset2018}{}}%
+Samset, B.
+H., M.
+Sand, C.
+J.
+Smith, S.
+E.
+Bauer, P.
+M.
+Forster, J.
+S.
+Fuglestvedt, S.
+Osprey, and C.‐F.
+Schleussner.
+2018.
+{``Climate Impacts from a Removal of Anthropogenic Aerosol Emissions.''} \emph{Geophysical Research Letters} 45 (2): 1020--29.
+\url{https://doi.org/10.1002/2017GL076079}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-Schwarz2020}{}}%
+Schwarz, M., D.
+Folini, S.
+Yang, R.
+P.
+Allan, and M.
+Wild.
+2020.
+{``Changes in Atmospheric Shortwave Absorption as Important Driver of Dimming and Brightening.''} \emph{Nature Geoscience} 13 (2): 110--15.
+\url{https://doi.org/10.1038/s41561-019-0528-y}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-Wild2009}{}}%
+Wild, Martin. 2009. {``Global Dimming and Brightening: A Review.''} \emph{Journal of Geophysical Research Atmospheres} 114 (12): 1--31. \url{https://doi.org/bcq}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-Wild2021}{}}%
+Wild, Martin, Stephan Wacker, Su Yang, and Arturo Sanchez-Lorenzo.
+2021.
+{``Evidence for Clear‐sky Dimming and Brightening in Central Europe.''} \emph{Geophysical Research Letters} 48 (6).
+\url{https://doi.org/10.1029/2020GL092216}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-Xia2007}{}}%
+Xia, Xiangao, Hongbin Chen, Zhanqing Li, Pucai Wang, and Jiankai Wang.
+2007.
+{``Significant Reduction of Surface Solar Irradiance Induced by Aerosols in a Suburban Region in Northeastern China.''} \emph{Journal of Geophysical Research Atmospheres} 112 (22): 1--9.
+\url{https://doi.org/cdtntw}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-Yamasoe2021}{}}%
+Yamasoe, Marcia Akemi, Nilton Manuel Évora Rosário, Samantha Novaes Santos Martins Almeida, and Martin Wild.
+2021.
+{``Fifty-Six Years of Surface Solar Radiation and Sunshine Duration over s{ã}o Paulo, Brazil: 1961--2016.''} \emph{Atmospheric Chemistry and Physics} 21 (9): 6593--603.
+\url{https://doi.org/10.5194/acp-21-6593-2021}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-Yang2021}{}}%
+Yang, Su, Zijiang Zhou, Yu Yu, and Martin Wild.
+2021.
+{``Cloud {``}Shrinking{''} and {``}Optical Thinning{''} in the {``}Dimming{''} Period and a Subsequent Recovery in the {``}Brightening{''} Period over China.''} \emph{Environmental Research Letters}, January.
+\url{https://doi.org/10.1088/1748-9326/abdf89}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-Yuan2021}{}}%
+Yuan, Menghan, Thomas Leirvik, and Martin Wild.
+2021.
+{``Global Trends in Downward Surface Solar Radiation from Spatial Interpolated Ground Observations During 1961-2019.''} \emph{Journal of Climate}, September, 1--56.
+\url{https://doi.org/10.1175/JCLI-D-21-0165.1}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-Zerefos2009}{}}%
+Zerefos, C.
+S., K.
+Eleftheratos, C.
+Meleti, S.
+Kazadzis, A.
+Romanou, C.
+Ichoku, G.
+Tselioudis, and A.
+Bais.
+2009.
+{``Solar Dimming and Brightening over Thessaloniki, Greece, and Beijing, China.''} \emph{Tellus B: Chemical and Physical Meteorology} 61 (4): 657.
+\url{https://doi.org/10.1111/j.1600-0889.2009.00425.x}.
 

### 2023-11-01 MDPI_submition.Rmd
Commit 2023-11-01 20:14
Commit 2023-11-01 21:08
Commit 2023-11-01 22:14

@@
     long: Solar downward radiation
-bibliography: references.bib
+bibliography: manualreferences.bib
 appendix:     appendix.tex
@@
 
+<!--
+Aerosols interactions
+
+#Yang2021  Cloud “shrinking” and “optical thinning” in the “dimming” period and a subsequent recovery in the “brightening” period over China
+
+Aerosols not only strengthen but also weaken the
+growth of clouds with respect to their coverage and
+optical thickness, depending on the levels of pollution
+and the associated amounts of aerosols as postulated
+in a conceptual framework (Wild 2009a, 2012, Yang
+et al 2012).
+
+...... cloud formation occurrences and constitution nuclei. cite .....
+
+#Wild2012 : simplify interaction nice graph
+
+-->
+
+# Observational data and methodology.
+
+The SDR data were measured with a Kipp & Zonen CM-21 pyranometer operating continuously at the Laboratory of Atmospheric Physics of the Aristotle University of Thessaloniki 
+($40^\circ\,38'\,$N, $22^\circ\,57'\,$E, $80\,$m\ a.s.l.)
+in the period from 
+<!-- \mbox{ -->
+`r min(ALL_1_daily_DESEAS$Date)`
+<!-- } -->
+to
+<!-- \mbox{ -->
+`r max(ALL_1_daily_DESEAS$Date)`.
+<!-- }. -->
+The monitoring site is located near the city centre, and we expect to be affected by the urban environment. 
+During the study period, the pyranometer has been independently calibrated three times at the Meteorologisches Observatorium Lindenberg, DWD, when it was verified the stability of the instrument to within better than $0.7\%$ relative to the initial calibration by the manufacturer.
+Along with SDR, the direct beam radiation (DNI) was also measured by a collocated Kipp & Zonen CHP-1 pyrheliometer, for the period
+<!-- \mbox{ -->
+`r min(ALL_1_daily_DESEAS[!is.na(DIR_att), Date])`
+<!-- } -->
+to
+<!-- \mbox{ -->
+`r max(ALL_1_daily_DESEAS[!is.na(DIR_att), Date])`.
+<!-- }. -->
+Although, we have performed a similar analysis to the DNI data the results are not presented here, as they lack the appropriate statistical significance, due to the sorter duration of the data. 
+However, the DNI data were used as auxiliary data, in the clear sky identification algorithm (CSid), which is discussed later, for the selection of the appropriate thresholds.
+It is noted that despite the capability of the CSid algorithm to use the DNI as a characterization parameter, we haven't utilized it here, to avoid any selection bias, due to unequal length of the two datasets.
+There are four distinct steps in the creation of the dataset analysed here:
+a)\ the acquisition of radiation measurements from the sensors,
+b)\ the data quality check,
+c)\ the identification of "clear sky" conditions from the radiometric data, and 
+d)\ the aggregation of data and trend analysis.
+
+<!--
+The horizon is clear of obstructions except for the azimuth angles between 35 and 120 where obstacles of up to 10 elevation block the direct irradiance in the morning for SZA larger than 80.
+These data points were excluded from the analysis, as
+-->
+
+<!-- Zerefos 
+Northern Greece has been selected because it is well known to be located at the crossroads of aerosols and various pollutants pathways originating in central and Eastern
+Europe (Lelieveld et al., 2002; Zerefos et al., 2002). Thessa-
+loniki, the major city in northern Greece, lies along the exit of air
+pollutants from Europe, in the path of the prevailing long-range
+transport (LRT) trajectories, as we move from central and east-
+ern Europe to the Mediterranean (Zerefos et. al., 2002). On the
+-->
+
+<!-- a) acquisition -->
+For the acquisition of radiometric data, the signal of the pyranometer is sampled with a rate of $1\,\text{Hz}$.
+The mean and the standard deviation of these samples are recorded every minute.
+The measurements are corrected for the zero offset ("dark signal" in volts).
+The "dark signal" is calculated by averaging all measurements recorded for a period of 
+$3\,\text{h}$, 
+before (morning) or after (evening) the Sun reaches an elevation angle of
+$-10^\circ$.
+The signal is converted to irradiance using a ramped value of the instrument’s sensitivity between calibrations.
+
+<!-- b) Radiation data quality check -->
+A manual screening was performed, to remove inconsistent and erroneous recordings that can occur stochastically or systematically, during the continuous operation of the instruments.
+The manual screening is aided by a radiation data quality assurance procedure, adjusted for the site, which is based on the methods of 
+Long and Shi\ [-@Long2008a; -@Long2006].
+Thus, problematic recordings have been excluded from further processing. 
+Although it is impossible to detect all false data, the large number of available data, and the aggregation scheme we used, ensures the good quality of the radiometric measurements used in this study.
+
+<!-- c) Clear sky identification -->
+In order to be able to estimate the effect of the sky condition on the long term variability of SDR, we created three datasets, by characterizing each one-minute measurement with a corresponding sky condition (i.e., all-sky, clear-sky and cloudy-sky).
+To identify the clear-sky conditions we used a method proposed by
+@Long2000
+and by
+@Reno2016,
+which was adapted and configured for the site, as the authors suggest.
+
+<!-- CS subjectivity -->
+We have to note, that the definition of what constitutes as clear or cloudy sky, has some subjectivity, in any method of characterization.
+As a result, the details of the definition are site specific, it relies on a combination of thresholds and comparisons with ideal actinometric models and statistical analysis on different signal metrics.
+The CSid algorithm was calibrated with the main focus, to identify the presence of clouds on the sky dome.
+Although the fine-tuning of the procedure, few marginal cases exist, that have been identified manually as false positive or false negative but cannot affect the final results of the study.
+
+```{r setup, echo=F, include=T, eval=TRUE}
+## Load Clear Sky id v14.2 variables
+MS <- data.frame(readRDS("~/CS_id/PARAMS/Clear_sky_id_Reno-Hansen_apply_v14_2.Rds"))
+# export all Clear Sky variables
+for (an in names(MS)) {
+assign(an, MS[[an]])
+}
+## get TSI slope for reference
+TSI <- data.table(read.csv("./figures/tbl_longterm_trends.dat",
+           strip.white = TRUE,
+           sep = ";",
+           header = TRUE))
+TSI <- TSI[grep("TSI", TSI$var), ]
+TSI$slope.sd           <- NULL
+TSI$slope.ConfInt_0.95 <- NULL
+TSI$slope.ConfInt_0.99 <- NULL
+TSI$var                <- NULL
+TSI$N                  <- NULL
+TSI$Obs                <- NULL
+setorder(TSI, -slope.stat_sig)
+rownames(TSI) <- NULL
+names(TSI) <- c("Trend [%/year]", "Trend p-value", "Sky condition", "Trend statistical signif. [%]")
+TSIslopeAll <- TSI[`Sky condition` == "All sky cond."]
+```
+
+<!-- # Detection of clear periods in GHI measurements for SDR trends. -->
+
+For completeness, we will provide below a brief overview of the clear sky identification algorithm (CSid), along with the site specific thresholds.
+To calculate the reference clear sky
+$\text{SDR}_\text{CSref}$ we used the $\text{SDR}_\text{Haurwitz}$ derived by
+the radiation model of @Haurwitz1945, adjusted for our site with a
+factor $a$ (Eq.\ \ref{eq:ahau}), resulted by an iterative optimization process, as described
+by @Long2000 and @Reno2016.
+The target of
+the optimization was the minimization of a function $f(a)$ (Eq.\ \ref{eq:minf}) and
+was accomplished with the algorithmic function "optimise", which is an implementation based on the work of @Brent1973, from the library "stats" of the R programming language [@RCT2023]. 
+The optimization and the selection of the clear sky reference model, was performed on SDR observations for the period 2016 - 2021.
+During the optimization, eight simple clear sky radiation models were tested (Daneshyar-Paltridge-Proctor, Kasten-Czeplak, Haurwitz, Berger-Duffie, Adnot-Bourges-Campana-Gicquel, Robledo-Soler, Kasten and Ineichen-Perez), with a wide range of factors.
+These models are described in more details by @Reno2012 and evaluated by @Reno2016.
+We found, that Haurwitz's model, adjusted with the factor $a = `r signif(alpha,3)`$ yields one of the lowest root mean squared errors (RMSE), 
+<!--[....rephrased....]{.comment-start id="3" author="A N" date="2023-07-15T20:32:50Z"}-->
+while the procedure, manages to characterize the majority of the data.
+<!--[]{.comment-end id="3"}--> 
+Thus, our clear sky reference is derived by the Eq.\ \ref{eq:ahau}.
+
+<!-- cost fuction -->
+\begin{equation}
+f(a) = \frac{1}{n}\sum_{i=1}^{n} ( \text{SDR}_{\text{CSid},i} - a \times \text{SDR}_{\text{testCSref},i} )^2 \label{eq:minf}
+\end{equation}
+where: $n$ is the total number of daylight records, $\text{SDR}_{\text{CSid},i}$ are the records identified as clear sky by CSid, $a$ is a hypothetical adjustment factor, and $\text{SDR}_{\text{testCSref},i}$ is any of the tested clear sky radiation models.
+
+<!-- adjusted Haurwitz -->
+\begin{equation}
+\text{SDR}_\text{CSref} = a \times \text{SDR}_\text{Haurwitz} = `r signif(alpha,3)` \times 1098 \times \cos(\theta) \times \exp \left( \frac{ - 0.057}{\cos(\theta)} \right) \label{eq:ahau}
+\end{equation}
+where: $\text{SDR}_\text{CSref}$ is the reference clear sky SDR, in $\text{w}\,\text{m}^{-2}$ and $\theta$ is the solar zenith angle (SZA).
+
+The criteria that were used to identify whether a measurement was taken under clear-sky conditions are presented below.
+A data point is flagged as "clear-sky" if all criteria are satisfied, otherwise it is considered to be "cloud-sky".
+Each criterion was applied for a running window of $`r nt`$ consecutive one-minute measurements, and the characterization is assigned to the central value of the window.
+Each parameter, was calculated both from the observations and the reference clear sky model, for each comparison.
+The allowable range of variation is defined by the model-derived value of the parameter multiplied by a factor plus an offset. 
+The factors and the offsets were determined empirically, by manual inspecting each filters performance on selected days, and adjusting them accordingly, during an iterative process.
+
+<!-- 1. MeanVIP -->
+<!-- ### Mean value of irradiance during the time period. -->
+
+<!--a) Mean of the measured $\overline{\text{SDR}}_i$ = \text{mean}_{n=i-5}^{n=i+5}[\text{SDR}_{n}]$ and running mean of the reference $\overline{\text{SDR}}_{\text{CSref},i}$ = \text{mean}_{n=i-5}^{n=i+5}[\text{SDR}_{\text{CSref},n}]$ (Eq.
+\ref{eq:MeanVIP}).-->
+
+a) Mean of the measured $\overline{\text{SDR}}_i$ (Eq. \ref{eq:MeanVIP}).
+\begin{equation}
+`r 1 - CS_ref_rm_VIP_RelLow` \times \overline{\text{SDR}}_{\text{CSref},i} - `r CS_ref_rm_VIP_LowOff`
+< \overline{\text{SDR}}_i <
+`r 1 + CS_ref_rm_VIP_RelUpp` \times \overline{\text{SDR}}_{\text{CSref},i} + `r CS_ref_rm_VIP_UppOff`
+\label{eq:MeanVIP}
+\end{equation}
+
+<!-- 2. MaxVIP -->
+<!-- ### Max value of irradiance during the time period. -->
+
+<!-- The running max measured value $M_{\text{SDR},i} = \max_{n=i-5}^{n=i+5}[\text{SDR}_{n}]$, is compared to a similar constructed value from the reference $M_{CSrefi} = \max_{n=i-5}^{n=i+5}[\text{SDR}_{\text{CSref},n}]$ (Eq.\ \ref{eq:MaxVIP}).
+-->
+
+b) Maximum measured value $M_{\text{}}$ (Eq.\ \ref{eq:MaxVIP}).
+\begin{equation}
+`r MaxVIP_fct` \times M_{\text{CSref},i} - `r MaxVIP_off_low`
+< M_{\text{}i} <
+`r MaxVIP_fct` \times M_{\text{CSref},i} + `r MaxVIP_off_upp`
+\label{eq:MaxVIP}
+\end{equation}
+
+<!-- 3. VIL -->
+<!-- ### Variability in irradiance by line length. -->
+c) Length $L_i$ of the sequential line segments, connecting the points of the $`r nt`$ SDR values (Eq. \ref{eq:VILeq}).
+\begin{equation}
+L_i = \sum_{i=1}^{n-1}\sqrt{\left ( \text{SDR}_{i+1} - \text{SDR}_{i}\right )^2 + \left ( t_{i+1} - t_i \right )^2}
+\label{eq:VILeq}
+\end{equation}
+\begin{equation}
+`r MinVIL_fct` \times L_{\text{CSref},i} - `r offVIL_dwl` < L_i < `r MaxVIL_fct` \times L_{\text{CSref},i} + `r offVIL_upl`
+\label{eq:VILcr}
+\end{equation}
+where: $t_i$ is the time each SDR measurement has been measured
+
+<!-- 4. VCT -->
+<!-- ### Variance of Changes in the Time series. -->
+
+<!-- Standard deviation of rate of change in irradiance. -->
+d) Standard deviation $\sigma_i$ of the slope ($s_i$) between the $`r nt`$ sequential points, normalized by the mean $\overline{\text{SDR}}_i$ (Eq.\ \ref{eq:VCT1}).
+<!-- \begin{gather}
+\bar{s} = \frac{1}{n-1} \sum_{i=1}^{n-1} s_i \label{eq:VCT2} \\
+\sigma_i = \frac {\sqrt{\frac{1}{n-1} \sum_{i=1}^{n-1} \left( s_i - \bar{s} \right)^2} } {\bar{G_i}} \label{eq:VCT3}
+\end{gather} -->
+\begin{gather}
+  \sigma_i = \frac {\sqrt{\frac{1}{n-1} \sum_{i=1}^{n-1} \left( s_i - \bar{s} \right)^2}} {\overline{\text{SDR}}_i} \label{eq:VCT1} \\
+  s_i = \frac{\text{SDR}_{i+1} - \text{SDR}_{i}}{t_{i+1} - t_i},\;\;   \bar{s} = \frac{1}{n-1} \sum_{i=1}^{n-1} s_i,\;\;\forall i \in \left \{ 1, 2, \ldots, n-1 \right \}\;\;
+\end{gather}
+For this criterion, $\sigma_i$ should be below a certain threshold (Eq.\ \ref{eq:VCTcr}):
+\begin{equation}
+  \sigma_i < `r offVCT` \label{eq:VCTcr}
+\end{equation}
+
+<!-- 5. VSM -->
+<!-- ### Variability in the Shape of the irradiance Measurements. -->
+
+e) Maximum difference $X_i$ between the change in measured irradiance and the change in clear sky irradiance over each measurement interval.
+\begin{gather}
+  X_i = \max{\left \{ \left | x_i - x_{\text{CSref},i} \right | \right \}} \label{eq:VSM3} \\
+  x_i = \text{SDR}_{i+1} - \text{SDR}_{i} \forall i \in \left \{ 1, 2, \ldots, n-1 \right \} \label{eq:VSM1} \\
+  x_{\text{CSref},i} = \text{SDR}_{\text{CSref},i+1} - \text{SDR}_{\text{CSref},i} \forall i \in \left \{ 1, 2, \ldots, n-1 \right \} \label{eq:VSM2}
+\end{gather}
+For this criterion, $X_i$ should be below a certain threshold (Eq.\ \ref{eq:VSMcr}):
+\begin{equation}
+  X_i < `r offVSM` \label{eq:VSMcr}
+\end{equation}
+
+<!-- ### Data selection -->
+Due to a significant measurement uncertainty near the horizon, we have to exclude all measurements with SZA greater than $`r 90-MIN_ELEVA`^\circ$. 
+Moreover, due to some obstructions around the site (hills and buildings), we excluded data with Azimuth angle between 
+$35^\circ$ and $120^\circ$ with SZA greater than $80^\circ$.
+On the latter instances, Sun is systematically, not visible by the instrument's location.
+<!-- 1au -->
+To make the measurements comparable throughout the dataset, we adjusted all one-minute radiometric values to the mean Sun - Earth distance.
+<!-- TSI -->
+Subsequently, we made all measurements relative to the Total Solar Irradiance (TSI) at $1\,\text{au}$, in order to compensate for the Sun's intensity variability, using a <!--homogenized--> time series of satellite TSI observations.
+The TSI data we use are part of the "NOAA Climate Data Record of Total Solar Irradiance" dataset [@Coddington2005]. 
+Where the initial daily values, were interpolated to match with the time step of our measurements. 
+The final dataset contains
+$`r SDR_N_All`$
+one-minute measurements, of which, 
+$`r signif(100 * SDR_N_Clear / SDR_N_All, 3)`\%$
+were identified as under clear-sky conditions and subsequently
+$`r signif(100 * SDR_N_Cloud / SDR_N_All, 3)`\%$
+as under cloud-sky conditions.
+<!-- The TSI data we use, is a combination of data from NOAA [@Coddington2005] (for `r as.Date(TSIinfo[TSI_Source == "NOAA", Start])` - `r as.Date(TSIinfo[TSI_Source == "NOAA", End])`) and adjusted data from @LASP2023 (for `r as.Date(TSIinfo[TSI_Source != "NOAA", Start])` - `r as.Date(TSIinfo[TSI_Source != "NOAA", End])`).-->
+
+<!-- aggregation -->
+In order to investigate the SDR trends, we implemented an appropriate aggregation scheme to the 1-minute data to derive a series in coarser timescale.
+To preserve the representativeness of the data we used the following criteria: 
+<!-- a) for daily mean values we exclude days with less than `r Daily_aggregation_N_lim` valid data points, -->
+a) for the daily mean values we accept days with more than `r 100 * All_daily_ratio_lim`% of the daytime measurements, present and valid,
+b) monthly values were computed from daily means
+only when at least `r Monthly_aggegation_N_lim` days were available.
+To create the daily and monthly climatological means, we averaged the data based on the day of year and calendar month, respectively.
+<!-- Seasonal -->
+For the seasonal means we averaged the mean daily values in each season (Winter: December - February, Spring: March - May, etc.).
+<!-- Deseasonalization -->
+Finally, each data set was deseasonalized by subtracting the corresponding climatological annual cycle (daily or monthly) from the actual data.
+To estimate SZA contribution to the SDR trends, the one-minute data were aggregated in $1^\circ$ SZA bins, separately for the morning and afternoon hours, and then were deseasonalized as mentioned above.
+
+# Results
+
+## Long-term SDR trends
+
 # Version
@@
 ## Journal Specific YAML variables
-
-\startlandscape
-```{r, mdpinames, echo=FALSE, results='asis'}
-mdpi_journals <- "acoustics, actuators, addictions, admsci, adolescents, aerobiology, aerospace, agriculture, agriengineering, agrochemicals, agronomy, ai, air, algorithms, allergies, alloys, analytica, analytics, anatomia, animals, antibiotics, antibodies, antioxidants, applbiosci, appliedchem, appliedmath, applmech, applmicrobiol, applnano, applsci, aquacj, architecture, arm, arthropoda, arts, asc, asi, astronomy, atmosphere, atoms, audiolres, automation, axioms, bacteria, batteries, bdcc, behavsci, beverages, biochem, bioengineering, biologics, biology, biomass, biomechanics, biomed, biomedicines, biomedinformatics, biomimetics, biomolecules, biophysica, biosensors, biotech, birds, bloods, blsf, brainsci, breath, buildings, businesses, cancers, carbon, cardiogenetics, catalysts, cells, ceramics, challenges, chemengineering, chemistry, chemosensors, chemproc, children, chips, cimb, civileng, cleantechnol, climate, clinpract, clockssleep, cmd, coasts, coatings, colloids, colorants, commodities, compounds, computation, computers, condensedmatter, conservation, constrmater, cosmetics, covid, crops, cryptography, crystals, csmf, ctn, curroncol, cyber, dairy, data, ddc, dentistry, dermato, dermatopathology, designs, devices, diabetology, diagnostics, dietetics, digital, disabilities, diseases, diversity, dna, drones, dynamics, earth, ebj, ecologies, econometrics, economies, education, ejihpe, electricity, electrochem, electronicmat, electronics, encyclopedia, endocrines, energies, eng, engproc, entomology, entropy, environments, environsciproc, epidemiologia, epigenomes, est, fermentation, fibers, fintech, fire, fishes, fluids, foods, forecasting, forensicsci, forests, foundations, fractalfract, fuels, future, futureinternet, futurepharmacol, futurephys, futuretransp, galaxies, games, gases, gastroent, gastrointestdisord, gels, genealogy, genes, geographies, geohazards, geomatics, geosciences, geotechnics, geriatrics, grasses, gucdd, hazardousmatters, healthcare, hearts, hemato, hematolrep, heritage, higheredu, highthroughput, histories, horticulturae, hospitals, humanities, humans, hydrobiology, hydrogen, hydrology, hygiene, idr, ijerph, ijfs, ijgi, ijms, ijns, ijpb, ijtm, ijtpp, ime, immuno, informatics, information, infrastructures, inorganics, insects, instruments, inventions, iot, j, jal, jcdd, jcm, jcp, jcs, jcto, jdb, jeta, jfb, jfmk, jimaging, jintelligence, jlpea, jmmp, jmp, jmse, jne, jnt, jof, joitmc, jor, journalmedia, jox, jpm, jrfm, jsan, jtaer, jvd, jzbg, kidneydial, kinasesphosphatases, knowledge, land, languages, laws, life, liquids, literature, livers, logics, logistics, lubricants, lymphatics, machines, macromol, magnetism, magnetochemistry, make, marinedrugs, materials, materproc, mathematics, mca, measurements, medicina, medicines, medsci, membranes, merits, metabolites, metals, meteorology, methane, metrology, micro, microarrays, microbiolres, micromachines, microorganisms, microplastics, minerals, mining, modelling, molbank, molecules, mps, msf, mti, muscles, nanoenergyadv, nanomanufacturing, nanomaterials, ncrna, ndt, network, neuroglia, neurolint, neurosci, nitrogen, notspecified, nri, nursrep, nutraceuticals, nutrients, obesities, oceans, ohbm, onco, oncopathology, optics, oral, organics, organoids, osteology, oxygen, parasites, parasitologia, particles, pathogens, pathophysiology, pediatrrep, pharmaceuticals, pharmaceutics, pharmacoepidemiology, pharmacy, philosophies, photochem, photonics, phycology, physchem, physics, physiologia, plants, plasma, platforms, pollutants, polymers, polysaccharides, poultry, powders, preprints, proceedings, processes, prosthesis, proteomes, psf, psych, psychiatryint, psychoactives, publications, quantumrep, quaternary, qubs, radiation, reactions, receptors, recycling, regeneration, religions, remotesensing, reports, reprodmed, resources, rheumato, risks, robotics, ruminants, safety, sci, scipharm, sclerosis, seeds, sensors, separations, sexes, signals, sinusitis, skins, smartcities, sna, societies, socsci, software, soilsystems, solar, solids, spectroscj, sports, standards, stats, std, stresses, surfaces, surgeries, suschem, sustainability, symmetry, synbio, systems, targets, taxonomy, technologies, telecom, test, textiles, thalassrep, thermo, tomography, tourismhosp, toxics, toxins, transplantology, transportation, traumacare, traumas, tropicalmed, universe, urbansci, uro, vaccines, vehicles, venereology, vetsci, vibration, virtualworlds, viruses, vision, waste, water, wem, wevj, wind, women, world, youth, zoonoticdis"
-
-knitr::kable(rticles::string_to_table(mdpi_journals, 8),
-             col.names = NULL,
-             format = "latex",
-             booktabs = TRUE, 
-             caption = "MDPI journal names.",
-             longtable = TRUE)
-
-```
-\finishlandscape
 
@@
 
-The text continues here. Proofs must be formatted as follows:
-
-Example of a proof:
-
-::: {.proof latex="[Proof of Theorem1]"}
-Text of the proof. Note that the phrase ``of Theorem 1'' is optional if it is 
-clear which theorem is being referred to.
-:::
-
 The text continues here.
@@
 discussion is unusually long or complex.
-
-# Patents
-
-This section is not mandatory, but may be added if there are patents resulting
-from the work reported in this manuscript.

### 2023-11-01 MDPI_submition.tex
Commit 2023-11-01 21:08
Commit 2023-11-01 22:14

@@
 
+\hypertarget{observational-data-and-methodology.}{%
+\section{Observational data and
+methodology.}\label{observational-data-and-methodology.}}
+
+The SDR data were measured with a Kipp \& Zonen CM-21 pyranometer
+operating continuously at the Laboratory of Atmospheric Physics of the
+Aristotle University of Thessaloniki (\(40^\circ\,38'\,\)N,
+\(22^\circ\,57'\,\)E, \(80\,\)m~a.s.l.) in the period from 1993-04-13 to
+2023-04-13. The monitoring site is located near the city centre, and we
+expect to be affected by the urban environment. During the study period,
+the pyranometer has been independently calibrated three times at the
+Meteorologisches Observatorium Lindenberg, DWD, when it was verified the
+stability of the instrument to within better than \(0.7\%\) relative to
+the initial calibration by the manufacturer. Along with SDR, the direct
+beam radiation (DNI) was also measured by a collocated Kipp \& Zonen
+CHP-1 pyrheliometer, for the period 2016-04-01 to 2023-04-13. Although,
+we have performed a similar analysis to the DNI data the results are not
+presented here, as they lack the appropriate statistical significance,
+due to the sorter duration of the data. However, the DNI data were used
+as auxiliary data, in the clear sky identification algorithm (CSid),
+which is discussed later, for the selection of the appropriate
+thresholds. It is noted that despite the capability of the CSid
+algorithm to use the DNI as a characterization parameter, we haven't
+utilized it here, to avoid any selection bias, due to unequal length of
+the two datasets. There are four distinct steps in the creation of the
+dataset analysed here: a)~the acquisition of radiation measurements from
+the sensors, b)~the data quality check, c)~the identification of ``clear
+sky'' conditions from the radiometric data, and d)~the aggregation of
+data and trend analysis.
+
+For the acquisition of radiometric data, the signal of the pyranometer
+is sampled with a rate of \(1\,\text{Hz}\). The mean and the standard
+deviation of these samples are recorded every minute. The measurements
+are corrected for the zero offset (``dark signal'' in volts). The ``dark
+signal'' is calculated by averaging all measurements recorded for a
+period of \(3\,\text{h}\), before (morning) or after (evening) the Sun
+reaches an elevation angle of \(-10^\circ\). The signal is converted to
+irradiance using a ramped value of the instrument's sensitivity between
+calibrations.
+
+A manual screening was performed, to remove inconsistent and erroneous
+recordings that can occur stochastically or systematically, during the
+continuous operation of the instruments. The manual screening is aided
+by a radiation data quality assurance procedure, adjusted for the site,
+which is based on the methods of Long and
+Shi~\citetext{\citeyear{Long2008a}; \citeyear{Long2006}}. Thus,
+problematic recordings have been excluded from further processing.
+Although it is impossible to detect all false data, the large number of
+available data, and the aggregation scheme we used, ensures the good
+quality of the radiometric measurements used in this study.
+
+In order to be able to estimate the effect of the sky condition on the
+long term variability of SDR, we created three datasets, by
+characterizing each one-minute measurement with a corresponding sky
+condition (i.e., all-sky, clear-sky and cloudy-sky). To identify the
+clear-sky conditions we used a method proposed by \citet{Long2000} and
+by \citet{Reno2016}, which was adapted and configured for the site, as
+the authors suggest.
+
+We have to note, that the definition of what constitutes as clear or
+cloudy sky, has some subjectivity, in any method of characterization. As
+a result, the details of the definition are site specific, it relies on
+a combination of thresholds and comparisons with ideal actinometric
+models and statistical analysis on different signal metrics. The CSid
+algorithm was calibrated with the main focus, to identify the presence
+of clouds on the sky dome. Although the fine-tuning of the procedure,
+few marginal cases exist, that have been identified manually as false
+positive or false negative but cannot affect the final results of the
+study.
+
+For completeness, we will provide below a brief overview of the clear
+sky identification algorithm (CSid), along with the site specific
+thresholds. To calculate the reference clear sky
+\(\text{SDR}_\text{CSref}\) we used the \(\text{SDR}_\text{Haurwitz}\)
+derived by the radiation model of \citet{Haurwitz1945}, adjusted for our
+site with a factor \(a\) (Eq.~\ref{eq:ahau}), resulted by an iterative
+optimization process, as described by \citet{Long2000} and
+\citet{Reno2016}. The target of the optimization was the minimization of
+a function \(f(a)\) (Eq.~\ref{eq:minf}) and was accomplished with the
+algorithmic function ``optimise'', which is an implementation based on
+the work of \citet{Brent1973}, from the library ``stats'' of the R
+programming language \citep{RCT2023}. The optimization and the selection
+of the clear sky reference model, was performed on SDR observations for
+the period 2016 - 2021. During the optimization, eight simple clear sky
+radiation models were tested (Daneshyar-Paltridge-Proctor,
+Kasten-Czeplak, Haurwitz, Berger-Duffie, Adnot-Bourges-Campana-Gicquel,
+Robledo-Soler, Kasten and Ineichen-Perez), with a wide range of factors.
+These models are described in more details by \citet{Reno2012} and
+evaluated by \citet{Reno2016}. We found, that Haurwitz's model, adjusted
+with the factor \(a = 0.965\) yields one of the lowest root mean squared
+errors (RMSE), while the procedure, manages to characterize the majority
+of the data. Thus, our clear sky reference is derived by the
+Eq.~\ref{eq:ahau}.
+
+\begin{equation}
+f(a) = \frac{1}{n}\sum_{i=1}^{n} ( \text{SDR}_{\text{CSid},i} - a \times \text{SDR}_{\text{testCSref},i} )^2 \label{eq:minf}
+\end{equation} where: \(n\) is the total number of daylight records,
+\(\text{SDR}_{\text{CSid},i}\) are the records identified as clear sky
+by CSid, \(a\) is a hypothetical adjustment factor, and
+\(\text{SDR}_{\text{testCSref},i}\) is any of the tested clear sky
+radiation models.
+
+\begin{equation}
+\text{SDR}_\text{CSref} = a \times \text{SDR}_\text{Haurwitz} = 0.965 \times 1098 \times \cos(\theta) \times \exp \left( \frac{ - 0.057}{\cos(\theta)} \right) \label{eq:ahau}
+\end{equation} where: \(\text{SDR}_\text{CSref}\) is the reference clear
+sky SDR, in \(\text{w}\,\text{m}^{-2}\) and \(\theta\) is the solar
+zenith angle (SZA).
+
+The criteria that were used to identify whether a measurement was taken
+under clear-sky conditions are presented below. A data point is flagged
+as ``clear-sky'' if all criteria are satisfied, otherwise it is
+considered to be ``cloud-sky''. Each criterion was applied for a running
+window of \(11\) consecutive one-minute measurements, and the
+characterization is assigned to the central value of the window. Each
+parameter, was calculated both from the observations and the reference
+clear sky model, for each comparison. The allowable range of variation
+is defined by the model-derived value of the parameter multiplied by a
+factor plus an offset. The factors and the offsets were determined
+empirically, by manual inspecting each filters performance on selected
+days, and adjusting them accordingly, during an iterative process.
+
+\begin{enumerate}
+\def\labelenumi{\alph{enumi})}
+\tightlist
+\item
+  Mean of the measured \(\overline{\text{SDR}}_i\) (Eq.
+  \ref{eq:MeanVIP}). \begin{equation}
+  0.91 \times \overline{\text{SDR}}_{\text{CSref},i} - 20
+  < \overline{\text{SDR}}_i <
+  1.095 \times \overline{\text{SDR}}_{\text{CSref},i} + 30
+  \label{eq:MeanVIP}
+  \end{equation}
+\end{enumerate}
+
+\begin{enumerate}
+\def\labelenumi{\alph{enumi})}
+\setcounter{enumi}{1}
+\tightlist
+\item
+  Maximum measured value \(M_{\text{}}\) (Eq.~\ref{eq:MaxVIP}).
+  \begin{equation}
+  1 \times M_{\text{CSref},i} - 75
+  < M_{\text{}i} <
+  1 \times M_{\text{CSref},i} + 75
+  \label{eq:MaxVIP}
+  \end{equation}
+\end{enumerate}
+
+\begin{enumerate}
+\def\labelenumi{\alph{enumi})}
+\setcounter{enumi}{2}
+\tightlist
+\item
+  Length \(L_i\) of the sequential line segments, connecting the points
+  of the \(11\) SDR values (Eq. \ref{eq:VILeq}). \begin{equation}
+  L_i = \sum_{i=1}^{n-1}\sqrt{\left ( \text{SDR}_{i+1} - \text{SDR}_{i}\right )^2 + \left ( t_{i+1} - t_i \right )^2}
+  \label{eq:VILeq}
+  \end{equation} \begin{equation}
+  1 \times L_{\text{CSref},i} - 5 < L_i < 1.3 \times L_{\text{CSref},i} + 13
+  \label{eq:VILcr}
+  \end{equation} where: \(t_i\) is the time each SDR measurement has
+  been measured
+\end{enumerate}
+
+\begin{enumerate}
+\def\labelenumi{\alph{enumi})}
+\setcounter{enumi}{3}
+\tightlist
+\item
+  Standard deviation \(\sigma_i\) of the slope (\(s_i\)) between the
+  \(11\) sequential points, normalized by the mean
+  \(\overline{\text{SDR}}_i\) (Eq.~\ref{eq:VCT1}). \begin{gather}
+    \sigma_i = \frac {\sqrt{\frac{1}{n-1} \sum_{i=1}^{n-1} \left( s_i - \bar{s} \right)^2}} {\overline{\text{SDR}}_i} \label{eq:VCT1} \\
+    s_i = \frac{\text{SDR}_{i+1} - \text{SDR}_{i}}{t_{i+1} - t_i},\;\;   \bar{s} = \frac{1}{n-1} \sum_{i=1}^{n-1} s_i,\;\;\forall i \in \left \{ 1, 2, \ldots, n-1 \right \}\;\;
+  \end{gather} For this criterion, \(\sigma_i\) should be below a
+  certain threshold (Eq.~\ref{eq:VCTcr}): \begin{equation}
+    \sigma_i < \ensuremath{1.1\times 10^{-4}} \label{eq:VCTcr}
+  \end{equation}
+\end{enumerate}
+
+\begin{enumerate}
+\def\labelenumi{\alph{enumi})}
+\setcounter{enumi}{4}
+\tightlist
+\item
+  Maximum difference \(X_i\) between the change in measured irradiance
+  and the change in clear sky irradiance over each measurement interval.
+  \begin{gather}
+    X_i = \max{\left \{ \left | x_i - x_{\text{CSref},i} \right | \right \}} \label{eq:VSM3} \\
+    x_i = \text{SDR}_{i+1} - \text{SDR}_{i} \forall i \in \left \{ 1, 2, \ldots, n-1 \right \} \label{eq:VSM1} \\
+    x_{\text{CSref},i} = \text{SDR}_{\text{CSref},i+1} - \text{SDR}_{\text{CSref},i} \forall i \in \left \{ 1, 2, \ldots, n-1 \right \} \label{eq:VSM2}
+  \end{gather} For this criterion, \(X_i\) should be below a certain
+  threshold (Eq.~\ref{eq:VSMcr}): \begin{equation}
+    X_i < 7.5 \label{eq:VSMcr}
+  \end{equation}
+\end{enumerate}
+
+Due to a significant measurement uncertainty near the horizon, we have
+to exclude all measurements with SZA greater than \(85^\circ\).
+Moreover, due to some obstructions around the site (hills and
+buildings), we excluded data with Azimuth angle between \(35^\circ\) and
+\(120^\circ\) with SZA greater than \(80^\circ\). On the latter
+instances, Sun is systematically, not visible by the instrument's
+location. To make the measurements comparable throughout the dataset, we
+adjusted all one-minute radiometric values to the mean Sun - Earth
+distance. Subsequently, we made all measurements relative to the Total
+Solar Irradiance (TSI) at \(1\,\text{au}\), in order to compensate for
+the Sun's intensity variability, using a time series of satellite TSI
+observations. The TSI data we use are part of the ``NOAA Climate Data
+Record of Total Solar Irradiance'' dataset \citep{Coddington2005}. Where
+the initial daily values, were interpolated to match with the time step
+of our measurements. The final dataset contains \(6589967\) one-minute
+measurements, of which, \(84.2\%\) were identified as under clear-sky
+conditions and subsequently \(15.8\%\) as under cloud-sky conditions.
+
+In order to investigate the SDR trends, we implemented an appropriate
+aggregation scheme to the 1-minute data to derive a series in coarser
+timescale. To preserve the representativeness of the data we used the
+following criteria: a) for the daily mean values we accept days with
+more than 50\% of the daytime measurements, present and valid, b)
+monthly values were computed from daily means only when at least 20 days
+were available. To create the daily and monthly climatological means, we
+averaged the data based on the day of year and calendar month,
+respectively. For the seasonal means we averaged the mean daily values
+in each season (Winter: December - February, Spring: March - May, etc.).
+Finally, each data set was deseasonalized by subtracting the
+corresponding climatological annual cycle (daily or monthly) from the
+actual data. To estimate SZA contribution to the SDR trends, the
+one-minute data were aggregated in \(1^\circ\) SZA bins, separately for
+the morning and afternoon hours, and then were deseasonalized as
+mentioned above.
+
+\hypertarget{results}{%
+\section{Results}\label{results}}
+
+\hypertarget{long-term-sdr-trends}{%
+\subsection{Long-term SDR trends}\label{long-term-sdr-trends}}
+
 \hypertarget{version}{%
@@
 variables}\label{journal-specific-yaml-variables}}
-
-\startlandscape
-
-\begin{longtable}[t]{llllllll}
-\caption{\label{tab:mdpinames}MDPI journal names.}\\
-\toprule
-acoustics & biomedinformatics & dentistry & galaxies & jcp & metabolites & philosophies & socsci\\
-actuators & biomimetics & dermato & games & jcs & metals & photochem & software\\
-addictions & biomolecules & dermatopathology & gases & jcto & meteorology & photonics & soilsystems\\
-admsci & biophysica & designs & gastroent & jdb & methane & phycology & solar\\
-adolescents & biosensors & devices & gastrointestdisord & jeta & metrology & physchem & solids\\
-\addlinespace
-aerobiology & biotech & diabetology & gels & jfb & micro & physics & spectroscj\\
-aerospace & birds & diagnostics & genealogy & jfmk & microarrays & physiologia & sports\\
-agriculture & bloods & dietetics & genes & jimaging & microbiolres & plants & standards\\
-agriengineering & blsf & digital & geographies & jintelligence & micromachines & plasma & stats\\
-agrochemicals & brainsci & disabilities & geohazards & jlpea & microorganisms & platforms & std\\
-\addlinespace
-agronomy & breath & diseases & geomatics & jmmp & microplastics & pollutants & stresses\\
-ai & buildings & diversity & geosciences & jmp & minerals & polymers & surfaces\\
-air & businesses & dna & geotechnics & jmse & mining & polysaccharides & surgeries\\
-algorithms & cancers & drones & geriatrics & jne & modelling & poultry & suschem\\
-allergies & carbon & dynamics & grasses & jnt & molbank & powders & sustainability\\
-\addlinespace
-alloys & cardiogenetics & earth & gucdd & jof & molecules & preprints & symmetry\\
-analytica & catalysts & ebj & hazardousmatters & joitmc & mps & proceedings & synbio\\
-analytics & cells & ecologies & healthcare & jor & msf & processes & systems\\
-anatomia & ceramics & econometrics & hearts & journalmedia & mti & prosthesis & targets\\
-animals & challenges & economies & hemato & jox & muscles & proteomes & taxonomy\\
-\addlinespace
-antibiotics & chemengineering & education & hematolrep & jpm & nanoenergyadv & psf & technologies\\
-antibodies & chemistry & ejihpe & heritage & jrfm & nanomanufacturing & psych & telecom\\
-antioxidants & chemosensors & electricity & higheredu & jsan & nanomaterials & psychiatryint & test\\
-applbiosci & chemproc & electrochem & highthroughput & jtaer & ncrna & psychoactives & textiles\\
-appliedchem & children & electronicmat & histories & jvd & ndt & publications & thalassrep\\
-\addlinespace
-appliedmath & chips & electronics & horticulturae & jzbg & network & quantumrep & thermo\\
-applmech & cimb & encyclopedia & hospitals & kidneydial & neuroglia & quaternary & tomography\\
-applmicrobiol & civileng & endocrines & humanities & kinasesphosphatases & neurolint & qubs & tourismhosp\\
-applnano & cleantechnol & energies & humans & knowledge & neurosci & radiation & toxics\\
-applsci & climate & eng & hydrobiology & land & nitrogen & reactions & toxins\\
-\addlinespace
-aquacj & clinpract & engproc & hydrogen & languages & notspecified & receptors & transplantology\\
-architecture & clockssleep & entomology & hydrology & laws & nri & recycling & transportation\\
-arm & cmd & entropy & hygiene & life & nursrep & regeneration & traumacare\\
-arthropoda & coasts & environments & idr & liquids & nutraceuticals & religions & traumas\\
-arts & coatings & environsciproc & ijerph & literature & nutrients & remotesensing & tropicalmed\\
-\addlinespace
-asc & colloids & epidemiologia & ijfs & livers & obesities & reports & universe\\
-asi & colorants & epigenomes & ijgi & logics & oceans & reprodmed & urbansci\\
-astronomy & commodities & est & ijms & logistics & ohbm & resources & uro\\
-atmosphere & compounds & fermentation & ijns & lubricants & onco & rheumato & vaccines\\
-atoms & computation & fibers & ijpb & lymphatics & oncopathology & risks & vehicles\\
-\addlinespace
-audiolres & computers & fintech & ijtm & machines & optics & robotics & venereology\\
-automation & condensedmatter & fire & ijtpp & macromol & oral & ruminants & vetsci\\
-axioms & conservation & fishes & ime & magnetism & organics & safety & vibration\\
-bacteria & constrmater & fluids & immuno & magnetochemistry & organoids & sci & virtualworlds\\
-batteries & cosmetics & foods & informatics & make & osteology & scipharm & viruses\\
-\addlinespace
-bdcc & covid & forecasting & information & marinedrugs & oxygen & sclerosis & vision\\
-behavsci & crops & forensicsci & infrastructures & materials & parasites & seeds & waste\\
-beverages & cryptography & forests & inorganics & materproc & parasitologia & sensors & water\\
-biochem & crystals & foundations & insects & mathematics & particles & separations & wem\\
-bioengineering & csmf & fractalfract & instruments & mca & pathogens & sexes & wevj\\
-\addlinespace
-biologics & ctn & fuels & inventions & measurements & pathophysiology & signals & wind\\
-biology & curroncol & future & iot & medicina & pediatrrep & sinusitis & women\\
-biomass & cyber & futureinternet & j & medicines & pharmaceuticals & skins & world\\
-biomechanics & dairy & futurepharmacol & jal & medsci & pharmaceutics & smartcities & youth\\
-biomed & data & futurephys & jcdd & membranes & pharmacoepidemiology & sna & zoonoticdis\\
-\addlinespace
-biomedicines & ddc & futuretransp & jcm & merits & pharmacy & societies & \\
-\bottomrule
-\end{longtable}
-\finishlandscape
 
@@
 
-The text continues here. Proofs must be formatted as follows:
-
-Example of a proof:
-
-\begin{proof}[Proof of Theorem1]
-Text of the proof. Note that the phrase ``of Theorem 1'\,' is optional
-if it is clear which theorem is being referred to.
-
-\end{proof}
-
 The text continues here.
@@
 discussion is unusually long or complex.
-
-\hypertarget{patents}{%
-\section{Patents}\label{patents}}
-
-This section is not mandatory, but may be added if there are patents
-resulting from the work reported in this manuscript.
 
@@
 \reftitle{References}
-\bibliography{references.bib}
+\bibliography{manualreferences.bib}
 

### 2023-11-02 MDPI_submition.Rmd
Commit 2023-11-02 06:14

@@
 
-For completeness, we will provide below a brief overview of the clear sky identification algorithm (CSid), along with the site specific thresholds.
+For completeness, we will provide below a brief overview of the 
+<!--
+[....Csid term is referenced above....]{.comment-start id="0" author="A N"
+date="2023-07-15T20:00:52Z"} 
+-->
+clear sky identification algorithm (CSid),
+<!--
+[]{.comment-end id="0"} 
+-->
+along with the site specific thresholds.
 To calculate the reference clear sky
@@
 
-The criteria that were used to identify whether a measurement was taken under clear-sky conditions are presented below.
-A data point is flagged as "clear-sky" if all criteria are satisfied, otherwise it is considered to be "cloud-sky".
-Each criterion was applied for a running window of $`r nt`$ consecutive one-minute measurements, and the characterization is assigned to the central value of the window.
-Each parameter, was calculated both from the observations and the reference clear sky model, for each comparison.
-The allowable range of variation is defined by the model-derived value of the parameter multiplied by a factor plus an offset. 
-The factors and the offsets were determined empirically, by manual inspecting each filters performance on selected days, and adjusting them accordingly, during an iterative process.
+The criteria that were used to identify whether a measurement was taken
+under clear-sky conditions are presented below.
+A data point is flagged
+as "clear-sky" if all criteria are satisfied, otherwise it is considered to be "cloud-sky".
+Each criterion was applied
+for a running window of $`r nt`$ consecutive one-minute measurements, and
+the characterization is assigned to the central value of the window.
+Each parameter, was calculated both from the observations and the
+reference clear sky model, for each comparison.
+The allowable range of variation is defined by the
+model-derived value of the parameter multiplied by a factor plus an
+offset. 
+<!--[.... .expanded the description ....]{.comment-start id="4"
+author="A N" date="2023-07-15T22:48:13Z"}
+-->
+The factors and the offsets were
+determined empirically, by manual inspecting each filters performance on
+selected days, and adjusting them accordingly, during an iterative
+process.
+<!--[[]{.comment-end id="4"}]{.insertion author="A N"
+date="2023-07-15T22:48:13Z"}
+-->
 
@@
 
+```{r echo=F, eval=TRUE}
+LT <- data.table(read.csv("./figures/tbl_longterm_trends.dat",
+               strip.white = TRUE,
+               sep = ";",
+               header = TRUE))
+LT <- LT[grep("SDR", LT$var), ]
+LT$slope.sd           <- NULL
+LT$slope.ConfInt_0.95 <- NULL
+LT$slope.ConfInt_0.99 <- NULL
+LT$var                <- NULL
+# LT$N                  <- NULL
+LT$Obs                <- NULL
+setorder(LT, -slope.stat_sig)
+rownames(LT) <- NULL
+names(LT)    <- c("Trend [%/year]", 
+                  "Trend p-value", 
+                  "Sky conditions",
+                  "Days",
+                  "Statistical signif. [%]")
+## no need all are 100%
+LTminSSig <- signif(min(LT$`Statistical signif. [%]`), 4)
+# LT$`Trend statistical signif. [%]` <- NULL
+## numeric precision
+LT$`Trend [%/year]` <- signif(LT$`Trend [%/year]`, 3)
+LT$`Trend p-value`  <- signif(LT$`Trend p-value`,  3)
+
+## some vars to use later
+alltrend   <- signif(LT[`Sky conditions` == "All sky cond.",    `Trend [%/year]`], 2)
+cleartrend <- signif(LT[`Sky conditions` == "Clear sky cond.",  `Trend [%/year]`], 2)
+cloudtrend <- signif(LT[`Sky conditions` == "Cloudy sky cond.", `Trend [%/year]`], 2)
+## rename sky conditions
+LT$`Sky conditions` <- sub(" cond.", "", LT$`Sky conditions`)
+
+## order columns
+LT1 <- LT[,c(
+      "Sky conditions",
+      "Trend [%/year]",
+      "Trend p-value",
+      "Days"
+      )]
+
+## order columns
+LT2 <- LT[,c(
+      "Sky conditions",
+      "Trend [%/year]",
+      "Statistical signif. [%]",
+      "Days"
+      )]
+```
+
+We calculated the linear SDR trends, from the departures of the mean daily values from the seasonal values, and for the three sky conditions (Table\ \ref{tab:trendtable}).
+In Figure\ \ref{fig:trendALL} we present only the time series under all-sky conditions; the plots for clear-sky and cloud-sky conditions, are very similar and are shown in the Appendix (Figures\ \ref{fig:trendCLEAR} and\  \ref{fig:trendCLOUD}).
+We observe a positive trend for all-sky conditions 
+($`r alltrend`\,\%/y$), a very close but lower trend for clear-skies ($`r cleartrend`\,\%/y$) and a negative weaker trend for cloud-skies ($`r cloudtrend`\,\%/y$).
+In the studied period, there is no significant break or change in the variability of the time series.
+Other studies for the European region report a change of the SDR slope, around 1980, a few years before the start of our records [@Wild2021; @Yuan2021; @Ohmura2009].
+It is interesting to note, that for the observation period, the trend of the TSI is
+$`r format(TSIslopeAll[1,1], scientific = T, digits = 2)`\,\%/y$, and thus we can not attribute any major effect on SDR trend to Solar variability.
+
+<!-- \footnotesize -->
+<!-- ```{r trendtable2, echo=F} -->
+<!-- panderOptions("table.alignment.default", "center") -->
+<!-- pander(LT1, -->
+<!--        caption = paste0("\\label{tab:trendtable2}Trends in SDR daily mean for different sky conditions for the period", year(min(ALL_1_daily_DESEAS$Date)), " - ", year(max(ALL_1_daily_DESEAS$Date)), ".
+For all cases the statistical significance is above $", LTminSSig, "\\,\\%$.")) -->
+<!-- ``` -->
+<!-- \normalsize -->
+
+```{r trendtable, echo=F}
+x_latex <- knitr::kable(LT2,
+                        format   = "latex",
+                        booktabs = T,
+                        escape   = T,
+                        format.args = list(digits = 4),
+                        linesep  = "",
+                        align    = "crr",
+                        caption  = paste0("\\label{tab:trendtable}Trends in SDR daily means for different sky conditions for the period ", year(min(ALL_1_daily_DESEAS$Date)), " - ", year(max(ALL_1_daily_DESEAS$Date)), "."))
+x_latex %>%
+    kable_styling(
+        latex_options = c("hold_position")
+    )
+        # position      =   "float_left"
+```
+
+```{r trendALL, echo=F, out.width='.70\\linewidth', fig.cap=paste0("Anomaly (\\%) of the daily all-sky SDR, relative to climatological values for ", year(min(ALL_1_daily_DESEAS$Date)), " - ", year(max(ALL_1_daily_DESEAS$Date)), ".
+The black line shows the long term linear trend.")}
+knitr::include_graphics("./images/LongtermTrends-2.png")
+```
+
+Although the year-to-year variability of the anomalies (Figure \ref{fig:trendALL} and Figures \ref{fig:trendCLEAR}, \ref{fig:trendCLOUD} on Appendix), shows a rather homogeneous behavior, plots of the cumulative sums (CUSUM) of the anomalies can reveal different structures within all the three sky conditions [@Regier2019].
+In the cases of all-sky and clear-sky (Figures \ref{fig:cusummonth-1} and \ref{fig:cusummonth-2}), we observe three macroscopic periods.
+A downward part from the start until about 2005, an relative steady part until about 2016 and finally, a steep upward part until the present.
+For cloud-sky (Figure\ \ref{fig:cusummonth-3}), we have a different sequence, it begins with a relative steady part until 1997, followed by a upward part until 2005, a long decline until 2020, with a small positive slope until the present.
+For a uniform trend, we would expect the CUSUMs of the anomalies to have a symmetric 'V' shape.
+This would indicate that the anomalies are evenly distributed around the climatological mean, and for a positive uniform trend, the first half to be below and the other half above the climatological mean.
+In our case, there is more complex evolution of the phenomenon.
+Another distinct feature of the CUSUMs, is the different pattern of the cloudy-sky dataset which peaks around the middle of the period.
+Although, it seems to exist a complementary relation to the CUSUMs of the clear- and all-sky cases, we can not assert that clouds are the main drive for this relation, due to the great difference in the number of observational data between the datasets (Table\ \ref{tab:trendtable}).
+
+```{r cusummonth, echo = FALSE, fig.cap='Running cumulative sum plots of the monthly SDR.', fig.subcap=c('All-skies.', 'Clear-skies.', 'Cloud-skies.'), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3}
+knitr::include_graphics(c(
+    "./images/CumulativeMonthlyCuSum-1.png",    # ALL sky
+    "./images/CumulativeMonthlyCuSum-5.png",    # CLEAR sky
+    "./images/CumulativeMonthlyCuSum-9.png"))   # Cloud sky
+```
+
+In order to investigate more, the features within the trends, we created another set of CUSUMs plots by subtracting the corresponding long term trend from the SDR anomaly data, prior to the CUSUM calculation (Figure\ \ref{fig:cusumnotrendmonthly}).
+With this approach periods when the CUSUMs diverge from zero can be interpreted as systematic variation of SDR from the climatological mean.
+When the CUSUM is increasing, the added values are above the climatological values of SDR trend and vice versa.
+Overall, for all- and clear-sky conditions (Figure\ \ref{fig:cusumnotrendmonthly-1} and\ \ref{fig:cusumnotrendmonthly-2}) we observe periods when the anomalies diverge from the climatological value, each lasting for several years.
+The pattern for the all-sky and clear-sky conditions is very similar, suggesting prevalence in clear skies over Thessaloniki.
+It is interesting that in the period 1993 - 2016 the anomalies of all-sky and clear-sky, have a high variability around zero.
+After 2016, the range of the variability is decreased at about one third of the prior period.
+For cloud-sky conditions (Figure\ \ref{fig:cusumnotrendmonthly-3}) the period 1997 - 2008 is dominated by positive CUSUMs suggesting a reduced effect of clouds on SDR.
+From 1997 to mid-2000s CUSUMs are increasing, likely due to a continuous decrease in the thickness of clouds, followed by a period of rapid increase (within 3 years) in cloud optical thickness lasting up to 2008.
+The following stable period spans for about 15 years up to 2021 when CUSUMs start increasing again.
+
+```{r cusumnotrendmonthly, echo = FALSE, fig.cap='Running cumulative sum plots of the monthly SDR after removing the long term trend from the monthly values.', fig.subcap=c('All-skies.', 'Clear-skies.', 'Cloud-skies.'), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3}
+knitr::include_graphics(c(
+    "./images/CumulativeMonthlyCuSumNOtrend-1.png",    # ALL sky
+    "./images/CumulativeMonthlyCuSumNOtrend-5.png",    # CLEAR sky
+    "./images/CumulativeMonthlyCuSumNOtrend-9.png"))   # Cloud sky
+```
+
 # Version
@@
 ## Journal Specific YAML variables
+
+\startlandscape
+```{r, mdpinames, echo=FALSE, results='asis'}
+mdpi_journals <- "acoustics, actuators, addictions, admsci, adolescents, aerobiology, aerospace, agriculture, agriengineering, agrochemicals, agronomy, ai, air, algorithms,  jmp, jmse, jne, jnt, jof, joitmc, jor, journalmedia, jox, jpm, jrfm, jsan, jtaer, jvd,  wem, wevj, wind, women, world, youth, zoonoticdis"
+
+knitr::kable(rticles::string_to_table(mdpi_journals, 8),
+             col.names = NULL,
+             format = "latex",
+             booktabs = TRUE, 
+             caption = "MDPI journal names.",
+             longtable = TRUE)
+
+```
+\finishlandscape
 

### 2023-11-02 MDPI_submition.tex
Commit 2023-11-02 06:14

@@
 
+We calculated the linear SDR trends, from the departures of the mean
+daily values from the seasonal values, and for the three sky conditions
+(Table~\ref{tab:trendtable}). In Figure~\ref{fig:trendALL} we present
+only the time series under all-sky conditions; the plots for clear-sky
+and cloud-sky conditions, are very similar and are shown in the Appendix
+(Figures~\ref{fig:trendCLEAR} and~ \ref{fig:trendCLOUD}). We observe a
+positive trend for all-sky conditions (\(0.38\,\%/y\)), a very close but
+lower trend for clear-skies (\(0.35\,\%/y\)) and a negative weaker trend
+for cloud-skies (\(-0.28\,\%/y\)). In the studied period, there is no
+significant break or change in the variability of the time series. Other
+studies for the European region report a change of the SDR slope, around
+1980, a few years before the start of our records
+\citep{Wild2021, Yuan2021, Ohmura2009}. It is interesting to note, that
+for the observation period, the trend of the TSI is \(-2e-04\,\%/y\),
+and thus we can not attribute any major effect on SDR trend to Solar
+variability.
+
+\begin{table}[!h]
+
+\caption{\label{tab:trendtable}\label{tab:trendtable}Trends in SDR daily means for different sky conditions for the period 1993 - 2023.}
+\centering
+\begin{tabular}[t]{crrc}
+\toprule
+Sky conditions & Trend [\%/year] & Statistical signif. [\%] & Days\\
+\midrule
+All sky & 0.376 & 100.00 & 10256\\
+Clear sky & 0.349 & 100.00 & 10256\\
+Cloudy sky & -0.276 & 99.97 & 5067\\
+\bottomrule
+\end{tabular}
+\end{table}
+
+\begin{figure}[h!]
+
+{\centering \includegraphics[width=.70\linewidth]{./images/LongtermTrends-2}
+
+}
+
+\caption{Anomaly (\%) of the daily all-sky SDR, relative to climatological values for 1993 - 2023. The black line shows the long term linear trend.}\label{fig:trendALL}
+\end{figure}
+
+Although the year-to-year variability of the anomalies (Figure
+\ref{fig:trendALL} and Figures \ref{fig:trendCLEAR},
+\ref{fig:trendCLOUD} on Appendix), shows a rather homogeneous behavior,
+plots of the cumulative sums (CUSUM) of the anomalies can reveal
+different structures within all the three sky conditions
+\citep{Regier2019}. In the cases of all-sky and clear-sky (Figures
+\ref{fig:cusummonth-1} and \ref{fig:cusummonth-2}), we observe three
+macroscopic periods. A downward part from the start until about 2005, an
+relative steady part until about 2016 and finally, a steep upward part
+until the present. For cloud-sky (Figure~\ref{fig:cusummonth-3}), we
+have a different sequence, it begins with a relative steady part until
+1997, followed by a upward part until 2005, a long decline until 2020,
+with a small positive slope until the present. For a uniform trend, we
+would expect the CUSUMs of the anomalies to have a symmetric `V' shape.
+This would indicate that the anomalies are evenly distributed around the
+climatological mean, and for a positive uniform trend, the first half to
+be below and the other half above the climatological mean. In our case,
+there is more complex evolution of the phenomenon. Another distinct
+feature of the CUSUMs, is the different pattern of the cloudy-sky
+dataset which peaks around the middle of the period. Although, it seems
+to exist a complementary relation to the CUSUMs of the clear- and
+all-sky cases, we can not assert that clouds are the main drive for this
+relation, due to the great difference in the number of observational
+data between the datasets (Table~\ref{tab:trendtable}).
+
+\begin{figure}[h!]
+
+{\centering \subfloat[All-skies.\label{fig:cusummonth-1}]{\includegraphics[width=.32\linewidth]{./images/CumulativeMonthlyCuSum-1} }\subfloat[Clear-skies.\label{fig:cusummonth-2}]{\includegraphics[width=.32\linewidth]{./images/CumulativeMonthlyCuSum-5} }\subfloat[Cloud-skies.\label{fig:cusummonth-3}]{\includegraphics[width=.32\linewidth]{./images/CumulativeMonthlyCuSum-9} }
+
+}
+
+\caption{Running cumulative sum plots of the monthly SDR.}\label{fig:cusummonth}
+\end{figure}
+
+In order to investigate more, the features within the trends, we created
+another set of CUSUMs plots by subtracting the corresponding long term
+trend from the SDR anomaly data, prior to the CUSUM calculation
+(Figure~\ref{fig:cusumnotrendmonthly}). With this approach periods when
+the CUSUMs diverge from zero can be interpreted as systematic variation
+of SDR from the climatological mean. When the CUSUM is increasing, the
+added values are above the climatological values of SDR trend and vice
+versa. Overall, for all- and clear-sky conditions
+(Figure~\ref{fig:cusumnotrendmonthly-1}
+and~\ref{fig:cusumnotrendmonthly-2}) we observe periods when the
+anomalies diverge from the climatological value, each lasting for
+several years. The pattern for the all-sky and clear-sky conditions is
+very similar, suggesting prevalence in clear skies over Thessaloniki. It
+is interesting that in the period 1993 - 2016 the anomalies of all-sky
+and clear-sky, have a high variability around zero. After 2016, the
+range of the variability is decreased at about one third of the prior
+period. For cloud-sky conditions
+(Figure~\ref{fig:cusumnotrendmonthly-3}) the period 1997 - 2008 is
+dominated by positive CUSUMs suggesting a reduced effect of clouds on
+SDR. From 1997 to mid-2000s CUSUMs are increasing, likely due to a
+continuous decrease in the thickness of clouds, followed by a period of
+rapid increase (within 3 years) in cloud optical thickness lasting up to
+2008. The following stable period spans for about 15 years up to 2021
+when CUSUMs start increasing again.
+
+\begin{figure}[h!]
+
+{\centering \subfloat[All-skies.\label{fig:cusumnotrendmonthly-1}]{\includegraphics[width=.32\linewidth]{./images/CumulativeMonthlyCuSumNOtrend-1} }\subfloat[Clear-skies.\label{fig:cusumnotrendmonthly-2}]{\includegraphics[width=.32\linewidth]{./images/CumulativeMonthlyCuSumNOtrend-5} }\subfloat[Cloud-skies.\label{fig:cusumnotrendmonthly-3}]{\includegraphics[width=.32\linewidth]{./images/CumulativeMonthlyCuSumNOtrend-9} }
+
+}
+
+\caption{Running cumulative sum plots of the monthly SDR after removing the long term trend from the monthly values.}\label{fig:cusumnotrendmonthly}
+\end{figure}
+
 \hypertarget{version}{%
@@
 variables}\label{journal-specific-yaml-variables}}
+
+\startlandscape
+
+\begin{longtable}[t]{llllllll}
+\caption{\label{tab:mdpinames}MDPI journal names.}\\
+\toprule
+acoustics & aerobiology & agronomy & jmse & jor & jsan & wind & \\
+actuators & aerospace & ai & jne & journalmedia & jtaer & women & \\
+addictions & agriculture & air & jnt & jox & jvd & world & \\
+admsci & agriengineering & algorithms & jof & jpm & wem & youth & \\
+adolescents & agrochemicals & jmp & joitmc & jrfm & wevj & zoonoticdis & \\
+\bottomrule
+\end{longtable}
+\finishlandscape
 

### 2023-11-02 MDPI_submition.Rmd
Commit 2023-11-02 07:09

@@
     long: Solar downward radiation
+header-includes:
+   # - \usepackage{subfig}
+   - \usepackage{subcaption}
+   - \captionsetup[sub]{position=bottom, labelfont={bf, small, stretch=1.17}, labelsep=space, textfont={small, stretch=1.5}, aboveskip=6pt,  belowskip=6pt, singlelinecheck=off, justification=justified}
 bibliography: manualreferences.bib
@@
 
-In order to investigate more, the features within the trends, we created another set of CUSUMs plots by subtracting the corresponding long term trend from the SDR anomaly data, prior to the CUSUM calculation (Figure\ \ref{fig:cusumnotrendmonthly}).
-With this approach periods when the CUSUMs diverge from zero can be interpreted as systematic variation of SDR from the climatological mean.
-When the CUSUM is increasing, the added values are above the climatological values of SDR trend and vice versa.
-Overall, for all- and clear-sky conditions (Figure\ \ref{fig:cusumnotrendmonthly-1} and\ \ref{fig:cusumnotrendmonthly-2}) we observe periods when the anomalies diverge from the climatological value, each lasting for several years.
-The pattern for the all-sky and clear-sky conditions is very similar, suggesting prevalence in clear skies over Thessaloniki.
-It is interesting that in the period 1993 - 2016 the anomalies of all-sky and clear-sky, have a high variability around zero.
-After 2016, the range of the variability is decreased at about one third of the prior period.
-For cloud-sky conditions (Figure\ \ref{fig:cusumnotrendmonthly-3}) the period 1997 - 2008 is dominated by positive CUSUMs suggesting a reduced effect of clouds on SDR.
-From 1997 to mid-2000s CUSUMs are increasing, likely due to a continuous decrease in the thickness of clouds, followed by a period of rapid increase (within 3 years) in cloud optical thickness lasting up to 2008.
-The following stable period spans for about 15 years up to 2021 when CUSUMs start increasing again.
-
-```{r cusumnotrendmonthly, echo = FALSE, fig.cap='Running cumulative sum plots of the monthly SDR after removing the long term trend from the monthly values.', fig.subcap=c('All-skies.', 'Clear-skies.', 'Cloud-skies.'), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3}
-knitr::include_graphics(c(
-    "./images/CumulativeMonthlyCuSumNOtrend-1.png",    # ALL sky
-    "./images/CumulativeMonthlyCuSumNOtrend-5.png",    # CLEAR sky
-    "./images/CumulativeMonthlyCuSumNOtrend-9.png"))   # Cloud sky
-```
-
 # Version

### 2023-11-02 MDPI_submition.tex
Commit 2023-11-02 07:09

@@
 
+\usepackage{subcaption}
+\captionsetup[sub]{position=bottom, labelfont={bf, small, stretch=1.17}, labelsep=space, textfont={small, stretch=1.5}, aboveskip=6pt,  belowskip=6pt, singlelinecheck=off, justification=justified}
 \usepackage{longtable}
@@
 
-In order to investigate more, the features within the trends, we created
-another set of CUSUMs plots by subtracting the corresponding long term
-trend from the SDR anomaly data, prior to the CUSUM calculation
-(Figure~\ref{fig:cusumnotrendmonthly}). With this approach periods when
-the CUSUMs diverge from zero can be interpreted as systematic variation
-of SDR from the climatological mean. When the CUSUM is increasing, the
-added values are above the climatological values of SDR trend and vice
-versa. Overall, for all- and clear-sky conditions
-(Figure~\ref{fig:cusumnotrendmonthly-1}
-and~\ref{fig:cusumnotrendmonthly-2}) we observe periods when the
-anomalies diverge from the climatological value, each lasting for
-several years. The pattern for the all-sky and clear-sky conditions is
-very similar, suggesting prevalence in clear skies over Thessaloniki. It
-is interesting that in the period 1993 - 2016 the anomalies of all-sky
-and clear-sky, have a high variability around zero. After 2016, the
-range of the variability is decreased at about one third of the prior
-period. For cloud-sky conditions
-(Figure~\ref{fig:cusumnotrendmonthly-3}) the period 1997 - 2008 is
-dominated by positive CUSUMs suggesting a reduced effect of clouds on
-SDR. From 1997 to mid-2000s CUSUMs are increasing, likely due to a
-continuous decrease in the thickness of clouds, followed by a period of
-rapid increase (within 3 years) in cloud optical thickness lasting up to
-2008. The following stable period spans for about 15 years up to 2021
-when CUSUMs start increasing again.
-
-\begin{figure}[h!]
-
-{\centering \subfloat[All-skies.\label{fig:cusumnotrendmonthly-1}]{\includegraphics[width=.32\linewidth]{./images/CumulativeMonthlyCuSumNOtrend-1} }\subfloat[Clear-skies.\label{fig:cusumnotrendmonthly-2}]{\includegraphics[width=.32\linewidth]{./images/CumulativeMonthlyCuSumNOtrend-5} }\subfloat[Cloud-skies.\label{fig:cusumnotrendmonthly-3}]{\includegraphics[width=.32\linewidth]{./images/CumulativeMonthlyCuSumNOtrend-9} }
-
-}
-
-\caption{Running cumulative sum plots of the monthly SDR after removing the long term trend from the monthly values.}\label{fig:cusumnotrendmonthly}
-\end{figure}
-
 \hypertarget{version}{%

### 2023-11-02 MDPI_submition.Rmd
Commit 2023-11-02 10:14

--- 2023-11-02 MDPI_submition.Rmd
@@
                   "Sky conditions",
-                  "Days",
+                  "Days with data",
                   "Statistical signif. [%]")
@@
       "Trend p-value",
-      "Days"
+      "Days with data"
       )]
@@
       "Statistical signif. [%]",
-      "Days"
+      "Days with data"
       )]
@@
                         format      = "latex",
-                        booktabs = T,
-                        escape   = T,
+                        booktabs    = TRUE,
+                        escape      = TRUE,
                         format.args = list(digits = 4),
                         linesep     = "",
-                        align    = "crr",
+                        align       = "crrr",
+                        centering   = FALSE,
+                        table.envir = "table",
+                        position    = "H",
                         caption     = paste0("\\label{tab:trendtable}Trends in SDR daily means for different sky conditions for the period ", year(min(ALL_1_daily_DESEAS$Date)), " - ", year(max(ALL_1_daily_DESEAS$Date)), "."))
+
+# x_latex %>%
+#     kable_styling(
+#         latex_options = c("hold_position")
+#     )
 x_latex %>% 
-    kable_styling(
-        latex_options = c("hold_position")
-    )
-        # position      =   "float_left"
+    kable_styling(full_width = TRUE) %>%
+    column_spec(1, width = "8em")
 ```
@@
 
-# Version
+<!-- ```{r cusum, echo = FALSE, fig.cap='Running cumulative sum plots of the daily SDR.', fig.subcap=c('All-sky conditions.', 'Clear-sky conditions.', 'Cloud-sky conditions.'), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3} -->
+<!-- knitr::include_graphics(c( -->
+<!--     "./images/CumulativeDailyCuSum-1.png",    # ALL sky -->
+<!--     "./images/CumulativeDailyCuSum-5.png",    # CLEAR sky -->
+<!--     "./images/CumulativeDailyCuSum-9.png"))   # Cloud sky -->
+<!-- ``` -->
 
-This Rmd-skeleton uses the mdpi Latex template published 2023-03-25. 
-However, the official template gets more frequently updated than the **rticles**
-package. Therefore, please make sure prior to paper submission, that you're 
-using the most recent .cls, .tex and .bst files 
-(available [here](http://www.mdpi.com/authors/latex)).
+<!-- ```{r cusumnotrend, echo = FALSE, fig.cap='Running cumulative sum plots of the daily SDR after removing the long term trend from the daily values.', fig.subcap=c('All-skies.', 'Clear-skies.', 'Cloud-skies.'), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3} -->
+<!-- knitr::include_graphics(c( -->
+<!--     "./images/CumulativeDailyCuSumNOtrend-1.png",    # ALL sky -->
+<!--     "./images/CumulativeDailyCuSumNOtrend-5.png",    # CLEAR sky -->
+<!--     "./images/CumulativeDailyCuSumNOtrend-9.png"))   # Cloud sky -->
+<!-- ``` -->
 
-# Article Header Information
+<!-- ```{r cusumSTD, echo = FALSE, fig.cap='two plots', fig.subcap=c('one plot', 'the other one'), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3} -->
+<!-- knitr::include_graphics(c( -->
+<!--     "./images/CumulativeDailyCuSumSTD-1.png",    # ALL sky -->
+<!--     "./images/CumulativeDailyCuSumSTD-2.png",    # CLEAR sky -->
+<!--     "./images/CumulativeDailyCuSumSTD-3.png"))   # Cloud sky -->
+<!-- ``` -->
 
-## Journal Specific YAML variables
+## Effects of the solar zenith angle on SDR.
 
-\startlandscape
-```{r, mdpinames, echo=FALSE, results='asis'}
-mdpi_journals <- "acoustics, actuators, addictions, admsci, adolescents, aerobiology, aerospace, agriculture, agriengineering, agrochemicals, agronomy, ai, air, algorithms,  jmp, jmse, jne, jnt, jof, joitmc, jor, journalmedia, jox, jpm, jrfm, jsan, jtaer, jvd,  wem, wevj, wind, women, world, youth, zoonoticdis"
+The solar zenith angle is a major driver of the attenuation of SDR reaching the ground, as SDR is affected by the enhancement of the radiation path in the atmosphere, especially in urban environment with human activities emitting aerosols [@Wang2021].
+In order to estimate the effect of the SZA on the SDR trends, we grouped the anomaly data in bins of $1^\circ$ SZA, and calculated the overall trend for each bin before noon and after noon (Figure\ \ref{fig:szatrends}).
+Although there are seasonal dependencies of the minimum SZA (see Appendix, Figure\ \ref{fig:SZAtrendSeason}), these dependencies would not be further examined here.
+For all-sky and clear-sky conditions the brightening effect of SDR (positive trend) is stronger for large SZAs (Figures\ \ref{fig:szatrends-1} and \ref{fig:szatrends-2}).
+The trends in the morning and afternoon hours are more or less consistent with small differences, which can be attributed to systematic diurnal variations of aerosols, particularly during the warm period of the year [@Wang2021].
+For cloud-sky conditions (Figure\ \ref{fig:szatrends-3}), we can not discern any significant dependence of the SDR trend with SZA.
+For SZAs $16^\circ$ - $50^\circ$, the trends range within about $\pm 0.2\,\%/y$, with a weak statistical significance.
+Between $50^\circ$ and $75^\circ$ SZA the trends for the period before noon are stronger and negative, possibly associated with stronger attenuation by clouds under oblique incidence angles.
 
-knitr::kable(rticles::string_to_table(mdpi_journals, 8),
-             col.names = NULL,
+```{r szatrends, echo = FALSE, fig.cap='Distribution of the SDR\'s long term trends by SZA before and after noon.
+The solid data points represent values with an acceptable statistical significance ($p<0.005$), the trends were computed with SZA bins of $1^\\circ$ for each day, before and after local noon.', fig.subcap=c('All-sky conditions.', 'Clear-sky conditions.', 'Cloud-sky conditions.'), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3}
+knitr::include_graphics("./images/SzaTrends-1.png")
+knitr::include_graphics("./images/SzaTrends-4.png" )
+knitr::include_graphics("./images/SzaTrends-7.png")
+```
+
+<!-- ```{r szatrendsmonthly, echo = FALSE, fig.cap='*** Montlhly *** Distribution of the SDR\'s long term trends by SZA before and after noon.
+The solid data points represent values with an acceptable statistical significance ($p<0.005$).', fig.subcap=c('All-sky conditions.', 'Clear-sky conditions.', 'Cloud-sky conditions.'), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3} -->
+<!-- knitr::include_graphics("./images/SzaTrendsMonthly-1.png") -->
+<!-- knitr::include_graphics("./images/SzaTrendsMonthly-4.png" ) -->
+<!-- knitr::include_graphics("./images/SzaTrendsMonthly-7.png") -->
+<!-- ``` -->
+
+## Long term trends by season of year
+
+```{r echo=F, eval=TRUE}
+## Prepare table and data
+spp <- read.csv("./figures/tbl_longterm_trends_season.dat",
+               strip.white = TRUE,
+               sep         = ";",
+               header      = TRUE)
+spp <- spp[grep("SDR" ,spp$var), ]
+spp$slope.sd           <- NULL
+spp$slope.ConfInt_0.95 <- NULL
+spp$slope.ConfInt_0.99 <- NULL
+spp$var                <- NULL
+rownames(spp)          <- NULL
+names(spp) <- c("Trend [%/year]", "Trend p-value", "Sky condition", "Season", "Statistical signif. [%]")
+## order by season
+target    <- c("Winter", "Spring", "Summer", "Autumn")
+new_order <- sapply(target, function(x,spp){which(spp$Season == x)}, spp = spp)
+spp        <- spp[new_order,]
+## order by sky condition
+spp <- spp[order(spp$`Sky condition`),]
+
+# ## put all sky first
+# A  <- which(spp$`Sky condition` == "All sky cond.")
+# B  <- which(spp$`Sky condition` != "All sky cond.")
+# spp <- spp[c(A,B),]
+## numeric precision
+spp$`Trend [%/year]`          <- signif(spp$`Trend [%/year]`, 3)
+spp$`Trend p-value`           <- signif(spp$`Trend p-value`,  3)
+spp$`Statistical signif. [%]` <- signif(spp$`Statistical signif. [%]`,  3)
+
+## rename sky conditions
+spp$`Sky condition` <- sub(" cond.", "", spp$`Sky condition`)
+```
+
+Similar to the long term trends discussed above, we have calculated the trend of the anomalies for the three different sky conditions, and for each season of the year, using the corresponding mean monthly values
+(Figure\ \ref{fig:seasonalALL} and Table\ \ref{tab:trendseasontable}).
+For all-sky conditions the trend in SDR in winter is the largest
+($`r signif(spp$"Trend [%/year]"[spp$Season == "Winter" & spp$"Sky condition" == "All sky"], 2)`\,\%/y$),
+followed by the trend in autumn
+($`r signif(spp$"Trend [%/year]"[spp$Season == "Autumn" & spp$"Sky condition" == "All sky"], 2)`\,\%/y$,
+value similar to the long term trend) both statistically significant above the $99\,\%$ confidence level. 
+In spring and summer, the trends are much smaller and less statistical significant.
+These seasonal differences indicate a possible relation of the trends to clouds during winter and autumn.
+For clear-skies, the trend in winter is $`r signif(spp$"Trend [%/year]"[spp$Season == "Winter" & spp$"Sky condition" == "Clear sky"], 2)`\,\%/y$, larger than for all-skies ($`r signif(spp$"Trend [%/year]"[spp$Season == "Winter" & spp$"Sky condition" == "All sky"], 2)`\,\%/y$), which is another indication of increasing cloud thickness.
+Moreover, the trends under clear- and cloud-sky conditions are almost complementary to each other, particularly for winter and autumn, where the signal is stronger.
+During spring and summer the statistical significance is very low and the actual trend too small for a meaningful comparison.
+
+```{r seasonalALL, echo=F, fig.cap="SDR trends by season of the year for the three sky conditions.
+Each point is a monthly mean calculated from daily means.
+The vertical scale is common for all graphs, each column of graphs corresponds to a sky condition and each row to a season of the year.", out.width="100%"}
+knitr::include_graphics("./images/SeasonalTrendsTogether3-2.png")
+```
+
+<!-- ```{r seasonalALL, echo=F, fig.cap="Trends by season for all-sky condition. Displays monthly means of daily means."} -->
+<!-- knitr::include_graphics("./images/SeasonalTrends-1.png") -->
+<!-- ``` -->
+
+<!-- ```{r seasonalCLEAR, echo=F, fig.cap="Trends by season for clear-sky condition. Displays monthly means of daily means."} -->
+<!-- knitr::include_graphics("./images/SeasonalTrends-2.png") -->
+<!-- ``` -->
+
+<!-- ```{r seasonalCLOUD, echo=F, fig.cap="Trends by season for Cloudy-sky condition. Displays monthly means of daily means."} -->
+<!-- knitr::include_graphics("./images/SeasonalTrends-3.png") -->
+<!-- ``` -->
+
+<!-- \footnotesize -->
+
+```{r trendseasontable, echo=F}
+## arrange columns
+spp <- spp[,c(
+      "Sky condition",
+      "Season",
+      "Trend [%/year]", 
+      # "Trend p-value", 
+      "Statistical signif. [%]"
+      )]
+
+rownames(spp) <- NULL
+# panderOptions("table.alignment.default", "left")
+# panderOptions("table.alignment.default", "center")
+# pander(spp,
+#        justify = c('center', 'right', 'right', 'center'),
+#        split.table = Inf,
+#        caption = "\\label{tab:trendseasontable}SDR linear trends of monthly anomalies for each season of the year.")
+
+## use kable
+x_latex <- knitr::kable(spp,
                         format   = "latex",
-             booktabs = TRUE, 
-             caption = "MDPI journal names.",
-             longtable = TRUE)
+                        booktabs = T,
+                        escape   = T,
+                        # format.args = list(digits = 3),
+                        linesep  = "",
+                        align    = "ccrr",
+                        caption  = "\\label{tab:trendseasontable}SDR linear trends of monthly anomalies for each season of the year.")
+
+x_latex %>%
+    collapse_rows(columns = 1, valign = "middle", latex_hline = "custom") %>%
+    kable_styling(latex_options = c("striped",
+                                    "hold_position")
+    )
+
+# position = "float_left"
 
 ```
-\finishlandscape
+<!-- \normalsize -->
 
-# Introduction
+# Conclusions
 
-The introduction should briefly place the study in a broad context and highlight
-why it is important. It should define the purpose of the work and its
-significance. The current state of the research field should be reviewed
-carefully and key publications cited. Please highlight controversial and
-diverging hypotheses when necessary. Finally, briefly mention the main aim of
-the work and highlight the principal conclusions. As far as possible, please
-keep the introduction comprehensible to scientists outside your particular
-field of research. Citing a journal paper [@bertrand-krajewski_distribution_1998;
-@leutnant_stormwater_2016]. And now citing a book reference @gujer_systems_2008.
-Some MDPI journals use Chicago and others use APA, this template should choose
-the correct citation format for you once you specify the journal in the YAML
-header.
+We demonstrate that in the period 
+`r paste(year(min(ALL_1_daily_DESEAS$Date)), " - ", year(max(ALL_1_daily_DESEAS$Date)))`, there is a brightening trend of SDR for all-sky conditions ($`r alltrend`\,\%/y$).
+On a previous study [@Bais2013], for the period 1993 - 2011 the trend was found to be $0.33\,\%/y$.
+The increase of this trend, indicates that the phenomenon and probably the causes, are still occurring.
+Accordingly, we found a similar trend for clear-sky conditions ($`r cleartrend`\,\%/y$) that further supports that the brightening can be independent of the cloud occurrences, although, cloud formation can also attenuate SDR.
+We have calculated the effect of TSI variability to be a negligible factor in SDR variability.
+Unfortunately, there is no available data for this period for the aerosols, in order to directly estimate their effect on SDR, which is the most probable main cause.
+The attenuation of SDR by aerosols over Europe have been proposed as major factor by @Wild2021.
+The dimming effect of SDR under cloud-sky condition ($`r cloudtrend`\,\%/y$), shows that clouds can also attenuate the amount of SDR reaching the ground, and changes on their occurrences can effect the net energy balance.
+Because we have no adequate data to investigate the long term changes of cloud formation in the region, we can not infer if the SDR trend we observe under cloud-sky is biased by changes in clouds and local climate.
+The observed brightening of SDR over Thessaloniki; is dependent on SZA (higher SZA, have stronger brightening trends), and also, with the seasons of the year, where Winter and Autumn show a significant strong trend, in contrast to Spring and Summer.
+<!-- to conclutions? -->
+Using CUSUMs of the trends for all- and clear-skies, we observed periods where the CUSUMs remain relative steady, 
+<!-- roughly around the middle of the study dataset,--> with a steep decline previous to that, and a steep incline after.
+This is an indication that the whole brightening phenomenon does not follow a smooth development over time.
 
-To use endnotes, change `endnotes: true` in the YAML header, then use
-`\endnote{This is an endnote.}`.
+Our findings are in agreement with other researchers for the region.
+Future observations from the collocated pyrheliometer, may provide us with the means to further investigate the variability of Solar radiation at ground level in Thessaloniki. 
+Also, more data (cloud occurrence, aerosols, atmospheric water vapour, etc.), will help us to determine the extent of SDR attenuation by each factor.
 
-# Materials and Methods
+<!--
+similar results with Evidence for Clear‐Sky Dimming and Brightening in Central Europe_Wild2021.pdf
+include in discussion
+-->
 
-Materials and Methods should be described with sufficient details to allow 
-others to replicate and build on published results. Please note that publication 
-of your manuscript implicates that you must make all materials, data, computer 
-code, and protocols associated with the publication available to readers. Please
-disclose at the submission stage any restrictions on the availability of 
-materials or information. New methods and protocols should be described in 
-detail while well-established methods can be briefly described and appropriately
-cited.
+--------------------
 
-Research manuscripts reporting large datasets that are deposited in a publicly 
-available database should specify where the data have been deposited and provide
-the relevant accession numbers. If the accession numbers have not yet been
-obtained at the time of submission, please state that they will be provided
-during review. They must be provided prior to publication.
+<!-- ## Figures, Tables and Schemes -->
 
-Interventionary studies involving animals or humans, and other studies require
-ethical approval must list the authority that provided approval and the
-corresponding ethical approval code.
+<!-- All figures and tables should be cited in the main text as Figure \ref{fig:fig1},  -->
+<!-- \ref{tab:tab1}, etc. To get cross-reference to figure generated by R chunks  -->
+<!-- include the `\\label{}` tag in the `fig.cap` attribute of the R chunk:  -->
+<!-- `fig.cap = "Fancy Caption\\label{fig:plot}"`. -->
 
-# Results
-
-This section may be divided by subheadings. It should provide a concise and
-precise description of the experimental results, their interpretation as well
-as the experimental conclusions that can be drawn.
-
-## Subsection Heading Here
-
-Subsection text here.
-
-### Subsubsection Heading Here
-
-Bulleted lists look like this:
-
-* First bullet
-* Second bullet
-* Third bullet
-
-Numbered lists can be added as follows:
-
-1. First item
-2. Second item
-3. Third item
-
-The text continues here.
-
-## Figures, Tables and Schemes
-
-All figures and tables should be cited in the main text as Figure \ref{fig:fig1}, 
-\ref{tab:tab1}, etc. To get cross-reference to figure generated by R chunks 
-include the `\\label{}` tag in the `fig.cap` attribute of the R chunk: 
-`fig.cap = "Fancy Caption\\label{fig:plot}"`.
-
-```{r fig1, echo=FALSE, fig.width=3, fig.cap="A figure added with a code chunk.\\label{fig:fig1}"}
-x = rnorm(10)
-y = rnorm(10)
-plot(x, y)
-```
+<!-- ```{r fig1, echo=FALSE, fig.width=3, fig.cap="A figure added with a code chunk.\\label{fig:fig1}"} -->
+<!-- x = rnorm(10) -->
+<!-- y = rnorm(10) -->
+<!-- plot(x, y) -->
+<!-- ``` -->
 
@@
 ```{r tab1, echo=FALSE}
-knitr::kable(mtcars[1:5, 1:3], format = "latex", 
+knitr::kable(mtcars[1:5, 1:3], 
+             format = "latex", 
              booktabs = TRUE, 
@@
 ```
-
-For a very wide table, landscape layouts are allowed.
-
-\startlandscape
-
-```{r tab2, echo=FALSE}
-df <- data.frame(`Title 1` = c("Entry 1", "Entry 2"),
-                 `Title 2` = c("Data", "Data"),
-                 `Title 3` = c("Data", "Data"),
-                 `Title 4` = c("This cell has some longer content that runs over
-                               two lines",
-                               "Data"))
-knitr::kable(df, format = "latex", 
-             booktabs = TRUE, 
-             caption = "This is a very wide table", 
-             align = 'ccc', centering = FALSE,
-             table.envir = "table", position = "H")
-```
-
-\finishlandscape
 
@@
 
-This is the example 2 of equation:
-
-\begin{adjustwidth}{-\extralength}{0cm}
-\begin{equation}
-a = b + c + d + e + f + g + h + i + j + k + l + m + n + o + p + q + r + s + t + 
-u + v + w + x + y + z
-\end{equation}
-\end{adjustwidth}
-
-Theorem-type environments (including propositions, lemmas, corollaries etc.) 
-can be formatted as follows:
-
-Example of a theorem:
-
-::: {.Theorem latex=true}
-Example text of a theorem
-:::
-
-The text continues here.
-
 # Discussion

### 2023-11-02 MDPI_submition.tex
Commit 2023-11-02 10:14

--- 2023-11-02 MDPI_submition.tex
@@
 
-\begin{table}[!h]
+\begin{table}[H]
 
 \caption{\label{tab:trendtable}\label{tab:trendtable}Trends in SDR daily means for different sky conditions for the period 1993 - 2023.}
-\centering
-\begin{tabular}[t]{crrc}
+\begin{tabu} to \linewidth {>{\centering\arraybackslash}p{8em}>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X}
 \toprule
-Sky conditions & Trend [\%/year] & Statistical signif. [\%] & Days\\
+Sky conditions & Trend [\%/year] & Statistical signif. [\%] & Days with data\\
 \midrule
@@
 \bottomrule
-\end{tabular}
+\end{tabu}
 \end{table}
@@
 
-\hypertarget{version}{%
-\section{Version}\label{version}}
+\hypertarget{effects-of-the-solar-zenith-angle-on-sdr.}{%
+\subsection{Effects of the solar zenith angle on
+SDR.}\label{effects-of-the-solar-zenith-angle-on-sdr.}}
 
-This Rmd-skeleton uses the mdpi Latex template published 2023-03-25.
-However, the official template gets more frequently updated than the
-\textbf{rticles} package. Therefore, please make sure prior to paper
-submission, that you're using the most recent .cls, .tex and .bst files
-(available \href{http://www.mdpi.com/authors/latex}{here}).
-
-\hypertarget{article-header-information}{%
-\section{Article Header Information}\label{article-header-information}}
-
-\hypertarget{journal-specific-yaml-variables}{%
-\subsection{Journal Specific YAML
-variables}\label{journal-specific-yaml-variables}}
-
-\startlandscape
-
-\begin{longtable}[t]{llllllll}
-\caption{\label{tab:mdpinames}MDPI journal names.}\\
-\toprule
-acoustics & aerobiology & agronomy & jmse & jor & jsan & wind & \\
-actuators & aerospace & ai & jne & journalmedia & jtaer & women & \\
-addictions & agriculture & air & jnt & jox & jvd & world & \\
-admsci & agriengineering & algorithms & jof & jpm & wem & youth & \\
-adolescents & agrochemicals & jmp & joitmc & jrfm & wevj & zoonoticdis & \\
-\bottomrule
-\end{longtable}
-\finishlandscape
-
-\hypertarget{introduction}{%
-\section{Introduction}\label{introduction}}
-
-The introduction should briefly place the study in a broad context and
-highlight why it is important. It should define the purpose of the work
-and its significance. The current state of the research field should be
-reviewed carefully and key publications cited. Please highlight
-controversial and diverging hypotheses when necessary. Finally, briefly
-mention the main aim of the work and highlight the principal
-conclusions. As far as possible, please keep the introduction
-comprehensible to scientists outside your particular field of research.
-Citing a journal paper
-\citep{bertrand-krajewski_distribution_1998, leutnant_stormwater_2016}.
-And now citing a book reference \citet{gujer_systems_2008}. Some MDPI
-journals use Chicago and others use APA, this template should choose the
-correct citation format for you once you specify the journal in the YAML
-header.
-
-To use endnotes, change \texttt{endnotes:\ true} in the YAML header,
-then use \texttt{\textbackslash{}endnote\{This\ is\ an\ endnote.\}}.
-
-\hypertarget{materials-and-methods}{%
-\section{Materials and Methods}\label{materials-and-methods}}
-
-Materials and Methods should be described with sufficient details to
-allow others to replicate and build on published results. Please note
-that publication of your manuscript implicates that you must make all
-materials, data, computer code, and protocols associated with the
-publication available to readers. Please disclose at the submission
-stage any restrictions on the availability of materials or information.
-New methods and protocols should be described in detail while
-well-established methods can be briefly described and appropriately
-cited.
-
-Research manuscripts reporting large datasets that are deposited in a
-publicly available database should specify where the data have been
-deposited and provide the relevant accession numbers. If the accession
-numbers have not yet been obtained at the time of submission, please
-state that they will be provided during review. They must be provided
-prior to publication.
-
-Interventionary studies involving animals or humans, and other studies
-require ethical approval must list the authority that provided approval
-and the corresponding ethical approval code.
-
-\hypertarget{results-1}{%
-\section{Results}\label{results-1}}
-
-This section may be divided by subheadings. It should provide a concise
-and precise description of the experimental results, their
-interpretation as well as the experimental conclusions that can be
-drawn.
-
-\hypertarget{subsection-heading-here}{%
-\subsection{Subsection Heading Here}\label{subsection-heading-here}}
-
-Subsection text here.
-
-\hypertarget{subsubsection-heading-here}{%
-\subsubsection{Subsubsection Heading
-Here}\label{subsubsection-heading-here}}
-
-Bulleted lists look like this:
-
-\begin{itemize}
-\tightlist
-\item
-  First bullet
-\item
-  Second bullet
-\item
-  Third bullet
-\end{itemize}
-
-Numbered lists can be added as follows:
-
-\begin{enumerate}
-\def\labelenumi{\arabic{enumi}.}
-\tightlist
-\item
-  First item
-\item
-  Second item
-\item
-  Third item
-\end{enumerate}
-
-The text continues here.
-
-\hypertarget{figures-tables-and-schemes}{%
-\subsection{Figures, Tables and
-Schemes}\label{figures-tables-and-schemes}}
-
-All figures and tables should be cited in the main text as Figure
-\ref{fig:fig1}, \ref{tab:tab1}, etc. To get cross-reference to figure
-generated by R chunks include the
-\texttt{\textbackslash{}\textbackslash{}label\{\}} tag in the
-\texttt{fig.cap} attribute of the R chunk:
-\texttt{fig.cap\ =\ "Fancy\ Caption\textbackslash{}\textbackslash{}label\{fig:plot\}"}.
+The solar zenith angle is a major driver of the attenuation of SDR
+reaching the ground, as SDR is affected by the enhancement of the
+radiation path in the atmosphere, especially in urban environment with
+human activities emitting aerosols \citep{Wang2021}. In order to
+estimate the effect of the SZA on the SDR trends, we grouped the anomaly
+data in bins of \(1^\circ\) SZA, and calculated the overall trend for
+each bin before noon and after noon (Figure~\ref{fig:szatrends}).
+Although there are seasonal dependencies of the minimum SZA (see
+Appendix, Figure~\ref{fig:SZAtrendSeason}), these dependencies would not
+be further examined here. For all-sky and clear-sky conditions the
+brightening effect of SDR (positive trend) is stronger for large SZAs
+(Figures~\ref{fig:szatrends-1} and \ref{fig:szatrends-2}). The trends in
+the morning and afternoon hours are more or less consistent with small
+differences, which can be attributed to systematic diurnal variations of
+aerosols, particularly during the warm period of the year
+\citep{Wang2021}. For cloud-sky conditions
+(Figure~\ref{fig:szatrends-3}), we can not discern any significant
+dependence of the SDR trend with SZA. For SZAs \(16^\circ\) -
+\(50^\circ\), the trends range within about \(\pm 0.2\,\%/y\), with a
+weak statistical significance. Between \(50^\circ\) and \(75^\circ\) SZA
+the trends for the period before noon are stronger and negative,
+possibly associated with stronger attenuation by clouds under oblique
+incidence angles.
 
@@
 
-{\centering \includegraphics[width=0.7\linewidth]{MDPI_submition_files/figure-latex/fig1-1}
+{\centering \subfloat[All-sky conditions.\label{fig:szatrends-1}]{\includegraphics[width=.32\linewidth]{./images/SzaTrends-1} }\subfloat[Clear-sky conditions.\label{fig:szatrends-2}]{\includegraphics[width=.32\linewidth]{./images/SzaTrends-4} }\subfloat[Cloud-sky conditions.\label{fig:szatrends-3}]{\includegraphics[width=.32\linewidth]{./images/SzaTrends-7} }
 
@@
 
-\caption{A figure added with a code chunk.\label{fig:fig1}}\label{fig:fig1}
+\caption{Distribution of the SDR's long term trends by SZA before and after noon.
+The solid data points represent values with an acceptable statistical significance ($p<0.005$), the trends were computed with SZA bins of $1^\circ$ for each day, before and after local noon.}\label{fig:szatrends}
 \end{figure}
+
+\hypertarget{long-term-trends-by-season-of-year}{%
+\subsection{Long term trends by season of
+year}\label{long-term-trends-by-season-of-year}}
+
+Similar to the long term trends discussed above, we have calculated the
+trend of the anomalies for the three different sky conditions, and for
+each season of the year, using the corresponding mean monthly values
+(Figure~\ref{fig:seasonalALL} and Table~\ref{tab:trendseasontable}). For
+all-sky conditions the trend in SDR in winter is the largest
+(\(0.69\,\%/y\)), followed by the trend in autumn (\(0.43\,\%/y\), value
+similar to the long term trend) both statistically significant above the
+\(99\,\%\) confidence level. In spring and summer, the trends are much
+smaller and less statistical significant. These seasonal differences
+indicate a possible relation of the trends to clouds during winter and
+autumn. For clear-skies, the trend in winter is \(0.83\,\%/y\), larger
+than for all-skies (\(0.69\,\%/y\)), which is another indication of
+increasing cloud thickness. Moreover, the trends under clear- and
+cloud-sky conditions are almost complementary to each other,
+particularly for winter and autumn, where the signal is stronger. During
+spring and summer the statistical significance is very low and the
+actual trend too small for a meaningful comparison.
+
+\begin{figure}[h!]
+
+{\centering \includegraphics[width=1\linewidth]{./images/SeasonalTrendsTogether3-2}
+
+}
+
+\caption{SDR trends by season of the year for the three sky conditions.
+Each point is a monthly mean calculated from daily means.
+The vertical scale is common for all graphs, each column of graphs corresponds to a sky condition and each row to a season of the year.}\label{fig:seasonalALL}
+\end{figure}
+
+\begin{table}[!h]
+
+\caption{\label{tab:trendseasontable}\label{tab:trendseasontable}SDR linear trends of monthly anomalies for each season of the year.}
+\centering
+\begin{tabular}[t]{ccrr}
+\toprule
+Sky condition & Season & Trend [\%/year] & Statistical signif. [\%]\\
+\midrule
+\cellcolor{gray!6}{} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{0.6860} & \cellcolor{gray!6}{99.9}\\
+
+ & Spring & 0.1450 & 81.8\\
+
+\cellcolor{gray!6}{} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{0.1200} & \cellcolor{gray!6}{89.4}\\
+
+\multirow{-4}{*}{\centering\arraybackslash All sky} & Autumn & 0.4310 & 99.3\\
+\cmidrule{1-4}
+\cellcolor{gray!6}{} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{0.8260} & \cellcolor{gray!6}{100.0}\\
+
+ & Spring & 0.0613 & 38.6\\
+
+\cellcolor{gray!6}{} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{-0.0307} & \cellcolor{gray!6}{25.9}\\
+
+\multirow{-4}{*}{\centering\arraybackslash Clear sky} & Autumn & 0.3670 & 97.2\\
+\cmidrule{1-4}
+\cellcolor{gray!6}{} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{-0.8820} & \cellcolor{gray!6}{98.9}\\
+
+ & Spring & -0.0991 & 37.2\\
+
+\cellcolor{gray!6}{} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{0.0444} & \cellcolor{gray!6}{21.5}\\
+
+\multirow{-4}{*}{\centering\arraybackslash Cloudy sky} & Autumn & -0.4000 & 89.5\\
+\bottomrule
+\end{tabular}
+\end{table}
+
+\hypertarget{conclusions}{%
+\section{Conclusions}\label{conclusions}}
+
+We demonstrate that in the period 1993 - 2023, there is a brightening
+trend of SDR for all-sky conditions (\(0.38\,\%/y\)). On a previous
+study \citep{Bais2013}, for the period 1993 - 2011 the trend was found
+to be \(0.33\,\%/y\). The increase of this trend, indicates that the
+phenomenon and probably the causes, are still occurring. Accordingly, we
+found a similar trend for clear-sky conditions (\(0.35\,\%/y\)) that
+further supports that the brightening can be independent of the cloud
+occurrences, although, cloud formation can also attenuate SDR. We have
+calculated the effect of TSI variability to be a negligible factor in
+SDR variability. Unfortunately, there is no available data for this
+period for the aerosols, in order to directly estimate their effect on
+SDR, which is the most probable main cause. The attenuation of SDR by
+aerosols over Europe have been proposed as major factor by
+\citet{Wild2021}. The dimming effect of SDR under cloud-sky condition
+(\(-0.28\,\%/y\)), shows that clouds can also attenuate the amount of
+SDR reaching the ground, and changes on their occurrences can effect the
+net energy balance. Because we have no adequate data to investigate the
+long term changes of cloud formation in the region, we can not infer if
+the SDR trend we observe under cloud-sky is biased by changes in clouds
+and local climate. The observed brightening of SDR over Thessaloniki; is
+dependent on SZA (higher SZA, have stronger brightening trends), and
+also, with the seasons of the year, where Winter and Autumn show a
+significant strong trend, in contrast to Spring and Summer. Using CUSUMs
+of the trends for all- and clear-skies, we observed periods where the
+CUSUMs remain relative steady, with a steep decline previous to that,
+and a steep incline after. This is an indication that the whole
+brightening phenomenon does not follow a smooth development over time.
+
+Our findings are in agreement with other researchers for the region.
+Future observations from the collocated pyrheliometer, may provide us
+with the means to further investigate the variability of Solar radiation
+at ground level in Thessaloniki. Also, more data (cloud occurrence,
+aerosols, atmospheric water vapour, etc.), will help us to determine the
+extent of SDR attenuation by each factor.
+
+\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}
 
@@
 
-For a very wide table, landscape layouts are allowed.
-
-\startlandscape
-
-\begin{table}[H]
-
-\caption{\label{tab:tab2}This is a very wide table}
-\begin{tabular}[t]{cccc}
-\toprule
-Title.1 & Title.2 & Title.3 & Title.4\\
-\midrule
-Entry 1 & Data & Data & This cell has some longer content that runs over
-                               two lines\\
-Entry 2 & Data & Data & Data\\
-\bottomrule
-\end{tabular}
-\end{table}
-
-\finishlandscape
-
 \hypertarget{formatting-of-mathematical-components}{%
@@
 punctuate equations as regular text.
-
-This is the example 2 of equation:
-
-\begin{adjustwidth}{-\extralength}{0cm}
-\begin{equation}
-a = b + c + d + e + f + g + h + i + j + k + l + m + n + o + p + q + r + s + t + 
-u + v + w + x + y + z
-\end{equation}
-\end{adjustwidth}
-
-Theorem-type environments (including propositions, lemmas, corollaries
-etc.) can be formatted as follows:
-
-Example of a theorem:
-
-\begin{Theorem}
-Example text of a theorem
-
-\end{Theorem}
-
-The text continues here.
 

### 2023-11-02 MDPI_submition.Rmd
Commit 2023-11-02 11:10

@@
 Thus, our clear sky reference is derived by the Eq.\ \ref{eq:ahau}.
-
 <!-- cost fuction -->
@@
 where: $n$ is the total number of daylight records, $\text{SDR}_{\text{CSid},i}$ are the records identified as clear sky by CSid, $a$ is a hypothetical adjustment factor, and $\text{SDR}_{\text{testCSref},i}$ is any of the tested clear sky radiation models.
-
 <!-- adjusted Haurwitz -->
@@
 -->
-
 <!-- 1. MeanVIP -->
 <!-- ### Mean value of irradiance during the time period. -->
-
 <!--a) Mean of the measured $\overline{\text{SDR}}_i$ = \text{mean}_{n=i-5}^{n=i+5}[\text{SDR}_{n}]$ and running mean of the reference $\overline{\text{SDR}}_{\text{CSref},i}$ = \text{mean}_{n=i-5}^{n=i+5}[\text{SDR}_{\text{CSref},n}]$ (Eq.
@@
                         format      = "latex",
-                        booktabs = T,
+                        booktabs    = TRUE,
                         escape      = T,
@@
                         align       = "ccrr",
+                        centering   = FALSE,
+                        table.envir = "table",
+                        # position    = "H",
                         caption  = "\\label{tab:trendseasontable}SDR linear trends of monthly anomalies for each season of the year.")
@@
     collapse_rows(columns = 1, valign = "middle", latex_hline = "custom") %>%
-    kable_styling(latex_options = c("striped",
-                                    "hold_position")
-    )
-
-# position = "float_left"
+    kable_styling(latex_options = c('striped', 'hold_position'), full_width = TRUE) %>%
+    # kable_styling(full_width = T) %>%
+    # kable_styling("striped")      %>%
+    column_spec(1, width = "8em")
 
@@
 
-When making tables using `kable`, it is suggested to use
-the `format="latex"` and `tabl.envir="table"` arguments
-to ensure table numbering and compatibility with the mdpi
-document class.
+<!-- When making tables using `kable`, it is suggested to use -->
+<!-- the `format="latex"` and `tabl.envir="table"` arguments -->
+<!-- to ensure table numbering and compatibility with the mdpi -->
+<!-- document class. -->
 
-```{r tab1, echo=FALSE}
-knitr::kable(mtcars[1:5, 1:3], 
-             format = "latex", 
-             booktabs = TRUE, 
-             caption = "This is a table caption. Tables should be placed in the 
-             main text near to the first time they are~cited.", 
-             align = 'ccc', centering = FALSE,
-             table.envir = "table", position = "H")
-```
+<!-- ```{r tab1, echo=FALSE} -->
+<!-- knitr::kable(mtcars[1:5, 1:3],  -->
+<!--              format = "latex",  -->
+<!--              booktabs = TRUE,  -->
+<!--              caption = "This is a table caption. Tables should be placed in the  -->
+<!--              main text near to the first time they are~cited.",  -->
+<!--              align = 'ccc', centering = FALSE, -->
+<!--              table.envir = "table", position = "H") -->
+<!-- ``` -->
 
 ## Formatting of Mathematical Components
-
-This is an example of an equation:
-
-$$
-a = 1.
-$$
 

### 2023-11-02 MDPI_submition.tex
Commit 2023-11-02 11:10

@@
 of the data. Thus, our clear sky reference is derived by the
-Eq.~\ref{eq:ahau}.
-
-\begin{equation}
+Eq.~\ref{eq:ahau}. \begin{equation}
 f(a) = \frac{1}{n}\sum_{i=1}^{n} ( \text{SDR}_{\text{CSid},i} - a \times \text{SDR}_{\text{testCSref},i} )^2 \label{eq:minf}
@@
 \(\text{SDR}_{\text{testCSref},i}\) is any of the tested clear sky
-radiation models.
-
-\begin{equation}
+radiation models. \begin{equation}
 \text{SDR}_\text{CSref} = a \times \text{SDR}_\text{Haurwitz} = 0.965 \times 1098 \times \cos(\theta) \times \exp \left( \frac{ - 0.057}{\cos(\theta)} \right) \label{eq:ahau}
@@
 
+\begin{verbatim}
+Usually it is recommended to use column_spec before collapse_rows, especially in LaTeX, to get a desired result. 
+\end{verbatim}
+
 \begin{table}[!h]
@@
 \caption{\label{tab:trendseasontable}\label{tab:trendseasontable}SDR linear trends of monthly anomalies for each season of the year.}
-\centering
-\begin{tabular}[t]{ccrr}
+\begin{tabu} to \linewidth {>{\centering\arraybackslash}p{8em}>{\centering}X>{\raggedleft}X>{\raggedleft}X}
 \toprule
@@
 
-When making tables using \texttt{kable}, it is suggested to use the
-\texttt{format="latex"} and \texttt{tabl.envir="table"} arguments to
-ensure table numbering and compatibility with the mdpi document class.
-
-\begin{table}[H]
-
-\caption{\label{tab:tab1}This is a table caption. Tables should be placed in the 
-             main text near to the first time they are~cited.}
-\begin{tabular}[t]{lccc}
-\toprule
-  & mpg & cyl & disp\\
-\midrule
-Mazda RX4 & 21.0 & 6 & 160\\
-Mazda RX4 Wag & 21.0 & 6 & 160\\
-Datsun 710 & 22.8 & 4 & 108\\
-Hornet 4 Drive & 21.4 & 6 & 258\\
-Hornet Sportabout & 18.7 & 8 & 360\\
-\bottomrule
-\end{tabular}
-\end{table}
-
 \hypertarget{formatting-of-mathematical-components}{%
@@
 Components}\label{formatting-of-mathematical-components}}
-
-This is an example of an equation:
-
-\[
-a = 1.
-\]
 

### 2023-11-02 MDPI_submition.Rmd
Commit 2023-11-02 12:14

@@
 \end{equation}
-
 <!-- 2. MaxVIP -->
 <!-- ### Max value of irradiance during the time period. -->
-
 <!-- The running max measured value $M_{\text{SDR},i} = \max_{n=i-5}^{n=i+5}[\text{SDR}_{n}]$, is compared to a similar constructed value from the reference $M_{CSrefi} = \max_{n=i-5}^{n=i+5}[\text{SDR}_{\text{CSref},n}]$ (Eq.\ \ref{eq:MaxVIP}).
 -->
-
 b) Maximum measured value $M_{\text{}}$ (Eq.\ \ref{eq:MaxVIP}).
@@
 \end{equation}
-
 <!-- 3. VIL -->
@@
 where: $t_i$ is the time each SDR measurement has been measured
-
 <!-- 4. VCT -->
 <!-- ### Variance of Changes in the Time series. -->
-
 <!-- Standard deviation of rate of change in irradiance. -->
@@
 \end{equation}
-
 <!-- 5. VSM -->
 <!-- ### Variability in the Shape of the irradiance Measurements. -->
-
 e) Maximum difference $X_i$ between the change in measured irradiance and the change in clear sky irradiance over each measurement interval.
@@
                         position    = "H",
-                        caption     = paste0("\\label{tab:trendtable}Trends in SDR daily means for different sky conditions for the period ", year(min(ALL_1_daily_DESEAS$Date)), " - ", year(max(ALL_1_daily_DESEAS$Date)), "."))
+                        caption     = paste0("Trends in SDR daily means for different sky conditions for the period ", year(min(ALL_1_daily_DESEAS$Date)), " - ", year(max(ALL_1_daily_DESEAS$Date)), "."))
 
@@
                         # position    = "H",
-                        caption  = "\\label{tab:trendseasontable}SDR linear trends of monthly anomalies for each season of the year.")
+                        caption  = "SDR linear trends of monthly anomalies for each season of the year.")
 
@@
 
-## Formatting of Mathematical Components
+<!-- ## Formatting of Mathematical Components -->
 
-If you want numbered equations use Latex and wrap in the equation environment:
+<!-- If you want numbered equations use Latex and wrap in the equation environment: -->
 
-\begin{equation}
-a = 1,
-\end{equation}
+<!-- \begin{equation} -->
+<!-- a = 1, -->
+<!-- \end{equation} -->
 
-the text following an equation need not be a new paragraph. Please punctuate 
-equations as regular text.
+<!-- the text following an equation need not be a new paragraph. Please punctuate  -->
+<!-- equations as regular text. -->
 
-```{comment}
-If the documentclass option "submit" is chosen, please insert a blank line before and after any math environment (equation and eqnarray environments).
+<!-- ```{comment} -->
+<!-- If the documentclass option "submit" is chosen, please insert a blank line before and after any math environment (equation and eqnarray environments).
 This ensures correct linenumbering.
 The blank line should be removed when the documentclass option is changed to "accept" because the text following an equation should not be a new paragraph.
-```
+-->
+<!-- ``` -->
 

### 2023-11-02 MDPI_submition.tex
Commit 2023-11-02 12:14

@@
   \end{equation}
-\end{enumerate}
-
-\begin{enumerate}
-\def\labelenumi{\alph{enumi})}
-\setcounter{enumi}{1}
-\tightlist
 \item
@@
   \end{equation}
-\end{enumerate}
-
-\begin{enumerate}
-\def\labelenumi{\alph{enumi})}
-\setcounter{enumi}{2}
-\tightlist
 \item
@@
   been measured
-\end{enumerate}
-
-\begin{enumerate}
-\def\labelenumi{\alph{enumi})}
-\setcounter{enumi}{3}
-\tightlist
 \item
@@
   \end{equation}
-\end{enumerate}
-
-\begin{enumerate}
-\def\labelenumi{\alph{enumi})}
-\setcounter{enumi}{4}
-\tightlist
 \item
@@
 
-\caption{\label{tab:trendtable}\label{tab:trendtable}Trends in SDR daily means for different sky conditions for the period 1993 - 2023.}
+\caption{\label{tab:trendtable}Trends in SDR daily means for different sky conditions for the period 1993 - 2023.}
 \begin{tabu} to \linewidth {>{\centering\arraybackslash}p{8em}>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X}
@@
 
-\caption{\label{tab:trendseasontable}\label{tab:trendseasontable}SDR linear trends of monthly anomalies for each season of the year.}
+\caption{\label{tab:trendseasontable}SDR linear trends of monthly anomalies for each season of the year.}
 \begin{tabu} to \linewidth {>{\centering\arraybackslash}p{8em}>{\centering}X>{\raggedleft}X>{\raggedleft}X}
@@
 \begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}
-
-\hypertarget{formatting-of-mathematical-components}{%
-\subsection{Formatting of Mathematical
-Components}\label{formatting-of-mathematical-components}}
-
-If you want numbered equations use Latex and wrap in the equation
-environment:
-
-\begin{equation}
-a = 1,
-\end{equation}
-
-the text following an equation need not be a new paragraph. Please
-punctuate equations as regular text.
 

### 2023-11-02 appendix.tex
Commit 2023-11-02 12:14

--- 2023-07-03 Article.tex
@@
-\documentclass[preprint, 3p,
-authoryear]{elsarticle} %review=doublespace preprint=single 5p=2 column
-%%% Begin My package additions %%%%%%%%%%%%%%%%%%%
+%% optional
+\appendixtitles{yes} %Leave argument "no" if all appendix headings stay EMPTY (then no dot is printed after "Appendix A"). If the appendix sections contain a heading then change the argument to "yes".
+\appendixsections{one} %Leave argument "multiple" if there are multiple sections.
+Then a counter is printed ("Appendix A").
+If there is only one appendix section then change the argument to "one" and no counter is printed ("Appendix").
+\appendix
+% \section{}
 
-\usepackage[hyphens]{url}
-
-  \journal{An awesome journal} % Sets Journal name
-
-\usepackage{graphicx}
-%%%%%%%%%%%%%%%% end my additions to header
-
-\usepackage[T1]{fontenc}
-\usepackage{lmodern}
-\usepackage{amssymb,amsmath}
-% TODO: Currently lineno needs to be loaded after amsmath because of conflict
-% https://github.com/latex-lineno/lineno/issues/5
-\usepackage{lineno} % add
-\usepackage{ifxetex,ifluatex}
-\usepackage{fixltx2e} % provides \textsubscript
-% use upquote if available, for straight quotes in verbatim environments
-\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
-\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
-  \usepackage[utf8]{inputenc}
-\else % if luatex or xelatex
-  \usepackage{fontspec}
-  \ifxetex
-    \usepackage{xltxtra,xunicode}
-  \fi
-  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
-  \newcommand{\euro}{€}
-\fi
-% use microtype if available
-\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
-\usepackage[]{natbib}
-\bibliographystyle{plainnat}
-
-\ifxetex
-  \usepackage[setpagesize=false, % page size defined by xetex
-              unicode=false, % unicode breaks when used with xetex
-              xetex]{hyperref}
-\else
-  \usepackage[unicode=true]{hyperref}
-\fi
-\hypersetup{breaklinks=true,
-            bookmarks=true,
-            pdfauthor={},
-            pdftitle={Trends of SDR in Thessaloniki},
-            colorlinks=false,
-            urlcolor=blue,
-            linkcolor=magenta,
-            pdfborder={0 0 0}}
-
-\setcounter{secnumdepth}{5}
-% Pandoc toggle for numbering sections (defaults to be off)
-
-% tightlist command for lists without linebreak
-\providecommand{\tightlist}{%
-  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
-
-% From pandoc table feature
-\usepackage{longtable,booktabs,array}
-\usepackage{calc} % for calculating minipage widths
-% Correct order of tables after \paragraph or \subparagraph
-\usepackage{etoolbox}
-\makeatletter
-\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
-\makeatother
-% Allow footnotes in longtable head/foot
-\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
-\makesavenoteenv{longtable}
-
-\usepackage{subcaption}
-
-\begin{document}
-
-\begin{frontmatter}
-
-  \title{Trends of SDR in Thessaloniki}
-    \author[Laboratory of Atmospheric Physics]{Athanasios Natsis%
-  \corref{cor1}%
-  }
-   \ead{natsisphysicist@gmail.com} 
-    \author[Laboratory of Atmospheric Physics]{Alkiviadis Bais%
-  %
-  }
-   \ead{abais@auth.gr} 
-      \affiliation[Laboratory of Atmospheric Physics]{
-    organization={Aristotle University of
-Thessaloniki},addressline={Campus Box
-149},city={Thessaloniki},postcode={54124},country={Greece},}
-    \cortext[cor1]{Corresponding author}
-
-  \begin{abstract}
-  Study of GHI and DNI radiation for `clear-sky' and all-sky conditions.
-  It consists of two paragraphs.
-  \end{abstract}
-    \begin{keyword}
-    GHI \sep SDR \sep 
-    Solar Brigthening/Dimming
-  \end{keyword}
-
- \end{frontmatter}
-
-\hypertarget{introduction.}{%
-\section{Introduction.}\label{introduction.}}
-
-The shortwave downward solar irradiance (SDR) at Earth's surface play a
-significant role, on it's climate. Changes of the SDR can be related to
-changes on Earth's energy budget, the mechanisms of climate change, and
-water and carbon cycle \citep{Wild2009}. Can also affect, solar and
-agricultural production, and living organisms. Studies of SDR
-variability, have identified some distinct SDR trends on different
-regions of the world and for different time periods. The term
-`brightening' is generally used to describe periods of positive SDR
-trend, and `dimming' for negative trend. There are many cases on the
-long term records of irradiance, showing a systematic change of SDR's
-trend slope, occurring roughly at the last decades of the 20th century.
-On multiple station in China, a dimming period was reported until about
-2000, followed by a brightening period \citep{Yang2021}. A similar
-pattern was identified, with the breaking point around 1980, for
-stations in Central Europe \citep{Wild2021} and Brazil
-\citep{Yamasoe2021}. Also, on global scale, an AI aided continental
-level spatial analysis, with data from multiple station, reach similar
-conclusions for the above regions and for the global trend
-\citep{Yuan2021}.
-
-There is a consensus, among researchers, that major factors of SDR
-attenuation is the interaction with atmospheric aerosols and clouds.
-Those interactions, among other factors, have been analysed with models
-\citep{Li2016, Samset2018}, showing the existence of feedback mechanisms
-between the two. Similar finds, have been showed in observational data
-\citep[ and references
-therein]{Schwarz2020, Ohvril2009, Zerefos2009, Xia2007}.
-
-Due to the variability of the phenomenon, and its contributing factors,
-there is a need to further investigate and monitor SDR, in different
-sites, to estimate its magnitude, and its relation to the local
-conditions. In this study, we examine the trends of SDR, with
-ground-based measurements at Thessaloniki, Greece for the period 1993 to
-2023, as derived from a CM-21 pyranometer. We reevaluated and extended
-the dataset used by \citet{Bais2013}, applying a different algorithm for
-the identification of clear-/cloud-sky instances
-\citep{Reno2016, Reno2012a}, and we derive the radiation trends for the
-period 1993 to 2023, under different sky conditions (all-sky, clear-sky
-and cloud-sky).
-
-\hypertarget{observational-data-and-methodology.}{%
-\section{Observational data and
-methodology.}\label{observational-data-and-methodology.}}
-
-The SDR is equivalent to the whole sky Global radiation (or Global
-Horizontal Irradiance, GHI), and was obtained with a horizontal leveled,
-CM-21 pyranometer, from 1993-04-12 to 2023-03-31. Also, we have
-measurements of direct beam radiation (DNI) obtained by a collocated
-CHP-1 pyrheliometer, from 2016-04-01 to 2023-03-31. Although, we
-performed a similar analysis to the DNI as for the SDR, the results are
-not presented here, as they lack the appropriate statistical
-significance, due to the sorter duration of the data. Moreover, the DNI
-was used as auxiliary data, with the clear sky identification algorithm
-(CSid), to aid with the selection of the appropriate thresholds. We have
-to note also, that despite the capability of the CSid to use the DNI as
-an characterization parameter, we haven't use it, to avoid any selection
-bias due to unequal availability of measurements.
-
-The SDR data were measured with a Kipp \& Zonen CM-21 pyranometer
-operating continuously at the Laboratory of Atmospheric Physics of the
-Aristotle University of Thessaloniki (\(40^\circ38'\)N,
-\(22^\circ57'\)E, 80m a.s.l.). The monitoring site is located near the
-city centre. During the study period the CM-21 pyranometer has been
-independently calibrated four times. There are three distinct steps to
-the creation of this dataset: a) the acquisition of radiation
-measurements from the sensors, b) a radiation data quality check, and c)
-the identification of ``clear sky'' conditions from the radiometric data
-d) data aggregation and trend analysis.
-
-For the acquisition of radiometric data, the signal of the broadband
-instruments is sampled with a rate of \(1 \text{Hz}\). The mean and its
-standard deviation are recorded for every minute. The measurements are
-corrected for the zero offset of the instrument signal. As reference of
-the ``dark signal'' is used the Sun elevation angle bellow
-\(-10^\circ\), for a period of \(3 \text{h}\). The signal is converted
-to radiation flux, using a ramped value of the instrument sensitivity,
-derived from the eight laboratory calibrations of the instrument, during
-the study period.
-
-A manual screening was performed, to remove inconsistencies and
-erroneous recordings, that can occur randomly or systematically, during
-the long continuous operation of the instrument. The manual screening is
-aided by a radiation data quality assurance procedure, adjusted for this
-site, based on methods of Long and
-Shi~\citetext{\citeyear{Long2008a}; \citeyear{Long2006}}. Thus,
-problematic recordings have been excluded from further processing.
-Although it is not possible to detect all the bad data, the large number
-of data and the aggregation scheme we use, can provide us with accurate
-radiometric measurements, for the scope of this study.
-
-In order to estimate the effect of the clouds on the SDR, we created
-three datasets, by characterizing each one-minute measurement with a
-corresponding sky condition. The all-sky conditions, containing all the
-valid measurements. The clear-sky conditions data, correspond to data
-where we have inferred that the sky was almost clear of clouds, and the
-remainder part the data was characterized with cloud sky conditions
-(cloud-sky). To identify the clear-sky conditions we used the method
-proposed by \citet{Long2000} and by \citet{Reno2016}, that was adapted
-and configured for the site.
-
-\ldots.
-
-cite iterative process
-
-\ldots\ldots{}
-
-We have to note, that the definition of what constitutes as clear or
-cloudy sky, has some subjectivity, due to the used method of
-characterization. As a result, the details of the definition are site
-specific, it relies on a combination of threshold of comparisons with
-ideal actinometric models and statistics on different signal behavior.
-
-The algorithm was calibrated with the main focus to identify the
-presence of clouds on the sky dome, although there are marginal cases
-that there will be false positives or false negatives identifications.
-
-As global radiation clear sky reference we are using the Haurwitz's
-model, adjusted for our site with a factor of 0.965 (Eq. \ref{eq:ahau}).
-The selection of a clear sky reference model, was based on SDR
-observation from the period 2016 -- 2021. Where, after an iterative
-optimization of eight simple models (Daneshyar--Paltridge--Proctor,
-Kasten--Czeplak, Haurwitz, Berger--Duffie,
-Adnot--Bourges--Campana--Gicquel, Robledo-Soler, Kasten and
-Ineichen-Perez) with different factors. We found, that Haurwitz's model,
-adjusted with a factor of 0.965 has the lower root mean squared error
-(RMSE). The tried models are described by \citet{Reno2012} and tested by
-\citet{Reno2016}. The iterative optimization method, for the selection
-of the reference is discussed by \citet{Long2000} and \citet{Reno2016}.
-
-\begin{equation}
-\text{SDR}_\text{Clear Sky} = 0.965 \times 1098 \times \cos( \text{SZA} ) \times \exp \left( \frac{ - 0.057}{\cos(\text{SZA})} \right) \label{eq:ahau}
-\end{equation}
-
-The following criteria and thresholds were used to identify clear-sky
-conditions. Each criterion was applied for a running window of \(11\)
-consecutive measurements/minutes, and the characterization is applied at
-the center value of the window. A data point is consider as under
-cloud-sky condition if it fails to pass any of the criteria, all the
-other data points are characterized as clear-sky.
-
-The mean of the measured value \(\overline{G}_i\) must be inside an
-envelope based on the reference model \(\text{SDR}_\text{Clear Sky}\)
-(Eq. \ref{eq:MeanVIP}).
-
-\begin{equation}
-0.91 \times \overline{\text{SDR}}_{i\text{Clear Sky}} - 20
-< \overline{G}_i <
-1.095 \times \overline{\text{SDR}}_{i\text{Clear Sky}} + 30
-\label{eq:MeanVIP}
-\end{equation}
-
-The running max measured value \(M_{Gi} = max[\text{SDR}_{i}]\), is
-compared to a similar constructed value from the reference
-\(M_{CSi} = max[\text{SDR}_{i\text{Clear Sky}}]\) (Eq. \ref{eq:MaxVIP}).
-
-\begin{equation}
-1 \times M_{CSi} - 75
-< M_{Gi} <
-1 \times M_{CSi} + 75
-\label{eq:MaxVIP}
-\end{equation}
-
-The length \(L\) (Eq. \ref{eq:VILeq}) of the sequence of line segments
-connecting the points of the SDR time series for the measured values
-\(L\) and similar for the reference \(L_{CS}\), must be within the
-limits of Eq. \ref{eq:VILcr}.
-
-\begin{equation}
-L = \sum_{i=1}^{n-1}\sqrt{\left ( \text{SDR}_{i+1} - \text{SDR}_{i}\right )^2 + \left ( t_{i+1} - t_i \right )^2}
-\label{eq:VILeq}
-\end{equation}
-
-\begin{equation}
-1 \times L_{CSi} - 5 < L_i < 1.3 \times L_{CSi} + 13
-\label{eq:VILcr}
-\end{equation}
-
-We calculate the standard deviation \(\sigma\) of the slope (\(s\))
-between sequential points in the time series, normalized by the average
-SDR during the time interval.
-
-\begin{equation}
-s_i = \frac{\text{SDR}_{i+1} - \text{SDR}_{i}}{t_{i+1} - t_i}, \forall i \in \left \{ 1, 2, \ldots, n-1 \right \},\;\;
-\bar{s} = \frac{1}{n-1} \sum_{i=1}^{n-1} s_i,\;\;
-\sigma_i = \frac {\sqrt{\frac{1}{n-1} \sum_{i=1}^{n-1} \left( s_i - \bar{s} \right)^2} } {\bar{G_i}} \label{eq:VCT3b}
-\end{equation}
-
-\begin{equation}
-\sigma_i < \ensuremath{1.1\times 10^{-4}}
-\label{eq:VCTcr}
-\end{equation}
-
-The maximum difference \(X\) between the change in measured irradiance
-and the change in clear sky irradiance over each measurement interval.
-
-\begin{gather}
-x_i = \text{SDR}_{i+1} - \text{SDR}_{i} \forall i \in \left \{ 1, 2, \ldots, n-1 \right \} \label{eq:VSM1} \\
-x_{CS,i} = \text{SDR}_{CS,i+1} - \text{SDR}_{CS,i} \forall i \in \left \{ 1, 2, \ldots, n-1 \right \} \label{eq:VSM2} \\
-X_i = \max{\left \{ \left | x_i - x_{CS,i} \right | \right \}} \label{eq:VSM3}
-\end{gather}
-
-\begin{equation}
-X_i < 7.5
-\label{eq:VSMcr}
-\end{equation}
-
-Due to a significant measurement uncertainty near, the horizon, we have
-to exclude all measurements with SZA greater than \(85^\circ\).
-Moreover, due to some obstructions around the site (hills and
-buildings), we excluded data with Azimuth angle between \(35^\circ\) and
-\(120^\circ\) with SZA greater than \(80^\circ\). On the latter
-instances, Sun is systematically, not visible by the instrument's
-location. To make the measurements comparable throughout the dataset, we
-adjusted all 1-minute radiometric values to the mean Sun - Earth
-distance. Subsequently, we made all measurements relative to the Total
-Solar Irradiance (TSI) at \(1 \text{au}\), in order to compensate for
-the Sun's intensity variability, using a homogenized time series of
-satellite TSI observations. The TSI data we use, comes from NOAA
-\citep{Coddington2005}, it overlaps with our data, and is interpolated
-to match with the time step of the data. As a result, we can present all
-radiation data as a fraction of Sun's TSI.
-
-\hypertarget{aggregation-of-radiometric-data}{%
-\subsubsection{Aggregation of radiometric
-data}\label{aggregation-of-radiometric-data}}
-
-Before further analysis and deseasonalization, we implement an
-appropriate aggregation scheme on the 1-minute data. To preserve the
-representativeness we use the following criteria: a) for daily mean
-values we exclude instances with less than 180 valid data points, b)
-accordingly, monthly values are computed by daily aggregated data, where
-months with less than 20 days are rejected. The daily seasonal values
-were indexed based on the day of year number, and the monthly by the
-corresponding calendar month. For the seasons of the year trends, we
-grouped the mean daily values by season (December - February: Winter,
-March - May: Spring, etc.). Finally, for each data set, we remove the
-natural occurring seasonal variation. This is done, by calculating the
-mean values for the appropriate time step, and subtracting the annual
-cycle from the actual data. In addition to the previous aggregation
-scheme, each dataset was aggregate in \(1^\circ\) SZA bins, separate for
-cases before and after local noon. This gave us an approximation of
-Sun's SZA contribution to the SDR brightening.
-
-Thus, for each data set we have obtained the relative departure of the
-seasonal mean. The statistical significance and aggregation scheme and
-filtering will be noted along with the corresponding results.
-
-\hypertarget{results}{%
-\section{Results}\label{results}}
-
-\hypertarget{long-term-sdr-trends}{%
-\subsection{Long-term SDR trends}\label{long-term-sdr-trends}}
-
-We calculated the SDR trends, from the mean daily values of the
-departure from the seasonal values, for the whole time series for the
-three sky conditions (Table \ref{tab:trendtable}).
+% copy output from Article.tex
 
@@
 
-{\centering \includegraphics[width=.70\linewidth]{./images/LongtermTrends-2}
+{\centering \includegraphics[width=0.7\linewidth]{./images/LongtermTrends-5}
 
@@
 
-\caption{Anomaly (\%) of the daily SDR relative to climatological values for 1993 - 2023  `r year(min(ALL_1_daily_DESEAS$Date))` to `r year(max(ALL_1_daily_DESEAS$Date))`.
-The black line shows the long term trend for all-sky conditions.}\label{fig:trendALL}
+\caption{Anomaly (\%) of the daily clear-sky SDR, relative to climatological values for1993 - 2023. The black line shows the long term linear trend for clear-sky conditions.}\label{fig:trendCLEAR}
 \end{figure}
-
-\footnotesize
-
-\begin{longtable}[]{@{}
-  >{\centering\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.2361}}
-  >{\centering\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.2222}}
-  >{\centering\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.2500}}@{}}
-\caption{\label{tab:trendtable}Trends of daily SDR means, for different
-sky conditions. All cases have a statistical significance above
-\(99.9\%\) \normalsize}\tabularnewline
-\toprule\noalign{}
-\begin{minipage}[b]{\linewidth}\centering
-Trend {[}\%/year{]}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
-Trend p-value
-\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
-Sky condition
-\end{minipage} \\
-\midrule\noalign{}
-\endfirsthead
-\toprule\noalign{}
-\begin{minipage}[b]{\linewidth}\centering
-Trend {[}\%/year{]}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
-Trend p-value
-\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
-Sky condition
-\end{minipage} \\
-\midrule\noalign{}
-\endhead
-\bottomrule\noalign{}
-\endlastfoot
-0.3996 & 8.665e-23 & All sky cond. \\
-0.3773 & 7.075e-18 & Clear sky cond. \\
-0.3594 & 3.318e-09 & Cloudy cond. \\
-\end{longtable}
-
-\hypertarget{long-term-trends-by-solar-zenith-angle.}{%
-\subsection{Long term trends by Solar zenith
-angle.}\label{long-term-trends-by-solar-zenith-angle.}}
-
-Analyzing the long term trends of SDR to bins of SZA, we can see the
-contribution of the geometry and time in a diurnal level (Figures
-\ref{fig:szatrends}a and \ref{fig:szatrends}b). Although there is a
-seasonal distribution of SZA than in not presented here.
 
@@
 
-{\centering \subfloat[Trend distribution for all-sky conditions.\label{fig:szatrends-1}]{\includegraphics[width=.35\linewidth]{./images/SzaTrends-13} }\subfloat[Trend distribution for cloud-sky conditions.\label{fig:szatrends-2}]{\includegraphics[width=.35\linewidth]{./images/SzaTrends-21} }
+{\centering \includegraphics[width=0.7\linewidth]{./images/LongtermTrends-8}
 
@@
 
-\caption{Distribution of the SDR's long term trends by SZA.}\label{fig:szatrends}
+\caption{Anomaly (\%) of the daily cloud-sky SDR, relative to climatological values for 1993 - 2023. The black line shows the long term linear trend for cloud-sky conditions.}\label{fig:trendCLOUD}
 \end{figure}
-
-\hypertarget{long-term-trends-by-season-of-year}{%
-\subsection{Long term trends by season of
-year}\label{long-term-trends-by-season-of-year}}
-
-Similar we have produced the deseasonalized trend for different sky
-conditions for each season of the year, using the corresponding mean
-monthly values. This can give us a better understanding of the annual
-variability of the trends.
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{./images/SeasonalTrendsTogether-1}
+{\centering \includegraphics[width=1\linewidth]{./images/SzaTrendsSeasTogether-2}
 
@@
 
-\caption{Trends by season for all condition. Displaying monthly means of daily means.}\label{fig:seasonalALL}
+\caption{Seasonla distribution of the SDR's long term trends by SZA before and after noon.
+The solid data points represent values with an acceptable statistical significance ($p<0.005$), the trends were computed with SZA bins of $1^\circ$ for each day, before and after local noon, all graphs have the same scale.
+Bins with $p<0.005$ or with less than 85 data points may be missing from the view.}\label{fig:SZAtrendSeason}
 \end{figure}
 
-\begin{longtable}[]{@{}
-  >{\centering\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2361}}
-  >{\centering\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2222}}
-  >{\centering\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2500}}
-  >{\centering\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1250}}@{}}
-\caption{\label{tab:trendseasontable}Trends of daily means by sky
-conditions for the seasons of the year. (continued
-below)}\tabularnewline
-\toprule\noalign{}
-\begin{minipage}[b]{\linewidth}\centering
-Trend {[}\%/year{]}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
-Trend p-value
-\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
-Sky condition
-\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
-Season
-\end{minipage} \\
-\midrule\noalign{}
-\endfirsthead
-\toprule\noalign{}
-\begin{minipage}[b]{\linewidth}\centering
-Trend {[}\%/year{]}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
-Trend p-value
-\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
-Sky condition
-\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
-Season
-\end{minipage} \\
-\midrule\noalign{}
-\endhead
-\bottomrule\noalign{}
-\endlastfoot
-0.7215 & 0.001209 & All sky cond. & Winter \\
-0.1537 & 0.1498 & All sky cond. & Spring \\
-0.128 & 0.08705 & All sky cond. & Summer \\
-0.4239 & 0.006659 & All sky cond. & Autumn \\
-0.8715 & 0.000269 & Clear sky cond. & Winter \\
-0.05881 & 0.6247 & Clear sky cond. & Spring \\
--0.02369 & 0.7972 & Clear sky cond. & Summer \\
-0.3583 & 0.02996 & Clear sky cond. & Autumn \\
--0.02581 & 0.9335 & Cloudy cond. & Winter \\
-0.4333 & 0.02937 & Cloudy cond. & Spring \\
-0.442 & 0.004988 & Cloudy cond. & Summer \\
-0.4776 & 0.01955 & Cloudy cond. & Autumn \\
-\end{longtable}
-
-\begin{longtable}[]{@{}
-  >{\centering\arraybackslash}p{(\columnwidth - 0\tabcolsep) * \real{0.4444}}@{}}
-\toprule\noalign{}
-\begin{minipage}[b]{\linewidth}\centering
-Trend statistical signif. {[}\%{]}
-\end{minipage} \\
-\midrule\noalign{}
-\endhead
-\bottomrule\noalign{}
-\endlastfoot
-99.88 \\
-85.02 \\
-91.29 \\
-99.33 \\
-99.97 \\
-37.53 \\
-20.28 \\
-97 \\
-6.646 \\
-97.06 \\
-99.5 \\
-98.04 \\
-\end{longtable}
-
-\hypertarget{consistency-of-the-trends}{%
-\subsection{Consistency of the trends}\label{consistency-of-the-trends}}
-
-A method to evaluate changes in the long term trend is to use the
-cumulative sum of the variable.
-
-Figure \ref{fig:cumsum}
-
-\ldots{} different in aggregation \ldots{}
-
-\begin{figure}[h!]
-
-{\centering \subfloat[one plot\label{fig:cumsum-1}]{\includegraphics[width=.35\linewidth]{./images/CumulativeDailyCumSum-1} }\subfloat[the other one\label{fig:cumsum-2}]{\includegraphics[width=.35\linewidth]{./images/CumulativeDailyCumSum-9} }
-
-}
-
-\caption{two plots}\label{fig:cumsum}
-\end{figure}
-
-\hypertarget{conclusions}{%
-\section{Conclusions}\label{conclusions}}
-
-Our result for all-sky condition (\(0.4\%/year\)), reaffirm the previous
-results of \citet{Bais2013} for the site. The increase of this trend,
-shows that the phenomenon and probably the causes have been amplified.
-
-\ldots.. Also, the trend of \(0.36\%/year\) for cloud-sky condition
-indicate the major part that clouds play on the \ldots..
-
-\ldots\ldots.
-
-About:
-
-\begin{itemize}
-\item
-  long terms trends
-\item
-  seasonal trends
-\item
-  effect of SZA
-\item
-  effect of clouds
-\item
-  Aerosols
-\end{itemize}
-
-similar results with Evidence for Clear‐Sky Dimming and Brightening in
-Central Europe\_Wild2021.pdf inclue in discusion
-
-\begin{longtable}[]{@{}c@{}}
-\toprule\noalign{}
-\endhead
-\bottomrule\noalign{}
-\endlastfoot
-\textbf{END} \\
-\end{longtable}
-
-\bibliography{Mainreferences.bib}
-
-\end{document}
+% \subsection{}
+% The appendix is an optional section that can contain details and data supplemental to the main text.
+For example, explanations of experimental details that would disrupt the flow of the main text, but nonetheless remain crucial to understanding and reproducing the research shown; figures of replicates for experiments of which representative data is shown in the main text can be added here if brief, or as Supplementary data.
+Mathematical proofs of results not central to the paper can be added as an appendix.
+%
+% \section{}
+% All appendix sections must be cited in the main text. In the appendixes, Figures, Tables, etc. should be labeled starting with `A', e.g., Figure A1, Figure A2, etc.

### 2023-11-02 Article.Rmd
Commit 2023-11-02 12:14
Commit 2023-11-02 13:10

@@
   Periods of positive tredns are refered as brightening periods and with negative tredns as dimming periods.
-  We studied 29 years of CHP1 data from Thessaloniki, Greece, under three sky conditions (clear sky, cloudy sky and all sky conditions), applying a cloud sky identification algorithm.
+  We studied 29 years of CM-21 data from Thessaloniki, Greece, under three sky conditions (clear sky, cloudy sky and all sky conditions), applying a cloud sky identification algorithm.
   We found a positive trend for all-sky and clear-sky conditions, and also, investigated the consistency of those trends, the effect of the solar zenith angle, and the variation of the trends for the seasons of the year.
@@
 numbersections:  true
+abbreviations:
+  - short: DNI
+    long: Direct beam/normal irradiance
+  - short: CSid
+    long: Clear sky identification algorithm
+  - short: CUSUM
+    long: Cumulative sum
+  - short: SDR 
+    long: Solar downward radiation
 header-includes:
@@
 It is interesting to note, that for the observation period, the trend of the TSI is
-$`r format(TSIslopeAll[1,1], scientific = T, digits = 2)`\,\%/y$, and thus we can not attribute any major effect on SDR trend to Solar variability.
+$`r format(TSIslopeAll[1,1], scientific = F, digits = 2)`\,\%/y$, and thus we can not attribute any major effect on SDR trend to Solar variability.
 
@@
                         format      = "latex",
-                        booktabs = T,
-                        escape   = T,
+                        booktabs    = TRUE,
+                        escape      = TRUE,
                         format.args = list(digits = 4),
                         linesep     = "",
-                        align    = "crr",
-                        caption  = paste0("\\label{tab:trendtable}Trends in SDR daily means for different sky conditions for the period ", year(min(ALL_1_daily_DESEAS$Date)), " - ", year(max(ALL_1_daily_DESEAS$Date)), "."))
+                        align       = "crrr",
+                        centering   = FALSE,
+                        table.envir = "table",
+                        position    = "H",
+                        caption     = paste0("Trends in SDR daily means for different sky conditions for the period ", year(min(ALL_1_daily_DESEAS$Date)), " - ", year(max(ALL_1_daily_DESEAS$Date)), "."))
+
+# x_latex %>%
+#     kable_styling(
+#         latex_options = c("hold_position")
+#     )
 x_latex %>% 
-    kable_styling(
-        latex_options = c("hold_position")
-    )
-        # position      =   "float_left"
+    kable_styling(full_width = TRUE) %>%
+    column_spec(1, width = "8em")
 ```
@@
                         format      = "latex",
-                        booktabs = T,
-                        escape   = T,
+                        booktabs    = TRUE,
+                        escape      = TRUE,
                         # format.args = list(digits = 3),
@@
                         align       = "ccrr",
-                        caption  = "\\label{tab:trendseasontable}SDR linear trends of monthly anomalies for each season of the year.")
+                        centering   = FALSE,
+                        table.envir = "table",
+                        # position    = "H",
+                        caption  = "SDR linear trends of monthly anomalies for each season of the year.")
+
+# x_latex %>%
+#     collapse_rows(columns = 1, valign = "middle", latex_hline = "custom") %>%
+#     kable_styling(latex_options = c("striped",
+#                                     "hold_position")
+#     )
 
@@
 
-```{r SZAtrendSeason, echo = FALSE, fig.cap='Seasonla distribution of the SDR\'s long term trends by SZA before and after noon.
+```{r SZAtrendSeason, echo = FALSE, fig.cap='Seasonal distribution of the SDR\'s long term trends by SZA before and after noon.
 The solid data points represent values with an acceptable statistical significance ($p<0.005$), the trends were computed with SZA bins of $1^\\circ$ for each day, before and after local noon, all graphs have the same scale.

### 2023-11-02 Article.tex
Commit 2023-11-02 13:10

@@
 of the data. Thus, our clear sky reference is derived by the
-Eq.~\ref{eq:ahau}.
-
-\begin{equation}
+Eq.~\ref{eq:ahau}. \begin{equation}
 f(a) = \frac{1}{n}\sum_{i=1}^{n} ( \text{SDR}_{\text{CSid},i} - a \times \text{SDR}_{\text{testCSref},i} )^2 \label{eq:minf}
-\end{equation}
-
-where: \(n\) is the total number of daylight records,
+\end{equation} where: \(n\) is the total number of daylight records,
 \(\text{SDR}_{\text{CSid},i}\) are the records identified as clear sky
@@
 \(\text{SDR}_{\text{testCSref},i}\) is any of the tested clear sky
-radiation models.
-
-\begin{equation}
+radiation models. \begin{equation}
 \text{SDR}_\text{CSref} = a \times \text{SDR}_\text{Haurwitz} = 0.965 \times 1098 \times \cos(\theta) \times \exp \left( \frac{ - 0.057}{\cos(\theta)} \right) \label{eq:ahau}
-\end{equation}
-
-where: \(\text{SDR}_\text{CSref}\) is the reference clear sky SDR, in
-\(\text{w}\,\text{m}^{-2}\) and \(\theta\) is the solar zenith angle
-(SZA).
+\end{equation} where: \(\text{SDR}_\text{CSref}\) is the reference clear
+sky SDR, in \(\text{w}\,\text{m}^{-2}\) and \(\theta\) is the solar
+zenith angle (SZA).
 
@@
 \citep{Wild2021, Yuan2021, Ohmura2009}. It is interesting to note, that
-for the observation period, the trend of the TSI is \(-2e-04\,\%/y\),
+for the observation period, the trend of the TSI is \(-0.0002\,\%/y\),
 and thus we can not attribute any major effect on SDR trend to Solar
@@
 
-\begin{table}[!h]
+\begin{table}[H]
 
-\caption{\label{tab:trendtable}\label{tab:trendtable}Trends in SDR daily means for different sky conditions for the period 1993 - 2023.}
-\centering
-\begin{tabular}[t]{crrc}
+\caption{\label{tab:trendtable}Trends in SDR daily means for different sky conditions for the period 1993 - 2023.}
+\begin{tabu} to \linewidth {>{\centering\arraybackslash}p{8em}>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X}
 \toprule
-Sky conditions & Trend [\%/year] & Statistical signif. [\%] & Days\\
+Sky conditions & Trend [\%/year] & Statistical signif. [\%] & Days with data\\
 \midrule
@@
 
+\begin{verbatim}
+Usually it is recommended to use column_spec before collapse_rows, especially in LaTeX, to get a desired result. 
+\end{verbatim}
+
 \begin{table}[!h]
 
-\caption{\label{tab:trendseasontable}\label{tab:trendseasontable}SDR linear trends of monthly anomalies for each season of the year.}
-\centering
-\begin{tabular}[t]{ccrr}
+\caption{\label{tab:trendseasontable}SDR linear trends of monthly anomalies for each season of the year.}
+\begin{tabu} to \linewidth {>{\centering\arraybackslash}p{8em}>{\centering}X>{\raggedleft}X>{\raggedleft}X}
 \toprule
@@
 
-\caption{Seasonla distribution of the SDR's long term trends by SZA before and after noon.
+\caption{Seasonal distribution of the SDR's long term trends by SZA before and after noon.
 The solid data points represent values with an acceptable statistical significance ($p<0.005$), the trends were computed with SZA bins of $1^\circ$ for each day, before and after local noon, all graphs have the same scale.

### 2023-11-02 MDPI_submition.Rmd
Commit 2023-11-02 13:10

@@
 abbreviations:
-  - short: MDPI
-    long: Multidisciplinary Digital Publishing Institute
-  - short: DOAJ
-    long: Directory of open access journals
-  - short: TLA
-    long: Three letter acronym
+  - short: DNI
+    long: Direct beam/normal irradiance
+  - short: CSid
+    long: Clear sky identification algorithm
+  - short: CUSUM
+    long: Cumulative sum
   - short: SDR 
     long: Solar downward radiation
+  - short: SZA
+    long: Solar zenith angle
+
 header-includes:
@@
                         booktabs    = TRUE,
-                        escape      = T,
+                        escape      = TRUE,
                         # format.args = list(digits = 3),
@@
                         caption  = "SDR linear trends of monthly anomalies for each season of the year.")
+
+# x_latex %>%
+#     collapse_rows(columns = 1, valign = "middle", latex_hline = "custom") %>%
+#     kable_styling(latex_options = c("striped",
+#                                     "hold_position")
+#     )
 

### 2023-11-02 MDPI_submition.tex
Commit 2023-11-02 13:10

@@
 are refered as brightening periods and with negative tredns as dimming
-periods. We studied 29 years of CHP1 data from Thessaloniki, Greece,
+periods. We studied 29 years of CM-21 data from Thessaloniki, Greece,
 under three sky conditions (clear sky, cloudy sky and all sky
@@
 \begin{tabular}{@{}ll}
-MDPI & Multidisciplinary Digital Publishing Institute \\
-DOAJ & Directory of open access journals \\
-TLA & Three letter acronym \\
+DNI & Direct beam/normal irradiance \\
+CSid & Clear sky identification algorithm \\
+CUSUM & Cumulative sum \\
 SDR & Solar downward radiation \\

### 2023-11-02 Article.Rmd
Commit 2023-11-02 14:14

@@
 #     text:         "Another author footnote."
+
 abstract: |
@@
   We indentified that there are some anomalies in the long term SDR trends, for all sky conditions.
-keywords: 
-  - GHI
-  - SDR
-  - Solar Brigthening/Dimming
+
+# back matter
+keywords: |
+  GHI; SDR; Solar Brigthening/Dimming.
+## list three to ten pertinent keywords specific to the article, yet reasonably common within the subject discipline
+
+acknowledgement: |
+  All sources of funding of the study should be disclosed. Please clearly 
+  indicate grants that you have received in support of your research work. 
+  Clearly state if you received funds for covering the costs to publish in open 
+  access.
+
+authorcontributions: |
+  For research articles with several authors, a short paragraph specifying their 
+  individual contributions must be provided. The following statements should be 
+  used ``X.X. and Y.Y. conceive and designed the experiments; X.X. performed the 
+  experiments; X.X. and Y.Y. analyzed the data; W.W. contributed 
+  reagents/materials/analysis tools; Y.Y. wrote the paper.'' Authorship must be
+  limited to those who have contributed substantially to the work reported.
+
+funding: |
+  Please add: ``This research received no external funding'' or ``This research 
+  was funded by NAME OF FUNDER grant number XXX.'' and  and ``The APC was funded 
+  by XXX''. Check carefully that the details given are accurate and use the 
+  standard spelling of funding agency names at 
+  \url{https://search.crossref.org/funding}, any errors may affect your future 
+  funding.
+
 journal:         "An awesome journal"
@@
     long: Solar downward radiation
+  - short: SZA
+    long: Solar zenith angle
+
 header-includes:
@@
    - \usepackage{subcaption}
+
 output: 
-  rticles::mdpi_article: default
+  rticles::mdpi_article:
+    extra_dependencies:  longtable 
+    keep_tex:            yes
   bookdown::word_document2: default
@@
 x_latex %>%
+    column_spec(1, width = "8em")                                         %>%
     collapse_rows(columns = 1, valign = "middle", latex_hline = "custom") %>%
-    kable_styling(latex_options = c('striped', 'hold_position'), full_width = TRUE) %>%
+    kable_styling(latex_options = c('striped', 'hold_position'), full_width = TRUE)
     # kable_styling(full_width = T) %>%
     # kable_styling("striped")      %>%
-    column_spec(1, width = "8em")
 

### 2023-11-02 Article_A.tex
Commit 2023-11-02 14:14

@@
 \usepackage{amsmath,amssymb}
-\usepackage{lmodern}
 \usepackage{iftex}
@@
 \else % if luatex or xetex
-  \usepackage{unicode-math}
+  \usepackage{unicode-math} % this also loads fontspec
   \defaultfontfeatures{Scale=MatchLowercase}
   \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
+\fi
+\usepackage{lmodern}
+\ifPDFTeX\else
+  % xetex/luatex font selection
 \fi
@@
 \usepackage{xcolor}
-\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
-\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
-\hypersetup{
-  pdftitle={Trends of SDR in Thessaloniki},
-  pdfkeywords={GHI, SDR, Solar Brigthening/Dimming},
-  hidelinks,
-  pdfcreator={LaTeX via pandoc}}
-\urlstyle{same} % disable monospaced font for URLs
 \usepackage[margin=1in]{geometry}
@@
 \usepackage{subcaption}
+\usepackage{booktabs}
+\usepackage{longtable}
+\usepackage{array}
+\usepackage{multirow}
+\usepackage{wrapfig}
+\usepackage{float}
+\usepackage{colortbl}
+\usepackage{pdflscape}
+\usepackage{tabu}
+\usepackage{threeparttable}
+\usepackage{threeparttablex}
+\usepackage[normalem]{ulem}
+\usepackage{makecell}
+\usepackage{xcolor}
 \ifLuaTeX
@@
 \fi
+\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
+\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
+\urlstyle{same}
+\hypersetup{
+  pdftitle={Trends of SDR in Thessaloniki},
+  pdfkeywords={GHI; SDR; Solar Brigthening/Dimming.},
+  hidelinks,
+  pdfcreator={LaTeX via pandoc}}
 
@@
 \author{true \and true}
-\date{2023-07-03}
+\date{2023-11-02}
 
@@
 \begin{abstract}
-Study of GHI and DNI radiation for `clear-sky' and all-sky conditions.
-It consists of two paragraphs.
+The shortwave downward solar irradiance (SDR) is an important factor that drives climate processes, production and can affect all living organisms.
+While monitoring the long term variability, there are observations of upward and downward SDR trends on different locations around the world for different time periods.
+Periods of positive tredns are refered as brightening periods and with negative tredns as dimming periods.
+We studied 29 years of CM-21 data from Thessaloniki, Greece, under three sky conditions (clear sky, cloudy sky and all sky conditions), applying a cloud sky identification algorithm.
+We found a positive trend for all-sky and clear-sky conditions, and also, investigated the consistency of those trends, the effect of the solar zenith angle, and the variation of the trends for the seasons of the year.
+We indentified that there are some anomalies in the long term SDR trends, for all sky conditions.
 \end{abstract}
@@
 
-The shortwave downward solar irradiance (SDR) at Earth's surface play a significant role, on it's climate.
+The shortwave downward solar irradiance (SDR) at Earth's surface play a significant role, on its climate.
 Changes of the SDR can be related to changes on Earth's energy budget, the mechanisms of climate change, and water and carbon cycle (Wild 2009).
-Can also affect, solar and agricultural production, and living organisms.
-Studies of SDR variability, have identified some distinct SDR trends on different regions of the world and for different time periods.
+Can also affect, solar and agricultural production, and all living organisms.
+Studies of SDR variability, have identified some distinct SDR trends on different regions of the world on different time periods.
 The term `brightening' is generally used to describe periods of positive SDR trend, and `dimming' for negative trend.
@@
 
-There is a consensus, among researchers, that major factors of SDR attenuation is the interaction with atmospheric aerosols and clouds.
+There is a consensus, among researchers, that the major factors of SDR attenuation is the interaction of Sun radiation with atmospheric aerosols and clouds.
 Those interactions, among other factors, have been analysed with models (Li et al. 2016; Samset et al. 2018), showing the existence of feedback mechanisms between the two.
@@
 
-Due to the variability of the phenomenon, and its contributing factors, there is a need to further investigate and monitor SDR, in different sites, to estimate its magnitude, and its relation to the local conditions.
+Due to the variability of the phenomenon, and its contributing factors, there is a constant need to investigate and monitor SDR, in different sites, to estimate its magnitude, and its relation to the local conditions.
 In this study, we examine the trends of SDR, with ground-based measurements at Thessaloniki, Greece for the period 1993 to 2023, as derived from a CM-21 pyranometer.
@@
 
-The SDR is equivalent to the whole sky Global radiation (or Global Horizontal Irradiance, GHI), and was obtained with a horizontal leveled, CM-21 pyranometer, from
-1993-04-12
+The SDR data were measured with a Kipp \& Zonen CM-21 pyranometer operating continuously at the Laboratory of Atmospheric Physics of the Aristotle University of Thessaloniki
+(\(40^\circ\,38'\,\)N, \(22^\circ\,57'\,\)E, \(80\,\)m~a.s.l.)
+in the period from
+1993-04-13
 to
-2023-03-31.
-Also, we have measurements of direct beam radiation (DNI) obtained by a collocated CHP-1 pyrheliometer, from
+2023-04-13.
+The monitoring site is located near the city centre, and we expect to be affected by the urban environment.
+During the study period, the pyranometer has been independently calibrated three times at the Meteorologisches Observatorium Lindenberg, DWD, when it was verified the stability of the instrument to within better than \(0.7\%\) relative to the initial calibration by the manufacturer.
+Along with SDR, the direct beam radiation (DNI) was also measured by a collocated Kipp \& Zonen CHP-1 pyrheliometer, for the period
 2016-04-01
 to
-2023-03-31.
-Although, we performed a similar analysis to the DNI as for the SDR, the results are not presented here, as they lack the appropriate statistical significance, due to the sorter duration of the data.
-Moreover, the DNI was used as auxiliary data, with the clear sky identification algorithm (CSid), to aid with the selection of the appropriate thresholds.
-We have to note also, that despite the capability of the CSid to use the DNI as an characterization parameter, we haven't use it, to avoid any selection bias due to unequal availability of measurements.
+2023-04-13.
+Although, we have performed a similar analysis to the DNI data the results are not presented here, as they lack the appropriate statistical significance, due to the sorter duration of the data.
+However, the DNI data were used as auxiliary data, in the clear sky identification algorithm (CSid), which is discussed later, for the selection of the appropriate thresholds.
+It is noted that despite the capability of the CSid algorithm to use the DNI as a characterization parameter, we haven't utilized it here, to avoid any selection bias, due to unequal length of the two datasets.
+There are four distinct steps in the creation of the dataset analysed here:
+a)~the acquisition of radiation measurements from the sensors,
+b)~the data quality check,
+c)~the identification of ``clear sky'' conditions from the radiometric data, and
+d)~the aggregation of data and trend analysis.
 
-The SDR data were measured with a Kipp \& Zonen CM-21 pyranometer operating continuously at the Laboratory of Atmospheric Physics of the Aristotle University of Thessaloniki (\(40^\circ38'\)N, \(22^\circ57'\)E, 80m a.s.l.).
-The monitoring site is located near the city centre.
-During the study period the CM-21 pyranometer has been independently calibrated four times.
-There are three distinct steps to the creation of this dataset:
-a) the acquisition of radiation measurements from the sensors,
-b) a radiation data quality check, and
-c) the identification of ``clear sky'' conditions from the radiometric data
-d) data aggregation and trend analysis.
+For the acquisition of radiometric data, the signal of the pyranometer is sampled with a rate of \(1\,\text{Hz}\).
+The mean and the standard deviation of these samples are recorded every minute.
+The measurements are corrected for the zero offset (``dark signal'' in volts).
+The ``dark signal'' is calculated by averaging all measurements recorded for a period of
+\(3\,\text{h}\),
+before (morning) or after (evening) the Sun reaches an elevation angle of
+\(-10^\circ\).
+The signal is converted to irradiance using a ramped value of the instrument's sensitivity between calibrations.
 
-For the acquisition of radiometric data,
-the signal of the broadband instruments is sampled with a rate of \(1 \text{Hz}\). The mean and its standard deviation are recorded for every minute.
-The measurements are corrected for the zero offset of the instrument signal.
-As reference of the ``dark signal'' is used the Sun elevation angle bellow \(-10^\circ\), for a period of \(3 \text{h}\).
-The signal is converted to radiation flux, using a ramped value of the instrument sensitivity, derived from the eight laboratory calibrations of the instrument, during the study period.
+A manual screening was performed, to remove inconsistent and erroneous recordings that can occur stochastically or systematically, during the continuous operation of the instruments.
+The manual screening is aided by a radiation data quality assurance procedure, adjusted for the site, which is based on the methods of
+Long and Shi~(2008, 2006).
+Thus, problematic recordings have been excluded from further processing.
+Although it is impossible to detect all false data, the large number of available data, and the aggregation scheme we used, ensures the good quality of the radiometric measurements used in this study.
 
-A manual screening was performed, to remove inconsistencies and erroneous recordings, that can occur randomly or systematically, during the long continuous operation of the instrument.
-The manual screening is aided by a radiation data quality assurance procedure, adjusted for this site, based on methods of Long and Shi~(2008, 2006).
-Thus, problematic recordings have been excluded from further processing.
-Although it is not possible to detect all the bad data, the large number of data and the aggregation scheme we use, can provide us with accurate radiometric measurements, for the scope of this study.
+In order to be able to estimate the effect of the sky condition on the long term variability of SDR, we created three datasets, by characterizing each one-minute measurement with a corresponding sky condition (i.e., all-sky, clear-sky and cloudy-sky).
+To identify the clear-sky conditions we used a method proposed by
+Long and Ackerman (2000)
+and by
+Reno and Hansen (2016),
+which was adapted and configured for the site, as the authors suggest.
 
-In order to estimate the effect of the clouds on the SDR, we created three datasets, by characterizing each one-minute measurement with a corresponding sky condition.
-The all-sky conditions, containing all the valid measurements.
-The clear-sky conditions data, correspond to data where we have inferred that the sky was almost clear of clouds, and the remainder part the data was characterized with cloud sky conditions (cloud-sky).
-To identify the clear-sky conditions we used the method proposed by C. N. Long and Ackerman (2000) and by Reno and Hansen (2016), that was adapted and configured for the site.
+We have to note, that the definition of what constitutes as clear or cloudy sky, has some subjectivity, in any method of characterization.
+As a result, the details of the definition are site specific, it relies on a combination of thresholds and comparisons with ideal actinometric models and statistical analysis on different signal metrics.
+The CSid algorithm was calibrated with the main focus, to identify the presence of clouds on the sky dome.
+Although the fine-tuning of the procedure, few marginal cases exist, that have been identified manually as false positive or false negative but cannot affect the final results of the study.
 
-\ldots.
+For completeness, we will provide below a brief overview of the
+clear sky identification algorithm (CSid),
+along with the site specific thresholds.
+To calculate the reference clear sky
+\(\text{SDR}_\text{CSref}\) we used the \(\text{SDR}_\text{Haurwitz}\) derived by
+the radiation model of Haurwitz (1945), adjusted for our site with a
+factor \(a\) (Eq.~\ref{eq:ahau}), resulted by an iterative optimization process, as described
+by Long and Ackerman (2000) and Reno and Hansen (2016).
+The target of
+the optimization was the minimization of a function \(f(a)\) (Eq.~\ref{eq:minf}) and
+was accomplished with the algorithmic function ``optimise'', which is an implementation based on the work of Brent (1973), from the library ``stats'' of the R programming language (R Core Team 2023).
+The optimization and the selection of the clear sky reference model, was performed on SDR observations for the period 2016 - 2021.
+During the optimization, eight simple clear sky radiation models were tested (Daneshyar-Paltridge-Proctor, Kasten-Czeplak, Haurwitz, Berger-Duffie, Adnot-Bourges-Campana-Gicquel, Robledo-Soler, Kasten and Ineichen-Perez), with a wide range of factors.
+These models are described in more details by Reno, Hansen, and Stein (2012b) and evaluated by Reno and Hansen (2016).
+We found, that Haurwitz's model, adjusted with the factor \(a = 0.965\) yields one of the lowest root mean squared errors (RMSE),
+while the procedure, manages to characterize the majority of the data.
+Thus, our clear sky reference is derived by the Eq.~\ref{eq:ahau}.
+\begin{equation}
+f(a) = \frac{1}{n}\sum_{i=1}^{n} ( \text{SDR}_{\text{CSid},i} - a \times \text{SDR}_{\text{testCSref},i} )^2 \label{eq:minf}
+\end{equation}
+where: \(n\) is the total number of daylight records, \(\text{SDR}_{\text{CSid},i}\) are the records identified as clear sky by CSid, \(a\) is a hypothetical adjustment factor, and \(\text{SDR}_{\text{testCSref},i}\) is any of the tested clear sky radiation models.
+\begin{equation}
+\text{SDR}_\text{CSref} = a \times \text{SDR}_\text{Haurwitz} = 0.965 \times 1098 \times \cos(\theta) \times \exp \left( \frac{ - 0.057}{\cos(\theta)} \right) \label{eq:ahau}
+\end{equation}
+where: \(\text{SDR}_\text{CSref}\) is the reference clear sky SDR, in \(\text{w}\,\text{m}^{-2}\) and \(\theta\) is the solar zenith angle (SZA).
 
-cite iterative process
+The criteria that were used to identify whether a measurement was taken
+under clear-sky conditions are presented below.
+A data point is flagged
+as ``clear-sky'' if all criteria are satisfied, otherwise it is considered to be ``cloud-sky''.
+Each criterion was applied
+for a running window of \(11\) consecutive one-minute measurements, and
+the characterization is assigned to the central value of the window.
+Each parameter, was calculated both from the observations and the
+reference clear sky model, for each comparison.
+The allowable range of variation is defined by the
+model-derived value of the parameter multiplied by a factor plus an
+offset.
+The factors and the offsets were
+determined empirically, by manual inspecting each filters performance on
+selected days, and adjusting them accordingly, during an iterative
+process.
 
-\ldots\ldots{}
-
-We have to note, that the definition of what constitutes as clear or cloudy sky, has some subjectivity, due to the used method of characterization.
-As a result, the details of the definition are site specific, it relies on a combination of threshold of comparisons with ideal actinometric models and statistics on different signal behavior.
-
-The algorithm was calibrated with the main focus to identify the presence of clouds on the sky dome, although there are marginal cases that there will be false positives or false negatives identifications.
-
-As global radiation clear sky reference we are using the Haurwitz's model, adjusted for our site with a factor of 0.965 (Eq. \ref{eq:ahau}).
-The selection of a clear sky reference model, was based on SDR observation from the period 2016 -- 2021.
-Where, after an iterative optimization of eight simple models (Daneshyar--Paltridge--Proctor, Kasten--Czeplak, Haurwitz, Berger--Duffie, Adnot--Bourges--Campana--Gicquel, Robledo-Soler, Kasten and Ineichen-Perez) with different factors.
-We found, that Haurwitz's model, adjusted with a factor of 0.965 has the lower root mean squared error (RMSE).
-The tried models are described by Reno, Hansen, and Stein (2012b) and tested by Reno and Hansen (2016).
-The iterative optimization method, for the selection of the reference is discussed by C. N. Long and Ackerman (2000) and Reno and Hansen (2016).
-
+\begin{enumerate}
+\def\labelenumi{\alph{enumi})}
+\tightlist
+\item
+  Mean of the measured \(\overline{\text{SDR}}_i\) (Eq. \ref{eq:MeanVIP}).
   \begin{equation}
-\text{SDR}_\text{Clear Sky} = 0.965 \times 1098 \times \cos( \text{SZA} ) \times \exp \left( \frac{ - 0.057}{\cos(\text{SZA})} \right) \label{eq:ahau}
-\end{equation}
-
-The following criteria and thresholds were used to identify clear-sky conditions.
-Each criterion was applied for a running window of \(11\) consecutive measurements/minutes, and the characterization is applied at the center value of the window.
-A data point is consider as under cloud-sky condition if it fails to pass any of the criteria, all the other data points are characterized as clear-sky.
-
-The mean of the measured value \(\overline{G}_i\) must be inside an envelope based on the reference model \(\text{SDR}_\text{Clear Sky}\) (Eq. \ref{eq:MeanVIP}).
-
-\begin{equation}
-0.91 \times \overline{\text{SDR}}_{i\text{Clear Sky}} - 20
-< \overline{G}_i <
-1.095 \times \overline{\text{SDR}}_{i\text{Clear Sky}} + 30
+  0.91 \times \overline{\text{SDR}}_{\text{CSref},i} - 20
+  < \overline{\text{SDR}}_i <
+  1.095 \times \overline{\text{SDR}}_{\text{CSref},i} + 30
   \label{eq:MeanVIP}
   \end{equation}
-
-The running max measured value \(M_{Gi} = max[\text{SDR}_{i}]\), is compared to a similar constructed value from the reference \(M_{CSi} = max[\text{SDR}_{i\text{Clear Sky}}]\) (Eq. \ref{eq:MaxVIP}).
-
+\item
+  Maximum measured value \(M_{\text{}}\) (Eq.~\ref{eq:MaxVIP}).
   \begin{equation}
-1 \times M_{CSi} - 75
-< M_{Gi} <
-1 \times M_{CSi} + 75
+  1 \times M_{\text{CSref},i} - 75
+  < M_{\text{}i} <
+  1 \times M_{\text{CSref},i} + 75
   \label{eq:MaxVIP}
   \end{equation}
-
-The length \(L\) (Eq. \ref{eq:VILeq}) of the sequence of line segments connecting the points of the SDR time
-series for the measured values \(L\) and similar for the reference \(L_{CS}\), must be within the limits of Eq. \ref{eq:VILcr}.
-
+\item
+  Length \(L_i\) of the sequential line segments, connecting the points of the \(11\) SDR values (Eq. \ref{eq:VILeq}).
   \begin{equation}
-L = \sum_{i=1}^{n-1}\sqrt{\left ( \text{SDR}_{i+1} - \text{SDR}_{i}\right )^2 + \left ( t_{i+1} - t_i \right )^2}
+  L_i = \sum_{i=1}^{n-1}\sqrt{\left ( \text{SDR}_{i+1} - \text{SDR}_{i}\right )^2 + \left ( t_{i+1} - t_i \right )^2}
   \label{eq:VILeq}
   \end{equation}
-
   \begin{equation}
-1 \times L_{CSi} - 5 < L_i < 1.3 \times L_{CSi} + 13
+  1 \times L_{\text{CSref},i} - 5 < L_i < 1.3 \times L_{\text{CSref},i} + 13
   \label{eq:VILcr}
   \end{equation}
+  where: \(t_i\) is the time each SDR measurement has been measured
+\item
+  Standard deviation \(\sigma_i\) of the slope (\(s_i\)) between the \(11\) sequential points, normalized by the mean \(\overline{\text{SDR}}_i\) (Eq.~\ref{eq:VCT1}).
+  \begin{gather}
+    \sigma_i = \frac {\sqrt{\frac{1}{n-1} \sum_{i=1}^{n-1} \left( s_i - \bar{s} \right)^2}} {\overline{\text{SDR}}_i} \label{eq:VCT1} \\
+    s_i = \frac{\text{SDR}_{i+1} - \text{SDR}_{i}}{t_{i+1} - t_i},\;\;   \bar{s} = \frac{1}{n-1} \sum_{i=1}^{n-1} s_i,\;\;\forall i \in \left \{ 1, 2, \ldots, n-1 \right \}\;\;
+  \end{gather}
+  For this criterion, \(\sigma_i\) should be below a certain threshold (Eq.~\ref{eq:VCTcr}):
+  \begin{equation}
+    \sigma_i < \ensuremath{1.1\times 10^{-4}} \label{eq:VCTcr}
+  \end{equation}
+\item
+  Maximum difference \(X_i\) between the change in measured irradiance and the change in clear sky irradiance over each measurement interval.
+  \begin{gather}
+    X_i = \max{\left \{ \left | x_i - x_{\text{CSref},i} \right | \right \}} \label{eq:VSM3} \\
+    x_i = \text{SDR}_{i+1} - \text{SDR}_{i} \forall i \in \left \{ 1, 2, \ldots, n-1 \right \} \label{eq:VSM1} \\
+    x_{\text{CSref},i} = \text{SDR}_{\text{CSref},i+1} - \text{SDR}_{\text{CSref},i} \forall i \in \left \{ 1, 2, \ldots, n-1 \right \} \label{eq:VSM2}
+  \end{gather}
+  For this criterion, \(X_i\) should be below a certain threshold (Eq.~\ref{eq:VSMcr}):
+  \begin{equation}
+    X_i < 7.5 \label{eq:VSMcr}
+  \end{equation}
+\end{enumerate}
 
-We calculate the standard deviation \(\sigma\) of the slope
-(\(s\)) between sequential points in the time series, normalized by the average SDR during the time interval.
-
-\begin{equation}
-s_i = \frac{\text{SDR}_{i+1} - \text{SDR}_{i}}{t_{i+1} - t_i}, \forall i \in \left \{ 1, 2, \ldots, n-1 \right \},\;\;
-\bar{s} = \frac{1}{n-1} \sum_{i=1}^{n-1} s_i,\;\;
-\sigma_i = \frac {\sqrt{\frac{1}{n-1} \sum_{i=1}^{n-1} \left( s_i - \bar{s} \right)^2} } {\bar{G_i}} \label{eq:VCT3b}
-\end{equation}
-
-\begin{equation}
-\sigma_i < \ensuremath{1.1\times 10^{-4}}
-\label{eq:VCTcr}
-\end{equation}
-
-The maximum difference \(X\) between the change in measured irradiance and the change in clear sky
-irradiance over each measurement interval.
-
-\begin{gather}
-x_i = \text{SDR}_{i+1} - \text{SDR}_{i} \forall i \in \left \{ 1, 2, \ldots, n-1 \right \} \label{eq:VSM1} \\
-x_{CS,i} = \text{SDR}_{CS,i+1} - \text{SDR}_{CS,i} \forall i \in \left \{ 1, 2, \ldots, n-1 \right \} \label{eq:VSM2} \\
-X_i = \max{\left \{ \left | x_i - x_{CS,i} \right | \right \}} \label{eq:VSM3}
-\end{gather}
-
-\begin{equation}
-X_i < 7.5
-\label{eq:VSMcr}
-\end{equation}
-
-Due to a significant measurement uncertainty near, the horizon, we have to exclude
-all measurements with SZA greater than \(85^\circ\).
+Due to a significant measurement uncertainty near the horizon, we have to exclude all measurements with SZA greater than \(85^\circ\).
 Moreover, due to some obstructions around the site (hills and buildings), we excluded data with Azimuth angle between
@@
 On the latter instances, Sun is systematically, not visible by the instrument's location.
-To make the measurements comparable throughout the dataset, we adjusted all 1-minute radiometric values to the mean Sun - Earth distance.
-Subsequently, we made all measurements relative to the Total Solar Irradiance (TSI) at \(1 \text{au}\), in order to compensate for the Sun's intensity variability, using a homogenized time series of satellite TSI observations.
-The TSI data we use, comes from NOAA (Coddington et al. 2005), it overlaps with our data, and is interpolated to match with the time step of the data.
-As a result, we can present all radiation data as a fraction of Sun's TSI.
+To make the measurements comparable throughout the dataset, we adjusted all one-minute radiometric values to the mean Sun - Earth distance.
+Subsequently, we made all measurements relative to the Total Solar Irradiance (TSI) at \(1\,\text{au}\), in order to compensate for the Sun's intensity variability, using a time series of satellite TSI observations.
+The TSI data we use are part of the ``NOAA Climate Data Record of Total Solar Irradiance'' dataset (Coddington et al. 2005).
+Where the initial daily values, were interpolated to match with the time step of our measurements.
+The final dataset contains
+\(6589967\)
+one-minute measurements, of which,
+\(84.2\%\)
+were identified as under clear-sky conditions and subsequently
+\(15.8\%\)
+as under cloud-sky conditions.
 
-\hypertarget{aggregation-of-radiometric-data}{%
-\subsubsection{Aggregation of radiometric data}\label{aggregation-of-radiometric-data}}
-
-Before further analysis and deseasonalization, we implement an appropriate aggregation scheme on the 1-minute data. To preserve the representativeness we use the following criteria:
-a) for daily mean values we exclude instances with less than 180 valid data points,
-b) accordingly, monthly values are computed by daily aggregated data, where months with less than 20 days are rejected.
-The daily seasonal values were indexed based on the day of year number, and the monthly by the corresponding calendar month.
-For the seasons of the year trends, we grouped the mean daily values by season (December - February: Winter, March - May: Spring, etc.).
-Finally, for each data set, we remove the natural occurring seasonal variation.
-This is done, by calculating the mean values for the appropriate time step, and subtracting the annual cycle from the actual data.
-In addition to the previous aggregation scheme, each dataset was aggregate in \(1^\circ\) SZA bins, separate for cases before and after local noon.
-This gave us an approximation of Sun's SZA contribution to the SDR brightening.
-
-Thus, for each data set we have obtained the relative departure of the seasonal mean.
-The statistical significance and aggregation scheme and filtering will be noted along with the corresponding results.
+In order to investigate the SDR trends, we implemented an appropriate aggregation scheme to the 1-minute data to derive a series in coarser timescale.
+To preserve the representativeness of the data we used the following criteria:
+a) for the daily mean values we accept days with more than 50\% of the daytime measurements, present and valid,
+b) monthly values were computed from daily means
+only when at least 20 days were available.
+To create the daily and monthly climatological means, we averaged the data based on the day of year and calendar month, respectively.
+For the seasonal means we averaged the mean daily values in each season (Winter: December - February, Spring: March - May, etc.).
+Finally, each data set was deseasonalized by subtracting the corresponding climatological annual cycle (daily or monthly) from the actual data.
+To estimate SZA contribution to the SDR trends, the one-minute data were aggregated in \(1^\circ\) SZA bins, separately for the morning and afternoon hours, and then were deseasonalized as mentioned above.
 
@@
 
-We calculated the SDR trends, from the mean daily values of the departure from the seasonal values, for the whole time series for the three sky conditions (Table \ref{tab:trendtable}).
+We calculated the linear SDR trends, from the departures of the mean daily values from the seasonal values, and for the three sky conditions (Table~\ref{tab:trendtable}).
+In Figure~\ref{fig:trendALL} we present only the time series under all-sky conditions; the plots for clear-sky and cloud-sky conditions, are very similar and are shown in the Appendix (Figures~\ref{fig:trendCLEAR} and~ \ref{fig:trendCLOUD}).
+We observe a positive trend for all-sky conditions
+(\(0.38\,\%/y\)), a very close but lower trend for clear-skies (\(0.35\,\%/y\)) and a negative weaker trend for cloud-skies (\(-0.28\,\%/y\)).
+In the studied period, there is no significant break or change in the variability of the time series.
+Other studies for the European region report a change of the SDR slope, around 1980, a few years before the start of our records (Wild et al.
+2021; Yuan, Leirvik, and Wild 2021; Ohmura 2009).
+It is interesting to note, that for the observation period, the trend of the TSI is
+\(-0.0002\,\%/y\), and thus we can not attribute any major effect on SDR trend to Solar variability.
+
+\begin{table}[H]
+
+\caption{(\#tab:trendtable)Trends in SDR daily means for different sky conditions for the period 1993 - 2023.}
+\begin{tabu} to \linewidth {>{\centering\arraybackslash}p{8em}>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X}
+\toprule
+Sky conditions & Trend [\%/year] & Statistical signif. [\%] & Days with data\\
+\midrule
+All sky & 0.376 & 100.00 & 10256\\
+Clear sky & 0.349 & 100.00 & 10256\\
+Cloudy sky & -0.276 & 99.97 & 5067\\
+\bottomrule
+\end{tabu}
+\end{table}
 
@@
 
-\caption{Anomaly (\%) of the daily SDR relative to climatological values for 1993 - 2023  `r year(min(ALL_1_daily_DESEAS$Date))` to `r year(max(ALL_1_daily_DESEAS$Date))`.
-The black line shows the long term trend for all-sky conditions.}(\#fig:trendALL)
+\caption{Anomaly (\%) of the daily all-sky SDR, relative to climatological values for 1993 - 2023. The black line shows the long term linear trend.}(\#fig:trendALL)
 \end{figure}
 
-\footnotesize
-
-\begin{longtable}[]{@{}
-  >{\centering\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.2361}}
-  >{\centering\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.2222}}
-  >{\centering\arraybackslash}p{(\columnwidth - 4\tabcolsep) * \real{0.2500}}@{}}
-\caption{\label{tab:trendtable}Trends of daily SDR means, for different sky conditions. All cases have a statistical significance above \(99.9\%\)
-\normalsize}\tabularnewline
-\toprule
-\begin{minipage}[b]{\linewidth}\centering
-Trend {[}\%/year{]}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
-Trend p-value
-\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
-Sky condition
-\end{minipage} \\
-\midrule
-\endfirsthead
-\toprule
-\begin{minipage}[b]{\linewidth}\centering
-Trend {[}\%/year{]}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
-Trend p-value
-\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
-Sky condition
-\end{minipage} \\
-\midrule
-\endhead
-0.3996 & 8.665e-23 & All sky cond. \\
-0.3773 & 7.075e-18 & Clear sky cond. \\
-0.3594 & 3.318e-09 & Cloudy cond. \\
-\bottomrule
-\end{longtable}
-
-\hypertarget{long-term-trends-by-solar-zenith-angle.}{%
-\subsection{Long term trends by Solar zenith angle.}\label{long-term-trends-by-solar-zenith-angle.}}
-
-Analyzing the long term trends of SDR to bins of SZA, we can see the contribution of the geometry and time in a diurnal level (Figures \ref{fig:szatrends}a and \ref{fig:szatrends}b).
-Although there is a seasonal distribution of SZA than in not presented here.
+Although the year-to-year variability of the anomalies (Figure \ref{fig:trendALL} and Figures \ref{fig:trendCLEAR}, \ref{fig:trendCLOUD} on Appendix), shows a rather homogeneous behavior, plots of the cumulative sums (CUSUM) of the anomalies can reveal different structures within all the three sky conditions (Regier, Briceño, and Boyer 2019).
+In the cases of all-sky and clear-sky (Figures \ref{fig:cusummonth-1} and \ref{fig:cusummonth-2}), we observe three macroscopic periods.
+A downward part from the start until about 2005, an relative steady part until about 2016 and finally, a steep upward part until the present.
+For cloud-sky (Figure~\ref{fig:cusummonth-3}), we have a different sequence, it begins with a relative steady part until 1997, followed by a upward part until 2005, a long decline until 2020, with a small positive slope until the present.
+For a uniform trend, we would expect the CUSUMs of the anomalies to have a symmetric `V' shape.
+This would indicate that the anomalies are evenly distributed around the climatological mean, and for a positive uniform trend, the first half to be below and the other half above the climatological mean.
+In our case, there is more complex evolution of the phenomenon.
+Another distinct feature of the CUSUMs, is the different pattern of the cloudy-sky dataset which peaks around the middle of the period.
+Although, it seems to exist a complementary relation to the CUSUMs of the clear- and all-sky cases, we can not assert that clouds are the main drive for this relation, due to the great difference in the number of observational data between the datasets (Table~\ref{tab:trendtable}).
 
@@
 
-{\centering \subfloat[Trend distribution for all-sky conditions.(\#fig:szatrends-1)]{\includegraphics[width=.35\linewidth]{./images/SzaTrends-13} }\subfloat[Trend distribution for cloud-sky conditions.(\#fig:szatrends-2)]{\includegraphics[width=.35\linewidth]{./images/SzaTrends-21} }
+{\centering \subfloat[All-skies.(\#fig:cusummonth-1)]{\includegraphics[width=.32\linewidth]{./images/CumulativeMonthlyCuSum-1} }\subfloat[Clear-skies.(\#fig:cusummonth-2)]{\includegraphics[width=.32\linewidth]{./images/CumulativeMonthlyCuSum-5} }\subfloat[Cloud-skies.(\#fig:cusummonth-3)]{\includegraphics[width=.32\linewidth]{./images/CumulativeMonthlyCuSum-9} }
 
@@
 
-\caption{Distribution of the SDR's long term trends by SZA.}(\#fig:szatrends)
+\caption{Running cumulative sum plots of the monthly SDR.}(\#fig:cusummonth)
+\end{figure}
+
+In order to investigate more, the features within the trends, we created another set of CUSUMs plots by subtracting the corresponding long term trend from the SDR anomaly data, prior to the CUSUM calculation (Figure~\ref{fig:cusumnotrendmonthly}).
+With this approach periods when the CUSUMs diverge from zero can be interpreted as systematic variation of SDR from the climatological mean.
+When the CUSUM is increasing, the added values are above the climatological values of SDR trend and vice versa.
+Overall, for all- and clear-sky conditions (Figure~\ref{fig:cusumnotrendmonthly-1} and~\ref{fig:cusumnotrendmonthly-2}) we observe periods when the anomalies diverge from the climatological value, each lasting for several years.
+The pattern for the all-sky and clear-sky conditions is very similar, suggesting prevalence in clear skies over Thessaloniki.
+It is interesting that in the period 1993 - 2016 the anomalies of all-sky and clear-sky, have a high variability around zero.
+After 2016, the range of the variability is decreased at about one third of the prior period.
+For cloud-sky conditions (Figure~\ref{fig:cusumnotrendmonthly-3}) the period 1997 - 2008 is dominated by positive CUSUMs suggesting a reduced effect of clouds on SDR.
+From 1997 to mid-2000s CUSUMs are increasing, likely due to a continuous decrease in the thickness of clouds, followed by a period of rapid increase (within 3 years) in cloud optical thickness lasting up to 2008.
+The following stable period spans for about 15 years up to 2021 when CUSUMs start increasing again.
+
+\begin{figure}[h!]
+
+{\centering \subfloat[All-skies.(\#fig:cusumnotrendmonthly-1)]{\includegraphics[width=.32\linewidth]{./images/CumulativeMonthlyCuSumNOtrend-1} }\subfloat[Clear-skies.(\#fig:cusumnotrendmonthly-2)]{\includegraphics[width=.32\linewidth]{./images/CumulativeMonthlyCuSumNOtrend-5} }\subfloat[Cloud-skies.(\#fig:cusumnotrendmonthly-3)]{\includegraphics[width=.32\linewidth]{./images/CumulativeMonthlyCuSumNOtrend-9} }
+
+}
+
+\caption{Running cumulative sum plots of the monthly SDR after removing the long term trend from the monthly values.}(\#fig:cusumnotrendmonthly)
+\end{figure}
+
+\hypertarget{effects-of-the-solar-zenith-angle-on-sdr.}{%
+\subsection{Effects of the solar zenith angle on SDR.}\label{effects-of-the-solar-zenith-angle-on-sdr.}}
+
+The solar zenith angle is a major driver of the attenuation of SDR reaching the ground, as SDR is affected by the enhancement of the radiation path in the atmosphere, especially in urban environment with human activities emitting aerosols (Wang et al.
+2021).
+In order to estimate the effect of the SZA on the SDR trends, we grouped the anomaly data in bins of \(1^\circ\) SZA, and calculated the overall trend for each bin before noon and after noon (Figure~\ref{fig:szatrends}).
+Although there are seasonal dependencies of the minimum SZA (see Appendix, Figure~\ref{fig:SZAtrendSeason}), these dependencies would not be further examined here.
+For all-sky and clear-sky conditions the brightening effect of SDR (positive trend) is stronger for large SZAs (Figures~\ref{fig:szatrends-1} and \ref{fig:szatrends-2}).
+The trends in the morning and afternoon hours are more or less consistent with small differences, which can be attributed to systematic diurnal variations of aerosols, particularly during the warm period of the year (Wang et al.
+2021).
+For cloud-sky conditions (Figure~\ref{fig:szatrends-3}), we can not discern any significant dependence of the SDR trend with SZA.
+For SZAs \(16^\circ\) - \(50^\circ\), the trends range within about \(\pm 0.2\,\%/y\), with a weak statistical significance.
+Between \(50^\circ\) and \(75^\circ\) SZA the trends for the period before noon are stronger and negative, possibly associated with stronger attenuation by clouds under oblique incidence angles.
+
+\begin{figure}[h!]
+
+{\centering \subfloat[All-sky conditions.(\#fig:szatrends-1)]{\includegraphics[width=.32\linewidth]{./images/SzaTrends-1} }\subfloat[Clear-sky conditions.(\#fig:szatrends-2)]{\includegraphics[width=.32\linewidth]{./images/SzaTrends-4} }\subfloat[Cloud-sky conditions.(\#fig:szatrends-3)]{\includegraphics[width=.32\linewidth]{./images/SzaTrends-7} }
+
+}
+
+\caption{Distribution of the SDR's long term trends by SZA before and after noon.
+The solid data points represent values with an acceptable statistical significance ($p<0.005$), the trends were computed with SZA bins of $1^\circ$ for each day, before and after local noon.}(\#fig:szatrends)
 \end{figure}
@@
 
-Similar we have produced the deseasonalized trend for different sky conditions for each season of the year, using the corresponding mean monthly values.
-This can give us a better understanding of the annual variability of the trends.
+Similar to the long term trends discussed above, we have calculated the trend of the anomalies for the three different sky conditions, and for each season of the year, using the corresponding mean monthly values
+(Figure~\ref{fig:seasonalALL} and Table~\ref{tab:trendseasontable}).
+For all-sky conditions the trend in SDR in winter is the largest
+(\(0.69\,\%/y\)),
+followed by the trend in autumn
+(\(0.43\,\%/y\),
+value similar to the long term trend) both statistically significant above the \(99\,\%\) confidence level.
+In spring and summer, the trends are much smaller and less statistical significant.
+These seasonal differences indicate a possible relation of the trends to clouds during winter and autumn.
+For clear-skies, the trend in winter is \(0.83\,\%/y\), larger than for all-skies (\(0.69\,\%/y\)), which is another indication of increasing cloud thickness.
+Moreover, the trends under clear- and cloud-sky conditions are almost complementary to each other, particularly for winter and autumn, where the signal is stronger.
+During spring and summer the statistical significance is very low and the actual trend too small for a meaningful comparison.
 
@@
 
-\caption{Trends by season for all condition. Displaying monthly means of daily means.}(\#fig:seasonalALL)
+\caption{SDR trends by season of the year for the three sky conditions.
+Each point is a monthly mean calculated from daily means.
+The vertical scale is common for all graphs, each column of graphs corresponds to a sky condition and each row to a season of the year.}(\#fig:seasonalALL)
 \end{figure}
 
-\begin{longtable}[]{@{}
-  >{\centering\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2361}}
-  >{\centering\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2222}}
-  >{\centering\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2500}}
-  >{\centering\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1250}}@{}}
-\caption{\label{tab:trendseasontable}Trends of daily means by sky conditions for the seasons of the year. (continued below)}\tabularnewline
+\begin{table}[!h]
+
+\caption{(\#tab:trendseasontable)SDR linear trends of monthly anomalies for each season of the year.}
+\begin{tabu} to \linewidth {>{\centering\arraybackslash}p{8em}>{\centering}X>{\raggedleft}X>{\raggedleft}X}{>{\centering\arraybackslash}p{8em}crr}
 \toprule
-\begin{minipage}[b]{\linewidth}\centering
-Trend {[}\%/year{]}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
-Trend p-value
-\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
-Sky condition
-\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
-Season
-\end{minipage} \\
+Sky condition & Season & Trend [\%/year] & Statistical signif. [\%]\\
 \midrule
-\endfirsthead
-\toprule
-\begin{minipage}[b]{\linewidth}\centering
-Trend {[}\%/year{]}
-\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
-Trend p-value
-\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
-Sky condition
-\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
-Season
-\end{minipage} \\
-\midrule
-\endhead
-0.7215 & 0.001209 & All sky cond. & Winter \\
-0.1537 & 0.1498 & All sky cond. & Spring \\
-0.128 & 0.08705 & All sky cond. & Summer \\
-0.4239 & 0.006659 & All sky cond. & Autumn \\
-0.8715 & 0.000269 & Clear sky cond. & Winter \\
-0.05881 & 0.6247 & Clear sky cond. & Spring \\
--0.02369 & 0.7972 & Clear sky cond. & Summer \\
-0.3583 & 0.02996 & Clear sky cond. & Autumn \\
--0.02581 & 0.9335 & Cloudy cond. & Winter \\
-0.4333 & 0.02937 & Cloudy cond. & Spring \\
-0.442 & 0.004988 & Cloudy cond. & Summer \\
-0.4776 & 0.01955 & Cloudy cond. & Autumn \\
+\cellcolor{gray!6}{} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{0.6860} & \cellcolor{gray!6}{99.9}\\
+
+ & Spring & 0.1450 & 81.8\\
+
+\cellcolor{gray!6}{} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{0.1200} & \cellcolor{gray!6}{89.4}\\
+
+\multirow{-4}{8em}{\centering\arraybackslash All sky} & Autumn & 0.4310 & 99.3\\
+\cmidrule{1-4}
+\cellcolor{gray!6}{} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{0.8260} & \cellcolor{gray!6}{100.0}\\
+
+ & Spring & 0.0613 & 38.6\\
+
+\cellcolor{gray!6}{} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{-0.0307} & \cellcolor{gray!6}{25.9}\\
+
+\multirow{-4}{8em}{\centering\arraybackslash Clear sky} & Autumn & 0.3670 & 97.2\\
+\cmidrule{1-4}
+\cellcolor{gray!6}{} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{-0.8820} & \cellcolor{gray!6}{98.9}\\
+
+ & Spring & -0.0991 & 37.2\\
+
+\cellcolor{gray!6}{} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{0.0444} & \cellcolor{gray!6}{21.5}\\
+
+\multirow{-4}{8em}{\centering\arraybackslash Cloudy sky} & Autumn & -0.4000 & 89.5\\
 \bottomrule
-\end{longtable}
-
-\begin{longtable}[]{@{}
-  >{\centering\arraybackslash}p{(\columnwidth - 0\tabcolsep) * \real{0.4444}}@{}}
-\toprule
-\begin{minipage}[b]{\linewidth}\centering
-Trend statistical signif. {[}\%{]}
-\end{minipage} \\
-\midrule
-\endhead
-99.88 \\
-85.02 \\
-91.29 \\
-99.33 \\
-99.97 \\
-37.53 \\
-20.28 \\
-97 \\
-6.646 \\
-97.06 \\
-99.5 \\
-98.04 \\
-\bottomrule
-\end{longtable}
-
-\hypertarget{consistency-of-the-trends}{%
-\subsection{Consistency of the trends}\label{consistency-of-the-trends}}
-
-A method to evaluate changes in the long term trend is to use the cumulative sum of the variable.
-
-Figure \ref{fig:cumsum}
-
-\ldots{} different in aggregation \ldots{}
-
-\begin{figure}[h!]
-
-{\centering \subfloat[one plot(\#fig:cumsum-1)]{\includegraphics[width=.35\linewidth]{./images/CumulativeDailyCumSum-1} }\subfloat[the other one(\#fig:cumsum-2)]{\includegraphics[width=.35\linewidth]{./images/CumulativeDailyCumSum-9} }
-
-}
-
-\caption{two plots}(\#fig:cumsum)
-\end{figure}
+\end{tabu}
+\end{table}
 
@@
 
-Our result for all-sky condition (\(0.4\%/year\)), reaffirm the previous results of Bais et al. (2013) for the site.
-The increase of this trend, shows that the phenomenon and probably the causes have been amplified.
+We demonstrate that in the period
+1993 - 2023, there is a brightening trend of SDR for all-sky conditions (\(0.38\,\%/y\)).
+On a previous study (Bais et al. 2013), for the period 1993 - 2011 the trend was found to be \(0.33\,\%/y\).
+The increase of this trend, indicates that the phenomenon and probably the causes, are still occurring.
+Accordingly, we found a similar trend for clear-sky conditions (\(0.35\,\%/y\)) that further supports that the brightening can be independent of the cloud occurrences, although, cloud formation can also attenuate SDR.
+We have calculated the effect of TSI variability to be a negligible factor in SDR variability.
+Unfortunately, there is no available data for this period for the aerosols, in order to directly estimate their effect on SDR, which is the most probable main cause.
+The attenuation of SDR by aerosols over Europe have been proposed as major factor by Wild et al. (2021).
+The dimming effect of SDR under cloud-sky condition (\(-0.28\,\%/y\)), shows that clouds can also attenuate the amount of SDR reaching the ground, and changes on their occurrences can effect the net energy balance.
+Because we have no adequate data to investigate the long term changes of cloud formation in the region, we can not infer if the SDR trend we observe under cloud-sky is biased by changes in clouds and local climate.
+The observed brightening of SDR over Thessaloniki; is dependent on SZA (higher SZA, have stronger brightening trends), and also, with the seasons of the year, where Winter and Autumn show a significant strong trend, in contrast to Spring and Summer.
+Using CUSUMs of the trends for all- and clear-skies, we observed periods where the CUSUMs remain relative steady,
+with a steep decline previous to that, and a steep incline after. This is an indication that the whole brightening phenomenon does not follow a smooth development over time.
 
-\ldots..
-Also, the trend of \(0.36\%/year\) for cloud-sky condition indicate the major part that clouds play on the
-\ldots..
+Our findings are in agreement with other researchers for the region.
+Future observations from the collocated pyrheliometer, may provide us with the means to further investigate the variability of Solar radiation at ground level in Thessaloniki.
+Also, more data (cloud occurrence, aerosols, atmospheric water vapour, etc.), will help us to determine the extent of SDR attenuation by each factor.
 
-\ldots\ldots.
-
-About:
-
-\begin{itemize}
-\item
-  long terms trends
-\item
-  seasonal trends
-\item
-  effect of SZA
-\item
-  effect of clouds
-\item
-  Aerosols
-\end{itemize}
-
-similar results with Evidence for Clear‐Sky Dimming and Brightening in Central Europe\_Wild2021.pdf
-inclue in discusion
-
-\begin{longtable}[]{@{}c@{}}
-\toprule
-\endhead
-\textbf{END} \\
-\bottomrule
-\end{longtable}
+\hypertarget{references}{%
+\section*{References}\label{references}}
+\addcontentsline{toc}{section}{References}
 
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-Brent1973}{}}%
+Brent, Richard P. 1973. {``Algorithms for Minimization Without Derivatives.''} \emph{PrenticeHall, Englewood Cliffs, NJ}.
+
 \leavevmode\vadjust pre{\hypertarget{ref-Coddington2005}{}}%
@@
 \url{https://doi.org/10.7289/V55B00C1}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-Haurwitz1945}{}}%
+Haurwitz, Bernhard. 1945. {``Insolation in {Relation} to {Cloudiness} and {Cloud} {Density}.''} \emph{Journal of Meteorology} 2 (September): 154--66.
 
@@
 \leavevmode\vadjust pre{\hypertarget{ref-Long2006}{}}%
-Long, C., and Y.
+Long, Charles N., and Y.
 Shi.
@@
 
+\leavevmode\vadjust pre{\hypertarget{ref-Ohmura2009}{}}%
+Ohmura, Atsumu. 2009. {``Observed Decadal Variations in Surface Solar Radiation and Their Causes.''} \emph{Jour} 114. \url{https://doi.org/10.1029/2008JD011290}.
+
 \leavevmode\vadjust pre{\hypertarget{ref-Ohvril2009}{}}%
@@
 \url{https://doi.org/10.1029/2008JD010644}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-RCT2023}{}}%
+R Core Team. 2023. \emph{R: A Language and Environment for Statistical Computing}. Vienna, Austria: R Foundation for Statistical Computing. \url{https://www.R-project.org/}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-Regier2019}{}}%
+Regier, Peter, Henry Briceño, and Joseph N.
+Boyer.
+2019.
+{``Analyzing and Comparing Complex Environmental Time Series Using a Cumulative Sums Approach.''} \emph{{MethodsX}} 6: 779--87.
+\url{https://doi.org/10.1016/j.mex.2019.03.014}.
 
@@
 2012a.
-{``{Global Horizontal Irradiance Clear Sky Models: Implementation and Analysis}.''} \emph{SANDIA REPORT Sand2012-2389 Unlimited Release Printed March 2012}, March, 1--66.
+{``{Global Horizontal Irradiance Clear Sky Models: Implementation and Analysis}.''} \emph{SANDIA REPORT SAND2012-2389 Unlimited Release Printed March 2012}, March, 1--66.
 
@@
 \url{https://doi.org/10.1038/s41561-019-0528-y}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-Wang2021}{}}%
+Wang, Yawen, Jiahua Zhang, Arturo Sanchez‐Lorenzo, Katsumasa Tanaka, Jörg Trentmann, Wenping Yuan, and Martin Wild.
+2021.
+{``Hourly Surface Observations Suggest Stronger Solar Dimming and Brightening at Sunrise and Sunset over China.''} \emph{Geophysical Research Letters} 48 (2).
+\url{https://doi.org/10.1029/2020GL091422}.
 
@@
 
+\hypertarget{appendix}{%
+\section*{Appendix}\label{appendix}}
+\addcontentsline{toc}{section}{Appendix}
+
+\begin{figure}[h!]
+
+{\centering \includegraphics[width=0.7\linewidth]{./images/LongtermTrends-5}
+
+}
+
+\caption{Anomaly (\%) of the daily clear-sky SDR, relative to climatological values for1993 - 2023. The black line shows the long term linear trend for clear-sky conditions.}(\#fig:trendCLEAR)
+\end{figure}
+
+\begin{figure}[h!]
+
+{\centering \includegraphics[width=0.7\linewidth]{./images/LongtermTrends-8}
+
+}
+
+\caption{Anomaly (\%) of the daily cloud-sky SDR, relative to climatological values for 1993 - 2023. The black line shows the long term linear trend for cloud-sky conditions.}(\#fig:trendCLOUD)
+\end{figure}
+
+\begin{figure}[h!]
+
+{\centering \includegraphics[width=1\linewidth]{./images/SzaTrendsSeasTogether-2}
+
+}
+
+\caption{Seasonal distribution of the SDR's long term trends by SZA before and after noon.
+The solid data points represent values with an acceptable statistical significance ($p<0.005$), the trends were computed with SZA bins of $1^\circ$ for each day, before and after local noon, all graphs have the same scale.
+Bins with $p<0.005$ or with less than 85 data points may be missing from the view.}(\#fig:SZAtrendSeason)
+\end{figure}
+
 \end{document}

### 2023-11-02 MDPI_submition.Rmd
Commit 2023-11-02 14:14

@@
 ---
-title: Full title of the paper (Capitalized)
+title:     "Trends of SDR in Thessaloniki"
 author:
@@
 #   Tel.: +XX-000-00-0000.
+
 # document options
@@
 status:  submit
+
 # front matter
@@
   A Simple summary goes here.
+
 abstract: |
@@
   We indentified that there are some anomalies in the long term SDR trends, for all sky conditions.
+
 # back matter
 keywords: |
-  GHI; SDR; Solar Brigthening/Dimming (list three to ten pertinent keywords specific 
-  to the article, yet reasonably common within the subject discipline.).
+  GHI; SDR; Solar Brigthening/Dimming.
+## list three to ten pertinent keywords specific to the article, yet reasonably common within the subject discipline
+
 acknowledgement: |
@@
   access.
+
 authorcontributions: |
@@
   limited to those who have contributed substantially to the work reported.
+
 funding: |
@@
   funding.
-institutionalreview: |
-  In this section, you should add the Institutional Review Board Statement and 
-  approval number, if relevant to your study. You might choose to exclude 
-  this statement if the study did not require ethical approval. Please note 
-  that the Editorial Office might ask you for further information. Please add 
-  “The study was conducted in accordance with the Declaration of Helsinki, 
-  and approved by the Institutional Review Board (or Ethics Committee) of 
-  NAME OF INSTITUTE (protocol code XXX and date of approval).” for studies 
-  involving humans. OR “The animal study protocol was approved by the 
-  Institutional Review Board (or Ethics Committee) of NAME OF INSTITUTE 
-  (protocol code XXX and date of approval).” for studies involving animals. 
-  OR “Ethical review and approval were waived for this study due to REASON 
-  (please provide a detailed justification).” OR “Not applicable” for
-   studies not involving humans or animals.
+
+# institutionalreview: |
+#   In this section, you should add the Institutional Review Board Statement and 
+#   approval number, if relevant to your study. You might choose to exclude 
+#   this statement if the study did not require ethical approval. Please note 
+#   that the Editorial Office might ask you for further information. Please add 
+#   “The study was conducted in accordance with the Declaration of Helsinki, 
+#   and approved by the Institutional Review Board (or Ethics Committee) of 
+#   NAME OF INSTITUTE (protocol code XXX and date of approval).” for studies 
+#   involving humans. OR “The animal study protocol was approved by the 
+#   Institutional Review Board (or Ethics Committee) of NAME OF INSTITUTE 
+#   (protocol code XXX and date of approval).” for studies involving animals. 
+#   OR “Ethical review and approval were waived for this study due to REASON 
+#   (please provide a detailed justification).” OR “Not applicable” for
+#    studies not involving humans or animals.
+
 # informedconsent: |
@@
 #   publish this paper'' if applicable.
+
 dataavailability: |
@@
   \url{https://www.mdpi.com/ethics}.
-conflictsofinterest: |
-  Declare conflicts of interest or state 'The authors declare no conflict of 
-  interest.' Authors must identify and declare any personal circumstances or
-  interest that may be perceived as inappropriately influencing the
-  representation or interpretation of reported research results. Any role of the
-  funding sponsors in the design of the study; in the collection, analyses or 
-  interpretation of data in the writing of the manuscript, or in the decision to 
-  publish the results must be declared in this section. If there is no role, 
-  please state 'The founding sponsors had no role in the design of the study; 
-  in the collection, analyses, or interpretation of data; in the writing of the 
-  manuscript, an in the decision to publish the results'.
+
+# conflictsofinterest: |
+#   Declare conflicts of interest or state 'The authors declare no conflict of 
+#   interest.' Authors must identify and declare any personal circumstances or
+#   interest that may be perceived as inappropriately influencing the
+#   representation or interpretation of reported research results. Any role of the
+#   funding sponsors in the design of the study; in the collection, analyses or 
+#   interpretation of data in the writing of the manuscript, or in the decision to 
+#   publish the results must be declared in this section. If there is no role, 
+#   please state 'The founding sponsors had no role in the design of the study; 
+#   in the collection, analyses, or interpretation of data; in the writing of the 
+#   manuscript, an in the decision to publish the results'.
+
 # sampleavailability: |
 #   Samples of the compounds ...... are available from the authors.
+
 supplementary: |
@@
  \linksupplementary{s1}, Figure S1: title; Table S1: title; Video S1: title.
+
 abbreviations:
@@
    - \captionsetup[sub]{position=bottom, labelfont={bf, small, stretch=1.17}, labelsep=space, textfont={small, stretch=1.5}, aboveskip=6pt,  belowskip=6pt, singlelinecheck=off, justification=justified}
+
 bibliography: manualreferences.bib
+
 appendix:     appendix.tex
+
 endnotes:     false
+
 output: 
@@
     keep_tex:            yes
-
+  bookdown::word_document2: default
+  bookdown::pdf_document2:
+    number_sections:  no
+    fig_caption:      yes
+    keep_tex:         no
+    latex_engine:     xelatex
+    toc:              no
+    toc_depth:        4
+    fig_width:        7
+    fig_height:       4.5
+    documentclass:    article
+    classoption:      a4paper,oneside
+    fontsize:         12pt
+    geometry:         "left=0.75in,right=0.75in,top=0.75in,bottom=0.75in"
+    link-citations:   yes
+    colorlinks:       yes
+    urlcolor:         blue
+    biblio-style:     apalike
 ---
@@
     kable_styling(latex_options = c('striped', 'hold_position'), full_width = TRUE) %>%
-    # kable_styling(full_width = T) %>%
-    # kable_styling("striped")      %>%
     column_spec(1, width = "8em")

### 2023-11-02 MDPI_submition.tex
Commit 2023-11-02 14:14

@@
 % Full title of the paper (Capitalized)
-\Title{Full title of the paper (Capitalized)}
+\Title{Trends of SDR in Thessaloniki}
 
 % MDPI internal command: Title for citation in the left column
-\TitleCitation{Full title of the paper (Capitalized)}
+\TitleCitation{Trends of SDR in Thessaloniki}
 
@@
 % Keywords
-\keyword{GHI; SDR; Solar Brigthening/Dimming (list three to ten
-pertinent keywords specific to the article, yet reasonably common within
-the subject discipline.).}
+\keyword{GHI; SDR; Solar Brigthening/Dimming.}
 
@@
 
-\institutionalreview{In this section, you should add the Institutional
-Review Board Statement and approval number, if relevant to your study.
-You might choose to exclude this statement if the study did not require
-ethical approval. Please note that the Editorial Office might ask you
-for further information. Please add ``The study was conducted in
-accordance with the Declaration of Helsinki, and approved by the
-Institutional Review Board (or Ethics Committee) of NAME OF INSTITUTE
-(protocol code XXX and date of approval).'' for studies involving
-humans. OR ``The animal study protocol was approved by the Institutional
-Review Board (or Ethics Committee) of NAME OF INSTITUTE (protocol code
-XXX and date of approval).'' for studies involving animals. OR ``Ethical
-review and approval were waived for this study due to REASON (please
-provide a detailed justification).'' OR ``Not applicable'' for studies
-not involving humans or animals.}
-
 \dataavailability{We encourage all authors of articles published in MDPI
@@
 
-\conflictsofinterest{Declare conflicts of interest or state `The authors
-declare no conflict of interest.' Authors must identify and declare any
-personal circumstances or interest that may be perceived as
-inappropriately influencing the representation or interpretation of
-reported research results. Any role of the funding sponsors in the
-design of the study; in the collection, analyses or interpretation of
-data in the writing of the manuscript, or in the decision to publish the
-results must be declared in this section. If there is no role, please
-state `The founding sponsors had no role in the design of the study; in
-the collection, analyses, or interpretation of data; in the writing of
-the manuscript, an in the decision to publish the results'.}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 SDR & Solar downward radiation \\
+SZA & Solar zenith angle \\
 \end{tabular}}

### 2023-11-02 Article.Rmd
Commit 2023-11-02 15:08

@@
   The shortwave downward solar irradiance (SDR) is an important factor that drives climate processes, production and can affect all living organisms. 
-  While monitoring the long term variability, there are observations of upward and downward SDR trends on different locations around the world for different time periods.
-  Periods of positive tredns are refered as brightening periods and with negative tredns as dimming periods.
-  We studied 29 years of CM-21 data from Thessaloniki, Greece, under three sky conditions (clear sky, cloudy sky and all sky conditions), applying a cloud sky identification algorithm.
-  We found a positive trend for all-sky and clear-sky conditions, and also, investigated the consistency of those trends, the effect of the solar zenith angle, and the variation of the trends for the seasons of the year.
-  We indentified that there are some anomalies in the long term SDR trends, for all sky conditions.
+  Observations of SDR at different locations around the world with different environmental characteristics have been used to investigate its long-term variability and trends at different time scales. 
+  Periods of positive trends are referred as brightening periods and of negative trends as dimming periods. 
+  Here we studied 29 years of pyranometer data in Thessaloniki, Greece, under three types of sky conditions (clear sky, cloudy sky and all sky). 
+  The clear-sky data were identified by applying a cloud screening algorithm.
+We found a positive trend of $0.38\,\%/\text{year}$ for all-sky, $0.35\,\%/\text{year}$ for clear-sky conditions, and $-0.28\,\%/\text{year}$ for cloudy conditions.
+  We have also investigated the consistency of these trends, the effect of the solar zenith angle, and their seasonal variability. 
+  Finally, we identified some anomalies in the long term SDR trends for all sky conditions.
 
@@
 
-The shortwave downward solar irradiance (SDR) at Earth's surface play a significant role, on its climate.
-Changes of the SDR can be related to changes on Earth's energy budget, the mechanisms of climate change, and water and carbon cycle [@Wild2009].
-Can also affect, solar and agricultural production, and all living organisms.
+The shortwave downward solar irradiance (SDR) at Earth's surface plays a significant role, on its climate.
+Changes of the SDR can be related to changes on Earth's energy budget, the mechanisms of climate change, and water and carbon cycles [@Wild2009].
+It can also affect solar and agricultural production, and all living organisms.
 Studies of SDR variability, have identified some distinct SDR trends on different regions of the world on different time periods.
 The term 'brightening' is generally used to describe periods of positive SDR trend, and 'dimming' for negative trend.
-There are many cases on the long term records of irradiance, showing a systematic change of SDR's trend slope, occurring roughly at the last decades of the 20th century.
+There are many cases in the long term records of irradiance, showing a systematic change of SDR's trend slope, occurring roughly at the last decades of the 20th century.
 <!-- Trends china -->
@@
 <!-- }. -->
-Although, we have performed a similar analysis to the DNI data the results are not presented here, as they lack the appropriate statistical significance, due to the sorter duration of the data. 
+Although, we have performed a similar analysis to the DNI data, the results are not presented here, as they lack the appropriate statistical significance, due to the sorter duration of the data. 
 However, the DNI data were used as auxiliary data, in the clear sky identification algorithm (CSid), which is discussed later, for the selection of the appropriate thresholds.
@@
 Thus, problematic recordings have been excluded from further processing. 
-Although it is impossible to detect all false data, the large number of available data, and the aggregation scheme we used, ensures the good quality of the radiometric measurements used in this study.
+Although it is impossible to detect all false data, the large number of available data, and the aggregation scheme we used, ensures the good quality of the radiation measurements used in this study.
 
 <!-- c) Clear sky identification -->
-In order to be able to estimate the effect of the sky condition on the long term variability of SDR, we created three datasets, by characterizing each one-minute measurement with a corresponding sky condition (i.e., all-sky, clear-sky and cloudy-sky).
+In order to be able to estimate the effect of the sky conditions on the long term variability of SDR, we created three datasets, by characterizing each one-minute measurement with a corresponding sky condition (i.e., all-sky, clear-sky and cloudy-sky).
 To identify the clear-sky conditions we used a method proposed by
@@
 @Reno2016,
-which was adapted and configured for the site, as the authors suggest.
+which were adapted and configured for the site, as the authors suggest.
 

### 2023-11-02 Article_A.tex
Commit 2023-11-02 15:08

@@
 The shortwave downward solar irradiance (SDR) is an important factor that drives climate processes, production and can affect all living organisms.
-While monitoring the long term variability, there are observations of upward and downward SDR trends on different locations around the world for different time periods.
-Periods of positive tredns are refered as brightening periods and with negative tredns as dimming periods.
-We studied 29 years of CM-21 data from Thessaloniki, Greece, under three sky conditions (clear sky, cloudy sky and all sky conditions), applying a cloud sky identification algorithm.
-We found a positive trend for all-sky and clear-sky conditions, and also, investigated the consistency of those trends, the effect of the solar zenith angle, and the variation of the trends for the seasons of the year.
-We indentified that there are some anomalies in the long term SDR trends, for all sky conditions.
+Observations of SDR at different locations around the world with different environmental characteristics have been used to investigate its long-term variability and trends at different time scales.
+Periods of positive trends are referred as brightening periods and of negative trends as dimming periods.
+Here we studied 29 years of pyranometer data in Thessaloniki, Greece, under three types of sky conditions (clear sky, cloudy sky and all sky).
+The clear-sky data were identified by applying a cloud screening algorithm.
+We found a positive trend of \(0.38\,\%/\text{year}\) for all-sky, \(0.35\,\%/\text{year}\) for clear-sky conditions, and \(-0.28\,\%/\text{year}\) for cloudy conditions.
+We have also investigated the consistency of these trends, the effect of the solar zenith angle, and their seasonal variability.
+Finally, we identified some anomalies in the long term SDR trends for all sky conditions.
 \end{abstract}
@@
 
-The shortwave downward solar irradiance (SDR) at Earth's surface play a significant role, on its climate.
-Changes of the SDR can be related to changes on Earth's energy budget, the mechanisms of climate change, and water and carbon cycle (Wild 2009).
-Can also affect, solar and agricultural production, and all living organisms.
+The shortwave downward solar irradiance (SDR) at Earth's surface plays a significant role, on its climate.
+Changes of the SDR can be related to changes on Earth's energy budget, the mechanisms of climate change, and water and carbon cycles (Wild 2009).
+It can also affect solar and agricultural production, and all living organisms.
 Studies of SDR variability, have identified some distinct SDR trends on different regions of the world on different time periods.
 The term `brightening' is generally used to describe periods of positive SDR trend, and `dimming' for negative trend.
-There are many cases on the long term records of irradiance, showing a systematic change of SDR's trend slope, occurring roughly at the last decades of the 20th century.
+There are many cases in the long term records of irradiance, showing a systematic change of SDR's trend slope, occurring roughly at the last decades of the 20th century.
 On multiple station in China, a dimming period was reported until about 2000, followed by a brightening period (Yang et al. 2021).
@@
 2023-04-13.
-Although, we have performed a similar analysis to the DNI data the results are not presented here, as they lack the appropriate statistical significance, due to the sorter duration of the data.
+Although, we have performed a similar analysis to the DNI data, the results are not presented here, as they lack the appropriate statistical significance, due to the sorter duration of the data.
 However, the DNI data were used as auxiliary data, in the clear sky identification algorithm (CSid), which is discussed later, for the selection of the appropriate thresholds.
@@
 Thus, problematic recordings have been excluded from further processing.
-Although it is impossible to detect all false data, the large number of available data, and the aggregation scheme we used, ensures the good quality of the radiometric measurements used in this study.
+Although it is impossible to detect all false data, the large number of available data, and the aggregation scheme we used, ensures the good quality of the radiation measurements used in this study.
 
-In order to be able to estimate the effect of the sky condition on the long term variability of SDR, we created three datasets, by characterizing each one-minute measurement with a corresponding sky condition (i.e., all-sky, clear-sky and cloudy-sky).
+In order to be able to estimate the effect of the sky conditions on the long term variability of SDR, we created three datasets, by characterizing each one-minute measurement with a corresponding sky condition (i.e., all-sky, clear-sky and cloudy-sky).
 To identify the clear-sky conditions we used a method proposed by
@@
 Reno and Hansen (2016),
-which was adapted and configured for the site, as the authors suggest.
+which were adapted and configured for the site, as the authors suggest.
 

### 2023-11-02 MDPI_submition.Rmd
Commit 2023-11-02 15:08

@@
 
+suppressMessages(
 x_latex %>%
@@
     column_spec(1, width = "8em")
+)
 

### 2023-11-02 MDPI_submition.tex
Commit 2023-11-02 15:08

@@
 factor that drives climate processes, production and can affect all
-living organisms. While monitoring the long term variability, there are
-observations of upward and downward SDR trends on different locations
-around the world for different time periods. Periods of positive tredns
-are refered as brightening periods and with negative tredns as dimming
-periods. We studied 29 years of CM-21 data from Thessaloniki, Greece,
-under three sky conditions (clear sky, cloudy sky and all sky
-conditions), applying a cloud sky identification algorithm. We found a
-positive trend for all-sky and clear-sky conditions, and also,
-investigated the consistency of those trends, the effect of the solar
-zenith angle, and the variation of the trends for the seasons of the
-year. We indentified that there are some anomalies in the long term SDR
-trends, for all sky conditions.}
+living organisms. Observations of SDR at different locations around the
+world with different environmental characteristics have been used to
+investigate its long-term variability and trends at different time
+scales. Periods of positive trends are referred as brightening periods
+and of negative trends as dimming periods. Here we studied 29 years of
+pyranometer data in Thessaloniki, Greece, under three types of sky
+conditions (clear sky, cloudy sky and all sky). The clear-sky data were
+identified by applying a cloud screening algorithm. We found a positive
+trend of \(0.38\,\%/\text{year}\) for all-sky, \(0.35\,\%/\text{year}\)
+for clear-sky conditions, and \(-0.28\,\%/\text{year}\) for cloudy
+conditions. We have also investigated the consistency of these trends,
+the effect of the solar zenith angle, and their seasonal variability.
+Finally, we identified some anomalies in the long term SDR trends for
+all sky conditions.}
 
@@
 
-The shortwave downward solar irradiance (SDR) at Earth's surface play a
+The shortwave downward solar irradiance (SDR) at Earth's surface plays a
 significant role, on its climate. Changes of the SDR can be related to
 changes on Earth's energy budget, the mechanisms of climate change, and
-water and carbon cycle \citep{Wild2009}. Can also affect, solar and
+water and carbon cycles \citep{Wild2009}. It can also affect solar and
 agricultural production, and all living organisms. Studies of SDR
@@
 is generally used to describe periods of positive SDR trend, and
-`dimming' for negative trend. There are many cases on the long term
+`dimming' for negative trend. There are many cases in the long term
 records of irradiance, showing a systematic change of SDR's trend slope,
@@
 CHP-1 pyrheliometer, for the period 2016-04-01 to 2023-04-13. Although,
-we have performed a similar analysis to the DNI data the results are not
-presented here, as they lack the appropriate statistical significance,
-due to the sorter duration of the data. However, the DNI data were used
-as auxiliary data, in the clear sky identification algorithm (CSid),
-which is discussed later, for the selection of the appropriate
-thresholds. It is noted that despite the capability of the CSid
-algorithm to use the DNI as a characterization parameter, we haven't
-utilized it here, to avoid any selection bias, due to unequal length of
-the two datasets. There are four distinct steps in the creation of the
-dataset analysed here: a)~the acquisition of radiation measurements from
-the sensors, b)~the data quality check, c)~the identification of ``clear
-sky'' conditions from the radiometric data, and d)~the aggregation of
-data and trend analysis.
+we have performed a similar analysis to the DNI data, the results are
+not presented here, as they lack the appropriate statistical
+significance, due to the sorter duration of the data. However, the DNI
+data were used as auxiliary data, in the clear sky identification
+algorithm (CSid), which is discussed later, for the selection of the
+appropriate thresholds. It is noted that despite the capability of the
+CSid algorithm to use the DNI as a characterization parameter, we
+haven't utilized it here, to avoid any selection bias, due to unequal
+length of the two datasets. There are four distinct steps in the
+creation of the dataset analysed here: a)~the acquisition of radiation
+measurements from the sensors, b)~the data quality check, c)~the
+identification of ``clear sky'' conditions from the radiometric data,
+and d)~the aggregation of data and trend analysis.
 
@@
 available data, and the aggregation scheme we used, ensures the good
-quality of the radiometric measurements used in this study.
+quality of the radiation measurements used in this study.
 
-In order to be able to estimate the effect of the sky condition on the
+In order to be able to estimate the effect of the sky conditions on the
 long term variability of SDR, we created three datasets, by
@@
 clear-sky conditions we used a method proposed by \citet{Long2000} and
-by \citet{Reno2016}, which was adapted and configured for the site, as
+by \citet{Reno2016}, which were adapted and configured for the site, as
 the authors suggest.
@@
 
-\begin{verbatim}
-Usually it is recommended to use column_spec before collapse_rows, especially in LaTeX, to get a desired result. 
-\end{verbatim}
-
 \begin{table}[!h]

### 2023-11-02 MDPI_submition.Rmd
Commit 2023-11-02 21:10

@@
 The term 'brightening' is generally used to describe periods of positive SDR trend, and 'dimming' for negative trend.
-There are many cases in the long term records of irradiance, showing a systematic change of SDR's trend slope, occurring roughly at the last decades of the 20th century.
+There are many cases in the long term records of irradiance, showing a systematic change in the magnitude of the trend, occurring roughly in the last decades of the 20th century.
 <!-- Trends china -->
-On multiple station in China, a dimming period was reported until about 2000, followed by a brightening period [@Yang2021]. 
+On multiple stations in China, a dimming period was reported until about 2000, followed by a brightening period [@Yang2021]. 
 <!-- Trends Europe Brazil -->
@@
 <!-- Trends Global -->
-Also, on global scale, an AI aided continental level spatial analysis, with data from multiple station, reach similar conclusions for the above regions and for the global trend [@Yuan2021].
+On global scale, an artificial Intelligence aided spatial analysis on continental level with data from multiple stations reach similar conclusions for these regions and for the global trend [@Yuan2021].
 
-There is a consensus, among researchers, that the major factors of SDR attenuation is the interaction of Sun radiation with atmospheric aerosols and clouds. 
+There is a consensus among researchers that the major factor affecting the variability of SDR attenuation is the interactions of solar radiation with atmospheric aerosols and clouds. 
 Those interactions, among other factors, have been analysed with models [@Li2016; @Samset2018], showing the existence of feedback mechanisms between the two.

### 2023-11-02 MDPI_submition.tex
Commit 2023-11-02 21:10

@@
 `dimming' for negative trend. There are many cases in the long term
-records of irradiance, showing a systematic change of SDR's trend slope,
-occurring roughly at the last decades of the 20th century. On multiple
-station in China, a dimming period was reported until about 2000,
-followed by a brightening period \citep{Yang2021}. A similar pattern was
-identified, with the breaking point around 1980, for stations in Central
-Europe \citep{Wild2021} and Brazil \citep{Yamasoe2021}. Also, on global
-scale, an AI aided continental level spatial analysis, with data from
-multiple station, reach similar conclusions for the above regions and
-for the global trend \citep{Yuan2021}.
+records of irradiance, showing a systematic change in the magnitude of
+the trend, occurring roughly in the last decades of the 20th century. On
+multiple stations in China, a dimming period was reported until about
+2000, followed by a brightening period \citep{Yang2021}. A similar
+pattern was identified, with the breaking point around 1980, for
+stations in Central Europe \citep{Wild2021} and Brazil
+\citep{Yamasoe2021}. On global scale, an artificial Intelligence aided
+spatial analysis on continental level with data from multiple stations
+reach similar conclusions for these regions and for the global trend
+\citep{Yuan2021}.
 
-There is a consensus, among researchers, that the major factors of SDR
-attenuation is the interaction of Sun radiation with atmospheric
-aerosols and clouds. Those interactions, among other factors, have been
-analysed with models \citep{Li2016, Samset2018}, showing the existence
-of feedback mechanisms between the two. Similar finds, have been showed
-in observational data \citep[ and references
+There is a consensus among researchers that the major factor affecting
+the variability of SDR attenuation is the interactions of solar
+radiation with atmospheric aerosols and clouds. Those interactions,
+among other factors, have been analysed with models
+\citep{Li2016, Samset2018}, showing the existence of feedback mechanisms
+between the two. Similar finds, have been showed in observational data
+\citep[ and references
 therein]{Schwarz2020, Ohvril2009, Zerefos2009, Xia2007}.

### 2023-11-03 MDPI_submition.Rmd
Commit 2023-11-03 10:14

@@
 Those interactions, among other factors, have been analysed with models [@Li2016; @Samset2018], showing the existence of feedback mechanisms between the two.
-Similar finds, have been showed in observational data [@Schwarz2020; @Ohvril2009; @Zerefos2009; @Xia2007 and references therein].
+Similar findings have been shown in observational data [@Schwarz2020; @Ohvril2009; @Zerefos2009; @Xia2007 and references therein].
 <!-- Sun variability: should cite that there is no significant trend found Wild2021 -->
 
-Due to the variability of the phenomenon, and its contributing factors, there is a constant need to investigate and monitor SDR, in different sites, to estimate its magnitude, and its relation to the local conditions.
-In this study, we examine the trends of SDR, with ground-based measurements at Thessaloniki, Greece for the period `r year(min(ALL_1_daily_DESEAS$Date))` to `r year(max(ALL_1_daily_DESEAS$Date))`, as derived from a CM-21 pyranometer.
+Due to the significant spatial and temporal variability of the trends, and the contributing factors, there is a constant need to monitor and investigate SDR in different sites in order to estimate the degree of variability,  and its relation to the local conditions.
+In this study, we examine the trends of SDR, using ground-based measurements at Thessaloniki, Greece, for the period `r year(min(ALL_1_daily_DESEAS$Date))` to `r year(max(ALL_1_daily_DESEAS$Date))`, as derived from a CM-21 pyranometer.
 <!-- goals -->
-We reevaluated and extended the dataset used by @Bais2013, applying a different algorithm for the identification of clear-/cloud-sky instances [@Reno2016; @Reno2012a], and we derive the radiation trends for the period `r year(min(ALL_1_daily_DESEAS$Date))` to `r year(max(ALL_1_daily_DESEAS$Date))`, under different sky conditions (all-sky, clear-sky and cloud-sky).
+We reevaluated and extended the dataset used by @Bais2013, we applied a different algorithm for the identification of clear-/cloud-sky instances [@Reno2016; @Reno2012a], and we derived the SDR trends for the period `r year(min(ALL_1_daily_DESEAS$Date))` to `r year(max(ALL_1_daily_DESEAS$Date))`, under different sky conditions (all-sky, clear-sky and cloud-sky).
+Finally, we investigated the dependence of the trends on solar zenith angle and season.
 
@@
 <!-- CS subjectivity -->
-We have to note, that the definition of what constitutes as clear or cloudy sky, has some subjectivity, in any method of characterization.
-As a result, the details of the definition are site specific, it relies on a combination of thresholds and comparisons with ideal actinometric models and statistical analysis on different signal metrics.
-The CSid algorithm was calibrated with the main focus, to identify the presence of clouds on the sky dome.
-Although the fine-tuning of the procedure, few marginal cases exist, that have been identified manually as false positive or false negative but cannot affect the final results of the study.
+We have to note, that the definition of clear or cloudy sky, has some subjectivity, in any method of characterization.
+As a result, the details of the definition are site specific, it relies on a combination of thresholds and comparisons with ideal radiation models and statistical analysis of different signal metrics.
+The CSid algorithm was calibrated with the main focus, to identify the presence of clouds on the sky.
+Despite the fine-tuning of the procedure, a few marginal cases exist, that have been identified manually as false positive or false negative but cannot affect the final results of the study.
 

### 2023-11-03 MDPI_submition.tex
Commit 2023-11-03 10:14

@@
 \citep{Li2016, Samset2018}, showing the existence of feedback mechanisms
-between the two. Similar finds, have been showed in observational data
+between the two. Similar findings have been shown in observational data
 \citep[ and references
@@
 
-Due to the variability of the phenomenon, and its contributing factors,
-there is a constant need to investigate and monitor SDR, in different
-sites, to estimate its magnitude, and its relation to the local
-conditions. In this study, we examine the trends of SDR, with
-ground-based measurements at Thessaloniki, Greece for the period 1993 to
-2023, as derived from a CM-21 pyranometer. We reevaluated and extended
-the dataset used by \citet{Bais2013}, applying a different algorithm for
-the identification of clear-/cloud-sky instances
-\citep{Reno2016, Reno2012a}, and we derive the radiation trends for the
+Due to the significant spatial and temporal variability of the trends,
+and the contributing factors, there is a constant need to monitor and
+investigate SDR in different sites in order to estimate the degree of
+variability, and its relation to the local conditions. In this study, we
+examine the trends of SDR, using ground-based measurements at
+Thessaloniki, Greece, for the period 1993 to 2023, as derived from a
+CM-21 pyranometer. We reevaluated and extended the dataset used by
+\citet{Bais2013}, we applied a different algorithm for the
+identification of clear-/cloud-sky instances
+\citep{Reno2016, Reno2012a}, and we derived the SDR trends for the
 period 1993 to 2023, under different sky conditions (all-sky, clear-sky
-and cloud-sky).
+and cloud-sky). Finally, we investigated the dependence of the trends on
+solar zenith angle and season.
 
@@
 
-We have to note, that the definition of what constitutes as clear or
-cloudy sky, has some subjectivity, in any method of characterization. As
-a result, the details of the definition are site specific, it relies on
-a combination of thresholds and comparisons with ideal actinometric
-models and statistical analysis on different signal metrics. The CSid
-algorithm was calibrated with the main focus, to identify the presence
-of clouds on the sky dome. Although the fine-tuning of the procedure,
-few marginal cases exist, that have been identified manually as false
-positive or false negative but cannot affect the final results of the
-study.
+We have to note, that the definition of clear or cloudy sky, has some
+subjectivity, in any method of characterization. As a result, the
+details of the definition are site specific, it relies on a combination
+of thresholds and comparisons with ideal radiation models and
+statistical analysis of different signal metrics. The CSid algorithm was
+calibrated with the main focus, to identify the presence of clouds on
+the sky. Despite the fine-tuning of the procedure, a few marginal cases
+exist, that have been identified manually as false positive or false
+negative but cannot affect the final results of the study.
 

### 2023-11-03 MDPI_submition.Rmd
Commit 2023-11-03 11:10

@@
 
-For completeness, we will provide below a brief overview of the 
-<!--
-[....Csid term is referenced above....]{.comment-start id="0" author="A N"
-date="2023-07-15T20:00:52Z"} 
--->
-clear sky identification algorithm (CSid),
-<!--
-[]{.comment-end id="0"} 
--->
-along with the site specific thresholds.
+For completeness, we will provide below a brief overview of the clear sky identification algorithm (CSid), along with the site specific thresholds.
 To calculate the reference clear sky
@@
 by @Long2000 and @Reno2016.
-The target of
-the optimization was the minimization of a function $f(a)$ (Eq.\ \ref{eq:minf}) and
+The target of the optimization was the minimization of a function $f(a)$ (Eq.\ \ref{eq:minf}) and
 was accomplished with the algorithmic function "optimise", which is an implementation based on the work of @Brent1973, from the library "stats" of the R programming language [@RCT2023]. 
@@
 \end{equation}
-where: $\text{SDR}_\text{CSref}$ is the reference clear sky SDR, in $\text{w}\,\text{m}^{-2}$ and $\theta$ is the solar zenith angle (SZA).
+where: $\text{SDR}_\text{CSref}$ is the reference clear sky SDR, in $\text{W}\,\text{m}^{-2}$ and $\theta$ is the solar zenith angle (SZA).
 
-The criteria that were used to identify whether a measurement was taken
-under clear-sky conditions are presented below.
-A data point is flagged
-as "clear-sky" if all criteria are satisfied, otherwise it is considered to be "cloud-sky".
-Each criterion was applied
-for a running window of $`r nt`$ consecutive one-minute measurements, and
-the characterization is assigned to the central value of the window.
-Each parameter, was calculated both from the observations and the
-reference clear sky model, for each comparison.
-The allowable range of variation is defined by the
-model-derived value of the parameter multiplied by a factor plus an
-offset. 
-<!--[.... .expanded the description ....]{.comment-start id="4"
-author="A N" date="2023-07-15T22:48:13Z"}
--->
-The factors and the offsets were
-determined empirically, by manual inspecting each filters performance on
-selected days, and adjusting them accordingly, during an iterative
-process.
-<!--[[]{.comment-end id="4"}]{.insertion author="A N"
-date="2023-07-15T22:48:13Z"}
--->
+The criteria that were used to identify whether a measurement was taken under clear-sky conditions are presented below.
+A data point is flagged as "clear-sky" if all criteria are satisfied, otherwise it is considered to be "cloud-sky".
+Each criterion was applied for a running window of $`r nt`$ consecutive one-minute measurements, and the characterization is assigned to the central value of the window.
+Each parameter, was calculated both from the observations and the reference clear sky model, for each comparison.
+The allowable range of variation is defined by the model-derived value of the parameter multiplied by a factor plus an offset. 
+The factors and the offsets were determined empirically, by manually inspecting each filter's performance on selected days and adjusting them accordingly during an iterative process.
+The criteria are:
 <!-- 1. MeanVIP -->
@@
 \begin{equation}
-`r 1 - CS_ref_rm_VIP_RelLow` \times \overline{\text{SDR}}_{\text{CSref},i} - `r CS_ref_rm_VIP_LowOff`
+`r 1 - CS_ref_rm_VIP_RelLow` \times \overline{\text{SDR}}_{\text{CSref},i} - `r CS_ref_rm_VIP_LowOff`\,Wm^{-2}
 < \overline{\text{SDR}}_i <
-`r 1 + CS_ref_rm_VIP_RelUpp` \times \overline{\text{SDR}}_{\text{CSref},i} + `r CS_ref_rm_VIP_UppOff`
+`r 1 + CS_ref_rm_VIP_RelUpp` \times \overline{\text{SDR}}_{\text{CSref},i} + `r CS_ref_rm_VIP_UppOff`\,Wm^{-2}
 \label{eq:MeanVIP}
@@
 -->
-b) Maximum measured value $M_{\text{}}$ (Eq.\ \ref{eq:MaxVIP}).
+b) Maximum measured value $M_{i}$ (Eq.\ \ref{eq:MaxVIP}).
 \begin{equation}
-`r MaxVIP_fct` \times M_{\text{CSref},i} - `r MaxVIP_off_low`
-< M_{\text{}i} <
-`r MaxVIP_fct` \times M_{\text{CSref},i} + `r MaxVIP_off_upp`
+`r MaxVIP_fct` \times M_{\text{CSref},i} - `r MaxVIP_off_low`\,Wm^{-2}
+< M_{i} <
+`r MaxVIP_fct` \times M_{\text{CSref},i} + `r MaxVIP_off_upp`\,Wm^{-2}
 \label{eq:MaxVIP}
@@
 \end{equation}
-where: $t_i$ is the time each SDR measurement has been measured
+where: $t_i$ is the time stamp of each SDR measurement.
 <!-- 4. VCT -->
@@
 <!-- ### Data selection -->
-Due to a significant measurement uncertainty near the horizon, we have to exclude all measurements with SZA greater than $`r 90-MIN_ELEVA`^\circ$. 
-Moreover, due to some obstructions around the site (hills and buildings), we excluded data with Azimuth angle between 
+Due to the significant measurement uncertainty when the Sun is near the horizon, we have excluded all measurements with SZA greater than $`r 90-MIN_ELEVA`^\circ$. 
+Moreover, due to some obstructions around the site (hills and buildings), we excluded data with azimuth angle between 
 $35^\circ$ and $120^\circ$ with SZA greater than $80^\circ$.
-On the latter instances, Sun is systematically, not visible by the instrument's location.
+On the latter instances, Sun is systematically not visible from the instrument's location.
 <!-- 1au -->
@@
 <!-- TSI -->
-Subsequently, we made all measurements relative to the Total Solar Irradiance (TSI) at $1\,\text{au}$, in order to compensate for the Sun's intensity variability, using a <!--homogenized--> time series of satellite TSI observations.
-The TSI data we use are part of the "NOAA Climate Data Record of Total Solar Irradiance" dataset [@Coddington2005]. 
-Where the initial daily values, were interpolated to match with the time step of our measurements. 
-The final dataset contains
+Subsequently, we made all measurements relative to the Total Solar Irradiance (TSI) at $1\,\text{au}$, in order to compensate for the Sun's intensity variability, using a time series of satellite TSI observations.
+The TSI data we used are part of the "NOAA Climate Data Record of Total Solar Irradiance" dataset [@Coddington2005]. 
+The initial daily values of this dataset were interpolated to match the time step of our measurements. 
+In the final dataset
 $`r SDR_N_All`$
@@
 $`r signif(100 * SDR_N_Clear / SDR_N_All, 3)`\%$
-were identified as under clear-sky conditions and subsequently
+were identified as under clear-sky conditions and
 $`r signif(100 * SDR_N_Cloud / SDR_N_All, 3)`\%$
@@
 <!-- aggregation -->
-In order to investigate the SDR trends, we implemented an appropriate aggregation scheme to the 1-minute data to derive a series in coarser timescale.
+In order to investigate the SDR trends, we implemented an appropriate aggregation scheme to the 1-minute data to derive a series in coarser time-scales.
 To preserve the representativeness of the data we used the following criteria: 
 <!-- a) for daily mean values we exclude days with less than `r Daily_aggregation_N_lim` valid data points, -->
-a) for the daily mean values we accept days with more than `r 100 * All_daily_ratio_lim`% of the daytime measurements, present and valid,
+a) for the daily mean values we accept days with more than `r 100 * All_daily_ratio_lim`% of the daytime measurements present and valid,
 b) monthly values were computed from daily means
@@
 Finally, each data set was deseasonalized by subtracting the corresponding climatological annual cycle (daily or monthly) from the actual data.
-To estimate SZA contribution to the SDR trends, the one-minute data were aggregated in $1^\circ$ SZA bins, separately for the morning and afternoon hours, and then were deseasonalized as mentioned above.
+To estimate the SZA effect on the SDR trends, the one-minute data were aggregated in $1^\circ$ SZA bins, separately for the morning and afternoon hours, and then were deseasonalized as mentioned above.
 

### 2023-11-03 MDPI_submition.tex
Commit 2023-11-03 11:10

@@
 \end{equation} where: \(\text{SDR}_\text{CSref}\) is the reference clear
-sky SDR, in \(\text{w}\,\text{m}^{-2}\) and \(\theta\) is the solar
+sky SDR, in \(\text{W}\,\text{m}^{-2}\) and \(\theta\) is the solar
 zenith angle (SZA).
@@
 factor plus an offset. The factors and the offsets were determined
-empirically, by manual inspecting each filters performance on selected
-days, and adjusting them accordingly, during an iterative process.
+empirically, by manually inspecting each filter's performance on
+selected days and adjusting them accordingly during an iterative
+process. The criteria are:
 
@@
   \ref{eq:MeanVIP}). \begin{equation}
-  0.91 \times \overline{\text{SDR}}_{\text{CSref},i} - 20
+  0.91 \times \overline{\text{SDR}}_{\text{CSref},i} - 20\,Wm^{-2}
   < \overline{\text{SDR}}_i <
-  1.095 \times \overline{\text{SDR}}_{\text{CSref},i} + 30
+  1.095 \times \overline{\text{SDR}}_{\text{CSref},i} + 30\,Wm^{-2}
   \label{eq:MeanVIP}
@@
 \item
-  Maximum measured value \(M_{\text{}}\) (Eq.~\ref{eq:MaxVIP}).
+  Maximum measured value \(M_{i}\) (Eq.~\ref{eq:MaxVIP}).
   \begin{equation}
-  1 \times M_{\text{CSref},i} - 75
-  < M_{\text{}i} <
-  1 \times M_{\text{CSref},i} + 75
+  1 \times M_{\text{CSref},i} - 75\,Wm^{-2}
+  < M_{i} <
+  1 \times M_{\text{CSref},i} + 75\,Wm^{-2}
   \label{eq:MaxVIP}
@@
   \label{eq:VILcr}
-  \end{equation} where: \(t_i\) is the time each SDR measurement has
-  been measured
+  \end{equation} where: \(t_i\) is the time stamp of each SDR
+  measurement.
 \item
@@
 
-Due to a significant measurement uncertainty near the horizon, we have
-to exclude all measurements with SZA greater than \(85^\circ\).
-Moreover, due to some obstructions around the site (hills and
-buildings), we excluded data with Azimuth angle between \(35^\circ\) and
-\(120^\circ\) with SZA greater than \(80^\circ\). On the latter
-instances, Sun is systematically, not visible by the instrument's
+Due to the significant measurement uncertainty when the Sun is near the
+horizon, we have excluded all measurements with SZA greater than
+\(85^\circ\). Moreover, due to some obstructions around the site (hills
+and buildings), we excluded data with azimuth angle between \(35^\circ\)
+and \(120^\circ\) with SZA greater than \(80^\circ\). On the latter
+instances, Sun is systematically not visible from the instrument's
 location. To make the measurements comparable throughout the dataset, we
@@
 the Sun's intensity variability, using a time series of satellite TSI
-observations. The TSI data we use are part of the ``NOAA Climate Data
-Record of Total Solar Irradiance'' dataset \citep{Coddington2005}. Where
-the initial daily values, were interpolated to match with the time step
-of our measurements. The final dataset contains \(6589967\) one-minute
+observations. The TSI data we used are part of the ``NOAA Climate Data
+Record of Total Solar Irradiance'' dataset \citep{Coddington2005}. The
+initial daily values of this dataset were interpolated to match the time
+step of our measurements. In the final dataset \(6589967\) one-minute
 measurements, of which, \(84.2\%\) were identified as under clear-sky
-conditions and subsequently \(15.8\%\) as under cloud-sky conditions.
+conditions and \(15.8\%\) as under cloud-sky conditions.
 
@@
 aggregation scheme to the 1-minute data to derive a series in coarser
-timescale. To preserve the representativeness of the data we used the
+time-scales. To preserve the representativeness of the data we used the
 following criteria: a) for the daily mean values we accept days with
-more than 50\% of the daytime measurements, present and valid, b)
-monthly values were computed from daily means only when at least 20 days
-were available. To create the daily and monthly climatological means, we
+more than 50\% of the daytime measurements present and valid, b) monthly
+values were computed from daily means only when at least 20 days were
+available. To create the daily and monthly climatological means, we
 averaged the data based on the day of year and calendar month,
@@
 corresponding climatological annual cycle (daily or monthly) from the
-actual data. To estimate SZA contribution to the SDR trends, the
+actual data. To estimate the SZA effect on the SDR trends, the
 one-minute data were aggregated in \(1^\circ\) SZA bins, separately for

### 2023-11-03 DHI_GHI_3_trends_consistency.R

@@ -1141,17 +1141,18 @@ for (adb in database) {
                               unlist(strsplit(avar, split = "_" ))[1:2]),
                             collapse = "_"))
 
-        if (DRAFT == TRUE) {
-            par("mar" = c(3, 5, 2,  1))
-        } else {
-            par("mar" = c(3, 5, 1, 0.5))
-        }
+        # if (DRAFT == TRUE) {
+        #     par("mar" = c(3, 5, 2,  1))
+        # } else {
+        #     par("mar" = c(3, 5, 1, 0.5))
+        # }
         par(pch = 19)
 
         ## scale fonts
         ccex <- ccex_sbs
         par(cex.lab = ccex, cex.axis = ccex, cex.main = ccex, cex = ccex)
 
+        par("mar" = c(3, 5, 1.7, 0.5))
         plot(1, type = "n",
              xlab = "",
              xlim = xlim, ylim = ylim,

### 2023-11-03 Article.Rmd
Commit 2023-11-03 14:14

@@
 <!-- Trends Global -->
-Also, on global scale, an AI aided continental level spatial analysis, with data from multiple station, reach similar conclusions for the above regions and for the global trend [@Yuan2021].
+On global scale, an artificial Intelligence aided spatial analysis on continental level with data from multiple stations reach similar conclusions for these regions and for the global trend [@Yuan2021].
 
-There is a consensus, among researchers, that the major factors of SDR attenuation is the interaction of Sun radiation with atmospheric aerosols and clouds. 
+There is a consensus among researchers that the major factor affecting the variability of SDR attenuation is the interactions of solar radiation with atmospheric aerosols and clouds. 
 Those interactions, among other factors, have been analysed with models [@Li2016; @Samset2018], showing the existence of feedback mechanisms between the two.
-Similar finds, have been showed in observational data [@Schwarz2020; @Ohvril2009; @Zerefos2009; @Xia2007 and references therein].
+Similar findings have been shown in observational data [@Schwarz2020; @Ohvril2009; @Zerefos2009; @Xia2007 and references therein].
 <!-- Sun variability: should cite that there is no significant trend found Wild2021 -->
 
-Due to the variability of the phenomenon, and its contributing factors, there is a constant need to investigate and monitor SDR, in different sites, to estimate its magnitude, and its relation to the local conditions.
-In this study, we examine the trends of SDR, with ground-based measurements at Thessaloniki, Greece for the period `r year(min(ALL_1_daily_DESEAS$Date))` to `r year(max(ALL_1_daily_DESEAS$Date))`, as derived from a CM-21 pyranometer.
+Due to the significant spatial and temporal variability of the trends, and the contributing factors, there is a constant need to monitor and investigate SDR in different sites in order to estimate the degree of variability,  and its relation to the local conditions.
+In this study, we examine the trends of SDR, using ground-based measurements at Thessaloniki, Greece, for the period `r year(min(ALL_1_daily_DESEAS$Date))` to `r year(max(ALL_1_daily_DESEAS$Date))`, as derived from a CM-21 pyranometer.
 <!-- goals -->
-We reevaluated and extended the dataset used by @Bais2013, applying a different algorithm for the identification of clear-/cloud-sky instances [@Reno2016; @Reno2012a], and we derive the radiation trends for the period `r year(min(ALL_1_daily_DESEAS$Date))` to `r year(max(ALL_1_daily_DESEAS$Date))`, under different sky conditions (all-sky, clear-sky and cloud-sky).
+We reevaluated and extended the dataset used by @Bais2013, we applied a different algorithm for the identification of clear-/cloud-sky instances [@Reno2016; @Reno2012a], and we derived the SDR trends for the period `r year(min(ALL_1_daily_DESEAS$Date))` to `r year(max(ALL_1_daily_DESEAS$Date))`, under different sky conditions (all-sky, clear-sky and cloud-sky).
+Finally, we investigated the dependence of the trends on solar zenith angle and season.
 

### 2023-11-03 Article_A.tex
Commit 2023-11-03 11:10
Commit 2023-11-03 14:14

@@
 \author{true \and true}
-\date{2023-11-02}
+\date{2023-11-03}
 
@@
 The term `brightening' is generally used to describe periods of positive SDR trend, and `dimming' for negative trend.
-There are many cases in the long term records of irradiance, showing a systematic change of SDR's trend slope, occurring roughly at the last decades of the 20th century.
-On multiple station in China, a dimming period was reported until about 2000, followed by a brightening period (Yang et al. 2021).
+There are many cases in the long term records of irradiance, showing a systematic change in the magnitude of the trend, occurring roughly in the last decades of the 20th century.
+On multiple stations in China, a dimming period was reported until about 2000, followed by a brightening period (Yang et al. 2021).
 A similar pattern was identified, with the breaking point around 1980, for stations in Central Europe (Wild et al. 2021) and Brazil (Yamasoe et al. 2021).
-Also, on global scale, an AI aided continental level spatial analysis, with data from multiple station, reach similar conclusions for the above regions and for the global trend (Yuan, Leirvik, and Wild 2021).
+On global scale, an artificial Intelligence aided spatial analysis on continental level with data from multiple stations reach similar conclusions for these regions and for the global trend (Yuan, Leirvik, and Wild 2021).
 
-There is a consensus, among researchers, that the major factors of SDR attenuation is the interaction of Sun radiation with atmospheric aerosols and clouds.
+There is a consensus among researchers that the major factor affecting the variability of SDR attenuation is the interactions of solar radiation with atmospheric aerosols and clouds.
 Those interactions, among other factors, have been analysed with models (Li et al. 2016; Samset et al. 2018), showing the existence of feedback mechanisms between the two.
-Similar finds, have been showed in observational data (Schwarz et al. 2020; Ohvril et al. 2009; Zerefos et al. 2009; Xia et al. 2007 and references therein).
+Similar findings have been shown in observational data (Schwarz et al. 2020; Ohvril et al. 2009; Zerefos et al. 2009; Xia et al. 2007 and references therein).
 
-Due to the variability of the phenomenon, and its contributing factors, there is a constant need to investigate and monitor SDR, in different sites, to estimate its magnitude, and its relation to the local conditions.
-In this study, we examine the trends of SDR, with ground-based measurements at Thessaloniki, Greece for the period 1993 to 2023, as derived from a CM-21 pyranometer.
+Due to the significant spatial and temporal variability of the trends, and the contributing factors, there is a constant need to monitor and investigate SDR in different sites in order to estimate the degree of variability, and its relation to the local conditions.
+In this study, we examine the trends of SDR, using ground-based measurements at Thessaloniki, Greece, for the period 1993 to 2023, as derived from a CM-21 pyranometer.
 We reevaluated and extended the dataset used by Bais et al.
-(2013), applying a different algorithm for the identification of clear-/cloud-sky instances (Reno and Hansen 2016; Reno, Hansen, and Stein 2012a), and we derive the radiation trends for the period 1993 to 2023, under different sky conditions (all-sky, clear-sky and cloud-sky).
+(2013), we applied a different algorithm for the identification of clear-/cloud-sky instances (Reno and Hansen 2016; Reno, Hansen, and Stein 2012a), and we derived the SDR trends for the period 1993 to 2023, under different sky conditions (all-sky, clear-sky and cloud-sky).
+Finally, we investigated the dependence of the trends on solar zenith angle and season.
 
@@
 
-We have to note, that the definition of what constitutes as clear or cloudy sky, has some subjectivity, in any method of characterization.
-As a result, the details of the definition are site specific, it relies on a combination of thresholds and comparisons with ideal actinometric models and statistical analysis on different signal metrics.
-The CSid algorithm was calibrated with the main focus, to identify the presence of clouds on the sky dome.
-Although the fine-tuning of the procedure, few marginal cases exist, that have been identified manually as false positive or false negative but cannot affect the final results of the study.
+We have to note, that the definition of clear or cloudy sky, has some subjectivity, in any method of characterization.
+As a result, the details of the definition are site specific, it relies on a combination of thresholds and comparisons with ideal radiation models and statistical analysis of different signal metrics.
+The CSid algorithm was calibrated with the main focus, to identify the presence of clouds on the sky.
+Despite the fine-tuning of the procedure, a few marginal cases exist, that have been identified manually as false positive or false negative but cannot affect the final results of the study.
 
-For completeness, we will provide below a brief overview of the
-clear sky identification algorithm (CSid),
-along with the site specific thresholds.
+For completeness, we will provide below a brief overview of the clear sky identification algorithm (CSid), along with the site specific thresholds.
 To calculate the reference clear sky
@@
 by Long and Ackerman (2000) and Reno and Hansen (2016).
-The target of
-the optimization was the minimization of a function \(f(a)\) (Eq.~\ref{eq:minf}) and
+The target of the optimization was the minimization of a function \(f(a)\) (Eq.~\ref{eq:minf}) and
 was accomplished with the algorithmic function ``optimise'', which is an implementation based on the work of Brent (1973), from the library ``stats'' of the R programming language (R Core Team 2023).
@@
 \end{equation}
-where: \(\text{SDR}_\text{CSref}\) is the reference clear sky SDR, in \(\text{w}\,\text{m}^{-2}\) and \(\theta\) is the solar zenith angle (SZA).
+where: \(\text{SDR}_\text{CSref}\) is the reference clear sky SDR, in \(\text{W}\,\text{m}^{-2}\) and \(\theta\) is the solar zenith angle (SZA).
 
-The criteria that were used to identify whether a measurement was taken
-under clear-sky conditions are presented below.
-A data point is flagged
-as ``clear-sky'' if all criteria are satisfied, otherwise it is considered to be ``cloud-sky''.
-Each criterion was applied
-for a running window of \(11\) consecutive one-minute measurements, and
-the characterization is assigned to the central value of the window.
-Each parameter, was calculated both from the observations and the
-reference clear sky model, for each comparison.
-The allowable range of variation is defined by the
-model-derived value of the parameter multiplied by a factor plus an
-offset.
-The factors and the offsets were
-determined empirically, by manual inspecting each filters performance on
-selected days, and adjusting them accordingly, during an iterative
-process.
+The criteria that were used to identify whether a measurement was taken under clear-sky conditions are presented below.
+A data point is flagged as ``clear-sky'' if all criteria are satisfied, otherwise it is considered to be ``cloud-sky''.
+Each criterion was applied for a running window of \(11\) consecutive one-minute measurements, and the characterization is assigned to the central value of the window.
+Each parameter, was calculated both from the observations and the reference clear sky model, for each comparison.
+The allowable range of variation is defined by the model-derived value of the parameter multiplied by a factor plus an offset.
+The factors and the offsets were determined empirically, by manually inspecting each filter's performance on selected days and adjusting them accordingly during an iterative process.
+The criteria are:
 
@@
   \begin{equation}
-  0.91 \times \overline{\text{SDR}}_{\text{CSref},i} - 20
+  0.91 \times \overline{\text{SDR}}_{\text{CSref},i} - 20\,Wm^{-2}
   < \overline{\text{SDR}}_i <
-  1.095 \times \overline{\text{SDR}}_{\text{CSref},i} + 30
+  1.095 \times \overline{\text{SDR}}_{\text{CSref},i} + 30\,Wm^{-2}
   \label{eq:MeanVIP}
@@
   \end{equation}
-  where: \(t_i\) is the time each SDR measurement has been measured
+  where: \(t_i\) is the time stamp of each SDR measurement.
 \item
@@
 
-Due to a significant measurement uncertainty near the horizon, we have to exclude all measurements with SZA greater than \(85^\circ\).
-Moreover, due to some obstructions around the site (hills and buildings), we excluded data with Azimuth angle between
+Due to the significant measurement uncertainty when the Sun is near the horizon, we have excluded all measurements with SZA greater than \(85^\circ\).
+Moreover, due to some obstructions around the site (hills and buildings), we excluded data with azimuth angle between
 \(35^\circ\) and \(120^\circ\) with SZA greater than \(80^\circ\).
-On the latter instances, Sun is systematically, not visible by the instrument's location.
+On the latter instances, Sun is systematically not visible from the instrument's location.
 To make the measurements comparable throughout the dataset, we adjusted all one-minute radiometric values to the mean Sun - Earth distance.
 Subsequently, we made all measurements relative to the Total Solar Irradiance (TSI) at \(1\,\text{au}\), in order to compensate for the Sun's intensity variability, using a time series of satellite TSI observations.
-The TSI data we use are part of the ``NOAA Climate Data Record of Total Solar Irradiance'' dataset (Coddington et al. 2005).
-Where the initial daily values, were interpolated to match with the time step of our measurements.
-The final dataset contains
+The TSI data we used are part of the ``NOAA Climate Data Record of Total Solar Irradiance'' dataset (Coddington et al. 2005).
+The initial daily values of this dataset were interpolated to match the time step of our measurements.
+In the final dataset
 \(6589967\)
@@
 \(84.2\%\)
-were identified as under clear-sky conditions and subsequently
+were identified as under clear-sky conditions and
 \(15.8\%\)
@@
 
-In order to investigate the SDR trends, we implemented an appropriate aggregation scheme to the 1-minute data to derive a series in coarser timescale.
+In order to investigate the SDR trends, we implemented an appropriate aggregation scheme to the 1-minute data to derive a series in coarser time-scales.
 To preserve the representativeness of the data we used the following criteria:
-a) for the daily mean values we accept days with more than 50\% of the daytime measurements, present and valid,
+a) for the daily mean values we accept days with more than 50\% of the daytime measurements present and valid,
 b) monthly values were computed from daily means
@@
 Finally, each data set was deseasonalized by subtracting the corresponding climatological annual cycle (daily or monthly) from the actual data.
-To estimate SZA contribution to the SDR trends, the one-minute data were aggregated in \(1^\circ\) SZA bins, separately for the morning and afternoon hours, and then were deseasonalized as mentioned above.
+To estimate the SZA effect on the SDR trends, the one-minute data were aggregated in \(1^\circ\) SZA bins, separately for the morning and afternoon hours, and then were deseasonalized as mentioned above.
 

### 2023-11-03 MDPI_submition.Rmd
Commit 2023-11-03 14:14

@@
 \begin{equation}
-  X_i < `r offVSM` \label{eq:VSMcr}
+  X_i < `r offVSM`\,Wm^{-2} \label{eq:VSMcr}
 \end{equation}
@@
 In the final dataset
-$`r SDR_N_All`$
-one-minute measurements, of which, 
+<!-- $`r SDR_N_All`$ -->
 $`r signif(100 * SDR_N_Clear / SDR_N_All, 3)`\%$
-were identified as under clear-sky conditions and
+of the data were identified as under clear-sky conditions and
 $`r signif(100 * SDR_N_Cloud / SDR_N_All, 3)`\%$

### 2023-11-03 MDPI_submition.tex
Commit 2023-11-03 14:14

@@
   threshold (Eq.~\ref{eq:VSMcr}): \begin{equation}
-    X_i < 7.5 \label{eq:VSMcr}
+    X_i < 7.5\,Wm^{-2} \label{eq:VSMcr}
   \end{equation}
@@
 initial daily values of this dataset were interpolated to match the time
-step of our measurements. In the final dataset \(6589967\) one-minute
-measurements, of which, \(84.2\%\) were identified as under clear-sky
-conditions and \(15.8\%\) as under cloud-sky conditions.
+step of our measurements. In the final dataset \(84.2\%\) of the data
+were identified as under clear-sky conditions and \(15.8\%\) as under
+cloud-sky conditions.
 

### 2023-11-03 MDPI_submition.Rmd
Commit 2023-11-03 15:09

@@
 
-We calculated the linear SDR trends, from the departures of the mean daily values from the seasonal values, and for the three sky conditions (Table\ \ref{tab:trendtable}).
+We calculated the linear SDR trends, from the departures of the mean daily values from the daily climatology and for the three sky conditions (Table\ \ref{tab:trendtable}).
 In Figure\ \ref{fig:trendALL} we present only the time series under all-sky conditions; the plots for clear-sky and cloud-sky conditions, are very similar and are shown in the Appendix (Figures\ \ref{fig:trendCLEAR} and\  \ref{fig:trendCLOUD}).
 We observe a positive trend for all-sky conditions 
-($`r alltrend`\,\%/y$), a very close but lower trend for clear-skies ($`r cleartrend`\,\%/y$) and a negative weaker trend for cloud-skies ($`r cloudtrend`\,\%/y$).
+($`r alltrend`\,\%/y$), a very close but smaller trend for clear-skies ($`r cleartrend`\,\%/y$) and a negative weaker trend for cloudy-skies ($`r cloudtrend`\,\%/y$).
 In the studied period, there is no significant break or change in the variability of the time series.
-Other studies for the European region report a change of the SDR slope, around 1980, a few years before the start of our records [@Wild2021; @Yuan2021; @Ohmura2009].
-It is interesting to note, that for the observation period, the trend of the TSI is
-$`r format(TSIslopeAll[1,1], scientific = F, digits = 2)`\,\%/y$, and thus we can not attribute any major effect on SDR trend to Solar variability.
+Other studies for the European region reported a change of the SDR slope, around 1980 [@Wild2021; @Yuan2021; @Ohmura2009], a few years before the start of our records. 
+It is interesting to note, that for the observations period, the trend of the TSI is
+$`r format(TSIslopeAll[1,1], scientific = F, digits = 2)`\,\%/y$, 
+and thus we can not attribute any major effect on SDR trend to Solar variability.
 
@@
 
-Although the year-to-year variability of the anomalies (Figure \ref{fig:trendALL} and Figures \ref{fig:trendCLEAR}, \ref{fig:trendCLOUD} on Appendix), shows a rather homogeneous behavior, plots of the cumulative sums (CUSUM) of the anomalies can reveal different structures within all the three sky conditions [@Regier2019].
+Although the year-to-year variability of the anomalies (Figure \ref{fig:trendALL} and Figures \ref{fig:trendCLEAR}, \ref{fig:trendCLOUD} in Appendix), shows a rather homogeneous behavior, plots of the cumulative sums (CUSUM) of the anomalies can reveal different structures within all the three sky conditions [@Regier2019].
 In the cases of all-sky and clear-sky (Figures \ref{fig:cusummonth-1} and \ref{fig:cusummonth-2}), we observe three macroscopic periods.
@@
 
-The solar zenith angle is a major driver of the attenuation of SDR reaching the ground, as SDR is affected by the enhancement of the radiation path in the atmosphere, especially in urban environment with human activities emitting aerosols [@Wang2021].
+The solar zenith angle is a major factor of SDR reaching the ground, due to the enhancement of the radiation path in the atmosphere, especially in urban environments with human activities emitting aerosols [@Wang2021].
 In order to estimate the effect of the SZA on the SDR trends, we grouped the anomaly data in bins of $1^\circ$ SZA, and calculated the overall trend for each bin before noon and after noon (Figure\ \ref{fig:szatrends}).
@@
 The trends in the morning and afternoon hours are more or less consistent with small differences, which can be attributed to systematic diurnal variations of aerosols, particularly during the warm period of the year [@Wang2021].
-For cloud-sky conditions (Figure\ \ref{fig:szatrends-3}), we can not discern any significant dependence of the SDR trend with SZA.
+For cloudy-sky conditions (Figure\ \ref{fig:szatrends-3}), we can not discern any significant dependence of the SDR trend with SZA.
 For SZAs $16^\circ$ - $50^\circ$, the trends range within about $\pm 0.2\,\%/y$, with a weak statistical significance.
@@
 
-Similar to the long term trends discussed above, we have calculated the trend of the anomalies for the three different sky conditions, and for each season of the year, using the corresponding mean monthly values
+Similarly to the long term trends discussed above, we have calculated the trend of the anomalies for the three different sky conditions, and for each season of the year, using the corresponding mean monthly values
 (Figure\ \ref{fig:seasonalALL} and Table\ \ref{tab:trendseasontable}).
@@
 ($`r signif(spp$"Trend [%/year]"[spp$Season == "Autumn" & spp$"Sky condition" == "All sky"], 2)`\,\%/y$,
-value similar to the long term trend) both statistically significant above the $99\,\%$ confidence level. 
-In spring and summer, the trends are much smaller and less statistical significant.
-These seasonal differences indicate a possible relation of the trends to clouds during winter and autumn.
+a value close to the long term trend) both statistically significant above the $99\,\%$ confidence level. 
+In spring and summer, the trends are much smaller and of lesser statistical significance.
+These seasonal differences indicate a possible relation of the trends in SDR to trends of clouds during winter and autumn.
 For clear-skies, the trend in winter is $`r signif(spp$"Trend [%/year]"[spp$Season == "Winter" & spp$"Sky condition" == "Clear sky"], 2)`\,\%/y$, larger than for all-skies ($`r signif(spp$"Trend [%/year]"[spp$Season == "Winter" & spp$"Sky condition" == "All sky"], 2)`\,\%/y$), which is another indication of increasing cloud thickness.

### 2023-11-03 MDPI_submition.tex
Commit 2023-11-03 15:09

@@
 We calculated the linear SDR trends, from the departures of the mean
-daily values from the seasonal values, and for the three sky conditions
+daily values from the daily climatology and for the three sky conditions
 (Table~\ref{tab:trendtable}). In Figure~\ref{fig:trendALL} we present
@@
 positive trend for all-sky conditions (\(0.38\,\%/y\)), a very close but
-lower trend for clear-skies (\(0.35\,\%/y\)) and a negative weaker trend
-for cloud-skies (\(-0.28\,\%/y\)). In the studied period, there is no
-significant break or change in the variability of the time series. Other
-studies for the European region report a change of the SDR slope, around
-1980, a few years before the start of our records
-\citep{Wild2021, Yuan2021, Ohmura2009}. It is interesting to note, that
-for the observation period, the trend of the TSI is \(-0.0002\,\%/y\),
-and thus we can not attribute any major effect on SDR trend to Solar
-variability.
+smaller trend for clear-skies (\(0.35\,\%/y\)) and a negative weaker
+trend for cloudy-skies (\(-0.28\,\%/y\)). In the studied period, there
+is no significant break or change in the variability of the time series.
+Other studies for the European region reported a change of the SDR
+slope, around 1980 \citep{Wild2021, Yuan2021, Ohmura2009}, a few years
+before the start of our records. It is interesting to note, that for the
+observations period, the trend of the TSI is \(-0.0002\,\%/y\), and thus
+we can not attribute any major effect on SDR trend to Solar variability.
 
@@
 \ref{fig:trendALL} and Figures \ref{fig:trendCLEAR},
-\ref{fig:trendCLOUD} on Appendix), shows a rather homogeneous behavior,
+\ref{fig:trendCLOUD} in Appendix), shows a rather homogeneous behavior,
 plots of the cumulative sums (CUSUM) of the anomalies can reveal
@@
 
-The solar zenith angle is a major driver of the attenuation of SDR
-reaching the ground, as SDR is affected by the enhancement of the
-radiation path in the atmosphere, especially in urban environment with
-human activities emitting aerosols \citep{Wang2021}. In order to
-estimate the effect of the SZA on the SDR trends, we grouped the anomaly
-data in bins of \(1^\circ\) SZA, and calculated the overall trend for
-each bin before noon and after noon (Figure~\ref{fig:szatrends}).
-Although there are seasonal dependencies of the minimum SZA (see
-Appendix, Figure~\ref{fig:SZAtrendSeason}), these dependencies would not
-be further examined here. For all-sky and clear-sky conditions the
-brightening effect of SDR (positive trend) is stronger for large SZAs
-(Figures~\ref{fig:szatrends-1} and \ref{fig:szatrends-2}). The trends in
-the morning and afternoon hours are more or less consistent with small
-differences, which can be attributed to systematic diurnal variations of
-aerosols, particularly during the warm period of the year
-\citep{Wang2021}. For cloud-sky conditions
-(Figure~\ref{fig:szatrends-3}), we can not discern any significant
-dependence of the SDR trend with SZA. For SZAs \(16^\circ\) -
-\(50^\circ\), the trends range within about \(\pm 0.2\,\%/y\), with a
+The solar zenith angle is a major factor of SDR reaching the ground, due
+to the enhancement of the radiation path in the atmosphere, especially
+in urban environments with human activities emitting aerosols
+\citep{Wang2021}. In order to estimate the effect of the SZA on the SDR
+trends, we grouped the anomaly data in bins of \(1^\circ\) SZA, and
+calculated the overall trend for each bin before noon and after noon
+(Figure~\ref{fig:szatrends}). Although there are seasonal dependencies
+of the minimum SZA (see Appendix, Figure~\ref{fig:SZAtrendSeason}),
+these dependencies would not be further examined here. For all-sky and
+clear-sky conditions the brightening effect of SDR (positive trend) is
+stronger for large SZAs (Figures~\ref{fig:szatrends-1} and
+\ref{fig:szatrends-2}). The trends in the morning and afternoon hours
+are more or less consistent with small differences, which can be
+attributed to systematic diurnal variations of aerosols, particularly
+during the warm period of the year \citep{Wang2021}. For cloudy-sky
+conditions (Figure~\ref{fig:szatrends-3}), we can not discern any
+significant dependence of the SDR trend with SZA. For SZAs \(16^\circ\)
+- \(50^\circ\), the trends range within about \(\pm 0.2\,\%/y\), with a
 weak statistical significance. Between \(50^\circ\) and \(75^\circ\) SZA
@@
 
-Similar to the long term trends discussed above, we have calculated the
-trend of the anomalies for the three different sky conditions, and for
-each season of the year, using the corresponding mean monthly values
+Similarly to the long term trends discussed above, we have calculated
+the trend of the anomalies for the three different sky conditions, and
+for each season of the year, using the corresponding mean monthly values
 (Figure~\ref{fig:seasonalALL} and Table~\ref{tab:trendseasontable}). For
 all-sky conditions the trend in SDR in winter is the largest
-(\(0.69\,\%/y\)), followed by the trend in autumn (\(0.43\,\%/y\), value
-similar to the long term trend) both statistically significant above the
-\(99\,\%\) confidence level. In spring and summer, the trends are much
-smaller and less statistical significant. These seasonal differences
-indicate a possible relation of the trends to clouds during winter and
-autumn. For clear-skies, the trend in winter is \(0.83\,\%/y\), larger
-than for all-skies (\(0.69\,\%/y\)), which is another indication of
-increasing cloud thickness. Moreover, the trends under clear- and
-cloud-sky conditions are almost complementary to each other,
-particularly for winter and autumn, where the signal is stronger. During
-spring and summer the statistical significance is very low and the
-actual trend too small for a meaningful comparison.
+(\(0.69\,\%/y\)), followed by the trend in autumn (\(0.43\,\%/y\), a
+value close to the long term trend) both statistically significant above
+the \(99\,\%\) confidence level. In spring and summer, the trends are
+much smaller and of lesser statistical significance. These seasonal
+differences indicate a possible relation of the trends in SDR to trends
+of clouds during winter and autumn. For clear-skies, the trend in winter
+is \(0.83\,\%/y\), larger than for all-skies (\(0.69\,\%/y\)), which is
+another indication of increasing cloud thickness. Moreover, the trends
+under clear- and cloud-sky conditions are almost complementary to each
+other, particularly for winter and autumn, where the signal is stronger.
+During spring and summer the statistical significance is very low and
+the actual trend too small for a meaningful comparison.
 

### 2023-11-03 MDPI_submition.Rmd
Commit 2023-11-03 19:09

@@
 For completeness, we will provide below a brief overview of the clear sky identification algorithm (CSid), along with the site specific thresholds.
-To calculate the reference clear sky
-$\text{SDR}_\text{CSref}$ we used the $\text{SDR}_\text{Haurwitz}$ derived by
-the radiation model of @Haurwitz1945, adjusted for our site with a
-factor $a$ (Eq.\ \ref{eq:ahau}), resulted by an iterative optimization process, as described
-by @Long2000 and @Reno2016.
+To calculate the reference clear sky $\text{SDR}_\text{CSref}$ we used the $\text{SDR}_\text{Haurwitz}$ derived by the radiation model of @Haurwitz1945, adjusted for our site with a factor $a$ (Eq.\ \ref{eq:ahau}), resulted by an iterative optimization process, as described by @Long2000 and @Reno2016.
 The target of the optimization was the minimization of a function $f(a)$ (Eq.\ \ref{eq:minf}) and
@@
 The optimization and the selection of the clear sky reference model, was performed on SDR observations for the period 2016 - 2021.
-During the optimization, eight simple clear sky radiation models were tested (Daneshyar-Paltridge-Proctor, Kasten-Czeplak, Haurwitz, Berger-Duffie, Adnot-Bourges-Campana-Gicquel, Robledo-Soler, Kasten and Ineichen-Perez), with a wide range of factors.
+During the optimization, eight simple clear sky radiation models were tested (namely, Daneshyar-Paltridge-Proctor, Kasten-Czeplak, Haurwitz, Berger-Duffie, Adnot-Bourges-Campana-Gicquel, Robledo-Soler, Kasten and Ineichen-Perez), with a wide range of factors.
 These models are described in more details by @Reno2012 and evaluated by @Reno2016.

### 2023-11-03 MDPI_submition.tex
Commit 2023-11-03 19:09

@@
 the period 2016 - 2021. During the optimization, eight simple clear sky
-radiation models were tested (Daneshyar-Paltridge-Proctor,
+radiation models were tested (namely, Daneshyar-Paltridge-Proctor,
 Kasten-Czeplak, Haurwitz, Berger-Duffie, Adnot-Bourges-Campana-Gicquel,

### 2023-11-04 Article.Rmd
Commit 2023-11-04 13:09

@@
 funding: |
-  Please add: ``This research received no external funding'' or ``This research 
-  was funded by NAME OF FUNDER grant number XXX.'' and  and ``The APC was funded 
-  by XXX''. Check carefully that the details given are accurate and use the 
-  standard spelling of funding agency names at 
-  \url{https://search.crossref.org/funding}, any errors may affect your future 
-  funding.
+  This research received no external funding.
 
@@
    - \usepackage{subcaption}
+   - \captionsetup[sub]{position=bottom, labelfont={bf, small, stretch=1.17}, labelsep=space, textfont={small, stretch=1.17}, aboveskip=6pt,  belowskip=-6pt, singlelinecheck=off, justification=justified}
 
@@
 #Wild2012 : simplify interaction nice graph
-
 -->
@@
 For completeness, we will provide below a brief overview of the clear sky identification algorithm (CSid), along with the site specific thresholds.
-To calculate the reference clear sky
-$\text{SDR}_\text{CSref}$ we used the $\text{SDR}_\text{Haurwitz}$ derived by
-the radiation model of @Haurwitz1945, adjusted for our site with a
-factor $a$ (Eq.\ \ref{eq:ahau}), resulted by an iterative optimization process, as described
-by @Long2000 and @Reno2016.
-The target of the optimization was the minimization of a function $f(a)$ (Eq.\ \ref{eq:minf}) and
-was accomplished with the algorithmic function "optimise", which is an implementation based on the work of @Brent1973, from the library "stats" of the R programming language [@RCT2023]. 
+To calculate the reference clear sky $\text{SDR}_\text{CSref}$ we used the $\text{SDR}_\text{Haurwitz}$ derived by the radiation model of @Haurwitz1945, adjusted for our site with a factor $a$ (Eq.\ \ref{eq:ahau}), resulted by an iterative optimization process, as described by @Long2000 and @Reno2016.
+The target of the optimization was the minimization of a function $f(a)$ (Eq.\ \ref{eq:minf}) and was accomplished with the algorithmic function "optimise", which is an implementation based on the work of @Brent1973, from the library "stats" of the R programming language [@RCT2023].
 The optimization and the selection of the clear sky reference model, was performed on SDR observations for the period 2016 - 2021.
-During the optimization, eight simple clear sky radiation models were tested (Daneshyar-Paltridge-Proctor, Kasten-Czeplak, Haurwitz, Berger-Duffie, Adnot-Bourges-Campana-Gicquel, Robledo-Soler, Kasten and Ineichen-Perez), with a wide range of factors.
+During the optimization, eight simple clear sky radiation models were tested (namely, Daneshyar-Paltridge-Proctor, Kasten-Czeplak, Haurwitz, Berger-Duffie, Adnot-Bourges-Campana-Gicquel, Robledo-Soler, Kasten and Ineichen-Perez), with a wide range of factors.
 These models are described in more details by @Reno2012 and evaluated by @Reno2016.
-We found, that Haurwitz's model, adjusted with the factor $a = `r signif(alpha,3)`$ yields one of the lowest root mean squared errors (RMSE), 
-<!--[....rephrased....]{.comment-start id="3" author="A N" date="2023-07-15T20:32:50Z"}-->
-while the procedure, manages to characterize the majority of the data.
-<!--[]{.comment-end id="3"}--> 
+We found, that Haurwitz's model, adjusted with the factor $a = `r signif(alpha,3)`$ yields one of the lowest root mean squared errors (RMSE), while the procedure, manages to characterize the majority of the data.
 Thus, our clear sky reference is derived by the Eq.\ \ref{eq:ahau}.
@@
 
-We calculated the linear SDR trends, from the departures of the mean daily values from the seasonal values, and for the three sky conditions (Table\ \ref{tab:trendtable}).
+We calculated the linear SDR trends, from the departures of the mean daily values from the daily climatology and for the three sky conditions (Table\ \ref{tab:trendtable}).
 In Figure\ \ref{fig:trendALL} we present only the time series under all-sky conditions; the plots for clear-sky and cloud-sky conditions, are very similar and are shown in the Appendix (Figures\ \ref{fig:trendCLEAR} and\  \ref{fig:trendCLOUD}).
 We observe a positive trend for all-sky conditions
-($`r alltrend`\,\%/y$), a very close but lower trend for clear-skies ($`r cleartrend`\,\%/y$) and a negative weaker trend for cloud-skies ($`r cloudtrend`\,\%/y$).
+($`r alltrend`\,\%/y$), a very close but smaller trend for clear-skies ($`r cleartrend`\,\%/y$) and a negative weaker trend for cloudy-skies ($`r cloudtrend`\,\%/y$).
 In the studied period, there is no significant break or change in the variability of the time series.
-Other studies for the European region report a change of the SDR slope, around 1980, a few years before the start of our records [@Wild2021; @Yuan2021; @Ohmura2009].
-It is interesting to note, that for the observation period, the trend of the TSI is
-$`r format(TSIslopeAll[1,1], scientific = F, digits = 2)`\,\%/y$, and thus we can not attribute any major effect on SDR trend to Solar variability.
+Other studies for the European region reported a change of the SDR slope, around 1980 [@Wild2021; @Yuan2021; @Ohmura2009], a few years before the start of our records.
+It is interesting to note, that for the observations period, the trend of the TSI is
+$`r format(TSIslopeAll[1,1], scientific = F, digits = 2)`\,\%/y$,
+and thus we can not attribute any major effect on SDR trend to Solar variability.
 
@@
 
-Although the year-to-year variability of the anomalies (Figure \ref{fig:trendALL} and Figures \ref{fig:trendCLEAR}, \ref{fig:trendCLOUD} on Appendix), shows a rather homogeneous behavior, plots of the cumulative sums (CUSUM) of the anomalies can reveal different structures within all the three sky conditions [@Regier2019].
-In the cases of all-sky and clear-sky (Figures \ref{fig:cusummonth-1} and \ref{fig:cusummonth-2}), we observe three macroscopic periods.
-A downward part from the start until about 2005, an relative steady part until about 2016 and finally, a steep upward part until the present.
-For cloud-sky (Figure\ \ref{fig:cusummonth-3}), we have a different sequence, it begins with a relative steady part until 1997, followed by a upward part until 2005, a long decline until 2020, with a small positive slope until the present.
+Although the year-to-year variability of the anomalies (Figure \ref{fig:trendALL} and Figures \ref{fig:trendCLEAR}, \ref{fig:trendCLOUD} in Appendix), shows a rather homogeneous behaviour, plots of the cumulative sums (CUSUM) [@Regier2019] of the anomalies can reveal different structures in the records of all three sky conditions.
+In the cases of all-sky and clear-sky conditions (Figures \ref{fig:cusummonth-1} and \ref{fig:cusummonth-2}), we observe three macroscopic periods.
+A downward part from the start until about 2005, a relatively steady part until about 2016 and, finally, a steep upward part until the present.
+For cloud-sky (Figure\ \ref{fig:cusummonth-3}), we have a different pattern; it begins with a relatively steady part until 1997, followed by an upward part until 2005, and a long decline until 2020, with a small positive slope until the present.
 For a uniform trend, we would expect the CUSUMs of the anomalies to have a symmetric 'V' shape.
 This would indicate that the anomalies are evenly distributed around the climatological mean, and for a positive uniform trend, the first half to be below and the other half above the climatological mean.
-In our case, there is more complex evolution of the phenomenon.
+In our case, there is a more complex evolution of the anomalies.
 Another distinct feature of the CUSUMs, is the different pattern of the cloudy-sky dataset which peaks around the middle of the period.
-Although, it seems to exist a complementary relation to the CUSUMs of the clear- and all-sky cases, we can not assert that clouds are the main drive for this relation, due to the great difference in the number of observational data between the datasets (Table\ \ref{tab:trendtable}).
+Although, there seems to exist a complementary relation to the CUSUMs of the clear- and all-sky cases, we can not assert that clouds are the main driver for this relation due to the great difference in the number of observational data between the two datasets (Table\ \ref{tab:trendtable}).
 
@@
 
-In order to investigate more, the features within the trends, we created another set of CUSUMs plots by subtracting the corresponding long term trend from the SDR anomaly data, prior to the CUSUM calculation (Figure\ \ref{fig:cusumnotrendmonthly}).
-With this approach periods when the CUSUMs diverge from zero can be interpreted as systematic variation of SDR from the climatological mean.
-When the CUSUM is increasing, the added values are above the climatological values of SDR trend and vice versa.
-Overall, for all- and clear-sky conditions (Figure\ \ref{fig:cusumnotrendmonthly-1} and\ \ref{fig:cusumnotrendmonthly-2}) we observe periods when the anomalies diverge from the climatological value, each lasting for several years.
+In order to investigate further the features of the CUSUMs, we created another set of CUSUM plots by subtracting the corresponding long term trend from the SDR anomaly data, prior to the CUSUM calculation (Figure\ \ref{fig:cusumnotrendmonthly}).
+With this approach periods when the CUSUMs diverge from zero can be interpreted as a systematic variation of SDR from the climatological mean. 
+When the CUSUM is increasing, the added values are above the climatological values of the SDR trend and vice versa.
+Overall, for all- and clear-sky conditions (Figures\ \ref{fig:cusumnotrendmonthly-1} and\ \ref{fig:cusumnotrendmonthly-2}) we observe periods when the anomalies diverge from the climatological value, each lasting for several years.
 The pattern for the all-sky and clear-sky conditions is very similar, suggesting prevalence in clear skies over Thessaloniki.
@@
 ($`r signif(spp$"Trend [%/year]"[spp$Season == "Autumn" & spp$"Sky condition" == "All sky"], 2)`\,\%/y$,
-value similar to the long term trend) both statistically significant above the $99\,\%$ confidence level. 
-In spring and summer, the trends are much smaller and less statistical significant.
-These seasonal differences indicate a possible relation of the trends to clouds during winter and autumn.
+a value close to the long term trend) both statistically significant above the $99\,\%$ confidence level.
+In spring and summer, the trends are much smaller and of lesser statistical significance.
+These seasonal differences indicate a possible relation of the trends in SDR to trends of clouds during winter and autumn.
 For clear-skies, the trend in winter is $`r signif(spp$"Trend [%/year]"[spp$Season == "Winter" & spp$"Sky condition" == "Clear sky"], 2)`\,\%/y$, larger than for all-skies ($`r signif(spp$"Trend [%/year]"[spp$Season == "Winter" & spp$"Sky condition" == "All sky"], 2)`\,\%/y$), which is another indication of increasing cloud thickness.

### 2023-11-04 MDPI_submition.Rmd
Commit 2023-11-04 13:09

@@
 funding: |
-  Please add: ``This research received no external funding'' or ``This research 
-  was funded by NAME OF FUNDER grant number XXX.'' and  and ``The APC was funded 
-  by XXX''. Check carefully that the details given are accurate and use the 
-  standard spelling of funding agency names at 
-  \url{https://search.crossref.org/funding}, any errors may affect your future 
-  funding.
+  This research received no external funding.
+
+# funding: |
+#   Please add: ``This research received no external funding'' or ``This research
+#   was funded by NAME OF FUNDER grant number XXX.'' and  and ``The APC was funded
+#   by XXX''. Check carefully that the details given are accurate and use the
+#   standard spelling of funding agency names at
+#   \url{https://search.crossref.org/funding}, any errors may affect your future
+#   funding.
 
@@
    - \usepackage{subcaption}
-   - \captionsetup[sub]{position=bottom, labelfont={bf, small, stretch=1.17}, labelsep=space, textfont={small, stretch=1.5}, aboveskip=6pt,  belowskip=6pt, singlelinecheck=off, justification=justified}
+   - \captionsetup[sub]{position=bottom, labelfont={bf, small, stretch=1.17}, labelsep=space, textfont={small, stretch=1.17}, aboveskip=6pt,  belowskip=-6pt, singlelinecheck=off, justification=justified}
+
+#  - \captionsetup[sub]{position=bottom, labelfont={bf, small, stretch=1.17}, labelsep=space, textfont={small, stretch=1.5}, aboveskip=6pt,  belowskip=6pt, singlelinecheck=off, justification=justified}
+#  \captionsetup[figure]{position=bottom, labelfont={bf, small, stretch=1.17}, labelsep=period, textfont={small, stretch=1.17}, aboveskip=6pt, belowskip=-6pt, singlelinecheck=off, justification=justified}
 
@@
 To calculate the reference clear sky $\text{SDR}_\text{CSref}$ we used the $\text{SDR}_\text{Haurwitz}$ derived by the radiation model of @Haurwitz1945, adjusted for our site with a factor $a$ (Eq.\ \ref{eq:ahau}), resulted by an iterative optimization process, as described by @Long2000 and @Reno2016.
-The target of the optimization was the minimization of a function $f(a)$ (Eq.\ \ref{eq:minf}) and
-was accomplished with the algorithmic function "optimise", which is an implementation based on the work of @Brent1973, from the library "stats" of the R programming language [@RCT2023]. 
+The target of the optimization was the minimization of a function $f(a)$ (Eq.\ \ref{eq:minf}) and was accomplished with the algorithmic function "optimise", which is an implementation based on the work of @Brent1973, from the library "stats" of the R programming language [@RCT2023].
 The optimization and the selection of the clear sky reference model, was performed on SDR observations for the period 2016 - 2021.
@@
 These models are described in more details by @Reno2012 and evaluated by @Reno2016.
-We found, that Haurwitz's model, adjusted with the factor $a = `r signif(alpha,3)`$ yields one of the lowest root mean squared errors (RMSE), 
-<!--[....rephrased....]{.comment-start id="3" author="A N" date="2023-07-15T20:32:50Z"}-->
-while the procedure, manages to characterize the majority of the data.
-<!--[]{.comment-end id="3"}--> 
+We found, that Haurwitz's model, adjusted with the factor $a = `r signif(alpha,3)`$ yields one of the lowest root mean squared errors (RMSE), while the procedure, manages to characterize the majority of the data.
 Thus, our clear sky reference is derived by the Eq.\ \ref{eq:ahau}.
@@
 
-Although the year-to-year variability of the anomalies (Figure \ref{fig:trendALL} and Figures \ref{fig:trendCLEAR}, \ref{fig:trendCLOUD} in Appendix), shows a rather homogeneous behavior, plots of the cumulative sums (CUSUM) of the anomalies can reveal different structures within all the three sky conditions [@Regier2019].
-In the cases of all-sky and clear-sky (Figures \ref{fig:cusummonth-1} and \ref{fig:cusummonth-2}), we observe three macroscopic periods.
-A downward part from the start until about 2005, an relative steady part until about 2016 and finally, a steep upward part until the present.
-For cloud-sky (Figure\ \ref{fig:cusummonth-3}), we have a different sequence, it begins with a relative steady part until 1997, followed by a upward part until 2005, a long decline until 2020, with a small positive slope until the present.
+Although the year-to-year variability of the anomalies (Figure \ref{fig:trendALL} and Figures \ref{fig:trendCLEAR}, \ref{fig:trendCLOUD} in Appendix), shows a rather homogeneous behaviour, plots of the cumulative sums (CUSUM) [@Regier2019] of the anomalies can reveal different structures in the records of all three sky conditions.
+In the cases of all-sky and clear-sky conditions (Figures \ref{fig:cusummonth-1} and \ref{fig:cusummonth-2}), we observe three macroscopic periods.
+A downward part from the start until about 2005, a relatively steady part until about 2016 and, finally, a steep upward part until the present.
+For cloud-sky (Figure\ \ref{fig:cusummonth-3}), we have a different pattern; it begins with a relatively steady part until 1997, followed by an upward part until 2005, and a long decline until 2020, with a small positive slope until the present.
 For a uniform trend, we would expect the CUSUMs of the anomalies to have a symmetric 'V' shape.
 This would indicate that the anomalies are evenly distributed around the climatological mean, and for a positive uniform trend, the first half to be below and the other half above the climatological mean.
-In our case, there is more complex evolution of the phenomenon.
+In our case, there is a more complex evolution of the anomalies.
 Another distinct feature of the CUSUMs, is the different pattern of the cloudy-sky dataset which peaks around the middle of the period.
-Although, it seems to exist a complementary relation to the CUSUMs of the clear- and all-sky cases, we can not assert that clouds are the main drive for this relation, due to the great difference in the number of observational data between the datasets (Table\ \ref{tab:trendtable}).
+Although, there seems to exist a complementary relation to the CUSUMs of the clear- and all-sky cases, we can not assert that clouds are the main driver for this relation due to the great difference in the number of observational data between the two datasets (Table\ \ref{tab:trendtable}).
 
@@
 
-In order to investigate more, the features within the trends, we created another set of CUSUMs plots by subtracting the corresponding long term trend from the SDR anomaly data, prior to the CUSUM calculation (Figure\ \ref{fig:cusumnotrendmonthly}).
-With this approach periods when the CUSUMs diverge from zero can be interpreted as systematic variation of SDR from the climatological mean.
-When the CUSUM is increasing, the added values are above the climatological values of SDR trend and vice versa.
-Overall, for all- and clear-sky conditions (Figure\ \ref{fig:cusumnotrendmonthly-1} and\ \ref{fig:cusumnotrendmonthly-2}) we observe periods when the anomalies diverge from the climatological value, each lasting for several years.
-The pattern for the all-sky and clear-sky conditions is very similar, suggesting prevalence in clear skies over Thessaloniki.
-It is interesting that in the period 1993 - 2016 the anomalies of all-sky and clear-sky, have a high variability around zero.
-After 2016, the range of the variability is decreased at about one third of the prior period.
+In order to investigate further the features of the CUSUMs, we created another set of CUSUM plots by subtracting the corresponding long term trend from the SDR anomaly data, prior to the CUSUM calculation (Figure\ \ref{fig:cusumnotrendmonthly}).
+With this approach periods when the CUSUMs diverge from zero can be interpreted as a systematic variation of SDR from the climatological mean. 
+When the CUSUM is increasing, the added values are above the climatological values of the SDR trend and vice versa.
+Overall, for all- and clear-sky conditions (Figures\ \ref{fig:cusumnotrendmonthly-1} and\ \ref{fig:cusumnotrendmonthly-2}) we observe periods when the anomalies diverge from the climatological value, each lasting for several years.
+The pattern in both datasets is very similar, suggesting prevalence in clear skies over Thessaloniki.
+It is interesting that in the period 1993 - 2016 the anomalies have a high variability around zero. After 2016, the range of the variability is decreased at about one third of the prior period.
 For cloud-sky conditions (Figure\ \ref{fig:cusumnotrendmonthly-3}) the period 1997 - 2008 is dominated by positive CUSUMs suggesting a reduced effect of clouds on SDR.

### 2023-11-04 MDPI_submition.tex
Commit 2023-11-04 13:09

@@
 \usepackage{subcaption}
-\captionsetup[sub]{position=bottom, labelfont={bf, small, stretch=1.17}, labelsep=space, textfont={small, stretch=1.5}, aboveskip=6pt,  belowskip=6pt, singlelinecheck=off, justification=justified}
+\captionsetup[sub]{position=bottom, labelfont={bf, small, stretch=1.17}, labelsep=space, textfont={small, stretch=1.17}, aboveskip=6pt,  belowskip=-6pt, singlelinecheck=off, justification=justified}
 \usepackage{longtable}
@@
 \ref{fig:trendALL} and Figures \ref{fig:trendCLEAR},
-\ref{fig:trendCLOUD} in Appendix), shows a rather homogeneous behavior,
-plots of the cumulative sums (CUSUM) of the anomalies can reveal
-different structures within all the three sky conditions
-\citep{Regier2019}. In the cases of all-sky and clear-sky (Figures
+\ref{fig:trendCLOUD} in Appendix), shows a rather homogeneous behaviour,
+plots of the cumulative sums (CUSUM) \citep{Regier2019} of the anomalies
+can reveal different structures in the records of all three sky
+conditions. In the cases of all-sky and clear-sky conditions (Figures
 \ref{fig:cusummonth-1} and \ref{fig:cusummonth-2}), we observe three
-macroscopic periods. A downward part from the start until about 2005, an
-relative steady part until about 2016 and finally, a steep upward part
-until the present. For cloud-sky (Figure~\ref{fig:cusummonth-3}), we
-have a different sequence, it begins with a relative steady part until
-1997, followed by a upward part until 2005, a long decline until 2020,
-with a small positive slope until the present. For a uniform trend, we
-would expect the CUSUMs of the anomalies to have a symmetric `V' shape.
-This would indicate that the anomalies are evenly distributed around the
-climatological mean, and for a positive uniform trend, the first half to
-be below and the other half above the climatological mean. In our case,
-there is more complex evolution of the phenomenon. Another distinct
-feature of the CUSUMs, is the different pattern of the cloudy-sky
-dataset which peaks around the middle of the period. Although, it seems
-to exist a complementary relation to the CUSUMs of the clear- and
-all-sky cases, we can not assert that clouds are the main drive for this
-relation, due to the great difference in the number of observational
-data between the datasets (Table~\ref{tab:trendtable}).
+macroscopic periods. A downward part from the start until about 2005, a
+relatively steady part until about 2016 and, finally, a steep upward
+part until the present. For cloud-sky (Figure~\ref{fig:cusummonth-3}),
+we have a different pattern; it begins with a relatively steady part
+until 1997, followed by an upward part until 2005, and a long decline
+until 2020, with a small positive slope until the present. For a uniform
+trend, we would expect the CUSUMs of the anomalies to have a symmetric
+`V' shape. This would indicate that the anomalies are evenly distributed
+around the climatological mean, and for a positive uniform trend, the
+first half to be below and the other half above the climatological mean.
+In our case, there is a more complex evolution of the anomalies. Another
+distinct feature of the CUSUMs, is the different pattern of the
+cloudy-sky dataset which peaks around the middle of the period.
+Although, there seems to exist a complementary relation to the CUSUMs of
+the clear- and all-sky cases, we can not assert that clouds are the main
+driver for this relation due to the great difference in the number of
+observational data between the two datasets
+(Table~\ref{tab:trendtable}).
 
@@
 
-In order to investigate more, the features within the trends, we created
-another set of CUSUMs plots by subtracting the corresponding long term
+In order to investigate further the features of the CUSUMs, we created
+another set of CUSUM plots by subtracting the corresponding long term
 trend from the SDR anomaly data, prior to the CUSUM calculation
 (Figure~\ref{fig:cusumnotrendmonthly}). With this approach periods when
-the CUSUMs diverge from zero can be interpreted as systematic variation
-of SDR from the climatological mean. When the CUSUM is increasing, the
-added values are above the climatological values of SDR trend and vice
-versa. Overall, for all- and clear-sky conditions
-(Figure~\ref{fig:cusumnotrendmonthly-1}
+the CUSUMs diverge from zero can be interpreted as a systematic
+variation of SDR from the climatological mean. When the CUSUM is
+increasing, the added values are above the climatological values of the
+SDR trend and vice versa. Overall, for all- and clear-sky conditions
+(Figures~\ref{fig:cusumnotrendmonthly-1}
 and~\ref{fig:cusumnotrendmonthly-2}) we observe periods when the
@@
 \supplementary{The following supporting information can be downloaded
-at:\\
-\linksupplementary{s1}, Figure S1: title; Table S1: title; Video S1:
+at: \linksupplementary{s1}, Figure S1: title; Table S1: title; Video S1:
 title.}
@@
 
-\funding{Please add:
-\texttt{This\ research\ received\ no\ external\ funding\textquotesingle{}\textquotesingle{}\ or}This
-research was funded by NAME OF FUNDER grant number XXX.'\,' and and
-``The APC was funded by XXX'\,'. Check carefully that the details given
-are accurate and use the standard spelling of funding agency names at
-\url{https://search.crossref.org/funding}, any errors may affect your
-future funding.}
+\funding{This research received no external funding.}
 

### 2023-11-04 DHI_GHI_3_trends_consistency.R

@@ -1141,18 +1141,17 @@ for (adb in database) {
                               unlist(strsplit(avar, split = "_" ))[1:2]),
                             collapse = "_"))
 
-        # if (DRAFT == TRUE) {
-        #     par("mar" = c(3, 5, 2,  1))
-        # } else {
-        #     par("mar" = c(3, 5, 1, 0.5))
-        # }
+        if (DRAFT == TRUE) {
+            par("mar" = c(3, 5, 2,  1))
+        } else {
+            par("mar" = c(3, 5, 3.4, 0.5))
+        }
         par(pch = 19)
 
         ## scale fonts
         ccex <- ccex_sbs
         par(cex.lab = ccex, cex.axis = ccex, cex.main = ccex, cex = ccex)
 
-        par("mar" = c(3, 5, 1.7, 0.5))
         plot(1, type = "n",
              xlab = "",
              xlim = xlim, ylim = ylim,

### 2023-11-04 MDPI_submition.Rmd
Commit 2023-11-04 14:14

@@
 The pattern in both datasets is very similar, suggesting prevalence in clear skies over Thessaloniki.
-It is interesting that in the period 1993 - 2016 the anomalies have a high variability around zero. After 2016, the range of the variability is decreased at about one third of the prior period.
-For cloud-sky conditions (Figure\ \ref{fig:cusumnotrendmonthly-3}) the period 1997 - 2008 is dominated by positive CUSUMs suggesting a reduced effect of clouds on SDR.
-From 1997 to mid-2000s CUSUMs are increasing, likely due to a continuous decrease in the thickness of clouds, followed by a period of rapid increase (within 3 years) in cloud optical thickness lasting up to 2008.
+It is interesting that in the period 1993 - 2016 the anomalies have a high variability around zero, while after 2016, the range of the variability is decreased to about one third of the prior period.
+For cloudy-sky conditions (Figure\ \ref{fig:cusumnotrendmonthly-3}) the period 1997 - 2008 is dominated by positive CUSUMs, suggesting a reduced effect of clouds on SDR.
+From 1997 to mid-2000s CUSUMs are increasing, likely due to a continuous decrease in the optical thickness of clouds, followed by a period of rapid increase (within 3 years) in cloud optical thickness lasting up to 2008.
 The following stable period spans for about 15 years up to 2021 when CUSUMs start increasing again.
@@
 These seasonal differences indicate a possible relation of the trends in SDR to trends of clouds during winter and autumn.
-For clear-skies, the trend in winter is $`r signif(spp$"Trend [%/year]"[spp$Season == "Winter" & spp$"Sky condition" == "Clear sky"], 2)`\,\%/y$, larger than for all-skies ($`r signif(spp$"Trend [%/year]"[spp$Season == "Winter" & spp$"Sky condition" == "All sky"], 2)`\,\%/y$), which is another indication of increasing cloud thickness.
-Moreover, the trends under clear- and cloud-sky conditions are almost complementary to each other, particularly for winter and autumn, where the signal is stronger.
+For clear-skies, the trend in winter is $`r signif(spp$"Trend [%/year]"[spp$Season == "Winter" & spp$"Sky condition" == "Clear sky"], 2)`\,\%/y$, larger than for all-skies ($`r signif(spp$"Trend [%/year]"[spp$Season == "Winter" & spp$"Sky condition" == "All sky"], 2)`\,\%/y$), which is another indication of a decreasing trend in cloud optical thickness.
+Moreover, the trends under clear- and cloudy-sky conditions are almost complementary to each other, particularly for winter and autumn, where the signal is stronger.
 During spring and summer the statistical significance is very low and the actual trend too small for a meaningful comparison.
 
-```{r seasonalALL, echo=F, fig.cap="SDR trends by season of the year for the three sky conditions.
-Each point is a monthly mean calculated from daily means.
-The vertical scale is common for all graphs, each column of graphs corresponds to a sky condition and each row to a season of the year.", out.width="100%"}
+```{r seasonalALL, echo=F, fig.cap="Linear trends (black lines) of monthly mean SDR season (columns of graphs) for the three sky conditions (rows of graphs).
+The vertical scale is common for all graphs.", out.width="100%"}
 knitr::include_graphics("./images/SeasonalTrendsTogether3-2.png")
@@
 
-We demonstrate that in the period
+We have demonstrated that in the period
 `r paste(year(min(ALL_1_daily_DESEAS$Date)), " - ", year(max(ALL_1_daily_DESEAS$Date)))`, there is a brightening trend of SDR for all-sky conditions ($`r alltrend`\,\%/y$).

### 2023-11-04 MDPI_submition.tex
Commit 2023-11-04 14:14

@@
 anomalies diverge from the climatological value, each lasting for
-several years. The pattern for the all-sky and clear-sky conditions is
-very similar, suggesting prevalence in clear skies over Thessaloniki. It
-is interesting that in the period 1993 - 2016 the anomalies of all-sky
-and clear-sky, have a high variability around zero. After 2016, the
-range of the variability is decreased at about one third of the prior
-period. For cloud-sky conditions
+several years. The pattern in both datasets is very similar, suggesting
+prevalence in clear skies over Thessaloniki. It is interesting that in
+the period 1993 - 2016 the anomalies have a high variability around
+zero, while after 2016, the range of the variability is decreased to
+about one third of the prior period. For cloudy-sky conditions
 (Figure~\ref{fig:cusumnotrendmonthly-3}) the period 1997 - 2008 is
-dominated by positive CUSUMs suggesting a reduced effect of clouds on
+dominated by positive CUSUMs, suggesting a reduced effect of clouds on
 SDR. From 1997 to mid-2000s CUSUMs are increasing, likely due to a
-continuous decrease in the thickness of clouds, followed by a period of
-rapid increase (within 3 years) in cloud optical thickness lasting up to
-2008. The following stable period spans for about 15 years up to 2021
-when CUSUMs start increasing again.
+continuous decrease in the optical thickness of clouds, followed by a
+period of rapid increase (within 3 years) in cloud optical thickness
+lasting up to 2008. The following stable period spans for about 15 years
+up to 2021 when CUSUMs start increasing again.
 
@@
 is \(0.83\,\%/y\), larger than for all-skies (\(0.69\,\%/y\)), which is
-another indication of increasing cloud thickness. Moreover, the trends
-under clear- and cloud-sky conditions are almost complementary to each
-other, particularly for winter and autumn, where the signal is stronger.
-During spring and summer the statistical significance is very low and
-the actual trend too small for a meaningful comparison.
+another indication of a decreasing trend in cloud optical thickness.
+Moreover, the trends under clear- and cloudy-sky conditions are almost
+complementary to each other, particularly for winter and autumn, where
+the signal is stronger. During spring and summer the statistical
+significance is very low and the actual trend too small for a meaningful
+comparison.
 
@@
 
-\caption{SDR trends by season of the year for the three sky conditions.
-Each point is a monthly mean calculated from daily means.
-The vertical scale is common for all graphs, each column of graphs corresponds to a sky condition and each row to a season of the year.}\label{fig:seasonalALL}
+\caption{Linear trends (black lines) of monthly mean SDR season (columns of graphs) for the three sky conditions (rows of graphs). The vertical scale is common for all graphs.}\label{fig:seasonalALL}
 \end{figure}
@@
 
-We demonstrate that in the period 1993 - 2023, there is a brightening
-trend of SDR for all-sky conditions (\(0.38\,\%/y\)). On a previous
-study \citep{Bais2013}, for the period 1993 - 2011 the trend was found
-to be \(0.33\,\%/y\). The increase of this trend, indicates that the
-phenomenon and probably the causes, are still occurring. Accordingly, we
-found a similar trend for clear-sky conditions (\(0.35\,\%/y\)) that
-further supports that the brightening can be independent of the cloud
-occurrences, although, cloud formation can also attenuate SDR. We have
-calculated the effect of TSI variability to be a negligible factor in
-SDR variability. Unfortunately, there is no available data for this
-period for the aerosols, in order to directly estimate their effect on
-SDR, which is the most probable main cause. The attenuation of SDR by
-aerosols over Europe have been proposed as major factor by
-\citet{Wild2021}. The dimming effect of SDR under cloud-sky condition
-(\(-0.28\,\%/y\)), shows that clouds can also attenuate the amount of
-SDR reaching the ground, and changes on their occurrences can effect the
-net energy balance. Because we have no adequate data to investigate the
-long term changes of cloud formation in the region, we can not infer if
-the SDR trend we observe under cloud-sky is biased by changes in clouds
-and local climate. The observed brightening of SDR over Thessaloniki; is
-dependent on SZA (higher SZA, have stronger brightening trends), and
-also, with the seasons of the year, where Winter and Autumn show a
-significant strong trend, in contrast to Spring and Summer. Using CUSUMs
-of the trends for all- and clear-skies, we observed periods where the
-CUSUMs remain relative steady, with a steep decline previous to that,
-and a steep incline after. This is an indication that the whole
-brightening phenomenon does not follow a smooth development over time.
+We have demonstrated that in the period 1993 - 2023, there is a
+brightening trend of SDR for all-sky conditions (\(0.38\,\%/y\)). On a
+previous study \citep{Bais2013}, for the period 1993 - 2011 the trend
+was found to be \(0.33\,\%/y\). The increase of this trend, indicates
+that the phenomenon and probably the causes, are still occurring.
+Accordingly, we found a similar trend for clear-sky conditions
+(\(0.35\,\%/y\)) that further supports that the brightening can be
+independent of the cloud occurrences, although, cloud formation can also
+attenuate SDR. We have calculated the effect of TSI variability to be a
+negligible factor in SDR variability. Unfortunately, there is no
+available data for this period for the aerosols, in order to directly
+estimate their effect on SDR, which is the most probable main cause. The
+attenuation of SDR by aerosols over Europe have been proposed as major
+factor by \citet{Wild2021}. The dimming effect of SDR under cloud-sky
+condition (\(-0.28\,\%/y\)), shows that clouds can also attenuate the
+amount of SDR reaching the ground, and changes on their occurrences can
+effect the net energy balance. Because we have no adequate data to
+investigate the long term changes of cloud formation in the region, we
+can not infer if the SDR trend we observe under cloud-sky is biased by
+changes in clouds and local climate. The observed brightening of SDR
+over Thessaloniki; is dependent on SZA (higher SZA, have stronger
+brightening trends), and also, with the seasons of the year, where
+Winter and Autumn show a significant strong trend, in contrast to Spring
+and Summer. Using CUSUMs of the trends for all- and clear-skies, we
+observed periods where the CUSUMs remain relative steady, with a steep
+decline previous to that, and a steep incline after. This is an
+indication that the whole brightening phenomenon does not follow a
+smooth development over time.
 

### 2023-11-04 DHI_GHI_3_trends_consistency.R
Commit 2023-11-04 19:09

@@ -853,7 +853,7 @@ CLOUD_1_D_monthly_DESEAS_notrend[, DIR_transp_cusum := cumsum(tidyr::replace_na(
 
 
 
-## MONTHLY daily SUSUM plots ---------------------------------------------------
+## MONTHLY daily CUSUM plots ---------------------------------------------------
 
 #'
 #' \newpage
@@ -960,109 +960,109 @@ for (adb in database) {
 
 
 
-##;        ## test plot for reference
-##;        ## linear model
-##;        lm1 <- lm(pdb[[paste0(avar,"_des")]] ~ pdb$Date)
-##;
-##;        plot(pdb$Date, pdb[[paste0(avar,"_des")]],
-##;             ylab = bquote("Monthly Anomaly [%]"),
-##;             cex = 0.3,
-##;             col = col)
-##;        abline(h = 0, lty = 2, lwd = 0.8)
-##;        title(paste(sub("_.*","",adb), "mean monthly values ",
-##;                    translate(avar) ), cex.main = 1)
-##;        abline(lm1, lwd = 2)
-##;
-##;
-##;
-##;        ## running mean
-##;        rm <- frollmean(pdb[[paste0(avar,"_des")]], round(running_mean_window_days),
-##;                        na.rm = TRUE, algo = "exact", align = "center")
-##;        points(pdb$Date, rm, col = "red", cex = 0.5)
-##;
-##;
-##;        ## LOESS curve
-##;        vec <- !is.na(pdb[[paste0(avar,"_des")]])
-##;        FTSE.lo3 <- loess.as(pdb$Date[vec], pdb[[paste0(avar,"_des")]][vec],
-##;                             degree = 1,
-##;                             criterion = c("aicc", "gcv")[2], user.span = NULL, plot = F)
-##;        FTSE.lo.predict3 <- predict(FTSE.lo3, pdb$Date)
-##;        lines(pdb$Date, FTSE.lo.predict3, col = "cyan", lwd = 2.5)
-##;
-##;
-##;        ## decorations
-##;        fit <- lm1[[1]]
-##;        legend('top', lty = 1, bty = "n", lwd = 2, cex = 1,
-##;               paste("Trend: ",
-##;                     if (fit[2] > 0) "+" else "-",
-##;                     signif(abs(fit[2]) * Days_of_year, 3),
-##;                     "% per year")
-##;        )
-##;
-##;
-##;        cat("\n\\newpage\n")
-##;        cat("\n\\footnotesize\n")
[…]

### 2023-11-04 MDPI_submition.Rmd
Commit 2023-11-04 19:09
Commit 2023-11-04 20:14

@@
 We have demonstrated that in the period
-`r paste(year(min(ALL_1_daily_DESEAS$Date)), " - ", year(max(ALL_1_daily_DESEAS$Date)))`, there is a brightening trend of SDR for all-sky conditions ($`r alltrend`\,\%/y$).
-On a previous study [@Bais2013], for the period 1993 - 2011 the trend was found to be $0.33\,\%/y$.
+`r paste(year(min(ALL_1_daily_DESEAS$Date)), " - ", year(max(ALL_1_daily_DESEAS$Date)))`, there is positive trend in SDR of for all-sky conditions ($`r alltrend`\,\%/y$).
+Α previous study [@Bais2013], for the period 1993 - 2011 the trend was found to be $0.33\,\%/y$.
 The increase of this trend, indicates that the phenomenon and probably the causes, are still occurring.

### 2023-11-04 MDPI_submition.tex
Commit 2023-11-04 20:14

@@
 
-We have demonstrated that in the period 1993 - 2023, there is a
-brightening trend of SDR for all-sky conditions (\(0.38\,\%/y\)). On a
-previous study \citep{Bais2013}, for the period 1993 - 2011 the trend
-was found to be \(0.33\,\%/y\). The increase of this trend, indicates
-that the phenomenon and probably the causes, are still occurring.
-Accordingly, we found a similar trend for clear-sky conditions
-(\(0.35\,\%/y\)) that further supports that the brightening can be
-independent of the cloud occurrences, although, cloud formation can also
-attenuate SDR. We have calculated the effect of TSI variability to be a
-negligible factor in SDR variability. Unfortunately, there is no
-available data for this period for the aerosols, in order to directly
-estimate their effect on SDR, which is the most probable main cause. The
-attenuation of SDR by aerosols over Europe have been proposed as major
-factor by \citet{Wild2021}. The dimming effect of SDR under cloud-sky
-condition (\(-0.28\,\%/y\)), shows that clouds can also attenuate the
-amount of SDR reaching the ground, and changes on their occurrences can
-effect the net energy balance. Because we have no adequate data to
-investigate the long term changes of cloud formation in the region, we
-can not infer if the SDR trend we observe under cloud-sky is biased by
-changes in clouds and local climate. The observed brightening of SDR
-over Thessaloniki; is dependent on SZA (higher SZA, have stronger
-brightening trends), and also, with the seasons of the year, where
-Winter and Autumn show a significant strong trend, in contrast to Spring
-and Summer. Using CUSUMs of the trends for all- and clear-skies, we
-observed periods where the CUSUMs remain relative steady, with a steep
-decline previous to that, and a steep incline after. This is an
-indication that the whole brightening phenomenon does not follow a
-smooth development over time.
+We have demonstrated that in the period 1993 - 2023, there is positive
+trend in SDR of for all-sky conditions (\(0.38\,\%/y\)). Α previous
+study \citep{Bais2013}, for the period 1993 - 2011 the trend was found
+to be \(0.33\,\%/y\). The increase of this trend, indicates that the
+phenomenon and probably the causes, are still occurring. Accordingly, we
+found a similar trend for clear-sky conditions (\(0.35\,\%/y\)) that
+further supports that the brightening can be independent of the cloud
+occurrences, although, cloud formation can also attenuate SDR. We have
+calculated the effect of TSI variability to be a negligible factor in
+SDR variability. Unfortunately, there is no available data for this
+period for the aerosols, in order to directly estimate their effect on
+SDR, which is the most probable main cause. The attenuation of SDR by
+aerosols over Europe have been proposed as major factor by
+\citet{Wild2021}. The dimming effect of SDR under cloud-sky condition
+(\(-0.28\,\%/y\)), shows that clouds can also attenuate the amount of
+SDR reaching the ground, and changes on their occurrences can effect the
+net energy balance. Because we have no adequate data to investigate the
+long term changes of cloud formation in the region, we can not infer if
+the SDR trend we observe under cloud-sky is biased by changes in clouds
+and local climate. The observed brightening of SDR over Thessaloniki; is
+dependent on SZA (higher SZA, have stronger brightening trends), and
+also, with the seasons of the year, where Winter and Autumn show a
+significant strong trend, in contrast to Spring and Summer. Using CUSUMs
+of the trends for all- and clear-skies, we observed periods where the
+CUSUMs remain relative steady, with a steep decline previous to that,
+and a steep incline after. This is an indication that the whole
+brightening phenomenon does not follow a smooth development over time.
 

### 2023-11-04 DHI_GHI_3_trends_consistency.R

@@ -392,463 +392,462 @@ CLOUD_1_D_monthly_DESEAS_notrend[, DIR_transp_cusum := cumsum(tidyr::replace_na(
 
 
 
-##; ## DAILY CUSUM plots -----------------------------------------------------------
-##;
-##; ## __ Daily CUSUM --------------------------------------------------------------
-##;
-##; #'
-##; #' \newpage
-##; #' \FloatBarrier
-##; #'
-##; #' ### Whole day daily cumulative sum
-##; #'
-##; #' All data in this section are referring to daily means.
-##; #'
-##; #+ echo=F, include=T
-##;
-##; # vars        <- c("GLB_att", "DIR_att", "DIR_transp")
-##; vars        <- c("GLB_att")
-##; database    <- c(  "ALL_1_daily_DESEAS",
-##;                  "CLEAR_1_daily_DESEAS",
-##;                  "CLOUD_1_daily_DESEAS")
-##;
-##; #+ CumulativeDailyCuSum, echo=F, include=T, results="asis"
-##; for (adb in database) {
-##;     DB  <- get(adb)
-##;
-##;     cat("\n\\newpage\n")
-##;     cat("\n#### Daily cum sums for", translate(sub("_.*", "", adb)), "\n\n")
-##;
-##;     for (avar in vars) {
-##;
-##;         ## Plot regular cusums
-##;         wcare <- c("Date", paste0(avar,"_des"), paste0(avar,"_cusum"))
-##;         pdb   <- DB[, ..wcare]
-##;         rm(DB)
-##;         xlim  <- range(pdb$Date)
-##;         ylim  <- range(pdb[[paste0(avar,"_cusum")]],na.rm = T)
-##;         col   <- get(paste0(c("col",
-##;                               unlist(strsplit(avar, split = "_" ))[1:2]),
-##;                             collapse = "_"))
-##;
-##;         if (DRAFT == TRUE) {
-##;             par("mar" = c(3, 5, 2,  1))
-##;         } else {
-##;             par("mar" = c(3, 5, 0.5, 0.5))
-##;         }
-##;         par(pch = 19)
-##;
-##;         ## scale fonts
-##;         ccex <- ccex_sbs
-##;         par(cex.lab = ccex, cex.axis = ccex, cex.main = ccex, cex = ccex)
-##;
-##;         plot(1, type = "n",
-##;              xlab = "",
-##;              xlim = xlim, ylim = ylim,
-##;              xaxt = "n",
[…]

### 2023-11-05 MDPI_submition.Rmd
Commit 2023-11-04 21:08
Commit 2023-11-05 11:09
Commit 2023-11-05 12:14

@@
 `r paste(year(min(ALL_1_daily_DESEAS$Date)), " - ", year(max(ALL_1_daily_DESEAS$Date)))`, there is positive trend in SDR of for all-sky conditions ($`r alltrend`\,\%/y$).
-Α previous study [@Bais2013], for the period 1993 - 2011 the trend was found to be $0.33\,\%/y$.
+A previous study [@Bais2013] for the period 1993 - 2011 the trend was found to be $0.33\,\%/y$.
 The increase of this trend, indicates that the phenomenon and probably the causes, are still occurring.
@@
 Because we have no adequate data to investigate the long term changes of cloud formation in the region, we can not infer if the SDR trend we observe under cloud-sky is biased by changes in clouds and local climate.
-The observed brightening of SDR over Thessaloniki; is dependent on SZA (higher SZA, have stronger brightening trends), and also, with the seasons of the year, where Winter and Autumn show a significant strong trend, in contrast to Spring and Summer.
-<!-- to conclutions? -->
-Using CUSUMs of the trends for all- and clear-skies, we observed periods where the CUSUMs remain relative steady,
-<!-- roughly around the middle of the study dataset,--> with a steep decline previous to that, and a steep incline after.
-This is an indication that the whole brightening phenomenon does not follow a smooth development over time.
 
-Our findings are in agreement with other researchers for the region.
-Future observations from the collocated pyrheliometer, may provide us with the means to further investigate the variability of Solar radiation at ground level in Thessaloniki.
-Also, more data (cloud occurrence, aerosols, atmospheric water vapour, etc.), will help us to determine the extent of SDR attenuation by each factor.
+The observed brightening on SDR over Thessaloniki is dependent on SZA (larger SZAs lead to stronger brightening).
+The trend is also dependent on season, with winter showing the strongest statistically significant trend, in contrast to spring and summer.
+Our findings are in agreement with other studies for the region.
+
+$`r signif(spp$"Trend [%/year]"[spp$Season == "Winter" & spp$"Sky condition" == "All sky"], 2)`\,\%/y$
+$`r signif(spp$"Trend [%/year]"[spp$Season == "Winter" & spp$"Sky condition" == "Clear sky"], 2)`\,\%/y$
+
+Using the CUSUMs of the monthly departures for all- and clear-skies, we observed periods where the CUSUMs remain relatively stable, with a steep decline before and a steep increase after. 
+This is an indication that the whole brightening effect does not follow a smooth development over time.
+
+Continued observations with a collocated pyrheliometer, which started in 2016, will allow us to further investigate the variability of solar radiation at ground level in Thessaloniki. 
+Also, additional data of cloudiness, aerosols, atmospheric water vapour, etc., will allow better attribution and quantification of the effects of these factors on SRD.
 

### 2023-11-05 MDPI_submition.tex
Commit 2023-11-04 21:08
Commit 2023-11-05 11:09
Commit 2023-11-05 12:14

@@
 We have demonstrated that in the period 1993 - 2023, there is positive
-trend in SDR of for all-sky conditions (\(0.38\,\%/y\)). Α previous
-study \citep{Bais2013}, for the period 1993 - 2011 the trend was found
-to be \(0.33\,\%/y\). The increase of this trend, indicates that the
+trend in SDR of for all-sky conditions (\(0.38\,\%/y\)). A previous
+study \citep{Bais2013} for the period 1993 - 2011 the trend was found to
+be \(0.33\,\%/y\). The increase of this trend, indicates that the
 phenomenon and probably the causes, are still occurring. Accordingly, we
@@
 the SDR trend we observe under cloud-sky is biased by changes in clouds
-and local climate. The observed brightening of SDR over Thessaloniki; is
-dependent on SZA (higher SZA, have stronger brightening trends), and
-also, with the seasons of the year, where Winter and Autumn show a
-significant strong trend, in contrast to Spring and Summer. Using CUSUMs
-of the trends for all- and clear-skies, we observed periods where the
-CUSUMs remain relative steady, with a steep decline previous to that,
-and a steep incline after. This is an indication that the whole
-brightening phenomenon does not follow a smooth development over time.
+and local climate.
 
-Our findings are in agreement with other researchers for the region.
-Future observations from the collocated pyrheliometer, may provide us
-with the means to further investigate the variability of Solar radiation
-at ground level in Thessaloniki. Also, more data (cloud occurrence,
-aerosols, atmospheric water vapour, etc.), will help us to determine the
-extent of SDR attenuation by each factor.
+The observed brightening on SDR over Thessaloniki is dependent on SZA
+(larger SZAs lead to stronger brightening). The trend is also dependent
+on season, with winter showing the strongest statistically significant
+trend, in contrast to spring and summer. Our findings are in agreement
+with other studies for the region.
+
+\(0.69\,\%/y\) \(0.83\,\%/y\)
+
+Using the CUSUMs of the monthly departures for all- and clear-skies, we
+observed periods where the CUSUMs remain relatively stable, with a steep
+decline before and a steep increase after. This is an indication that
+the whole brightening effect does not follow a smooth development over
+time.
+
+Continued observations with a collocated pyrheliometer, which started in
+2016, will allow us to further investigate the variability of solar
+radiation at ground level in Thessaloniki. Also, additional data of
+cloudiness, aerosols, atmospheric water vapour, etc., will allow better
+attribution and quantification of the effects of these factors on SRD.
 

### 2023-11-05 MDPI_submition.Rmd
Commit 2023-11-05 13:09

@@
 The attenuation of SDR by aerosols over Europe have been proposed as major factor by @Wild2021.
-The dimming effect of SDR under cloud-sky condition ($`r cloudtrend`\,\%/y$), shows that clouds can also attenuate the amount of SDR reaching the ground, and changes on their occurrences can effect the net energy balance.
-Because we have no adequate data to investigate the long term changes of cloud formation in the region, we can not infer if the SDR trend we observe under cloud-sky is biased by changes in clouds and local climate.
+The dimming effect on SDR under cloudy-sky conditions ($`r cloudtrend`\,\%/y$),suggests that cloud optical thickness is decreasing during this period.
+Because we have no adequate data to investigate the long term changes of cloud thickness in the region, we can not verify if the negative SDR trend we observe under under cloudy-skies can be attributed solely to changes in clouds.
 
 The observed brightening on SDR over Thessaloniki is dependent on SZA (larger SZAs lead to stronger brightening).
-The trend is also dependent on season, with winter showing the strongest statistically significant trend, in contrast to spring and summer.
+The trend is also dependent on season, with winter showing the strongest statistically significant trend of
+$`r signif(spp$"Trend [%/year]"[spp$Season == "Winter" & spp$"Sky condition" == "All sky"], 2)`$
+and
+$`r signif(spp$"Trend [%/year]"[spp$Season == "Winter" & spp$"Sky condition" == "Clear sky"], 2)`\,\%/y$
+for all- and clear-skies, respectively, in contrast to spring and summer.
+The trends for autumn are also significant but smaller (
+$`r signif(spp$"Trend [%/year]"[spp$Season == "Autumn" & spp$"Sky condition" == "All sky"], 2)`$
+and
+$`r signif(spp$"Trend [%/year]"[spp$Season == "Autumn" & spp$"Sky condition" == "Clear sky"], 2)`\,\%/y$
+for all- and clear-skies, respectively).
 Our findings are in agreement with other studies for the region.
-
-$`r signif(spp$"Trend [%/year]"[spp$Season == "Winter" & spp$"Sky condition" == "All sky"], 2)`\,\%/y$
-$`r signif(spp$"Trend [%/year]"[spp$Season == "Winter" & spp$"Sky condition" == "Clear sky"], 2)`\,\%/y$
 

### 2023-11-05 MDPI_submition.tex
Commit 2023-11-05 13:09

@@
 aerosols over Europe have been proposed as major factor by
-\citet{Wild2021}. The dimming effect of SDR under cloud-sky condition
-(\(-0.28\,\%/y\)), shows that clouds can also attenuate the amount of
-SDR reaching the ground, and changes on their occurrences can effect the
-net energy balance. Because we have no adequate data to investigate the
-long term changes of cloud formation in the region, we can not infer if
-the SDR trend we observe under cloud-sky is biased by changes in clouds
-and local climate.
+\citet{Wild2021}. The dimming effect on SDR under cloudy-sky conditions
+(\(-0.28\,\%/y\)),suggests that cloud optical thickness is decreasing
+during this period. Because we have no adequate data to investigate the
+long term changes of cloud thickness in the region, we can not verify if
+the negative SDR trend we observe under under cloudy-skies can be
+attributed solely to changes in clouds.
 
@@
 on season, with winter showing the strongest statistically significant
-trend, in contrast to spring and summer. Our findings are in agreement
-with other studies for the region.
-
-\(0.69\,\%/y\) \(0.83\,\%/y\)
+trend of \(0.69\) and \(0.83\,\%/y\) for all- and clear-skies,
+respectively, in contrast to spring and summer. The trends for autumn
+are also significant but smaller ( \(0.43\) and \(0.37\,\%/y\) for all-
+and clear-skies, respectively). Our findings are in agreement with other
+studies for the region.
 

### 2023-11-05 DHI_GHI_1_longterm_trends.R

@@ -1076,13 +1076,13 @@ for (avar in vars) {
             plot.new()
             text(x = 0.05, y = 0.5,
                  adj  = c(0.5,0.5),
-                 srt  = 90, "Summer", cex = 0.9, font = 2)
+                 srt  = 90, "Spring", cex = 0.9, font = 2)
         }
         if (i == 15) {
             plot.new()
             text(x = 0.25, y = 0.5,
                  adj  = c(0.5,0.5),
-                 srt  = 90, "Summer", cex = 0.9, font = 2)
+                 srt  = 90, "Spring", cex = 0.9, font = 2)
         }
 
 
@@ -1090,13 +1090,13 @@ for (avar in vars) {
             plot.new()
             text(x = 0.05, y = 0.5,
                  adj  = c(0.5,0.5),
-                 srt  = 90, "Autumn", cex = 0.9, font = 2)
+                 srt  = 90, "Summer", cex = 0.9, font = 2)
         }
         if (i == 20) {
             plot.new()
             text(x = 0.25, y = 0.5,
                  adj  = c(0.5,0.5),
-                 srt  = 90, "Autumn", cex = 0.9, font = 2)
+                 srt  = 90, "Summer", cex = 0.9, font = 2)
         }
 
 
@@ -1104,13 +1104,13 @@ for (avar in vars) {
             plot.new()
             text(x = 0.05, y = 0.5,
                  adj  = c(0.5,0.5),
-                 srt  = 90, "Spring", cex = 0.9, font = 2)
+                 srt  = 90, "Automn", cex = 0.9, font = 2)
         }
         if (i == 25) {
             plot.new()
             text(x = 0.25, y = 0.5,
                  adj  = c(0.5,0.5),
-                 srt  = 90, "Spring", cex = 0.9, font = 2)
+                 srt  = 90, "Automn", cex = 0.9, font = 2)
         }
 
 

### 2023-11-05 MDPI_submition.Rmd
Commit 2023-11-05 19:09

@@
       Campus Box 149, 54124 Thessaloniki, Greece
-    email: natsisphysicist@gmail.com
+    email: natsisphysicist@gmail.com (A.N.); abais@auth.gr (A.B.)
   - num: 2

### 2023-11-05 MDPI_submition.tex
Commit 2023-11-05 19:09

@@
 Atmospheric Physics, Campus Box 149, 54124 Thessaloniki,
-Greece; \href{mailto:natsisphysicist@gmail.com}{\nolinkurl{natsisphysicist@gmail.com}}\\
+Greece; \href{mailto:natsisphysicist@gmail.com}{\nolinkurl{natsisphysicist@gmail.com}}
+(A.N.); \href{mailto:abais@auth.gr}{\nolinkurl{abais@auth.gr}} (A.B.)\\
 $^{2}$ \quad Your department Street, City, Country; \\

### 2023-11-05 MDPI_submition.Rmd
Commit 2023-11-05 20:14

@@
 # front matter
-simplesummary: |
-  A Simple summary goes here.
+# simplesummary: |
+#   A Simple summary goes here.
 
@@
 keywords: |
-  GHI; SDR; Solar Brigthening/Dimming.
+  GHI; SDR; Solar Brigthening/Dimming, aerosols, clouds.
 ## list three to ten pertinent keywords specific to the article, yet reasonably common within the subject discipline
@@
 We have demonstrated that in the period
-`r paste(year(min(ALL_1_daily_DESEAS$Date)), " - ", year(max(ALL_1_daily_DESEAS$Date)))`, there is positive trend in SDR of for all-sky conditions ($`r alltrend`\,\%/y$).
-A previous study [@Bais2013] for the period 1993 - 2011 the trend was found to be $0.33\,\%/y$.
-The increase of this trend, indicates that the phenomenon and probably the causes, are still occurring.
-Accordingly, we found a similar trend for clear-sky conditions ($`r cleartrend`\,\%/y$) that further supports that the brightening can be independent of the cloud occurrences, although, cloud formation can also attenuate SDR.
-We have calculated the effect of TSI variability to be a negligible factor in SDR variability.
-Unfortunately, there is no available data for this period for the aerosols, in order to directly estimate their effect on SDR, which is the most probable main cause.
+`r paste(year(min(ALL_1_daily_DESEAS$Date)), " - ", year(max(ALL_1_daily_DESEAS$Date)))`, there is a positive trend in SDR of ($`r alltrend`\,\%/y$) (brightening) (positive trend) in Thessaloniki, Greece, under all-sky conditions.
+A previous study [@Bais2013] for the period 1993 - 2011 found also a positive trend of $0.33\,\%/y$.
+The increase of this trend indicates that the brightening of SDR continues and is probably caused by continuing decreases in aerosol optical depth and the optical thickness of clouds over the area.
+Moreover, we found a similar trend under clear-sky conditions ($`r cleartrend`\,\%/y$) that further supports the assumption that the brightening is caused mainly by decreasing aerosols.
+Unfortunately, for the entire period there is no available data for the aerosols, in order to quantify their effect on SDR. 
+However, @Siomos2020 have shown that aerosol optical depth over Thessaloniki is decreasing constantly at least up to 2018.
 The attenuation of SDR by aerosols over Europe have been proposed as major factor by @Wild2021.
-The dimming effect on SDR under cloudy-sky conditions ($`r cloudtrend`\,\%/y$),suggests that cloud optical thickness is decreasing during this period.
-Because we have no adequate data to investigate the long term changes of cloud thickness in the region, we can not verify if the negative SDR trend we observe under under cloudy-skies can be attributed solely to changes in clouds.
+The dimming effect on SDR under cloudy-sky conditions ($`r cloudtrend`\,\%/y$), suggests that cloud optical thickness is decreasing during this period.
+Because we have no adequate data to investigate the long term changes of cloud thickness in the region, we cannot verify if the negative SDR trend we observe under under cloudy-skies can be attributed solely to changes in clouds.
 

### 2023-11-05 MDPI_submition.tex
Commit 2023-11-05 20:14

@@
 % Simple summary
-\simplesumm{A Simple summary goes here.}
 
@@
 % Keywords
-\keyword{GHI; SDR; Solar Brigthening/Dimming.}
+\keyword{GHI; SDR; Solar Brigthening/Dimming, aerosols, clouds.}
 
@@
 
-We have demonstrated that in the period 1993 - 2023, there is positive
-trend in SDR of for all-sky conditions (\(0.38\,\%/y\)). A previous
-study \citep{Bais2013} for the period 1993 - 2011 the trend was found to
-be \(0.33\,\%/y\). The increase of this trend, indicates that the
-phenomenon and probably the causes, are still occurring. Accordingly, we
-found a similar trend for clear-sky conditions (\(0.35\,\%/y\)) that
-further supports that the brightening can be independent of the cloud
-occurrences, although, cloud formation can also attenuate SDR. We have
-calculated the effect of TSI variability to be a negligible factor in
-SDR variability. Unfortunately, there is no available data for this
-period for the aerosols, in order to directly estimate their effect on
-SDR, which is the most probable main cause. The attenuation of SDR by
-aerosols over Europe have been proposed as major factor by
-\citet{Wild2021}. The dimming effect on SDR under cloudy-sky conditions
-(\(-0.28\,\%/y\)),suggests that cloud optical thickness is decreasing
-during this period. Because we have no adequate data to investigate the
-long term changes of cloud thickness in the region, we can not verify if
-the negative SDR trend we observe under under cloudy-skies can be
-attributed solely to changes in clouds.
+We have demonstrated that in the period 1993 - 2023, there is a positive
+trend in SDR of (\(0.38\,\%/y\)) (brightening) (positive trend) in
+Thessaloniki, Greece, under all-sky conditions. A previous study
+\citep{Bais2013} for the period 1993 - 2011 found also a positive trend
+of \(0.33\,\%/y\). The increase of this trend indicates that the
+brightening of SDR continues and is probably caused by continuing
+decreases in aerosol optical depth and the optical thickness of clouds
+over the area. Moreover, we found a similar trend under clear-sky
+conditions (\(0.35\,\%/y\)) that further supports the assumption that
+the brightening is caused mainly by decreasing aerosols. Unfortunately,
+for the entire period there is no available data for the aerosols, in
+order to quantify their effect on SDR. However, \citet{Siomos2020} have
+shown that aerosol optical depth over Thessaloniki is decreasing
+constantly at least up to 2018. The attenuation of SDR by aerosols over
+Europe have been proposed as major factor by \citet{Wild2021}. The
+dimming effect on SDR under cloudy-sky conditions (\(-0.28\,\%/y\)),
+suggests that cloud optical thickness is decreasing during this period.
+Because we have no adequate data to investigate the long term changes of
+cloud thickness in the region, we cannot verify if the negative SDR
+trend we observe under under cloudy-skies can be attributed solely to
+changes in clouds.
 

### 2023-11-05 MDPI_submission.tex
Commit 2023-11-05 21:09

--- 2023-11-05 MDPI_submition.tex
@@
 Natsis$^{1,*}$\href{https://orcid.org/0000-0002-5199-4119}
-{\orcidicon}, Alkiviadis Bais$^{2,}$}
+{\orcidicon}, Alkiviadis Bais$^{1}$}
 
@@
 (A.N.); \href{mailto:abais@auth.gr}{\nolinkurl{abais@auth.gr}} (A.B.)\\
-$^{2}$ \quad Your department Street, City, Country; \\
 }
@@
 is generally used to describe periods of positive SDR trend, and
-`dimming' for negative trend. There are many cases in the long term
-records of irradiance, showing a systematic change in the magnitude of
-the trend, occurring roughly in the last decades of the 20th century. On
-multiple stations in China, a dimming period was reported until about
-2000, followed by a brightening period \citep{Yang2021}. A similar
-pattern was identified, with the breaking point around 1980, for
-stations in Central Europe \citep{Wild2021} and Brazil
-\citep{Yamasoe2021}. On global scale, an artificial Intelligence aided
-spatial analysis on continental level with data from multiple stations
-reach similar conclusions for these regions and for the global trend
-\citep{Yuan2021}.
+`dimming' for negative trend \citep{Wild2009}. There are many cases in
+the long term records of irradiance, showing a systematic change in the
+magnitude of the trend, occurring roughly in the last decades of the
+20th century. On multiple stations in China, a dimming period was
+reported until about 2000, followed by a brightening period
+\citep{Yang2021}. A similar pattern was identified, with the breaking
+point around 1980, for stations in Central Europe \citep{Wild2021} and
+Brazil \citep{Yamasoe2021}. On global scale, an artificial Intelligence
+aided spatial analysis on continental level with data from multiple
+stations reach similar conclusions for these regions and for the global
+trend \citep{Yuan2021}.
 
@@
 the work of \citet{Brent1973}, from the library ``stats'' of the R
-programming language \citep{RCT2023}. The optimization and the selection
-of the clear sky reference model, was performed on SDR observations for
-the period 2016 - 2021. During the optimization, eight simple clear sky
-radiation models were tested (namely, Daneshyar-Paltridge-Proctor,
-Kasten-Czeplak, Haurwitz, Berger-Duffie, Adnot-Bourges-Campana-Gicquel,
-Robledo-Soler, Kasten and Ineichen-Perez), with a wide range of factors.
-These models are described in more details by \citet{Reno2012} and
-evaluated by \citet{Reno2016}. We found, that Haurwitz's model, adjusted
-with the factor \(a = 0.965\) yields one of the lowest root mean squared
-errors (RMSE), while the procedure, manages to characterize the majority
-of the data. Thus, our clear sky reference is derived by the
-Eq.~\ref{eq:ahau}. \begin{equation}
+programming language \citep{RCT2023}. \begin{equation}
 f(a) = \frac{1}{n}\sum_{i=1}^{n} ( \text{SDR}_{\text{CSid},i} - a \times \text{SDR}_{\text{testCSref},i} )^2 \label{eq:minf}
@@
 \(\text{SDR}_{\text{testCSref},i}\) is any of the tested clear sky
-radiation models. \begin{equation}
+radiation models. The optimization and the selection of the clear sky
+reference model, was performed on SDR observations for the period 2016 -
+2021. During the optimization, eight simple clear sky radiation models
+were tested (namely, Daneshyar-Paltridge-Proctor, Kasten-Czeplak,
+Haurwitz, Berger-Duffie, Adnot-Bourges-Campana-Gicquel, Robledo-Soler,
+Kasten and Ineichen-Perez), with a wide range of factors. These models
+are described in more details by \citet{Reno2012} and evaluated by
+\citet{Reno2016}. We found, that Haurwitz's model, adjusted with the
+factor \(a = 0.965\) yields one of the lowest root mean squared errors
+(RMSE), while the procedure, manages to characterize the majority of the
+data. Thus, our clear sky reference is derived by the Eq.~\ref{eq:ahau}.
+\begin{equation}
 \text{SDR}_\text{CSref} = a \times \text{SDR}_\text{Haurwitz} = 0.965 \times 1098 \times \cos(\theta) \times \exp \left( \frac{ - 0.057}{\cos(\theta)} \right) \label{eq:ahau}
@@
 
-\begin{center}\rule{0.5\linewidth}{0.5pt}\end{center}
-
-\hypertarget{discussion}{%
-\section{Discussion}\label{discussion}}
-
-Authors should discuss the results and how they can be interpreted in
-perspective of previous studies and of the working hypotheses. The
-findings and their implications should be discussed in the broadest
-context possible. Future research directions may also be highlighted.
-
-\hypertarget{conclusion}{%
-\section{Conclusion}\label{conclusion}}
-
-This section is not mandatory, but can be added to the manuscript if the
-discussion is unusually long or complex.
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %% optional
-\supplementary{The following supporting information can be downloaded
-at: \linksupplementary{s1}, Figure S1: title; Table S1: title; Video S1:
-title.}
 

### 2023-11-06 Build_159/MDPI_submission_B159.Rmd
Commit 2023-11-06 09:11

--- 2023-11-05 MDPI_submition.Rmd
@@
   - name: Alkiviadis Bais
-    affil: 2, #\dagger, # \ddagger
+    affil: 1 #, \dagger, # \ddagger
 affiliation:
@@
     email: natsisphysicist@gmail.com (A.N.); abais@auth.gr (A.B.)
-  - num: 2
-    address: |
-      Your department
-      Street, City, Country
+  # - num: 2
+  #   address: |
+  #     Your department
+  #     Street, City, Country
 #     email: abais@auth.gr
@@
 
-supplementary: |
- The following supporting information can be downloaded at:
- \linksupplementary{s1}, Figure S1: title; Table S1: title; Video S1: title.
+# supplementary: |
+#  The following supporting information can be downloaded at:
+#  \linksupplementary{s1}, Figure S1: title; Table S1: title; Video S1: title.
 
@@
 Studies of SDR variability, have identified some distinct SDR trends on different regions of the world on different time periods.
-The term 'brightening' is generally used to describe periods of positive SDR trend, and 'dimming' for negative trend.
+The term 'brightening' is generally used to describe periods of positive SDR trend, and 'dimming' for negative trend [@Wild2009].
 There are many cases in the long term records of irradiance, showing a systematic change in the magnitude of the trend, occurring roughly in the last decades of the 20th century.
@@
 The target of the optimization was the minimization of a function $f(a)$ (Eq.\ \ref{eq:minf}) and was accomplished with the algorithmic function "optimise", which is an implementation based on the work of @Brent1973, from the library "stats" of the R programming language [@RCT2023].
+<!-- cost function -->
+\begin{equation}
+f(a) = \frac{1}{n}\sum_{i=1}^{n} ( \text{SDR}_{\text{CSid},i} - a \times \text{SDR}_{\text{testCSref},i} )^2 \label{eq:minf}
+\end{equation}
+where: $n$ is the total number of daylight records, $\text{SDR}_{\text{CSid},i}$ are the records identified as clear sky by CSid, $a$ is a hypothetical adjustment factor, and $\text{SDR}_{\text{testCSref},i}$ is any of the tested clear sky radiation models.
 The optimization and the selection of the clear sky reference model, was performed on SDR observations for the period 2016 - 2021.
@@
 Thus, our clear sky reference is derived by the Eq.\ \ref{eq:ahau}.
-<!-- cost fuction -->
-\begin{equation}
-f(a) = \frac{1}{n}\sum_{i=1}^{n} ( \text{SDR}_{\text{CSid},i} - a \times \text{SDR}_{\text{testCSref},i} )^2 \label{eq:minf}
-\end{equation}
-where: $n$ is the total number of daylight records, $\text{SDR}_{\text{CSid},i}$ are the records identified as clear sky by CSid, $a$ is a hypothetical adjustment factor, and $\text{SDR}_{\text{testCSref},i}$ is any of the tested clear sky radiation models.
 <!-- adjusted Haurwitz -->
@@
 
---------------------
+<!-- ------------------------------------------------------------ -->
 
@@
 
-# Discussion
+<!-- # Discussion -->
 
-Authors should discuss the results and how they can be interpreted in
-perspective of previous studies and of the working hypotheses. The findings and
-their implications should be discussed in the broadest context possible. Future
-research directions may also be highlighted.
+<!-- Authors should discuss the results and how they can be interpreted in -->
+<!-- perspective of previous studies and of the working hypotheses. The findings and -->
+<!-- their implications should be discussed in the broadest context possible. Future -->
+<!-- research directions may also be highlighted. -->
 
-# Conclusion
+<!-- # Conclusion -->
 
-This section is not mandatory, but can be added to the manuscript if the
-discussion is unusually long or complex.
+<!-- This section is not mandatory, but can be added to the manuscript if the -->
+<!-- discussion is unusually long or complex. -->

### 2023-11-08 MDPI_submission.Rmd
Commit 2023-11-08 12:12

--- 2023-11-06 Build_159/MDPI_submission_B159.Rmd
@@
 ---
-title:     "Trends of SDR in Thessaloniki"
+title:     "Trends from 30-year observations of downward solar irradiance in Thessaloniki, Greece"
 author:
@@
   Periods of positive trends are referred as brightening periods and of negative trends as dimming periods.
-  Here we studied 29 years of pyranometer data in Thessaloniki, Greece, under three types of sky conditions (clear sky, cloudy sky and all sky).
+  Here we studied 30 years of pyranometer data in Thessaloniki, Greece, under three types of sky conditions (clear sky, cloudy sky and all sky).
   The clear-sky data were identified by applying a cloud screening algorithm.
@@
 where: $n$ is the total number of daylight records, $\text{SDR}_{\text{CSid},i}$ are the records identified as clear sky by CSid, $a$ is a hypothetical adjustment factor, and $\text{SDR}_{\text{testCSref},i}$ is any of the tested clear sky radiation models.
+
 The optimization and the selection of the clear sky reference model, was performed on SDR observations for the period 2016 - 2021.
@@
 
-```{r cusummonth, echo = FALSE, fig.cap='Running cumulative sum plots of the monthly SDR.', fig.subcap=c('All-skies.', 'Clear-skies.', 'Cloud-skies.'), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3}
+```{r cusummonth, echo = FALSE, fig.cap='Running cumulative sum plots of the monthly SDR.', fig.subcap=c('All skies.', 'Clear skies.', 'Cloudy skies.'), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3}
 knitr::include_graphics(c(
@@
 
-```{r cusumnotrendmonthly, echo = FALSE, fig.cap='Running cumulative sum plots of the monthly SDR after removing the long term trend from the monthly values.', fig.subcap=c('All-skies.', 'Clear-skies.', 'Cloud-skies.'), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3}
+```{r cusumnotrendmonthly, echo = FALSE, fig.cap='Running cumulative sum plots of the monthly SDR after removing the long term trend from the monthly values.', fig.subcap=c('All skies.', 'Clear skies.', 'Cloudy skies.'), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3}
 knitr::include_graphics(c(
@@
 ```{r szatrends, echo = FALSE, fig.cap='Distribution of the SDR\'s long term trends by SZA before and after noon.
-The solid data points represent values with an acceptable statistical significance ($p<0.005$), the trends were computed with SZA bins of $1^\\circ$ for each day, before and after local noon.', fig.subcap=c('All-sky conditions.', 'Clear-sky conditions.', 'Cloud-sky conditions.'), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3}
+The solid data points represent values with an acceptable statistical significance ($p<0.005$), the trends were computed with SZA bins of $1^\\circ$ for each day, before and after local noon.', fig.subcap=c('All skies.', 'Clear skies.', 'Cloudy skies.'), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3}
 knitr::include_graphics("./images/SzaTrends-1.png")

### 2023-11-08 MDPI_submission.tex
Commit 2023-11-08 12:12

@@
 % Full title of the paper (Capitalized)
-\Title{Trends of SDR in Thessaloniki}
+\Title{Trends from 30-year observations of downward solar irradiance in
+Thessaloniki, Greece}
 
 % MDPI internal command: Title for citation in the left column
-\TitleCitation{Trends of SDR in Thessaloniki}
+\TitleCitation{Trends from 30-year observations of downward solar
+irradiance in Thessaloniki, Greece}
 
@@
 scales. Periods of positive trends are referred as brightening periods
-and of negative trends as dimming periods. Here we studied 29 years of
+and of negative trends as dimming periods. Here we studied 30 years of
 pyranometer data in Thessaloniki, Greece, under three types of sky
@@
 \(\text{SDR}_{\text{testCSref},i}\) is any of the tested clear sky
-radiation models. The optimization and the selection of the clear sky
-reference model, was performed on SDR observations for the period 2016 -
-2021. During the optimization, eight simple clear sky radiation models
-were tested (namely, Daneshyar-Paltridge-Proctor, Kasten-Czeplak,
-Haurwitz, Berger-Duffie, Adnot-Bourges-Campana-Gicquel, Robledo-Soler,
-Kasten and Ineichen-Perez), with a wide range of factors. These models
-are described in more details by \citet{Reno2012} and evaluated by
+radiation models.
+
+The optimization and the selection of the clear sky reference model, was
+performed on SDR observations for the period 2016 - 2021. During the
+optimization, eight simple clear sky radiation models were tested
+(namely, Daneshyar-Paltridge-Proctor, Kasten-Czeplak, Haurwitz,
+Berger-Duffie, Adnot-Bourges-Campana-Gicquel, Robledo-Soler, Kasten and
+Ineichen-Perez), with a wide range of factors. These models are
+described in more details by \citet{Reno2012} and evaluated by
 \citet{Reno2016}. We found, that Haurwitz's model, adjusted with the
@@
 
-{\centering \subfloat[All-skies.\label{fig:cusummonth-1}]{\includegraphics[width=.32\linewidth]{./images/CumulativeMonthlyCuSum-1} }\subfloat[Clear-skies.\label{fig:cusummonth-2}]{\includegraphics[width=.32\linewidth]{./images/CumulativeMonthlyCuSum-5} }\subfloat[Cloud-skies.\label{fig:cusummonth-3}]{\includegraphics[width=.32\linewidth]{./images/CumulativeMonthlyCuSum-9} }
+{\centering \subfloat[All skies.\label{fig:cusummonth-1}]{\includegraphics[width=.32\linewidth]{./images/CumulativeMonthlyCuSum-1} }\subfloat[Clear skies.\label{fig:cusummonth-2}]{\includegraphics[width=.32\linewidth]{./images/CumulativeMonthlyCuSum-5} }\subfloat[Cloudy skies.\label{fig:cusummonth-3}]{\includegraphics[width=.32\linewidth]{./images/CumulativeMonthlyCuSum-9} }
 
@@
 
-{\centering \subfloat[All-skies.\label{fig:cusumnotrendmonthly-1}]{\includegraphics[width=.32\linewidth]{./images/CumulativeMonthlyCuSumNOtrend-1} }\subfloat[Clear-skies.\label{fig:cusumnotrendmonthly-2}]{\includegraphics[width=.32\linewidth]{./images/CumulativeMonthlyCuSumNOtrend-5} }\subfloat[Cloud-skies.\label{fig:cusumnotrendmonthly-3}]{\includegraphics[width=.32\linewidth]{./images/CumulativeMonthlyCuSumNOtrend-9} }
+{\centering \subfloat[All skies.\label{fig:cusumnotrendmonthly-1}]{\includegraphics[width=.32\linewidth]{./images/CumulativeMonthlyCuSumNOtrend-1} }\subfloat[Clear skies.\label{fig:cusumnotrendmonthly-2}]{\includegraphics[width=.32\linewidth]{./images/CumulativeMonthlyCuSumNOtrend-5} }\subfloat[Cloudy skies.\label{fig:cusumnotrendmonthly-3}]{\includegraphics[width=.32\linewidth]{./images/CumulativeMonthlyCuSumNOtrend-9} }
 
@@
 
-{\centering \subfloat[All-sky conditions.\label{fig:szatrends-1}]{\includegraphics[width=.32\linewidth]{./images/SzaTrends-1} }\subfloat[Clear-sky conditions.\label{fig:szatrends-2}]{\includegraphics[width=.32\linewidth]{./images/SzaTrends-4} }\subfloat[Cloud-sky conditions.\label{fig:szatrends-3}]{\includegraphics[width=.32\linewidth]{./images/SzaTrends-7} }
+{\centering \subfloat[All skies.\label{fig:szatrends-1}]{\includegraphics[width=.32\linewidth]{./images/SzaTrends-1} }\subfloat[Clear skies.\label{fig:szatrends-2}]{\includegraphics[width=.32\linewidth]{./images/SzaTrends-4} }\subfloat[Cloudy skies.\label{fig:szatrends-3}]{\includegraphics[width=.32\linewidth]{./images/SzaTrends-7} }
 

### 2023-11-08 MDPI_submission.Rmd
Commit 2023-11-08 13:18

@@
 <!-- goals -->
-We reevaluated and extended the dataset used by @Bais2013, we applied a different algorithm for the identification of clear-/cloud-sky instances [@Reno2016; @Reno2012a], and we derived the SDR trends for the period `r year(min(ALL_1_daily_DESEAS$Date))` to `r year(max(ALL_1_daily_DESEAS$Date))`, under different sky conditions (all-sky, clear-sky and cloud-sky).
+We reevaluated and extended the dataset used by @Bais2013, we applied a different algorithm for the identification of clear-/cloud-sky instances [@Reno2016; @Reno2012a], and we derived the SDR trends for the period of study, under different sky conditions (all-sky, clear-sky and cloud-sky).
 Finally, we investigated the dependence of the trends on solar zenith angle and season.
@@
 ```{r trendtable, echo=F}
+LT2$`Sky conditions` <- sub(" sky", " skies", LT2$`Sky conditions`)
 x_latex <- knitr::kable(LT2,
@@
 
-```{r trendALL, echo=F, out.width='.70\\linewidth', fig.cap=paste0("Anomaly (\\%) of the daily all-sky SDR, relative to climatological values for ", year(min(ALL_1_daily_DESEAS$Date)), " - ", year(max(ALL_1_daily_DESEAS$Date)), ".
+```{r trendALL, echo=F, out.width='.70\\linewidth', fig.cap=paste0("Anomalies (\\%) of the daily all-sky SDR, relative to climatological values for ", year(min(ALL_1_daily_DESEAS$Date)), " - ", year(max(ALL_1_daily_DESEAS$Date)), ".
 The black line shows the long term linear trend.")}
@@
       )]
+spp2 <- spp
+## rename conditions
+spp2$`Sky condition` <- sub(" sky", " skies", spp2$`Sky condition`)
 
-rownames(spp) <- NULL
+rownames(spp2) <- NULL
 # panderOptions("table.alignment.default", "left")
@@
 ## use kable
-x_latex <- knitr::kable(spp,
+x_latex <- knitr::kable(spp2,
                         format      = "latex",

### 2023-11-08 MDPI_submission.tex
Commit 2023-11-08 13:18

@@
 \citep{Reno2016, Reno2012a}, and we derived the SDR trends for the
-period 1993 to 2023, under different sky conditions (all-sky, clear-sky
-and cloud-sky). Finally, we investigated the dependence of the trends on
+period of study, under different sky conditions (all-sky, clear-sky and
+cloud-sky). Finally, we investigated the dependence of the trends on
 solar zenith angle and season.
@@
 \midrule
-All sky & 0.376 & 100.00 & 10256\\
-Clear sky & 0.349 & 100.00 & 10256\\
-Cloudy sky & -0.276 & 99.97 & 5067\\
+All skies & 0.376 & 100.00 & 10256\\
+Clear skies & 0.349 & 100.00 & 10256\\
+Cloudy skies & -0.276 & 99.97 & 5067\\
 \bottomrule
@@
 
-\caption{Anomaly (\%) of the daily all-sky SDR, relative to climatological values for 1993 - 2023. The black line shows the long term linear trend.}\label{fig:trendALL}
+\caption{Anomalies (\%) of the daily all-sky SDR, relative to climatological values for 1993 - 2023. The black line shows the long term linear trend.}\label{fig:trendALL}
 \end{figure}
@@
 
-\multirow{-4}{*}{\centering\arraybackslash All sky} & Autumn & 0.4310 & 99.3\\
+\multirow{-4}{*}{\centering\arraybackslash All skies} & Autumn & 0.4310 & 99.3\\
 \cmidrule{1-4}
@@
 
-\multirow{-4}{*}{\centering\arraybackslash Clear sky} & Autumn & 0.3670 & 97.2\\
+\multirow{-4}{*}{\centering\arraybackslash Clear skies} & Autumn & 0.3670 & 97.2\\
 \cmidrule{1-4}
@@
 
-\multirow{-4}{*}{\centering\arraybackslash Cloudy sky} & Autumn & -0.4000 & 89.5\\
+\multirow{-4}{*}{\centering\arraybackslash Cloudy skies} & Autumn & -0.4000 & 89.5\\
 \bottomrule

### 2023-11-08 appendix.tex
Commit 2023-11-05 21:09
Commit 2023-11-08 13:18

@@
 %% optional
-\appendixtitles{yes} %Leave argument "no" if all appendix headings stay EMPTY (then no dot is printed after "Appendix A"). If the appendix sections contain a heading then change the argument to "yes".
+\appendixtitles{no} %Leave argument "no" if all appendix headings stay EMPTY (then no dot is printed after "Appendix A"). If the appendix sections contain a heading then change the argument to "yes".
 \appendixsections{one} %Leave argument "multiple" if there are multiple sections.
@@
 
-\caption{Anomaly (\%) of the daily clear-sky SDR, relative to climatological values for1993 - 2023. The black line shows the long term linear trend for clear-sky conditions.}\label{fig:trendCLEAR}
+\caption{Anomalies (\%) of the daily clear-sky SDR, relative to climatological values for1993 - 2023. The black line shows the long term linear trend for clear-sky conditions.}\label{fig:trendCLEAR}
 \end{figure}
@@
 
-\caption{Anomaly (\%) of the daily cloud-sky SDR, relative to climatological values for 1993 - 2023. The black line shows the long term linear trend for cloud-sky conditions.}\label{fig:trendCLOUD}
+\caption{Anomalies (\%) of the daily cloud-sky SDR, relative to climatological values for 1993 - 2023. The black line shows the long term linear trend for cloud-sky conditions.}\label{fig:trendCLOUD}
 \end{figure}

### 2023-11-08 DHI_GHI_1_longterm_trends.R

@@ -270,7 +270,7 @@ for (DBn in dbs) {
                  cex.main = 0.8,
                  yaxt     = "n",
                  xlab     = "",
-                 ylab     = bquote("Anomaly [%]")
+                 ylab     = bquote("Anomalies [%]")
             )
             # y axis
             axis(2, pretty(dataset[[avar]]), las = 2 )
@@ -383,7 +383,7 @@ for (DBn in dbs) {
                  cex.main = 0.8,
                  yaxt     = "n",
                  xlab     = "",
-                 ylab     = bquote("Anomaly [%]")
+                 ylab     = bquote("Anomalies [%]")
             )
             axis(2, pretty(dataset[[avar]]), las = 2 )
 

### 2023-11-08 MDPI_submission.Rmd
Commit 2023-11-08 15:20

@@
   We have also investigated the consistency of these trends, the effect of the solar zenith angle, and their seasonal variability.
-  Finally, we identified some anomalies in the long term SDR trends for all sky conditions.
+  Finally, we identified some anomalies in the long term SDR trends for all sky conditions by examining the cumulative sums of monthly anomalies from the climatological mean.
 
@@
 keywords: |
-  GHI; SDR; Solar Brigthening/Dimming, aerosols, clouds.
+  GHI; SDR; solar radiation; Solar Brigthening/Dimming; aerosols; clouds.
 ## list three to ten pertinent keywords specific to the article, yet reasonably common within the subject discipline
@@
 
-```{r cusummonth, echo = FALSE, fig.cap='Running cumulative sum plots of the monthly SDR.', fig.subcap=c('All skies.', 'Clear skies.', 'Cloudy skies.'), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3}
+```{r cusummonth, echo = FALSE, fig.cap='Cumulative sum plots of the monthly SDR anomalies in (%) for different sky conditions.', fig.subcap=c('All skies.', 'Clear skies.', 'Cloudy skies.'), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3}
 knitr::include_graphics(c(
@@
 
-```{r cusumnotrendmonthly, echo = FALSE, fig.cap='Running cumulative sum plots of the monthly SDR after removing the long term trend from the monthly values.', fig.subcap=c('All skies.', 'Clear skies.', 'Cloudy skies.'), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3}
+```{r cusumnotrendmonthly, echo = FALSE, fig.cap='Cumulative sum plots of monthly SDR anomalies in (%) for different sky conditions after removing the long-term linear trend.', fig.subcap=c('All skies.', 'Clear skies.', 'Cloudy skies.'), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3}
 knitr::include_graphics(c(
@@
 
-```{r szatrends, echo = FALSE, fig.cap='Distribution of the SDR\'s long term trends by SZA before and after noon.
-The solid data points represent values with an acceptable statistical significance ($p<0.005$), the trends were computed with SZA bins of $1^\\circ$ for each day, before and after local noon.', fig.subcap=c('All skies.', 'Clear skies.', 'Cloudy skies.'), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3}
+```{r szatrends, echo = FALSE, fig.cap='Long term trends of SDR as a function of SZA separately form morning and afternoon periods.
+Solid shapes  represent statistically significant trends ($p < 0.005$).', fig.subcap=c('All skies.', 'Clear skies.', 'Cloudy skies.'), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3}
 knitr::include_graphics("./images/SzaTrends-1.png")
-knitr::include_graphics("./images/SzaTrends-4.png" )
+knitr::include_graphics("./images/SzaTrends-4.png")
 knitr::include_graphics("./images/SzaTrends-7.png")
@@
 
-## Long term trends by season of year
+## Long term trends by season
 
@@
 
-```{r seasonalALL, echo=F, fig.cap="Linear trends (black lines) of monthly mean SDR season (columns of graphs) for the three sky conditions (rows of graphs).
-The vertical scale is common for all graphs.", out.width="100%"}
+```{r seasonalALL, echo=F, fig.cap="Linear trends (black lines) of monthly mean anomalies of SDR by season (rows of plots) for the three sky conditions (columns of plots).", out.width="100%"}
 knitr::include_graphics("./images/SeasonalTrendsTogether3-2.png")

### 2023-11-08 MDPI_submission.tex
Commit 2023-11-08 15:20

@@
 Finally, we identified some anomalies in the long term SDR trends for
-all sky conditions.}
+all sky conditions by examining the cumulative sums of monthly anomalies
+from the climatological mean.}
 
 % Keywords
-\keyword{GHI; SDR; Solar Brigthening/Dimming, aerosols, clouds.}
+\keyword{GHI; SDR; solar radiation; Solar Brigthening/Dimming; aerosols;
+clouds.}
 
@@
 
-\caption{Running cumulative sum plots of the monthly SDR.}\label{fig:cusummonth}
+\caption{Cumulative sum plots of the monthly SDR anomalies in (\%) for different sky conditions.}\label{fig:cusummonth}
 \end{figure}
@@
 
-\caption{Running cumulative sum plots of the monthly SDR after removing the long term trend from the monthly values.}\label{fig:cusumnotrendmonthly}
+\caption{Cumulative sum plots of monthly SDR anomalies in (\%) for different sky conditions after removing the long-term linear trend.}\label{fig:cusumnotrendmonthly}
 \end{figure}
@@
 
-\caption{Distribution of the SDR's long term trends by SZA before and after noon.
-The solid data points represent values with an acceptable statistical significance ($p<0.005$), the trends were computed with SZA bins of $1^\circ$ for each day, before and after local noon.}\label{fig:szatrends}
+\caption{Long term trends of SDR as a function of SZA separately form morning and afternoon periods. Solid shapes  represent statistically significant trends ($p < 0.005$).}\label{fig:szatrends}
 \end{figure}
 
-\hypertarget{long-term-trends-by-season-of-year}{%
-\subsection{Long term trends by season of
-year}\label{long-term-trends-by-season-of-year}}
+\hypertarget{long-term-trends-by-season}{%
+\subsection{Long term trends by
+season}\label{long-term-trends-by-season}}
 
@@
 
-\caption{Linear trends (black lines) of monthly mean SDR season (columns of graphs) for the three sky conditions (rows of graphs). The vertical scale is common for all graphs.}\label{fig:seasonalALL}
+\caption{Linear trends (black lines) of monthly mean anomalies of SDR by season (rows of plots) for the three sky conditions (columns of plots).}\label{fig:seasonalALL}
 \end{figure}

### 2023-11-08 MDPI_submission.Rmd
Commit 2023-11-08 16:10

@@
   GHI; SDR; solar radiation; Solar Brigthening/Dimming; aerosols; clouds.
-## list three to ten pertinent keywords specific to the article, yet reasonably common within the subject discipline
 
-acknowledgement: |
-  All sources of funding of the study should be disclosed. Please clearly
-  indicate grants that you have received in support of your research work.
-  Clearly state if you received funds for covering the costs to publish in open
-  access.
+# acknowledgement: |
+#   All sources of funding of the study should be disclosed. Please clearly
+#   indicate grants that you have received in support of your research work.
+#   Clearly state if you received funds for covering the costs to publish in open
+#   access.
 
@@
 dataavailability: |
-  We encourage all authors of articles published in MDPI journals to share
-  their research data. In this section, please provide details regarding where
-  data supporting reported results can be found, including links to publicly
-  archived datasets analyzed or generated during the study. Where no new data
-  were created, or where data is unavailable due to privacy or ethical
-  re-strictions, a statement is still required. Suggested Data Availability
-  Statements are available in section “MDPI Research Data Policies” at
-  \url{https://www.mdpi.com/ethics}.
+  Data as daily sums are available through the WRDC database \url{http://http://wrdc.mgo.rssi.ru}.
+  One minute data are available on request from the corresponding author. The data are not publicly available for protection against unmonitored commercial use.
 

### 2023-11-08 MDPI_submission.tex
Commit 2023-11-08 16:10

@@
 
-\dataavailability{We encourage all authors of articles published in MDPI
-journals to share their research data. In this section, please provide
-details regarding where data supporting reported results can be found,
-including links to publicly archived datasets analyzed or generated
-during the study. Where no new data were created, or where data is
-unavailable due to privacy or ethical re-strictions, a statement is
-still required. Suggested Data Availability Statements are available in
-section ``MDPI Research Data Policies'' at
-\url{https://www.mdpi.com/ethics}.}
-
-\acknowledgments{All sources of funding of the study should be
-disclosed. Please clearly indicate grants that you have received in
-support of your research work. Clearly state if you received funds for
-covering the costs to publish in open access.}
+\dataavailability{Data as daily sums are available through the WRDC
+database \url{http://http://wrdc.mgo.rssi.ru}. One minute data are
+available on request from the corresponding author. The data are not
+publicly available for protection against unmonitored commercial use.}
 

### 2023-11-08 DHI_GHI_2_sza_trends.R

@@ -403,7 +403,9 @@ vars <- c("GLB_att_des")
 #+ SzaTrends, echo=F, include=T, results = "asis"
 for (avar in vars) {
     ## ALL - CS
+    lec <- 0
     for (type in unique(szatrends$DATA)) {
+        lec <- lec + 1
 
         cat("\n\\newpage\n\n")
         cat("\n#### ", translate(type), translate(avar) , "from daily\n\n")
@@ -479,7 +481,8 @@ for (avar in vars) {
                 # title(paste(translate(avar), awename, "for", translate(type)), cex.main =  .8 * ccex)
                 title(paste(translate(avar), "trend", "for", translate(type)), cex.main =  .8 * ccex)
             } else {
-                legend("bottomleft", 0, translate(type),
+                # translate(type)
+                legend("bottomright", 0, paste0("(", letters[lec], ")"),
                        bty   = "n",
                        cex   = .8 * ccex,
                        xjust = 0.5,      # 0.5 means center adjusted

### 2023-11-08 appendix.tex
Commit 2023-11-08 17:10

@@
 
-\caption{Seasonla distribution of the SDR's long term trends by SZA before and after noon.
-The solid data points represent values with an acceptable statistical significance ($p<0.005$), the trends were computed with SZA bins of $1^\circ$ for each day, before and after local noon, all graphs have the same scale.
-Bins with $p<0.005$ or with less than 85 data points may be missing from the view.}\label{fig:SZAtrendSeason}
+\caption{Seasonal distribution of the SDR's long term trends by SZA before and after noon.
+The solid data points represent values with an acceptable statistical significance ($p<0.005$), the trends were computed with SZA bins of $1^\circ$, before and after local noon.
+Bins where $p<0.005$ or with less than 85 data points may be missing from view.}\label{fig:SZAtrendSeason}
 \end{figure}

### 2023-11-08 DHI_GHI_2_sza_trends.R

@@ -1135,6 +1135,7 @@ for (ase in seasons) {
                      "CLEAR_2_bySeason_daily_DESEAS",
                      "CLOUD_2_bySeason_daily_DESEAS")
     Seasons     <- c("Winter", "Spring", "Summer", "Autumn")
+    lec         <- 0
 
     ## the order must be predefined to match
     expanded <- expand.grid(Dataset = dbs, Seasons = Seasons, stringsAsFactors = FALSE)
@@ -1230,6 +1231,7 @@ for (ase in seasons) {
         ## actual plots
         if (! i %in% c(6,11,16,21,10,15,20,25)) {
 
+            lec <- lec + 1
             par("mar" = c(0,0,0,0))
 
             kk       <- expanded[ 1, ]
@@ -1342,6 +1344,17 @@ for (ase in seasons) {
                    pch = ppm$pch,
                    col = 3,
                    cex = ccex)
+
+            ## tag plots with letters
+            legend("bottomleft", 0, paste0("(",letters[lec],")"),
+                   cex   = 1.1,
+                   bty   = "n",
+                   xjust = 0.5,      # 0.5 means center adjusted
+                   yjust = 0.5,      # 0.5 means center adjusted
+                   x.intersp = -0.5, # adjust character interspacing as you like to effect box width
+                   y.intersp =  0.2, # adjust character interspacing to effect box height
+                   adj = c(0, 0.5))  # adjust string position (default values used here)
+
 
             ## legend on any of the plots
             if (i %in% c(7)) {

### 2023-11-08 MDPI_submission.Rmd
Commit 2023-11-08 19:09

@@
 
-authorcontributions: |
-  For research articles with several authors, a short paragraph specifying their
-  individual contributions must be provided. The following statements should be
-  used ``X.X. and Y.Y. conceive and designed the experiments; X.X. performed the
-  experiments; X.X. and Y.Y. analyzed the data; W.W. contributed
-  reagents/materials/analysis tools; Y.Y. wrote the paper.'' Authorship must be
-  limited to those who have contributed substantially to the work reported.
+# authorcontributions: |
+#   For research articles with several authors, a short paragraph specifying their
+#   individual contributions must be provided. The following statements should be
+#   used ``X.X. and Y.Y. conceive and designed the experiments; X.X. performed the
+#   experiments; X.X. and Y.Y. analyzed the data; W.W. contributed
+#   reagents/materials/analysis tools; Y.Y. wrote the paper.'' Authorship must be
+#   limited to those who have contributed substantially to the work reported.
 

### 2023-11-08 MDPI_submission.tex
Commit 2023-11-08 19:09

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\authorcontributions{For research articles with several authors, a short
-paragraph specifying their individual contributions must be provided.
-The following statements should be used ``X.X. and Y.Y. conceive and
-designed the experiments; X.X. performed the experiments; X.X. and Y.Y.
-analyzed the data; W.W. contributed reagents/materials/analysis tools;
-Y.Y. wrote the paper.'\,' Authorship must be limited to those who have
-contributed substantially to the work reported.}
 

### 2023-11-08 appendix.tex
Commit 2023-11-08 20:14

@@
 
+\section[\appendixname~\thesection]{} % test
+
+\  \
+
 % copy output from Article.tex
@@
 
-{\centering \includegraphics[width=0.7\linewidth]{./images/LongtermTrends-5}
-
-}
+{\centering \includegraphics[width=0.7\linewidth]{./images/LongtermTrends-5}}
 
@@
 
+\  \
+
 \begin{figure}[h!]
 
-{\centering \includegraphics[width=0.7\linewidth]{./images/LongtermTrends-8}
-
-}
+{\centering \includegraphics[width=0.7\linewidth]{./images/LongtermTrends-8}}
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{./images/SzaTrendsSeasTogether-2}
+{\centering \includegraphics[width=1\linewidth]{./images/SzaTrendsSeasTogether-2}}
 
-}
+\caption{Long term trends of SDR as a function of SZA separately form morning and afternoon
+periods, by season (rows of plots) for the three sky conditions (columns of plots).
+Solid shapes represent statistical significant trends ($p<0.005$).
+Cases where $p<0.005$ or with less than $85$ observations may be missing from view.}\label{fig:SZAtrendSeason}
+\end{figure}
 
-\caption{Seasonal distribution of the SDR's long term trends by SZA before and after noon.
-The solid data points represent values with an acceptable statistical significance ($p<0.005$), the trends were computed with SZA bins of $1^\circ$, before and after local noon.
-Bins where $p<0.005$ or with less than 85 data points may be missing from view.}\label{fig:SZAtrendSeason}
-\end{figure}
+\ \
 

### 2023-11-09 MDPI_submission.Rmd
Commit 2023-11-08 20:14
Commit 2023-11-09 04:14

@@
    - \captionsetup[sub]{position=bottom, labelfont={bf, small, stretch=1.17}, labelsep=space, textfont={small, stretch=1.17}, aboveskip=6pt,  belowskip=-6pt, singlelinecheck=off, justification=justified}
-
-#  - \captionsetup[sub]{position=bottom, labelfont={bf, small, stretch=1.17}, labelsep=space, textfont={small, stretch=1.5}, aboveskip=6pt,  belowskip=6pt, singlelinecheck=off, justification=justified}
-#  \captionsetup[figure]{position=bottom, labelfont={bf, small, stretch=1.17}, labelsep=period, textfont={small, stretch=1.17}, aboveskip=6pt, belowskip=-6pt, singlelinecheck=off, justification=justified}
+   - \usepackage{placeins}
 
@@
 -->
-
-<!-- ------------------------------------------------------------ -->
-
-<!-- ## Figures, Tables and Schemes -->
-
-<!-- All figures and tables should be cited in the main text as Figure \ref{fig:fig1},  -->
-<!-- \ref{tab:tab1}, etc. To get cross-reference to figure generated by R chunks  -->
-<!-- include the `\\label{}` tag in the `fig.cap` attribute of the R chunk:  -->
-<!-- `fig.cap = "Fancy Caption\\label{fig:plot}"`. -->
-
-<!-- ```{r fig1, echo=FALSE, fig.width=3, fig.cap="A figure added with a code chunk.\\label{fig:fig1}"} -->
-<!-- x = rnorm(10) -->
-<!-- y = rnorm(10) -->
-<!-- plot(x, y) -->
-<!-- ``` -->
-
-<!-- When making tables using `kable`, it is suggested to use -->
-<!-- the `format="latex"` and `tabl.envir="table"` arguments -->
-<!-- to ensure table numbering and compatibility with the mdpi -->
-<!-- document class. -->
-
-<!-- ```{r tab1, echo=FALSE} -->
-<!-- knitr::kable(mtcars[1:5, 1:3],  -->
-<!--              format = "latex",  -->
-<!--              booktabs = TRUE,  -->
-<!--              caption = "This is a table caption. Tables should be placed in the  -->
-<!--              main text near to the first time they are~cited.",  -->
-<!--              align = 'ccc', centering = FALSE, -->
-<!--              table.envir = "table", position = "H") -->
-<!-- ``` -->
-
-<!-- ## Formatting of Mathematical Components -->
-
-<!-- If you want numbered equations use Latex and wrap in the equation environment: -->
-
-<!-- \begin{equation} -->
-<!-- a = 1, -->
-<!-- \end{equation} -->
-
-<!-- the text following an equation need not be a new paragraph. Please punctuate  -->
-<!-- equations as regular text. -->
-
-<!-- ```{comment} -->
-<!-- If the documentclass option "submit" is chosen, please insert a blank line before and after any math environment (equation and eqnarray environments).
-This ensures correct linenumbering.
-The blank line should be removed when the documentclass option is changed to "accept" because the text following an equation should not be a new paragraph.
--->
-<!-- ``` -->
-
-<!-- # Discussion -->
-
-<!-- Authors should discuss the results and how they can be interpreted in -->
-<!-- perspective of previous studies and of the working hypotheses. The findings and -->
-<!-- their implications should be discussed in the broadest context possible. Future -->
-<!-- research directions may also be highlighted. -->
-
-<!-- # Conclusion -->
-
-<!-- This section is not mandatory, but can be added to the manuscript if the -->
-<!-- discussion is unusually long or complex. -->

### 2023-11-09 MDPI_submission.Rmd
Commit 2023-11-09 05:11

@@
 
-```{r cusummonth, echo = FALSE, fig.cap='Cumulative sum plots of the monthly SDR anomalies in (%) for different sky conditions.', fig.subcap=c('All skies.', 'Clear skies.', 'Cloudy skies.'), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3}
-knitr::include_graphics(c(
-    "./images/CumulativeMonthlyCuSum-1.png",    # ALL sky
-    "./images/CumulativeMonthlyCuSum-5.png",    # CLEAR sky
-    "./images/CumulativeMonthlyCuSum-9.png"))   # Cloud sky
-```
+\begin{figure}[h!]
+    \begin{adjustwidth}{-\extralength}{0cm}
+        {\centering 
+        \subfloat[All skies.\label{fig:cusummonth-1}]
+            {\includegraphics[width=.32\linewidth]{./images/CumulativeMonthlyCuSum-1}}\hfill
+        \subfloat[Clear skies.\label{fig:cusummonth-2}]
+            {\includegraphics[width=.32\linewidth]{./images/CumulativeMonthlyCuSum-5}}\hfill
+        \subfloat[Cloudy skies.\label{fig:cusummonth-3}]
+            {\includegraphics[width=.32\linewidth]{./images/CumulativeMonthlyCuSum-9}}\hfill
+        }
+\caption{Cumulative sum plots of the monthly SDR anomalies in (\%) for different sky conditions.}\label{fig:cusummonth}
+\end{adjustwidth}
+\end{figure}
+
+<!-- ```{r cusummonth, echo = FALSE, fig.cap='Cumulative sum plots of the monthly SDR anomalies in (%) for different sky conditions.', fig.subcap=c('All skies.', 'Clear skies.', 'Cloudy skies.'), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3} -->
+<!-- knitr::include_graphics(c( -->
+<!--     "./images/CumulativeMonthlyCuSum-1.png",    # ALL sky -->
+<!--     "./images/CumulativeMonthlyCuSum-5.png",    # CLEAR sky -->
+<!--     "./images/CumulativeMonthlyCuSum-9.png"))   # Cloud sky -->
+<!-- ``` -->
 
@@
 
-```{r cusumnotrendmonthly, echo = FALSE, fig.cap='Cumulative sum plots of monthly SDR anomalies in (%) for different sky conditions after removing the long-term linear trend.', fig.subcap=c('All skies.', 'Clear skies.', 'Cloudy skies.'), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3}
-knitr::include_graphics(c(
-    "./images/CumulativeMonthlyCuSumNOtrend-1.png",    # ALL sky
-    "./images/CumulativeMonthlyCuSumNOtrend-5.png",    # CLEAR sky
-    "./images/CumulativeMonthlyCuSumNOtrend-9.png"))   # Cloud sky
-```
+\begin{figure}[h!]
+    \begin{adjustwidth}{-\extralength}{0cm}
+        {\centering 
+            \subfloat[All skies.\label{fig:cusumnotrendmonthly-1}]
+                {\includegraphics[width=.32\linewidth]{./images/CumulativeMonthlyCuSumNOtrend-1} }\hfill
+            \subfloat[Clear skies.\label{fig:cusumnotrendmonthly-2}]
+                {\includegraphics[width=.32\linewidth]{./images/CumulativeMonthlyCuSumNOtrend-5} }\hfill
+            \subfloat[Cloudy skies.\label{fig:cusumnotrendmonthly-3}]
+                {\includegraphics[width=.32\linewidth]{./images/CumulativeMonthlyCuSumNOtrend-9} }
+        }
+\caption{Cumulative sum plots of monthly SDR anomalies in (\%) for different sky conditions after removing the long-term linear trend.}\label{fig:cusumnotrendmonthly}
+\end{adjustwidth}
+\end{figure}
+
+<!-- ```{r cusumnotrendmonthly, echo = FALSE, fig.cap='Cumulative sum plots of monthly SDR anomalies in (%) for different sky conditions after removing the long-term linear trend.', fig.subcap=c('All skies.', 'Clear skies.', 'Cloudy skies.'), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3} -->
+<!-- knitr::include_graphics(c( -->
+<!--     "./images/CumulativeMonthlyCuSumNOtrend-1.png",    # ALL sky -->
+<!--     "./images/CumulativeMonthlyCuSumNOtrend-5.png",    # CLEAR sky -->
+<!--     "./images/CumulativeMonthlyCuSumNOtrend-9.png"))   # Cloud sky -->
+<!-- ``` -->
 
@@
 Between $50^\circ$ and $75^\circ$ SZA the trends for the period before noon are stronger and negative, possibly associated with stronger attenuation by clouds under oblique incidence angles.
+
+\begin{figure}[h!]
+    \begin{adjustwidth}{-\extralength}{0cm}
+
+{\centering \subfloat[All skies.\label{fig:szatrends-1}]{\includegraphics[width=.32\linewidth]{./images/SzaTrends-1} }\subfloat[Clear skies.\label{fig:szatrends-2}]{\includegraphics[width=.32\linewidth]{./images/SzaTrends-4} }\subfloat[Cloudy skies.\label{fig:szatrends-3}]{\includegraphics[width=.32\linewidth]{./images/SzaTrends-7} }
+
+}
+
+\caption{Long term trends of SDR as a function of SZA separately form morning and afternoon periods. Solid shapes  represent statistically significant trends ($p < 0.005$).}\label{fig:szatrends}
+\end{adjustwidth}
+\end{figure}
 

### 2023-11-09 MDPI_submission.tex
Commit 2023-11-09 04:14
Commit 2023-11-09 05:11

@@
 \captionsetup[sub]{position=bottom, labelfont={bf, small, stretch=1.17}, labelsep=space, textfont={small, stretch=1.17}, aboveskip=6pt,  belowskip=-6pt, singlelinecheck=off, justification=justified}
+\usepackage{placeins}
 \usepackage{longtable}
@@
 \begin{figure}[h!]
-
-{\centering \subfloat[All skies.\label{fig:cusummonth-1}]{\includegraphics[width=.32\linewidth]{./images/CumulativeMonthlyCuSum-1} }\subfloat[Clear skies.\label{fig:cusummonth-2}]{\includegraphics[width=.32\linewidth]{./images/CumulativeMonthlyCuSum-5} }\subfloat[Cloudy skies.\label{fig:cusummonth-3}]{\includegraphics[width=.32\linewidth]{./images/CumulativeMonthlyCuSum-9} }
-
+    \begin{adjustwidth}{-\extralength}{0cm}
+        {\centering 
+        \subfloat[All skies.\label{fig:cusummonth-1}]
+            {\includegraphics[width=.32\linewidth]{./images/CumulativeMonthlyCuSum-1}}\hfill
+        \subfloat[Clear skies.\label{fig:cusummonth-2}]
+            {\includegraphics[width=.32\linewidth]{./images/CumulativeMonthlyCuSum-5}}\hfill
+        \subfloat[Cloudy skies.\label{fig:cusummonth-3}]
+            {\includegraphics[width=.32\linewidth]{./images/CumulativeMonthlyCuSum-9}}\hfill
         }
-
 \caption{Cumulative sum plots of the monthly SDR anomalies in (\%) for different sky conditions.}\label{fig:cusummonth}
+\end{adjustwidth}
 \end{figure}
@@
 \begin{figure}[h!]
-
-{\centering \subfloat[All skies.\label{fig:cusumnotrendmonthly-1}]{\includegraphics[width=.32\linewidth]{./images/CumulativeMonthlyCuSumNOtrend-1} }\subfloat[Clear skies.\label{fig:cusumnotrendmonthly-2}]{\includegraphics[width=.32\linewidth]{./images/CumulativeMonthlyCuSumNOtrend-5} }\subfloat[Cloudy skies.\label{fig:cusumnotrendmonthly-3}]{\includegraphics[width=.32\linewidth]{./images/CumulativeMonthlyCuSumNOtrend-9} }
-
+    \begin{adjustwidth}{-\extralength}{0cm}
+        {\centering 
+            \subfloat[All skies.\label{fig:cusumnotrendmonthly-1}]
+                {\includegraphics[width=.32\linewidth]{./images/CumulativeMonthlyCuSumNOtrend-1} }\hfill
+            \subfloat[Clear skies.\label{fig:cusumnotrendmonthly-2}]
+                {\includegraphics[width=.32\linewidth]{./images/CumulativeMonthlyCuSumNOtrend-5} }\hfill
+            \subfloat[Cloudy skies.\label{fig:cusumnotrendmonthly-3}]
+                {\includegraphics[width=.32\linewidth]{./images/CumulativeMonthlyCuSumNOtrend-9} }
         }
-
 \caption{Cumulative sum plots of monthly SDR anomalies in (\%) for different sky conditions after removing the long-term linear trend.}\label{fig:cusumnotrendmonthly}
+\end{adjustwidth}
 \end{figure}
@@
 incidence angles.
+
+\begin{figure}[h!]
+    \begin{adjustwidth}{-\extralength}{0cm}
+
+{\centering \subfloat[All skies.\label{fig:szatrends-1}]{\includegraphics[width=.32\linewidth]{./images/SzaTrends-1} }\subfloat[Clear skies.\label{fig:szatrends-2}]{\includegraphics[width=.32\linewidth]{./images/SzaTrends-4} }\subfloat[Cloudy skies.\label{fig:szatrends-3}]{\includegraphics[width=.32\linewidth]{./images/SzaTrends-7} }
+
+}
+
+\caption{Long term trends of SDR as a function of SZA separately form morning and afternoon periods. Solid shapes  represent statistically significant trends ($p < 0.005$).}\label{fig:szatrends}
+\end{adjustwidth}
+\end{figure}
 

### 2023-11-09 MDPI_submission.tex
Commit 2023-11-09 06:10

@@
     \begin{adjustwidth}{-\extralength}{0cm}
-
-{\centering \subfloat[All skies.\label{fig:szatrends-1}]{\includegraphics[width=.32\linewidth]{./images/SzaTrends-1} }\subfloat[Clear skies.\label{fig:szatrends-2}]{\includegraphics[width=.32\linewidth]{./images/SzaTrends-4} }\subfloat[Cloudy skies.\label{fig:szatrends-3}]{\includegraphics[width=.32\linewidth]{./images/SzaTrends-7} }
-
+        {\centering 
+            \subfloat[All skies.\label{fig:szatrends-1}]
+                {\includegraphics[width=.32\linewidth]{./images/SzaTrends-1}}\hfill
+            \subfloat[Clear skies.\label{fig:szatrends-2}]
+                {\includegraphics[width=.32\linewidth]{./images/SzaTrends-4}}\hfill
+            \subfloat[Cloudy skies.\label{fig:szatrends-3}]
+                {\includegraphics[width=.32\linewidth]{./images/SzaTrends-7}}
         }
-
-\caption{Long term trends of SDR as a function of SZA separately form morning and afternoon periods. Solid shapes  represent statistically significant trends ($p < 0.005$).}\label{fig:szatrends}
+        \caption{Long term trends of SDR as a function of SZA separately form morning and afternoon periods.
+Solid shapes  represent statistically significant trends ($p < 0.005$).}\label{fig:szatrends}
     \end{adjustwidth}
-\end{figure}
-
-\begin{figure}[h!]
-
-{\centering \subfloat[All skies.\label{fig:szatrends-1}]{\includegraphics[width=.32\linewidth]{./images/SzaTrends-1} }\subfloat[Clear skies.\label{fig:szatrends-2}]{\includegraphics[width=.32\linewidth]{./images/SzaTrends-4} }\subfloat[Cloudy skies.\label{fig:szatrends-3}]{\includegraphics[width=.32\linewidth]{./images/SzaTrends-7} }
-
-}
-
-\caption{Long term trends of SDR as a function of SZA separately form morning and afternoon periods. Solid shapes  represent statistically significant trends ($p < 0.005$).}\label{fig:szatrends}
 \end{figure}
@@
 \begin{figure}[h!]
-
-{\centering \includegraphics[width=1\linewidth]{./images/SeasonalTrendsTogether3-2}
-
+    \begin{adjustwidth}{-\extralength}{0cm}
+        {\centering 
+            \includegraphics[width=1\linewidth]{./images/SeasonalTrendsTogether3-2} 
         }
-
         \caption{Linear trends (black lines) of monthly mean anomalies of SDR by season (rows of plots) for the three sky conditions (columns of plots).}\label{fig:seasonalALL}
+    \end{adjustwidth}
 \end{figure}

### 2023-11-09 gather_test/MDPI_submission.Rmd
Commit 2023-11-09 06:10

--- 2023-11-09 MDPI_submission.Rmd
@@
     \begin{adjustwidth}{-\extralength}{0cm}
-
-{\centering \subfloat[All skies.\label{fig:szatrends-1}]{\includegraphics[width=.32\linewidth]{./images/SzaTrends-1} }\subfloat[Clear skies.\label{fig:szatrends-2}]{\includegraphics[width=.32\linewidth]{./images/SzaTrends-4} }\subfloat[Cloudy skies.\label{fig:szatrends-3}]{\includegraphics[width=.32\linewidth]{./images/SzaTrends-7} }
-
+        {\centering 
+            \subfloat[All skies.\label{fig:szatrends-1}]
+                {\includegraphics[width=.32\linewidth]{./images/SzaTrends-1}}\hfill
+            \subfloat[Clear skies.\label{fig:szatrends-2}]
+                {\includegraphics[width=.32\linewidth]{./images/SzaTrends-4}}\hfill
+            \subfloat[Cloudy skies.\label{fig:szatrends-3}]
+                {\includegraphics[width=.32\linewidth]{./images/SzaTrends-7}}
         }
-
-\caption{Long term trends of SDR as a function of SZA separately form morning and afternoon periods. Solid shapes  represent statistically significant trends ($p < 0.005$).}\label{fig:szatrends}
+        \caption{Long term trends of SDR as a function of SZA separately form morning and afternoon periods.
+Solid shapes  represent statistically significant trends ($p < 0.005$).}\label{fig:szatrends}
     \end{adjustwidth}
@@
 
-```{r szatrends, echo = FALSE, fig.cap='Long term trends of SDR as a function of SZA separately form morning and afternoon periods.
-Solid shapes  represent statistically significant trends ($p < 0.005$).', fig.subcap=c('All skies.', 'Clear skies.', 'Cloudy skies.'), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3}
-knitr::include_graphics("./images/SzaTrends-1.png")
-knitr::include_graphics("./images/SzaTrends-4.png")
-knitr::include_graphics("./images/SzaTrends-7.png")
-```
+<!-- ```{r szatrends, echo = FALSE, fig.cap='Long term trends of SDR as a function of SZA separately form morning and afternoon periods.
+Solid shapes  represent statistically significant trends ($p < 0.005$).', fig.subcap=c('All skies.', 'Clear skies.', 'Cloudy skies.'), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3} -->
+<!-- knitr::include_graphics("./images/SzaTrends-1.png") -->
+<!-- knitr::include_graphics("./images/SzaTrends-4.png") -->
+<!-- knitr::include_graphics("./images/SzaTrends-7.png") -->
+<!-- ``` -->
 
@@
 target    <- c("Winter", "Spring", "Summer", "Autumn")
-new_order <- sapply(target, function(x,spp){which(spp$Season == x)}, spp = spp)
+new_order <- sapply(target, function(x, spp){which(spp$Season == x)}, spp = spp)
 spp        <- spp[new_order,]
@@
 
-```{r seasonalALL, echo=F, fig.cap="Linear trends (black lines) of monthly mean anomalies of SDR by season (rows of plots) for the three sky conditions (columns of plots).", out.width="100%"}
-knitr::include_graphics("./images/SeasonalTrendsTogether3-2.png")
-```
+\begin{figure}[h!]
+    \begin{adjustwidth}{-\extralength}{0cm}
+        {\centering 
+            \includegraphics[width=1\linewidth]{./images/SeasonalTrendsTogether3-2} 
+        }
+        \caption{Linear trends (black lines) of monthly mean anomalies of SDR by season (rows of plots) for the three sky conditions (columns of plots).}\label{fig:seasonalALL}
+    \end{adjustwidth}
+\end{figure}
+
+<!-- ```{r seasonalALL, echo=F, fig.cap="Linear trends (black lines) of monthly mean anomalies of SDR by season (rows of plots) for the three sky conditions (columns of plots).", out.width="100%"} -->
+<!-- knitr::include_graphics("./images/SeasonalTrendsTogether3-2.png") -->
+<!-- ``` -->
 

### 2023-11-09 gather_include.sh

@@ -1,0 +1,38 @@
+#!/bin/bash
+## created on 2023-11-09
+
+#### enter description here
+
+## Repo must only one level deep no rucursive folders
+
+## Source file with paths in REPO
+SOURCE="$1"
+## Repository of include files
+REPO="$2"
+## Location to copy include files
+TARGET="$3"
+
+[ ! -f "$SOURCE" ] && echo && echo "NOT A FILE $SOURCE"   && exit 1
+[ ! -d "$REPO"   ] && echo && echo "NOT A FOLDER $REPO"   && exit 1
+[ ! -d "$TARGET" ] && echo && echo "NOT A FOLDER $TARGET" && exit 1
+
+mkdir -p "$TARGET/$REPO"
+
+## get included paths
+sed -e :a -re 's/<!--.*?-->//g;/<!--/N;//ba' "$SOURCE" |
+    sed '/^[[:blank:]]*#/d;s/#.*//' |
+    grep "$REPO/"                   |
+    sed 's/.*\('"$REPO"'\)/\1/'     |
+    sed 's/\..*//'                  |
+    sed 's/}.*//'                   |
+    sort -u                         | 
+    while read line; do 
+        echo "Processing: $line"
+        # ls "$line"*
+        cp -v "$line"* "$TARGET/$REPO"
+    done
+
+cp -v  "$SOURCE"       "$TARGET"
+cp -vr "./Definitions" "$TARGET"
+
+exit 0 

### 2023-11-09 MDPI_submission.Rmd
Commit 2023-11-09 07:20

@@
 dataavailability: |
-  Data as daily sums are available through the WRDC database \url{http://http://wrdc.mgo.rssi.ru}.
+  Data as daily sums are available through the WRDC database \url{http://wrdc.mgo.rssi.ru}.
   One minute data are available on request from the corresponding author. The data are not publicly available for protection against unmonitored commercial use.

### 2023-11-09 MDPI_submission.Rmd
Commit 2023-11-09 09:22

@@
     affil: 1 #, \dagger, # \ddagger
+  - name: Alkiviadis Bais
+    affil: 1 #, \dagger, # \ddagger
+
 affiliation:
@@
 We found a positive trend of $0.38\,\%/\text{year}$ for all-sky, $0.35\,\%/\text{year}$ for clear-sky conditions, and $-0.28\,\%/\text{year}$ for cloudy conditions.
-  We have also investigated the consistency of these trends, the effect of the solar zenith angle, and their seasonal variability.
+  We have also investigated the consistency of these trends, their seasonal variability, and the effect of the solar zenith angle. 
+  We have found that for all-sky and clear-sky conditions the SDR trend is positive in winter (0.7 and 0.8 %/y, respectively) and autumn (~0.4 %/y), while under cloudy skies the trend is negative (-0.9 %/y in winter and -0.4 %/y in autumn).
+  In spring and summer the trend is very close to zero, irrespective of sky conditions. 
+  The SDR trend is increasing with increasing solar zenith angle, except under cloudy skies where the trend is highly variable and close to zero.
   Finally, we identified some anomalies in the long term SDR trends for all sky conditions by examining the cumulative sums of monthly anomalies from the climatological mean.

### 2023-11-09 Build_163/MDPI_submission_B163.Rmd
Commit 2023-11-09 19:08

--- 2023-11-09 MDPI_submission.Rmd
@@
     affil: 1 #, \dagger, # \ddagger
-  - name: Alkiviadis Bais
+  - name: Charikleia Meleti
     affil: 1 #, \dagger, # \ddagger
@@
       Campus Box 149, 54124 Thessaloniki, Greece
-    email: natsisphysicist@gmail.com (A.N.); abais@auth.gr (A.B.)
+#    email: natsisphysicist@gmail.com (A.N.); abais@auth.gr (A.B.); meleti@auth.gr (C.M.)
   # - num: 2
@@
 authorcitation: |
-  Natsis, A.; Bais, A.
+  Natsis, A.; Bais, A.; Meleti, C.
 # firstnote to eighthnote
@@
   We have also investigated the consistency of these trends, their seasonal variability, and the effect of the solar zenith angle. 
-  We have found that for all-sky and clear-sky conditions the SDR trend is positive in winter (0.7 and 0.8 %/y, respectively) and autumn (~0.4 %/y), while under cloudy skies the trend is negative (-0.9 %/y in winter and -0.4 %/y in autumn).
+  We have found that for all-sky and clear-sky conditions the SDR trend is positive in winter ($0.7$ and $0.8\,\%/\text{year}$, respectively) and autumn ($~0.4\,\%/\text{year}$), while under cloudy skies the trend is negative ($-0.9\,\%/\text{year}$ in winter and $-0.4\,\%/\text{year}$ in autumn).
   In spring and summer the trend is very close to zero, irrespective of sky conditions. 

### 2023-11-09 Build_163/MDPI_submission_B163.tex
Commit 2023-11-09 19:08

--- 2023-11-09 MDPI_submission.tex
@@
 Natsis$^{1,*}$\href{https://orcid.org/0000-0002-5199-4119}
-{\orcidicon}, Alkiviadis Bais$^{1}$}
+{\orcidicon}, Alkiviadis Bais$^{1}$, Charikleia Meleti$^{1}$}
 
@@
 % MDPI internal command: Authors, for metadata in PDF
-\AuthorNames{Athanasios Natsis, Alkiviadis Bais}
+\AuthorNames{Athanasios Natsis, Alkiviadis Bais, Charikleia Meleti}
 
@@
 % If this is a Chicago style journal: Lastname, Firstname, Firstname Lastname, and Firstname Lastname.
-\AuthorCitation{Natsis, A.; Bais, A.}
+\AuthorCitation{Natsis, A.; Bais, A.; Meleti, C.}
 
@@
 $^{1}$ \quad Aristotle University of Thessaloniki - Laboratory of
-Atmospheric Physics, Campus Box 149, 54124 Thessaloniki,
-Greece; \href{mailto:natsisphysicist@gmail.com}{\nolinkurl{natsisphysicist@gmail.com}}
-(A.N.); \href{mailto:abais@auth.gr}{\nolinkurl{abais@auth.gr}} (A.B.)\\
+Atmospheric Physics, Campus Box 149, 54124 Thessaloniki, Greece; \\
 }
@@
 conditions. We have also investigated the consistency of these trends,
-the effect of the solar zenith angle, and their seasonal variability.
-Finally, we identified some anomalies in the long term SDR trends for
-all sky conditions by examining the cumulative sums of monthly anomalies
-from the climatological mean.}
+their seasonal variability, and the effect of the solar zenith angle. We
+have found that for all-sky and clear-sky conditions the SDR trend is
+positive in winter (\(0.7\) and \(0.8\,\%/\text{year}\), respectively)
+and autumn (\(~0.4\,\%/\text{year}\)), while under cloudy skies the
+trend is negative (\(-0.9\,\%/\text{year}\) in winter and
+\(-0.4\,\%/\text{year}\) in autumn). In spring and summer the trend is
+very close to zero, irrespective of sky conditions. The SDR trend is
+increasing with increasing solar zenith angle, except under cloudy skies
+where the trend is highly variable and close to zero. Finally, we
+identified some anomalies in the long term SDR trends for all sky
+conditions by examining the cumulative sums of monthly anomalies from
+the climatological mean.}
 
@@
 \dataavailability{Data as daily sums are available through the WRDC
-database \url{http://http://wrdc.mgo.rssi.ru}. One minute data are
-available on request from the corresponding author. The data are not
-publicly available for protection against unmonitored commercial use.}
+database \url{http://wrdc.mgo.rssi.ru}. One minute data are available on
+request from the corresponding author. The data are not publicly
+available for protection against unmonitored commercial use.}
 

### 2023-11-09 SUBMISSION/appendix.tex
Commit 2023-11-09 07:20
Commit 2023-11-09 19:08

--- 2023-11-08 appendix.tex
@@
-%% optional
-\appendixtitles{no} %Leave argument "no" if all appendix headings stay EMPTY (then no dot is printed after "Appendix A"). If the appendix sections contain a heading then change the argument to "yes".
-\appendixsections{one} %Leave argument "multiple" if there are multiple sections.
+% % \renewcommand\thefigure{\thesection.\arabic{figure}} %% my
+% % \renewcommand\thefigure{\thesection.\arabic{figure}}
+%
+% %% From R rticle MDPI template
+% %% optional
+% \appendixtitles{no} %Leave argument "no" if all appendix headings stay EMPTY (then no dot is printed after "Appendix A").
+If the appendix sections contain a heading then change the argument to "yes".
+% \appendixsections{one} %Leave argument "multiple" if there are multiple sections.
 Then a counter is printed ("Appendix A").
 If there is only one appendix section then change the argument to "one" and no counter is printed ("Appendix").
+%
+% % \appendix
+% % \section{}
+%
+% \appendix
+% \setcounter{secnumdepth}{0}
+% \section{Appendix}
+%
+% \setcounter{figure}{0}    %% my
+
+%% From raw MDPI template
+%% Optional
+\appendixtitles{no} % Leave argument "no" if all appendix headings stay EMPTY (then no dot is printed after "Appendix A"). If the appendix sections contain a heading then change the argument to "yes".
+\appendixstart
 \appendix
-% \section{}
+\section[\appendixname~\thesection]{}
 
-\section[\appendixname~\thesection]{} % test
-
-\  \
-
-% copy output from Article.tex
+% copy output from Rmd
 
 \begin{figure}[h!]
+    {\centering
+        \includegraphics[width=0.7\linewidth]{./images/LongtermTrends-5}
 
-{\centering \includegraphics[width=0.7\linewidth]{./images/LongtermTrends-5}}
-
-\caption{Anomalies (\%) of the daily clear-sky SDR, relative to climatological values for1993 - 2023. The black line shows the long term linear trend for clear-sky conditions.}\label{fig:trendCLEAR}
-\end{figure}
-
-\  \
-
-\begin{figure}[h!]
-
-{\centering \includegraphics[width=0.7\linewidth]{./images/LongtermTrends-8}}
-
-\caption{Anomalies (\%) of the daily cloud-sky SDR, relative to climatological values for 1993 - 2023. The black line shows the long term linear trend for cloud-sky conditions.}\label{fig:trendCLOUD}
+    }
+    \caption{Anomalies (\%) of the daily clear-sky SDR, relative to climatological values for1993 - 2023.
+The black line shows the long term linear trend for clear-sky conditions.}\label{fig:trendCLEAR}
 \end{figure}
@@
 \begin{figure}[h!]
+    {\centering
+        \includegraphics[width=0.7\linewidth]{./images/LongtermTrends-8}
 
-{\centering \includegraphics[width=1\linewidth]{./images/SzaTrendsSeasTogether-2}}
-
-\caption{Long term trends of SDR as a function of SZA separately form morning and afternoon
-periods, by season (rows of plots) for the three sky conditions (columns of plots).
-Solid shapes represent statistical significant trends ($p<0.005$).
-Cases where $p<0.005$ or with less than $85$ observations may be missing from view.}\label{fig:SZAtrendSeason}
+    }
+    \caption{Anomalies (\%) of the daily cloud-sky SDR, relative to climatological values for 1993 - 2023.
+The black line shows the long term linear trend for cloud-sky conditions.}\label{fig:trendCLOUD}
 \end{figure}
 
-\ \
+\begin{figure}[h!]
+    \begin{adjustwidth}{-\extralength}{0cm}
+        {\centering
+            \includegraphics[width=1.0\linewidth]{./images/SzaTrendsSeasTogether-2}
+        }
+        \caption{Long term trends of SDR as a function of SZA separately form morning
+               and afternoon periods, by season (rows of plots) for the three sky
+               conditions (columns of plots).
+               Solid shapes represent statistical significant trends ($p<0.005$).
+               Cases where $p<0.005$ or with less than $85$ observations may be
+               missing from view.}\label{fig:SZAtrendSeason}
+    \end{adjustwidth}
+\end{figure}
+
+\FloatBarrier
 

### 2023-11-10 MDPI_submission.Rmd
Commit 2023-11-10 16:17

@@
   - name:  Athanasios Natsis
-    affil: 1,* # \ddagger
+    affil: 1 # \ddagger
     orcid: 0000-0002-5199-4119
   - name: Alkiviadis Bais
-    affil: 1 #, \dagger, # \ddagger
-  - name: Alkiviadis Bais
+    affil: 1,* #, \dagger, # \ddagger
+  - name: Charikleia Meleti
     affil: 1 #, \dagger, # \ddagger
@@
       Campus Box 149, 54124 Thessaloniki, Greece
-    email: natsisphysicist@gmail.com (A.N.); abais@auth.gr (A.B.)
+    email: natsisphysicist@gmail.com (A.N.); abais@auth.gr (A.B.); meleti@auth.gr (C.M.)
   # - num: 2
@@
 correspondence: |
-  natsisphysicist@gmail.com;
+  abais@auth.gr
+# natsisphysicist@gmail.com;
 #   Tel.: +XX-000-00-0000.

### 2023-11-24 MDPI_submission.Rmd
Commit 2023-11-24 10:14

@@
 ## load data from trends analysis
-load("./data/common_data_14_2.Rda")
+source("./DHI_GHI_00_raw_data.R")
+load(common_data)
 load("./data/Input_1_longterm_trends.Rda")

### 2023-11-24 MDPI_submission.tex
Commit 2023-11-24 10:14

@@
 \Author{Athanasios
-Natsis$^{1,*}$\href{https://orcid.org/0000-0002-5199-4119}
-{\orcidicon}, Alkiviadis Bais$^{1}$}
+Natsis$^{1}$\href{https://orcid.org/0000-0002-5199-4119}
+{\orcidicon}, Alkiviadis Bais$^{1,*}$, Charikleia Meleti$^{1}$}
 
@@
 Greece; \href{mailto:natsisphysicist@gmail.com}{\nolinkurl{natsisphysicist@gmail.com}}
-(A.N.); \href{mailto:abais@auth.gr}{\nolinkurl{abais@auth.gr}} (A.B.)\\
+(A.N.); \href{mailto:abais@auth.gr}{\nolinkurl{abais@auth.gr}} (A.B.);
+\href{mailto:meleti@auth.gr}{\nolinkurl{meleti@auth.gr}} (C.M.)\\
 }
@@
 % Contact information of the corresponding author
-\corres{Correspondence: \href{mailto:natsisphysicist@gmail.com}{\nolinkurl{natsisphysicist@gmail.com}};}
+\corres{Correspondence: \href{mailto:abais@auth.gr}{\nolinkurl{abais@auth.gr}}}
 
@@
 initial daily values of this dataset were interpolated to match the time
-step of our measurements. In the final dataset \(84.2\%\) of the data
-were identified as under clear-sky conditions and \(15.8\%\) as under
-cloud-sky conditions.
+step of our measurements. In the final dataset \(40\%\) of the data were
+identified as under clear-sky conditions and \(60\%\) as under cloud-sky
+conditions.
 
@@
 positive trend for all-sky conditions (\(0.38\,\%/y\)), a very close but
-smaller trend for clear-skies (\(0.35\,\%/y\)) and a negative weaker
-trend for cloudy-skies (\(-0.28\,\%/y\)). In the studied period, there
+smaller trend for clear-skies (\(0.16\,\%/y\)) and a negative weaker
+trend for cloudy-skies (\(0.022\,\%/y\)). In the studied period, there
 is no significant break or change in the variability of the time series.
@@
 \midrule
-All skies & 0.376 & 100.00 & 10256\\
-Clear skies & 0.349 & 100.00 & 10256\\
-Cloudy skies & -0.276 & 99.97 & 5067\\
+All skies & 0.3760 & 100.00 & 10256\\
+Clear skies & 0.1650 & 100.00 & 7995\\
+Cloudy skies & 0.0225 & 32.57 & 10226\\
 \bottomrule
@@
 of clouds during winter and autumn. For clear-skies, the trend in winter
-is \(0.83\,\%/y\), larger than for all-skies (\(0.69\,\%/y\)), which is
+is \(0.7\,\%/y\), larger than for all-skies (\(0.69\,\%/y\)), which is
 another indication of a decreasing trend in cloud optical thickness.
@@
 \midrule
-\cellcolor{gray!6}{} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{0.6860} & \cellcolor{gray!6}{99.9}\\
+\cellcolor{gray!6}{} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{0.686000} & \cellcolor{gray!6}{99.900}\\
 
- & Spring & 0.1450 & 81.8\\
+ & Spring & 0.145000 & 81.800\\
 
-\cellcolor{gray!6}{} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{0.1200} & \cellcolor{gray!6}{89.4}\\
+\cellcolor{gray!6}{} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{0.120000} & \cellcolor{gray!6}{89.400}\\
 
-\multirow{-4}{*}{\centering\arraybackslash All skies} & Autumn & 0.4310 & 99.3\\
+\multirow{-4}{*}{\centering\arraybackslash All skies} & Autumn & 0.431000 & 99.300\\
 \cmidrule{1-4}
-\cellcolor{gray!6}{} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{0.8260} & \cellcolor{gray!6}{100.0}\\
+\cellcolor{gray!6}{} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{0.700000} & \cellcolor{gray!6}{100.000}\\
 
- & Spring & 0.0613 & 38.6\\
+ & Spring & 0.000266 & 0.153\\
 
-\cellcolor{gray!6}{} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{-0.0307} & \cellcolor{gray!6}{25.9}\\
+\cellcolor{gray!6}{} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{-0.172000} & \cellcolor{gray!6}{98.500}\\
 
-\multirow{-4}{*}{\centering\arraybackslash Clear skies} & Autumn & 0.3670 & 97.2\\
+\multirow{-4}{*}{\centering\arraybackslash Clear skies} & Autumn & 0.131000 & 74.000\\
 \cmidrule{1-4}
-\cellcolor{gray!6}{} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{-0.8820} & \cellcolor{gray!6}{98.9}\\
+\cellcolor{gray!6}{} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{0.074600} & \cellcolor{gray!6}{32.300}\\
 
- & Spring & -0.0991 & 37.2\\
+ & Spring & 0.046000 & 26.200\\
 
-\cellcolor{gray!6}{} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{0.0444} & \cellcolor{gray!6}{21.5}\\
+\cellcolor{gray!6}{} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{-0.038900} & \cellcolor{gray!6}{26.200}\\
 
-\multirow{-4}{*}{\centering\arraybackslash Cloudy skies} & Autumn & -0.4000 & 89.5\\
+\multirow{-4}{*}{\centering\arraybackslash Cloudy skies} & Autumn & -0.049000 & 23.300\\
 \bottomrule
@@
 over the area. Moreover, we found a similar trend under clear-sky
-conditions (\(0.35\,\%/y\)) that further supports the assumption that
+conditions (\(0.16\,\%/y\)) that further supports the assumption that
 the brightening is caused mainly by decreasing aerosols. Unfortunately,
@@
 Europe have been proposed as major factor by \citet{Wild2021}. The
-dimming effect on SDR under cloudy-sky conditions (\(-0.28\,\%/y\)),
+dimming effect on SDR under cloudy-sky conditions (\(0.022\,\%/y\)),
 suggests that cloud optical thickness is decreasing during this period.
@@
 on season, with winter showing the strongest statistically significant
-trend of \(0.69\) and \(0.83\,\%/y\) for all- and clear-skies,
+trend of \(0.69\) and \(0.7\,\%/y\) for all- and clear-skies,
 respectively, in contrast to spring and summer. The trends for autumn
-are also significant but smaller ( \(0.43\) and \(0.37\,\%/y\) for all-
+are also significant but smaller ( \(0.43\) and \(0.13\,\%/y\) for all-
 and clear-skies, respectively). Our findings are in agreement with other

### 2023-11-24 DHI_GHI_00_raw_data.R

@@ -21,13 +21,17 @@ source("./DHI_GHI_0_variables.R")
 
 #  Run data construction  ------------------------------------------------------
 
+# _ Set Importan variables for all the project  --------------------------------
+
 D_14_2 <- FALSE
 D_14   <- FALSE
 D_13   <- FALSE
+D_15   <- FALSE
 
-D_14_2 <- TRUE
+# D_14_2 <- TRUE
 # D_14   <- TRUE
 # D_13   <- TRUE
+D_15   <- TRUE
 
 TEST <- TRUE
 TEST <- FALSE
@@ -42,6 +46,13 @@ if (D_14_2) {
     common_data <- common_data_14_2
     CS_file     <- CS_file_14_2
     inpatern    <- "Clear_sky_id_Reno-Hansen_apply_v14_2_[0-9]{4}.Rds"
+}
+
+## Test for missing data
+if (D_15) {
+    common_data <- common_data_15
+    CS_file     <- CS_file_15
+    inpatern    <- "Clear_sky_id_Reno-Hansen_apply_v15_[0-9]{4}.Rds"
 }
 
 ## new implementation with corrected limits
@@ -63,6 +74,7 @@ dir.create(dirname(common_data), showWarnings = FALSE)
 dir.create("./figures",          showWarnings = FALSE)
 dir.create("./images",           showWarnings = FALSE)
 dir.create("./runtime",          showWarnings = FALSE)
+
 
 
 ##_  Check if we need to run data export  --------------------------------------
@@ -155,7 +167,7 @@ if (havetorun) {
     }
 
 
-    ##_  Skip data ranges for CM-21  -------------------------------------------
+    ## __ Skip data ranges for CM-21 --------------------------------------------
     for (as in nrow(SKIP_cm21)) {
         skip <- SKIP_cm21[as,]
         DATA[ Date >= skip$From & Date <= skip$Until, wattGLB    := NA ]
@@ -219,8 +231,8 @@ if (havetorun) {
         DATA[!QCF_GLB %in% keepQF, wattGLB := NA]
     }
 
-    ##_  Keep data characterized as 'TRUE' by Radiation Quality control v14 ----
-    if (D_14 | D_14_2) {
+    #__  Keep data characterized as 'TRUE' by Radiation Quality control v14 ----
+    if (D_14 | D_14_2 | D_15) {
         DATA[QCF_DIR == FALSE, wattDIR := NA]
[…]

### 2023-11-24 Article_A.tex
Commit 2023-11-24 14:14

@@
 \usepackage{subcaption}
+\captionsetup[sub]{position=bottom, labelfont={bf, small, stretch=1.17}, labelsep=space, textfont={small, stretch=1.17}, aboveskip=6pt,  belowskip=-6pt, singlelinecheck=off, justification=justified}
 \usepackage{booktabs}
@@
 \author{true \and true}
-\date{2023-11-03}
+\date{2023-11-24}
 
@@
 For completeness, we will provide below a brief overview of the clear sky identification algorithm (CSid), along with the site specific thresholds.
-To calculate the reference clear sky
-\(\text{SDR}_\text{CSref}\) we used the \(\text{SDR}_\text{Haurwitz}\) derived by
-the radiation model of Haurwitz (1945), adjusted for our site with a
-factor \(a\) (Eq.~\ref{eq:ahau}), resulted by an iterative optimization process, as described
-by Long and Ackerman (2000) and Reno and Hansen (2016).
-The target of the optimization was the minimization of a function \(f(a)\) (Eq.~\ref{eq:minf}) and
-was accomplished with the algorithmic function ``optimise'', which is an implementation based on the work of Brent (1973), from the library ``stats'' of the R programming language (R Core Team 2023).
+To calculate the reference clear sky \(\text{SDR}_\text{CSref}\) we used the \(\text{SDR}_\text{Haurwitz}\) derived by the radiation model of Haurwitz (1945), adjusted for our site with a factor \(a\) (Eq.~\ref{eq:ahau}), resulted by an iterative optimization process, as described by Long and Ackerman (2000) and Reno and Hansen (2016).
+The target of the optimization was the minimization of a function \(f(a)\) (Eq.~\ref{eq:minf}) and was accomplished with the algorithmic function ``optimise'', which is an implementation based on the work of Brent (1973), from the library ``stats'' of the R programming language (R Core Team 2023).
 The optimization and the selection of the clear sky reference model, was performed on SDR observations for the period 2016 - 2021.
-During the optimization, eight simple clear sky radiation models were tested (Daneshyar-Paltridge-Proctor, Kasten-Czeplak, Haurwitz, Berger-Duffie, Adnot-Bourges-Campana-Gicquel, Robledo-Soler, Kasten and Ineichen-Perez), with a wide range of factors.
+During the optimization, eight simple clear sky radiation models were tested (namely, Daneshyar-Paltridge-Proctor, Kasten-Czeplak, Haurwitz, Berger-Duffie, Adnot-Bourges-Campana-Gicquel, Robledo-Soler, Kasten and Ineichen-Perez), with a wide range of factors.
 These models are described in more details by Reno, Hansen, and Stein (2012b) and evaluated by Reno and Hansen (2016).
-We found, that Haurwitz's model, adjusted with the factor \(a = 0.965\) yields one of the lowest root mean squared errors (RMSE),
-while the procedure, manages to characterize the majority of the data.
+We found, that Haurwitz's model, adjusted with the factor \(a = 0.965\) yields one of the lowest root mean squared errors (RMSE), while the procedure, manages to characterize the majority of the data.
 Thus, our clear sky reference is derived by the Eq.~\ref{eq:ahau}.
@@
   \begin{equation}
-    X_i < 7.5 \label{eq:VSMcr}
+    X_i < 7.5\,Wm^{-2} \label{eq:VSMcr}
   \end{equation}
@@
 In the final dataset
-\(6589967\)
-one-minute measurements, of which,
 \(84.2\%\)
-were identified as under clear-sky conditions and
+of the data were identified as under clear-sky conditions and
 \(15.8\%\)
@@
 
-We calculated the linear SDR trends, from the departures of the mean daily values from the seasonal values, and for the three sky conditions (Table~\ref{tab:trendtable}).
+We calculated the linear SDR trends, from the departures of the mean daily values from the daily climatology and for the three sky conditions (Table~\ref{tab:trendtable}).
 In Figure~\ref{fig:trendALL} we present only the time series under all-sky conditions; the plots for clear-sky and cloud-sky conditions, are very similar and are shown in the Appendix (Figures~\ref{fig:trendCLEAR} and~ \ref{fig:trendCLOUD}).
 We observe a positive trend for all-sky conditions
-(\(0.38\,\%/y\)), a very close but lower trend for clear-skies (\(0.35\,\%/y\)) and a negative weaker trend for cloud-skies (\(-0.28\,\%/y\)).
+(\(0.38\,\%/y\)), a very close but smaller trend for clear-skies (\(0.35\,\%/y\)) and a negative weaker trend for cloudy-skies (\(-0.28\,\%/y\)).
 In the studied period, there is no significant break or change in the variability of the time series.
-Other studies for the European region report a change of the SDR slope, around 1980, a few years before the start of our records (Wild et al.
-2021; Yuan, Leirvik, and Wild 2021; Ohmura 2009).
-It is interesting to note, that for the observation period, the trend of the TSI is
-\(-0.0002\,\%/y\), and thus we can not attribute any major effect on SDR trend to Solar variability.
+Other studies for the European region reported a change of the SDR slope, around 1980 (Wild et al. 2021; Yuan, Leirvik, and Wild 2021; Ohmura 2009), a few years before the start of our records.
+It is interesting to note, that for the observations period, the trend of the TSI is
+\(-0.0002\,\%/y\),
+and thus we can not attribute any major effect on SDR trend to Solar variability.
 
@@
 
-Although the year-to-year variability of the anomalies (Figure \ref{fig:trendALL} and Figures \ref{fig:trendCLEAR}, \ref{fig:trendCLOUD} on Appendix), shows a rather homogeneous behavior, plots of the cumulative sums (CUSUM) of the anomalies can reveal different structures within all the three sky conditions (Regier, Briceño, and Boyer 2019).
-In the cases of all-sky and clear-sky (Figures \ref{fig:cusummonth-1} and \ref{fig:cusummonth-2}), we observe three macroscopic periods.
-A downward part from the start until about 2005, an relative steady part until about 2016 and finally, a steep upward part until the present.
-For cloud-sky (Figure~\ref{fig:cusummonth-3}), we have a different sequence, it begins with a relative steady part until 1997, followed by a upward part until 2005, a long decline until 2020, with a small positive slope until the present.
+Although the year-to-year variability of the anomalies (Figure \ref{fig:trendALL} and Figures \ref{fig:trendCLEAR}, \ref{fig:trendCLOUD} in Appendix), shows a rather homogeneous behaviour, plots of the cumulative sums (CUSUM) (Regier, Briceño, and Boyer 2019) of the anomalies can reveal different structures in the records of all three sky conditions.
+In the cases of all-sky and clear-sky conditions (Figures \ref{fig:cusummonth-1} and \ref{fig:cusummonth-2}), we observe three macroscopic periods.
+A downward part from the start until about 2005, a relatively steady part until about 2016 and, finally, a steep upward part until the present.
+For cloud-sky (Figure~\ref{fig:cusummonth-3}), we have a different pattern; it begins with a relatively steady part until 1997, followed by an upward part until 2005, and a long decline until 2020, with a small positive slope until the present.
 For a uniform trend, we would expect the CUSUMs of the anomalies to have a symmetric `V' shape.
 This would indicate that the anomalies are evenly distributed around the climatological mean, and for a positive uniform trend, the first half to be below and the other half above the climatological mean.
-In our case, there is more complex evolution of the phenomenon.
+In our case, there is a more complex evolution of the anomalies.
 Another distinct feature of the CUSUMs, is the different pattern of the cloudy-sky dataset which peaks around the middle of the period.
-Although, it seems to exist a complementary relation to the CUSUMs of the clear- and all-sky cases, we can not assert that clouds are the main drive for this relation, due to the great difference in the number of observational data between the datasets (Table~\ref{tab:trendtable}).
+Although, there seems to exist a complementary relation to the CUSUMs of the clear- and all-sky cases, we can not assert that clouds are the main driver for this relation due to the great difference in the number of observational data between the two datasets (Table~\ref{tab:trendtable}).
 
@@
 
-In order to investigate more, the features within the trends, we created another set of CUSUMs plots by subtracting the corresponding long term trend from the SDR anomaly data, prior to the CUSUM calculation (Figure~\ref{fig:cusumnotrendmonthly}).
-With this approach periods when the CUSUMs diverge from zero can be interpreted as systematic variation of SDR from the climatological mean.
-When the CUSUM is increasing, the added values are above the climatological values of SDR trend and vice versa.
-Overall, for all- and clear-sky conditions (Figure~\ref{fig:cusumnotrendmonthly-1} and~\ref{fig:cusumnotrendmonthly-2}) we observe periods when the anomalies diverge from the climatological value, each lasting for several years.
+In order to investigate further the features of the CUSUMs, we created another set of CUSUM plots by subtracting the corresponding long term trend from the SDR anomaly data, prior to the CUSUM calculation (Figure~\ref{fig:cusumnotrendmonthly}).
+With this approach periods when the CUSUMs diverge from zero can be interpreted as a systematic variation of SDR from the climatological mean.
+When the CUSUM is increasing, the added values are above the climatological values of the SDR trend and vice versa.
+Overall, for all- and clear-sky conditions (Figures~\ref{fig:cusumnotrendmonthly-1} and~\ref{fig:cusumnotrendmonthly-2}) we observe periods when the anomalies diverge from the climatological value, each lasting for several years.
 The pattern for the all-sky and clear-sky conditions is very similar, suggesting prevalence in clear skies over Thessaloniki.
@@
 
-The solar zenith angle is a major driver of the attenuation of SDR reaching the ground, as SDR is affected by the enhancement of the radiation path in the atmosphere, especially in urban environment with human activities emitting aerosols (Wang et al.
+The solar zenith angle is a major factor of SDR reaching the ground, due to the enhancement of the radiation path in the atmosphere, especially in urban environments with human activities emitting aerosols (Wang et al.
 2021).
@@
 2021).
-For cloud-sky conditions (Figure~\ref{fig:szatrends-3}), we can not discern any significant dependence of the SDR trend with SZA.
+For cloudy-sky conditions (Figure~\ref{fig:szatrends-3}), we can not discern any significant dependence of the SDR trend with SZA.
 For SZAs \(16^\circ\) - \(50^\circ\), the trends range within about \(\pm 0.2\,\%/y\), with a weak statistical significance.
@@
 
-Similar to the long term trends discussed above, we have calculated the trend of the anomalies for the three different sky conditions, and for each season of the year, using the corresponding mean monthly values
+Similarly to the long term trends discussed above, we have calculated the trend of the anomalies for the three different sky conditions, and for each season of the year, using the corresponding mean monthly values
 (Figure~\ref{fig:seasonalALL} and Table~\ref{tab:trendseasontable}).
@@
 (\(0.43\,\%/y\),
-value similar to the long term trend) both statistically significant above the \(99\,\%\) confidence level.
-In spring and summer, the trends are much smaller and less statistical significant.
-These seasonal differences indicate a possible relation of the trends to clouds during winter and autumn.
+a value close to the long term trend) both statistically significant above the \(99\,\%\) confidence level.
+In spring and summer, the trends are much smaller and of lesser statistical significance.
+These seasonal differences indicate a possible relation of the trends in SDR to trends of clouds during winter and autumn.
 For clear-skies, the trend in winter is \(0.83\,\%/y\), larger than for all-skies (\(0.69\,\%/y\)), which is another indication of increasing cloud thickness.

### 2023-11-29 SUBMISSION_01/MDPI_submission.tex
Commit 2023-11-29 11:29

--- 2023-11-09 Build_163/MDPI_submission_B163.tex
@@
 \Author{Athanasios
-Natsis$^{1,*}$\href{https://orcid.org/0000-0002-5199-4119}
-{\orcidicon}, Alkiviadis Bais$^{1}$, Charikleia Meleti$^{1}$}
+Natsis$^{1}$\href{https://orcid.org/0000-0002-5199-4119}
+{\orcidicon}, Alkiviadis Bais$^{1,*}$, Charikleia Meleti$^{1}$}
 
@@
 $^{1}$ \quad Aristotle University of Thessaloniki - Laboratory of
-Atmospheric Physics, Campus Box 149, 54124 Thessaloniki, Greece; \\
+Atmospheric Physics, Campus Box 149, 54124 Thessaloniki,
+Greece; \href{mailto:natsisphysicist@gmail.com}{\nolinkurl{natsisphysicist@gmail.com}}
+(A.N.); \href{mailto:abais@auth.gr}{\nolinkurl{abais@auth.gr}} (A.B.);
+\href{mailto:meleti@auth.gr}{\nolinkurl{meleti@auth.gr}} (C.M.)\\
 }

### 2023-11-29 DHI_GHI_1_longterm_trends.R
Commit 2023-11-29 13:10

@@ -82,6 +82,7 @@ panderOptions("table.split.table",        120   )
 ## Functions from `https://github.com/thanasisn/IStillBreakStuff/tree/main/FUNCTIONS/R`
 source("~/CODE/FUNCTIONS/R/sumNA.R")
 source("~/CODE/FUNCTIONS/R/linear_fit_stats.R")
+source("~/CODE/FUNCTIONS/R/cor_test_stats.R")
 source("~/CODE/FUNCTIONS/R/trig_deg.R")
 source("~/CODE/FUNCTIONS/R/data.R")
 
@@ -246,11 +247,14 @@ for (DBn in dbs) {
 
             ## linear model by day step
             lm1 <- lm(dataset[[avar]] ~ dataset$Date)
+            ## correlation test
+            cor1 <- cor.test(x = dataset[[avar]], y = as.numeric(dataset$Date), method = 'pearson')
 
             ## capture lm for table
             gather <- rbind(gather,
                             data.frame(
                                 linear_fit_stats(lm1, confidence_interval = Daily_confidence_limit),
+                                cor_test_stats(cor1),
                                 DATA      = DBn,
                                 var       = avar,
                                 N         = sum(!is.na(dataset[[avar]]))
@@ -359,22 +363,20 @@ for (DBn in dbs) {
 
             ## linear model by day step
             lm1  <- lm(dataset[[avar]] ~ dataset$Date)
-
+            ## correlation test
             cor1 <- cor.test(x = dataset[[avar]], y = as.numeric(dataset$Date), method = 'pearson')
 
-
-            data.frame(cor1)
-
-            stop("D")
             ## capture lm for table
             gather <- rbind(gather,
                             data.frame(
                                 linear_fit_stats(lm1, confidence_interval = Daily_confidence_limit),
+                                cor_test_stats(cor1),
                                 DATA      = DBn,
                                 var       = avar,
                                 N         = sum(!is.na(dataset[[avar]])),
                                 Obs       = sum(dataset[[sub("_des","_N", avar)]], na.rm = TRUE)
-                            ))
+                            )
+            )
 
             par("mar" = c(3, 4, 2, 1))
 
@@ -1276,13 +1278,16 @@ for (DBn in dbs) {
             if (sum(!is.na(dataset[[avar]])) <= 1) next()
 
             ## linear model counting years
-            lm2 <- lm( dataset[[avar]] ~ dataset$Year )
+            lm2 <- lm(dataset[[avar]] ~ dataset$Year)
+            ## correlation test
[…]

### 2023-11-29 MDPI_submission.Rmd
Commit 2023-11-29 14:14

@@
 TSI$Obs                <- NULL
+
+TSI$cor.null_value     <- NULL
+TSI$cor.alternative    <- NULL
+TSI$cor.method         <- NULL
+TSI$cor.data_name      <- NULL
+TSI$cor.p              <- NULL
+care <- grep("^cor\\.", names(TSI), value = T )
+for (an in care) {
+    TSI[, (an) := NULL ]
+}
+
 setorder(TSI, -slope.stat_sig)

### 2023-11-29 MDPI_submission.Rmd
Commit 2023-11-29 18:14

@@
 TSI$cor.p              <- NULL
-care <- grep("^cor\\.", names(TSI), value = T )
+TSI$conf.level         <- NULL
+care <- grep("^cor\\.", names(TSI), value = T)
 for (an in care) {

### 2023-11-29 MDPI_submission.Rmd
Commit 2023-11-29 19:09

@@
 rownames(LT) <- NULL
+
 names(LT)    <- c("Trend [%/year]",

### 2023-11-29 MDPI_submission.Rmd
Commit 2023-11-29 20:14

@@
 TSI$Obs                <- NULL
+TSI$Rsqrd              <- NULL
+TSI$RsqrdAdj           <- NULL
 
@@
 rownames(LT) <- NULL
+
+LT$cor.df           <- NULL
+LT$cor.null_value   <- NULL
+LT$cor.alternative  <- NULL
+LT$cor.method       <- NULL
+LT$cor.data_name    <- NULL
+LT$cor.p            <- NULL
 

### 2023-11-29 MDPI_submission.Rmd
Commit 2023-11-29 21:09

@@
 LT$cor.p            <- NULL
-
-names(LT)    <- c("Trend [%/year]",
-                  "Trend p-value",
-                  "Sky conditions",
-                  "Days with data",
-                  "Statistical signif. [%]")
-## no need all are 100%
-LTminSSig <- signif(min(LT$`Statistical signif. [%]`), 4)
-# LT$`Trend statistical signif. [%]` <- NULL
-## numeric precision
-LT$`Trend [%/year]` <- signif(LT$`Trend [%/year]`, 3)
-LT$`Trend p-value`  <- signif(LT$`Trend p-value`,  3)
+LT$conf.level       <- NULL
 
 ## some vars to use later
-alltrend   <- signif(LT[`Sky conditions` == "All sky cond.",    `Trend [%/year]`], 2)
-cleartrend <- signif(LT[`Sky conditions` == "Clear sky cond.",  `Trend [%/year]`], 2)
-cloudtrend <- signif(LT[`Sky conditions` == "Cloudy sky cond.", `Trend [%/year]`], 2)
+alltrend   <- signif(LT[DATA == "All sky cond.",    slope], 2)
+cleartrend <- signif(LT[DATA == "Clear sky cond.",  slope], 2)
+cloudtrend <- signif(LT[DATA == "Cloudy sky cond.", slope], 2)
+LTminSSig  <- signif(min(LT$slope.stat_sig), 4)
+
+## numeric precision of table
+LT$slope    <- signif(LT$slope,   3)
+LT$slope.p  <- signif(LT$slope.p, 3)
 ## rename sky conditions
-LT$`Sky conditions` <- sub(" cond.", "", LT$`Sky conditions`)
+LT$DATA <- sub(" cond.", "", LT$DATA)
+
+# 
+# names(LT)    <- c("Trend [%/year]",
+#                   "Trend p-value",
+#                   "Sky conditions",
+#                   "Days with data",
+#                   "Statistical signif. [%]")
+
+## no need all are 100%
+
+# LT$`Trend statistical signif. [%]` <- NULL
 
 ## order columns
-LT1 <- LT[,c(
-      "Sky conditions",
-      "Trend [%/year]",
-      "Trend p-value",
-      "Days with data"
-      )]
+# LT1 <- LT[,c(
+#       "Sky conditions",
+#       "Trend [%/year]",
+#       "Trend p-value",
+#       "Days with data"
+#       )]
 
 ## order columns
-LT2 <- LT[,c(
-      "Sky conditions",
-      "Trend [%/year]",
-      "Statistical signif. [%]",
-      "Days with data"
-      )]
+# LT2 <- LT[,c(
+#       "Sky conditions",
+#       "Trend [%/year]",
+#       "Statistical signif. [%]",
+#       "Days with data"
+#       )]
+LT2 <- LT
 ```
@@
 spp$var                <- NULL
+
+spp$cor.df             <- NULL
+spp$cor.null_value     <- NULL
+spp$cor.alternative    <- NULL
+spp$cor.method         <- NULL
+spp$cor.data_name      <- NULL
+spp$cor.p              <- NULL
+spp$conf.level         <- NULL
+
 rownames(spp)          <- NULL
-names(spp) <- c("Trend [%/year]", "Trend p-value", "Sky condition", "Season", "Statistical signif. [%]")
+
 ## order by season
@@
 ## order by sky condition
-spp <- spp[order(spp$`Sky condition`),]
+spp <- spp[order(spp$DATA), ]
+
+## numeric precision
+spp$slope          <- signif(spp$slope,          3)
+spp$slope.p        <- signif(spp$slope.p,        3)
+spp$slope.stat_sig <- signif(spp$slope.stat_sig, 3)
+
+## rename sky conditions
+spp$DATA <- sub(" cond.", "", spp$DATA)
+## rename conditions
+spp$DATA <- sub(" sky", " skies", spp$DATA)
+
+# names(spp) <- c("Trend [%/year]", "Trend p-value", "Sky condition", "Season", "Statistical signif. [%]")
 
@@
 # spp <- spp[c(A,B),]
-## numeric precision
-spp$`Trend [%/year]`          <- signif(spp$`Trend [%/year]`, 3)
-spp$`Trend p-value`           <- signif(spp$`Trend p-value`,  3)
-spp$`Statistical signif. [%]` <- signif(spp$`Statistical signif. [%]`,  3)
 
-## rename sky conditions
-spp$`Sky condition` <- sub(" cond.", "", spp$`Sky condition`)
 ```
@@
 ## arrange columns
-spp <- spp[,c(
-      "Sky condition",
-      "Season",
-      "Trend [%/year]",
-      # "Trend p-value",
-      "Statistical signif. [%]"
-      )]
+# spp <- spp[,c(
+#       "Sky condition",
+#       "Season",
+#       "Trend [%/year]",
+#       # "Trend p-value",
+#       "Statistical signif. [%]"
+#       )]
+
 spp2 <- spp
-## rename conditions
-spp2$`Sky condition` <- sub(" sky", " skies", spp2$`Sky condition`)
 

### 2023-11-29 Article.Rmd
Commit 2023-11-29 22:14

@@
   Periods of positive trends are referred as brightening periods and of negative trends as dimming periods.
-  Here we studied 29 years of pyranometer data in Thessaloniki, Greece, under three types of sky conditions (clear sky, cloudy sky and all sky).
+  Here we studied 30 years of pyranometer data in Thessaloniki, Greece, under three types of sky conditions (clear sky, cloudy sky and all sky).
   The clear-sky data were identified by applying a cloud screening algorithm.
 We found a positive trend of $0.38\,\%/\text{year}$ for all-sky, $0.35\,\%/\text{year}$ for clear-sky conditions, and $-0.28\,\%/\text{year}$ for cloudy conditions.
-  We have also investigated the consistency of these trends, the effect of the solar zenith angle, and their seasonal variability.
-  Finally, we identified some anomalies in the long term SDR trends for all sky conditions.
+  We have also investigated the consistency of these trends, their seasonal variability, and the effect of the solar zenith angle. 
+  We have found that for all-sky and clear-sky conditions the SDR trend is positive in winter ($0.7$ and $0.8\,\%/\text{year}$, respectively) and autumn ($~0.4\,\%/\text{year}$), while under cloudy skies the trend is negative ($-0.9\,\%/\text{year}$ in winter and $-0.4\,\%/\text{year}$ in autumn).
+  In spring and summer the trend is very close to zero, irrespective of sky conditions. 
+  The SDR trend is increasing with increasing solar zenith angle, except under cloudy skies where the trend is highly variable and close to zero.
+  Finally, we identified some anomalies in the long term SDR trends for all sky conditions by examining the cumulative sums of monthly anomalies from the climatological mean.
 
@@
 keywords: |
-  GHI; SDR; Solar Brigthening/Dimming.
-## list three to ten pertinent keywords specific to the article, yet reasonably common within the subject discipline
+  GHI; SDR; solar radiation; Solar Brigthening/Dimming; aerosols; clouds.
 
-acknowledgement: |
-  All sources of funding of the study should be disclosed. Please clearly
-  indicate grants that you have received in support of your research work.
-  Clearly state if you received funds for covering the costs to publish in open
-  access.
+# acknowledgement: |
+#   All sources of funding of the study should be disclosed. Please clearly
+#   indicate grants that you have received in support of your research work.
+#   Clearly state if you received funds for covering the costs to publish in open
+#   access.
 
-authorcontributions: |
-  For research articles with several authors, a short paragraph specifying their
-  individual contributions must be provided. The following statements should be
-  used ``X.X. and Y.Y. conceive and designed the experiments; X.X. performed the
-  experiments; X.X. and Y.Y. analyzed the data; W.W. contributed
-  reagents/materials/analysis tools; Y.Y. wrote the paper.'' Authorship must be
-  limited to those who have contributed substantially to the work reported.
+# authorcontributions: |
+#   For research articles with several authors, a short paragraph specifying their
+#   individual contributions must be provided. The following statements should be
+#   used ``X.X. and Y.Y. conceive and designed the experiments; X.X. performed the
+#   experiments; X.X. and Y.Y. analyzed the data; W.W. contributed
+#   reagents/materials/analysis tools; Y.Y. wrote the paper.'' Authorship must be
+#   limited to those who have contributed substantially to the work reported.
 
@@
    - \captionsetup[sub]{position=bottom, labelfont={bf, small, stretch=1.17}, labelsep=space, textfont={small, stretch=1.17}, aboveskip=6pt,  belowskip=-6pt, singlelinecheck=off, justification=justified}
+   - \usepackage{placeins}
 
@@
 TSI$Obs                <- NULL
+TSI$Rsqrd              <- NULL
+TSI$RsqrdAdj           <- NULL
+
+TSI$cor.null_value     <- NULL
+TSI$cor.alternative    <- NULL
+TSI$cor.method         <- NULL
+TSI$cor.data_name      <- NULL
+TSI$cor.p              <- NULL
+TSI$conf.level         <- NULL
+care <- grep("^cor\\.", names(TSI), value = T)
+for (an in care) {
+    TSI[, (an) := NULL ]
+}
+
 setorder(TSI, -slope.stat_sig)
@@
 The target of the optimization was the minimization of a function $f(a)$ (Eq.\ \ref{eq:minf}) and was accomplished with the algorithmic function "optimise", which is an implementation based on the work of @Brent1973, from the library "stats" of the R programming language [@RCT2023].
+<!-- cost function -->
+\begin{equation}
+f(a) = \frac{1}{n}\sum_{i=1}^{n} ( \text{SDR}_{\text{CSid},i} - a \times \text{SDR}_{\text{testCSref},i} )^2 \label{eq:minf}
+\end{equation}
+where: $n$ is the total number of daylight records, $\text{SDR}_{\text{CSid},i}$ are the records identified as clear sky by CSid, $a$ is a hypothetical adjustment factor, and $\text{SDR}_{\text{testCSref},i}$ is any of the tested clear sky radiation models.
+
 The optimization and the selection of the clear sky reference model, was performed on SDR observations for the period 2016 - 2021.
@@
 rownames(LT) <- NULL
-names(LT)    <- c("Trend [%/year]",
-                  "Trend p-value",
-                  "Sky conditions",
-                  "Days with data",
-                  "Statistical signif. [%]")
-## no need all are 100%
-LTminSSig <- signif(min(LT$`Statistical signif. [%]`), 4)
-# LT$`Trend statistical signif. [%]` <- NULL
-## numeric precision
-LT$`Trend [%/year]` <- signif(LT$`Trend [%/year]`, 3)
-LT$`Trend p-value`  <- signif(LT$`Trend p-value`,  3)
+
+LT$cor.df           <- NULL
+LT$cor.null_value   <- NULL
+LT$cor.alternative  <- NULL
+LT$cor.method       <- NULL
+LT$cor.data_name    <- NULL
+LT$cor.p            <- NULL
+LT$conf.level       <- NULL
 
 ## some vars to use later
-alltrend   <- signif(LT[`Sky conditions` == "All sky cond.",    `Trend [%/year]`], 2)
-cleartrend <- signif(LT[`Sky conditions` == "Clear sky cond.",  `Trend [%/year]`], 2)
-cloudtrend <- signif(LT[`Sky conditions` == "Cloudy sky cond.", `Trend [%/year]`], 2)
+alltrend   <- signif(LT[DATA == "All sky cond.",    slope], 2)
+cleartrend <- signif(LT[DATA == "Clear sky cond.",  slope], 2)
+cloudtrend <- signif(LT[DATA == "Cloudy sky cond.", slope], 2)
+LTminSSig  <- signif(min(LT$slope.stat_sig), 4)
+
+## numeric precision of table
+LT$slope    <- signif(LT$slope,   3)
+LT$slope.p  <- signif(LT$slope.p, 3)
 ## rename sky conditions
-LT$`Sky conditions` <- sub(" cond.", "", LT$`Sky conditions`)
+LT$DATA <- sub(" cond.", "", LT$DATA)
+LT$DATA <- sub(" sky", " skies", LT$DATA)
+
+# 
+# names(LT)    <- c("Trend [%/year]",
+#                   "Trend p-value",
+#                   "Sky conditions",
+#                   "Days with data",
+#                   "Statistical signif. [%]")
+
+## no need all are 100%
+
+# LT$`Trend statistical signif. [%]` <- NULL
 
@@
 ```{r trendtable, echo=F}
+
 x_latex <- knitr::kable(LT2,
@@
 Overall, for all- and clear-sky conditions (Figures\ \ref{fig:cusumnotrendmonthly-1} and\ \ref{fig:cusumnotrendmonthly-2}) we observe periods when the anomalies diverge from the climatological value, each lasting for several years.
-The pattern for the all-sky and clear-sky conditions is very similar, suggesting prevalence in clear skies over Thessaloniki.
-It is interesting that in the period 1993 - 2016 the anomalies of all-sky and clear-sky, have a high variability around zero.
-After 2016, the range of the variability is decreased at about one third of the prior period.
-For cloud-sky conditions (Figure\ \ref{fig:cusumnotrendmonthly-3}) the period 1997 - 2008 is dominated by positive CUSUMs suggesting a reduced effect of clouds on SDR.
-From 1997 to mid-2000s CUSUMs are increasing, likely due to a continuous decrease in the thickness of clouds, followed by a period of rapid increase (within 3 years) in cloud optical thickness lasting up to 2008.
+The pattern in both datasets is very similar, suggesting prevalence in clear skies over Thessaloniki.
+It is interesting that in the period 1993 - 2016 the anomalies have a high variability around zero, while after 2016, the range of the variability is decreased to about one third of the prior period.
+For cloudy-sky conditions (Figure\ \ref{fig:cusumnotrendmonthly-3}) the period 1997 - 2008 is dominated by positive CUSUMs, suggesting a reduced effect of clouds on SDR.
+From 1997 to mid-2000s CUSUMs are increasing, likely due to a continuous decrease in the optical thickness of clouds, followed by a period of rapid increase (within 3 years) in cloud optical thickness lasting up to 2008.
 The following stable period spans for about 15 years up to 2021 when CUSUMs start increasing again.
@@
 spp$var                <- NULL
+
+spp$cor.df             <- NULL
+spp$cor.null_value     <- NULL
+spp$cor.alternative    <- NULL
+spp$cor.method         <- NULL
+spp$cor.data_name      <- NULL
+spp$cor.p              <- NULL
+spp$conf.level         <- NULL
+
 rownames(spp)          <- NULL
-names(spp) <- c("Trend [%/year]", "Trend p-value", "Sky condition", "Season", "Statistical signif. [%]")
+
 ## order by season
 target    <- c("Winter", "Spring", "Summer", "Autumn")
-new_order <- sapply(target, function(x,spp){which(spp$Season == x)}, spp = spp)
+new_order <- sapply(target, function(x, spp){which(spp$Season == x)}, spp = spp)
 spp       <- spp[new_order,]
 ## order by sky condition
-spp <- spp[order(spp$`Sky condition`),]
+spp <- spp[order(spp$DATA), ]
+
+## numeric precision
+spp$slope          <- signif(spp$slope,          3)
+spp$slope.p        <- signif(spp$slope.p,        3)
+spp$slope.stat_sig <- signif(spp$slope.stat_sig, 3)
+
+## rename sky conditions
+spp$DATA <- sub(" cond.", "", spp$DATA)
+## rename conditions
+spp$DATA <- sub(" sky", " skies", spp$DATA)
+
+spp <- rename(spp, "Trend [%/year]" = slope)
+spp <- rename(spp, "Sky condition"  = DATA )
+
+# names(spp) <- c("Trend [%/year]", "Trend p-value", "Sky condition", "Season", "Statistical signif. [%]")
 
@@
 For all-sky conditions the trend in SDR in winter is the largest
-($`r signif(spp$"Trend [%/year]"[spp$Season == "Winter" & spp$"Sky condition" == "All sky"], 2)`\,\%/y$),
+($`r signif(spp$"Trend [%/year]"[spp$Season == "Winter" & spp$"Sky condition" == "All skies"], 2)`\,\%/y$),
 followed by the trend in autumn
-($`r signif(spp$"Trend [%/year]"[spp$Season == "Autumn" & spp$"Sky condition" == "All sky"], 2)`\,\%/y$,
+($`r signif(spp$"Trend [%/year]"[spp$Season == "Autumn" & spp$"Sky condition" == "All skies"], 2)`\,\%/y$,
 a value close to the long term trend) both statistically significant above the $99\,\%$ confidence level.
@@
 These seasonal differences indicate a possible relation of the trends in SDR to trends of clouds during winter and autumn.
-For clear-skies, the trend in winter is $`r signif(spp$"Trend [%/year]"[spp$Season == "Winter" & spp$"Sky condition" == "Clear sky"], 2)`\,\%/y$, larger than for all-skies ($`r signif(spp$"Trend [%/year]"[spp$Season == "Winter" & spp$"Sky condition" == "All sky"], 2)`\,\%/y$), which is another indication of increasing cloud thickness.
-Moreover, the trends under clear- and cloud-sky conditions are almost complementary to each other, particularly for winter and autumn, where the signal is stronger.
+For clear-skies, the trend in winter is $`r signif(spp$"Trend [%/year]"[spp$Season == "Winter" & spp$"Sky condition" == "Clear skies"], 2)`\,\%/y$, larger than for all-skies ($`r signif(spp$"Trend [%/year]"[spp$Season == "Winter" & spp$"Sky condition" == "All skies"], 2)`\,\%/y$), which is another indication of a decreasing trend in cloud optical thickness.
+Moreover, the trends under clear- and cloudy-sky conditions are almost complementary to each other, particularly for winter and autumn, where the signal is stronger.
 During spring and summer the statistical significance is very low and the actual trend too small for a meaningful comparison.
 
-```{r seasonalALL, echo=F, fig.cap="SDR trends by season of the year for the three sky conditions.
-Each point is a monthly mean calculated from daily means.
-The vertical scale is common for all graphs, each column of graphs corresponds to a sky condition and each row to a season of the year.", out.width="100%"}
-knitr::include_graphics("./images/SeasonalTrendsTogether3-2.png")
-```
+\begin{figure}[h!]
+    \begin{adjustwidth}{-\extralength}{0cm}
+        {\centering 
+            \includegraphics[width=1\linewidth]{./images/SeasonalTrendsTogether3-2} 
+        }
+        \caption{Linear trends (black lines) of monthly mean anomalies of SDR by season (rows of plots) for the three sky conditions (columns of plots).}\label{fig:seasonalALL}
+    \end{adjustwidth}
+\end{figure}
+
+<!-- ```{r seasonalALL, echo=F, fig.cap="Linear trends (black lines) of monthly mean anomalies of SDR by season (rows of plots) for the three sky conditions (columns of plots).", out.width="100%"} -->
+<!-- knitr::include_graphics("./images/SeasonalTrendsTogether3-2.png") -->
+<!-- ``` -->
 
@@
 
-<!-- \footnotesize -->
+```{r trendseasontable, echo=F}
 
-```{r trendseasontable, echo=F}
 ## arrange columns
-spp <- spp[,c(
-      "Sky condition",
-      "Season",
-      "Trend [%/year]",
-      # "Trend p-value",
-      "Statistical signif. [%]"
-      )]
+# spp <- spp[,c(
+#       "Sky condition",
+#       "Season",
+#       "Trend [%/year]",
+#       # "Trend p-value",
+#       "Statistical signif. [%]"
+#       )]
 
-rownames(spp) <- NULL
+spp2 <- spp
+
+rownames(spp2) <- NULL
 # panderOptions("table.alignment.default", "left")
@@
 ```
-<!-- \normalsize -->
 
@@
 
-We demonstrate that in the period
-`r paste(year(min(ALL_1_daily_DESEAS$Date)), " - ", year(max(ALL_1_daily_DESEAS$Date)))`, there is a brightening trend of SDR for all-sky conditions ($`r alltrend`\,\%/y$).
-On a previous study [@Bais2013], for the period 1993 - 2011 the trend was found to be $0.33\,\%/y$.
-The increase of this trend, indicates that the phenomenon and probably the causes, are still occurring.
-Accordingly, we found a similar trend for clear-sky conditions ($`r cleartrend`\,\%/y$) that further supports that the brightening can be independent of the cloud occurrences, although, cloud formation can also attenuate SDR.
-We have calculated the effect of TSI variability to be a negligible factor in SDR variability.
-Unfortunately, there is no available data for this period for the aerosols, in order to directly estimate their effect on SDR, which is the most probable main cause.
+We have demonstrated that in the period
+`r paste(year(min(ALL_1_daily_DESEAS$Date)), " - ", year(max(ALL_1_daily_DESEAS$Date)))`, there is a positive trend in SDR of ($`r alltrend`\,\%/y$) (brightening) (positive trend) in Thessaloniki, Greece, under all-sky conditions.
+A previous study [@Bais2013] for the period 1993 - 2011 found also a positive trend of $0.33\,\%/y$.
+The increase of this trend indicates that the brightening of SDR continues and is probably caused by continuing decreases in aerosol optical depth and the optical thickness of clouds over the area.
+Moreover, we found a similar trend under clear-sky conditions ($`r cleartrend`\,\%/y$) that further supports the assumption that the brightening is caused mainly by decreasing aerosols.
+Unfortunately, for the entire period there is no available data for the aerosols, in order to quantify their effect on SDR. 
+However, @Siomos2020 have shown that aerosol optical depth over Thessaloniki is decreasing constantly at least up to 2018.
 The attenuation of SDR by aerosols over Europe have been proposed as major factor by @Wild2021.
-The dimming effect of SDR under cloud-sky condition ($`r cloudtrend`\,\%/y$), shows that clouds can also attenuate the amount of SDR reaching the ground, and changes on their occurrences can effect the net energy balance.
-Because we have no adequate data to investigate the long term changes of cloud formation in the region, we can not infer if the SDR trend we observe under cloud-sky is biased by changes in clouds and local climate.
-The observed brightening of SDR over Thessaloniki; is dependent on SZA (higher SZA, have stronger brightening trends), and also, with the seasons of the year, where Winter and Autumn show a significant strong trend, in contrast to Spring and Summer.
-<!-- to conclutions? -->
-Using CUSUMs of the trends for all- and clear-skies, we observed periods where the CUSUMs remain relative steady,
-<!-- roughly around the middle of the study dataset,--> with a steep decline previous to that, and a steep incline after.
-This is an indication that the whole brightening phenomenon does not follow a smooth development over time.
+The dimming effect on SDR under cloudy-sky conditions ($`r cloudtrend`\,\%/y$), suggests that cloud optical thickness is decreasing during this period.
+Because we have no adequate data to investigate the long term changes of cloud thickness in the region, we cannot verify if the negative SDR trend we observe under under cloudy-skies can be attributed solely to changes in clouds.
 
-Our findings are in agreement with other researchers for the region.
-Future observations from the collocated pyrheliometer, may provide us with the means to further investigate the variability of Solar radiation at ground level in Thessaloniki.
-Also, more data (cloud occurrence, aerosols, atmospheric water vapour, etc.), will help us to determine the extent of SDR attenuation by each factor.
+The observed brightening on SDR over Thessaloniki is dependent on SZA (larger SZAs lead to stronger brightening).
+The trend is also dependent on season, with winter showing the strongest statistically significant trend of
+$`r signif(spp$"Trend [%/year]"[spp$Season == "Winter" & spp$"Sky condition" == "All skies"], 2)`$
+and
+$`r signif(spp$"Trend [%/year]"[spp$Season == "Winter" & spp$"Sky condition" == "Clear skies"], 2)`\,\%/y$
+for all- and clear-skies, respectively, in contrast to spring and summer.
+The trends for autumn are also significant but smaller (
+$`r signif(spp$"Trend [%/year]"[spp$Season == "Autumn" & spp$"Sky condition" == "All skies"], 2)`$
+and
+$`r signif(spp$"Trend [%/year]"[spp$Season == "Autumn" & spp$"Sky condition" == "Clear skies"], 2)`\,\%/y$
+for all- and clear-skies, respectively).
+Our findings are in agreement with other studies for the region.
+
+Using the CUSUMs of the monthly departures for all- and clear-skies, we observed periods where the CUSUMs remain relatively stable, with a steep decline before and a steep increase after. 
+This is an indication that the whole brightening effect does not follow a smooth development over time.
+
+Continued observations with a collocated pyrheliometer, which started in 2016, will allow us to further investigate the variability of solar radiation at ground level in Thessaloniki. 
+Also, additional data of cloudiness, aerosols, atmospheric water vapour, etc., will allow better attribution and quantification of the effects of these factors on SRD.
 

### 2023-11-29 MDPI_submission.Rmd
Commit 2023-11-29 22:14

@@
 LT$conf.level       <- NULL
+LT$cor.conf_int_1   <- NULL
+LT$cor.conf_int_2   <- NULL
 
@@
 LT$DATA <- sub(" cond.", "", LT$DATA)
+LT$DATA <- sub(" sky", " skies", LT$DATA)
 
@@
 ```{r trendtable, echo=F}
-LT2$`Sky conditions` <- sub(" sky", " skies", LT2$`Sky conditions`)
+
 x_latex <- knitr::kable(LT2,
@@
 spp$conf.level         <- NULL
+spp$cor.conf_int_1     <- NULL
+spp$cor.conf_int_2     <- NULL
 
@@
 
+spp <- rename(spp, "Trend [%/year]" = slope)
+spp <- rename(spp, "Sky condition"  = DATA )
+
 # names(spp) <- c("Trend [%/year]", "Trend p-value", "Sky condition", "Season", "Statistical signif. [%]")
@@
 These seasonal differences indicate a possible relation of the trends in SDR to trends of clouds during winter and autumn.
-For clear-skies, the trend in winter is $`r signif(spp$"Trend [%/year]"[spp$Season == "Winter" & spp$"Sky condition" == "Clear sky"], 2)`\,\%/y$, larger than for all-skies ($`r signif(spp$"Trend [%/year]"[spp$Season == "Winter" & spp$"Sky condition" == "All sky"], 2)`\,\%/y$), which is another indication of a decreasing trend in cloud optical thickness.
+For clear-skies, the trend in winter is $`r signif(spp$"Trend [%/year]"[spp$Season == "Winter" & spp$"Sky condition" == "Clear skies"], 2)`\,\%/y$, larger than for all-skies ($`r signif(spp$"Trend [%/year]"[spp$Season == "Winter" & spp$"Sky condition" == "All skies"], 2)`\,\%/y$), which is another indication of a decreasing trend in cloud optical thickness.
 Moreover, the trends under clear- and cloudy-sky conditions are almost complementary to each other, particularly for winter and autumn, where the signal is stronger.
@@
 
-<!-- \footnotesize -->
+```{r trendseasontable, echo=F}
 
-```{r trendseasontable, echo=F}
 ## arrange columns
@@
 The trend is also dependent on season, with winter showing the strongest statistically significant trend of
-$`r signif(spp$"Trend [%/year]"[spp$Season == "Winter" & spp$"Sky condition" == "All sky"], 2)`$
+$`r signif(spp$"Trend [%/year]"[spp$Season == "Winter" & spp$"Sky condition" == "All skies"], 2)`$
 and
-$`r signif(spp$"Trend [%/year]"[spp$Season == "Winter" & spp$"Sky condition" == "Clear sky"], 2)`\,\%/y$
+$`r signif(spp$"Trend [%/year]"[spp$Season == "Winter" & spp$"Sky condition" == "Clear skies"], 2)`\,\%/y$
 for all- and clear-skies, respectively, in contrast to spring and summer.
 The trends for autumn are also significant but smaller (
-$`r signif(spp$"Trend [%/year]"[spp$Season == "Autumn" & spp$"Sky condition" == "All sky"], 2)`$
+$`r signif(spp$"Trend [%/year]"[spp$Season == "Autumn" & spp$"Sky condition" == "All skies"], 2)`$
 and
-$`r signif(spp$"Trend [%/year]"[spp$Season == "Autumn" & spp$"Sky condition" == "Clear sky"], 2)`\,\%/y$
+$`r signif(spp$"Trend [%/year]"[spp$Season == "Autumn" & spp$"Sky condition" == "Clear skies"], 2)`\,\%/y$
 for all- and clear-skies, respectively).

### 2023-11-29 MDPI_submission.tex
Commit 2023-11-29 22:14

--- 2023-11-09 MDPI_submission.tex
@@
 \caption{\label{tab:trendtable}Trends in SDR daily means for different sky conditions for the period 1993 - 2023.}
-\begin{tabu} to \linewidth {>{\centering\arraybackslash}p{8em}>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X}
+\begin{tabu} to \linewidth {>{\centering\arraybackslash}p{8em}>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X>{\centering}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X>{\centering}X>{\raggedleft}X>{\raggedleft}X}
 \toprule
-Sky conditions & Trend [\%/year] & Statistical signif. [\%] & Days with data\\
+slope & slope.p & Rsqrd & RsqrdAdj & cor.t & cor.esimate & cor.conf\_int\_1 & cor.conf\_int\_2 & DATA & N & slope.stat\_sig\\
 \midrule
-All skies & 0.376 & 100.00 & 10256\\
-Clear skies & 0.349 & 100.00 & 10256\\
-Cloudy skies & -0.276 & 99.97 & 5067\\
+0.376 & 0.00000 & 0.008308 & 0.008211 & 9.268 & 0.09115 & 0.07192 & 0.11031 & All skies & 10256 & 100.00\\
+0.349 & 0.00000 & 0.006183 & 0.006087 & 7.987 & 0.07864 & 0.05937 & 0.09784 & Clear skies & 10256 & 100.00\\
+-0.276 & 0.00026 & 0.002631 & 0.002434 & -3.655 & -0.05129 & -0.07872 & -0.02379 & Cloudy skies & 5067 & 99.97\\
 \bottomrule
@@
 \caption{\label{tab:trendseasontable}SDR linear trends of monthly anomalies for each season of the year.}
-\begin{tabu} to \linewidth {>{\centering\arraybackslash}p{8em}>{\centering}X>{\raggedleft}X>{\raggedleft}X}
+\begin{tabu} to \linewidth {>{\centering\arraybackslash}p{8em}>{\centering}X>{\raggedleft}X>{\raggedleft}X>{\centering}X>{\centering}X>{\raggedleft}X>{\raggedleft}X>{\centering}X}
 \toprule
-Sky condition & Season & Trend [\%/year] & Statistical signif. [\%]\\
+Trend [\%/year] & slope.p & cor.t & cor.esimate & cor.conf\_int\_1 & cor.conf\_int\_2 & Sky condition & Season & slope.stat\_sig\\
 \midrule
-\cellcolor{gray!6}{} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{0.6860} & \cellcolor{gray!6}{99.9}\\
-
- & Spring & 0.1450 & 81.8\\
-
-\cellcolor{gray!6}{} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{0.1200} & \cellcolor{gray!6}{89.4}\\
-
-\multirow{-4}{*}{\centering\arraybackslash All skies} & Autumn & 0.4310 & 99.3\\
-\cmidrule{1-4}
-\cellcolor{gray!6}{} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{0.8260} & \cellcolor{gray!6}{100.0}\\
-
- & Spring & 0.0613 & 38.6\\
-
-\cellcolor{gray!6}{} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{-0.0307} & \cellcolor{gray!6}{25.9}\\
-
-\multirow{-4}{*}{\centering\arraybackslash Clear skies} & Autumn & 0.3670 & 97.2\\
-\cmidrule{1-4}
-\cellcolor{gray!6}{} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{-0.8820} & \cellcolor{gray!6}{98.9}\\
-
- & Spring & -0.0991 & 37.2\\
-
-\cellcolor{gray!6}{} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{0.0444} & \cellcolor{gray!6}{21.5}\\
-
-\multirow{-4}{*}{\centering\arraybackslash Cloudy skies} & Autumn & -0.4000 & 89.5\\
+\cellcolor{gray!6}{0.6860} & \cellcolor{gray!6}{0.001300} & \cellcolor{gray!6}{3.5748458} & \cellcolor{gray!6}{0.5598043} & \cellcolor{gray!6}{0.2499438} & \cellcolor{gray!6}{0.7656559} & \cellcolor{gray!6}{All skies} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{99.9}\\
+\cmidrule{1-9}
+0.1450 & 0.182000 & 1.3668258 & 0.2460126 & -0.1186724 & 0.5522150 & All skies & Spring & 81.8\\
+\cmidrule{1-9}
+\cellcolor{gray!6}{0.1200} & \cellcolor{gray!6}{0.106000} & \cellcolor{gray!6}{1.6689762} & \cellcolor{gray!6}{0.3007995} & \cellcolor{gray!6}{-0.0666976} & \cellcolor{gray!6}{0.5964339} & \cellcolor{gray!6}{All skies} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{89.4}\\
+\cmidrule{1-9}
+0.4310 & 0.006650 & 2.9317564 & 0.4846362 & 0.1506747 & 0.7193136 & All skies & Autumn & 99.3\\
+\cmidrule{1-9}
+\cellcolor{gray!6}{0.8260} & \cellcolor{gray!6}{0.000259} & \cellcolor{gray!6}{4.1794536} & \cellcolor{gray!6}{0.6198226} & \cellcolor{gray!6}{0.3341758} & \cellcolor{gray!6}{0.8011848} & \cellcolor{gray!6}{Clear skies} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{100.0}\\
+\cmidrule{1-9}
+0.0613 & 0.614000 & 0.5095289 & 0.0941964 & -0.2691265 & 0.4340490 & Clear skies & Spring & 38.6\\
+\cmidrule{1-9}
+\cellcolor{gray!6}{-0.0307} & \cellcolor{gray!6}{0.741000} & \cellcolor{gray!6}{-0.3343454} & \cellcolor{gray!6}{-0.0630596} & \cellcolor{gray!6}{-0.4139251} & \cellcolor{gray!6}{0.3041187} & \cellcolor{gray!6}{Clear skies} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{25.9}\\
+\cmidrule{1-9}
+0.3670 & 0.028300 & 2.3126418 & 0.4004713 & 0.0469803 & 0.6648218 & Clear skies & Autumn & 97.2\\
+\cmidrule{1-9}
+\cellcolor{gray!6}{-0.8820} & \cellcolor{gray!6}{0.011100} & \cellcolor{gray!6}{-2.7197359} & \cellcolor{gray!6}{-0.4571342} & \cellcolor{gray!6}{-0.7018198} & \cellcolor{gray!6}{-0.1159631} & \cellcolor{gray!6}{Cloudy skies} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{98.9}\\
+\cmidrule{1-9}
+-0.0991 & 0.628000 & -0.4892643 & -0.0904814 & -0.4310031 & 0.2725985 & Cloudy skies & Spring & 37.2\\
+\cmidrule{1-9}
+\cellcolor{gray!6}{0.0444} & \cellcolor{gray!6}{0.785000} & \cellcolor{gray!6}{0.2750824} & \cellcolor{gray!6}{0.0519156} & \cellcolor{gray!6}{-0.3142309} & \cellcolor{gray!6}{0.4046170} & \cellcolor{gray!6}{Cloudy skies} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{21.5}\\
+\cmidrule{1-9}
+-0.4000 & 0.105000 & -1.6769099 & -0.3020992 & -0.5973542 & 0.0652743 & Cloudy skies & Autumn & 89.5\\
 \bottomrule

### 2023-11-29 DHI_GHI_1_longterm_trends.R

@@ -1313,8 +1313,8 @@ pprint           <- gather_seas[ , ..wecare]
 
 pprint[, slope.stat_sig     := 100 * (1 - slope.p)]
 pprint[, slope.t            := NULL               ]
-pprint[, Rsqrd              := NULL               ]
-pprint[, RsqrdAdj           := NULL               ]
+# pprint[, Rsqrd              := NULL               ]
+# pprint[, RsqrdAdj           := NULL               ]
 pprint[, N                  := NULL               ]
 pprint[, slope.ConfInt_0.99 := NULL               ]
 pprint[, slope.ConfInt_0.95 := NULL               ]

### 2023-11-29 DHI_GHI_1_longterm_trends.R
Commit 2023-11-29 23:10

@@ -536,6 +536,12 @@ pprint <- pprint[ , ..wecare]
 pprint[, DATA               := translate(DATA)                  ]
 pprint[, var                := translate(var)                   ]
 
+pprint$cor.data_name   <- NULL
+pprint$cor.null_value  <- NULL
+pprint$cor.method      <- NULL
+pprint$cor.alternative <- NULL
+pprint$cor.p           <- NULL
+
 
 
 # \scriptsize
@@ -1320,6 +1326,12 @@ pprint[, slope.ConfInt_0.99 := NULL               ]
 pprint[, slope.ConfInt_0.95 := NULL               ]
 pprint[, DATA               := translate(DATA)    ]
 pprint[, var                := translate(var)     ]
+
+pprint$cor.data_name   <- NULL
+pprint$cor.null_value  <- NULL
+pprint$cor.method      <- NULL
+pprint$cor.alternative <- NULL
+pprint$cor.p           <- NULL
 
 setorder(pprint, DATA, var)
 

### 2023-11-30 MDPI_submission.tex
Commit 2023-11-30 06:14

@@
 
+\begin{verbatim}
+Warning in set(x, j = name, value = value): Column 'cor.null_value' does not
+exist to remove
+\end{verbatim}
+
+\begin{verbatim}
+Warning in set(x, j = name, value = value): Column 'cor.alternative' does not
+exist to remove
+\end{verbatim}
+
+\begin{verbatim}
+Warning in set(x, j = name, value = value): Column 'cor.method' does not exist
+to remove
+\end{verbatim}
+
+\begin{verbatim}
+Warning in set(x, j = name, value = value): Column 'cor.data_name' does not
+exist to remove
+\end{verbatim}
+
+\begin{verbatim}
+Warning in set(x, j = name, value = value): Column 'cor.p' does not exist to
+remove
+\end{verbatim}
+
 For completeness, we will provide below a brief overview of the clear
@@
 
+\begin{verbatim}
+Warning in set(x, j = name, value = value): Column 'cor.null_value' does not
+exist to remove
+\end{verbatim}
+
+\begin{verbatim}
+Warning in set(x, j = name, value = value): Column 'cor.alternative' does not
+exist to remove
+\end{verbatim}
+
+\begin{verbatim}
+Warning in set(x, j = name, value = value): Column 'cor.method' does not exist
+to remove
+\end{verbatim}
+
+\begin{verbatim}
+Warning in set(x, j = name, value = value): Column 'cor.data_name' does not
+exist to remove
+\end{verbatim}
+
+\begin{verbatim}
+Warning in set(x, j = name, value = value): Column 'cor.p' does not exist to
+remove
+\end{verbatim}
+
 We calculated the linear SDR trends, from the departures of the mean
@@
 \caption{\label{tab:trendtable}Trends in SDR daily means for different sky conditions for the period 1993 - 2023.}
-\begin{tabu} to \linewidth {>{\centering\arraybackslash}p{8em}>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X>{\centering}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X>{\centering}X>{\raggedleft}X>{\raggedleft}X}
+\begin{tabu} to \linewidth {>{\centering\arraybackslash}p{8em}>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X>{\centering}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X>{\centering}X}
 \toprule
-slope & slope.p & Rsqrd & RsqrdAdj & cor.t & cor.esimate & cor.conf\_int\_1 & cor.conf\_int\_2 & DATA & N & slope.stat\_sig\\
+slope & slope.p & Rsqrd & RsqrdAdj & cor.t & cor.esimate & DATA & N & slope.stat\_sig\\
 \midrule
-0.376 & 0.00000 & 0.008308 & 0.008211 & 9.268 & 0.09115 & 0.07192 & 0.11031 & All skies & 10256 & 100.00\\
-0.349 & 0.00000 & 0.006183 & 0.006087 & 7.987 & 0.07864 & 0.05937 & 0.09784 & Clear skies & 10256 & 100.00\\
--0.276 & 0.00026 & 0.002631 & 0.002434 & -3.655 & -0.05129 & -0.07872 & -0.02379 & Cloudy skies & 5067 & 99.97\\
+0.376 & 0.00000 & 0.008308 & 0.008211 & 9.268 & 0.09115 & All skies & 10256 & 100.00\\
+0.349 & 0.00000 & 0.006183 & 0.006087 & 7.987 & 0.07864 & Clear skies & 10256 & 100.00\\
+-0.276 & 0.00026 & 0.002631 & 0.002434 & -3.655 & -0.05129 & Cloudy skies & 5067 & 99.97\\
 \bottomrule
@@
 \toprule
-Trend [\%/year] & slope.p & cor.t & cor.esimate & cor.conf\_int\_1 & cor.conf\_int\_2 & Sky condition & Season & slope.stat\_sig\\
+Trend [\%/year] & slope.p & Rsqrd & RsqrdAdj & cor.t & cor.esimate & Sky condition & Season & slope.stat\_sig\\
 \midrule
-\cellcolor{gray!6}{0.6860} & \cellcolor{gray!6}{0.001300} & \cellcolor{gray!6}{3.5748458} & \cellcolor{gray!6}{0.5598043} & \cellcolor{gray!6}{0.2499438} & \cellcolor{gray!6}{0.7656559} & \cellcolor{gray!6}{All skies} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{99.9}\\
+\cellcolor{gray!6}{0.6860} & \cellcolor{gray!6}{0.001300} & \cellcolor{gray!6}{0.3133809} & \cellcolor{gray!6}{0.2888588} & \cellcolor{gray!6}{3.5748458} & \cellcolor{gray!6}{0.5598043} & \cellcolor{gray!6}{All skies} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{99.9}\\
 \cmidrule{1-9}
-0.1450 & 0.182000 & 1.3668258 & 0.2460126 & -0.1186724 & 0.5522150 & All skies & Spring & 81.8\\
+0.1450 & 0.182000 & 0.0605222 & 0.0281264 & 1.3668258 & 0.2460126 & All skies & Spring & 81.8\\
 \cmidrule{1-9}
-\cellcolor{gray!6}{0.1200} & \cellcolor{gray!6}{0.106000} & \cellcolor{gray!6}{1.6689762} & \cellcolor{gray!6}{0.3007995} & \cellcolor{gray!6}{-0.0666976} & \cellcolor{gray!6}{0.5964339} & \cellcolor{gray!6}{All skies} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{89.4}\\
+\cellcolor{gray!6}{0.1200} & \cellcolor{gray!6}{0.106000} & \cellcolor{gray!6}{0.0904804} & \cellcolor{gray!6}{0.0579975} & \cellcolor{gray!6}{1.6689762} & \cellcolor{gray!6}{0.3007995} & \cellcolor{gray!6}{All skies} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{89.4}\\
 \cmidrule{1-9}
-0.4310 & 0.006650 & 2.9317564 & 0.4846362 & 0.1506747 & 0.7193136 & All skies & Autumn & 99.3\\
+0.4310 & 0.006650 & 0.2348722 & 0.2075462 & 2.9317564 & 0.4846362 & All skies & Autumn & 99.3\\
 \cmidrule{1-9}
-\cellcolor{gray!6}{0.8260} & \cellcolor{gray!6}{0.000259} & \cellcolor{gray!6}{4.1794536} & \cellcolor{gray!6}{0.6198226} & \cellcolor{gray!6}{0.3341758} & \cellcolor{gray!6}{0.8011848} & \cellcolor{gray!6}{Clear skies} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{100.0}\\
+\cellcolor{gray!6}{0.8260} & \cellcolor{gray!6}{0.000259} & \cellcolor{gray!6}{0.3841800} & \cellcolor{gray!6}{0.3621864} & \cellcolor{gray!6}{4.1794536} & \cellcolor{gray!6}{0.6198226} & \cellcolor{gray!6}{Clear skies} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{100.0}\\
 \cmidrule{1-9}
-0.0613 & 0.614000 & 0.5095289 & 0.0941964 & -0.2691265 & 0.4340490 & Clear skies & Spring & 38.6\\
+0.0613 & 0.614000 & 0.0088730 & -0.0253038 & 0.5095289 & 0.0941964 & Clear skies & Spring & 38.6\\
 \cmidrule{1-9}
-\cellcolor{gray!6}{-0.0307} & \cellcolor{gray!6}{0.741000} & \cellcolor{gray!6}{-0.3343454} & \cellcolor{gray!6}{-0.0630596} & \cellcolor{gray!6}{-0.4139251} & \cellcolor{gray!6}{0.3041187} & \cellcolor{gray!6}{Clear skies} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{25.9}\\
+\cellcolor{gray!6}{-0.0307} & \cellcolor{gray!6}{0.741000} & \cellcolor{gray!6}{0.0039765} & \cellcolor{gray!6}{-0.0315958} & \cellcolor{gray!6}{-0.3343454} & \cellcolor{gray!6}{-0.0630596} & \cellcolor{gray!6}{Clear skies} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{25.9}\\
 \cmidrule{1-9}
-0.3670 & 0.028300 & 2.3126418 & 0.4004713 & 0.0469803 & 0.6648218 & Clear skies & Autumn & 97.2\\
+0.3670 & 0.028300 & 0.1603773 & 0.1303908 & 2.3126418 & 0.4004713 & Clear skies & Autumn & 97.2\\
 \cmidrule{1-9}
-\cellcolor{gray!6}{-0.8820} & \cellcolor{gray!6}{0.011100} & \cellcolor{gray!6}{-2.7197359} & \cellcolor{gray!6}{-0.4571342} & \cellcolor{gray!6}{-0.7018198} & \cellcolor{gray!6}{-0.1159631} & \cellcolor{gray!6}{Cloudy skies} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{98.9}\\
+\cellcolor{gray!6}{-0.8820} & \cellcolor{gray!6}{0.011100} & \cellcolor{gray!6}{0.2089717} & \cellcolor{gray!6}{0.1807207} & \cellcolor{gray!6}{-2.7197359} & \cellcolor{gray!6}{-0.4571342} & \cellcolor{gray!6}{Cloudy skies} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{98.9}\\
 \cmidrule{1-9}
--0.0991 & 0.628000 & -0.4892643 & -0.0904814 & -0.4310031 & 0.2725985 & Cloudy skies & Spring & 37.2\\
+-0.0991 & 0.628000 & 0.0081869 & -0.0260136 & -0.4892643 & -0.0904814 & Cloudy skies & Spring & 37.2\\
 \cmidrule{1-9}
-\cellcolor{gray!6}{0.0444} & \cellcolor{gray!6}{0.785000} & \cellcolor{gray!6}{0.2750824} & \cellcolor{gray!6}{0.0519156} & \cellcolor{gray!6}{-0.3142309} & \cellcolor{gray!6}{0.4046170} & \cellcolor{gray!6}{Cloudy skies} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{21.5}\\
+\cellcolor{gray!6}{0.0444} & \cellcolor{gray!6}{0.785000} & \cellcolor{gray!6}{0.0026952} & \cellcolor{gray!6}{-0.0329228} & \cellcolor{gray!6}{0.2750824} & \cellcolor{gray!6}{0.0519156} & \cellcolor{gray!6}{Cloudy skies} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{21.5}\\
 \cmidrule{1-9}
--0.4000 & 0.105000 & -1.6769099 & -0.3020992 & -0.5973542 & 0.0652743 & Cloudy skies & Autumn & 89.5\\
+-0.4000 & 0.105000 & 0.0912639 & 0.0588091 & -1.6769099 & -0.3020992 & Cloudy skies & Autumn & 89.5\\
 \bottomrule

### 2023-11-30 MDPI_submission.Rmd
Commit 2023-11-30 07:10

@@
 )
+
+knit_hooks$set(extrawide = function(before, options, envir) {
+    if (before) {
+        return("\\begin{adjustwidth}{-\extralength}{0cm}")
+    }
+
+    if (!before) {
+        return("\\end{adjustwidth}")
+    }
+
+    # else {
+    #     
+    #     output <- vector(mode = "character", length = options$fig.num + 1)
+    #     
+    #     for (i in 1:options$fig.num) {
+    #         output[i] <- sprintf("\\includegraphics{%s}", fig_path(number = i))
+    #     }
+    #     
+    #     output[i+1] <- "\\end{adjustwidth}"
+    #     return(paste(output, collapse = ""))
+    # }
+})
+
 ```
@@
 
-TSI$cor.null_value     <- NULL
-TSI$cor.alternative    <- NULL
-TSI$cor.method         <- NULL
-TSI$cor.data_name      <- NULL
-TSI$cor.p              <- NULL
+# TSI$cor.null_value     <- NULL
+# TSI$cor.alternative    <- NULL
+# TSI$cor.method         <- NULL
+# TSI$cor.data_name      <- NULL
+# TSI$cor.p              <- NULL
 TSI$conf.level         <- NULL
@@
 LT$cor.df           <- NULL
-LT$cor.null_value   <- NULL
-LT$cor.alternative  <- NULL
-LT$cor.method       <- NULL
-LT$cor.data_name    <- NULL
-LT$cor.p            <- NULL
+# LT$cor.null_value   <- NULL
+# LT$cor.alternative  <- NULL
+# LT$cor.method       <- NULL
+# LT$cor.data_name    <- NULL
+# LT$cor.p            <- NULL
 LT$conf.level       <- NULL

### 2023-11-30 MDPI_submission.tex
Commit 2023-11-30 07:10

@@
 
-\begin{verbatim}
-Warning in set(x, j = name, value = value): Column 'cor.null_value' does not
-exist to remove
-\end{verbatim}
-
-\begin{verbatim}
-Warning in set(x, j = name, value = value): Column 'cor.alternative' does not
-exist to remove
-\end{verbatim}
-
-\begin{verbatim}
-Warning in set(x, j = name, value = value): Column 'cor.method' does not exist
-to remove
-\end{verbatim}
-
-\begin{verbatim}
-Warning in set(x, j = name, value = value): Column 'cor.data_name' does not
-exist to remove
-\end{verbatim}
-
-\begin{verbatim}
-Warning in set(x, j = name, value = value): Column 'cor.p' does not exist to
-remove
-\end{verbatim}
-
 For completeness, we will provide below a brief overview of the clear
@@
 \subsection{Long-term SDR trends}\label{long-term-sdr-trends}}
-
-\begin{verbatim}
-Warning in set(x, j = name, value = value): Column 'cor.null_value' does not
-exist to remove
-\end{verbatim}
-
-\begin{verbatim}
-Warning in set(x, j = name, value = value): Column 'cor.alternative' does not
-exist to remove
-\end{verbatim}
-
-\begin{verbatim}
-Warning in set(x, j = name, value = value): Column 'cor.method' does not exist
-to remove
-\end{verbatim}
-
-\begin{verbatim}
-Warning in set(x, j = name, value = value): Column 'cor.data_name' does not
-exist to remove
-\end{verbatim}
-
-\begin{verbatim}
-Warning in set(x, j = name, value = value): Column 'cor.p' does not exist to
-remove
-\end{verbatim}
 

### 2023-12-01 DHI_GHI_3_trends_consistency.R
Commit 2023-12-01 12:14

@@ -52,8 +52,6 @@ knitr::opts_chunk$set(dev        = c("pdf", "png"))
 knitr::opts_chunk$set(out.width  = "100%"   )
 knitr::opts_chunk$set(fig.align  = "center" )
 knitr::opts_chunk$set(cache      =  FALSE   )  ## !! breaks calculations
-# knitr::opts_chunk$set(fig.pos    = '!h'    )
-warning("Don't use cache it breaks computations")
 
 #+ include=F, echo=F
 ## __ Set environment ----------------------------------------------------------
@@ -66,14 +64,12 @@ if (!interactive()) {
     filelock::lock(paste0("./runtime/", basename(sub("\\.R$",".lock", Script.Name))), timeout = 0)
 }
 
-
 #+ echo=F, include=T
 library(data.table, quietly = TRUE, warn.conflicts = FALSE)
 library(pander,     quietly = TRUE, warn.conflicts = FALSE)
 library(lubridate,  quietly = TRUE, warn.conflicts = FALSE)
 library(ggplot2,    quietly = TRUE, warn.conflicts = FALSE)
 library(fANCOVA,    quietly = TRUE, warn.conflicts = FALSE)
-
 
 panderOptions("table.alignment.default", "right")
 panderOptions("table.split.table",        120   )
@@ -85,20 +81,31 @@ source("~/CODE/FUNCTIONS/R/linear_fit_stats.R")
 source("~/CODE/FUNCTIONS/R/trig_deg.R")
 source("~/CODE/FUNCTIONS/R/data.R")
 
-
 ## __ Source initial scripts ---------------------------------------------------
 # source("./DHI_GHI_0_data_input.R")
 source("./DHI_GHI_0_variables.R")
 
 ## This use data from 1 !!!
 ## check previous steps
-if (! file.exists(I1_longterm) |
+if (!file.exists(I1_longterm) |
     file.mtime(I1_longterm) < file.mtime("./DHI_GHI_0_variables.R") |
     file.mtime(I1_longterm) < file.mtime("./DHI_GHI_00_raw_data.R") |
-    file.mtime(I1_longterm) < file.mtime("./DHI_GHI_01_Input_longterm.R") )
+    file.mtime(I1_longterm) < file.mtime("./DHI_GHI_01_Input_longterm.R"))
 {
     # source("./DHI_GHI_01_Input_longterm.R")
-    tryCatch(source("./DHI_GHI_01_Input_longterm.R"), exit=function(cond) {
[…]

### 2023-12-01 MDPI_submission.Rmd
Commit 2023-11-30 13:08
Commit 2023-12-01 13:10

@@
 source("./DHI_GHI_00_raw_data.R")
-load(common_data)
+# load(common_data)
 load("./data/Input_1_longterm_trends.Rda")
@@
 
-knit_hooks$set(extrawide = function(before, options, envir) {
+knitr::knit_hooks$set(extrawide = function(before, options, envir) {
     if (before) {
-        return("\\begin{adjustwidth}{-\extralength}{0cm}")
+        return("\\begin{adjustwidth}{-\\extralength}{0cm}")
     }
@@
 The measurements are corrected for the zero offset ("dark signal" in volts).
-The "dark signal" is calculated by averaging all measurements recorded for a period of
-$3\,\text{h}$,
-before (morning) or after (evening) the Sun reaches an elevation angle of
-$-10^\circ$.
+The "dark signal" is calculated by averaging all measurements recorded for a period of $3\,\text{h}$, before (morning) or after (evening) the Sun reaches an elevation angle of $-10^\circ$.
 The signal is converted to irradiance using a ramped value of the instrument’s sensitivity between calibrations.
@@
 Moreover, due to some obstructions around the site (hills and buildings), we excluded data with azimuth angle between
-$35^\circ$ and $120^\circ$ with SZA greater than $80^\circ$.
+$`r FIBais_Az_1`^\circ$ and $`r FIBais_Az_2`^\circ$ with SZA greater than $`r 90 -  FIBais_Elev`^\circ$.
 On the latter instances, Sun is systematically not visible from the instrument's location.
@@
 $`r signif(100 * SDR_N_Clear / SDR_N_All, 3)`\%$
-of the data were identified as under clear-sky conditions and
+of the 1-min data were identified as under clear-sky conditions and
 $`r signif(100 * SDR_N_Cloud / SDR_N_All, 3)`\%$

### 2023-12-01 MDPI_submission.tex
Commit 2023-12-01 13:10

@@
 \(85^\circ\). Moreover, due to some obstructions around the site (hills
-and buildings), we excluded data with azimuth angle between \(35^\circ\)
-and \(120^\circ\) with SZA greater than \(80^\circ\). On the latter
+and buildings), we excluded data with azimuth angle between \(58^\circ\)
+and \(120^\circ\) with SZA greater than \(78^\circ\). On the latter
 instances, Sun is systematically not visible from the instrument's
@@
 initial daily values of this dataset were interpolated to match the time
-step of our measurements. In the final dataset \(84.2\%\) of the data
-were identified as under clear-sky conditions and \(15.8\%\) as under
+step of our measurements. In the final dataset \(40.1\%\) of the data
+were identified as under clear-sky conditions and \(59.9\%\) as under
 cloud-sky conditions.
@@
 positive trend for all-sky conditions (\(0.38\,\%/y\)), a very close but
-smaller trend for clear-skies (\(0.35\,\%/y\)) and a negative weaker
-trend for cloudy-skies (\(-0.28\,\%/y\)). In the studied period, there
+smaller trend for clear-skies (\(0.16\,\%/y\)) and a negative weaker
+trend for cloudy-skies (\(0.019\,\%/y\)). In the studied period, there
 is no significant break or change in the variability of the time series.
@@
 before the start of our records. It is interesting to note, that for the
-observations period, the trend of the TSI is \(-0.0002\,\%/y\), and thus
-we can not attribute any major effect on SDR trend to Solar variability.
+observations period, the trend of the TSI is \(-0.00022\,\%/y\), and
+thus we can not attribute any major effect on SDR trend to Solar
+variability.
 
@@
 \midrule
-0.376 & 0.00000 & 0.008308 & 0.008211 & 9.268 & 0.09115 & All skies & 10256 & 100.00\\
-0.349 & 0.00000 & 0.006183 & 0.006087 & 7.987 & 0.07864 & Clear skies & 10256 & 100.00\\
--0.276 & 0.00026 & 0.002631 & 0.002434 & -3.655 & -0.05129 & Cloudy skies & 5067 & 99.97\\
+0.376 & 0.00e+00 & 0.0083302 & 0.0082334 & 9.279 & 0.091270 & All skies & 10252 & 100.00\\
+0.156 & 7.00e-07 & 0.0030928 & 0.0029676 & 4.971 & 0.055613 & Clear skies & 7967 & 100.00\\
+0.019 & 7.23e-01 & 0.0000123 & -0.0000858 & 0.354 & 0.003506 & Cloudy skies & 10194 & 27.66\\
 \bottomrule
@@
 \midrule
-\cellcolor{gray!6}{0.6860} & \cellcolor{gray!6}{0.001300} & \cellcolor{gray!6}{0.3133809} & \cellcolor{gray!6}{0.2888588} & \cellcolor{gray!6}{3.5748458} & \cellcolor{gray!6}{0.5598043} & \cellcolor{gray!6}{All skies} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{99.9}\\
+\cellcolor{gray!6}{0.6870} & \cellcolor{gray!6}{1.33e-03} & \cellcolor{gray!6}{0.3121839} & \cellcolor{gray!6}{0.2876190} & \cellcolor{gray!6}{3.5649059} & \cellcolor{gray!6}{0.5587342} & \cellcolor{gray!6}{All skies} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{99.90}\\
 \cmidrule{1-9}
-0.1450 & 0.182000 & 0.0605222 & 0.0281264 & 1.3668258 & 0.2460126 & All skies & Spring & 81.8\\
+0.1460 & 1.78e-01 & 0.0616942 & 0.0293388 & 1.3808581 & 0.2483832 & All skies & Spring & 82.20\\
 \cmidrule{1-9}
-\cellcolor{gray!6}{0.1200} & \cellcolor{gray!6}{0.106000} & \cellcolor{gray!6}{0.0904804} & \cellcolor{gray!6}{0.0579975} & \cellcolor{gray!6}{1.6689762} & \cellcolor{gray!6}{0.3007995} & \cellcolor{gray!6}{All skies} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{89.4}\\
+\cellcolor{gray!6}{0.1230} & \cellcolor{gray!6}{9.61e-02} & \cellcolor{gray!6}{0.0957525} & \cellcolor{gray!6}{0.0634580} & \cellcolor{gray!6}{1.7219100} & \cellcolor{gray!6}{0.3094390} & \cellcolor{gray!6}{All skies} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{90.40}\\
 \cmidrule{1-9}
-0.4310 & 0.006650 & 0.2348722 & 0.2075462 & 2.9317564 & 0.4846362 & All skies & Autumn & 99.3\\
+0.4300 & 6.67e-03 & 0.2347186 & 0.2073871 & 2.9305032 & 0.4844777 & All skies & Autumn & 99.30\\
 \cmidrule{1-9}
-\cellcolor{gray!6}{0.8260} & \cellcolor{gray!6}{0.000259} & \cellcolor{gray!6}{0.3841800} & \cellcolor{gray!6}{0.3621864} & \cellcolor{gray!6}{4.1794536} & \cellcolor{gray!6}{0.6198226} & \cellcolor{gray!6}{Clear skies} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{100.0}\\
+\cellcolor{gray!6}{0.6990} & \cellcolor{gray!6}{8.51e-05} & \cellcolor{gray!6}{0.4293113} & \cellcolor{gray!6}{0.4089296} & \cellcolor{gray!6}{4.5895019} & \cellcolor{gray!6}{0.6552185} & \cellcolor{gray!6}{Clear skies} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{100.00}\\
 \cmidrule{1-9}
-0.0613 & 0.614000 & 0.0088730 & -0.0253038 & 0.5095289 & 0.0941964 & Clear skies & Spring & 38.6\\
+-0.0265 & 8.46e-01 & 0.0013302 & -0.0331066 & -0.1965409 & -0.0364724 & Clear skies & Spring & 15.40\\
 \cmidrule{1-9}
-\cellcolor{gray!6}{-0.0307} & \cellcolor{gray!6}{0.741000} & \cellcolor{gray!6}{0.0039765} & \cellcolor{gray!6}{-0.0315958} & \cellcolor{gray!6}{-0.3343454} & \cellcolor{gray!6}{-0.0630596} & \cellcolor{gray!6}{Clear skies} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{25.9}\\
+\cellcolor{gray!6}{-0.1810} & \cellcolor{gray!6}{8.08e-03} & \cellcolor{gray!6}{0.2250803} & \cellcolor{gray!6}{0.1974046} & \cellcolor{gray!6}{-2.8518021} & \cellcolor{gray!6}{-0.4744263} & \cellcolor{gray!6}{Clear skies} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{99.20}\\
 \cmidrule{1-9}
-0.3670 & 0.028300 & 0.1603773 & 0.1303908 & 2.3126418 & 0.4004713 & Clear skies & Autumn & 97.2\\
+0.1260 & 2.57e-01 & 0.0455591 & 0.0114719 & 1.1560907 & 0.2134457 & Clear skies & Autumn & 74.30\\
 \cmidrule{1-9}
-\cellcolor{gray!6}{-0.8820} & \cellcolor{gray!6}{0.011100} & \cellcolor{gray!6}{0.2089717} & \cellcolor{gray!6}{0.1807207} & \cellcolor{gray!6}{-2.7197359} & \cellcolor{gray!6}{-0.4571342} & \cellcolor{gray!6}{Cloudy skies} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{98.9}\\
+\cellcolor{gray!6}{0.0944} & \cellcolor{gray!6}{6.13e-01} & \cellcolor{gray!6}{0.0092345} & \cellcolor{gray!6}{-0.0261500} & \cellcolor{gray!6}{0.5108573} & \cellcolor{gray!6}{0.0960962} & \cellcolor{gray!6}{Cloudy skies} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{38.70}\\
 \cmidrule{1-9}
--0.0991 & 0.628000 & 0.0081869 & -0.0260136 & -0.4892643 & -0.0904814 & Cloudy skies & Spring & 37.2\\
+0.0161 & 9.09e-01 & 0.0004533 & -0.0340138 & 0.1146809 & 0.0212909 & Cloudy skies & Spring & 9.05\\
 \cmidrule{1-9}
-\cellcolor{gray!6}{0.0444} & \cellcolor{gray!6}{0.785000} & \cellcolor{gray!6}{0.0026952} & \cellcolor{gray!6}{-0.0329228} & \cellcolor{gray!6}{0.2750824} & \cellcolor{gray!6}{0.0519156} & \cellcolor{gray!6}{Cloudy skies} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{21.5}\\
+\cellcolor{gray!6}{-0.0283} & \cellcolor{gray!6}{8.07e-01} & \cellcolor{gray!6}{0.0021748} & \cellcolor{gray!6}{-0.0334619} & \cellcolor{gray!6}{-0.2470344} & \cellcolor{gray!6}{-0.0466343} & \cellcolor{gray!6}{Cloudy skies} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{19.30}\\
 \cmidrule{1-9}
--0.4000 & 0.105000 & 0.0912639 & 0.0588091 & -1.6769099 & -0.3020992 & Cloudy skies & Autumn & 89.5\\
+-0.0427 & 7.98e-01 & 0.0023737 & -0.0332559 & -0.2581092 & -0.0487201 & Cloudy skies & Autumn & 20.20\\
 \bottomrule
@@
 Europe have been proposed as major factor by \citet{Wild2021}. The
-dimming effect on SDR under cloudy-sky conditions (\(-0.28\,\%/y\)),
+dimming effect on SDR under cloudy-sky conditions (\(0.019\,\%/y\)),
 suggests that cloud optical thickness is decreasing during this period.

### 2023-12-01 MDPI_submission.Rmd
Commit 2023-12-01 14:14

@@
 $`r signif(100 * SDR_N_Clear / SDR_N_All, 3)`\%$
-of the 1-min data were identified as under clear-sky conditions and
+of the 1-minute data were identified as under clear-sky conditions and
 $`r signif(100 * SDR_N_Cloud / SDR_N_All, 3)`\%$

### 2023-12-01 DHI_GHI_01_Input_longterm.R

@@ -73,6 +73,7 @@ ALL_1_daily_mean <-
                 HOR_att_sd    = sd(  HOR_att,      na.rm = T),
                 DIR_transp_sd = sd(  DIR_transp,   na.rm = T),
                 tsi1au_att_sd = sd(  tsi_1au_comb, na.rm = T),
+                Daylenth      = dlllllllllll
                 doy           = yday(Date),
                 GLB_att_N     = sum(!is.na(GLB_att)),
                 HOR_att_N     = sum(!is.na(HOR_att)),
@@ -113,6 +114,9 @@ CLOUD_1_daily_mean <-
                   DIR_att_N     = sum(!is.na(DIR_att))),
                by = .( Date = Day ) ]
 
+
+
+stop()
 
 
 ## _ Margin of error for confidence interval -----------------------------------

### 2023-12-01 DHI_GHI_01_Input_longterm.R
Commit 2023-12-01 18:14

@@ -120,6 +120,9 @@ CLOUD_1_daily_mean <-
 
 range(CLOUD_1_daily_mean[, GLB_att_N / DayLength], breaks = 100)
 
+range(CLEAR_1_daily_mean[, GLB_att_N / DayLength], breaks = 100)
+
+
 test <- DATA_Clear[, sum(!is.na(GLB_att)) / max(DayLength) ,
                   by = Day]
 

### 2023-12-02 MDPI_submission.Rmd
Commit 2023-12-02 22:14

@@
 ## load data from trends analysis
-source("./DHI_GHI_00_raw_data.R")
+# source("./DHI_GHI_00_raw_data.R")
 # load(common_data)

### 2023-12-02 MDPI_submission.tex
Commit 2023-12-01 18:14
Commit 2023-12-02 22:14

@@
 initial daily values of this dataset were interpolated to match the time
-step of our measurements. In the final dataset \(40.1\%\) of the data
-were identified as under clear-sky conditions and \(59.9\%\) as under
+step of our measurements. In the final dataset \(NA\%\) of the 1-minute
+data were identified as under clear-sky conditions and \(NA\%\) as under
 cloud-sky conditions.
@@
 positive trend for all-sky conditions (\(0.38\,\%/y\)), a very close but
-smaller trend for clear-skies (\(0.16\,\%/y\)) and a negative weaker
-trend for cloudy-skies (\(0.019\,\%/y\)). In the studied period, there
-is no significant break or change in the variability of the time series.
+smaller trend for clear-skies (\(0.041\,\%/y\)) and a negative weaker
+trend for cloudy-skies (\(0.34\,\%/y\)). In the studied period, there is
+no significant break or change in the variability of the time series.
 Other studies for the European region reported a change of the SDR
@@
 \toprule
-slope & slope.p & Rsqrd & RsqrdAdj & cor.t & cor.esimate & DATA & N & slope.stat\_sig\\
+slope & slope.p & Rsqrd & RsqrdAdj & cor.t & cor.estimate & DATA & N & slope.stat\_sig\\
 \midrule
-0.376 & 0.00e+00 & 0.0083302 & 0.0082334 & 9.279 & 0.091270 & All skies & 10252 & 100.00\\
-0.156 & 7.00e-07 & 0.0030928 & 0.0029676 & 4.971 & 0.055613 & Clear skies & 7967 & 100.00\\
-0.019 & 7.23e-01 & 0.0000123 & -0.0000858 & 0.354 & 0.003506 & Cloudy skies & 10194 & 27.66\\
+0.3760 & 0.0000 & 0.008339 & 0.0082421 & 9.284 & 0.09132 & All skies & 10251 & 100.00\\
+0.3430 & 0.0000 & 0.005312 & 0.0051756 & 6.234 & 0.07289 & Cloudy skies & 7279 & 100.00\\
+0.0406 & 0.0141 & 0.001195 & 0.0009971 & 2.455 & 0.03458 & Clear skies & 5036 & 98.59\\
 \bottomrule
@@
 of clouds during winter and autumn. For clear-skies, the trend in winter
-is \(0.7\,\%/y\), larger than for all-skies (\(0.69\,\%/y\)), which is
+is \(0.5\,\%/y\), larger than for all-skies (\(0.69\,\%/y\)), which is
 another indication of a decreasing trend in cloud optical thickness.
@@
 \toprule
-Trend [\%/year] & slope.p & Rsqrd & RsqrdAdj & cor.t & cor.esimate & Sky condition & Season & slope.stat\_sig\\
+Trend [\%/year] & slope.p & Rsqrd & RsqrdAdj & cor.t & cor.estimate & Sky condition & Season & slope.stat\_sig\\
 \midrule
-\cellcolor{gray!6}{0.6870} & \cellcolor{gray!6}{1.33e-03} & \cellcolor{gray!6}{0.3121839} & \cellcolor{gray!6}{0.2876190} & \cellcolor{gray!6}{3.5649059} & \cellcolor{gray!6}{0.5587342} & \cellcolor{gray!6}{All skies} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{99.90}\\
+\cellcolor{gray!6}{0.6870} & \cellcolor{gray!6}{0.001330} & \cellcolor{gray!6}{0.3121839} & \cellcolor{gray!6}{0.2876190} & \cellcolor{gray!6}{3.564906} & \cellcolor{gray!6}{0.5587342} & \cellcolor{gray!6}{All skies} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{99.9}\\
 \cmidrule{1-9}
-0.1460 & 1.78e-01 & 0.0616942 & 0.0293388 & 1.3808581 & 0.2483832 & All skies & Spring & 82.20\\
+0.1460 & 0.178000 & 0.0616942 & 0.0293388 & 1.380858 & 0.2483832 & All skies & Spring & 82.2\\
 \cmidrule{1-9}
-\cellcolor{gray!6}{0.1230} & \cellcolor{gray!6}{9.61e-02} & \cellcolor{gray!6}{0.0957525} & \cellcolor{gray!6}{0.0634580} & \cellcolor{gray!6}{1.7219100} & \cellcolor{gray!6}{0.3094390} & \cellcolor{gray!6}{All skies} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{90.40}\\
+\cellcolor{gray!6}{0.1230} & \cellcolor{gray!6}{0.096100} & \cellcolor{gray!6}{0.0957525} & \cellcolor{gray!6}{0.0634580} & \cellcolor{gray!6}{1.721910} & \cellcolor{gray!6}{0.3094390} & \cellcolor{gray!6}{All skies} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{90.4}\\
 \cmidrule{1-9}
-0.4300 & 6.67e-03 & 0.2347186 & 0.2073871 & 2.9305032 & 0.4844777 & All skies & Autumn & 99.30\\
+0.4300 & 0.006670 & 0.2347063 & 0.2073744 & 2.930403 & 0.4844650 & All skies & Autumn & 99.3\\
 \cmidrule{1-9}
-\cellcolor{gray!6}{0.6990} & \cellcolor{gray!6}{8.51e-05} & \cellcolor{gray!6}{0.4293113} & \cellcolor{gray!6}{0.4089296} & \cellcolor{gray!6}{4.5895019} & \cellcolor{gray!6}{0.6552185} & \cellcolor{gray!6}{Clear skies} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{100.00}\\
+\cellcolor{gray!6}{0.4990} & \cellcolor{gray!6}{0.000166} & \cellcolor{gray!6}{0.4025975} & \cellcolor{gray!6}{0.3812617} & \cellcolor{gray!6}{4.343912} & \cellcolor{gray!6}{0.6345057} & \cellcolor{gray!6}{Clear skies} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{100.0}\\
 \cmidrule{1-9}
--0.0265 & 8.46e-01 & 0.0013302 & -0.0331066 & -0.1965409 & -0.0364724 & Clear skies & Spring & 15.40\\
+-0.0865 & 0.315000 & 0.0348099 & 0.0015275 & -1.022690 & -0.1865741 & Clear skies & Spring & 68.5\\
 \cmidrule{1-9}
-\cellcolor{gray!6}{-0.1810} & \cellcolor{gray!6}{8.08e-03} & \cellcolor{gray!6}{0.2250803} & \cellcolor{gray!6}{0.1974046} & \cellcolor{gray!6}{-2.8518021} & \cellcolor{gray!6}{-0.4744263} & \cellcolor{gray!6}{Clear skies} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{99.20}\\
+\cellcolor{gray!6}{-0.2180} & \cellcolor{gray!6}{0.000398} & \cellcolor{gray!6}{0.3659586} & \cellcolor{gray!6}{0.3433142} & \cellcolor{gray!6}{-4.020094} & \cellcolor{gray!6}{-0.6049451} & \cellcolor{gray!6}{Clear skies} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{100.0}\\
 \cmidrule{1-9}
-0.1260 & 2.57e-01 & 0.0455591 & 0.0114719 & 1.1560907 & 0.2134457 & Clear skies & Autumn & 74.30\\
+0.1090 & 0.221000 & 0.0529350 & 0.0191112 & 1.251009 & 0.2300761 & Clear skies & Autumn & 77.9\\
 \cmidrule{1-9}
-\cellcolor{gray!6}{0.0944} & \cellcolor{gray!6}{6.13e-01} & \cellcolor{gray!6}{0.0092345} & \cellcolor{gray!6}{-0.0261500} & \cellcolor{gray!6}{0.5108573} & \cellcolor{gray!6}{0.0960962} & \cellcolor{gray!6}{Cloudy skies} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{38.70}\\
+\cellcolor{gray!6}{0.5200} & \cellcolor{gray!6}{0.009580} & \cellcolor{gray!6}{0.2164332} & \cellcolor{gray!6}{0.1884487} & \cellcolor{gray!6}{2.781012} & \cellcolor{gray!6}{0.4652238} & \cellcolor{gray!6}{Cloudy skies} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{99.0}\\
 \cmidrule{1-9}
-0.0161 & 9.09e-01 & 0.0004533 & -0.0340138 & 0.1146809 & 0.0212909 & Cloudy skies & Spring & 9.05\\
+0.1780 & 0.122000 & 0.0806332 & 0.0489309 & 1.594819 & 0.2839598 & Cloudy skies & Spring & 87.8\\
 \cmidrule{1-9}
-\cellcolor{gray!6}{-0.0283} & \cellcolor{gray!6}{8.07e-01} & \cellcolor{gray!6}{0.0021748} & \cellcolor{gray!6}{-0.0334619} & \cellcolor{gray!6}{-0.2470344} & \cellcolor{gray!6}{-0.0466343} & \cellcolor{gray!6}{Cloudy skies} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{19.30}\\
+\cellcolor{gray!6}{0.1940} & \cellcolor{gray!6}{0.036800} & \cellcolor{gray!6}{0.1466137} & \cellcolor{gray!6}{0.1161356} & \cellcolor{gray!6}{2.193277} & \cellcolor{gray!6}{0.3829017} & \cellcolor{gray!6}{Cloudy skies} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{96.3}\\
 \cmidrule{1-9}
--0.0427 & 7.98e-01 & 0.0023737 & -0.0332559 & -0.2581092 & -0.0487201 & Cloudy skies & Autumn & 20.20\\
+0.3470 & 0.100000 & 0.0934930 & 0.0611177 & 1.699350 & 0.3057662 & Cloudy skies & Autumn & 90.0\\
 \bottomrule
@@
 over the area. Moreover, we found a similar trend under clear-sky
-conditions (\(0.16\,\%/y\)) that further supports the assumption that
+conditions (\(0.041\,\%/y\)) that further supports the assumption that
 the brightening is caused mainly by decreasing aerosols. Unfortunately,
@@
 Europe have been proposed as major factor by \citet{Wild2021}. The
-dimming effect on SDR under cloudy-sky conditions (\(0.019\,\%/y\)),
+dimming effect on SDR under cloudy-sky conditions (\(0.34\,\%/y\)),
 suggests that cloud optical thickness is decreasing during this period.
@@
 on season, with winter showing the strongest statistically significant
-trend of \(0.69\) and \(0.7\,\%/y\) for all- and clear-skies,
+trend of \(0.69\) and \(0.5\,\%/y\) for all- and clear-skies,
 respectively, in contrast to spring and summer. The trends for autumn
-are also significant but smaller ( \(0.43\) and \(0.13\,\%/y\) for all-
+are also significant but smaller ( \(0.43\) and \(0.11\,\%/y\) for all-
 and clear-skies, respectively). Our findings are in agreement with other

### 2023-12-03 MDPI_submission.Rmd
Commit 2023-12-03 03:09

@@
 <!-- a) for daily mean values we exclude days with less than `r Daily_aggregation_N_lim` valid data points, -->
-a) for the daily mean values we accept days with more than `r 100 * All_daily_ratio_lim`% of the daytime measurements present and valid,
-b) monthly values were computed from daily means
+a) we accept only days with more than `r 100 * All_daily_ratio_lim`% of the daytime measurements present and valid,
+b) on the dataset of clear- and cloudy-skies, we included only days with more than
+`r 100 * Clear_daily_ratio_lim`% of the daytime measurements identified as clear or cloudy respectively,
+c) monthly values were computed from daily means
 only when at least `r Monthly_aggegation_N_lim` days were available.
@@
 LT$slope.p  <- signif(LT$slope.p, 3)
+LT$Rsqrd    <- signif(LT$Rsqrd,   3)
+
+LT$RsqrdAdj <- NULL
+LT$cor.t    <- NULL
+
 ## rename sky conditions
@@
 LT$DATA <- sub(" sky", " skies", LT$DATA)
+
+LT <- rename(LT, "Trend [%/year]" = slope)
+LT <- rename(LT, "Sky condition"  = DATA )
+LT <- rename(LT, "Days"           = N    )
 
@@
         }
-        \caption{Long term trends of SDR as a function of SZA separately form morning and afternoon periods.
+        \caption{Long term trends of SDR as a function of SZA separately from morning and afternoon periods.
 Solid shapes  represent statistically significant trends ($p < 0.005$).}\label{fig:szatrends}
@@
 
-<!-- ```{r szatrends, echo = FALSE, fig.cap='Long term trends of SDR as a function of SZA separately form morning and afternoon periods.
+<!-- ```{r szatrends, echo = FALSE, fig.cap='Long term trends of SDR as a function of SZA separately from morning and afternoon periods.
 Solid shapes  represent statistically significant trends ($p < 0.005$).', fig.subcap=c('All skies.', 'Clear skies.', 'Cloudy skies.'), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3} -->
@@
 spp$slope.stat_sig <- signif(spp$slope.stat_sig, 3)
+spp$Rsqrd          <- signif(spp$Rsqrd,          3)
+spp$RsqrdAdj <- NULL
+spp$cor.t    <- NULL
 

### 2023-12-03 MDPI_submission.tex
Commit 2023-12-03 03:09

@@
 time-scales. To preserve the representativeness of the data we used the
-following criteria: a) for the daily mean values we accept days with
-more than 50\% of the daytime measurements present and valid, b) monthly
+following criteria: a) we accept only days with more than 50\% of the
+daytime measurements present and valid, b) on the dataset of clear- and
+cloudy-skies, we included only days with more than 30\% of the daytime
+measurements identified as clear or cloudy respectively, c) monthly
 values were computed from daily means only when at least 20 days were
@@
 \caption{\label{tab:trendtable}Trends in SDR daily means for different sky conditions for the period 1993 - 2023.}
-\begin{tabu} to \linewidth {>{\centering\arraybackslash}p{8em}>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X>{\centering}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X>{\centering}X}
+\begin{tabu} to \linewidth {>{\centering\arraybackslash}p{8em}>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X>{\centering}X>{\raggedleft}X>{\raggedleft}X}
 \toprule
-slope & slope.p & Rsqrd & RsqrdAdj & cor.t & cor.estimate & DATA & N & slope.stat\_sig\\
+Trend [\%/year] & slope.p & Rsqrd & cor.estimate & Sky condition & Days & slope.stat\_sig\\
 \midrule
-0.3760 & 0.0000 & 0.008339 & 0.0082421 & 9.284 & 0.09132 & All skies & 10251 & 100.00\\
-0.3430 & 0.0000 & 0.005312 & 0.0051756 & 6.234 & 0.07289 & Cloudy skies & 7279 & 100.00\\
-0.0406 & 0.0141 & 0.001195 & 0.0009971 & 2.455 & 0.03458 & Clear skies & 5036 & 98.59\\
+0.3760 & 0.0000 & 0.00834 & 0.09132 & All skies & 10251 & 100.00\\
+0.3430 & 0.0000 & 0.00531 & 0.07289 & Cloudy skies & 7279 & 100.00\\
+0.0406 & 0.0141 & 0.00120 & 0.03458 & Clear skies & 5036 & 98.59\\
 \bottomrule
@@
 \caption{\label{tab:trendseasontable}SDR linear trends of monthly anomalies for each season of the year.}
-\begin{tabu} to \linewidth {>{\centering\arraybackslash}p{8em}>{\centering}X>{\raggedleft}X>{\raggedleft}X>{\centering}X>{\centering}X>{\raggedleft}X>{\raggedleft}X>{\centering}X}
+\begin{tabu} to \linewidth {>{\centering\arraybackslash}p{8em}>{\centering}X>{\raggedleft}X>{\raggedleft}X>{\centering}X>{\centering}X>{\raggedleft}X}
 \toprule
-Trend [\%/year] & slope.p & Rsqrd & RsqrdAdj & cor.t & cor.estimate & Sky condition & Season & slope.stat\_sig\\
+Trend [\%/year] & slope.p & Rsqrd & cor.estimate & Sky condition & Season & slope.stat\_sig\\
 \midrule
-\cellcolor{gray!6}{0.6870} & \cellcolor{gray!6}{0.001330} & \cellcolor{gray!6}{0.3121839} & \cellcolor{gray!6}{0.2876190} & \cellcolor{gray!6}{3.564906} & \cellcolor{gray!6}{0.5587342} & \cellcolor{gray!6}{All skies} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{99.9}\\
-\cmidrule{1-9}
-0.1460 & 0.178000 & 0.0616942 & 0.0293388 & 1.380858 & 0.2483832 & All skies & Spring & 82.2\\
-\cmidrule{1-9}
-\cellcolor{gray!6}{0.1230} & \cellcolor{gray!6}{0.096100} & \cellcolor{gray!6}{0.0957525} & \cellcolor{gray!6}{0.0634580} & \cellcolor{gray!6}{1.721910} & \cellcolor{gray!6}{0.3094390} & \cellcolor{gray!6}{All skies} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{90.4}\\
-\cmidrule{1-9}
-0.4300 & 0.006670 & 0.2347063 & 0.2073744 & 2.930403 & 0.4844650 & All skies & Autumn & 99.3\\
-\cmidrule{1-9}
-\cellcolor{gray!6}{0.4990} & \cellcolor{gray!6}{0.000166} & \cellcolor{gray!6}{0.4025975} & \cellcolor{gray!6}{0.3812617} & \cellcolor{gray!6}{4.343912} & \cellcolor{gray!6}{0.6345057} & \cellcolor{gray!6}{Clear skies} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{100.0}\\
-\cmidrule{1-9}
--0.0865 & 0.315000 & 0.0348099 & 0.0015275 & -1.022690 & -0.1865741 & Clear skies & Spring & 68.5\\
-\cmidrule{1-9}
-\cellcolor{gray!6}{-0.2180} & \cellcolor{gray!6}{0.000398} & \cellcolor{gray!6}{0.3659586} & \cellcolor{gray!6}{0.3433142} & \cellcolor{gray!6}{-4.020094} & \cellcolor{gray!6}{-0.6049451} & \cellcolor{gray!6}{Clear skies} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{100.0}\\
-\cmidrule{1-9}
-0.1090 & 0.221000 & 0.0529350 & 0.0191112 & 1.251009 & 0.2300761 & Clear skies & Autumn & 77.9\\
-\cmidrule{1-9}
-\cellcolor{gray!6}{0.5200} & \cellcolor{gray!6}{0.009580} & \cellcolor{gray!6}{0.2164332} & \cellcolor{gray!6}{0.1884487} & \cellcolor{gray!6}{2.781012} & \cellcolor{gray!6}{0.4652238} & \cellcolor{gray!6}{Cloudy skies} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{99.0}\\
-\cmidrule{1-9}
-0.1780 & 0.122000 & 0.0806332 & 0.0489309 & 1.594819 & 0.2839598 & Cloudy skies & Spring & 87.8\\
-\cmidrule{1-9}
-\cellcolor{gray!6}{0.1940} & \cellcolor{gray!6}{0.036800} & \cellcolor{gray!6}{0.1466137} & \cellcolor{gray!6}{0.1161356} & \cellcolor{gray!6}{2.193277} & \cellcolor{gray!6}{0.3829017} & \cellcolor{gray!6}{Cloudy skies} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{96.3}\\
-\cmidrule{1-9}
-0.3470 & 0.100000 & 0.0934930 & 0.0611177 & 1.699350 & 0.3057662 & Cloudy skies & Autumn & 90.0\\
+\cellcolor{gray!6}{0.6870} & \cellcolor{gray!6}{0.001330} & \cellcolor{gray!6}{0.3120} & \cellcolor{gray!6}{0.5587342} & \cellcolor{gray!6}{All skies} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{99.9}\\
+\cmidrule{1-7}
+0.1460 & 0.178000 & 0.0617 & 0.2483832 & All skies & Spring & 82.2\\
+\cmidrule{1-7}
+\cellcolor{gray!6}{0.1230} & \cellcolor{gray!6}{0.096100} & \cellcolor{gray!6}{0.0958} & \cellcolor{gray!6}{0.3094390} & \cellcolor{gray!6}{All skies} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{90.4}\\
+\cmidrule{1-7}
+0.4300 & 0.006670 & 0.2350 & 0.4844650 & All skies & Autumn & 99.3\\
+\cmidrule{1-7}
+\cellcolor{gray!6}{0.4990} & \cellcolor{gray!6}{0.000166} & \cellcolor{gray!6}{0.4030} & \cellcolor{gray!6}{0.6345057} & \cellcolor{gray!6}{Clear skies} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{100.0}\\
+\cmidrule{1-7}
+-0.0865 & 0.315000 & 0.0348 & -0.1865741 & Clear skies & Spring & 68.5\\
+\cmidrule{1-7}
+\cellcolor{gray!6}{-0.2180} & \cellcolor{gray!6}{0.000398} & \cellcolor{gray!6}{0.3660} & \cellcolor{gray!6}{-0.6049451} & \cellcolor{gray!6}{Clear skies} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{100.0}\\
+\cmidrule{1-7}
+0.1090 & 0.221000 & 0.0529 & 0.2300761 & Clear skies & Autumn & 77.9\\
+\cmidrule{1-7}
+\cellcolor{gray!6}{0.5200} & \cellcolor{gray!6}{0.009580} & \cellcolor{gray!6}{0.2160} & \cellcolor{gray!6}{0.4652238} & \cellcolor{gray!6}{Cloudy skies} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{99.0}\\
+\cmidrule{1-7}
+0.1780 & 0.122000 & 0.0806 & 0.2839598 & Cloudy skies & Spring & 87.8\\
+\cmidrule{1-7}
+\cellcolor{gray!6}{0.1940} & \cellcolor{gray!6}{0.036800} & \cellcolor{gray!6}{0.1470} & \cellcolor{gray!6}{0.3829017} & \cellcolor{gray!6}{Cloudy skies} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{96.3}\\
+\cmidrule{1-7}
+0.3470 & 0.100000 & 0.0935 & 0.3057662 & Cloudy skies & Autumn & 90.0\\
 \bottomrule

### 2023-12-03 MDPI_submission.Rmd
Commit 2023-12-03 09:10
Commit 2023-12-03 20:14

@@
 ## number of measurements for each dataset
-SDR_N_All   <- sum(  ALL_1_daily_DESEAS[,GLB_att_N])
-SDR_N_Clear <- sum(CLEAR_1_daily_DESEAS[,GLB_att_N])
-SDR_N_Cloud <- sum(CLOUD_1_daily_DESEAS[,GLB_att_N])
+SDR_N_All   <- sum(  ALL_1_daily_DESEAS[,GLB_att_N], na.rm = T)
+SDR_N_Clear <- sum(CLEAR_1_daily_DESEAS[,GLB_att_N], na.rm = T)
+SDR_N_Cloud <- sum(CLOUD_1_daily_DESEAS[,GLB_att_N], na.rm = T)
 ## TSI data span info
@@
 ## numeric precision of table
-LT$slope    <- signif(LT$slope,   3)
-LT$slope.p  <- signif(LT$slope.p, 3)
-LT$Rsqrd    <- signif(LT$Rsqrd,   3)
+LT$slope    <- signif(LT$slope,   4)
+LT$slope.p  <- signif(LT$slope.p, 4)
+LT$Rsqrd    <- signif(LT$Rsqrd,   4)
 
@@
 
-LT <- rename(LT, "Trend [%/year]" = slope)
-LT <- rename(LT, "Sky condition"  = DATA )
+LT <- rename(LT, "Trend [%/year]" = slope       )
+LT <- rename(LT, "Sky conditions" = DATA        )
 LT <- rename(LT, "Days"           = N           )
+LT <- rename(LT, "Pearson cor."   = cor.estimate)
+LT <- rename(LT, "Trend p-value"  = slope.p     )
 
@@
 ## order columns
-# LT1 <- LT[,c(
-#       "Sky conditions",
-#       "Trend [%/year]",
-#       "Trend p-value",
-#       "Days with data"
-#       )]
+LT1 <- LT[,c(
+      "Sky conditions",
+      "Trend [%/year]",
+      "Trend p-value",
+      "Pearson cor.",
+      "Days"
+      )]
 
@@
 
-x_latex <- knitr::kable(LT2,
+x_latex <- knitr::kable(LT1,
                         format      = "latex",
@@
                         escape      = TRUE,
-                        format.args = list(digits = 4),
+                        format.args = list(digits = 3),
                         linesep     = "",
-                        align       = "crrr",
+                        align       = "crrrr",
                         centering   = FALSE,

### 2023-12-03 MDPI_submission.tex
Commit 2023-12-03 20:14

@@
 initial daily values of this dataset were interpolated to match the time
-step of our measurements. In the final dataset \(NA\%\) of the 1-minute
-data were identified as under clear-sky conditions and \(NA\%\) as under
-cloud-sky conditions.
+step of our measurements. In the final dataset \(35.7\%\) of the
+1-minute data were identified as under clear-sky conditions and
+\(55.3\%\) as under cloud-sky conditions.
 
@@
 \caption{\label{tab:trendtable}Trends in SDR daily means for different sky conditions for the period 1993 - 2023.}
-\begin{tabu} to \linewidth {>{\centering\arraybackslash}p{8em}>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X>{\centering}X>{\raggedleft}X>{\raggedleft}X}
+\begin{tabu} to \linewidth {>{\centering\arraybackslash}p{8em}>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X}
 \toprule
-Trend [\%/year] & slope.p & Rsqrd & cor.estimate & Sky condition & Days & slope.stat\_sig\\
+Sky conditions & Trend [\%/year] & Trend p-value & Pearson cor. & Days\\
 \midrule
-0.3760 & 0.0000 & 0.00834 & 0.09132 & All skies & 10251 & 100.00\\
-0.3430 & 0.0000 & 0.00531 & 0.07289 & Cloudy skies & 7279 & 100.00\\
-0.0406 & 0.0141 & 0.00120 & 0.03458 & Clear skies & 5036 & 98.59\\
+All skies & 0.3765 & 0.0000 & 0.0913 & 10251\\
+Cloudy skies & 0.3432 & 0.0000 & 0.0729 & 7279\\
+Clear skies & 0.0406 & 0.0141 & 0.0346 & 5036\\
 \bottomrule

### 2023-12-04 MDPI_submission.Rmd
Commit 2023-12-04 10:14

@@
 SDR_N_Cloud <- sum(CLOUD_1_daily_DESEAS[,GLB_att_N], na.rm = T)
+
+# CLEAR_1_daily_DESEAS[!is.na(GLB_att), .(sum(GLB_att_N), .N)]
+
 ## TSI data span info

### 2023-12-04 MDPI_submission.tex
Commit 2023-12-04 10:14

@@
 All skies & 0.3765 & 0.0000 & 0.0913 & 10251\\
-Cloudy skies & 0.3432 & 0.0000 & 0.0729 & 7279\\
-Clear skies & 0.0406 & 0.0141 & 0.0346 & 5036\\
+Cloudy skies & 0.3434 & 0.0000 & 0.0729 & 7278\\
+Clear skies & 0.0408 & 0.0135 & 0.0348 & 5035\\
 \bottomrule
@@
 \cmidrule{1-7}
-\cellcolor{gray!6}{0.4990} & \cellcolor{gray!6}{0.000166} & \cellcolor{gray!6}{0.4030} & \cellcolor{gray!6}{0.6345057} & \cellcolor{gray!6}{Clear skies} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{100.0}\\
+\cellcolor{gray!6}{0.4980} & \cellcolor{gray!6}{0.000158} & \cellcolor{gray!6}{0.4050} & \cellcolor{gray!6}{0.6361307} & \cellcolor{gray!6}{Clear skies} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{100.0}\\
 \cmidrule{1-7}
@@
 \cmidrule{1-7}
-\cellcolor{gray!6}{0.1940} & \cellcolor{gray!6}{0.036800} & \cellcolor{gray!6}{0.1470} & \cellcolor{gray!6}{0.3829017} & \cellcolor{gray!6}{Cloudy skies} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{96.3}\\
+\cellcolor{gray!6}{0.1940} & \cellcolor{gray!6}{0.036000} & \cellcolor{gray!6}{0.1480} & \cellcolor{gray!6}{0.3843557} & \cellcolor{gray!6}{Cloudy skies} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{96.4}\\
 \cmidrule{1-7}

### 2023-12-04 DHI_GHI_01_Input_longterm.R

@@ -9,6 +9,13 @@ source("~/CODE/FUNCTIONS/R/data.R")
 source("./DHI_GHI_0_variables.R")
 Script.Name <- "DHI_GHI_01_Input_longterm.R"
 
+if (!interactive()) {
+    pdf( file = paste0("./runtime/",  basename(sub("\\.R$",".pdf", Script.Name))))
+    sink(file = paste0("./runtime/",  basename(sub("\\.R$",".out", Script.Name))), split = TRUE)
+    filelock::lock(paste0("./runtime/", basename(sub("\\.R$",".lock", Script.Name))), timeout = 0)
+}
+
+
 ##  Prepare raw data if needed  ------------------------------------------------
 ## check previous steps
 if (
@@ -50,6 +57,7 @@ DATA_Cloud[,.N]
 DATA_all[,length(unique(as.Date(Date)))]
 DATA_Clear[,length(unique(as.Date(Date)))]
 DATA_Cloud[,length(unique(as.Date(Date)))]
+
 
 
 
@@ -119,17 +127,22 @@ CLOUD_1_daily_mean <-
 hist(CLEAR_1_daily_mean[, GLB_att_N / DayLength], breaks = 100)
 abline(v = Clear_daily_ratio_lim, col = "red")
 
-CLEAR_1_daily_mean[GLB_att_N / DayLength < Clear_daily_ratio_lim, GLB_att    := NA]
-CLEAR_1_daily_mean[GLB_att_N / DayLength < Clear_daily_ratio_lim, GLB_att_N  := NA]
-CLEAR_1_daily_mean[GLB_att_N / DayLength < Clear_daily_ratio_lim, GLB_att_sd := NA]
-
 hist(CLOUD_1_daily_mean[, GLB_att_N / DayLength], breaks = 100)
 abline(v = Cloud_daily_ratio_lim, col = "red")
 
-CLOUD_1_daily_mean[GLB_att_N / DayLength < Clear_daily_ratio_lim, GLB_att    := NA]
-CLOUD_1_daily_mean[GLB_att_N / DayLength < Clear_daily_ratio_lim, GLB_att_N  := NA]
-CLOUD_1_daily_mean[GLB_att_N / DayLength < Clear_daily_ratio_lim, GLB_att_sd := NA]
+# ## proper way to apply filter
[…]

### 2023-12-04 DHI_GHI_02_Input_szatrends.R

@@ -10,7 +10,13 @@ require(zoo)
 source("~/CODE/FUNCTIONS/R/trig_deg.R")
 source("~/CODE/FUNCTIONS/R/data.R")
 source("./DHI_GHI_0_variables.R")
+Script.Name <- "DHI_GHI_02_Input_szatrends.R"
 
+if (!interactive()) {
+    pdf( file = paste0("./runtime/",  basename(sub("\\.R$",".pdf", Script.Name))))
+    sink(file = paste0("./runtime/",  basename(sub("\\.R$",".out", Script.Name))), split = TRUE)
+    filelock::lock(paste0("./runtime/", basename(sub("\\.R$",".lock", Script.Name))), timeout = 0)
+}
 
 ##  Prepare raw data if needed  ------------------------------------------------
 ## check previous steps

### 2023-12-04 MDPI_submission.tex
Commit 2023-12-04 12:07

@@
 All skies & 0.3765 & 0.0000 & 0.0913 & 10251\\
+Clear skies & 0.0408 & 0.0135 & 0.0348 & 5035\\
 Cloudy skies & 0.3434 & 0.0000 & 0.0729 & 7278\\
-Clear skies & 0.0408 & 0.0135 & 0.0348 & 5035\\
 \bottomrule

### 2023-12-04 DHI_GHI_01_Input_longterm.R

@@ -123,12 +123,26 @@ CLOUD_1_daily_mean <-
 
 
 
+hist(ALL_1_daily_mean$GLB_att,   breaks = 100)
+
+hist(CLEAR_1_daily_mean$GLB_att, breaks = 100)
+
+hist(CLOUD_1_daily_mean$GLB_att, breaks = 100)
+
+
+
+
 ## _ Exclude days with few data for Clear and cloud  ---------------------------
 hist(CLEAR_1_daily_mean[, GLB_att_N / DayLength], breaks = 100)
 abline(v = Clear_daily_ratio_lim, col = "red")
 
 hist(CLOUD_1_daily_mean[, GLB_att_N / DayLength], breaks = 100)
 abline(v = Cloud_daily_ratio_lim, col = "red")
+
+
+CLEAR_1_daily_mean[!is.na(GLB_att) & GLB_att_N / DayLength > Clear_daily_ratio_lim, sum(GLB_att_N) ]
+CLOUD_1_daily_mean[!is.na(GLB_att) & GLB_att_N / DayLength > Cloud_daily_ratio_lim, sum(GLB_att_N) ]
+
 
 # ## proper way to apply filter
 # CLEAR_1_daily_mean[GLB_att_N / DayLength < Clear_daily_ratio_lim, GLB_att    := NA]
@@ -217,6 +231,16 @@ CLOUD_1_daily_seas <-
                           DIR_att_N_seas     = sum(!is.na(DIR_att))  ),
                        by = .( doy ) ]
 
+
+
+plot(ALL_1_daily_seas[, GLB_att_seas, doy],pch= 1, cex = 1,
+     main = "ALL_1_daily_seas[, GLB_att_seas, doy]")
+
+plot(CLEAR_1_daily_seas[, GLB_att_seas, doy],pch= 1, cex = 1,
+     main = "CLEAR_1_daily_seas[, GLB_att_seas, doy]")
+
+plot(CLOUD_1_daily_seas[, GLB_att_seas, doy],pch= 1, cex = 1,
+     main = "CLOUD_1_daily_seas[, GLB_att_seas, doy]")
 
 
 ## _ Margin of error for confidence interval on seasonal data ------------------
@@ -333,6 +357,13 @@ CLOUD_1_monthly_daily_mean <-
                        ) ]
 
 
+hist(ALL_1_monthly_daily_mean$GLB_att, breaks = 100)
+
[…]

### 2023-12-04 DHI_GHI_02_Input_szatrends.R

@@ -2,9 +2,6 @@
 
 ## Data input for this paper
 
-## to force a rebuild of the dataset remove stored
-# file.remove(common_data)
-
 require(data.table)
 require(zoo)
 source("~/CODE/FUNCTIONS/R/trig_deg.R")

### 2023-12-04 MDPI_submission.tex
Commit 2023-12-04 15:12

@@
 initial daily values of this dataset were interpolated to match the time
-step of our measurements. In the final dataset \(35.7\%\) of the
-1-minute data were identified as under clear-sky conditions and
-\(55.3\%\) as under cloud-sky conditions.
+step of our measurements. In the final dataset \(38\%\) of the 1-minute
+data were identified as under clear-sky conditions and \(57.8\%\) as
+under cloud-sky conditions.
 
@@
 daytime measurements present and valid, b) on the dataset of clear- and
-cloudy-skies, we included only days with more than 30\% of the daytime
+cloudy-skies, we included only days with more than 20\% of the daytime
 measurements identified as clear or cloudy respectively, c) monthly
@@
 positive trend for all-sky conditions (\(0.38\,\%/y\)), a very close but
-smaller trend for clear-skies (\(0.041\,\%/y\)) and a negative weaker
+smaller trend for clear-skies (\(0.024\,\%/y\)) and a negative weaker
 trend for cloudy-skies (\(0.34\,\%/y\)). In the studied period, there is
@@
 \midrule
-All skies & 0.3765 & 0.0000 & 0.0913 & 10251\\
-Clear skies & 0.0408 & 0.0135 & 0.0348 & 5035\\
-Cloudy skies & 0.3434 & 0.0000 & 0.0729 & 7278\\
+All skies & 0.3765 & 0.000 & 0.0913 & 10251\\
+Clear skies & 0.0244 & 0.179 & 0.0176 & 5804\\
+Cloudy skies & 0.3416 & 0.000 & 0.0725 & 8113\\
 \bottomrule
@@
 \midrule
-\cellcolor{gray!6}{0.6870} & \cellcolor{gray!6}{0.001330} & \cellcolor{gray!6}{0.3120} & \cellcolor{gray!6}{0.5587342} & \cellcolor{gray!6}{All skies} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{99.9}\\
+\cellcolor{gray!6}{0.6870} & \cellcolor{gray!6}{1.33e-03} & \cellcolor{gray!6}{0.3120} & \cellcolor{gray!6}{0.5587342} & \cellcolor{gray!6}{All skies} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{99.9}\\
 \cmidrule{1-7}
-0.1460 & 0.178000 & 0.0617 & 0.2483832 & All skies & Spring & 82.2\\
+0.1460 & 1.78e-01 & 0.0617 & 0.2483832 & All skies & Spring & 82.2\\
 \cmidrule{1-7}
-\cellcolor{gray!6}{0.1230} & \cellcolor{gray!6}{0.096100} & \cellcolor{gray!6}{0.0958} & \cellcolor{gray!6}{0.3094390} & \cellcolor{gray!6}{All skies} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{90.4}\\
+\cellcolor{gray!6}{0.1230} & \cellcolor{gray!6}{9.61e-02} & \cellcolor{gray!6}{0.0958} & \cellcolor{gray!6}{0.3094390} & \cellcolor{gray!6}{All skies} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{90.4}\\
 \cmidrule{1-7}
-0.4300 & 0.006670 & 0.2350 & 0.4844650 & All skies & Autumn & 99.3\\
+0.4300 & 6.67e-03 & 0.2350 & 0.4844650 & All skies & Autumn & 99.3\\
 \cmidrule{1-7}
-\cellcolor{gray!6}{0.4980} & \cellcolor{gray!6}{0.000158} & \cellcolor{gray!6}{0.4050} & \cellcolor{gray!6}{0.6361307} & \cellcolor{gray!6}{Clear skies} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{100.0}\\
+\cellcolor{gray!6}{0.5050} & \cellcolor{gray!6}{4.16e-04} & \cellcolor{gray!6}{0.3640} & \cellcolor{gray!6}{0.6033797} & \cellcolor{gray!6}{Clear skies} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{100.0}\\
 \cmidrule{1-7}
--0.0865 & 0.315000 & 0.0348 & -0.1865741 & Clear skies & Spring & 68.5\\
+-0.1360 & 1.71e-01 & 0.0635 & -0.2520277 & Clear skies & Spring & 82.9\\
 \cmidrule{1-7}
-\cellcolor{gray!6}{-0.2180} & \cellcolor{gray!6}{0.000398} & \cellcolor{gray!6}{0.3660} & \cellcolor{gray!6}{-0.6049451} & \cellcolor{gray!6}{Clear skies} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{100.0}\\
+\cellcolor{gray!6}{-0.2310} & \cellcolor{gray!6}{3.01e-05} & \cellcolor{gray!6}{0.4690} & \cellcolor{gray!6}{-0.6846238} & \cellcolor{gray!6}{Clear skies} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{100.0}\\
 \cmidrule{1-7}
-0.1090 & 0.221000 & 0.0529 & 0.2300761 & Clear skies & Autumn & 77.9\\
+0.0432 & 5.79e-01 & 0.0111 & 0.1054619 & Clear skies & Autumn & 42.1\\
 \cmidrule{1-7}
-\cellcolor{gray!6}{0.5200} & \cellcolor{gray!6}{0.009580} & \cellcolor{gray!6}{0.2160} & \cellcolor{gray!6}{0.4652238} & \cellcolor{gray!6}{Cloudy skies} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{99.0}\\
+\cellcolor{gray!6}{0.4720} & \cellcolor{gray!6}{2.01e-02} & \cellcolor{gray!6}{0.1780} & \cellcolor{gray!6}{0.4221299} & \cellcolor{gray!6}{Cloudy skies} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{98.0}\\
 \cmidrule{1-7}
-0.1780 & 0.122000 & 0.0806 & 0.2839598 & Cloudy skies & Spring & 87.8\\
+0.1780 & 1.25e-01 & 0.0792 & 0.2814813 & Cloudy skies & Spring & 87.5\\
 \cmidrule{1-7}
-\cellcolor{gray!6}{0.1940} & \cellcolor{gray!6}{0.036000} & \cellcolor{gray!6}{0.1480} & \cellcolor{gray!6}{0.3843557} & \cellcolor{gray!6}{Cloudy skies} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{96.4}\\
+\cellcolor{gray!6}{0.2130} & \cellcolor{gray!6}{2.20e-02} & \cellcolor{gray!6}{0.1740} & \cellcolor{gray!6}{0.4167034} & \cellcolor{gray!6}{Cloudy skies} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{97.8}\\
 \cmidrule{1-7}
-0.3470 & 0.100000 & 0.0935 & 0.3057662 & Cloudy skies & Autumn & 90.0\\
+0.4140 & 2.46e-02 & 0.1680 & 0.4094891 & Cloudy skies & Autumn & 97.5\\
 \bottomrule
@@
 over the area. Moreover, we found a similar trend under clear-sky
-conditions (\(0.041\,\%/y\)) that further supports the assumption that
+conditions (\(0.024\,\%/y\)) that further supports the assumption that
 the brightening is caused mainly by decreasing aerosols. Unfortunately,
@@
 respectively, in contrast to spring and summer. The trends for autumn
-are also significant but smaller ( \(0.43\) and \(0.11\,\%/y\) for all-
+are also significant but smaller ( \(0.43\) and \(0.043\,\%/y\) for all-
 and clear-skies, respectively). Our findings are in agreement with other

### 2023-12-04 MDPI_submission.tex
Commit 2023-12-04 21:25

@@
 \(22^\circ\,57'\,\)E, \(80\,\)m~a.s.l.) in the period from 1993-04-13 to
-2023-04-13. The monitoring site is located near the city centre, and we
+2023-04-12. The monitoring site is located near the city centre, and we
 expect to be affected by the urban environment. During the study period,
@@
 beam radiation (DNI) was also measured by a collocated Kipp \& Zonen
-CHP-1 pyrheliometer, for the period 2016-04-01 to 2023-04-13. Although,
+CHP-1 pyrheliometer, for the period 2016-04-01 to 2023-04-12. Although,
 we have performed a similar analysis to the DNI data, the results are
@@
 before the start of our records. It is interesting to note, that for the
-observations period, the trend of the TSI is \(-0.00022\,\%/y\), and
+observations period, the trend of the TSI is \(-0.00024\,\%/y\), and
 thus we can not attribute any major effect on SDR trend to Solar
@@
 \midrule
-All skies & 0.3765 & 0.000 & 0.0913 & 10251\\
-Clear skies & 0.0244 & 0.179 & 0.0176 & 5804\\
+All skies & 0.3756 & 0.000 & 0.0911 & 10250\\
+Clear skies & 0.0239 & 0.189 & 0.0172 & 5803\\
 Cloudy skies & 0.3416 & 0.000 & 0.0725 & 8113\\
@@
 \cmidrule{1-7}
-0.1460 & 1.78e-01 & 0.0617 & 0.2483832 & All skies & Spring & 82.2\\
+0.1360 & 2.19e-01 & 0.0515 & 0.2270307 & All skies & Spring & 78.1\\
 \cmidrule{1-7}
@@
 \cmidrule{1-7}
--0.1360 & 1.71e-01 & 0.0635 & -0.2520277 & Clear skies & Spring & 82.9\\
+-0.1450 & 1.49e-01 & 0.0706 & -0.2656359 & Clear skies & Spring & 85.1\\
 \cmidrule{1-7}

### 2023-12-05 MDPI_submission.tex
Commit 2023-12-05 21:10

@@
 \(22^\circ\,57'\,\)E, \(80\,\)m~a.s.l.) in the period from 1993-04-13 to
-2023-04-12. The monitoring site is located near the city centre, and we
+2023-04-13. The monitoring site is located near the city centre, and we
 expect to be affected by the urban environment. During the study period,
@@
 beam radiation (DNI) was also measured by a collocated Kipp \& Zonen
-CHP-1 pyrheliometer, for the period 2016-04-01 to 2023-04-12. Although,
+CHP-1 pyrheliometer, for the period 2016-04-01 to 2023-04-13. Although,
 we have performed a similar analysis to the DNI data, the results are
@@
 initial daily values of this dataset were interpolated to match the time
-step of our measurements. In the final dataset \(38\%\) of the 1-minute
-data were identified as under clear-sky conditions and \(57.8\%\) as
-under cloud-sky conditions.
+step of our measurements. In the final dataset \(23.3\%\) of the
+1-minute data were identified as under clear-sky conditions and \(43\%\)
+as under cloud-sky conditions.
 
@@
 daytime measurements present and valid, b) on the dataset of clear- and
-cloudy-skies, we included only days with more than 20\% of the daytime
+cloudy-skies, we included only days with more than 60\% of the daytime
 measurements identified as clear or cloudy respectively, c) monthly

### 2023-12-06 MDPI_submission.Rmd
Commit 2023-12-04 12:07
Commit 2023-12-06 21:11

@@
 # LT$cor.p            <- NULL
-LT$conf.level       <- NULL
+LT$cor.conf_level   <- NULL
 LT$cor.conf_int_1   <- NULL
@@
 LT$DATA <- sub(" sky", " skies", LT$DATA)
+
+LT <- LT[order(LT$DATA),]
 
@@
 spp$cor.p              <- NULL
-spp$conf.level         <- NULL
+spp$cor.conf.level     <- NULL
 spp$cor.conf_int_1     <- NULL

### 2023-12-06 MDPI_submission.tex
Commit 2023-12-06 21:11

--- 2023-12-04 MDPI_submission.tex
@@
 
+\begin{verbatim}
+Warning in set(x, j = name, value = value): Column 'conf.level' does not exist
+to remove
+\end{verbatim}
+
 For completeness, we will provide below a brief overview of the clear
@@
 daytime measurements present and valid, b) on the dataset of clear- and
-cloudy-skies, we included only days with more than 20\% of the daytime
+cloudy-skies, we included only days with more than 70\% of the daytime
 measurements identified as clear or cloudy respectively, c) monthly
@@
 positive trend for all-sky conditions (\(0.38\,\%/y\)), a very close but
-smaller trend for clear-skies (\(0.024\,\%/y\)) and a negative weaker
-trend for cloudy-skies (\(0.34\,\%/y\)). In the studied period, there is
+smaller trend for clear-skies (\(0.097\,\%/y\)) and a negative weaker
+trend for cloudy-skies (\(0.41\,\%/y\)). In the studied period, there is
 no significant break or change in the variability of the time series.
@@
 \midrule
-All skies & 0.3765 & 0.000 & 0.0913 & 10251\\
-Clear skies & 0.0244 & 0.179 & 0.0176 & 5804\\
-Cloudy skies & 0.3416 & 0.000 & 0.0725 & 8113\\
+All skies & 0.377 & 0 & 0.0913 & 10251\\
+Clear skies & 0.097 & 0 & 0.1381 & 2689\\
+Cloudy skies & 0.408 & 0 & 0.0810 & 4937\\
 \bottomrule
@@
 of clouds during winter and autumn. For clear-skies, the trend in winter
-is \(0.5\,\%/y\), larger than for all-skies (\(0.69\,\%/y\)), which is
+is \(0.36\,\%/y\), larger than for all-skies (\(0.69\,\%/y\)), which is
 another indication of a decreasing trend in cloud optical thickness.
@@
 \caption{\label{tab:trendseasontable}SDR linear trends of monthly anomalies for each season of the year.}
-\begin{tabu} to \linewidth {>{\centering\arraybackslash}p{8em}>{\centering}X>{\raggedleft}X>{\raggedleft}X>{\centering}X>{\centering}X>{\raggedleft}X}
+\begin{tabu} to \linewidth {>{\centering\arraybackslash}p{8em}>{\centering}X>{\raggedleft}X>{\raggedleft}X>{\centering}X>{\centering}X>{\raggedleft}X>{\raggedleft}X}
 \toprule
-Trend [\%/year] & slope.p & Rsqrd & cor.estimate & Sky condition & Season & slope.stat\_sig\\
+Trend [\%/year] & slope.p & Rsqrd & cor.estimate & cor.conf\_level & Sky condition & Season & slope.stat\_sig\\
 \midrule
-\cellcolor{gray!6}{0.6870} & \cellcolor{gray!6}{1.33e-03} & \cellcolor{gray!6}{0.3120} & \cellcolor{gray!6}{0.5587342} & \cellcolor{gray!6}{All skies} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{99.9}\\
-\cmidrule{1-7}
-0.1460 & 1.78e-01 & 0.0617 & 0.2483832 & All skies & Spring & 82.2\\
-\cmidrule{1-7}
-\cellcolor{gray!6}{0.1230} & \cellcolor{gray!6}{9.61e-02} & \cellcolor{gray!6}{0.0958} & \cellcolor{gray!6}{0.3094390} & \cellcolor{gray!6}{All skies} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{90.4}\\
-\cmidrule{1-7}
-0.4300 & 6.67e-03 & 0.2350 & 0.4844650 & All skies & Autumn & 99.3\\
-\cmidrule{1-7}
-\cellcolor{gray!6}{0.5050} & \cellcolor{gray!6}{4.16e-04} & \cellcolor{gray!6}{0.3640} & \cellcolor{gray!6}{0.6033797} & \cellcolor{gray!6}{Clear skies} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{100.0}\\
-\cmidrule{1-7}
--0.1360 & 1.71e-01 & 0.0635 & -0.2520277 & Clear skies & Spring & 82.9\\
-\cmidrule{1-7}
-\cellcolor{gray!6}{-0.2310} & \cellcolor{gray!6}{3.01e-05} & \cellcolor{gray!6}{0.4690} & \cellcolor{gray!6}{-0.6846238} & \cellcolor{gray!6}{Clear skies} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{100.0}\\
-\cmidrule{1-7}
-0.0432 & 5.79e-01 & 0.0111 & 0.1054619 & Clear skies & Autumn & 42.1\\
-\cmidrule{1-7}
-\cellcolor{gray!6}{0.4720} & \cellcolor{gray!6}{2.01e-02} & \cellcolor{gray!6}{0.1780} & \cellcolor{gray!6}{0.4221299} & \cellcolor{gray!6}{Cloudy skies} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{98.0}\\
-\cmidrule{1-7}
-0.1780 & 1.25e-01 & 0.0792 & 0.2814813 & Cloudy skies & Spring & 87.5\\
-\cmidrule{1-7}
-\cellcolor{gray!6}{0.2130} & \cellcolor{gray!6}{2.20e-02} & \cellcolor{gray!6}{0.1740} & \cellcolor{gray!6}{0.4167034} & \cellcolor{gray!6}{Cloudy skies} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{97.8}\\
-\cmidrule{1-7}
-0.4140 & 2.46e-02 & 0.1680 & 0.4094891 & Cloudy skies & Autumn & 97.5\\
+\cellcolor{gray!6}{0.6870} & \cellcolor{gray!6}{0.001330} & \cellcolor{gray!6}{0.31200} & \cellcolor{gray!6}{0.5587342} & \cellcolor{gray!6}{0.95} & \cellcolor{gray!6}{All skies} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{99.9}\\
+\cmidrule{1-8}
+0.1460 & 0.178000 & 0.06170 & 0.2483832 & 0.95 & All skies & Spring & 82.2\\
+\cmidrule{1-8}
+\cellcolor{gray!6}{0.1230} & \cellcolor{gray!6}{0.096100} & \cellcolor{gray!6}{0.09580} & \cellcolor{gray!6}{0.3094390} & \cellcolor{gray!6}{0.95} & \cellcolor{gray!6}{All skies} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{90.4}\\
+\cmidrule{1-8}
+0.4300 & 0.006670 & 0.23500 & 0.4844650 & 0.95 & All skies & Autumn & 99.3\\
+\cmidrule{1-8}
+\cellcolor{gray!6}{0.3570} & \cellcolor{gray!6}{0.007260} & \cellcolor{gray!6}{0.23000} & \cellcolor{gray!6}{0.4800278} & \cellcolor{gray!6}{0.95} & \cellcolor{gray!6}{Clear skies} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{99.3}\\
+\cmidrule{1-8}
+0.0317 & 0.721000 & 0.00445 & 0.0667338 & 0.95 & Clear skies & Spring & 27.9\\
+\cmidrule{1-8}
+\cellcolor{gray!6}{-0.0931} & \cellcolor{gray!6}{0.007770} & \cellcolor{gray!6}{0.22700} & \cellcolor{gray!6}{-0.4764646} & \cellcolor{gray!6}{0.95} & \cellcolor{gray!6}{Clear skies} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{99.2}\\
+\cmidrule{1-8}
+0.1100 & 0.308000 & 0.03710 & 0.1926085 & 0.95 & Clear skies & Autumn & 69.2\\
+\cmidrule{1-8}
+\cellcolor{gray!6}{0.8250} & \cellcolor{gray!6}{0.000123} & \cellcolor{gray!6}{0.41500} & \cellcolor{gray!6}{0.6439160} & \cellcolor{gray!6}{0.95} & \cellcolor{gray!6}{Cloudy skies} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{100.0}\\
+\cmidrule{1-8}
+0.0942 & 0.425000 & 0.02210 & 0.1485860 & 0.95 & Cloudy skies & Spring & 57.5\\
+\cmidrule{1-8}
+\cellcolor{gray!6}{-0.0983} & \cellcolor{gray!6}{0.414000} & \cellcolor{gray!6}{0.02400} & \cellcolor{gray!6}{-0.1548507} & \cellcolor{gray!6}{0.95} & \cellcolor{gray!6}{Cloudy skies} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{58.6}\\
+\cmidrule{1-8}
+0.1950 & 0.460000 & 0.01960 & 0.1401398 & 0.95 & Cloudy skies & Autumn & 54.0\\
 \bottomrule
@@
 over the area. Moreover, we found a similar trend under clear-sky
-conditions (\(0.024\,\%/y\)) that further supports the assumption that
+conditions (\(0.097\,\%/y\)) that further supports the assumption that
 the brightening is caused mainly by decreasing aerosols. Unfortunately,
@@
 Europe have been proposed as major factor by \citet{Wild2021}. The
-dimming effect on SDR under cloudy-sky conditions (\(0.34\,\%/y\)),
+dimming effect on SDR under cloudy-sky conditions (\(0.41\,\%/y\)),
 suggests that cloud optical thickness is decreasing during this period.
@@
 on season, with winter showing the strongest statistically significant
-trend of \(0.69\) and \(0.5\,\%/y\) for all- and clear-skies,
+trend of \(0.69\) and \(0.36\,\%/y\) for all- and clear-skies,
 respectively, in contrast to spring and summer. The trends for autumn
-are also significant but smaller ( \(0.43\) and \(0.043\,\%/y\) for all-
+are also significant but smaller ( \(0.43\) and \(0.11\,\%/y\) for all-
 and clear-skies, respectively). Our findings are in agreement with other

### 2023-12-06 MDPI_submission.tex
Commit 2023-12-06 23:10

--- 2023-12-04 MDPI_submission.tex
@@
 negative but cannot affect the final results of the study.
+
+\begin{verbatim}
+Warning in set(x, j = name, value = value): Column 'conf.level' does not exist
+to remove
+\end{verbatim}
 
@@
 before the start of our records. It is interesting to note, that for the
-observations period, the trend of the TSI is \(-0.00024\,\%/y\), and
+observations period, the trend of the TSI is \(-0.00022\,\%/y\), and
 thus we can not attribute any major effect on SDR trend to Solar
@@
 \midrule
-All skies & 0.3756 & 0.000 & 0.0911 & 10250\\
-Clear skies & 0.0239 & 0.189 & 0.0172 & 5803\\
+All skies & 0.3765 & 0.000 & 0.0913 & 10251\\
+Clear skies & 0.0244 & 0.179 & 0.0176 & 5804\\
 Cloudy skies & 0.3416 & 0.000 & 0.0725 & 8113\\
@@
 \caption{\label{tab:trendseasontable}SDR linear trends of monthly anomalies for each season of the year.}
-\begin{tabu} to \linewidth {>{\centering\arraybackslash}p{8em}>{\centering}X>{\raggedleft}X>{\raggedleft}X>{\centering}X>{\centering}X>{\raggedleft}X}
+\begin{tabu} to \linewidth {>{\centering\arraybackslash}p{8em}>{\centering}X>{\raggedleft}X>{\raggedleft}X>{\centering}X>{\centering}X>{\raggedleft}X>{\raggedleft}X}
 \toprule
-Trend [\%/year] & slope.p & Rsqrd & cor.estimate & Sky condition & Season & slope.stat\_sig\\
+Trend [\%/year] & slope.p & Rsqrd & cor.estimate & cor.conf\_level & Sky condition & Season & slope.stat\_sig\\
 \midrule
-\cellcolor{gray!6}{0.6870} & \cellcolor{gray!6}{1.33e-03} & \cellcolor{gray!6}{0.3120} & \cellcolor{gray!6}{0.5587342} & \cellcolor{gray!6}{All skies} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{99.9}\\
-\cmidrule{1-7}
-0.1360 & 2.19e-01 & 0.0515 & 0.2270307 & All skies & Spring & 78.1\\
-\cmidrule{1-7}
-\cellcolor{gray!6}{0.1230} & \cellcolor{gray!6}{9.61e-02} & \cellcolor{gray!6}{0.0958} & \cellcolor{gray!6}{0.3094390} & \cellcolor{gray!6}{All skies} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{90.4}\\
-\cmidrule{1-7}
-0.4300 & 6.67e-03 & 0.2350 & 0.4844650 & All skies & Autumn & 99.3\\
-\cmidrule{1-7}
-\cellcolor{gray!6}{0.5050} & \cellcolor{gray!6}{4.16e-04} & \cellcolor{gray!6}{0.3640} & \cellcolor{gray!6}{0.6033797} & \cellcolor{gray!6}{Clear skies} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{100.0}\\
-\cmidrule{1-7}
--0.1450 & 1.49e-01 & 0.0706 & -0.2656359 & Clear skies & Spring & 85.1\\
-\cmidrule{1-7}
-\cellcolor{gray!6}{-0.2310} & \cellcolor{gray!6}{3.01e-05} & \cellcolor{gray!6}{0.4690} & \cellcolor{gray!6}{-0.6846238} & \cellcolor{gray!6}{Clear skies} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{100.0}\\
-\cmidrule{1-7}
-0.0432 & 5.79e-01 & 0.0111 & 0.1054619 & Clear skies & Autumn & 42.1\\
-\cmidrule{1-7}
-\cellcolor{gray!6}{0.4720} & \cellcolor{gray!6}{2.01e-02} & \cellcolor{gray!6}{0.1780} & \cellcolor{gray!6}{0.4221299} & \cellcolor{gray!6}{Cloudy skies} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{98.0}\\
-\cmidrule{1-7}
-0.1780 & 1.25e-01 & 0.0792 & 0.2814813 & Cloudy skies & Spring & 87.5\\
-\cmidrule{1-7}
-\cellcolor{gray!6}{0.2130} & \cellcolor{gray!6}{2.20e-02} & \cellcolor{gray!6}{0.1740} & \cellcolor{gray!6}{0.4167034} & \cellcolor{gray!6}{Cloudy skies} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{97.8}\\
-\cmidrule{1-7}
-0.4140 & 2.46e-02 & 0.1680 & 0.4094891 & Cloudy skies & Autumn & 97.5\\
+\cellcolor{gray!6}{0.6870} & \cellcolor{gray!6}{1.33e-03} & \cellcolor{gray!6}{0.3120} & \cellcolor{gray!6}{0.5587342} & \cellcolor{gray!6}{0.95} & \cellcolor{gray!6}{All skies} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{99.9}\\
+\cmidrule{1-8}
+0.1460 & 1.78e-01 & 0.0617 & 0.2483832 & 0.95 & All skies & Spring & 82.2\\
+\cmidrule{1-8}
+\cellcolor{gray!6}{0.1230} & \cellcolor{gray!6}{9.61e-02} & \cellcolor{gray!6}{0.0958} & \cellcolor{gray!6}{0.3094390} & \cellcolor{gray!6}{0.95} & \cellcolor{gray!6}{All skies} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{90.4}\\
+\cmidrule{1-8}
+0.4300 & 6.67e-03 & 0.2350 & 0.4844650 & 0.95 & All skies & Autumn & 99.3\\
+\cmidrule{1-8}
+\cellcolor{gray!6}{0.5050} & \cellcolor{gray!6}{4.16e-04} & \cellcolor{gray!6}{0.3640} & \cellcolor{gray!6}{0.6033797} & \cellcolor{gray!6}{0.95} & \cellcolor{gray!6}{Clear skies} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{100.0}\\
+\cmidrule{1-8}
+-0.1360 & 1.71e-01 & 0.0635 & -0.2520277 & 0.95 & Clear skies & Spring & 82.9\\
+\cmidrule{1-8}
+\cellcolor{gray!6}{-0.2310} & \cellcolor{gray!6}{3.01e-05} & \cellcolor{gray!6}{0.4690} & \cellcolor{gray!6}{-0.6846238} & \cellcolor{gray!6}{0.95} & \cellcolor{gray!6}{Clear skies} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{100.0}\\
+\cmidrule{1-8}
+0.0432 & 5.79e-01 & 0.0111 & 0.1054619 & 0.95 & Clear skies & Autumn & 42.1\\
+\cmidrule{1-8}
+\cellcolor{gray!6}{0.4720} & \cellcolor{gray!6}{2.01e-02} & \cellcolor{gray!6}{0.1780} & \cellcolor{gray!6}{0.4221299} & \cellcolor{gray!6}{0.95} & \cellcolor{gray!6}{Cloudy skies} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{98.0}\\
+\cmidrule{1-8}
+0.1780 & 1.25e-01 & 0.0792 & 0.2814813 & 0.95 & Cloudy skies & Spring & 87.5\\
+\cmidrule{1-8}
+\cellcolor{gray!6}{0.2130} & \cellcolor{gray!6}{2.20e-02} & \cellcolor{gray!6}{0.1740} & \cellcolor{gray!6}{0.4167034} & \cellcolor{gray!6}{0.95} & \cellcolor{gray!6}{Cloudy skies} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{97.8}\\
+\cmidrule{1-8}
+0.4140 & 2.46e-02 & 0.1680 & 0.4094891 & 0.95 & Cloudy skies & Autumn & 97.5\\
 \bottomrule

### 2023-12-07 MDPI_submission.Rmd
Commit 2023-12-07 11:10

@@
 spp$cor.conf_int_2     <- NULL
+spp$cor.conf_level     <- NULL
 

### 2023-12-07 MDPI_submission.tex
Commit 2023-12-07 16:14

@@
 initial daily values of this dataset were interpolated to match the time
-step of our measurements. In the final dataset \(38\%\) of the 1-minute
-data were identified as under clear-sky conditions and \(57.8\%\) as
-under cloud-sky conditions.
+step of our measurements. In the final dataset \(23.7\%\) of the
+1-minute data were identified as under clear-sky conditions and \(41\%\)
+as under cloud-sky conditions.
 
@@
 \caption{\label{tab:trendseasontable}SDR linear trends of monthly anomalies for each season of the year.}
-\begin{tabu} to \linewidth {>{\centering\arraybackslash}p{8em}>{\centering}X>{\raggedleft}X>{\raggedleft}X>{\centering}X>{\centering}X>{\raggedleft}X>{\raggedleft}X}
+\begin{tabu} to \linewidth {>{\centering\arraybackslash}p{8em}>{\centering}X>{\raggedleft}X>{\raggedleft}X>{\centering}X>{\centering}X>{\raggedleft}X}
 \toprule
-Trend [\%/year] & slope.p & Rsqrd & cor.estimate & cor.conf\_level & Sky condition & Season & slope.stat\_sig\\
+Trend [\%/year] & slope.p & Rsqrd & cor.estimate & Sky condition & Season & slope.stat\_sig\\
 \midrule
-\cellcolor{gray!6}{0.6870} & \cellcolor{gray!6}{1.33e-03} & \cellcolor{gray!6}{0.3120} & \cellcolor{gray!6}{0.5587342} & \cellcolor{gray!6}{0.95} & \cellcolor{gray!6}{All skies} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{99.9}\\
-\cmidrule{1-8}
-0.1460 & 1.78e-01 & 0.0617 & 0.2483832 & 0.95 & All skies & Spring & 82.2\\
-\cmidrule{1-8}
-\cellcolor{gray!6}{0.1230} & \cellcolor{gray!6}{9.61e-02} & \cellcolor{gray!6}{0.0958} & \cellcolor{gray!6}{0.3094390} & \cellcolor{gray!6}{0.95} & \cellcolor{gray!6}{All skies} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{90.4}\\
-\cmidrule{1-8}
-0.4300 & 6.67e-03 & 0.2350 & 0.4844650 & 0.95 & All skies & Autumn & 99.3\\
-\cmidrule{1-8}
-\cellcolor{gray!6}{0.5050} & \cellcolor{gray!6}{4.16e-04} & \cellcolor{gray!6}{0.3640} & \cellcolor{gray!6}{0.6033797} & \cellcolor{gray!6}{0.95} & \cellcolor{gray!6}{Clear skies} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{100.0}\\
-\cmidrule{1-8}
--0.1360 & 1.71e-01 & 0.0635 & -0.2520277 & 0.95 & Clear skies & Spring & 82.9\\
-\cmidrule{1-8}
-\cellcolor{gray!6}{-0.2310} & \cellcolor{gray!6}{3.01e-05} & \cellcolor{gray!6}{0.4690} & \cellcolor{gray!6}{-0.6846238} & \cellcolor{gray!6}{0.95} & \cellcolor{gray!6}{Clear skies} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{100.0}\\
-\cmidrule{1-8}
-0.0432 & 5.79e-01 & 0.0111 & 0.1054619 & 0.95 & Clear skies & Autumn & 42.1\\
-\cmidrule{1-8}
-\cellcolor{gray!6}{0.4720} & \cellcolor{gray!6}{2.01e-02} & \cellcolor{gray!6}{0.1780} & \cellcolor{gray!6}{0.4221299} & \cellcolor{gray!6}{0.95} & \cellcolor{gray!6}{Cloudy skies} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{98.0}\\
-\cmidrule{1-8}
-0.1780 & 1.25e-01 & 0.0792 & 0.2814813 & 0.95 & Cloudy skies & Spring & 87.5\\
-\cmidrule{1-8}
-\cellcolor{gray!6}{0.2130} & \cellcolor{gray!6}{2.20e-02} & \cellcolor{gray!6}{0.1740} & \cellcolor{gray!6}{0.4167034} & \cellcolor{gray!6}{0.95} & \cellcolor{gray!6}{Cloudy skies} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{97.8}\\
-\cmidrule{1-8}
-0.4140 & 2.46e-02 & 0.1680 & 0.4094891 & 0.95 & Cloudy skies & Autumn & 97.5\\
+\cellcolor{gray!6}{0.6870} & \cellcolor{gray!6}{1.33e-03} & \cellcolor{gray!6}{0.3120} & \cellcolor{gray!6}{0.5587342} & \cellcolor{gray!6}{All skies} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{99.9}\\
+\cmidrule{1-7}
+0.1460 & 1.78e-01 & 0.0617 & 0.2483832 & All skies & Spring & 82.2\\
+\cmidrule{1-7}
+\cellcolor{gray!6}{0.1230} & \cellcolor{gray!6}{9.61e-02} & \cellcolor{gray!6}{0.0958} & \cellcolor{gray!6}{0.3094390} & \cellcolor{gray!6}{All skies} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{90.4}\\
+\cmidrule{1-7}
+0.4300 & 6.67e-03 & 0.2350 & 0.4844650 & All skies & Autumn & 99.3\\
+\cmidrule{1-7}
+\cellcolor{gray!6}{0.5050} & \cellcolor{gray!6}{4.16e-04} & \cellcolor{gray!6}{0.3640} & \cellcolor{gray!6}{0.6033797} & \cellcolor{gray!6}{Clear skies} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{100.0}\\
+\cmidrule{1-7}
+-0.1360 & 1.71e-01 & 0.0635 & -0.2520277 & Clear skies & Spring & 82.9\\
+\cmidrule{1-7}
+\cellcolor{gray!6}{-0.2310} & \cellcolor{gray!6}{3.01e-05} & \cellcolor{gray!6}{0.4690} & \cellcolor{gray!6}{-0.6846238} & \cellcolor{gray!6}{Clear skies} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{100.0}\\
+\cmidrule{1-7}
+0.0432 & 5.79e-01 & 0.0111 & 0.1054619 & Clear skies & Autumn & 42.1\\
+\cmidrule{1-7}
+\cellcolor{gray!6}{0.4720} & \cellcolor{gray!6}{2.01e-02} & \cellcolor{gray!6}{0.1780} & \cellcolor{gray!6}{0.4221299} & \cellcolor{gray!6}{Cloudy skies} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{98.0}\\
+\cmidrule{1-7}
+0.1780 & 1.25e-01 & 0.0792 & 0.2814813 & Cloudy skies & Spring & 87.5\\
+\cmidrule{1-7}
+\cellcolor{gray!6}{0.2130} & \cellcolor{gray!6}{2.20e-02} & \cellcolor{gray!6}{0.1740} & \cellcolor{gray!6}{0.4167034} & \cellcolor{gray!6}{Cloudy skies} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{97.8}\\
+\cmidrule{1-7}
+0.4140 & 2.46e-02 & 0.1680 & 0.4094891 & Cloudy skies & Autumn & 97.5\\
 \bottomrule

### 2023-12-07 MDPI_submission.Rmd
Commit 2023-12-07 19:09

@@
 # TSI$cor.p              <- NULL
-TSI$conf.level         <- NULL
+TSI$cor.conf_level      <- NULL
 care <- grep("^cor\\.", names(TSI), value = T)

### 2023-12-07 DHI_GHI_00_raw_data.R

@@ -315,7 +315,7 @@ if (havetorun) {
     #
 
     warning("REMOVING CLOUD ENHANCEMENT")
-    DATA <- DATA[wattGLB < cosde(SZA) * TSIextEARTH_comb * 0.8 + 20]
+    DATA <- DATA[wattGLB < cosde(SZA) * TSIextEARTH_comb * 0.8 + 30]
 
 
     ##_ Select only CM-21 flags for trends -------------------------------------

### 2023-12-07 DHI_GHI_1_longterm_trends.R

@@ -389,12 +389,18 @@ for (DBn in dbs) {
 
             par("mar" = c(3, 4, 2, 1))
 
+            if (grepl("wattGLB", yvar)) {
+                acol <- "green"
+            } else {
+                acol <- get(paste0(c("col", unlist(strsplit(yvar, split = "_"))[1:2]),
+                                   collapse = "_"))
+            }
+
+
             ## plot data
             plot(dataset$Date, dataset[[avar]],
-                 pch  = ".",
-                 col  = get(paste0(c("col",
-                                     unlist(strsplit(avar, split = "_"))[1:2]),
-                                   collapse = "_")),
+                 pch      = ".",
+                 col      = acol,
                  cex      = 2,
                  main     = paste(translate(DBn), translate(avar)),
                  cex.main = 0.8,

### 2023-12-08 MDPI_submission.tex
Commit 2023-12-08 11:08

@@
 
-\begin{verbatim}
-Warning in set(x, j = name, value = value): Column 'conf.level' does not exist
-to remove
-\end{verbatim}
-
 For completeness, we will provide below a brief overview of the clear
@@
 (Figures~\ref{fig:trendCLEAR} and~ \ref{fig:trendCLOUD}). We observe a
-positive trend for all-sky conditions (\(0.38\,\%/y\)), a very close but
-smaller trend for clear-skies (\(0.024\,\%/y\)) and a negative weaker
-trend for cloudy-skies (\(0.34\,\%/y\)). In the studied period, there is
+positive trend for all-sky conditions (\(0.33\,\%/y\)), a very close but
+smaller trend for clear-skies (\(0.097\,\%/y\)) and a negative weaker
+trend for cloudy-skies (\(0.22\,\%/y\)). In the studied period, there is
 no significant break or change in the variability of the time series.
@@
 \midrule
-All skies & 0.3765 & 0.000 & 0.0913 & 10251\\
-Clear skies & 0.0244 & 0.179 & 0.0176 & 5804\\
-Cloudy skies & 0.3416 & 0.000 & 0.0725 & 8113\\
+All skies & 0.326 & 0.00000 & 0.0776 & 10251\\
+Clear skies & 0.097 & 0.00000 & 0.1381 & 2689\\
+Cloudy skies & 0.221 & 0.00261 & 0.0439 & 4707\\
 \bottomrule
@@
 all-sky conditions the trend in SDR in winter is the largest
-(\(0.69\,\%/y\)), followed by the trend in autumn (\(0.43\,\%/y\), a
+(\(0.66\,\%/y\)), followed by the trend in autumn (\(0.4\,\%/y\), a
 value close to the long term trend) both statistically significant above
@@
 of clouds during winter and autumn. For clear-skies, the trend in winter
-is \(0.5\,\%/y\), larger than for all-skies (\(0.69\,\%/y\)), which is
+is \(0.36\,\%/y\), larger than for all-skies (\(0.66\,\%/y\)), which is
 another indication of a decreasing trend in cloud optical thickness.
@@
 \midrule
-\cellcolor{gray!6}{0.6870} & \cellcolor{gray!6}{1.33e-03} & \cellcolor{gray!6}{0.3120} & \cellcolor{gray!6}{0.5587342} & \cellcolor{gray!6}{All skies} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{99.9}\\
+\cellcolor{gray!6}{0.6640} & \cellcolor{gray!6}{1.95e-03} & \cellcolor{gray!6}{0.29400} & \cellcolor{gray!6}{0.5425947} & \cellcolor{gray!6}{All skies} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{99.8}\\
 \cmidrule{1-7}
-0.1460 & 1.78e-01 & 0.0617 & 0.2483832 & All skies & Spring & 82.2\\
+0.0691 & 5.25e-01 & 0.01410 & 0.1186549 & All skies & Spring & 47.5\\
 \cmidrule{1-7}
-\cellcolor{gray!6}{0.1230} & \cellcolor{gray!6}{9.61e-02} & \cellcolor{gray!6}{0.0958} & \cellcolor{gray!6}{0.3094390} & \cellcolor{gray!6}{All skies} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{90.4}\\
+\cellcolor{gray!6}{0.0492} & \cellcolor{gray!6}{5.41e-01} & \cellcolor{gray!6}{0.01350} & \cellcolor{gray!6}{0.1162612} & \cellcolor{gray!6}{All skies} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{45.9}\\
 \cmidrule{1-7}
-0.4300 & 6.67e-03 & 0.2350 & 0.4844650 & All skies & Autumn & 99.3\\
+0.4020 & 1.31e-02 & 0.20100 & 0.4479191 & All skies & Autumn & 98.7\\
 \cmidrule{1-7}
-\cellcolor{gray!6}{0.5050} & \cellcolor{gray!6}{4.16e-04} & \cellcolor{gray!6}{0.3640} & \cellcolor{gray!6}{0.6033797} & \cellcolor{gray!6}{Clear skies} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{100.0}\\
+\cellcolor{gray!6}{0.3570} & \cellcolor{gray!6}{7.26e-03} & \cellcolor{gray!6}{0.23000} & \cellcolor{gray!6}{0.4800278} & \cellcolor{gray!6}{Clear skies} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{99.3}\\
 \cmidrule{1-7}
--0.1360 & 1.71e-01 & 0.0635 & -0.2520277 & Clear skies & Spring & 82.9\\
+0.0318 & 7.21e-01 & 0.00448 & 0.0669349 & Clear skies & Spring & 27.9\\
 \cmidrule{1-7}
-\cellcolor{gray!6}{-0.2310} & \cellcolor{gray!6}{3.01e-05} & \cellcolor{gray!6}{0.4690} & \cellcolor{gray!6}{-0.6846238} & \cellcolor{gray!6}{Clear skies} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{100.0}\\
+\cellcolor{gray!6}{-0.0933} & \cellcolor{gray!6}{7.66e-03} & \cellcolor{gray!6}{0.22800} & \cellcolor{gray!6}{-0.4772665} & \cellcolor{gray!6}{Clear skies} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{99.2}\\
 \cmidrule{1-7}
-0.0432 & 5.79e-01 & 0.0111 & 0.1054619 & Clear skies & Autumn & 42.1\\
+0.1100 & 3.08e-01 & 0.03710 & 0.1926085 & Clear skies & Autumn & 69.2\\
 \cmidrule{1-7}
-\cellcolor{gray!6}{0.4720} & \cellcolor{gray!6}{2.01e-02} & \cellcolor{gray!6}{0.1780} & \cellcolor{gray!6}{0.4221299} & \cellcolor{gray!6}{Cloudy skies} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{98.0}\\
+\cellcolor{gray!6}{0.7800} & \cellcolor{gray!6}{3.61e-04} & \cellcolor{gray!6}{0.37000} & \cellcolor{gray!6}{0.6083684} & \cellcolor{gray!6}{Cloudy skies} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{100.0}\\
 \cmidrule{1-7}
-0.1780 & 1.25e-01 & 0.0792 & 0.2814813 & Cloudy skies & Spring & 87.5\\
+-0.1870 & 1.60e-01 & 0.06700 & -0.2588055 & Cloudy skies & Spring & 84.0\\
 \cmidrule{1-7}
-\cellcolor{gray!6}{0.2130} & \cellcolor{gray!6}{2.20e-02} & \cellcolor{gray!6}{0.1740} & \cellcolor{gray!6}{0.4167034} & \cellcolor{gray!6}{Cloudy skies} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{97.8}\\
+\cellcolor{gray!6}{-0.6630} & \cellcolor{gray!6}{8.91e-05} & \cellcolor{gray!6}{0.42800} & \cellcolor{gray!6}{-0.6538402} & \cellcolor{gray!6}{Cloudy skies} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{100.0}\\
 \cmidrule{1-7}
-0.4140 & 2.46e-02 & 0.1680 & 0.4094891 & Cloudy skies & Autumn & 97.5\\
+0.0985 & 7.20e-01 & 0.00465 & 0.0681684 & Cloudy skies & Autumn & 28.0\\
 \bottomrule
@@
 We have demonstrated that in the period 1993 - 2023, there is a positive
-trend in SDR of (\(0.38\,\%/y\)) (brightening) (positive trend) in
+trend in SDR of (\(0.33\,\%/y\)) (brightening) (positive trend) in
 Thessaloniki, Greece, under all-sky conditions. A previous study
@@
 Europe have been proposed as major factor by \citet{Wild2021}. The
-dimming effect on SDR under cloudy-sky conditions (\(0.34\,\%/y\)),
+dimming effect on SDR under cloudy-sky conditions (\(0.22\,\%/y\)),
 suggests that cloud optical thickness is decreasing during this period.
@@
 on season, with winter showing the strongest statistically significant
-trend of \(0.69\) and \(0.5\,\%/y\) for all- and clear-skies,
+trend of \(0.66\) and \(0.36\,\%/y\) for all- and clear-skies,
 respectively, in contrast to spring and summer. The trends for autumn
-are also significant but smaller ( \(0.43\) and \(0.043\,\%/y\) for all-
+are also significant but smaller ( \(0.4\) and \(0.11\,\%/y\) for all-
 and clear-skies, respectively). Our findings are in agreement with other

### 2023-12-08 DHI_GHI_0_variables.R
Commit 2023-12-08 15:09

@@ -94,8 +94,8 @@ Monthly_aggegation_N_lim  <-  20
 SZA_aggregation_N_lim     <-   4
 
 All_daily_ratio_lim       <- 0.5
-Clear_daily_ratio_lim     <- 0.2 ## keep both the same!!
-Cloud_daily_ratio_lim     <- 0.2
+Clear_daily_ratio_lim     <- 0.6 ## keep both the same!!
+Cloud_daily_ratio_lim     <- 0.6
 
 FIBais_Az_1               <-  58
 FIBais_Az_2               <- 120

### 2023-12-09 DHI_GHI_01_Input_longterm.R
Commit 2023-12-09 21:10

@@ -52,6 +52,12 @@ DATA_all  [, TYPE := NULL]
 DATA_Clear[, TYPE := NULL]
 DATA_Cloud[, TYPE := NULL]
 
+
+warning("REMOVING CLOUD ENHANCEMENT")
+DATA_Cloud <- DATA_Cloud[wattGLB < cosde(SZA) * TSIextEARTH_comb * 0.8 + 30 ]
+
+
+
 DATA_all[,.N]
 DATA_Clear[,.N]
 DATA_Cloud[,.N]

### 2023-12-09 DHI_GHI_03_Input_consistency.R

@@ -53,6 +53,10 @@ DATA_Clear[, TYPE := NULL]
 DATA_Cloud[, TYPE := NULL]
 
 
+warning("REMOVING CLOUD ENHANCEMENT")
+DATA_Cloud <- DATA_Cloud[wattGLB < cosde(SZA) * TSIextEARTH_comb * 0.8 + 30 ]
+
+
 
 
 # ......................................................................... ----

### 2023-12-09 MDPI_submission.tex
Commit 2023-12-09 22:10

--- 2023-12-07 MDPI_submission.tex
@@
 initial daily values of this dataset were interpolated to match the time
-step of our measurements. In the final dataset \(23.7\%\) of the
-1-minute data were identified as under clear-sky conditions and \(41\%\)
-as under cloud-sky conditions.
+step of our measurements. In the final dataset \(23.3\%\) of the
+1-minute data were identified as under clear-sky conditions and
+\(40.7\%\) as under cloud-sky conditions.
 
@@
 positive trend for all-sky conditions (\(0.38\,\%/y\)), a very close but
-smaller trend for clear-skies (\(0.024\,\%/y\)) and a negative weaker
-trend for cloudy-skies (\(0.34\,\%/y\)). In the studied period, there is
+smaller trend for clear-skies (\(0.097\,\%/y\)) and a negative weaker
+trend for cloudy-skies (\(0.25\,\%/y\)). In the studied period, there is
 no significant break or change in the variability of the time series.
@@
 \midrule
-All skies & 0.3765 & 0.000 & 0.0913 & 10251\\
-Clear skies & 0.0244 & 0.179 & 0.0176 & 5804\\
-Cloudy skies & 0.3416 & 0.000 & 0.0725 & 8113\\
+All skies & 0.377 & 0.000000 & 0.0913 & 10251\\
+Clear skies & 0.097 & 0.000000 & 0.1381 & 2689\\
+Cloudy skies & 0.250 & 0.000655 & 0.0494 & 4747\\
 \bottomrule
@@
 \midrule
-\cellcolor{gray!6}{0.6870} & \cellcolor{gray!6}{1.33e-03} & \cellcolor{gray!6}{0.3120} & \cellcolor{gray!6}{0.5587342} & \cellcolor{gray!6}{All skies} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{99.9}\\
+\cellcolor{gray!6}{0.6870} & \cellcolor{gray!6}{0.001330} & \cellcolor{gray!6}{0.31200} & \cellcolor{gray!6}{0.5587342} & \cellcolor{gray!6}{All skies} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{99.9}\\
 \cmidrule{1-7}
-0.1460 & 1.78e-01 & 0.0617 & 0.2483832 & All skies & Spring & 82.2\\
+0.1460 & 0.178000 & 0.06170 & 0.2483832 & All skies & Spring & 82.2\\
 \cmidrule{1-7}
-\cellcolor{gray!6}{0.1230} & \cellcolor{gray!6}{9.61e-02} & \cellcolor{gray!6}{0.0958} & \cellcolor{gray!6}{0.3094390} & \cellcolor{gray!6}{All skies} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{90.4}\\
+\cellcolor{gray!6}{0.1230} & \cellcolor{gray!6}{0.096100} & \cellcolor{gray!6}{0.09580} & \cellcolor{gray!6}{0.3094390} & \cellcolor{gray!6}{All skies} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{90.4}\\
 \cmidrule{1-7}
-0.4300 & 6.67e-03 & 0.2350 & 0.4844650 & All skies & Autumn & 99.3\\
+0.4300 & 0.006670 & 0.23500 & 0.4844650 & All skies & Autumn & 99.3\\
 \cmidrule{1-7}
-\cellcolor{gray!6}{0.5050} & \cellcolor{gray!6}{4.16e-04} & \cellcolor{gray!6}{0.3640} & \cellcolor{gray!6}{0.6033797} & \cellcolor{gray!6}{Clear skies} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{100.0}\\
+\cellcolor{gray!6}{0.3570} & \cellcolor{gray!6}{0.007260} & \cellcolor{gray!6}{0.23000} & \cellcolor{gray!6}{0.4800278} & \cellcolor{gray!6}{Clear skies} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{99.3}\\
 \cmidrule{1-7}
--0.1360 & 1.71e-01 & 0.0635 & -0.2520277 & Clear skies & Spring & 82.9\\
+0.0317 & 0.721000 & 0.00445 & 0.0667338 & Clear skies & Spring & 27.9\\
 \cmidrule{1-7}
-\cellcolor{gray!6}{-0.2310} & \cellcolor{gray!6}{3.01e-05} & \cellcolor{gray!6}{0.4690} & \cellcolor{gray!6}{-0.6846238} & \cellcolor{gray!6}{Clear skies} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{100.0}\\
+\cellcolor{gray!6}{-0.0931} & \cellcolor{gray!6}{0.007770} & \cellcolor{gray!6}{0.22700} & \cellcolor{gray!6}{-0.4764646} & \cellcolor{gray!6}{Clear skies} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{99.2}\\
 \cmidrule{1-7}
-0.0432 & 5.79e-01 & 0.0111 & 0.1054619 & Clear skies & Autumn & 42.1\\
+0.1100 & 0.308000 & 0.03710 & 0.1926085 & Clear skies & Autumn & 69.2\\
 \cmidrule{1-7}
-\cellcolor{gray!6}{0.4720} & \cellcolor{gray!6}{2.01e-02} & \cellcolor{gray!6}{0.1780} & \cellcolor{gray!6}{0.4221299} & \cellcolor{gray!6}{Cloudy skies} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{98.0}\\
+\cellcolor{gray!6}{0.7890} & \cellcolor{gray!6}{0.000262} & \cellcolor{gray!6}{0.38400} & \cellcolor{gray!6}{0.6194510} & \cellcolor{gray!6}{Cloudy skies} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{100.0}\\
 \cmidrule{1-7}
-0.1780 & 1.25e-01 & 0.0792 & 0.2814813 & Cloudy skies & Spring & 87.5\\
+-0.1460 & 0.261000 & 0.04330 & -0.2081141 & Cloudy skies & Spring & 73.9\\
 \cmidrule{1-7}
-\cellcolor{gray!6}{0.2130} & \cellcolor{gray!6}{2.20e-02} & \cellcolor{gray!6}{0.1740} & \cellcolor{gray!6}{0.4167034} & \cellcolor{gray!6}{Cloudy skies} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{97.8}\\
+\cellcolor{gray!6}{-0.5680} & \cellcolor{gray!6}{0.000483} & \cellcolor{gray!6}{0.35800} & \cellcolor{gray!6}{-0.5979999} & \cellcolor{gray!6}{Cloudy skies} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{100.0}\\
 \cmidrule{1-7}
-0.4140 & 2.46e-02 & 0.1680 & 0.4094891 & Cloudy skies & Autumn & 97.5\\
+0.1070 & 0.701000 & 0.00536 & 0.0732067 & Cloudy skies & Autumn & 29.9\\
 \bottomrule
@@
 Europe have been proposed as major factor by \citet{Wild2021}. The
-dimming effect on SDR under cloudy-sky conditions (\(0.34\,\%/y\)),
+dimming effect on SDR under cloudy-sky conditions (\(0.25\,\%/y\)),
 suggests that cloud optical thickness is decreasing during this period.

### 2023-12-10 MDPI_submission.tex
Commit 2023-12-10 07:10

@@
 step of our measurements. In the final dataset \(23.3\%\) of the
-1-minute data were identified as under clear-sky conditions and
-\(40.7\%\) as under cloud-sky conditions.
+1-minute data were identified as under clear-sky conditions and \(43\%\)
+as under cloud-sky conditions.
 
@@
 smaller trend for clear-skies (\(0.097\,\%/y\)) and a negative weaker
-trend for cloudy-skies (\(0.25\,\%/y\)). In the studied period, there is
+trend for cloudy-skies (\(0.41\,\%/y\)). In the studied period, there is
 no significant break or change in the variability of the time series.
@@
 \midrule
-All skies & 0.377 & 0.000000 & 0.0913 & 10251\\
-Clear skies & 0.097 & 0.000000 & 0.1381 & 2689\\
-Cloudy skies & 0.250 & 0.000655 & 0.0494 & 4747\\
+All skies & 0.377 & 0 & 0.0913 & 10251\\
+Clear skies & 0.097 & 0 & 0.1381 & 2689\\
+Cloudy skies & 0.408 & 0 & 0.0810 & 4937\\
 \bottomrule
@@
 \cmidrule{1-7}
-\cellcolor{gray!6}{0.7890} & \cellcolor{gray!6}{0.000262} & \cellcolor{gray!6}{0.38400} & \cellcolor{gray!6}{0.6194510} & \cellcolor{gray!6}{Cloudy skies} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{100.0}\\
+\cellcolor{gray!6}{0.8250} & \cellcolor{gray!6}{0.000123} & \cellcolor{gray!6}{0.41500} & \cellcolor{gray!6}{0.6439160} & \cellcolor{gray!6}{Cloudy skies} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{100.0}\\
 \cmidrule{1-7}
--0.1460 & 0.261000 & 0.04330 & -0.2081141 & Cloudy skies & Spring & 73.9\\
+0.0942 & 0.425000 & 0.02210 & 0.1485860 & Cloudy skies & Spring & 57.5\\
 \cmidrule{1-7}
-\cellcolor{gray!6}{-0.5680} & \cellcolor{gray!6}{0.000483} & \cellcolor{gray!6}{0.35800} & \cellcolor{gray!6}{-0.5979999} & \cellcolor{gray!6}{Cloudy skies} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{100.0}\\
+\cellcolor{gray!6}{-0.0983} & \cellcolor{gray!6}{0.414000} & \cellcolor{gray!6}{0.02400} & \cellcolor{gray!6}{-0.1548507} & \cellcolor{gray!6}{Cloudy skies} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{58.6}\\
 \cmidrule{1-7}
-0.1070 & 0.701000 & 0.00536 & 0.0732067 & Cloudy skies & Autumn & 29.9\\
+0.1950 & 0.460000 & 0.01960 & 0.1401398 & Cloudy skies & Autumn & 54.0\\
 \bottomrule
@@
 Europe have been proposed as major factor by \citet{Wild2021}. The
-dimming effect on SDR under cloudy-sky conditions (\(0.25\,\%/y\)),
+dimming effect on SDR under cloudy-sky conditions (\(0.41\,\%/y\)),
 suggests that cloud optical thickness is decreasing during this period.

### 2023-12-10 DHI_GHI_01_Input_longterm.R

@@ -52,9 +52,9 @@ DATA_all  [, TYPE := NULL]
 DATA_Clear[, TYPE := NULL]
 DATA_Cloud[, TYPE := NULL]
 
-
-warning("REMOVING CLOUD ENHANCEMENT")
-DATA_Cloud <- DATA_Cloud[wattGLB < cosde(SZA) * TSIextEARTH_comb * 0.8 + 30 ]
+#
+# warning("REMOVING CLOUD ENHANCEMENT")
+# DATA_Cloud <- DATA_Cloud[wattGLB < cosde(SZA) * TSIextEARTH_comb * 0.8 + 30 ]
 
 
 

### 2023-12-10 DHI_GHI_02_Input_szatrends.R

@@ -49,9 +49,9 @@ DATA_all  [, TYPE := NULL]
 DATA_Clear[, TYPE := NULL]
 DATA_Cloud[, TYPE := NULL]
 
-
-warning("REMOVING CLOUD ENHANCEMENT")
-DATA_Cloud <- DATA_Cloud[wattGLB < cosde(SZA) * TSIextEARTH_comb * 0.8 + 30 ]
+#
+# warning("REMOVING CLOUD ENHANCEMENT")
+# DATA_Cloud <- DATA_Cloud[wattGLB < cosde(SZA) * TSIextEARTH_comb * 0.8 + 30 ]
 
 
 

### 2023-12-12 Build_165/MDPI_submission_B165.Rmd
Commit 2023-12-13 00:18

--- 2023-12-07 MDPI_submission.Rmd
@@
 spp$cor.t          <- NULL
+spp$slope.stat_sig <- NULL
 

### 2023-12-12 Build_167/MDPI_submission_B167.tex
Commit 2023-12-13 01:11

--- 2023-12-10 MDPI_submission.tex
@@
 \midrule
-All skies & 0.377 & 0 & 0.0913 & 10251\\
+All skies & 0.376 & 0 & 0.0911 & 10250\\
 Clear skies & 0.097 & 0 & 0.1381 & 2689\\
@@
 \caption{\label{tab:trendseasontable}SDR linear trends of monthly anomalies for each season of the year.}
-\begin{tabu} to \linewidth {>{\centering\arraybackslash}p{8em}>{\centering}X>{\raggedleft}X>{\raggedleft}X>{\centering}X>{\centering}X>{\raggedleft}X}
+\begin{tabu} to \linewidth {>{\centering\arraybackslash}p{8em}>{\centering}X>{\raggedleft}X>{\raggedleft}X>{\centering}X>{\centering}X>{\raggedleft}X>{\raggedleft}X>{\centering}X>{\centering}X>{\raggedleft}X}
 \toprule
-Trend [\%/year] & slope.p & Rsqrd & cor.estimate & Sky condition & Season & slope.stat\_sig\\
+Trend [\%/year] & slope.p & Rsqrd & cor.estimate & Sky condition & Season & N\_eff & t\_eff & t\_eff\_cri & conf\_2.5 & conf\_97.5\\
 \midrule
-\cellcolor{gray!6}{0.6870} & \cellcolor{gray!6}{0.001330} & \cellcolor{gray!6}{0.31200} & \cellcolor{gray!6}{0.5587342} & \cellcolor{gray!6}{All skies} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{99.9}\\
-\cmidrule{1-7}
-0.1460 & 0.178000 & 0.06170 & 0.2483832 & All skies & Spring & 82.2\\
-\cmidrule{1-7}
-\cellcolor{gray!6}{0.1230} & \cellcolor{gray!6}{0.096100} & \cellcolor{gray!6}{0.09580} & \cellcolor{gray!6}{0.3094390} & \cellcolor{gray!6}{All skies} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{90.4}\\
-\cmidrule{1-7}
-0.4300 & 0.006670 & 0.23500 & 0.4844650 & All skies & Autumn & 99.3\\
-\cmidrule{1-7}
-\cellcolor{gray!6}{0.3570} & \cellcolor{gray!6}{0.007260} & \cellcolor{gray!6}{0.23000} & \cellcolor{gray!6}{0.4800278} & \cellcolor{gray!6}{Clear skies} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{99.3}\\
-\cmidrule{1-7}
-0.0317 & 0.721000 & 0.00445 & 0.0667338 & Clear skies & Spring & 27.9\\
-\cmidrule{1-7}
-\cellcolor{gray!6}{-0.0931} & \cellcolor{gray!6}{0.007770} & \cellcolor{gray!6}{0.22700} & \cellcolor{gray!6}{-0.4764646} & \cellcolor{gray!6}{Clear skies} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{99.2}\\
-\cmidrule{1-7}
-0.1100 & 0.308000 & 0.03710 & 0.1926085 & Clear skies & Autumn & 69.2\\
-\cmidrule{1-7}
-\cellcolor{gray!6}{0.8250} & \cellcolor{gray!6}{0.000123} & \cellcolor{gray!6}{0.41500} & \cellcolor{gray!6}{0.6439160} & \cellcolor{gray!6}{Cloudy skies} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{100.0}\\
-\cmidrule{1-7}
-0.0942 & 0.425000 & 0.02210 & 0.1485860 & Cloudy skies & Spring & 57.5\\
-\cmidrule{1-7}
-\cellcolor{gray!6}{-0.0983} & \cellcolor{gray!6}{0.414000} & \cellcolor{gray!6}{0.02400} & \cellcolor{gray!6}{-0.1548507} & \cellcolor{gray!6}{Cloudy skies} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{58.6}\\
-\cmidrule{1-7}
-0.1950 & 0.460000 & 0.01960 & 0.1401398 & Cloudy skies & Autumn & 54.0\\
+\cellcolor{gray!6}{0.6870} & \cellcolor{gray!6}{0.001330} & \cellcolor{gray!6}{0.31200} & \cellcolor{gray!6}{0.5587342} & \cellcolor{gray!6}{All skies} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{24.08252} & \cellcolor{gray!6}{1e-07} & \cellcolor{gray!6}{2.063524} & \cellcolor{gray!6}{-0.0011} & \cellcolor{gray!6}{0.0015253}\\
+\cmidrule{1-11}
+0.1360 & 0.219000 & 0.05150 & 0.2270307 & All skies & Spring & 44.44554 & 1e-07 & 2.014797 & -0.0011 & 0.0015253\\
+\cmidrule{1-11}
+\cellcolor{gray!6}{0.1230} & \cellcolor{gray!6}{0.096100} & \cellcolor{gray!6}{0.09580} & \cellcolor{gray!6}{0.3094390} & \cellcolor{gray!6}{All skies} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{23.37502} & \cellcolor{gray!6}{0e+00} & \cellcolor{gray!6}{2.066823} & \cellcolor{gray!6}{-0.0011} & \cellcolor{gray!6}{0.0015253}\\
+\cmidrule{1-11}
+0.4300 & 0.006670 & 0.23500 & 0.4844650 & All skies & Autumn & 21.43236 & 1e-07 & 2.077062 & -0.0011 & 0.0015253\\
+\cmidrule{1-11}
+\cellcolor{gray!6}{0.3570} & \cellcolor{gray!6}{0.007260} & \cellcolor{gray!6}{0.23000} & \cellcolor{gray!6}{0.4800278} & \cellcolor{gray!6}{Clear skies} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{18.55573} & \cellcolor{gray!6}{0e+00} & \cellcolor{gray!6}{2.096421} & \cellcolor{gray!6}{-0.0011} & \cellcolor{gray!6}{0.0015253}\\
+\cmidrule{1-11}
+0.0317 & 0.721000 & 0.00445 & 0.0667338 & Clear skies & Spring & 21.33345 & 0e+00 & 2.077636 & -0.0011 & 0.0015253\\
+\cmidrule{1-11}
+\cellcolor{gray!6}{-0.0931} & \cellcolor{gray!6}{0.007770} & \cellcolor{gray!6}{0.22700} & \cellcolor{gray!6}{-0.4764646} & \cellcolor{gray!6}{Clear skies} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{25.12298} & \cellcolor{gray!6}{0e+00} & \cellcolor{gray!6}{2.059027} & \cellcolor{gray!6}{-0.0011} & \cellcolor{gray!6}{0.0015253}\\
+\cmidrule{1-11}
+0.1100 & 0.308000 & 0.03710 & 0.1926085 & Clear skies & Autumn & 23.96832 & 0e+00 & 2.064043 & -0.0011 & 0.0015253\\
+\cmidrule{1-11}
+\cellcolor{gray!6}{0.8250} & \cellcolor{gray!6}{0.000123} & \cellcolor{gray!6}{0.41500} & \cellcolor{gray!6}{0.6439160} & \cellcolor{gray!6}{Cloudy skies} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{14.68318} & \cellcolor{gray!6}{1e-07} & \cellcolor{gray!6}{2.135462} & \cellcolor{gray!6}{-0.0011} & \cellcolor{gray!6}{0.0015253}\\
+\cmidrule{1-11}
+0.0942 & 0.425000 & 0.02210 & 0.1485860 & Cloudy skies & Spring & 35.40650 & 1e-07 & 2.029275 & -0.0011 & 0.0015253\\
+\cmidrule{1-11}
+\cellcolor{gray!6}{-0.0983} & \cellcolor{gray!6}{0.414000} & \cellcolor{gray!6}{0.02400} & \cellcolor{gray!6}{-0.1548507} & \cellcolor{gray!6}{Cloudy skies} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{37.66778} & \cellcolor{gray!6}{1e-07} & \cellcolor{gray!6}{2.024981} & \cellcolor{gray!6}{-0.0011} & \cellcolor{gray!6}{0.0015253}\\
+\cmidrule{1-11}
+0.1950 & 0.460000 & 0.01960 & 0.1401398 & Cloudy skies & Autumn & 28.41812 & 2e-07 & 2.047050 & -0.0011 & 0.0015253\\
 \bottomrule

### 2023-12-12 appendix.tex
Commit 2023-11-08 21:10
Commit 2023-11-09 04:14
Commit 2023-11-09 05:11
Commit 2023-11-09 18:14
Commit 2023-12-13 01:11

@@
-%% optional
-\appendixtitles{no} %Leave argument "no" if all appendix headings stay EMPTY (then no dot is printed after "Appendix A"). If the appendix sections contain a heading then change the argument to "yes".
-\appendixsections{one} %Leave argument "multiple" if there are multiple sections.
+% % \renewcommand\thefigure{\thesection.\arabic{figure}} %% my
+% % \renewcommand\thefigure{\thesection.\arabic{figure}}
+%
+% %% From R rticle MDPI template
+% %% optional
+% \appendixtitles{no} %Leave argument "no" if all appendix headings stay EMPTY (then no dot is printed after "Appendix A").
+If the appendix sections contain a heading then change the argument to "yes".
+% \appendixsections{one} %Leave argument "multiple" if there are multiple sections.
 Then a counter is printed ("Appendix A").
 If there is only one appendix section then change the argument to "one" and no counter is printed ("Appendix").
+%
+% % \appendix
+% % \section{}
+%
+% \appendix
+% \setcounter{secnumdepth}{0}
+% \section{Appendix}
+%
+% \setcounter{figure}{0}    %% my
+
+%% From raw MDPI template
+%% Optional
+\appendixtitles{no} % Leave argument "no" if all appendix headings stay EMPTY (then no dot is printed after "Appendix A"). If the appendix sections contain a heading then change the argument to "yes".
+\appendixstart
 \appendix
-% \section{}
+\section[\appendixname~\thesection]{}
 
-\section[\appendixname~\thesection]{} % test
-
-\  \
-
-% copy output from Article.tex
+% copy output from Rmd
 
 \begin{figure}[h!]
+    {\centering
+        \includegraphics[width=0.7\linewidth]{./images/LongtermTrends-7}
 
-{\centering \includegraphics[width=0.7\linewidth]{./images/LongtermTrends-5}}
-
-\caption{Anomalies (\%) of the daily clear-sky SDR, relative to climatological values for1993 - 2023. The black line shows the long term linear trend for clear-sky conditions.}\label{fig:trendCLEAR}
-\end{figure}
-
-\  \
-
-\begin{figure}[h!]
-
-{\centering \includegraphics[width=0.7\linewidth]{./images/LongtermTrends-8}}
-
-\caption{Anomalies (\%) of the daily cloud-sky SDR, relative to climatological values for 1993 - 2023. The black line shows the long term linear trend for cloud-sky conditions.}\label{fig:trendCLOUD}
+    }
+    \caption{Anomalies (\%) of the daily clear-sky SDR, relative to climatological values for1993 - 2023.
+The black line shows the long term linear trend for clear-sky conditions.}\label{fig:trendCLEAR}
 \end{figure}
@@
 \begin{figure}[h!]
+    {\centering
+        \includegraphics[width=0.7\linewidth]{./images/LongtermTrends-10}
 
-{\centering \includegraphics[width=1\linewidth]{./images/SzaTrendsSeasTogether-2}}
-
-\caption{Long term trends of SDR as a function of SZA separately form morning and afternoon
-periods, by season (rows of plots) for the three sky conditions (columns of plots).
-Solid shapes represent statistical significant trends ($p<0.005$).
-Cases where $p<0.005$ or with less than $85$ observations may be missing from view.}\label{fig:SZAtrendSeason}
+    }
+    \caption{Anomalies (\%) of the daily cloud-sky SDR, relative to climatological values for 1993 - 2023.
+The black line shows the long term linear trend for cloud-sky conditions.}\label{fig:trendCLOUD}
 \end{figure}
 
-\ \
+\begin{figure}[h!]
+    \begin{adjustwidth}{-\extralength}{0cm}
+        {\centering
+            \includegraphics[width=1.0\linewidth]{./images/SzaTrendsSeasTogether-2}
+        }
+        \caption{Long term trends of SDR as a function of SZA separately form morning
+               and afternoon periods, by season (rows of plots) for the three sky
+               conditions (columns of plots).
+               Solid shapes represent statistical significant trends ($p<0.005$).
+               Cases where $p<0.005$ or with less than $85$ observations may be
+               missing from view.}\label{fig:SZAtrendSeason}
+    \end{adjustwidth}
+\end{figure}
+
+\FloatBarrier
 

### 2023-12-13 MDPI_submission.Rmd
Commit 2023-12-13 14:14

@@
 
-setorder(TSI, -slope.stat_sig)
-rownames(TSI) <- NULL
-names(TSI) <- c("Trend [%/year]", "Trend p-value", "Sky condition", "Trend statistical signif. [%]")
-TSIslopeAll <- TSI[`Sky condition` == "All sky cond."]
+# setorder(TSI, -slope.stat_sig)
+# rownames(TSI) <- NULL
+# names(TSI) <- c("Trend [%/year]", "Trend p-value", "Sky condition", "Trend statistical signif. [%]")
+# 
+# TSIslopeAll <- TSI[`Sky condition` == "All sky cond."]
+TSIslopeAll <- TSI[DATA == "All sky cond."]
 ```
@@
 
-# 
-# names(LT)    <- c("Trend [%/year]",
-#                   "Trend p-value",
-#                   "Sky conditions",
-#                   "Days with data",
-#                   "Statistical signif. [%]")
-
-## no need all are 100%
+LT$slope.stat_sig  <- NULL
+LT$`Trend p-value` <- NULL
+LT$N_eff           <- NULL
 
@@
 ## order columns
-LT1 <- LT[,c(
-      "Sky conditions",
-      "Trend [%/year]",
-      "Trend p-value",
-      "Pearson cor.",
-      "Days"
-      )]
+# LT1 <- LT[,c(
+#       "Sky conditions",
+#       "Trend [%/year]",
+#       "Trend p-value",
+#       "Pearson cor.",
+#       "Days"
+#       )]
+LT1 <- LT
 

### 2023-12-13 MDPI_submission.tex
Commit 2023-12-13 14:14

@@
 \caption{\label{tab:trendtable}Trends in SDR daily means for different sky conditions for the period 1993 - 2023.}
-\begin{tabu} to \linewidth {>{\centering\arraybackslash}p{8em}>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X}
+\begin{tabu} to \linewidth {>{\centering\arraybackslash}p{8em}>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X>{\centering}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X>{\centering}X}
 \toprule
-Sky conditions & Trend [\%/year] & Trend p-value & Pearson cor. & Days\\
+Trend [\%/year] & Rsqrd & Pearson cor. & Sky conditions & Days & t\_eff & t\_eff\_cri & conf\_2.5 & conf\_97.5 & mean\_clima & ChangeWpY\\
 \midrule
-All skies & 0.377 & 0 & 0.0913 & 10251\\
-Clear skies & 0.097 & 0 & 0.1381 & 2689\\
-Cloudy skies & 0.408 & 0 & 0.0810 & 4937\\
+0.376 & 0.00830 & 0.0911 & All skies & 10250 & 4.727 & 1.96 & 0.000811 & 0.001246 & 387 & 1.45\\
+0.097 & 0.01908 & 0.1381 & Clear skies & 2689 & 0.476 & 1.96 & 0.000194 & 0.000338 & 515 & 0.50\\
+0.408 & 0.00657 & 0.0810 & Cloudy skies & 4937 & 3.566 & 1.96 & 0.000733 & 0.001499 & 290 & 1.18\\
 \bottomrule

### 2023-12-15 MDPI_submission.tex
Commit 2023-12-15 13:08

@@
 \midrule
-0.376 & 0.00830 & 0.0911 & All skies & 10250 & 4.727 & 1.96 & 0.000811 & 0.001246 & 387 & 1.45\\
-0.097 & 0.01908 & 0.1381 & Clear skies & 2689 & 0.476 & 1.96 & 0.000194 & 0.000338 & 515 & 0.50\\
-0.408 & 0.00657 & 0.0810 & Cloudy skies & 4937 & 3.566 & 1.96 & 0.000733 & 0.001499 & 290 & 1.18\\
+0.3756 & 0.00830 & 0.0911 & All skies & 10250 & 4.727 & 1.96 & 0.000811 & 0.001246 & 387 & 1.452\\
+0.0972 & 0.01912 & 0.1383 & Clear skies & 2684 & 0.476 & 1.96 & 0.000194 & 0.000338 & 515 & 0.501\\
+0.4077 & 0.00657 & 0.0810 & Cloudy skies & 4937 & 3.566 & 1.96 & 0.000733 & 0.001499 & 290 & 1.181\\
 \bottomrule

### 2023-12-15 DHI_GHI_1_longterm_trends.R

@@ -151,6 +151,19 @@ LOESS_CRITERIO <-  c("aicc", "gcv")[1]
 
 
 
+ALL_1_daily_DESEAS[           GLB_att_des == 0]
+CLOUD_1_daily_DESEAS[         GLB_att_des == 0]
+# test <- CLEAR_1_daily_DESEAS[ GLB_att_des == 0]
+# test <- rm.cols.dups.DT(test)
+CLEAR_1_daily_DESEAS[         GLB_att_des == 0, GLB_att_des := NA]
+
+
+
+stopifnot(ALL_1_daily_DESEAS[   GLB_att_des == 0, .N ] == 0)
+stopifnot(CLOUD_1_daily_DESEAS[ GLB_att_des == 0, .N ] == 0)
+stopifnot(CLEAR_1_daily_DESEAS[ GLB_att_des == 0, .N ] == 0)
+
+
 
 #+ echo=F, include=T
 #'
@@ -287,31 +300,31 @@ for (DBn in dbs) {
 
             ## _ Arima tests  --------------------------------------------------
 
-            aamodelo <- auto.arima(dd[, avar])
-            summary(aamodelo)
-
-            amodelo  <- arima(dd[, avar], order = c(1,0,0), method = "ML")
-            summary(amodelo)
-            ddd <- summary(amodelo)
-
-            myforecast <- forecast(aamodelo,level = c(95),h=500)
-            plot(myforecast)
-            abline(amodelo, col = "red")
-
-            sarima(dd[, avar], 1,0,0, details = F, method = "ML", no.constant = T)
-            sarima(dd[, avar], 1,0,0, details = F, method = "ML")
-
-
-
-            armodel <- ar.ols(dd[, avar], na.action = na.exclude, order.max = 1, demean = F, intercept = T )
-            summary(armodel)
-            armodel
-
-            lmtest::coeftest(amodelo)
-            lmtest::coeftest(lm1) * Days_of_year
-
-
-stop("tess")
+#             aamodelo <- auto.arima(dd[, avar])
+#             summary(aamodelo)
+#
+#             amodelo  <- arima(dd[, avar], order = c(1,0,0), method = "ML")
+#             summary(amodelo)
+#             ddd <- summary(amodelo)
+#
+#             myforecast <- forecast(aamodelo,level = c(95),h=500)
+#             plot(myforecast)
+#             abline(amodelo, col = "red")
+#
[…]

### 2023-12-16 MDPI_submission.Rmd
Commit 2023-12-16 12:14

@@
 
-# Observational data and methodology.
+# Data and methodology
 
 The SDR data were measured with a Kipp & Zonen CM-21 pyranometer operating continuously at the Laboratory of Atmospheric Physics of the Aristotle University of Thessaloniki
-($40^\circ\,38'\,$N, $22^\circ\,57'\,$E, $80\,$m\ a.s.l.)
-in the period from
-<!-- \mbox{ -->
+($40^\circ\,38'\,$N, $22^\circ\,57'\,$E, $80\,$m\ a.s.l.) since 1993.
+Here we use data for the period from
 `r min(ALL_1_daily_DESEAS$Date)`
-<!-- } -->
 to
-<!-- \mbox{ -->
 `r max(ALL_1_daily_DESEAS$Date)`.
-<!-- }. -->
-The monitoring site is located near the city centre, and we expect to be affected by the urban environment.
-During the study period, the pyranometer has been independently calibrated three times at the Meteorologisches Observatorium Lindenberg, DWD, when it was verified the stability of the instrument to within better than $0.7\%$ relative to the initial calibration by the manufacturer.
-Along with SDR, the direct beam radiation (DNI) was also measured by a collocated Kipp & Zonen CHP-1 pyrheliometer, for the period
-<!-- \mbox{ -->
-`r min(ALL_1_daily_DESEAS[!is.na(DIR_att), Date])`
-<!-- } -->
-to
-<!-- \mbox{ -->
-`r max(ALL_1_daily_DESEAS[!is.na(DIR_att), Date])`.
-<!-- }. -->
-Although, we have performed a similar analysis to the DNI data, the results are not presented here, as they lack the appropriate statistical significance, due to the sorter duration of the data.
-However, the DNI data were used as auxiliary data, in the clear sky identification algorithm (CSid), which is discussed later, for the selection of the appropriate thresholds.
-It is noted that despite the capability of the CSid algorithm to use the DNI as a characterization parameter, we haven't utilized it here, to avoid any selection bias, due to unequal length of the two datasets.
-There are four distinct steps in the creation of the dataset analysed here:
+The monitoring site is located
+near the city centre thus we expect that measurements are affected by the urban environment.
+During the study period, the pyranometer has been
+independently calibrated three times at the Meteorologisches
+Observatorium Lindenberg, DWD, verifying that stability of the
+instrument's sensitivity is better than $0.7\%$ relative to the initial
+calibration by the manufacturer. 
+Along with SDR, the direct beam
+radiation (DNI) is also measured with a collocated Kipp & Zonen CHP-1 pyrheliometer since 2016-04-01. 
+The DNI data were used as auxiliary data
+to support the selection of appropriate thresholds in the clear sky
+identification algorithm (CSid), which is discussed later. It is noted
+that the limited dataset of DNI was not used for the identification of
+clear sky cases in the entire SDR series to avoid any selection bias due
+to unequal length of the two datasets.
+There are four distinct steps in the creation of the dataset analyzed here:
 a)\ the acquisition of radiation measurements from the sensors,
@@
 
-<!--
-The horizon is clear of obstructions except for the azimuth angles between 35 and 120 where obstacles of up to 10 elevation block the direct irradiance in the morning for SZA larger than 80.
-These data points were excluded from the analysis, as
--->
-
-<!-- Zerefos
-Northern Greece has been selected because it is well known to be located at the crossroads of aerosols and various pollutants pathways originating in central and Eastern
-Europe (Lelieveld et al., 2002; Zerefos et al., 2002). Thessa-
-loniki, the major city in northern Greece, lies along the exit of air
-pollutants from Europe, in the path of the prevailing long-range
-transport (LRT) trajectories, as we move from central and east-
-ern Europe to the Mediterranean (Zerefos et. al., 2002). On the
--->
-
 <!-- a) acquisition -->
-For the acquisition of radiometric data, the signal of the pyranometer is sampled with a rate of $1\,\text{Hz}$.
-The mean and the standard deviation of these samples are recorded every minute.
-The measurements are corrected for the zero offset ("dark signal" in volts).
-The "dark signal" is calculated by averaging all measurements recorded for a period of $3\,\text{h}$, before (morning) or after (evening) the Sun reaches an elevation angle of $-10^\circ$.
-The signal is converted to irradiance using a ramped value of the instrument’s sensitivity between calibrations.
+For the acquisition of radiometric data, the signal of the pyranometer is sampled at a rate of $1\,\text{Hz}$.
+The mean and the standard deviation of these samples are calculated and recorded every minute.
+The measurements are corrected for the zero offset ("dark signal" in volts), which is calculated by averaging all measurements recorded for a period of $3\,\text{h}$, before (morning) or after (evening) the Sun reaches an elevation angle of $-10^\circ$.
+The signal is converted to irradiance using a ramped value of the instrument’s sensitivity between  subsequent calibrations.
 
@@
 Thus, problematic recordings have been excluded from further processing.
-Although it is impossible to detect all false data, the large number of available data, and the aggregation scheme we used, ensures the good quality of the radiation measurements used in this study.
+Although it is impossible to detect all false data, the large number of available data, and the aggregation scheme we used, ensures the quality of the radiation measurements used in this study.
 
@@
 
-<!-- # Detection of clear periods in GHI measurements for SDR trends. -->
+## The clear sky identification algorithm
 

### 2023-12-16 MDPI_submission.tex
Commit 2023-12-16 12:14

@@
 
-\hypertarget{observational-data-and-methodology.}{%
-\section{Observational data and
-methodology.}\label{observational-data-and-methodology.}}
+\hypertarget{data-and-methodology}{%
+\section{Data and methodology}\label{data-and-methodology}}
 
@@
 Aristotle University of Thessaloniki (\(40^\circ\,38'\,\)N,
-\(22^\circ\,57'\,\)E, \(80\,\)m~a.s.l.) in the period from 1993-04-13 to
-2023-04-12. The monitoring site is located near the city centre, and we
-expect to be affected by the urban environment. During the study period,
-the pyranometer has been independently calibrated three times at the
-Meteorologisches Observatorium Lindenberg, DWD, when it was verified the
-stability of the instrument to within better than \(0.7\%\) relative to
-the initial calibration by the manufacturer. Along with SDR, the direct
-beam radiation (DNI) was also measured by a collocated Kipp \& Zonen
-CHP-1 pyrheliometer, for the period 2016-04-01 to 2023-04-12. Although,
-we have performed a similar analysis to the DNI data, the results are
-not presented here, as they lack the appropriate statistical
-significance, due to the sorter duration of the data. However, the DNI
-data were used as auxiliary data, in the clear sky identification
-algorithm (CSid), which is discussed later, for the selection of the
-appropriate thresholds. It is noted that despite the capability of the
-CSid algorithm to use the DNI as a characterization parameter, we
-haven't utilized it here, to avoid any selection bias, due to unequal
-length of the two datasets. There are four distinct steps in the
-creation of the dataset analysed here: a)~the acquisition of radiation
-measurements from the sensors, b)~the data quality check, c)~the
-identification of ``clear sky'' conditions from the radiometric data,
-and d)~the aggregation of data and trend analysis.
+\(22^\circ\,57'\,\)E, \(80\,\)m~a.s.l.) since 1993. Here we use data for
+the period from 1993-04-13 to 2023-04-12. The monitoring site is located
+near the city centre thus we expect that measurements are affected by
+the urban environment. During the study period, the pyranometer has been
+independently calibrated three times at the Meteorologisches
+Observatorium Lindenberg, DWD, verifying that stability of the
+instrument's sensitivity is better than \(0.7\%\) relative to the
+initial calibration by the manufacturer. Along with SDR, the direct beam
+radiation (DNI) is also measured with a collocated Kipp \& Zonen CHP-1
+pyrheliometer since 2016-04-01. The DNI data were used as auxiliary data
+to support the selection of appropriate thresholds in the clear sky
+identification algorithm (CSid), which is discussed later. It is noted
+that the limited dataset of DNI was not used for the identification of
+clear sky cases in the entire SDR series to avoid any selection bias due
+to unequal length of the two datasets. There are four distinct steps in
+the creation of the dataset analyzed here: a)~the acquisition of
+radiation measurements from the sensors, b)~the data quality check,
+c)~the identification of ``clear sky'' conditions from the radiometric
+data, and d)~the aggregation of data and trend analysis.
 
 For the acquisition of radiometric data, the signal of the pyranometer
-is sampled with a rate of \(1\,\text{Hz}\). The mean and the standard
-deviation of these samples are recorded every minute. The measurements
-are corrected for the zero offset (``dark signal'' in volts). The ``dark
-signal'' is calculated by averaging all measurements recorded for a
+is sampled at a rate of \(1\,\text{Hz}\). The mean and the standard
+deviation of these samples are calculated and recorded every minute. The
+measurements are corrected for the zero offset (``dark signal'' in
+volts), which is calculated by averaging all measurements recorded for a
 period of \(3\,\text{h}\), before (morning) or after (evening) the Sun
@@
 irradiance using a ramped value of the instrument's sensitivity between
-calibrations.
+subsequent calibrations.
 
@@
 Although it is impossible to detect all false data, the large number of
-available data, and the aggregation scheme we used, ensures the good
-quality of the radiation measurements used in this study.
+available data, and the aggregation scheme we used, ensures the quality
+of the radiation measurements used in this study.
 
@@
 negative but cannot affect the final results of the study.
+
+\hypertarget{the-clear-sky-identification-algorithm}{%
+\subsection{The clear sky identification
+algorithm}\label{the-clear-sky-identification-algorithm}}
 

### 2023-12-16 MDPI_submission.Rmd
Commit 2023-12-16 13:09

@@
 
-<!--
-Η εισαγωγή είναι εντάξει. Ίσως θα μπορούσες να βάλεις στη 2η
-παράγραφο ακόμη 2-3 σειρές για τις τάσεις του αεροζόλ και των νεφών
-από τις εργασίες και να καταλήξεις με κάποια στοιχεία για το αεροζόλ
-στη Θεσσαλονίκη από εργασίες δικές μας. Για τα νέφη δεν έχω υπόψη μου
-καμία εργασία αλλά ίσως θα μπορούσε να ειπωθεί κάτι χρησιμοποιώντας
-παγκόσμιες εκτιμήσεις για την περιοχή της Μεσογείου.
-π.χ. https://link.springer.com/article/10.1007/s00382-022-06438-3
--->
-
 The shortwave downward solar irradiance (SDR) at Earth's surface plays a significant role, on its climate.
@@
 
+<!-- ### Data selection -->
+Due to the significant measurement uncertainty when the Sun is near the horizon, we have excluded all measurements with SZA greater than $`r 90-MIN_ELEVA`^\circ$.
+Moreover, due to obstructions around the site (hills and buildings) which block the direct irradiance, we excluded data with azimuth angle in the range
+$`r FIBais_Az_1`^{\circ}$ - $`r FIBais_Az_2`^{\circ}$ and with SZA greater than $`r 90 -  FIBais_Elev`^{\circ}$.
+<!-- 1au -->
+To make the measurements comparable throughout the dataset, we adjusted all one-minute data to the mean Sun - Earth distance.
+<!-- TSI -->
+Subsequently, we adjusted all measurements to the Total Solar Irradiance (TSI) at $1\,\text{au}$, in order to compensate for the Sun's intensity variability, using a time series of satellite TSI observations.
+The TSI data we used are part of the "NOAA Climate Data Record of Total Solar Irradiance" dataset [@Coddington2005].
+The initial daily values of this dataset were interpolated to match the time step of our measurements.
+
 <!-- c) Clear sky identification -->
-In order to be able to estimate the effect of the sky conditions on the long term variability of SDR, we created three datasets, by characterizing each one-minute measurement with a corresponding sky condition (i.e., all-sky, clear-sky and cloudy-sky).
-To identify the clear-sky conditions we used a method proposed by
+In order to estimate the effect of the sky conditions on the long term variability of SDR, we created three datasets by characterizing each one-minute measurement with a corresponding sky-condition flag (i.e., all-sky, clear-sky and cloudy-sky).
+To identify the clear-cases we used a method proposed by
 @Long2000
@@
 @Reno2016,
-which were adapted and configured for the site, as the authors suggest.
-
+which were adapted and configured for the site of Thessaloniki.
 <!-- CS subjectivity -->
-We have to note, that the definition of clear or cloudy sky, has some subjectivity, in any method of characterization.
-As a result, the details of the definition are site specific, it relies on a combination of thresholds and comparisons with ideal radiation models and statistical analysis of different signal metrics.
-The CSid algorithm was calibrated with the main focus, to identify the presence of clouds on the sky.
-Despite the fine-tuning of the procedure, a few marginal cases exist, that have been identified manually as false positive or false negative but cannot affect the final results of the study.
+We note that all methods have some subjectivity in the definition of clear or cloudy sky cases.
+As a result, the details of the definition are site specific and they rely on a combination of thresholds and comparisons with ideal radiation models and statistical analysis of different signal metrics.
+The CSid algorithm was calibrated with the main focus, to identify the presence of clouds. Despite the fine-tuning
+of the procedure, in a few marginal cases false positive or false
+negative results were identified by manual inspection. However, due to
+their small number, they cannot affect the final results of the study.
+For completeness, we provide below a brief overview of the clear-sky
+identification algorithm (CSid), along with the site specific
+thresholds.
 
@@
 
-For completeness, we will provide below a brief overview of the clear sky identification algorithm (CSid), along with the site specific thresholds.
-To calculate the reference clear sky $\text{SDR}_\text{CSref}$ we used the $\text{SDR}_\text{Haurwitz}$ derived by the radiation model of @Haurwitz1945, adjusted for our site with a factor $a$ (Eq.\ \ref{eq:ahau}), resulted by an iterative optimization process, as described by @Long2000 and @Reno2016.
+To calculate the reference clear-sky $\text{SDR}_\text{CSref}$ we used the $\text{SDR}_\text{Haurwitz}$ derived by the radiation model of @Haurwitz1945, adjusted for our site with a factor $a$ (Eq.\ \ref{eq:ahau}), estimated through an iterative optimization process, as described by @Long2000 and @Reno2016.
 The target of the optimization was the minimization of a function $f(a)$ (Eq.\ \ref{eq:minf}) and was accomplished with the algorithmic function "optimise", which is an implementation based on the work of @Brent1973, from the library "stats" of the R programming language [@RCT2023].
@@
 \end{equation}
-where: $n$ is the total number of daylight records, $\text{SDR}_{\text{CSid},i}$ are the records identified as clear sky by CSid, $a$ is a hypothetical adjustment factor, and $\text{SDR}_{\text{testCSref},i}$ is any of the tested clear sky radiation models.
+where: $n$ is the total number of daylight data, $\text{SDR}_{\text{CSid},i}$ are the data identified as clear sky by CSid, $a$ is a site-specific adjustment factor, and $\text{SDR}_{\text{testCSref},i}$ is the SDR derived by any of the
+tested clear-sky radiation models.
 
@@
 These models are described in more details by @Reno2012 and evaluated by @Reno2016.
-We found, that Haurwitz's model, adjusted with the factor $a = `r signif(alpha,3)`$ yields one of the lowest root mean squared errors (RMSE), while the procedure, manages to characterize the majority of the data.
+We found, that Haurwitz's model, adjusted with the factor $a = `r signif(alpha,3)`$ yields one of the lowest root mean squared errors (RMSE), while the procedure manages to characterize successfully the majority of the data.
 Thus, our clear sky reference is derived by the Eq.\ \ref{eq:ahau}.
@@
 \end{equation}
-where: $\text{SDR}_\text{CSref}$ is the reference clear sky SDR, in $\text{W}\,\text{m}^{-2}$ and $\theta$ is the solar zenith angle (SZA).
+where $\theta$ is the solar zenith angle (SZA).
 
 The criteria that were used to identify whether a measurement was taken under clear-sky conditions are presented below.
-A data point is flagged as "clear-sky" if all criteria are satisfied, otherwise it is considered to be "cloud-sky".
-Each criterion was applied for a running window of $`r nt`$ consecutive one-minute measurements, and the characterization is assigned to the central value of the window.
+A data point is flagged as "clear-sky" if all criteria are satisfied; otherwise it is considered as "cloud-sky".
+Each criterion was applied for a running window of $`r nt`$ consecutive one-minute measurements, and the characterization was assigned to the central datum of the window.
 Each parameter, was calculated both from the observations and the reference clear sky model, for each comparison.
@@
 The factors and the offsets were determined empirically, by manually inspecting each filter's performance on selected days and adjusting them accordingly during an iterative process.
-The criteria are:
+The criteria are
+listed below, together with the range of values within which the
+respective parameter should fall in order to raise the clear-sky flag:
 <!-- 1. MeanVIP -->
@@
 
-<!-- ### Data selection -->
-Due to the significant measurement uncertainty when the Sun is near the horizon, we have excluded all measurements with SZA greater than $`r 90-MIN_ELEVA`^\circ$.
-Moreover, due to some obstructions around the site (hills and buildings), we excluded data with azimuth angle between
-$`r FIBais_Az_1`^\circ$ and $`r FIBais_Az_2`^\circ$ with SZA greater than $`r 90 -  FIBais_Elev`^\circ$.
-On the latter instances, Sun is systematically not visible from the instrument's location.
-<!-- 1au -->
-To make the measurements comparable throughout the dataset, we adjusted all one-minute radiometric values to the mean Sun - Earth distance.
-<!-- TSI -->
-Subsequently, we made all measurements relative to the Total Solar Irradiance (TSI) at $1\,\text{au}$, in order to compensate for the Sun's intensity variability, using a time series of satellite TSI observations.
-The TSI data we used are part of the "NOAA Climate Data Record of Total Solar Irradiance" dataset [@Coddington2005].
-The initial daily values of this dataset were interpolated to match the time step of our measurements.
 In the final dataset
@@
 
-<!-- aggregation -->
-In order to investigate the SDR trends, we implemented an appropriate aggregation scheme to the 1-minute data to derive a series in coarser time-scales.
+## Aggregation of data and statistical approach
+
+In order to investigate the SDR trends which are the main focus of the
+study, we implemented an appropriate aggregation scheme to the
+one-minute data to derive a series in coarser time-scales.
 To preserve the representativeness of the data we used the following criteria:
@@
 
-```{r trendALL, echo=F, out.width='.70\\linewidth', fig.cap=paste0("Anomalies (\\%) of the daily all-sky SDR, relative to climatological values for ", year(min(ALL_1_daily_DESEAS$Date)), " - ", year(max(ALL_1_daily_DESEAS$Date)), ".
-The black line shows the long term linear trend.")}
+```{r trendALL, echo=F, out.width='.70\\linewidth', fig.cap=paste0("Anomalies (\\%) of the daily all-sky SDR from the climatological mean for the period", year(min(ALL_1_daily_DESEAS$Date)), " - ", year(max(ALL_1_daily_DESEAS$Date)), ".
+The black line is the long term linear trend.")}
 knitr::include_graphics("./images/LongtermTrends-2.png")

### 2023-12-16 MDPI_submission.tex
Commit 2023-12-16 13:09

@@
 
-In order to be able to estimate the effect of the sky conditions on the
-long term variability of SDR, we created three datasets, by
-characterizing each one-minute measurement with a corresponding sky
-condition (i.e., all-sky, clear-sky and cloudy-sky). To identify the
-clear-sky conditions we used a method proposed by \citet{Long2000} and
-by \citet{Reno2016}, which were adapted and configured for the site, as
-the authors suggest.
+Due to the significant measurement uncertainty when the Sun is near the
+horizon, we have excluded all measurements with SZA greater than
+\(85^\circ\). Moreover, due to obstructions around the site (hills and
+buildings) which block the direct irradiance, we excluded data with
+azimuth angle in the range \(58^{\circ}\) - \(120^{\circ}\) and with SZA
+greater than \(78^{\circ}\). To make the measurements comparable
+throughout the dataset, we adjusted all one-minute data to the mean Sun
+- Earth distance. Subsequently, we adjusted all measurements to the
+Total Solar Irradiance (TSI) at \(1\,\text{au}\), in order to compensate
+for the Sun's intensity variability, using a time series of satellite
+TSI observations. The TSI data we used are part of the ``NOAA Climate
+Data Record of Total Solar Irradiance'' dataset \citep{Coddington2005}.
+The initial daily values of this dataset were interpolated to match the
+time step of our measurements.
 
-We have to note, that the definition of clear or cloudy sky, has some
-subjectivity, in any method of characterization. As a result, the
-details of the definition are site specific, it relies on a combination
-of thresholds and comparisons with ideal radiation models and
-statistical analysis of different signal metrics. The CSid algorithm was
-calibrated with the main focus, to identify the presence of clouds on
-the sky. Despite the fine-tuning of the procedure, a few marginal cases
-exist, that have been identified manually as false positive or false
-negative but cannot affect the final results of the study.
+In order to estimate the effect of the sky conditions on the long term
+variability of SDR, we created three datasets by characterizing each
+one-minute measurement with a corresponding sky-condition flag (i.e.,
+all-sky, clear-sky and cloudy-sky). To identify the clear-cases we used
+a method proposed by \citet{Long2000} and by \citet{Reno2016}, which
+were adapted and configured for the site of Thessaloniki. We note that
+all methods have some subjectivity in the definition of clear or cloudy
+sky cases. As a result, the details of the definition are site specific
+and they rely on a combination of thresholds and comparisons with ideal
+radiation models and statistical analysis of different signal metrics.
+The CSid algorithm was calibrated with the main focus, to identify the
+presence of clouds. Despite the fine-tuning of the procedure, in a few
+marginal cases false positive or false negative results were identified
+by manual inspection. However, due to their small number, they cannot
+affect the final results of the study. For completeness, we provide
+below a brief overview of the clear-sky identification algorithm (CSid),
+along with the site specific thresholds.
 
@@
 
-For completeness, we will provide below a brief overview of the clear
-sky identification algorithm (CSid), along with the site specific
-thresholds. To calculate the reference clear sky
-\(\text{SDR}_\text{CSref}\) we used the \(\text{SDR}_\text{Haurwitz}\)
-derived by the radiation model of \citet{Haurwitz1945}, adjusted for our
-site with a factor \(a\) (Eq.~\ref{eq:ahau}), resulted by an iterative
-optimization process, as described by \citet{Long2000} and
-\citet{Reno2016}. The target of the optimization was the minimization of
-a function \(f(a)\) (Eq.~\ref{eq:minf}) and was accomplished with the
-algorithmic function ``optimise'', which is an implementation based on
-the work of \citet{Brent1973}, from the library ``stats'' of the R
-programming language \citep{RCT2023}. \begin{equation}
+To calculate the reference clear-sky \(\text{SDR}_\text{CSref}\) we used
+the \(\text{SDR}_\text{Haurwitz}\) derived by the radiation model of
+\citet{Haurwitz1945}, adjusted for our site with a factor \(a\)
+(Eq.~\ref{eq:ahau}), estimated through an iterative optimization
+process, as described by \citet{Long2000} and \citet{Reno2016}. The
+target of the optimization was the minimization of a function \(f(a)\)
+(Eq.~\ref{eq:minf}) and was accomplished with the algorithmic function
+``optimise'', which is an implementation based on the work of
+\citet{Brent1973}, from the library ``stats'' of the R programming
+language \citep{RCT2023}. \begin{equation}
 f(a) = \frac{1}{n}\sum_{i=1}^{n} ( \text{SDR}_{\text{CSid},i} - a \times \text{SDR}_{\text{testCSref},i} )^2 \label{eq:minf}
-\end{equation} where: \(n\) is the total number of daylight records,
-\(\text{SDR}_{\text{CSid},i}\) are the records identified as clear sky
-by CSid, \(a\) is a hypothetical adjustment factor, and
-\(\text{SDR}_{\text{testCSref},i}\) is any of the tested clear sky
-radiation models.
+\end{equation} where: \(n\) is the total number of daylight data,
+\(\text{SDR}_{\text{CSid},i}\) are the data identified as clear sky by
+CSid, \(a\) is a site-specific adjustment factor, and
+\(\text{SDR}_{\text{testCSref},i}\) is the SDR derived by any of the
+tested clear-sky radiation models.
 
@@
 factor \(a = 0.965\) yields one of the lowest root mean squared errors
-(RMSE), while the procedure, manages to characterize the majority of the
-data. Thus, our clear sky reference is derived by the Eq.~\ref{eq:ahau}.
-\begin{equation}
+(RMSE), while the procedure manages to characterize successfully the
+majority of the data. Thus, our clear sky reference is derived by the
+Eq.~\ref{eq:ahau}. \begin{equation}
 \text{SDR}_\text{CSref} = a \times \text{SDR}_\text{Haurwitz} = 0.965 \times 1098 \times \cos(\theta) \times \exp \left( \frac{ - 0.057}{\cos(\theta)} \right) \label{eq:ahau}
-\end{equation} where: \(\text{SDR}_\text{CSref}\) is the reference clear
-sky SDR, in \(\text{W}\,\text{m}^{-2}\) and \(\theta\) is the solar
-zenith angle (SZA).
+\end{equation} where \(\theta\) is the solar zenith angle (SZA).
 
@@
 under clear-sky conditions are presented below. A data point is flagged
-as ``clear-sky'' if all criteria are satisfied, otherwise it is
-considered to be ``cloud-sky''. Each criterion was applied for a running
+as ``clear-sky'' if all criteria are satisfied; otherwise it is
+considered as ``cloud-sky''. Each criterion was applied for a running
 window of \(11\) consecutive one-minute measurements, and the
-characterization is assigned to the central value of the window. Each
+characterization was assigned to the central datum of the window. Each
 parameter, was calculated both from the observations and the reference
@@
 selected days and adjusting them accordingly during an iterative
-process. The criteria are:
+process. The criteria are listed below, together with the range of
+values within which the respective parameter should fall in order to
+raise the clear-sky flag:
 
@@
 
-Due to the significant measurement uncertainty when the Sun is near the
-horizon, we have excluded all measurements with SZA greater than
-\(85^\circ\). Moreover, due to some obstructions around the site (hills
-and buildings), we excluded data with azimuth angle between \(58^\circ\)
-and \(120^\circ\) with SZA greater than \(78^\circ\). On the latter
-instances, Sun is systematically not visible from the instrument's
-location. To make the measurements comparable throughout the dataset, we
-adjusted all one-minute radiometric values to the mean Sun - Earth
-distance. Subsequently, we made all measurements relative to the Total
-Solar Irradiance (TSI) at \(1\,\text{au}\), in order to compensate for
-the Sun's intensity variability, using a time series of satellite TSI
-observations. The TSI data we used are part of the ``NOAA Climate Data
-Record of Total Solar Irradiance'' dataset \citep{Coddington2005}. The
-initial daily values of this dataset were interpolated to match the time
-step of our measurements. In the final dataset \(23.3\%\) of the
-1-minute data were identified as under clear-sky conditions and \(43\%\)
-as under cloud-sky conditions.
+In the final dataset \(23.3\%\) of the 1-minute data were identified as
+under clear-sky conditions and \(43\%\) as under cloud-sky conditions.
 
-In order to investigate the SDR trends, we implemented an appropriate
-aggregation scheme to the 1-minute data to derive a series in coarser
-time-scales. To preserve the representativeness of the data we used the
-following criteria: a) we accept only days with more than 50\% of the
-daytime measurements present and valid, b) on the dataset of clear- and
-cloudy-skies, we included only days with more than 60\% of the daytime
-measurements identified as clear or cloudy respectively, c) monthly
-values were computed from daily means only when at least 20 days were
-available. To create the daily and monthly climatological means, we
-averaged the data based on the day of year and calendar month,
-respectively. For the seasonal means we averaged the mean daily values
-in each season (Winter: December - February, Spring: March - May, etc.).
-Finally, each data set was deseasonalized by subtracting the
-corresponding climatological annual cycle (daily or monthly) from the
-actual data. To estimate the SZA effect on the SDR trends, the
-one-minute data were aggregated in \(1^\circ\) SZA bins, separately for
-the morning and afternoon hours, and then were deseasonalized as
-mentioned above.
+\hypertarget{aggregation-of-data-and-statistical-approach}{%
+\subsection{Aggregation of data and statistical
+approach}\label{aggregation-of-data-and-statistical-approach}}
+
+In order to investigate the SDR trends which are the main focus of the
+study, we implemented an appropriate aggregation scheme to the
+one-minute data to derive a series in coarser time-scales. To preserve
+the representativeness of the data we used the following criteria: a) we
+accept only days with more than 50\% of the daytime measurements present
+and valid, b) on the dataset of clear- and cloudy-skies, we included
+only days with more than 60\% of the daytime measurements identified as
+clear or cloudy respectively, c) monthly values were computed from daily
+means only when at least 20 days were available. To create the daily and
+monthly climatological means, we averaged the data based on the day of
+year and calendar month, respectively. For the seasonal means we
+averaged the mean daily values in each season (Winter: December -
+February, Spring: March - May, etc.). Finally, each data set was
+deseasonalized by subtracting the corresponding climatological annual
+cycle (daily or monthly) from the actual data. To estimate the SZA
+effect on the SDR trends, the one-minute data were aggregated in
+\(1^\circ\) SZA bins, separately for the morning and afternoon hours,
+and then were deseasonalized as mentioned above.
 
@@
 
-\caption{Anomalies (\%) of the daily all-sky SDR, relative to climatological values for 1993 - 2023. The black line shows the long term linear trend.}\label{fig:trendALL}
+\caption{Anomalies (\%) of the daily all-sky SDR from the climatological mean for the period1993 - 2023. The black line is the long term linear trend.}\label{fig:trendALL}
 \end{figure}

### 2023-12-16 MDPI_submission.Rmd
Commit 2023-12-16 14:14

@@
 
+In the final [dataset $23.3\%$ of the 1-minute data were
+identified as under clear-sky conditions and]{.mark} $43\%$ [as]{.mark}
+under cloud-sky conditions.
+
 ## Aggregation of data and statistical approach
@@
 <!-- a) for daily mean values we exclude days with less than `r Daily_aggregation_N_lim` valid data points, -->
-a) we accept only days with more than `r 100 * All_daily_ratio_lim`% of the daytime measurements present and valid,
-b) on the dataset of clear- and cloudy-skies, we included only days with more than
-`r 100 * Clear_daily_ratio_lim`% of the daytime measurements identified as clear or cloudy respectively,
-c) monthly values were computed from daily means
-only when at least `r Monthly_aggegation_N_lim` days were available.
+a) we excluded all days with less than `r 100 * All_daily_ratio_lim`% of the expected daytime measurements,
+b) daily means for the clear-sky and cloudy-sky datasets
+were calculated only for days with more than
+`r 100 * Clear_daily_ratio_lim`% of the expected daytime
+measurements identified as clear or cloudy respectively,
+c) monthly means were computed from daily means only when at least `r Monthly_aggegation_N_lim` days of the month were available.
 To create the daily and monthly climatological means, we averaged the data based on the day of year and calendar month, respectively.
@@
 Finally, each data set was deseasonalized by subtracting the corresponding climatological annual cycle (daily or monthly) from the actual data.
-To estimate the SZA effect on the SDR trends, the one-minute data were aggregated in $1^\circ$ SZA bins, separately for the morning and afternoon hours, and then were deseasonalized as mentioned above.
+Finally, to estimate the SZA effect on the SDR trends, the one-minute data were aggregated in $1^\circ$ SZA bins, separately for the morning and afternoon hours, and then were deseasonalized as mentioned above.
 
@@
 
-We calculated the linear SDR trends, from the departures of the mean daily values from the daily climatology and for the three sky conditions (Table\ \ref{tab:trendtable}).
-In Figure\ \ref{fig:trendALL} we present only the time series under all-sky conditions; the plots for clear-sky and cloud-sky conditions, are very similar and are shown in the Appendix (Figures\ \ref{fig:trendCLEAR} and\  \ref{fig:trendCLOUD}).
+We calculated the linear trends of SDR, from the departures of the mean daily values from the daily climatology and for the three sky conditions (Table\ \ref{tab:trendtable}).
+In Figure\ \ref{fig:trendALL} we present only the time series under all-sky conditions; the plots for clear-sky and cloud-sky conditions, are shown in the Appendix (Figures\ \ref{fig:trendCLEAR} and\  \ref{fig:trendCLOUD}).
 We observe a positive trend for all-sky conditions
@@
 
-The solar zenith angle is a major factor of SDR reaching the ground, due to the enhancement of the radiation path in the atmosphere, especially in urban environments with human activities emitting aerosols [@Wang2021].
-In order to estimate the effect of the SZA on the SDR trends, we grouped the anomaly data in bins of $1^\circ$ SZA, and calculated the overall trend for each bin before noon and after noon (Figure\ \ref{fig:szatrends}).
+The solar zenith angle is a major factor affecting the SDR, since
+increases in SZA leads to enhancement of the radiation path in the
+atmosphere, especially in urban environments with human activities
+emitting aerosols 
+[@Wang2021].
+In order to estimate the effect of SZA on the SDR trends, we grouped the anomaly data in bins of $1^\circ$ SZA, and calculated the overall trend for each bin before noon and after noon (Figure\ \ref{fig:szatrends}).
 Although there are seasonal dependencies of the minimum SZA (see Appendix, Figure\ \ref{fig:SZAtrendSeason}), these dependencies would not be further examined here.
@@
 
-Similarly to the long term trends discussed above, we have calculated the trend of the anomalies for the three different sky conditions, and for each season of the year, using the corresponding mean monthly values
+Similarly to the long term trends from daily means of SDR discussed above, we have calculated the trend of the anomalies for the three different sky conditions, and for each season of the year, using the corresponding mean monthly values
 (Figure\ \ref{fig:seasonalALL} and Table\ \ref{tab:trendseasontable}).

### 2023-12-16 MDPI_submission.tex
Commit 2023-12-16 14:14

@@
 
+In the final \hl{dataset \(23.3\%\) of the 1-minute data were identified
+as under clear-sky conditions and} \(43\%\) \hl{as} under cloud-sky
+conditions.
+
 \hypertarget{aggregation-of-data-and-statistical-approach}{%
@@
 the representativeness of the data we used the following criteria: a) we
-accept only days with more than 50\% of the daytime measurements present
-and valid, b) on the dataset of clear- and cloudy-skies, we included
-only days with more than 60\% of the daytime measurements identified as
-clear or cloudy respectively, c) monthly values were computed from daily
-means only when at least 20 days were available. To create the daily and
-monthly climatological means, we averaged the data based on the day of
-year and calendar month, respectively. For the seasonal means we
-averaged the mean daily values in each season (Winter: December -
-February, Spring: March - May, etc.). Finally, each data set was
-deseasonalized by subtracting the corresponding climatological annual
-cycle (daily or monthly) from the actual data. To estimate the SZA
-effect on the SDR trends, the one-minute data were aggregated in
-\(1^\circ\) SZA bins, separately for the morning and afternoon hours,
-and then were deseasonalized as mentioned above.
+excluded all days with less than 50\% of the expected daytime
+measurements, b) daily means for the clear-sky and cloudy-sky datasets
+were calculated only for days with more than 60\% of the expected
+daytime measurements identified as clear or cloudy respectively, c)
+monthly means were computed from daily means only when at least 20 days
+of the month were available. To create the daily and monthly
+climatological means, we averaged the data based on the day of year and
+calendar month, respectively. For the seasonal means we averaged the
+mean daily values in each season (Winter: December - February, Spring:
+March - May, etc.). Finally, each data set was deseasonalized by
+subtracting the corresponding climatological annual cycle (daily or
+monthly) from the actual data. Finally, to estimate the SZA effect on
+the SDR trends, the one-minute data were aggregated in \(1^\circ\) SZA
+bins, separately for the morning and afternoon hours, and then were
+deseasonalized as mentioned above.
 
@@
 
-We calculated the linear SDR trends, from the departures of the mean
+We calculated the linear trends of SDR, from the departures of the mean
 daily values from the daily climatology and for the three sky conditions
@@
 only the time series under all-sky conditions; the plots for clear-sky
-and cloud-sky conditions, are very similar and are shown in the Appendix
+and cloud-sky conditions, are shown in the Appendix
 (Figures~\ref{fig:trendCLEAR} and~ \ref{fig:trendCLOUD}). We observe a
@@
 
-The solar zenith angle is a major factor of SDR reaching the ground, due
-to the enhancement of the radiation path in the atmosphere, especially
-in urban environments with human activities emitting aerosols
-\citep{Wang2021}. In order to estimate the effect of the SZA on the SDR
-trends, we grouped the anomaly data in bins of \(1^\circ\) SZA, and
-calculated the overall trend for each bin before noon and after noon
-(Figure~\ref{fig:szatrends}). Although there are seasonal dependencies
-of the minimum SZA (see Appendix, Figure~\ref{fig:SZAtrendSeason}),
-these dependencies would not be further examined here. For all-sky and
-clear-sky conditions the brightening effect of SDR (positive trend) is
-stronger for large SZAs (Figures~\ref{fig:szatrends-1} and
-\ref{fig:szatrends-2}). The trends in the morning and afternoon hours
-are more or less consistent with small differences, which can be
-attributed to systematic diurnal variations of aerosols, particularly
-during the warm period of the year \citep{Wang2021}. For cloudy-sky
-conditions (Figure~\ref{fig:szatrends-3}), we can not discern any
-significant dependence of the SDR trend with SZA. For SZAs \(16^\circ\)
-- \(50^\circ\), the trends range within about \(\pm 0.2\,\%/y\), with a
+The solar zenith angle is a major factor affecting the SDR, since
+increases in SZA leads to enhancement of the radiation path in the
+atmosphere, especially in urban environments with human activities
+emitting aerosols \citep{Wang2021}. In order to estimate the effect of
+SZA on the SDR trends, we grouped the anomaly data in bins of
+\(1^\circ\) SZA, and calculated the overall trend for each bin before
+noon and after noon (Figure~\ref{fig:szatrends}). Although there are
+seasonal dependencies of the minimum SZA (see Appendix,
+Figure~\ref{fig:SZAtrendSeason}), these dependencies would not be
+further examined here. For all-sky and clear-sky conditions the
+brightening effect of SDR (positive trend) is stronger for large SZAs
+(Figures~\ref{fig:szatrends-1} and \ref{fig:szatrends-2}). The trends in
+the morning and afternoon hours are more or less consistent with small
+differences, which can be attributed to systematic diurnal variations of
+aerosols, particularly during the warm period of the year
+\citep{Wang2021}. For cloudy-sky conditions
+(Figure~\ref{fig:szatrends-3}), we can not discern any significant
+dependence of the SDR trend with SZA. For SZAs \(16^\circ\) -
+\(50^\circ\), the trends range within about \(\pm 0.2\,\%/y\), with a
 weak statistical significance. Between \(50^\circ\) and \(75^\circ\) SZA
@@
 
-Similarly to the long term trends discussed above, we have calculated
-the trend of the anomalies for the three different sky conditions, and
-for each season of the year, using the corresponding mean monthly values
-(Figure~\ref{fig:seasonalALL} and Table~\ref{tab:trendseasontable}). For
-all-sky conditions the trend in SDR in winter is the largest
-(\(0.69\,\%/y\)), followed by the trend in autumn (\(0.43\,\%/y\), a
-value close to the long term trend) both statistically significant above
-the \(99\,\%\) confidence level. In spring and summer, the trends are
-much smaller and of lesser statistical significance. These seasonal
-differences indicate a possible relation of the trends in SDR to trends
-of clouds during winter and autumn. For clear-skies, the trend in winter
-is \(0.36\,\%/y\), larger than for all-skies (\(0.69\,\%/y\)), which is
-another indication of a decreasing trend in cloud optical thickness.
-Moreover, the trends under clear- and cloudy-sky conditions are almost
-complementary to each other, particularly for winter and autumn, where
-the signal is stronger. During spring and summer the statistical
-significance is very low and the actual trend too small for a meaningful
-comparison.
+Similarly to the long term trends from daily means of SDR discussed
+above, we have calculated the trend of the anomalies for the three
+different sky conditions, and for each season of the year, using the
+corresponding mean monthly values (Figure~\ref{fig:seasonalALL} and
+Table~\ref{tab:trendseasontable}). For all-sky conditions the trend in
+SDR in winter is the largest (\(0.69\,\%/y\)), followed by the trend in
+autumn (\(0.43\,\%/y\), a value close to the long term trend) both
+statistically significant above the \(99\,\%\) confidence level. In
+spring and summer, the trends are much smaller and of lesser statistical
+significance. These seasonal differences indicate a possible relation of
+the trends in SDR to trends of clouds during winter and autumn. For
+clear-skies, the trend in winter is \(0.36\,\%/y\), larger than for
+all-skies (\(0.69\,\%/y\)), which is another indication of a decreasing
+trend in cloud optical thickness. Moreover, the trends under clear- and
+cloudy-sky conditions are almost complementary to each other,
+particularly for winter and autumn, where the signal is stronger. During
+spring and summer the statistical significance is very low and the
+actual trend too small for a meaningful comparison.
 

### 2023-12-17 MDPI_submission.Rmd
Commit 2023-12-16 15:10
Commit 2023-12-17 20:14

@@
 
-In the final [dataset $23.3\%$ of the 1-minute data were
-identified as under clear-sky conditions and]{.mark} $43\%$ [as]{.mark}
-under cloud-sky conditions.
-
 ## Aggregation of data and statistical approach
@@
 [@Wang2021].
-In order to estimate the effect of SZA on the SDR trends, we grouped the anomaly data in bins of $1^\circ$ SZA, and calculated the overall trend for each bin before noon and after noon (Figure\ \ref{fig:szatrends}).
-Although there are seasonal dependencies of the minimum SZA (see Appendix, Figure\ \ref{fig:SZAtrendSeason}), these dependencies would not be further examined here.
-For all-sky and clear-sky conditions the brightening effect of SDR (positive trend) is stronger for large SZAs (Figures\ \ref{fig:szatrends-1} and \ref{fig:szatrends-2}).
+In order to estimate the effect of SZA on the SDR trends, we grouped the data in bins of $1^\circ$ SZA, and calculated the overall trend for each bin, separately for the daily
+periods before noon and after noon (Figure\ \ref{fig:szatrends}).
+Although there are seasonal dependencies of the minimum SZA (see Appendix, Figure\ \ref{fig:SZAtrendSeason}), these dependencies are not discussed further.
+
+For all-sky conditions the brightening effect of SDR (positive trend) is
+stronger for large SZAs (Figures\ \ref{fig:szatrends-1} and \ref{fig:szatrends-2}).
 The trends in the morning and afternoon hours are more or less consistent with small differences, which can be attributed to systematic diurnal variations of aerosols, particularly during the warm period of the year [@Wang2021].
+
 For cloudy-sky conditions (Figure\ \ref{fig:szatrends-3}), we can not discern any significant dependence of the SDR trend with SZA.

### 2023-12-17 MDPI_submission.tex
Commit 2023-12-16 15:10
Commit 2023-12-17 20:14

@@
 
-In the final \hl{dataset \(23.3\%\) of the 1-minute data were identified
-as under clear-sky conditions and} \(43\%\) \hl{as} under cloud-sky
-conditions.
-
 \hypertarget{aggregation-of-data-and-statistical-approach}{%
@@
 emitting aerosols \citep{Wang2021}. In order to estimate the effect of
-SZA on the SDR trends, we grouped the anomaly data in bins of
-\(1^\circ\) SZA, and calculated the overall trend for each bin before
-noon and after noon (Figure~\ref{fig:szatrends}). Although there are
-seasonal dependencies of the minimum SZA (see Appendix,
-Figure~\ref{fig:SZAtrendSeason}), these dependencies would not be
-further examined here. For all-sky and clear-sky conditions the
-brightening effect of SDR (positive trend) is stronger for large SZAs
-(Figures~\ref{fig:szatrends-1} and \ref{fig:szatrends-2}). The trends in
-the morning and afternoon hours are more or less consistent with small
-differences, which can be attributed to systematic diurnal variations of
-aerosols, particularly during the warm period of the year
-\citep{Wang2021}. For cloudy-sky conditions
-(Figure~\ref{fig:szatrends-3}), we can not discern any significant
-dependence of the SDR trend with SZA. For SZAs \(16^\circ\) -
-\(50^\circ\), the trends range within about \(\pm 0.2\,\%/y\), with a
-weak statistical significance. Between \(50^\circ\) and \(75^\circ\) SZA
-the trends for the period before noon are stronger and negative,
-possibly associated with stronger attenuation by clouds under oblique
-incidence angles.
+SZA on the SDR trends, we grouped the data in bins of \(1^\circ\) SZA,
+and calculated the overall trend for each bin, separately for the daily
+periods before noon and after noon (Figure~\ref{fig:szatrends}).
+Although there are seasonal dependencies of the minimum SZA (see
+Appendix, Figure~\ref{fig:SZAtrendSeason}), these dependencies are not
+discussed further.
+
+For all-sky conditions the brightening effect of SDR (positive trend) is
+stronger for large SZAs (Figures~\ref{fig:szatrends-1} and
+\ref{fig:szatrends-2}). The trends in the morning and afternoon hours
+are more or less consistent with small differences, which can be
+attributed to systematic diurnal variations of aerosols, particularly
+during the warm period of the year \citep{Wang2021}.
+
+For cloudy-sky conditions (Figure~\ref{fig:szatrends-3}), we can not
+discern any significant dependence of the SDR trend with SZA. For SZAs
+\(16^\circ\) - \(50^\circ\), the trends range within about
+\(\pm 0.2\,\%/y\), with a weak statistical significance. Between
+\(50^\circ\) and \(75^\circ\) SZA the trends for the period before noon
+are stronger and negative, possibly associated with stronger attenuation
+by clouds under oblique incidence angles.
 

### 2023-12-17 DHI_GHI_1_longterm_trends.R

@@ -1165,7 +1165,7 @@ for (avar in vars) {
 
 
 
-## ____ by season in a tight grid ----------------------------------------------
+## ____ tight grid by season ----------------------------------------------
 
 #'
 #' ### Grid of trends for each season of year from daily
@@ -1176,9 +1176,12 @@ for (avar in vars) {
 {
     vars        <- c("GLB_att_des", "wattGLB")
     avar        <- vars[1]
-    dbs         <- c(  "ALL_1_D_bySeason_DESEAS",
-                       "CLEAR_1_D_bySeason_DESEAS",
-                       "CLOUD_1_D_bySeason_DESEAS")
+    # dbs         <- c("ALL_1_D_bySeason_DESEAS",
+    #                  "CLEAR_1_D_bySeason_DESEAS",
+    #                  "CLOUD_1_D_bySeason_DESEAS")
+    dbs         <- c("ALL_1_D_bySeason_DESEAS",
+                     "CLEAR_1_D_bySeason_DESEAS",
+                     "CLOUD_1_D_bySeason_DESEAS")
     Seasons     <- c("Winter", "Spring", "Summer", "Autumn")
     lec         <- 0
 

### 2023-12-17 MDPI_submission.Rmd
Commit 2023-12-17 21:10

@@
 
-# Introduction.
+# Introduction
 
@@
 In Figure\ \ref{fig:trendALL} we present only the time series under all-sky conditions; the plots for clear-sky and cloud-sky conditions, are shown in the Appendix (Figures\ \ref{fig:trendCLEAR} and\  \ref{fig:trendCLOUD}).
+In the studied period, there is no
+significant break or change in the variability pattern of the time
+series.
 We observe a positive trend for all-sky conditions
@@
 
-## Effects of the solar zenith angle on SDR.
+## Effects of the solar zenith angle on SDR
 
@@
 For all-sky conditions the brightening effect of SDR (positive trend) is
-stronger for large SZAs (Figures\ \ref{fig:szatrends-1} and \ref{fig:szatrends-2}).
+stronger for large SZAs (Figures\ \ref{fig:szatrends-1} and \ref{fig:szatrends-2})
+ranging from about $0.1\,\%/y$ to
+about $0.7\,\%/y$.
 The trends in the morning and afternoon hours are more or less consistent with small differences, which can be attributed to systematic diurnal variations of aerosols, particularly during the warm period of the year [@Wang2021].

### 2023-12-17 MDPI_submission.tex
Commit 2023-12-17 21:10

@@
 
-\hypertarget{introduction.}{%
-\section{Introduction.}\label{introduction.}}
+\hypertarget{introduction}{%
+\section{Introduction}\label{introduction}}
 
@@
 and cloud-sky conditions, are shown in the Appendix
-(Figures~\ref{fig:trendCLEAR} and~ \ref{fig:trendCLOUD}). We observe a
-positive trend for all-sky conditions (\(0.38\,\%/y\)), a very close but
-smaller trend for clear-skies (\(0.097\,\%/y\)) and a negative weaker
-trend for cloudy-skies (\(0.41\,\%/y\)). In the studied period, there is
-no significant break or change in the variability of the time series.
-Other studies for the European region reported a change of the SDR
-slope, around 1980 \citep{Wild2021, Yuan2021, Ohmura2009}, a few years
-before the start of our records. It is interesting to note, that for the
+(Figures~\ref{fig:trendCLEAR} and~ \ref{fig:trendCLOUD}). In the studied
+period, there is no significant break or change in the variability
+pattern of the time series. We observe a positive trend for all-sky
+conditions (\(0.38\,\%/y\)), a very close but smaller trend for
+clear-skies (\(0.097\,\%/y\)) and a negative weaker trend for
+cloudy-skies (\(0.41\,\%/y\)). In the studied period, there is no
+significant break or change in the variability of the time series. Other
+studies for the European region reported a change of the SDR slope,
+around 1980 \citep{Wild2021, Yuan2021, Ohmura2009}, a few years before
+the start of our records. It is interesting to note, that for the
 observations period, the trend of the TSI is \(-0.00024\,\%/y\), and
@@
 
-\hypertarget{effects-of-the-solar-zenith-angle-on-sdr.}{%
+\hypertarget{effects-of-the-solar-zenith-angle-on-sdr}{%
 \subsection{Effects of the solar zenith angle on
-SDR.}\label{effects-of-the-solar-zenith-angle-on-sdr.}}
+SDR}\label{effects-of-the-solar-zenith-angle-on-sdr}}
 
@@
 stronger for large SZAs (Figures~\ref{fig:szatrends-1} and
-\ref{fig:szatrends-2}). The trends in the morning and afternoon hours
-are more or less consistent with small differences, which can be
-attributed to systematic diurnal variations of aerosols, particularly
-during the warm period of the year \citep{Wang2021}.
+\ref{fig:szatrends-2}) ranging from about \(0.1\,\%/y\) to about
+\(0.7\,\%/y\). The trends in the morning and afternoon hours are more or
+less consistent with small differences, which can be attributed to
+systematic diurnal variations of aerosols, particularly during the warm
+period of the year \citep{Wang2021}.
 

### 2023-12-17 DHI_GHI_01_Input_longterm.R

@@ -989,10 +989,205 @@ setorder(CLEAR_1_D_bySeason_DESEAS, Yqrt)
 setorder(CLOUD_1_D_bySeason_DESEAS, Yqrt)
 
 
+rm(ALL_1_daily_mean,   ALL_1_daily_seas,
+   CLEAR_1_daily_mean, CLEAR_1_daily_seas,
+   CLOUD_1_daily_mean, CLOUD_1_daily_seas)
+dummy <- gc()
+
+
+
+
+
+In the studied period, there is no
+significant break or change in the variability pattern of the time
+series.
+
+
+
+## Season of year monthly aggregation --------------------------------------------
+
+## Quarter of year with one month shift to include December in the next years winter
+ALL_1_monthly_daily_mean[, season_Yqrt := as.yearqtr(as.yearmon(paste(year(Date), month(Date), sep = "-")) + 1/12)]
+CLEAR_1_monthly_daily_mean[, season_Yqrt := as.yearqtr(as.yearmon(paste(year(Date), month(Date), sep = "-")) + 1/12)]
+CLOUD_1_monthly_daily_mean[, season_Yqrt := as.yearqtr(as.yearmon(paste(year(Date), month(Date), sep = "-")) + 1/12)]
+stop("DD")
+## Flag seasons using quarters
+ALL_1_daily_mean[season_Yqrt %% 1 == 0   , Season := "Winter"]
+ALL_1_daily_mean[season_Yqrt %% 1 == 0.25, Season := "Spring"]
+ALL_1_daily_mean[season_Yqrt %% 1 == 0.50, Season := "Summer"]
+ALL_1_daily_mean[season_Yqrt %% 1 == 0.75, Season := "Autumn"]
+CLEAR_1_daily_mean[season_Yqrt %% 1 == 0   , Season := "Winter"]
+CLEAR_1_daily_mean[season_Yqrt %% 1 == 0.25, Season := "Spring"]
+CLEAR_1_daily_mean[season_Yqrt %% 1 == 0.50, Season := "Summer"]
+CLEAR_1_daily_mean[season_Yqrt %% 1 == 0.75, Season := "Autumn"]
+CLOUD_1_daily_mean[season_Yqrt %% 1 == 0   , Season := "Winter"]
+CLOUD_1_daily_mean[season_Yqrt %% 1 == 0.25, Season := "Spring"]
[…]

### 2023-12-18 MDPI_submission.Rmd
Commit 2023-12-18 10:14

@@
 LT$cor.t    <- NULL
+LT$Rsqrd    <- NULL
 
@@
 spp$RsqrdAdj       <- NULL
+spp$Rsqrd          <- NULL
 spp$cor.t          <- NULL

### 2023-12-18 DHI_GHI_3_trends_consistency.R

@@ -270,6 +270,7 @@ dsets    <- c(  "ALL_1_D_monthly_DESEAS",
               "CLEAR_1_D_monthly_DESEAS",
               "CLOUD_1_D_monthly_DESEAS")
 
+
 for (ad in dsets) {
     DBS  <- get(ad)
     type <- sub("_.*", "", ad)
@@ -412,6 +413,8 @@ database    <- c(  "ALL_1_daily_DESEAS",
                  "CLEAR_1_daily_DESEAS",
                  "CLOUD_1_daily_DESEAS")
 
+
+
 #+ CumulativeDailyCuSum, echo=F, include=T, results="asis"
 for (adb in database) {
     DB  <- get(adb)
@@ -476,7 +479,7 @@ for (adb in database) {
         ## daily from other DT
         lines(pdb$Date, pdb[[paste0(avar,"_cusum")]], col = col, lwd = 2)
 
-        points(pdb$Date, pdb[[paste0(avar,"_cusum")]], col = "red", lwd = 2)
+        # points(pdb$Date, pdb[[paste0(avar,"_cusum")]], col = "red", lwd = 2)
 
 
         if (DRAFT == T) {
@@ -877,6 +880,20 @@ database    <- c(  "ALL_1_D_monthly_DESEAS",
                  "CLOUD_1_D_monthly_DESEAS")
 lec         <- 0
 
+for (adb in database) {
+    DB    <- get(adb)
+    avar  <- "GLB_att"
+    TYPE  <- sub("_.*", "", adb)
+    wcare <- c("Date", avar ,paste0(avar,"_sd"), paste0(avar,"_des"), paste0(avar,"_cusum"))
+    # DB[, ..wcare]
+    # paste0("Export_",TYPE, "_1_D_monthly_DESEAS.csv")
+    write.csv(file = paste0("Export_",TYPE, "_1_D_monthly_DESEAS.csv"),
+              x = DB[, ..wcare],
+              row.names = F)
+}
+
+
+
 #+ CumulativeMonthlyCuSum, echo=F, include=T, results="asis"
 for (adb in database) {
     DB  <- get(adb)
@@ -940,6 +957,10 @@ for (adb in database) {
 
         ## daily from other DT
         lines(pdb$Date, pdb[[paste0(avar,"_cusum")]], col = col, lwd = 2)
+
+        #test
+        points(pdb$Date, pdb[[paste0(avar,"_cusum")]], col = "red")
+
 
[…]

### 2023-12-18 MDPI_submission.Rmd
Commit 2023-12-18 11:10

@@
 measurements identified as clear or cloudy respectively,
-c) monthly means were computed from daily means only when at least `r Monthly_aggegation_N_lim` days of the month were available.
-To create the daily and monthly climatological means, we averaged the data based on the day of year and calendar month, respectively.
-<!-- Seasonal -->
-For the seasonal means we averaged the mean daily values in each season (Winter: December - February, Spring: March - May, etc.).
-<!-- Deseasonalization -->
-Finally, each data set was deseasonalized by subtracting the corresponding climatological annual cycle (daily or monthly) from the actual data.
-Finally, to estimate the SZA effect on the SDR trends, the one-minute data were aggregated in $1^\circ$ SZA bins, separately for the morning and afternoon hours, and then were deseasonalized as mentioned above.
+c) monthly means were computed from daily means, for all-skies dataset 
+means were computed only when at least `r Monthly_aggegation_N_lim` 
+days of the month were available.
+Seasonal means were derived by averaging the
+monthly mean values in each season (winter: December - February, spring:
+March - May, etc.).
+The daily and monthly climatological means were
+derived by averaging the data for each day of year and calendar month,
+respectively. The daily and monthly datasets were deseasonalized by
+subtracting the corresponding climatological annual cycle (daily or
+monthly) from the actual data. Finally, to estimate the SZA effect on
+the SDR trends, the one-minute data were aggregated in $1^{\circ}$ SZA
+bins, separately for the morning and afternoon hours.
+
+The linear trends were calculated using an autoregressive model
+<!-- ([ARIMA......)]{.mark} and were reported together with the 2σ errors. -->
+<!-- [For the trends derived from monthly mean data we used an effective -->
+<!-- number of observations to calculate the standard error (Sander et al., -->
+<!-- 2000) in order to assess the significance of the trends.]{.mark} -->
+@Santer2000
 
@@
 about $0.7\,\%/y$.
-The trends in the morning and afternoon hours are more or less consistent with small differences, which can be attributed to systematic diurnal variations of aerosols, particularly during the warm period of the year [@Wang2021].
+The trends in the morning and afternoon hours are more or
+less consistent with small differences at small SZAs which can be
+attributed to effects on clear sky SDR from systematic diurnal patterns
+of aerosols during the warm period of the year [@Wang2021]. 
+Note that SZAs less than $35^\circ$ can only occur during the warm period of the
+year when clear skies are more frequent.
 
-For cloudy-sky conditions (Figure\ \ref{fig:szatrends-3}), we can not discern any significant dependence of the SDR trend with SZA.
-For SZAs $16^\circ$ - $50^\circ$, the trends range within about $\pm 0.2\,\%/y$, with a weak statistical significance.
+For cloudy-sky conditions (Figure\ \ref{fig:szatrends-3}), we can not discern any significant dependence of the SDR trend with SZA as the variability of irradiance is
+dominated by the cloud effects.
+For SZAs $25^\circ$ - $75^\circ$, the trends range within about $\pm 0.2\,\%/y$, 
+the trends are variable and for most SZAs insignificant ranging between
+$0$ and $0.2%\,\%/y$.
 Between $50^\circ$ and $75^\circ$ SZA the trends for the period before noon are stronger and negative, possibly associated with stronger attenuation by clouds under oblique incidence angles.
@@
 ## Prepare table and data
-spp <- read.csv("./figures/tbl_longterm_trends_season.dat",
+
+# spp <- read.csv("./figures/tbl_longterm_trends_season.dat",
+#                strip.white = TRUE,
+#                sep         = ";",
+#                header      = TRUE)
+spp <- read.csv("./figures/tbl_longterm_trends_season_byM.dat",
                strip.white = TRUE,
@@
                header      = TRUE)
+
 spp <- spp[grep("SDR" ,spp$var), ]
@@
 spp$slope.stat_sig <- NULL
+spp$N_eff          <- NULL
 
@@
         {\centering 
-            \includegraphics[width=1\linewidth]{./images/SeasonalTrendsTogether3-2} 
+            \includegraphics[width=1\linewidth]{./images/SeasonalMTrendsTogether3-2}   % Seasonal from Monthly
         }
@@
 \end{figure}
+
+            <!-- % \includegraphics[width=1\linewidth]{./images/SeasonalTrendsTogether3-2}  % Seasonal from Daily -->
 

### 2023-12-18 MDPI_submission.tex
Commit 2023-12-18 11:10

@@
 daytime measurements identified as clear or cloudy respectively, c)
-monthly means were computed from daily means only when at least 20 days
-of the month were available. To create the daily and monthly
-climatological means, we averaged the data based on the day of year and
-calendar month, respectively. For the seasonal means we averaged the
-mean daily values in each season (Winter: December - February, Spring:
-March - May, etc.). Finally, each data set was deseasonalized by
+monthly means were computed from daily means, for all-skies dataset
+means were computed only when at least 20 days of the month were
+available. Seasonal means were derived by averaging the monthly mean
+values in each season (winter: December - February, spring: March - May,
+etc.). The daily and monthly climatological means were derived by
+averaging the data for each day of year and calendar month,
+respectively. The daily and monthly datasets were deseasonalized by
 subtracting the corresponding climatological annual cycle (daily or
 monthly) from the actual data. Finally, to estimate the SZA effect on
-the SDR trends, the one-minute data were aggregated in \(1^\circ\) SZA
-bins, separately for the morning and afternoon hours, and then were
-deseasonalized as mentioned above.
+the SDR trends, the one-minute data were aggregated in \(1^{\circ}\) SZA
+bins, separately for the morning and afternoon hours.
+
+The linear trends were calculated using an autoregressive model
+\citet{Santer2000}
 
@@
 \caption{\label{tab:trendtable}Trends in SDR daily means for different sky conditions for the period 1993 - 2023.}
-\begin{tabu} to \linewidth {>{\centering\arraybackslash}p{8em}>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X>{\centering}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X>{\centering}X}
+\begin{tabu} to \linewidth {>{\centering\arraybackslash}p{8em}>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X>{\centering}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X}
 \toprule
-Trend [\%/year] & Rsqrd & Pearson cor. & Sky conditions & Days & t\_eff & t\_eff\_cri & conf\_2.5 & conf\_97.5 & mean\_clima & ChangeWpY\\
+Trend [\%/year] & Pearson cor. & Sky conditions & Days & t\_eff & t\_eff\_cri & conf\_2.5 & conf\_97.5 & mean\_clima & ChangeWpY\\
 \midrule
-0.3756 & 0.00830 & 0.0911 & All skies & 10250 & 4.727 & 1.96 & 0.000811 & 0.001246 & 387 & 1.452\\
-0.0972 & 0.01912 & 0.1383 & Clear skies & 2684 & 0.476 & 1.96 & 0.000194 & 0.000338 & 515 & 0.501\\
-0.4077 & 0.00657 & 0.0810 & Cloudy skies & 4937 & 3.566 & 1.96 & 0.000733 & 0.001499 & 290 & 1.181\\
+0.3756 & 0.0911 & All skies & 10250 & 4.727 & 1.96 & 0.000811 & 0.001246 & 387 & 1.452\\
+0.0972 & 0.1383 & Clear skies & 2684 & 0.476 & 1.96 & 0.000194 & 0.000338 & 515 & 0.501\\
+0.4077 & 0.0810 & Cloudy skies & 4937 & 3.566 & 1.96 & 0.000733 & 0.001499 & 290 & 1.181\\
 \bottomrule
@@
 \(0.7\,\%/y\). The trends in the morning and afternoon hours are more or
-less consistent with small differences, which can be attributed to
-systematic diurnal variations of aerosols, particularly during the warm
-period of the year \citep{Wang2021}.
+less consistent with small differences at small SZAs which can be
+attributed to effects on clear sky SDR from systematic diurnal patterns
+of aerosols during the warm period of the year \citep{Wang2021}. Note
+that SZAs less than \(35^\circ\) can only occur during the warm period
+of the year when clear skies are more frequent.
 
 For cloudy-sky conditions (Figure~\ref{fig:szatrends-3}), we can not
-discern any significant dependence of the SDR trend with SZA. For SZAs
-\(16^\circ\) - \(50^\circ\), the trends range within about
-\(\pm 0.2\,\%/y\), with a weak statistical significance. Between
+discern any significant dependence of the SDR trend with SZA as the
+variability of irradiance is dominated by the cloud effects. For SZAs
+\(25^\circ\) - \(75^\circ\), the trends range within about
+\(\pm 0.2\,\%/y\), the trends are variable and for most SZAs
+insignificant ranging between \(0\) and \(0.2%\,\%/y\). Between
 \(50^\circ\) and \(75^\circ\) SZA the trends for the period before noon
@@
 Table~\ref{tab:trendseasontable}). For all-sky conditions the trend in
-SDR in winter is the largest (\(0.69\,\%/y\)), followed by the trend in
-autumn (\(0.43\,\%/y\), a value close to the long term trend) both
+SDR in winter is the largest (\(0.7\,\%/y\)), followed by the trend in
+autumn (\(0.42\,\%/y\), a value close to the long term trend) both
 statistically significant above the \(99\,\%\) confidence level. In
@@
 the trends in SDR to trends of clouds during winter and autumn. For
-clear-skies, the trend in winter is \(0.36\,\%/y\), larger than for
-all-skies (\(0.69\,\%/y\)), which is another indication of a decreasing
+clear-skies, the trend in winter is \(0.4\,\%/y\), larger than for
+all-skies (\(0.7\,\%/y\)), which is another indication of a decreasing
 trend in cloud optical thickness. Moreover, the trends under clear- and
@@
 
+\begin{verbatim}
+        <!-- % \includegraphics[width=1\linewidth]{./images/SeasonalTrendsTogether3-2}  % Seasonal from Daily -->
+\end{verbatim}
+
 \begin{table}[!h]
@@
 \caption{\label{tab:trendseasontable}SDR linear trends of monthly anomalies for each season of the year.}
-\begin{tabu} to \linewidth {>{\centering\arraybackslash}p{8em}>{\centering}X>{\raggedleft}X>{\raggedleft}X>{\centering}X>{\centering}X>{\raggedleft}X>{\raggedleft}X>{\centering}X>{\centering}X>{\raggedleft}X}
+\begin{tabu} to \linewidth {>{\centering\arraybackslash}p{8em}>{\centering}X>{\raggedleft}X>{\raggedleft}X>{\centering}X>{\centering}X>{\raggedleft}X>{\raggedleft}X>{\centering}X}
 \toprule
-Trend [\%/year] & slope.p & Rsqrd & cor.estimate & Sky condition & Season & N\_eff & t\_eff & t\_eff\_cri & conf\_2.5 & conf\_97.5\\
+Trend [\%/year] & slope.p & cor.estimate & Sky condition & Season & t\_eff & t\_eff\_cri & conf\_2.5 & conf\_97.5\\
 \midrule
-\cellcolor{gray!6}{0.6870} & \cellcolor{gray!6}{0.001330} & \cellcolor{gray!6}{0.31200} & \cellcolor{gray!6}{0.5587342} & \cellcolor{gray!6}{All skies} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{24.08252} & \cellcolor{gray!6}{1e-07} & \cellcolor{gray!6}{2.063524} & \cellcolor{gray!6}{-0.0011} & \cellcolor{gray!6}{0.0015253}\\
-\cmidrule{1-11}
-0.1360 & 0.219000 & 0.05150 & 0.2270307 & All skies & Spring & 44.44554 & 1e-07 & 2.014797 & -0.0011 & 0.0015253\\
-\cmidrule{1-11}
-\cellcolor{gray!6}{0.1230} & \cellcolor{gray!6}{0.096100} & \cellcolor{gray!6}{0.09580} & \cellcolor{gray!6}{0.3094390} & \cellcolor{gray!6}{All skies} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{23.37502} & \cellcolor{gray!6}{0e+00} & \cellcolor{gray!6}{2.066823} & \cellcolor{gray!6}{-0.0011} & \cellcolor{gray!6}{0.0015253}\\
-\cmidrule{1-11}
-0.4300 & 0.006670 & 0.23500 & 0.4844650 & All skies & Autumn & 21.43236 & 1e-07 & 2.077062 & -0.0011 & 0.0015253\\
-\cmidrule{1-11}
-\cellcolor{gray!6}{0.3570} & \cellcolor{gray!6}{0.007260} & \cellcolor{gray!6}{0.23000} & \cellcolor{gray!6}{0.4800278} & \cellcolor{gray!6}{Clear skies} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{18.55573} & \cellcolor{gray!6}{0e+00} & \cellcolor{gray!6}{2.096421} & \cellcolor{gray!6}{-0.0011} & \cellcolor{gray!6}{0.0015253}\\
-\cmidrule{1-11}
-0.0317 & 0.721000 & 0.00445 & 0.0667338 & Clear skies & Spring & 21.33345 & 0e+00 & 2.077636 & -0.0011 & 0.0015253\\
-\cmidrule{1-11}
-\cellcolor{gray!6}{-0.0931} & \cellcolor{gray!6}{0.007770} & \cellcolor{gray!6}{0.22700} & \cellcolor{gray!6}{-0.4764646} & \cellcolor{gray!6}{Clear skies} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{25.12298} & \cellcolor{gray!6}{0e+00} & \cellcolor{gray!6}{2.059027} & \cellcolor{gray!6}{-0.0011} & \cellcolor{gray!6}{0.0015253}\\
-\cmidrule{1-11}
-0.1100 & 0.308000 & 0.03710 & 0.1926085 & Clear skies & Autumn & 23.96832 & 0e+00 & 2.064043 & -0.0011 & 0.0015253\\
-\cmidrule{1-11}
-\cellcolor{gray!6}{0.8250} & \cellcolor{gray!6}{0.000123} & \cellcolor{gray!6}{0.41500} & \cellcolor{gray!6}{0.6439160} & \cellcolor{gray!6}{Cloudy skies} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{14.68318} & \cellcolor{gray!6}{1e-07} & \cellcolor{gray!6}{2.135462} & \cellcolor{gray!6}{-0.0011} & \cellcolor{gray!6}{0.0015253}\\
-\cmidrule{1-11}
-0.0942 & 0.425000 & 0.02210 & 0.1485860 & Cloudy skies & Spring & 35.40650 & 1e-07 & 2.029275 & -0.0011 & 0.0015253\\
-\cmidrule{1-11}
-\cellcolor{gray!6}{-0.0983} & \cellcolor{gray!6}{0.414000} & \cellcolor{gray!6}{0.02400} & \cellcolor{gray!6}{-0.1548507} & \cellcolor{gray!6}{Cloudy skies} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{37.66778} & \cellcolor{gray!6}{1e-07} & \cellcolor{gray!6}{2.024981} & \cellcolor{gray!6}{-0.0011} & \cellcolor{gray!6}{0.0015253}\\
-\cmidrule{1-11}
-0.1950 & 0.460000 & 0.01960 & 0.1401398 & Cloudy skies & Autumn & 28.41812 & 2e-07 & 2.047050 & -0.0011 & 0.0015253\\
+\cellcolor{gray!6}{0.7000} & \cellcolor{gray!6}{0.002770} & \cellcolor{gray!6}{0.5353286} & \cellcolor{gray!6}{All skies} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{2e-07} & \cellcolor{gray!6}{2.060774} & \cellcolor{gray!6}{-0.0011} & \cellcolor{gray!6}{0.0015253}\\
+\cmidrule{1-9}
+0.1070 & 0.371000 & 0.1663281 & All skies & Spring & 1e-07 & 2.011940 & -0.0011 & 0.0015253\\
+\cmidrule{1-9}
+\cellcolor{gray!6}{0.1050} & \cellcolor{gray!6}{0.175000} & \cellcolor{gray!6}{0.2543074} & \cellcolor{gray!6}{All skies} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{0e+00} & \cellcolor{gray!6}{2.050358} & \cellcolor{gray!6}{-0.0011} & \cellcolor{gray!6}{0.0015253}\\
+\cmidrule{1-9}
+0.4230 & 0.009390 & 0.4663313 & All skies & Autumn & 1e-07 & 2.074143 & -0.0011 & 0.0015253\\
+\cmidrule{1-9}
+\cellcolor{gray!6}{0.3950} & \cellcolor{gray!6}{0.000515} & \cellcolor{gray!6}{0.5956771} & \cellcolor{gray!6}{Clear skies} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{0e+00} & \cellcolor{gray!6}{2.116842} & \cellcolor{gray!6}{-0.0011} & \cellcolor{gray!6}{0.0015253}\\
+\cmidrule{1-9}
+0.0591 & 0.497000 & 0.1267072 & Clear skies & Spring & 0e+00 & 2.069495 & -0.0011 & 0.0015253\\
+\cmidrule{1-9}
+\cellcolor{gray!6}{-0.0540} & \cellcolor{gray!6}{0.106000} & \cellcolor{gray!6}{-0.3013716} & \cellcolor{gray!6}{Clear skies} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{0e+00} & \cellcolor{gray!6}{2.035625} & \cellcolor{gray!6}{-0.0011} & \cellcolor{gray!6}{0.0015253}\\
+\cmidrule{1-9}
+0.0539 & 0.366000 & 0.1712676 & Clear skies & Autumn & 0e+00 & 2.119437 & -0.0011 & 0.0015253\\
+\cmidrule{1-9}
+\cellcolor{gray!6}{0.7590} & \cellcolor{gray!6}{0.000665} & \cellcolor{gray!6}{0.5861420} & \cellcolor{gray!6}{Cloudy skies} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{1e-07} & \cellcolor{gray!6}{2.077321} & \cellcolor{gray!6}{-0.0011} & \cellcolor{gray!6}{0.0015253}\\
+\cmidrule{1-9}
+0.0621 & 0.593000 & 0.0997924 & Cloudy skies & Spring & 1e-07 & 2.026109 & -0.0011 & 0.0015253\\
+\cmidrule{1-9}
+\cellcolor{gray!6}{-0.0789} & \cellcolor{gray!6}{0.560000} & \cellcolor{gray!6}{-0.1109085} & \cellcolor{gray!6}{Cloudy skies} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{1e-07} & \cellcolor{gray!6}{2.013800} & \cellcolor{gray!6}{-0.0011} & \cellcolor{gray!6}{0.0015253}\\
+\cmidrule{1-9}
+0.1910 & 0.384000 & 0.1648275 & Cloudy skies & Autumn & 1e-07 & 2.055579 & -0.0011 & 0.0015253\\
 \bottomrule
@@
 on season, with winter showing the strongest statistically significant
-trend of \(0.69\) and \(0.36\,\%/y\) for all- and clear-skies,
+trend of \(0.7\) and \(0.4\,\%/y\) for all- and clear-skies,
 respectively, in contrast to spring and summer. The trends for autumn
-are also significant but smaller ( \(0.43\) and \(0.11\,\%/y\) for all-
+are also significant but smaller ( \(0.42\) and \(0.054\,\%/y\) for all-
 and clear-skies, respectively). Our findings are in agreement with other

### 2023-12-18 DHI_GHI_3_trends_consistency.R

@@ -958,8 +958,8 @@ for (adb in database) {
         ## daily from other DT
         lines(pdb$Date, pdb[[paste0(avar,"_cusum")]], col = col, lwd = 2)
 
-        #test
-        points(pdb$Date, pdb[[paste0(avar,"_cusum")]], col = "red")
+        # #test
+        # points(pdb$Date, pdb[[paste0(avar,"_cusum")]], col = "red")
 
 
         if (DRAFT == T) {

### 2023-12-18 MDPI_submission.Rmd
Commit 2023-12-18 12:14

@@
 $0$ and $0.2%\,\%/y$.
-Between $50^\circ$ and $75^\circ$ SZA the trends for the period before noon are stronger and negative, possibly associated with stronger attenuation by clouds under oblique incidence angles.
+The smooth increase of the trend at SZAs larger than
+$\sim 75^{\circ}$ is probably associated with stronger attenuation by
+clouds under oblique incidence angles which results in smaller
+variability.
 

### 2023-12-18 MDPI_submission.tex
Commit 2023-12-18 12:14

@@
 \(\pm 0.2\,\%/y\), the trends are variable and for most SZAs
-insignificant ranging between \(0\) and \(0.2%\,\%/y\). Between
-\(50^\circ\) and \(75^\circ\) SZA the trends for the period before noon
-are stronger and negative, possibly associated with stronger attenuation
-by clouds under oblique incidence angles.
+insignificant ranging between \(0\) and \(0.2%\,\%/y\). The smooth
+increase of the trend at SZAs larger than \(\sim 75^{\circ}\) is
+probably associated with stronger attenuation by clouds under oblique
+incidence angles which results in smaller variability.
 

### 2023-12-18 DHI_GHI_1_longterm_trends.R
Commit 2023-12-18 13:10

@@ -274,7 +274,6 @@ vars <- c( "GLB_att_des")
 
 dbs         <- c("CLEAR_1_daily_DESEAS",
                  "ALL_1_daily_DESEAS",
-
                  "CLOUD_1_daily_DESEAS")
 ## gather trends
 gather <- data.frame()
@@ -292,7 +291,7 @@ for (DBn in dbs) {
             ## linear model by day step
             lm1 <- lm(dataset[[avar]] ~ dataset$Date)
             d   <- summary(lm1)$coefficients
-            cat("lm:      ", lm1$coefficients[2] * Days_of_year, "+/-", d[2,2] * Days_of_year,"\n")
+            cat("lm:      ", lm1$coefficients[2] * Days_of_year, "+/-", d[2,2] * Days_of_year,"\n\n")
             ## correlation test
             cor1 <- cor.test(x = dataset[[avar]], y = as.numeric(dataset$Date), method = 'pearson')
 
@@ -304,36 +303,28 @@ for (DBn in dbs) {
 
             ## _ Arima tests  --------------------------------------------------
 
-            amodelo  <- arima(dd[, avar], order = c(1,0,0), method = "ML")
-            amod_test <- lmtest::coeftest(amodelo)
-            cat("My arima:", round(amod_test[1,], 4), "\n")
+            # amodelo   <- arima(dd[, avar], order = c(1, 0, 0), method = "ML")
+            # amod_test <- lmtest::coeftest(amodelo)
+            # cat("My arima:", round(amod_test[1,], 4), "\n")
+
+            ## _ auto reggression arima Tourpali -------------------------------
+            ## create a time variable (with lag of 1 day ?)
+            dataset[, ts := (year(Date) - min(year(Date))) + ( doy - 1 ) / Hmisc::yearDays(Date) ]
+            tmodelo <- arima(x = dataset[[avar]], order = c(1,0,0), xreg = dataset$ts, method = "ML")
+
+            ## trend per year with auto correlation
+            Tres <- data.frame(t(lmtest::coeftest(tmodelo)[3,]))
+            Tint <- data.frame(t(lmtest::coeftest(tmodelo)[2,]))
+            names(Tres) <- paste0("Tmod_", names(Tres))
+
+            cat("Tourpali:", paste(round(Tres, 4)), "\n\n")
 
 
-            ## tourpal
[…]

### 2023-12-18 MDPI_submission.Rmd
Commit 2023-12-18 19:10

--- 2023-12-18 MDPI_submission.Rmd
@@
 
-The linear trends were calculated using an autoregressive model
-<!-- ([ARIMA......)]{.mark} and were reported together with the 2σ errors. -->
+The linear trends were calculated using an first order autoregressive model
+(arima) with lag-1, and were reported together with the $2\sigma$ errors (Table\ \ref{tab:trendtable}).
+
 <!-- [For the trends derived from monthly mean data we used an effective -->
@@
 ## numeric precision of table
-LT$slope    <- signif(LT$slope,   4)
-LT$slope.p  <- signif(LT$slope.p, 4)
-LT$Rsqrd    <- signif(LT$Rsqrd,   4)
+LT$slope         <- signif(LT$slope,         3)
+LT$slope.p       <- signif(LT$slope.p,       3)
+LT$Rsqrd         <- signif(LT$Rsqrd,         3)
+LT$Tmod_Estimate <- signif(LT$Tmod_Estimate, 3)
+LT$cor.estimate  <- signif(LT$cor.estimate,  3)
+LT$Tmod_Std..Error <- signif(LT$Tmod_Std..Error * 2, 3)
+LT$ChangeWpY       <- signif(LT$ChangeWpY, 3)
 
@@
 
-LT <- rename(LT, "Trend [%/year]" = slope       )
+# LT <- rename(LT, "Trend [%/year]" = slope        )
 LT <- rename(LT, "Sky conditions"  = DATA         )
 LT <- rename(LT, "Days"            = N            )
-LT <- rename(LT, "Pearson cor."   = cor.estimate)
+LT <- rename(LT, "Pearson cor."    = cor.estimate )
 LT <- rename(LT, "Trend p-value"   = slope.p      )
+LT <- rename(LT, "Trend [%/year]"  = Tmod_Estimate) # Arima slope
+LT <- rename(LT, "Trend S.E."      = Tmod_Std..Error )
+LT <- rename(LT, "Trend [W/year]"  = ChangeWpY )
 
@@
 LT$N_eff           <- NULL
-
-# LT$`Trend statistical signif. [%]` <- NULL
+LT$slope           <- NULL
+LT$t_eff           <- NULL
+LT$t_eff_cri       <- NULL
+LT$conf_2.5        <- NULL
+LT$conf_97.5       <- NULL
+LT$mean_clima      <- NULL
+LT$Tmod_z.value    <- NULL
+LT$Tmod_Pr...z..   <- NULL
 
 ## order columns
-# LT1 <- LT[,c(
-#       "Sky conditions",
-#       "Trend [%/year]",
-#       "Trend p-value",
-#       "Pearson cor.",
-#       "Days"
-#       )]
-LT1 <- LT
+LT1 <- LT[,c(
+      "Sky conditions",
+      "Trend [%/year]",
+      "Trend S.E.",
+      "Pearson cor.",
+      "Trend [W/year]",
+      "Days"
+      )]
+# LT1 <- LT
 
@@
 
+# names(LT1)
+# "Sky conditions", "Trend [%/year]", "Trend S.E.", "Pearson cor.", "Trend [W/year]", "Days" 
 x_latex <- knitr::kable(LT1,
@@
                         booktabs    = TRUE,
-                        escape      = TRUE,
+                        escape      = FALSE,
+                        col.names   = c("Sky conditions", "Trend [\\%/year]", "Trend S.E. ($2\\sigma$)", "Pearson cor.", "Trend [W/year]", "Days"),
                         format.args = list(digits = 3),
                         linesep     = "",
-                        align       = "crrrr",
+                        align       = "crrrrr",
                         centering   = FALSE,
@@
 dominated by the cloud effects.
-For SZAs $25^\circ$ - $75^\circ$, the trends range within about $\pm 0.2\,\%/y$, 
-the trends are variable and for most SZAs insignificant ranging between
-$0$ and $0.2%\,\%/y$.
-Between $50^\circ$ and $75^\circ$ SZA the trends for the period before noon are stronger and negative, possibly associated with stronger attenuation by clouds under oblique incidence angles.
+For SZAs $25^\circ$ -- $75^\circ$, the trends range within about $\pm 0.2\,\%/y$,
+the trends are variable and for most SZAs insignificant ranging between $0$ and $0.2\,\%/y$.
+The smooth increase of the trend at SZAs larger than $75^\circ$ is probably associated with stronger attenuation by
+clouds under oblique incidence angles which results in smaller
+variability.
 

### 2023-12-18 MDPI_submission.tex
Commit 2023-12-18 19:10

--- 2023-12-18 MDPI_submission.tex
@@
 
-The linear trends were calculated using an autoregressive model
+The linear trends were calculated using an first order autoregressive
+model (arima) with lag-1, and were reported together with the
+\(2\sigma\) errors (Table~\ref{tab:trendtable}).
+
 \citet{Santer2000}
@@
 \caption{\label{tab:trendtable}Trends in SDR daily means for different sky conditions for the period 1993 - 2023.}
-\begin{tabu} to \linewidth {>{\centering\arraybackslash}p{8em}>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X>{\centering}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X}
+\begin{tabu} to \linewidth {>{\centering\arraybackslash}p{8em}>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X}
 \toprule
-Trend [\%/year] & Pearson cor. & Sky conditions & Days & t\_eff & t\_eff\_cri & conf\_2.5 & conf\_97.5 & mean\_clima & ChangeWpY\\
+Sky conditions & Trend [\%/year] & Trend S.E. ($2\sigma$) & Pearson cor. & Trend [W/year] & Days\\
 \midrule
-0.3756 & 0.0911 & All skies & 10250 & 4.727 & 1.96 & 0.000811 & 0.001246 & 387 & 1.452\\
-0.0972 & 0.1383 & Clear skies & 2684 & 0.476 & 1.96 & 0.000194 & 0.000338 & 515 & 0.501\\
-0.4077 & 0.0810 & Cloudy skies & 4937 & 3.566 & 1.96 & 0.000733 & 0.001499 & 290 & 1.181\\
+All skies & 0.3760 & 0.1210 & 0.0911 & 1.450 & 10250\\
+Clear skies & 0.0972 & 0.0327 & 0.1380 & 0.501 & 2684\\
+Cloudy skies & 0.4080 & 0.1770 & 0.0810 & 1.180 & 4937\\
 \bottomrule
@@
 variability of irradiance is dominated by the cloud effects. For SZAs
-\(25^\circ\) - \(75^\circ\), the trends range within about
+\(25^\circ\) -- \(75^\circ\), the trends range within about
 \(\pm 0.2\,\%/y\), the trends are variable and for most SZAs
-insignificant ranging between \(0\) and \(0.2%\,\%/y\). Between
-\(50^\circ\) and \(75^\circ\) SZA the trends for the period before noon
-are stronger and negative, possibly associated with stronger attenuation
-by clouds under oblique incidence angles.
+insignificant ranging between \(0\) and \(0.2\,\%/y\). The smooth
+increase of the trend at SZAs larger than \(75^\circ\) is probably
+associated with stronger attenuation by clouds under oblique incidence
+angles which results in smaller variability.
 
@@
 \end{figure}
-
-\begin{verbatim}
-        <!-- % \includegraphics[width=1\linewidth]{./images/SeasonalTrendsTogether3-2}  % Seasonal from Daily -->
-\end{verbatim}
 

### 2023-12-18 MDPI_submission.Rmd
Commit 2023-12-18 20:14

@@
 alltrend   <- signif(LT[DATA == "All sky cond.",    slope], 2)
-cleartrend <- signif(LT[DATA == "Clear sky cond.",  slope], 2)
+cleartrend <- signif(LT[DATA == "Clear sky cond.",  slope], 1)
 cloudtrend <- signif(LT[DATA == "Cloudy sky cond.", slope], 2)
@@
 series.
-We observe a positive trend for all-sky conditions
-($`r alltrend`\,\%/y$), a very close but smaller trend for clear-skies ($`r cleartrend`\,\%/y$) and a negative weaker trend for cloudy-skies ($`r cloudtrend`\,\%/y$).
-In the studied period, there is no significant break or change in the variability of the time series.
-Other studies for the European region reported a change of the SDR slope, around 1980 [@Wild2021; @Yuan2021; @Ohmura2009], a few years before the start of our records.
-It is interesting to note, that for the observations period, the trend of the TSI is
-$`r format(TSIslopeAll[1,1], scientific = F, digits = 2)`\,\%/y$,
-and thus we can not attribute any major effect on SDR trend to Solar variability.
-
-<!-- \footnotesize -->
-<!-- ```{r trendtable2, echo=F} -->
-<!-- panderOptions("table.alignment.default", "center") -->
-<!-- pander(LT1, -->
-<!--        caption = paste0("\\label{tab:trendtable2}Trends in SDR daily mean for different sky conditions for the period", year(min(ALL_1_daily_DESEAS$Date)), " - ", year(max(ALL_1_daily_DESEAS$Date)), ".
-For all cases the statistical significance is above $", LTminSSig, "\\,\\%$.")) -->
-<!-- ``` -->
-<!-- \normalsize -->
+The linear trends in all three datasets are positive and around
+$0.4\,\%/y$ for all-sky and cloudy-sky conditions, while for clear-skies the
+trend is much smaller (\~$`r cloudtrend`\,\%/y$). 
+The linear trends were calculated
+taking into account the autocorrelation of the time series and all three
+are statistically significant at least at the $95\,\%$ confidence level, as
+they are larger than the corresponding $2\sigma$ errors.
+The clear-sky trend is
+very small suggesting a small effect from aerosols and water vapor which
+are the dominant factors of the SDR variability. 
+In contrast the large
+positive trend of SDR under cloudy skies can be attributed to reduction
+in cloud cover and/or optical depth. 
+Lack of continuous cloud
+observations that could support these findings does not allow drawing
+firm conclusions. However, there are indications that the total
+cloud-cover as inferred from the ERA5 analysis for the grid point of
+Thessaloniki is decreasing over the period of study.
+
+<!-- ($`r alltrend`\,\%/y$), a very close but smaller trend for clear-skies ($`r cleartrend`\,\%/y$) and a negative weaker trend for cloudy-skies ($`r cloudtrend`\,\%/y$). -->
+The all-sky trend is similar to the one reported in @Bais2013
+from a ten-year shorter dataset suggesting that the tendency of SDR in
+Thessaloniki is systematic.
+Other studies for the European region reported a change in the SDR trend
+around 1980 from negative to positive
+with comparable magnitude [@Wild2021; @Yuan2021; @Ohmura2009], 
+well before the start of our records.
+However, the
+trends reported here for the three datasets are in accordance with the
+widely accepted solar radiation brightening over Europe.
+For the period
+of our observations the trend in the TSI is negligible
+($`r format(TSIslopeAll[1,1], scientific = F, digits = 2)`\,\%/y$),
+and thus we cannot attribute any significant effect
+on the SDR trend to solar variability.
 
 ```{r trendtable, echo=F}
-
 # names(LT1)
@@
 dominated by the cloud effects.
-For SZAs $25^\circ$ -- $75^\circ$, the trends range within about $\pm 0.2\,\%/y$,
+For SZAs $25^\circ$ -- $75^\circ$,
 the trends are variable and for most SZAs insignificant ranging between $0$ and $0.2\,\%/y$.
-The smooth increase of the trend at SZAs larger than $75^\circ$ is probably associated with stronger attenuation by
+The smooth increase of the trend at SZAs larger than $\sim 75^\circ$ is probably associated with stronger attenuation by
 clouds under oblique incidence angles which results in smaller
@@
 spp$Rsqrd          <- signif(spp$Rsqrd,          3)
+spp$cor.estimate   <- signif(spp$cor.estimate,   3)
 spp$RsqrdAdj       <- NULL
@@
 spp$N_eff          <- NULL
+spp$conf_2.5       <- NULL
+spp$conf_97.5      <- NULL
+spp$t_eff          <- NULL
+spp$t_eff_cri      <- NULL
 
@@
 
-spp <- rename(spp, "Trend [%/year]" = slope)
+spp <- rename(spp, "Trend [%/year]" = slope       )
 spp <- rename(spp, "Sky condition"  = DATA        )
-
-# names(spp) <- c("Trend [%/year]", "Trend p-value", "Sky condition", "Season", "Statistical signif. [%]")
+spp <- rename(spp, "Pearson cor."   = cor.estimate)
+
+## arrange columns
+spp <- spp[,c(
+      "Sky condition",
+      "Season",
+      "Trend [%/year]",
+      "Pearson cor.",
+      "Trend p-value"
+      )]
 
@@
 
-Similarly to the long term trends from daily means of SDR discussed above, we have calculated the trend of the anomalies for the three different sky conditions, and for each season of the year, using the corresponding mean monthly values
+Similarly to the long term trends from daily means of SDR discussed above, we have calculated the trend for the three sky conditions, and for each season of the year, using the corresponding mean monthly anomalies
 (Figure\ \ref{fig:seasonalALL} and Table\ \ref{tab:trendseasontable}).
@@
 
-## arrange columns
-# spp <- spp[,c(
-#       "Sky condition",
-#       "Season",
-#       "Trend [%/year]",
-#       # "Trend p-value",
-#       "Statistical signif. [%]"
-#       )]
-
 spp2 <- spp
@@
                         linesep     = "",
-                        align       = "ccrr",
+                        align       = "ccrrr",
                         centering   = FALSE,

### 2023-12-18 MDPI_submission.tex
Commit 2023-12-18 20:14

@@
 period, there is no significant break or change in the variability
-pattern of the time series. We observe a positive trend for all-sky
-conditions (\(0.38\,\%/y\)), a very close but smaller trend for
-clear-skies (\(0.097\,\%/y\)) and a negative weaker trend for
-cloudy-skies (\(0.41\,\%/y\)). In the studied period, there is no
-significant break or change in the variability of the time series. Other
-studies for the European region reported a change of the SDR slope,
-around 1980 \citep{Wild2021, Yuan2021, Ohmura2009}, a few years before
-the start of our records. It is interesting to note, that for the
-observations period, the trend of the TSI is \(-0.00024\,\%/y\), and
-thus we can not attribute any major effect on SDR trend to Solar
+pattern of the time series. The linear trends in all three datasets are
+positive and around \(0.4\,\%/y\) for all-sky and cloudy-sky conditions,
+while for clear-skies the trend is much smaller
+(\textasciitilde{}\(0.41\,\%/y\)). The linear trends were calculated
+taking into account the autocorrelation of the time series and all three
+are statistically significant at least at the \(95\,\%\) confidence
+level, as they are larger than the corresponding \(2\sigma\) errors. The
+clear-sky trend is very small suggesting a small effect from aerosols
+and water vapor which are the dominant factors of the SDR variability.
+In contrast the large positive trend of SDR under cloudy skies can be
+attributed to reduction in cloud cover and/or optical depth. Lack of
+continuous cloud observations that could support these findings does not
+allow drawing firm conclusions. However, there are indications that the
+total cloud-cover as inferred from the ERA5 analysis for the grid point
+of Thessaloniki is decreasing over the period of study.
+
+The all-sky trend is similar to the one reported in \citet{Bais2013}
+from a ten-year shorter dataset suggesting that the tendency of SDR in
+Thessaloniki is systematic. Other studies for the European region
+reported a change in the SDR trend around 1980 from negative to positive
+with comparable magnitude \citep{Wild2021, Yuan2021, Ohmura2009}, well
+before the start of our records. However, the trends reported here for
+the three datasets are in accordance with the widely accepted solar
+radiation brightening over Europe. For the period of our observations
+the trend in the TSI is negligible (\(-0.00024\,\%/y\)), and thus we
+cannot attribute any significant effect on the SDR trend to solar
 variability.
@@
 variability of irradiance is dominated by the cloud effects. For SZAs
-\(25^\circ\) -- \(75^\circ\), the trends range within about
-\(\pm 0.2\,\%/y\), the trends are variable and for most SZAs
+\(25^\circ\) -- \(75^\circ\), the trends are variable and for most SZAs
 insignificant ranging between \(0\) and \(0.2\,\%/y\). The smooth
-increase of the trend at SZAs larger than \(75^\circ\) is probably
+increase of the trend at SZAs larger than \(\sim 75^\circ\) is probably
 associated with stronger attenuation by clouds under oblique incidence
@@
 Similarly to the long term trends from daily means of SDR discussed
-above, we have calculated the trend of the anomalies for the three
-different sky conditions, and for each season of the year, using the
-corresponding mean monthly values (Figure~\ref{fig:seasonalALL} and
+above, we have calculated the trend for the three sky conditions, and
+for each season of the year, using the corresponding mean monthly
+anomalies (Figure~\ref{fig:seasonalALL} and
 Table~\ref{tab:trendseasontable}). For all-sky conditions the trend in
@@
 \caption{\label{tab:trendseasontable}SDR linear trends of monthly anomalies for each season of the year.}
-\begin{tabu} to \linewidth {>{\centering\arraybackslash}p{8em}>{\centering}X>{\raggedleft}X>{\raggedleft}X>{\centering}X>{\centering}X>{\raggedleft}X>{\raggedleft}X>{\centering}X}
+\begin{tabu} to \linewidth {>{\centering\arraybackslash}p{8em}>{\centering}X>{\raggedleft}X>{\raggedleft}X>{\centering}X}
 \toprule
-Trend [\%/year] & slope.p & cor.estimate & Sky condition & Season & t\_eff & t\_eff\_cri & conf\_2.5 & conf\_97.5\\
+Trend [\%/year] & slope.p & Pearson cor. & Sky condition & Season\\
 \midrule
-\cellcolor{gray!6}{0.7000} & \cellcolor{gray!6}{0.002770} & \cellcolor{gray!6}{0.5353286} & \cellcolor{gray!6}{All skies} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{2e-07} & \cellcolor{gray!6}{2.060774} & \cellcolor{gray!6}{-0.0011} & \cellcolor{gray!6}{0.0015253}\\
-\cmidrule{1-9}
-0.1070 & 0.371000 & 0.1663281 & All skies & Spring & 1e-07 & 2.011940 & -0.0011 & 0.0015253\\
-\cmidrule{1-9}
-\cellcolor{gray!6}{0.1050} & \cellcolor{gray!6}{0.175000} & \cellcolor{gray!6}{0.2543074} & \cellcolor{gray!6}{All skies} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{0e+00} & \cellcolor{gray!6}{2.050358} & \cellcolor{gray!6}{-0.0011} & \cellcolor{gray!6}{0.0015253}\\
-\cmidrule{1-9}
-0.4230 & 0.009390 & 0.4663313 & All skies & Autumn & 1e-07 & 2.074143 & -0.0011 & 0.0015253\\
-\cmidrule{1-9}
-\cellcolor{gray!6}{0.3950} & \cellcolor{gray!6}{0.000515} & \cellcolor{gray!6}{0.5956771} & \cellcolor{gray!6}{Clear skies} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{0e+00} & \cellcolor{gray!6}{2.116842} & \cellcolor{gray!6}{-0.0011} & \cellcolor{gray!6}{0.0015253}\\
-\cmidrule{1-9}
-0.0591 & 0.497000 & 0.1267072 & Clear skies & Spring & 0e+00 & 2.069495 & -0.0011 & 0.0015253\\
-\cmidrule{1-9}
-\cellcolor{gray!6}{-0.0540} & \cellcolor{gray!6}{0.106000} & \cellcolor{gray!6}{-0.3013716} & \cellcolor{gray!6}{Clear skies} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{0e+00} & \cellcolor{gray!6}{2.035625} & \cellcolor{gray!6}{-0.0011} & \cellcolor{gray!6}{0.0015253}\\
-\cmidrule{1-9}
-0.0539 & 0.366000 & 0.1712676 & Clear skies & Autumn & 0e+00 & 2.119437 & -0.0011 & 0.0015253\\
-\cmidrule{1-9}
-\cellcolor{gray!6}{0.7590} & \cellcolor{gray!6}{0.000665} & \cellcolor{gray!6}{0.5861420} & \cellcolor{gray!6}{Cloudy skies} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{1e-07} & \cellcolor{gray!6}{2.077321} & \cellcolor{gray!6}{-0.0011} & \cellcolor{gray!6}{0.0015253}\\
-\cmidrule{1-9}
-0.0621 & 0.593000 & 0.0997924 & Cloudy skies & Spring & 1e-07 & 2.026109 & -0.0011 & 0.0015253\\
-\cmidrule{1-9}
-\cellcolor{gray!6}{-0.0789} & \cellcolor{gray!6}{0.560000} & \cellcolor{gray!6}{-0.1109085} & \cellcolor{gray!6}{Cloudy skies} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{1e-07} & \cellcolor{gray!6}{2.013800} & \cellcolor{gray!6}{-0.0011} & \cellcolor{gray!6}{0.0015253}\\
-\cmidrule{1-9}
-0.1910 & 0.384000 & 0.1648275 & Cloudy skies & Autumn & 1e-07 & 2.055579 & -0.0011 & 0.0015253\\
+\cellcolor{gray!6}{0.7000} & \cellcolor{gray!6}{0.002770} & \cellcolor{gray!6}{0.5350} & \cellcolor{gray!6}{All skies} & \cellcolor{gray!6}{Winter}\\
+\cmidrule{1-5}
+0.1070 & 0.371000 & 0.1660 & All skies & Spring\\
+\cmidrule{1-5}
+\cellcolor{gray!6}{0.1050} & \cellcolor{gray!6}{0.175000} & \cellcolor{gray!6}{0.2540} & \cellcolor{gray!6}{All skies} & \cellcolor{gray!6}{Summer}\\
+\cmidrule{1-5}
+0.4230 & 0.009390 & 0.4660 & All skies & Autumn\\
+\cmidrule{1-5}
+\cellcolor{gray!6}{0.3950} & \cellcolor{gray!6}{0.000515} & \cellcolor{gray!6}{0.5960} & \cellcolor{gray!6}{Clear skies} & \cellcolor{gray!6}{Winter}\\
+\cmidrule{1-5}
+0.0591 & 0.497000 & 0.1270 & Clear skies & Spring\\
+\cmidrule{1-5}
+\cellcolor{gray!6}{-0.0540} & \cellcolor{gray!6}{0.106000} & \cellcolor{gray!6}{-0.3010} & \cellcolor{gray!6}{Clear skies} & \cellcolor{gray!6}{Summer}\\
+\cmidrule{1-5}
+0.0539 & 0.366000 & 0.1710 & Clear skies & Autumn\\
+\cmidrule{1-5}
+\cellcolor{gray!6}{0.7590} & \cellcolor{gray!6}{0.000665} & \cellcolor{gray!6}{0.5860} & \cellcolor{gray!6}{Cloudy skies} & \cellcolor{gray!6}{Winter}\\
+\cmidrule{1-5}
+0.0621 & 0.593000 & 0.0998 & Cloudy skies & Spring\\
+\cmidrule{1-5}
+\cellcolor{gray!6}{-0.0789} & \cellcolor{gray!6}{0.560000} & \cellcolor{gray!6}{-0.1110} & \cellcolor{gray!6}{Cloudy skies} & \cellcolor{gray!6}{Summer}\\
+\cmidrule{1-5}
+0.1910 & 0.384000 & 0.1650 & Cloudy skies & Autumn\\
 \bottomrule
@@
 over the area. Moreover, we found a similar trend under clear-sky
-conditions (\(0.097\,\%/y\)) that further supports the assumption that
-the brightening is caused mainly by decreasing aerosols. Unfortunately,
-for the entire period there is no available data for the aerosols, in
-order to quantify their effect on SDR. However, \citet{Siomos2020} have
-shown that aerosol optical depth over Thessaloniki is decreasing
-constantly at least up to 2018. The attenuation of SDR by aerosols over
-Europe have been proposed as major factor by \citet{Wild2021}. The
-dimming effect on SDR under cloudy-sky conditions (\(0.41\,\%/y\)),
-suggests that cloud optical thickness is decreasing during this period.
-Because we have no adequate data to investigate the long term changes of
-cloud thickness in the region, we cannot verify if the negative SDR
-trend we observe under under cloudy-skies can be attributed solely to
-changes in clouds.
+conditions (\(0.1\,\%/y\)) that further supports the assumption that the
+brightening is caused mainly by decreasing aerosols. Unfortunately, for
+the entire period there is no available data for the aerosols, in order
+to quantify their effect on SDR. However, \citet{Siomos2020} have shown
+that aerosol optical depth over Thessaloniki is decreasing constantly at
+least up to 2018. The attenuation of SDR by aerosols over Europe have
+been proposed as major factor by \citet{Wild2021}. The dimming effect on
+SDR under cloudy-sky conditions (\(0.41\,\%/y\)), suggests that cloud
+optical thickness is decreasing during this period. Because we have no
+adequate data to investigate the long term changes of cloud thickness in
+the region, we cannot verify if the negative SDR trend we observe under
+under cloudy-skies can be attributed solely to changes in clouds.
 

### 2023-12-18 MDPI_submission.Rmd
Commit 2023-12-18 22:14

@@
 spp <- rename(spp, "Trend [%/year]" = slope       )
+spp <- rename(spp, "Trend p-value"  = slope.p     )
 spp <- rename(spp, "Sky condition"  = DATA        )
@@
 
-Similarly to the long term trends from daily means of SDR discussed above, we have calculated the trend for the three sky conditions, and for each season of the year, using the corresponding mean monthly anomalies
+Similarly to the long term trends from daily means of SDR discussed above, we have calculated the trend for the three sky conditions and for each season of the year, using the corresponding mean monthly anomalies
 (Figure\ \ref{fig:seasonalALL} and Table\ \ref{tab:trendseasontable}).
@@
 ($`r signif(spp$"Trend [%/year]"[spp$Season == "Autumn" & spp$"Sky condition" == "All skies"], 2)`\,\%/y$,
-a value close to the long term trend) both statistically significant above the $99\,\%$ confidence level.
+a value close to the long term trend) both statistically significant at the $99\,\%$ confidence level.
 In spring and summer, the trends are much smaller and of lesser statistical significance.

### 2023-12-18 MDPI_submission.tex
Commit 2023-12-18 22:14

@@
 Similarly to the long term trends from daily means of SDR discussed
-above, we have calculated the trend for the three sky conditions, and
-for each season of the year, using the corresponding mean monthly
-anomalies (Figure~\ref{fig:seasonalALL} and
-Table~\ref{tab:trendseasontable}). For all-sky conditions the trend in
-SDR in winter is the largest (\(0.7\,\%/y\)), followed by the trend in
-autumn (\(0.42\,\%/y\), a value close to the long term trend) both
-statistically significant above the \(99\,\%\) confidence level. In
-spring and summer, the trends are much smaller and of lesser statistical
-significance. These seasonal differences indicate a possible relation of
-the trends in SDR to trends of clouds during winter and autumn. For
-clear-skies, the trend in winter is \(0.4\,\%/y\), larger than for
-all-skies (\(0.7\,\%/y\)), which is another indication of a decreasing
-trend in cloud optical thickness. Moreover, the trends under clear- and
-cloudy-sky conditions are almost complementary to each other,
-particularly for winter and autumn, where the signal is stronger. During
-spring and summer the statistical significance is very low and the
-actual trend too small for a meaningful comparison.
+above, we have calculated the trend for the three sky conditions and for
+each season of the year, using the corresponding mean monthly anomalies
+(Figure~\ref{fig:seasonalALL} and Table~\ref{tab:trendseasontable}). For
+all-sky conditions the trend in SDR in winter is the largest
+(\(0.7\,\%/y\)), followed by the trend in autumn (\(0.42\,\%/y\), a
+value close to the long term trend) both statistically significant at
+the \(99\,\%\) confidence level. In spring and summer, the trends are
+much smaller and of lesser statistical significance. These seasonal
+differences indicate a possible relation of the trends in SDR to trends
+of clouds during winter and autumn. For clear-skies, the trend in winter
+is \(0.4\,\%/y\), larger than for all-skies (\(0.7\,\%/y\)), which is
+another indication of a decreasing trend in cloud optical thickness.
+Moreover, the trends under clear- and cloudy-sky conditions are almost
+complementary to each other, particularly for winter and autumn, where
+the signal is stronger. During spring and summer the statistical
+significance is very low and the actual trend too small for a meaningful
+comparison.
 
@@
 \caption{\label{tab:trendseasontable}SDR linear trends of monthly anomalies for each season of the year.}
-\begin{tabu} to \linewidth {>{\centering\arraybackslash}p{8em}>{\centering}X>{\raggedleft}X>{\raggedleft}X>{\centering}X}
+\begin{tabu} to \linewidth {>{\centering\arraybackslash}p{8em}>{\centering}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X}
 \toprule
-Trend [\%/year] & slope.p & Pearson cor. & Sky condition & Season\\
+Sky condition & Season & Trend [\%/year] & Pearson cor. & Trend p-value\\
 \midrule
-\cellcolor{gray!6}{0.7000} & \cellcolor{gray!6}{0.002770} & \cellcolor{gray!6}{0.5350} & \cellcolor{gray!6}{All skies} & \cellcolor{gray!6}{Winter}\\
+\cellcolor{gray!6}{} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{0.7000} & \cellcolor{gray!6}{0.5350} & \cellcolor{gray!6}{0.002770}\\
+
+ & Spring & 0.1070 & 0.1660 & 0.371000\\
+
+\cellcolor{gray!6}{} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{0.1050} & \cellcolor{gray!6}{0.2540} & \cellcolor{gray!6}{0.175000}\\
+
+\multirow{-4}{*}{\centering\arraybackslash All skies} & Autumn & 0.4230 & 0.4660 & 0.009390\\
 \cmidrule{1-5}
-0.1070 & 0.371000 & 0.1660 & All skies & Spring\\
+\cellcolor{gray!6}{} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{0.3950} & \cellcolor{gray!6}{0.5960} & \cellcolor{gray!6}{0.000515}\\
+
+ & Spring & 0.0591 & 0.1270 & 0.497000\\
+
+\cellcolor{gray!6}{} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{-0.0540} & \cellcolor{gray!6}{-0.3010} & \cellcolor{gray!6}{0.106000}\\
+
+\multirow{-4}{*}{\centering\arraybackslash Clear skies} & Autumn & 0.0539 & 0.1710 & 0.366000\\
 \cmidrule{1-5}
-\cellcolor{gray!6}{0.1050} & \cellcolor{gray!6}{0.175000} & \cellcolor{gray!6}{0.2540} & \cellcolor{gray!6}{All skies} & \cellcolor{gray!6}{Summer}\\
-\cmidrule{1-5}
-0.4230 & 0.009390 & 0.4660 & All skies & Autumn\\
-\cmidrule{1-5}
-\cellcolor{gray!6}{0.3950} & \cellcolor{gray!6}{0.000515} & \cellcolor{gray!6}{0.5960} & \cellcolor{gray!6}{Clear skies} & \cellcolor{gray!6}{Winter}\\
-\cmidrule{1-5}
-0.0591 & 0.497000 & 0.1270 & Clear skies & Spring\\
-\cmidrule{1-5}
-\cellcolor{gray!6}{-0.0540} & \cellcolor{gray!6}{0.106000} & \cellcolor{gray!6}{-0.3010} & \cellcolor{gray!6}{Clear skies} & \cellcolor{gray!6}{Summer}\\
-\cmidrule{1-5}
-0.0539 & 0.366000 & 0.1710 & Clear skies & Autumn\\
-\cmidrule{1-5}
-\cellcolor{gray!6}{0.7590} & \cellcolor{gray!6}{0.000665} & \cellcolor{gray!6}{0.5860} & \cellcolor{gray!6}{Cloudy skies} & \cellcolor{gray!6}{Winter}\\
-\cmidrule{1-5}
-0.0621 & 0.593000 & 0.0998 & Cloudy skies & Spring\\
-\cmidrule{1-5}
-\cellcolor{gray!6}{-0.0789} & \cellcolor{gray!6}{0.560000} & \cellcolor{gray!6}{-0.1110} & \cellcolor{gray!6}{Cloudy skies} & \cellcolor{gray!6}{Summer}\\
-\cmidrule{1-5}
-0.1910 & 0.384000 & 0.1650 & Cloudy skies & Autumn\\
+\cellcolor{gray!6}{} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{0.7590} & \cellcolor{gray!6}{0.5860} & \cellcolor{gray!6}{0.000665}\\
+
+ & Spring & 0.0621 & 0.0998 & 0.593000\\
+
+\cellcolor{gray!6}{} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{-0.0789} & \cellcolor{gray!6}{-0.1110} & \cellcolor{gray!6}{0.560000}\\
+
+\multirow{-4}{*}{\centering\arraybackslash Cloudy skies} & Autumn & 0.1910 & 0.1650 & 0.384000\\
 \bottomrule

### 2023-12-19 MDPI_submission.Rmd
Commit 2023-12-19 09:12

@@
 
-To calculate the reference clear-sky $\text{SDR}_\text{CSref}$ we used the $\text{SDR}_\text{Haurwitz}$ derived by the radiation model of @Haurwitz1945, adjusted for our site with a factor $a$ (Eq.\ \ref{eq:ahau}), estimated through an iterative optimization process, as described by @Long2000 and @Reno2016.
+To calculate the reference clear-sky $\text{SDR}_\text{CSref}$ we used the $\text{SDR}_\text{Haurwitz}$ derived by the radiation model of @Haurwitz1945 (Eq.\ \ref{eq:hau}), adjusted for our site:
+<!-- adjusted Haurwitz -->
+\begin{equation}
+\text{SDR}_\text{Haurwitz} = 1098 \times \cos(\theta) \times \exp \left( \frac{ - 0.057}{\cos(\theta)} \right) \label{eq:hau}
+\end{equation}
+where $\theta$ is the solar zenith angle (SZA).
+
+The adjustment was made with a factor $a$ (Eq.\ \ref{eq:ahau}), which was estimated through an iterative optimization process, as described by @Long2000 and @Reno2016.
 The target of the optimization was the minimization of a function $f(a)$ (Eq.\ \ref{eq:minf}) and was accomplished with the algorithmic function "optimise", which is an implementation based on the work of @Brent1973, from the library "stats" of the R programming language [@RCT2023].
@@
 We found, that Haurwitz's model, adjusted with the factor $a = `r signif(alpha,3)`$ yields one of the lowest root mean squared errors (RMSE), while the procedure manages to characterize successfully the majority of the data.
-Thus, our clear sky reference is derived by the Eq.\ \ref{eq:ahau}.
-<!-- adjusted Haurwitz -->
+Thus, our clear sky reference is derived by the Eq.\ \ref{eq:ahau}:
 \begin{equation}
@@
 \end{equation}
-where $\theta$ is the solar zenith angle (SZA).
 
@@
 $0.4\,\%/y$ for all-sky and cloudy-sky conditions, while for clear-skies the
-trend is much smaller (\~$`r cloudtrend`\,\%/y$). 
+trend is much smaller (\~$`r cleartrend`\,\%/y$). 
 The linear trends were calculated
@@
 positive trend of SDR under cloudy skies can be attributed to reduction
-in cloud cover and/or optical depth. 
+in cloud cover and/or cloud optical depth. 
 Lack of continuous cloud
@@
 
-```{r trendALL, echo=F, out.width='.70\\linewidth', fig.cap=paste0("Anomalies (\\%) of the daily all-sky SDR from the climatological mean for the period", year(min(ALL_1_daily_DESEAS$Date)), " - ", year(max(ALL_1_daily_DESEAS$Date)), ".
+```{r trendALL, echo=F, out.width='.70\\linewidth', fig.cap=paste0("Anomalies (\\%) of the daily all-sky SDR from the climatological mean for the period ", year(min(ALL_1_daily_DESEAS$Date)), " -- ", year(max(ALL_1_daily_DESEAS$Date)), ".
 The black line is the long term linear trend.")}
@@
 Although the year-to-year variability of the anomalies (Figure \ref{fig:trendALL} and Figures \ref{fig:trendCLEAR}, \ref{fig:trendCLOUD} in Appendix), shows a rather homogeneous behaviour, plots of the cumulative sums (CUSUM) [@Regier2019] of the anomalies can reveal different structures in the records of all three sky conditions.
-In the cases of all-sky and clear-sky conditions (Figures \ref{fig:cusummonth-1} and \ref{fig:cusummonth-2}), we observe three macroscopic periods.
+For time series with a uniform trend, we would expect
+the CUSUMs of the anomalies to have a symmetric 'V' shape centered
+around the middle of the series. This would indicate that the anomalies
+are evenly distributed around the climatological mean, and for a
+positive uniform trend, the first half is below and the second half
+above the climatological mean.
+In our case, there is a more complex
+evolution of the anomalies.
+For all-skies (Figure \ref{fig:cusummonth-1}),
+we observe three
+rather distinct periods: a) a downward part between the start of the
+datasets and about 2000, denoting that all anomalies are negative thus
+below the climatology; b) a relatively steady part lasting for almost 20
+years suggesting little variability in SDR anomalies; and c) a steep
+upward part to present indicating anomalies above the climatology.
+
+(Figures \ref{fig:cusummonth-1} and \ref{fig:cusummonth-2}), we observe three macroscopic periods.
 A downward part from the start until about 2005, a relatively steady part until about 2016 and, finally, a steep upward part until the present.
@@
 
-We have demonstrated that in the period
-`r paste(year(min(ALL_1_daily_DESEAS$Date)), " - ", year(max(ALL_1_daily_DESEAS$Date)))`, there is a positive trend in SDR of ($`r alltrend`\,\%/y$) (brightening) (positive trend) in Thessaloniki, Greece, under all-sky conditions.
-A previous study [@Bais2013] for the period 1993 - 2011 found also a positive trend of $0.33\,\%/y$.
-The increase of this trend indicates that the brightening of SDR continues and is probably caused by continuing decreases in aerosol optical depth and the optical thickness of clouds over the area.
+We have analyzed a 30 year dataset of SRD measurements in Thessaloniki Greece 
+(`r paste(year(min(ALL_1_daily_DESEAS$Date)), " -- ", year(max(ALL_1_daily_DESEAS$Date)))`) aiming to identify the long term variability of
+solar irradiance under different sky conditions.
+Under all-sky conditions 
+there is a positive trend in SDR of $`r alltrend`\,\%/y$ (brightening). 
+A previous study [@Bais2013] for the period 1993 -- 2011 reported also a positive trend, of $0.33\,\%/y$.
+The slight increase of this trend indicates that the brightening of SDR continues and is probably caused by continuing decreases in aerosol optical depth and the optical thickness of clouds over the area.
 Moreover, we found a similar trend under clear-sky conditions ($`r cleartrend`\,\%/y$) that further supports the assumption that the brightening is caused mainly by decreasing aerosols.
@@
 
-Using the CUSUMs of the monthly departures for all- and clear-skies, we observed periods where the CUSUMs remain relatively stable, with a steep decline before and a steep increase after. 
-This is an indication that the whole brightening effect does not follow a smooth development over time.
+Using the CUSUMs of the monthly departures for all- and cloudy-skies, we observed a 20 year period starting around 2000 where the CUSUMs remain relatively stable, with a steep decline before and a steep increase after.
+The rather smooth course of the CUSUMs suggests that no important
+change in the SDR pattern has occurred in the entire record.
 

### 2023-12-19 MDPI_submission.tex
Commit 2023-12-19 09:12

@@
 \(22^\circ\,57'\,\)E, \(80\,\)m~a.s.l.) since 1993. Here we use data for
-the period from 1993-04-13 to 2023-04-12. The monitoring site is located
+the period from 1993-04-13 to 2023-04-13. The monitoring site is located
 near the city centre thus we expect that measurements are affected by
@@
 the \(\text{SDR}_\text{Haurwitz}\) derived by the radiation model of
-\citet{Haurwitz1945}, adjusted for our site with a factor \(a\)
-(Eq.~\ref{eq:ahau}), estimated through an iterative optimization
-process, as described by \citet{Long2000} and \citet{Reno2016}. The
-target of the optimization was the minimization of a function \(f(a)\)
-(Eq.~\ref{eq:minf}) and was accomplished with the algorithmic function
-``optimise'', which is an implementation based on the work of
-\citet{Brent1973}, from the library ``stats'' of the R programming
-language \citep{RCT2023}. \begin{equation}
+\citet{Haurwitz1945} (Eq.~\ref{eq:hau}), adjusted for our site:
+\begin{equation}
+\text{SDR}_\text{Haurwitz} = 1098 \times \cos(\theta) \times \exp \left( \frac{ - 0.057}{\cos(\theta)} \right) \label{eq:hau}
+\end{equation} where \(\theta\) is the solar zenith angle (SZA).
+
+The adjustment was made with a factor \(a\) (Eq.~\ref{eq:ahau}), which
+was estimated through an iterative optimization process, as described by
+\citet{Long2000} and \citet{Reno2016}. The target of the optimization
+was the minimization of a function \(f(a)\) (Eq.~\ref{eq:minf}) and was
+accomplished with the algorithmic function ``optimise'', which is an
+implementation based on the work of \citet{Brent1973}, from the library
+``stats'' of the R programming language \citep{RCT2023}.
+\begin{equation}
 f(a) = \frac{1}{n}\sum_{i=1}^{n} ( \text{SDR}_{\text{CSid},i} - a \times \text{SDR}_{\text{testCSref},i} )^2 \label{eq:minf}
@@
 majority of the data. Thus, our clear sky reference is derived by the
-Eq.~\ref{eq:ahau}. \begin{equation}
+Eq.~\ref{eq:ahau}: \begin{equation}
 \text{SDR}_\text{CSref} = a \times \text{SDR}_\text{Haurwitz} = 0.965 \times 1098 \times \cos(\theta) \times \exp \left( \frac{ - 0.057}{\cos(\theta)} \right) \label{eq:ahau}
-\end{equation} where \(\theta\) is the solar zenith angle (SZA).
+\end{equation}
 
@@
 while for clear-skies the trend is much smaller
-(\textasciitilde{}\(0.41\,\%/y\)). The linear trends were calculated
+(\textasciitilde{}\(0.1\,\%/y\)). The linear trends were calculated
 taking into account the autocorrelation of the time series and all three
@@
 In contrast the large positive trend of SDR under cloudy skies can be
-attributed to reduction in cloud cover and/or optical depth. Lack of
-continuous cloud observations that could support these findings does not
-allow drawing firm conclusions. However, there are indications that the
-total cloud-cover as inferred from the ERA5 analysis for the grid point
-of Thessaloniki is decreasing over the period of study.
+attributed to reduction in cloud cover and/or cloud optical depth. Lack
+of continuous cloud observations that could support these findings does
+not allow drawing firm conclusions. However, there are indications that
+the total cloud-cover as inferred from the ERA5 analysis for the grid
+point of Thessaloniki is decreasing over the period of study.
 
@@
 
-\caption{Anomalies (\%) of the daily all-sky SDR from the climatological mean for the period1993 - 2023. The black line is the long term linear trend.}\label{fig:trendALL}
+\caption{Anomalies (\%) of the daily all-sky SDR from the climatological mean for the period 1993 -- 2023. The black line is the long term linear trend.}\label{fig:trendALL}
 \end{figure}
@@
 can reveal different structures in the records of all three sky
-conditions. In the cases of all-sky and clear-sky conditions (Figures
-\ref{fig:cusummonth-1} and \ref{fig:cusummonth-2}), we observe three
-macroscopic periods. A downward part from the start until about 2005, a
-relatively steady part until about 2016 and, finally, a steep upward
-part until the present. For cloud-sky (Figure~\ref{fig:cusummonth-3}),
-we have a different pattern; it begins with a relatively steady part
-until 1997, followed by an upward part until 2005, and a long decline
-until 2020, with a small positive slope until the present. For a uniform
-trend, we would expect the CUSUMs of the anomalies to have a symmetric
-`V' shape. This would indicate that the anomalies are evenly distributed
-around the climatological mean, and for a positive uniform trend, the
-first half to be below and the other half above the climatological mean.
-In our case, there is a more complex evolution of the anomalies. Another
-distinct feature of the CUSUMs, is the different pattern of the
-cloudy-sky dataset which peaks around the middle of the period.
-Although, there seems to exist a complementary relation to the CUSUMs of
-the clear- and all-sky cases, we can not assert that clouds are the main
-driver for this relation due to the great difference in the number of
-observational data between the two datasets
-(Table~\ref{tab:trendtable}).
+conditions. For time series with a uniform trend, we would expect the
+CUSUMs of the anomalies to have a symmetric `V' shape centered around
+the middle of the series. This would indicate that the anomalies are
+evenly distributed around the climatological mean, and for a positive
+uniform trend, the first half is below and the second half above the
+climatological mean. In our case, there is a more complex evolution of
+the anomalies. For all-skies (Figure \ref{fig:cusummonth-1}) we observe
+three rather distinct periods: a) a downward part between the start of
+the datasets and about 2000, denoting that all anomalies are negative
+thus below the climatology; b) a relatively steady part lasting for
+almost 20 years suggesting little variability in SDR anomalies; and c) a
+steep upward part to present indicating anomalies above the climatology.
+
+(Figures \ref{fig:cusummonth-1} and \ref{fig:cusummonth-2}), we observe
+three macroscopic periods. A downward part from the start until about
+2005, a relatively steady part until about 2016 and, finally, a steep
+upward part until the present. For cloud-sky
+(Figure~\ref{fig:cusummonth-3}), we have a different pattern; it begins
+with a relatively steady part until 1997, followed by an upward part
+until 2005, and a long decline until 2020, with a small positive slope
+until the present. For a uniform trend, we would expect the CUSUMs of
+the anomalies to have a symmetric `V' shape. This would indicate that
+the anomalies are evenly distributed around the climatological mean, and
+for a positive uniform trend, the first half to be below and the other
+half above the climatological mean. In our case, there is a more complex
+evolution of the anomalies. Another distinct feature of the CUSUMs, is
+the different pattern of the cloudy-sky dataset which peaks around the
+middle of the period. Although, there seems to exist a complementary
+relation to the CUSUMs of the clear- and all-sky cases, we can not
+assert that clouds are the main driver for this relation due to the
+great difference in the number of observational data between the two
+datasets (Table~\ref{tab:trendtable}).
 
@@
 
-We have demonstrated that in the period 1993 - 2023, there is a positive
-trend in SDR of (\(0.38\,\%/y\)) (brightening) (positive trend) in
-Thessaloniki, Greece, under all-sky conditions. A previous study
-\citep{Bais2013} for the period 1993 - 2011 found also a positive trend
-of \(0.33\,\%/y\). The increase of this trend indicates that the
-brightening of SDR continues and is probably caused by continuing
-decreases in aerosol optical depth and the optical thickness of clouds
-over the area. Moreover, we found a similar trend under clear-sky
-conditions (\(0.1\,\%/y\)) that further supports the assumption that the
-brightening is caused mainly by decreasing aerosols. Unfortunately, for
-the entire period there is no available data for the aerosols, in order
-to quantify their effect on SDR. However, \citet{Siomos2020} have shown
-that aerosol optical depth over Thessaloniki is decreasing constantly at
-least up to 2018. The attenuation of SDR by aerosols over Europe have
-been proposed as major factor by \citet{Wild2021}. The dimming effect on
-SDR under cloudy-sky conditions (\(0.41\,\%/y\)), suggests that cloud
-optical thickness is decreasing during this period. Because we have no
-adequate data to investigate the long term changes of cloud thickness in
-the region, we cannot verify if the negative SDR trend we observe under
-under cloudy-skies can be attributed solely to changes in clouds.
+We have analyzed a 30 year dataset of SRD measurements in Thessaloniki
+Greece (1993 -- 2023) aiming to identify the long term variability of
+solar irradiance under different sky conditions. Under all-sky
+conditions there is a positive trend in SDR of \(0.38\,\%/y\)
+(brightening). A previous study \citep{Bais2013} for the period 1993 --
+2011 found also a positive trend of \(0.33\,\%/y\). The increase of this
+trend indicates that the brightening of SDR continues and is probably
+caused by continuing decreases in aerosol optical depth and the optical
+thickness of clouds over the area. Moreover, we found a similar trend
+under clear-sky conditions (\(0.1\,\%/y\)) that further supports the
+assumption that the brightening is caused mainly by decreasing aerosols.
+Unfortunately, for the entire period there is no available data for the
+aerosols, in order to quantify their effect on SDR. However,
+\citet{Siomos2020} have shown that aerosol optical depth over
+Thessaloniki is decreasing constantly at least up to 2018. The
+attenuation of SDR by aerosols over Europe have been proposed as major
+factor by \citet{Wild2021}. The dimming effect on SDR under cloudy-sky
+conditions (\(0.41\,\%/y\)), suggests that cloud optical thickness is
+decreasing during this period. Because we have no adequate data to
+investigate the long term changes of cloud thickness in the region, we
+cannot verify if the negative SDR trend we observe under under
+cloudy-skies can be attributed solely to changes in clouds.
 
@@
 
-Using the CUSUMs of the monthly departures for all- and clear-skies, we
-observed periods where the CUSUMs remain relatively stable, with a steep
-decline before and a steep increase after. This is an indication that
-the whole brightening effect does not follow a smooth development over
-time.
+Using the CUSUMs of the monthly departures for all- and cloudy-skies, we
+observed a 20 year period starting around 2000 where the CUSUMs remain
+relatively stable, with a steep decline before and a steep increase
+after. The rather smooth course of the CUSUMs suggests that no important
+change in the SDR pattern has occurred in the entire record.
 

### 2023-12-19 Build_169/MDPI_submission_B169.Rmd
Commit 2023-12-19 11:09

--- 2023-12-19 MDPI_submission.Rmd
@@
                         escape      = FALSE,
-                        col.names   = c("Sky conditions", "Trend [\\%/year]", "Trend S.E. ($2\\sigma$)", "Pearson cor.", "Trend [W/year]", "Days"),
+                        col.names   = c("Sky conditions", "Trend [\\%/year]", "Trend S.E. ($2\\sigma$)", "Pearson correl.", "Trend [W/m\\textsuperscript{2}/year]", "Days"),
                         format.args = list(digits = 3),
@@
 
-Although the year-to-year variability of the anomalies (Figure \ref{fig:trendALL} and Figures \ref{fig:trendCLEAR}, \ref{fig:trendCLOUD} in Appendix), shows a rather homogeneous behaviour, plots of the cumulative sums (CUSUM) [@Regier2019] of the anomalies can reveal different structures in the records of all three sky conditions.
+Although the year-to-year variability of the anomalies (Figure \ref{fig:trendALL} and Figures\ \ref{fig:trendCLEAR}, \ref{fig:trendCLOUD} in Appendix), shows a rather homogeneous behaviour, plots of the cumulative sums (CUSUM) [@Regier2019] of the anomalies can reveal different structures in the records of all three sky conditions.
 For time series with a uniform trend, we would expect
@@
 evolution of the anomalies.
-For all-skies (Figure \ref{fig:cusummonth-1}),
+For all-skies (Figure\ \ref{fig:cusummonth-1}),
 we observe three
@@
 years suggesting little variability in SDR anomalies; and c) a steep
-upward part to present indicating anomalies above the climatology.
-
-(Figures \ref{fig:cusummonth-1} and \ref{fig:cusummonth-2}), we observe three macroscopic periods.
-A downward part from the start until about 2005, a relatively steady part until about 2016 and, finally, a steep upward part until the present.
-For cloud-sky (Figure\ \ref{fig:cusummonth-3}), we have a different pattern; it begins with a relatively steady part until 1997, followed by an upward part until 2005, and a long decline until 2020, with a small positive slope until the present.
-For a uniform trend, we would expect the CUSUMs of the anomalies to have a symmetric 'V' shape.
-This would indicate that the anomalies are evenly distributed around the climatological mean, and for a positive uniform trend, the first half to be below and the other half above the climatological mean.
-In our case, there is a more complex evolution of the anomalies.
-Another distinct feature of the CUSUMs, is the different pattern of the cloudy-sky dataset which peaks around the middle of the period.
-Although, there seems to exist a complementary relation to the CUSUMs of the clear- and all-sky cases, we can not assert that clouds are the main driver for this relation due to the great difference in the number of observational data between the two datasets (Table\ \ref{tab:trendtable}).
+upward part to present indicating anomalies above the climatology. The
+CUSUMs for cloudy-skies (Figure\ \ref{fig:cusummonth-3}), show a similar behavior with some short-term
+differences that do not change the overall pattern.
+For clear skies
+(Figure\ \ref{fig:cusummonth-2}), a monotonic downward tendency is evident until 2004,
+suggesting that the anomalies are all negative. After 2004 the anomalies
+turn to positive at a fast rate for about five years and at a slower
+rate thereafter.
 
@@
 
-In order to investigate further the features of the CUSUMs, we created another set of CUSUM plots by subtracting the corresponding long term trend from the SDR anomaly data, prior to the CUSUM calculation (Figure\ \ref{fig:cusumnotrendmonthly}).
+In order to unveil further the features of the variability of the three
+datasets, Figure\ \ref{fig:cusumnotrendmonthly}
+of the CUSUMs, we created another set of CUSUM plots by subtracting the corresponding long term trend from the SDR anomaly data, prior to the CUSUM calculation (Figure\ \ref{fig:cusumnotrendmonthly}).
 With this approach periods when the CUSUMs diverge from zero can be interpreted as a systematic variation of SDR from the climatological mean. 
@@
 spp <- spp[grep("SDR" ,spp$var), ]
-spp$slope.sd           <- NULL
-spp$slope.ConfInt_0.95 <- NULL
-spp$slope.ConfInt_0.99 <- NULL
-spp$var                <- NULL
-
-spp$cor.df             <- NULL
-spp$cor.null_value     <- NULL
+spp$N_eff              <- NULL
+spp$Rsqrd              <- NULL
+spp$RsqrdAdj           <- NULL
+spp$conf_2.5           <- NULL
+spp$conf_97.5          <- NULL
 spp$cor.alternative    <- NULL
-spp$cor.method         <- NULL
-spp$cor.data_name      <- NULL
-spp$cor.p              <- NULL
 spp$cor.conf.level     <- NULL
@@
 spp$cor.conf_level     <- NULL
-
+spp$cor.data_name      <- NULL
+spp$cor.df             <- NULL
+spp$cor.method         <- NULL
+spp$cor.null_value     <- NULL
+spp$cor.p              <- NULL
+spp$cor.t              <- NULL
+spp$slope.ConfInt_0.95 <- NULL
+spp$slope.ConfInt_0.99 <- NULL
+spp$slope.sd           <- NULL
+spp$t_eff              <- NULL
+spp$t_eff_cri          <- NULL
+spp$var                <- NULL
 rownames(spp)          <- NULL
@@
 ## numeric precision
+spp$Tlm_Std..Error  <- signif(spp$Tlm_Std..Error * 2, 3)
 spp$slope           <- signif(spp$slope,          3)
 spp$slope.p         <- signif(spp$slope.p,        3)
-spp$slope.stat_sig <- signif(spp$slope.stat_sig, 3)
-spp$Rsqrd          <- signif(spp$Rsqrd,          3)
+# spp$slope.stat_sig  <- signif(spp$slope.stat_sig, 3)
+# spp$Rsqrd           <- signif(spp$Rsqrd,          3)
 spp$cor.estimate    <- signif(spp$cor.estimate,   3)
-spp$RsqrdAdj       <- NULL
-spp$Rsqrd          <- NULL
-spp$cor.t          <- NULL
-spp$slope.stat_sig <- NULL
-spp$N_eff          <- NULL
-spp$conf_2.5       <- NULL
-spp$conf_97.5      <- NULL
-spp$t_eff          <- NULL
-spp$t_eff_cri      <- NULL
+
+# spp$Tlm_Pr...t..
 
@@
 spp <- rename(spp, "Sky condition"   = DATA           )
-spp <- rename(spp, "Pearson cor."   = cor.estimate)
+spp <- rename(spp, "Pearson correl." = cor.estimate   )
+spp <- rename(spp, "Trend S.E."      = Tlm_Std..Error )
 
@@
       "Trend [%/year]",
-      "Pearson cor.",
+      "Trend S.E.",
+      "Pearson correl.",
       "Trend p-value"
@@
 In spring and summer, the trends are much smaller and of lesser statistical significance.
-These seasonal differences indicate a possible relation of the trends in SDR to trends of clouds during winter and autumn.
+These seasonal differences indicate a possible relation of the trends in SDR to trends in clouds during winter and autumn.
 For clear-skies, the trend in winter is $`r signif(spp$"Trend [%/year]"[spp$Season == "Winter" & spp$"Sky condition" == "Clear skies"], 2)`\,\%/y$, larger than for all-skies ($`r signif(spp$"Trend [%/year]"[spp$Season == "Winter" & spp$"Sky condition" == "All skies"], 2)`\,\%/y$), which is another indication of a decreasing trend in cloud optical thickness.
@@
 
-rownames(spp2) <- NULL
+# names(spp2)
+# col.names   = c("Sky conditions", "Trend [\\%/year]", "Trend S.E. ($2\\sigma$)", "Pearson correl.", "Trend [W/m\\textsuperscript{2}/year]", "Days"),
+
+# sub("Trend S.E.", "Trend S.E. ($2\\sigma$)", names(spp2))
+
 # panderOptions("table.alignment.default", "left")
@@
                         booktabs    = TRUE,
-                        escape      = TRUE,
+                        escape      = FALSE,
+                        col.names   = c("Sky condition",
+                                        "Season",
+                                        "Trend [\\%/year]",
+                                        "Trend S.E. ($2\\sigma$)",
+                                        "Pearson correl.",
+                                        "Trend p-value" ),
                         # format.args = list(digits = 3),
                         linesep     = "",
-                        align       = "ccrrr",
+                        align       = "ccrrrr",
                         centering   = FALSE,
@@
 The slight increase of this trend indicates that the brightening of SDR continues and is probably caused by continuing decreases in aerosol optical depth and the optical thickness of clouds over the area.
-Moreover, we found a similar trend under clear-sky conditions ($`r cleartrend`\,\%/y$) that further supports the assumption that the brightening is caused mainly by decreasing aerosols.
+A smaller trend has
+been found under clear-sky conditions ($`r cleartrend`\,\%/y$) that further supports the assumption that the brightening is caused mainly by decreasing aerosols.
 Unfortunately, for the entire period there is no available data for the aerosols, in order to quantify their effect on SDR. 

### 2023-12-19 Build_169/MDPI_submission_B169.tex
Commit 2023-12-19 11:09

--- 2023-12-19 MDPI_submission.tex
@@
 radiation brightening over Europe. For the period of our observations
-the trend in the TSI is negligible (\(-0.00024\,\%/y\)), and thus we
+the trend in the TSI is negligible (\(-0.00022\,\%/y\)), and thus we
 cannot attribute any significant effect on the SDR trend to solar
@@
 \toprule
-Sky conditions & Trend [\%/year] & Trend S.E. ($2\sigma$) & Pearson cor. & Trend [W/year] & Days\\
+Sky conditions & Trend [\%/year] & Trend S.E. ($2\sigma$) & Pearson correl. & Trend [W/m\textsuperscript{2}/year] & Days\\
 \midrule
-All skies & 0.3760 & 0.1210 & 0.0911 & 1.450 & 10250\\
+All skies & 0.3760 & 0.1210 & 0.0913 & 1.460 & 10251\\
 Clear skies & 0.0972 & 0.0327 & 0.1380 & 0.501 & 2684\\
@@
 Although the year-to-year variability of the anomalies (Figure
-\ref{fig:trendALL} and Figures \ref{fig:trendCLEAR},
+\ref{fig:trendALL} and Figures~\ref{fig:trendCLEAR},
 \ref{fig:trendCLOUD} in Appendix), shows a rather homogeneous behaviour,
@@
 climatological mean. In our case, there is a more complex evolution of
-the anomalies. For all-skies (Figure \ref{fig:cusummonth-1}) we observe
+the anomalies. For all-skies (Figure~\ref{fig:cusummonth-1}), we observe
 three rather distinct periods: a) a downward part between the start of
@@
 steep upward part to present indicating anomalies above the climatology.
-
-(Figures \ref{fig:cusummonth-1} and \ref{fig:cusummonth-2}), we observe
-three macroscopic periods. A downward part from the start until about
-2005, a relatively steady part until about 2016 and, finally, a steep
-upward part until the present. For cloud-sky
-(Figure~\ref{fig:cusummonth-3}), we have a different pattern; it begins
-with a relatively steady part until 1997, followed by an upward part
-until 2005, and a long decline until 2020, with a small positive slope
-until the present. For a uniform trend, we would expect the CUSUMs of
-the anomalies to have a symmetric `V' shape. This would indicate that
-the anomalies are evenly distributed around the climatological mean, and
-for a positive uniform trend, the first half to be below and the other
-half above the climatological mean. In our case, there is a more complex
-evolution of the anomalies. Another distinct feature of the CUSUMs, is
-the different pattern of the cloudy-sky dataset which peaks around the
-middle of the period. Although, there seems to exist a complementary
-relation to the CUSUMs of the clear- and all-sky cases, we can not
-assert that clouds are the main driver for this relation due to the
-great difference in the number of observational data between the two
-datasets (Table~\ref{tab:trendtable}).
+The CUSUMs for cloudy-skies (Figure~\ref{fig:cusummonth-3}), show a
+similar behavior with some short-term differences that do not change the
+overall pattern. For clear skies (Figure~\ref{fig:cusummonth-2}), a
+monotonic downward tendency is evident until 2004, suggesting that the
+anomalies are all negative. After 2004 the anomalies turn to positive at
+a fast rate for about five years and at a slower rate thereafter.
 
@@
 
-In order to investigate further the features of the CUSUMs, we created
+In order to unveil further the features of the variability of the three
+datasets, Figure~\ref{fig:cusumnotrendmonthly} of the CUSUMs, we created
 another set of CUSUM plots by subtracting the corresponding long term
@@
 differences indicate a possible relation of the trends in SDR to trends
-of clouds during winter and autumn. For clear-skies, the trend in winter
+in clouds during winter and autumn. For clear-skies, the trend in winter
 is \(0.4\,\%/y\), larger than for all-skies (\(0.7\,\%/y\)), which is
@@
 \caption{\label{tab:trendseasontable}SDR linear trends of monthly anomalies for each season of the year.}
-\begin{tabu} to \linewidth {>{\centering\arraybackslash}p{8em}>{\centering}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X}
+\begin{tabu} to \linewidth {>{\centering\arraybackslash}p{8em}>{\centering}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X}
 \toprule
-Sky condition & Season & Trend [\%/year] & Pearson cor. & Trend p-value\\
+Sky condition & Season & Trend [\%/year] & Trend S.E. ($2\sigma$) & Pearson correl. & Trend p-value\\
 \midrule
-\cellcolor{gray!6}{} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{0.7000} & \cellcolor{gray!6}{0.5350} & \cellcolor{gray!6}{0.002770}\\
+\cellcolor{gray!6}{} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{0.7000} & \cellcolor{gray!6}{0.4250} & \cellcolor{gray!6}{0.5350} & \cellcolor{gray!6}{0.002770}\\
 
- & Spring & 0.1070 & 0.1660 & 0.371000\\
+ & Spring & 0.1070 & 0.2350 & 0.1660 & 0.371000\\
 
-\cellcolor{gray!6}{} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{0.1050} & \cellcolor{gray!6}{0.2540} & \cellcolor{gray!6}{0.175000}\\
+\cellcolor{gray!6}{} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{0.1050} & \cellcolor{gray!6}{0.1520} & \cellcolor{gray!6}{0.2540} & \cellcolor{gray!6}{0.175000}\\
 
-\multirow{-4}{*}{\centering\arraybackslash All skies} & Autumn & 0.4230 & 0.4660 & 0.009390\\
-\cmidrule{1-5}
-\cellcolor{gray!6}{} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{0.3950} & \cellcolor{gray!6}{0.5960} & \cellcolor{gray!6}{0.000515}\\
+\multirow{-4}{*}{\centering\arraybackslash All skies} & Autumn & 0.4230 & 0.3030 & 0.4660 & 0.009390\\
+\cmidrule{1-6}
+\cellcolor{gray!6}{} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{0.3950} & \cellcolor{gray!6}{0.2010} & \cellcolor{gray!6}{0.5960} & \cellcolor{gray!6}{0.000515}\\
 
- & Spring & 0.0591 & 0.1270 & 0.497000\\
+ & Spring & 0.0591 & 0.1720 & 0.1270 & 0.497000\\
 
-\cellcolor{gray!6}{} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{-0.0540} & \cellcolor{gray!6}{-0.3010} & \cellcolor{gray!6}{0.106000}\\
+\cellcolor{gray!6}{} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{-0.0540} & \cellcolor{gray!6}{0.0646} & \cellcolor{gray!6}{-0.3010} & \cellcolor{gray!6}{0.106000}\\
 
-\multirow{-4}{*}{\centering\arraybackslash Clear skies} & Autumn & 0.0539 & 0.1710 & 0.366000\\
-\cmidrule{1-5}
-\cellcolor{gray!6}{} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{0.7590} & \cellcolor{gray!6}{0.5860} & \cellcolor{gray!6}{0.000665}\\
+\multirow{-4}{*}{\centering\arraybackslash Clear skies} & Autumn & 0.0539 & 0.1170 & 0.1710 & 0.366000\\
+\cmidrule{1-6}
+\cellcolor{gray!6}{} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{0.7590} & \cellcolor{gray!6}{0.3970} & \cellcolor{gray!6}{0.5860} & \cellcolor{gray!6}{0.000665}\\
 
- & Spring & 0.0621 & 0.0998 & 0.593000\\
+ & Spring & 0.0621 & 0.2300 & 0.0998 & 0.593000\\
 
-\cellcolor{gray!6}{} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{-0.0789} & \cellcolor{gray!6}{-0.1110} & \cellcolor{gray!6}{0.560000}\\
+\cellcolor{gray!6}{} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{-0.0789} & \cellcolor{gray!6}{0.2670} & \cellcolor{gray!6}{-0.1110} & \cellcolor{gray!6}{0.560000}\\
 
-\multirow{-4}{*}{\centering\arraybackslash Cloudy skies} & Autumn & 0.1910 & 0.1650 & 0.384000\\
+\multirow{-4}{*}{\centering\arraybackslash Cloudy skies} & Autumn & 0.1910 & 0.4320 & 0.1650 & 0.384000\\
 \bottomrule
@@
 (brightening). A previous study \citep{Bais2013} for the period 1993 --
-2011 found also a positive trend of \(0.33\,\%/y\). The increase of this
-trend indicates that the brightening of SDR continues and is probably
-caused by continuing decreases in aerosol optical depth and the optical
-thickness of clouds over the area. Moreover, we found a similar trend
-under clear-sky conditions (\(0.1\,\%/y\)) that further supports the
-assumption that the brightening is caused mainly by decreasing aerosols.
-Unfortunately, for the entire period there is no available data for the
-aerosols, in order to quantify their effect on SDR. However,
-\citet{Siomos2020} have shown that aerosol optical depth over
-Thessaloniki is decreasing constantly at least up to 2018. The
+2011 reported also a positive trend, of \(0.33\,\%/y\). The slight
+increase of this trend indicates that the brightening of SDR continues
+and is probably caused by continuing decreases in aerosol optical depth
+and the optical thickness of clouds over the area. A smaller trend has
+been found under clear-sky conditions (\(0.1\,\%/y\)) that further
+supports the assumption that the brightening is caused mainly by
+decreasing aerosols. Unfortunately, for the entire period there is no
+available data for the aerosols, in order to quantify their effect on
+SDR. However, \citet{Siomos2020} have shown that aerosol optical depth
+over Thessaloniki is decreasing constantly at least up to 2018. The
 attenuation of SDR by aerosols over Europe have been proposed as major

### 2023-12-19 MDPI_submission.Rmd
Commit 2023-12-19 12:14

@@
 ## numeric precision of table
-LT$slope         <- signif(LT$slope,         3)
+LT$slope           <- signif(LT$slope,               2)
 LT$slope.p         <- signif(LT$slope.p,             3)
-LT$Rsqrd         <- signif(LT$Rsqrd,         3)
-LT$Tmod_Estimate <- signif(LT$Tmod_Estimate, 3)
-LT$cor.estimate  <- signif(LT$cor.estimate,  3)
-LT$Tmod_Std..Error <- signif(LT$Tmod_Std..Error * 2, 3)
-LT$ChangeWpY       <- signif(LT$ChangeWpY, 3)
+LT$Rsqrd           <- signif(LT$Rsqrd,               2)
+LT$Tmod_Estimate   <- signif(LT$Tmod_Estimate,       2)
+LT$cor.estimate    <- signif(LT$cor.estimate,        2)
+LT$Tmod_Std..Error <- signif(LT$Tmod_Std..Error * 2, 2)
+LT$ChangeWpY       <- signif(LT$ChangeWpY,           2)
+LT$slope <- format(LT$slope, 3, drop0trailing = T)
 
@@
 # LT1 <- LT
+
+signif(LT1$`Trend [%/year]` , 3)
+round(LT1$`Trend [%/year]` , 3)
+format(LT1$`Trend [%/year]` , 3, drop0trailing = T)
+as.character(format(LT1$`Trend [%/year]` , 3))
 
@@
                         position    = "H",
+                        drop0trailing = T,
                         caption     = paste0("Trends in SDR daily means for different sky conditions for the period ", year(min(ALL_1_daily_DESEAS$Date)), " - ", year(max(ALL_1_daily_DESEAS$Date)), "."))
@@
 ## numeric precision
-spp$slope          <- signif(spp$slope,          3)
-spp$slope.p        <- signif(spp$slope.p,        3)
-spp$slope.stat_sig <- signif(spp$slope.stat_sig, 3)
-spp$Rsqrd          <- signif(spp$Rsqrd,          3)
-spp$cor.estimate   <- signif(spp$cor.estimate,   3)
-spp$RsqrdAdj       <- NULL
-spp$Rsqrd          <- NULL
-spp$cor.t          <- NULL
-spp$slope.stat_sig <- NULL
-spp$N_eff          <- NULL
-spp$conf_2.5       <- NULL
-spp$conf_97.5      <- NULL
-spp$t_eff          <- NULL
-spp$t_eff_cri      <- NULL
+spp$Tlm_Std..Error  <- round(spp$Tlm_Std..Error * 2, 2)
+spp$slope           <- round(spp$slope,              2)
+spp$slope.p         <- round(spp$slope.p,            3)
+spp$cor.estimate    <- round(spp$cor.estimate,       2)
+# spp$slope.stat_sig  <- signif(spp$slope.stat_sig, 3)
+# spp$Rsqrd           <- signif(spp$Rsqrd,          3)
+
+# spp$Tlm_Pr...t..
 

### 2023-12-19 MDPI_submission.tex
Commit 2023-12-19 12:14

@@
 
+\begin{verbatim}
+[1] 0.380 0.097 0.410
+\end{verbatim}
+
+\begin{verbatim}
+[1] 0.380 0.097 0.410
+\end{verbatim}
+
+\begin{verbatim}
+[1] "0.38"  "0.097" "0.41" 
+\end{verbatim}
+
+\begin{verbatim}
+[1] "0.380" "0.097" "0.410"
+\end{verbatim}
+
 We calculated the linear trends of SDR, from the departures of the mean
@@
 \toprule
-Sky conditions & Trend [\%/year] & Trend S.E. ($2\sigma$) & Pearson cor. & Trend [W/year] & Days\\
+Sky conditions & Trend [\%/year] & Trend S.E. ($2\sigma$) & Pearson correl. & Trend [W/m\textsuperscript{2}/year] & Days\\
 \midrule
-All skies & 0.3760 & 0.1210 & 0.0911 & 1.450 & 10250\\
-Clear skies & 0.0972 & 0.0327 & 0.1380 & 0.501 & 2684\\
-Cloudy skies & 0.4080 & 0.1770 & 0.0810 & 1.180 & 4937\\
+All skies & 0.380 & 0.120 & 0.091 & 1.5 & 10251\\
+Clear skies & 0.097 & 0.033 & 0.140 & 0.5 & 2684\\
+Cloudy skies & 0.410 & 0.180 & 0.081 & 1.2 & 4937\\
 \bottomrule
@@
 \caption{\label{tab:trendseasontable}SDR linear trends of monthly anomalies for each season of the year.}
-\begin{tabu} to \linewidth {>{\centering\arraybackslash}p{8em}>{\centering}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X}
+\begin{tabu} to \linewidth {>{\centering\arraybackslash}p{8em}>{\centering}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X}
 \toprule
-Sky condition & Season & Trend [\%/year] & Pearson cor. & Trend p-value\\
+Sky condition & Season & Trend [\%/year] & Trend S.E. ($2\sigma$) & Pearson correl. & Trend p-value\\
 \midrule
-\cellcolor{gray!6}{} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{0.7000} & \cellcolor{gray!6}{0.5350} & \cellcolor{gray!6}{0.002770}\\
+\cellcolor{gray!6}{} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{0.70} & \cellcolor{gray!6}{0.43} & \cellcolor{gray!6}{0.54} & \cellcolor{gray!6}{0.003}\\
 
- & Spring & 0.1070 & 0.1660 & 0.371000\\
+ & Spring & 0.11 & 0.24 & 0.17 & 0.371\\
 
-\cellcolor{gray!6}{} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{0.1050} & \cellcolor{gray!6}{0.2540} & \cellcolor{gray!6}{0.175000}\\
+\cellcolor{gray!6}{} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{0.11} & \cellcolor{gray!6}{0.15} & \cellcolor{gray!6}{0.25} & \cellcolor{gray!6}{0.175}\\
 
-\multirow{-4}{*}{\centering\arraybackslash All skies} & Autumn & 0.4230 & 0.4660 & 0.009390\\
-\cmidrule{1-5}
-\cellcolor{gray!6}{} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{0.3950} & \cellcolor{gray!6}{0.5960} & \cellcolor{gray!6}{0.000515}\\
+\multirow{-4}{*}{\centering\arraybackslash All skies} & Autumn & 0.42 & 0.30 & 0.47 & 0.009\\
+\cmidrule{1-6}
+\cellcolor{gray!6}{} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{0.40} & \cellcolor{gray!6}{0.20} & \cellcolor{gray!6}{0.60} & \cellcolor{gray!6}{0.001}\\
 
- & Spring & 0.0591 & 0.1270 & 0.497000\\
+ & Spring & 0.06 & 0.17 & 0.13 & 0.497\\
 
-\cellcolor{gray!6}{} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{-0.0540} & \cellcolor{gray!6}{-0.3010} & \cellcolor{gray!6}{0.106000}\\
+\cellcolor{gray!6}{} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{-0.05} & \cellcolor{gray!6}{0.06} & \cellcolor{gray!6}{-0.30} & \cellcolor{gray!6}{0.106}\\
 
-\multirow{-4}{*}{\centering\arraybackslash Clear skies} & Autumn & 0.0539 & 0.1710 & 0.366000\\
-\cmidrule{1-5}
-\cellcolor{gray!6}{} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{0.7590} & \cellcolor{gray!6}{0.5860} & \cellcolor{gray!6}{0.000665}\\
+\multirow{-4}{*}{\centering\arraybackslash Clear skies} & Autumn & 0.05 & 0.12 & 0.17 & 0.366\\
+\cmidrule{1-6}
+\cellcolor{gray!6}{} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{0.76} & \cellcolor{gray!6}{0.40} & \cellcolor{gray!6}{0.59} & \cellcolor{gray!6}{0.001}\\
 
- & Spring & 0.0621 & 0.0998 & 0.593000\\
+ & Spring & 0.06 & 0.23 & 0.10 & 0.593\\
 
-\cellcolor{gray!6}{} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{-0.0789} & \cellcolor{gray!6}{-0.1110} & \cellcolor{gray!6}{0.560000}\\
+\cellcolor{gray!6}{} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{-0.08} & \cellcolor{gray!6}{0.27} & \cellcolor{gray!6}{-0.11} & \cellcolor{gray!6}{0.560}\\
 
-\multirow{-4}{*}{\centering\arraybackslash Cloudy skies} & Autumn & 0.1910 & 0.1650 & 0.384000\\
+\multirow{-4}{*}{\centering\arraybackslash Cloudy skies} & Autumn & 0.19 & 0.43 & 0.16 & 0.384\\
 \bottomrule
@@
 respectively, in contrast to spring and summer. The trends for autumn
-are also significant but smaller ( \(0.42\) and \(0.054\,\%/y\) for all-
+are also significant but smaller ( \(0.42\) and \(0.05\,\%/y\) for all-
 and clear-skies, respectively). Our findings are in agreement with other

### 2023-12-19 DHI_GHI_1_longterm_trends.R
Commit 2023-12-19 14:14

@@ -422,7 +422,6 @@ for (DBn in dbs) {
             # ylab = bquote("Deseas." ~ .(translate(avar)) ~ "[" ~ Watt/m^2 ~ "]" ) )
 
 
-
             if (DRAFT == TRUE) {
                 ## Running mean
                 first <- head(which(!is.na(dataset[[avar]])),1)
@@ -449,12 +448,12 @@ for (DBn in dbs) {
 
             ## display trend on graph
             fit <- lm1[[1]]
-
+signif(Tres[2], 2)
             legend("top", lty = 1, bty = "n", lwd = 2, cex = 1,
                    paste("Trend: ",
                          if (fit[2] > 0) "+" else "-",
-                         signif(abs(fit[2]) * Days_of_year, 3),
-                         "%/y" )
+                         signif(abs(fit[2]) * Days_of_year, 2),
+                         "±", signif(2 * Tres[2], 2) ,"%/y" )
             )
 
     }
@@ -1721,16 +1720,16 @@ write_dat(pprint,
 
             abline(lm2)
 
-
             ## decorations
             fit <- lm2[[1]]
+            pm <- lmtest::coeftest(lm2)[2,2]
 
-            legend("top", lty = 1, bty = "n", lwd = 1, cex = 0.8,
-                   inset = 0.1,
+            legend("bottom", lty = 1, bty = "n", lwd = 1, cex = 0.8,
+                   inset = 0.03, # o.1
                    paste("Trend: ",
                          if (fit[2] > 0) '+' else '-',
                          signif(abs(fit[2]), 2),
-                         "%/y")
+                         "±", signif(2 * pm, 2), "%/y")
             )
 
 

### 2023-12-19 MDPI_submission.Rmd
Commit 2023-12-19 15:11

@@
 
-## Aggregation of data and statistical approach
+## Aggregation of data and statistical approach {#aggregationstatistical}
 
@@
 The linear trends were calculated using an first order autoregressive model
-(arima) with lag-1, and were reported together with the $2\sigma$ errors (Table\ \ref{tab:trendtable}).
+ with lag of 1 day with fitting method of the "maximum likelihood" [@Gardner1980; @Jones1980], by implementing the function "arima" from the library "stats" of the R programming language [@RCT2023].
+The trends were reported together with the $2\sigma$ errors (Table\ \ref{tab:trendtable}).
 
@@
 <!-- 2000) in order to assess the significance of the trends.]{.mark} -->
-@Santer2000
+<!-- @Santer2000 -->
 
@@
 alltrend   <- signif(LT[DATA == "All sky cond.",    slope], 2)
-cleartrend <- signif(LT[DATA == "Clear sky cond.",  slope], 1)
+cleartrend <- signif(LT[DATA == "Clear sky cond.",  slope], 2)
 cloudtrend <- signif(LT[DATA == "Cloudy sky cond.", slope], 2)
@@
 
-signif(LT1$`Trend [%/year]` , 3)
-round(LT1$`Trend [%/year]` , 3)
-format(LT1$`Trend [%/year]` , 3, drop0trailing = T)
-as.character(format(LT1$`Trend [%/year]` , 3))
+# signif(LT1$`Trend [%/year]` , 3)
+# round(LT1$`Trend [%/year]` , 3)
+# format(LT1$`Trend [%/year]` , 3, drop0trailing = T)
+# as.character(format(LT1$`Trend [%/year]` , 3))
 
@@
 $0.4\,\%/y$ for all-sky and cloudy-sky conditions, while for clear-skies the
-trend is much smaller (\~$`r cleartrend`\,\%/y$). 
+trend is much smaller (\~$`r signif(cleartrend,1)`\,\%/y$). 
 The linear trends were calculated
@@
 datasets, Figure\ \ref{fig:cusumnotrendmonthly}
-of the CUSUMs, we created another set of CUSUM plots by subtracting the corresponding long term trend from the SDR anomaly data, prior to the CUSUM calculation (Figure\ \ref{fig:cusumnotrendmonthly}).
-With this approach periods when the CUSUMs diverge from zero can be interpreted as a systematic variation of SDR from the climatological mean. 
-When the CUSUM is increasing, the added values are above the climatological values of the SDR trend and vice versa.
-Overall, for all- and clear-sky conditions (Figures\ \ref{fig:cusumnotrendmonthly-1} and\ \ref{fig:cusumnotrendmonthly-2}) we observe periods when the anomalies diverge from the climatological value, each lasting for several years.
-The pattern in both datasets is very similar, suggesting prevalence in clear skies over Thessaloniki.
-It is interesting that in the period 1993 - 2016 the anomalies have a high variability around zero, while after 2016, the range of the variability is decreased to about one third of the prior period.
-For cloudy-sky conditions (Figure\ \ref{fig:cusumnotrendmonthly-3}) the period 1997 - 2008 is dominated by positive CUSUMs, suggesting a reduced effect of clouds on SDR.
-From 1997 to mid-2000s CUSUMs are increasing, likely due to a continuous decrease in the optical thickness of clouds, followed by a period of rapid increase (within 3 years) in cloud optical thickness lasting up to 2008.
-The following stable period spans for about 15 years up to 2021 when CUSUMs start increasing again.
+presents another set of CUSUM plots using anomalies
+after the long-term linear trend is removed. With this approach, periods
+when the CUSUMs diverge from zero can be interpreted as a systematic
+variation of SDR from the climatological mean. When the CUSUM is
+increasing, the anomalies values are above the climatology and vice
+versa. 
+Overall, for all- and cloudy-sky conditions (Figures\ \ref{fig:cusumnotrendmonthly-1} and\ \ref{fig:cusumnotrendmonthly-3}) we
+observe periods with anomalies diverging from the climatological values,
+each lasting for several years. These fluctuations are probably within
+the natural variability and no distinct changes are identified. The
+pattern in both datasets is similar, suggesting prevalence in cloudy
+skies over Thessaloniki.
+For clear skies (Figure\ \ref{fig:cusumnotrendmonthly-2}) the distinct change
+in 2004 is now clearer. The most likely reason for this change is the
+monotonic reduction of aerosols in Thessaloniki. At that year there a
+change in the rate of decrease in aerosol optical depth as illustrated
+in Figure 7 of @Siomos2020. This abrupt change in CUSUMs lasts
+until about 2010 when the anomalies become again variable.
 
@@
 
-For all-sky conditions the brightening effect of SDR (positive trend) is
-stronger for large SZAs (Figures\ \ref{fig:szatrends-1} and \ref{fig:szatrends-2})
+For all-sky conditions the brightening effect of SDR (positive trend) increases with SZAs (Figures\ \ref{fig:szatrends-1})
 ranging from about $0.1\,\%/y$ to
-about $0.7\,\%/y$.
+about $0.7\,\%/y$ for the statistically significant trends.
 The trends in the morning and afternoon hours are more or
@@
 attributed to effects on clear sky SDR from systematic diurnal patterns
-of aerosols during the warm period of the year [@Wang2021]. 
-Note that SZAs less than $35^\circ$ can only occur during the warm period of the
-year when clear skies are more frequent.
+of aerosols during the warm period of the year, consistently with the results reported for China by @Wang2021. 
+Note that SZAs less than $25^\circ$ can only occur during the warm period of the year around noon when clear skies are more
+frequent. The increasing of the trend with SZA is likely caused by the
+increased attenuation of SDR with SZA. The effect is larger when aerosol
+and/or cloud layers are optically thicker, therefore, decreases in
+aerosol and clouds through the study period will result in larger
+positive trends of SDR at larger SZAs.
+
+Under clear skies (Figures\ \ref{fig:szatrends-2}), the trends are smaller and less variable,
+ranging between $0.1$ and $0.15\,\%/y$ up to $77^\circ$ SZA.
+At higher SZAs and in
+the afternoon hours there a sharp increase in the trend up to $0.3\,\%/y$,
+which may have been caused by the long path length of radiation through
+the atmosphere as discussed above for the all-sky conditions. The small
 
@@
 (Figure\ \ref{fig:seasonalALL} and Table\ \ref{tab:trendseasontable}).
+<!-- all sky -->
 For all-sky conditions the trend in SDR in winter is the largest
@@
 ($`r signif(spp$"Trend [%/year]"[spp$Season == "Autumn" & spp$"Sky condition" == "All skies"], 2)`\,\%/y$,
-a value close to the long term trend) both statistically significant at the $99\,\%$ confidence level.
+a value close to the long term trend) both statistically significant at the $95\,\%$ confidence level.
 In spring and summer, the trends are much smaller and of lesser statistical significance.
 These seasonal differences indicate a possible relation of the trends in SDR to trends in clouds during winter and autumn.
-For clear-skies, the trend in winter is $`r signif(spp$"Trend [%/year]"[spp$Season == "Winter" & spp$"Sky condition" == "Clear skies"], 2)`\,\%/y$, larger than for all-skies ($`r signif(spp$"Trend [%/year]"[spp$Season == "Winter" & spp$"Sky condition" == "All skies"], 2)`\,\%/y$), which is another indication of a decreasing trend in cloud optical thickness.
-Moreover, the trends under clear- and cloudy-sky conditions are almost complementary to each other, particularly for winter and autumn, where the signal is stronger.
-During spring and summer the statistical significance is very low and the actual trend too small for a meaningful comparison.
+<!-- clear sky -->
+For clear-skies, the trend in winter is $`r signif(spp$"Trend [%/year]"[spp$Season == "Winter" & spp$"Sky condition" == "Clear skies"], 2)`\,\%/y$, and is 
+associated with the decreasing trend in
+aerosol optical depth [@Siomos2020], and almost half of that for
+all-skies, which is another indication of a decreasing trend in cloud
+optical thickness. 
+<!-- clear sky -->
+In other seasons the clear-sky trend is still very small
+(below $0.1\,\%/y$). 
+<!-- cloud dsky -->
+Finally, for cloud-skies the
+winter trend is the largest 
+($`r signif(spp$"Trend [%/year]"[spp$Season == "Winter" & spp$"Sky condition" == "Cloudy skies"], 2)`\,\%/y$)
+and greater than for all-skies,
+followed by a much smaller trend in autumn 
+($`r signif(spp$"Trend [%/year]"[spp$Season == "Autumn" & spp$"Sky condition" == "Cloudy skies"], 2)`\,\%/y$).
+<!-- ($`r signif(spp$"Trend [%/year]"[spp$Season == "Winter" & spp$"Sky condition" == "All skies"], 2)`\,\%/y$) -->
+
+The trends under clear- and cloudy-sky conditions are of the same
+direction and it would be expected that their sum is similar to the
+all-sky trend. This does not happen, especially for winter, likely due
+to the way the monthly means for clear and cloudy skies were calculated.
+Daily means were calculated only when at least $60\,\%$ of the clear- or
+cloudy-sky data were available (see sect. \ref{aggregationstatistical}).
 
@@
 A smaller trend has
-been found under clear-sky conditions ($`r cleartrend`\,\%/y$) that further supports the assumption that the brightening is caused mainly by decreasing aerosols.
-Unfortunately, for the entire period there is no available data for the aerosols, in order to quantify their effect on SDR. 
-However, @Siomos2020 have shown that aerosol optical depth over Thessaloniki is decreasing constantly at least up to 2018.
-The attenuation of SDR by aerosols over Europe have been proposed as major factor by @Wild2021.
-The dimming effect on SDR under cloudy-sky conditions ($`r cloudtrend`\,\%/y$), suggests that cloud optical thickness is decreasing during this period.
-Because we have no adequate data to investigate the long term changes of cloud thickness in the region, we cannot verify if the negative SDR trend we observe under under cloudy-skies can be attributed solely to changes in clouds.
+been found under clear-sky conditions ($`r cleartrend`\,\%/y$) which supports the
+notion that part of the brightening is caused by decreasing aerosols.
+@Siomos2020 have shown that aerosol optical depth over
+Thessaloniki is decreasing constantly at least up to 2018.
+Unfortunately, for this study aerosol data for the entire period were
+not available in order to quantify their effect on SDR. The attenuation
+of SDR by aerosols over Europe has been proposed as major factor by @Wild2021.
+The brightening effect on SDR under cloudy-sky conditions
+($`r cloudtrend`\,\%/y$), suggests that cloud optical thickness is also decreasing
+during this period.
+As long term data of cloud optical thickness are
+also not available for the region, we cannot draw quantitative
+conclusions.
 
@@
 and
-$`r signif(spp$"Trend [%/year]"[spp$Season == "Winter" & spp$"Sky condition" == "Clear skies"], 2)`\,\%/y$
-for all- and clear-skies, respectively, in contrast to spring and summer.
+$`r signif(spp$"Trend [%/year]"[spp$Season == "Winter" & spp$"Sky condition" == "Cloudy skies"], 2)`\,\%/y$
+for all- and cloudy-skies, respectively, in contrast to spring and summer.
 The trends for autumn are also significant but smaller (
@@
 and
-$`r signif(spp$"Trend [%/year]"[spp$Season == "Autumn" & spp$"Sky condition" == "Clear skies"], 2)`\,\%/y$
-for all- and clear-skies, respectively).
+$`r signif(spp$"Trend [%/year]"[spp$Season == "Autumn" & spp$"Sky condition" == "Cloudy skies"], 2)`\,\%/y$
+for all- and cloudy-skies, respectively).
 Our findings are in agreement with other studies for the region.
+The trend for clear skies is largest in winter
+($`r signif(spp$"Trend [%/year]"[spp$Season == "Winter" & spp$"Sky condition" == "Clear skies"], 2)`$) 
+and negligible in spring, summer and autumn.
 

### 2023-12-19 MDPI_submission.tex
Commit 2023-12-19 15:11

@@
 
-\hypertarget{aggregation-of-data-and-statistical-approach}{%
+\hypertarget{aggregationstatistical}{%
 \subsection{Aggregation of data and statistical
-approach}\label{aggregation-of-data-and-statistical-approach}}
+approach}\label{aggregationstatistical}}
 
@@
 The linear trends were calculated using an first order autoregressive
-model (arima) with lag-1, and were reported together with the
+model with lag of 1 day with fitting method of the ``maximum
+likelihood'' \citep{Gardner1980, Jones1980}, by implementing the
+function ``arima'' from the library ``stats'' of the R programming
+language \citep{RCT2023}. The trends were reported together with the
 \(2\sigma\) errors (Table~\ref{tab:trendtable}).
-
-\citet{Santer2000}
 
@@
 \subsection{Long-term SDR trends}\label{long-term-sdr-trends}}
-
-\begin{verbatim}
-[1] 0.380 0.097 0.410
-\end{verbatim}
-
-\begin{verbatim}
-[1] 0.380 0.097 0.410
-\end{verbatim}
-
-\begin{verbatim}
-[1] "0.38"  "0.097" "0.41" 
-\end{verbatim}
-
-\begin{verbatim}
-[1] "0.380" "0.097" "0.410"
-\end{verbatim}
 
@@
 In order to unveil further the features of the variability of the three
-datasets, Figure~\ref{fig:cusumnotrendmonthly} of the CUSUMs, we created
-another set of CUSUM plots by subtracting the corresponding long term
-trend from the SDR anomaly data, prior to the CUSUM calculation
-(Figure~\ref{fig:cusumnotrendmonthly}). With this approach periods when
-the CUSUMs diverge from zero can be interpreted as a systematic
-variation of SDR from the climatological mean. When the CUSUM is
-increasing, the added values are above the climatological values of the
-SDR trend and vice versa. Overall, for all- and clear-sky conditions
+datasets, Figure~\ref{fig:cusumnotrendmonthly} presents another set of
+CUSUM plots using anomalies after the long-term linear trend is removed.
+With this approach, periods when the CUSUMs diverge from zero can be
+interpreted as a systematic variation of SDR from the climatological
+mean. When the CUSUM is increasing, the anomalies values are above the
+climatology and vice versa. Overall, for all- and cloudy-sky conditions
 (Figures~\ref{fig:cusumnotrendmonthly-1}
-and~\ref{fig:cusumnotrendmonthly-2}) we observe periods when the
-anomalies diverge from the climatological value, each lasting for
-several years. The pattern in both datasets is very similar, suggesting
-prevalence in clear skies over Thessaloniki. It is interesting that in
-the period 1993 - 2016 the anomalies have a high variability around
-zero, while after 2016, the range of the variability is decreased to
-about one third of the prior period. For cloudy-sky conditions
-(Figure~\ref{fig:cusumnotrendmonthly-3}) the period 1997 - 2008 is
-dominated by positive CUSUMs, suggesting a reduced effect of clouds on
-SDR. From 1997 to mid-2000s CUSUMs are increasing, likely due to a
-continuous decrease in the optical thickness of clouds, followed by a
-period of rapid increase (within 3 years) in cloud optical thickness
-lasting up to 2008. The following stable period spans for about 15 years
-up to 2021 when CUSUMs start increasing again.
+and~\ref{fig:cusumnotrendmonthly-3}) we observe periods with anomalies
+diverging from the climatological values, each lasting for several
+years. These fluctuations are probably within the natural variability
+and no distinct changes are identified. The pattern in both datasets is
+similar, suggesting prevalence in cloudy skies over Thessaloniki. For
+clear skies (Figure~\ref{fig:cusumnotrendmonthly-2}) the distinct change
+in 2004 is now clearer. The most likely reason for this change is the
+monotonic reduction of aerosols in Thessaloniki. At that year there a
+change in the rate of decrease in aerosol optical depth as illustrated
+in Figure 7 of \citet{Siomos2020}. This abrupt change in CUSUMs lasts
+until about 2010 when the anomalies become again variable.
 
@@
 
-For all-sky conditions the brightening effect of SDR (positive trend) is
-stronger for large SZAs (Figures~\ref{fig:szatrends-1} and
-\ref{fig:szatrends-2}) ranging from about \(0.1\,\%/y\) to about
-\(0.7\,\%/y\). The trends in the morning and afternoon hours are more or
-less consistent with small differences at small SZAs which can be
-attributed to effects on clear sky SDR from systematic diurnal patterns
-of aerosols during the warm period of the year \citep{Wang2021}. Note
-that SZAs less than \(35^\circ\) can only occur during the warm period
-of the year when clear skies are more frequent.
+For all-sky conditions the brightening effect of SDR (positive trend)
+increases with SZAs (Figures~\ref{fig:szatrends-1}) ranging from about
+\(0.1\,\%/y\) to about \(0.7\,\%/y\) for the statistically significant
+trends. The trends in the morning and afternoon hours are more or less
+consistent with small differences at small SZAs which can be attributed
+to effects on clear sky SDR from systematic diurnal patterns of aerosols
+during the warm period of the year, consistently with the results
+reported for China by \citet{Wang2021}. Note that SZAs less than
+\(25^\circ\) can only occur during the warm period of the year around
+noon when clear skies are more frequent. The increasing of the trend
+with SZA is likely caused by the increased attenuation of SDR with SZA.
+The effect is larger when aerosol and/or cloud layers are optically
+thicker, therefore, decreases in aerosol and clouds through the study
+period will result in larger positive trends of SDR at larger SZAs.
+
+Under clear skies (Figures~\ref{fig:szatrends-2}), the trends are
+smaller and less variable, ranging between \(0.1\) and \(0.15\,\%/y\) up
+to \(77^\circ\) SZA. At higher SZAs and in the afternoon hours there a
+sharp increase in the trend up to \(0.3\,\%/y\), which may have been
+caused by the long path length of radiation through the atmosphere as
+discussed above for the all-sky conditions. The small
 
@@
 value close to the long term trend) both statistically significant at
-the \(99\,\%\) confidence level. In spring and summer, the trends are
+the \(95\,\%\) confidence level. In spring and summer, the trends are
 much smaller and of lesser statistical significance. These seasonal
@@
 in clouds during winter and autumn. For clear-skies, the trend in winter
-is \(0.4\,\%/y\), larger than for all-skies (\(0.7\,\%/y\)), which is
-another indication of a decreasing trend in cloud optical thickness.
-Moreover, the trends under clear- and cloudy-sky conditions are almost
-complementary to each other, particularly for winter and autumn, where
-the signal is stronger. During spring and summer the statistical
-significance is very low and the actual trend too small for a meaningful
-comparison.
+is \(0.4\,\%/y\), and is associated with the decreasing trend in aerosol
+optical depth \citep{Siomos2020}, and almost half of that for all-skies,
+which is another indication of a decreasing trend in cloud optical
+thickness. In other seasons the clear-sky trend is still very small
+(below \(0.1\,\%/y\)). Finally, for cloud-skies the winter trend is the
+largest (\(0.76\,\%/y\)) and greater than for all-skies, followed by a
+much smaller trend in autumn (\(0.19\,\%/y\)).
+
+The trends under clear- and cloudy-sky conditions are of the same
+direction and it would be expected that their sum is similar to the
+all-sky trend. This does not happen, especially for winter, likely due
+to the way the monthly means for clear and cloudy skies were calculated.
+Daily means were calculated only when at least \(60\,\%\) of the clear-
+or cloudy-sky data were available (see sect.
+\ref{aggregationstatistical}).
 
@@
 and the optical thickness of clouds over the area. A smaller trend has
-been found under clear-sky conditions (\(0.1\,\%/y\)) that further
-supports the assumption that the brightening is caused mainly by
-decreasing aerosols. Unfortunately, for the entire period there is no
-available data for the aerosols, in order to quantify their effect on
-SDR. However, \citet{Siomos2020} have shown that aerosol optical depth
-over Thessaloniki is decreasing constantly at least up to 2018. The
-attenuation of SDR by aerosols over Europe have been proposed as major
-factor by \citet{Wild2021}. The dimming effect on SDR under cloudy-sky
+been found under clear-sky conditions (\(0.097\,\%/y\)) which supports
+the notion that part of the brightening is caused by decreasing
+aerosols. \citet{Siomos2020} have shown that aerosol optical depth over
+Thessaloniki is decreasing constantly at least up to 2018.
+Unfortunately, for this study aerosol data for the entire period were
+not available in order to quantify their effect on SDR. The attenuation
+of SDR by aerosols over Europe has been proposed as major factor by
+\citet{Wild2021}. The brightening effect on SDR under cloudy-sky
 conditions (\(0.41\,\%/y\)), suggests that cloud optical thickness is
-decreasing during this period. Because we have no adequate data to
-investigate the long term changes of cloud thickness in the region, we
-cannot verify if the negative SDR trend we observe under under
-cloudy-skies can be attributed solely to changes in clouds.
+also decreasing during this period. As long term data of cloud optical
+thickness are also not available for the region, we cannot draw
+quantitative conclusions.
 
@@
 on season, with winter showing the strongest statistically significant
-trend of \(0.7\) and \(0.4\,\%/y\) for all- and clear-skies,
+trend of \(0.7\) and \(0.76\,\%/y\) for all- and cloudy-skies,
 respectively, in contrast to spring and summer. The trends for autumn
-are also significant but smaller ( \(0.42\) and \(0.05\,\%/y\) for all-
-and clear-skies, respectively). Our findings are in agreement with other
-studies for the region.
+are also significant but smaller ( \(0.42\) and \(0.19\,\%/y\) for all-
+and cloudy-skies, respectively). Our findings are in agreement with
+other studies for the region. The trend for clear skies is largest in
+winter (\(0.4\)) and negligible in spring, summer and autumn.
 

### 2023-12-19 MDPI_submission.Rmd
Commit 2023-12-19 16:14

@@
 the atmosphere as discussed above for the all-sky conditions. The small
+differences in the trend between morning and afternoon between $35$ and
+$60^\circ$ SZA is likely a result of less attenuation of SDR in the morning
+hours due to lesser amounts of aerosols and shallower boundary layer.
 
 For cloudy-sky conditions (Figure\ \ref{fig:szatrends-3}), we can not discern any significant dependence of the SDR trend with SZA as the variability of irradiance is
-dominated by the cloud effects.
-For SZAs $25^\circ$ -- $75^\circ$,
-the trends are variable and for most SZAs insignificant ranging between $0$ and $0.2\,\%/y$.
-The smooth increase of the trend at SZAs larger than $\sim 75^\circ$ is probably associated with stronger attenuation by
-clouds under oblique incidence angles which results in smaller
-variability.
+dominated by the cloud effects leading to insignificant
+trends. Statistically significant trends appear only in the afternoon and
+for SZAs larger than $60^\circ$.
+The sharp increase of the trend at SZAs
+larger than $\sim{75}^{\circ}$, observed also for clear skies, is
+probably associated with stronger attenuation by clouds under oblique
+incidence angles, which result also in smaller variability.
 

### 2023-12-19 MDPI_submission.tex
Commit 2023-12-19 16:14

@@
 caused by the long path length of radiation through the atmosphere as
-discussed above for the all-sky conditions. The small
+discussed above for the all-sky conditions. The small differences in the
+trend between morning and afternoon between \(35\) and \(60^\circ\) SZA
+is likely a result of less attenuation of SDR in the morning hours due
+to lesser amounts of aerosols and shallower boundary layer.
 
@@
 discern any significant dependence of the SDR trend with SZA as the
-variability of irradiance is dominated by the cloud effects. For SZAs
-\(25^\circ\) -- \(75^\circ\), the trends are variable and for most SZAs
-insignificant ranging between \(0\) and \(0.2\,\%/y\). The smooth
-increase of the trend at SZAs larger than \(\sim 75^\circ\) is probably
-associated with stronger attenuation by clouds under oblique incidence
-angles which results in smaller variability.
+variability of irradiance is dominated by the cloud effects leading to
+insignificant trends. Statistically significant trends appear only in
+the afternoon and for SZAs larger than \(60^\circ\). The sharp increase
+of the trend at SZAs larger than \(\sim{75}^{\circ}\), observed also for
+clear skies, is probably associated with stronger attenuation by clouds
+under oblique incidence angles, which result also in smaller
+variability.
 

### 2023-12-19 MDPI_submission.Rmd
Commit 2023-12-19 17:11

@@
 The manual screening is aided by a radiation data quality assurance procedure, adjusted for the site, which is based on the methods of
-Long and Shi\ [-@Long2008a; -@Long2006].
+Long and Shi\ [@Long2006; @Long2008a].
 Thus, problematic recordings have been excluded from further processing.
@@
 
-<!-- ### Data selection -->
+<!-- Data selection -->
 Due to the significant measurement uncertainty when the Sun is near the horizon, we have excluded all measurements with SZA greater than $`r 90-MIN_ELEVA`^\circ$.
@@
 In order to estimate the effect of the sky conditions on the long term variability of SDR, we created three datasets by characterizing each one-minute measurement with a corresponding sky-condition flag (i.e., all-sky, clear-sky and cloudy-sky).
-To identify the clear-cases we used a method proposed by
-@Long2000
-and by
-@Reno2016,
-which were adapted and configured for the site of Thessaloniki.
+To identify the clear-cases we used the method proposed by
+@Reno2016, which requires the definition of some site specific parameters.
+These parameters, were produced by an iterative process, as the original authors proposed and it are discussed in the next section.
+
 <!-- CS subjectivity -->
@@
 
-```{r trendALL, echo=F, out.width='.70\\linewidth', fig.cap=paste0("Anomalies (\\%) of the daily all-sky SDR from the climatological mean for the period ", year(min(ALL_1_daily_DESEAS$Date)), " -- ", year(max(ALL_1_daily_DESEAS$Date)), ".
+```{r trendALL, echo=F, out.width='.75\\linewidth', fig.cap=paste0("Anomalies (\\%) of the daily all-sky SDR from the climatological mean for the period ", year(min(ALL_1_daily_DESEAS$Date)), " -- ", year(max(ALL_1_daily_DESEAS$Date)), ".
 The black line is the long term linear trend.")}
@@
 The trend for clear skies is largest in winter
-($`r signif(spp$"Trend [%/year]"[spp$Season == "Winter" & spp$"Sky condition" == "Clear skies"], 2)`$) 
+($`r signif(spp$"Trend [%/year]"[spp$Season == "Winter" & spp$"Sky condition" == "Clear skies"], 2)`\,\%/y$) 
 and negligible in spring, summer and autumn.

### 2023-12-19 MDPI_submission.tex
Commit 2023-12-19 17:11

@@
 which is based on the methods of Long and
-Shi~\citetext{\citeyear{Long2008a}; \citeyear{Long2006}}. Thus,
-problematic recordings have been excluded from further processing.
-Although it is impossible to detect all false data, the large number of
-available data, and the aggregation scheme we used, ensures the quality
-of the radiation measurements used in this study.
+Shi~\citep{Long2006, Long2008a}. Thus, problematic recordings have been
+excluded from further processing. Although it is impossible to detect
+all false data, the large number of available data, and the aggregation
+scheme we used, ensures the quality of the radiation measurements used
+in this study.
 
@@
 all-sky, clear-sky and cloudy-sky). To identify the clear-cases we used
-a method proposed by \citet{Long2000} and by \citet{Reno2016}, which
-were adapted and configured for the site of Thessaloniki. We note that
-all methods have some subjectivity in the definition of clear or cloudy
-sky cases. As a result, the details of the definition are site specific
-and they rely on a combination of thresholds and comparisons with ideal
-radiation models and statistical analysis of different signal metrics.
-The CSid algorithm was calibrated with the main focus, to identify the
-presence of clouds. Despite the fine-tuning of the procedure, in a few
-marginal cases false positive or false negative results were identified
-by manual inspection. However, due to their small number, they cannot
-affect the final results of the study. For completeness, we provide
-below a brief overview of the clear-sky identification algorithm (CSid),
-along with the site specific thresholds.
+the method proposed by \citet{Reno2016}, which requires the definition
+of some site specific parameters. These parameters, were produced by an
+iterative process, as the original authors proposed and it are discussed
+in the next section.
+
+We note that all methods have some subjectivity in the definition of
+clear or cloudy sky cases. As a result, the details of the definition
+are site specific and they rely on a combination of thresholds and
+comparisons with ideal radiation models and statistical analysis of
+different signal metrics. The CSid algorithm was calibrated with the
+main focus, to identify the presence of clouds. Despite the fine-tuning
+of the procedure, in a few marginal cases false positive or false
+negative results were identified by manual inspection. However, due to
+their small number, they cannot affect the final results of the study.
+For completeness, we provide below a brief overview of the clear-sky
+identification algorithm (CSid), along with the site specific
+thresholds.
 
@@
 
-{\centering \includegraphics[width=.70\linewidth]{./images/LongtermTrends-2}
+{\centering \includegraphics[width=.75\linewidth]{./images/LongtermTrends-2}
 
@@
 other studies for the region. The trend for clear skies is largest in
-winter (\(0.4\)) and negligible in spring, summer and autumn.
+winter (\(0.4\,\%/y\)) and negligible in spring, summer and autumn.
 

### 2023-12-19 DHI_GHI_2_sza_trends.R

@@ -580,7 +580,7 @@ for (avar in vars) {
             ## empty plot
             plot(1, type = "n",
                  xlab = "",
-                 ylab = awename,
+                 ylab = "",
                  xlim = xlim,
                  ylim = ylim,
                  yaxt = "n")
@@ -593,6 +593,9 @@ for (avar in vars) {
                       tcl = -0.25)
             title(xlab = bquote("Solar zenith angle (SZA)"),
                   line = 2.5)
+
+            title(ylab = awename,
+                  line = 3.3)
 
             ## zero line
             abline(h = 0, lty = 3)

### 2023-12-19 MDPI_submission.Rmd
Commit 2023-12-19 19:11

@@
 LT$Tmod_Std..Error <- signif(LT$Tmod_Std..Error * 2, 2)
-LT$ChangeWpY       <- signif(LT$ChangeWpY,           2)
+LT$ChangeWpY       <- signif(LT$ChangeWpY,           3)
 LT$slope <- format(LT$slope, 3, drop0trailing = T)
@@
 The black line is the long term linear trend.")}
-knitr::include_graphics("./images/LongtermTrends-2.png")
+knitr::include_graphics("./images/LongtermTrends-1.png")
 ```
@@
 conclusions.
+@Lozano2023 reported a Cloud Radiative Forcing (CRF) trend of $1.22\,W/m^2/y$ for the SDR for the city of Granada Spain ($37.16^\circ\,$N, $3.61^\circ\,$E, $680\,$m\ a.s.l.). 
+ <!-- (37.16◦ N, 3.61◦ W, 680 m a.s.l.), 37° 9' 35.9994" N  3° 36' 35.9994" W-->
+We found a comparable result, for the same parameter, using the long term trend of all- and clear-skies, as follows $CRF = 1.46 - 0.501 = `r 1.46 - 0.501`\,W/m^2/y$.
 

### 2023-12-19 MDPI_submission.tex
Commit 2023-12-19 19:11

@@
 \midrule
-All skies & 0.380 & 0.120 & 0.091 & 1.5 & 10251\\
-Clear skies & 0.097 & 0.033 & 0.140 & 0.5 & 2684\\
-Cloudy skies & 0.410 & 0.180 & 0.081 & 1.2 & 4937\\
+All skies & 0.380 & 0.120 & 0.091 & 1.460 & 10251\\
+Clear skies & 0.097 & 0.033 & 0.140 & 0.501 & 2684\\
+Cloudy skies & 0.410 & 0.180 & 0.081 & 1.180 & 4937\\
 \bottomrule
@@
 
-{\centering \includegraphics[width=.75\linewidth]{./images/LongtermTrends-2}
+{\centering \includegraphics[width=.75\linewidth]{./images/LongtermTrends-1}
 
@@
 thickness are also not available for the region, we cannot draw
-quantitative conclusions.
+quantitative conclusions. \citet{Lozano2023} reported a Cloud Radiative
+Forcing (CRF) trend of \(1.22\,W/m^2/y\) for the SDR for the city of
+Granada Spain (\(37.16^\circ\,\)N, \(3.61^\circ\,\)E,
+\(680\,\)m~a.s.l.). We found a comparable result, for the same
+parameter, using the long term trend of all- and clear-skies, as follows
+\(CRF = 1.46 - 0.501 = 0.959\,W/m^2/y\).
 

### 2023-12-19 Build_172/MDPI_submission_B172.Rmd
Commit 2023-12-19 20:08

--- 2023-12-19 MDPI_submission.Rmd
@@
 conclusions.
-@Lozano2023 reported a Cloud Radiative Forcing (CRF) trend of $1.22\,W/m^2/y$ for the SDR for the city of Granada Spain ($37.16^\circ\,$N, $3.61^\circ\,$E, $680\,$m\ a.s.l.). 
- <!-- (37.16◦ N, 3.61◦ W, 680 m a.s.l.), 37° 9' 35.9994" N  3° 36' 35.9994" W-->
-We found a comparable result, for the same parameter, using the long term trend of all- and clear-skies, as follows $CRF = 1.46 - 0.501 = `r 1.46 - 0.501`\,W/m^2/y$.
+@Lozano2023 reported a Cloud Radiative Forcing (CRF) SDR trend of $1.22\,W/m^2/y$ for the city of Granada Spain ($37^\circ\,10'\,$N, $3^\circ\,37'\,$E, $680\,$m\ a.s.l.). 
+ <!-- (37.16◦ N, 3.61◦ W, 680 m a.s.l.), 37° 9' 35.9994" N  3° 36' 35.9994" W    ($37^\circ\,10'\,$N, $3^\circ\,37'\,$E, $680\,$m\ a.s.l.)-->
+We found a comparable result, for the same parameter, using the long term trend of all- and clear-skies, as follows $\text{CRF} = 1.46 - 0.501 = `r 1.46 - 0.501`\,W/m^2/y$.
 

### 2023-12-19 Build_172/MDPI_submission_B172.tex
Commit 2023-12-19 20:08

--- 2023-12-19 MDPI_submission.tex
@@
 quantitative conclusions. \citet{Lozano2023} reported a Cloud Radiative
-Forcing (CRF) trend of \(1.22\,W/m^2/y\) for the SDR for the city of
-Granada Spain (\(37.16^\circ\,\)N, \(3.61^\circ\,\)E,
-\(680\,\)m~a.s.l.). We found a comparable result, for the same
-parameter, using the long term trend of all- and clear-skies, as follows
-\(CRF = 1.46 - 0.501 = 0.959\,W/m^2/y\).
+Forcing (CRF) SDR trend of \(1.22\,W/m^2/y\) for the city of Granada
+Spain (\(37^\circ\,10'\,\)N, \(3^\circ\,37'\,\)E, \(680\,\)m~a.s.l.). We
+found a comparable result, for the same parameter, using the long term
+trend of all- and clear-skies, as follows
+\(\text{CRF} = 1.46 - 0.501 = 0.959\,W/m^2/y\).
 

### 2023-12-20 MDPI_submission.Rmd
Commit 2023-12-20 11:14

@@
 abstract: |
-  The shortwave downward solar irradiance (SDR) is an important factor that drives climate processes, production and can affect all living organisms.
+  The shortwave downward solar irradiance (SDR) is an important factor that drives climate processes, energy production and can affect all living organisms. 
   Observations of SDR at different locations around the world with different environmental characteristics have been used to investigate its long-term variability and trends at different time scales.
 Periods of positive trends are referred as brightening periods and of negative trends as dimming periods.
-  Here we studied 30 years of pyranometer data in Thessaloniki, Greece, under three types of sky conditions (clear sky, cloudy sky and all sky).
+  In this study we have used 30 years of pyranometer data in Thessaloniki, Greece, to investigate the variability of SDR under three types of sky conditions (clear-, cloudy- and all-sky). 
   The clear-sky data were identified by applying a cloud screening algorithm. 
-We found a positive trend of $0.38\,\%/\text{year}$ for all-sky, $0.35\,\%/\text{year}$ for clear-sky conditions, and $-0.28\,\%/\text{year}$ for cloudy conditions.
-  We have also investigated the consistency of these trends, their seasonal variability, and the effect of the solar zenith angle. 
-  We have found that for all-sky and clear-sky conditions the SDR trend is positive in winter ($0.7$ and $0.8\,\%/\text{year}$, respectively) and autumn ($~0.4\,\%/\text{year}$), while under cloudy skies the trend is negative ($-0.9\,\%/\text{year}$ in winter and $-0.4\,\%/\text{year}$ in autumn).
-  In spring and summer the trend is very close to zero, irrespective of sky conditions. 
-  The SDR trend is increasing with increasing solar zenith angle, except under cloudy skies where the trend is highly variable and close to zero.
-  Finally, we identified some anomalies in the long term SDR trends for all sky conditions by examining the cumulative sums of monthly anomalies from the climatological mean.
+  We have found a positive trend of $0.38\,\%/\text{year}$ for all-sky, $\sim 0.1\,\%/\text{year}$ for clear-sky, and $-0.41\,\%/\text{year}$ for cloudy conditions.
+  The consistency of these trends, their seasonal variability, and the effect of the solar zenith angle have also been investigated.
+
+  We have found that for all conditions the SDR trend is stronger in winter ($0.7$, $0.4$ and $0.76\,\%/\text{year}$, for all- clear- and cloudy skies respectively).
+For all- and cloudy-skis, in autumn there are the next larger trends ($0.42$ and $0.19\,\%/\text{year}$, respectively), the rest cases bare a negligable seasonal trend.
+  The SDR trend is increasing with increasing solar zenith angle (SZA) for all-skies, from about $0.2$ to $0.6\,\%/\text{year}$.
+For clear- and cloudy-skies remains low, about $0.1$ -- $0.2\,\%/\text{year}\,\%/\text{year}$ for all SZAs below $\sim 75^\circ$, and rizes rapidly for the larger SZAs.
+
+  Finally, we are discussing shorter-term variations in SDR anomalies by examining the patterns of the cumulative sums of monthly anomalies from the climatological mean, both before and after removing the long-term trend.
 
@@
 <!-- Trends Global -->
-On global scale, an artificial Intelligence aided spatial analysis on continental level with data from multiple stations reach similar conclusions for these regions and for the global trend [@Yuan2021].
+On global scale, artificial Intelligence aided spatial analysis on continental level with data from multiple stations reach similar conclusions for these regions and for the global trend [@Yuan2021].
 
@@
 <!-- goals -->
-We reevaluated and extended the dataset used by @Bais2013, we applied a different algorithm for the identification of clear-/cloud-sky instances [@Reno2016; @Reno2012a], and we derived the SDR trends for the period of study, under different sky conditions (all-sky, clear-sky and cloud-sky).
+We re-evaluated and extended the dataset used by @Bais2013, we applied a different algorithm for the identification of clear-/cloud-sky instances [@Reno2016; @Reno2012a], and we derived the SDR trends for the period of study, under different sky conditions (all-sky, clear-sky and cloud-sky).
 Finally, we investigated the dependence of the trends on solar zenith angle and season.
@@
 to unequal length of the two datasets.
-There are four distinct steps in the creation of the dataset analyzed here:
+There are four distinct steps in the creation of the dataset analysed here:
 a)\ the acquisition of radiation measurements from the sensors,
@@
 @Reno2016, which requires the definition of some site specific parameters.
-These parameters, were produced by an iterative process, as the original authors proposed and it are discussed in the next section.
+These parameters, were produced by an iterative process, as the original authors proposed, and are discussed in the next section.
 
@@
 We note that all methods have some subjectivity in the definition of clear or cloudy sky cases.
-As a result, the details of the definition are site specific and they rely on a combination of thresholds and comparisons with ideal radiation models and statistical analysis of different signal metrics.
+As a result, the details of the definition are site specific, and they rely on a combination of thresholds and comparisons with ideal radiation models and statistical analysis of different signal metrics.
 The CSid algorithm was calibrated with the main focus, to identify the presence of clouds. Despite the fine-tuning
@@
 
-The linear trends were calculated using an first order autoregressive model
+The linear trends were calculated using a first order autoregressive model
  with lag of 1 day with fitting method of the "maximum likelihood" [@Gardner1980; @Jones1980], by implementing the function "arima" from the library "stats" of the R programming language [@RCT2023].
@@
 The clear-sky trend is
-very small suggesting a small effect from aerosols and water vapor which
+very small suggesting a small effect from aerosols and water vapour which
 are the dominant factors of the SDR variability. 
@@
 For time series with a uniform trend, we would expect
-the CUSUMs of the anomalies to have a symmetric 'V' shape centered
+the CUSUMs of the anomalies to have a symmetric 'V' shape centred
 around the middle of the series. This would indicate that the anomalies
@@
 upward part to present indicating anomalies above the climatology. The
-CUSUMs for cloudy-skies (Figure\ \ref{fig:cusummonth-3}), show a similar behavior with some short-term
+CUSUMs for cloudy-skies (Figure\ \ref{fig:cusummonth-3}), show a similar behaviour with some short-term
 differences that do not change the overall pattern.
@@
 The trends under clear- and cloudy-sky conditions are of the same
-direction and it would be expected that their sum is similar to the
+direction, and it would be expected that their sum is similar to the
 all-sky trend. This does not happen, especially for winter, likely due
@@
 
-We have analyzed a 30 year dataset of SRD measurements in Thessaloniki Greece 
+We have analysed a 30-year dataset of SRD measurements in Thessaloniki Greece 
 (`r paste(year(min(ALL_1_daily_DESEAS$Date)), " -- ", year(max(ALL_1_daily_DESEAS$Date)))`) aiming to identify the long term variability of
@@
 not available in order to quantify their effect on SDR. The attenuation
-of SDR by aerosols over Europe has been proposed as major factor by @Wild2021.
+of SDR by aerosols over Europe have been proposed as major factor by @Wild2021.
 The brightening effect on SDR under cloudy-sky conditions
@@
 
-Using the CUSUMs of the monthly departures for all- and cloudy-skies, we observed a 20 year period starting around 2000 where the CUSUMs remain relatively stable, with a steep decline before and a steep increase after.
+Using the CUSUMs of the monthly departures for all- and cloudy-skies, we observed a 20-year period starting around 2000 where the CUSUMs remain relatively stable, with a steep decline before and a steep increase after.
 The rather smooth course of the CUSUMs suggests that no important

### 2023-12-20 MDPI_submission.tex
Commit 2023-12-20 11:14

@@
 \abstract{The shortwave downward solar irradiance (SDR) is an important
-factor that drives climate processes, production and can affect all
-living organisms. Observations of SDR at different locations around the
-world with different environmental characteristics have been used to
+factor that drives climate processes, energy production and can affect
+all living organisms. Observations of SDR at different locations around
+the world with different environmental characteristics have been used to
 investigate its long-term variability and trends at different time
 scales. Periods of positive trends are referred as brightening periods
-and of negative trends as dimming periods. Here we studied 30 years of
-pyranometer data in Thessaloniki, Greece, under three types of sky
-conditions (clear sky, cloudy sky and all sky). The clear-sky data were
-identified by applying a cloud screening algorithm. We found a positive
-trend of \(0.38\,\%/\text{year}\) for all-sky, \(0.35\,\%/\text{year}\)
-for clear-sky conditions, and \(-0.28\,\%/\text{year}\) for cloudy
-conditions. We have also investigated the consistency of these trends,
-their seasonal variability, and the effect of the solar zenith angle. We
-have found that for all-sky and clear-sky conditions the SDR trend is
-positive in winter (\(0.7\) and \(0.8\,\%/\text{year}\), respectively)
-and autumn (\(~0.4\,\%/\text{year}\)), while under cloudy skies the
-trend is negative (\(-0.9\,\%/\text{year}\) in winter and
-\(-0.4\,\%/\text{year}\) in autumn). In spring and summer the trend is
-very close to zero, irrespective of sky conditions. The SDR trend is
-increasing with increasing solar zenith angle, except under cloudy skies
-where the trend is highly variable and close to zero. Finally, we
-identified some anomalies in the long term SDR trends for all sky
-conditions by examining the cumulative sums of monthly anomalies from
-the climatological mean.}
+and of negative trends as dimming periods. In this study we have used 30
+years of pyranometer data in Thessaloniki, Greece, to investigate the
+variability of SDR under three types of sky conditions (clear-, cloudy-
+and all-sky). The clear-sky data were identified by applying a cloud
+screening algorithm. We have found a positive trend of
+\(0.38\,\%/\text{year}\) for all-sky, \(\sim 0.1\,\%/\text{year}\) for
+clear-sky, and \(-0.41\,\%/\text{year}\) for cloudy conditions. The
+consistency of these trends, their seasonal variability, and the effect
+of the solar zenith angle have also been investigated.
+
+We have found that for all conditions the SDR trend is stronger in
+winter (\(0.7\), \(0.4\) and \(0.76\,\%/\text{year}\), for all- clear-
+and cloudy skies respectively). For all- and cloudy-skis, in autumn
+there are the next larger trends (\(0.42\) and \(0.19\,\%/\text{year}\),
+respectively), the rest cases bare a negligable seasonal trend. The SDR
+trend is increasing with increasing solar zenith angle (SZA) for
+all-skies, from about \(0.2\) to \(0.6\,\%/\text{year}\). For clear- and
+cloudy-skies remains low, about \(0.1\) --
+\(0.2\,\%/\text{year}\,\%/\text{year}\) for all SZAs below
+\(\sim 75^\circ\), and rizes rapidly for the larger SZAs.
+
+Finally, we are discussing shorter-term variations in SDR anomalies by
+examining the patterns of the cumulative sums of monthly anomalies from
+the climatological mean, both before and after removing the long-term
+trend.}
 
@@
 point around 1980, for stations in Central Europe \citep{Wild2021} and
-Brazil \citep{Yamasoe2021}. On global scale, an artificial Intelligence
+Brazil \citep{Yamasoe2021}. On global scale, artificial Intelligence
 aided spatial analysis on continental level with data from multiple
@@
 Thessaloniki, Greece, for the period 1993 to 2023, as derived from a
-CM-21 pyranometer. We reevaluated and extended the dataset used by
+CM-21 pyranometer. We re-evaluated and extended the dataset used by
 \citet{Bais2013}, we applied a different algorithm for the
@@
 to unequal length of the two datasets. There are four distinct steps in
-the creation of the dataset analyzed here: a)~the acquisition of
+the creation of the dataset analysed here: a)~the acquisition of
 radiation measurements from the sensors, b)~the data quality check,
@@
 of some site specific parameters. These parameters, were produced by an
-iterative process, as the original authors proposed and it are discussed
+iterative process, as the original authors proposed, and are discussed
 in the next section.
@@
 clear or cloudy sky cases. As a result, the details of the definition
-are site specific and they rely on a combination of thresholds and
+are site specific, and they rely on a combination of thresholds and
 comparisons with ideal radiation models and statistical analysis of
@@
 
-The linear trends were calculated using an first order autoregressive
+The linear trends were calculated using a first order autoregressive
 model with lag of 1 day with fitting method of the ``maximum
@@
 clear-sky trend is very small suggesting a small effect from aerosols
-and water vapor which are the dominant factors of the SDR variability.
+and water vapour which are the dominant factors of the SDR variability.
 In contrast the large positive trend of SDR under cloudy skies can be
@@
 conditions. For time series with a uniform trend, we would expect the
-CUSUMs of the anomalies to have a symmetric `V' shape centered around
-the middle of the series. This would indicate that the anomalies are
-evenly distributed around the climatological mean, and for a positive
-uniform trend, the first half is below and the second half above the
+CUSUMs of the anomalies to have a symmetric `V' shape centred around the
+middle of the series. This would indicate that the anomalies are evenly
+distributed around the climatological mean, and for a positive uniform
+trend, the first half is below and the second half above the
 climatological mean. In our case, there is a more complex evolution of
@@
 The CUSUMs for cloudy-skies (Figure~\ref{fig:cusummonth-3}), show a
-similar behavior with some short-term differences that do not change the
-overall pattern. For clear skies (Figure~\ref{fig:cusummonth-2}), a
+similar behaviour with some short-term differences that do not change
+the overall pattern. For clear skies (Figure~\ref{fig:cusummonth-2}), a
 monotonic downward tendency is evident until 2004, suggesting that the
@@
 
-We have analyzed a 30 year dataset of SRD measurements in Thessaloniki
+We have analysed a 30-year dataset of SRD measurements in Thessaloniki
 Greece (1993 -- 2023) aiming to identify the long term variability of
@@
 not available in order to quantify their effect on SDR. The attenuation
-of SDR by aerosols over Europe has been proposed as major factor by
+of SDR by aerosols over Europe have been proposed as major factor by
 \citet{Wild2021}. The brightening effect on SDR under cloudy-sky
@@
 Using the CUSUMs of the monthly departures for all- and cloudy-skies, we
-observed a 20 year period starting around 2000 where the CUSUMs remain
+observed a 20-year period starting around 2000 where the CUSUMs remain
 relatively stable, with a steep decline before and a steep increase

### 2023-12-20 appendix.tex
Commit 2023-12-19 17:11
Commit 2023-12-19 18:21
Commit 2023-12-20 11:14

@@
     {\centering
-        \includegraphics[width=0.7\linewidth]{./images/LongtermTrends-7}
+        \includegraphics[width=0.75\linewidth]{./images/LongtermTrends-2}
 
     }
-    \caption{Anomalies (\%) of the daily clear-sky SDR, relative to climatological values for1993 - 2023.
+    \caption{Anomalies (\%) of the daily clear-sky SDR, relative to climatological values for 1993 -- 2023.
 The black line shows the long term linear trend for clear-sky conditions.}\label{fig:trendCLEAR}
@@
     {\centering
-        \includegraphics[width=0.7\linewidth]{./images/LongtermTrends-10}
+        \includegraphics[width=0.75\linewidth]{./images/LongtermTrends-3}
 
     }
-    \caption{Anomalies (\%) of the daily cloud-sky SDR, relative to climatological values for 1993 - 2023.
+    \caption{Anomalies (\%) of the daily cloud-sky SDR, relative to climatological values for 1993 -- 2023.
 The black line shows the long term linear trend for cloud-sky conditions.}\label{fig:trendCLOUD}

### 2023-12-20 MDPI_submission.Rmd
Commit 2023-12-20 12:10

@@
   The clear-sky data were identified by applying a cloud screening algorithm. 
-  We have found a positive trend of $0.38\,\%/\text{year}$ for all-sky, $\sim 0.1\,\%/\text{year}$ for clear-sky, and $-0.41\,\%/\text{year}$ for cloudy conditions.
+  We have found a positive trend of $0.38\,\%/\text{year}$ for all-sky, $\sim 0.1\,\%/\text{year}$ for clear-sky, and $0.41\,\%/\text{year}$ for cloudy conditions.
   The consistency of these trends, their seasonal variability, and the effect of the solar zenith angle have also been investigated.
-
-  We have found that for all conditions the SDR trend is stronger in winter ($0.7$, $0.4$ and $0.76\,\%/\text{year}$, for all- clear- and cloudy skies respectively).
-For all- and cloudy-skis, in autumn there are the next larger trends ($0.42$ and $0.19\,\%/\text{year}$, respectively), the rest cases bare a negligable seasonal trend.
-  The SDR trend is increasing with increasing solar zenith angle (SZA) for all-skies, from about $0.2$ to $0.6\,\%/\text{year}$.
-For clear- and cloudy-skies remains low, about $0.1$ -- $0.2\,\%/\text{year}\,\%/\text{year}$ for all SZAs below $\sim 75^\circ$, and rizes rapidly for the larger SZAs.
-
+  Under all three sky categories the SDR trend is stronger in winter with $0.7$, $0.4$ and $0.76\,\%/\text{year}$, respectively for all-, clear- and cloudy-skies.
+In automn, there are the next larger trends of $0.42$ and $0.19\,\%/\text{year}$, for all- and cloudy-skies respectively.
+  The rest seasonal trends are significant smaller, close to zero, with clear- and cloudy-skies having a negative trend for summer.
+  The SDR trend is increasing with increasing solar zenith angle, except under cloudy skies where the trend is highly variable and close to zero.
   Finally, we are discussing shorter-term variations in SDR anomalies by examining the patterns of the cumulative sums of monthly anomalies from the climatological mean, both before and after removing the long-term trend.
@@
 It can also affect solar and agricultural production, and all living organisms.
-Studies of SDR variability, have identified some distinct SDR trends on different regions of the world on different time periods.
+Studies of SDR variability have identified some distinct SDR trends on different regions of the world on different time periods.
 The term 'brightening' is generally used to describe periods of positive SDR trend, and 'dimming' for negative trend [@Wild2009].
@@
 <!-- Trends Europe Brazil -->
-A similar pattern was identified, with the breaking point around 1980, for stations in Central Europe [@Wild2021] and Brazil [@Yamasoe2021].
+A similar pattern was identified, with a breaking point around 1980, for stations in Central Europe [@Wild2021] and Brazil [@Yamasoe2021].
 <!-- Trends Global -->
-On global scale, artificial Intelligence aided spatial analysis on continental level with data from multiple stations reach similar conclusions for these regions and for the global trend [@Yuan2021].
+On global scale, an Artificial Intelligence aided spatial analysis on continental level with data from multiple stations reached similar conclusions for these regions and for the global trend [@Yuan2021].
 
@@
 Those interactions, among other factors, have been analysed with models [@Li2016; @Samset2018], showing the existence of feedback mechanisms between the two.
-Similar findings have been shown in observational data [@Schwarz2020; @Ohvril2009; @Zerefos2009; @Xia2007 and references therein].
-<!-- Sun variability: should cite that there is no significant trend found Wild2021 -->
+Similar findings have been from the analysis of observations at other locations [@Schwarz2020; @Ohvril2009; @Zerefos2009; @Xia2007 and references therein].
+In the Mediterranean region aerosols have been recognized as an important factor affecting the penetration of solar radiation to the surface [@Fountoulakis2016; @Siomos2018; @Gkikas2013; @Lozano2021].
+These studies investigated the long-term trend in aerosol optical depth, which has been found to decrease in the last three decades, the transport and composition of aerosols, and their radiative effects.
 
 Due to the significant spatial and temporal variability of the trends, and the contributing factors, there is a constant need to monitor and investigate SDR in different sites in order to estimate the degree of variability, and its relation to the local conditions.
-In this study, we examine the trends of SDR, using ground-based measurements at Thessaloniki, Greece, for the period `r year(min(ALL_1_daily_DESEAS$Date))` to `r year(max(ALL_1_daily_DESEAS$Date))`, as derived from a CM-21 pyranometer.
+In this study, we examine the trends of SDR, using ground-based measurements at Thessaloniki, Greece, for the period `r year(min(ALL_1_daily_DESEAS$Date))` to `r year(max(ALL_1_daily_DESEAS$Date))`.
 <!-- goals -->
@@
 The SDR data were measured with a Kipp & Zonen CM-21 pyranometer operating continuously at the Laboratory of Atmospheric Physics of the Aristotle University of Thessaloniki
-($40^\circ\,38'\,$N, $22^\circ\,57'\,$E, $80\,$m\ a.s.l.) since 1993.
+($40^\circ\,38'\,$N, $22^\circ\,57'\,$E, $80\,$m\ a.s.l.).
 Here we use data for the period from
@@
 The monitoring site is located
-near the city centre thus we expect that measurements are affected by the urban environment.
+near the city centre thus we expect that measurements are affected by the urban environment, mainly by aerosols.
 During the study period, the pyranometer has been
 independently calibrated three times at the Meteorologisches
-Observatorium Lindenberg, DWD, verifying that stability of the
+Observatorium Lindenberg, DWD, verifying that the stability of the
 instrument's sensitivity is better than $0.7\%$ relative to the initial
@@
 The DNI data were used as auxiliary data
-to support the selection of appropriate thresholds in the clear sky
-identification algorithm (CSid), which is discussed later. It is noted
+to support the selection of appropriate thresholds in the clear-sky
+identification algorithm (CSid), which is discussed in sect. \ref{CDIDalgorithm}. 
+It is noted
 that the limited dataset of DNI was not used for the identification of
@@
 b)\ the data quality check,
-c)\ the identification of "clear sky" conditions from the radiometric data, and
+c)\ the identification of "clear sky" conditions from the SDR data, and
 d)\ the aggregation of data and trend analysis.
@@
 <!-- Data selection -->
-Due to the significant measurement uncertainty when the Sun is near the horizon, we have excluded all measurements with SZA greater than $`r 90-MIN_ELEVA`^\circ$.
+Due to the significant measurement uncertainty when the Sun is near the horizon, we have excluded all measurements with solar zenith angle (SZA) greater than $`r 90-MIN_ELEVA`^\circ$.
 Moreover, due to obstructions around the site (hills and buildings) which block the direct irradiance, we excluded data with azimuth angle in the range
@@
 @Reno2016, which requires the definition of some site specific parameters.
-These parameters, were produced by an iterative process, as the original authors proposed, and are discussed in the next section.
+These parameters were determined by an iterative process, as the original authors proposed, and are discussed in the next section.
 
@@
 of the procedure, in a few marginal cases false positive or false
-negative results were identified by manual inspection. However, due to
+negative results were identified by manual inspection. 
+However, due to
 their small number, they cannot affect the final results of the study.
-For completeness, we provide below a brief overview of the clear-sky
-identification algorithm (CSid), along with the site specific
-thresholds.
+For completeness, we provide below a brief overview of the CSid algorithm, along with the site specific thresholds.
 
@@
 
-## The clear sky identification algorithm
+## The clear sky identification algorithm {#CDIDalgorithm}
 
@@
 \end{equation}
-where $\theta$ is the solar zenith angle (SZA).
+where $\theta$ is the SZA.
 

### 2023-12-20 MDPI_submission.tex
Commit 2023-12-20 12:10

@@
 \(0.38\,\%/\text{year}\) for all-sky, \(\sim 0.1\,\%/\text{year}\) for
-clear-sky, and \(-0.41\,\%/\text{year}\) for cloudy conditions. The
+clear-sky, and \(0.41\,\%/\text{year}\) for cloudy conditions. The
 consistency of these trends, their seasonal variability, and the effect
-of the solar zenith angle have also been investigated.
-
-We have found that for all conditions the SDR trend is stronger in
-winter (\(0.7\), \(0.4\) and \(0.76\,\%/\text{year}\), for all- clear-
-and cloudy skies respectively). For all- and cloudy-skis, in autumn
-there are the next larger trends (\(0.42\) and \(0.19\,\%/\text{year}\),
-respectively), the rest cases bare a negligable seasonal trend. The SDR
-trend is increasing with increasing solar zenith angle (SZA) for
-all-skies, from about \(0.2\) to \(0.6\,\%/\text{year}\). For clear- and
-cloudy-skies remains low, about \(0.1\) --
-\(0.2\,\%/\text{year}\,\%/\text{year}\) for all SZAs below
-\(\sim 75^\circ\), and rizes rapidly for the larger SZAs.
-
+of the solar zenith angle have also been investigated. Under all three
+sky categories the SDR trend is stronger in winter with \(0.7\), \(0.4\)
+and \(0.76\,\%/\text{year}\), respectively for all-, clear- and
+cloudy-skies. In automn, there are the next larger trends of \(0.42\)
+and \(0.19\,\%/\text{year}\), for all- and cloudy-skies respectively.
+The rest seasonal trends are significant smaller, close to zero, with
+clear- and cloudy-skies having a negative trend for summer. The SDR
+trend is increasing with increasing solar zenith angle, except under
+cloudy skies where the trend is highly variable and close to zero.
 Finally, we are discussing shorter-term variations in SDR anomalies by
@@
 agricultural production, and all living organisms. Studies of SDR
-variability, have identified some distinct SDR trends on different
+variability have identified some distinct SDR trends on different
 regions of the world on different time periods. The term `brightening'
@@
 reported until about 2000, followed by a brightening period
-\citep{Yang2021}. A similar pattern was identified, with the breaking
+\citep{Yang2021}. A similar pattern was identified, with a breaking
 point around 1980, for stations in Central Europe \citep{Wild2021} and
-Brazil \citep{Yamasoe2021}. On global scale, artificial Intelligence
+Brazil \citep{Yamasoe2021}. On global scale, an Artificial Intelligence
 aided spatial analysis on continental level with data from multiple
-stations reach similar conclusions for these regions and for the global
-trend \citep{Yuan2021}.
+stations reached similar conclusions for these regions and for the
+global trend \citep{Yuan2021}.
 
@@
 \citep{Li2016, Samset2018}, showing the existence of feedback mechanisms
-between the two. Similar findings have been shown in observational data
-\citep[ and references
-therein]{Schwarz2020, Ohvril2009, Zerefos2009, Xia2007}.
+between the two. Similar findings have been from the analysis of
+observations at other locations \citep[ and references
+therein]{Schwarz2020, Ohvril2009, Zerefos2009, Xia2007}. In the
+Mediterranean region aerosols have been recognized as an important
+factor affecting the penetration of solar radiation to the surface
+\citep{Fountoulakis2016, Siomos2018, Gkikas2013, Lozano2021}. These
+studies investigated the long-term trend in aerosol optical depth, which
+has been found to decrease in the last three decades, the transport and
+composition of aerosols, and their radiative effects.
 
@@
 examine the trends of SDR, using ground-based measurements at
-Thessaloniki, Greece, for the period 1993 to 2023, as derived from a
-CM-21 pyranometer. We re-evaluated and extended the dataset used by
-\citet{Bais2013}, we applied a different algorithm for the
-identification of clear-/cloud-sky instances
+Thessaloniki, Greece, for the period 1993 to 2023. We re-evaluated and
+extended the dataset used by \citet{Bais2013}, we applied a different
+algorithm for the identification of clear-/cloud-sky instances
 \citep{Reno2016, Reno2012a}, and we derived the SDR trends for the
@@
 Aristotle University of Thessaloniki (\(40^\circ\,38'\,\)N,
-\(22^\circ\,57'\,\)E, \(80\,\)m~a.s.l.) since 1993. Here we use data for
-the period from 1993-04-13 to 2023-04-13. The monitoring site is located
-near the city centre thus we expect that measurements are affected by
-the urban environment. During the study period, the pyranometer has been
-independently calibrated three times at the Meteorologisches
-Observatorium Lindenberg, DWD, verifying that stability of the
-instrument's sensitivity is better than \(0.7\%\) relative to the
-initial calibration by the manufacturer. Along with SDR, the direct beam
-radiation (DNI) is also measured with a collocated Kipp \& Zonen CHP-1
-pyrheliometer since 2016-04-01. The DNI data were used as auxiliary data
-to support the selection of appropriate thresholds in the clear sky
-identification algorithm (CSid), which is discussed later. It is noted
-that the limited dataset of DNI was not used for the identification of
-clear sky cases in the entire SDR series to avoid any selection bias due
-to unequal length of the two datasets. There are four distinct steps in
-the creation of the dataset analysed here: a)~the acquisition of
-radiation measurements from the sensors, b)~the data quality check,
-c)~the identification of ``clear sky'' conditions from the radiometric
-data, and d)~the aggregation of data and trend analysis.
+\(22^\circ\,57'\,\)E, \(80\,\)m~a.s.l.). Here we use data for the period
+from 1993-04-13 to 2023-04-13. The monitoring site is located near the
+city centre thus we expect that measurements are affected by the urban
+environment, mainly by aerosols. During the study period, the
+pyranometer has been independently calibrated three times at the
+Meteorologisches Observatorium Lindenberg, DWD, verifying that the
+stability of the instrument's sensitivity is better than \(0.7\%\)
+relative to the initial calibration by the manufacturer. Along with SDR,
+the direct beam radiation (DNI) is also measured with a collocated Kipp
+\& Zonen CHP-1 pyrheliometer since 2016-04-01. The DNI data were used as
+auxiliary data to support the selection of appropriate thresholds in the
+clear-sky identification algorithm (CSid), which is discussed in sect.
+\ref{CDIDalgorithm}. It is noted that the limited dataset of DNI was not
+used for the identification of clear sky cases in the entire SDR series
+to avoid any selection bias due to unequal length of the two datasets.
+There are four distinct steps in the creation of the dataset analysed
+here: a)~the acquisition of radiation measurements from the sensors,
+b)~the data quality check, c)~the identification of ``clear sky''
+conditions from the SDR data, and d)~the aggregation of data and trend
+analysis.
 
@@
 Due to the significant measurement uncertainty when the Sun is near the
-horizon, we have excluded all measurements with SZA greater than
-\(85^\circ\). Moreover, due to obstructions around the site (hills and
-buildings) which block the direct irradiance, we excluded data with
-azimuth angle in the range \(58^{\circ}\) - \(120^{\circ}\) and with SZA
-greater than \(78^{\circ}\). To make the measurements comparable
-throughout the dataset, we adjusted all one-minute data to the mean Sun
-- Earth distance. Subsequently, we adjusted all measurements to the
-Total Solar Irradiance (TSI) at \(1\,\text{au}\), in order to compensate
-for the Sun's intensity variability, using a time series of satellite
-TSI observations. The TSI data we used are part of the ``NOAA Climate
-Data Record of Total Solar Irradiance'' dataset \citep{Coddington2005}.
-The initial daily values of this dataset were interpolated to match the
-time step of our measurements.
+horizon, we have excluded all measurements with solar zenith angle (SZA)
+greater than \(85^\circ\). Moreover, due to obstructions around the site
+(hills and buildings) which block the direct irradiance, we excluded
+data with azimuth angle in the range \(58^{\circ}\) - \(120^{\circ}\)
+and with SZA greater than \(78^{\circ}\). To make the measurements
+comparable throughout the dataset, we adjusted all one-minute data to
+the mean Sun - Earth distance. Subsequently, we adjusted all
+measurements to the Total Solar Irradiance (TSI) at \(1\,\text{au}\), in
+order to compensate for the Sun's intensity variability, using a time
+series of satellite TSI observations. The TSI data we used are part of
+the ``NOAA Climate Data Record of Total Solar Irradiance'' dataset
+\citep{Coddington2005}. The initial daily values of this dataset were
+interpolated to match the time step of our measurements.
 
@@
 the method proposed by \citet{Reno2016}, which requires the definition
-of some site specific parameters. These parameters, were produced by an
+of some site specific parameters. These parameters were determined by an
 iterative process, as the original authors proposed, and are discussed
@@
 their small number, they cannot affect the final results of the study.
-For completeness, we provide below a brief overview of the clear-sky
-identification algorithm (CSid), along with the site specific
-thresholds.
+For completeness, we provide below a brief overview of the CSid
+algorithm, along with the site specific thresholds.
 
-\hypertarget{the-clear-sky-identification-algorithm}{%
+\hypertarget{CDIDalgorithm}{%
 \subsection{The clear sky identification
-algorithm}\label{the-clear-sky-identification-algorithm}}
+algorithm}\label{CDIDalgorithm}}
 
@@
 \text{SDR}_\text{Haurwitz} = 1098 \times \cos(\theta) \times \exp \left( \frac{ - 0.057}{\cos(\theta)} \right) \label{eq:hau}
-\end{equation} where \(\theta\) is the solar zenith angle (SZA).
+\end{equation} where \(\theta\) is the SZA.
 

### 2023-12-20 MDPI_submission.Rmd
Commit 2023-12-20 13:17

@@
 Those interactions, among other factors, have been analysed with models [@Li2016; @Samset2018], showing the existence of feedback mechanisms between the two.
-Similar findings have been from the analysis of observations at other locations [@Schwarz2020; @Ohvril2009; @Zerefos2009; @Xia2007 and references therein].
+Similar findings have been shown from the analysis of observations at other locations [@Schwarz2020; @Ohvril2009; @Zerefos2009; @Xia2007 and references therein].
 In the Mediterranean region aerosols have been recognized as an important factor affecting the penetration of solar radiation to the surface [@Fountoulakis2016; @Siomos2018; @Gkikas2013; @Lozano2021].
@@
 @Reno2016, which requires the definition of some site specific parameters.
-These parameters were determined by an iterative process, as the original authors proposed, and are discussed in the next section.
+These parameters were determined by an iterative process, as the original authors proposed and are discussed in the next section.
 
@@
 We note that all methods have some subjectivity in the definition of clear or cloudy sky cases.
-As a result, the details of the definition are site specific, and they rely on a combination of thresholds and comparisons with ideal radiation models and statistical analysis of different signal metrics.
+As a result, the details of the definition are site specific and they rely on a combination of thresholds and comparisons with ideal radiation models and statistical analysis of different signal metrics.
 The CSid algorithm was calibrated with the main focus, to identify the presence of clouds. Despite the fine-tuning
@@
 \end{equation}
-where: $n$ is the total number of daylight data, $\text{SDR}_{\text{CSid},i}$ are the data identified as clear sky by CSid, $a$ is a site-specific adjustment factor, and $\text{SDR}_{\text{testCSref},i}$ is the SDR derived by any of the
+where: $n$ is the total number of daylight data, $\text{SDR}_{\text{CSid},i}$ are the data identified as clear-sky by CSid, $a$ is a site-specific adjustment factor, and $\text{SDR}_{\text{testCSref},i}$ is the SDR derived by any of the
 tested clear-sky radiation models.
@@
 During the optimization, eight simple clear sky radiation models were tested (namely, Daneshyar-Paltridge-Proctor, Kasten-Czeplak, Haurwitz, Berger-Duffie, Adnot-Bourges-Campana-Gicquel, Robledo-Soler, Kasten and Ineichen-Perez), with a wide range of factors.
-These models are described in more details by @Reno2012 and evaluated by @Reno2016.
+These models are described in more detail by @Reno2012 and are evaluated by @Reno2016.
 We found, that Haurwitz's model, adjusted with the factor $a = `r signif(alpha,3)`$ yields one of the lowest root mean squared errors (RMSE), while the procedure manages to characterize successfully the majority of the data.
@@
 Each criterion was applied for a running window of $`r nt`$ consecutive one-minute measurements, and the characterization was assigned to the central datum of the window.
-Each parameter, was calculated both from the observations and the reference clear sky model, for each comparison.
+Each parameter was calculated from the observations in comparison to the reference clear sky model.
 The allowable range of variation is defined by the model-derived value of the parameter multiplied by a factor plus an offset.
@@
 In order to investigate the SDR trends which are the main focus of the
-study, we implemented an appropriate aggregation scheme to the
-one-minute data to derive a series in coarser time-scales.
+study, we implemented an aggregation scheme to the
+one-minute data to derive series in coarser time-scales.
 To preserve the representativeness of the data we used the following criteria:
@@
 measurements identified as clear or cloudy respectively,
-c) monthly means were computed from daily means, for all-skies dataset 
-means were computed only when at least `r Monthly_aggegation_N_lim` 
-days of the month were available.
+c) monthly means were computed from daily means. For the all-skies dataset monthly means were computed only when at least `r Monthly_aggegation_N_lim` 
+days were available.
 Seasonal means were derived by averaging the
@@
 The linear trends were calculated using a first order autoregressive model
- with lag of 1 day with fitting method of the "maximum likelihood" [@Gardner1980; @Jones1980], by implementing the function "arima" from the library "stats" of the R programming language [@RCT2023].
-The trends were reported together with the $2\sigma$ errors (Table\ \ref{tab:trendtable}).
+ with lag of 1 day using the "maximum likelihood" fitting method [@Gardner1980; @Jones1980], by implementing the function "arima" from the library "stats" of the R programming language [@RCT2023].
+The trends were reported together with the $2\sigma$ errors.
 
@@
 
-We calculated the linear trends of SDR, from the departures of the mean daily values from the daily climatology and for the three sky conditions (Table\ \ref{tab:trendtable}).
+We calculated the linear trends of SDR, from the departures of the mean daily values from the daily climatology and for the three sky conditions.
+These are presented in Table\ \ref{tab:trendtable} which contains also the $2\sigma$ standard error, the Pearson's correlation coefficient R and the trend in absolute units.
 In Figure\ \ref{fig:trendALL} we present only the time series under all-sky conditions; the plots for clear-sky and cloud-sky conditions, are shown in the Appendix (Figures\ \ref{fig:trendCLEAR} and\  \ref{fig:trendCLOUD}).
@@
 are statistically significant at least at the $95\,\%$ confidence level, as
-they are larger than the corresponding $2\sigma$ errors.
+they are larger than the corresponding $2\sigma$ errors, despite the small values of R, which is due to the large variability of the daily values.
 The clear-sky trend is
 very small suggesting a small effect from aerosols and water vapour which
-are the dominant factors of the SDR variability. 
+are the dominant factors of the SDR variability [@Fountoulakis2016; @Siomos2018; @Yu2022].
 In contrast the large
@@
 in cloud cover and/or cloud optical depth. 
-Lack of continuous cloud
-observations that could support these findings does not allow drawing
+Lack of continuous observations of cloud optical thickness that could support these findings does not allow drawing
 firm conclusions. However, there are indications that the total

### 2023-12-20 MDPI_submission.tex
Commit 2023-12-20 13:17

@@
 \citep{Li2016, Samset2018}, showing the existence of feedback mechanisms
-between the two. Similar findings have been from the analysis of
+between the two. Similar findings have been shown from the analysis of
 observations at other locations \citep[ and references
@@
 of some site specific parameters. These parameters were determined by an
-iterative process, as the original authors proposed, and are discussed
-in the next section.
+iterative process, as the original authors proposed and are discussed in
+the next section.
 
@@
 clear or cloudy sky cases. As a result, the details of the definition
-are site specific, and they rely on a combination of thresholds and
+are site specific and they rely on a combination of thresholds and
 comparisons with ideal radiation models and statistical analysis of
@@
 \end{equation} where: \(n\) is the total number of daylight data,
-\(\text{SDR}_{\text{CSid},i}\) are the data identified as clear sky by
+\(\text{SDR}_{\text{CSid},i}\) are the data identified as clear-sky by
 CSid, \(a\) is a site-specific adjustment factor, and
@@
 Ineichen-Perez), with a wide range of factors. These models are
-described in more details by \citet{Reno2012} and evaluated by
+described in more detail by \citet{Reno2012} and are evaluated by
 \citet{Reno2016}. We found, that Haurwitz's model, adjusted with the
@@
 characterization was assigned to the central datum of the window. Each
-parameter, was calculated both from the observations and the reference
-clear sky model, for each comparison. The allowable range of variation
-is defined by the model-derived value of the parameter multiplied by a
-factor plus an offset. The factors and the offsets were determined
-empirically, by manually inspecting each filter's performance on
-selected days and adjusting them accordingly during an iterative
-process. The criteria are listed below, together with the range of
-values within which the respective parameter should fall in order to
-raise the clear-sky flag:
+parameter was calculated from the observations in comparison to the
+reference clear sky model. The allowable range of variation is defined
+by the model-derived value of the parameter multiplied by a factor plus
+an offset. The factors and the offsets were determined empirically, by
+manually inspecting each filter's performance on selected days and
+adjusting them accordingly during an iterative process. The criteria are
+listed below, together with the range of values within which the
+respective parameter should fall in order to raise the clear-sky flag:
 
@@
 In order to investigate the SDR trends which are the main focus of the
-study, we implemented an appropriate aggregation scheme to the
-one-minute data to derive a series in coarser time-scales. To preserve
-the representativeness of the data we used the following criteria: a) we
-excluded all days with less than 50\% of the expected daytime
-measurements, b) daily means for the clear-sky and cloudy-sky datasets
-were calculated only for days with more than 60\% of the expected
-daytime measurements identified as clear or cloudy respectively, c)
-monthly means were computed from daily means, for all-skies dataset
-means were computed only when at least 20 days of the month were
-available. Seasonal means were derived by averaging the monthly mean
-values in each season (winter: December - February, spring: March - May,
-etc.). The daily and monthly climatological means were derived by
-averaging the data for each day of year and calendar month,
+study, we implemented an aggregation scheme to the one-minute data to
+derive series in coarser time-scales. To preserve the representativeness
+of the data we used the following criteria: a) we excluded all days with
+less than 50\% of the expected daytime measurements, b) daily means for
+the clear-sky and cloudy-sky datasets were calculated only for days with
+more than 60\% of the expected daytime measurements identified as clear
+or cloudy respectively, c) monthly means were computed from daily means.
+For the all-skies dataset monthly means were computed only when at least
+20 days were available. Seasonal means were derived by averaging the
+monthly mean values in each season (winter: December - February, spring:
+March - May, etc.). The daily and monthly climatological means were
+derived by averaging the data for each day of year and calendar month,
 respectively. The daily and monthly datasets were deseasonalized by
@@
 The linear trends were calculated using a first order autoregressive
-model with lag of 1 day with fitting method of the ``maximum
-likelihood'' \citep{Gardner1980, Jones1980}, by implementing the
-function ``arima'' from the library ``stats'' of the R programming
-language \citep{RCT2023}. The trends were reported together with the
-\(2\sigma\) errors (Table~\ref{tab:trendtable}).
+model with lag of 1 day using the ``maximum likelihood'' fitting method
+\citep{Gardner1980, Jones1980}, by implementing the function ``arima''
+from the library ``stats'' of the R programming language
+\citep{RCT2023}. The trends were reported together with the \(2\sigma\)
+errors.
 
@@
 We calculated the linear trends of SDR, from the departures of the mean
-daily values from the daily climatology and for the three sky conditions
-(Table~\ref{tab:trendtable}). In Figure~\ref{fig:trendALL} we present
-only the time series under all-sky conditions; the plots for clear-sky
-and cloud-sky conditions, are shown in the Appendix
-(Figures~\ref{fig:trendCLEAR} and~ \ref{fig:trendCLOUD}). In the studied
-period, there is no significant break or change in the variability
-pattern of the time series. The linear trends in all three datasets are
-positive and around \(0.4\,\%/y\) for all-sky and cloudy-sky conditions,
-while for clear-skies the trend is much smaller
-(\textasciitilde{}\(0.1\,\%/y\)). The linear trends were calculated
-taking into account the autocorrelation of the time series and all three
-are statistically significant at least at the \(95\,\%\) confidence
-level, as they are larger than the corresponding \(2\sigma\) errors. The
-clear-sky trend is very small suggesting a small effect from aerosols
-and water vapour which are the dominant factors of the SDR variability.
-In contrast the large positive trend of SDR under cloudy skies can be
-attributed to reduction in cloud cover and/or cloud optical depth. Lack
-of continuous cloud observations that could support these findings does
-not allow drawing firm conclusions. However, there are indications that
-the total cloud-cover as inferred from the ERA5 analysis for the grid
-point of Thessaloniki is decreasing over the period of study.
+daily values from the daily climatology and for the three sky
+conditions. These are presented in Table~\ref{tab:trendtable} which
+contains also the \(2\sigma\) standard error, the Pearson's correlation
+coefficient R and the trend in absolute units. In
+Figure~\ref{fig:trendALL} we present only the time series under all-sky
+conditions; the plots for clear-sky and cloud-sky conditions, are shown
+in the Appendix (Figures~\ref{fig:trendCLEAR} and~
+\ref{fig:trendCLOUD}). In the studied period, there is no significant
+break or change in the variability pattern of the time series. The
+linear trends in all three datasets are positive and around
+\(0.4\,\%/y\) for all-sky and cloudy-sky conditions, while for
+clear-skies the trend is much smaller (\textasciitilde{}\(0.1\,\%/y\)).
+The linear trends were calculated taking into account the
+autocorrelation of the time series and all three are statistically
+significant at least at the \(95\,\%\) confidence level, as they are
+larger than the corresponding \(2\sigma\) errors, despite the small
+values of R, which is due to the large variability of the daily values.
+The clear-sky trend is very small suggesting a small effect from
+aerosols and water vapour which are the dominant factors of the SDR
+variability \citep{Fountoulakis2016, Siomos2018, Yu2022}. In contrast
+the large positive trend of SDR under cloudy skies can be attributed to
+reduction in cloud cover and/or cloud optical depth. Lack of continuous
+observations of cloud optical thickness that could support these
+findings does not allow drawing firm conclusions. However, there are
+indications that the total cloud-cover as inferred from the ERA5
+analysis for the grid point of Thessaloniki is decreasing over the
+period of study.
 

### 2023-12-20 MDPI_submission.Rmd
Commit 2023-12-20 14:13

@@
 There is a consensus among researchers that the major factor affecting the variability of SDR attenuation is the interactions of solar radiation with atmospheric aerosols and clouds.
-Those interactions, among other factors, have been analysed with models [@Li2016; @Samset2018], showing the existence of feedback mechanisms between the two.
+Those interactions, among other factors, have been analyzed with models [@Li2016; @Samset2018], showing the existence of feedback mechanisms between the two.
 Similar findings have been shown from the analysis of observations at other locations [@Schwarz2020; @Ohvril2009; @Zerefos2009; @Xia2007 and references therein].
@@
 The monitoring site is located
-near the city centre thus we expect that measurements are affected by the urban environment, mainly by aerosols.
+near the city center thus we expect that measurements are affected by the urban environment, mainly by aerosols.
 During the study period, the pyranometer has been
@@
 to unequal length of the two datasets.
-There are four distinct steps in the creation of the dataset analysed here:
+There are four distinct steps in the creation of the dataset analyzed here:
 a)\ the acquisition of radiation measurements from the sensors,
@@
 The clear-sky trend is
-very small suggesting a small effect from aerosols and water vapour which
+very small suggesting a small effect from aerosols and water vapor which
 are the dominant factors of the SDR variability [@Fountoulakis2016; @Siomos2018; @Yu2022].
-In contrast the large
+In contrast, the large
 positive trend of SDR under cloudy skies can be attributed to reduction
@@
 Thessaloniki is decreasing over the period of study.
+From the difference between all-sky and clear-sky SDR trends, expressed in $W/m^2/y$ using the long-term mean of the respective datasets, the radiative effect of clouds is estimated to $0.96\,W/m^2/y$.
+This estimate is similar to the cloud radiative forcing of $1.22\,W/m^2/y$ reported for Granada, Spain [@Lozano2023].
 
@@
 
-Although the year-to-year variability of the anomalies (Figure \ref{fig:trendALL} and Figures\ \ref{fig:trendCLEAR}, \ref{fig:trendCLOUD} in Appendix), shows a rather homogeneous behaviour, plots of the cumulative sums (CUSUM) [@Regier2019] of the anomalies can reveal different structures in the records of all three sky conditions.
+Although the year-to-year variability of the anomalies (Figure \ref{fig:trendALL} and Figures\ \ref{fig:trendCLEAR}, \ref{fig:trendCLOUD} in Appendix), shows a rather homogeneous behavior, plots of the cumulative sums (CUSUM) [@Regier2019] of the anomalies can reveal different structures in the records of all three sky conditions.
 For time series with a uniform trend, we would expect
-the CUSUMs of the anomalies to have a symmetric 'V' shape centred
+the CUSUMs of the anomalies to have a symmetric 'V' shape centered
 around the middle of the series. This would indicate that the anomalies
@@
 upward part to present indicating anomalies above the climatology. The
-CUSUMs for cloudy-skies (Figure\ \ref{fig:cusummonth-3}), show a similar behaviour with some short-term
+CUSUMs for cloudy-skies (Figure\ \ref{fig:cusummonth-3}), show a similar behavior with some short-term
 differences that do not change the overall pattern.
@@
 Note that SZAs less than $25^\circ$ can only occur during the warm period of the year around noon when clear skies are more
-frequent. The increasing of the trend with SZA is likely caused by the
+frequent. 
+The increasing trend with SZA is likely caused by the
 increased attenuation of SDR with SZA. The effect is larger when aerosol
@@
 the atmosphere as discussed above for the all-sky conditions. The small
-differences in the trend between morning and afternoon between $35$ and
+differences in the trend between morning and afternoon between $35^\circ$ and
 $60^\circ$ SZA is likely a result of less attenuation of SDR in the morning
@@
 
-For cloudy-sky conditions (Figure\ \ref{fig:szatrends-3}), we can not discern any significant dependence of the SDR trend with SZA as the variability of irradiance is
+For cloudy-sky conditions (Figure\ \ref{fig:szatrends-3}), we cannot discern any significant dependence of the SDR trend with SZA as the variability of irradiance is
 dominated by the cloud effects leading to insignificant
@@
 
-## Long term trends by season
+## Long term SDR trends by season
 
@@
 (Figure\ \ref{fig:seasonalALL} and Table\ \ref{tab:trendseasontable}).
+Table\ \ref{tab:trendseasontable} contains also the $2\sigma$ standard error, the Pearson's correlation coefficient R and the corresponding p-value.
+The winter linear trends exhibit generally the largest R values ranging between $0.54$ and $0.60\,\%/y$. 
 <!-- all sky -->
@@
 <!-- clear sky -->
-For clear-skies, the trend in winter is $`r signif(spp$"Trend [%/year]"[spp$Season == "Winter" & spp$"Sky condition" == "Clear skies"], 2)`\,\%/y$, and is 
+For clear-skies, the trend in winter is $`r signif(spp$"Trend [%/year]"[spp$Season == "Winter" & spp$"Sky condition" == "Clear skies"], 2)`\,\%/y$ and is 
 associated with the decreasing trend in
-aerosol optical depth [@Siomos2020], and almost half of that for
+aerosol optical depth [@Siomos2020]. Moreover, it is almost half of that for
 all-skies, which is another indication of a decreasing trend in cloud
@@
 <!-- clear sky -->
-In other seasons the clear-sky trend is still very small
+In other seasons the clear-sky trend is very small
 (below $0.1\,\%/y$). 
 <!-- cloud dsky -->
-Finally, for cloud-skies the
+Finally, for cloudy-skies the
 winter trend is the largest 
@@
 
-We have analysed a 30-year dataset of SRD measurements in Thessaloniki Greece 
+We have analyzed a 30-year dataset of SRD measurements in Thessaloniki Greece 
 (`r paste(year(min(ALL_1_daily_DESEAS$Date)), " -- ", year(max(ALL_1_daily_DESEAS$Date)))`) aiming to identify the long term variability of
@@
 there is a positive trend in SDR of $`r alltrend`\,\%/y$ (brightening). 
-A previous study [@Bais2013] for the period 1993 -- 2011 reported also a positive trend, of $0.33\,\%/y$.
-The slight increase of this trend indicates that the brightening of SDR continues and is probably caused by continuing decreases in aerosol optical depth and the optical thickness of clouds over the area.
+A previous study [@Bais2013] for the period 1993 -- 2011 reported also a positive trend of $0.33\,\%/y$.
+The slight increase of this trend indicates that the brightening of SDR continues and is likely caused by continuing decreases in aerosol optical depth and the optical thickness of clouds over the area.
 A smaller trend has
@@
 Thessaloniki is decreasing constantly at least up to 2018.
+The attenuation
+of SDR by aerosols over Europe have been proposed as major factor by @Wild2021.
 Unfortunately, for this study aerosol data for the entire period were
-not available in order to quantify their effect on SDR. The attenuation
-of SDR by aerosols over Europe have been proposed as major factor by @Wild2021.
+not available in order to quantify their effect on SDR. 
 The brightening effect on SDR under cloudy-sky conditions
@@
 conclusions.
-@Lozano2023 reported a Cloud Radiative Forcing (CRF) SDR trend of $1.22\,W/m^2/y$ for the city of Granada Spain ($37^\circ\,10'\,$N, $3^\circ\,37'\,$E, $680\,$m\ a.s.l.). 
+<!-- @Lozano2023 reported a Cloud Radiative Forcing (CRF) SDR trend of $1.22\,W/m^2/y$ for the city of Granada Spain ($37^\circ\,10'\,$N, $3^\circ\,37'\,$E, $680\,$m\ a.s.l.).  -->
  <!-- (37.16◦ N, 3.61◦ W, 680 m a.s.l.), 37° 9' 35.9994" N  3° 36' 35.9994" W    ($37^\circ\,10'\,$N, $3^\circ\,37'\,$E, $680\,$m\ a.s.l.)-->
-We found a comparable result, for the same parameter, using the long term trend of all- and clear-skies, as follows $\text{CRF} = 1.46 - 0.501 = `r 1.46 - 0.501`\,W/m^2/y$.
+<!-- We found a comparable result, for the same parameter, using the long term trend of all- and clear-skies, as follows $\text{CRF} = 1.46 - 0.501 = `r 1.46 - 0.501`\,W/m^2/y$. -->
 
@@
 for all- and cloudy-skies, respectively, in contrast to spring and summer.
-The trends for autumn are also significant but smaller (
-$`r signif(spp$"Trend [%/year]"[spp$Season == "Autumn" & spp$"Sky condition" == "All skies"], 2)`$
+The trends for autumn are also significant but smaller ($`r signif(spp$"Trend [%/year]"[spp$Season == "Autumn" & spp$"Sky condition" == "All skies"], 2)`$
 and
@@
 for all- and cloudy-skies, respectively).
-Our findings are in agreement with other studies for the region.
+<!-- Our findings are in agreement with other studies for the region. -->
 The trend for clear skies is largest in winter
@@
 Also, additional data of cloudiness, aerosols, atmospheric water vapour, etc., will allow better attribution and quantification of the effects of these factors on SRD.
-
-<!--
-similar results with Evidence for Clear‐Sky Dimming and Brightening in Central Europe_Wild2021.pdf
-include in discussion
--->

### 2023-12-20 MDPI_submission.tex
Commit 2023-12-20 14:13

@@
 radiation with atmospheric aerosols and clouds. Those interactions,
-among other factors, have been analysed with models
+among other factors, have been analyzed with models
 \citep{Li2016, Samset2018}, showing the existence of feedback mechanisms
@@
 from 1993-04-13 to 2023-04-13. The monitoring site is located near the
-city centre thus we expect that measurements are affected by the urban
+city center thus we expect that measurements are affected by the urban
 environment, mainly by aerosols. During the study period, the
@@
 to avoid any selection bias due to unequal length of the two datasets.
-There are four distinct steps in the creation of the dataset analysed
+There are four distinct steps in the creation of the dataset analyzed
 here: a)~the acquisition of radiation measurements from the sensors,
@@
 The clear-sky trend is very small suggesting a small effect from
-aerosols and water vapour which are the dominant factors of the SDR
-variability \citep{Fountoulakis2016, Siomos2018, Yu2022}. In contrast
+aerosols and water vapor which are the dominant factors of the SDR
+variability \citep{Fountoulakis2016, Siomos2018, Yu2022}. In contrast,
 the large positive trend of SDR under cloudy skies can be attributed to
@@
 analysis for the grid point of Thessaloniki is decreasing over the
-period of study.
+period of study. From the difference between all-sky and clear-sky SDR
+trends, expressed in \(W/m^2/y\) using the long-term mean of the
+respective datasets, the radiative effect of clouds is estimated to
+\(0.96\,W/m^2/y\). This estimate is similar to the cloud radiative
+forcing of \(1.22\,W/m^2/y\) reported for Granada, Spain
+\citep{Lozano2023}.
 
@@
 \ref{fig:trendALL} and Figures~\ref{fig:trendCLEAR},
-\ref{fig:trendCLOUD} in Appendix), shows a rather homogeneous behaviour,
+\ref{fig:trendCLOUD} in Appendix), shows a rather homogeneous behavior,
 plots of the cumulative sums (CUSUM) \citep{Regier2019} of the anomalies
@@
 conditions. For time series with a uniform trend, we would expect the
-CUSUMs of the anomalies to have a symmetric `V' shape centred around the
-middle of the series. This would indicate that the anomalies are evenly
-distributed around the climatological mean, and for a positive uniform
-trend, the first half is below and the second half above the
+CUSUMs of the anomalies to have a symmetric `V' shape centered around
+the middle of the series. This would indicate that the anomalies are
+evenly distributed around the climatological mean, and for a positive
+uniform trend, the first half is below and the second half above the
 climatological mean. In our case, there is a more complex evolution of
@@
 The CUSUMs for cloudy-skies (Figure~\ref{fig:cusummonth-3}), show a
-similar behaviour with some short-term differences that do not change
-the overall pattern. For clear skies (Figure~\ref{fig:cusummonth-2}), a
+similar behavior with some short-term differences that do not change the
+overall pattern. For clear skies (Figure~\ref{fig:cusummonth-2}), a
 monotonic downward tendency is evident until 2004, suggesting that the
@@
 \(25^\circ\) can only occur during the warm period of the year around
-noon when clear skies are more frequent. The increasing of the trend
-with SZA is likely caused by the increased attenuation of SDR with SZA.
-The effect is larger when aerosol and/or cloud layers are optically
-thicker, therefore, decreases in aerosol and clouds through the study
-period will result in larger positive trends of SDR at larger SZAs.
+noon when clear skies are more frequent. The increasing trend with SZA
+is likely caused by the increased attenuation of SDR with SZA. The
+effect is larger when aerosol and/or cloud layers are optically thicker,
+therefore, decreases in aerosol and clouds through the study period will
+result in larger positive trends of SDR at larger SZAs.
 
@@
 discussed above for the all-sky conditions. The small differences in the
-trend between morning and afternoon between \(35\) and \(60^\circ\) SZA
-is likely a result of less attenuation of SDR in the morning hours due
-to lesser amounts of aerosols and shallower boundary layer.
+trend between morning and afternoon between \(35^\circ\) and
+\(60^\circ\) SZA is likely a result of less attenuation of SDR in the
+morning hours due to lesser amounts of aerosols and shallower boundary
+layer.
 
-For cloudy-sky conditions (Figure~\ref{fig:szatrends-3}), we can not
+For cloudy-sky conditions (Figure~\ref{fig:szatrends-3}), we cannot
 discern any significant dependence of the SDR trend with SZA as the
@@
 
-\hypertarget{long-term-trends-by-season}{%
-\subsection{Long term trends by
-season}\label{long-term-trends-by-season}}
+\hypertarget{long-term-sdr-trends-by-season}{%
+\subsection{Long term SDR trends by
+season}\label{long-term-sdr-trends-by-season}}
 
@@
 each season of the year, using the corresponding mean monthly anomalies
-(Figure~\ref{fig:seasonalALL} and Table~\ref{tab:trendseasontable}). For
-all-sky conditions the trend in SDR in winter is the largest
-(\(0.7\,\%/y\)), followed by the trend in autumn (\(0.42\,\%/y\), a
-value close to the long term trend) both statistically significant at
-the \(95\,\%\) confidence level. In spring and summer, the trends are
-much smaller and of lesser statistical significance. These seasonal
-differences indicate a possible relation of the trends in SDR to trends
-in clouds during winter and autumn. For clear-skies, the trend in winter
-is \(0.4\,\%/y\), and is associated with the decreasing trend in aerosol
-optical depth \citep{Siomos2020}, and almost half of that for all-skies,
-which is another indication of a decreasing trend in cloud optical
-thickness. In other seasons the clear-sky trend is still very small
-(below \(0.1\,\%/y\)). Finally, for cloud-skies the winter trend is the
-largest (\(0.76\,\%/y\)) and greater than for all-skies, followed by a
-much smaller trend in autumn (\(0.19\,\%/y\)).
+(Figure~\ref{fig:seasonalALL} and Table~\ref{tab:trendseasontable}).
+Table~\ref{tab:trendseasontable} contains also the \(2\sigma\) standard
+error, the Pearson's correlation coefficient R and the corresponding
+p-value. The winter linear trends exhibit generally the largest R values
+ranging between \(0.54\) and \(0.60\,\%/y\). For all-sky conditions the
+trend in SDR in winter is the largest (\(0.7\,\%/y\)), followed by the
+trend in autumn (\(0.42\,\%/y\), a value close to the long term trend)
+both statistically significant at the \(95\,\%\) confidence level. In
+spring and summer, the trends are much smaller and of lesser statistical
+significance. These seasonal differences indicate a possible relation of
+the trends in SDR to trends in clouds during winter and autumn. For
+clear-skies, the trend in winter is \(0.4\,\%/y\) and is associated with
+the decreasing trend in aerosol optical depth \citep{Siomos2020}.
+Moreover, it is almost half of that for all-skies, which is another
+indication of a decreasing trend in cloud optical thickness. In other
+seasons the clear-sky trend is very small (below \(0.1\,\%/y\)).
+Finally, for cloudy-skies the winter trend is the largest
+(\(0.76\,\%/y\)) and greater than for all-skies, followed by a much
+smaller trend in autumn (\(0.19\,\%/y\)).
 
@@
 
-We have analysed a 30-year dataset of SRD measurements in Thessaloniki
+We have analyzed a 30-year dataset of SRD measurements in Thessaloniki
 Greece (1993 -- 2023) aiming to identify the long term variability of
@@
 (brightening). A previous study \citep{Bais2013} for the period 1993 --
-2011 reported also a positive trend, of \(0.33\,\%/y\). The slight
+2011 reported also a positive trend of \(0.33\,\%/y\). The slight
 increase of this trend indicates that the brightening of SDR continues
-and is probably caused by continuing decreases in aerosol optical depth
+and is likely caused by continuing decreases in aerosol optical depth
 and the optical thickness of clouds over the area. A smaller trend has
@@
 aerosols. \citet{Siomos2020} have shown that aerosol optical depth over
-Thessaloniki is decreasing constantly at least up to 2018.
-Unfortunately, for this study aerosol data for the entire period were
-not available in order to quantify their effect on SDR. The attenuation
-of SDR by aerosols over Europe have been proposed as major factor by
-\citet{Wild2021}. The brightening effect on SDR under cloudy-sky
-conditions (\(0.41\,\%/y\)), suggests that cloud optical thickness is
-also decreasing during this period. As long term data of cloud optical
+Thessaloniki is decreasing constantly at least up to 2018. The
+attenuation of SDR by aerosols over Europe have been proposed as major
+factor by \citet{Wild2021}. Unfortunately, for this study aerosol data
+for the entire period were not available in order to quantify their
+effect on SDR. The brightening effect on SDR under cloudy-sky conditions
+(\(0.41\,\%/y\)), suggests that cloud optical thickness is also
+decreasing during this period. As long term data of cloud optical
 thickness are also not available for the region, we cannot draw
-quantitative conclusions. \citet{Lozano2023} reported a Cloud Radiative
-Forcing (CRF) SDR trend of \(1.22\,W/m^2/y\) for the city of Granada
-Spain (\(37^\circ\,10'\,\)N, \(3^\circ\,37'\,\)E, \(680\,\)m~a.s.l.). We
-found a comparable result, for the same parameter, using the long term
-trend of all- and clear-skies, as follows
-\(\text{CRF} = 1.46 - 0.501 = 0.959\,W/m^2/y\).
+quantitative conclusions.
 
@@
 respectively, in contrast to spring and summer. The trends for autumn
-are also significant but smaller ( \(0.42\) and \(0.19\,\%/y\) for all-
-and cloudy-skies, respectively). Our findings are in agreement with
-other studies for the region. The trend for clear skies is largest in
+are also significant but smaller (\(0.42\) and \(0.19\,\%/y\) for all-
+and cloudy-skies, respectively). The trend for clear skies is largest in
 winter (\(0.4\,\%/y\)) and negligible in spring, summer and autumn.

### 2023-12-20 DHI_GHI_1_longterm_trends.R

@@ -272,8 +272,7 @@ vars <- c("DIR_att_des", "GLB_att_des", "tsi1au_att", "near_tcc_att", "near_tcc_
 # test
 vars <- c( "GLB_att_des")
 
-dbs         <- c("CLEAR_1_daily_DESEAS",
-                 "ALL_1_daily_DESEAS",
+dbs         <- c("ALL_1_daily_DESEAS",                        "CLEAR_1_daily_DESEAS",
                  "CLOUD_1_daily_DESEAS")
 ## gather trends
 gather <- data.frame()

### 2023-12-20 MDPI_submission.Rmd
Commit 2023-12-20 16:09

@@
   Under all three sky categories the SDR trend is stronger in winter with $0.7$, $0.4$ and $0.76\,\%/\text{year}$, respectively for all-, clear- and cloudy-skies. 
-In automn, there are the next larger trends of $0.42$ and $0.19\,\%/\text{year}$, for all- and cloudy-skies respectively.
-  The rest seasonal trends are significant smaller, close to zero, with clear- and cloudy-skies having a negative trend for summer.
+  The next larger seasonal trends are in automn $0.42$ and $0.19\,\%/\text{year}$, for all- and cloudy-skies respectively. 
+  The rest seasonal trends are significant smaller, close to zero, with a negative values in summer, for clear- and cloudy-skies.
   The SDR trend is increasing with increasing solar zenith angle, except under cloudy skies where the trend is highly variable and close to zero.
@@
     long: Direct beam/normal irradiance
+  - short: ERA5
+    long: ECMWF Reanalysis v5
   - short: CSid
@@
 \begin{equation}
-\text{SDR}_\text{Haurwitz} = 1098 \times \cos(\theta) \times \exp \left( \frac{ - 0.057}{\cos(\theta)} \right) \label{eq:hau}
+\text{SDR}_\text{Haurwitz} = 1098 \times \cos(\theta) \times \exp \left( \frac{ - 0.059}{\cos(\theta)} \right) \label{eq:hau}
 \end{equation}
@@
 <!-- $`r SDR_N_All`$ -->
-$`r signif(100 * SDR_N_Clear / SDR_N_All, 3)`\%$
-of the 1-minute data were identified as under clear-sky conditions and
-$`r signif(100 * SDR_N_Cloud / SDR_N_All, 3)`\%$
-as under cloud-sky conditions.
+<!-- $`r signif(100 * SDR_N_Clear / SDR_N_All, 3)`\%$ -->
+$26\,\%$ of the days were identified as under clear-sky conditions and
+<!-- $`r signif(100 * SDR_N_Cloud / SDR_N_All, 3)`\%$ -->
+$48\,\%$ as under cloud-sky conditions.
+The remaining $26\,\%$ of the data correspond to mixed cases and were not analyzed as a separate group.
 <!-- The TSI data we use, is a combination of data from NOAA [@Coddington2005] (for `r as.Date(TSIinfo[TSI_Source == "NOAA", Start])` - `r as.Date(TSIinfo[TSI_Source == "NOAA", End])`) and adjusted data from @LASP2023 (for `r as.Date(TSIinfo[TSI_Source != "NOAA", Start])` - `r as.Date(TSIinfo[TSI_Source != "NOAA", End])`).-->
@@
 The attenuation
-of SDR by aerosols over Europe have been proposed as major factor by @Wild2021.
+of SDR by aerosols over Europe has been proposed as major factor by @Wild2021.
 Unfortunately, for this study aerosol data for the entire period were

### 2023-12-20 MDPI_submission.tex
Commit 2023-12-20 16:09

@@
 and \(0.76\,\%/\text{year}\), respectively for all-, clear- and
-cloudy-skies. In automn, there are the next larger trends of \(0.42\)
-and \(0.19\,\%/\text{year}\), for all- and cloudy-skies respectively.
-The rest seasonal trends are significant smaller, close to zero, with
-clear- and cloudy-skies having a negative trend for summer. The SDR
-trend is increasing with increasing solar zenith angle, except under
-cloudy skies where the trend is highly variable and close to zero.
-Finally, we are discussing shorter-term variations in SDR anomalies by
-examining the patterns of the cumulative sums of monthly anomalies from
-the climatological mean, both before and after removing the long-term
+cloudy-skies. The next larger seasonal trends are in automn \(0.42\) and
+\(0.19\,\%/\text{year}\), for all- and cloudy-skies respectively. The
+rest seasonal trends are significant smaller, close to zero, with a
+negative values in summer, for clear- and cloudy-skies. The SDR trend is
+increasing with increasing solar zenith angle, except under cloudy skies
+where the trend is highly variable and close to zero. Finally, we are
+discussing shorter-term variations in SDR anomalies by examining the
+patterns of the cumulative sums of monthly anomalies from the
+climatological mean, both before and after removing the long-term
 trend.}
@@
 \begin{equation}
-\text{SDR}_\text{Haurwitz} = 1098 \times \cos(\theta) \times \exp \left( \frac{ - 0.057}{\cos(\theta)} \right) \label{eq:hau}
+\text{SDR}_\text{Haurwitz} = 1098 \times \cos(\theta) \times \exp \left( \frac{ - 0.059}{\cos(\theta)} \right) \label{eq:hau}
 \end{equation} where \(\theta\) is the SZA.
@@
 
-In the final dataset \(23.3\%\) of the 1-minute data were identified as
-under clear-sky conditions and \(43\%\) as under cloud-sky conditions.
+In the final dataset \(26\,\%\) of the days were identified as under
+clear-sky conditions and \(48\,\%\) as under cloud-sky conditions. The
+remaining \(26\,\%\) of the data correspond to mixed cases and were not
+analyzed as a separate group.
 
@@
 Thessaloniki is decreasing constantly at least up to 2018. The
-attenuation of SDR by aerosols over Europe have been proposed as major
+attenuation of SDR by aerosols over Europe has been proposed as major
 factor by \citet{Wild2021}. Unfortunately, for this study aerosol data
@@
 DNI & Direct beam/normal irradiance \\
+ERA5 & ECMWF Reanalysis v5 \\
 CSid & Clear sky identification algorithm \\

### 2023-12-24 MDPI_submission.tex
Commit 2023-12-24 18:14

@@
 respective datasets, the radiative effect of clouds is estimated to
-\(0.96\,W/m^2/y\). This estimate is similar to the cloud radiative
+\(0.96\,W/m^3/y\). This estimate is similar to the cloud radiative
 forcing of \(1.22\,W/m^2/y\) reported for Granada, Spain
@@
         }
-        \caption{Long term trends of SDR as a function of SZA separately from morning and afternoon periods.
+        \caption{Long term trends of daily SDR as a function of SZA for (a) all-sky, (b) clear-sky and (c) cloudy-sky conditions, separately for morning and afternoon periods.
 Solid shapes represent statistically significant trends ($p < 0.005$).}\label{fig:szatrends}
@@
         }
-        \caption{Linear trends (black lines) of monthly mean anomalies of SDR by season (rows of plots) for the three sky conditions (columns of plots).}\label{fig:seasonalALL}
+        \caption{Monthly mean anomalies of SDR by season (rows of plots) for the three sky conditions (columns of plots). The black lines represent the linear trends.}\label{fig:seasonalALL}
     \end{adjustwidth}
@@
 
-\caption{\label{tab:trendseasontable}SDR linear trends of monthly anomalies for each season of the year.}
+\caption{\label{tab:trendseasontable}SDR linear trends of monthly anomalies for each season of the year and related statistical parameters.}
 \begin{tabu} to \linewidth {>{\centering\arraybackslash}p{8em}>{\centering}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X}

### 2023-12-26 manuscript.v5/applsci-2743235.tex
Commit 2023-12-26 20:14

%  LaTeX support: latex@mdpi.com
%  For support, please attach all files needed for compiling as well as the log file, and specify your operating system, LaTeX version, and LaTeX editor.

%=================================================================
% pandoc conditionals added to preserve backwards compatibility with previous versions of rticles

\documentclass[applsci,article,accept,moreauthors,pdftex]{Definitions/mdpi}

%% Some pieces required from the pandoc template
\setlist[itemize]{leftmargin=*,labelsep=5.8mm}
\setlist[enumerate]{leftmargin=*,labelsep=4.9mm}

%--------------------
% Class Options:
%--------------------

%---------
% article
%---------
% The default type of manuscript is "article", but can be replaced by:
% abstract, addendum, article, book, bookreview, briefreport, casereport, comment, commentary, communication, conferenceproceedings, correction, conferencereport, entry, expressionofconcern, extendedabstract, datadescriptor, editorial, essay, erratum, hypothesis, interestingimage, obituary, opinion, projectreport, reply, retraction, review, perspective, protocol, shortnote, studyprotocol, systematicreview, supfile, technicalnote, viewpoint, guidelines, registeredreport, tutorial
% supfile = supplementary materials

%----------
% submit
%----------
% The class option "submit" will be changed to "accept" by the Editorial Office when the paper is accepted. This will only make changes to the frontpage (e.g., the logo of the journal will get visible), the headings, and the copyright information. Also, line numbering will be removed. Journal info and pagination for accepted papers will also be assigned by the Editorial Office.

%------------------
% moreauthors
%------------------
% If there is only one author the class option oneauthor should be used. Otherwise use the class option moreauthors.

%---------
% pdftex
%---------
% The option pdftex is for use with pdfLaTeX. Remove "pdftex" for (1) compiling with LaTeX & dvi2pdf (if eps figures are used) or for (2) compiling with XeLaTeX.

%=================================================================
% MDPI internal commands - do not modify
\firstpage{1}
\makeatletter
\setcounter{page}{\@firstpage}
\makeatother
\pubvolume{1}
\issuenum{1}
\articlenumber{0}
\pubyear{2024}
\copyrightyear{2023}
\externaleditor{Academic Editor: }
\datereceived{13 November 2023 }
\daterevised{20 December 2023 } % Comment out if no revised date
\dateaccepted{ }
\datepublished{ }
%\datecorrected{} % For corrected papers: "Corrected: XXX" date in the original paper.
%\dateretracted{} % For corrected papers: "Retracted: XXX" date in the original paper.
\hreflink{https://doi.org/} % If needed use \linebreak
%\doinum{}
%\pdfoutput=1 % Uncommented for upload to arXiv.org

%=================================================================
% Add packages and commands here. The following packages are loaded in our class file: fontenc, inputenc, calc, indentfirst, fancyhdr, graphicx, epstopdf, lastpage, ifthen, float, amsmath, amssymb, lineno, setspace, enumitem, mathpazo, booktabs, titlesec, etoolbox, tabto, xcolor, colortbl, soul, multirow, microtype, tikz, totcount, changepage, attrib, upgreek, array, tabularx, pbox, ragged2e, tocloft, marginnote, marginfix, enotez, amsthm, natbib, hyperref, cleveref, scrextend, url, geometry, newfloat, caption, draftwatermark, seqsplit
% cleveref: load \crefname definitions after \begin{document}

%=================================================================
% Please use the following mathematics environments: Theorem, Lemma, Corollary, Proposition, Characterization, Property, Problem, Example, ExamplesandDefinitions, Hypothesis, Remark, Definition, Notation, Assumption
%% For proofs, please use the proof environment (the amsthm package is loaded by the MDPI class).

%=================================================================
% Full title of the paper (Capitalized)
\Title{Trends from 30-Year Observations of Downward Solar Irradiance in
Thessaloniki, \colorbox{green}{Greece}}%MDPI: Important note:
%1. The paper was edited by our English editor, please check the whole text and confirm if your meaning is retained.
%2. Do not delete any comment we left for you and reply to each comment so that we can understand your meaning clearly.
%3. Please directly correct on this version. If you need to revise somewhere in your paper, please highlight the revisions and track changes to make us known.
%4. Please finish the proofreading based on this version.
%5.Please confirm and revise all the comments with “Confirmed”, “OK”, “Revised”, “It should be italic”; “I confirm”; “I confirm xx is correct”; “I have checked and revised all.” , etc.“
%6. Please note that at this stage (the manuscript has been accepted in the current form), we will not accept authorship or content changes to the main text.
%(Thank you for your cooperation in advance.)

% MDPI internal command: Title for citation in the left column
\TitleCitation{Trends from 30-Year Observations of Downward Solar Irradiance in
Thessaloniki, Greece}

% Author Orchid ID: enter ID or remove command
%\newcommand{\orcidauthorA}{0000-0000-0000-000X} % Add \orcidA{} behind the author's name
%\newcommand{\orcidauthorB}{0000-0000-0000-000X} % Add \orcidB{} behind the author's name

% Authors, for the paper (add full first names)
\Author{Athanasios
Natsis \href{https://orcid.org/0000-0002-5199-4119}
{\orcidicon}, Alkiviadis Bais * and Charikleia Meleti}

%\longauthorlist{yes}

% MDPI internal command: Authors, for metadata in PDF
\AuthorNames{Athanasios Natsis, Alkiviadis Bais, Charikleia Meleti}

% MDPI internal command: Authors, for citation in the left column
%\AuthorCitation{Lastname, F.; Lastname, F.; Lastname, F.}
% If this is a Chicago style journal: Lastname, Firstname, Firstname Lastname, and Firstname Lastname.
\AuthorCitation{Natsis, A.; Bais, A.; Meleti, C.}

% Affiliations / Addresses (Add [1] after \address if there is only one affiliation.)
\address[1]{Laboratory of
Atmospheric Physics, Aristotle University of Thessaloniki, Campus Box 149, \mbox{54124 Thessaloniki,
Greece;} {{natsisphysicist@gmail.com}}
(A.N.); 
{{meleti@auth.gr}} (C.M.)\\
}

% Contact information of the corresponding author
\corres{\hangafter=1 \hangindent=1.0em \hspace{-1em}Correspondence: {{abais@auth.gr}}}

% Current address and/or shared authorship

% The commands \thirdnote{} till \eighthnote{} are available for further notes

% Simple summary

%\conference{} % An extended version of a conference paper

% Abstract (Do not insert blank lines, i.e. \\)
\abstract{The shortwave downward solar irradiance (SDR) is an important
factor that drives climate processes and energy production and can affect
all living organisms. Observations of SDR at different locations around
the world with different environmental characteristics have been used to
investigate its long-term variability and trends at different time
scales. Periods of positive trends are referred to as brightening periods
and of negative trends as dimming periods. In this study we have used 30
years of pyranometer data in Thessaloniki, Greece, to investigate the
variability of SDR under three types of sky conditions (clear-, cloudy-
and all-sky). The clear-sky data were identified by applying a cloud
screening algorithm. We have found a positive trend of
\(0.38\%/\text{year}\) for all-sky, $\sim$0.1\%/{year} for
clear-sky, and \(0.41\%/\text{year}\) for cloudy conditions. The
consistency of these trends, their seasonal variability, and the effect
of the solar zenith angle have also been investigated. Under all three
sky categories, the SDR trend is stronger in winter, with \(0.7\), \(0.4\),
and \(0.76\%/\text{year}\), respectively, for all-, clear-, and
cloudy-sky conditions. The next larger seasonal trends are in autumn---\(0.42\) and
\(0.19\%/\text{year}\), for all and cloudy skies, respectively. The
rest of the seasonal trends are significant smaller, close to zero, with a
negative values in summer, for clear and cloudy skies. The SDR trend is
increasing with increasing solar zenith angle, except under cloudy skies,
where the trend is highly variable and close to zero. Finally, we
discuss shorter-term variations in SDR anomalies by examining the
patterns of the cumulative sums of monthly anomalies from the
climatological mean, both before and after removing the long-term
trend.}

% Keywords
\keyword{GHI; SDR; solar radiation; solar brigthening/dimming; aerosols;
clouds}

% The fields PACS, MSC, and JEL may be left empty or commented out if not applicable
%\PACS{J0101}
%\MSC{}
%\JEL{}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Diversity
%\LSID{\url{http://}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Applied Sciences

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Data

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Toxins

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Encyclopedia

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Only for the journal Advances in Respiratory Medicine
%\addhighlights{yes}
%\renewcommand{\addhighlights}{%

%\noindent This is an obligatory section in “Advances in Respiratory Medicine”, whose goal is to increase the discoverability and readability of the article via search engines and other scholars. Highlights should not be a copy of the abstract, but a simple text allowing the reader to quickly and simplified find out what the article is about and what can be cited from it. Each of these parts should be devoted up to 2~bullet points.\vspace{3pt}\\
%\textbf{What are the main findings?}
% \begin{itemize}[labelsep=2.5mm,topsep=-3pt]
% \item First bullet.
% \item Second bullet.
% \end{itemize}\vspace{3pt}
%\textbf{What is the implication of the main finding?}
% \begin{itemize}[labelsep=2.5mm,topsep=-3pt]
% \item First bullet.
% \item Second bullet.
% \end{itemize}
%}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% tightlist command for lists without linebreak
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\usepackage{subcaption}
\captionsetup[sub]{position=bottom, labelfont={bf, small, stretch=1.17}, labelsep=space, textfont={small, stretch=1.17}, aboveskip=6pt,  belowskip=-6pt, singlelinecheck=off, justification=justified}
\usepackage{placeins}
\usepackage{longtable}
\usepackage{booktabs}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

The shortwave downward solar irradiance (SDR) at Earth's surface plays a
significant role on its climate. Changes in the SDR can be related to
changes in Earth's energy budget, the mechanisms of climate change, and
water and carbon cycles \citep{Wild2009}. It can also affect solar and
agricultural production and all living organisms. Studies of SDR
variability have identified some distinct SDR trends in different
regions of the world in different time periods. The term `brightening'
is generally used to describe periods of positive SDR trend, and
`dimming' for periods of negative trend \citep{Wild2009}. There are many cases in
the long-term records of irradiance showing a systematic change in the
magnitude of the trend, occurring roughly in the last decades of the
20th century. At multiple stations in China, a dimming period was
reported until about 2000, followed by a brightening period
\citep{Yang2021}. A similar pattern was identified, with a breaking
point around 1980, for stations in Central Europe \citep{Wild2021} and
Brazil~\citep{Yamasoe2021}. On global scale, an artificial intelligence
aided spatial analysis on the continental level with data from multiple
stations reached similar conclusions for these regions and for the
global trend \citep{Yuan2021}.

There is a consensus among researchers that the major factor affecting
the variability of SDR attenuation is the interactions of solar
radiation with atmospheric aerosols and clouds. Those interactions,
among other factors, have been analyzed with models
\citep{Li2016, Samset2018}, showing the existence of feedback mechanisms
between the two. Similar findings have been shown from the analysis of
observations at other \hl{locations} %MDPI: We revised citation. Please confirm.
 \citep{Schwarz2020, Ohvril2009, Zerefos2009, Xia2007} [and references
therein]. In the
Mediterranean region, aerosols have been recognized as an important
factor affecting the penetration of solar radiation at the surface
\citep{Fountoulakis2016, Siomos2018, Gkikas2013, Lozano2021}. These
studies investigated the long-term trend in aerosol optical depth, which
has been found to decrease in the last three decades, the transport and
composition of aerosols, and their radiative effects.

Due to the significant spatial and temporal variability of the trends
and the contributing factors, there is a constant need to monitor and
investigate SDR at different sites in order to estimate the degree of
variability and its relation to the local conditions. In this study, we
examine the trends of SDR using ground-based measurements at
Thessaloniki, Greece, for the period from 1993 to 2023. We re-evaluated and
extended the dataset used by \citet{Bais2013}, we applied a different
algorithm for the identification of clear-/cloud-sky instances
\citep{Reno2016, Reno2012a}, and we derived the SDR trends for the
period of study under different sky conditions (all-sky, clear-sky, and
cloudy-sky). Finally, we investigated the dependence of the trends on
solar zenith angle and season.

\hypertarget{data-and-methodology}{%
\section{Data and Methodology}\label{data-and-methodology}}

The SDR data were measured with a Kipp \& Zonen CM-21 pyranometer
operating continuously at the Laboratory of Atmospheric Physics of the
Aristotle University of Thessaloniki (\(40^\circ\,38'\,\)N,
\(22^\circ\,57'\,\)E, \(80\,\)m~a.s.l.). Here, we used data for the period
from 13~April 1993 to 13 April 2023. The monitoring site was located near the
city center, thus we expect that measurements were affected by the urban
environment, mainly by aerosols. During the study period, the
pyranometer was independently calibrated three times at the
Meteorologisches Observatorium Lindenberg, DWD, verifying that the
stability of the instrument's sensitivity was better than \(0.7\%\)
relative to the initial calibration by the manufacturer. Along with SDR,
the direct beam radiation (DNI) was also measured with a collocated Kipp
\& Zonen CHP-1 pyrheliometer since 1 April 2016. The DNI data were used as
auxiliary data to support the selection of appropriate thresholds in the
clear-sky identification algorithm (CSid), which is discussed in Section
\ref{CDIDalgorithm}. It is noted that the limited dataset of DNI was not
used for the identification of clear-sky cases in the entire SDR series
to avoid any selection bias due to the unequal length of the two datasets.
There are four distinct steps in the creation of the dataset analyzed
here: (a)~the acquisition of radiation measurements from the sensors,
(b)~the data quality check, (c)~the identification of ``clear sky''
conditions from the SDR data, and (d)~the aggregation of data and trend
analysis.

For the acquisition of radiometric data, the signal of the pyranometer
was sampled at a rate of \(1\,\text{Hz}\). The mean and the standard
deviation of these samples were calculated and recorded every minute. The
measurements were corrected for the zero offset (``dark signal'' in
volts), which was calculated by averaging all measurements recorded for a
period of \(3\,\text{h}\), before (morning) or after (evening) the Sun
reaches an elevation angle of \(-10^\circ\). The signal was converted to
irradiance using a ramped value of the instrument's sensitivity between
subsequent calibrations.

A manual screening was performed to remove inconsistent and erroneous
recordings that can occur stochastically or systematically during the
continuous operation of the instruments. The manual screening was aided
by a radiation data quality assurance procedure, adjusted for the site,
which was based on the methods of Long and
Shi~\citep{Long2006, Long2008a}. Thus, problematic recordings have been
excluded from further processing. Although it is impossible to detect
all false data, the large number of available data, and the aggregation
scheme we used, ensures the quality of the radiation measurements used
in this study.

Due to the significant measurement uncertainty when the Sun is near the
horizon, we have excluded all measurements with solar zenith angle (SZA)
greater than \(85^\circ\). Moreover, due to obstructions around the site
(hills and buildings) that block the direct irradiance, we excluded
data with azimuth angle in the range \(58^{\circ}\)--\(120^{\circ}\)
and with SZA greater than \(78^{\circ}\). To make the measurements
comparable throughout the dataset, we adjusted all one-minute data to
the mean Sun--Earth distance. Subsequently, we adjusted all
measurements to the total solar irradiance (TSI) at \(1\,\text{au}\) in
order to compensate for the Sun's intensity variability using a time
series of satellite TSI observations. The TSI data we used are part of
the `NOAA Climate Data Record of Total Solar Irradiance' dataset
\citep{Coddington2005}. The initial daily values of this dataset were
interpolated to match the time step of our measurements.

In order to estimate the effect of the sky conditions on the long-term
variability of SDR, we created three datasets by characterizing each
one-minute measurement with a corresponding sky-condition flag (i.e.,
all-sky, clear-sky, and cloudy-sky). To identify the clear cases, we used
the method proposed by \citet{Reno2016}, which requires the definition
of some site specific parameters. These parameters were determined by an
iterative process, as the original authors proposed, and are discussed in
the next section.

We note that all methods have some subjectivity in the definition of
clear or cloudy sky cases. As a result, the details of the definition
are site specific, and they rely on a combination of thresholds and
comparisons with ideal radiation models and statistical analysis of
different signal metrics. The CSid algorithm was calibrated with the
main focus to identify the presence of clouds. Despite the fine-tuning
of the procedure, in a few marginal cases, false positive or false
negative results were identified by manual inspection. However, due to
their small number, they did not affect the final results of the study.
For completeness, we provide below a brief overview of the CSid
algorithm, along with the site-specific thresholds.

\hypertarget{CDIDalgorithm}{%
\subsection{The Clear Sky Identification
Algorithm}\label{CDIDalgorithm}}

To calculate the reference clear-sky \(\text{SDR}_\text{CSref}\), we used
the \(\text{SDR}_\text{Haurwitz}\) derived by the radiation model of
\citet{Haurwitz1945} (Equation~\eqref{eq:hau}), adjusted for our site:
\begin{equation}
\text{SDR}_\text{Haurwitz} = 1098 \times \cos(\theta) \times \exp \left( \frac{ - 0.059}{\cos(\theta)} \right) \label{eq:hau}
\end{equation} where \(\theta\) is the SZA.

The adjustment was made with a factor \(a\) (Equation~\eqref{eq:ahau}), which
was estimated through an iterative optimization process, as described by
\citet{Long2000} and \citet{Reno2016}. The target of the optimization
was the minimization of a function \(f(a)\) (Equation~\eqref{eq:minf}) and was
accomplished with the algorithmic function `optimise', which is an
implementation based on the work of \citet{Brent1973}, from the library
`stats' of the R programming language \citep{RCT2023}.
\begin{equation}
f(a) = \frac{1}{n}\sum_{i=1}^{n} ( \text{SDR}_{\text{CSid},i} - a \times \text{SDR}_{\text{testCSref},i} )^2 \label{eq:minf}
\end{equation} where \(n\) is the total number of daylight data,
\(\text{SDR}_{\text{CSid},i}\) are the data identified as clear-sky by
CSid, \(a\) is a site-specific adjustment factor, and
\(\text{SDR}_{\text{testCSref},i}\) is the SDR derived by any of the
tested clear-sky radiation models.

The optimization and the selection of the clear-sky reference model was
performed on SDR observations for the period 2016--2021. During the
optimization, eight simple clear-sky radiation models were tested
(namely, Daneshyar--Paltridge--Proctor, Kasten--Czeplak, Haurwitz,
Berger--Duffie, Adnot--Bourges--Campana--Gicquel, Robledo--Soler, Kasten, and
Ineichen--Perez) with a wide range of factors. These models are
described in more detail by \citet{Reno2012} and are evaluated by
\citet{Reno2016}. We found that Haurwitz's model, adjusted with the
factor \(a = 0.965\), yields one of the lowest root mean squared errors
(RMSE), and the procedure manages to successfully characterize the
majority of the data. Thus, our clear sky reference is derived by
Equation~\eqref{eq:ahau}: \begin{equation}
\text{SDR}_\text{CSref} = a \times \text{SDR}_\text{Haurwitz} = 0.965 \times 1098 \times \cos(\theta) \times \exp \left( \frac{ - 0.057}{\cos(\theta)} \right) \label{eq:ahau}
\end{equation}

The criteria that were used to identify whether a measurement was taken
under clear-sky conditions are presented below. A data point is flagged
as `clear-sky' if all criteria are satisfied; otherwise, it is
considered as `cloud-sky'. Each criterion was applied for a running
window of \(11\) consecutive one-minute measurements, and the
characterization was assigned to the central datum of the window. Each
parameter was calculated from the observations in comparison to the
reference clear-sky model. The allowable range of variation is defined
by the model-derived value of the parameter multiplied by a factor plus
an offset. The factors and the offsets were determined empirically by
manually inspecting each filter's performance on selected days and
adjusting them accordingly during an iterative process. The criteria are
listed below, together with the range of values within which the
respective parameter should fall in order to raise the clear-sky flag:

\begin{enumerate}
\def\labelenumi{(\alph{enumi})}
\tightlist
\item
  \hl{Mean} %MDPI: We removed italics of units. Please confirm.
 of the measured \(\overline{\text{SDR}}_i\) (Equation~\eqref{eq:MeanVIP}): \begin{equation}
  0.91 \times \overline{\text{SDR}}_{\text{CSref},i} - 20\,\text{Wm}^{-2}
  < \overline{\text{SDR}}_i <
  1.095 \times \overline{\text{SDR}}_{\text{CSref},i} + 30\,\text{Wm}^{-2}
  \label{eq:MeanVIP}
  \end{equation}
\item
  Maximum measured value \(M_{i}\) (Equation~\eqref{eq:MaxVIP}):
  \begin{equation}
  1 \times M_{\text{CSref},i} - 75\,\text{Wm}^{-2}
  < M_{i} <
  1 \times M_{\text{CSref},i} + 75\,\text{Wm}^{-2}
  \label{eq:MaxVIP}
  \end{equation}
\item
  Length \(L_i\) of the sequential line segments, connecting the points
  of the \(11\) SDR values (Equation~\eqref{eq:VILeq}): \begin{equation}
  L_i = \sum_{i=1}^{n-1}\sqrt{\left ( \text{SDR}_{i+1} - \text{SDR}_{i}\right )^2 + \left ( t_{i+1} - t_i \right )^2}
  \label{eq:VILeq}
  \end{equation} \begin{equation}
  1 \times L_{\text{CSref},i} - 5 < L_i < 1.3 \times L_{\text{CSref},i} + 13
  \label{eq:VILcr}
  \end{equation} where \(t_i\) is the time stamp of each SDR
  measurement.
\item
  Standard deviation \(\sigma_i\) of the slope (\(s_i\)) between the
  \(11\) sequential points, normalized by the mean
  \(\overline{\text{SDR}}_i\) (Equation~\eqref{eq:VCT1}): \begin{gather}
    \sigma_i = \frac {\sqrt{\frac{1}{n-1} \sum_{i=1}^{n-1} \left( s_i - \bar{s} \right)^2}} {\overline{\text{SDR}}_i} \label{eq:VCT1} \\
    s_i = \frac{\text{SDR}_{i+1} - \text{SDR}_{i}}{t_{i+1} - t_i},\;\;   \bar{s} = \frac{1}{n-1} \sum_{i=1}^{n-1} s_i,\;\;\forall i \in \left \{ 1, 2, \ldots, n-1 \right \}\;\;
  \end{gather} For this criterion, \(\sigma_i\) should be below a
  certain threshold (Equation~\eqref{eq:VCTcr}): \begin{equation}
    \sigma_i < \ensuremath{1.1\times 10^{-4}} \label{eq:VCTcr}
  \end{equation}
\item
  Maximum difference \(X_i\) between the change in measured irradiance
  and the change in clear sky irradiance over each measurement interval:
  \begin{gather}
    X_i = \max{\left \{ \left | x_i - x_{\text{CSref},i} \right | \right \}} \label{eq:VSM3} \\
    x_i = \text{SDR}_{i+1} - \text{SDR}_{i} \forall i \in \left \{ 1, 2, \ldots, n-1 \right \} \label{eq:VSM1} \\
    x_{\text{CSref},i} = \text{SDR}_{\text{CSref},i+1} - \text{SDR}_{\text{CSref},i} \forall i \in \left \{ 1, 2, \ldots, n-1 \right \} \label{eq:VSM2}
  \end{gather} For this criterion, \(X_i\) should be below a certain
  threshold (Equation~\eqref{eq:VSMcr}): \begin{equation}
    X_i < 7.5\,\text{Wm}^{-2} \label{eq:VSMcr}
  \end{equation}
\end{enumerate}

In the final dataset, \(26\%\) of the days were identified as under
clear-sky conditions and \(48\%\) as under cloud-sky conditions. The
remaining \(26\%\) of the data correspond to mixed cases and were not
analyzed as a separate group.

\hypertarget{aggregationstatistical}{%
\subsection{Aggregation of Data and Statistical
Approach}\label{aggregationstatistical}}

In order to investigate the SDR trends that are the main focus of the
study, we implemented an aggregation scheme to the one-minute data to
derive series in coarser time scales. To preserve the representativeness
of the data, we used the following criteria: (a)~we excluded all days with
less than 50\% of the expected daytime measurements, (b)~daily means for
the clear-sky and cloudy-sky datasets were calculated only for days with
more than 60\% of the expected daytime measurements identified as clear
or cloudy, respectively, and (c) monthly means were computed from daily means.
For the all-skies dataset, monthly means were computed only when at least
20 days were available. Seasonal means were derived by averaging the
monthly mean values in each season (winter: December--February, spring:
March--May, etc.). The daily and monthly climatological means were
derived by averaging the data for each day of the year and calendar month,
respectively. The daily and monthly datasets were deseasonalized by
subtracting the corresponding climatological annual cycle (daily or
monthly) from the actual data. Finally, to estimate the SZA effect on
the SDR trends, the one-minute data were aggregated in \(1^{\circ}\) SZA
bins, separately for the morning and afternoon hours.

The linear trends were calculated using a first-order autoregressive
model with lag of 1~day using the `maximum likelihood' fitting method
\citep{Gardner1980, Jones1980} by implementing the function `arima'
from the library `stats' of the R programming language
\citep{RCT2023}. The trends were reported together with the \(2\sigma\)
errors.

\hypertarget{results}{%
\section{Results}\label{results}}

\hypertarget{long-term-sdr-trends}{%
\subsection{Long-Term SDR Trends}\label{long-term-sdr-trends}}

We calculated the linear trends of SDR from the departures of the mean
daily values from the daily climatology and for the three sky
conditions. These are presented in Table~\ref{tab:trendtable}, which also contains the \(2\sigma\) standard error, the Pearson's correlation
coefficient R, and the trend in absolute units. In
Figure~\ref{fig:trendALL}, we present only the time series under all-sky
conditions; the plots for clear-sky and cloud-sky conditions, are shown
in \mbox{\hl{Appendix} %MDPI: We revised citation. Please confirm.
 \ref{app1}} \mbox{(Figures~\ref{fig:trendCLEAR} and~
\ref{fig:trendCLOUD}).} In the studied period, there is no significant
break or change in the variability pattern of the time series. The
linear trends in all three datasets are positive and \hl{around} %MDPI: We removed italics of units. Please confirm. The same below.
\(0.4\%/\)y for all-sky and cloudy-sky conditions, whereas for
clear-skies the trend is much smaller (\textasciitilde{}\(0.1\%/\)y).
The linear trends were calculated taking into account the
autocorrelation of the time series, and all three are statistically
significant at least at the \(95\%\) confidence level, as they are
larger than the corresponding \(2\sigma\) errors, despite the small
values of R, which is due to the large variability of the daily values.
The clear-sky trend is very small, suggesting a small effect from
aerosols and water vapor, which are the dominant factors of the SDR
variability \citep{Fountoulakis2016, Siomos2018, Yu2022}. In contrast,
the large positive trend of SDR under cloudy skies can be attributed to
reduction in cloud cover and/or cloud optical depth. Lack of continuous
observations of cloud optical thickness that could support these
findings does not allow drawing firm conclusions. However, there are
indications that the total cloud cover as inferred from the ERA5
analysis for the grid point of Thessaloniki is decreasing over the
period of study. From the difference between all-sky and clear-sky SDR
trends, expressed in W/m$^2$/y using the long-term mean of the
respective datasets, the radiative effect of clouds is estimated to
0.96 W/m$^2$/y. This estimate is similar to the cloud radiative
forcing of 1.22 W/m$^2$/y  reported for Granada, Spain
\citep{Lozano2023}.

The all-sky trend is similar to the one reported in \citet{Bais2013}
from a ten-year shorter dataset, suggesting that the tendency of SDR in
Thessaloniki is systematic. Other studies for the European region
reported a change in the SDR trend around 1980 from negative to positive
with comparable magnitude \citep{Wild2021, Yuan2021, Ohmura2009}, well
before the start of our records. However, the trends reported here for
the three datasets are in accordance with the widely accepted solar
radiation brightening over Europe. For the period of our observations,
the trend in the TSI is negligible (\(-0.00022\%/\)y), and thus we
cannot attribute any significant effect on the SDR trend to solar
variability.

\begin{table}[H]

\caption{\label{tab:trendtable}Trends in SDR daily means for different sky conditions for the period 1993--2023.}
\begin{tabu} to \linewidth {>{\centering\arraybackslash}p{8em}>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X}
\toprule
\multirow{-1}{*}{\textbf{Sky Conditions}} & \textbf{Trend [\%/year]} & \textbf{Trend S.E. (\boldmath$2\sigma$)} & \textbf{Pearson Correl.} & \textbf{Trend [W/m\textsuperscript{2}/year]} & \multirow{-1}{*}{\textbf{Days}}\\
\midrule
All skies & 0.380 & 0.120 & 0.091 & 1.460 & 10251\\
Clear skies & 0.097 & 0.033 & 0.140 & 0.501 & 2684\\
Cloudy skies & 0.410 & 0.180 & 0.081 & 1.180 & 4937\\
\bottomrule
\end{tabu}
\end{table}

\vspace{-6pt}

\begin{figure}[H]

{ \includegraphics[width=.75\linewidth]{./images/LongtermTrends-1}

}

\caption{Anomalies (\%) of the daily all-sky SDR from the climatological mean for the period 1993--2023. The black line is the long-term linear trend.}\label{fig:trendALL}
\end{figure}

Although the year-to-year variability of the anomalies (Figure
\ref{fig:trendALL} and \mbox{Figures~\ref{fig:trendCLEAR} and
\ref{fig:trendCLOUD}} in \mbox{Appendix \ref{app1})} shows a rather homogeneous behavior,
plots of the cumulative sums (CUSUM) \citep{Regier2019} of the anomalies
can reveal different structures in the records of all three sky
conditions. For time series with a uniform trend, we would expect the
CUSUMs of the anomalies to have a symmetric `V' shape centered around
the middle of the series. This would indicate that the anomalies are
evenly distributed around the climatological mean, and, for a positive
uniform trend, the first half is below and the second half above the
climatological mean. In our case, there is a more complex evolution of
the anomalies. For all skies (Figure~\ref{fig:cusummonth}a), we observe
three rather distinct periods: (a) a downward part between the start of
the datasets and about 2000, denoting that all anomalies are negative,
thus below the climatology; (b) a relatively steady part lasting for
almost 20 years, suggesting little variability in SDR anomalies; and (c) a
steep upward part to the present, indicating anomalies above the climatology.
The CUSUMs for cloudy skies (Figure~\ref{fig:cusummonth}c) show a
similar behavior with some short-term differences that do not change the
overall pattern. For clear skies (Figure~\ref{fig:cusummonth}b), a
monotonic downward tendency is evident until 2004, suggesting that the
anomalies are all negative. After 2004, the anomalies turn positive at
a fast rate for about five years and at a slower rate thereafter.

\begin{figure}[H]
    \begin{adjustwidth}{-\extralength}{0cm}
        {\centering 
       % \subfloat[\label{fig:cusummonth-1}]
            {\includegraphics[width=.32\linewidth]{./images/CumulativeMonthlyCuSum-1}}\hfill
       % \subfloat[\label{fig:cusummonth-2}]
            {\includegraphics[width=.32\linewidth]{./images/CumulativeMonthlyCuSum-5}}\hfill
        \subfloat[\label{fig:cusummonth-3}]
            {\includegraphics[width=.32\linewidth]{./images/CumulativeMonthlyCuSum-9}}\hfill
        }
\end{adjustwidth}
\caption{\hl{Cumulative} %MDPI: We moved the subfigure explanations into the figure caption. Please confirm.
 sum plots of the monthly SDR anomalies in (\%) for different sky conditions: (\textbf{a})~all skies; (\textbf{b}) clear skies; (\textbf{c}) cloudy skies.}\label{fig:cusummonth}

\end{figure}

In order to unveil further the features of the variability of the three
datasets, Figure~\ref{fig:cusumnotrendmonthly} presents another set of
CUSUM plots using anomalies after the long-term linear trend is removed.
With this approach, periods when the CUSUMs diverge from zero can be
interpreted as a systematic variation of SDR from the climatological
mean. When the CUSUM is increasing, the anomalies values are above the
climatology, and vice versa. Overall, for all- and cloudy-sky conditions
(Figure~\ref{fig:cusumnotrendmonthly}a,c), we observe periods with anomalies
diverging from the climatological values, each lasting for several
years. These fluctuations are probably within the natural variability,
and no distinct changes are identified. The pattern in both datasets is
similar, suggesting prevalence in cloudy skies over Thessaloniki. For
clear skies (Figure~\ref{fig:cusumnotrendmonthly}b), the distinct change
in 2004 is now clearer. The most likely reason for this change is the
monotonic reduction of aerosols in Thessaloniki. In that year, there was a
change in the rate of decrease in aerosol optical depth, as illustrated
in Figure 7 of \citet{Siomos2020}. This abrupt change in CUSUMs lasted
until about 2010, when the anomalies become again variable.

\begin{figure}[H]
    \begin{adjustwidth}{-\extralength}{0cm}
        {\centering 
          %  \subfloat[All skies.\label{fig:cusumnotrendmonthly-1}]
                {\includegraphics[width=.32\linewidth]{./images/CumulativeMonthlyCuSumNOtrend-1} }\hfill
           % \subfloat[Clear skies.\label{fig:cusumnotrendmonthly-2}]
                {\includegraphics[width=.32\linewidth]{./images/CumulativeMonthlyCuSumNOtrend-5} }\hfill
           % \subfloat[Cloudy skies.\label{fig:cusumnotrendmonthly-3}]
                {\includegraphics[width=.32\linewidth]{./images/CumulativeMonthlyCuSumNOtrend-9} }
        }

\end{adjustwidth}
        \caption{\hl{Cumulative} %MDPI: We moved the subfigure explanations into the figure caption. Please confirm.
 sum plots of monthly SDR anomalies in (\%) for different sky conditions after removing the long-term linear trend: (\textbf{a})~all skies; (\textbf{b}) clear skies; (\textbf{c}) cloudy skies.}\label{fig:cusumnotrendmonthly}
\end{figure}

\hypertarget{effects-of-the-solar-zenith-angle-on-sdr}{%
\subsection{Effects of the Solar Zenith Angle on
SDR}\label{effects-of-the-solar-zenith-angle-on-sdr}}

The solar zenith angle is a major factor affecting the SDR, as
increases in SZA leads to enhancement of the radiation path in the
atmosphere, especially in urban environments with human activities
emitting aerosols \citep{Wang2021}. In order to estimate the effect of
SZA on the SDR trends, we grouped the data in bins of \(1^\circ\) SZA
and calculated the overall trend for each bin separately for the daily
periods before noon and after noon (Figure~\ref{fig:szatrends}).
Although there are seasonal dependencies of the minimum SZA (see
Appendix \ref{app1}, Figure~\ref{fig:SZAtrendSeason}), these dependencies are not
discussed further.

For all-sky conditions, the brightening effect of SDR (positive trend)
increases with SZAs (Figure~\ref{fig:szatrends}a), ranging from about
\(0.1\%/\)y to about \(0.7\%/\)y for the statistically significant
trends. The trends in the morning and afternoon hours are more or less
consistent with small differences at small SZAs, which can be attributed
to effects on clear sky SDR from systematic diurnal patterns of aerosols
during the warm period of the year, consistent with the results
reported for China by \citet{Wang2021}. Note that SZAs less than
\(25^\circ\) can only occur during the warm period of the year around
noon when clear skies are more frequent. The increasing trend with SZA
is likely caused by the increased attenuation of SDR with SZA. The
effect is larger when aerosol and/or cloud layers are optically thicker;
therefore, decreases in aerosol and clouds through the study period will
result in larger positive trends of SDR at larger SZAs.

Under clear skies (Figure~\ref{fig:szatrends}b), the trends are
smaller and less variable, ranging between \(0.1\) and \(0.15\%/\)y up
to \(77^\circ\) SZA. At higher SZAs and in the afternoon hours, there is a
sharp increase in the trend up to \(0.3\%/\)y, which may have been
caused by the long path length of radiation through the atmosphere as
discussed above for the all-sky conditions. The small differences in the
trend between morning and afternoon between \(35^\circ\) and
\(60^\circ\) SZA is likely a result of less attenuation of SDR in the
morning hours due to lesser amounts of aerosols and a shallower boundary
layer.

For cloudy-sky conditions (Figure~\ref{fig:szatrends}c), we cannot
discern any significant dependence of the SDR trend with SZA, as the
variability of irradiance is dominated by the cloud effects leading to
insignificant trends. Statistically significant trends appear only in
the afternoon and for SZAs larger than \(60^\circ\). The sharp increase
of the trend at SZAs larger than $\sim${75}$^{\circ}$, observed also for
clear skies, is probably associated with stronger attenuation by clouds
under oblique incidence angles, which also result in smaller
variability.

\begin{figure}[H]
    \begin{adjustwidth}{-\extralength}{0cm}
        {\centering 
          %  \subfloat[All skies.\label{fig:szatrends-1}]
                {\includegraphics[width=.32\linewidth]{./images/SzaTrends-1}}\hfill
         %   \subfloat[Clear skies.\label{fig:szatrends-2}]
                {\includegraphics[width=.32\linewidth]{./images/SzaTrends-4}}\hfill
           % \subfloat[Cloudy skies.\label{fig:szatrends-3}]
                {\includegraphics[width=.32\linewidth]{./images/SzaTrends-7}}
        }

    \end{adjustwidth}
        \caption{\hl{Long} %MDPI: We removed duplicate explanation in figure.
 -term trends of daily SDR as a function of SZA for (\textbf{a}) all-sky, (\textbf{b}) clear-sky and (\textbf{c}) cloudy-sky conditions, separately for morning and afternoon periods. Solid shapes represent statistically significant trends ($p < 0.005$).}\label{fig:szatrends}
\end{figure}

\hypertarget{long-term-sdr-trends-by-season}{%
\subsection{Long-Term SDR Trends by
Season}\label{long-term-sdr-trends-by-season}}

Similarly to the long term trends from daily means of SDR discussed
above, we have calculated the trend for the three sky conditions and for
each season of the year using the corresponding mean monthly anomalies
(Figure~\ref{fig:seasonalALL} and Table~\ref{tab:trendseasontable}).
Table~\ref{tab:trendseasontable} also contains the \(2\sigma\) standard
error, the Pearson's correlation coefficient R, and the corresponding
p-value. The winter linear trends generally exhibit the largest R values,
ranging between \(0.54\) and \(0.60\%/\)y. For all-sky conditions, the
trend in SDR in winter is the largest (\(0.7\%/\)y), followed by the
trend in autumn (\(0.42\%/\)y, a value close to the long-term trend),
both statistically significant at the \(95\%\) confidence level. In
spring and summer, the trends are much smaller and of lesser statistical
significance. These seasonal differences indicate a possible relation of
the trends in SDR to trends in clouds during winter and autumn. For
clear skies, the trend in winter is \(0.4\%/\)y and is associated with
the decreasing trend in aerosol optical depth \citep{Siomos2020}.
Moreover, it is almost half of that for all skies, which is another
indication of a decreasing trend in cloud optical thickness. In other
seasons, the clear-sky trend is very small (below \(0.1\%/\)y).
Finally, for cloudy skies, the winter trend is the largest
(\(0.76\%/\)y) and greater than for all skies, followed by a much
smaller trend in autumn (\(0.19\%/\)y).

The trends under clear- and cloudy-sky conditions are in the same
direction, and it would be expected that their sum is similar to the
all-sky trend. This does not happen, especially for winter, likely due
to the way the monthly means for clear and cloudy skies were calculated.
Daily means were calculated only when at least \(60\%\) of the clear-
or cloudy-sky data were available (see Section \ref{aggregationstatistical}).

\begin{table}[H]

\caption{\label{tab:trendseasontable}SDR linear trends of monthly anomalies for each season of the year and related statistical parameters.}
\begin{tabu} to \linewidth {>{\centering\arraybackslash}p{8em}>{\centering}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X}
\toprule
\multirow{-1}{*}{\textbf{Sky Condition}} & \multirow{-1}{*}{\textbf{Season}} & \textbf{Trend [\%/year]} & \textbf{Trend S.E. (\boldmath$2\sigma$)} & \textbf{Pearson Correl.} & \textbf{Trend \emph{p}-Value}\\
\midrule
\cellcolor{gray!6}{} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{0.70} & \cellcolor{gray!6}{0.43} & \cellcolor{gray!6}{0.54} & \cellcolor{gray!6}{0.003}\\

 & Spring & 0.11 & 0.24 & 0.17 & 0.371\\

\cellcolor{gray!6}{} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{0.11} & \cellcolor{gray!6}{0.15} & \cellcolor{gray!6}{0.25} & \cellcolor{gray!6}{0.175}\\

\multirow{-4}{*}{\centering\arraybackslash All skies} & Autumn & 0.42 & 0.30 & 0.47 & 0.009\\
\cmidrule{1-6}
\cellcolor{gray!6}{} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{0.40} & \cellcolor{gray!6}{0.20} & \cellcolor{gray!6}{0.60} & \cellcolor{gray!6}{0.001}\\

 & Spring & 0.06 & 0.17 & 0.13 & 0.497\\

\cellcolor{gray!6}{} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{$-$0.05} & \cellcolor{gray!6}{0.06} & \cellcolor{gray!6}{$-$0.30} & \cellcolor{gray!6}{0.106}\\

\multirow{-4}{*}{\centering\arraybackslash Clear skies} & Autumn & 0.05 & 0.12 & 0.17 & 0.366\\
%\cmidrule{1-6}
\bottomrule
\end{tabu}
\end{table}

\begin{table}[H]\ContinuedFloat

\caption{\label{tab:trendseasontable}\emph{Cont.}}
\begin{tabu} to \linewidth {>{\centering\arraybackslash}p{8em}>{\centering}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X}
\toprule
\multirow{-1}{*}{\textbf{Sky Condition}} & \multirow{-1}{*}{\textbf{Season}} & \textbf{Trend [\%/year]} & \textbf{Trend S.E. (\boldmath$2\sigma$)} & \textbf{Pearson Correl.} & \textbf{Trend \emph{p}-Value}\\
\midrule
\cellcolor{gray!6}{} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{0.76} & \cellcolor{gray!6}{0.40} & \cellcolor{gray!6}{0.59} & \cellcolor{gray!6}{0.001}\\

 & Spring & 0.06 & 0.23 & 0.10 & 0.593\\

\cellcolor{gray!6}{} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{$-$0.08} & \cellcolor{gray!6}{0.27} & \cellcolor{gray!6}{$-$0.11} & \cellcolor{gray!6}{0.560}\\

\multirow{-4}{*}{\centering\arraybackslash Cloudy skies} & Autumn & 0.19 & 0.43 & 0.16 & 0.384\\
\bottomrule
\end{tabu}
\end{table}

\vspace{-7pt}

\begin{figure}[H]
    \begin{adjustwidth}{-\extralength}{0cm}
        {\centering 
            \includegraphics[width=1\linewidth]{./images/SeasonalMTrendsTogether3-2}   % Seasonal from Monthly
        }

    \end{adjustwidth}
 \caption{\colorbox{red}{Monthly} %MDPI: Please add the explanation for subfifgures in the figure.
 mean anomalies of SDR by season (rows of plots) for the three sky conditions (columns of plots). The black lines represent the linear trends.}\label{fig:seasonalALL}
\end{figure}

\hypertarget{conclusions}{%
\section{\colorbox{green}{Conclusions}\label{conclusions}}} %MDPI: Please make sure no figure in paper is repeated

We have analyzed a 30-year dataset of SRD measurements in Thessaloniki,
Greece (1993--2023), aiming to identify the long-term variability of
solar irradiance under different sky conditions. Under all-sky
conditions, there is a positive trend in SDR of \(0.38\%/\)y
(brightening). A previous study \citep{Bais2013} for the period 1993--2011 also reported a positive trend of \(0.33\%/\)y. The slight
increase in this trend indicates that the brightening of SDR continues
and is likely caused by continuing decreases in aerosol optical depth
and the optical thickness of clouds over the area. A smaller trend has
been found under clear-sky conditions (\(0.097\%/\)y), which supports
the notion that part of the brightening is caused by decreasing
aerosols. \citet{Siomos2020} showed that aerosol optical depth over
Thessaloniki was decreasing constantly, at least up until 2018. The
attenuation of SDR by aerosols over Europe has been proposed as major
factor by \citet{Wild2021}. Unfortunately, for this study, aerosol data
for the entire period were not available in order to quantify their
effect on SDR. The brightening effect on SDR under cloudy-sky conditions
(\(0.41\%/\)y) suggests that cloud optical thickness is also
decreasing during this period. As long-term data of cloud optical
thickness are also not available for the region, we cannot draw
quantitative~conclusions.

The observed brightening on SDR over Thessaloniki is dependent on SZA
(larger SZAs lead to stronger brightening). The trend is also dependent
on season, with winter showing the strongest statistically significant
trend of \(0.7\) and \(0.76\%/\)y for all- and cloudy-skies,
respectively, in contrast to spring and summer. The trends for autumn
are also significant but smaller (\(0.42\) and \(0.19\%/\)y for all-
and cloudy-skies, respectively). The trend for clear skies is largest in
winter (\(0.4\%/\)y) and negligible in spring, summer, and autumn.

Using the CUSUMs of the monthly departures for all and cloudy skies, we
observed a 20-year period starting around 2000 where the CUSUMs remain
relatively stable, with a steep decline before and a steep increase
after. The rather smooth course of the CUSUMs suggests that no important
change in the SDR pattern has occurred in the entire record.

Continued observations with a collocated pyrheliometer, which started in
2016, will allow us to further investigate the variability of solar
radiation at ground level in Thessaloniki. Also, additional data of
cloudiness, aerosols, atmospheric water vapor, etc., will allow better
attribution and quantification of the effects of these factors on SRD.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\vspace{6pt}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% optional

% Only for the journal Methods and Protocols:
% If you wish to submit a video article, please do so with any other supplementary material.
% \supplementary{The following supporting information can be downloaded at: \linksupplementary{s1}, Figure S1: title; Table S1: title; Video S1: title. A supporting video article is available at doi: link.}
\authorcontributions{\hl{  } %MDPI: For research articles with several authors, a short paragraph specifying their individual contributions must be provided. The following statements should be used ``Conceptualization, X.X. and Y.Y.; methodology, X.X.; software, X.X.; validation, X.X., Y.Y. and Z.Z.; formal analysis, X.X.; investigation, X.X.; resources, X.X.; data curation, X.X.; writing---original draft preparation, X.X.; writing---review and editing, X.X.; visualization, X.X.; supervision, X.X.; project administration, X.X.; funding acquisition, Y.Y. All authors have read and agreed to the published version of the manuscript.'', please turn to the  \href{http://img.mdpi.org/data/contributor-role-instruction.pdf}{CRediT taxonomy} for the term explanation. Authorship must be limited to those who have contributed substantially to the work~reported.
}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\funding{This research received no external funding.}

\institutionalreview{\hl{  } %MDPI: In this section, you should add the Institutional Review Board Statement and approval number, if relevant to your study. You might choose to exclude this statement if the study did not require ethical approval. Please note that the Editorial Office might ask you for further information. Please add “The study was conducted in accordance with the Declaration of Helsinki, and approved by the Institutional Review Board (or Ethics Committee) of NAME OF INSTITUTE (protocol code XXX and date of approval).” for studies involving humans. OR “The animal study protocol was approved by the Institutional Review Board (or Ethics Committee) of NAME OF INSTITUTE (protocol code XXX and date of approval).” for studies involving animals. OR “Ethical review and approval were waived for this study due to REASON (please provide a detailed justification).” OR “Not applicable” for studies not involving humans or animals.
}

\informedconsent{\hl{  } %MDPI: Any research article describing a study involving humans should contain this statement. Please add ``Informed consent was obtained from all subjects involved in the study.'' OR ``Patient consent was waived due to REASON (please provide a detailed justification).'' OR ``Not applicable'' for studies not involving humans. You might also choose to exclude this statement if the study did not involve humans.

%Written informed consent for publication must be obtained from participating patients who can be identified (including by the patients themselves). Please state ``Written informed consent has been obtained from the patient(s) to publish this paper'' if applicable.
}

\dataavailability{Data as daily sums are available through the WRDC
database, \url{http://wrdc.mgo.rssi.ru}. One-minute data are available on
request from the corresponding author. The data are not publicly
available for protection against unmonitored commercial use.}

\conflictsofinterest{\hl{  } %MDPI: Declare conflicts of interest or state ``The authors declare no conflict of interest.'' Authors must identify and declare any personal circumstances or interest that may be perceived as inappropriately influencing the representation or interpretation of reported research results. Any role of the funders in the design of the study; in the collection, analyses or interpretation of data; in the writing of the manuscript; or in the decision to publish the results must be declared in this section. If there is no role, please state ``The funders had no role in the design of the study; in the collection, analyses, or interpretation of data; in the writing of the manuscript; or in the decision to publish the results''.
} 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Optional

%% Only for journal Encyclopedia

\abbreviations{Abbreviations}{
The following abbreviations are used in this manuscript:\\

\noindent
\begin{tabular}{@{}ll}
DNI & Direct beam/normal irradiance \\
ERA5 & ECMWF Reanalysis v5 \\
CSid & Clear sky identification algorithm \\
CUSUM & Cumulative sum \\
SDR & Solar downward radiation \\
SZA & Solar zenith angle \\
\end{tabular}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Optional
%\input{"appendix.tex"}

% % \renewcommand\thefigure{\thesection.\arabic{figure}} %% my
% % \renewcommand\thefigure{\thesection.\arabic{figure}}
%
% %% From R rticle MDPI template
% %% optional
% \appendixtitles{no} %Leave argument "no" if all appendix headings stay EMPTY (then no dot is printed after "Appendix A"). If the appendix sections contain a heading then change the argument to "yes".
% \appendixsections{one} %Leave argument "multiple" if there are multiple sections. Then a counter is printed ("Appendix A"). If there is only one appendix section then change the argument to "one" and no counter is printed ("Appendix").
%
% % \appendix
% % \section{}
%
% \appendix
% \setcounter{secnumdepth}{0}
% \section{Appendix}
%
% \setcounter{figure}{0}    %% my

%% From raw MDPI template
%% Optional
\appendixtitles{no} % Leave argument "no" if all appendix headings stay EMPTY (then no dot is printed after "Appendix A"). If the appendix sections contain a heading then change the argument to "yes".
\appendixstart
\appendix
\section[\appendixname~\thesection]{}

\label{app1}

% copy output from Rmd

\begin{figure}[H]
    {
        \includegraphics[width=0.75\linewidth]{./images/LongtermTrends-2}

    }
    \caption{Anomalies (\%) of the daily clear-sky SDR, relative to climatological values for 1993--2023. The black line shows the long-term linear trend for clear-sky conditions.}\label{fig:trendCLEAR}
\end{figure}

\begin{figure}[H]
    {
        \includegraphics[width=0.75\linewidth]{./images/LongtermTrends-3}

    }
    \caption{Anomalies (\%) of the daily cloud-sky SDR, relative to climatological values for 1993--2023. The black line shows the long-term linear trend for cloud-sky conditions.}\label{fig:trendCLOUD}
\end{figure}

\vspace{-6pt}

\begin{figure}[H]
    \begin{adjustwidth}{-\extralength}{0cm}
        {\centering
            \includegraphics[width=1.0\linewidth]{./images/SzaTrendsSeasTogether-2}
        }
   \end{adjustwidth}
        \caption{\colorbox{red}{Long} %MDPI: Please add the explanation for subfifgures in the figure.
 -term trends of SDR as a function of SZA separately form morning
               and afternoon periods, by season (rows of plots) for the three sky
               conditions (columns of plots).
               Solid shapes represent statistical significant trends ($p<0.005$).
               Cases where $p<0.005$ or with less than $85$~observations may be
               missing from view.}\label{fig:SZAtrendSeason}

\end{figure}

\FloatBarrier

% \subsection{}
% The appendix is an optional section that can contain details and data supplemental to the main text. For example, explanations of experimental details that would disrupt the flow of the main text, but nonetheless remain crucial to understanding and reproducing the research shown; figures of replicates for experiments of which representative data is shown in the main text can be added here if brief, or as Supplementary data. Mathematical proofs of results not central to the paper can be added as an appendix.
%
% \section{}
% All appendix sections must be cited in the main text. In the appendixes, Figures, Tables, etc. should be labeled starting with `A', e.g., Figure A1, Figure A2, etc.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{adjustwidth}{-\extralength}{0cm}

%\printendnotes[custom] % Un-comment to print a list of endnotes

\reftitle{References}
%\bibliography{manualreferences.bib}
\begin{thebibliography}{999}

\bibitem[Wild(2009)]{Wild2009}
Wild, M.
\newblock Global dimming and brightening: A review.
\newblock {\em J. Geophys. Res. Atmos.} {\bf 2009}, {\em
  114},~1--31.
\newblock {\url{https://doi.org/10/bcq}}.

\bibitem[Yang et~al.(2021)Yang, Zhou, Yu, and Wild]{Yang2021}
Yang, S.; Zhou, Z.; Yu, Y.; Wild, M.
\newblock Cloud {\textquotedblleft}shrinking{\textquotedblright} and
  {\textquotedblleft}optical thinning{\textquotedblright} in the
  {\textquotedblleft}dimming{\textquotedblright} period and a subsequent
  recovery in the {\textquotedblleft}brightening{\textquotedblright} period
  over China.
\newblock {\em Environ. Res. Lett.} {\bf 2021}, \emph{16}, 034013.
\newblock {\url{https://doi.org/10.1088/1748-9326/abdf89}}.

\bibitem[Wild et~al.(2021)Wild, Wacker, Yang, and Sanchez-Lorenzo]{Wild2021}
Wild, M.; Wacker, S.; Yang, S.; Sanchez-Lorenzo, A.
\newblock Evidence for Clear‐Sky Dimming and Brightening in Central Europe.
\newblock {\em Geophys. Res. Lett.} {\bf 2021}, {\em 48}, e2020GL092216.
\newblock {\url{https://doi.org/10.1029/2020GL092216}}.

\bibitem[Yamasoe et~al.(2021)Yamasoe, Ros{\'{a}}rio, Almeida, and
  Wild]{Yamasoe2021}
Yamasoe, M.A.; Ros{\'{a}}rio, N.M.{\'{E}}.; Almeida, S.N.S.M.; Wild, M.
\newblock Fifty-six years of surface solar radiation and sunshine duration over
  S{\~{a}}o Paulo, Brazil: 1961--2016.
\newblock {\em Atmos. Chem. Phys.} {\bf 2021}, {\em
  21},~6593--6603.
\newblock {\url{https://doi.org/10.5194/acp-21-6593-2021}}.

\bibitem[Yuan et~al.(2021)Yuan, Leirvik, and Wild]{Yuan2021}
Yuan, M.; Leirvik, T.; Wild, M.
\newblock Global trends in downward surface solar radiation from spatial
  interpolated ground observations during 1961--2019.
\newblock {\em J. Clim.} {\bf 2021}, \emph{34}, 9501--9521.
\newblock {\url{https://doi.org/10.1175/JCLI-D-21-0165.1}}.

\bibitem[Li et~al.(2016)Li, Lau, Ramanathan, Wu, Ding, Manoj, Liu, Qian, Li,
  Zhou, Fan, Rosenfeld, Ming, Wang, Huang, Wang, Xu, Lee, Cribb, Zhang, Yang,
  Zhao, Takemura, Wang, Xia, Yin, Zhang, Guo, Zhai, Sugimoto, Babu, and
  Brasseur]{Li2016}
Li, Z.; Lau, W.K.; Ramanathan, V.; Wu, G.; Ding, Y.; Manoj, M.G.; Liu, J.;
  Qian, Y.; Li, J.; Zhou, T.;  et~al.
\newblock Aerosol and monsoon climate interactions over Asia.
\newblock {\em Rev. Geophys.} {\bf 2016}, {\em 54},~866--929.
\newblock {\url{https://doi.org/10.1002/2015RG000500}}.

\bibitem[Samset et~al.(2018)Samset, Sand, Smith, Bauer, Forster, Fuglestvedt,
  Osprey, and Schleussner]{Samset2018}
Samset, B.H.; Sand, M.; Smith, C.J.; Bauer, S.E.; Forster, P.M.; Fuglestvedt,
  J.S.; Osprey, S.; Schleussner, C.
\newblock Climate Impacts From a Removal of Anthropogenic Aerosol Emissions.
\newblock {\em Geophys. Res. Lett.} {\bf 2018}, {\em 45},~1020--1029.
\newblock {\url{https://doi.org/10.1002/2017GL076079}}.

\bibitem[Schwarz et~al.(2020)Schwarz, Folini, Yang, Allan, and
  Wild]{Schwarz2020}
Schwarz, M.; Folini, D.; Yang, S.; Allan, R.P.; Wild, M.
\newblock Changes in atmospheric shortwave absorption as important driver of
  dimming and brightening.
\newblock {\em Nat. Geosci.} {\bf 2020}, {\em 13},~110--115.
\newblock {\url{https://doi.org/10.1038/s41561-019-0528-y}}.

\bibitem[Ohvril et~al.(2009)Ohvril, Teral, Neiman, Kannel, Uustare, Tee,
  Russak, Okulov, Jõeveer, Kallis, Ohvril, Terez, Terez, Gushchin, Abakumova,
  Gorbarenko, Tsvetkov, and Laulainen]{Ohvril2009}
Ohvril, H.; Teral, H.; Neiman, L.; Kannel, M.; Uustare, M.; Tee, M.; Russak,
  V.; Okulov, O.; Jõeveer, A.; Kallis, A.;  et~al.
\newblock Global dimming and brightening versus atmospheric column
  transparency, Europe, 1906–2007.
\newblock {\em J. Geophys. Res.} {\bf 2009}, {\em 114}.
\newblock {\url{https://doi.org/10.1029/2008JD010644}}.

\bibitem[Zerefos et~al.(2009)Zerefos, Eleftheratos, Meleti, Kazadzis, Romanou,
  Ichoku, Tselioudis, and Bais]{Zerefos2009}
Zerefos, C.S.; Eleftheratos, K.; Meleti, C.; Kazadzis, S.; Romanou, A.; Ichoku,
  C.; Tselioudis, G.; Bais, A.
\newblock Solar dimming and brightening over Thessaloniki, Greece, and Beijing,
  China.
\newblock {\em Tellus B Chem. Phys. Meteorol.} {\bf 2009}, {\em
  61},~657.
\newblock {\url{https://doi.org/10.1111/j.1600-0889.2009.00425.x}}.

\bibitem[Xia et~al.(2007)Xia, Chen, Li, Wang, and Wang]{Xia2007}
Xia, X.; Chen, H.; Li, Z.; Wang, P.; Wang, J.
\newblock Significant reduction of surface solar irradiance induced by aerosols
  in a suburban region in northeastern China.
\newblock {\em J. Geophys. Res. Atmos.} {\bf 2007}, {\em
  112},~1--9.
\newblock {\url{https://doi.org/10/cdtntw}}.

\bibitem[Fountoulakis et~al.(2016)Fountoulakis, Redondas, Bais,
  Rodriguez-Franco, Fragkos, and Cede]{Fountoulakis2016}
Fountoulakis, I.; Redondas, A.; Bais, A.F.; Rodriguez-Franco, J.J.; Fragkos,
  K.; Cede, A.
\newblock Dead time effect on the Brewer measurements: Correction and estimated
  uncertainties.
\newblock {\em Atmos. Meas. Tech.} {\bf 2016}, {\em
  9},~1799--1816.
\newblock {\url{https://doi.org/10/gcc32t}}.

\bibitem[Siomos et~al.(2018)Siomos, Balis, Voudouri, Giannakaki, Filioglou,
  Amiridis, Papayannis, and Fragkos]{Siomos2018}
Siomos, N.; Balis, D.S.; Voudouri, K.A.; Giannakaki, E.; Filioglou, M.;
  Amiridis, V.; Papayannis, A.; Fragkos, K.
\newblock Are {EARLINET} and {AERONET} climatologies consistent? The case of
  Thessaloniki, Greece.
\newblock {\em Atmos. Chem. Phys.} {\bf 2018}, {\em
  18},~11885--11903.
\newblock {\url{https://doi.org/10.5194/acp-18-11885-2018}}.

\bibitem[Gkikas et~al.(2013)Gkikas, Hatzianastassiou, Mihalopoulos, Katsoulis,
  Kazadzis, Pey, Querol, and Torres]{Gkikas2013}
Gkikas, A.; Hatzianastassiou, N.; Mihalopoulos, N.; Katsoulis, V.; Kazadzis,
  S.; Pey, J.; Querol, X.; Torres, O.
\newblock The regime of intense desert dust episodes in the Mediterranean based
  on contemporary satellite observations and ground measurements.
\newblock {\em Atmos. Chem. Phys.} {\bf 2013}, {\em
  13},~12135--12154.
\newblock {\url{https://doi.org/10.5194/acp-13-12135-2013}}.

\bibitem[Lozano et~al.(2021)Lozano, Sánchez-Hernández, Guerrero-Rascado,
  Alados, and Foyo-Moreno]{Lozano2021}
Lozano, I.L.; Sánchez-Hernández, G.; Guerrero-Rascado, J.L.; Alados, I.;
  Foyo-Moreno, I.
\newblock Aerosol radiative effects in photosynthetically active radiation and
  total irradiance at a Mediterranean site from an 11-year database.
\newblock {\em Atmos. Res.} {\bf 2021}, {\em 255},~105538.
\newblock {\url{https://doi.org/10.1016/j.atmosres.2021.105538}}.

\bibitem[Bais et~al.(2013)Bais, Drosoglou, Meleti, Tourpali, and
  Kouremeti]{Bais2013}
Bais, A.F.; Drosoglou, T.; Meleti, C.; Tourpali, K.; Kouremeti, N.
\newblock Changes in surface shortwave solar irradiance from 1993 to 2011 at
  Thessaloniki (Greece).
\newblock {\em Int. J. Climatol.} {\bf 2013}, {\em
  33},~2871--2876.
\newblock {\url{https://doi.org/10/f5dzz5}}.

\bibitem[Reno and Hansen(2016)]{Reno2016}
Reno, M.J.; Hansen, C.W.
\newblock Identification of periods of clear sky irradiance in time series of
  GHI measurements.
\newblock {\em Renew. Energy} {\bf 2016}, {\em 90},~520--531.
\newblock {\url{https://doi.org/10/gq3sbg}}.

\bibitem[Reno et~al.(2012)Reno, Hansen, and Stein]{Reno2012a}
\colorbox{green}{Reno,} %MDPI: Refs. 18 and 26 are duplicated. Please remove duplicated references and rearrange all the references to appear in numerical order. Please ensure that there are no duplicated references.
 M.J.; Hansen, C.W.; Stein, J.S.
\newblock {\emph{Global Horizontal Irradiance Clear Sky Models: Implementation and
  Analysis}};
\newblock \hl{Technical report;}  %MDPI: Please confirm if it can be removed. Please add the name of the publisher and their location.
 2012.

\bibitem[Long and Shi(2006)]{Long2006}
Long, C.N.; Shi, Y.
\newblock \emph{The QCRad Value Added Product: Surface Radiation Measurement Quality
  Control Testing, Including Climatology Configurable Limits};
\newblock \hl{Technical Report DOE/SC-ARM/TR-074;} %MDPI: Please confirm if it can be removed. Please add the name of the publisher and their location.
  2006.

\bibitem[Long and Shi(2008)]{Long2008a}
Long, C.N.; Shi, Y.
\newblock An Automated Quality Assessment and Control Algorithm for Surface
  Radiation Measurements.
\newblock {\em  Open Atmos. Sci. J.} {\bf 2008}, \emph{2}, 23--37.

\bibitem[Coddington et~al.(2005)Coddington, Lean, Lindholm, Pilewskie, Snow,
  and {NOAA CDR Program}]{Coddington2005}
Coddington, O.; Lean, J.L.; Lindholm, D.; Pilewskie, P.; Snow, M.; {NOAA CDR
  Program}.
\newblock \emph{{NOAA} Climate Data Record ({CDR}) of Total Solar Irradiance ({TSI}),
  {NRLTSI} Version 2}; \hl{{D}aily;} %MDPI: Please confirm if it can be removed. Please add the name of the publisher and their location.
  2005.
\newblock {\url{https://doi.org/10.7289/V55B00C1}}.

\bibitem[Haurwitz(1945)]{Haurwitz1945}
Haurwitz, B.
\newblock Insolation in {Relation} to {Cloudiness} and {Cloud} {Density}.
\newblock {\em J. Meteorol.} {\bf 1945}, {\em 2},~154--166.

\bibitem[Long and Ackerman(2000)]{Long2000}
Long, C.N.; Ackerman, T.P.
\newblock Identification of clear skies from broadband pyranometer measurements
  and calculation of downwelling shortwave cloud effects.
\newblock {\em J. Geophys. Res. Atmos.} {\bf 2000}, {\em
  105},~15609--15626.
\newblock {\url{https://doi.org/10.1029/2000jd900077}}.

\bibitem[Brent(1973)]{Brent1973}
Brent, R.P.
\newblock \emph{Algorithms for Minimization without Derivatives};
\newblock {\em PrenticeHall: Englewood Cliffs, NJ, USA,} {1973}.

\bibitem[{R Core Team}(2023)]{RCT2023}
{R Core Team}.
\newblock {\em R: A Language and Environment for Statistical Computing};
\newblock R Foundation for Statistical Computing: Vienna, Austria,  2023.

\bibitem[Reno et~al.(2012)Reno, Hansen, and Stein]{Reno2012}
\colorbox{red}{Reno,} 
 M.J.; Hansen, C.W.; Stein, J.S.
\newblock \emph{Global Horizontal Irradiance Clear Sky Models: Implementation and
  Analysis};
\newblock \hl{Technical Report SAND2012-2389, 1039404;} %MDPI: Please confirm if it can be removed. Please add the name of the publisher and their location.
  2012.
\newblock {\url{https://doi.org/10/gq5npv}}.

\bibitem[Gardner et~al.(1980)Gardner, Harvey, and Phillips]{Gardner1980}
Gardner, G.; Harvey, A.C.; Phillips, G.D.A.
\newblock Algorithm {AS} 154: An Algorithm for Exact Maximum Likelihood
  Estimation of Autoregressive-Moving Average Models by Means of Kalman
  Filtering.
\newblock {\em Appl. Stat.} {\bf 1980}, {\em 29},~311.
\newblock {\url{https://doi.org/10.2307/2346910}}.

\bibitem[Jones(1980)]{Jones1980}
Jones, R.H.
\newblock Maximum Likelihood Fitting of {ARMA} Models to Time Series With
  Missing Observations.
\newblock {\em Technometrics} {\bf 1980}, {\em 22},~389--395.
\newblock {\url{https://doi.org/10.1080/00401706.1980.10486171}}.

\bibitem[Yu et~al.(2022)Yu, Zhang, Wang, Qin, Jiang, and Li]{Yu2022}
Yu, L.; Zhang, M.; Wang, L.; Qin, W.; Jiang, D.; Li, J.
\newblock Variability of surface solar radiation under clear skies over
  Qinghai-Tibet Plateau: Role of aerosols and water vapor.
\newblock {\em Atmos. Environ.} {\bf 2022}, {\em 287},~119286.
\newblock {\url{https://doi.org/10.1016/j.atmosenv.2022.119286}}.

\bibitem[Lozano et~al.(2023)Lozano, Alados, and Foyo-Moreno]{Lozano2023}
Lozano, I.L.; Alados, I.; Foyo-Moreno, I.
\newblock Analysis of the solar radiation/atmosphere interaction at a
  Mediterranean site: The role of clouds.
\newblock {\em Atmos. Res.} {\bf 2023}, {\em 296},~107072.
\newblock {\url{https://doi.org/10.1016/j.atmosres.2023.107072}}.

\bibitem[Ohmura(2009)]{Ohmura2009}
Ohmura, A.
\newblock Observed decadal variations in surface solar radiation and their
  causes.
\newblock {\em J. Geophys. Res. Atmos.} {\bf 2009}, {\em 114}.
\newblock {\url{https://doi.org/10.1029/2008JD011290}}.

\bibitem[Regier et~al.(2019)Regier, Brice{\~{n}}o, and Boyer]{Regier2019}
Regier, P.; Brice{\~{n}}o, H.; Boyer, J.N.
\newblock Analyzing and comparing complex environmental time series using a
  cumulative sums approach.
\newblock {\em {MethodsX}} {\bf 2019}, {\em 6},~779--787.
\newblock {\url{https://doi.org/10.1016/j.mex.2019.03.014}}.

\bibitem[Siomos et~al.(2020)Siomos, Fountoulakis, Natsis, Drosoglou, and
  Bais]{Siomos2020}
Siomos, N.; Fountoulakis, I.; Natsis, A.; Drosoglou, T.; Bais, A.
\newblock Automated Aerosol Classification from Spectral {UV} Measurements
  Using Machine Learning Clustering.
\newblock {\em Remote Sens.} {\bf 2020}, {\em 12},~965.
\newblock {\url{https://doi.org/10.3390/rs12060965}}.

\bibitem[Wang et~al.(2021)Wang, Zhang, Sanchez‐Lorenzo, Tanaka, Trentmann,
  Yuan, and Wild]{Wang2021}
Wang, Y.; Zhang, J.; Sanchez‐Lorenzo, A.; Tanaka, K.; Trentmann, J.; Yuan,
  W.; Wild, M.
\newblock Hourly Surface Observations Suggest Stronger Solar Dimming and
  Brightening at Sunrise and Sunset Over China.
\newblock {\em Geophys. Res. Lett.} {\bf 2021}, {\em 48}, e2020GL091422.
\newblock {\url{https://doi.org/10.1029/2020GL091422}}.

\end{thebibliography}

% If authors have biography, please use the format below
%\section*{Short Biography of Authors}
%\bio
%{\raisebox{-0.35cm}{\includegraphics[width=3.5cm,height=5.3cm,clip,keepaspectratio]{Definitions/author1.pdf}}}
%{\textbf{Firstname Lastname} Biography of first author}
%
%\bio
%{\raisebox{-0.35cm}{\includegraphics[width=3.5cm,height=5.3cm,clip,keepaspectratio]{Definitions/author2.jpg}}}
%{\textbf{Firstname Lastname} Biography of second author}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% for journal Sci
%\reviewreports{\\
%Reviewer 1 comments and authors’ response\\
%Reviewer 2 comments and authors’ response\\
%Reviewer 3 comments and authors’ response
%}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\PublishersNote{}
\end{adjustwidth}

\end{document}

### 2023-12-26 MDPI_submission.Rmd
Commit 2023-12-26 21:09

@@
 ---
-title:     "Trends from 30-year observations of downward solar irradiance in Thessaloniki, Greece"
+title:     "Trends from 30-Year Observations of Downward Solar Irradiance in Thessaloniki, Greece"
 author:
@@
 
-# front matter
-# simplesummary: |
-#   A Simple summary goes here.
-
 abstract: |
-  The shortwave downward solar irradiance (SDR) is an important factor that drives climate processes, energy production and can affect all living organisms. 
+  The shortwave downward solar irradiance (SDR) is an important factor that drives climate processes and energy production and can affect all living organisms. 
   Observations of SDR at different locations around the world with different environmental characteristics have been used to investigate its long-term variability and trends at different time scales. 
-Periods of positive trends are referred as brightening periods and of negative trends as dimming periods.
+  Periods of positive trends are referred to as brightening periods and of negative trends as dimming periods. 
   In this study we have used 30 years of pyranometer data in Thessaloniki, Greece, to investigate the variability of SDR under three types of sky conditions (clear-, cloudy- and all-sky). 
@@
   The consistency of these trends, their seasonal variability, and the effect of the solar zenith angle have also been investigated.
-  Under all three sky categories the SDR trend is stronger in winter with $0.7$, $0.4$ and $0.76\,\%/\text{year}$, respectively for all-, clear- and cloudy-skies. 
-  The next larger seasonal trends are in automn $0.42$ and $0.19\,\%/\text{year}$, for all- and cloudy-skies respectively. 
-  The rest seasonal trends are significant smaller, close to zero, with a negative values in summer, for clear- and cloudy-skies.
-  The SDR trend is increasing with increasing solar zenith angle, except under cloudy skies where the trend is highly variable and close to zero.
-  Finally, we are discussing shorter-term variations in SDR anomalies by examining the patterns of the cumulative sums of monthly anomalies from the climatological mean, both before and after removing the long-term trend.
-
-# back matter
+  Under all three sky categories, the SDR trend is stronger in winter, with $0.7$, $0.4$, and $0.76\,\%/\text{year}$, respectively, for all-, clear-, and cloudy-sky conditions. 
+  The next larger seasonal trends are in autumn $0.42$ and $0.19\,\%/\text{year}$, for all and cloudy skies, respectively. 
+  The rest of the seasonal trends are significant smaller, close to zero, with a negative values in summer, for clear and cloudy skies.
+  The SDR trend is increasing with increasing solar zenith angle, except under cloudy skies, where the trend is highly variable and close to zero.
+  Finally, we discuss shorter-term variations in SDR anomalies by examining the patterns of the cumulative sums of monthly anomalies from the climatological mean, both before and after removing the long-term trend.
+
 keywords: |
-  GHI; SDR; solar radiation; Solar Brigthening/Dimming; aerosols; clouds.
+  GHI; SDR; solar radiation; solar brigthening/dimming; aerosols; clouds
 
@@
 
-The shortwave downward solar irradiance (SDR) at Earth's surface plays a significant role, on its climate.
-Changes of the SDR can be related to changes on Earth's energy budget, the mechanisms of climate change, and water and carbon cycles [@Wild2009].
-It can also affect solar and agricultural production, and all living organisms.
-Studies of SDR variability have identified some distinct SDR trends on different regions of the world on different time periods.
-The term 'brightening' is generally used to describe periods of positive SDR trend, and 'dimming' for negative trend [@Wild2009].
-There are many cases in the long term records of irradiance, showing a systematic change in the magnitude of the trend, occurring roughly in the last decades of the 20th century.
+The shortwave downward solar irradiance (SDR) at Earth's surface plays a significant role on its climate.
+Changes in the SDR can be related to changes in Earth's energy budget, the mechanisms of climate change, and water and carbon cycles [@Wild2009].
+It can also affect solar and agricultural production and all living organisms.
+Studies of SDR variability have identified some distinct SDR trends in different regions of the world in different time periods.
+The term 'brightening' is generally used to describe periods of positive SDR trend, and 'dimming' for periods of negative trend [@Wild2009].
+There are many cases in the long-term records of irradiance showing a systematic change in the magnitude of the trend, occurring roughly in the last decades of the 20th century.
 <!-- Trends china -->
-On multiple stations in China, a dimming period was reported until about 2000, followed by a brightening period [@Yang2021].
+At multiple stations in China, a dimming period was reported until about 2000, followed by a brightening period [@Yang2021].
 <!-- Trends Europe Brazil -->
@@
 <!-- Trends Global -->
-On global scale, an Artificial Intelligence aided spatial analysis on continental level with data from multiple stations reached similar conclusions for these regions and for the global trend [@Yuan2021].
+On global scale, an artificial intelligence aided spatial analysis on the continental level with data from multiple stations reached similar conclusions for these regions and for the global trend [@Yuan2021].
 
@@
 Those interactions, among other factors, have been analyzed with models [@Li2016; @Samset2018], showing the existence of feedback mechanisms between the two.
-Similar findings have been shown from the analysis of observations at other locations [@Schwarz2020; @Ohvril2009; @Zerefos2009; @Xia2007 and references therein].
-In the Mediterranean region aerosols have been recognized as an important factor affecting the penetration of solar radiation to the surface [@Fountoulakis2016; @Siomos2018; @Gkikas2013; @Lozano2021].
+Similar findings have been shown from the analysis of observations at other locations [@Schwarz2020; @Ohvril2009; @Zerefos2009; @Xia2007] [and references therein].
+In the Mediterranean region aerosols have been recognized as an important factor affecting the penetration of solar radiation at the surface [@Fountoulakis2016; @Siomos2018; @Gkikas2013; @Lozano2021].
 These studies investigated the long-term trend in aerosol optical depth, which has been found to decrease in the last three decades, the transport and composition of aerosols, and their radiative effects.
 
-Due to the significant spatial and temporal variability of the trends, and the contributing factors, there is a constant need to monitor and investigate SDR in different sites in order to estimate the degree of variability, and its relation to the local conditions.
-In this study, we examine the trends of SDR, using ground-based measurements at Thessaloniki, Greece, for the period `r year(min(ALL_1_daily_DESEAS$Date))` to `r year(max(ALL_1_daily_DESEAS$Date))`.
+Due to the significant spatial and temporal variability of the trends and the contributing factors, there is a constant need to monitor and investigate SDR at different sites in order to estimate the degree of variability, and its relation to the local conditions.
+In this study, we examine the trends of SDR, using ground-based measurements at Thessaloniki, Greece, for the period from `r year(min(ALL_1_daily_DESEAS$Date))` to `r year(max(ALL_1_daily_DESEAS$Date))`.
 <!-- goals -->
-We re-evaluated and extended the dataset used by @Bais2013, we applied a different algorithm for the identification of clear-/cloud-sky instances [@Reno2016; @Reno2012a], and we derived the SDR trends for the period of study, under different sky conditions (all-sky, clear-sky and cloud-sky).
+We re-evaluated and extended the dataset used by @Bais2013, we applied a different algorithm for the identification of clear-/cloud-sky instances [@Reno2016; @Reno2012a], and we derived the SDR trends for the period of study, under different sky conditions (all-sky, clear-sky, and cloudy-sky).
 Finally, we investigated the dependence of the trends on solar zenith angle and season.
@@
 
-# Data and methodology
+# Data and Methodology
 
@@
 ($40^\circ\,38'\,$N, $22^\circ\,57'\,$E, $80\,$m\ a.s.l.).
-Here we use data for the period from
+Here, we used data for the period from
 `r min(ALL_1_daily_DESEAS$Date)`
@@
 `r max(ALL_1_daily_DESEAS$Date)`.
-The monitoring site is located
-near the city center thus we expect that measurements are affected by the urban environment, mainly by aerosols.
+The monitoring site was located
+near the city center, thus we expect that measurements are affected by the urban environment, mainly by aerosols.
 During the study period, the pyranometer has been
@@
                         # position    = "H",
-                        caption  = "SDR linear trends of monthly anomalies for each season of the year.")
+                        caption  = "SDR linear trends of monthly anomalies for each season of the year and related statistical parameters.")
 

### 2023-12-26 MDPI_submission.tex
Commit 2023-12-26 21:09

@@
 % Full title of the paper (Capitalized)
-\Title{Trends from 30-year observations of downward solar irradiance in
+\Title{Trends from 30-Year Observations of Downward Solar Irradiance in
 Thessaloniki, Greece}
@@
 % MDPI internal command: Title for citation in the left column
-\TitleCitation{Trends from 30-year observations of downward solar
-irradiance in Thessaloniki, Greece}
+\TitleCitation{Trends from 30-Year Observations of Downward Solar
+Irradiance in Thessaloniki, Greece}
 
@@
 \abstract{The shortwave downward solar irradiance (SDR) is an important
-factor that drives climate processes, energy production and can affect
-all living organisms. Observations of SDR at different locations around
-the world with different environmental characteristics have been used to
-investigate its long-term variability and trends at different time
-scales. Periods of positive trends are referred as brightening periods
-and of negative trends as dimming periods. In this study we have used 30
-years of pyranometer data in Thessaloniki, Greece, to investigate the
-variability of SDR under three types of sky conditions (clear-, cloudy-
-and all-sky). The clear-sky data were identified by applying a cloud
-screening algorithm. We have found a positive trend of
+factor that drives climate processes and energy production and can
+affect all living organisms. Observations of SDR at different locations
+around the world with different environmental characteristics have been
+used to investigate its long-term variability and trends at different
+time scales. Periods of positive trends are referred to as brightening
+periods and of negative trends as dimming periods. In this study we have
+used 30 years of pyranometer data in Thessaloniki, Greece, to
+investigate the variability of SDR under three types of sky conditions
+(clear-, cloudy- and all-sky). The clear-sky data were identified by
+applying a cloud screening algorithm. We have found a positive trend of
 \(0.38\,\%/\text{year}\) for all-sky, \(\sim 0.1\,\%/\text{year}\) for
@@
 of the solar zenith angle have also been investigated. Under all three
-sky categories the SDR trend is stronger in winter with \(0.7\), \(0.4\)
-and \(0.76\,\%/\text{year}\), respectively for all-, clear- and
-cloudy-skies. The next larger seasonal trends are in automn \(0.42\) and
-\(0.19\,\%/\text{year}\), for all- and cloudy-skies respectively. The
-rest seasonal trends are significant smaller, close to zero, with a
-negative values in summer, for clear- and cloudy-skies. The SDR trend is
-increasing with increasing solar zenith angle, except under cloudy skies
-where the trend is highly variable and close to zero. Finally, we are
-discussing shorter-term variations in SDR anomalies by examining the
-patterns of the cumulative sums of monthly anomalies from the
-climatological mean, both before and after removing the long-term
+sky categories, the SDR trend is stronger in winter, with \(0.7\),
+\(0.4\), and \(0.76\,\%/\text{year}\), respectively, for all-, clear-,
+and cloudy-sky conditions. The next larger seasonal trends are in autumn
+\(0.42\) and \(0.19\,\%/\text{year}\), for all and cloudy skies,
+respectively. The rest of the seasonal trends are significant smaller,
+close to zero, with a negative values in summer, for clear and cloudy
+skies. The SDR trend is increasing with increasing solar zenith angle,
+except under cloudy skies, where the trend is highly variable and close
+to zero. Finally, we discuss shorter-term variations in SDR anomalies by
+examining the patterns of the cumulative sums of monthly anomalies from
+the climatological mean, both before and after removing the long-term
 trend.}
@@
 % Keywords
-\keyword{GHI; SDR; solar radiation; Solar Brigthening/Dimming; aerosols;
-clouds.}
+\keyword{GHI; SDR; solar radiation; solar brigthening/dimming; aerosols;
+clouds}
 
@@
 The shortwave downward solar irradiance (SDR) at Earth's surface plays a
-significant role, on its climate. Changes of the SDR can be related to
-changes on Earth's energy budget, the mechanisms of climate change, and
+significant role on its climate. Changes in the SDR can be related to
+changes in Earth's energy budget, the mechanisms of climate change, and
 water and carbon cycles \citep{Wild2009}. It can also affect solar and
-agricultural production, and all living organisms. Studies of SDR
-variability have identified some distinct SDR trends on different
-regions of the world on different time periods. The term `brightening'
+agricultural production and all living organisms. Studies of SDR
+variability have identified some distinct SDR trends in different
+regions of the world in different time periods. The term `brightening'
 is generally used to describe periods of positive SDR trend, and
-`dimming' for negative trend \citep{Wild2009}. There are many cases in
-the long term records of irradiance, showing a systematic change in the
-magnitude of the trend, occurring roughly in the last decades of the
-20th century. On multiple stations in China, a dimming period was
+`dimming' for periods of negative trend \citep{Wild2009}. There are many
+cases in the long-term records of irradiance showing a systematic change
+in the magnitude of the trend, occurring roughly in the last decades of
+the 20th century. At multiple stations in China, a dimming period was
 reported until about 2000, followed by a brightening period
@@
 point around 1980, for stations in Central Europe \citep{Wild2021} and
-Brazil \citep{Yamasoe2021}. On global scale, an Artificial Intelligence
-aided spatial analysis on continental level with data from multiple
+Brazil \citep{Yamasoe2021}. On global scale, an artificial intelligence
+aided spatial analysis on the continental level with data from multiple
 stations reached similar conclusions for these regions and for the
@@
 between the two. Similar findings have been shown from the analysis of
-observations at other locations \citep[ and references
-therein]{Schwarz2020, Ohvril2009, Zerefos2009, Xia2007}. In the
-Mediterranean region aerosols have been recognized as an important
-factor affecting the penetration of solar radiation to the surface
-\citep{Fountoulakis2016, Siomos2018, Gkikas2013, Lozano2021}. These
-studies investigated the long-term trend in aerosol optical depth, which
-has been found to decrease in the last three decades, the transport and
-composition of aerosols, and their radiative effects.
+observations at other locations
+\citep{Schwarz2020, Ohvril2009, Zerefos2009, Xia2007} {[}and references
+therein{]}. In the Mediterranean region aerosols have been recognized as
+an important factor affecting the penetration of solar radiation at the
+surface \citep{Fountoulakis2016, Siomos2018, Gkikas2013, Lozano2021}.
+These studies investigated the long-term trend in aerosol optical depth,
+which has been found to decrease in the last three decades, the
+transport and composition of aerosols, and their radiative effects.
 
@@
 and the contributing factors, there is a constant need to monitor and
-investigate SDR in different sites in order to estimate the degree of
+investigate SDR at different sites in order to estimate the degree of
 variability, and its relation to the local conditions. In this study, we
 examine the trends of SDR, using ground-based measurements at
-Thessaloniki, Greece, for the period 1993 to 2023. We re-evaluated and
-extended the dataset used by \citet{Bais2013}, we applied a different
-algorithm for the identification of clear-/cloud-sky instances
+Thessaloniki, Greece, for the period from 1993 to 2023. We re-evaluated
+and extended the dataset used by \citet{Bais2013}, we applied a
+different algorithm for the identification of clear-/cloud-sky instances
 \citep{Reno2016, Reno2012a}, and we derived the SDR trends for the
-period of study, under different sky conditions (all-sky, clear-sky and
-cloud-sky). Finally, we investigated the dependence of the trends on
+period of study, under different sky conditions (all-sky, clear-sky, and
+cloudy-sky). Finally, we investigated the dependence of the trends on
 solar zenith angle and season.
@@
 \hypertarget{data-and-methodology}{%
-\section{Data and methodology}\label{data-and-methodology}}
+\section{Data and Methodology}\label{data-and-methodology}}
 
@@
 Aristotle University of Thessaloniki (\(40^\circ\,38'\,\)N,
-\(22^\circ\,57'\,\)E, \(80\,\)m~a.s.l.). Here we use data for the period
-from 1993-04-13 to 2023-04-13. The monitoring site is located near the
-city center thus we expect that measurements are affected by the urban
-environment, mainly by aerosols. During the study period, the
+\(22^\circ\,57'\,\)E, \(80\,\)m~a.s.l.). Here, we used data for the
+period from 1993-04-13 to 2023-04-13. The monitoring site was located
+near the city center, thus we expect that measurements are affected by
+the urban environment, mainly by aerosols. During the study period, the
 pyranometer has been independently calibrated three times at the
@@
 respective datasets, the radiative effect of clouds is estimated to
-\(0.96\,W/m^3/y\). This estimate is similar to the cloud radiative
+\(0.96\,W/m^2/y\). This estimate is similar to the cloud radiative
 forcing of \(1.22\,W/m^2/y\) reported for Granada, Spain

### 2023-12-26 manuscript.v5/applsci-2743235.tex
Commit 2023-12-26 21:09

@@
 Natsis \href{https://orcid.org/0000-0002-5199-4119}
-{\orcidicon}, Alkiviadis Bais * and Charikleia Meleti}
+{\orcidicon}, Alkiviadis Bais \href{https://orcid.org/0000-0003-3899-2001}
+{\orcidicon}* and Charikleia Meleti}
 

### 2023-12-26 MDPI_submission.Rmd
Commit 2023-12-26 22:14

@@
 The monitoring site was located
-near the city center, thus we expect that measurements are affected by the urban environment, mainly by aerosols.
-During the study period, the pyranometer has been
+near the city center, thus we expect that measurements were affected by the urban environment, mainly by aerosols.
+During the study period, the pyranometer was
 independently calibrated three times at the Meteorologisches
 Observatorium Lindenberg, DWD, verifying that the stability of the
-instrument's sensitivity is better than $0.7\%$ relative to the initial
+instrument's sensitivity was better than $0.7\%$ relative to the initial
 calibration by the manufacturer. 
 Along with SDR, the direct beam
-radiation (DNI) is also measured with a collocated Kipp & Zonen CHP-1 pyrheliometer since 2016-04-01. 
+radiation (DNI) was also measured with a collocated Kipp & Zonen CHP-1 pyrheliometer since 2016-04-01. 
 The DNI data were used as auxiliary data
 to support the selection of appropriate thresholds in the clear-sky
-identification algorithm (CSid), which is discussed in sect. \ref{CDIDalgorithm}. 
+identification algorithm (CSid), which is discussed in Section \ref{CDIDalgorithm}. 
 It is noted
 that the limited dataset of DNI was not used for the identification of
-clear sky cases in the entire SDR series to avoid any selection bias due
-to unequal length of the two datasets.
+clear-sky cases in the entire SDR series to avoid any selection bias due
+to the unequal length of the two datasets.
 There are four distinct steps in the creation of the dataset analyzed here:
-a)\ the acquisition of radiation measurements from the sensors,
-b)\ the data quality check,
-c)\ the identification of "clear sky" conditions from the SDR data, and
-d)\ the aggregation of data and trend analysis.
+(a)\ the acquisition of radiation measurements from the sensors,
+(b)\ the data quality check,
+(c)\ the identification of "clear sky" conditions from the SDR data, and
+(d)\ the aggregation of data and trend analysis.
 
 <!-- a) acquisition -->
-For the acquisition of radiometric data, the signal of the pyranometer is sampled at a rate of $1\,\text{Hz}$.
-The mean and the standard deviation of these samples are calculated and recorded every minute.
-The measurements are corrected for the zero offset ("dark signal" in volts), which is calculated by averaging all measurements recorded for a period of $3\,\text{h}$, before (morning) or after (evening) the Sun reaches an elevation angle of $-10^\circ$.
-The signal is converted to irradiance using a ramped value of the instrument’s sensitivity between subsequent calibrations.
+For the acquisition of radiometric data, the signal of the pyranometer was sampled at a rate of $1\,\text{Hz}$.
+The mean and the standard deviation of these samples were calculated and recorded every minute.
+The measurements were corrected for the zero offset ("dark signal" in volts), which was calculated by averaging all measurements recorded for a period of $3\,\text{h}$, before (morning) or after (evening) the Sun reaches an elevation angle of $-10^\circ$.
+The signal was converted to irradiance using a ramped value of the instrument’s sensitivity between subsequent calibrations.
 
@@
 A manual screening was performed, to remove inconsistent and erroneous recordings that can occur stochastically or systematically, during the continuous operation of the instruments.
-The manual screening is aided by a radiation data quality assurance procedure, adjusted for the site, which is based on the methods of
+The manual screening was aided by a radiation data quality assurance procedure, adjusted for the site, which was based on the methods of
 Long and Shi\ [@Long2006; @Long2008a].
@@
 Due to the significant measurement uncertainty when the Sun is near the horizon, we have excluded all measurements with solar zenith angle (SZA) greater than $`r 90-MIN_ELEVA`^\circ$.
-Moreover, due to obstructions around the site (hills and buildings) which block the direct irradiance, we excluded data with azimuth angle in the range
-$`r FIBais_Az_1`^{\circ}$ - $`r FIBais_Az_2`^{\circ}$ and with SZA greater than $`r 90 -  FIBais_Elev`^{\circ}$.
+Moreover, due to obstructions around the site (hills and buildings) that block the direct irradiance, we excluded data with azimuth angle in the range
+$`r FIBais_Az_1`^{\circ}$--$`r FIBais_Az_2`^{\circ}$ and with SZA greater than $`r 90 - FIBais_Elev`^{\circ}$.
 <!-- 1au -->
-To make the measurements comparable throughout the dataset, we adjusted all one-minute data to the mean Sun - Earth distance.
+To make the measurements comparable throughout the dataset, we adjusted all one-minute data to the mean Sun--Earth distance.
 <!-- TSI -->

### 2023-12-26 MDPI_submission.tex
Commit 2023-12-26 22:14

@@
 \citep{Schwarz2020, Ohvril2009, Zerefos2009, Xia2007} {[}and references
-therein{]}. In the Mediterranean region aerosols have been recognized as
-an important factor affecting the penetration of solar radiation at the
+therein{]}.\\
+In the Mediterranean region aerosols have been recognized as an
+important factor affecting the penetration of solar radiation at the
 surface \citep{Fountoulakis2016, Siomos2018, Gkikas2013, Lozano2021}.
@@
 
-Due to the significant spatial and temporal variability of the trends,
+Due to the significant spatial and temporal variability of the trends
 and the contributing factors, there is a constant need to monitor and
@@
 period from 1993-04-13 to 2023-04-13. The monitoring site was located
-near the city center, thus we expect that measurements are affected by
+near the city center, thus we expect that measurements were affected by
 the urban environment, mainly by aerosols. During the study period, the
-pyranometer has been independently calibrated three times at the
+pyranometer was independently calibrated three times at the
 Meteorologisches Observatorium Lindenberg, DWD, verifying that the
-stability of the instrument's sensitivity is better than \(0.7\%\)
+stability of the instrument's sensitivity was better than \(0.7\%\)
 relative to the initial calibration by the manufacturer. Along with SDR,
-the direct beam radiation (DNI) is also measured with a collocated Kipp
+the direct beam radiation (DNI) was also measured with a collocated Kipp
 \& Zonen CHP-1 pyrheliometer since 2016-04-01. The DNI data were used as
 auxiliary data to support the selection of appropriate thresholds in the
-clear-sky identification algorithm (CSid), which is discussed in sect.
+clear-sky identification algorithm (CSid), which is discussed in Section
 \ref{CDIDalgorithm}. It is noted that the limited dataset of DNI was not
-used for the identification of clear sky cases in the entire SDR series
-to avoid any selection bias due to unequal length of the two datasets.
-There are four distinct steps in the creation of the dataset analyzed
-here: a)~the acquisition of radiation measurements from the sensors,
-b)~the data quality check, c)~the identification of ``clear sky''
-conditions from the SDR data, and d)~the aggregation of data and trend
-analysis.
+used for the identification of clear-sky cases in the entire SDR series
+to avoid any selection bias due to the unequal length of the two
+datasets. There are four distinct steps in the creation of the dataset
+analyzed here: (a)~the acquisition of radiation measurements from the
+sensors, (b)~the data quality check, (c)~the identification of ``clear
+sky'' conditions from the SDR data, and (d)~the aggregation of data and
+trend analysis.
 
 For the acquisition of radiometric data, the signal of the pyranometer
-is sampled at a rate of \(1\,\text{Hz}\). The mean and the standard
-deviation of these samples are calculated and recorded every minute. The
-measurements are corrected for the zero offset (``dark signal'' in
-volts), which is calculated by averaging all measurements recorded for a
-period of \(3\,\text{h}\), before (morning) or after (evening) the Sun
-reaches an elevation angle of \(-10^\circ\). The signal is converted to
+was sampled at a rate of \(1\,\text{Hz}\). The mean and the standard
+deviation of these samples were calculated and recorded every minute.
+The measurements were corrected for the zero offset (``dark signal'' in
+volts), which was calculated by averaging all measurements recorded for
+a period of \(3\,\text{h}\), before (morning) or after (evening) the Sun
+reaches an elevation angle of \(-10^\circ\). The signal was converted to
 irradiance using a ramped value of the instrument's sensitivity between
@@
 recordings that can occur stochastically or systematically, during the
-continuous operation of the instruments. The manual screening is aided
+continuous operation of the instruments. The manual screening was aided
 by a radiation data quality assurance procedure, adjusted for the site,
-which is based on the methods of Long and
+which was based on the methods of Long and
 Shi~\citep{Long2006, Long2008a}. Thus, problematic recordings have been
@@
 greater than \(85^\circ\). Moreover, due to obstructions around the site
-(hills and buildings) which block the direct irradiance, we excluded
-data with azimuth angle in the range \(58^{\circ}\) - \(120^{\circ}\)
-and with SZA greater than \(78^{\circ}\). To make the measurements
-comparable throughout the dataset, we adjusted all one-minute data to
-the mean Sun - Earth distance. Subsequently, we adjusted all
-measurements to the Total Solar Irradiance (TSI) at \(1\,\text{au}\), in
-order to compensate for the Sun's intensity variability, using a time
-series of satellite TSI observations. The TSI data we used are part of
-the ``NOAA Climate Data Record of Total Solar Irradiance'' dataset
-\citep{Coddington2005}. The initial daily values of this dataset were
-interpolated to match the time step of our measurements.
+(hills and buildings) that block the direct irradiance, we excluded data
+with azimuth angle in the range \(58^{\circ}\)--\(120^{\circ}\) and with
+SZA greater than \(78^{\circ}\). To make the measurements comparable
+throughout the dataset, we adjusted all one-minute data to the mean
+Sun--Earth distance. Subsequently, we adjusted all measurements to the
+Total Solar Irradiance (TSI) at \(1\,\text{au}\), in order to compensate
+for the Sun's intensity variability, using a time series of satellite
+TSI observations. The TSI data we used are part of the ``NOAA Climate
+Data Record of Total Solar Irradiance'' dataset \citep{Coddington2005}.
+The initial daily values of this dataset were interpolated to match the
+time step of our measurements.
 

### 2023-12-27 DHI_GHI_1_longterm_trends.R
Commit 2023-12-27 09:11

@@ -1288,13 +1288,13 @@ for (avar in vars) {
             plot.new()
             text(x = 0.05, y = 0.5,
                  adj  = c(0.5,0.5),
-                 srt  = 90, "Automn", cex = 0.9, font = 2)
+                 srt  = 90, "Autumn", cex = 0.9, font = 2)
         }
         if (i == 25) {
             plot.new()
             text(x = 0.25, y = 0.5,
                  adj  = c(0.5,0.5),
-                 srt  = 90, "Automn", cex = 0.9, font = 2)
+                 srt  = 90, "Autumn", cex = 0.9, font = 2)
         }
 
 
@@ -1643,13 +1643,13 @@ write_dat(pprint,
             plot.new()
             text(x = 0.05, y = 0.5,
                  adj  = c(0.5,0.5),
-                 srt  = 90, "Automn", cex = 0.9, font = 2)
+                 srt  = 90, "Autumn", cex = 0.9, font = 2)
         }
         if (i == 25) {
             plot.new()
             text(x = 0.25, y = 0.5,
                  adj  = c(0.5,0.5),
-                 srt  = 90, "Automn", cex = 0.9, font = 2)
+                 srt  = 90, "Autumn", cex = 0.9, font = 2)
         }
 
 

### 2023-12-27 manuscript.v5/applsci-2743235.tex
Commit 2023-12-27 10:14

--- 2023-12-26 manuscript.v5/applsci-2743235.tex
@@
 Natsis \href{https://orcid.org/0000-0002-5199-4119}
-{\orcidicon}, Alkiviadis Bais * and Charikleia Meleti}
+{\orcidicon}, \colorbox{magenta}{Alkiviadis Bais} \href{https://orcid.org/0000-0003-3899-2001} %Natsis: added orcid  
+{\orcidicon}* and Charikleia Meleti}
 
@@
 observations at other \hl{locations} %MDPI: We revised citation. Please confirm. 
+%Natsis: Confirmed
  \citep{Schwarz2020, Ohvril2009, Zerefos2009, Xia2007} [and references
@@
   \hl{Mean} %MDPI: We removed italics of units. Please confirm. 
+  %Natsis: Confirmed
  of the measured \(\overline{\text{SDR}}_i\) (Equation~\eqref{eq:MeanVIP}): \begin{equation}
@@
 in \mbox{\hl{Appendix} %MDPI: We revised citation. Please confirm. 
+	%Natsis: confirmed
  \ref{app1}} \mbox{(Figures~\ref{fig:trendCLEAR} and~
@@
 linear trends in all three datasets are positive and \hl{around} %MDPI: We removed italics of units. Please confirm. The same below.  
+%Natsis: confirmed
 \(0.4\%/\)y for all-sky and cloudy-sky conditions, whereas for
@@
             {\includegraphics[width=.32\linewidth]{./images/CumulativeMonthlyCuSum-5}}\hfill
-        \subfloat[\label{fig:cusummonth-3}]
+       % \subfloat[\label{fig:cusummonth-3}]  %Natsis: commented this line
             {\includegraphics[width=.32\linewidth]{./images/CumulativeMonthlyCuSum-9}}\hfill
@@
 \end{adjustwidth}
-\caption{\hl{Cumulative} %MDPI: We moved the subfigure explanations into the figure caption. Please confirm.
+\caption{\hl{Cumulative} %MDPI: We moved the subfigure explanations into the figure caption. Please confirm.  %Natsis: confirmed
  sum plots of the monthly SDR anomalies in (\%) for different sky conditions: (\textbf{a})~all skies; (\textbf{b}) clear skies; (\textbf{c}) cloudy skies.}\label{fig:cusummonth}
@@
 \end{adjustwidth}
-        \caption{\hl{Cumulative} %MDPI: We moved the subfigure explanations into the figure caption. Please confirm.
+        \caption{\hl{Cumulative} %MDPI: We moved the subfigure explanations into the figure caption. Please confirm.  %Natsis: confirmed
  sum plots of monthly SDR anomalies in (\%) for different sky conditions after removing the long-term linear trend: (\textbf{a})~all skies; (\textbf{b}) clear skies; (\textbf{c}) cloudy skies.}\label{fig:cusumnotrendmonthly}
@@
     \end{adjustwidth}
-        \caption{\hl{Long} %MDPI: We removed duplicate explanation in figure.
+        \caption{\hl{Long}%MDPI: We removed duplicate explanation in figure. %Natsis: confirmed
  -term trends of daily SDR as a function of SZA for (\textbf{a}) all-sky, (\textbf{b}) clear-sky and (\textbf{c}) cloudy-sky conditions, separately for morning and afternoon periods.
@@
         {\centering 
-            \includegraphics[width=1\linewidth]{./images/SeasonalMTrendsTogether3-2}   % Seasonal from Monthly
+            \includegraphics[width=1\linewidth]{./images/SeasonalMTrendsTogether3-2}   %Natsis: Image update Automn -> Autumn
         }
@@
     \end{adjustwidth}
- \caption{\colorbox{red}{Monthly} %MDPI: Please add the explanation for subfifgures in the figure.
- mean anomalies of SDR by season (rows of plots) for the three sky conditions (columns of plots). The black lines represent the linear trends.}\label{fig:seasonalALL}
+ \caption{Time series of \colorbox{red}{monthly} %MDPI: Please add the explanation for subfifgures in the figure.  %Natsis: Done that, and also rephrased the caption
+ mean anomalies of SDR by season (rows of plots) for the three sky conditions (columns of plots).
+The black lines represent the linear trends, (\textbf{a}-\textbf{c}) winter, (\textbf{d}-\textbf{f}) spring, (\textbf{g}-\textbf{i}) summer, (\textbf{j}-\textbf{i}) autumn, and also (\textbf{a}, \textbf{d}, \textbf{g}, \textbf{i}) all-skies, (\textbf{b}, \textbf{e}, \textbf{h}, \textbf{k}) clear-skies and (\textbf{c}, \textbf{f}, \textbf{i}, \textbf{j}) cloudy-skies.}\label{fig:seasonalALL}
 \end{figure}
@@
 You might also choose to exclude this statement if the study did not involve humans.
-
 %Written informed consent for publication must be obtained from participating patients who can be identified (including by the patients themselves).
 Please state ``Written informed consent has been obtained from the patient(s) to publish this paper'' if applicable.
-}
+Not applicable} %Natsis: section added
 
@@
 If there is no role, please state ``The funders had no role in the design of the study; in the collection, analyses, or interpretation of data; in the writing of the manuscript; or in the decision to publish the results''.
-} 
+The authors declare no conflict of interest.} %Natsis: section added
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2023-12-27 MDPI_submission.Rmd
Commit 2023-12-27 11:10

@@
 <!-- goals -->
-We re-evaluated and extended the dataset used by @Bais2013, we applied a different algorithm for the identification of clear-/cloud-sky instances [@Reno2016; @Reno2012a], and we derived the SDR trends for the period of study, under different sky conditions (all-sky, clear-sky, and cloudy-sky).
+We re-evaluated and extended the dataset used by @Bais2013, we applied a different algorithm for the identification of clear-/cloud-sky instances [@Reno2016; @Reno2012], and we derived the SDR trends for the period of study, under different sky conditions (all-sky, clear-sky, and cloudy-sky).
 Finally, we investigated the dependence of the trends on solar zenith angle and season.

### 2023-12-27 MDPI_submission.tex
Commit 2023-12-27 11:10

@@
 \citep{Schwarz2020, Ohvril2009, Zerefos2009, Xia2007} {[}and references
-therein{]}.\\
-In the Mediterranean region aerosols have been recognized as an
-important factor affecting the penetration of solar radiation at the
+therein{]}. In the Mediterranean region aerosols have been recognized as
+an important factor affecting the penetration of solar radiation at the
 surface \citep{Fountoulakis2016, Siomos2018, Gkikas2013, Lozano2021}.
@@
 different algorithm for the identification of clear-/cloud-sky instances
-\citep{Reno2016, Reno2012a}, and we derived the SDR trends for the
-period of study, under different sky conditions (all-sky, clear-sky, and
+\citep{Reno2016, Reno2012}, and we derived the SDR trends for the period
+of study, under different sky conditions (all-sky, clear-sky, and
 cloudy-sky). Finally, we investigated the dependence of the trends on

### 2023-12-27 manuscript.v5/applsci-2743235.tex
Commit 2023-12-27 11:10

@@
 algorithm for the identification of clear-/cloud-sky instances
-\citep{Reno2016, Reno2012a}, and we derived the SDR trends for the
+\citep{Reno2016, Reno2012}, %Natsis: citation edited
+and we derived the SDR trends for the
 period of study under different sky conditions (all-sky, clear-sky, and
@@
  mean anomalies of SDR by season (rows of plots) for the three sky conditions (columns of plots).
-The black lines represent the linear trends, (\textbf{a}-\textbf{c}) winter, (\textbf{d}-\textbf{f}) spring, (\textbf{g}-\textbf{i}) summer, (\textbf{j}-\textbf{i}) autumn, and also (\textbf{a}, \textbf{d}, \textbf{g}, \textbf{i}) all-skies, (\textbf{b}, \textbf{e}, \textbf{h}, \textbf{k}) clear-skies and (\textbf{c}, \textbf{f}, \textbf{i}, \textbf{j}) cloudy-skies.}\label{fig:seasonalALL}
+The black lines represent the linear trends, (\textbf{a}-\textbf{c}) winter, (\textbf{d}-\textbf{f}) spring, (\textbf{g}-\textbf{i}) summer, (\textbf{j}-\textbf{l}) autumn, and also (\textbf{a}, \textbf{d}, \textbf{g}, \textbf{j}) all-skies, (\textbf{b}, \textbf{e}, \textbf{h}, \textbf{k}) clear-skies and (\textbf{c}, \textbf{f}, \textbf{i}, \textbf{l}) cloudy-skies.}\label{fig:seasonalALL}
 \end{figure}
@@
 Authorship must be limited to those who have contributed substantially to the work~reported.
-}
+Conceptualization, methodology, software, data analysis and original writing---draft preparation A.N.; writing---review and editing, supervision, A.B.; data curation, C.M.
+All authors have read and agreed to the published version of the manuscript.} %Natsis: section added
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 OR “Ethical review and approval were waived for this study due to REASON (please provide a detailed justification).” OR “Not applicable” for studies not involving humans or animals.
-}
+Not applicable}  %Natsis: section added
 
@@
    \end{adjustwidth}
-        \caption{\colorbox{red}{Long} %MDPI: Please add the explanation for subfifgures in the figure.
+        \caption{\colorbox{red}{Long}%MDPI: Please add the explanation for subfifgures in the figure.
  -term trends of SDR as a function of SZA separately form morning
@@
                conditions (columns of plots).
-               Solid shapes represent statistical significant trends ($p<0.005$).
+               Solid shapes represent statistical significant trends ($p<0.005$), (\textbf{a}-\textbf{c}) winter, (\textbf{d}-\textbf{f}) spring, (\textbf{g}-\textbf{i}) summer, (\textbf{j}-\textbf{l}) autumn, and also (\textbf{a}, \textbf{d}, \textbf{g}, \textbf{j}) all-skies, (\textbf{b}, \textbf{e}, \textbf{h}, \textbf{k}) clear-skies and (\textbf{c}, \textbf{f}, \textbf{i}, \textbf{l}) cloudy-skies.
                Cases where $p<0.005$ or with less than $85$~observations may be

### 2023-12-27 DHI_GHI_3_trends_consistency.R

@@ -413,8 +413,6 @@ database    <- c(  "ALL_1_daily_DESEAS",
                  "CLEAR_1_daily_DESEAS",
                  "CLOUD_1_daily_DESEAS")
 
-
-
 #+ CumulativeDailyCuSum, echo=F, include=T, results="asis"
 for (adb in database) {
     DB  <- get(adb)
@@ -695,7 +693,7 @@ for (adb in database) {
                         translate(avar) ), cex.main = 1)
         } else {
             legend("bottomleft", 0, translate(sub("_.*","",adb)),
-                   cex   = 1.1,
+                   cex   = 2,
                    bty   = "n",
                    xjust = 0.5,      # 0.5 means center adjusted
                    yjust = 0.5,      # 0.5 means center adjusted
@@ -974,15 +972,15 @@ for (adb in database) {
             ## tag the corner
             ## translate(sub("_.*","",adb))
             legend("bottomleft", 0, paste0("(",letters[lec],")"),
-                   cex   = 1.1,
+                   # cex   = 1.1,
+                   cex   = 2,
                    bty   = "n",
                    xjust = 0.5,      # 0.5 means center adjusted
                    yjust = 0.5,      # 0.5 means center adjusted
                    x.intersp = -0.5, # adjust character interspacing as you like to effect box width
                    y.intersp =  0.2, # adjust character interspacing to effect box height
                    adj = c(0, 0.5))  # adjust string position (default values used here)
-            # cex = 1.5,      # change cex if you like (not used here)
-            # text.font = 2)  # bold the text if you like (not used here)
+            # text.font = 2  # bold the text if you like (not used here)
             par("mar" = c(3,4,2,1))
         }
 
@@ -1204,7 +1202,7 @@ for (adb in database) {
             ## tag plot with type
             ## translate(sub("_.*","",adb))
             legend("bottomleft", 0, paste0("(",letters[lec],")"),
-                   cex   = 1.1,
+                   cex   = 2,
                    bty   = "n",
                    xjust = 0.5,      # 0.5 means center adjusted
                    yjust = 0.5,      # 0.5 means center adjusted

### 2023-12-27 MDPI_submission.Rmd
Commit 2023-12-27 13:09

@@
 Here, we used data for the period from
-`r min(ALL_1_daily_DESEAS$Date)`
+`r strftime(min(ALL_1_daily_DESEAS$Date), "%d %B %Y")`
 to
-`r max(ALL_1_daily_DESEAS$Date)`.
+`r strftime(max(ALL_1_daily_DESEAS$Date), "%d %B %Y")`.
 The monitoring site was located
@@
 Along with SDR, the direct beam
-radiation (DNI) was also measured with a collocated Kipp & Zonen CHP-1 pyrheliometer since 2016-04-01. 
+radiation (DNI) was also measured with a collocated Kipp & Zonen CHP-1 pyrheliometer since 1 April 2016. 
 The DNI data were used as auxiliary data
@@
 Subsequently, we adjusted all measurements to the Total Solar Irradiance (TSI) at $1\,\text{au}$, in order to compensate for the Sun's intensity variability, using a time series of satellite TSI observations.
-The TSI data we used are part of the "NOAA Climate Data Record of Total Solar Irradiance" dataset [@Coddington2005].
+The TSI data we used are part of the 'NOAA Climate Data Record of Total Solar Irradiance' dataset [@Coddington2005].
 The initial daily values of this dataset were interpolated to match the time step of our measurements.
@@
 <!-- c) Clear sky identification -->
-In order to estimate the effect of the sky conditions on the long term variability of SDR, we created three datasets by characterizing each one-minute measurement with a corresponding sky-condition flag (i.e., all-sky, clear-sky and cloudy-sky).
-To identify the clear-cases we used the method proposed by
+In order to estimate the effect of the sky conditions on the long term variability of SDR, we created three datasets by characterizing each one-minute measurement with a corresponding sky-condition flag (i.e., all-sky, clear-sky, and cloudy-sky).
+To identify the clear cases we used the method proposed by
 @Reno2016, which requires the definition of some site specific parameters.
@@
 As a result, the details of the definition are site specific, and they rely on a combination of thresholds and comparisons with ideal radiation models and statistical analysis of different signal metrics.
-The CSid algorithm was calibrated with the main focus, to identify the presence of clouds. Despite the fine-tuning
+The CSid algorithm was calibrated with the main focus to identify the presence of clouds. Despite the fine-tuning
 of the procedure, in a few marginal cases false positive or false
@@
 However, due to
-their small number, they cannot affect the final results of the study.
+their small number, they did not affect the final results of the study.
 For completeness, we provide below a brief overview of the CSid algorithm, along with the site specific thresholds.

### 2023-12-27 MDPI_submission.tex
Commit 2023-12-27 13:09

@@
 \(22^\circ\,57'\,\)E, \(80\,\)m~a.s.l.). Here, we used data for the
-period from 1993-04-13 to 2023-04-13. The monitoring site was located
-near the city center, thus we expect that measurements were affected by
-the urban environment, mainly by aerosols. During the study period, the
-pyranometer was independently calibrated three times at the
+period from 13 April 1993 to 13 April 2023. The monitoring site was
+located near the city center, thus we expect that measurements were
+affected by the urban environment, mainly by aerosols. During the study
+period, the pyranometer was independently calibrated three times at the
 Meteorologisches Observatorium Lindenberg, DWD, verifying that the
@@
 the direct beam radiation (DNI) was also measured with a collocated Kipp
-\& Zonen CHP-1 pyrheliometer since 2016-04-01. The DNI data were used as
-auxiliary data to support the selection of appropriate thresholds in the
-clear-sky identification algorithm (CSid), which is discussed in Section
-\ref{CDIDalgorithm}. It is noted that the limited dataset of DNI was not
-used for the identification of clear-sky cases in the entire SDR series
-to avoid any selection bias due to the unequal length of the two
+\& Zonen CHP-1 pyrheliometer since 1 April 2016. The DNI data were used
+as auxiliary data to support the selection of appropriate thresholds in
+the clear-sky identification algorithm (CSid), which is discussed in
+Section \ref{CDIDalgorithm}. It is noted that the limited dataset of DNI
+was not used for the identification of clear-sky cases in the entire SDR
+series to avoid any selection bias due to the unequal length of the two
 datasets. There are four distinct steps in the creation of the dataset
@@
 for the Sun's intensity variability, using a time series of satellite
-TSI observations. The TSI data we used are part of the ``NOAA Climate
-Data Record of Total Solar Irradiance'' dataset \citep{Coddington2005}.
+TSI observations. The TSI data we used are part of the `NOAA Climate
+Data Record of Total Solar Irradiance' dataset \citep{Coddington2005}.
 The initial daily values of this dataset were interpolated to match the
@@
 one-minute measurement with a corresponding sky-condition flag (i.e.,
-all-sky, clear-sky and cloudy-sky). To identify the clear-cases we used
+all-sky, clear-sky, and cloudy-sky). To identify the clear cases we used
 the method proposed by \citet{Reno2016}, which requires the definition
@@
 different signal metrics. The CSid algorithm was calibrated with the
-main focus, to identify the presence of clouds. Despite the fine-tuning
+main focus to identify the presence of clouds. Despite the fine-tuning
 of the procedure, in a few marginal cases false positive or false
 negative results were identified by manual inspection. However, due to
-their small number, they cannot affect the final results of the study.
+their small number, they did not affect the final results of the study.
 For completeness, we provide below a brief overview of the CSid

### 2023-12-27 MDPI_submission.Rmd
Commit 2023-12-27 23:09

@@
   - name: Charikleia Meleti
-    affil: 1 #, \dagger, # \ddagger
-
+    affil: 1
 affiliation:
@@
     email: natsisphysicist@gmail.com (A.N.); abais@auth.gr (A.B.); meleti@auth.gr (C.M.)
-  # - num: 2
-  #   address: |
-  #     Your department
-  #     Street, City, Country
-#     email: abais@auth.gr
-# author citation list in chicago format
 authorcitation: |
   Natsis, A.; Bais, A.; Meleti, C.
-# firstnote to eighthnote
-# firstnote: |
-#   Current address: Updated affiliation
-# secondnote: |
-#   These authors contributed equally to this work.
 correspondence: |
   abais@auth.gr
-# natsisphysicist@gmail.com;
-#   Tel.: +XX-000-00-0000.
 
@@
   This research received no external funding.
-
-# funding: |
-#   Please add: ``This research received no external funding'' or ``This research
-#   was funded by NAME OF FUNDER grant number XXX.'' and  and ``The APC was funded
-#   by XXX''. Check carefully that the details given are accurate and use the
-#   standard spelling of funding agency names at
-#   \url{https://search.crossref.org/funding}, any errors may affect your future
-#   funding.
 
@@
 bibliography: manualreferences.bib
-
 appendix:     appendix.tex
-
 endnotes:     false
-
 output:
@@
 
-# CLEAR_1_daily_DESEAS[!is.na(GLB_att), .(sum(GLB_att_N), .N)]
-
 ## TSI data span info
@@
 Similar findings have been shown from the analysis of observations at other locations [@Schwarz2020; @Ohvril2009; @Zerefos2009; @Xia2007] [and references therein].
-In the Mediterranean region aerosols have been recognized as an important factor affecting the penetration of solar radiation at the surface [@Fountoulakis2016; @Siomos2018; @Gkikas2013; @Lozano2021].
+In the Mediterranean region, aerosols have been recognized as an important factor affecting the penetration of solar radiation at the surface [@Fountoulakis2016; @Siomos2018; @Gkikas2013; @Lozano2021].
 These studies investigated the long-term trend in aerosol optical depth, which has been found to decrease in the last three decades, the transport and composition of aerosols, and their radiative effects.
@@
 
-...... cloud formation occurrences and constitution nuclei. cite .....
-
 #Wild2012 : simplify interaction nice graph
@@
 Here, we used data for the period from
-`r strftime(min(ALL_1_daily_DESEAS$Date), "%d %B %Y")`
+`r strftime(min(ALL_1_daily_DESEAS$Date), "%d\\ %B %Y")`
 to
-`r strftime(max(ALL_1_daily_DESEAS$Date), "%d %B %Y")`.
+`r strftime(max(ALL_1_daily_DESEAS$Date), "%d\\ %B %Y")`.
 The monitoring site was located
@@
 <!-- c) Clear sky identification -->
-In order to estimate the effect of the sky conditions on the long term variability of SDR, we created three datasets by characterizing each one-minute measurement with a corresponding sky-condition flag (i.e., all-sky, clear-sky, and cloudy-sky).
-To identify the clear cases we used the method proposed by
+In order to estimate the effect of the sky conditions on the long-term variability of SDR, we created three datasets by characterizing each one-minute measurement with a corresponding sky-condition flag (i.e., all-sky, clear-sky, and cloudy-sky).
+To identify the clear cases, we used the method proposed by
 @Reno2016, which requires the definition of some site specific parameters.
-These parameters were determined by an iterative process, as the original authors proposed and are discussed in the next section.
+These parameters were determined by an iterative process, as the original authors proposed, and are discussed in the next section.
 
@@
 The CSid algorithm was calibrated with the main focus to identify the presence of clouds. Despite the fine-tuning
-of the procedure, in a few marginal cases false positive or false
+of the procedure, in a few marginal cases, false positive or false
 negative results were identified by manual inspection. 
@@
 their small number, they did not affect the final results of the study.
-For completeness, we provide below a brief overview of the CSid algorithm, along with the site specific thresholds.
+For completeness, we provide below a brief overview of the CSid algorithm, along with the site-specific thresholds.
 
@@
 
-## The clear sky identification algorithm {#CDIDalgorithm}
-
-To calculate the reference clear-sky $\text{SDR}_\text{CSref}$ we used the $\text{SDR}_\text{Haurwitz}$ derived by the radiation model of @Haurwitz1945 (Eq.\ \ref{eq:hau}), adjusted for our site:
+## The Clear Sky Identification Algorithm {#CDIDalgorithm}
+
+To calculate the reference clear-sky $\text{SDR}_\text{CSref}$, we used the $\text{SDR}_\text{Haurwitz}$ derived by the radiation model of @Haurwitz1945 (Equation\ (\ref{eq:hau})), adjusted for our site:
 <!-- adjusted Haurwitz -->
@@
 
-The adjustment was made with a factor $a$ (Eq.\ \ref{eq:ahau}), which was estimated through an iterative optimization process, as described by @Long2000 and @Reno2016.
-The target of the optimization was the minimization of a function $f(a)$ (Eq.\ \ref{eq:minf}) and was accomplished with the algorithmic function "optimise", which is an implementation based on the work of @Brent1973, from the library "stats" of the R programming language [@RCT2023].
+The adjustment was made with a factor $a$ (Equation\ (\ref{eq:ahau})), which was estimated through an iterative optimization process, as described by @Long2000 and @Reno2016.
+The target of the optimization was the minimization of a function $f(a)$ (Equation\ (\ref{eq:minf})) and was accomplished with the algorithmic function 'optimise', which is an implementation based on the work of @Brent1973, from the library 'stats' of the R programming language [@RCT2023].
 <!-- cost function -->
@@
 \end{equation}
-where: $n$ is the total number of daylight data, $\text{SDR}_{\text{CSid},i}$ are the data identified as clear-sky by CSid, $a$ is a site-specific adjustment factor, and $\text{SDR}_{\text{testCSref},i}$ is the SDR derived by any of the
+where $n$ is the total number of daylight data, $\text{SDR}_{\text{CSid},i}$ are the data identified as clear-sky by CSid, $a$ is a site-specific adjustment factor, and $\text{SDR}_{\text{testCSref},i}$ is the SDR derived by any of the
 tested clear-sky radiation models.
 
-The optimization and the selection of the clear sky reference model, was performed on SDR observations for the period 2016 - 2021.
-During the optimization, eight simple clear sky radiation models were tested (namely, Daneshyar-Paltridge-Proctor, Kasten-Czeplak, Haurwitz, Berger-Duffie, Adnot-Bourges-Campana-Gicquel, Robledo-Soler, Kasten and Ineichen-Perez), with a wide range of factors.
+The optimization and the selection of the clear-sky reference model was performed on SDR observations for the period 2016--2021.
+During the optimization, eight simple clear-sky radiation models were tested (namely, Daneshyar--Paltridge--Proctor, Kasten--Czeplak, Haurwitz, Berger--Duffie, Adnot--Bourges--Campana--Gicquel, Robledo--Soler, Kasten, and Ineichen--Perez), with a wide range of factors.
 These models are described in more detail by @Reno2012 and are evaluated by @Reno2016.
-We found, that Haurwitz's model, adjusted with the factor $a = `r signif(alpha,3)`$ yields one of the lowest root mean squared errors (RMSE), while the procedure manages to characterize successfully the majority of the data.
-Thus, our clear sky reference is derived by the Eq.\ \ref{eq:ahau}:
+We found that Haurwitz's model, adjusted with the factor $a = `r signif(alpha,3)`$, yields one of the lowest root mean squared errors (RMSE), and the procedure manages to successfully characterize the majority of the data.
+Thus, our clear sky reference is derived by Equation\ (\ref{eq:ahau}):
 \begin{equation}
@@
 <!-- ### Variability in irradiance by line length. -->
-c) Length $L_i$ of the sequential line segments, connecting the points of the $`r nt`$ SDR values (Eq. \ref{eq:VILeq}).
+c) Length $L_i$ of the sequential line segments, connecting the points of the $`r nt`$ SDR values (Equation \ref{eq:VILeq}).
 \begin{equation}

### 2023-12-27 MDPI_submission.tex
Commit 2023-12-27 23:09

@@
 \citep{Schwarz2020, Ohvril2009, Zerefos2009, Xia2007} {[}and references
-therein{]}. In the Mediterranean region aerosols have been recognized as
-an important factor affecting the penetration of solar radiation at the
-surface \citep{Fountoulakis2016, Siomos2018, Gkikas2013, Lozano2021}.
-These studies investigated the long-term trend in aerosol optical depth,
-which has been found to decrease in the last three decades, the
-transport and composition of aerosols, and their radiative effects.
+therein{]}. In the Mediterranean region, aerosols have been recognized
+as an important factor affecting the penetration of solar radiation at
+the surface
+\citep{Fountoulakis2016, Siomos2018, Gkikas2013, Lozano2021}. These
+studies investigated the long-term trend in aerosol optical depth, which
+has been found to decrease in the last three decades, the transport and
+composition of aerosols, and their radiative effects.
 
@@
 \(22^\circ\,57'\,\)E, \(80\,\)m~a.s.l.). Here, we used data for the
-period from 13 April 1993 to 13 April 2023. The monitoring site was
+period from 13~April 1993 to 13~April 2023. The monitoring site was
 located near the city center, thus we expect that measurements were
@@
 
-In order to estimate the effect of the sky conditions on the long term
+In order to estimate the effect of the sky conditions on the long-term
 variability of SDR, we created three datasets by characterizing each
 one-minute measurement with a corresponding sky-condition flag (i.e.,
-all-sky, clear-sky, and cloudy-sky). To identify the clear cases we used
-the method proposed by \citet{Reno2016}, which requires the definition
-of some site specific parameters. These parameters were determined by an
-iterative process, as the original authors proposed and are discussed in
-the next section.
+all-sky, clear-sky, and cloudy-sky). To identify the clear cases, we
+used the method proposed by \citet{Reno2016}, which requires the
+definition of some site specific parameters. These parameters were
+determined by an iterative process, as the original authors proposed,
+and are discussed in the next section.
 
@@
 main focus to identify the presence of clouds. Despite the fine-tuning
-of the procedure, in a few marginal cases false positive or false
+of the procedure, in a few marginal cases, false positive or false
 negative results were identified by manual inspection. However, due to
@@
 For completeness, we provide below a brief overview of the CSid
-algorithm, along with the site specific thresholds.
+algorithm, along with the site-specific thresholds.
 
 \hypertarget{CDIDalgorithm}{%
-\subsection{The clear sky identification
-algorithm}\label{CDIDalgorithm}}
+\subsection{The Clear Sky Identification
+Algorithm}\label{CDIDalgorithm}}
 
-To calculate the reference clear-sky \(\text{SDR}_\text{CSref}\) we used
-the \(\text{SDR}_\text{Haurwitz}\) derived by the radiation model of
-\citet{Haurwitz1945} (Eq.~\ref{eq:hau}), adjusted for our site:
-\begin{equation}
+To calculate the reference clear-sky \(\text{SDR}_\text{CSref}\), we
+used the \(\text{SDR}_\text{Haurwitz}\) derived by the radiation model
+of \citet{Haurwitz1945} (Equation~(\ref{eq:hau})), adjusted for our
+site: \begin{equation}
 \text{SDR}_\text{Haurwitz} = 1098 \times \cos(\theta) \times \exp \left( \frac{ - 0.059}{\cos(\theta)} \right) \label{eq:hau}
@@
 
-The adjustment was made with a factor \(a\) (Eq.~\ref{eq:ahau}), which
-was estimated through an iterative optimization process, as described by
-\citet{Long2000} and \citet{Reno2016}. The target of the optimization
-was the minimization of a function \(f(a)\) (Eq.~\ref{eq:minf}) and was
-accomplished with the algorithmic function ``optimise'', which is an
-implementation based on the work of \citet{Brent1973}, from the library
-``stats'' of the R programming language \citep{RCT2023}.
-\begin{equation}
+The adjustment was made with a factor \(a\) (Equation~(\ref{eq:ahau})),
+which was estimated through an iterative optimization process, as
+described by \citet{Long2000} and \citet{Reno2016}. The target of the
+optimization was the minimization of a function \(f(a)\)
+(Equation~(\ref{eq:minf})) and was accomplished with the algorithmic
+function `optimise', which is an implementation based on the work of
+\citet{Brent1973}, from the library `stats' of the R programming
+language \citep{RCT2023}. \begin{equation}
 f(a) = \frac{1}{n}\sum_{i=1}^{n} ( \text{SDR}_{\text{CSid},i} - a \times \text{SDR}_{\text{testCSref},i} )^2 \label{eq:minf}
-\end{equation} where: \(n\) is the total number of daylight data,
+\end{equation} where \(n\) is the total number of daylight data,
 \(\text{SDR}_{\text{CSid},i}\) are the data identified as clear-sky by
@@
 
-The optimization and the selection of the clear sky reference model, was
-performed on SDR observations for the period 2016 - 2021. During the
-optimization, eight simple clear sky radiation models were tested
-(namely, Daneshyar-Paltridge-Proctor, Kasten-Czeplak, Haurwitz,
-Berger-Duffie, Adnot-Bourges-Campana-Gicquel, Robledo-Soler, Kasten and
-Ineichen-Perez), with a wide range of factors. These models are
-described in more detail by \citet{Reno2012} and are evaluated by
-\citet{Reno2016}. We found, that Haurwitz's model, adjusted with the
-factor \(a = 0.965\) yields one of the lowest root mean squared errors
-(RMSE), while the procedure manages to characterize successfully the
-majority of the data. Thus, our clear sky reference is derived by the
-Eq.~\ref{eq:ahau}: \begin{equation}
+The optimization and the selection of the clear-sky reference model was
+performed on SDR observations for the period 2016--2021. During the
+optimization, eight simple clear-sky radiation models were tested
+(namely, Daneshyar--Paltridge--Proctor, Kasten--Czeplak, Haurwitz,
+Berger--Duffie, Adnot--Bourges--Campana--Gicquel, Robledo--Soler,
+Kasten, and Ineichen--Perez), with a wide range of factors. These models
+are described in more detail by \citet{Reno2012} and are evaluated by
+\citet{Reno2016}. We found that Haurwitz's model, adjusted with the
+factor \(a = 0.965\), yields one of the lowest root mean squared errors
+(RMSE), and the procedure manages to successfully characterize the
+majority of the data. Thus, our clear sky reference is derived by
+Equation~(\ref{eq:ahau}): \begin{equation}
 \text{SDR}_\text{CSref} = a \times \text{SDR}_\text{Haurwitz} = 0.965 \times 1098 \times \cos(\theta) \times \exp \left( \frac{ - 0.057}{\cos(\theta)} \right) \label{eq:ahau}
@@
   Length \(L_i\) of the sequential line segments, connecting the points
-  of the \(11\) SDR values (Eq. \ref{eq:VILeq}). \begin{equation}
+  of the \(11\) SDR values (Equation \ref{eq:VILeq}). \begin{equation}
   L_i = \sum_{i=1}^{n-1}\sqrt{\left ( \text{SDR}_{i+1} - \text{SDR}_{i}\right )^2 + \left ( t_{i+1} - t_i \right )^2}

### 2023-12-27 MDPI_submission.Rmd
Commit 2023-12-28 00:14

@@
 The adjustment was made with a factor $a$ (Equation\ (\ref{eq:ahau})), which was estimated through an iterative optimization process, as described by @Long2000 and @Reno2016.
-The target of the optimization was the minimization of a function $f(a)$ (Equation\ (\ref{eq:minf})) and was accomplished with the algorithmic function 'optimise', which is an implementation based on the work of @Brent1973, from the library 'stats' of the R programming language [@RCT2023].
+The target of the optimization was the minimization of a function $f(a)$ (Equation\ \ref{eq:minf}) and was accomplished with the algorithmic function 'optimise', which is an implementation based on the work of @Brent1973, from the library 'stats' of the R programming language [@RCT2023].
 <!-- cost function -->
@@
 We found that Haurwitz's model, adjusted with the factor $a = `r signif(alpha,3)`$, yields one of the lowest root mean squared errors (RMSE), and the procedure manages to successfully characterize the majority of the data.
-Thus, our clear sky reference is derived by Equation\ (\ref{eq:ahau}):
+Thus, our clear sky reference is derived by Equation\ \ref{eq:ahau}:
 \begin{equation}

### 2023-12-27 MDPI_submission.tex
Commit 2023-12-28 00:14

@@
 
+\vspace{+6pt}
+
 \hypertarget{introduction}{%
@@
 investigate SDR at different sites in order to estimate the degree of
-variability, and its relation to the local conditions. In this study, we
-examine the trends of SDR, using ground-based measurements at
+variability and its relation to the local conditions. In this study, we
+examine the trends of SDR using ground-based measurements at
 Thessaloniki, Greece, for the period from 1993 to 2023. We re-evaluated
@@
 \citep{Reno2016, Reno2012}, and we derived the SDR trends for the period
-of study, under different sky conditions (all-sky, clear-sky, and
+of study under different sky conditions (all-sky, clear-sky, and
 cloudy-sky). Finally, we investigated the dependence of the trends on
@@
 
-A manual screening was performed, to remove inconsistent and erroneous
-recordings that can occur stochastically or systematically, during the
+A manual screening was performed to remove inconsistent and erroneous
+recordings that can occur stochastically or systematically during the
 continuous operation of the instruments. The manual screening was aided
@@
 Sun--Earth distance. Subsequently, we adjusted all measurements to the
-Total Solar Irradiance (TSI) at \(1\,\text{au}\), in order to compensate
-for the Sun's intensity variability, using a time series of satellite
-TSI observations. The TSI data we used are part of the `NOAA Climate
-Data Record of Total Solar Irradiance' dataset \citep{Coddington2005}.
-The initial daily values of this dataset were interpolated to match the
-time step of our measurements.
+total solar irradiance (TSI) at \(1\,\text{au}\) in order to compensate
+for the Sun's intensity variability using a time series of satellite TSI
+observations. The TSI data we used are part of the `NOAA Climate Data
+Record of Total Solar Irradiance' dataset \citep{Coddington2005}. The
+initial daily values of this dataset were interpolated to match the time
+step of our measurements.
 
@@
 Berger--Duffie, Adnot--Bourges--Campana--Gicquel, Robledo--Soler,
-Kasten, and Ineichen--Perez), with a wide range of factors. These models
+Kasten, and Ineichen--Perez) with a wide range of factors. These models
 are described in more detail by \citet{Reno2012} and are evaluated by
@@
 under clear-sky conditions are presented below. A data point is flagged
-as ``clear-sky'' if all criteria are satisfied; otherwise it is
-considered as ``cloud-sky''. Each criterion was applied for a running
+as `clear-sky' if all criteria are satisfied; otherwise, it is
+considered as `cloud-sky'. Each criterion was applied for a running
 window of \(11\) consecutive one-minute measurements, and the
@@
 parameter was calculated from the observations in comparison to the
-reference clear sky model. The allowable range of variation is defined
+reference clear-sky model. The allowable range of variation is defined
 by the model-derived value of the parameter multiplied by a factor plus
-an offset. The factors and the offsets were determined empirically, by
+an offset. The factors and the offsets were determined empirically by
 manually inspecting each filter's performance on selected days and
@@
 \begin{enumerate}
-\def\labelenumi{\alph{enumi})}
+\def\labelenumi{(\alph{enumi})}
 \tightlist
 \item
-  Mean of the measured \(\overline{\text{SDR}}_i\) (Eq.
-  \ref{eq:MeanVIP}). \begin{equation}
+  Mean of the measured \(\overline{\text{SDR}}_i\)
+  (Equation~(\ref{eq:MeanVIP})): \begin{equation}
   0.91 \times \overline{\text{SDR}}_{\text{CSref},i} - 20\,Wm^{-2}
@@
 \item
-  Maximum measured value \(M_{i}\) (Eq.~\ref{eq:MaxVIP}).
+  Maximum measured value \(M_{i}\) (Equation~(\ref{eq:MaxVIP})):
   \begin{equation}
@@
   Length \(L_i\) of the sequential line segments, connecting the points
-  of the \(11\) SDR values (Equation \ref{eq:VILeq}). \begin{equation}
+  of the \(11\) SDR values (Equation~(\ref{eq:VILeq})): \begin{equation}
   L_i = \sum_{i=1}^{n-1}\sqrt{\left ( \text{SDR}_{i+1} - \text{SDR}_{i}\right )^2 + \left ( t_{i+1} - t_i \right )^2}
@@
   \label{eq:VILcr}
-  \end{equation} where: \(t_i\) is the time stamp of each SDR
+  \end{equation} where \(t_i\) is the time stamp of each SDR
   measurement.
@@
   \(11\) sequential points, normalized by the mean
-  \(\overline{\text{SDR}}_i\) (Eq.~\ref{eq:VCT1}). \begin{gather}
+  \(\overline{\text{SDR}}_i\) (Equation~(\ref{eq:VCT1})): \begin{gather}
     \sigma_i = \frac {\sqrt{\frac{1}{n-1} \sum_{i=1}^{n-1} \left( s_i - \bar{s} \right)^2}} {\overline{\text{SDR}}_i} \label{eq:VCT1} \\
@@
   \end{gather} For this criterion, \(\sigma_i\) should be below a
-  certain threshold (Eq.~\ref{eq:VCTcr}): \begin{equation}
+  certain threshold (Equation~(\ref{eq:VCTcr})): \begin{equation}
     \sigma_i < \ensuremath{1.1\times 10^{-4}} \label{eq:VCTcr}
@@
   Maximum difference \(X_i\) between the change in measured irradiance
-  and the change in clear sky irradiance over each measurement interval.
+  and the change in clear sky irradiance over each measurement interval:
   \begin{gather}
@@
   \end{gather} For this criterion, \(X_i\) should be below a certain
-  threshold (Eq.~\ref{eq:VSMcr}): \begin{equation}
+  threshold (Equation~(\ref{eq:VSMcr})): \begin{equation}
     X_i < 7.5\,Wm^{-2} \label{eq:VSMcr}
@@
 \hypertarget{aggregationstatistical}{%
-\subsection{Aggregation of data and statistical
-approach}\label{aggregationstatistical}}
+\subsection{Aggregation of Data and Statistical
+Approach}\label{aggregationstatistical}}
 
-In order to investigate the SDR trends which are the main focus of the
+In order to investigate the SDR trends that are the main focus of the
 study, we implemented an aggregation scheme to the one-minute data to
-derive series in coarser time-scales. To preserve the representativeness
-of the data we used the following criteria: a) we excluded all days with
-less than 50\% of the expected daytime measurements, b) daily means for
-the clear-sky and cloudy-sky datasets were calculated only for days with
-more than 60\% of the expected daytime measurements identified as clear
-or cloudy respectively, c) monthly means were computed from daily means.
-For the all-skies dataset monthly means were computed only when at least
-20 days were available. Seasonal means were derived by averaging the
-monthly mean values in each season (winter: December - February, spring:
-March - May, etc.). The daily and monthly climatological means were
-derived by averaging the data for each day of year and calendar month,
-respectively. The daily and monthly datasets were deseasonalized by
-subtracting the corresponding climatological annual cycle (daily or
-monthly) from the actual data. Finally, to estimate the SZA effect on
-the SDR trends, the one-minute data were aggregated in \(1^{\circ}\) SZA
-bins, separately for the morning and afternoon hours.
+derive series in coarser time scales. To preserve the representativeness
+of the data, we used the following criteria: (a) we excluded all days
+with less than 50\% of the expected daytime measurements, (b) daily
+means for the clear-sky and cloudy-sky datasets were calculated only for
+days with more than 60\% of the expected daytime measurements identified
+as clear or cloudy, respectively, and (c) monthly means were computed
+from daily means. For the all-skies dataset, monthly means were computed
+only when at least 20 days were available. Seasonal means were derived
+by averaging the monthly mean values in each season (winter:
+December--February, spring: March--May, etc.). The daily and monthly
+climatological means were derived by averaging the data for each day of
+the year and calendar month, respectively. The daily and monthly
+datasets were deseasonalized by subtracting the corresponding
+climatological annual cycle (daily or monthly) from the actual data.
+Finally, to estimate the SZA effect on the SDR trends, the one-minute
+data were aggregated in \(1^{\circ}\) SZA bins, separately for the
+morning and afternoon hours.
 
-The linear trends were calculated using a first order autoregressive
-model with lag of 1 day using the ``maximum likelihood'' fitting method
-\citep{Gardner1980, Jones1980}, by implementing the function ``arima''
-from the library ``stats'' of the R programming language
-\citep{RCT2023}. The trends were reported together with the \(2\sigma\)
-errors.
+The linear trends were calculated using a first-order autoregressive
+model with lag of 1 day using the `maximum likelihood' fitting method
+\citep{Gardner1980, Jones1980} by implementing the function `arima' from
+the library `stats' of the R programming language \citep{RCT2023}. The
+trends were reported together with the \(2\sigma\) errors.
 
@@
 \hypertarget{long-term-sdr-trends}{%
-\subsection{Long-term SDR trends}\label{long-term-sdr-trends}}
+\subsection{Long-Term SDR Trends}\label{long-term-sdr-trends}}
 
-We calculated the linear trends of SDR, from the departures of the mean
+We calculated the linear trends of SDR from the departures of the mean
 daily values from the daily climatology and for the three sky
-conditions. These are presented in Table~\ref{tab:trendtable} which
-contains also the \(2\sigma\) standard error, the Pearson's correlation
-coefficient R and the trend in absolute units. In
-Figure~\ref{fig:trendALL} we present only the time series under all-sky
+conditions. These are presented in Table~\ref{tab:trendtable}, which
+also contains the \(2\sigma\) standard error, the Pearson's correlation
+coefficient R, and the trend in absolute units. In
+Figure~\ref{fig:trendALL}, we present only the time series under all-sky
 conditions; the plots for clear-sky and cloud-sky conditions, are shown
-in the Appendix (Figures~\ref{fig:trendCLEAR} and~
-\ref{fig:trendCLOUD}). In the studied period, there is no significant
-break or change in the variability pattern of the time series. The
-linear trends in all three datasets are positive and around
-\(0.4\,\%/y\) for all-sky and cloudy-sky conditions, while for
-clear-skies the trend is much smaller (\textasciitilde{}\(0.1\,\%/y\)).
-The linear trends were calculated taking into account the
-autocorrelation of the time series and all three are statistically
-significant at least at the \(95\,\%\) confidence level, as they are
-larger than the corresponding \(2\sigma\) errors, despite the small
-values of R, which is due to the large variability of the daily values.
-The clear-sky trend is very small suggesting a small effect from
-aerosols and water vapor which are the dominant factors of the SDR
-variability \citep{Fountoulakis2016, Siomos2018, Yu2022}. In contrast,
-the large positive trend of SDR under cloudy skies can be attributed to
-reduction in cloud cover and/or cloud optical depth. Lack of continuous
+in Appendix (Figures~\ref{fig:trendCLEAR} and~ \ref{fig:trendCLOUD}). In
+the studied period, there is no significant break or change in the
+variability pattern of the time series. The linear trends in all three
+datasets are positive and around \(0.4\,\%/y\) for all-sky and
+cloudy-sky conditions, whereas for clear-skies the trend is much smaller
+(\textasciitilde{}\(0.1\,\%/y\)). The linear trends were calculated
+taking into account the autocorrelation of the time series, and all
+three are statistically significant at least at the \(95\,\%\)
+confidence level, as they are larger than the corresponding \(2\sigma\)
+errors, despite the small values of R, which is due to the large
+variability of the daily values. The clear-sky trend is very small,
+suggesting a small effect from aerosols and water vapor, which are the
+dominant factors of the SDR variability
+\citep{Fountoulakis2016, Siomos2018, Yu2022}. In contrast, the large
+positive trend of SDR under cloudy skies can be attributed to reduction
+in cloud cover and/or cloud optical depth. Lack of continuous
 observations of cloud optical thickness that could support these
 findings does not allow drawing firm conclusions. However, there are
-indications that the total cloud-cover as inferred from the ERA5
+indications that the total cloud cover as inferred from the ERA5
 analysis for the grid point of Thessaloniki is decreasing over the
@@
 The all-sky trend is similar to the one reported in \citet{Bais2013}
-from a ten-year shorter dataset suggesting that the tendency of SDR in
+from a ten-year shorter dataset, suggesting that the tendency of SDR in
 Thessaloniki is systematic. Other studies for the European region
@@
 the three datasets are in accordance with the widely accepted solar
-radiation brightening over Europe. For the period of our observations
+radiation brightening over Europe. For the period of our observations,
 the trend in the TSI is negligible (\(-0.00022\,\%/y\)), and thus we

### 2023-12-27 MDPI_submission.Rmd
Commit 2023-12-28 01:10

--- 2023-12-27 MDPI_submission.Rmd
@@
 
+\vspace{+6pt}
+
 # Introduction
@@
 
-Due to the significant spatial and temporal variability of the trends and the contributing factors, there is a constant need to monitor and investigate SDR at different sites in order to estimate the degree of variability, and its relation to the local conditions.
-In this study, we examine the trends of SDR, using ground-based measurements at Thessaloniki, Greece, for the period from `r year(min(ALL_1_daily_DESEAS$Date))` to `r year(max(ALL_1_daily_DESEAS$Date))`.
+Due to the significant spatial and temporal variability of the trends and the contributing factors, there is a constant need to monitor and investigate SDR at different sites in order to estimate the degree of variability and its relation to the local conditions.
+In this study, we examine the trends of SDR using ground-based measurements at Thessaloniki, Greece, for the period from `r year(min(ALL_1_daily_DESEAS$Date))` to `r year(max(ALL_1_daily_DESEAS$Date))`.
 <!-- goals -->
-We re-evaluated and extended the dataset used by @Bais2013, we applied a different algorithm for the identification of clear-/cloud-sky instances [@Reno2016; @Reno2012], and we derived the SDR trends for the period of study, under different sky conditions (all-sky, clear-sky, and cloudy-sky).
+We re-evaluated and extended the dataset used by @Bais2013, we applied a different algorithm for the identification of clear-/cloud-sky instances [@Reno2016; @Reno2012], and we derived the SDR trends for the period of study under different sky conditions (all-sky, clear-sky, and cloudy-sky).
 Finally, we investigated the dependence of the trends on solar zenith angle and season.
@@
 <!-- b) Radiation data quality check -->
-A manual screening was performed, to remove inconsistent and erroneous recordings that can occur stochastically or systematically, during the continuous operation of the instruments.
+A manual screening was performed to remove inconsistent and erroneous recordings that can occur stochastically or systematically during the continuous operation of the instruments.
 The manual screening was aided by a radiation data quality assurance procedure, adjusted for the site, which was based on the methods of
@@
 <!-- TSI -->
-Subsequently, we adjusted all measurements to the Total Solar Irradiance (TSI) at $1\,\text{au}$, in order to compensate for the Sun's intensity variability, using a time series of satellite TSI observations.
+Subsequently, we adjusted all measurements to the total solar irradiance (TSI) at $1\,\text{au}$ in order to compensate for the Sun's intensity variability using a time series of satellite TSI observations.
 The TSI data we used are part of the 'NOAA Climate Data Record of Total Solar Irradiance' dataset [@Coddington2005].
@@
 The optimization and the selection of the clear-sky reference model was performed on SDR observations for the period 2016--2021.
-During the optimization, eight simple clear-sky radiation models were tested (namely, Daneshyar--Paltridge--Proctor, Kasten--Czeplak, Haurwitz, Berger--Duffie, Adnot--Bourges--Campana--Gicquel, Robledo--Soler, Kasten, and Ineichen--Perez), with a wide range of factors.
+During the optimization, eight simple clear-sky radiation models were tested (namely, Daneshyar--Paltridge--Proctor, Kasten--Czeplak, Haurwitz, Berger--Duffie, Adnot--Bourges--Campana--Gicquel, Robledo--Soler, Kasten, and Ineichen--Perez) with a wide range of factors.
 These models are described in more detail by @Reno2012 and are evaluated by @Reno2016.
@@
 The criteria that were used to identify whether a measurement was taken under clear-sky conditions are presented below.
-A data point is flagged as "clear-sky" if all criteria are satisfied; otherwise it is considered as "cloud-sky".
+A data point is flagged as 'clear-sky' if all criteria are satisfied; otherwise, it is considered as 'cloud-sky'.
 Each criterion was applied for a running window of $`r nt`$ consecutive one-minute measurements, and the characterization was assigned to the central datum of the window.
-Each parameter was calculated from the observations in comparison to the reference clear sky model.
+Each parameter was calculated from the observations in comparison to the reference clear-sky model.
 The allowable range of variation is defined by the model-derived value of the parameter multiplied by a factor plus an offset.
-The factors and the offsets were determined empirically, by manually inspecting each filter's performance on selected days and adjusting them accordingly during an iterative process.
+The factors and the offsets were determined empirically by manually inspecting each filter's performance on selected days and adjusting them accordingly during an iterative process.
 The criteria are
@@
 respective parameter should fall in order to raise the clear-sky flag:
+
 <!-- 1. MeanVIP -->
 <!-- ### Mean value of irradiance during the time period. -->
-<!--a) Mean of the measured $\overline{\text{SDR}}_i$ = \text{mean}_{n=i-5}^{n=i+5}[\text{SDR}_{n}]$ and running mean of the reference $\overline{\text{SDR}}_{\text{CSref},i}$ = \text{mean}_{n=i-5}^{n=i+5}[\text{SDR}_{\text{CSref},n}]$ (Eq.
-\ref{eq:MeanVIP}).-->
-
-a) Mean of the measured $\overline{\text{SDR}}_i$ (Eq. \ref{eq:MeanVIP}).
+(a) Mean of the measured $\overline{\text{SDR}}_i$ (Equation\ (\ref{eq:MeanVIP})):
 \begin{equation}
@@
 <!-- ### Max value of irradiance during the time period. -->
-<!-- The running max measured value $M_{\text{SDR},i} = \max_{n=i-5}^{n=i+5}[\text{SDR}_{n}]$, is compared to a similar constructed value from the reference $M_{CSrefi} = \max_{n=i-5}^{n=i+5}[\text{SDR}_{\text{CSref},n}]$ (Eq.\ \ref{eq:MaxVIP}).
--->
-b) Maximum measured value $M_{i}$ (Eq.\ \ref{eq:MaxVIP}).
+(b) Maximum measured value $M_{i}$ (Equation\ (\ref{eq:MaxVIP})):
 \begin{equation}
@@
 <!-- ### Variability in irradiance by line length. -->
-c) Length $L_i$ of the sequential line segments, connecting the points of the $`r nt`$ SDR values (Equation \ref{eq:VILeq}).
+(c) Length $L_i$ of the sequential line segments, connecting the points of the $`r nt`$ SDR values (Equation\ (\ref{eq:VILeq})):
 \begin{equation}
@@
 \end{equation}
-where: $t_i$ is the time stamp of each SDR measurement.
+where $t_i$ is the time stamp of each SDR measurement.
 <!-- 4. VCT -->
@@
 <!-- Standard deviation of rate of change in irradiance. -->
-d) Standard deviation $\sigma_i$ of the slope ($s_i$) between the $`r nt`$ sequential points, normalized by the mean $\overline{\text{SDR}}_i$ (Eq.\ \ref{eq:VCT1}).
+(d) Standard deviation $\sigma_i$ of the slope ($s_i$) between the $`r nt`$ sequential points, normalized by the mean $\overline{\text{SDR}}_i$ (Equation\ (\ref{eq:VCT1})):
 <!-- \begin{gather}
@@
 \end{gather}
-For this criterion, $\sigma_i$ should be below a certain threshold (Eq.\ \ref{eq:VCTcr}):
+For this criterion, $\sigma_i$ should be below a certain threshold (Equation\ (\ref{eq:VCTcr})):
 \begin{equation}
@@
 <!-- ### Variability in the Shape of the irradiance Measurements. -->
-e) Maximum difference $X_i$ between the change in measured irradiance and the change in clear sky irradiance over each measurement interval.
+(e) Maximum difference $X_i$ between the change in measured irradiance and the change in clear sky irradiance over each measurement interval:
 \begin{gather}
@@
 \end{gather}
-For this criterion, $X_i$ should be below a certain threshold (Eq.\ \ref{eq:VSMcr}):
+For this criterion, $X_i$ should be below a certain threshold (Equation\ (\ref{eq:VSMcr})):
 \begin{equation}
@@
 
-## Aggregation of data and statistical approach {#aggregationstatistical}
-
-In order to investigate the SDR trends which are the main focus of the
+## Aggregation of Data and Statistical Approach {#aggregationstatistical}
+
+In order to investigate the SDR trends that are the main focus of the
 study, we implemented an aggregation scheme to the
-one-minute data to derive series in coarser time-scales.
-To preserve the representativeness of the data we used the following criteria:
-<!-- a) for daily mean values we exclude days with less than `r Daily_aggregation_N_lim` valid data points, -->
-a) we excluded all days with less than `r 100 * All_daily_ratio_lim`% of the expected daytime measurements,
-b) daily means for the clear-sky and cloudy-sky datasets
+one-minute data to derive series in coarser time scales.
+To preserve the representativeness of the data, we used the following criteria:
+(a) we excluded all days with less than `r 100 * All_daily_ratio_lim`% of the expected daytime measurements,
+(b) daily means for the clear-sky and cloudy-sky datasets
 were calculated only for days with more than
 `r 100 * Clear_daily_ratio_lim`% of the expected daytime
-measurements identified as clear or cloudy respectively,
-c) monthly means were computed from daily means. For the all-skies dataset monthly means were computed only when at least `r Monthly_aggegation_N_lim` 
+measurements identified as clear or cloudy, respectively, and
+(c) monthly means were computed from daily means. For the all-skies dataset, monthly means were computed only when at least `r Monthly_aggegation_N_lim` 
 days were available.
 Seasonal means were derived by averaging the
-monthly mean values in each season (winter: December - February, spring:
-March - May, etc.).
+monthly mean values in each season (winter: December--February, spring:
+March--May, etc.).
 The daily and monthly climatological means were
-derived by averaging the data for each day of year and calendar month,
+derived by averaging the data for each day of the year and calendar month,
 respectively. The daily and monthly datasets were deseasonalized by
@@
 
-The linear trends were calculated using a first order autoregressive model
- with lag of 1 day using the "maximum likelihood" fitting method [@Gardner1980; @Jones1980], by implementing the function "arima" from the library "stats" of the R programming language [@RCT2023].
+The linear trends were calculated using a first-order autoregressive model
+ with lag of 1 day using the 'maximum likelihood' fitting method [@Gardner1980; @Jones1980] by implementing the function 'arima' from the library 'stats' of the R programming language [@RCT2023].
 The trends were reported together with the $2\sigma$ errors.
 
-<!-- [For the trends derived from monthly mean data we used an effective -->
-<!-- number of observations to calculate the standard error (Sander et al., -->
-<!-- 2000) in order to assess the significance of the trends.]{.mark} -->
-<!-- @Santer2000 -->
-
 # Results
 
-## Long-term SDR trends
+## Long-Term SDR Trends
 
@@
 
-We calculated the linear trends of SDR, from the departures of the mean daily values from the daily climatology and for the three sky conditions.
-These are presented in Table\ \ref{tab:trendtable} which contains also the $2\sigma$ standard error, the Pearson's correlation coefficient R and the trend in absolute units.
-In Figure\ \ref{fig:trendALL} we present only the time series under all-sky conditions; the plots for clear-sky and cloud-sky conditions, are shown in the Appendix (Figures\ \ref{fig:trendCLEAR} and\  \ref{fig:trendCLOUD}).
+We calculated the linear trends of SDR from the departures of the mean daily values from the daily climatology and for the three sky conditions.
+These are presented in Table\ \ref{tab:trendtable}, which also contains the $2\sigma$ standard error, the Pearson's correlation coefficient R, and the trend in absolute units.
+In Figure\ \ref{fig:trendALL}, we present only the time series under all-sky conditions; the plots for clear-sky and cloud-sky conditions, are shown in Appendix (Figures\ \ref{fig:trendCLEAR} and\  \ref{fig:trendCLOUD}).
 In the studied period, there is no
@@
 The linear trends in all three datasets are positive and around
-$0.4\,\%/y$ for all-sky and cloudy-sky conditions, while for clear-skies the
+$0.4\,\%/y$ for all-sky and cloudy-sky conditions, whereas for clear-skies the
 trend is much smaller (\~$`r signif(cleartrend,1)`\,\%/y$). 
 The linear trends were calculated
-taking into account the autocorrelation of the time series and all three
+taking into account the autocorrelation of the time series, and all three
 are statistically significant at least at the $95\,\%$ confidence level, as
@@
 The clear-sky trend is
-very small suggesting a small effect from aerosols and water vapor which
+very small, suggesting a small effect from aerosols and water vapor, which
 are the dominant factors of the SDR variability [@Fountoulakis2016; @Siomos2018; @Yu2022].
@@
 firm conclusions. However, there are indications that the total
-cloud-cover as inferred from the ERA5 analysis for the grid point of
+cloud cover as inferred from the ERA5 analysis for the grid point of
 Thessaloniki is decreasing over the period of study.
@@
 The all-sky trend is similar to the one reported in @Bais2013
-from a ten-year shorter dataset suggesting that the tendency of SDR in
+from a ten-year shorter dataset, suggesting that the tendency of SDR in
 Thessaloniki is systematic.
@@
 For the period
-of our observations the trend in the TSI is negligible
+of our observations, the trend in the TSI is negligible
 ($`r format(TSIslopeAll[1,1], scientific = F, digits = 2)`\,\%/y$),
@@
                         drop0trailing = T,
-                        caption     = paste0("Trends in SDR daily means for different sky conditions for the period ", year(min(ALL_1_daily_DESEAS$Date)), " - ", year(max(ALL_1_daily_DESEAS$Date)), "."))
+                        caption     = paste0("Trends in SDR daily means for different sky conditions for the period ", year(min(ALL_1_daily_DESEAS$Date)), "--", year(max(ALL_1_daily_DESEAS$Date)), "."))
 
@@
 
-Although the year-to-year variability of the anomalies (Figure \ref{fig:trendALL} and Figures\ \ref{fig:trendCLEAR}, \ref{fig:trendCLOUD} in Appendix), shows a rather homogeneous behavior, plots of the cumulative sums (CUSUM) [@Regier2019] of the anomalies can reveal different structures in the records of all three sky conditions.
+Although the year-to-year variability of the anomalies (Figure \ref{fig:trendALL} and Figures\ \ref{fig:trendCLEAR} and \ref{fig:trendCLOUD} in Appendix), shows a rather homogeneous behavior, plots of the cumulative sums (CUSUM) [@Regier2019] of the anomalies can reveal different structures in the records of all three sky conditions.
 For time series with a uniform trend, we would expect
@@
 evolution of the anomalies.
-For all-skies (Figure\ \ref{fig:cusummonth-1}),
+For all skies (Figure\ \ref{fig:cusummonth-1}),
 we observe three
 rather distinct periods: a) a downward part between the start of the
-datasets and about 2000, denoting that all anomalies are negative thus
+datasets and about 2000, denoting that all anomalies are negative, thus
 below the climatology; b) a relatively steady part lasting for almost 20
 years suggesting little variability in SDR anomalies; and c) a steep
-upward part to present indicating anomalies above the climatology. The
+upward part to the present indicating anomalies above the climatology. The
 CUSUMs for cloudy-skies (Figure\ \ref{fig:cusummonth-3}), show a similar behavior with some short-term

### 2023-12-27 MDPI_submission.tex
Commit 2023-12-28 01:10

@@
 
-\caption{\label{tab:trendtable}Trends in SDR daily means for different sky conditions for the period 1993 - 2023.}
+\caption{\label{tab:trendtable}Trends in SDR daily means for different sky conditions for the period 1993--2023.}
 \begin{tabu} to \linewidth {>{\centering\arraybackslash}p{8em}>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X}
@@
 Although the year-to-year variability of the anomalies (Figure
-\ref{fig:trendALL} and Figures~\ref{fig:trendCLEAR},
+\ref{fig:trendALL} and Figures~\ref{fig:trendCLEAR} and
 \ref{fig:trendCLOUD} in Appendix), shows a rather homogeneous behavior,
@@
 climatological mean. In our case, there is a more complex evolution of
-the anomalies. For all-skies (Figure~\ref{fig:cusummonth-1}), we observe
+the anomalies. For all skies (Figure~\ref{fig:cusummonth-1}), we observe
 three rather distinct periods: a) a downward part between the start of
-the datasets and about 2000, denoting that all anomalies are negative
+the datasets and about 2000, denoting that all anomalies are negative,
 thus below the climatology; b) a relatively steady part lasting for
 almost 20 years suggesting little variability in SDR anomalies; and c) a
-steep upward part to present indicating anomalies above the climatology.
-The CUSUMs for cloudy-skies (Figure~\ref{fig:cusummonth-3}), show a
-similar behavior with some short-term differences that do not change the
-overall pattern. For clear skies (Figure~\ref{fig:cusummonth-2}), a
-monotonic downward tendency is evident until 2004, suggesting that the
-anomalies are all negative. After 2004 the anomalies turn to positive at
-a fast rate for about five years and at a slower rate thereafter.
+steep upward part to the present indicating anomalies above the
+climatology. The CUSUMs for cloudy-skies
+(Figure~\ref{fig:cusummonth-3}), show a similar behavior with some
+short-term differences that do not change the overall pattern. For clear
+skies (Figure~\ref{fig:cusummonth-2}), a monotonic downward tendency is
+evident until 2004, suggesting that the anomalies are all negative.
+After 2004 the anomalies turn to positive at a fast rate for about five
+years and at a slower rate thereafter.
 

### 2023-12-28 MDPI_submission.Rmd
Commit 2023-12-28 19:09

@@
 
-Although the year-to-year variability of the anomalies (Figure \ref{fig:trendALL} and Figures\ \ref{fig:trendCLEAR} and \ref{fig:trendCLOUD} in Appendix), shows a rather homogeneous behavior, plots of the cumulative sums (CUSUM) [@Regier2019] of the anomalies can reveal different structures in the records of all three sky conditions.
+Although the year-to-year variability of the anomalies (Figure \ref{fig:trendALL} and Figures\ \ref{fig:trendCLEAR} and \ref{fig:trendCLOUD} in Appendix) shows a rather homogeneous behavior, plots of the cumulative sums (CUSUM) [@Regier2019] of the anomalies can reveal different structures in the records of all three sky conditions.
 For time series with a uniform trend, we would expect
@@
 around the middle of the series. This would indicate that the anomalies
-are evenly distributed around the climatological mean, and for a
+are evenly distributed around the climatological mean, and, for a
 positive uniform trend, the first half is below and the second half
@@
 we observe three
-rather distinct periods: a) a downward part between the start of the
+rather distinct periods: (a) a downward part between the start of the
 datasets and about 2000, denoting that all anomalies are negative, thus
-below the climatology; b) a relatively steady part lasting for almost 20
-years suggesting little variability in SDR anomalies; and c) a steep
-upward part to the present indicating anomalies above the climatology. The
-CUSUMs for cloudy-skies (Figure\ \ref{fig:cusummonth-3}), show a similar behavior with some short-term
+below the climatology; (b) a relatively steady part lasting for almost 20
+years suggesting little variability in SDR anomalies; and (c) a steep
+upward part to the present, indicating anomalies above the climatology. The
+CUSUMs for cloudy skies (Figure\ \ref{fig:cusummonth-3}) show a similar behavior with some short-term
 differences that do not change the overall pattern.
@@
 (Figure\ \ref{fig:cusummonth-2}), a monotonic downward tendency is evident until 2004,
-suggesting that the anomalies are all negative. After 2004 the anomalies
-turn to positive at a fast rate for about five years and at a slower
+suggesting that the anomalies are all negative. After 2004, the anomalies
+turn positive at a fast rate for about five years and at a slower
 rate thereafter.
@@
 in 2004 is now clearer. The most likely reason for this change is the
-monotonic reduction of aerosols in Thessaloniki. At that year there a
-change in the rate of decrease in aerosol optical depth as illustrated
-in Figure 7 of @Siomos2020. This abrupt change in CUSUMs lasts
+monotonic reduction of aerosols in Thessaloniki. In that year, there was a
+change in the rate of decrease in aerosol optical depth, as illustrated
+in Figure 7 of @Siomos2020.
+This abrupt change in CUSUMs lasted
 until about 2010 when the anomalies become again variable.
@@
 
-## Effects of the solar zenith angle on SDR
+## Effects of the Solar Zenith Angle on SDR
 

### 2023-12-28 MDPI_submission.tex
Commit 2023-12-28 19:09

@@
 \ref{fig:trendALL} and Figures~\ref{fig:trendCLEAR} and
-\ref{fig:trendCLOUD} in Appendix), shows a rather homogeneous behavior,
+\ref{fig:trendCLOUD} in Appendix) shows a rather homogeneous behavior,
 plots of the cumulative sums (CUSUM) \citep{Regier2019} of the anomalies
@@
 the middle of the series. This would indicate that the anomalies are
-evenly distributed around the climatological mean, and for a positive
+evenly distributed around the climatological mean, and, for a positive
 uniform trend, the first half is below and the second half above the
@@
 the anomalies. For all skies (Figure~\ref{fig:cusummonth-1}), we observe
-three rather distinct periods: a) a downward part between the start of
+three rather distinct periods: (a) a downward part between the start of
 the datasets and about 2000, denoting that all anomalies are negative,
-thus below the climatology; b) a relatively steady part lasting for
-almost 20 years suggesting little variability in SDR anomalies; and c) a
-steep upward part to the present indicating anomalies above the
-climatology. The CUSUMs for cloudy-skies
-(Figure~\ref{fig:cusummonth-3}), show a similar behavior with some
-short-term differences that do not change the overall pattern. For clear
-skies (Figure~\ref{fig:cusummonth-2}), a monotonic downward tendency is
+thus below the climatology; (b) a relatively steady part lasting for
+almost 20 years suggesting little variability in SDR anomalies; and (c)
+a steep upward part to the present, indicating anomalies above the
+climatology. The CUSUMs for cloudy skies (Figure~\ref{fig:cusummonth-3})
+show a similar behavior with some short-term differences that do not
+change the overall pattern. For clear skies
+(Figure~\ref{fig:cusummonth-2}), a monotonic downward tendency is
 evident until 2004, suggesting that the anomalies are all negative.
-After 2004 the anomalies turn to positive at a fast rate for about five
+After 2004, the anomalies turn positive at a fast rate for about five
 years and at a slower rate thereafter.
@@
 in 2004 is now clearer. The most likely reason for this change is the
-monotonic reduction of aerosols in Thessaloniki. At that year there a
-change in the rate of decrease in aerosol optical depth as illustrated
-in Figure 7 of \citet{Siomos2020}. This abrupt change in CUSUMs lasts
-until about 2010 when the anomalies become again variable.
+monotonic reduction of aerosols in Thessaloniki. In that year, there was
+a change in the rate of decrease in aerosol optical depth, as
+illustrated in Figure 7 of \citet{Siomos2020}. This abrupt change in
+CUSUMs lasted until about 2010 when the anomalies become again variable.
 

### 2023-12-28 MDPI_submission.Rmd
Commit 2023-12-28 20:14

@@
 each lasting for several years. These fluctuations are probably within
-the natural variability and no distinct changes are identified. The
+the natural variability, and no distinct changes are identified. The
 pattern in both datasets is similar, suggesting prevalence in cloudy
@@
 This abrupt change in CUSUMs lasted
-until about 2010 when the anomalies become again variable.
+until about 2010, when the anomalies become again variable.
 
@@
 
-Under clear skies (Figures\ \ref{fig:szatrends-2}), the trends are smaller and less variable,
+Under clear skies (Figure\ \ref{fig:szatrends-2}), the trends are smaller and less variable,
 ranging between $0.1$ and $0.15\,\%/y$ up to $77^\circ$ SZA.
 At higher SZAs and in
-the afternoon hours there a sharp increase in the trend up to $0.3\,\%/y$,
+the afternoon hours, there is a sharp increase in the trend up to $0.3\,\%/y$,
 which may have been caused by the long path length of radiation through
@@
 $60^\circ$ SZA is likely a result of less attenuation of SDR in the morning
-hours due to lesser amounts of aerosols and shallower boundary layer.
+hours due to lesser amounts of aerosols and a shallower boundary layer.
 

### 2023-12-28 MDPI_submission.tex
Commit 2023-12-28 20:14

@@
 diverging from the climatological values, each lasting for several
-years. These fluctuations are probably within the natural variability
+years. These fluctuations are probably within the natural variability,
 and no distinct changes are identified. The pattern in both datasets is
@@
 illustrated in Figure 7 of \citet{Siomos2020}. This abrupt change in
-CUSUMs lasted until about 2010 when the anomalies become again variable.
+CUSUMs lasted until about 2010, when the anomalies become again
+variable.
 
@@
 \hypertarget{effects-of-the-solar-zenith-angle-on-sdr}{%
-\subsection{Effects of the solar zenith angle on
+\subsection{Effects of the Solar Zenith Angle on
 SDR}\label{effects-of-the-solar-zenith-angle-on-sdr}}
@@
 
-Under clear skies (Figures~\ref{fig:szatrends-2}), the trends are
-smaller and less variable, ranging between \(0.1\) and \(0.15\,\%/y\) up
-to \(77^\circ\) SZA. At higher SZAs and in the afternoon hours there a
+Under clear skies (Figure~\ref{fig:szatrends-2}), the trends are smaller
+and less variable, ranging between \(0.1\) and \(0.15\,\%/y\) up to
+\(77^\circ\) SZA. At higher SZAs and in the afternoon hours, there is a
 sharp increase in the trend up to \(0.3\,\%/y\), which may have been
@@
 \(60^\circ\) SZA is likely a result of less attenuation of SDR in the
-morning hours due to lesser amounts of aerosols and shallower boundary
+morning hours due to lesser amounts of aerosols and a shallower boundary
 layer.

### 2024-01-01 MDPI_submission.Rmd
Commit 2024-01-01 18:12

@@
       )]
-# LT1 <- LT
-
-# signif(LT1$`Trend [%/year]` , 3)
-# round(LT1$`Trend [%/year]` , 3)
-# format(LT1$`Trend [%/year]` , 3, drop0trailing = T)
-# as.character(format(LT1$`Trend [%/year]` , 3))
-
-## order columns
-# LT2 <- LT[,c(
-#       "Sky conditions",
-#       "Trend [%/year]",
-#       "Statistical signif. [%]",
-#       "Days with data"
-#       )]
-LT2 <- LT
 ```
@@
 These are presented in Table\ \ref{tab:trendtable}, which also contains the $2\sigma$ standard error, the Pearson's correlation coefficient R, and the trend in absolute units.
-In Figure\ \ref{fig:trendALL}, we present only the time series under all-sky conditions; the plots for clear-sky and cloud-sky conditions, are shown in Appendix (Figures\ \ref{fig:trendCLEAR} and\  \ref{fig:trendCLOUD}).
+In Figure\ \ref{fig:trendALL}, we present only the time series under all-sky conditions; the plots for clear-sky and cloud-sky conditions, are shown in Appendix \ref{appendix} (Figures\ \ref{fig:trendCLEAR} and\  \ref{fig:trendCLOUD}).
 In the studied period, there is no
@@
                         escape      = FALSE,
-                        col.names   = c("Sky conditions", "Trend [\\%/year]", "Trend S.E. ($2\\sigma$)", "Pearson correl.", "Trend [W/m\\textsuperscript{2}/year]", "Days"),
+                        col.names   = c("Sky Conditions", "Trend [\\%/year]", "Trend S.E. ($2\\sigma$)", "Pearson Correl.", "Trend [W/m\\textsuperscript{2}/year]", "Days"),
                         format.args = list(digits = 3),
@@
 
-# x_latex %>%
-#     kable_styling(
-#         latex_options = c("hold_position")
-#     )
 x_latex                              %>%
+    row_spec(0,bold=TRUE)            %>%
     kable_styling(full_width = TRUE) %>%
@@
 
-```{r trendALL, echo=F, out.width='.75\\linewidth', fig.cap=paste0("Anomalies (\\%) of the daily all-sky SDR from the climatological mean for the period ", year(min(ALL_1_daily_DESEAS$Date)), " -- ", year(max(ALL_1_daily_DESEAS$Date)), ".
-The black line is the long term linear trend.")}
+```{r trendALL, echo=F, out.width='.75\\linewidth', fig.cap=paste0("Anomalies (\\%) of the daily all-sky SDR from the climatological mean for the period ", year(min(ALL_1_daily_DESEAS$Date)), "--", year(max(ALL_1_daily_DESEAS$Date)), ".
+The black line is the long-term linear trend.")}
 knitr::include_graphics("./images/LongtermTrends-1.png")
@@
 
-Although the year-to-year variability of the anomalies (Figure \ref{fig:trendALL} and Figures\ \ref{fig:trendCLEAR} and \ref{fig:trendCLOUD} in Appendix) shows a rather homogeneous behavior, plots of the cumulative sums (CUSUM) [@Regier2019] of the anomalies can reveal different structures in the records of all three sky conditions.
+Although the year-to-year variability of the anomalies (Figures\ \ref{fig:trendALL}, \ref{fig:trendCLEAR} and \ref{fig:trendCLOUD} in Appendix \ref{appendix}) shows a rather homogeneous behavior, plots of the cumulative sums (CUSUM) [@Regier2019] of the anomalies can reveal different structures in the records of all three sky conditions.
 For time series with a uniform trend, we would expect
@@
 evolution of the anomalies.
-For all skies (Figure\ \ref{fig:cusummonth-1}),
+For all skies (Figure\ \ref{fig:cusummonth}a),
 we observe three
@@
 below the climatology; (b) a relatively steady part lasting for almost 20
-years suggesting little variability in SDR anomalies; and (c) a steep
+years, suggesting little variability in SDR anomalies; and (c) a steep
 upward part to the present, indicating anomalies above the climatology. The
-CUSUMs for cloudy skies (Figure\ \ref{fig:cusummonth-3}) show a similar behavior with some short-term
+CUSUMs for cloudy skies (Figure\ \ref{fig:cusummonth}c) show a similar behavior with some short-term
 differences that do not change the overall pattern.
 For clear skies
-(Figure\ \ref{fig:cusummonth-2}), a monotonic downward tendency is evident until 2004,
+(Figure\ \ref{fig:cusummonth}b), a monotonic downward tendency is evident until 2004,
 suggesting that the anomalies are all negative. After 2004, the anomalies
@@
         {\centering 
-        \subfloat[All skies.\label{fig:cusummonth-1}]
             {\includegraphics[width=.32\linewidth]{./images/CumulativeMonthlyCuSum-1}}\hfill
-        \subfloat[Clear skies.\label{fig:cusummonth-2}]
             {\includegraphics[width=.32\linewidth]{./images/CumulativeMonthlyCuSum-5}}\hfill
-        \subfloat[Cloudy skies.\label{fig:cusummonth-3}]
-            {\includegraphics[width=.32\linewidth]{./images/CumulativeMonthlyCuSum-9}}\hfill
+            {\includegraphics[width=.32\linewidth]{./images/CumulativeMonthlyCuSum-9}}
         }
-\caption{Cumulative sum plots of the monthly SDR anomalies in (\%) for different sky conditions.}\label{fig:cusummonth}
     \end{adjustwidth}
+\caption{Cumulative sum plots of the monthly SDR anomalies in (\%) for different sky conditions: (a) all skies; (b) clear skies; (c) cloudy skies.}\label{fig:cusummonth}
 \end{figure}
 
-<!-- ```{r cusummonth, echo = FALSE, fig.cap='Cumulative sum plots of the monthly SDR anomalies in (%) for different sky conditions.', fig.subcap=c('All skies.', 'Clear skies.', 'Cloudy skies.'), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3} -->
-<!-- knitr::include_graphics(c( -->
-<!--     "./images/CumulativeMonthlyCuSum-1.png",    # ALL sky -->
-<!--     "./images/CumulativeMonthlyCuSum-5.png",    # CLEAR sky -->
-<!--     "./images/CumulativeMonthlyCuSum-9.png"))   # Cloud sky -->
-<!-- ``` -->
+\vspace{-6pt}
 
@@
         {\centering 
-            \subfloat[All skies.\label{fig:cusumnotrendmonthly-1}]
+            % \subfloat[All skies.\label{fig:cusumnotrendmonthly-1}]
                 {\includegraphics[width=.32\linewidth]{./images/CumulativeMonthlyCuSumNOtrend-1} }\hfill
-            \subfloat[Clear skies.\label{fig:cusumnotrendmonthly-2}]
+            % \subfloat[Clear skies.\label{fig:cusumnotrendmonthly-2}]
                 {\includegraphics[width=.32\linewidth]{./images/CumulativeMonthlyCuSumNOtrend-5} }\hfill
-            \subfloat[Cloudy skies.\label{fig:cusumnotrendmonthly-3}]
+            % \subfloat[Cloudy skies.\label{fig:cusumnotrendmonthly-3}]
                 {\includegraphics[width=.32\linewidth]{./images/CumulativeMonthlyCuSumNOtrend-9} }
         }
-        \caption{Cumulative sum plots of monthly SDR anomalies in (\%) for different sky conditions after removing the long-term linear trend.}\label{fig:cusumnotrendmonthly}
+        \caption{Cumulative sum plots of monthly SDR anomalies in (\%) for different sky conditions after removing the long-term linear trend: (a) all skies; (b) clear skies; (c) cloudy skies.}\label{fig:cusumnotrendmonthly}
 \end{adjustwidth}
 \end{figure}
-
-<!-- ```{r cusumnotrendmonthly, echo = FALSE, fig.cap='Cumulative sum plots of monthly SDR anomalies in (%) for different sky conditions after removing the long-term linear trend.', fig.subcap=c('All skies.', 'Clear skies.', 'Cloudy skies.'), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3} -->
-<!-- knitr::include_graphics(c( -->
-<!--     "./images/CumulativeMonthlyCuSumNOtrend-1.png",    # ALL sky -->
-<!--     "./images/CumulativeMonthlyCuSumNOtrend-5.png",    # CLEAR sky -->
-<!--     "./images/CumulativeMonthlyCuSumNOtrend-9.png"))   # Cloud sky -->
-<!-- ``` -->
-
-<!-- ```{r cusum, echo = FALSE, fig.cap='Running cumulative sum plots of the daily SDR.', fig.subcap=c('All-sky conditions.', 'Clear-sky conditions.', 'Cloud-sky conditions.'), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3} -->
-<!-- knitr::include_graphics(c( -->
-<!--     "./images/CumulativeDailyCuSum-1.png",    # ALL sky -->
-<!--     "./images/CumulativeDailyCuSum-5.png",    # CLEAR sky -->
-<!--     "./images/CumulativeDailyCuSum-9.png"))   # Cloud sky -->
-<!-- ``` -->
-
-<!-- ```{r cusumnotrend, echo = FALSE, fig.cap='Running cumulative sum plots of the daily SDR after removing the long term trend from the daily values.', fig.subcap=c('All-skies.', 'Clear-skies.', 'Cloud-skies.'), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3} -->
-<!-- knitr::include_graphics(c( -->
-<!--     "./images/CumulativeDailyCuSumNOtrend-1.png",    # ALL sky -->
-<!--     "./images/CumulativeDailyCuSumNOtrend-5.png",    # CLEAR sky -->
-<!--     "./images/CumulativeDailyCuSumNOtrend-9.png"))   # Cloud sky -->
-<!-- ``` -->
-
-<!-- ```{r cusumSTD, echo = FALSE, fig.cap='two plots', fig.subcap=c('one plot', 'the other one'), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3} -->
-<!-- knitr::include_graphics(c( -->
-<!--     "./images/CumulativeDailyCuSumSTD-1.png",    # ALL sky -->
-<!--     "./images/CumulativeDailyCuSumSTD-2.png",    # CLEAR sky -->
-<!--     "./images/CumulativeDailyCuSumSTD-3.png"))   # Cloud sky -->
-<!-- ``` -->
 

### 2024-01-01 MDPI_submission.tex
Commit 2024-01-01 18:12

@@
 conditions; the plots for clear-sky and cloud-sky conditions, are shown
-in Appendix (Figures~\ref{fig:trendCLEAR} and~ \ref{fig:trendCLOUD}). In
-the studied period, there is no significant break or change in the
-variability pattern of the time series. The linear trends in all three
-datasets are positive and around \(0.4\,\%/y\) for all-sky and
-cloudy-sky conditions, whereas for clear-skies the trend is much smaller
-(\textasciitilde{}\(0.1\,\%/y\)). The linear trends were calculated
-taking into account the autocorrelation of the time series, and all
-three are statistically significant at least at the \(95\,\%\)
-confidence level, as they are larger than the corresponding \(2\sigma\)
-errors, despite the small values of R, which is due to the large
-variability of the daily values. The clear-sky trend is very small,
-suggesting a small effect from aerosols and water vapor, which are the
-dominant factors of the SDR variability
-\citep{Fountoulakis2016, Siomos2018, Yu2022}. In contrast, the large
-positive trend of SDR under cloudy skies can be attributed to reduction
-in cloud cover and/or cloud optical depth. Lack of continuous
+in Appendix \ref{appendix} (Figures~\ref{fig:trendCLEAR} and~
+\ref{fig:trendCLOUD}). In the studied period, there is no significant
+break or change in the variability pattern of the time series. The
+linear trends in all three datasets are positive and around
+\(0.4\,\%/y\) for all-sky and cloudy-sky conditions, whereas for
+clear-skies the trend is much smaller (\textasciitilde{}\(0.1\,\%/y\)).
+The linear trends were calculated taking into account the
+autocorrelation of the time series, and all three are statistically
+significant at least at the \(95\,\%\) confidence level, as they are
+larger than the corresponding \(2\sigma\) errors, despite the small
+values of R, which is due to the large variability of the daily values.
+The clear-sky trend is very small, suggesting a small effect from
+aerosols and water vapor, which are the dominant factors of the SDR
+variability \citep{Fountoulakis2016, Siomos2018, Yu2022}. In contrast,
+the large positive trend of SDR under cloudy skies can be attributed to
+reduction in cloud cover and/or cloud optical depth. Lack of continuous
 observations of cloud optical thickness that could support these
@@
 \toprule
-Sky conditions & Trend [\%/year] & Trend S.E. ($2\sigma$) & Pearson correl. & Trend [W/m\textsuperscript{2}/year] & Days\\
+\textbf{Sky Conditions} & \textbf{Trend [\%/year]} & \textbf{Trend S.E. ($2\sigma$)} & \textbf{Pearson Correl.} & \textbf{Trend [W/m\textsuperscript{2}/year]} & \textbf{Days}\\
 \midrule
@@
 
-\caption{Anomalies (\%) of the daily all-sky SDR from the climatological mean for the period 1993 -- 2023. The black line is the long term linear trend.}\label{fig:trendALL}
+\caption{Anomalies (\%) of the daily all-sky SDR from the climatological mean for the period 1993--2023. The black line is the long-term linear trend.}\label{fig:trendALL}
 \end{figure}
 
-Although the year-to-year variability of the anomalies (Figure
-\ref{fig:trendALL} and Figures~\ref{fig:trendCLEAR} and
-\ref{fig:trendCLOUD} in Appendix) shows a rather homogeneous behavior,
-plots of the cumulative sums (CUSUM) \citep{Regier2019} of the anomalies
-can reveal different structures in the records of all three sky
-conditions. For time series with a uniform trend, we would expect the
-CUSUMs of the anomalies to have a symmetric `V' shape centered around
-the middle of the series. This would indicate that the anomalies are
-evenly distributed around the climatological mean, and, for a positive
-uniform trend, the first half is below and the second half above the
-climatological mean. In our case, there is a more complex evolution of
-the anomalies. For all skies (Figure~\ref{fig:cusummonth-1}), we observe
-three rather distinct periods: (a) a downward part between the start of
-the datasets and about 2000, denoting that all anomalies are negative,
-thus below the climatology; (b) a relatively steady part lasting for
-almost 20 years suggesting little variability in SDR anomalies; and (c)
-a steep upward part to the present, indicating anomalies above the
-climatology. The CUSUMs for cloudy skies (Figure~\ref{fig:cusummonth-3})
-show a similar behavior with some short-term differences that do not
-change the overall pattern. For clear skies
-(Figure~\ref{fig:cusummonth-2}), a monotonic downward tendency is
-evident until 2004, suggesting that the anomalies are all negative.
-After 2004, the anomalies turn positive at a fast rate for about five
-years and at a slower rate thereafter.
+Although the year-to-year variability of the anomalies
+(Figures~\ref{fig:trendALL}, \ref{fig:trendCLEAR} and
+\ref{fig:trendCLOUD} in Appendix \ref{appendix}) shows a rather
+homogeneous behavior, plots of the cumulative sums (CUSUM)
+\citep{Regier2019} of the anomalies can reveal different structures in
+the records of all three sky conditions. For time series with a uniform
+trend, we would expect the CUSUMs of the anomalies to have a symmetric
+`V' shape centered around the middle of the series. This would indicate
+that the anomalies are evenly distributed around the climatological
+mean, and, for a positive uniform trend, the first half is below and the
+second half above the climatological mean. In our case, there is a more
+complex evolution of the anomalies. For all skies
+(Figure~\ref{fig:cusummonth}a), we observe three rather distinct
+periods: (a) a downward part between the start of the datasets and about
+2000, denoting that all anomalies are negative, thus below the
+climatology; (b) a relatively steady part lasting for almost 20 years,
+suggesting little variability in SDR anomalies; and (c) a steep upward
+part to the present, indicating anomalies above the climatology. The
+CUSUMs for cloudy skies (Figure~\ref{fig:cusummonth}c) show a similar
+behavior with some short-term differences that do not change the overall
+pattern. For clear skies (Figure~\ref{fig:cusummonth}b), a monotonic
+downward tendency is evident until 2004, suggesting that the anomalies
+are all negative. After 2004, the anomalies turn positive at a fast rate
+for about five years and at a slower rate thereafter.
 
@@
         {\centering 
-        \subfloat[All skies.\label{fig:cusummonth-1}]
             {\includegraphics[width=.32\linewidth]{./images/CumulativeMonthlyCuSum-1}}\hfill
-        \subfloat[Clear skies.\label{fig:cusummonth-2}]
             {\includegraphics[width=.32\linewidth]{./images/CumulativeMonthlyCuSum-5}}\hfill
-        \subfloat[Cloudy skies.\label{fig:cusummonth-3}]
-            {\includegraphics[width=.32\linewidth]{./images/CumulativeMonthlyCuSum-9}}\hfill
+            {\includegraphics[width=.32\linewidth]{./images/CumulativeMonthlyCuSum-9}}
         }
-\caption{Cumulative sum plots of the monthly SDR anomalies in (\%) for different sky conditions.}\label{fig:cusummonth}
     \end{adjustwidth}
+\caption{Cumulative sum plots of the monthly SDR anomalies in (\%) for different sky conditions: (a) all skies; (b) clear skies; (c) cloudy skies.}\label{fig:cusummonth}
 \end{figure}
+
+\vspace{-6pt}
 
@@
         {\centering 
-            \subfloat[All skies.\label{fig:cusumnotrendmonthly-1}]
+            % \subfloat[All skies.\label{fig:cusumnotrendmonthly-1}]
                 {\includegraphics[width=.32\linewidth]{./images/CumulativeMonthlyCuSumNOtrend-1} }\hfill
-            \subfloat[Clear skies.\label{fig:cusumnotrendmonthly-2}]
+            % \subfloat[Clear skies.\label{fig:cusumnotrendmonthly-2}]
                 {\includegraphics[width=.32\linewidth]{./images/CumulativeMonthlyCuSumNOtrend-5} }\hfill
-            \subfloat[Cloudy skies.\label{fig:cusumnotrendmonthly-3}]
+            % \subfloat[Cloudy skies.\label{fig:cusumnotrendmonthly-3}]
                 {\includegraphics[width=.32\linewidth]{./images/CumulativeMonthlyCuSumNOtrend-9} }
         }
-        \caption{Cumulative sum plots of monthly SDR anomalies in (\%) for different sky conditions after removing the long-term linear trend.}\label{fig:cusumnotrendmonthly}
+        \caption{Cumulative sum plots of monthly SDR anomalies in (\%) for different sky conditions after removing the long-term linear trend: (a) all skies; (b) clear skies; (c) cloudy skies.}\label{fig:cusumnotrendmonthly}
 \end{adjustwidth}

### 2024-01-01 MDPI_submission.Rmd
Commit 2024-01-01 19:12

@@
 
-# institutionalreview: |
-#   In this section, you should add the Institutional Review Board Statement and
-#   approval number, if relevant to your study. You might choose to exclude
-#   this statement if the study did not require ethical approval. Please note
-#   that the Editorial Office might ask you for further information. Please add
-#   “The study was conducted in accordance with the Declaration of Helsinki,
-#   and approved by the Institutional Review Board (or Ethics Committee) of
-#   NAME OF INSTITUTE (protocol code XXX and date of approval).” for studies
-#   involving humans. OR “The animal study protocol was approved by the
-#   Institutional Review Board (or Ethics Committee) of NAME OF INSTITUTE
-#   (protocol code XXX and date of approval).” for studies involving animals.
-#   OR “Ethical review and approval were waived for this study due to REASON
-#   (please provide a detailed justification).” OR “Not applicable” for
-#    studies not involving humans or animals.
-
-# informedconsent: |
-#   Any research article describing a study involving humans should contain this
-#   statement. Please add ``Informed consent was obtained from all subjects
-#   involved in the study.'' OR ``Patient consent was waived due to REASON
-#   (please provide a detailed justification).'' OR ``Not applicable'' for
-#   studies not involving humans. You might also choose to exclude this statement
-#   if the study did not involve humans.
-#
-#   Written informed consent for publication must be obtained from participating
-#   patients who can be identified (including by the patients themselves). Please
-#   state ``Written informed consent has been obtained from the patient(s) to
-#   publish this paper'' if applicable.
+institutionalreview: |
+  Not applicable.
+
+informedconsent: |
+  Not applicable.
 
@@
 
-# conflictsofinterest: |
-#   Declare conflicts of interest or state 'The authors declare no conflict of
-#   interest.' Authors must identify and declare any personal circumstances or
-#   interest that may be perceived as inappropriately influencing the
-#   representation or interpretation of reported research results. Any role of the
-#   funding sponsors in the design of the study; in the collection, analyses or
-#   interpretation of data in the writing of the manuscript, or in the decision to
-#   publish the results must be declared in this section. If there is no role,
-#   please state 'The founding sponsors had no role in the design of the study;
-#   in the collection, analyses, or interpretation of data; in the writing of the
-#   manuscript, an in the decision to publish the results'.
+conflictsofinterest: |
+  The authors declare no conflict of interest.
 
@@
 ## load data from trends analysis
-# source("./DHI_GHI_00_raw_data.R")
-# load(common_data)
 load("./data/Input_1_longterm_trends.Rda")
@@
     # else {
-    #     
     #     output <- vector(mode = "character", length = options$fig.num + 1)
@@
 Finally, we investigated the dependence of the trends on solar zenith angle and season.
-
-<!--
-Aerosols interactions
-
-#Yang2021  Cloud “shrinking” and “optical thinning” in the “dimming” period and a subsequent recovery in the “brightening” period over China
-
-Aerosols not only strengthen but also weaken the
-growth of clouds with respect to their coverage and
-optical thickness, depending on the levels of pollution
-and the associated amounts of aerosols as postulated
-in a conceptual framework (Wild 2009a, 2012, Yang
-et al 2012).
-
-#Wild2012 : simplify interaction nice graph
--->
 
@@
 
-# TSI$cor.null_value     <- NULL
-# TSI$cor.alternative    <- NULL
-# TSI$cor.method         <- NULL
-# TSI$cor.data_name      <- NULL
-# TSI$cor.p              <- NULL
 TSI$cor.conf_level      <- NULL
@@
 variation of SDR from the climatological mean. When the CUSUM is
-increasing, the anomalies values are above the climatology and vice
+increasing, the anomalies values are above the climatology, and vice
 versa. 
-Overall, for all- and cloudy-sky conditions (Figures\ \ref{fig:cusumnotrendmonthly-1} and\ \ref{fig:cusumnotrendmonthly-3}) we
+Overall, for all- and cloudy-sky conditions (Figure\ \ref{fig:cusumnotrendmonthly}a,c), we
 observe periods with anomalies diverging from the climatological values,
@@
 skies over Thessaloniki.
-For clear skies (Figure\ \ref{fig:cusumnotrendmonthly-2}) the distinct change
+For clear skies (Figure\ \ref{fig:cusumnotrendmonthly}b), the distinct change
 in 2004 is now clearer. The most likely reason for this change is the
@@
         }
+    \end{adjustwidth}
     \caption{Cumulative sum plots of monthly SDR anomalies in (\%) for different sky conditions after removing the long-term linear trend: (a) all skies; (b) clear skies; (c) cloudy skies.}\label{fig:cusumnotrendmonthly}
-\end{adjustwidth}
 \end{figure}
@@
 
-The solar zenith angle is a major factor affecting the SDR, since
+The solar zenith angle is a major factor affecting the SDR, as
 increases in SZA leads to enhancement of the radiation path in the
 atmosphere, especially in urban environments with human activities
-emitting aerosols 
-[@Wang2021].
-In order to estimate the effect of SZA on the SDR trends, we grouped the data in bins of $1^\circ$ SZA, and calculated the overall trend for each bin, separately for the daily
+emitting aerosols [@Wang2021].
+In order to estimate the effect of SZA on the SDR trends, we grouped the data in 
+bins of $1^\circ$ SZA and calculated the overall trend for each bin separately for the daily
 periods before noon and after noon (Figure\ \ref{fig:szatrends}).
-Although there are seasonal dependencies of the minimum SZA (see Appendix, Figure\ \ref{fig:SZAtrendSeason}), these dependencies are not discussed further.
-
-For all-sky conditions the brightening effect of SDR (positive trend) increases with SZAs (Figures\ \ref{fig:szatrends-1})
+Although there are seasonal dependencies of the minimum SZA (see Appendix \ref{appendix}, Figure\ \ref{fig:SZAtrendSeason}), these dependencies are not discussed further.
+
+For all-sky conditions, the brightening effect of SDR (positive trend) increases with SZAs (Figure\ \ref{fig:szatrends}a),
 ranging from about $0.1\,\%/y$ to
@@
 The trends in the morning and afternoon hours are more or
-less consistent with small differences at small SZAs which can be
+less consistent with small differences at small SZAs, which can be
 attributed to effects on clear sky SDR from systematic diurnal patterns
-of aerosols during the warm period of the year, consistently with the results reported for China by @Wang2021. 
-Note that SZAs less than $25^\circ$ can only occur during the warm period of the year around noon when clear skies are more
-frequent. 
+of aerosols during the warm period of the year, consistent with the 
+results reported for China by @Wang2021. 
+Note that SZAs less than $25^\circ$ can only occur during the warm period
+of the year around noon when clear skies are more frequent. 
 The increasing trend with SZA is likely caused by the
 increased attenuation of SDR with SZA. The effect is larger when aerosol
-and/or cloud layers are optically thicker, therefore, decreases in
+and/or cloud layers are optically thicker; therefore, decreases in
 aerosol and clouds through the study period will result in larger
@@
 
-Under clear skies (Figure\ \ref{fig:szatrends-2}), the trends are smaller and less variable,
+Under clear skies (Figure\ \ref{fig:szatrends}b), the trends are smaller
+and less variable,
 ranging between $0.1$ and $0.15\,\%/y$ up to $77^\circ$ SZA.
@@
 
-For cloudy-sky conditions (Figure\ \ref{fig:szatrends-3}), we cannot discern any significant dependence of the SDR trend with SZA as the variability of irradiance is
+For cloudy-sky conditions (Figure\ \ref{fig:szatrends}c), we cannot discern any significant dependence of the SDR trend with SZA, as the variability of irradiance is
 dominated by the cloud effects leading to insignificant
@@
 probably associated with stronger attenuation by clouds under oblique
-incidence angles, which result also in smaller variability.
+incidence angles, which also result in smaller variability.
 
@@
         {\centering 
-            \subfloat[All skies.\label{fig:szatrends-1}]
+            % \subfloat[All skies.\label{fig:szatrends-1}]
                 {\includegraphics[width=.32\linewidth]{./images/SzaTrends-1}}\hfill
-            \subfloat[Clear skies.\label{fig:szatrends-2}]
+            % \subfloat[Clear skies.\label{fig:szatrends-2}]
                 {\includegraphics[width=.32\linewidth]{./images/SzaTrends-4}}\hfill
-            \subfloat[Cloudy skies.\label{fig:szatrends-3}]
+            % \subfloat[Cloudy skies.\label{fig:szatrends-3}]
                 {\includegraphics[width=.32\linewidth]{./images/SzaTrends-7}}
         }
-        \caption{Long term trends of daily SDR as a function of SZA for (a) all-sky, (b) clear-sky and (c) cloudy-sky conditions, separately for morning and afternoon periods.
+    \end{adjustwidth}
+    \caption{Long-term trends of daily SDR as a function of SZA for (a) all-sky, (b) clear-sky and (c) cloudy-sky conditions, separately for morning and afternoon periods.
 Solid shapes represent statistically significant trends ($p < 0.005$).}\label{fig:szatrends}
-    \end{adjustwidth}
 \end{figure}
 
-<!-- ```{r szatrends, echo = FALSE, fig.cap='Long term trends of SDR as a function of SZA separately from morning and afternoon periods.
-Solid shapes  represent statistically significant trends ($p < 0.005$).', fig.subcap=c('All skies.', 'Clear skies.', 'Cloudy skies.'), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3} -->
-<!-- knitr::include_graphics("./images/SzaTrends-1.png") -->
-<!-- knitr::include_graphics("./images/SzaTrends-4.png") -->
-<!-- knitr::include_graphics("./images/SzaTrends-7.png") -->
-<!-- ``` -->
-
-<!-- ```{r szatrendsmonthly, echo = FALSE, fig.cap='*** Montlhly *** Distribution of the SDR\'s long term trends by SZA before and after noon.
-The solid data points represent values with an acceptable statistical significance ($p<0.005$).', fig.subcap=c('All-sky conditions.', 'Clear-sky conditions.', 'Cloud-sky conditions.'), out.width='.32\\linewidth', fig.asp=1, fig.ncol = 3} -->
-<!-- knitr::include_graphics("./images/SzaTrendsMonthly-1.png") -->
-<!-- knitr::include_graphics("./images/SzaTrendsMonthly-4.png" ) -->
-<!-- knitr::include_graphics("./images/SzaTrendsMonthly-7.png") -->
-<!-- ``` -->
-
-## Long term SDR trends by season
+## Long-Term SDR Trends by Season
 
@@
 spp$cor.estimate    <- round(spp$cor.estimate,       2)
-# spp$slope.stat_sig  <- signif(spp$slope.stat_sig, 3)
-# spp$Rsqrd           <- signif(spp$Rsqrd,          3)
-
-# spp$Tlm_Pr...t..
 
@@
 
-Similarly to the long term trends from daily means of SDR discussed above, we have calculated the trend for the three sky conditions and for each season of the year, using the corresponding mean monthly anomalies
+Similarly to the long term trends from daily means of SDR discussed above,
+we have calculated the trend for the three sky conditions and for each 
+season of the year, using the corresponding mean monthly anomalies
 (Figure\ \ref{fig:seasonalALL} and Table\ \ref{tab:trendseasontable}).
-Table\ \ref{tab:trendseasontable} contains also the $2\sigma$ standard error, the Pearson's correlation coefficient R and the corresponding p-value.
-The winter linear trends exhibit generally the largest R values ranging between $0.54$ and $0.60\,\%/y$. 
+Table\ \ref{tab:trendseasontable} also contains the $2\sigma$ standard error,
+the Pearson's correlation coefficient R and the corresponding p-value.
+The winter linear trends exhibit generally the largest R values ranging
+between $0.54$ and $0.60\,\%/y$. 
 <!-- all sky -->
-For all-sky conditions the trend in SDR in winter is the largest
+For all-sky conditions, the trend in SDR in winter is the largest
 ($`r signif(spp$"Trend [%/year]"[spp$Season == "Winter" & spp$"Sky condition" == "All skies"], 2)`\,\%/y$),
@@
 <!-- clear sky -->
-In other seasons the clear-sky trend is very small
-(below $0.1\,\%/y$). 
+In other seasons the clear-sky trend is very small (below $0.1\,\%/y$). 
 <!-- cloud dsky -->
-Finally, for cloudy-skies the
-winter trend is the largest 
+Finally, for cloudy-skies the winter trend is the largest 
 ($`r signif(spp$"Trend [%/year]"[spp$Season == "Winter" & spp$"Sky condition" == "Cloudy skies"], 2)`\,\%/y$)
-and greater than for all-skies,
-followed by a much smaller trend in autumn 
+and greater than for all-skies, followed by a much smaller trend in autumn 
 ($`r signif(spp$"Trend [%/year]"[spp$Season == "Autumn" & spp$"Sky condition" == "Cloudy skies"], 2)`\,\%/y$).
-<!-- ($`r signif(spp$"Trend [%/year]"[spp$Season == "Winter" & spp$"Sky condition" == "All skies"], 2)`\,\%/y$) -->
 
@@
 
-<!-- % \includegraphics[width=1\linewidth]{./images/SeasonalTrendsTogether3-2}  % Seasonal from Daily -->
-
-<!-- ```{r seasonalALL, echo=F, fig.cap="Linear trends (black lines) of monthly mean anomalies of SDR by season (rows of plots) for the three sky conditions (columns of plots).", out.width="100%"} -->
-<!-- knitr::include_graphics("./images/SeasonalTrendsTogether3-2.png") -->
-<!-- ``` -->
-
-<!-- ```{r seasonalALL, echo=F, fig.cap="Trends by season for all-sky condition. Displays monthly means of daily means."} -->
-<!-- knitr::include_graphics("./images/SeasonalTrends-1.png") -->
-<!-- ``` -->
-
-<!-- ```{r seasonalCLEAR, echo=F, fig.cap="Trends by season for clear-sky condition. Displays monthly means of daily means."} -->
-<!-- knitr::include_graphics("./images/SeasonalTrends-2.png") -->
-<!-- ``` -->
-
-<!-- ```{r seasonalCLOUD, echo=F, fig.cap="Trends by season for Cloudy-sky condition. Displays monthly means of daily means."} -->
-<!-- knitr::include_graphics("./images/SeasonalTrends-3.png") -->
-<!-- ``` -->
-
 ```{r trendseasontable, echo=F}
-
 spp2 <- spp
@@
 
-# sub("Trend S.E.", "Trend S.E. ($2\\sigma$)", names(spp2))
-
-# panderOptions("table.alignment.default", "left")
-# panderOptions("table.alignment.default", "center")
-# pander(spp,
-#        justify = c('center', 'right', 'right', 'center'),
-#        split.table = Inf,
-#        caption = "\\label{tab:trendseasontable}SDR linear trends of monthly anomalies for each season of the year.")
-
-## use kable
 x_latex <- knitr::kable(spp2,
@@
                         caption  = "SDR linear trends of monthly anomalies for each season of the year and related statistical parameters.")
-
-# x_latex %>%
-#     collapse_rows(columns = 1, valign = "middle", latex_hline = "custom") %>%
-#     kable_styling(latex_options = c("striped",
-#                                     "hold_position")
-#     )
 

### 2024-01-01 MDPI_submission.tex
Commit 2024-01-01 19:12

@@
 mean. When the CUSUM is increasing, the anomalies values are above the
-climatology and vice versa. Overall, for all- and cloudy-sky conditions
-(Figures~\ref{fig:cusumnotrendmonthly-1}
-and~\ref{fig:cusumnotrendmonthly-3}) we observe periods with anomalies
-diverging from the climatological values, each lasting for several
-years. These fluctuations are probably within the natural variability,
-and no distinct changes are identified. The pattern in both datasets is
-similar, suggesting prevalence in cloudy skies over Thessaloniki. For
-clear skies (Figure~\ref{fig:cusumnotrendmonthly-2}) the distinct change
-in 2004 is now clearer. The most likely reason for this change is the
-monotonic reduction of aerosols in Thessaloniki. In that year, there was
-a change in the rate of decrease in aerosol optical depth, as
-illustrated in Figure 7 of \citet{Siomos2020}. This abrupt change in
-CUSUMs lasted until about 2010, when the anomalies become again
-variable.
+climatology, and vice versa. Overall, for all- and cloudy-sky conditions
+(Figure~\ref{fig:cusumnotrendmonthly}a,c), we observe periods with
+anomalies diverging from the climatological values, each lasting for
+several years. These fluctuations are probably within the natural
+variability, and no distinct changes are identified. The pattern in both
+datasets is similar, suggesting prevalence in cloudy skies over
+Thessaloniki. For clear skies (Figure~\ref{fig:cusumnotrendmonthly}b),
+the distinct change in 2004 is now clearer. The most likely reason for
+this change is the monotonic reduction of aerosols in Thessaloniki. In
+that year, there was a change in the rate of decrease in aerosol optical
+depth, as illustrated in Figure 7 of \citet{Siomos2020}. This abrupt
+change in CUSUMs lasted until about 2010, when the anomalies become
+again variable.
 
@@
 
-The solar zenith angle is a major factor affecting the SDR, since
-increases in SZA leads to enhancement of the radiation path in the
-atmosphere, especially in urban environments with human activities
-emitting aerosols \citep{Wang2021}. In order to estimate the effect of
-SZA on the SDR trends, we grouped the data in bins of \(1^\circ\) SZA,
-and calculated the overall trend for each bin, separately for the daily
-periods before noon and after noon (Figure~\ref{fig:szatrends}).
-Although there are seasonal dependencies of the minimum SZA (see
-Appendix, Figure~\ref{fig:SZAtrendSeason}), these dependencies are not
-discussed further.
+The solar zenith angle is a major factor affecting the SDR, as increases
+in SZA leads to enhancement of the radiation path in the atmosphere,
+especially in urban environments with human activities emitting aerosols
+\citep{Wang2021}. In order to estimate the effect of SZA on the SDR
+trends, we grouped the data in bins of \(1^\circ\) SZA and calculated
+the overall trend for each bin separately for the daily periods before
+noon and after noon (Figure~\ref{fig:szatrends}). Although there are
+seasonal dependencies of the minimum SZA (see Appendix \ref{appendix},
+Figure~\ref{fig:SZAtrendSeason}), these dependencies are not discussed
+further.
 
-For all-sky conditions the brightening effect of SDR (positive trend)
-increases with SZAs (Figures~\ref{fig:szatrends-1}) ranging from about
+For all-sky conditions, the brightening effect of SDR (positive trend)
+increases with SZAs (Figure~\ref{fig:szatrends}a), ranging from about
 \(0.1\,\%/y\) to about \(0.7\,\%/y\) for the statistically significant
 trends. The trends in the morning and afternoon hours are more or less
-consistent with small differences at small SZAs which can be attributed
+consistent with small differences at small SZAs, which can be attributed
 to effects on clear sky SDR from systematic diurnal patterns of aerosols
-during the warm period of the year, consistently with the results
-reported for China by \citet{Wang2021}. Note that SZAs less than
-\(25^\circ\) can only occur during the warm period of the year around
-noon when clear skies are more frequent. The increasing trend with SZA
-is likely caused by the increased attenuation of SDR with SZA. The
-effect is larger when aerosol and/or cloud layers are optically thicker,
-therefore, decreases in aerosol and clouds through the study period will
-result in larger positive trends of SDR at larger SZAs.
+during the warm period of the year, consistent with the results reported
+for China by \citet{Wang2021}. Note that SZAs less than \(25^\circ\) can
+only occur during the warm period of the year around noon when clear
+skies are more frequent. The increasing trend with SZA is likely caused
+by the increased attenuation of SDR with SZA. The effect is larger when
+aerosol and/or cloud layers are optically thicker; therefore, decreases
+in aerosol and clouds through the study period will result in larger
+positive trends of SDR at larger SZAs.
 
-Under clear skies (Figure~\ref{fig:szatrends-2}), the trends are smaller
+Under clear skies (Figure~\ref{fig:szatrends}b), the trends are smaller
 and less variable, ranging between \(0.1\) and \(0.15\,\%/y\) up to
@@
 
-For cloudy-sky conditions (Figure~\ref{fig:szatrends-3}), we cannot
-discern any significant dependence of the SDR trend with SZA as the
+For cloudy-sky conditions (Figure~\ref{fig:szatrends}c), we cannot
+discern any significant dependence of the SDR trend with SZA, as the
 variability of irradiance is dominated by the cloud effects leading to
@@
 clear skies, is probably associated with stronger attenuation by clouds
-under oblique incidence angles, which result also in smaller
+under oblique incidence angles, which also result in smaller
 variability.
@@
         {\centering 
-            \subfloat[All skies.\label{fig:szatrends-1}]
+            % \subfloat[All skies.\label{fig:szatrends-1}]
                 {\includegraphics[width=.32\linewidth]{./images/SzaTrends-1}}\hfill
-            \subfloat[Clear skies.\label{fig:szatrends-2}]
+            % \subfloat[Clear skies.\label{fig:szatrends-2}]
                 {\includegraphics[width=.32\linewidth]{./images/SzaTrends-4}}\hfill
-            \subfloat[Cloudy skies.\label{fig:szatrends-3}]
+            % \subfloat[Cloudy skies.\label{fig:szatrends-3}]
                 {\includegraphics[width=.32\linewidth]{./images/SzaTrends-7}}
         }
-        \caption{Long term trends of daily SDR as a function of SZA for (a) all-sky, (b) clear-sky and (c) cloudy-sky conditions, separately for morning and afternoon periods.
+    \end{adjustwidth}
+    \caption{Long-term trends of daily SDR as a function of SZA for (a) all-sky, (b) clear-sky and (c) cloudy-sky conditions, separately for morning and afternoon periods.
 Solid shapes represent statistically significant trends ($p < 0.005$).}\label{fig:szatrends}
-    \end{adjustwidth}
 \end{figure}
@@
 \hypertarget{long-term-sdr-trends-by-season}{%
-\subsection{Long term SDR trends by
-season}\label{long-term-sdr-trends-by-season}}
+\subsection{Long-Term SDR Trends by
+Season}\label{long-term-sdr-trends-by-season}}
 
@@
 (Figure~\ref{fig:seasonalALL} and Table~\ref{tab:trendseasontable}).
-Table~\ref{tab:trendseasontable} contains also the \(2\sigma\) standard
+Table~\ref{tab:trendseasontable} also contains the \(2\sigma\) standard
 error, the Pearson's correlation coefficient R and the corresponding
 p-value. The winter linear trends exhibit generally the largest R values
-ranging between \(0.54\) and \(0.60\,\%/y\). For all-sky conditions the
+ranging between \(0.54\) and \(0.60\,\%/y\). For all-sky conditions, the
 trend in SDR in winter is the largest (\(0.7\,\%/y\)), followed by the
@@
 
+\institutionalreview{Not applicable.}
+
+\informedconsent{Not applicable.}
+
 \dataavailability{Data as daily sums are available through the WRDC
@@
 available for protection against unmonitored commercial use.}
+
+\conflictsofinterest{The authors declare no conflict of interest.}
 

### 2024-01-02 MDPI_submission.Rmd
Commit 2024-01-02 16:13

@@
 
-```{r trendtable, echo=F}
+```{r trendtable, echo=F, fig.pos="H"}
 # names(LT1)
@@
     row_spec(0,bold=TRUE)            %>%
-    kable_styling(full_width = TRUE) %>%
+    kable_styling(full_width = TRUE, 
+                  repeat_header_continued = T) %>%
     column_spec(1, width = "8em")
@@
 we have calculated the trend for the three sky conditions and for each 
-season of the year, using the corresponding mean monthly anomalies
+season of the year using the corresponding mean monthly anomalies
 (Figure\ \ref{fig:seasonalALL} and Table\ \ref{tab:trendseasontable}).
 Table\ \ref{tab:trendseasontable} also contains the $2\sigma$ standard error,
-the Pearson's correlation coefficient R and the corresponding p-value.
-The winter linear trends exhibit generally the largest R values ranging
+the Pearson's correlation coefficient R, and the corresponding p-value.
+The winter linear trends generally exhibit the largest R values, ranging
 between $0.54$ and $0.60\,\%/y$. 
@@
 ($`r signif(spp$"Trend [%/year]"[spp$Season == "Autumn" & spp$"Sky condition" == "All skies"], 2)`\,\%/y$,
-a value close to the long term trend) both statistically significant at the $95\,\%$ confidence level.
+a value close to the long-term trend), both statistically significant at the $95\,\%$ confidence level.
 In spring and summer, the trends are much smaller and of lesser statistical significance.
@@
 <!-- clear sky -->
-For clear-skies, the trend in winter is $`r signif(spp$"Trend [%/year]"[spp$Season == "Winter" & spp$"Sky condition" == "Clear skies"], 2)`\,\%/y$ and is 
+For clear skies, the trend in winter is $`r signif(spp$"Trend [%/year]"[spp$Season == "Winter" & spp$"Sky condition" == "Clear skies"], 2)`\,\%/y$ and is 
 associated with the decreasing trend in
 aerosol optical depth [@Siomos2020]. Moreover, it is almost half of that for
-all-skies, which is another indication of a decreasing trend in cloud
+all skies, which is another indication of a decreasing trend in cloud
 optical thickness. 
 <!-- clear sky -->
-In other seasons the clear-sky trend is very small (below $0.1\,\%/y$). 
+In other seasons, the clear-sky trend is very small (below $0.1\,\%/y$). 
 <!-- cloud dsky -->
-Finally, for cloudy-skies the winter trend is the largest 
+Finally, for cloudy skies, the winter trend is the largest 
 ($`r signif(spp$"Trend [%/year]"[spp$Season == "Winter" & spp$"Sky condition" == "Cloudy skies"], 2)`\,\%/y$)
-and greater than for all-skies, followed by a much smaller trend in autumn 
+and greater than for all skies, followed by a much smaller trend in autumn 
 ($`r signif(spp$"Trend [%/year]"[spp$Season == "Autumn" & spp$"Sky condition" == "Cloudy skies"], 2)`\,\%/y$).
 
-The trends under clear- and cloudy-sky conditions are of the same
+The trends under clear- and cloudy-sky conditions are in the same
 direction, and it would be expected that their sum is similar to the
@@
 Daily means were calculated only when at least $60\,\%$ of the clear- or
-cloudy-sky data were available (see sect. \ref{aggregationstatistical}).
-
-\begin{figure}[h!]
-    \begin{adjustwidth}{-\extralength}{0cm}
-        {\centering 
-            \includegraphics[width=1\linewidth]{./images/SeasonalMTrendsTogether3-2}   % Seasonal from Monthly
-        }
-        \caption{Monthly mean anomalies of SDR by season (rows of plots) for the three sky conditions (columns of plots). The black lines represent the linear trends.}\label{fig:seasonalALL}
-    \end{adjustwidth}
-\end{figure}
+cloudy-sky data were available (see Section \ref{aggregationstatistical}).
 
 ```{r trendseasontable, echo=F}
-spp2 <- spp
-
-# names(spp2)
-# col.names   = c("Sky conditions", "Trend [\\%/year]", "Trend S.E. ($2\\sigma$)", "Pearson correl.", "Trend [W/m\\textsuperscript{2}/year]", "Days"),
-
-x_latex <- knitr::kable(spp2,
+x_latex <- knitr::kable(spp,
                         format      = "latex",
@@
 ```
+
+\begin{figure}[h!]
+    \begin{adjustwidth}{-\extralength}{0cm}
+        {\centering 
+            \includegraphics[width=1\linewidth]{./images/SeasonalMTrendsTogether3-2}   % Seasonal from Monthly
+        }
+        \caption{Monthly mean anomalies of SDR by season (rows of plots) for the three sky conditions (columns of plots). The black lines represent the linear trends.}\label{fig:seasonalALL}
+    \end{adjustwidth}
+\end{figure}
 

### 2024-01-02 MDPI_submission.tex
Commit 2024-01-02 16:13

@@
 above, we have calculated the trend for the three sky conditions and for
-each season of the year, using the corresponding mean monthly anomalies
+each season of the year using the corresponding mean monthly anomalies
 (Figure~\ref{fig:seasonalALL} and Table~\ref{tab:trendseasontable}).
 Table~\ref{tab:trendseasontable} also contains the \(2\sigma\) standard
-error, the Pearson's correlation coefficient R and the corresponding
-p-value. The winter linear trends exhibit generally the largest R values
-ranging between \(0.54\) and \(0.60\,\%/y\). For all-sky conditions, the
-trend in SDR in winter is the largest (\(0.7\,\%/y\)), followed by the
-trend in autumn (\(0.42\,\%/y\), a value close to the long term trend)
-both statistically significant at the \(95\,\%\) confidence level. In
-spring and summer, the trends are much smaller and of lesser statistical
-significance. These seasonal differences indicate a possible relation of
-the trends in SDR to trends in clouds during winter and autumn. For
-clear-skies, the trend in winter is \(0.4\,\%/y\) and is associated with
-the decreasing trend in aerosol optical depth \citep{Siomos2020}.
-Moreover, it is almost half of that for all-skies, which is another
-indication of a decreasing trend in cloud optical thickness. In other
-seasons the clear-sky trend is very small (below \(0.1\,\%/y\)).
-Finally, for cloudy-skies the winter trend is the largest
-(\(0.76\,\%/y\)) and greater than for all-skies, followed by a much
-smaller trend in autumn (\(0.19\,\%/y\)).
+error, the Pearson's correlation coefficient R, and the corresponding
+p-value. The winter linear trends generally exhibit the largest R
+values, ranging between \(0.54\) and \(0.60\,\%/y\). For all-sky
+conditions, the trend in SDR in winter is the largest (\(0.7\,\%/y\)),
+followed by the trend in autumn (\(0.42\,\%/y\), a value close to the
+long-term trend), both statistically significant at the \(95\,\%\)
+confidence level. In spring and summer, the trends are much smaller and
+of lesser statistical significance. These seasonal differences indicate
+a possible relation of the trends in SDR to trends in clouds during
+winter and autumn. For clear skies, the trend in winter is \(0.4\,\%/y\)
+and is associated with the decreasing trend in aerosol optical depth
+\citep{Siomos2020}. Moreover, it is almost half of that for all skies,
+which is another indication of a decreasing trend in cloud optical
+thickness. In other seasons, the clear-sky trend is very small (below
+\(0.1\,\%/y\)). Finally, for cloudy skies, the winter trend is the
+largest (\(0.76\,\%/y\)) and greater than for all skies, followed by a
+much smaller trend in autumn (\(0.19\,\%/y\)).
 
-The trends under clear- and cloudy-sky conditions are of the same
+The trends under clear- and cloudy-sky conditions are in the same
 direction, and it would be expected that their sum is similar to the
@@
 Daily means were calculated only when at least \(60\,\%\) of the clear-
-or cloudy-sky data were available (see sect.
+or cloudy-sky data were available (see Section
 \ref{aggregationstatistical}).
-
-\begin{figure}[h!]
-    \begin{adjustwidth}{-\extralength}{0cm}
-        {\centering 
-            \includegraphics[width=1\linewidth]{./images/SeasonalMTrendsTogether3-2}   % Seasonal from Monthly
-        }
-        \caption{Monthly mean anomalies of SDR by season (rows of plots) for the three sky conditions (columns of plots). The black lines represent the linear trends.}\label{fig:seasonalALL}
-    \end{adjustwidth}
-\end{figure}
 
@@
 \end{table}
+
+\begin{figure}[h!]
+    \begin{adjustwidth}{-\extralength}{0cm}
+        {\centering 
+            \includegraphics[width=1\linewidth]{./images/SeasonalMTrendsTogether3-2}   % Seasonal from Monthly
+        }
+        \caption{Monthly mean anomalies of SDR by season (rows of plots) for the three sky conditions (columns of plots). The black lines represent the linear trends.}\label{fig:seasonalALL}
+    \end{adjustwidth}
+\end{figure}
 

### 2024-01-02 MDPI_submission.Rmd
Commit 2024-01-02 17:13

@@
    - \usepackage{placeins}
+   - \usepackage{float}
 
@@
 
-```{r trendtable, echo=F, fig.pos="H"}
+```{r trendtable, echo=F}
 # names(LT1)
@@
 x_latex                              %>%
-    row_spec(0,bold=TRUE)            %>%
-    kable_styling(full_width = TRUE, 
-                  repeat_header_continued = T) %>%
+    row_spec(0,bold = TRUE)          %>%
+    kable_styling(full_width = TRUE) %>%
     column_spec(1, width = "8em")
@@
 
-```{r trendseasontable, echo=F}
+```{r trendseasontable, echo=F, fig.pos='!h'}
 x_latex <- knitr::kable(spp,
@@
                         booktabs    = TRUE,
+                        longtable = T, 
                         escape      = FALSE,
-                        col.names   = c("Sky condition",
+                        col.names   = c("Sky Condition",
                                         "Season",
@@
                                         "Trend S.E. ($2\\sigma$)",
-                                        "Pearson correl.",
-                                        "Trend p-value" ),
-                        # format.args = list(digits = 3),
+                                        "Pearson Correl.",
+                                        "Trend p-Value" ),
                         linesep     = "",
@@
                         table.envir = "table",
-                        # position    = "H",
                         caption     = "SDR linear trends of monthly anomalies for each season of the year and related statistical parameters.")
@@
     collapse_rows(columns = 1, valign = "middle", latex_hline = "custom")           %>%
-    kable_styling(latex_options = c('striped', 'hold_position'), full_width = TRUE) %>%
+    row_spec(0, bold = TRUE)                                                        %>%
+    kable_styling(latex_options = c('striped', 
+                                    'hold_position', 
+                                    "repeat_header"),
+                  repeat_header_continued = TRUE,
+                  full_width = TRUE) %>%
     column_spec(1, width = "8em")
@@
         }
-        \caption{Monthly mean anomalies of SDR by season (rows of plots) for the three sky conditions (columns of plots). The black lines represent the linear trends.}\label{fig:seasonalALL}
     \end{adjustwidth}
+    \caption{Time serries of monthly mean anomalies of SDR by season (rows of plots) for the three sky conditions (columns of plots).
+The black lines represent the linear trends, (a-c) winter}\label{fig:seasonalALL}
 \end{figure}
+
+\FloatBarrier
 
@@
 We have analyzed a 30-year dataset of SRD measurements in Thessaloniki Greece 
-(`r paste(year(min(ALL_1_daily_DESEAS$Date)), " -- ", year(max(ALL_1_daily_DESEAS$Date)))`) aiming to identify the long term variability of
+(`r paste(year(min(ALL_1_daily_DESEAS$Date)), " -- ", year(max(ALL_1_daily_DESEAS$Date)))`), aiming to identify the long-term variability of
 solar irradiance under different sky conditions.
@@
 and
-$`r signif(spp$"Trend [%/year]"[spp$Season == "Winter" & spp$"Sky condition" == "Cloudy skies"], 2)`\,\%/y$
+$`r signif(spp$"Trend [%/year]"[spp$Season == "Winter" & spp$"Sky Condition" == "Cloudy skies"], 2)`\,\%/y$
 for all- and cloudy-skies, respectively, in contrast to spring and summer.
-The trends for autumn are also significant but smaller ($`r signif(spp$"Trend [%/year]"[spp$Season == "Autumn" & spp$"Sky condition" == "All skies"], 2)`$
+The trends for autumn are also significant but smaller ($`r signif(spp$"Trend [%/year]"[spp$Season == "Autumn" & spp$"Sky Condition" == "All skies"], 2)`$
 and
-$`r signif(spp$"Trend [%/year]"[spp$Season == "Autumn" & spp$"Sky condition" == "Cloudy skies"], 2)`\,\%/y$
+$`r signif(spp$"Trend [%/year]"[spp$Season == "Autumn" & spp$"Sky Condition" == "Cloudy skies"], 2)`\,\%/y$
 for all- and cloudy-skies, respectively).
@@
 The trend for clear skies is largest in winter
-($`r signif(spp$"Trend [%/year]"[spp$Season == "Winter" & spp$"Sky condition" == "Clear skies"], 2)`\,\%/y$) 
+($`r signif(spp$"Trend [%/year]"[spp$Season == "Winter" & spp$"Sky Condition" == "Clear skies"], 2)`\,\%/y$) 
 and negligible in spring, summer and autumn.

### 2024-01-02 MDPI_submission.tex
Commit 2024-01-02 17:13

@@
 \usepackage{placeins}
+\usepackage{float}
 \usepackage{longtable}
@@
 
-\begin{table}[!h]
-
-\caption{\label{tab:trendseasontable}SDR linear trends of monthly anomalies for each season of the year and related statistical parameters.}
-\begin{tabu} to \linewidth {>{\centering\arraybackslash}p{8em}>{\centering}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X}
+\begin{longtabu} to \linewidth {>{\centering\arraybackslash}p{8em}>{\centering}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X}
+\caption{\label{tab:trendseasontable}SDR linear trends of monthly anomalies for each season of the year and related statistical parameters.}\\
 \toprule
-Sky condition & Season & Trend [\%/year] & Trend S.E. ($2\sigma$) & Pearson correl. & Trend p-value\\
+\textbf{Sky Condition} & \textbf{Season} & \textbf{Trend [\%/year]} & \textbf{Trend S.E. ($2\sigma$)} & \textbf{Pearson Correl.} & \textbf{Trend p-Value}\\
 \midrule
+\endfirsthead
+\caption[]{SDR linear trends of monthly anomalies for each season of the year and related statistical parameters. \textit{(continued)}}\\
+\toprule
+\textbf{Sky Condition} & \textbf{Season} & \textbf{Trend [\%/year]} & \textbf{Trend S.E. ($2\sigma$)} & \textbf{Pearson Correl.} & \textbf{Trend p-Value}\\
+\midrule
+\endhead
+\midrule
+\multicolumn{6}{r@{}}{\textit{(continued \ldots)}}\
+\endfoot
+\bottomrule
+\endlastfoot
 \cellcolor{gray!6}{} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{0.70} & \cellcolor{gray!6}{0.43} & \cellcolor{gray!6}{0.54} & \cellcolor{gray!6}{0.003}\\
@@
 
-\multirow{-4}{*}{\centering\arraybackslash Cloudy skies} & Autumn & 0.19 & 0.43 & 0.16 & 0.384\\
-\bottomrule
-\end{tabu}
-\end{table}
+\multirow{-4}{*}{\centering\arraybackslash Cloudy skies} & Autumn & 0.19 & 0.43 & 0.16 & 0.384\\*
+\end{longtabu}
+
+\emph{\}\{\centering\arraybackslash Cloudy skies\} \& Autumn \& 0.19 \&
+0.43 \& 0.16 \& 0.384\textbackslash{}} \textbackslash end\{longtabu\}
 
@@
 We have analyzed a 30-year dataset of SRD measurements in Thessaloniki
-Greece (1993 -- 2023) aiming to identify the long term variability of
+Greece (1993 -- 2023), aiming to identify the long-term variability of
 solar irradiance under different sky conditions. Under all-sky
@@
 on season, with winter showing the strongest statistically significant
-trend of \(0.7\) and \(0.76\,\%/y\) for all- and cloudy-skies,
-respectively, in contrast to spring and summer. The trends for autumn
-are also significant but smaller (\(0.42\) and \(0.19\,\%/y\) for all-
-and cloudy-skies, respectively). The trend for clear skies is largest in
-winter (\(0.4\,\%/y\)) and negligible in spring, summer and autumn.
+trend of \(0.7\) and \(\,\%/y\) for all- and cloudy-skies, respectively,
+in contrast to spring and summer. The trends for autumn are also
+significant but smaller (\$\$ and \(\,\%/y\) for all- and cloudy-skies,
+respectively). The trend for clear skies is largest in winter
+(\(\,\%/y\)) and negligible in spring, summer and autumn.
 

### 2024-01-02 MDPI_submission.Rmd
Commit 2024-01-02 18:13

@@
 
-# authorcontributions: |
-#   For research articles with several authors, a short paragraph specifying their
-#   individual contributions must be provided. The following statements should be
-#   used ``X.X. and Y.Y. conceive and designed the experiments; X.X. performed the
-#   experiments; X.X. and Y.Y. analyzed the data; W.W. contributed
-#   reagents/materials/analysis tools; Y.Y. wrote the paper.'' Authorship must be
-#   limited to those who have contributed substantially to the work reported.
+authorcontributions: |
+  Conceptualization, methodology, software, data analysis and original writing---draft preparation A.N.; writing---review and editing, supervision, A.B.; data curation, C.M.
+All authors have read and agreed to the published version of the manuscript.
 
@@
     \end{adjustwidth}
-\caption{Cumulative sum plots of the monthly SDR anomalies in (\%) for different sky conditions: (a) all skies; (b) clear skies; (c) cloudy skies.}\label{fig:cusummonth}
+\caption{Cumulative sum plots of the monthly SDR anomalies in (\%) for different sky conditions: (\textbf{a})\ all skies; (\textbf{b}) clear skies; (\textbf{c}) cloudy skies.}\label{fig:cusummonth}
 \end{figure}
@@
     \end{adjustwidth}
-    \caption{Cumulative sum plots of monthly SDR anomalies in (\%) for different sky conditions after removing the long-term linear trend: (a) all skies; (b) clear skies; (c) cloudy skies.}\label{fig:cusumnotrendmonthly}
+    \caption{Cumulative sum plots of monthly SDR anomalies in (\%) for different sky conditions after removing the long-term linear trend: (\textbf{a})\ all skies; (\textbf{b}) clear skies; (\textbf{c}) cloudy skies.}\label{fig:cusumnotrendmonthly}
 \end{figure}
@@
 
-```{r trendseasontable, echo=F, fig.pos='!h'}
+```{r trendseasontable, echo=F, out.extra = ''}
 x_latex <- knitr::kable(spp,
@@
                         booktabs    = TRUE,
-                        longtable = T, 
                         escape      = FALSE,
@@
                         table.envir = "table",
+                        potition    = 'H',
                         caption     = "SDR linear trends of monthly anomalies for each season of the year and related statistical parameters.")
@@
 
-\begin{figure}[h!]
-    \begin{adjustwidth}{-\extralength}{0cm}
-        {\centering 
-            \includegraphics[width=1\linewidth]{./images/SeasonalMTrendsTogether3-2}   % Seasonal from Monthly
-        }
-    \end{adjustwidth}
-    \caption{Time serries of monthly mean anomalies of SDR by season (rows of plots) for the three sky conditions (columns of plots).
-The black lines represent the linear trends, (a-c) winter}\label{fig:seasonalALL}
-\end{figure}
+<!-- \begin{figure}[h!] -->
+<!--     \begin{adjustwidth}{-\extralength}{0cm} -->
+<!--         {\centering  -->
+<!--             \includegraphics[width=1\linewidth]{./images/SeasonalMTrendsTogether3-2}   % Seasonal from Monthly -->
+<!--         } -->
+<!--     \end{adjustwidth} -->
+<!--     \caption{Time serries of monthly mean anomalies of SDR by season (rows of plots) for the three sky conditions (columns of plots).
+The black lines represent the linear trends, (\textbf{a}--\textbf{c}) winter, (\textbf{d}--\textbf{f}) spring, (\textbf{g}--\textbf{i}) summer, (\textbf{j}--\textbf{l}) autumn, and also (\textbf{a}, \textbf{d}, \textbf{g}, \textbf{j}) all-skies, (\textbf{b}, \textbf{e}, \textbf{h}, \textbf{k}) clear-skies and (\textbf{c}, \textbf{f}, \textbf{i}, \textbf{l}) cloudy-skies.}\label{fig:seasonalALL} -->
+<!-- \end{figure} -->
+
+\vspace{+100pt}
 
@@
 and
-$`r signif(spp$"Trend [%/year]"[spp$Season == "Winter" & spp$"Sky Condition" == "Cloudy skies"], 2)`\,\%/y$
+$`r signif(spp$"Trend [%/year]"[spp$Season == "Winter" & spp$"Sky condition" == "Cloudy skies"], 2)`\,\%/y$
 for all- and cloudy-skies, respectively, in contrast to spring and summer.
-The trends for autumn are also significant but smaller ($`r signif(spp$"Trend [%/year]"[spp$Season == "Autumn" & spp$"Sky Condition" == "All skies"], 2)`$
+The trends for autumn are also significant but smaller ($`r signif(spp$"Trend [%/year]"[spp$Season == "Autumn" & spp$"Sky condition" == "All skies"], 2)`$
 and
-$`r signif(spp$"Trend [%/year]"[spp$Season == "Autumn" & spp$"Sky Condition" == "Cloudy skies"], 2)`\,\%/y$
+$`r signif(spp$"Trend [%/year]"[spp$Season == "Autumn" & spp$"Sky condition" == "Cloudy skies"], 2)`\,\%/y$
 for all- and cloudy-skies, respectively).
@@
 The trend for clear skies is largest in winter
-($`r signif(spp$"Trend [%/year]"[spp$Season == "Winter" & spp$"Sky Condition" == "Clear skies"], 2)`\,\%/y$) 
+($`r signif(spp$"Trend [%/year]"[spp$Season == "Winter" & spp$"Sky condition" == "Clear skies"], 2)`\,\%/y$) 
 and negligible in spring, summer and autumn.
 
-Using the CUSUMs of the monthly departures for all- and cloudy-skies, we observed a 20-year period starting around 2000 where the CUSUMs remain relatively stable, with a steep decline before and a steep increase after.
+Using the CUSUMs of the monthly departures for all and cloudy skies, we observed a 20-year period starting around 2000 where the CUSUMs remain relatively stable, with a steep decline before and a steep increase after.
 The rather smooth course of the CUSUMs suggests that no important
@@
 Continued observations with a collocated pyrheliometer, which started in 2016, will allow us to further investigate the variability of solar radiation at ground level in Thessaloniki. 
-Also, additional data of cloudiness, aerosols, atmospheric water vapour, etc., will allow better attribution and quantification of the effects of these factors on SRD.
+Also, additional data of cloudiness, aerosols, atmospheric water vapour, etc., will allow better attribution and quantification of the effects of these factors on SDR.

### 2024-01-02 MDPI_submission.tex
Commit 2024-01-02 18:13

--- 2024-01-02 MDPI_submission.tex
@@
 \toprule
-Sky condition & Season & Trend [\%/year] & Trend S.E. ($2\sigma$) & Pearson correl. & Trend p-value\\
+\textbf{Sky Condition} & \textbf{Season} & \textbf{Trend [\%/year]} & \textbf{Trend S.E. ($2\sigma$)} & \textbf{Pearson Correl.} & \textbf{Trend p-Value}\\
 \midrule
@@
 
-\begin{figure}[h!]
-    \begin{adjustwidth}{-\extralength}{0cm}
-        {\centering 
-            \includegraphics[width=1\linewidth]{./images/SeasonalMTrendsTogether3-2}   % Seasonal from Monthly
-        }
-        \caption{Monthly mean anomalies of SDR by season (rows of plots) for the three sky conditions (columns of plots). The black lines represent the linear trends.}\label{fig:seasonalALL}
-    \end{adjustwidth}
-\end{figure}
+\FloatBarrier
 
@@
 
-Using the CUSUMs of the monthly departures for all- and cloudy-skies, we
+Using the CUSUMs of the monthly departures for all and cloudy skies, we
 observed a 20-year period starting around 2000 where the CUSUMs remain
@@
 cloudiness, aerosols, atmospheric water vapour, etc., will allow better
-attribution and quantification of the effects of these factors on SRD.
+attribution and quantification of the effects of these factors on SDR.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\authorcontributions{Conceptualization, methodology, software, data
+analysis and original writing---draft preparation A.N.; writing---review
+and editing, supervision, A.B.; data curation, C.M. All authors have
+read and agreed to the published version of the manuscript.}
 

### 2024-01-02 MDPI_submission.Rmd
Commit 2024-01-02 19:12

@@
 
-<!-- \begin{figure}[h!] -->
-<!--     \begin{adjustwidth}{-\extralength}{0cm} -->
-<!--         {\centering  -->
-<!--             \includegraphics[width=1\linewidth]{./images/SeasonalMTrendsTogether3-2}   % Seasonal from Monthly -->
-<!--         } -->
-<!--     \end{adjustwidth} -->
-<!--     \caption{Time serries of monthly mean anomalies of SDR by season (rows of plots) for the three sky conditions (columns of plots).
-The black lines represent the linear trends, (\textbf{a}--\textbf{c}) winter, (\textbf{d}--\textbf{f}) spring, (\textbf{g}--\textbf{i}) summer, (\textbf{j}--\textbf{l}) autumn, and also (\textbf{a}, \textbf{d}, \textbf{g}, \textbf{j}) all-skies, (\textbf{b}, \textbf{e}, \textbf{h}, \textbf{k}) clear-skies and (\textbf{c}, \textbf{f}, \textbf{i}, \textbf{l}) cloudy-skies.}\label{fig:seasonalALL} -->
-<!-- \end{figure} -->
-
-\vspace{+100pt}
+\begin{figure}[h!]
+    \begin{adjustwidth}{-\extralength}{0cm}
+        {\centering 
+            \includegraphics[width=1\linewidth]{./images/SeasonalMTrendsTogether3-2}   % Seasonal from Monthly
+        }
+    \end{adjustwidth}
+    \caption{Time serries of monthly mean anomalies of SDR by season (rows of plots) for the three sky conditions (columns of plots).
+The black lines represent the linear trends, (\textbf{a}--\textbf{c}) winter, (\textbf{d}--\textbf{f}) spring, (\textbf{g}--\textbf{i}) summer, (\textbf{j}--\textbf{l}) autumn, and also (\textbf{a}, \textbf{d}, \textbf{g}, \textbf{j}) all-skies, (\textbf{b}, \textbf{e}, \textbf{h}, \textbf{k}) clear-skies and (\textbf{c}, \textbf{f}, \textbf{i}, \textbf{l}) cloudy-skies.}\label{fig:seasonalALL}
+\end{figure}
 

### 2024-01-02 MDPI_submission.tex
Commit 2024-01-02 19:12

@@
 
+\begin{figure}[h!]
+    \begin{adjustwidth}{-\extralength}{0cm}
+        {\centering 
+            \includegraphics[width=1\linewidth]{./images/SeasonalMTrendsTogether3-2}   % Seasonal from Monthly
+        }
+    \end{adjustwidth}
+    \caption{Time serries of monthly mean anomalies of SDR by season (rows of plots) for the three sky conditions (columns of plots).
+The black lines represent the linear trends, (\textbf{a}--\textbf{c}) winter, (\textbf{d}--\textbf{f}) spring, (\textbf{g}--\textbf{i}) summer, (\textbf{j}--\textbf{l}) autumn, and also (\textbf{a}, \textbf{d}, \textbf{g}, \textbf{j}) all-skies, (\textbf{b}, \textbf{e}, \textbf{h}, \textbf{k}) clear-skies and (\textbf{c}, \textbf{f}, \textbf{i}, \textbf{l}) cloudy-skies.}\label{fig:seasonalALL}
+\end{figure}
+
 \FloatBarrier

### 2024-01-02 MDPI_submission.Rmd
Commit 2024-01-02 21:12

--- 2024-01-02 MDPI_submission.Rmd
@@
 
-\vspace{+100pt}
+\FloatBarrier
 
-\FloatBarrier
+\ \
+
+\ \
+
+\ \
+
+\ \
+
+\ \
+
+\ \
+
+\ \
+
+\ \
+
+\ \
+
+\ \
+
+\ \
+
+\ \
+
+\ \
+
+\ \
+
+\ \
+
+\ \
+
+\ \
 
@@
 
-We have analyzed a 30-year dataset of SRD measurements in Thessaloniki Greece 
-(`r paste(year(min(ALL_1_daily_DESEAS$Date)), " -- ", year(max(ALL_1_daily_DESEAS$Date)))`), aiming to identify the long-term variability of
-solar irradiance under different sky conditions.
-Under all-sky conditions 
-there is a positive trend in SDR of $`r alltrend`\,\%/y$ (brightening). 
-A previous study [@Bais2013] for the period 1993 -- 2011 reported also a positive trend of $0.33\,\%/y$.
-The slight increase of this trend indicates that the brightening of SDR continues and is likely caused by continuing decreases in aerosol optical depth and the optical thickness of clouds over the area.
-A smaller trend has
-been found under clear-sky conditions ($`r cleartrend`\,\%/y$) which supports the
-notion that part of the brightening is caused by decreasing aerosols.
-@Siomos2020 have shown that aerosol optical depth over
-Thessaloniki is decreasing constantly at least up to 2018.
-The attenuation
-of SDR by aerosols over Europe has been proposed as major factor by @Wild2021.
-Unfortunately, for this study aerosol data for the entire period were
+We have analyzed a 30-year dataset of SRD measurements in Thessaloniki, Greece 
+(`r paste(year(min(ALL_1_daily_DESEAS$Date)), " -- ", year(max(ALL_1_daily_DESEAS$Date)))`), aiming to identify the long-term 
+variability of solar irradiance under different sky conditions.
+Under all-sky conditions, there is a positive trend in SDR of 
+$`r alltrend`\,\%/y$ (brightening). 
+A previous study [@Bais2013] for the period 1993--2011 also reported a
+positive trend of $0.33\,\%/y$.
+The slight increase in this trend indicates that the brightening of SDR
+continues and is likely caused by continuing decreases in aerosol optical
+depth and the optical thickness of clouds over the area.
+A smaller trend has been found under clear-sky conditions
+($`r cleartrend`\,\%/y$), which supports the notion that part of the
+brightening is caused by decreasing aerosols.
+@Siomos2020 showed that aerosol optical depth over Thessaloniki
+was decreasing constantly, at least up until 2018.
+The attenuation of SDR by aerosols over Europe has been proposed as
+major factor by @Wild2021.
+Unfortunately, for this study, aerosol data for the entire period were
 not available in order to quantify their effect on SDR. 
 The brightening effect on SDR under cloudy-sky conditions
-($`r cloudtrend`\,\%/y$), suggests that cloud optical thickness is also decreasing
-during this period.
-As long term data of cloud optical thickness are
+($`r cloudtrend`\,\%/y$) suggests that cloud optical thickness 
+is also decreasing during this period.
+As long-term data of cloud optical thickness are
 also not available for the region, we cannot draw quantitative
@@
 
-The observed brightening on SDR over Thessaloniki is dependent on SZA (larger SZAs lead to stronger brightening).
-The trend is also dependent on season, with winter showing the strongest statistically significant trend of
+The observed brightening on SDR over Thessaloniki is dependent on
+SZA (larger SZAs lead to stronger brightening).
+The trend is also dependent on season, with winter showing the 
+strongest statistically significant trend of
 $`r signif(spp$"Trend [%/year]"[spp$Season == "Winter" & spp$"Sky condition" == "All skies"], 2)`$

### 2024-01-02 MDPI_submission.tex
Commit 2024-01-02 21:12

--- 2024-01-02 MDPI_submission.tex
@@
 
+~~
+
+~~
+
+~~
+
+~~
+
+~~
+
+~~
+
+~~
+
+~~
+
+~~
+
+~~
+
+~~
+
+~~
+
+~~
+
+~~
+
+~~
+
+~~
+
+~~
+
 \hypertarget{conclusions}{%
@@
 
-We have analyzed a 30-year dataset of SRD measurements in Thessaloniki
+We have analyzed a 30-year dataset of SRD measurements in Thessaloniki,
 Greece (1993 -- 2023), aiming to identify the long-term variability of
 solar irradiance under different sky conditions. Under all-sky
-conditions there is a positive trend in SDR of \(0.38\,\%/y\)
-(brightening). A previous study \citep{Bais2013} for the period 1993 --
-2011 reported also a positive trend of \(0.33\,\%/y\). The slight
-increase of this trend indicates that the brightening of SDR continues
+conditions, there is a positive trend in SDR of \(0.38\,\%/y\)
+(brightening). A previous study \citep{Bais2013} for the period
+1993--2011 also reported a positive trend of \(0.33\,\%/y\). The slight
+increase in this trend indicates that the brightening of SDR continues
 and is likely caused by continuing decreases in aerosol optical depth
 and the optical thickness of clouds over the area. A smaller trend has
-been found under clear-sky conditions (\(0.097\,\%/y\)) which supports
+been found under clear-sky conditions (\(0.097\,\%/y\)), which supports
 the notion that part of the brightening is caused by decreasing
-aerosols. \citet{Siomos2020} have shown that aerosol optical depth over
-Thessaloniki is decreasing constantly at least up to 2018. The
+aerosols. \citet{Siomos2020} showed that aerosol optical depth over
+Thessaloniki was decreasing constantly, at least up until 2018. The
 attenuation of SDR by aerosols over Europe has been proposed as major
-factor by \citet{Wild2021}. Unfortunately, for this study aerosol data
+factor by \citet{Wild2021}. Unfortunately, for this study, aerosol data
 for the entire period were not available in order to quantify their
 effect on SDR. The brightening effect on SDR under cloudy-sky conditions
-(\(0.41\,\%/y\)), suggests that cloud optical thickness is also
-decreasing during this period. As long term data of cloud optical
+(\(0.41\,\%/y\)) suggests that cloud optical thickness is also
+decreasing during this period. As long-term data of cloud optical
 thickness are also not available for the region, we cannot draw

### 2024-01-02 MDPI_submission.Rmd
Commit 2024-01-02 22:03

@@
 
-<!-- \begin{figure}[h!] -->
-<!--     \begin{adjustwidth}{-\extralength}{0cm} -->
-<!--         {\centering  -->
-<!--             \includegraphics[width=1\linewidth]{./images/SeasonalMTrendsTogether3-2}   % Seasonal from Monthly -->
-<!--         } -->
-<!--     \end{adjustwidth} -->
-<!--     \caption{Time serries of monthly mean anomalies of SDR by season (rows of plots) for the three sky conditions (columns of plots).
-The black lines represent the linear trends, (\textbf{a}--\textbf{c}) winter, (\textbf{d}--\textbf{f}) spring, (\textbf{g}--\textbf{i}) summer, (\textbf{j}--\textbf{l}) autumn, and also (\textbf{a}, \textbf{d}, \textbf{g}, \textbf{j}) all-skies, (\textbf{b}, \textbf{e}, \textbf{h}, \textbf{k}) clear-skies and (\textbf{c}, \textbf{f}, \textbf{i}, \textbf{l}) cloudy-skies.}\label{fig:seasonalALL} -->
-<!-- \end{figure} -->
-
-\FloatBarrier
-
-\ \
-
-\ \
-
-\ \
-
-\ \
-
-\ \
-
-\ \
-
-\ \
-
-\ \
-
-\ \
-
-\ \
-
-\ \
-
-\ \
-
-\ \
-
-\ \
-
-\ \
-
-\ \
-
-\ \
+\begin{figure}[h!]
+    \begin{adjustwidth}{-\extralength}{0cm}
+        {\centering
+            \includegraphics[width=1\linewidth]{./images/SeasonalMTrendsTogether3-2}   % Seasonal from Monthly
+        }
+    \end{adjustwidth}
+    \caption{Time serries of monthly mean anomalies of SDR by season (rows of plots) for the three sky conditions (columns of plots).
+The black lines represent the linear trends, (\textbf{a}--\textbf{c}) winter, (\textbf{d}--\textbf{f}) spring, (\textbf{g}--\textbf{i}) summer, (\textbf{j}--\textbf{l}) autumn, and also (\textbf{a}, \textbf{d}, \textbf{g}, \textbf{j}) all-skies, (\textbf{b}, \textbf{e}, \textbf{h}, \textbf{k}) clear-skies and (\textbf{c}, \textbf{f}, \textbf{i}, \textbf{l}) cloudy-skies.}\label{fig:seasonalALL}
+\end{figure}
 
@@
 conclusions.
-<!-- @Lozano2023 reported a Cloud Radiative Forcing (CRF) SDR trend of $1.22\,W/m^2/y$ for the city of Granada Spain ($37^\circ\,10'\,$N, $3^\circ\,37'\,$E, $680\,$m\ a.s.l.).  -->
- <!-- (37.16◦ N, 3.61◦ W, 680 m a.s.l.), 37° 9' 35.9994" N  3° 36' 35.9994" W    ($37^\circ\,10'\,$N, $3^\circ\,37'\,$E, $680\,$m\ a.s.l.)-->
-<!-- We found a comparable result, for the same parameter, using the long term trend of all- and clear-skies, as follows $\text{CRF} = 1.46 - 0.501 = `r 1.46 - 0.501`\,W/m^2/y$. -->
 
@@
 ($`r signif(spp$"Trend [%/year]"[spp$Season == "Winter" & spp$"Sky condition" == "Clear skies"], 2)`\,\%/y$) 
-and negligible in spring, summer and autumn.
+and negligible in spring, summer, and autumn.
 

### 2024-01-02 MDPI_submission.tex
Commit 2024-01-02 22:03

@@
 
-\FloatBarrier
-
-~~
-
-~~
-
-~~
-
-~~
-
-~~
-
-~~
-
-~~
-
-~~
-
-~~
-
-~~
-
-~~
-
-~~
-
-~~
-
-~~
-
-~~
-
-~~
-
-~~
+\begin{figure}[h!]
+    \begin{adjustwidth}{-\extralength}{0cm}
+        {\centering
+            \includegraphics[width=1\linewidth]{./images/SeasonalMTrendsTogether3-2}   % Seasonal from Monthly
+        }
+    \end{adjustwidth}
+    \caption{Time serries of monthly mean anomalies of SDR by season (rows of plots) for the three sky conditions (columns of plots).
+The black lines represent the linear trends, (\textbf{a}--\textbf{c}) winter, (\textbf{d}--\textbf{f}) spring, (\textbf{g}--\textbf{i}) summer, (\textbf{j}--\textbf{l}) autumn, and also (\textbf{a}, \textbf{d}, \textbf{g}, \textbf{j}) all-skies, (\textbf{b}, \textbf{e}, \textbf{h}, \textbf{k}) clear-skies and (\textbf{c}, \textbf{f}, \textbf{i}, \textbf{l}) cloudy-skies.}\label{fig:seasonalALL}
+\end{figure}
 
@@
 and cloudy-skies, respectively). The trend for clear skies is largest in
-winter (\(0.4\,\%/y\)) and negligible in spring, summer and autumn.
+winter (\(0.4\,\%/y\)) and negligible in spring, summer, and autumn.
 

### 2024-01-02 appendix.tex
Commit 2024-01-01 18:12
Commit 2024-01-02 22:03

@@
 \section[\appendixname~\thesection]{}
+\label{appendix}
 
@@
     }
-    \caption{Anomalies (\%) of the daily clear-sky SDR, relative to climatological values for 1993 -- 2023.
-The black line shows the long term linear trend for clear-sky conditions.}\label{fig:trendCLEAR}
+    \caption{Anomalies (\%) of the daily clear-sky SDR, relative to climatological values for 1993--2023.
+The black line shows the long-term linear trend for clear-sky conditions.}\label{fig:trendCLEAR}
 \end{figure}
@@
     }
-    \caption{Anomalies (\%) of the daily cloud-sky SDR, relative to climatological values for 1993 -- 2023.
-The black line shows the long term linear trend for cloud-sky conditions.}\label{fig:trendCLOUD}
+    \caption{Anomalies (\%) of the daily cloud-sky SDR, relative to climatological values for 1993--2023.
+The black line shows the long-term linear trend for cloud-sky conditions.}\label{fig:trendCLOUD}
 \end{figure}
@@
         }
-        \caption{Long term trends of SDR as a function of SZA separately form morning
+    \end{adjustwidth}
+    \caption{Long-term trends of SDR as a function of SZA separately form morning
            and afternoon periods, by season (rows of plots) for the three sky
            conditions (columns of plots).
-               Solid shapes represent statistical significant trends ($p<0.005$).
+           Solid shapes represent statistical significant trends ($p<0.005$), (\textbf{a}--\textbf{c}) winter, (\textbf{d}--\textbf{f}) spring, (\textbf{g}--\textbf{i}) summer, (\textbf{j}--\textbf{l}) autumn, and also (\textbf{a}, \textbf{d}, \textbf{g}, \textbf{j}) all-skies, (\textbf{b}, \textbf{e}, \textbf{h}, \textbf{k}) clear-skies and (\textbf{c}, \textbf{f}, \textbf{i}, \textbf{l}) cloudy-skies.
            Cases where $p<0.005$ or with less than $85$ observations may be
            missing from view.}\label{fig:SZAtrendSeason}
-    \end{adjustwidth}
 \end{figure}

### 2024-01-07 Build_174/MDPI_submission_B174.Rmd
Commit 2024-01-07 21:08

--- 2023-12-19 MDPI_submission.Rmd
@@
 abstract: |
-  The shortwave downward solar irradiance (SDR) is an important factor that drives climate processes, production and can affect all living organisms.
+  The shortwave downward solar irradiance (SDR) is an important factor that drives climate processes, energy production and can affect all living organisms. 
   Observations of SDR at different locations around the world with different environmental characteristics have been used to investigate its long-term variability and trends at different time scales.
 Periods of positive trends are referred as brightening periods and of negative trends as dimming periods.
-  Here we studied 30 years of pyranometer data in Thessaloniki, Greece, under three types of sky conditions (clear sky, cloudy sky and all sky).
+  In this study we have used 30 years of pyranometer data in Thessaloniki, Greece, to investigate the variability of SDR under three types of sky conditions (clear-, cloudy- and all-sky). 
   The clear-sky data were identified by applying a cloud screening algorithm. 
-We found a positive trend of $0.38\,\%/\text{year}$ for all-sky, $0.35\,\%/\text{year}$ for clear-sky conditions, and $-0.28\,\%/\text{year}$ for cloudy conditions.
-  We have also investigated the consistency of these trends, their seasonal variability, and the effect of the solar zenith angle. 
-  We have found that for all-sky and clear-sky conditions the SDR trend is positive in winter ($0.7$ and $0.8\,\%/\text{year}$, respectively) and autumn ($~0.4\,\%/\text{year}$), while under cloudy skies the trend is negative ($-0.9\,\%/\text{year}$ in winter and $-0.4\,\%/\text{year}$ in autumn).
-  In spring and summer the trend is very close to zero, irrespective of sky conditions. 
+  We have found a positive trend of $0.38\,\%/\text{year}$ for all-sky, $\sim 0.1\,\%/\text{year}$ for clear-sky, and $0.41\,\%/\text{year}$ for cloudy conditions.
+  The consistency of these trends, their seasonal variability, and the effect of the solar zenith angle have also been investigated.
+  Under all three sky categories the SDR trend is stronger in winter with $0.7$, $0.4$ and $0.76\,\%/\text{year}$, respectively for all-, clear- and cloudy-skies. 
+  The next larger seasonal trends are in automn $0.42$ and $0.19\,\%/\text{year}$, for all- and cloudy-skies respectively. 
+  The rest seasonal trends are significant smaller, close to zero, with a negative values in summer, for clear- and cloudy-skies.
   The SDR trend is increasing with increasing solar zenith angle, except under cloudy skies where the trend is highly variable and close to zero.
-  Finally, we identified some anomalies in the long term SDR trends for all sky conditions by examining the cumulative sums of monthly anomalies from the climatological mean.
+  Finally, we are discussing shorter-term variations in SDR anomalies by examining the patterns of the cumulative sums of monthly anomalies from the climatological mean, both before and after removing the long-term trend.
 
@@
 <!-- Trends Europe Brazil -->
-A similar pattern was identified, with the breaking point around 1980, for stations in Central Europe [@Wild2021] and Brazil [@Yamasoe2021].
+A similar pattern was identified, with a breaking point around 1980, for stations in Central Europe [@Wild2021] and Brazil [@Yamasoe2021].
 <!-- Trends Global -->
-On global scale, an artificial Intelligence aided spatial analysis on continental level with data from multiple stations reach similar conclusions for these regions and for the global trend [@Yuan2021].
+On global scale, an Artificial Intelligence aided spatial analysis on continental level with data from multiple stations reached similar conclusions for these regions and for the global trend [@Yuan2021].
 
 There is a consensus among researchers that the major factor affecting the variability of SDR attenuation is the interactions of solar radiation with atmospheric aerosols and clouds.
-Those interactions, among other factors, have been analysed with models [@Li2016; @Samset2018], showing the existence of feedback mechanisms between the two.
-Similar findings have been shown in observational data [@Schwarz2020; @Ohvril2009; @Zerefos2009; @Xia2007 and references therein].
-<!-- Sun variability: should cite that there is no significant trend found Wild2021 -->
+Those interactions, among other factors, have been analyzed with models [@Li2016; @Samset2018], showing the existence of feedback mechanisms between the two.
+Similar findings have been shown from the analysis of observations at other locations [@Schwarz2020; @Ohvril2009; @Zerefos2009; @Xia2007 and references therein].
+In the Mediterranean region aerosols have been recognized as an important factor affecting the penetration of solar radiation to the surface [@Fountoulakis2016; @Siomos2018; @Gkikas2013; @Lozano2021].
+These studies investigated the long-term trend in aerosol optical depth, which has been found to decrease in the last three decades, the transport and composition of aerosols, and their radiative effects.
 
 Due to the significant spatial and temporal variability of the trends, and the contributing factors, there is a constant need to monitor and investigate SDR in different sites in order to estimate the degree of variability, and its relation to the local conditions.
-In this study, we examine the trends of SDR, using ground-based measurements at Thessaloniki, Greece, for the period `r year(min(ALL_1_daily_DESEAS$Date))` to `r year(max(ALL_1_daily_DESEAS$Date))`, as derived from a CM-21 pyranometer.
+In this study, we examine the trends of SDR, using ground-based measurements at Thessaloniki, Greece, for the period `r year(min(ALL_1_daily_DESEAS$Date))` to `r year(max(ALL_1_daily_DESEAS$Date))`.
 <!-- goals -->
-We reevaluated and extended the dataset used by @Bais2013, we applied a different algorithm for the identification of clear-/cloud-sky instances [@Reno2016; @Reno2012a], and we derived the SDR trends for the period of study, under different sky conditions (all-sky, clear-sky and cloud-sky).
+We re-evaluated and extended the dataset used by @Bais2013, we applied a different algorithm for the identification of clear-/cloud-sky instances [@Reno2016; @Reno2012a], and we derived the SDR trends for the period of study, under different sky conditions (all-sky, clear-sky and cloud-sky).
 Finally, we investigated the dependence of the trends on solar zenith angle and season.
@@
 The monitoring site is located
-near the city centre thus we expect that measurements are affected by the urban environment.
+near the city center thus we expect that measurements are affected by the urban environment, mainly by aerosols.
 During the study period, the pyranometer has been
 independently calibrated three times at the Meteorologisches
-Observatorium Lindenberg, DWD, verifying that stability of the
+Observatorium Lindenberg, DWD, verifying that the stability of the
 instrument's sensitivity is better than $0.7\%$ relative to the initial
@@
 @Reno2016, which requires the definition of some site specific parameters.
-These parameters, were produced by an iterative process, as the original authors proposed and it are discussed in the next section.
+These parameters were determined by an iterative process, as the original authors proposed and are discussed in the next section.
 
@@
 We note that all methods have some subjectivity in the definition of clear or cloudy sky cases.
-As a result, the details of the definition are site specific and they rely on a combination of thresholds and comparisons with ideal radiation models and statistical analysis of different signal metrics.
+As a result, the details of the definition are site specific, and they rely on a combination of thresholds and comparisons with ideal radiation models and statistical analysis of different signal metrics.
 The CSid algorithm was calibrated with the main focus, to identify the presence of clouds. Despite the fine-tuning
 of the procedure, in a few marginal cases false positive or false
-negative results were identified by manual inspection. However, due to
+negative results were identified by manual inspection. 
+However, due to
 their small number, they cannot affect the final results of the study.
-For completeness, we provide below a brief overview of the clear-sky
-identification algorithm (CSid), along with the site specific
-thresholds.
+For completeness, we provide below a brief overview of the CSid algorithm, along with the site specific thresholds.
 
@@
 \begin{equation}
-\text{SDR}_\text{Haurwitz} = 1098 \times \cos(\theta) \times \exp \left( \frac{ - 0.057}{\cos(\theta)} \right) \label{eq:hau}
+\text{SDR}_\text{Haurwitz} = 1098 \times \cos(\theta) \times \exp \left( \frac{ - 0.059}{\cos(\theta)} \right) \label{eq:hau}
 \end{equation}
-where $\theta$ is the solar zenith angle (SZA).
+where $\theta$ is the SZA.
 
@@
 
-The linear trends were calculated using an first order autoregressive model
- with lag of 1 day with fitting method of the "maximum likelihood" [@Gardner1980; @Jones1980], by implementing the function "arima" from the library "stats" of the R programming language [@RCT2023].
-The trends were reported together with the $2\sigma$ errors (Table\ \ref{tab:trendtable}).
+The linear trends were calculated using a first order autoregressive model
+ with lag of 1 day using the "maximum likelihood" fitting method [@Gardner1980; @Jones1980], by implementing the function "arima" from the library "stats" of the R programming language [@RCT2023].
+The trends were reported together with the $2\sigma$ errors.
 
@@
 are statistically significant at least at the $95\,\%$ confidence level, as
-they are larger than the corresponding $2\sigma$ errors.
+they are larger than the corresponding $2\sigma$ errors, despite the small values of R, which is due to the large variability of the daily values.
 The clear-sky trend is
 very small suggesting a small effect from aerosols and water vapor which
-are the dominant factors of the SDR variability. 
-In contrast the large
+are the dominant factors of the SDR variability [@Fountoulakis2016; @Siomos2018; @Yu2022].
+In contrast, the large
 positive trend of SDR under cloudy skies can be attributed to reduction
 in cloud cover and/or cloud optical depth. 
-Lack of continuous cloud
-observations that could support these findings does not allow drawing
+Lack of continuous observations of cloud optical thickness that could support these findings does not allow drawing
 firm conclusions. However, there are indications that the total
@@
 
-Although the year-to-year variability of the anomalies (Figure \ref{fig:trendALL} and Figures\ \ref{fig:trendCLEAR}, \ref{fig:trendCLOUD} in Appendix), shows a rather homogeneous behaviour, plots of the cumulative sums (CUSUM) [@Regier2019] of the anomalies can reveal different structures in the records of all three sky conditions.
+Although the year-to-year variability of the anomalies (Figure \ref{fig:trendALL} and Figures\ \ref{fig:trendCLEAR}, \ref{fig:trendCLOUD} in Appendix), shows a rather homogeneous behavior, plots of the cumulative sums (CUSUM) [@Regier2019] of the anomalies can reveal different structures in the records of all three sky conditions.
 For time series with a uniform trend, we would expect
@@
 
-We have analyzed a 30 year dataset of SRD measurements in Thessaloniki Greece 
+We have analyzed a 30-year dataset of SRD measurements in Thessaloniki Greece 
 (`r paste(year(min(ALL_1_daily_DESEAS$Date)), " -- ", year(max(ALL_1_daily_DESEAS$Date)))`) aiming to identify the long term variability of
@@
 Thessaloniki is decreasing constantly at least up to 2018.
+The attenuation
+of SDR by aerosols over Europe has been proposed as major factor by @Wild2021.
 Unfortunately, for this study aerosol data for the entire period were
-not available in order to quantify their effect on SDR. The attenuation
-of SDR by aerosols over Europe has been proposed as major factor by @Wild2021.
+not available in order to quantify their effect on SDR. 
 The brightening effect on SDR under cloudy-sky conditions
@@
 conclusions.
-@Lozano2023 reported a Cloud Radiative Forcing (CRF) trend of $1.22\,W/m^2/y$ for the SDR for the city of Granada Spain ($37.16^\circ\,$N, $3.61^\circ\,$E, $680\,$m\ a.s.l.). 
- <!-- (37.16◦ N, 3.61◦ W, 680 m a.s.l.), 37° 9' 35.9994" N  3° 36' 35.9994" W-->
-We found a comparable result, for the same parameter, using the long term trend of all- and clear-skies, as follows $CRF = 1.46 - 0.501 = `r 1.46 - 0.501`\,W/m^2/y$.
+<!-- @Lozano2023 reported a Cloud Radiative Forcing (CRF) SDR trend of $1.22\,W/m^2/y$ for the city of Granada Spain ($37^\circ\,10'\,$N, $3^\circ\,37'\,$E, $680\,$m\ a.s.l.).  -->
+ <!-- (37.16◦ N, 3.61◦ W, 680 m a.s.l.), 37° 9' 35.9994" N  3° 36' 35.9994" W    ($37^\circ\,10'\,$N, $3^\circ\,37'\,$E, $680\,$m\ a.s.l.)-->
+<!-- We found a comparable result, for the same parameter, using the long term trend of all- and clear-skies, as follows $\text{CRF} = 1.46 - 0.501 = `r 1.46 - 0.501`\,W/m^2/y$. -->
 

### 2024-01-07 Build_174/MDPI_submission_B174.tex
Commit 2024-01-07 21:08

--- 2023-12-19 MDPI_submission.tex
@@
 \abstract{The shortwave downward solar irradiance (SDR) is an important
-factor that drives climate processes, production and can affect all
-living organisms. Observations of SDR at different locations around the
-world with different environmental characteristics have been used to
+factor that drives climate processes, energy production and can affect
+all living organisms. Observations of SDR at different locations around
+the world with different environmental characteristics have been used to
 investigate its long-term variability and trends at different time
 scales. Periods of positive trends are referred as brightening periods
-and of negative trends as dimming periods. Here we studied 30 years of
-pyranometer data in Thessaloniki, Greece, under three types of sky
-conditions (clear sky, cloudy sky and all sky). The clear-sky data were
-identified by applying a cloud screening algorithm. We found a positive
-trend of \(0.38\,\%/\text{year}\) for all-sky, \(0.35\,\%/\text{year}\)
-for clear-sky conditions, and \(-0.28\,\%/\text{year}\) for cloudy
-conditions. We have also investigated the consistency of these trends,
-their seasonal variability, and the effect of the solar zenith angle. We
-have found that for all-sky and clear-sky conditions the SDR trend is
-positive in winter (\(0.7\) and \(0.8\,\%/\text{year}\), respectively)
-and autumn (\(~0.4\,\%/\text{year}\)), while under cloudy skies the
-trend is negative (\(-0.9\,\%/\text{year}\) in winter and
-\(-0.4\,\%/\text{year}\) in autumn). In spring and summer the trend is
-very close to zero, irrespective of sky conditions. The SDR trend is
+and of negative trends as dimming periods. In this study we have used 30
+years of pyranometer data in Thessaloniki, Greece, to investigate the
+variability of SDR under three types of sky conditions (clear-, cloudy-
+and all-sky). The clear-sky data were identified by applying a cloud
+screening algorithm. We have found a positive trend of
+\(0.38\,\%/\text{year}\) for all-sky, \(\sim 0.1\,\%/\text{year}\) for
+clear-sky, and \(0.41\,\%/\text{year}\) for cloudy conditions. The
+consistency of these trends, their seasonal variability, and the effect
+of the solar zenith angle have also been investigated. Under all three
+sky categories the SDR trend is stronger in winter with \(0.7\), \(0.4\)
+and \(0.76\,\%/\text{year}\), respectively for all-, clear- and
+cloudy-skies. The next larger seasonal trends are in automn \(0.42\) and
+\(0.19\,\%/\text{year}\), for all- and cloudy-skies respectively. The
+rest seasonal trends are significant smaller, close to zero, with a
+negative values in summer, for clear- and cloudy-skies. The SDR trend is
 increasing with increasing solar zenith angle, except under cloudy skies
-where the trend is highly variable and close to zero. Finally, we
-identified some anomalies in the long term SDR trends for all sky
-conditions by examining the cumulative sums of monthly anomalies from
-the climatological mean.}
+where the trend is highly variable and close to zero. Finally, we are
+discussing shorter-term variations in SDR anomalies by examining the
+patterns of the cumulative sums of monthly anomalies from the
+climatological mean, both before and after removing the long-term
+trend.}
 
@@
 reported until about 2000, followed by a brightening period
-\citep{Yang2021}. A similar pattern was identified, with the breaking
+\citep{Yang2021}. A similar pattern was identified, with a breaking
 point around 1980, for stations in Central Europe \citep{Wild2021} and
-Brazil \citep{Yamasoe2021}. On global scale, an artificial Intelligence
+Brazil \citep{Yamasoe2021}. On global scale, an Artificial Intelligence
 aided spatial analysis on continental level with data from multiple
-stations reach similar conclusions for these regions and for the global
-trend \citep{Yuan2021}.
+stations reached similar conclusions for these regions and for the
+global trend \citep{Yuan2021}.
 
@@
 radiation with atmospheric aerosols and clouds. Those interactions,
-among other factors, have been analysed with models
+among other factors, have been analyzed with models
 \citep{Li2016, Samset2018}, showing the existence of feedback mechanisms
-between the two. Similar findings have been shown in observational data
-\citep[ and references
-therein]{Schwarz2020, Ohvril2009, Zerefos2009, Xia2007}.
+between the two. Similar findings have been shown from the analysis of
+observations at other locations \citep[ and references
+therein]{Schwarz2020, Ohvril2009, Zerefos2009, Xia2007}. In the
+Mediterranean region aerosols have been recognized as an important
+factor affecting the penetration of solar radiation to the surface
+\citep{Fountoulakis2016, Siomos2018, Gkikas2013, Lozano2021}. These
+studies investigated the long-term trend in aerosol optical depth, which
+has been found to decrease in the last three decades, the transport and
+composition of aerosols, and their radiative effects.
 
@@
 examine the trends of SDR, using ground-based measurements at
-Thessaloniki, Greece, for the period 1993 to 2023, as derived from a
-CM-21 pyranometer. We reevaluated and extended the dataset used by
-\citet{Bais2013}, we applied a different algorithm for the
-identification of clear-/cloud-sky instances
+Thessaloniki, Greece, for the period 1993 to 2023. We re-evaluated and
+extended the dataset used by \citet{Bais2013}, we applied a different
+algorithm for the identification of clear-/cloud-sky instances
 \citep{Reno2016, Reno2012a}, and we derived the SDR trends for the
@@
 Aristotle University of Thessaloniki (\(40^\circ\,38'\,\)N,
-\(22^\circ\,57'\,\)E, \(80\,\)m~a.s.l.) since 1993. Here we use data for
-the period from 1993-04-13 to 2023-04-13. The monitoring site is located
-near the city centre thus we expect that measurements are affected by
-the urban environment. During the study period, the pyranometer has been
-independently calibrated three times at the Meteorologisches
-Observatorium Lindenberg, DWD, verifying that stability of the
-instrument's sensitivity is better than \(0.7\%\) relative to the
-initial calibration by the manufacturer. Along with SDR, the direct beam
-radiation (DNI) is also measured with a collocated Kipp \& Zonen CHP-1
-pyrheliometer since 2016-04-01. The DNI data were used as auxiliary data
-to support the selection of appropriate thresholds in the clear sky
-identification algorithm (CSid), which is discussed later. It is noted
-that the limited dataset of DNI was not used for the identification of
-clear sky cases in the entire SDR series to avoid any selection bias due
-to unequal length of the two datasets. There are four distinct steps in
-the creation of the dataset analyzed here: a)~the acquisition of
-radiation measurements from the sensors, b)~the data quality check,
-c)~the identification of ``clear sky'' conditions from the radiometric
-data, and d)~the aggregation of data and trend analysis.
+\(22^\circ\,57'\,\)E, \(80\,\)m~a.s.l.). Here we use data for the period
+from 1993-04-13 to 2023-04-13. The monitoring site is located near the
+city center thus we expect that measurements are affected by the urban
+environment, mainly by aerosols. During the study period, the
+pyranometer has been independently calibrated three times at the
+Meteorologisches Observatorium Lindenberg, DWD, verifying that the
+stability of the instrument's sensitivity is better than \(0.7\%\)
+relative to the initial calibration by the manufacturer. Along with SDR,
+the direct beam radiation (DNI) is also measured with a collocated Kipp
+\& Zonen CHP-1 pyrheliometer since 2016-04-01. The DNI data were used as
+auxiliary data to support the selection of appropriate thresholds in the
+clear-sky identification algorithm (CSid), which is discussed in sect.
+\ref{CDIDalgorithm}. It is noted that the limited dataset of DNI was not
+used for the identification of clear sky cases in the entire SDR series
+to avoid any selection bias due to unequal length of the two datasets.
+There are four distinct steps in the creation of the dataset analyzed
+here: a)~the acquisition of radiation measurements from the sensors,
+b)~the data quality check, c)~the identification of ``clear sky''
+conditions from the SDR data, and d)~the aggregation of data and trend
+analysis.
 
@@
 the method proposed by \citet{Reno2016}, which requires the definition
-of some site specific parameters. These parameters, were produced by an
-iterative process, as the original authors proposed and it are discussed
-in the next section.
+of some site specific parameters. These parameters were determined by an
+iterative process, as the original authors proposed and are discussed in
+the next section.
 
@@
 \begin{equation}
-\text{SDR}_\text{Haurwitz} = 1098 \times \cos(\theta) \times \exp \left( \frac{ - 0.057}{\cos(\theta)} \right) \label{eq:hau}
-\end{equation} where \(\theta\) is the solar zenith angle (SZA).
+\text{SDR}_\text{Haurwitz} = 1098 \times \cos(\theta) \times \exp \left( \frac{ - 0.059}{\cos(\theta)} \right) \label{eq:hau}
+\end{equation} where \(\theta\) is the SZA.
 
@@
 
-The linear trends were calculated using an first order autoregressive
-model with lag of 1 day with fitting method of the ``maximum
-likelihood'' \citep{Gardner1980, Jones1980}, by implementing the
-function ``arima'' from the library ``stats'' of the R programming
-language \citep{RCT2023}. The trends were reported together with the
-\(2\sigma\) errors (Table~\ref{tab:trendtable}).
+The linear trends were calculated using a first order autoregressive
+model with lag of 1 day using the ``maximum likelihood'' fitting method
+\citep{Gardner1980, Jones1980}, by implementing the function ``arima''
+from the library ``stats'' of the R programming language
+\citep{RCT2023}. The trends were reported together with the \(2\sigma\)
+errors.
 
@@
 We calculated the linear trends of SDR, from the departures of the mean
-daily values from the daily climatology and for the three sky conditions
-(Table~\ref{tab:trendtable}). In Figure~\ref{fig:trendALL} we present
-only the time series under all-sky conditions; the plots for clear-sky
-and cloud-sky conditions, are shown in the Appendix
-(Figures~\ref{fig:trendCLEAR} and~ \ref{fig:trendCLOUD}). In the studied
-period, there is no significant break or change in the variability
-pattern of the time series. The linear trends in all three datasets are
-positive and around \(0.4\,\%/y\) for all-sky and cloudy-sky conditions,
-while for clear-skies the trend is much smaller
-(\textasciitilde{}\(0.1\,\%/y\)). The linear trends were calculated
-taking into account the autocorrelation of the time series and all three
-are statistically significant at least at the \(95\,\%\) confidence
-level, as they are larger than the corresponding \(2\sigma\) errors. The
-clear-sky trend is very small suggesting a small effect from aerosols
-and water vapor which are the dominant factors of the SDR variability.
-In contrast the large positive trend of SDR under cloudy skies can be
-attributed to reduction in cloud cover and/or cloud optical depth. Lack
-of continuous cloud observations that could support these findings does
-not allow drawing firm conclusions. However, there are indications that
-the total cloud-cover as inferred from the ERA5 analysis for the grid
-point of Thessaloniki is decreasing over the period of study.
+daily values from the daily climatology and for the three sky
+conditions. These are presented in Table~\ref{tab:trendtable} which
+contains also the \(2\sigma\) standard error, the Pearson's correlation
+coefficient R and the trend in absolute units. In
+Figure~\ref{fig:trendALL} we present only the time series under all-sky
+conditions; the plots for clear-sky and cloud-sky conditions, are shown
+in the Appendix (Figures~\ref{fig:trendCLEAR} and~
+\ref{fig:trendCLOUD}). In the studied period, there is no significant
+break or change in the variability pattern of the time series. The
+linear trends in all three datasets are positive and around
+\(0.4\,\%/y\) for all-sky and cloudy-sky conditions, while for
+clear-skies the trend is much smaller (\textasciitilde{}\(0.1\,\%/y\)).
+The linear trends were calculated taking into account the
+autocorrelation of the time series and all three are statistically
+significant at least at the \(95\,\%\) confidence level, as they are
+larger than the corresponding \(2\sigma\) errors, despite the small
+values of R, which is due to the large variability of the daily values.
+The clear-sky trend is very small suggesting a small effect from
+aerosols and water vapor which are the dominant factors of the SDR
+variability \citep{Fountoulakis2016, Siomos2018, Yu2022}. In contrast,
+the large positive trend of SDR under cloudy skies can be attributed to
+reduction in cloud cover and/or cloud optical depth. Lack of continuous
+observations of cloud optical thickness that could support these
+findings does not allow drawing firm conclusions. However, there are
+indications that the total cloud-cover as inferred from the ERA5
+analysis for the grid point of Thessaloniki is decreasing over the
+period of study. From the difference between all-sky and clear-sky SDR
+trends, expressed in \(W/m^2/y\) using the long-term mean of the
+respective datasets, the radiative effect of clouds is estimated to
+\(0.96\,W/m^2/y\). This estimate is similar to the cloud radiative
+forcing of \(1.22\,W/m^2/y\) reported for Granada, Spain
+\citep{Lozano2023}.
 
@@
 each season of the year, using the corresponding mean monthly anomalies
-(Figure~\ref{fig:seasonalALL} and Table~\ref{tab:trendseasontable}). For
-all-sky conditions the trend in SDR in winter is the largest
-(\(0.7\,\%/y\)), followed by the trend in autumn (\(0.42\,\%/y\), a
-value close to the long term trend) both statistically significant at
-the \(95\,\%\) confidence level. In spring and summer, the trends are
-much smaller and of lesser statistical significance. These seasonal
-differences indicate a possible relation of the trends in SDR to trends
-in clouds during winter and autumn. For clear-skies, the trend in winter
-is \(0.4\,\%/y\), and is associated with the decreasing trend in aerosol
-optical depth \citep{Siomos2020}, and almost half of that for all-skies,
-which is another indication of a decreasing trend in cloud optical
-thickness. In other seasons the clear-sky trend is still very small
-(below \(0.1\,\%/y\)). Finally, for cloud-skies the winter trend is the
-largest (\(0.76\,\%/y\)) and greater than for all-skies, followed by a
-much smaller trend in autumn (\(0.19\,\%/y\)).
+(Figure~\ref{fig:seasonalALL} and Table~\ref{tab:trendseasontable}).
+Table~\ref{tab:trendseasontable} contains also the \(2\sigma\) standard
+error, the Pearson's correlation coefficient R and the corresponding
+p-value. The winter linear trends exhibit generally the largest R values
+ranging between \(0.54\) and \(0.60\,\%/y\). For all-sky conditions the
+trend in SDR in winter is the largest (\(0.7\,\%/y\)), followed by the
+trend in autumn (\(0.42\,\%/y\), a value close to the long term trend)
+both statistically significant at the \(95\,\%\) confidence level. In
+spring and summer, the trends are much smaller and of lesser statistical
+significance. These seasonal differences indicate a possible relation of
+the trends in SDR to trends in clouds during winter and autumn. For
+clear-skies, the trend in winter is \(0.4\,\%/y\) and is associated with
+the decreasing trend in aerosol optical depth \citep{Siomos2020}.
+Moreover, it is almost half of that for all-skies, which is another
+indication of a decreasing trend in cloud optical thickness. In other
+seasons the clear-sky trend is very small (below \(0.1\,\%/y\)).
+Finally, for cloudy-skies the winter trend is the largest
+(\(0.76\,\%/y\)) and greater than for all-skies, followed by a much
+smaller trend in autumn (\(0.19\,\%/y\)).
 
 The trends under clear- and cloudy-sky conditions are of the same
-direction and it would be expected that their sum is similar to the
+direction, and it would be expected that their sum is similar to the
 all-sky trend. This does not happen, especially for winter, likely due
@@
 
-We have analyzed a 30 year dataset of SRD measurements in Thessaloniki
+We have analyzed a 30-year dataset of SRD measurements in Thessaloniki
 Greece (1993 -- 2023) aiming to identify the long term variability of
@@
 aerosols. \citet{Siomos2020} have shown that aerosol optical depth over
-Thessaloniki is decreasing constantly at least up to 2018.
-Unfortunately, for this study aerosol data for the entire period were
-not available in order to quantify their effect on SDR. The attenuation
-of SDR by aerosols over Europe has been proposed as major factor by
-\citet{Wild2021}. The brightening effect on SDR under cloudy-sky
-conditions (\(0.41\,\%/y\)), suggests that cloud optical thickness is
-also decreasing during this period. As long term data of cloud optical
+Thessaloniki is decreasing constantly at least up to 2018. The
+attenuation of SDR by aerosols over Europe has been proposed as major
+factor by \citet{Wild2021}. Unfortunately, for this study aerosol data
+for the entire period were not available in order to quantify their
+effect on SDR. The brightening effect on SDR under cloudy-sky conditions
+(\(0.41\,\%/y\)), suggests that cloud optical thickness is also
+decreasing during this period. As long term data of cloud optical
 thickness are also not available for the region, we cannot draw
-quantitative conclusions. \citet{Lozano2023} reported a Cloud Radiative
-Forcing (CRF) trend of \(1.22\,W/m^2/y\) for the SDR for the city of
-Granada Spain (\(37.16^\circ\,\)N, \(3.61^\circ\,\)E,
-\(680\,\)m~a.s.l.). We found a comparable result, for the same
-parameter, using the long term trend of all- and clear-skies, as follows
-\(CRF = 1.46 - 0.501 = 0.959\,W/m^2/y\).
+quantitative conclusions.
 

### 2024-09-13 MDPI_submission.Rmd
Commit 2024-09-13 10:14

@@
 
-We have analyzed a 30-year dataset of SRD measurements in Thessaloniki, Greece 
+We have analyzed a 30-year dataset of SDR measurements in Thessaloniki, Greece 
 (`r paste(year(min(ALL_1_daily_DESEAS$Date)), " -- ", year(max(ALL_1_daily_DESEAS$Date)))`), aiming to identify the long-term 

### 2024-11-02 appendix.tex
Commit 2024-11-02 22:14

@@
     \end{adjustwidth}
-    \caption{Long-term trends of SDR as a function of SZA separately form morning
+    \caption{Long-term trends of SDR as a function of SZA separately for morning
            and afternoon periods, by season (rows of plots) for the three sky
